### 2023-02-17 theory.tex
first commit

\documentclass[11pt]{article}
\setlength{\parindent}{10pt}
\usepackage [square,colon]{natbib}
\usepackage[utf8]{inputenc}
\usepackage[british]{babel}
\usepackage {comment}
\usepackage[flushleft]{threeparttable}
\renewcommand{\familydefault}{\sfdefault}
\usepackage[a4paper, margin=1.5cm]{geometry}
\usepackage{doi}
\usepackage{subcaption}
\renewcommand{\rmdefault}{phv} % Arial
\renewcommand{\sfdefault}{phv} % Arial
\usepackage[modulo]{lineno}
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\usepackage{amsmath}
\usepackage{cases}
\usepackage{float}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{authblk}
\usepackage{marginnote}
\usepackage[T1, OT1]{fontenc}
\usepackage{wrapfig}
\linenumbers
\errorcontextlines=3
\usepackage{hyperref}
\usepackage{lscape}
\linenumbers
\usepackage[explicit]{titlesec}
\usepackage{setspace}
\usepackage{xcolor}   % Make the colors used for the URL available
\hypersetup{ % Remove ugly boxes around URLs
  colorlinks, linkcolor={black!50!black}, citecolor={black!50!black}, urlcolor={black!80!black}
}

\usepackage[export]{adjustbox}
\usepackage{bibentry}
\usepackage{tikz}
\usepackage[inline,shortlabels]{enumitem}
\usepackage[all]{nowidow}

\newcommand{\aref}[1]{\textbf{Reference #1}}
\newcommand{\TODO}[1]{\textbf{TODO #1}}
\newcommand{\ian}[1]{{\textbf{\color{blue}Ian says:} \color{blue} #1} }
\newcommand{\m}{$\,\mathrm{m}$\,}
\newcommand{\cm}{$\,\mathrm{cm}$\,}
\newcommand{\mma}{$\,\mathrm{mm  \, a^{-1}}$\,}
\newcommand{\mmma}{$\,\mathrm{m^3\, a^{-1}}$\,}
\newcommand{\unit}[1]{$\mathrm{#1}$}

\newenvironment{laysummary}{
  \chapter{\centering \textbf{Plain Language Summary}}
}

\newenvironment{abstract_}{
  \chapter{\centering \textbf{Abstract}}
}

\author[1]{Ian Delaney}
\affil[1]{Institut des dynamiques de la surface terrestre (IDYST), Universit\'{e} de Lausanne, B\^{a}timent G\'{e}opolis, CH-1015 Lausanne}
\title{Sediment transport capacity can vary more in subglacial systems compared to fluvial ones}

\begin{document}
\maketitle

\ian{check for tense}

\paragraph{Key findings:}
\begin{enumerate}
\item Sediment transport capacity in pressurized subglacial channels  responds differently to changing water discharge in open fluvial channels.
\item Sediment transport capacity varies far more in subglacial channels compared to fluvial ones, even though changing channel width  in fluvial channels accommodates some of the difference.
\item Differences in the sediment transport capacity response to changing water discharge could impact processes at proglacial margins and the interpretation of sediment transport records from glaciers. 
\end{enumerate}

\abstract % 150 words
Sediment transport capacity depends largely on the velocity of water flowing through a channel and the channel width over which to mobilize sediment.
In fluvial channels, changing water discharge can be accommodated by changing water depth and width, as well as water velocity.
In subglacial channels over short time scales, however, water flows through a conduit with fixed size and is pressurized by the ice above.
As a result, changing water discharge only may result in modifying water velocity.
Parameterizations of flow in pressurized, subglacial channels and  subaerial fluvial channels show that sediment transport capacity varies  more in response to changing water discharge in subglacial channels compared to fluvial ones.
This characteristic of sediment transport capacity in subglacial systems holds true  across a wide range of channel geometries and flow conditions.
The increased variability in sediment transport capacity impacts the interpretation of records of sediment transport from glacierized catchments compared to fluvial ones.

\vspace{0.5cm}

\laysummary % 200 words
Sediment transport capacity responds differently to changing water discharge in subglacial and fluvial channels.
In both subglacial and fluvial channels, sediment transport capacity, to the first order, depends on 1) the channel width over-which to mobilize sediment and 2) the water's shear stress on sediment at the channel bottom to mobilize it.
The shear stress depends largely on water velocity.
In fluvial channels, changing water discharge will change the water depth, water velocity (controlling sediment transport capacity),  and channel width (which dictates the span over which sediment transport  will occur).
Conversely, in subglacial environments, on short times scales of hours to days, subglacial channels only experience a change in water velocity following changing water discharge, because the water flows through a subglacial conduit with a fixed geometry.
The fixed conduit geometry also means that the channel width, over-which sediment can be mobilized, does not respond to changing water discharge.
As a result, sediment transport capacity varies far more in subglacial channels compared to fluvial ones, even though changing channel width can accommodate some sediment transport variability in fluvial channels.
This increased variability has several important implications for sediment transport in glacial systems compared to fluvial ones.

%% TC:endignore
\begin{spacing}{1}
  \section{Introduction}

  Glaciers are well known for the quantities of sediment that they expel from their termini \citep{hallet1996}.
  Changing glacier dynamics, geomorphology, and hydrology in cold regions has prompted a recent focus on subglacial sediment transport processes \citep[e.g.][]{zhang2022}.
  To date, numerical models consider sediment discharge from glaciers as a function of the availability of sediment, either from the erosion of bedrock or the existence of sediment and the transport capacity of this sediment by fluvial processes below the glacier \citep{creyts2013,brinkerhoff2017,beaud2018,delaney2019}.
  These models often represent sediment transport processes similar to rivers, with the main difference being that water flows through a subglacial conduit, pressurized by the ice above \citep{rothlisberger1972}.

  When an adequate quantity of sediment is present as to be transported fluvially (i.e. in a ``transport-limited'' regime), shear-stress between the water and the sediment that it flows over is a first-order control on the transport of sediment \citep{shields1936}.
  Shear stress increases with the water velocity and roughness of the channel.
  Greater shear stress will mobilize increased quantities of sediment, along with larger sizes of sediment clasts.
  The fundamental role of this shear-stress is demonstrated in sediment transport relationships such as \citet{shields1936}.
  Sediment transport equations such as \citet{meyer1948}  and \citet{engelund1967} use shear stress, or an approximation thereof, to evaluate sediment transport capacity per unit width.

  While sediment transport over a unit width of the channel depends directly on shear-stress \citep[e.g.][]{meyer1948}, the total sediment transport capacity across a channel's cross depends on the integrated shear-stress over the entire channel bed (Figure~\ref{fig:cartoon}).
  Furthermore, as water discharge changes, the channel's cross-sectional area, as well as the water velocity, accommodates the variations.
  As a result, channel geometry plays a key role in sediment transport in both fluvial channels and subglacial ones \citep[e.g.][]{church2006,delaney2019}.

  However, in response to changing water discharge, the behavior of subglacial channels diverges from fluvial ones.
  % In fluvial channels, the channel width and height may evolve with changing flow conditions, but the underlying channel geometry responds on a far longer timescale than water discharge of the water discharge \citep[Figure~\ref{fig:cartoon}\,b and d; e.g. ][]{church2006}.
  In subglacial channels, however, the water is pressurized in a subglacial channel and the size of this channel responds to the roughness of the channel walls and the velocity of the water flowing through the channel \citep[Figure~\ref{fig:cartoon}\,a and c; e.g. ][]{rothlisberger1972}.
  Here, diurnal variations in water flux, for instance, may cause water to flow through a conduit whose size does not respond to the same timescales as the increase in water flux \citep{rothlisberger1972}.
  With respect to sediment transport, this means that increased water discharge in subglacial channels will result in greater water velocity and thus sediment transport capacity.
  At the same time, the fixed channel width in subglacial channels means that the channel width over which to mobilize sediment does not evolve over the same timescales as water discharge \citep[e.g.][]{werder2010}.

  Conversely, in fluvial systems with open-channel flow, a commensurate increase in water discharge will be accommodated by both a reduced increase in water velocity, along with increased channel width and/or depth (Figure~\ref{fig:cartoon}\,b and d).
  The water velocity change might be lower, but the width over which sediment can be mobilized increases given by discharge to width scaling relationships \citep{yalin1992}.
  Thus, the systematic response in sediment transport capacity with respect to changing water discharge is fundamentally different in the two systems.

  % While these differences between open-channel and pressurized flow have been well documented \cite{}, their implications on sediment transport are less well explored \citep[e.g.][]{walder1994}.
  The different processes in subglacial and fluvial systems have been implicitly included in a wide range of available models and relationships that transport sediment in both glacial and fluvial environments \citep[e.g.][]{yalin1992,walder1994,beaud2018,wickert2019}.
  Furthermore, models of fluvial transport of sediment in glacial and fluvial systems have been used to understand the relationship between sediment dynamics and climate \citep[e.g.][]{tucker1997,delaney2020}
  Yet, the differing relationship between variations in water discharge and sediment discharge in subglacial and fluvial systems  has been less well discussed.

  To compare the differences in the response of sediment transport capacity to water discharge, this manuscript leverages parameterizations of subglacial hydraulics and fluvial hydraulics.
  Parameterizations are applied to explore the differences in the relationship between sediment transport capacity and water discharge and the impact of water discharge variability on sediment transport capacity in the two systems.
  Lastly, the implications for sediment transport processes in subglacial systems, compared to fluvial ones, are discussed.

  \begin{center}
    \begin{figure}[H]
      \includegraphics[width=0.9\linewidth]{Cartoon.pdf}
      \caption{Cartoon of the situation with some parameters. Velocity $v_g$ and $v_f$ magnitude are shown by arrow length and marks at the channel boundary. $S_g$ and $S_f$ represent the cross-sectional area in glacial and fluvial channels.  Note that in the fluvial channel parameterization, rectangular channel shapes are implemented (Section~\ref{sect:fluv}).} 
      \label{fig:cartoon}
    \end{figure}
  \end{center}

  \section{Methodology}
  \label{sect:meth}
  This section contains the relationships between water discharge, subglacial water velocity, and channel geometry in both fluvial and subglacial channels.
  Both parameterizations are to represent shear stress, upon which sediment transport commonly depends \citep[][]{shields1936}, and integrate this value across the channel bed.
  Note that both parameterizations evaluate shear stress, as opposed to sediment transport capacity, as shear stress omits a grain size parameter and the choice of a sediment transport relationship \citep{shields1936}.

  \begin{table}[H]
    \centering
    \caption{Variables, parameters, and constants }
    \begin{tabular}{ l  c  c c }
      Name &Symbol&  Value&Units \\ \hline
      \textbf{Variables}  & & & \\
      Channel hydraulic diameter (glacier, fluvial) &  $D_{hg},\,D_{hf}$&  & $\mathrm{m}$     \\
      Width of channel floor (glacial, fluvial) & $w_{cg},w_{cf}$&  & $\mathrm{m}$     \\
      Channel cross-sectional area (glacier, fluvial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
      Water discharge (instantaneous) & $Q_w$& & $\mathrm{m^{3}\,s^{-1}}$ \\
      Representative water discharge & $Q_{w}^*$& & $\mathrm{m^{3}\,s^{-1}}$ \\
      % Gradient of hydraulic potential (glacier) &$\Psi$ & & $\mathrm{Pa\, m^{-1}}$\\
      Representative hydraulic gradient  &$\Psi^*$ & & $\mathrm{Pa\, m^{-1}}$\\
      Water velocity (glacier, fluvial)  & $v_g,\,v_{f}$& & $\mathrm{m\,s^{-1}}$ \\
      Shear stress (glacier, fluvial) & $\tau_g,\,\tau_f$&& $\mathrm{Pa \, m^{-1}}$ \\
      Width-integrated shear-stress (glacier, fluvial) & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$ \\
      Reynolds number &$R_e$& & $\mathrm{(-)}$\\
      Variable &$\Gamma$&&$\mathrm{(-)}$\\
      Variability of a variable ($\Gamma$) &$\chi$& &$\mathrm{(-)}$\\
           &&&\\

      \textbf{Parameters and Constants}  & & &\\
      Gravitational constant&$g$& $-9.81$&$\mathrm{m\,s^{-2}}$\\
      Density of water & $\rho_w$& $1000$ & $\mathrm{kg\,m^{-3}}$ \\
      Density of ice & $\rho_i$& $900$ & $\mathrm{kg\,m^{-3}}$ \\
      Kinematic viscosity of water &$\nu$& $10^{-4}$& $\mathrm{m^2\,s^{-1}}$\\
      Hooke angle of channel & $\beta$ & $\frac{\pi}{2}$ & \unit{rad}\\
      Flotation fraction & $f_f$&$0.7$& $\mathrm{(-)}$\\
      Friction factor (glacier) & $f_r$ & $4$ & $\mathrm{(-)}$ \\
      Friction factor (fluvial) & $f_p$ & $3$ & $\mathrm{(-)}$\\
      Gradient of glacier surface & $\nabla z_{sg}$ &$0.25$& $\mathrm{(-)}$\\
      Gradient of channel bed (fluvial) &$\nabla z_{c}$ &$0.05$& $\mathrm{(-)}$\\
      Fluvial channel factor & $k$ &$3$ & $\mathrm{m^{-2}\, s}$\\
      Channel geometry exponent &$e$& $\frac{1}{2}$&$\mathrm{(-)}$ \\
      \hline
    \end{tabular}
    \label{table:vpm}
  \end{table}

  \subsection{Subglacial channel  parameterization}
  \label{sect:sub_mode}
  To parameterize the shear stress of water flowing across sediments below a glacier, the subglacial channel parameterization evaluates the channel geometry and the velocity of the flowing water. 
  To accomplish this,  a version of hydraulics parameterization presented in \citet{delaney2019} is used.
  The parameterization uses the assumption that the water is transported through subglacial channels \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972}, and that a channel and geometry will respond to a representative water discharge over a certain time period $Q_{w}^*$, using the Darcy-Weisbach formulation for water-flow through a pipe to describe this relationship \citep[e.g.][]{rothlisberger1972,clarke2003,werder2013}.
  Here, it is assumed that the hydraulic diameter of the subglacial channel, $D_h$, responds to a representative water discharge, $Q_{w}^*$, and a representative gradient of the hydraulic potential, $\Psi^*$:
  \begin{linenomath*}
    \begin{equation}
      \label{eq:DW}
      D_h = \big(s\, f_r\,\rho_w\, \frac{Q_w^{*\,2}}{\Psi^*}\big)^{\frac{1}{5}}~~.
    \end{equation}
  \end{linenomath*}
  % 
  The representative hydraulic gradient is based upon the pressurized flow in Equation~\ref{eq:DW},
  \begin{equation}
    \label{eq:psi}
    \Psi^*= f_f \,  \rho_i \, g (\nabla z_{sg} - \nabla z_c) +  \rho_i \, g \, \nabla z_c,
  \end{equation}
  % 
  \noindent
  where, $f_f$ is the flotation fraction, $\rho_i$ is density of ice $\nabla z_{sg}$ is the surface slope of the glacier and $\nabla z_{c}$ is the glacier bed slope.

  In Equation~\ref{eq:DW}, $f_r$ is the Darcy-Weisbach friction factor and $\rho_w$ is the density of water (Table \ref{table:vpm}). $s$ is a factor accounting for channel geometry \citep{hooke1990}, calculated as:
  \begin{equation}
    \label{eq:Hf}
    s = \frac{2\,(\beta -\sin \beta)^2}{(\frac{\beta}{2}\,+\,\sin \frac{\beta}{2})^4},
  \end{equation}
  where $\beta$ is the central angle of the circular segment that comprises the channel (the so-called Hooke angle).
  Smaller values of $\beta$ result in shallow, wide channels, and $\beta =\pi$ corresponds to a semi-circle.
  The width of the flat channel floor $w_c$ is given by
  \begin{equation}
    \label{eq:dh2wc}
    w_{cg} = 2  \sin \frac{\beta}{2} \sqrt{\frac{2\, S}{\beta -\sin \beta}},
  \end{equation}
  % 
  where $S$ is the cross-sectional area of the channel given by
  \begin{equation}
    \label{eq:dh2S}
    S_g =  \frac{D_h^2}{2}~ \frac{\Big(\frac{\beta}{2} \,+ \, \sin \frac{\beta}{2}\Big)^2  }{\beta\,-\,\sin \beta}.
  \end{equation}

  The shear stress, $\tau$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
  \begin{equation}
    \label{eq:tau}
    \tau_g=\frac{1}{8}\,f_r\,\rho_w\,v_g^2,
  \end{equation}
  % 
  where $v_g = \frac{Q_w}{S_g}$ is the water velocity.
  Here, $Q_w$ represents the instantaneous water discharge, not used to evaluate the hydraulic gradient of the channel in Equation~\ref{eq:DW}.
  The channel size is usually not in equilibrium with the quantity of water flowing through it.
  With this formulation, the processes in an R-channel can be represented algebraically \citep{delaney2019}.

  For purposes discussed below, we evaluate the width integrated shear stress as $\tau_{tg}=w_{cg}\,\tau_g $.

  \subsection{Fluvial channel  parameterization}
  \label{sect:fluv}

  The fluvial channel parameterization implements the hydraulics parameterization presented in \citet{tucker1997}.
  Here, the assuming the conservation of mass, sufficiently wide channels so that the hydraulic radius is consistent with flow depth, uniform flow, and the Darcy-Weisbach relationship,
  the shear stress $\tau_f$ at the river bed is represented as
  \begin{linenomath*}
    \begin{equation}
      \label{eq:DW_tau}
      \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q_w}{w_{cf}} \Big)^{\frac{2}{3}} \,\nabla z_{c}^{\frac{2}{3}},
    \end{equation}
  \end{linenomath*}
  where $\nabla z_{c}$ is the channel slope, and $f_p$ is the friction factor for fluvial systems.
  Channel width $w_{cf}$ is 
  \begin{equation}
    \label{eq:wcf}
    w_{cf} = k \, Q_w^e,
  \end{equation}
  % 
  where $k$ is a constant and $e$ is an exponent commonly equal to $\frac{1}{2}$ \citep{leopold1953}. Cross-sectional area is 
  \begin{equation}
    \label{eq:Sf}
    S_f= \frac{Q_w}{v_f}.
  \end{equation}
  As in Section~\ref{sect:sub_mode}, the width integrated shear stress is $\tau_{tf}=w_{cf}\,\tau_f$.
  Hydraulic diameter of the channel is given by $D_{hf} = \frac{4\,w_{cf}\,d}{2\,d+w_{cf}}$, with flow depth $d$ determined from knowledge of $S_f$ and $w_{cf}$ in a rectangular channel.

  \subsection{Implementation}

  The formulation for the glacial system above requires inputs of the hydraulic gradient, water discharge, representative water discharge, Hooke angle $\beta$ \citep{hooke1990}, and friction factor $f_r$ to evaluate channel shape.
  Values of $\beta$ and $f_r$ are chosen so that they produce reasonable values of water velocity, in accordance with available measurements through dye-tracing \citep[Section~\ref{sect:sub_mode}, Figure~\ref{fig:model_outs}; e.g.][]{werder2010}.

  In the fluvial system, the water is not pressurized, so only the bed slope is needed. Friction factor $f_f$ and channel shape factor $k$ are assigned (Section~\ref{sect:fluv}).

  The Reynolds number $R_e$ is given as 
  \begin{equation}
    \label{eq:re}
    R_e\,=\, v \,\frac{D_h}{\nu},
  \end{equation}
  % 
  \noindent where $\nu$ is the kinematic viscosity of water, and  $v$ is velocity of water, $v_g$ or $v_f$.

  The variance $ \chi$ of a variable $\Gamma$ is given by 
  \begin{equation}
    \label{eq:var}
    \chi(\Gamma) \,=\, 1 - \frac{\mathrm{min}(\Gamma)}{\mathrm{max}(\Gamma)}\,\,.
  \end{equation}
  % 
  \noindent Below we examine variables, $\Gamma$, water discharge ($Q_w$) water velocity ($v$, $v_f$), shear stress ($\tau$, $\tau_f$), and integrated shear stress ($\tau_t$, $\tau_{tf}$).

  \section{Results}
  First, the subglacial hydraulics parameterization is applied to a representative water discharge (Equation~\ref{eq:DW}) of $12$\,\unit{m}$^{3}$\,\unit{s}$^{-1}$.
  Both setups are then forced with a water discharge of between $8$ and $16$ \,\unit{m}$^{3}$\,\unit{s}$^{-1}$, to represent diurnal variations in water discharge.

  \begin{center}
    \begin{figure}[H]
      \includegraphics[width=0.9\linewidth]{model_outputs.pdf}
      \caption{Response of channel width (a), hydraulic diameter (b),  Reynolds number  (c),  water velocity (d), shear-stress (e), and width-integrated shear-stress (f)  to different water discharge values. The glacial system is in blue and the fluvial system is in green. in response to variable water discharge. Horizontal bars in d, e, and f represent the range of values in the  output. Gray space represents water discharge below the mean value. }
      \label{fig:model_outs}
    \end{figure}
  \end{center}

  Across the hydrological forcing, variable water discharge results in a substantial increase in the range of water velocities in the glacial system compared to the fluvial system (Figure~\ref{fig:model_outs}).
  While water velocity in the glacier system varies by  $0.5$ (Equation~\ref{eq:var}) over the discharge values, in the fluvial system the water velocity only varies by $0.08$.
  Similarly, shear stress at the bed varies by $0.75$ across the range of water discharge values, whereas shear stress across the fluvial system varies by roughly $0.17$\,\%. This increased variability  in shear stress between the two systems occurs as the water velocity is raised to the power of $2$ in Equation~\ref{eq:tau}. 
  These differences in both shear stress and water velocity demonstrate key differences in the sediment transport capacity in the glacier and fluvial systems.
  Changing channel geometry accommodates evolving water discharge in the fluvial system (Equation~\ref{eq:wcf}), whereas in the subglacial case, increased water flux can only be accommodated by an increase in velocity (Equation~\ref{eq:DW}).

  However, the variable channel width in the fluvial case results in an evolving surface over which to transport sediment in response to changing water discharge (Equation~\ref{eq:wcf}).
  The range of values between width-integrated shear stress in the fluvial system increases to $0.44$\,\%, whereas, in the glacial system, remains at $0.75$\,\%, compared to the shear stress.
  Indeed, the relative  difference in width-integrated shear stress is less between the fluvial and glacier cases, compared to the water velocity and shear stress outputs.
  The reduced variability in width-integrated shear stress results from the additional width across the fluvial channel that accommodates a portion of the sediment transport capacity.

  Next, the setup is run $25,000$ times with randomly selected channel geometries ($k=(2,6)$, $\beta=(\frac{\pi}{10},\pi)$), friction factors ($f_i=(0.001,20)$, $f_p=(0.001,20)$) and water discharges (maximum and minimum values selected between $5$ to $500$ \,\mmma). $Q_w^*$ in Equation ~\ref{eq:DW} is determined as the midpoint in the water discharge range. Parameter combinations are only selected if their maximum and minimum velocities are between $0.2$ and $3$\,\unit{m}\,\unit{s}$^{-1}$, resulting in roughly $10,600$ adequate parameter combinations.

  \begin{center}
    \begin{figure}[H]
      \includegraphics[width=0.9\linewidth]{multi_run.pdf}
      \caption{Ratio of range of a variable $\Gamma$ for glacial and fluvial systems for water velocity  (a), shear-stress (b), and  width-integrated shear-stress (c) in response to mean water discharge,
        Range  $\chi$ for each variable is given in Equation~\ref{eq:var}. Values greater (less) than $1$ shown in white (gray) are parameter combinations that vary more in a subglacial (fluvial) system. Plots (d-f) show the response of variable $\Gamma$ to the variability in water discharge ($\chi(Q_w)$) in the parameter combination. 
        Histograms (g-i) show the frequency of variations across variables in the parameterization.}
      \label{fig:range}
    \end{figure}
  \end{center}

  Across the accepted model runs, variability in water velocity and shear stress remains substantially higher across the subglacial model compared to fluvial ones (Figure~\ref{fig:range}).
  Water velocity, shear stress, and width-integrated shear stress all have a variability ratio ($\frac{\chi(\Gamma_{g})}{\chi(\Gamma_f)}$) greater than $1$ in the subglacial  system compared to fluvial one (Figure~S1).

  The minimum ratio of  water velocity, $\frac{\chi(v_{g})}{\chi(v_{f})}$, and shear-stress, $\frac{\chi(\tau_{g})}{\chi(\tau_{f})}$, is around $2$, showing that these factors are substantially more variable in subglacial channels compared to fluvial ones.
  The width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, spans between $1.02$ and $2.4$, showing  accommodation of the reduced shear stress by increasing the channel width in the fluvial system.
  However, no runs experienced width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, less than $1$.

  Runs with the smallest variability in discharge ($\chi(Q_w)$) have the greatest variability in water velocity, shear stress, and width-integrated shear stress in subglacial systems compared to fluvial ones (Figure~\ref{fig:range} and S1).
  Yet, the greatest difference in the two systems at the smallest variations of water discharge still results in greater variability in the subglacial system compared to the fluvial one (Figure S1).

  \section{Discussion}
  \subsection{Assumptions}

  The direct comparison between the systems remains difficult as they exist with different hydraulic potentials, and the response of channel width to water discharge varies considerably based on the values chosen for $\beta$ and $k$.

  The subglacial parameterization assumes that the friction factor does not evolve between the formation of the channel and the sediment transport conditions examined.
  The role of this effect is uncertain, and most numerical models assume a steady friction factor \citep[e.g.][]{werder2013}. 
  However, the friction factor has been documented to evolve in time \citep{pohle2022}.
  Furthermore, both subglacial and fluvial parameterizations here assume that the distribution of velocity and shear stress of water across the channel bed is homogeneous (Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}).

  The subglacial parameterization also assumes that subglacial water is pressurized, which is a commonly held assumption in subglacial hydrology \citep[e.g.][]{gimbert2016}.
  While the subglacial channel geometry need not be in a steady state for the water to remain pressurized, the subglacial parameterization assumes that the channel geometry must vary minimally over the timescales of varying  water discharge \citep[e.g.][]{nanni2020}.
  Note that in R-channel formulations with changing melt input, the variability in sediment transport capacity could be somewhat smaller than the parameterizations here.
  Here, peaks in the hydrograph would occur as the R-channel is opening due to pressure melt, thus causing a decrease in velocity compared to the results here \citep{rothlisberger1972}.
  A similar effect could occur with creep closure of the subglacial channel at the hydrograph minima.

  The fluvial channel parameterization (Section~\ref{sect:fluv}) assumes that sediment transported across all channel widths, as opposed bankfull width when most sediment transport occurs \citep{wolman1960}.
  As a result, the variability in the fluvial system across the range of discharge values could be underestimated in the parameterization (Figure~\ref{fig:range}).

  \subsection{Geomorphic Implications}
  \label{sect:GI}
  The consistent relationship between water discharge variability and relative variability of the water velocity, shear stress, and width-integrated shear stress results from the fixed geometry in the glacier system and variable geometry in the fluvial system as they respond to changes in water discharge (Equations~\ref{eq:tau} and \ref{eq:Sf}).
  Increased variability in sediment transport capacity with respect to discharge has several implications for sediment transport in subglacial systems compared to fluvial ones.
  Shear stress is scaled to the $\frac{3}{2}$ power  and relies on shear stress exceeding a threshold in sediment transport relationships  such as \citet{meyer1948}.
  In the total sediment transport relationship \citet{engelund1967}, the shear stress is scaled to $\frac{5}{2}$ power.
  In both cases, the exponent greater than $1$ magnifies the variability in sediment transport beyond the highly variable shear stress and width-integrated shear stress described above in subglacial environments (Figure~\ref{fig:model_outs}\, b and d; Figure~S1).

  \subsubsection{Sediment transport subglacially and at glacier margins}

  The response of sediment discharge caused by the greater variability in sediment transport in subglacial channels will be greatest in completely transport-limited regimes \citep{kasmalkar2019}.
  In supply-limited systems, the changing sediment transport capacity (Figure~\ref{fig:model_outs}) would have a minimal effect on sediment discharge due to the absence of sediment available for transport \citep{delaney2019}.
  Yet, reaching the threshold of motion for sediment more frequently may result in additional sediment transport and the tendency for glaciers to transport sediment in a supply-limited case \citep[e.g.][]{herman2015}.

  The effect of pressurized channel flow could be most pronounced during precipitation or flood events when subglacial conduits have not adapted to the water discharge flowing through them, resulting in a very large increase in sediment transport capacity \citep{delaney2019}.
  Precipitation events have been documented to result in exceptionally high sediment discharge from glaciers \citep{cowan1988,delaney2018}.
  Increased sediment discharge may result from changes to sediment access below the glacier, yet rapid increases in water velocity from water flowing through a small conduit, not conditioned by the amount of water flow, will cause a large increase in sediment transport capacity (Section~\ref{sect:sub_mode}).

  Increased variability in sediment transport also impacts sediment export from glacier margins into proglacial areas \citep[e.g.][]{lane2016,delaney2017,perolo2018}.
  Here, in the transition to open-channel flow as the water leaves the glacier, velocities at the highest water discharges could generally decrease (Figure~\ref{fig:model_outs}, blue and green horizontal bars), potentially resulting in sediment deposition once sediment enters the proglacial area.
 Note that the transition from pressurized to open channel flow could occur above the glacier's terminus, when cavity closure rates are reduced and water can flow freely \citep{egli2021b}, resulting in intermittent deposition and remobilization of sediment below the glacier terminus \citep{perolo2018}.

  In large catchments and further downstream of proglacial margins, water discharge is generally far less variable than at the tops of catchments, where glaciers lie \citep[c.f.][]{costa2017,vanas2017,delaney2018,hasholt2018}.
  Here, external hydrological factors further downstream of glaciers could result in smaller variations in sediment transport capacity in many fluvial systems compared to subglacial ones.
  The parameterization here shows that the smallest variability in water discharge has the largest difference in relative variability from subglacial to proglacial systems (Figure~\ref{fig:model_outs}\, d,\,e,\,f), while the variability in the subglacial system remains higher (Figure~S1).
  Therefore, at locations such as the Greenland Ice Sheet, with smaller diurnal variations in water discharge compared to smaller catchments in the Alps \citep[c.f.][]{delaney2018,hasholt2018}, the variability in sediment transport capacity in the transition from the ice sheet to the proglacial river could be more dramatic than alpine systems as water leaves the glacier and moves downstream.

  \subsubsection{Records of sediment discharge  from glaciers}

  Large variations in sediment transport capacity below glaciers may result in sediment mobilization and deposition processes in close spatial and temporal proximity to each other below the glacier \citep{gimbert2016,perolo2018}, in response to both evolving channel geometry and water discharge with accumulating with melt down-glacier \citep{beaud2018,delaney2019}.
  Increasing variability in sediment transport down-glacier is likely because of spatial discontinuities in sediment availability, increasing water discharge down the glacier, and the transport times of subglacial sediment \citep{williams1989,delaney2019}.
  The rapid increase and decreases in sediment transport capacity over a diurnal cycle in a subglacial system, for instance, compounds the variability already present in fluvial systems \citep{williams1989,jerolmack2010}.
  As a result, variations in sediment transport capacity could be responsible for the fluctuations in sediment discharge (``flushings'') from glaciers that are not attributed to variations in water discharge \citep[e.g.][]{richards2003,swift2021}, but could be the result of reduced conduit size and therefore increased sediment transport capacity.

   Records of sediment discharge have been used to establish the relationship, or lack thereof, between sediment discharge and climate in glacial systems \citep[e.g.][]{koppes2009a,willenbring2016,mariotti2021}.
  The variability in subglacial sediment transport capacity presented here applies most generally to short-time scales responsible for the size of subglacial channels.
  Yet, variability in sediment transport from glaciers may add to the already large amount of noise in the fluvial transport of sediment from the glacier to the sink  \citep{castletort2003,jerolmack2010,romans2016}, should the glacier be land-terminating with a proglacial river connecting the glacier to the sink of sediment.
   Therefore, identifying climatic signals in sediment transport from transport-limited glaciers may require higher thresholds of climatic perturbation compared to fluvial systems \citep{tofelde2021}, given the amount of noise present.
  For instance, sediment discharge records from two glaciers in the Swiss Alps show that $40$ to $50$\% of a season's sediment discharge occurs when water discharge is below the $75^{\mathrm{th}}$ quantile of the season \citep{delaney2018}, possibly because sediment transport capacity relies on the size of the glacier conduit, in addition to water discharge (Section~\ref{sect:sub_mode}).
  This suggests that a larger perturbation in water discharge might be needed to significantly alter the sediment discharge from the glacier compared to a fluvial system, should the subglacial sediment transport be in a largely transport-limited regime.

  Additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to fluvial systems \citep{jansson2005,ganti2016}. 
  If a glacier expels sediment in a transport-limited regime, the pressurized nature of subglacial water flow means sediment discharge capacity responds to the thickness of the ice, controlling the closure of the channel, and the surface slope of the glacier \citep{rothlisberger1972,shreve1972}, controlling water velocity, in addition to water discharge (Section~\ref{sect:sub_mode}).
  Conversely, sediment transport in most fluvial systems increases with water discharge and bed slope \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wong2006}. 
  Furthermore, if the glacier is in the supply-limited case, sediment discharge record will also represents additional processes of sliding  \citep{herman2015,seguinot2021}, adding additional processes represented in the sediment discharge signal \citep{delaney2019}.

  \section{Conclusions}

  Divergent processes occur in sediment transport capacity's response to water discharge that exists between subglacial and fluvial systems due to their fixed and variable channel geometry, respectively.
  Parameterization of these processes shows that variability in water velocity and shear stress in response to variable water discharge is much higher in subglacial channels than in fluvial ones.
  The variability in shear stress is reduced when the evolving channel width is accounted for in fluvial systems, yet, the variability in width-integrated shear stress is still higher in subglacial systems.
  When shear stress and width integrated shear stress are linked to sediment transport capacity \citep[e.g.][]{shields1936,meyer1948,engelund1967}, sediment transport capacity can vary far more in subglacial systems compared to fluvial ones.

  A subglacial channel's greater variability in sediment transport capacity is a process that limits geomorphic comparison with fluvial systems, given impacts on the mobilization, transport, and deposition of sediment below glaciers.
  This variability in sediment transport capacity could cause sediment deposition for sediment transport in glacier margins, where water flow transitions from pressurized subglacial flow to open-channel and fluvial.
  Sediment transport in glacier systems responds to a large number of factors such as ice thickness and glacier surface slope, which may respond to climate differently than water discharge. 
  The inconsistent response of subglacial and fluvial systems to water discharge could adds additional noise to sediment transport signals in glacial systems compared to fluvial ones.

  \section{Acknowledgments}

  I was funded by SNSF Project No. PZ00P2\_202024.
  The julia code used herein is included in the supplementary material.

\end{spacing}

\bibliographystyle{apa}
\bibliography{Paperlib}

\section{Supplement}

\begin{center}
  \begin{figure}[H]
    \includegraphics[width=0.9\linewidth]{multi_run_vars.pdf}
    \caption{Variability $\chi$ in velocity ($v_g$, $v_f$; a) shear stress ($\tau_g$, $\tau_f$; b) and width-integrated shear stress ($\tau_{tg}$, $\tau_{tf}$; b)  in glacial and fluvial systems. }
    \label{fig:gammas}
  \end{figure}
\end{center}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

% LocalWords:  pre todo phillips CVODE decadal refreezing Mauro eq%  LocalWords:  dQs
% LocalWords:  dx ms ca DF Hm meltwater Griessee gridded ehrbar ng
% LocalWords:  koppes costa bendixen delaney herman hallet creyts ian
% LocalWords:  farinotti stott rickenman delaney2019 plagerized
% LocalWords:  fischer zekollari clarke raymond larour delaneyinprep
% LocalWords:  guillon exner wolman engelund Weisbach

  As a result, the parameterization here is most relevant for relatively wide rivers, where channel width evolves greatly with water discharge (Equation~\ref{eq:wcf}).
  % This is probably the case in many proglacial rivers and glacierized catchments where the divergent behavior of subglacial and fluvial channels is most relevant (Figure~\ref{sect:GI}) \ian{Find reference about the geometry of proglacial river}.

### 2023-02-17 theory.tex
Updates from Overleaf
Small bits

@@
 
-\ian{check for tense}
-
         \paragraph{Key findings:}
         \begin{enumerate}
-\item Sediment transport capacity in pressurized subglacial channels  responds differently to changing water discharge in open fluvial channels.
-\item Sediment transport capacity varies far more in subglacial channels compared to fluvial ones, even though changing channel width  in fluvial channels accommodates some of the difference.
+        \item Sediment transport capacity varies more in pressurized subglacial channels compared to open fluvial ones, while changing channel width  in fluvial channels accommodates some of the difference.
         \item Differences in the sediment transport capacity response to changing water discharge could impact processes at proglacial margins and the interpretation of sediment transport records from glaciers.
@@
           However, in response to changing water discharge, the behavior of subglacial channels diverges from fluvial ones.
-  % In fluvial channels, the channel width and height may evolve with changing flow conditions, but the underlying channel geometry responds on a far longer timescale than water discharge of the water discharge \citep[Figure~\ref{fig:cartoon}\,b and d; e.g.
-][]{church2006}.
           In subglacial channels, however, the water is pressurized in a subglacial channel and the size of this channel responds to the roughness of the channel walls and the velocity of the water flowing through the channel \citep[Figure~\ref{fig:cartoon}\,a and c; e.g.
@@
 
-  Next, the setup is run $25,000$ times with randomly selected channel geometries ($k=(2,6)$, $\beta=(\frac{\pi}{10},\pi)$), friction factors ($f_i=(0.001,20)$, $f_p=(0.001,20)$) and water discharges (maximum and minimum values selected between $5$ to $500$ \,\mmma).
+          Next, the parameterizations are run $25,000$ times with randomly selected channel geometries ($k=(2,6)$, $\beta=(\frac{\pi}{10},\pi)$), friction factors ($f_i=(0.001,20)$, $f_p=(0.001,20)$) and water discharges (maximum and minimum values selected between $5$ to $500$ \,\mmma).
 $Q_w^*$ in Equation ~\ref{eq:DW} is determined as the midpoint in the water discharge range.
@@
 Plots (d-f) show the response of variable $\Gamma$ to the variability in water discharge ($\chi(Q_w)$) in the parameter combination.
-        Histograms (g-i) show the frequency of variations across variables in the parameterization.}
+                Histograms (g-i) show the frequency of relative variations across variables in the parameterization.}
               \label{fig:range}
@@
 
-  The subglacial parameterization assumes that the friction factor does not evolve between the formation of the channel and the sediment transport conditions examined.
-  The role of this effect is uncertain, and most numerical models assume a steady friction factor \citep[e.g.][]{werder2013}. 
-  However, the friction factor has been documented to evolve in time \citep{pohle2022}.
-  Furthermore, both subglacial and fluvial parameterizations here assume that the distribution of velocity and shear stress of water across the channel bed is homogeneous (Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}).
-
           The subglacial parameterization also assumes that subglacial water is pressurized, which is a commonly held assumption in subglacial hydrology \citep[e.g.][]{gimbert2016}.
           While the subglacial channel geometry need not be in a steady state for the water to remain pressurized, the subglacial parameterization assumes that the channel geometry must vary minimally over the timescales of varying  water discharge \citep[e.g.][]{nanni2020}.
-  Note that in R-channel formulations with changing melt input, the variability in sediment transport capacity could be somewhat smaller than the parameterizations here.
-  Here, peaks in the hydrograph would occur as the R-channel is opening due to pressure melt, thus causing a decrease in velocity compared to the results here \citep{rothlisberger1972}.
+
+          In R-channel formulations with changing melt input, the variability in sediment transport capacity could be somewhat smaller than the parameterizations here.
+          Here, peaks in the hydrograph could cause additional opening of the R-channel  due to pressure melt, thus causing a decrease in velocity compared to the results here \citep{rothlisberger1972}.
           A similar effect could occur with creep closure of the subglacial channel at the hydrograph minima.
+          Additionally, at high water pressures, glacier uplift might change the access of subglacial water to sediment \citep{iken1993}.
 
-  The fluvial channel parameterization (Section~\ref{sect:fluv}) assumes that sediment transported across all channel widths, as opposed bankfull width when most sediment transport occurs \citep{wolman1960}.
+          Both subglacial and fluvial parameterizations here assume that the distribution of velocity and shear stress of water across the channel bed is homogeneous (Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}).
+          The fluvial parameterization assumes that sediment is transported across all channel widths, as opposed bankfull width when much sediment transport occurs \citep{wolman1960}.
           As a result, the variability in the fluvial system across the range of discharge values could be underestimated in the parameterization (Figure~\ref{fig:range}).
@@
           In supply-limited systems, the changing sediment transport capacity (Figure~\ref{fig:model_outs}) would have a minimal effect on sediment discharge due to the absence of sediment available for transport \citep{delaney2019}.
-  Yet, reaching the threshold of motion for sediment more frequently may result in additional sediment transport and the tendency for glaciers to transport sediment in a supply-limited case \citep[e.g.][]{herman2015}.
+          Yet, reaching the threshold of motion for sediment more frequently may result in additional sediment transport and the tendency for glaciers to transport sediment in a supply-limited regime \citep[e.g.][]{herman2015}.
 
-  The effect of pressurized channel flow could be most pronounced during precipitation or flood events when subglacial conduits have not adapted to the water discharge flowing through them, resulting in a very large increase in sediment transport capacity \citep{delaney2019}.
-  Precipitation events have been documented to result in exceptionally high sediment discharge from glaciers \citep{cowan1988,delaney2018}.
+          The effect of pressurized channel flow could be most pronounced during precipitation or flood events when subglacial conduits have not adapted to the water discharge flowing through them, resulting in a very large increase in sediment transport capacity and discharge \citep[e.g.][]{cowan1988,delaney2019}.
          Increased sediment discharge may result from changes to sediment access below the glacier, yet rapid increases in water velocity from water flowing through a small conduit, not conditioned by the amount of water flow, will cause a large increase in sediment transport capacity (Section~\ref{sect:sub_mode}).
 
-  Increased variability in sediment transport also impacts sediment export from glacier margins into proglacial areas \citep[e.g.][]{lane2016,delaney2017,perolo2018}.
+          Increased variability in sediment transport also impacts sediment export from glacier margins into proglacial areas \citep[e.g.][]{delaney2017,perolo2018}.
           Here, in the transition to open-channel flow as the water leaves the glacier, velocities at the highest water discharges could generally decrease (Figure~\ref{fig:model_outs}, blue and green horizontal bars), potentially resulting in sediment deposition once sediment enters the proglacial area.
@@
 
+                    Additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to fluvial systems \citep{jansson2005,ganti2016}.
+    For instance, if a glacier expels sediment in a transport-limited regime, the pressurized nature of subglacial water flow means sediment discharge capacity responds to the thickness of the ice, controlling the closure of the channel, and the surface slope of the glacier \citep{rothlisberger1972,shreve1972}, controlling water velocity, in addition to water discharge (Section~\ref{sect:sub_mode}).
+          Conversely, sediment transport in most fluvial systems increases with water discharge and bed slope \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wong2006}. 
+          Furthermore, if the glacier is in the supply-limited case, the sediment discharge record will also represent additional processes of sliding  \citep{herman2015,seguinot2021}, adding additional processes represented in the sediment discharge signal \citep{delaney2019}.
+
            Records of sediment discharge have been used to establish the relationship, or lack thereof, between sediment discharge and climate in glacial systems \citep[e.g.][]{koppes2009a,willenbring2016,mariotti2021}.
           The variability in subglacial sediment transport capacity presented here applies most generally to short-time scales responsible for the size of subglacial channels.
-  Yet, variability in sediment transport from glaciers may add to the already large amount of noise in the fluvial transport of sediment from the glacier to the sink  \citep{castletort2003,jerolmack2010,romans2016}, should the glacier be land-terminating with a proglacial river connecting the glacier to the sink of sediment.
+          Yet, variability in sediment transport from glaciers may add to the already large amount of noise in the fluvial transport of sediment from the glacier to the sink  \citep{castletort2003,jerolmack2010,romans2016}, especially if the glacier is land-terminating with a proglacial river connecting the glacier to the sink of sediment.
            Therefore, identifying climatic signals in sediment transport from transport-limited glaciers may require higher thresholds of climatic perturbation compared to fluvial systems \citep{tofelde2021}, given the amount of noise present.
           For instance, sediment discharge records from two glaciers in the Swiss Alps show that $40$ to $50$\% of a season's sediment discharge occurs when water discharge is below the $75^{\mathrm{th}}$ quantile of the season \citep{delaney2018}, possibly because sediment transport capacity relies on the size of the glacier conduit, in addition to water discharge (Section~\ref{sect:sub_mode}).
-  This suggests that a larger perturbation in water discharge might be needed to significantly alter the sediment discharge from the glacier compared to a fluvial system, should the subglacial sediment transport be in a largely transport-limited regime.
-
-  Additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to fluvial systems \citep{jansson2005,ganti2016}. 
-  If a glacier expels sediment in a transport-limited regime, the pressurized nature of subglacial water flow means sediment discharge capacity responds to the thickness of the ice, controlling the closure of the channel, and the surface slope of the glacier \citep{rothlisberger1972,shreve1972}, controlling water velocity, in addition to water discharge (Section~\ref{sect:sub_mode}).
-  Conversely, sediment transport in most fluvial systems increases with water discharge and bed slope \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wong2006}. 
-  Furthermore, if the glacier is in the supply-limited case, sediment discharge record will also represents additional processes of sliding  \citep{herman2015,seguinot2021}, adding additional processes represented in the sediment discharge signal \citep{delaney2019}.
+          This suggests that a larger climate perturbation might be needed to significantly alter the sediment discharge capacity from the glaciers compared to fluvial systems, to overcome the increased variability in sediment transport capacity.
 
@@
 
-  Divergent processes occur in sediment transport capacity's response to water discharge that exists between subglacial and fluvial systems due to their fixed and variable channel geometry, respectively.
-  Parameterization of these processes shows that variability in water velocity and shear stress in response to variable water discharge is much higher in subglacial channels than in fluvial ones.
-  The variability in shear stress is reduced when the evolving channel width is accounted for in fluvial systems, yet, the variability in width-integrated shear stress is still higher in subglacial systems.
-  When shear stress and width integrated shear stress are linked to sediment transport capacity \citep[e.g.][]{shields1936,meyer1948,engelund1967}, sediment transport capacity can vary far more in subglacial systems compared to fluvial ones.
+            Application of parameterizations for subglacial and fluvial water flow shows that divergent processes occur in sediment transport capacity's response to water discharge, due to their fixed and variable channel geometry, respectively.
+            Results show that the impact of variable water discharge on in water velocity and shear stress, first-order controls on sediment mobilization, is much higher in subglacial channels than in fluvial ones.
+              The variability in shear stress is reduced when the evolving channel width is accounted for in fluvial systems.
+    Even so, variability in shear stress across a channel's width is still higher in subglacial channels, compared to fluvial ones.
 
-  A subglacial channel's greater variability in sediment transport capacity is a process that limits geomorphic comparison with fluvial systems, given impacts on the mobilization, transport, and deposition of sediment below glaciers.
-  This variability in sediment transport capacity could cause sediment deposition for sediment transport in glacier margins, where water flow transitions from pressurized subglacial flow to open-channel and fluvial.
-  Sediment transport in glacier systems responds to a large number of factors such as ice thickness and glacier surface slope, which may respond to climate differently than water discharge. 
-  The inconsistent response of subglacial and fluvial systems to water discharge could adds additional noise to sediment transport signals in glacial systems compared to fluvial ones.
+    These divergent characteristics between glacial and fluvial channels show that records of sediment transport downstream of glaciers capture multiple processes together.
+    Such processes may compound at glacier margins, where water flow transitions from pressurized subglacial flow to open-channel flow.
+      The inconsistent response of subglacial and fluvial systems to water discharge could add  noise to sediment transport signals in glacial systems compared to fluvial ones.
+        In addition, variations in sediment transport capacity in glacier systems respond to a large number of factors such as channel size, ice thickness, and glacier surface slope, which may react to climate forcing differently than water discharge alone.
+      The differences in the systems' response to changing water discharge should be considered when examining sediment transport processes in glacierized catchments.
 
@@
 
-\bibliographystyle{apa}
-\bibliography{Paperlib}
+        \bibliographystyle{apalike}
+        \bibliography{Paperlib.bib}
 
@@
             \includegraphics[width=0.9\linewidth]{multi_run_vars.pdf}
-    \caption{Variability $\chi$ in velocity ($v_g$, $v_f$; a) shear stress ($\tau_g$, $\tau_f$; b) and width-integrated shear stress ($\tau_{tg}$, $\tau_{tf}$; b)  in glacial and fluvial systems. }
+            \caption{Variability $\chi$ in velocity ($v_g$, $v_f$; a) shear stress ($\tau_g$, $\tau_f$; b) and width-integrated shear stress ($\tau_{tg}$, $\tau_{tf}$; b)  in glacial and fluvial systems.
+}
             \label{fig:gammas}
@@
           % This is probably the case in many proglacial rivers and glacierized catchments where the divergent behavior of subglacial and fluvial channels is most relevant (Figure~\ref{sect:GI}) \ian{Find reference about the geometry of proglacial river}.
+
+Qbs_out Qws_out

### 2023-02-21 theory.tex
small bits and edits

@@
         \affil[1]{Institut des dynamiques de la surface terrestre (IDYST), Universit\'{e} de Lausanne, B\^{a}timent G\'{e}opolis, CH-1015 Lausanne}
-        \title{Sediment transport capacity can vary more in subglacial systems compared to fluvial ones}
+        \title{Sediment transport capacity varies more in subglacial systems compared to fluvial ones}
 
@@
         \begin{enumerate}
-        \item Sediment transport capacity varies more in pressurized subglacial channels compared to open fluvial ones, while changing channel width  in fluvial channels accommodates some of the difference.
+        \item Sediment transport capacity varies more in pressurized subglacial channels compared to open fluvial ones, in response to changing water discharge due to fixed conduit size of their channels.
         \item Differences in the sediment transport capacity response to changing water discharge could impact processes at proglacial margins and the interpretation of sediment transport records from glaciers.
@@
         Sediment transport capacity depends largely on the velocity of water flowing through a channel and the channel width over which to mobilize sediment.
-        In fluvial channels, changing water discharge can be accommodated by changing water depth and width, as well as water velocity.
-        In subglacial channels over short time scales, however, water flows through a conduit with fixed size and is pressurized by the ice above.
-        As a result, changing water discharge only may result in modifying water velocity.
-        Parameterizations of flow in pressurized, subglacial channels and  subaerial fluvial channels show that sediment transport capacity varies  more in response to changing water discharge in subglacial channels compared to fluvial ones.
-        This characteristic of sediment transport capacity in subglacial systems holds true  across a wide range of channel geometries and flow conditions.
-        The increased variability in sediment transport capacity impacts the interpretation of records of sediment transport from glacierized catchments compared to fluvial ones.
+        In fluvial channels, changing water discharge can be accommodated by water depth, width, and water velocity.
+        In subglacial channels, over short time scales, however, water flows through a conduit with a fixed size, and the glacier ice above pressurizes the water.
+        As a result, changing water discharge only may result in modifying water velocity, not the channel's shape.
+        Parameterizations of flow in both pressurized, subglacial channels and  open, fluvial channels show that sediment transport capacity varies  more in response to changing water discharge in subglacial channels compared to fluvial ones.
+        This characteristic of increased variability in sediment transport capacity in subglacial systems holds true  across a wide range of channel geometries and flow conditions.
+        The increased variability in sediment transport capacity impacts sediment transport processes and the interpretation of records of sediment transport from glacierized catchments.
 
@@
           Glaciers are well known for the quantities of sediment that they expel from their termini \citep{hallet1996}.
-          Changing glacier dynamics, geomorphology, and hydrology in cold regions has prompted a recent focus on subglacial sediment transport processes \citep[e.g.][]{zhang2022}.
-          To date, numerical models consider sediment discharge from glaciers as a function of the availability of sediment, either from the erosion of bedrock or the existence of sediment and the transport capacity of this sediment by fluvial processes below the glacier \citep{creyts2013,brinkerhoff2017,beaud2018,delaney2019}.
+          Changing glacier dynamics, geomorphology, and hydrology  has prompted a recent focus on  sediment transport processes in cold regions \citep[e.g.][]{zhang2022}.          
+          Numerical models consider the subglacial component of sediment discharge as a function of the availability of sediment, from the erosion of bedrock or the prior existence of material, and the sediment transport capacity by fluvial processes below the glacier \citep{creyts2013,brinkerhoff2017,beaud2018,delaney2019}.
           These models often represent sediment transport processes similar to rivers, with the main difference being that water flows through a subglacial conduit, pressurized by the ice above \citep{rothlisberger1972}.
 
-          When an adequate quantity of sediment is present as to be transported fluvially (i.e.
-in a ``transport-limited'' regime), shear-stress between the water and the sediment that it flows over is a first-order control on the transport of sediment \citep{shields1936}.
+          Both subglacially and fluvially, shear stress between the water and the sediment that it flows over is a first-order control on the transport of sediment \citep{shields1936} if an adequate quantity of sediment is present as to be transported fluvially (i.e.
+in a ``transport-limited'' regime).
           Shear stress increases with the water velocity and roughness of the channel.
-          Greater shear stress will mobilize increased quantities of sediment, along with larger sizes of sediment clasts.
-          The fundamental role of this shear-stress is demonstrated in sediment transport relationships such as \citet{shields1936}.
-          Sediment transport equations such as \citet{meyer1948}  and \citet{engelund1967} use shear stress, or an approximation thereof, to evaluate sediment transport capacity per unit width.
-
-          While sediment transport over a unit width of the channel depends directly on shear-stress \citep[e.g.][]{meyer1948}, the total sediment transport capacity across a channel's cross depends on the integrated shear-stress over the entire channel bed (Figure~\ref{fig:cartoon}).
-          Furthermore, as water discharge changes, the channel's cross-sectional area, as well as the water velocity, accommodates the variations.
+          Greater shear stress will mobilize increased quantities of sediment, along with larger sediment clasts.
+          Sediment transport relationships such as \citet{shields1936}, \citet{meyer1948},  and \citet{engelund1967} use shear stress, or an approximation thereof, to evaluate sediment transport capacity per unit width.
+          The total sediment transport capacity across a channel's cross depends on the integrated shear stress over the entire channel bed (Figure~\ref{fig:cartoon}).
           As a result, channel geometry plays a key role in sediment transport in both fluvial channels and subglacial ones \citep[e.g.][]{church2006,delaney2019}.
 
-          However, in response to changing water discharge, the behavior of subglacial channels diverges from fluvial ones.
-          In subglacial channels, however, the water is pressurized in a subglacial channel and the size of this channel responds to the roughness of the channel walls and the velocity of the water flowing through the channel \citep[Figure~\ref{fig:cartoon}\,a and c; e.g.
+          However, in response to changing water discharge, the behavior of subglacial channels differs from fluvial ones.
+          In subglacial channels, the water is pressurized in a subglacial channel and the size of this channel responds to the roughness of the channel walls and the velocity of the water flowing through the channel \citep[Figure~\ref{fig:cartoon}\,a and c; e.g.
 ][]{rothlisberger1972}.
@@
           Conversely, in fluvial systems with open-channel flow, a commensurate increase in water discharge will be accommodated by both a reduced increase in water velocity, along with increased channel width and/or depth (Figure~\ref{fig:cartoon}\,b and d).
-          The water velocity change might be lower, but the width over which sediment can be mobilized increases given by discharge to width scaling relationships \citep{yalin1992}.
-          Thus, the systematic response in sediment transport capacity with respect to changing water discharge is fundamentally different in the two systems.
+          The change in water velocity could be lower, but the width over which sediment can be mobilized changes \citep{leopold1953}.
+          Thus, the systematic response in sediment transport capacity with respect to changing water discharge is  different in the two systems.
 
-          % While these differences between open-channel and pressurized flow have been well documented \cite{}, their implications on sediment transport are less well explored \citep[e.g.][]{walder1994}.
-          The different processes in subglacial and fluvial systems have been implicitly included in a wide range of available models and relationships that transport sediment in both glacial and fluvial environments \citep[e.g.][]{yalin1992,walder1994,beaud2018,wickert2019}.
+          The different processes in subglacial and fluvial systems have been implicitly included in a wide range of available models and relationships that transport sediment in both glacial and fluvial environments \citep[e.g.][]{walder1994,tucker1997,beaud2018,wickert2019}.
           Furthermore, models of fluvial transport of sediment in glacial and fluvial systems have been used to understand the relationship between sediment dynamics and climate \citep[e.g.][]{tucker1997,delaney2020}
@@
               \includegraphics[width=0.9\linewidth]{Cartoon.pdf}
-              \caption{Cartoon of the situation with some parameters.
+              \caption{Cartoon of the situation with various parameters.
 Velocity $v_g$ and $v_f$ magnitude are shown by arrow length and marks at the channel boundary.
@@
           This section contains the relationships between water discharge, subglacial water velocity, and channel geometry in both fluvial and subglacial channels.
-          Both parameterizations are to represent shear stress, upon which sediment transport commonly depends \citep[][]{shields1936}, and integrate this value across the channel bed.
-          Note that both parameterizations evaluate shear stress, as opposed to sediment transport capacity, as shear stress omits a grain size parameter and the choice of a sediment transport relationship \citep{shields1936}.
+          Because sediment transport largely depends on shear stress \citep{shields1936}, both parameterizations calculated this value and integrate  it across the channel bed.
+          The parameterizations evaluate shear stress, as opposed to sediment transport capacity, as shear stress omits a grain size parameter and the choice of a sediment transport relationship \citep{shields1936}.
 
@@
               Friction factor (fluvial) & $f_p$ & $3$ & $\mathrm{(-)}$\\
-              Gradient of glacier surface & $\nabla z_{sg}$ &$0.25$& $\mathrm{(-)}$\\
-              Gradient of channel bed (fluvial) &$\nabla z_{c}$ &$0.05$& $\mathrm{(-)}$\\
+              Gradient of glacier surface & $\frac{\paratial z_{sg}}{\partial x}$ &$0.25$& $\mathrm{(-)}$\\
+              Gradient of channel bed (fluvial) &$\frac{\paratial z_c}{\partial x}$ &$0.05$& $\mathrm{(-)}$\\
               Fluvial channel factor & $k$ &$3$ & $\mathrm{m^{-2}\, s}$\\
@@
             \label{eq:psi}
-            \Psi^*= f_f \,  \rho_i \, g (\nabla z_{sg} - \nabla z_c) +  \rho_i \, g \, \nabla z_c,
+            \Psi^*= f_f \,  \rho_i \, g (\frac{\partial  z_{sg}}{\partial x} - \frac{\partial z_c}{\partial x}) +  \rho_i \, g \, \frac{\partial z_c}{\partial x},
           \end{equation}
@@
           \noindent
-          where, $f_f$ is the flotation fraction, $\rho_i$ is density of ice $\nabla z_{sg}$ is the surface slope of the glacier and $\nabla z_{c}$ is the glacier bed slope.
+          where, $f_f$ is the flotation fraction, $\rho_i$ is density of ice $\frac{\partial z_{sg}}{\partial x}$ is the surface slope of the glacier and $\frac{\partial z_c}{\partial x}$ is the glacier bed slope.
 
@@
           where $\beta$ is the central angle of the circular segment that comprises the channel (the so-called Hooke angle).
-          Smaller values of $\beta$ result in shallow, wide channels, and $\beta =\pi$ corresponds to a semi-circle.
+          Smaller values of $\beta$ result in shallow, wide channels. $\beta =\pi$ corresponds to a semi-circle.
           The width of the flat channel floor $w_c$ is given by
@@
               \label{eq:DW_tau}
-              \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q_w}{w_{cf}} \Big)^{\frac{2}{3}} \,\nabla z_{c}^{\frac{2}{3}},
+              \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q_w}{w_{cf}} \Big)^{\frac{2}{3}} \,\frac{\partial z_c}{\partial x}^{\frac{2}{3}},
             \end{equation}
           \end{linenomath*}
-          where $\nabla z_{c}$ is the channel slope, and $f_p$ is the friction factor for fluvial systems.
+          where $\frac{\paratial z_c}{\partial x}$ is the channel slope, and $f_p$ is the friction factor for fluvial systems.
           Channel width $w_{cf}$ is 
@@
 
-          The formulation for the glacial system above requires inputs of the hydraulic gradient, water discharge, representative water discharge, Hooke angle $\beta$ \citep{hooke1990}, and friction factor $f_r$ to evaluate channel shape.
+          The formulation for the glacial system above requires inputs of the hydraulic gradient, water discharge, representative water discharge, Hooke angle $\beta$ \citep{hooke1990}, and friction factor $f_r$.
           Values of $\beta$ and $f_r$ are chosen so that they produce reasonable values of water velocity, in accordance with available measurements through dye-tracing \citep[Section~\ref{sect:sub_mode}, Figure~\ref{fig:model_outs}; e.g.][]{werder2010}.
@@
           % 
-          \noindent Below we examine variables, $\Gamma$, water discharge ($Q_w$) water velocity ($v$, $v_f$), shear stress ($\tau$, $\tau_f$), and integrated shear stress ($\tau_t$, $\tau_{tf}$).
+          \noindent Below we examine variables, $\Gamma$, as water discharge ($Q_w$) water velocity ($v$, $v_f$), shear stress ($\tau$, $\tau_f$), and integrated shear stress ($\tau_t$, $\tau_{tf}$).
 
           \section{Results}
-          First, the subglacial hydraulics parameterization is applied to a representative water discharge (Equation~\ref{eq:DW}) of $12$\,\unit{m}$^{3}$\,\unit{s}$^{-1}$.
-          Both setups are then forced with a water discharge of between $8$ and $16$ \,\unit{m}$^{3}$\,\unit{s}$^{-1}$, to represent diurnal variations in water discharge.
+          First, the subglacial hydraulics parameterization is applied to a representative water discharge ($Q_w^*$; Equation~\ref{eq:DW}) of $12$\,\unit{m}$^{3}$\,\unit{s}$^{-1}$.
+          Both parameterizations are forced with a water discharge ($Q_w$) varying between $8$ and $16$ \,\unit{m}$^{3}$\,\unit{s}$^{-1}$, to represent a range of sediment transport conditions, for instance over diurnal fluctuation in water discharge.
+The fluvial channel width factor $k$ is $3$\,\unit{m}$^{2}$\,\unit{s}$^{-1}$, and the Hooke angle $\beta$ is $\frac{\pi}{2}$.
 
@@
 The glacial system is in blue and the fluvial system is in green.
-in response to variable water discharge.
 Horizontal bars in d, e, and f represent the range of values in the  output.
@@
           Across the hydrological forcing, variable water discharge results in a substantial increase in the range of water velocities in the glacial system compared to the fluvial system (Figure~\ref{fig:model_outs}).
-          While water velocity in the glacier system varies by  $0.5$ (Equation~\ref{eq:var}) over the discharge values, in the fluvial system the water velocity only varies by $0.08$.
-          Similarly, shear stress at the bed varies by $0.75$ across the range of water discharge values, whereas shear stress across the fluvial system varies by roughly $0.17$\,\%.
+          While water velocity in the glacier system varies by  $0.5$ (Equation~\ref{eq:var}) over the discharge values, in the fluvial system the water velocity only varies by $0.11$.
+          Similarly, shear stress at the glacier bed varies by $0.75$ across the range of water discharge values, whereas shear stress across the fluvial system varies by roughly $0.21$.
 This increased variability  in shear stress between the two systems occurs as the water velocity is raised to the power of $2$ in Equation~\ref{eq:tau}.
-          These differences in both shear stress and water velocity demonstrate key differences in the sediment transport capacity in the glacier and fluvial systems.
+
           Changing channel geometry accommodates evolving water discharge in the fluvial system (Equation~\ref{eq:wcf}), whereas in the subglacial case, increased water flux can only be accommodated by an increase in velocity (Equation~\ref{eq:DW}).
-
-          However, the variable channel width in the fluvial case results in an evolving surface over which to transport sediment in response to changing water discharge (Equation~\ref{eq:wcf}).
-          The range of values between width-integrated shear stress in the fluvial system increases to $0.44$\,\%, whereas, in the glacial system, remains at $0.75$\,\%, compared to the shear stress.
+          The range of values between width-integrated shear stress in the fluvial system increases to $0.44$, whereas, variability in the glacial system lies at $0.83$ (Figure~\ref{fig:range}).
           Indeed, the relative  difference in width-integrated shear stress is less between the fluvial and glacier cases, compared to the water velocity and shear stress outputs.
@@
           The minimum ratio of  water velocity, $\frac{\chi(v_{g})}{\chi(v_{f})}$, and shear-stress, $\frac{\chi(\tau_{g})}{\chi(\tau_{f})}$, is around $2$, showing that these factors are substantially more variable in subglacial channels compared to fluvial ones.
-          The width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, spans between $1.02$ and $2.4$, showing  accommodation of the reduced shear stress by increasing the channel width in the fluvial system.
-          However, no runs experienced width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, less than $1$.
+          The width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, spans between $1.02$ and $2.4$ because changing channel width accommodates the reduced shear stress in the fluvial system.
+          No runs experienced width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, less than $1$.
 
@@
 
-          The direct comparison between the systems remains difficult as they exist with different hydraulic potentials, and the response of channel width to water discharge varies considerably based on the values chosen for $\beta$ and $k$.
+          The direct comparison between the systems remains difficult as they have with different hydraulic potentials, and the response of channel width to water discharge varies considerably based on the values chosen for $\beta$ and $k$.
 
-          The subglacial parameterization also assumes that subglacial water is pressurized, which is a commonly held assumption in subglacial hydrology \citep[e.g.][]{gimbert2016}.
+          The subglacial parameterization assumes that subglacial water is pressurized, which is a commonly held assumption in subglacial hydrology, though observations suggest that this may not always be the case \citep[e.g.][]{gimbert2016}.
           While the subglacial channel geometry need not be in a steady state for the water to remain pressurized, the subglacial parameterization assumes that the channel geometry must vary minimally over the timescales of varying  water discharge \citep[e.g.][]{nanni2020}.
 
-          In R-channel formulations with changing melt input, the variability in sediment transport capacity could be somewhat smaller than the parameterizations here.
-          Here, peaks in the hydrograph could cause additional opening of the R-channel  due to pressure melt, thus causing a decrease in velocity compared to the results here \citep{rothlisberger1972}.
+          In R-channel formulations, peaks in the hydrograph could cause additional opening of the conduit due to pressure melt, thus causing a decrease in velocity compared to the results here \citep{rothlisberger1972}.
           A similar effect could occur with creep closure of the subglacial channel at the hydrograph minima.
-          Additionally, at high water pressures, glacier uplift might change the access of subglacial water to sediment \citep{iken1993}.
+          Additionally, at high water pressures, glacier uplift and bed separation \citep{andrews2014} might slightly modify the width of the subglacial channel.
 
           Both subglacial and fluvial parameterizations here assume that the distribution of velocity and shear stress of water across the channel bed is homogeneous (Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}).
-          The fluvial parameterization assumes that sediment is transported across all channel widths, as opposed bankfull width when much sediment transport occurs \citep{wolman1960}.
+          The fluvial parameterization assumes that sediment is transported across all channel widths, as opposed to bankfull width when much sediment transport occurs \citep{wolman1960}.
           As a result, the variability in the fluvial system across the range of discharge values could be underestimated in the parameterization (Figure~\ref{fig:range}).
@@
 
-          \subsubsection{Sediment transport subglacially and at glacier margins}
+          \subsubsection{Sediment transport below glaciers and at their margins}
 
-          The response of sediment discharge caused by the greater variability in sediment transport in subglacial channels will be greatest in completely transport-limited regimes \citep{kasmalkar2019}.
+          The response of sediment discharge caused by the greater variability in sediment transport in subglacial channels will be greatest in completely transport-limited regimes \citep[e.g.][]{kasmalkar2019}.
           In supply-limited systems, the changing sediment transport capacity (Figure~\ref{fig:model_outs}) would have a minimal effect on sediment discharge due to the absence of sediment available for transport \citep{delaney2019}.
@@
 
-          In large catchments and further downstream of proglacial margins, water discharge is generally far less variable than at the tops of catchments, where glaciers lie \citep[c.f.][]{costa2017,vanas2017,delaney2018,hasholt2018}.
+          In large catchments and further downstream of proglacial margins, water discharge is generally less variable than at the tops of catchments, where glaciers lie \citep[c.f.][]{costa2017,vanas2017,delaney2018,hasholt2018}.
           Here, external hydrological factors further downstream of glaciers could result in smaller variations in sediment transport capacity in many fluvial systems compared to subglacial ones.
-          The parameterization here shows that the smallest variability in water discharge has the largest difference in relative variability from subglacial to proglacial systems (Figure~\ref{fig:model_outs}\, d,\,e,\,f), while the variability in the subglacial system remains higher (Figure~S1).
-          Therefore, at locations such as the Greenland Ice Sheet, with smaller diurnal variations in water discharge compared to smaller catchments in the Alps \citep[c.f.][]{delaney2018,hasholt2018}, the variability in sediment transport capacity in the transition from the ice sheet to the proglacial river could be more dramatic than alpine systems as water leaves the glacier and moves downstream.
+          The parameterization here shows that the smallest variability in water discharge has the largest difference in relative variability from subglacial to proglacial systems (Figure~\ref{fig:range}\, d,\,e,\,f), while the variability in the subglacial system remains higher (Figure~S1).
+          Therefore, at locations such as the Greenland Ice Sheet, with smaller diurnal variations in water discharge compared to many alpine catchments \citep[c.f.][]{delaney2018,hasholt2018}, the variability in sediment transport capacity in the transition from the ice sheet to the proglacial river could be more dramatic than alpine systems as water leaves the glacier and moves downstream.
+          Thus the disparity in sediment transport capacity between the subglacial and fluvial systems could be most pronounced in these catchments, potentially resulting in temporally mismatched periods between when glaciers introduce sediment to these river systems  and when the river mobilizes it (Figures~\ref{fig:range}~ and~\ref{fig:model_outs}).
 
@@
                     Additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to fluvial systems \citep{jansson2005,ganti2016}.
-    For instance, if a glacier expels sediment in a transport-limited regime, the pressurized nature of subglacial water flow means sediment discharge capacity responds to the thickness of the ice, controlling the closure of the channel, and the surface slope of the glacier \citep{rothlisberger1972,shreve1972}, controlling water velocity, in addition to water discharge (Section~\ref{sect:sub_mode}).
-          Conversely, sediment transport in most fluvial systems increases with water discharge and bed slope \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wong2006}. 
-          Furthermore, if the glacier is in the supply-limited case, the sediment discharge record will also represent additional processes of sliding  \citep{herman2015,seguinot2021}, adding additional processes represented in the sediment discharge signal \citep{delaney2019}.
+    For instance, if a glacier expels sediment in a transport-limited regime, the pressurized nature of subglacial water flow means sediment discharge capacity responds to the thickness of the ice, controlling the closure of the channel, and the surface slope of the glacier, controlling water velocity, in addition to water discharge \citep[Section~\ref{sect:sub_mode}; ] []{rothlisberger1972,shreve1972}.
+          Conversely, sediment transport in most fluvial systems increases with water discharge and bed slope, with the bed slope, probably being more stable over long time periods \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wong2006}.
+          Furthermore, if the glacier is in a supply-limited case, the sediment discharge record will also represent additional processes of sliding  \citep{herman2015,seguinot2021}, adding additional processes represented in the sediment discharge signal \citep{delaney2019}.
 
@@
           The variability in subglacial sediment transport capacity presented here applies most generally to short-time scales responsible for the size of subglacial channels.
-          Yet, variability in sediment transport from glaciers may add to the already large amount of noise in the fluvial transport of sediment from the glacier to the sink  \citep{castletort2003,jerolmack2010,romans2016}, especially if the glacier is land-terminating with a proglacial river connecting the glacier to the sink of sediment.
-           Therefore, identifying climatic signals in sediment transport from transport-limited glaciers may require higher thresholds of climatic perturbation compared to fluvial systems \citep{tofelde2021}, given the amount of noise present.
+        Yet, identifying climatic signals in sediment transport from transport-limited glaciers may require higher thresholds of climatic perturbation compared to fluvial systems \citep{tofelde2021}, given the potential amount of additional noise \citep{castletort2003,jerolmack2010,romans2016}.
           For instance, sediment discharge records from two glaciers in the Swiss Alps show that $40$ to $50$\% of a season's sediment discharge occurs when water discharge is below the $75^{\mathrm{th}}$ quantile of the season \citep{delaney2018}, possibly because sediment transport capacity relies on the size of the glacier conduit, in addition to water discharge (Section~\ref{sect:sub_mode}).
@@
 
-            Application of parameterizations for subglacial and fluvial water flow shows that divergent processes occur in sediment transport capacity's response to water discharge, due to their fixed and variable channel geometry, respectively.
-            Results show that the impact of variable water discharge on in water velocity and shear stress, first-order controls on sediment mobilization, is much higher in subglacial channels than in fluvial ones.
+            Application of parameterizations for subglacial and fluvial water flow shows that variable processes occur in sediment transport capacity's response to water discharge, due to their fixed and variable channel geometry, respectively.
+            Results show that the impact of variable water discharge on water velocity and shear stress, first-order controls on sediment mobilization, is much higher in subglacial channels than in fluvial ones.
               The variability in shear stress is reduced when the evolving channel width is accounted for in fluvial systems.
@@
 
-    These divergent characteristics between glacial and fluvial channels show that records of sediment transport downstream of glaciers capture multiple processes together.
+    These different characteristics between glacial and fluvial channels show that records of sediment transport downstream of glaciers capture multiple processes together.
     Such processes may compound at glacier margins, where water flow transitions from pressurized subglacial flow to open-channel flow.

### 2023-02-21 theory.tex
new figs

@@
       Friction factor (fluvial) & $f_p$ & $3$ & $\mathrm{(-)}$\\
-              Gradient of glacier surface & $\frac{\paratial z_{sg}}{\partial x}$ &$0.25$& $\mathrm{(-)}$\\
-              Gradient of channel bed (fluvial) &$\frac{\paratial z_c}{\partial x}$ &$0.05$& $\mathrm{(-)}$\\
+      Gradient of glacier surface & $\frac{\partial z_{sg}}{\partial x}$ &$0.25$& $\mathrm{(-)}$\\
+      Gradient of channel bed (fluvial) &$\frac{\partial z_c}{\partial x}$ &$0.05$& $\mathrm{(-)}$\\
       Fluvial channel factor & $k$ &$3$ & $\mathrm{m^{-2}\, s}$\\
@@
   \end{linenomath*}
-          where $\frac{\paratial z_c}{\partial x}$ is the channel slope, and $f_p$ is the friction factor for fluvial systems.
+  where $\frac{\partial z_c}{\partial x}$ is the channel slope, and $f_p$ is the friction factor for fluvial systems.
   Channel width $w_{cf}$ is 
@@
     \includegraphics[width=0.9\linewidth]{multi_run_vars.pdf}
-            \caption{Variability $\chi$ in velocity ($v_g$, $v_f$; a) shear stress ($\tau_g$, $\tau_f$; b) and width-integrated shear stress ($\tau_{tg}$, $\tau_{tf}$; b)  in glacial and fluvial systems.
-}
+    \caption{Variability $\chi$ in velocity ($v_g$, $v_f$; a) shear stress ($\tau_g$, $\tau_f$; b) and width-integrated shear stress ($\tau_{tg}$, $\tau_{tf}$; b)  in glacial and fluvial systems. }
     \label{fig:gammas}

### 2023-02-28 theory.tex
small updates
Updates from Overleaf

@@
     \begin{figure}[H]
-      \includegraphics[width=0.9\linewidth]{Cartoon.pdf}
+      \includegraphics[width=0.65\linewidth]{Cartoon.pdf}
       \caption{Cartoon of the situation with various parameters.
@@
 
+  \ian{check units on fluvial channel factor}
   \begin{table}[H]
@@
   To parameterize the shear stress of water flowing across sediments below a glacier, the subglacial channel parameterization evaluates the channel geometry and the velocity of the flowing water. 
-  To accomplish this, a version of hydraulics parameterization presented in \citet{delaney2019} is used.
+  To accomplish this, the hydraulics parameterization presented in \citet{delaney2019} is used.
   The parameterization uses the assumption that the water is transported through subglacial channels \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972}, and that a channel and geometry will respond to a representative water discharge over a certain time period $Q_{w}^*$, using the Darcy-Weisbach formulation for water-flow through a pipe to describe this relationship \citep[e.g.][]{rothlisberger1972,clarke2003,werder2013}.
@@
     \begin{figure}[H]
-      \includegraphics[width=0.9\linewidth]{model_outputs.pdf}
+      \includegraphics[width=0.7\linewidth]{model_outputs.pdf}
       \caption{Response of channel width (a), hydraulic diameter (b),  Reynolds number  (c),  water velocity (d), shear-stress (e), and width-integrated shear-stress (f)  to different water discharge values.
@@
     \begin{figure}[H]
-      \includegraphics[width=0.9\linewidth]{multi_run.pdf}
+      \includegraphics[width=0.7\linewidth]{multi_run.pdf}
       \caption{Ratio of range of a variable $\Gamma$ for glacial and fluvial systems for water velocity  (a), shear-stress (b), and  width-integrated shear-stress (c) in response to mean water discharge,
@@
 
-  These different characteristics between glacial and fluvial channels show that records of sediment transport downstream of glaciers capture multiple processes together.
-  Such processes may compound at glacier margins, where water flow transitions from pressurized subglacial flow to open-channel flow.
-  The inconsistent response of subglacial and fluvial systems to water discharge could add  noise to sediment transport signals in glacial systems compared to fluvial ones.
-  In addition, variations in sediment transport capacity in glacier systems respond to a large number of factors such as channel size, ice thickness, and glacier surface slope, which may react to climate forcing differently than water discharge alone.
-  The differences in the systems' response to changing water discharge should be considered when examining sediment transport processes in glacierized catchments.
+  These different characteristics between glacial and fluvial channels show that records of sediment transport downstream of glaciers capture multiple fluvial processes together.
+  Such processes might be particularly relevant at glacier margins, where water transitions from pressurized subglacial flow to open-channel flow.
+  The inconsistent response of sediment transport in subglacial and fluvial systems to changing water discharge could add noise to sediment transport signals between the systems.
+  In addition, the subglacial parameterization shows that variations in sediment transport capacity in glacier systems respond to a large number of factors such as channel size, ice thickness, and glacier surface slope, which may react to climate forcing differently than water discharge alone.
+  The differences in fluvial and subglacial channels' response to changing water discharge should be considered when examining sediment transport processes in glacierized catchments.
 
@@
   \begin{figure}[H]
-    \includegraphics[width=0.9\linewidth]{multi_run_vars.pdf}
-    \caption{Variability $\chi$ in velocity ($v_g$, $v_f$; a) shear stress ($\tau_g$, $\tau_f$; b) and width-integrated shear stress ($\tau_{tg}$, $\tau_{tf}$; b)  in glacial and fluvial systems. }
+    \includegraphics[width=0.7\linewidth]{multi_run_vars.pdf}
+    \caption{Variability $\chi$ in velocity ($v_g$, $v_f$; a) shear stress ($\tau_g$, $\tau_f$; b) and width-integrated shear stress ($\tau_{tg}$, $\tau_{tf}$; c)  in glacial and fluvial systems. }
     \label{fig:gammas}

### 2023-03-06 theory.tex
MJ first edits

@@
 \affil[1]{Institut des dynamiques de la surface terrestre (IDYST), Universit\'{e} de Lausanne, B\^{a}timent G\'{e}opolis, CH-1015 Lausanne}
-\title{Sediment transport capacity varies more in subglacial systems compared to fluvial ones}
+\title{Comparison of sediment transport capacity in subglacial and fluvial systems }
 
@@
 \begin{enumerate}
-\item Sediment transport capacity varies more in pressurized subglacial channels compared to open fluvial ones, in response to changing water discharge due to fixed conduit size of their channels.
-\item Differences in the sediment transport capacity response to changing water discharge could impact processes at proglacial margins and the interpretation of sediment transport records from glaciers.
+\item Sediment transport capacity varies more in pressurized subglacial channels compared to open fluvial ones, in response to changing water discharge due to the fixed size of their channels.
+I can't fully undertand this.
+Too much information is overlain here.
+Separate concepts and give a more linear explanation.
+\item  As water discharge changes, differences in the sediment transport capacity response of fluvial vs subglacial channels  could substantially  impact ________ what kind?
+processes at proglacial margins as well as the interpretation of sediment transport records in glacial environments (?).
 \end{enumerate}
@@
 \abstract % 150 words
-Sediment transport capacity depends largely on the velocity of water flowing through a channel and the channel width over which to mobilize sediment.
-In fluvial channels, changing water discharge can be accommodated by water depth, width, and water velocity.
-In subglacial channels, over short time scales, however, water flows through a conduit with a fixed size, and the glacier ice above pressurizes the water.
-As a result, changing water discharge only may result in modifying water velocity, not the channel's shape.
-Parameterizations of flow in both pressurized, subglacial channels and  open, fluvial channels show that sediment transport capacity varies  more in response to changing water discharge in subglacial channels compared to fluvial ones.
-This characteristic of increased variability in sediment transport capacity in subglacial systems holds true  across a wide range of channel geometries and flow conditions.
+Sediment transport capacity depends largely on the velocity of water flowing through a channel and the channel width over which the sediment is mobilized.
+In fluvial channels, changing water discharge can be accommodated by changes in? water depth, width of water flow [ is this width of water flow? ], and water velocity.
+In subglacial channels, however, water flows through a conduit with a fixed size and shape?, and the water is pressurized by the overlying glacier ice.
+As a result, changing water discharge may result in modifying the water velocity, but not the  shape of the channel.
+Parameterization of _______ indicates that a calculated?
+observed?
+increased variability in sediment transport capacity in subglacial systems occurs across a wide range of channel geometries and flow conditions.
 The increased variability in sediment transport capacity impacts sediment transport processes and the interpretation of records of sediment transport from glacierized catchments.
+This last sentence seems repetitive.
+I'm not fully understanding the results.
+Suggest revise after a complete update of the manuscript.
 
@@
 \laysummary % 200 words
-Sediment transport capacity responds differently to changing water discharge in subglacial and fluvial channels.
-In both subglacial and fluvial channels, sediment transport capacity, to the first order, depends on 1) the channel width over-which to mobilize sediment and 2) the water's shear stress on sediment at the channel bottom to mobilize it.
+It is well known that sediment transport capacity responds differently to changing water discharge in subglacial and fluvial channels.
+In both subglacial and fluvial channels, sediment transport capacity depends, to the first order, on 1) the channel width over-which the sediment is mobilized  and 2) the shear stress applied by flowing water on sediment at the channel bottom.
 The shear stress depends largely on water velocity.
-In fluvial channels, changing water discharge will change the water depth, water velocity (controlling sediment transport capacity),  and channel width (which dictates the span over which sediment transport  will occur).
+In fluvial channels, changing water discharge will change the water depth, the water velocity (controlling sediment transport capacity),  and channel width (which dictates the span over which sediment transport  will occur).
 Conversely, in subglacial environments, on short times scales of hours to days, subglacial channels only experience a change in water velocity following changing water discharge, because the water flows through a subglacial conduit with a fixed geometry.
@@
 This increased variability has several important implications for sediment transport in glacial systems compared to fluvial ones.
+This is hard to follow. Suggest you rewrite after greater clarity in the rest of the paper.
 
@@
 
-  Glaciers are well known for the quantities of sediment that they expel from their termini \citep{hallet1996}.
-  Changing glacier dynamics, geomorphology, and hydrology  has prompted a recent focus on  sediment transport processes in cold regions \citep[e.g.][]{zhang2022}.          
-  Numerical models consider the subglacial component of sediment discharge as a function of the availability of sediment, from the erosion of bedrock or the prior existence of material, and the sediment transport capacity by fluvial processes below the glacier \citep{creyts2013,brinkerhoff2017,beaud2018,delaney2019}.
-  These models often represent sediment transport processes similar to rivers, with the main difference being that water flows through a subglacial conduit, pressurized by the ice above \citep{rothlisberger1972}.
+  Glaciers expel large quantities of sediment  from their termini \citep{hallet1996}. NEEDS MORE REFERENCES! Current 
+  changes in glacier dynamics, geomorphology, and hydrology  have prompted numerous  recent studies of  sediment transport processes in cold regions, such as ________, _________, _____ \citep[e.g.][]{zhang2022}.
+NEEDS MORE REFERENCES to fit with geographic sites!
+Get the observations and measurements well stated, first.
+  Numerical models integrated with glacial monitoring data  consider evaluate?
+the subglacial component of sediment discharge ___________at glacial termini?
+as a function of the availability of sediment from  erosion of bedrock or  prior existence of material and the sediment transport capacity generated by fluvial processes below the glacier ice?
+\citep{creyts2013,brinkerhoff2017,beaud2018,delaney2019}.
+  These models  essentially represent subglacial sediment transport processes as similar to sediment transport processes in rivers, with the main difference being that when water flows through a subglacial conduit, it is pressurized by the ice above \citep{rothlisberger1972}.
 
-  Both subglacially and fluvially, shear stress between the water and the sediment that it flows over is a first-order control on the transport of sediment \citep{shields1936} if an adequate quantity of sediment is present as to be transported fluvially (i.e.
+ Shear stress between the water and the sediment that it flows over has a first-order control on the transport of sediment in both  subglacialand fluvial systems, \citep{shields1936} , but only if an adequate quantity of sediment is present so as to be transported fluvially (i.e.
 in a ``transport-limited'' regime).
-  Shear stress increases with the water velocity and roughness of the channel.
-  Greater shear stress will mobilize increased quantities of sediment, along with larger sediment clasts.
-  Sediment transport relationships such as \citet{shields1936}, \citet{meyer1948},  and \citet{engelund1967} use shear stress, or an approximation thereof, to evaluate sediment transport capacity per unit width.
-  The total sediment transport capacity across a channel's cross depends on the integrated shear stress over the entire channel bed (Figure~\ref{fig:cartoon}).
-  As a result, channel geometry plays a key role in sediment transport in both fluvial channels and subglacial ones \citep[e.g.][]{church2006,delaney2019}.
+Confusing ; fluvial is used in too many contexts.
+How can you clarify this?
+ between the water and the sediment that it flows over : this is an awkward phrasing. Is this correct: shear stress applied by the flowing water on the channel bed?
+  The shear stress increases with increasing water velocity and roughness of the channel.
+  Greater shear stress will mobilize increased quantities of sediment, along with larger sediment clasts. Confusing - need to specify grain size here. what is meant by clast?
+  Sediment transport relationships described by  \citet{shields1936}, \citet{meyer1948},  and \citet{engelund1967} , for example, use shear stress, or an approximation thereof [ is this needed here?, confusing], to evaluate sediment transport capacity per unit width of ______?.
+  The total sediment transport capacity across a channel's cross section depends on the integrated shear stress over the entire channel bed (Figure~\ref{fig:cartoon}).
+Here, the channel cross section refers to _______ (site a, Fig.
+___) and the channel bed refers to ____ (site b, Fig.
+__)
+Thus,  channel geometry plays a key role in sediment transport in both fluvial channels and subglacial ones because _________________ state the reason why \citep[e.g.][]{church2006,delaney2019}.
 
-  However, in response to changing water discharge, the behavior of subglacial channels differs from fluvial ones.
-  In subglacial channels, the water is pressurized in a subglacial channel and the size of this channel responds to the roughness of the channel walls and the velocity of the water flowing through the channel \citep[Figure~\ref{fig:cartoon}\,a and c; e.g.
+In a regime that is not transport limited [???
+How is this different from the previous paragraph?]  the behavior of _________ what process?
+in subglacial channels differs from that in fluvial channels  in response to changing water discharge, Again, not following here...
+is there changing water discharge in the paragraph above?
+
+  In subglacial environments?
+systems?
+, the water is pressurized and the changing?
+size of thechannel responds to the roughness of the walls and the water velocity   \citep[Figure~\ref{fig:cartoon}\,a and c; e.g.
 ][]{rothlisberger1972}.
-  Here, diurnal variations in water flux, for instance, may cause water to flow through a conduit whose size does not respond to the same timescales as the increase in water flux \citep{rothlisberger1972}.
-  With respect to sediment transport, this means that increased water discharge in subglacial channels will result in greater water velocity and thus sediment transport capacity.
-  At the same time, the fixed channel width in subglacial channels means that the channel width over which to mobilize sediment does not evolve over the same timescales as water discharge \citep[e.g.][]{werder2010}.
+you have used "channel" many times in this sentence - confusing.
+ Diurnal variations in water flux may cause water to flow through a conduit whose size does not respond at the same timescale(s) as the increase in water flux \citep{rothlisberger1972}.
+This means that increased water discharge in subglacial channels can result in greater water velocity and thus sediment transport capacity.
+  At the same time, a relatively  fixed channel width  means that the distance over sediment is mobilized does not evolve over the same timescale(s) as water discharge \citep[e.g.][]{werder2010}.
 
-  Conversely, in fluvial systems with open-channel flow, a commensurate increase in water discharge will be accommodated by both a reduced increase in water velocity, along with increased channel width and/or depth (Figure~\ref{fig:cartoon}\,b and d).
-  The change in water velocity could be lower, but the width over which sediment can be mobilized changes \citep{leopold1953}.
-  Thus, the systematic response in sediment transport capacity with respect to changing water discharge is  different in the two systems.
+In fluvial systems with open-channel flow [why is this specified as compared with subglacial system?
+needed?], a commensurate [relative to ___?, delete?] increase in water discharge will be accommodated by  a reduced increase [whoa - hard to figure out] a reduction?
+in water velocity along with an increase in channel width and/or depth (Figure~\ref{fig:cartoon}\,b and d).
+  The change in water velocity could be lower [not following here], but the width over which sediment can be mobilized changes \citep{leopold1953}.
+[not following here, think through a clear explanation even if this requires more sentences.
+]
+  Thus, the systematic response in sediment transport capacity with respect to changing water discharge is  different in the two systems because ______________.
 
-  The different processes in subglacial and fluvial systems have been implicitly included in a wide range of available models and relationships that transport sediment in both glacial and fluvial environments \citep[e.g.][]{walder1994,tucker1997,beaud2018,wickert2019}.
-  Furthermore, models of fluvial transport of sediment in glacial and fluvial systems have been used to understand the relationship between sediment dynamics and climate \citep[e.g.][]{tucker1997,delaney2020}
+ Differences in the processes of ________________in subglacial and fluvial systems have been implicitly included in a wide range of available models that evaluate sediment transport relationships.
+\citep[e.g.][]{walder1994,tucker1997,beaud2018,wickert2019}.
+  Furthermore, models of \textbf{fluvial} [again, two uses of fluvial in this sentence is confusing] transport of sediment in glacial and \textbf{fluvial} systems investigate the relationship between sediment dynamics and climate change?
+\citep[e.g.][]{tucker1997,delaney2020}
   Yet, the differing relationship between variations in water discharge and sediment discharge in subglacial and fluvial systems has been less well discussed.
 
-  To compare the differences in the response of sediment transport capacity to water discharge, this manuscript leverages parameterizations of subglacial hydraulics and fluvial hydraulics.
-  Parameterizations are applied to explore the differences in the relationship between sediment transport capacity and water discharge and the impact of water discharge variability on sediment transport capacity in the two systems.
+ This manuscript leverages parameterizations of subglacial hydraulics and fluvial hydraulics to explore  differences in the relationships among sediment transport capacity and water discharge.
+It also explores  the impact of water discharge variability on sediment transport capacity in idealized?
+subglacial and fluvial  two systems.
   Lastly, the implications for sediment transport processes in subglacial systems, compared to fluvial ones, are discussed.
@@
       \caption{Cartoon of the situation with various parameters.
-Velocity $v_g$ and $v_f$ magnitude are shown by arrow length and marks at the channel boundary.
-$S_g$ and $S_f$ represent the cross-sectional area in glacial and fluvial channels.
+Water?
+Velocity magnitude in the glacial, $v_g$ , and fluvial (river?) ______?
+,$v_f$ ,systems are shown by arrow length and lines of ______greatest value?
+marks at the channel boundary.
+$S_g$ and $S_f$ represent the cross-sectional area in glacial and fluvial (river?) channels.
 Note that in the fluvial channel parameterization, rectangular channel shapes are implemented (Section~\ref{sect:fluv}).}
@@
 
-  \section{Methodology}
+  \section{Methods}
   \label{sect:meth}
-  This section contains the relationships between water discharge, subglacial water velocity, and channel geometry in both fluvial and subglacial channels.
-  Because sediment transport largely depends on shear stress \citep{shields1936}, both parameterizations calculated this value and integrate  it across the channel bed.
-  The parameterizations evaluate shear stress, as opposed to sediment transport capacity, as shear stress omits a grain size parameter and the choice of a sediment transport relationship \citep{shields1936}.
+ Relationships among water discharge, subglacial water velocity, and channel geometry in both fluvial (river?) and subglacial channels (Table ___) are described with subglacial and fluvial(word?) parameterization models.
+  Because sediment transport largely depends on shear stress applied by water flow \citep{shields1936}, both parameterizations calculatethis value and integrate  it across the channel bed (Fig. __).
+  The evaluation of shear stress  omits a grain size parameter and, thus, leaves open the selection of a sediment transport relationship  \citep{shields1936}.
+and an eventual sediment transport capacity.
+[is this what you mean?]
 
@@
   To accomplish this, the hydraulics parameterization presented in \citet{delaney2019} is used.
-  The parameterization uses the assumption that the water is transported through subglacial channels \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972}, and that a channel and geometry will respond to a representative water discharge over a certain time period $Q_{w}^*$, using the Darcy-Weisbach formulation for water-flow through a pipe to describe this relationship \citep[e.g.][]{rothlisberger1972,clarke2003,werder2013}.
-  Here, it is assumed that the hydraulic diameter of the subglacial channel, $D_h$, responds to a representative water discharge, $Q_{w}^*$, and a representative gradient of the hydraulic potential, $\Psi^*$:
+Here, it is assumed that the water is transported through subglacial channels \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972}, and that the?
+channel  and geometry shape and size?
+will respond to a representative water discharge over a certain time period $Q_{w}^*$, using the Darcy-Weisbach formulation for water-flow through a pipe  \citep[e.g.][]{rothlisberger1972,clarke2003,werder2013}.
+  It is also assumed??
+More, specifically,  the hydraulic diameter of the subglacial channel, $D_h$, responds to a representative water discharge, $Q_{w}^*$, and a representative gradient of the hydraulic potential, $\Psi^*$ such that:
   \begin{linenomath*}
@@
   \end{equation}
-  where $\beta$ is the central angle of the circular segment that comprises the channel (the so-called Hooke angle).
-  Smaller values of $\beta$ result in shallow, wide channels. $\beta =\pi$ corresponds to a semi-circle.
-  The width of the flat channel floor $w_c$ is given by
+  where $\beta$ is the central angle of the circular segment that comprises the channel (the so-called Hooke angle). Note that $\beta =\pi$ corresponds to a semi-circle and
+  smaller values of $\beta$ result in shallow, wide channels. 
+  The width of the flat channel floor $w_c$ , where beta equals _____? is given by
   \begin{equation}
@@
   where $v_g = \frac{Q_w}{S_g}$ is the water velocity.
-  Here, $Q_w$ represents the instantaneous water discharge, not used to evaluate the hydraulic gradient of the channel in Equation~\ref{eq:DW}.
-  The channel size is usually not in equilibrium with the quantity of water flowing through it.
-  With this formulation, the processes in an R-channel can be represented algebraically \citep{delaney2019}.
+  Here, $Q_w$ represents the instantaneous water discharge; it is not used ?? to evaluate the hydraulic gradient of the channel in Equation~\ref{eq:DW}.
+  It is assumed that?? the channel size is usually not in equilibrium with the quantity of water flowing through it.
+  With this formulation, the processes in an R-channel can be represented algebraically \citep{delaney2019}. What is an "R-channel"? doesn't seem to be defined above.
 
@@
 
-  The fluvial channel parameterization implements the hydraulics parameterization presented in \citet{tucker1997}.
-  Here, the assuming the conservation of mass, sufficiently wide channels so that the hydraulic radius is consistent with flow depth, uniform flow, and the Darcy-Weisbach relationship,
-  the shear stress $\tau_f$ at the river bed is represented as
+ To parameterize?? describe? the shear stress of water flowing across sediments in the  fluvial channel  the hydraulics parameterization presented in \citet{tucker1997} is implemented.
+  Here, it is assumed that there is  conservation of mass and a sufficiently wide channel so that the hydraulic radius, ____,  is consistent with flow depth, uniform flow, and the Darcy-Weisbach relationship.
+  The shear stress $\tau_f$ at the river bed is represented as
   \begin{linenomath*}
@@
   % 
-  where $k$ is a constant and $e$ is an exponent commonly equal to $\frac{1}{2}$ \citep{leopold1953}. Cross-sectional area is 
+  where $k$ is a constant and $e$ is an exponent commonly equal to $\frac{1}{2}$ \citep{leopold1953}.
+  Cross-sectional area, $ S_f$, is 
   \begin{equation}
     \label{eq:Sf}
-    S_f= \frac{Q_w}{v_f}.
+    S_f = \frac{Q_w}{v_f},
   \end{equation}
+  %
+  where $v_f$ is water velocity, given by rearranging Equation \ref{eq:tau} as
+ \begin{equation}
+    \label{eq:vf}
+    v_f = \sqrt{\frac{8\,\tau_f}{f_p\,\rho_w}}.
+  \end{equation}
+  %
   As in Section~\ref{sect:sub_mode}, the width integrated shear stress is $\tau_{tf}=w_{cf}\,\tau_f$.
@@
 
-  The formulation for the glacial system above requires inputs of the hydraulic gradient, water discharge, representative water discharge, Hooke angle $\beta$ \citep{hooke1990}, and friction factor $f_r$.
-  Values of $\beta$ and $f_r$ are chosen so that they produce reasonable values of water velocity, in accordance with available measurements through dye-tracing \citep[Section~\ref{sect:sub_mode}, Figure~\ref{fig:model_outs}; e.g.][]{werder2010}.
+  The formulation for the _______ of the glacial system , eqn _____, requires inputs of the hydraulic gradient, ___, water discharge, ____, representative water discharge, ____,  Hooke angle $\beta$ \citep{hooke1990}, and friction factor, $f_r$.
+Help the reader find these quantities and equations!
+  Values of $\beta$ and $f_r$ are chosen so that they produce reasonable values of water velocity, ___________ m/s, in accordance with  measurements made through dye-tracing \citep[Section~\ref{sect:sub_mode}, Figure~\ref{fig:model_outs}; e.g.][]{werder2010}.
+in xyz glacial environments.
+Give the reader context for this!
 
-  In the fluvial system, the water is not pressurized, so only the bed slope is needed. Friction factor $f_f$ and channel shape factor $k$ are assigned (Section~\ref{sect:fluv}).
+  In the fluvial river?
+system, the water is not pressurized, so only the bed slope is needed to define the _____________, eqn ______.
+The friction factor, $f_f$ ,and channel shape factor, $k$, are assigned (Section~\ref{sect:fluv}), using xyz measurements from abc stream?
+river?
+environments.
 
-  The Reynolds number $R_e$ is given as 
+  The Reynolds number $R_e$, which gives an evaluation of the ________________, is given as 
   \begin{equation}
@@
   % 
-  \noindent Below we examine variables, $\Gamma$, as water discharge ($Q_w$) water velocity ($v$, $v_f$), shear stress ($\tau$, $\tau_f$), and integrated shear stress ($\tau_t$, $\tau_{tf}$).
+  \noindent Below we examine variables, $\Gamma$, as water discharge ($Q_w$) water velocity ($v$, $v_f$), shear stress ($\tau$, $\tau_f$), and integrated shear stress ($\tau_t$, $\tau_{tf}$) with the purpose of evaluating __________________.
 

### 2023-03-07 theory.tex
polishing

--- 2023-02-28 theory.tex
@@
 \newcommand{\ian}[1]{{\textbf{\color{blue}Ian says:} \color{blue} #1} }
+\newcommand{\marie}[1]{{\textbf{\color{green}Marie says:} \color{green} #1} }
 \newcommand{\m}{$\,\mathrm{m}$\,}
@@
 \newenvironment{abstract_}{
-  \chapter{\centering \textbf{Abstract}}
+  \chapter{\cemuchntering \textbf{Abstract}}
 }
@@
 \author[1]{Ian Delaney}
-\affil[1]{Institut des dynamiques de la surface terrestre (IDYST), Universit\'{e} de Lausanne, B\^{a}timent G\'{e}opolis, CH-1015 Lausanne}
-\title{Sediment transport capacity varies more in subglacial systems compared to fluvial ones}
+\affil[1]{Institut des dynamiques de la surface terrestre (IDYST), Universit\'{e} de Lausanne
+
+ B\^{a}timent G\'{e}opolis, CH-1015 Lausanne}
+\title{Comparison of sediment transport capacity variations in subglacial and subaerial systems }
 
@@
 \begin{enumerate}
-\item Sediment transport capacity varies more in pressurized subglacial channels compared to open fluvial ones, in response to changing water discharge due to fixed conduit size of their channels.
-\item Differences in the sediment transport capacity response to changing water discharge could impact processes at proglacial margins and the interpretation of sediment transport records from glaciers.
+\item The sediment carrying capacity of subglacial water varies more in response to changing water discharge compared to subaerial channels, as subglacial channels are pressurized by the ice above.
+\item The slowly evolving size of subglacial channels means that variations in water discharge must be accommodated by changing the velocity of water flowing through the channel.
+%\marie{I can't fully understand this. Too much information is overlain here. Separate concepts and give a more linear explanation.} 
+\item  As water discharge changes, differences in the sediment transport capacity response of subaerial vs subglacial channels  could substantially  impact sediment transport processes at proglacial margins as well as the interpretation of sediment transport records in glacial environments.
 \end{enumerate}
@@
 \abstract % 150 words
-Sediment transport capacity depends largely on the velocity of water flowing through a channel and the channel width over which to mobilize sediment.
-In fluvial channels, changing water discharge can be accommodated by water depth, width, and water velocity.
-In subglacial channels, over short time scales, however, water flows through a conduit with a fixed size, and the glacier ice above pressurizes the water.
-As a result, changing water discharge only may result in modifying water velocity, not the channel's shape.
-Parameterizations of flow in both pressurized, subglacial channels and  open, fluvial channels show that sediment transport capacity varies  more in response to changing water discharge in subglacial channels compared to fluvial ones.
-This characteristic of increased variability in sediment transport capacity in subglacial systems holds true  across a wide range of channel geometries and flow conditions.
+Sediment transport capacity depends mainly on the velocity of water flowing through a channel and the channel width over which the sediment is mobilized.
+In subaerial channels, changing water discharge can be accommodated by changes in the water depth, the width of the channel that  water flows over, and water velocity.
+In subglacial channels, however, water flows through a conduit with a fixed size and shape, and the water is pressurized by the overlying glacier ice.
+As a result, changing water discharge may modify the water velocity, but not the  shape of the channel.
+Parameterization of channel dynamics indicates that increased variability in sediment transport capacity in subglacial systems occurs across a wide range of channel geometries and flow conditions.
 The increased variability in sediment transport capacity impacts sediment transport processes and the interpretation of records of sediment transport from glacierized catchments.
+%\marie{This last sentence seems repetitive.
+I do not fully understand the results.
+Suggest revision after a complete update of the manuscript}
 
@@
 \laysummary % 200 words
-Sediment transport capacity responds differently to changing water discharge in subglacial and fluvial channels.
-In both subglacial and fluvial channels, sediment transport capacity, to the first order, depends on 1) the channel width over-which to mobilize sediment and 2) the water's shear stress on sediment at the channel bottom to mobilize it.
+It is well known that sediment transport capacity responds differently to changing water discharge in subglacial and subaerial channels.
+In both subglacial and subaerial channels, sediment transport capacity depends, in the first order, on 1) the channel width over-which the sediment is mobilized  and 2) the shear stress applied by flowing water on sediment at the channel bottom.
 The shear stress depends largely on water velocity.
-In fluvial channels, changing water discharge will change the water depth, water velocity (controlling sediment transport capacity),  and channel width (which dictates the span over which sediment transport  will occur).
+In subaerial channels, changing water discharge will change the water depth, the water velocity (controlling sediment transport capacity),  and channel width (which dictates the span over which sediment transport  will occur).
 Conversely, in subglacial environments, on short times scales of hours to days, subglacial channels only experience a change in water velocity following changing water discharge, because the water flows through a subglacial conduit with a fixed geometry.
 The fixed conduit geometry also means that the channel width, over-which sediment can be mobilized, does not respond to changing water discharge.
-As a result, sediment transport capacity varies far more in subglacial channels compared to fluvial ones, even though changing channel width can accommodate some sediment transport variability in fluvial channels.
-This increased variability has several important implications for sediment transport in glacial systems compared to fluvial ones.
+As a result, sediment transport capacity varies far more in subglacial channels compared to subaerial ones, even though changing channel width can accommodate some sediment transport variability in subaerial channels.
+This increased variability has several important implications for sediment transport in glacial systems compared to subaerial ones.
+%\maire{This is hard to follow. Suggest you rewrite after greater clarity in the rest of the paper.}
 
@@
 
-  Glaciers are well known for the quantities of sediment that they expel from their termini \citep{hallet1996}.
-  Changing glacier dynamics, geomorphology, and hydrology  has prompted a recent focus on  sediment transport processes in cold regions \citep[e.g.][]{zhang2022}.          
-  Numerical models consider the subglacial component of sediment discharge as a function of the availability of sediment, from the erosion of bedrock or the prior existence of material, and the sediment transport capacity by fluvial processes below the glacier \citep{creyts2013,brinkerhoff2017,beaud2018,delaney2019}.
-  These models often represent sediment transport processes similar to rivers, with the main difference being that water flows through a subglacial conduit, pressurized by the ice above \citep{rothlisberger1972}.
+  Glaciers expel large quantities of sediment from their termini \citep{hallet1996}.
+  Changes in glacier dynamics, geomorphology, and hydrology  have prompted numerous  recent studies of  sediment transport processes in cold regions \citep[e.g.][]{zhang2022}.
+  Increased sediment discharge at both the glacier and regional scale \citep{costa2017,delaney2020,li2022,vergara2022}, creates an imperative to examine the processes controlling sediment discharge from glacierized catchments.
 
-  Both subglacially and fluvially, shear stress between the water and the sediment that it flows over is a first-order control on the transport of sediment \citep{shields1936} if an adequate quantity of sediment is present as to be transported fluvially (i.e.
-in a ``transport-limited'' regime).
-  Shear stress increases with the water velocity and roughness of the channel.
-  Greater shear stress will mobilize increased quantities of sediment, along with larger sediment clasts.
-  Sediment transport relationships such as \citet{shields1936}, \citet{meyer1948},  and \citet{engelund1967} use shear stress, or an approximation thereof, to evaluate sediment transport capacity per unit width.
-  The total sediment transport capacity across a channel's cross depends on the integrated shear stress over the entire channel bed (Figure~\ref{fig:cartoon}).
-  As a result, channel geometry plays a key role in sediment transport in both fluvial channels and subglacial ones \citep[e.g.][]{church2006,delaney2019}.
+Over long periods, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from below glaciers \citep[c.f.][]{hallet1979,iverson1993,herman2021}.
+ On shorter time periods, pressurized water transports available sediment from the glacier \citep{creyts2013,beaud2018b,delaney2019}, should enough sediment be present below the glacier (i.e.
+in a "transport"-limited regime).
 
-  However, in response to changing water discharge, the behavior of subglacial channels differs from fluvial ones.
-  In subglacial channels, the water is pressurized in a subglacial channel and the size of this channel responds to the roughness of the channel walls and the velocity of the water flowing through the channel \citep[Figure~\ref{fig:cartoon}\,a and c; e.g.
-][]{rothlisberger1972}.
-  Here, diurnal variations in water flux, for instance, may cause water to flow through a conduit whose size does not respond to the same timescales as the increase in water flux \citep{rothlisberger1972}.
-  With respect to sediment transport, this means that increased water discharge in subglacial channels will result in greater water velocity and thus sediment transport capacity.
-  At the same time, the fixed channel width in subglacial channels means that the channel width over which to mobilize sediment does not evolve over the same timescales as water discharge \citep[e.g.][]{werder2010}.
+In a "transport"-limited regime, sediment discharge responds to the sediment transport capacity or the amount of sediment that could be carried by the water.
+Sediment transport capacity depends on the shear stress between water and sediment it flows  over \citep{shields1936,meyer1948,engelund1967} and the width of the channel $w$ over which to mobilize sediment.
+The shear stress  across the channel width $\tau_t$ responds to the velocity of water $v$ flowing through at channel so that
+\begin{linenomath*}
+    \begin{equation}
+      \label{eq:tau_t}
+        \tau_t \propto w\, v^2,
+        \end{equation}
+  \end{linenomath*}
+  % 
+  where $w$ is the width of the channel.
+  Following the conservation of mass, the velocity of the water flowing through a rectangular channel is given as 
+  \begin{linenomath*}
+    \begin{equation}
+      \label{eq:v}
+  v = \frac{Q_w}{S(w)},
+  \end{equation}
+  \end{linenomath*}
+%
+where $S$ is the channel's cross-sectional area that responds to channel width $w$, and $Q_w$ is water discharge.
 
-  Conversely, in fluvial systems with open-channel flow, a commensurate increase in water discharge will be accommodated by both a reduced increase in water velocity, along with increased channel width and/or depth (Figure~\ref{fig:cartoon}\,b and d).
-  The change in water velocity could be lower, but the width over which sediment can be mobilized changes \citep{leopold1953}.
-  Thus, the systematic response in sediment transport capacity with respect to changing water discharge is  different in the two systems.
+In subaerial channels, such as rivers with open channel flow, the depth $d$ and width $w$ of the channel evolve with changing water discharge $Q_w$ \citep{leopold1953}.
+During a flood with greater water discharge $Q_w$, for instance, cross-sectional area $S$ increases with channel width $w$ along with velocity $v$.
+As a result, the shear stress across the channel $\tau_t$ will increase in response to both width $w$ and $v$.
 
-  The different processes in subglacial and fluvial systems have been implicitly included in a wide range of available models and relationships that transport sediment in both glacial and fluvial environments \citep[e.g.][]{walder1994,tucker1997,beaud2018,wickert2019}.
-  Furthermore, models of fluvial transport of sediment in glacial and fluvial systems have been used to understand the relationship between sediment dynamics and climate \citep[e.g.][]{tucker1997,delaney2020}
-  Yet, the differing relationship between variations in water discharge and sediment discharge in subglacial and fluvial systems has been less well discussed.
+Yet, the response of water velocity to changing water discharge in subglacial channels is different than subaerial ones.
+In subglacial systems, the water is pressurized by the ice above \citep{shreve1972}.
+The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by frictional heating of water flowing through the channel \citep{rothlisberger1972}.
+As a result, the subglacial channel size can evolve on a longer timescale than variations in water discharge \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
+Changes in water discharge $Q_w$ are mainly accommodated in water velocity $v$, because the width $w$ and depth $d$ of the channel are fixed (Figure~\ref{fig:cartoon}).
+It follows that greater water discharge $Q_w$ will cause the shear stress across the channel $\tau_t$ to primarily respond to the water velocity $v$, as opposed to the subaerial system where width $w$ may evolve as well.
 
-  To compare the differences in the response of sediment transport capacity to water discharge, this manuscript leverages parameterizations of subglacial hydraulics and fluvial hydraulics.
-  Parameterizations are applied to explore the differences in the relationship between sediment transport capacity and water discharge and the impact of water discharge variability on sediment transport capacity in the two systems.
-  Lastly, the implications for sediment transport processes in subglacial systems, compared to fluvial ones, are discussed.
+As a result, sediment mobilization in subaerial and subglacial systems responds differently to changing water discharge.
+These differences have been implicitly included in a wide range of available models that  quantify sediment transport from these environments in both subglacial and subaerial systems  \citep[e.g.][]{walder1994,tucker1997,creyts2013,wickert2019,delaney2019}.
+  The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \citep[e.g.][]{lane2016}.
+  Additionally, the variable response of sediment transport capacity to water discharge in the two systems may impact the interpretation of records of sediment transport in glacier compared to subaerial systems \citep[e.g.][]{muller1968,richards2003}.
+  Yet, the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial systems has been minimally discussed.
+
+ This manuscript leverages parameterizations of subglacial and subaerial hydraulics to explore differences in the relationships between  water discharge, channel geometry, and water velocity. 
+ With these parameterizations, the manuscript discusses the impact of water discharge variability on sediment transport capacity in idealized subglacial and subaerial systems.
+  Lastly, the implications for sediment transport processes in subglacial systems, compared to subaerial ones, are discussed.
 
@@
       \caption{Cartoon of the situation with various parameters.
-Velocity $v_g$ and $v_f$ magnitude are shown by arrow length and marks at the channel boundary.
-$S_g$ and $S_f$ represent the cross-sectional area in glacial and fluvial channels.
-Note that in the fluvial channel parameterization, rectangular channel shapes are implemented (Section~\ref{sect:fluv}).}
+Water velocity magnitude in the subglacial, $v_g$ , and subaerial ,$v_f$ , systems are shown by arrow length \marie{and lines of xxx greatest value} marks at the channel boundary.
+$S_g$ and $S_f$ represent the cross-sectional area in glacial and subaerial channels.
+Note that in the subaerial channel parameterization, rectangular channel shapes are implemented (Section~\ref{sect:fluv}).}
       \label{fig:cartoon}
@@
 
-  \section{Methodology}
+  \section{Methods}
   \label{sect:meth}
-  This section contains the relationships between water discharge, subglacial water velocity, and channel geometry in both fluvial and subglacial channels.
-  Because sediment transport largely depends on shear stress \citep{shields1936}, both parameterizations calculated this value and integrate  it across the channel bed.
-  The parameterizations evaluate shear stress, as opposed to sediment transport capacity, as shear stress omits a grain size parameter and the choice of a sediment transport relationship \citep{shields1936}.
+ Relationships among water discharge, subglacial water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}) are described with subglacial and subaerial parameterizations.
+  Because sediment transport largely depends on shear stress applied by water flow \citep{shields1936}, both parameterizations calculate this value and integrate  it across the channel bed (Fig.
+\ref{fig:cartoon}).
+  The evaluation of shear stress  omits a grain size parameter and, thus, leaves open the selection of a sediment transport relationship  \citep{shields1936} and thus the calculation of sediment transport capacity.
 
-  \ian{check units on fluvial channel factor}
+  \ian{check units on subaerial channel factor}
   \begin{table}[H]
@@
       \textbf{Variables}  & & & \\
-      Channel hydraulic diameter (glacier, fluvial) &  $D_{hg},\,D_{hf}$&  & $\mathrm{m}$     \\
-      Width of channel floor (glacial, fluvial) & $w_{cg},w_{cf}$&  & $\mathrm{m}$     \\
-      Channel cross-sectional area (glacier, fluvial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
+      Channel hydraulic diameter (glacier, subaerial) &  $D_{hg},\,D_{hf}$&  & $\mathrm{m}$     \\
+      Width of channel floor (glacial, subaerial) & $w_{cg},w_{cf}$&  & $\mathrm{m}$     \\
+      Channel cross-sectional area (glacier, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
       Water discharge (instantaneous) & $Q_w$& & $\mathrm{m^{3}\,s^{-1}}$ \\
@@
       Representative hydraulic gradient  &$\Psi^*$ & & $\mathrm{Pa\, m^{-1}}$\\
-      Water velocity (glacier, fluvial)  & $v_g,\,v_{f}$& & $\mathrm{m\,s^{-1}}$ \\
-      Shear stress (glacier, fluvial) & $\tau_g,\,\tau_f$&& $\mathrm{Pa \, m^{-1}}$ \\
-      Width-integrated shear-stress (glacier, fluvial) & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$ \\
+      Water velocity (glacier, subaerial)  & $v_g,\,v_{f}$& & $\mathrm{m\,s^{-1}}$ \\
+      Shear stress (glacier, subaerial) & $\tau_g,\,\tau_f$&& $\mathrm{Pa \, m^{-1}}$ \\
+      Width-integrated shear-stress (glacier, subaerial) & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$ \\
       Reynolds number &$R_e$& & $\mathrm{(-)}$\\
@@
       Friction factor (glacier) & $f_r$ & $4$ & $\mathrm{(-)}$ \\
-      Friction factor (fluvial) & $f_p$ & $3$ & $\mathrm{(-)}$\\
+      Friction factor (subaerial) & $f_p$ & $3$ & $\mathrm{(-)}$\\
       Gradient of glacier surface & $\frac{\partial z_{sg}}{\partial x}$ &$0.25$& $\mathrm{(-)}$\\
-      Gradient of channel bed (fluvial) &$\frac{\partial z_c}{\partial x}$ &$0.05$& $\mathrm{(-)}$\\
-      Fluvial channel factor & $k$ &$3$ & $\mathrm{m^{-2}\, s}$\\
+      Gradient of channel bed (subaerial) &$\frac{\partial z_c}{\partial x}$ &$0.05$& $\mathrm{(-)}$\\
+      Subaerial channel factor & $k$ &$3$ & $\mathrm{m^{-2}\, s}$\\
       Channel geometry exponent &$e$& $\frac{1}{2}$&$\mathrm{(-)}$ \\
@@
   To accomplish this, the hydraulics parameterization presented in \citet{delaney2019} is used.
-  The parameterization uses the assumption that the water is transported through subglacial channels \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972}, and that a channel and geometry will respond to a representative water discharge over a certain time period $Q_{w}^*$, using the Darcy-Weisbach formulation for water-flow through a pipe to describe this relationship \citep[e.g.][]{rothlisberger1972,clarke2003,werder2013}.
-  Here, it is assumed that the hydraulic diameter of the subglacial channel, $D_h$, responds to a representative water discharge, $Q_{w}^*$, and a representative gradient of the hydraulic potential, $\Psi^*$:
+Here, it is assumed that the water is transported through subglacial channels \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972}, and that the channel  and geometry will respond to a representative water discharge over a certain time period $Q_{w}^*$, using the Darcy-Weisbach formulation for water-flow through a pipe  \citep[e.g.][]{rothlisberger1972,clarke2003,werder2013}.
+  Thus, the hydraulic diameter of the subglacial channel, $D_h$, responds to a representative water discharge, $Q_{w}^*$, and a representative gradient of the hydraulic potential, $\Psi^*$ such that:
   \begin{linenomath*}
@@
   \end{equation}
-  where $\beta$ is the central angle of the circular segment that comprises the channel (the so-called Hooke angle).
-  Smaller values of $\beta$ result in shallow, wide channels. $\beta =\pi$ corresponds to a semi-circle.
-  The width of the flat channel floor $w_c$ is given by
-  \begin{equation}
+  where $\beta$ is the central angle of the circular segment that comprises the channel (the so-called Hooke angle). Note that $\beta =\pi$ corresponds to a semi-circle and
+  smaller values of $\beta$ result in shallow, wide channels. 
+  The width of the flat channel floor $w_c$ , given $beta$ is   \begin{equation}
     \label{eq:dh2wc}
@@
   where $v_g = \frac{Q_w}{S_g}$ is the water velocity.
-  Here, $Q_w$ represents the instantaneous water discharge, not used to evaluate the hydraulic gradient of the channel in Equation~\ref{eq:DW}.
-  The channel size is usually not in equilibrium with the quantity of water flowing through it.
-  With this formulation, the processes in an R-channel can be represented algebraically \citep{delaney2019}.
+  Here, $Q_w$ represents the instantaneous water discharge, rather than $Q_w^*$, which is used to evaluate the channel geometry.
+  With this formulation, the processes in an R-channel, through which water flows subglacially, can be represented algebraically \citep{rothlisberger1972,delaney2019}.
 
-  For purposes discussed below, we evaluate the width integrated shear stress as $\tau_{tg}=w_{cg}\,\tau_g $.
+  For purposes discussed below, width integrated shear stress is represented as $\tau_{tg}=w_{cg}\,\tau_g $.
 
-  \subsection{Fluvial channel  parameterization}
+  \subsection{Subaerial channel  parameterization}
   \label{sect:fluv}
 
-  The fluvial channel parameterization implements the hydraulics parameterization presented in \citet{tucker1997}.
-  Here, the assuming the conservation of mass, sufficiently wide channels so that the hydraulic radius is consistent with flow depth, uniform flow, and the Darcy-Weisbach relationship,
-  the shear stress $\tau_f$ at the river bed is represented as
+ To parameterize the shear stress of water flowing across sediments in the subaerial channel,  the hydraulics parameterization presented in \citet{tucker1997} is implemented.
+  Here, it is assumed that there is conservation of mass and a sufficiently wide channel so that the hydraulic radius is consistent with flow depth, uniform flow, and the Darcy-Weisbach relationship.
+  The shear stress $\tau_f$ at the river bed is represented as
   \begin{linenomath*}
@@
   \end{linenomath*}
-  where $\frac{\partial z_c}{\partial x}$ is the channel slope, and $f_p$ is the friction factor for fluvial systems.
+  where $\frac{\partial z_c}{\partial x}$ is the channel slope, and $f_p$ is the friction factor for subaerial systems.
   Channel width $w_{cf}$ is 
@@
 
-  The formulation for the glacial system above requires inputs of the hydraulic gradient, water discharge, representative water discharge, Hooke angle $\beta$ \citep{hooke1990}, and friction factor $f_r$.
-  Values of $\beta$ and $f_r$ are chosen so that they produce reasonable values of water velocity, in accordance with available measurements through dye-tracing \citep[Section~\ref{sect:sub_mode}, Figure~\ref{fig:model_outs}; e.g.][]{werder2010}.
+The formulation of the glacial system (Section~\ref{sect:sub_mode}) requires inputs of the hydraulic gradient (Equation \ref{eq:psi}; $\frac{\partial z_{sg}}{\partial x}$, $\frac{\partial z_{c}}{\partial x}$), water discharge, $Q_w$ representative water discharge, $Q_w^*$,  Hooke angle (Equation~\ref{eq:Hf}; $\beta$), and friction factor (Equation~\ref{eq:DW}; $f_r$).
+\marie{Help the reader find these quantities and equations}
+  Values of $\beta$ and $f_r$ are chosen so that they produce reasonable values of water velocity ($0.6-- 1.2$\,\unit{m}\,\unit{s}$^{-1}$) in accordance with measurements made through dye-tracing experiments to examine subglacial hydraulic conditions \citep[Section~\ref{sect:sub_mode}, Figure~\ref{fig:model_outs}; e.g.][]{werder2010}.
 
-  In the fluvial system, the water is not pressurized, so only the bed slope is needed. Friction factor $f_f$ and channel shape factor $k$ are assigned (Section~\ref{sect:fluv}).
+  In the subaerial system, the water is not pressurized, so only the bed slope, $\frac{\partial z_c}{\partial x}$, is needed in Equation~\ref{eq:DW_tau}.
+The friction factor, $f_f$, and channel shape factor, $k$, are assigned such that they produce reasonable values for water velocity and channel width  (Section~\ref{sect:fluv}).
 
-  The Reynolds number $R_e$ is given as 
+  The Reynolds number $R_e$, which gives an evaluation of the degree of turbulence of the water, is given as 
   \begin{equation}
@@
   % 
-  \noindent Below we examine variables, $\Gamma$, as water discharge ($Q_w$) water velocity ($v$, $v_f$), shear stress ($\tau$, $\tau_f$), and integrated shear stress ($\tau_t$, $\tau_{tf}$).
+  \noindent Below we examine variables, $\Gamma$, as water discharge ($Q_w$) water velocity ($v$, $v_f$), shear stress ($\tau$, $\tau_f$), and integrated shear stress ($\tau_t$, $\tau_{tf}$) with the purpose of evaluating total shear stress across the glacier bed.
 
@@
   Both parameterizations are forced with a water discharge ($Q_w$) varying between $8$ and $16$ \,\unit{m}$^{3}$\,\unit{s}$^{-1}$, to represent a range of sediment transport conditions, for instance over diurnal fluctuation in water discharge.
-The fluvial channel width factor $k$ is $3$\,\unit{m}$^{2}$\,\unit{s}$^{-1}$, and the Hooke angle $\beta$ is $\frac{\pi}{2}$.
+The subaerial channel width factor $k$ is $3$\,\unit{m}$^{2}$\,\unit{s}$^{-1}$, and the Hooke angle $\beta$ is $\frac{\pi}{2}$.
 
@@
       \caption{Response of channel width (a), hydraulic diameter (b),  Reynolds number  (c),  water velocity (d), shear-stress (e), and width-integrated shear-stress (f)  to different water discharge values.
-The glacial system is in blue and the fluvial system is in green.
+The glacial system is in blue and the subaerial system is in green.
 Horizontal bars in d, e, and f represent the range of values in the  output.
@@
 
-  Across the hydrological forcing, variable water discharge results in a substantial increase in the range of water velocities in the glacial system compared to the fluvial system (Figure~\ref{fig:model_outs}).
-  While water velocity in the glacier system varies by  $0.5$ (Equation~\ref{eq:var}) over the discharge values, in the fluvial system the water velocity only varies by $0.11$.
-  Similarly, shear stress at the glacier bed varies by $0.75$ across the range of water discharge values, whereas shear stress across the fluvial system varies by roughly $0.21$.
+  Across the hydrological forcing, variable water discharge results in a substantial increase in the range of water velocities in the glacial system compared to the subaerial system (Figure~\ref{fig:model_outs}).
+  While water velocity in the glacier system varies by  $0.5$ (Equation~\ref{eq:var}) over the discharge values, in the subaerial system the water velocity only varies by $0.11$.
+  Similarly, shear stress at the glacier bed varies by $0.75$ across the range of water discharge values, whereas shear stress across the subaerial system varies by roughly $0.21$.
 This increased variability  in shear stress between the two systems occurs as the water velocity is raised to the power of $2$ in Equation~\ref{eq:tau}.
 
-  Changing channel geometry accommodates evolving water discharge in the fluvial system (Equation~\ref{eq:wcf}), whereas in the subglacial case, increased water flux can only be accommodated by an increase in velocity (Equation~\ref{eq:DW}).
-  The range of values between width-integrated shear stress in the fluvial system increases to $0.44$, whereas, variability in the glacial system lies at $0.83$ (Figure~\ref{fig:range}).
-  Indeed, the relative  difference in width-integrated shear stress is less between the fluvial and glacier cases, compared to the water velocity and shear stress outputs.
-  The reduced variability in width-integrated shear stress results from the additional width across the fluvial channel that accommodates a portion of the sediment transport capacity.
+  Changing channel geometry accommodates evolving water discharge in the subaerial system (Equation~\ref{eq:wcf}), whereas in the subglacial case, increased water flux can only be accommodated by an increase in velocity (Equation~\ref{eq:DW}).
+  The range of values between width-integrated shear stress in the subaerial system increases to $0.44$, whereas, variability in the glacial system lies at $0.83$ (Figure~\ref{fig:range}).
+  Indeed, the relative  difference in width-integrated shear stress is less between the subaerial and glacier cases, compared to the water velocity and shear stress outputs.
+  The reduced variability in width-integrated shear stress results from the additional width across the subaerial channel that accommodates a portion of the sediment transport capacity.
 
@@
       \includegraphics[width=0.7\linewidth]{multi_run.pdf}
-      \caption{Ratio of range of a variable $\Gamma$ for glacial and fluvial systems for water velocity  (a), shear-stress (b), and  width-integrated shear-stress (c) in response to mean water discharge,
+      \caption{Ratio of range of a variable $\Gamma$ for glacial and subaerial systems for water velocity  (a), shear-stress (b), and  width-integrated shear-stress (c) in response to mean water discharge,
         Range  $\chi$ for each variable is given in Equation~\ref{eq:var}.
-Values greater (less) than $1$ shown in white (gray) are parameter combinations that vary more in a subglacial (fluvial) system.
+Values greater (less) than $1$ shown in white (gray) are parameter combinations that vary more in a subglacial (subaerial) system.
 Plots (d-f) show the response of variable $\Gamma$ to the variability in water discharge ($\chi(Q_w)$) in the parameter combination.
@@
 
-  Across the accepted model runs, variability in water velocity and shear stress remains substantially higher across the subglacial model compared to fluvial ones (Figure~\ref{fig:range}).
-  Water velocity, shear stress, and width-integrated shear stress all have a variability ratio ($\frac{\chi(\Gamma_{g})}{\chi(\Gamma_f)}$) greater than $1$ in the subglacial  system compared to fluvial one (Figure~S1).
+  Across the accepted model runs, variability in water velocity and shear stress remains substantially higher across the subglacial model compared to subaerial ones (Figure~\ref{fig:range}).
+  Water velocity, shear stress, and width-integrated shear stress all have a variability ratio ($\frac{\chi(\Gamma_{g})}{\chi(\Gamma_f)}$) greater than $1$ in the subglacial  system compared to subaerial one (Figure~S1).
 
-  The minimum ratio of  water velocity, $\frac{\chi(v_{g})}{\chi(v_{f})}$, and shear-stress, $\frac{\chi(\tau_{g})}{\chi(\tau_{f})}$, is around $2$, showing that these factors are substantially more variable in subglacial channels compared to fluvial ones.
-  The width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, spans between $1.02$ and $2.4$ because changing channel width accommodates the reduced shear stress in the fluvial system.
+  The minimum ratio of  water velocity, $\frac{\chi(v_{g})}{\chi(v_{f})}$, and shear-stress, $\frac{\chi(\tau_{g})}{\chi(\tau_{f})}$, is around $2$, showing that these factors are substantially more variable in subglacial channels compared to subaerial ones.
+  The width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, spans between $1.02$ and $2.4$ because changing channel width accommodates the reduced shear stress in the subaerial system.
   No runs experienced width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, less than $1$.
 
-  Runs with the smallest variability in discharge ($\chi(Q_w)$) have the greatest variability in water velocity, shear stress, and width-integrated shear stress in subglacial systems compared to fluvial ones (Figure~\ref{fig:range} and S1).
-  Yet, the greatest difference in the two systems at the smallest variations of water discharge still results in greater variability in the subglacial system compared to the fluvial one (Figure S1).
+  Runs with the smallest variability in discharge ($\chi(Q_w)$) have the greatest variability in water velocity, shear stress, and width-integrated shear stress in subglacial systems compared to subaerial ones (Figure~\ref{fig:range} and S1).
+  Yet, the greatest difference in the two systems at the smallest variations of water discharge still results in greater variability in the subglacial system compared to the subaerial one (Figure S1).
 
@@
 
-  The direct comparison between the systems remains difficult as they have with different hydraulic potentials, and the response of channel width to water discharge varies considerably based on the values chosen for $\beta$ and $k$.
+  The direct comparison between the systems remains difficult as they have with different hydraulic potentials, and the response of channel width to water discharge varies considerably based on the values chosen for channel shape factors $\beta$ and $k$.
 
@@
 
-  Both subglacial and fluvial parameterizations here assume that the distribution of velocity and shear stress of water across the channel bed is homogeneous (Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}).
-  The fluvial parameterization assumes that sediment is transported across all channel widths, as opposed to bankfull width when much sediment transport occurs \citep{wolman1960}.
-  As a result, the variability in the fluvial system across the range of discharge values could be underestimated in the parameterization (Figure~\ref{fig:range}).
+  Both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress of water across the channel bed is homogeneous \citep[Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}][]{wainwright2015}.
+  The subaerial parameterization assumes that sediment is transported across all channel widths, as opposed to large flows such as floods when much sediment transport occurs \citep{wolman1960}.
+  As a result, the variability in the subaerial system across the range of discharge values could be underestimated in the parameterization (Figure~\ref{fig:range}).
 
@@
   \label{sect:GI}
-  The consistent relationship between water discharge variability and relative variability of the water velocity, shear stress, and width-integrated shear stress results from the fixed geometry in the glacier system and variable geometry in the fluvial system as they respond to changes in water discharge (Equations~\ref{eq:tau} and \ref{eq:Sf}).
-  Increased variability in sediment transport capacity with respect to discharge has several implications for sediment transport in subglacial systems compared to fluvial ones.
-  Shear stress is scaled to the $\frac{3}{2}$ power  and relies on shear stress exceeding a threshold in sediment transport relationships  such as \citet{meyer1948}.
+  The consistent relationship between water discharge variability and relative variability of the water velocity, shear stress, and width-integrated shear stress results from the fixed geometry in the glacier system and variable geometry in the subaerial system as they respond to changes in water discharge (Equations~\ref{eq:tau} and \ref{eq:Sf}).
+  Increased variability in sediment transport capacity with respect to discharge has several implications for sediment transport in subglacial systems compared to subaerial ones.
+  Shear stress is scaled to the $\frac{3}{2}$ power and relies on shear stress exceeding a threshold in sediment transport relationships  such as in \citet{meyer1948}.
   In the total sediment transport relationship \citet{engelund1967}, the shear stress is scaled to $\frac{5}{2}$ power.
-  In both cases, the exponent greater than $1$ magnifies the variability in sediment transport beyond the highly variable shear stress and width-integrated shear stress described above in subglacial environments (Figure~\ref{fig:model_outs}\, b and d; Figure~S1).
+  In both cases, the exponent greater than $1$ of shear stress in the sediment transport relationship magnifies the variability in sediment transport beyond the highly variable shear stress and width-integrated shear stress described above in subglacial environments (Figure~\ref{fig:model_outs}\, b and d; Figure~S1).
 
@@
   In supply-limited systems, the changing sediment transport capacity (Figure~\ref{fig:model_outs}) would have a minimal effect on sediment discharge due to the absence of sediment available for transport \citep{delaney2019}.
-  Yet, reaching the threshold of motion for sediment more frequently may result in additional sediment transport and the tendency for glaciers to transport sediment in a supply-limited regime \citep[e.g.][]{herman2015}.
+  Yet, reaching the threshold of motion for sediment  may result in additional sediment transport and the tendency for glaciers to transport sediment in a supply-limited regime whereby sediment discharge scales with glacier sliding  \citep[e.g.][]{herman2015,koppes2015}.
 
   The effect of pressurized channel flow could be most pronounced during precipitation or flood events when subglacial conduits have not adapted to the water discharge flowing through them, resulting in a very large increase in sediment transport capacity and discharge \citep[e.g.][]{cowan1988,delaney2019}.
-  Increased sediment discharge may result from changes to sediment access below the glacier, yet rapid increases in water velocity from water flowing through a small conduit, not conditioned by the amount of water flow, will cause a large increase in sediment transport capacity (Section~\ref{sect:sub_mode}).
+  Increased sediment discharge may result from changes to sediment access below the glacier, yet rapid increases in water velocity from water flowing through a small channel will cause a large increase in sediment transport capacity (Section~\ref{sect:sub_mode}).
 
-  Increased variability in sediment transport also impacts sediment export from glacier margins into proglacial areas \citep[e.g.][]{delaney2017,perolo2018}.
-  Here, in the transition to open-channel flow as the water leaves the glacier, velocities at the highest water discharges could generally decrease (Figure~\ref{fig:model_outs}, blue and green horizontal bars), potentially resulting in sediment deposition once sediment enters the proglacial area.
-  Note that the transition from pressurized to open channel flow could occur above the glacier's terminus, when cavity closure rates are reduced and water can flow freely \citep{egli2021b}, resulting in intermittent deposition and remobilization of sediment below the glacier terminus \citep{perolo2018}.
+  Increased variability in sediment transport may also impact sediment export from the subglacial environment into proglacial areas \citep[e.g.][]{delaney2017,perolo2018}.
+  Here, velocities at the highest water discharges could generally decrease (Figure~\ref{fig:model_outs}, blue and green horizontal bars), in the transition of pressurized subglacial flow to subaerial open-channel flow as the water leaves the glacier.
+The decrease in water velocity potentially results in sediment deposition once sediment enters the proglacial area.
+  Note that the transition from pressurized to open channel flow could occur above glacier termini when cavity closure rates are reduced and water can flow freely \citep{egli2021b}. 
+  Periodic sediment deposition in these areas may even result in the transition between pressurized and subaerial flow near the glacier terminus \citep{perolo2018}.
 
   In large catchments and further downstream of proglacial margins, water discharge is generally less variable than at the tops of catchments, where glaciers lie \citep[c.f.][]{costa2017,vanas2017,delaney2018,hasholt2018}.
-  Here, external hydrological factors further downstream of glaciers could result in smaller variations in sediment transport capacity in many fluvial systems compared to subglacial ones.
+  Here, external hydrological factors further downstream of glaciers could result in smaller variations in sediment transport capacity in many subaerial systems compared to subglacial ones.
   The parameterization here shows that the smallest variability in water discharge has the largest difference in relative variability from subglacial to proglacial systems (Figure~\ref{fig:range}\, d,\,e,\,f), while the variability in the subglacial system remains higher (Figure~S1).
   Therefore, at locations such as the Greenland Ice Sheet, with smaller diurnal variations in water discharge compared to many alpine catchments \citep[c.f.][]{delaney2018,hasholt2018}, the variability in sediment transport capacity in the transition from the ice sheet to the proglacial river could be more dramatic than alpine systems as water leaves the glacier and moves downstream.
-  Thus the disparity in sediment transport capacity between the subglacial and fluvial systems could be most pronounced in these catchments, potentially resulting in temporally mismatched periods between when glaciers introduce sediment to these river systems  and when the river mobilizes it (Figures~\ref{fig:range}~ and~\ref{fig:model_outs}).
+  Thus the disparity in sediment transport capacity between the subglacial and subaerial systems could be most pronounced in these catchments, potentially resulting in temporally mismatched periods between when glaciers introduce sediment to these river systems  and when the river mobilizes it (Figures~\ref{fig:range}~ and~\ref{fig:model_outs}).
 
@@
   Increasing variability in sediment transport down-glacier is likely because of spatial discontinuities in sediment availability, increasing water discharge down the glacier, and the transport times of subglacial sediment \citep{williams1989,delaney2019}.
-  The rapid increase and decreases in sediment transport capacity over a diurnal cycle in a subglacial system, for instance, compounds the variability already present in fluvial systems \citep{williams1989,jerolmack2010}.
-  As a result, variations in sediment transport capacity could be responsible for the fluctuations in sediment discharge (``flushings'') from glaciers that are not attributed to variations in water discharge \citep[e.g.][]{richards2003,swift2021}, but could be the result of reduced conduit size and therefore increased sediment transport capacity.
+  The rapid increase and decreases in sediment transport capacity over a diurnal cycle in a subglacial system, for instance, compounds the variability already present in subaerial systems \citep{williams1989,jerolmack2010}.
+  As a result, variations in sediment transport capacity could be responsible for the fluctuations in sediment discharge (``flushings'') from glaciers that are not attributed to variations in water discharge \citep[e.g.][]{richards2003,swift2021}.
+  Aside from the sporadic nature of these events, variations could be the result of increased water velocity resulting from reduced channel size and therefore increased sediment transport capacity.
 
-  Additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to fluvial systems \citep{jansson2005,ganti2016}. 
+  Additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to subaerial systems \citep{jansson2005,ganti2016}. 
   For instance, if a glacier expels sediment in a transport-limited regime, the pressurized nature of subglacial water flow means sediment discharge capacity responds to the thickness of the ice, controlling the closure of the channel, and the surface slope of the glacier, controlling water velocity, in addition to water discharge \citep[Section~\ref{sect:sub_mode}; ] []{rothlisberger1972,shreve1972}.
-  Conversely, sediment transport in most fluvial systems increases with water discharge and bed slope, with the bed slope, probably being more stable over long time periods \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wong2006}.
-  Furthermore, if the glacier is in a supply-limited case, the sediment discharge record will also represent additional processes of sliding  \citep{herman2015,seguinot2021}, adding additional processes represented in the sediment discharge signal \citep{delaney2019}.
+  Conversely, sediment transport in most subaerial systems increases with water discharge and bed slope, with the bed slope, probably being more stable over long time periods \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wong2006}.
+  Furthermore, if no sediment is available and the glacier is in a supply-limited case, then the sediment discharge record will also represent additional processes of sliding  \citep{herman2015,seguinot2021}, adding additional processes represented in the sediment discharge signal \citep{delaney2019}.
 
@@
   The variability in subglacial sediment transport capacity presented here applies most generally to short-time scales responsible for the size of subglacial channels.
-  Yet, identifying climatic signals in sediment transport from transport-limited glaciers may require higher thresholds of climatic perturbation compared to fluvial systems \citep{tofelde2021}, given the potential amount of additional noise \citep{castletort2003,jerolmack2010,romans2016}.
+  Yet, identifying climatic signals in sediment transport from transport-limited glaciers may require higher thresholds of climatic perturbation compared to subaerial systems \citep{tofelde2021}, given the potential amount of additional noise \citep{castletort2003,jerolmack2010,romans2016}.
   For instance, sediment discharge records from two glaciers in the Swiss Alps show that $40$ to $50$\% of a season's sediment discharge occurs when water discharge is below the $75^{\mathrm{th}}$ quantile of the season \citep{delaney2018}, possibly because sediment transport capacity relies on the size of the glacier conduit, in addition to water discharge (Section~\ref{sect:sub_mode}).
-  This suggests that a larger climate perturbation might be needed to significantly alter the sediment discharge capacity from the glaciers compared to fluvial systems, to overcome the increased variability in sediment transport capacity.
+  This suggests that a larger climate perturbation might be needed to significantly alter the sediment discharge capacity from the glaciers compared to subaerial systems, to overcome the noise resulting from increased variability in sediment transport capacity.
 
@@
 
-  Application of parameterizations for subglacial and fluvial water flow shows that variable processes occur in sediment transport capacity's response to water discharge, due to their fixed and variable channel geometry, respectively.
-  Results show that the impact of variable water discharge on water velocity and shear stress, first-order controls on sediment mobilization, is much higher in subglacial channels than in fluvial ones.
-  The variability in shear stress is reduced when the evolving channel width is accounted for in fluvial systems.
-  Even so, variability in shear stress across a channel's width is still higher in subglacial channels, compared to fluvial ones.
+  \ian{simplify as did in introduction}
 
-  These different characteristics between glacial and fluvial channels show that records of sediment transport downstream of glaciers capture multiple fluvial processes together.
+  Application of parameterizations for subglacial and subaerial water flow shows that variable processes occur in sediment transport capacity's response to water discharge, due to their fixed and variable channel geometry, respectively.
+  Results show that the impact of variable water discharge on water velocity and shear stress, first-order controls on sediment mobilization, is much higher in subglacial channels than in subaerial ones.
+  The variability in shear stress is reduced when the evolving channel width is accounted for in subaerial systems.
+  Even so, variability in shear stress across a channel's width is still higher in subglacial channels, compared to subaerial ones.
+
+  These different characteristics between glacial and subaerial channels show that records of sediment transport downstream of glaciers capture multiple subaerial processes together.
   Such processes might be particularly relevant at glacier margins, where water transitions from pressurized subglacial flow to open-channel flow.
-  The inconsistent response of sediment transport in subglacial and fluvial systems to changing water discharge could add noise to sediment transport signals between the systems.
+  The inconsistent response of sediment transport in subglacial and subaerial systems to changing water discharge could add noise to sediment transport signals between the systems.
   In addition, the subglacial parameterization shows that variations in sediment transport capacity in glacier systems respond to a large number of factors such as channel size, ice thickness, and glacier surface slope, which may react to climate forcing differently than water discharge alone.
-  The differences in fluvial and subglacial channels' response to changing water discharge should be considered when examining sediment transport processes in glacierized catchments.
+  The differences in subaerial and subglacial channels' response to changing water discharge should be considered when examining sediment transport processes in glacierized catchments.
 
@@
     \includegraphics[width=0.7\linewidth]{multi_run_vars.pdf}
-    \caption{Variability $\chi$ in velocity ($v_g$, $v_f$; a) shear stress ($\tau_g$, $\tau_f$; b) and width-integrated shear stress ($\tau_{tg}$, $\tau_{tf}$; c)  in glacial and fluvial systems. }
+    \caption{Variability $\chi$ in velocity ($v_g$, $v_f$; a) shear stress ($\tau_g$, $\tau_f$; b) and width-integrated shear stress ($\tau_{tg}$, $\tau_{tf}$; c)  in glacial and subaerial systems. }
     \label{fig:gammas}
@@
 % LocalWords:  guillon exner wolman engelund Weisbach
-
-As a result, the parameterization here is most relevant for relatively wide rivers, where channel width evolves greatly with water discharge (Equation~\ref{eq:wcf}).
-% This is probably the case in many proglacial rivers and glacierized catchments where the divergent behavior of subglacial and fluvial channels is most relevant (Figure~\ref{sect:GI}) \ian{Find reference about the geometry of proglacial river}.
-
-Qbs_out Qws_out

### 2023-03-09 theory.tex
update

@@
   B\^{a}timent G\'{e}opolis, CH-1015 Lausanne}
-\title{Comparison of sediment transport capacity variations in subglacial and subaerial systems }
+
+\title{Comparing variations in sediment transport capacity between subglacial and subaerial systems }
 
@@
 \begin{enumerate}
-\item The sediment carrying capacity of subglacial water varies more in response to changing water discharge compared to subaerial channels, as subglacial channels are pressurized by the ice above.
-\item The slowly evolving size of subglacial channels means that variations in water discharge must be accommodated by changing the velocity of water flowing through the channel.
-%\marie{I can't fully understand this. Too much information is overlain here. Separate concepts and give a more linear explanation.} 
-\item  As water discharge changes, differences in the sediment transport capacity response of subaerial vs subglacial channels  could substantially  impact sediment transport processes at proglacial margins as well as the interpretation of sediment transport records in glacial environments.
+\item The slowly evolving size of subglacial channels means that variations in water discharge must be accommodated by changing the velocity of water flowing through the channel, as opposed to increasing channel area.
+
+\item Increased variability in water velocity results in greater variability in sediment transport capacity in subglacial systems compared to subaerial ones, although the variable width of subaerial channels reduces some of the total sediment transport capacity across the channel.
+
+\item  As water discharge changes, differences in the sediment transport capacity response of subaerial and subglacial channels could impact sediment transport processes at proglacial margins as well as the interpretation of sediment transport records in glacial environments.
 \end{enumerate}
@@
 \abstract % 150 words
-Sediment transport capacity depends mainly on the velocity of water flowing through a channel and the channel width over which the sediment is mobilized.
+Sediment transport capacity depends mainly on the shear stress or velocity of water flowing through a channel and the channel width over which the sediment is mobilized.
 In subaerial channels, changing water discharge can be accommodated by changes in the water depth, the width of the channel that  water flows over, and water velocity.
-In subglacial channels, however, water flows through a conduit with a fixed size and shape, and the water is pressurized by the overlying glacier ice.
+In subglacial channels, however, water flows through a conduit with a slowly evolving size, because the water is pressurized by the glacier ice.
 As a result, changing water discharge may modify the water velocity, but not the  shape of the channel.
-Parameterization of channel dynamics indicates that increased variability in sediment transport capacity in subglacial systems occurs across a wide range of channel geometries and flow conditions.
-The increased variability in sediment transport capacity impacts sediment transport processes and the interpretation of records of sediment transport from glacierized catchments.
-%\marie{This last sentence seems repetitive.
-I do not fully understand the results.
-Suggest revision after a complete update of the manuscript}
+Parameterizations of channel dynamics indicate that greater variability in sediment transport capacity in subglacial systems occurs across a wide range of channel geometries and flow conditions.
+The increased variability in water velocity and sediment transport capacity impacts sediment transport processes and the interpretation of records of sediment transport from glacierized catchments.
 
@@
 \laysummary % 200 words
-It is well known that sediment transport capacity responds differently to changing water discharge in subglacial and subaerial channels.
-In both subglacial and subaerial channels, sediment transport capacity depends, in the first order, on 1) the channel width over-which the sediment is mobilized  and 2) the shear stress applied by flowing water on sediment at the channel bottom.
-The shear stress depends largely on water velocity.
+In both subglacial and subaerial channels, sediment transport capacity depends, in the first order, on 1) the channel width over which the sediment is mobilized  and 2) the shear stress applied by flowing water on sediment at the channel bottom, depending largely on the water velocity.
 In subaerial channels, changing water discharge will change the water depth, the water velocity (controlling sediment transport capacity),  and channel width (which dictates the span over which sediment transport  will occur).
@@
 The fixed conduit geometry also means that the channel width, over-which sediment can be mobilized, does not respond to changing water discharge.
-As a result, sediment transport capacity varies far more in subglacial channels compared to subaerial ones, even though changing channel width can accommodate some sediment transport variability in subaerial channels.
-This increased variability has several important implications for sediment transport in glacial systems compared to subaerial ones.
-%\maire{This is hard to follow. Suggest you rewrite after greater clarity in the rest of the paper.}
+Parameterizations of the processes show that sediment transport capacity varies far more in subglacial channels compared to subaerial ones, even though changing channel width can accommodate some sediment transport variability in subaerial channels.
+This increased variability in sediment transport capacity in subglacial systems, compared to subaerial ones, has several important implications for sediment transport in glacierized catchments and the interpretation of sediment discharge records.
 
@@
   Changes in glacier dynamics, geomorphology, and hydrology  have prompted numerous  recent studies of  sediment transport processes in cold regions \citep[e.g.][]{zhang2022}.
-  Increased sediment discharge at both the glacier and regional scale \citep{costa2017,delaney2020,li2022,vergara2022}, creates an imperative to examine the processes controlling sediment discharge from glacierized catchments.
+  Recent papers examining increased sediment discharge form glacierized regions at both the glacier and regional scale \citep{costa2017,delaney2020,li2022,vergara2022}, creates an imperative to examine the processes controlling sediment discharge from glacierized catchments.
 
-Over long periods, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from below glaciers \citep[c.f.][]{hallet1979,iverson1993,herman2021}.
- On shorter time periods, pressurized water transports available sediment from the glacier \citep{creyts2013,beaud2018b,delaney2019}, should enough sediment be present below the glacier (i.e.
-in a "transport"-limited regime).
+  Over long periods, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from below glaciers \citep[c.f.][]{hallet1979,iverson2012,ugelvig2018}.
+  On shorter time periods, pressurized water transports available sediment from the glacier \citep{walder1994,creyts2013,beaud2018b}, should enough sediment be present below the glacier (i.e.
+in a transport-limited regime).
 
-In a "transport"-limited regime, sediment discharge responds to the sediment transport capacity or the amount of sediment that could be carried by the water.
-Sediment transport capacity depends on the shear stress between water and sediment it flows  over \citep{shields1936,meyer1948,engelund1967} and the width of the channel $w$ over which to mobilize sediment.
+  In a transport-limited regime, sediment discharge responds to the sediment transport capacity or the amount of sediment that could be carried by the water.
+Sediment transport capacity depends on the shear stress between water and sediment it flows  over \citep{shields1936,meyer1948,engelund1967} and the width of the channel $w_c$ over which to mobilize sediment.
 The shear stress  across the channel width $\tau_t$ responds to the velocity of water $v$ flowing through at channel so that
@@
       \label{eq:tau_t}
-        \tau_t \propto w\, v^2,
+      \tau_t \propto w_c\, v^2,
     \end{equation}
@@
   % 
-  where $w$ is the width of the channel.
+  where $w_c$ is the width of the channel.
   Following the conservation of mass, the velocity of the water flowing through a rectangular channel is given as 
@@
       \label{eq:v}
-  v = \frac{Q_w}{S(w)},
+      v = \frac{Q_w}{S},
     \end{equation}
@@
   % 
-where $S$ is the channel's cross-sectional area that responds to channel width $w$, and $Q_w$ is water discharge.
+  where $Q_w$ is water discharge,  and the denominator is the channel's cross sectional area $S$.
 
-In subaerial channels, such as rivers with open channel flow, the depth $d$ and width $w$ of the channel evolve with changing water discharge $Q_w$ \citep{leopold1953}.
-During a flood with greater water discharge $Q_w$, for instance, cross-sectional area $S$ increases with channel width $w$ along with velocity $v$.
-As a result, the shear stress across the channel $\tau_t$ will increase in response to both width $w$ and $v$.
+  In subaerial channels,with open channel flow, the cross sectional area $S$ of the channel evolves with changing water discharge $Q_w$  by changing the width of the channel along with the depth \citep{leopold1953}.
+During a flood with greater water discharge $Q_w$, for instance, the cross sectional area $S$ increases as the channel widens and deepens, along with water velocity $v$.
+As a result, the shear stress across the channel $\tau_t$ will increase in response to both the channel width  and water velocity $v$.
 
@@
   As a result, the subglacial channel size can evolve on a longer timescale than variations in water discharge \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
-Changes in water discharge $Q_w$ are mainly accommodated in water velocity $v$, because the width $w$ and depth $d$ of the channel are fixed (Figure~\ref{fig:cartoon}).
-It follows that greater water discharge $Q_w$ will cause the shear stress across the channel $\tau_t$ to primarily respond to the water velocity $v$, as opposed to the subaerial system where width $w$ may evolve as well.
+  In turn, changes in water discharge $Q_w$ are mainly accommodated with water velocity $v$, because the size of the channel $S$ is fixed (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
+  It follows that greater water discharge $Q_w$ will cause the shear stress across the channel $\tau_t$ to respond to the water velocity $v$, as opposed to the subaerial system where width evolves as well.
 
   As a result, sediment mobilization in subaerial and subglacial systems responds differently to changing water discharge.
-These differences have been implicitly included in a wide range of available models that  quantify sediment transport from these environments in both subglacial and subaerial systems  \citep[e.g.][]{walder1994,tucker1997,creyts2013,wickert2019,delaney2019}.
+  These differences have been implicitly included in a wide range of available models that  quantify sediment transport from these environments in both subglacial and subaerial systems  \citep[e.g.][]{walder1994,tucker1997,creyts2013,wickert2019,hewitt2019}.
   The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \citep[e.g.][]{lane2016}.
-  Additionally, the variable response of sediment transport capacity to water discharge in the two systems may impact the interpretation of records of sediment transport in glacier compared to subaerial systems \citep[e.g.][]{muller1968,richards2003}.
+  Additionally, the variable response of sediment transport capacity to water discharge in the two systems may impact the interpretation of records of sediment transport in glacier compared to subaerial systems \citep[e.g.][]{muller1968,richards2003,ganti2016}.
   Yet, the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial systems has been minimally discussed.
@@
   This manuscript leverages parameterizations of subglacial and subaerial hydraulics to explore differences in the relationships between  water discharge, channel geometry, and water velocity. 
- With these parameterizations, the manuscript discusses the impact of water discharge variability on sediment transport capacity in idealized subglacial and subaerial systems.
+  With these parameterizations, the manuscript examines the impact of water discharge variability on sediment transport capacity in idealized subglacial and subaerial systems.
   Lastly, the implications for sediment transport processes in subglacial systems, compared to subaerial ones, are discussed.
@@
     \begin{figure}[H]
-      \includegraphics[width=0.65\linewidth]{Cartoon.pdf}
+      \includegraphics[width=0.65\linewidth]{Cartoon_text.pdf}
       \caption{Cartoon of the situation with various parameters.
-Water velocity magnitude in the subglacial, $v_g$ , and subaerial ,$v_f$ , systems are shown by arrow length \marie{and lines of xxx greatest value} marks at the channel boundary.
-$S_g$ and $S_f$ represent the cross-sectional area in glacial and subaerial channels.
+Water velocity magnitude in the subglacial, $v_g$ , and subaerial ,$v_f$ , systems are shown by arrow length marks at the channel boundary.
+$S_g$ and $S_f$ represent the cross-sectional area in subglacial and subaerial channels.
+The subglacial channel widths $w_{cg}$ remain steady, while the subaerial channel widths evolve $w_{cf}$.
+Text below panels shows different variables response to high and low water discharge ($Q_{wh}$ and $Q_{wl}$, respectively).
+        Relationships between the variables presented here are given in Equations~\ref{eq:tau_t} and~\ref{eq:v}.
         Note that in the subaerial channel parameterization, rectangular channel shapes are implemented (Section~\ref{sect:fluv}).} 
@@
 
-  \ian{check units on subaerial channel factor}
   \begin{table}[H]
@@
       Representative water discharge & $Q_{w}^*$& & $\mathrm{m^{3}\,s^{-1}}$ \\
-      % Gradient of hydraulic potential (glacier) &$\Psi$ & & $\mathrm{Pa\, m^{-1}}$\\
       Representative hydraulic gradient  &$\Psi^*$ & & $\mathrm{Pa\, m^{-1}}$\\
@@
       Gradient of channel bed (subaerial) &$\frac{\partial z_c}{\partial x}$ &$0.05$& $\mathrm{(-)}$\\
-      Subaerial channel factor & $k$ &$3$ & $\mathrm{m^{-2}\, s}$\\
+      Subaerial channel factor & $k$ &$3$ & $\mathrm{s\,m^{-2}}$\\
       Channel geometry exponent &$e$& $\frac{1}{2}$&$\mathrm{(-)}$ \\
@@
   The formulation of the glacial system (Section~\ref{sect:sub_mode}) requires inputs of the hydraulic gradient (Equation \ref{eq:psi}; $\frac{\partial z_{sg}}{\partial x}$, $\frac{\partial z_{c}}{\partial x}$), water discharge, $Q_w$ representative water discharge, $Q_w^*$,  Hooke angle (Equation~\ref{eq:Hf}; $\beta$), and friction factor (Equation~\ref{eq:DW}; $f_r$).
-\marie{Help the reader find these quantities and equations}
-  Values of $\beta$ and $f_r$ are chosen so that they produce reasonable values of water velocity ($0.6-- 1.2$\,\unit{m}\,\unit{s}$^{-1}$) in accordance with measurements made through dye-tracing experiments to examine subglacial hydraulic conditions \citep[Section~\ref{sect:sub_mode}, Figure~\ref{fig:model_outs}; e.g.][]{werder2010}.
 
@@
 The subaerial channel width factor $k$ is $3$\,\unit{m}$^{2}$\,\unit{s}$^{-1}$, and the Hooke angle $\beta$ is $\frac{\pi}{2}$.
+  Values of Hooke angle $\beta$ and subglacial friction factor $f_r$ are chosen so that they produce reasonable values of water velocity ($0.6-- 1.2$\,\unit{m}\,\unit{s}$^{-1}$) in accordance with measurements made through dye-tracing experiments to examine subglacial hydraulic conditions \citep[Section~\ref{sect:sub_mode}, Figure~\ref{fig:model_outs}; e.g.][]{werder2010}.
 
@@
 
-  Both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress of water across the channel bed is homogeneous \citep[Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}][]{wainwright2015}.
+  Both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress of water across the channel bed is homogeneous \citep[Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}][]{yager2018}.
   The subaerial parameterization assumes that sediment is transported across all channel widths, as opposed to large flows such as floods when much sediment transport occurs \citep{wolman1960}.
@@
   \label{sect:GI}
-  The consistent relationship between water discharge variability and relative variability of the water velocity, shear stress, and width-integrated shear stress results from the fixed geometry in the glacier system and variable geometry in the subaerial system as they respond to changes in water discharge (Equations~\ref{eq:tau} and \ref{eq:Sf}).
+  The relationship between water discharge variability and relative variability of the water velocity, shear stress, and width-integrated shear stress results from the fixed geometry in the glacier system and variable geometry in the subaerial system as they respond to changes in water discharge (Equations~\ref{eq:tau_t} and \ref{eq:v}).
   Increased variability in sediment transport capacity with respect to discharge has several implications for sediment transport in subglacial systems compared to subaerial ones.
@@
   The response of sediment discharge caused by the greater variability in sediment transport in subglacial channels will be greatest in completely transport-limited regimes \citep[e.g.][]{kasmalkar2019}.
-  In supply-limited systems, the changing sediment transport capacity (Figure~\ref{fig:model_outs}) would have a minimal effect on sediment discharge due to the absence of sediment available for transport \citep{delaney2019}.
-  Yet, reaching the threshold of motion for sediment  may result in additional sediment transport and the tendency for glaciers to transport sediment in a supply-limited regime whereby sediment discharge scales with glacier sliding  \citep[e.g.][]{herman2015,koppes2015}.
+  In supply-limited systems or partially supply-limited systems, the changing sediment transport capacity (Figure~\ref{fig:model_outs}) would have a minimal effect on sediment discharge due to the absence of sediment available for transport \citep{delaney2019}.
+  Yet, reaching the threshold of motion for sediment frequently  may result in additional sediment transport and the tendency for glaciers to transport sediment in a supply-limited regime whereby sediment discharge scales with glacier sliding  \citep[e.g.][]{herman2015,koppes2015}.
+  Sediment exhaustion by this process may also explain the dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, as opposed to glacier melt \citep{overeem2017}.
 
-  The effect of pressurized channel flow could be most pronounced during precipitation or flood events when subglacial conduits have not adapted to the water discharge flowing through them, resulting in a very large increase in sediment transport capacity and discharge \citep[e.g.][]{cowan1988,delaney2019}.
+  The effect of pressurized channel flow could be most pronounced during precipitation or flood events when subglacial conduits have not adapted to the water discharge flowing through them, resulting in a  large increase in sediment transport capacity and discharge \citep[e.g.][]{cowan1988,delaney2019}.
   Increased sediment discharge may result from changes to sediment access below the glacier, yet rapid increases in water velocity from water flowing through a small channel will cause a large increase in sediment transport capacity (Section~\ref{sect:sub_mode}).
@@
   Note that the transition from pressurized to open channel flow could occur above glacier termini when cavity closure rates are reduced and water can flow freely \citep{egli2021b}. 
-  Periodic sediment deposition in these areas may even result in the transition between pressurized and subaerial flow near the glacier terminus \citep{perolo2018}.
+  Periodic sediment deposition in sediment deposition may even result in the transition between pressurized and subaerial flow near the glacier terminus \citep{perolo2018}.
 
-  In large catchments and further downstream of proglacial margins, water discharge is generally less variable than at the tops of catchments, where glaciers lie \citep[c.f.][]{costa2017,vanas2017,delaney2018,hasholt2018}.
+  In large catchments and further downstream of proglacial margins, water discharge is generally less variable than at the tops of catchments, where glaciers lie \citep[c.f.][]{swift2005,riihimaki2005,costa2017,vanas2017,hasholt2018}.
   Here, external hydrological factors further downstream of glaciers could result in smaller variations in sediment transport capacity in many subaerial systems compared to subglacial ones.
-  The parameterization here shows that the smallest variability in water discharge has the largest difference in relative variability from subglacial to proglacial systems (Figure~\ref{fig:range}\, d,\,e,\,f), while the variability in the subglacial system remains higher (Figure~S1).
-  Therefore, at locations such as the Greenland Ice Sheet, with smaller diurnal variations in water discharge compared to many alpine catchments \citep[c.f.][]{delaney2018,hasholt2018}, the variability in sediment transport capacity in the transition from the ice sheet to the proglacial river could be more dramatic than alpine systems as water leaves the glacier and moves downstream.
+
+  The parameterization here shows that the smallest variability in water discharge has the largest difference in relative variability in sediment transport parameters (Figure~\ref{fig:range}\, d,\,e,\,f).
+  At locations such as the Greenland Ice Sheet, smaller diurnal variations in water discharge compared to many alpine catchments \citep[c.f.][]{swift2005,riihimaki2005,hasholt2018}.
+  Therefore, the variability in sediment transport capacity in the transition from the ice sheet to the proglacial river could be more dramatic in Greenland than in alpine systems as water leaves the glacier and moves downstream.
   Thus the disparity in sediment transport capacity between the subglacial and subaerial systems could be most pronounced in these catchments, potentially resulting in temporally mismatched periods between when glaciers introduce sediment to these river systems  and when the river mobilizes it (Figures~\ref{fig:range}~ and~\ref{fig:model_outs}).
@@
 
-  Large variations in sediment transport capacity below glaciers may result in sediment mobilization and deposition processes in close spatial and temporal proximity to each other below the glacier \citep{gimbert2016,perolo2018}, in response to both evolving channel geometry and water discharge with accumulating with melt down-glacier \citep{beaud2018,delaney2019}.
+  Large variations in sediment transport capacity below glaciers may result in sediment mobilization and deposition processes in close spatial and temporal proximity to each other below the glacier \citep{gimbert2016,perolo2018}.
+  This sediment mobilization and deposition would occur in response to both evolving channel geometry and water discharge with accumulating with melt down-glacier \citep{beaud2018,delaney2019}.
   Increasing variability in sediment transport down-glacier is likely because of spatial discontinuities in sediment availability, increasing water discharge down the glacier, and the transport times of subglacial sediment \citep{williams1989,delaney2019}.
@@
   As a result, variations in sediment transport capacity could be responsible for the fluctuations in sediment discharge (``flushings'') from glaciers that are not attributed to variations in water discharge \citep[e.g.][]{richards2003,swift2021}.
-  Aside from the sporadic nature of these events, variations could be the result of increased water velocity resulting from reduced channel size and therefore increased sediment transport capacity.
+  Aside from the sporadic nature of these events, variations could be the result of increased water velocity resulting from reduced channel size, not water discharge, therefore increasing sediment transport capacity.
 
   Additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to subaerial systems \citep{jansson2005,ganti2016}. 
-  For instance, if a glacier expels sediment in a transport-limited regime, the pressurized nature of subglacial water flow means sediment discharge capacity responds to the thickness of the ice, controlling the closure of the channel, and the surface slope of the glacier, controlling water velocity, in addition to water discharge \citep[Section~\ref{sect:sub_mode}; ] []{rothlisberger1972,shreve1972}.
-  Conversely, sediment transport in most subaerial systems increases with water discharge and bed slope, with the bed slope, probably being more stable over long time periods \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wong2006}.
-  Furthermore, if no sediment is available and the glacier is in a supply-limited case, then the sediment discharge record will also represent additional processes of sliding  \citep{herman2015,seguinot2021}, adding additional processes represented in the sediment discharge signal \citep{delaney2019}.
+  For instance, sediment expelled from a glacier in a transport-limited regime reacts, in addition to water discharge, to sediment discharge capacity defined by the thickness of the ice, controlling the closure of the channel, and the surface slope of the glacier, controlling water velocity \citep[Section~\ref{sect:sub_mode}; ] []{rothlisberger1972,shreve1972,delaney2022,stevens2022}.
+  Conversely, sediment transport in most subaerial systems increases with water discharge and bed slope, with the bed slope, probably being more stable over long time periods \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wong2006,wickert2019}.
+  Furthermore, if no sediment is available and the glacier is in a supply-limited case, then the sediment discharge record will also represent additional processes of bedrock erosion sliding and water pressure variations  \citep{iverson2012,herman2015}.
 
@@
   The variability in subglacial sediment transport capacity presented here applies most generally to short-time scales responsible for the size of subglacial channels.
-  Yet, identifying climatic signals in sediment transport from transport-limited glaciers may require higher thresholds of climatic perturbation compared to subaerial systems \citep{tofelde2021}, given the potential amount of additional noise \citep{castletort2003,jerolmack2010,romans2016}.
-  For instance, sediment discharge records from two glaciers in the Swiss Alps show that $40$ to $50$\% of a season's sediment discharge occurs when water discharge is below the $75^{\mathrm{th}}$ quantile of the season \citep{delaney2018}, possibly because sediment transport capacity relies on the size of the glacier conduit, in addition to water discharge (Section~\ref{sect:sub_mode}).
+  Yet, identifying climatic signals in sediment transport from transport-limited glaciers may require higher thresholds of climatic perturbation compared to subaerial systems \citep{tofelde2021}, in addition to the stochastic nature of erosion and deposition in fluvial evironments \citep{castletort2003,jerolmack2010,romans2016}.
+  For instance, sediment discharge records from two glaciers in the Swiss Alps show that $40$ to $50$\% of a season's sediment discharge occurs when water discharge is below the $75^{\mathrm{th}}$ quantile of the season \citep{delaney2018}, possibly in part because sediment transport capacity relies on the size of the glacier conduit, in addition to water discharge (Section~\ref{sect:sub_mode}).
   This suggests that a larger climate perturbation might be needed to significantly alter the sediment discharge capacity from the glaciers compared to subaerial systems, to overcome the noise resulting from increased variability in sediment transport capacity.
@@
 
-  \ian{simplify as did in introduction}
+  The pressurized nature of subglacial channels means that changing water discharge is accommodated by altering water velocity and shear stress, upon which sediment transport depends.
+  This occurs because the size of subglacial channels evolves slowly compared to variations in water discharge.
+  Conversely, in subaerial systems changing water discharge modulates the water velocity as well as the channel width over which sediment can be transported.
 
-  Application of parameterizations for subglacial and subaerial water flow shows that variable processes occur in sediment transport capacity's response to water discharge, due to their fixed and variable channel geometry, respectively.
-  Results show that the impact of variable water discharge on water velocity and shear stress, first-order controls on sediment mobilization, is much higher in subglacial channels than in subaerial ones.
-  The variability in shear stress is reduced when the evolving channel width is accounted for in subaerial systems.
-  Even so, variability in shear stress across a channel's width is still higher in subglacial channels, compared to subaerial ones.
+  Parameterizations of subglacial and subaerial water flow show that subglacial systems exhibit increased variability in sediment transport capacity as they respond change water discharge.
+  The variability in sediment transport capacity is reduced when the evolving channel width is accounted for in subaerial systems.
+  Even so, variability in sediment transport capacity across a channel's width is consistently higher in subglacial channels, compared to subaerial ones.
 
-  These different characteristics between glacial and subaerial channels show that records of sediment transport downstream of glaciers capture multiple subaerial processes together.
-  Such processes might be particularly relevant at glacier margins, where water transitions from pressurized subglacial flow to open-channel flow.
-  The inconsistent response of sediment transport in subglacial and subaerial systems to changing water discharge could add noise to sediment transport signals between the systems.
-  In addition, the subglacial parameterization shows that variations in sediment transport capacity in glacier systems respond to a large number of factors such as channel size, ice thickness, and glacier surface slope, which may react to climate forcing differently than water discharge alone.
-  The differences in subaerial and subglacial channels' response to changing water discharge should be considered when examining sediment transport processes in glacierized catchments.
+  These different characteristics between glacial and subaerial channels show that records of sediment transport downstream of glaciers capture both subglacial and subaerial processes together.
+  Such processes might be particularly relevant at glacier margins, where water transitions from pressurized subglacial flow to open-channel subaerial flow.
+  The inconsistent response of sediment transport in subglacial and subaerial systems to changing water discharge could add uncertainty in evaluating  sediment transport signals in subglacial systems.
+  In addition, the subglacial parameterization shows that sediment transport capacity in glacier systems respond to a large number of factors such as channel size, ice thickness, and glacier surface slope, which may react to climate forcing differently than water discharge alone.
+  The differences in subglacial and subaerial channels' response to changing water discharge should be considered when examining sediment transport processes in glacierized catchments.
 

### 2023-03-09 theory.tex
small bits

@@
   % 
-  where $Q_w$ is water discharge,  and the denominator is the channel's cross sectional area $S$.
+  where $Q_w$ is water discharge,  and the denominator is the channel's cross-sectional area $S$.
 
-  In subaerial channels,with open channel flow, the cross sectional area $S$ of the channel evolves with changing water discharge $Q_w$  by changing the width of the channel along with the depth \citep{leopold1953}.
-During a flood with greater water discharge $Q_w$, for instance, the cross sectional area $S$ increases as the channel widens and deepens, along with water velocity $v$.
+  In subaerial channels, with open channel flow, the cross-sectional area $S$ of the channel evolves with changing water discharge $Q_w$  by changing the width of the channel along with the depth \citep{leopold1953}.
+During a flood with greater water discharge $Q_w$, for instance, the cross-sectional area $S$ increases as the channel widens and deepens, along with water velocity $v$.
 As a result, the shear stress across the channel $\tau_t$ will increase in response to both the channel width  and water velocity $v$.
@@
   \label{sect:meth}
-  Relationships among water discharge, subglacial water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}) are described with subglacial and subaerial parameterizations.
+    Parameterizations below represent relationships among water discharge, subglacial water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
   Because sediment transport largely depends on shear stress applied by water flow \citep{shields1936}, both parameterizations calculate this value and integrate  it across the channel bed (Fig.
 \ref{fig:cartoon}).
-  The evaluation of shear stress  omits a grain size parameter and, thus, leaves open the selection of a sediment transport relationship  \citep{shields1936} and thus the calculation of sediment transport capacity.
+  The evaluation of shear stress  omits a grain size parameter and, thus, leaves open the selection of a sediment transport relationship  \citep[e.g.][]{shields1936,meyer1948} to  calculate sediment transport capacity.
 
@@
   \label{sect:sub_mode}
-  To parameterize the shear stress of water flowing across sediments below a glacier, the subglacial channel parameterization evaluates the channel geometry and the velocity of the flowing water. 
-  To accomplish this, the hydraulics parameterization presented in \citet{delaney2019} is used.
+  To evaluate the shear stress of water flowing across sediments below a glacier, the subglacial channel parameterization evaluates the channel geometry and the velocity of the flowing water.
+  To accomplish this, the hydraulics model presented in \citet{delaney2019} is used.
+
   Here, it is assumed that the water is transported through subglacial channels \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972}, and that the channel  and geometry will respond to a representative water discharge over a certain time period $Q_{w}^*$, using the Darcy-Weisbach formulation for water-flow through a pipe  \citep[e.g.][]{rothlisberger1972,clarke2003,werder2013}.
@@
   smaller values of $\beta$ result in shallow, wide channels. 
-  The width of the flat channel floor $w_c$ , given $beta$ is   \begin{equation}
+  The width of the flat channel floor $w_c$, given angle $beta$ is 
+  \begin{equation}
     \label{eq:dh2wc}
@@
   The inconsistent response of sediment transport in subglacial and subaerial systems to changing water discharge could add uncertainty in evaluating  sediment transport signals in subglacial systems.
-  In addition, the subglacial parameterization shows that sediment transport capacity in glacier systems respond to a large number of factors such as channel size, ice thickness, and glacier surface slope, which may react to climate forcing differently than water discharge alone.
-  The differences in subglacial and subaerial channels' response to changing water discharge should be considered when examining sediment transport processes in glacierized catchments.
+  In addition, the subglacial parameterization shows that sediment transport capacity in glacier systems responds to a large number of factors such as channel size, ice thickness, and glacier surface slope, which may react to climate forcing differently than water discharge alone.
+  The differences in subglacial and subaerial channels' response to changing hydrology should be considered when examining sediment transport processes in glacierized catchments.
 

### 2023-03-10 theory.tex
small bits.. and new fig

@@
   Changes in glacier dynamics, geomorphology, and hydrology  have prompted numerous  recent studies of  sediment transport processes in cold regions \citep[e.g.][]{zhang2022}.
-  Recent papers examining increased sediment discharge form glacierized regions at both the glacier and regional scale \citep{costa2017,delaney2020,li2022,vergara2022}, creates an imperative to examine the processes controlling sediment discharge from glacierized catchments.
+  Recent papers examining increased sediment discharge form glacierized regions at both the glacier and regional scale \citep{costa2017,delaney2020,li2022,vergara2022}, create an imperative to examine the processes controlling sediment discharge from glacierized catchments.
+\ian{more details about paper... why have they not examined the processes more?}
 
@@
   % 
-  where $Q_w$ is water discharge,  and the denominator is the channel's cross-sectional area $S$.
+  where $Q_w$ is water discharge,  and $S$ is the channel's cross-sectional area.
 
@@
   The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by frictional heating of water flowing through the channel \citep{rothlisberger1972}.
-  As a result, the subglacial channel size can evolve on a longer timescale than variations in water discharge \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
+  As a result, the subglacial channel size evolves over weeks or months, whereas water discharge varies over hours \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
   In turn, changes in water discharge $Q_w$ are mainly accommodated with water velocity $v$, because the size of the channel $S$ is fixed (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
@@
     \begin{figure}[H]
-      \includegraphics[width=0.65\linewidth]{Cartoon_text.pdf}
+      \includegraphics[width=0.65\linewidth]{Cartoon.pdf}
       \caption{Cartoon of the situation with various parameters.
@@
 The subglacial channel widths $w_{cg}$ remain steady, while the subaerial channel widths evolve $w_{cf}$.
-Text below panels shows different variables response to high and low water discharge ($Q_{wh}$ and $Q_{wl}$, respectively).
-        Relationships between the variables presented here are given in Equations~\ref{eq:tau_t} and~\ref{eq:v}.
         Note that in the subaerial channel parameterization, rectangular channel shapes are implemented (Section~\ref{sect:fluv}).} 
@@
   \section{Conclusions}
-
-  The pressurized nature of subglacial channels means that changing water discharge is accommodated by altering water velocity and shear stress, upon which sediment transport depends.
+  Contrary to subaerial channels that can alter their channel width and water velocity in response to changing water discharge,
+   pressurized subglacial channels accommodate changing water discharge by altering water velocity and shear stress, upon which sediment transport depends.
   This occurs because the size of subglacial channels evolves slowly compared to variations in water discharge.
-  Conversely, in subaerial systems changing water discharge modulates the water velocity as well as the channel width over which sediment can be transported.
 

### 2023-03-10 theory.tex
intro update

@@
   Changes in glacier dynamics, geomorphology, and hydrology  have prompted numerous  recent studies of  sediment transport processes in cold regions \citep[e.g.][]{zhang2022}.
-  Recent papers examining increased sediment discharge form glacierized regions at both the glacier and regional scale \citep{costa2017,delaney2020,li2022,vergara2022}, create an imperative to examine the processes controlling sediment discharge from glacierized catchments.
-\ian{more details about paper... why have they not examined the processes more?}
+  Increases in sediment transport  have been observed in Greenland \citep{bendixen2017}, the European Alps \citep{costa2017}, the Himalayas \citep{li2021} and the Andes \citep{vergara2022}.
+  In some of these regions, increased water discharge and glacier melt have been attributed to the greater sediment transport \citep{bendixen2017,costa2017,li2021}.
+  In other environments, enhanced access of melt water to subglacial sediment at high elevations was attributed as the cause of increased sediment transport \citep{delaney2020,vergara2022}.
+  The observed changes to these systems create an imperative to examine the processes controlling sediment discharge from glacierized catchments. 
+  Yet, the relationship between water discharge and sediment transport in subglacial environments remains minimally discussed and not contrasted with the subaerial river systems that these glaciers feed.
 
   Over long periods, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from below glaciers \citep[c.f.][]{hallet1979,iverson2012,ugelvig2018}.
-  On shorter time periods, pressurized water transports available sediment from the glacier \citep{walder1994,creyts2013,beaud2018b}, should enough sediment be present below the glacier (i.e.
+  On shorter time periods, pressurized water transports sediment from the glacier \citep{walder1994,creyts2013,beaud2018}, should enough sediment be present below the glacier (i.e.
 in a transport-limited regime).
@@
   In a transport-limited regime, sediment discharge responds to the sediment transport capacity or the amount of sediment that could be carried by the water.
-Sediment transport capacity depends on the shear stress between water and sediment it flows  over \citep{shields1936,meyer1948,engelund1967} and the width of the channel $w_c$ over which to mobilize sediment.
+Sediment transport capacity depends on the shear stress between water and sediment it flows  over \citep{shields1936,meyer1948,engelund1967} and the width of the channel bottom $w_c$ over which to mobilize sediment.
 The shear stress  across the channel width $\tau_t$ responds to the velocity of water $v$ flowing through at channel so that
@@
 During a flood with greater water discharge $Q_w$, for instance, the cross-sectional area $S$ increases as the channel widens and deepens, along with water velocity $v$.
-As a result, the shear stress across the channel $\tau_t$ will increase in response to both the channel width  and water velocity $v$.
+As a result, the shear stress across the channel $\tau_t$ will increase in response to both the channel width $w_c$  and water velocity $v$.
 
@@
       \caption{Cartoon of the situation with various parameters.
-Water velocity magnitude in the subglacial, $v_g$ , and subaerial ,$v_f$ , systems are shown by arrow length marks at the channel boundary.
+Water velocity magnitude in the subglacial, $v_g$ , and subaerial ,$v_f$ , systems are shown by arrow length.
 $S_g$ and $S_f$ represent the cross-sectional area in subglacial and subaerial channels.
-The subglacial channel widths $w_{cg}$ remain steady, while the subaerial channel widths evolve $w_{cf}$.
+The subglacial channel widths $w_{cg}$ remain steady, while the subaerial channel widths evolve $w_{cf}$, denoted by thick black line.
+Shear stress $\tau$ is responsible for the mobilization of sediment and increases with the number of makers at the channel bed.
         Note that in the subaerial channel parameterization, rectangular channel shapes are implemented (Section~\ref{sect:fluv}).} 

### 2023-03-14 theory.tex
Changes to key points and abstract...

@@
 \begin{enumerate}
-\item The slowly evolving size of subglacial channels means that variations in water discharge must be accommodated by changing the velocity of water flowing through the channel, as opposed to increasing channel area.
+\item Variations in water discharge in subglacial systems are accommodated by water velocity of water, given the slow evolution of channel size.
 
-\item Increased variability in water velocity results in greater variability in sediment transport capacity in subglacial systems compared to subaerial ones, although the variable width of subaerial channels reduces some of the total sediment transport capacity across the channel.
+\item  Greater variability in water velocity in subglacial systems causes sediment transport capacity to vary more compared to subaerial systems.
 
-\item  As water discharge changes, differences in the sediment transport capacity response of subaerial and subglacial channels could impact sediment transport processes at proglacial margins as well as the interpretation of sediment transport records in glacial environments.
+% although the evolving width of subaerial channels tempers some difference in total sediment transport capacity between the two systems.
+
+\item The systems' divergence in transport capacity response impacts sediment dynamics at glacier margins and interpretation of sediment records.
 \end{enumerate}
@@
 \abstract % 150 words
-Sediment transport capacity depends mainly on the shear stress or velocity of water flowing through a channel and the channel width over which the sediment is mobilized.
-In subaerial channels, changing water discharge can be accommodated by changes in the water depth, the width of the channel that  water flows over, and water velocity.
-In subglacial channels, however, water flows through a conduit with a slowly evolving size, because the water is pressurized by the glacier ice.
-As a result, changing water discharge may modify the water velocity, but not the  shape of the channel.
-Parameterizations of channel dynamics indicate that greater variability in sediment transport capacity in subglacial systems occurs across a wide range of channel geometries and flow conditions.
-The increased variability in water velocity and sediment transport capacity impacts sediment transport processes and the interpretation of records of sediment transport from glacierized catchments.
+Sediment transport capacity depends mainly on the shear stress or velocity of water flowing through a channel and the channel width over which to mobilize sediment.
+In subaerial channels, changing water discharge can be accommodated by rapid changes in the width of the channel over which water flows, the water depth, and the water velocity.
+In subglacial channels, however, water flows through a conduit whose size evolves  slowly, because the water is pressurized by the glacier ice above.
+As a result, variations in water discharge may modify the water velocity, but not the channel's width.
+Here, parameterizations of subglacial and subaerial channels indicate that sediment transport capacity varies more in subglacial systems compared with subaerial systems, across a wide range of channel geometries and flow conditions.
+Large variations in subglacial water sediment transport capacity may impact sediment transport processes and the interpretation of records of sediment transport from glacierized catchments.
 
@@
 \laysummary % 200 words
-In both subglacial and subaerial channels, sediment transport capacity depends, in the first order, on 1) the channel width over which the sediment is mobilized  and 2) the shear stress applied by flowing water on sediment at the channel bottom, depending largely on the water velocity.
-In subaerial channels, changing water discharge will change the water depth, the water velocity (controlling sediment transport capacity),  and channel width (which dictates the span over which sediment transport  will occur).
-Conversely, in subglacial environments, on short times scales of hours to days, subglacial channels only experience a change in water velocity following changing water discharge, because the water flows through a subglacial conduit with a fixed geometry.
-The fixed conduit geometry also means that the channel width, over-which sediment can be mobilized, does not respond to changing water discharge.
-Parameterizations of the processes show that sediment transport capacity varies far more in subglacial channels compared to subaerial ones, even though changing channel width can accommodate some sediment transport variability in subaerial channels.
-This increased variability in sediment transport capacity in subglacial systems, compared to subaerial ones, has several important implications for sediment transport in glacierized catchments and the interpretation of sediment discharge records.
+In both subglacial and subaerial channels, sediment transport capacity depends mainly on 1) the channel width over which the sediment is mobilized  and 2) the shear stress applied by flowing water on sediment at the channel bottom; the shear stress depends largely on the water velocity.
+In subaerial channels, changing water discharge will  change the water depth,  the channel width (which dictates the distance  over which sediment transport  occurs), and the  water velocity (which controls the sediment transport capacity).
+ In subglacial environments, however, on times scales of hours to weeks, changing water discharge will only alter  water velocity, not channel area, because the subglacial conduit maintains a largely fixed geometry.
+The fixed conduit geometry also means that the channel width over which the sediment is  mobilized also does not respond to changing water discharge.
+Parameterizations of these processes show that sediment transport capacity varies far more in subglacial systems compared to subaerial systems, even though changing channel width can accommodate some sediment transport variability in subaerial channels.
+The increased variability in sediment transport capacity in subglacial systems, compared to subaerial ones, has several important implications for sediment transport in glacierized catchments and the interpretation of sediment discharge records.
 
@@
   Changes in glacier dynamics, geomorphology, and hydrology  have prompted numerous  recent studies of  sediment transport processes in cold regions \citep[e.g.][]{zhang2022}.
-  Increases in sediment transport  have been observed in Greenland \citep{bendixen2017}, the European Alps \citep{costa2017}, the Himalayas \citep{li2021} and the Andes \citep{vergara2022}.
+  Increases in sediment transport  have been observed in Greenland \citep{bendixen2017}, the European Alps \citep{costa2017}, the Himalayas \citep{li2021}, and the Andes \citep{vergara2022}.
   In some of these regions, increased water discharge and glacier melt have been attributed to the greater sediment transport \citep{bendixen2017,costa2017,li2021}.
-  In other environments, enhanced access of melt water to subglacial sediment at high elevations was attributed as the cause of increased sediment transport \citep{delaney2020,vergara2022}.
-  The observed changes to these systems create an imperative to examine the processes controlling sediment discharge from glacierized catchments. 
+  In other cases, enhanced access of meltwater to subglacial sediment at high elevations was attributed as the cause of increased sediment transport \citep{delaney2020,vergara2022}.
+  Observed changes to sediment transport in glacierized catchments create an imperative to examine the processes controlling sediment discharge here. 
   Yet, the relationship between water discharge and sediment transport in subglacial environments remains minimally discussed and not contrasted with the subaerial river systems that these glaciers feed.
@@
   Over long periods, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from below glaciers \citep[c.f.][]{hallet1979,iverson2012,ugelvig2018}.
-  On shorter time periods, pressurized water transports sediment from the glacier \citep{walder1994,creyts2013,beaud2018}, should enough sediment be present below the glacier (i.e.
+  On shorter time periods, pressurized water transports sediment from the below glacier \citep{walder1994,creyts2013,beaud2018}, should enough sediment be present below the glacier (i.e.
 in a transport-limited regime).
@@
   The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by frictional heating of water flowing through the channel \citep{rothlisberger1972}.
-  As a result, the subglacial channel size evolves over weeks or months, whereas water discharge varies over hours \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
+  As a result, the subglacial channel size evolves over days to months, whereas water discharge varies over hours \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
   In turn, changes in water discharge $Q_w$ are mainly accommodated with water velocity $v$, because the size of the channel $S$ is fixed (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
@@
   As a result, sediment mobilization in subaerial and subglacial systems responds differently to changing water discharge.
-  These differences have been implicitly included in a wide range of available models that  quantify sediment transport from these environments in both subglacial and subaerial systems  \citep[e.g.][]{walder1994,tucker1997,creyts2013,wickert2019,hewitt2019}.
+  These differences have been implicitly included in a wide range of available models that  quantify sediment transport from these catchments in both subglacial and subaerial systems  \citep[e.g.][]{walder1994,tucker1997,creyts2013,wickert2019,hewitt2019}.
   The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \citep[e.g.][]{lane2016}.
-  Additionally, the variable response of sediment transport capacity to water discharge in the two systems may impact the interpretation of records of sediment transport in glacier compared to subaerial systems \citep[e.g.][]{muller1968,richards2003,ganti2016}.
+  Additionally, the variable response of sediment transport capacity to water discharge in the two systems may impact the interpretation of records of sediment transport in glacial compared to subaerial systems \citep[e.g.][]{muller1968,richards2003,ganti2016}.
   Yet, the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial systems has been minimally discussed.
@@
       \caption{Cartoon of the situation with various parameters.
-Water velocity magnitude in the subglacial, $v_g$ , and subaerial ,$v_f$ , systems are shown by arrow length.
+Water velocity magnitude in the subglacial, $v_g$, and subaerial, $v_f$, systems are shown by arrow length.
 $S_g$ and $S_f$ represent the cross-sectional area in subglacial and subaerial channels.
-The subglacial channel widths $w_{cg}$ remain steady, while the subaerial channel widths evolve $w_{cf}$, denoted by thick black line.
+The subglacial channel widths $w_{cg}$ remain steady, while the subaerial channel widths evolve $w_{cf}$, denoted by the thick black line.
 Shear stress $\tau$ is responsible for the mobilization of sediment and increases with the number of makers at the channel bed.
@@
   \label{sect:meth}
-    Parameterizations below represent relationships among water discharge, subglacial water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
-  Because sediment transport largely depends on shear stress applied by water flow \citep{shields1936}, both parameterizations calculate this value and integrate  it across the channel bed (Fig.
-\ref{fig:cartoon}).
+    Parameterizations below represent relationships amongst water discharge, subglacial water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
+  Because sediment transport largely depends on shear stress applied by water flow \citep{shields1936}, both parameterizations calculate this value and integrate  it across the channel bed (Figure \ref{fig:cartoon}).
   The evaluation of shear stress omits a grain size parameter and, thus, leaves open the selection of a sediment transport relationship  \citep[e.g.][]{shields1936,meyer1948} to  calculate sediment transport capacity.
+  As a result, relative relationships in shear stress, as opposed to sediment transport capacity, are examined.
 
@@
   % 
-  where $v_g = \frac{Q_w}{S_g}$ is the water velocity.
+  where  the water velocity $v_g$ is $v_g = \frac{Q_w}{S_g}$.
   Here, $Q_w$ represents the instantaneous water discharge, rather than $Q_w^*$, which is used to evaluate the channel geometry.
@@
 
-  For purposes discussed below, width integrated shear stress is represented as $\tau_{tg}=w_{cg}\,\tau_g $.
+    Width integrated shear stress is represented as $\tau_{tg}=w_{cg}\,\tau_g $.
 
@@
   % 
-  \noindent Below we examine variables, $\Gamma$, as water discharge ($Q_w$) water velocity ($v$, $v_f$), shear stress ($\tau$, $\tau_f$), and integrated shear stress ($\tau_t$, $\tau_{tf}$) with the purpose of evaluating total shear stress across the glacier bed.
+  \noindent Below we examine variables, $\Gamma$, as water discharge ($Q_w$) water velocity ($v_g$, $v_f$), shear stress ($\tau_g$, $\tau_f$), and integrated shear stress ($\tau_{tg}$, $\tau_{tf}$) with the purpose of evaluating total shear stress across the glacier bed.
 
   \section{Results}
-  First, the subglacial hydraulics parameterization is applied to a representative water discharge ($Q_w^*$; Equation~\ref{eq:DW}) of $12$\,\unit{m}$^{3}$\,\unit{s}$^{-1}$.
-  Both parameterizations are forced with a water discharge ($Q_w$) varying between $8$ and $16$ \,\unit{m}$^{3}$\,\unit{s}$^{-1}$, to represent a range of sediment transport conditions, for instance over diurnal fluctuation in water discharge.
+  Initially, both parameterizations are forced with a water discharge ($Q_w$) varying between $8$ and $16$ \,\unit{m}$^{3}$\,\unit{s}$^{-1}$, to represent a range of sediment transport conditions, for instance over diurnal fluctuation in water discharge.
 The subaerial channel width factor $k$ is $3$\,\unit{m}$^{2}$\,\unit{s}$^{-1}$, and the Hooke angle $\beta$ is $\frac{\pi}{2}$.
-  Values of Hooke angle $\beta$ and subglacial friction factor $f_r$ are chosen so that they produce reasonable values of water velocity ($0.6-- 1.2$\,\unit{m}\,\unit{s}$^{-1}$) in accordance with measurements made through dye-tracing experiments to examine subglacial hydraulic conditions \citep[Section~\ref{sect:sub_mode}, Figure~\ref{fig:model_outs}; e.g.][]{werder2010}.
+  Values of Hooke angle $\beta$ and subglacial friction factor $f_r$ are chosen so that they produce reasonable values of water velocity ($0.6- 1.2$\,\unit{m}\,\unit{s}$^{-1}$) in accordance with measurements made through dye-tracing experiments to examine subglacial hydraulic conditions \citep[Section~\ref{sect:sub_mode}, Figure~\ref{fig:model_outs}; e.g.][]{werder2010}.
+  The subglacial representative water discharge ($Q_w^*$; Equation~\ref{eq:DW}) of $12$\,\unit{m}$^{3}$\,\unit{s}$^{-1}$ is applied to establish the hydraulic diameter and shape of the channel.
 
@@
 
-  The direct comparison between the systems remains difficult as they have with different hydraulic potentials, and the response of channel width to water discharge varies considerably based on the values chosen for channel shape factors $\beta$ and $k$.
-
   The subglacial parameterization assumes that subglacial water is pressurized, which is a commonly held assumption in subglacial hydrology, though observations suggest that this may not always be the case \citep[e.g.][]{gimbert2016}.
-  While the subglacial channel geometry need not be in a steady state for the water to remain pressurized, the subglacial parameterization assumes that the channel geometry must vary minimally over the timescales of varying  water discharge \citep[e.g.][]{nanni2020}.
+  The subglacial channel geometry need not be in equilibrium with the hydrological conditions for the water to remain pressurized.
+  Yet the subglacial parameterization assumes that the channel geometry must vary minimally over the timescales of varying  water discharge \citep[e.g.][]{nanni2020}.
 
@@
   A similar effect could occur with creep closure of the subglacial channel at the hydrograph minima.
-  Additionally, at high water pressures, glacier uplift and bed separation \citep{andrews2014} might slightly modify the width of the subglacial channel.
+  Additionally, at high water pressures, glacier uplift and bed separation \citep{andrews2014} might modify the width of the subglacial channel.
 
@@
   The response of sediment discharge caused by the greater variability in sediment transport in subglacial channels will be greatest in completely transport-limited regimes \citep[e.g.][]{kasmalkar2019}.
-  In supply-limited systems or partially supply-limited systems, the changing sediment transport capacity (Figure~\ref{fig:model_outs}) would have a minimal effect on sediment discharge due to the absence of sediment available for transport \citep{delaney2019}.
-  Yet, reaching the threshold of motion for sediment frequently  may result in additional sediment transport and the tendency for glaciers to transport sediment in a supply-limited regime whereby sediment discharge scales with glacier sliding  \citep[e.g.][]{herman2015,koppes2015}.
+  In supply-limited systems, the changing sediment transport capacity (Figure~\ref{fig:model_outs}) would have a minimal effect on sediment discharge due to the absence of sediment available for transport \citep{delaney2019}.
+  Yet, reaching the threshold of motion for sediment frequently, across a range of discharges,  may result in additional sediment transport and the tendency for glaciers to transport sediment in a supply-limited regime whereby sediment discharge scales with glacier sliding  \citep[e.g.][]{herman2015,koppes2015}.
   Sediment exhaustion by this process may also explain the dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, as opposed to glacier melt \citep{overeem2017}.
@@
 The decrease in water velocity potentially results in sediment deposition once sediment enters the proglacial area.
-  Note that the transition from pressurized to open channel flow could occur above glacier termini when cavity closure rates are reduced and water can flow freely \citep{egli2021b}. 
-  Periodic sediment deposition in sediment deposition may even result in the transition between pressurized and subaerial flow near the glacier terminus \citep{perolo2018}.
+  Periodic sediment deposition in sediment deposition may even result in the transition between pressurized and subaerial flow below the glacier terminus \citep{perolo2018}, especially when cavity closure rates of the ice are slow \citep{egli2021b}.
 
-  In large catchments and further downstream of proglacial margins, water discharge is generally less variable than at the tops of catchments, where glaciers lie \citep[c.f.][]{swift2005,riihimaki2005,costa2017,vanas2017,hasholt2018}.
+  In large catchments and further downstream of proglacial margins, water discharge is generally less variable than at the tops of catchments, where glaciers lie in alpine settings \citep[c.f.][]{swift2005,riihimaki2005,costa2017}.
   Here, external hydrological factors further downstream of glaciers could result in smaller variations in sediment transport capacity in many subaerial systems compared to subglacial ones.
@@
   The parameterization here shows that the smallest variability in water discharge has the largest difference in relative variability in sediment transport parameters (Figure~\ref{fig:range}\, d,\,e,\,f).
-  At locations such as the Greenland Ice Sheet, smaller diurnal variations in water discharge compared to many alpine catchments \citep[c.f.][]{swift2005,riihimaki2005,hasholt2018}.
+  At locations such as the Greenland Ice Sheet, smaller diurnal variations in water discharge compared to many alpine catchments \citep[c.f.][]{swift2005,riihimaki2005,vanas2017,hasholt2018}.
   Therefore, the variability in sediment transport capacity in the transition from the ice sheet to the proglacial river could be more dramatic in Greenland than in alpine systems as water leaves the glacier and moves downstream.
-  Thus the disparity in sediment transport capacity between the subglacial and subaerial systems could be most pronounced in these catchments, potentially resulting in temporally mismatched periods between when glaciers introduce sediment to these river systems  and when the river mobilizes it (Figures~\ref{fig:range}~ and~\ref{fig:model_outs}).
+  Thus the disparity in sediment transport capacity between the subglacial and subaerial systems could be most pronounced in larger catchments, potentially resulting in temporally mismatched periods between when glaciers introduce sediment to these river systems  and when the river mobilizes it (Figures~\ref{fig:model_outs}~ and~\ref{fig:range}).
 
@@
   Large variations in sediment transport capacity below glaciers may result in sediment mobilization and deposition processes in close spatial and temporal proximity to each other below the glacier \citep{gimbert2016,perolo2018}.
-  This sediment mobilization and deposition would occur in response to both evolving channel geometry and water discharge with accumulating with melt down-glacier \citep{beaud2018,delaney2019}.
-  Increasing variability in sediment transport down-glacier is likely because of spatial discontinuities in sediment availability, increasing water discharge down the glacier, and the transport times of subglacial sediment \citep{williams1989,delaney2019}.
-  The rapid increase and decreases in sediment transport capacity over a diurnal cycle in a subglacial system, for instance, compounds the variability already present in subaerial systems \citep{williams1989,jerolmack2010}.
+  The rapid increase and decreases in sediment transport capacity over a diurnal cycle in a subglacial system may compound the variability already present in subaerial systems \citep{williams1989,jerolmack2010}.
   As a result, variations in sediment transport capacity could be responsible for the fluctuations in sediment discharge (``flushings'') from glaciers that are not attributed to variations in water discharge \citep[e.g.][]{richards2003,swift2021}.
-  Aside from the sporadic nature of these events, variations could be the result of increased water velocity resulting from reduced channel size, not water discharge, therefore increasing sediment transport capacity.
+  Aside from the sporadic nature of these events, the observed flushings could result from increases in water velocity resulting from reduced channel size, not water discharge, which increase sediment transport capacity.
 
@@
   For instance, sediment expelled from a glacier in a transport-limited regime reacts, in addition to water discharge, to sediment discharge capacity defined by the thickness of the ice, controlling the closure of the channel, and the surface slope of the glacier, controlling water velocity \citep[Section~\ref{sect:sub_mode}; ] []{rothlisberger1972,shreve1972,delaney2022,stevens2022}.
-  Conversely, sediment transport in most subaerial systems increases with water discharge and bed slope, with the bed slope, probably being more stable over long time periods \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wong2006,wickert2019}.
-  Furthermore, if no sediment is available and the glacier is in a supply-limited case, then the sediment discharge record will also represent additional processes of bedrock erosion sliding and water pressure variations  \citep{iverson2012,herman2015}.
+  Conversely, sediment transport in most subaerial systems increases with water discharge and bed slope, with the bed slope, probably being more stable over yearly to century-scale time periods \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wong2006,wickert2019}.
 
@@
   The variability in subglacial sediment transport capacity presented here applies most generally to short-time scales responsible for the size of subglacial channels.
-  Yet, identifying climatic signals in sediment transport from transport-limited glaciers may require higher thresholds of climatic perturbation compared to subaerial systems \citep{tofelde2021}, in addition to the stochastic nature of erosion and deposition in fluvial evironments \citep{castletort2003,jerolmack2010,romans2016}.
-  For instance, sediment discharge records from two glaciers in the Swiss Alps show that $40$ to $50$\% of a season's sediment discharge occurs when water discharge is below the $75^{\mathrm{th}}$ quantile of the season \citep{delaney2018}, possibly in part because sediment transport capacity relies on the size of the glacier conduit, in addition to water discharge (Section~\ref{sect:sub_mode}).
-  This suggests that a larger climate perturbation might be needed to significantly alter the sediment discharge capacity from the glaciers compared to subaerial systems, to overcome the noise resulting from increased variability in sediment transport capacity.
+  Yet, identifying climatic signals in sediment transport from transport-limited glaciers may require higher thresholds of climatic perturbation compared to subaerial systems \citep{tofelde2021}, in addition to the stochastic nature of erosion and deposition in fluvial environments \citep{castletort2003,jerolmack2010,romans2016}.
+  For instance, sediment discharge records from two glaciers in the Swiss Alps show that $40$ to $50$\% of a season's sediment discharge occurs when water discharge is below the $75^{\mathrm{th}}$ quantile of the season \citep{delaney2018}, possibly in part because sediment transport capacity responds the size of the glacier conduit, in addition to water discharge (Section~\ref{sect:sub_mode}).
+  This suggests that a larger  perturbation of water discharge might be needed to significantly alter the sediment discharge capacity from the glaciers compared to subaerial systems, to overcome the noise resulting from increased variability in sediment transport capacity.
+Furthermore, if no sediment is available and the glacier is in a supply-limited case, then the sediment discharge record will also represent additional processes of bedrock erosion from sliding and water pressure variations  \citep{iverson2012,herman2015}.
 
@@
 
-  Parameterizations of subglacial and subaerial water flow show that subglacial systems exhibit increased variability in sediment transport capacity as they respond change water discharge.
+  Parameterizations of subglacial and subaerial water flow show that subglacial systems exhibit increased variability in sediment transport capacity as they respond  to changes in water discharge.
   The variability in sediment transport capacity is reduced when the evolving channel width is accounted for in subaerial systems.
@@
     \includegraphics[width=0.7\linewidth]{multi_run_vars.pdf}
-    \caption{Variability $\chi$ in velocity ($v_g$, $v_f$; a) shear stress ($\tau_g$, $\tau_f$; b) and width-integrated shear stress ($\tau_{tg}$, $\tau_{tf}$; c)  in glacial and subaerial systems. }
+    \caption{Variability $\chi$ in velocity ($v_g$, $v_f$; a) shear stress ($\tau_g$, $\tau_f$; b) and width-integrated shear stress ($\tau_{tg}$, $\tau_{tf}$; c)  in glacial and subaerial systems.
+Across the parameter space in Figure~\ref{fig:range}.
+}
     \label{fig:gammas}

### 2023-03-15 theory.tex
more small changes

@@
 \begin{enumerate}
-\item Variations in water discharge in subglacial systems are accommodated by water velocity of water, given the slow evolution of channel size.
+\item Water discharge variations in subglacial systems are accommodated by water velocity, not channel size, given the slow evolution of channels.
 
@@
   To parameterize the shear stress of water flowing across sediments in the subaerial channel,  the hydraulics parameterization presented in \citet{tucker1997} is implemented.
-  Here, it is assumed that there is conservation of mass and a sufficiently wide channel so that the hydraulic radius is consistent with flow depth, uniform flow, and the Darcy-Weisbach relationship.
+  Here, it is assumed that there is a conservation of mass and a sufficiently wide channel so that the hydraulic radius is consistent with flow depth, uniform flow, and the Darcy-Weisbach relationship.
   The shear stress $\tau_f$ at the river bed is represented as
@@
       \includegraphics[width=0.7\linewidth]{multi_run.pdf}
-      \caption{Ratio of range of a variable $\Gamma$ for glacial and subaerial systems for water velocity  (a), shear-stress (b), and  width-integrated shear-stress (c) in response to mean water discharge,
-        Range  $\chi$ for each variable is given in Equation~\ref{eq:var}.
-Values greater (less) than $1$ shown in white (gray) are parameter combinations that vary more in a subglacial (subaerial) system.
-Plots (d-f) show the response of variable $\Gamma$ to the variability in water discharge ($\chi(Q_w)$) in the parameter combination.
+      \caption{Ratio of range of water velocity, shear-stress, and width integrated shear-stress for glacial and subaerial systems in response to mean water discharge (a, b, c), and  variability in water discharge (d, e, f).
+        Range  $\chi$ for each variable is given in Equation~\ref{eq:var}. Values greater than $1$ shown in white space are parameter combinations that vary more in a subglacial  system.
         Histograms (g-i) show the frequency of relative variations across variables in the parameterization.}
@@
 
-  The minimum ratio of  water velocity, $\frac{\chi(v_{g})}{\chi(v_{f})}$, and shear-stress, $\frac{\chi(\tau_{g})}{\chi(\tau_{f})}$, is around $2$, showing that these factors are substantially more variable in subglacial channels compared to subaerial ones.
-  The width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, spans between $1.02$ and $2.4$ because changing channel width accommodates the reduced shear stress in the subaerial system.
+  The minimum ratio of  water velocity, $\frac{\chi(v_{g})}{\chi(v_{f})}$, and shear-stress, $\frac{\chi(\tau_{g})}{\chi(\tau_{f})}$, is around $2$, showing that these factors are substantially more variable in subglacial channels compared to subaerial ones (Figure~\ref{fig:range} g, h).
+  The width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, spans between $1.02$ and $2.4$ because changing channel width accommodates some of the reduced shear stress variability in subaerial systems (Figure~\ref{fig:range} i).
   No runs experienced width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, less than $1$.
@@
   Runs with the smallest variability in discharge ($\chi(Q_w)$) have the greatest variability in water velocity, shear stress, and width-integrated shear stress in subglacial systems compared to subaerial ones (Figure~\ref{fig:range} and S1).
-  Yet, the greatest difference in the two systems at the smallest variations of water discharge still results in greater variability in the subglacial system compared to the subaerial one (Figure S1).
+  Yet, the greatest difference in the two systems at the smallest variations of water discharge still results in greater variability in the subglacial system compared to the subaerial one (Figure \ref{fig:range}).
 

### 2023-03-20 theory.tex
draft to GK

@@
   As a result, sediment mobilization in subaerial and subglacial systems responds differently to changing water discharge.
-  These differences have been implicitly included in a wide range of available models that  quantify sediment transport from these catchments in both subglacial and subaerial systems  \citep[e.g.][]{walder1994,tucker1997,creyts2013,wickert2019,hewitt2019}.
+  These differences have been implicitly included in a wide range of available models that  quantify sediment transport from catchments in both subglacial and subaerial systems  \citep[e.g.][]{walder1994,tucker1997,creyts2013,wickert2019,hewitt2019}.
   The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \citep[e.g.][]{lane2016}.
@@
   Across the accepted model runs, variability in water velocity and shear stress remains substantially higher across the subglacial model compared to subaerial ones (Figure~\ref{fig:range}).
-  Water velocity, shear stress, and width-integrated shear stress all have a variability ratio ($\frac{\chi(\Gamma_{g})}{\chi(\Gamma_f)}$) greater than $1$ in the subglacial  system compared to subaerial one (Figure~S1).
+  Water velocity, shear stress, and width-integrated shear stress all have a variability ratio ($\frac{\chi(\Gamma_{g})}{\chi(\Gamma_f)}$) greater than $1$ in the subglacial  system compared to subaerial one (Figure~S \ref{fig:gammas}).
 
@@
 
-  Runs with the smallest variability in discharge ($\chi(Q_w)$) have the greatest variability in water velocity, shear stress, and width-integrated shear stress in subglacial systems compared to subaerial ones (Figure~\ref{fig:range} and S1).
+  Runs with the smallest variability in discharge ($\chi(Q_w)$) have the greatest variability in water velocity, shear stress, and width-integrated shear stress in subglacial systems compared to subaerial ones (Figure~\ref{fig:range} and S \ref{fig:gammas}).
   Yet, the greatest difference in the two systems at the smallest variations of water discharge still results in greater variability in the subglacial system compared to the subaerial one (Figure \ref{fig:range}).
@@
   In the total sediment transport relationship \citet{engelund1967}, the shear stress is scaled to $\frac{5}{2}$ power.
-  In both cases, the exponent greater than $1$ of shear stress in the sediment transport relationship magnifies the variability in sediment transport beyond the highly variable shear stress and width-integrated shear stress described above in subglacial environments (Figure~\ref{fig:model_outs}\, b and d; Figure~S1).
+  In both cases, the exponent greater than $1$ of shear stress in the sediment transport relationship magnifies the variability in sediment transport beyond the highly variable shear stress and width-integrated shear stress described above in subglacial environments (Figure~\ref{fig:model_outs}\, b and d; Figure~S~\ref{fig:gammas}).
 
@@
 
-  In large catchments and further downstream of proglacial margins, water discharge is generally less variable than at the tops of catchments, where glaciers lie in alpine settings \citep[c.f.][]{swift2005,riihimaki2005,costa2017}.
+  In large catchments and further downstream of proglacial margins, water discharge is generally less variable than at the tops of catchments, where glaciers lie in alpine settings \citep[c.f.][]{swift2005,riihimaki2005,costa2017,vanas2017}.
   Here, external hydrological factors further downstream of glaciers could result in smaller variations in sediment transport capacity in many subaerial systems compared to subglacial ones.
 
-  The parameterization here shows that the smallest variability in water discharge has the largest difference in relative variability in sediment transport parameters (Figure~\ref{fig:range}\, d,\,e,\,f).
+  The parameterization here shows that the smallest variability in water discharge has the largest difference in relative variability in sediment transport metrics (Figure~\ref{fig:range}\, d,\,e,\,f).
   At locations such as the Greenland Ice Sheet, smaller diurnal variations in water discharge compared to many alpine catchments \citep[c.f.][]{swift2005,riihimaki2005,vanas2017,hasholt2018}.
@@
 
+\renewcommand{\figurename}{Figure S}
+\setcounter{figure}{0}
+
 \begin{center}

### 2023-03-20 theory.tex
Updates from Overleaf

--- 2023-03-15 theory.tex
@@
 
-  Here, it is assumed that the water is transported through subglacial channels \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972}, and that the channel  and geometry will respond to a representative water discharge over a certain time period $Q_{w}^*$, using the Darcy-Weisbach formulation for water-flow through a pipe  \citep[e.g.][]{rothlisberger1972,clarke2003,werder2013}.
+  Here, it is assumed that the water is transported through subglacial channels \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972}, and that the channel  geometry will respond to a representative water discharge over a certain time period $Q_{w}^*$, using the Darcy-Weisbach formulation for water-flow through a pipe  \citep[e.g.][]{rothlisberger1972,clarke2003,werder2013}.
   Thus, the hydraulic diameter of the subglacial channel, $D_h$, responds to a representative water discharge, $Q_{w}^*$, and a representative gradient of the hydraulic potential, $\Psi^*$ such that:
@@
 
-  In the subaerial system, the water is not pressurized, so only the bed slope, $\frac{\partial z_c}{\partial x}$, is needed in Equation~\ref{eq:DW_tau}.
+  In the subaerial system,  only the bed slope, $\frac{\partial z_c}{\partial x}$, is needed in Equation~\ref{eq:DW_tau}.
 The friction factor, $f_f$, and channel shape factor, $k$, are assigned such that they produce reasonable values for water velocity and channel width  (Section~\ref{sect:fluv}).
@@
   \section{Discussion}
-  \subsection{Assumptions}
-
-  The subglacial parameterization assumes that subglacial water is pressurized, which is a commonly held assumption in subglacial hydrology, though observations suggest that this may not always be the case \citep[e.g.][]{gimbert2016}.
-  The subglacial channel geometry need not be in equilibrium with the hydrological conditions for the water to remain pressurized.
-  Yet the subglacial parameterization assumes that the channel geometry must vary minimally over the timescales of varying  water discharge \citep[e.g.][]{nanni2020}.
-
-  In R-channel formulations, peaks in the hydrograph could cause additional opening of the conduit due to pressure melt, thus causing a decrease in velocity compared to the results here \citep{rothlisberger1972}.
-  A similar effect could occur with creep closure of the subglacial channel at the hydrograph minima.
-  Additionally, at high water pressures, glacier uplift and bed separation \citep{andrews2014} might modify the width of the subglacial channel.
-
-  Both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress of water across the channel bed is homogeneous \citep[Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}][]{yager2018}.
-  The subaerial parameterization assumes that sediment is transported across all channel widths, as opposed to large flows such as floods when much sediment transport occurs \citep{wolman1960}.
-  As a result, the variability in the subaerial system across the range of discharge values could be underestimated in the parameterization (Figure~\ref{fig:range}).
 
@@
   In both cases, the exponent greater than $1$ of shear stress in the sediment transport relationship magnifies the variability in sediment transport beyond the highly variable shear stress and width-integrated shear stress described above in subglacial environments (Figure~\ref{fig:model_outs}\, b and d; Figure~S1).
+
+  Both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress of water across the channel bed is homogeneous \citep[Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}][]{yager2018}.
+  The subaerial parameterization assumes that sediment is transported across all channel widths, as opposed to large flows such as floods when much sediment transport occurs \citep{wolman1960}.
+  As a result, the variability in the subaerial system across the range of discharge values could be underestimated in the parameterization (Figure~\ref{fig:range}).
 
@@
 % LocalWords:  guillon exner wolman engelund Weisbach
+
+  Both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress of water across the channel bed is homogeneous \citep[Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}][]{yager2018}.
+  The subaerial parameterization assumes that sediment is transported across all channel widths, as opposed to large flows such as floods when much sediment transport occurs \citep{wolman1960}.
+  As a result, the variability in the subaerial system across the range of discharge values could be underestimated in the parameterization (Figure~\ref{fig:range}).
+
+  \subsection{Assumptions}
+
+  The subglacial parameterization assumes that subglacial water is pressurized, which is a commonly held assumption in subglacial hydrology, though observations suggest that this may not always be the case \citep[e.g.][]{gimbert2016}.
+  The subglacial channel geometry need not be in equilibrium with the hydrological conditions for the water to remain pressurized.
+  Yet the subglacial parameterization assumes that the channel geometry must vary minimally over the timescales of varying  water discharge \citep[e.g.][]{nanni2020}.

### 2023-03-20 theory.tex

--- 2023-03-20 theory.tex
@@
   In both cases, the exponent greater than $1$ of shear stress in the sediment transport relationship magnifies the variability in sediment transport beyond the highly variable shear stress and width-integrated shear stress described above in subglacial environments (Figure~\ref{fig:model_outs}\, b and d; Figure~S~\ref{fig:gammas}).
+
+  Both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress of water across the channel bed is homogeneous \citep[Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}][]{yager2018}.
+  The subaerial parameterization assumes that sediment is transported across all channel widths, as opposed to large flows such as floods when much sediment transport occurs \citep{wolman1960}.
+  As a result, the variability in the subaerial system across the range of discharge values could be underestimated in the parameterization (Figure~\ref{fig:range}).
 

### 2023-03-20 theory.tex
Pushing back to mac

@@
 \newcommand{\mmma}{$\,\mathrm{m^3\, a^{-1}}$\,}
+\newcommand{\mmms}{$\,\mathrm{m^3\, s^{-1}}$\,}
 \newcommand{\unit}[1]{$\mathrm{#1}$}
@@
   Across the hydrological forcing, variable water discharge results in a substantial increase in the range of water velocities in the glacial system compared to the subaerial system (Figure~\ref{fig:model_outs}).
-  While water velocity in the glacier system varies by  $0.5$ (Equation~\ref{eq:var}) over the discharge values, in the subaerial system the water velocity only varies by $0.11$.
-  Similarly, shear stress at the glacier bed varies by $0.75$ across the range of water discharge values, whereas shear stress across the subaerial system varies by roughly $0.21$.
-This increased variability  in shear stress between the two systems occurs as the water velocity is raised to the power of $2$ in Equation~\ref{eq:tau}.
+  Water velocity in the glacier system vary by  $0.5$ (Equation~\ref{eq:var}) over the discharge values, in the subaerial system the water velocity only varies by $0.11$ (Figure~\ref{fig:model_outs}\,e).
+  Shear stress at the glacier bed varies by $0.75$ across the range of water discharge values, whereas shear stress across the subaerial system varies by roughly $0.21$ (Figure~\ref{fig:model_outs}\,f).
+The increased variability in shear stress relative to water velocity between the two systems occurs as the water velocity is raised to the power of $2$ in Equation~\ref{eq:tau}.
 
-  Changing channel geometry accommodates evolving water discharge in the subaerial system (Equation~\ref{eq:wcf}), whereas in the subglacial case, increased water flux can only be accommodated by an increase in velocity (Equation~\ref{eq:DW}).
-  The range of values between width-integrated shear stress in the subaerial system increases to $0.44$, whereas, variability in the glacial system lies at $0.83$ (Figure~\ref{fig:range}).
-  Indeed, the relative  difference in width-integrated shear stress is less between the subaerial and glacier cases, compared to the water velocity and shear stress outputs.
-  The reduced variability in width-integrated shear stress results from the additional width across the subaerial channel that accommodates a portion of the sediment transport capacity.
+  In the subaerial systems, changing channel geometry accommodates evolving water discharge  (Equation~\ref{eq:wcf}), whereas in the subglacial case, increased water flux can only result in changing water velocity (Equations~\ref{eq:v}~ and ~\ref{eq:DW}).
+  As a result, in the subaerial system, variability in the shear stress $\tau_f$ and width-integrated shear stress $\tau_{tf}$ increases from $0.21$ to $0.44$.
+Variability in width-integrated shear stress  in the glacial system $\tau_{tg}$ is $0.83$ (Figure~\ref{fig:range}).
+  The reduced variability in width-integrated shear stress results from the additional width across the subaerial channel that accommodates some variations in water discharge.
+  Indeed, the relative  difference in width-integrated shear stress is less between the subaerial and glacier cases, compared to the water velocity and shear stress outputs (Figure~\ref{fig:model_outs}\,e--g).
 
-  Next, the parameterizations are run $25,000$ times with randomly selected channel geometries ($k=(2,6)$, $\beta=(\frac{\pi}{10},\pi)$), friction factors ($f_i=(0.001,20)$, $f_p=(0.001,20)$) and water discharges (maximum and minimum values selected between $5$ to $500$ \,\mmma).
-$Q_w^*$ in Equation ~\ref{eq:DW} is determined as the midpoint in the water discharge range.
-Parameter combinations are only selected if their maximum and minimum velocities are between $0.2$ and $3$\,\unit{m}\,\unit{s}$^{-1}$, resulting in roughly $10,600$ adequate parameter combinations.
+  Next, the parameterizations are run $25,000$ times with randomly selected channel geometries ($k=(2,6)$, $\beta=(\frac{\pi}{10},\pi)$), friction factors ($f_i=(0.001,20)$, $f_p=(0.001,20)$) and water discharges (maximum and minimum values selected between $5$ to $500$ \,\mmms).
+$Q_w^*$ in Equation ~\ref{eq:DW} is determined from the mean in the water discharge range.
+Parameter combinations are only selected if their maximum and minimum velocities for both the subglacial and subaerial systems are between $0.2$ and $3$\,\unit{m}\,\unit{s}$^{-1}$, resulting in roughly $10,600$ adequate parameter combinations.
 
@@
 
-  The minimum ratio of  water velocity, $\frac{\chi(v_{g})}{\chi(v_{f})}$, and shear-stress, $\frac{\chi(\tau_{g})}{\chi(\tau_{f})}$, is around $2$, showing that these factors are substantially more variable in subglacial channels compared to subaerial ones (Figure~\ref{fig:range} g, h).
-  The width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, spans between $1.02$ and $2.4$ because changing channel width accommodates some of the reduced shear stress variability in subaerial systems (Figure~\ref{fig:range} i).
-  No runs experienced width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, less than $1$.
+  The minimum ratio of  water velocity, $\frac{\chi(v_{g})}{\chi(v_{f})}$, and shear-stress, $\frac{\chi(\tau_{g})}{\chi(\tau_{f})}$, is greater than $1.5$.
+  These results show that increased variability occurs across a wide range of subaerial and subglacial channel configurations (Figure~\ref{fig:range} g, h).
+The width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, spans between $1.02$ and $2.4$, a smaller amount of relative variability compared to  water velocity $\frac{\chi(v_{g})}{\chi(v_{f})}$ or shear stress $\frac{\chi(\tau_{g})}{\chi(\tau_{f})}$ (Figure~\ref{fig:range}).
+Consistent with the individual case (Figure ~\ref{fig:model_outs}), the subaerial systems' reduce variability in width-integrated shear stress occurs  as the evolving channel width accommodates some of the reduced shear stress variability in subaerial systems (Figure~\ref{fig:range} i).
+  Importantly though, no runs experienced relative variability in width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, less than $1$, showing that the subglacial has consistently high variability in width integrated shear stress compared to subaerial systems.
 
-  Runs with the smallest variability in discharge ($\chi(Q_w)$) have the greatest variability in water velocity, shear stress, and width-integrated shear stress in subglacial systems compared to subaerial ones (Figure~\ref{fig:range} and S \ref{fig:gammas}).
+  Runs with the smallest variability in discharge ($\chi(Q_w)$) have the greatest variability in water velocity, shear stress, and width-integrated shear stress in subglacial systems compared to subaerial ones (Figure~\ref{fig:range}\,d--f and S \ref{fig:gammas}).
   Yet, the greatest difference in the two systems at the smallest variations of water discharge still results in greater variability in the subglacial system compared to the subaerial one (Figure \ref{fig:range}).
+\ian{extra sentence is needed here}
 

### 2023-03-21 theory.tex
small change

@@
   Water velocity, shear stress, and width-integrated shear stress all have a variability ratio ($\frac{\chi(\Gamma_{g})}{\chi(\Gamma_f)}$) greater than $1$ in the subglacial  system compared to subaerial one (Figure~S \ref{fig:gammas}).
+  Runs with the smallest variability in discharge ($\chi(Q_w)$) have the greatest variability in water velocity, shear stress, and width-integrated shear stress in subglacial systems compared to subaerial ones (Figure~\ref{fig:range}\,d--f and S \ref{fig:gammas}).
 
@@
 Consistent with the individual case (Figure ~\ref{fig:model_outs}), the subaerial systems' reduce variability in width-integrated shear stress occurs  as the evolving channel width accommodates some of the reduced shear stress variability in subaerial systems (Figure~\ref{fig:range} i).
-  Importantly though, no runs experienced relative variability in width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, less than $1$, showing that the subglacial has consistently high variability in width integrated shear stress compared to subaerial systems.
-
-  Runs with the smallest variability in discharge ($\chi(Q_w)$) have the greatest variability in water velocity, shear stress, and width-integrated shear stress in subglacial systems compared to subaerial ones (Figure~\ref{fig:range}\,d--f and S \ref{fig:gammas}).
-  Yet, the greatest difference in the two systems at the smallest variations of water discharge still results in greater variability in the subglacial system compared to the subaerial one (Figure \ref{fig:range}).
-\ian{extra sentence is needed here}
+  Importantly though, no runs experienced relative variability in width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, less than $1$, showing that the subglacial has consistently high variability in width integrated shear stress compared to subaerial systems across the range of tested situations.
 
@@
   In supply-limited systems, the changing sediment transport capacity (Figure~\ref{fig:model_outs}) would have a minimal effect on sediment discharge due to the absence of sediment available for transport \citep{delaney2019}.
-  Yet, reaching the threshold of motion for sediment frequently, across a range of discharges,  may result in additional sediment transport and the tendency for glaciers to transport sediment in a supply-limited regime whereby sediment discharge scales with glacier sliding  \citep[e.g.][]{herman2015,koppes2015}.
+  Yet, in the subglacial system reaching the threshold of motion for sediment frequently, across a range of discharges,  may result in additional sediment transport.
+  The variable sediment transport capacity may result in a tendency for glaciers to evacuate sediment and transition to a supply-limited regime whereby sediment discharge scales with glacier sliding  \citep[e.g.][]{herman2015,koppes2015}.
   Sediment exhaustion by this process may also explain the dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, as opposed to glacier melt \citep{overeem2017}.
@@
   The effect of pressurized channel flow could be most pronounced during precipitation or flood events when subglacial conduits have not adapted to the water discharge flowing through them, resulting in a large increase in sediment transport capacity and discharge \citep[e.g.][]{cowan1988,delaney2019}.
-  Increased sediment discharge may result from changes to sediment access below the glacier, yet rapid increases in water velocity from water flowing through a small channel will cause a large increase in sediment transport capacity (Section~\ref{sect:sub_mode}).
+  Exceptional sediment discharge during floods may result from changes to sediment access below the glacier, yet rapid increases in water velocity from water flowing through a small channel will also cause a large increase in sediment transport capacity, not possible in a subaerial system (Section~\ref{sect:sub_mode}).
 
@@
   Large variations in sediment transport capacity below glaciers may result in sediment mobilization and deposition processes in close spatial and temporal proximity to each other below the glacier \citep{gimbert2016,perolo2018}.
-  The rapid increase and decreases in sediment transport capacity over a diurnal cycle in a subglacial system may compound the variability already present in subaerial systems \citep{williams1989,jerolmack2010}.
+  The rapid increase and decreases in sediment transport capacity, for example over a diurnal cycle, in a subglacial system may compound the variability in sediment transport already present in subaerial systems \citep{williams1989,jerolmack2010}.
   As a result, variations in sediment transport capacity could be responsible for the fluctuations in sediment discharge (``flushings'') from glaciers that are not attributed to variations in water discharge \citep[e.g.][]{richards2003,swift2021}.

### 2023-03-23 Delaney2023.tex
GRL template added

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
% 
% This file includes commands and instructions
% given in the order necessary to produce a final output that will
% satisfy AGU requirements, including customized APA reference formatting.
% 
% You may copy this file and give it your
% article name, and enter your text.
% 
% 
% Step 1: Set the \documentclass
% 
%

%% To submit your paper:
\documentclass[draft]{agujournal2019}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage{hyphenat}
\usepackage{float}
\usepackage{multirow}
\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\newcolumntype{C}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}

\linenumbers
\sloppy
\hbadness=99999 
%%%%%%% 
% As of 2018 we recommend use of the TrackChanges package to mark revisions.
% The trackchanges package adds five new LaTeX commands:
% 
% \note[editor]{The note}
% \annote[editor]{Text to annotate}{The note}
% \add[editor]{Text to add}
% \remove[editor]{Text to remove}
% \change[editor]{Text to remove}{Text to add}
% 
% complete documentation is here: http://trackchanges.sourceforge.net/
%%%%%%%

\draftfalse

%% Enter journal name below.
%% Choose from this list of Journals:
% 
% JGR: Atmospheres
% JGR: Biogeosciences
% JGR: Earth Surface
% JGR: Oceans
% JGR: Planets
% JGR: Solid Earth
% JGR: Space Physics
% Global Biogeochemical Cycles
% Geophysical Research Letters
% Paleoceanography and Paleoclimatology
% Radio Science
% Reviews of Geophysics
% Tectonics
% Space Weather
% Water Resources Research
% Geochemistry, Geophysics, Geosystems
% Journal of Advances in Modeling Earth Systems (JAMES)
% Earth's Future
% Earth and Space Science
% Geohealth
% 
% ie, \journalname{Water Resources Research}

\journalname{Geophysical Research Letters}

\newcommand{\aref}[1]{\textbf{Reference #1}}
\newcommand{\TODO}[1]{\textbf{TODO #1}}
\newcommand{\ian}[1]{{\textbf{\color{blue}Ian says:} \color{blue} #1} }
\newcommand{\leif}[1]{{\textbf{\color{red}Leif says:} \color{red} #1} }
\newcommand{\db}{\textit{DEBRIS}\,}
\newcommand{\cl}{\textit{CLEAN}\,}
\newcommand{\m}{$\,\mathrm{m}$\,}
\newcommand{\cm}{$\,\mathrm{cm}$\,}
\newcommand{\mma}{$\,\mathrm{mm  \, a^{-1}}$\,}
\newcommand{\mmma}{$\,\mathrm{m^3\, a^{-1}}$\,}
\newcommand{\mmms}{$\,\mathrm{m^3\, s^{-1}}$\,}
\newcommand{\unit}[1]{$\mathrm{#1}$}

\newenvironment{laysummary}{
  \chapter{\centering \textbf{Plain Language Summary}}
}

\begin{document}
%% ------------------------------------------------------------------------ %%
% Title
% 
% (A title should be specific, informative, and brief. Use
% abbreviations only if they are defined in the abstract. Titles that
% start with general keywords then specific terms are optimized in
% searches)
% 
%% ------------------------------------------------------------------------ %%

% Example: \title{This is a test title}

\title{Comparing variations in sediment transport capacity between subglacial and subaerial systems }

%% ------------------------------------------------------------------------ %%
% 
% AUTHORS AND AFFILIATIONS
%

% Authors are individuals who have significantly contributed to the
% research and preparation of the article. Group authors are allowed, if
% each author in the group is separately identified in an appendix.)

% List authors by first name or initial followed by last name and
% separated by commas. Use \affil{} to number affiliations, and
% \thanks{} for author notes.
% Additional author notes should be indicated with \thanks{} (for
% example, for current addresses).

% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antarctica}, B. C. Author\affil{2,3}, and D. E.
% Author\affil{3,4}\thanks{Also funded by Monsanto.}}

\authors{Ian Delaney\affil{1}}

\affiliation{1}{Institut des dynamiques de la surface terrestre (IDYST), Universit\'{e} de Lausanne, B\^{a}timent G\'{e}opolis, CH-1015 Lausanne }

% \affiliation{3}{Third Affiliation}
% \affiliation{4}{Fourth Affiliation}

% \affiliation{=number=}{=Affiliation Address=}
% (repeat as many times as is necessary)

%% Corresponding Author:
% Corresponding author mailing address and e-mail address:

% (include name and email addresses of the corresponding author.  More
% than one corresponding author is allowed in this LaTeX file and for
% publication; but only one corresponding author is allowed in our
% editorial system.)

% Example: \correspondingauthor{First and Last Name}{email@address.edu}

\correspondingauthor{Ian Delaney}{ianarburua.delaney@unil.ch}

%% Keypoints, final entry on title page.

% List up to three key points (at least one is required)
% Key Points summarize the main points and conclusions of the article
% Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences

% Example:
% \begin{keypoints}
% \item	List up to three key points (at least one is required)
% \item	Key Points summarize the main points and conclusions of the article
% \item	Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
% \end{keypoints}

\begin{keypoints}
\item Water discharge variations in subglacial channels are accommodated by water velocity, not channel size, given their slow evolution.

\item  Greater variability in water velocity in subglacial systems causes sediment transport capacity to vary more compared to subaerial systems.

\item The systems' divergence in transport capacity response may impact sediment dynamics at glacier margins and interpretation of sediment records.
\end{keypoints}

%% ------------------------------------------------------------------------ %%
% 
% ABSTRACT and PLAIN LANGUAGE SUMMARY
% 
% A good Abstract will begin with a short description of the problem
% being addressed, briefly describe the new data or analyses, then
% briefly states the main conclusion(s) and how they are supported and\cite <e.g.,>[and others]{lewin76}
% uncertainties.

% The Plain Language Summary should be written for a broad audience,
% including journalists and the science-interested public, that will not have 
% a background in your field.
% 
% A Plain Language Summary is required in GRL, JGR: Planets, JGR: Biogeosciences,
% JGR: Oceans, G-Cubed, Reviews of Geophysics, and JAMES.
% see http://sharingscience.agu.org/creating-plain-language-summary/)
% 
%% ------------------------------------------------------------------------ %%

%% \begin{abstract} starts the second page

\begin{abstract} % 150 words
Sediment transport capacity depends mainly on the shear stress or velocity of water flowing through a channel and the channel width over which to mobilize sediment.
In subaerial channels, changing water discharge can be accommodated by rapid changes in the width of the channel over which water flows, the water depth, and the water velocity.
In subglacial channels, however, water flows through a conduit whose size evolves  slowly, because the water is pressurized by the glacier ice above.
As a result, variations in water discharge may modify the water velocity, but not the channel's width.
Here, parameterizations of subglacial and subaerial channels indicate that sediment transport capacity varies more in subglacial systems compared with subaerial systems, across a wide range of channel geometries and flow conditions.
Large variations in subglacial water sediment transport capacity may impact sediment transport processes and the interpretation of records of sediment transport from glacierized catchments.

\end{abstract}

\section*{Plain language summary} % 200 words
In both subglacial and subaerial channels, sediment transport capacity depends mainly on 1) the channel width over which the sediment is mobilized  and 2) the shear stress applied by flowing water on sediment at the channel bottom; the shear stress depends largely on the water velocity.
In subaerial channels, changing water discharge will  change the water depth,  the channel width (which dictates the distance  over which sediment transport  occurs), and the  water velocity (which controls the sediment transport capacity).
 In subglacial environments, however, on times scales of hours to weeks, changing water discharge will only alter  water velocity, not channel area, because the subglacial conduit maintains a largely fixed geometry.
The fixed conduit geometry also means that the channel width over which the sediment is  mobilized also does not respond to changing water discharge.
Parameterizations of these processes show that sediment transport capacity varies far more in subglacial systems compared to subaerial systems, even though changing channel width can accommodate some sediment transport variability in subaerial channels.
The increased variability in sediment transport capacity in subglacial systems, compared to subaerial ones, has several important implications for sediment transport in glacierized catchments and the interpretation of sediment discharge records.

%% ------------------------------------------------------------------------ %%
% 
% TEXT
% 
%% ------------------------------------------------------------------------ %%

%%% Suggested section heads:
\section{Introduction}
\label{sect:intro}

  Glaciers expel large quantities of sediment from their termini \cite{hallet1996}.
  Changes in glacier dynamics, geomorphology, and hydrology  have prompted numerous  recent studies of  sediment transport processes in cold regions \cite<e.g.>[]{zhang2022}.
  Increases in sediment transport  have been observed in Greenland \cite{bendixen2017}, the European Alps \cite{costa2017}, the Himalayas \cite{li2021}, and the Andes \cite{vergara2022}.
  In some of these regions, increased water discharge and glacier melt have been attributed to the greater sediment transport \cite{bendixen2017,costa2017,li2021}.
  In other cases, enhanced access of meltwater to subglacial sediment at high elevations was attributed as the cause of increased sediment transport \cite{delaney2020,vergara2022}.
  Observed changes to sediment transport in glacierized catchments create an imperative to examine the processes controlling sediment discharge here. 
  Yet, the relationship between water discharge and sediment transport in subglacial environments remains minimally discussed and not contrasted with the subaerial river systems that these glaciers feed.

  Over long periods, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from below glaciers \cite<c.f.>{hallet1979,iverson2012,ugelvig2018}. 
  On shorter time periods, pressurized water transports sediment from the below glacier \cite{walder1994,creyts2013,beaud2018}, should enough sediment be present below the glacier (i.e. in a transport-limited regime).

  In a transport-limited regime, sediment discharge responds to the sediment transport capacity or the amount of sediment that could be carried by the water. Sediment transport capacity depends on the shear stress between water and sediment it flows  over \cite{shields1936,meyer1948,engelund1967} and the width of the channel bottom $w_c$ over which to mobilize sediment. The shear stress  across the channel width $\tau_t$ responds to the velocity of water $v$ flowing through at channel so that 
  \begin{linenomath*}
    \begin{equation}
      \label{eq:tau_t}
      \tau_t \propto w_c\, v^2,
    \end{equation}
  \end{linenomath*}
  % 
  where $w_c$ is the width of the channel.
  Following the conservation of mass, the velocity of the water flowing through a rectangular channel is given as 
  \begin{linenomath*}
    \begin{equation}
      \label{eq:v}
      v = \frac{Q_w}{S},
    \end{equation}
  \end{linenomath*}
  % 
  where $Q_w$ is water discharge,  and $S$ is the channel's cross-sectional area.

  In subaerial channels, with open channel flow, the cross-sectional area $S$ of the channel evolves with changing water discharge $Q_w$  by changing the width of the channel along with the depth \cite{leopold1953}. During a flood with greater water discharge $Q_w$, for instance, the cross-sectional area $S$ increases as the channel widens and deepens, along with water velocity $v$. As a result, the shear stress across the channel $\tau_t$ will increase in response to both the channel width $w_c$  and water velocity $v$.

  Yet, the response of water velocity to changing water discharge in subglacial channels is different than subaerial ones.
  In subglacial systems, the water is pressurized by the ice above \cite{shreve1972}.
  The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by frictional heating of water flowing through the channel \cite{rothlisberger1972}.
  As a result, the subglacial channel size evolves over days to months, whereas water discharge varies over hours \cite<e.g.>[]{iken1986,andrews2014,nanni2020}.
  In turn, changes in water discharge $Q_w$ are mainly accommodated with water velocity $v$, because the size of the channel $S$ is largely fixed (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
  It follows that greater water discharge $Q_w$ will cause the shear stress across the channel $\tau_t$ to respond to the water velocity $v$, as opposed to the subaerial system where width evolves as well.

  As a result, sediment mobilization in subaerial and subglacial systems responds differently to changing water discharge.
  These differences have been implicitly included in a wide range of available models that  quantify sediment transport from catchments in both subglacial and subaerial systems  \cite<e.g.>[]{walder1994,tucker1997,creyts2013,wickert2019,hewitt2019}.
  The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \cite<e.g.>[]{lane2016}.
  Additionally, the variable response of sediment transport capacity to water discharge in the two systems may impact the interpretation of records of sediment transport in glacial compared to subaerial systems \cite<e.g.>[]{muller1968,richards2003,ganti2016}.
  Yet, the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial systems has been minimally discussed.

  This manuscript leverages parameterizations of subglacial and subaerial hydraulics to explore differences in the relationships between  water discharge, channel geometry, and water velocity. 
  With these parameterizations, the manuscript examines the impact of water discharge variability on sediment transport capacity in idealized subglacial and subaerial systems.
  Lastly, the manuscript discusses the implications for sediment transport processes in subglacial systems, compared to subaerial ones.

  \begin{center}
    \begin{figure}[H]
      \includegraphics[width=0.65\linewidth]{Cartoon.pdf}
      \caption{Cartoon of the situation with various parameters. Water velocity magnitude in the subglacial, $v_g$, and subaerial, $v_f$, systems are shown by arrow length. $S_g$ and $S_f$ represent the cross-sectional area in subglacial and subaerial channels. The subglacial channel widths $w_{cg}$ remain steady, while the subaerial channel widths evolve $w_{cf}$, denoted by the thick black line. Shear stress $\tau$ is responsible for the mobilization of sediment and increases with the number of makers at the channel bed. 
        Note that in the subaerial channel parameterization, rectangular channel shapes are implemented (Section~\ref{sect:fluv}).} 
      \label{fig:cartoon}
    \end{figure}
  \end{center}

\section{Methods}
\label{sect:meth}
    Parameterizations below represent relationships amongst water discharge, subglacial water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
  Because sediment transport largely depends on shear stress applied by water flow \cite{shields1936}, both parameterizations calculate this value and integrate  it across the channel bed (Figure \ref{fig:cartoon}).
  The evaluation of shear stress omits a grain size parameter and, thus, leaves open the selection of a sediment transport relationship  \cite<e.g.>[]{shields1936,meyer1948} to  calculate sediment transport capacity.
  As a result, relative relationships in shear stress, as opposed to sediment transport capacity, are examined.

  \begin{table}[H]
    \centering
    \caption{Variables, parameters, and constants }
    \begin{tabular}{ l  c  c c }
      Name &Symbol&  Value&Units \\ \hline
      \textbf{Variables}  & & & \\
      Channel hydraulic diameter (glacier, subaerial) &  $D_{hg},\,D_{hf}$&  & $\mathrm{m}$     \\
      Width of channel floor (glacial, subaerial) & $w_{cg},w_{cf}$&  & $\mathrm{m}$     \\
      Channel cross-sectional area (glacier, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
      Water discharge (instantaneous) & $Q_w$& & $\mathrm{m^{3}\,s^{-1}}$ \\
      Representative water discharge & $Q_{w}^*$& & $\mathrm{m^{3}\,s^{-1}}$ \\
      Representative hydraulic gradient  &$\Psi^*$ & & $\mathrm{Pa\, m^{-1}}$\\
      Water velocity (glacier, subaerial)  & $v_g,\,v_{f}$& & $\mathrm{m\,s^{-1}}$ \\
      Shear stress (glacier, subaerial) & $\tau_g,\,\tau_f$&& $\mathrm{Pa \, m^{-1}}$ \\
      Width-integrated shear-stress (glacier, subaerial) & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$ \\
      Reynolds number &$R_e$& & $\mathrm{(-)}$\\
      Variable &$\Gamma$&&$\mathrm{(-)}$\\
      Variability of a variable ($\Gamma$) &$\chi$& &$\mathrm{(-)}$\\
           &&&\\

      \textbf{Parameters and Constants}  & & &\\
      Gravitational constant&$g$& $-9.81$&$\mathrm{m\,s^{-2}}$\\
      Density of water & $\rho_w$& $1000$ & $\mathrm{kg\,m^{-3}}$ \\
      Density of ice & $\rho_i$& $900$ & $\mathrm{kg\,m^{-3}}$ \\
      Kinematic viscosity of water &$\nu$& $10^{-4}$& $\mathrm{m^2\,s^{-1}}$\\
      Hooke angle of channel & $\beta$ & $\frac{\pi}{2}$ & \unit{rad}\\
      Flotation fraction & $f_f$&$0.7$& $\mathrm{(-)}$\\
      Friction factor (glacier) & $f_r$ & $4$ & $\mathrm{(-)}$ \\
      Friction factor (subaerial) & $f_p$ & $3$ & $\mathrm{(-)}$\\
      Gradient of glacier surface & $\frac{\partial z_{sg}}{\partial x}$ &$0.25$& $\mathrm{(-)}$\\
      Gradient of channel bed (subaerial) &$\frac{\partial z_c}{\partial x}$ &$0.05$& $\mathrm{(-)}$\\
      Subaerial channel factor & $k$ &$3$ & $\mathrm{s\,m^{-2}}$\\
      Channel geometry exponent &$e$& $\frac{1}{2}$&$\mathrm{(-)}$ \\
      \hline
    \end{tabular}
    \label{table:vpm}
  \end{table}

  \subsection{Subglacial channel  parameterization}
  \label{sect:sub_mode}
  To evaluate the shear stress of water flowing across sediments below a glacier, the subglacial channel parameterization evaluates the channel geometry and the velocity of the flowing water.
  To accomplish this, the hydraulics model presented in \citeA{delaney2019} is used.

  Here, it is assumed that the water is transported through subglacial channels \cite<Figure~\ref{fig:cartoon}; >{rothlisberger1972}, and that the channel  geometry will respond to a representative water discharge over a certain time period $Q_{w}^*$, using the Darcy-Weisbach formulation for water-flow through a pipe  \cite<e.g.>[]{rothlisberger1972,clarke2003,werder2013}.
  Thus, the hydraulic diameter of the subglacial channel, $D_h$, responds to a representative water discharge, $Q_{w}^*$, and a representative gradient of the hydraulic potential, $\Psi^*$ such that:
  \begin{linenomath*}
    \begin{equation}
      \label{eq:DW}
      D_h = \big(s\, f_r\,\rho_w\, \frac{Q_w^{*\,2}}{\Psi^*}\big)^{\frac{1}{5}}~~.
    \end{equation}
  \end{linenomath*}
  % 
  The representative hydraulic gradient is based upon the pressurized flow in Equation~\ref{eq:DW},
  \begin{equation}
    \label{eq:psi}
    \Psi^*= f_f \,  \rho_i \, g (\frac{\partial  z_{sg}}{\partial x} - \frac{\partial z_c}{\partial x}) +  \rho_i \, g \, \frac{\partial z_c}{\partial x},
  \end{equation}
  % 
  \noindent
  where, $f_f$ is the flotation fraction, $\rho_i$ is density of ice $\frac{\partial z_{sg}}{\partial x}$ is the surface slope of the glacier and $\frac{\partial z_c}{\partial x}$ is the glacier bed slope.

  In Equation~\ref{eq:DW}, $f_r$ is the Darcy-Weisbach friction factor and $\rho_w$ is the density of water (Table \ref{table:vpm}). $s$ is a factor accounting for channel geometry \cite{hooke1990}, calculated as:
  \begin{equation}
    \label{eq:Hf}
    s = \frac{2\,(\beta -\sin \beta)^2}{(\frac{\beta}{2}\,+\,\sin \frac{\beta}{2})^4},
  \end{equation}
  where $\beta$ is the central angle of the circular segment that comprises the channel (the so-called Hooke angle). Note that $\beta =\pi$ corresponds to a semi-circle and
  smaller values of $\beta$ result in shallow, wide channels. 
  The width of the flat channel floor $w_c$, given angle $beta$ is 
  \begin{equation}
    \label{eq:dh2wc}
    w_{cg} = 2  \sin \frac{\beta}{2} \sqrt{\frac{2\, S}{\beta -\sin \beta}},
  \end{equation}
  % 
  where $S$ is the cross-sectional area of the channel given by
  \begin{equation}
    \label{eq:dh2S}
    S_g =  \frac{D_h^2}{2}~ \frac{\Big(\frac{\beta}{2} \,+ \, \sin \frac{\beta}{2}\Big)^2  }{\beta\,-\,\sin \beta}.
  \end{equation}

  The shear stress, $\tau$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
  \begin{equation}
    \label{eq:tau}
    \tau_g=\frac{1}{8}\,f_r\,\rho_w\,v_g^2,
  \end{equation}
  % 
  where  the water velocity $v_g$ is $v_g = \frac{Q_w}{S_g}$.
  Here, $Q_w$ represents the instantaneous water discharge, rather than $Q_w^*$, which is used to evaluate the channel geometry.
  With this formulation, the processes in an R-channel, through which water flows subglacially, can be represented algebraically \cite{rothlisberger1972,delaney2019}.

    Width integrated shear stress is represented as $\tau_{tg}=w_{cg}\,\tau_g $.

  \subsection{Subaerial channel  parameterization}
  \label{sect:fluv}

  To parameterize the shear stress of water flowing across sediments in the subaerial channel,  the hydraulics parameterization presented in \citeA{tucker1997} is implemented.
  Here, it is assumed that there is a conservation of mass and a sufficiently wide channel so that the hydraulic radius is consistent with flow depth, uniform flow, and the Darcy-Weisbach relationship.
  The shear stress $\tau_f$ at the river bed is represented as
  \begin{linenomath*}
    \begin{equation}
      \label{eq:DW_tau}
      \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q_w}{w_{cf}} \Big)^{\frac{2}{3}} \,\frac{\partial z_c}{\partial x}^{\frac{2}{3}},
    \end{equation}
  \end{linenomath*}
  where $\frac{\partial z_c}{\partial x}$ is the channel slope, and $f_p$ is the friction factor for subaerial systems.
  Channel width $w_{cf}$ is 
  \begin{equation}
    \label{eq:wcf}
    w_{cf} = k \, Q_w^e,
  \end{equation}
  % 
  where $k$ is a constant and $e$ is an exponent commonly equal to $\frac{1}{2}$ \cite{leopold1953}.
  Cross-sectional area, $ S_f$, is 
  \begin{equation}
    \label{eq:Sf}
    S_f = \frac{Q_w}{v_f},
  \end{equation}
  % 
  where $v_f$ is water velocity, given by rearranging Equation \ref{eq:tau} as
  \begin{equation}
    \label{eq:vf}
    v_f = \sqrt{\frac{8\,\tau_f}{f_p\,\rho_w}}.
  \end{equation}
  % 
  As in Section~\ref{sect:sub_mode}, the width integrated shear stress is $\tau_{tf}=w_{cf}\,\tau_f$.
  Hydraulic diameter of the channel is given by $D_{hf} = \frac{4\,w_{cf}\,d}{2\,d+w_{cf}}$, with flow depth $d$ determined from knowledge of $S_f$ and $w_{cf}$ in a rectangular channel.

  \subsection{Implementation}

  The formulation of the glacial system (Section~\ref{sect:sub_mode}) requires inputs of the hydraulic gradient (Equation \ref{eq:psi}; $\frac{\partial z_{sg}}{\partial x}$, $\frac{\partial z_{c}}{\partial x}$), water discharge, $Q_w$ representative water discharge, $Q_w^*$,  Hooke angle (Equation~\ref{eq:Hf}; $\beta$), and friction factor (Equation~\ref{eq:DW}; $f_r$).

  In the subaerial system,  only the bed slope, $\frac{\partial z_c}{\partial x}$, is needed in Equation~\ref{eq:DW_tau}. The friction factor, $f_f$, and channel shape factor, $k$, are assigned such that they produce reasonable values for water velocity and channel width  (Section~\ref{sect:fluv}).

  The Reynolds number $R_e$, which gives an evaluation of the degree of turbulence of the water, is given as 
  \begin{equation}
    \label{eq:re}
    R_e\,=\, v \,\frac{D_h}{\nu},
  \end{equation}
  % 
  \noindent where $\nu$ is the kinematic viscosity of water, and  $v$ is velocity of water, $v_g$ or $v_f$.

  The variance $ \chi$ of a variable $\Gamma$ is given by 
  \begin{equation}
    \label{eq:var}
    \chi(\Gamma) \,=\, 1 - \frac{\mathrm{min}(\Gamma)}{\mathrm{max}(\Gamma)}\,\,.
  \end{equation}
  % 
  \noindent Below we examine variables, $\Gamma$, as water discharge ($Q_w$) water velocity ($v_g$, $v_f$), shear stress ($\tau_g$, $\tau_f$), and integrated shear stress ($\tau_{tg}$, $\tau_{tf}$) with the purpose of evaluating total shear stress across the glacier bed.

\section{Results}
  Initially, both parameterizations are forced with a water discharge ($Q_w$) varying between $8$ and $16$ \,\unit{m}$^{3}$\,\unit{s}$^{-1}$, to represent a range of sediment transport conditions, for instance over diurnal fluctuation in water discharge. The subaerial channel width factor $k$ is $3$\,\unit{m}$^{2}$\,\unit{s}$^{-1}$, and the Hooke angle $\beta$ is $\frac{\pi}{2}$.
  Values of Hooke angle $\beta$ and subglacial friction factor $f_r$ are chosen so that they produce reasonable values of water velocity ($0.6- 1.2$\,\unit{m}\,\unit{s}$^{-1}$) in accordance with measurements made through dye-tracing experiments to examine subglacial hydraulic conditions \cite<Section~\ref{sect:sub_mode}, Figure~\ref{fig:model_outs}; e.g.>{werder2010}.
  The subglacial representative water discharge ($Q_w^*$; Equation~\ref{eq:DW}) of $12$\,\unit{m}$^{3}$\,\unit{s}$^{-1}$ is applied to establish the hydraulic diameter and shape of the channel.

  \begin{center}
    \begin{figure}[H]
      \includegraphics[width=0.7\linewidth]{model_outputs.pdf}
      \caption{Response of channel width (a), hydraulic diameter (b),  Reynolds number  (c),  water velocity (d), shear-stress (e), and width-integrated shear-stress (f)  to different water discharge values. The glacial system is in blue and the subaerial system is in green.  Horizontal bars in d, e, and f represent the range of values in the  output. Gray space represents water discharge below the mean value. }
      \label{fig:model_outs}
    \end{figure}
  \end{center}

  Across the hydrological forcing, variable water discharge results in a substantial increase in the range of water velocities in the glacial system compared to the subaerial system (Figure~\ref{fig:model_outs}).
  Water velocity in the glacier system vary by  $0.5$ (Equation~\ref{eq:var}) over the discharge values, in the subaerial system the water velocity only varies by $0.11$ (Figure~\ref{fig:model_outs}\,e).
  Shear stress at the glacier bed varies by $0.75$ across the range of water discharge values, whereas shear stress across the subaerial system varies by roughly $0.21$ (Figure~\ref{fig:model_outs}\,f). The increased variability in shear stress relative to water velocity between the two systems occurs as the water velocity is raised to the power of $2$ in Equation~\ref{eq:tau}.

  In the subaerial systems, changing channel geometry accommodates evolving water discharge  (Equation~\ref{eq:wcf}), whereas in the subglacial case, increased water flux can only result in changing water velocity (Equations~\ref{eq:v}~ and ~\ref{eq:DW}).
  As a result, in the subaerial system, variability in the shear stress $\tau_f$ and width-integrated shear stress $\tau_{tf}$ increases from $0.21$ to $0.44$. Variability in width-integrated shear stress  in the glacial system $\tau_{tg}$ is $0.83$ (Figure~\ref{fig:range}).
  The reduced variability in width-integrated shear stress results from the additional width across the subaerial channel that accommodates some variations in water discharge.
  Indeed, the relative  difference in width-integrated shear stress is less between the subaerial and glacier cases, compared to the water velocity and shear stress outputs (Figure~\ref{fig:model_outs}\,e--g).

  Next, the parameterizations are run $25,000$ times with randomly selected channel geometries ($k=(2,6)$, $\beta=(\frac{\pi}{10},\pi)$), friction factors ($f_i=(0.001,20)$, $f_p=(0.001,20)$) and water discharges (maximum and minimum values selected between $5$ to $500$ \,\mmms). $Q_w^*$ in Equation ~\ref{eq:DW} is determined from the mean in the water discharge range. Parameter combinations are only selected if their maximum and minimum velocities for both the subglacial and subaerial systems are between $0.2$ and $3$\,\unit{m}\,\unit{s}$^{-1}$, resulting in roughly $10,600$ adequate parameter combinations.

  \begin{center}
    \begin{figure}[H]
      \includegraphics[width=0.7\linewidth]{multi_run.pdf}
      \caption{Ratio of range of water velocity, shear-stress, and width integrated shear-stress for glacial and subaerial systems in response to mean water discharge (a, b, c), and  variability in water discharge (d, e, f).  
        Range  $\chi$ for each variable is given in Equation~\ref{eq:var}. Values greater than $1$ shown in white space are parameter combinations that vary more in a subglacial  system.
        Histograms (g-i) show the frequency of relative variations across variables in the parameterization.}
      \label{fig:range}
    \end{figure}
  \end{center}

  Across the accepted model runs, variability in water velocity and shear stress remains substantially higher across the subglacial model compared to subaerial ones (Figure~\ref{fig:range}).
  Water velocity, shear stress, and width-integrated shear stress all have a variability ratio ($\frac{\chi(\Gamma_{g})}{\chi(\Gamma_f)}$) greater than $1$ in the subglacial  system compared to subaerial one (Figure~S \ref{fig:gammas}).
  Runs with the smallest variability in discharge ($\chi(Q_w)$) have the greatest variability in water velocity, shear stress, and width-integrated shear stress in subglacial systems compared to subaerial ones (Figure~\ref{fig:range}\,d--f and S \ref{fig:gammas}).

  The minimum ratio of  water velocity, $\frac{\chi(v_{g})}{\chi(v_{f})}$, and shear-stress, $\frac{\chi(\tau_{g})}{\chi(\tau_{f})}$, is greater than $1.5$.
  These results show that increased variability occurs across a wide range of subaerial and subglacial channel configurations (Figure~\ref{fig:range} g, h).
The width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, spans between $1.02$ and $2.4$, a smaller amount of relative variability compared to  water velocity $\frac{\chi(v_{g})}{\chi(v_{f})}$ or shear stress $\frac{\chi(\tau_{g})}{\chi(\tau_{f})}$ (Figure~\ref{fig:range}).
Consistent with the individual case (Figure ~\ref{fig:model_outs}), the subaerial systems' reduce variability in width-integrated shear stress occurs  as the evolving channel width accommodates some of the reduced shear stress variability in subaerial systems (Figure~\ref{fig:range} i).
  Importantly though, no runs experienced relative variability in width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, less than $1$, showing that the subglacial has consistently high variability in width integrated shear stress compared to subaerial systems across the range of tested situations.

  \section{Discussion}

  \subsection{Geomorphic Implications}
  \label{sect:GI}
  The relationship between water discharge variability and relative variability of the water velocity, shear stress, and width-integrated shear stress results from the fixed geometry in the glacier system and variable geometry in the subaerial system as they respond to changes in water discharge (Equations~\ref{eq:tau_t} and \ref{eq:v}).
  Increased variability in sediment transport capacity with respect to discharge has several implications for sediment transport in subglacial systems compared to subaerial ones.
  Shear stress is scaled to the $\frac{3}{2}$ power and relies on shear stress exceeding a threshold in sediment transport relationships  such as in \citeA{meyer1948}.
  In the total sediment transport relationship \citeA{engelund1967}, the shear stress is scaled to $\frac{5}{2}$ power.
  In both cases, the exponent greater than $1$ of shear stress in the sediment transport relationship magnifies the variability in sediment transport beyond the highly variable shear stress and width-integrated shear stress described above in subglacial environments (Figure~\ref{fig:model_outs}\, b and d; Figure~S~\ref{fig:gammas}).

  Both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress of water across the channel bed is homogeneous \cite<Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}; >{yager2018}. 
  The subaerial parameterization assumes that sediment is transported across all channel widths, as opposed to large flows such as floods when much sediment transport occurs \cite{wolman1960}.
  As a result, the variability in the subaerial system across the range of discharge values could be underestimated in the parameterization (Figure~\ref{fig:range}).

  \subsubsection{Sediment transport below glaciers and at their margins}

  The response of sediment discharge caused by the greater variability in sediment transport in subglacial channels will be greatest in completely transport-limited regimes \cite<e.g.>[]{kasmalkar2019}.
  In supply-limited systems, the changing sediment transport capacity (Figure~\ref{fig:model_outs}) would have a minimal effect on sediment discharge due to the absence of sediment available for transport \cite{delaney2019}.
  Yet, in the subglacial system reaching the threshold of motion for sediment frequently, across a range of discharges,  may result in additional sediment transport.
  The variable sediment transport capacity may result in a tendency for glaciers to evacuate sediment and transition to a supply-limited regime whereby sediment discharge scales with glacier sliding  \cite<e.g.>[]{herman2015,koppes2015}.
  Sediment exhaustion by this process may also explain the dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, as opposed to glacier melt \cite{overeem2017}.

  The effect of pressurized channel flow could be most pronounced during precipitation or flood events when subglacial conduits have not adapted to the water discharge flowing through them, resulting in a large increase in sediment transport capacity and discharge \cite<e.g.>[]{cowan1988,delaney2019}.
  Exceptional sediment discharge during floods may result from changes to sediment access below the glacier, yet rapid increases in water velocity from water flowing through a small channel will also cause a large increase in sediment transport capacity, not possible in a subaerial system (Section~\ref{sect:sub_mode}).

  Increased variability in sediment transport may also impact sediment export from the subglacial environment into proglacial areas \cite<e.g.>[]{delaney2017,perolo2018}.
  Here, velocities at the highest water discharges could generally decrease (Figure~\ref{fig:model_outs}, blue and green horizontal bars), in the transition of pressurized subglacial flow to subaerial open-channel flow as the water leaves the glacier. The decrease in water velocity potentially results in sediment deposition once sediment enters the proglacial area.
  Periodic sediment deposition in sediment deposition may even result in the transition between pressurized and subaerial flow below the glacier terminus \cite{perolo2018}, especially when cavity closure rates of the ice are slow \cite{egli2021b}.

  In large catchments and further downstream of proglacial margins, water discharge is generally less variable than at the tops of catchments, where glaciers lie in alpine settings \cite<c.f.>{swift2005,riihimaki2005,costa2017,vanas2017}.
  Here, external hydrological factors further downstream of glaciers could result in smaller variations in sediment transport capacity in many subaerial systems compared to subglacial ones.

  The parameterization here shows that the smallest variability in water discharge has the largest difference in relative variability in sediment transport metrics (Figure~\ref{fig:range}\, d,\,e,\,f).
  At locations such as the Greenland Ice Sheet, smaller diurnal variations in water discharge compared to many alpine catchments \cite<c.f.>{swift2005,riihimaki2005,vanas2017,hasholt2018}.
  Therefore, the variability in sediment transport capacity in the transition from the ice sheet to the proglacial river could be more dramatic in Greenland than in alpine systems as water leaves the glacier and moves downstream.
  Thus the disparity in sediment transport capacity between the subglacial and subaerial systems could be most pronounced in larger catchments, potentially resulting in temporally mismatched periods between when glaciers introduce sediment to these river systems  and when the river mobilizes it (Figures~\ref{fig:model_outs}~ and~\ref{fig:range}).

  \subsubsection{Records of sediment discharge  from glaciers}

  Large variations in sediment transport capacity below glaciers may result in sediment mobilization and deposition processes in close spatial and temporal proximity to each other below the glacier \cite{gimbert2016,perolo2018}.
  The rapid increase and decreases in sediment transport capacity, for example over a diurnal cycle, in a subglacial system may compound the variability in sediment transport already present in subaerial systems \cite{williams1989,jerolmack2010}.
  As a result, variations in sediment transport capacity could be responsible for the fluctuations in sediment discharge (``flushings'') from glaciers that are not attributed to variations in water discharge \cite<e.g.>[]{richards2003,swift2021}.
  Aside from the sporadic nature of these events, the observed flushings could result from increases in water velocity resulting from reduced channel size, not water discharge, which increase sediment transport capacity.

  Additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to subaerial systems \cite{jansson2005,ganti2016}. 
  For instance, sediment transport capacity from a glacier reacts to the ice thickness, controlling the closure of the channel, and the surface slope of the glacier, controlling water velocity \cite<Section~\ref{sect:sub_mode}; >{rothlisberger1972,shreve1972,delaney2022,stevens2022}, in addition to water discharge.
  Conversely, sediment transport in most subaerial systems increases with water discharge and bed slope, with the bed slope probably being more stable over yearly to century-scale time periods \cite<Section~\ref{sect:fluv}; e.g.>{muller1968,whipple1999,wong2006,wickert2019}.

  Records of sediment discharge have been used to establish the relationship, or lack thereof, between sediment discharge and climate in glacial systems \cite<e.g.>[]{koppes2009a,willenbring2016,mariotti2021}.
  The variability in subglacial sediment transport capacity presented here applies most generally to short-time scales responsible for the size of subglacial channels.
  Yet, identifying climatic signals in sediment transport from transport-limited glaciers may require higher thresholds of climatic perturbation compared to subaerial systems \cite{tofelde2021}.
  This higher threshold of climatic perturbation may persist in addition to the stochastic nature of erosion and deposition in fluvial environments \cite{castletort2003,jerolmack2010,romans2016}.
  For instance, sediment discharge records from two glaciers in the Swiss Alps show that $40$ to $50$\% of a season's sediment discharge occurs when water discharge is below the $75^{\mathrm{th}}$ quantile of the season \cite{delaney2018}.
  Such a large quantity of sediment transported at lower water discharges may occur in part because sediment transport capacity responds the size of the glacier conduit, in addition to water discharge (Section~\ref{sect:sub_mode}).

  This suggests that a larger  perturbation of water discharge might be needed to significantly alter the sediment discharge capacity from the glaciers compared to subaerial systems, to overcome the noise resulting from increased variability in sediment transport capacity.
Furthermore, if no sediment is available and the glacier is in a supply-limited case, then the sediment discharge record will also represent additional processes of bedrock erosion from sliding and water pressure variations  \cite{iverson2012,herman2015}.

\section{Conclusions}
  Subaerial channels can alter their channel width and water velocity in response to changing water discharge.
  Conversely, pressurized subglacial channels accommodate changing water discharge by altering water velocity and shear stress, upon which sediment transport depends.
  This occurs because the size of subglacial channels evolves slowly compared to variations in water discharge.

  Parameterizations of subglacial and subaerial water flow show that subglacial systems exhibit increased variability in sediment transport capacity as they respond  to changes in water discharge.
  The variability in sediment transport capacity is reduced when the evolving channel width is accounted for in subaerial systems.
  Even so, variability in sediment transport capacity across a channel's width is consistently higher in subglacial channels compared to subaerial ones.

  These different characteristics between glacial and subaerial channels show that records of sediment transport downstream of glaciers capture both subglacial and subaerial processes together.
  Such processes might be particularly relevant at glacier margins, where water transitions from pressurized subglacial flow to open-channel subaerial flow.
  The inconsistent response of sediment transport in subglacial and subaerial systems to changing water discharge could add uncertainty in evaluating  sediment transport signals in subglacial systems.
  In addition, the subglacial parameterization shows that sediment transport capacity in glacier systems responds to a large number of factors such as channel size, ice thickness, and glacier surface slope, which may react to climate forcing differently than water discharge alone. 
  The differences in subglacial and subaerial channels' response to changing hydrology should be considered when examining sediment transport processes in glacierized catchments.

  \section*{Open Research}
  \noindent
  The julia code used herein is included in the supplementary material.

\acknowledgments
I was funded by SNSF Project No. PZ00P2\_202024.

%% ------------------------------------------------------------------------ %%
%% References and Citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
% 
% \bibliography{<name of your .bib file>} don't specify the file extension
% 
% don't specify bibliographystylei

% In the References section, cite the data/software described in the Availability Statement (this includes primary and processed data used for your research). For details on data/software citation as well as examples, see the Data & Software Citation section of the Data & Software for Authors guidance
% https://www.agu.org/Publish-with-AGU/Publish/Author-Resources/Data-and-Software-for-Authors#citation

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\bibliography{Paperlib.bib}
\begin{thebibliography}{}

\bibitem [\protect \citeauthoryear {%
Andrews%
\ \protect \BOthers {.}}{%
Andrews%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2014}}%
}]{%
andrews2014}
\APACinsertmetastar {%
andrews2014}%
\begin{APACrefauthors}%
Andrews, L\BPBI C.%
, Catania, G\BPBI A.%
, Hoffman, M\BPBI J.%
, Gulley, J\BPBI D.%
, L{\"u}thi, M\BPBI P.%
, Ryser, C.%
\BDBL {}Neumann, T\BPBI A.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2014}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Direct observations of evolving subglacial drainage
  beneath the {G}reenland {I}ce {S}heet} {Direct observations of evolving
  subglacial drainage beneath the {G}reenland {I}ce {S}heet}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Nature}{514}{7520}{80}.
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Beaud%
, Flowers%
\BCBL {}\ \BBA {} Venditti%
}{%
Beaud%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2018}}%
}]{%
beaud2018}
\APACinsertmetastar {%
beaud2018}%
\begin{APACrefauthors}%
Beaud, F.%
, Flowers, G.%
\BCBL {}\ \BBA {} Venditti, J\BPBI G.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2018}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Modeling sediment transport in ice-walled subglacial
  channels and its implications for esker formation and pro-glacial sediment
  yields} {Modeling sediment transport in ice-walled subglacial channels and
  its implications for esker formation and pro-glacial sediment yields}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Journal of Geophysical Research: Earth
  Surface}{123}{}{1--56}.
\newblock
\begin{APACrefDOI} \doi{10.1029/2018JF004779} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Bendixen%
\ \protect \BOthers {.}}{%
Bendixen%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2017}}%
}]{%
bendixen2017}
\APACinsertmetastar {%
bendixen2017}%
\begin{APACrefauthors}%
Bendixen, M.%
, Iversen, L\BPBI L.%
, Bj{\o}rk, A\BPBI A.%
, Elberling, B.%
, Westergaard-Nielsen, A.%
, Overeem, I.%
\BDBL {}others%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2017}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Delta progradation in {G}reenland driven by increasing
  glacial mass loss} {Delta progradation in {G}reenland driven by increasing
  glacial mass loss}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Nature}{550}{7674}{101}.
\newblock
\begin{APACrefDOI} \doi{10.1038/nature23873} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Castelltort%
\ \BBA {} {Van Den Driessche}%
}{%
Castelltort%
\ \BBA {} {Van Den Driessche}%
}{%
{\protect \APACyear {2003}}%
}]{%
castletort2003}
\APACinsertmetastar {%
castletort2003}%
\begin{APACrefauthors}%
Castelltort, S.%
\BCBT {}\ \BBA {} {Van Den Driessche}, J.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2003}{}{}.
\newblock
{\BBOQ}\APACrefatitle {{How plausible are high-frequency sediment supply-driven
  cycles in the stratigraphic record?}} {{How plausible are high-frequency
  sediment supply-driven cycles in the stratigraphic record?}}{\BBCQ}
\newblock
\APACjournalVolNumPages{Sedimentary Geology}{157}{1}{3-13}.
\newblock
\begin{APACrefURL}
  \url{https://www.sciencedirect.com/science/article/pii/S0037073803000666}
  \end{APACrefURL}
\newblock
\begin{APACrefDOI} \doi{10.1016/S0037-0738(03)00066-6} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Clarke%
}{%
Clarke%
}{%
{\protect \APACyear {2003}}%
}]{%
clarke2003}
\APACinsertmetastar {%
clarke2003}%
\begin{APACrefauthors}%
Clarke, G\BPBI K\BPBI C.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2003}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Hydraulics of subglacial outburst floods: new insights
  from the {S}pring--{H}utter formulation} {Hydraulics of subglacial outburst
  floods: new insights from the {S}pring--{H}utter formulation}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Journal of Glaciology}{49}{165}{299--313}.
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Costa%
\ \protect \BOthers {.}}{%
Costa%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2018}}%
}]{%
costa2017}
\APACinsertmetastar {%
costa2017}%
\begin{APACrefauthors}%
Costa, A.%
, Molnar, P.%
, Stutenbecker, L.%
, Bakker, M.%
, Silva, T\BPBI A.%
, Schlunegger, F.%
\BDBL {}Girardclos, S.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2018}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Temperature signal in suspended sediment export from an
  {A}lpine catchment} {Temperature signal in suspended sediment export from an
  {A}lpine catchment}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Hydrology and Earth System Sciences}{22}{1}{509-528}.
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Cowan%
, Powell%
\BCBL {}\ \BBA {} Smith%
}{%
Cowan%
\ \protect \BOthers {.}}{%
{\protect \APACyear {1988}}%
}]{%
cowan1988}
\APACinsertmetastar {%
cowan1988}%
\begin{APACrefauthors}%
Cowan, E.%
, Powell, R.%
\BCBL {}\ \BBA {} Smith, N.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{1988}{}{}.
\newblock
{\BBOQ}\APACrefatitle {{Rainstorm-induced event sedimentation at the tidewater
  front of a temperate glacier}} {{Rainstorm-induced event sedimentation at the
  tidewater front of a temperate glacier}}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Geology}{16}{5}{409--412}.
\newblock
\begin{APACrefDOI}
  \doi{https://doi.org/10.1130/0091-7613(1988)016%3C0409:RIESAT%3E2.3.CO;2}
  \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Creyts%
, Clarke%
\BCBL {}\ \BBA {} Church%
}{%
Creyts%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2013}}%
}]{%
creyts2013}
\APACinsertmetastar {%
creyts2013}%
\begin{APACrefauthors}%
Creyts, T\BPBI T.%
, Clarke, G\BPBI K\BPBI C.%
\BCBL {}\ \BBA {} Church, M.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2013}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Evolution of subglacial overdeepenings in response to
  sediment redistribution and glaciohydraulic supercooling} {Evolution of
  subglacial overdeepenings in response to sediment redistribution and
  glaciohydraulic supercooling}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Journal of Geophysical Research: Earth
  Surface}{118}{2}{423--446}.
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Delaney%
\ \BBA {} Adhikari%
}{%
Delaney%
\ \BBA {} Adhikari%
}{%
{\protect \APACyear {2020}}%
}]{%
delaney2020}
\APACinsertmetastar {%
delaney2020}%
\begin{APACrefauthors}%
Delaney, I.%
\BCBT {}\ \BBA {} Adhikari, S.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2020}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Increased subglacial sediment discharge during century
  scale glacier retreat: consideration of ice dynamics, glacial erosion and
  fluvial sediment transport} {Increased subglacial sediment discharge during
  century scale glacier retreat: consideration of ice dynamics, glacial erosion
  and fluvial sediment transport}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Geophyiscal Research Letters}{}{}{e2019GL085672}.
\newblock
\begin{APACrefDOI} \doi{10.1029/2019GL085672} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Delaney%
\ \BBA {} Anderson%
}{%
Delaney%
\ \BBA {} Anderson%
}{%
{\protect \APACyear {2022}}%
}]{%
delaney2022}
\APACinsertmetastar {%
delaney2022}%
\begin{APACrefauthors}%
Delaney, I.%
\BCBT {}\ \BBA {} Anderson, L\BPBI S.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2022}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Debris Cover Limits Subglacial Erosion and Promotes Till
  Accumulation} {Debris cover limits subglacial erosion and promotes till
  accumulation}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Geophysical Research Letters}{49}{16}{e2022GL099049}.
\newblock
\APACrefnote{e2022GL099049 2022GL099049}
\newblock
\begin{APACrefDOI} \doi{10.1029/2022GL099049} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Delaney%
, Bauder%
, Huss%
\BCBL {}\ \BBA {} Weidmann%
}{%
Delaney%
, Bauder%
, Huss%
\BCBL {}\ \BBA {} Weidmann%
}{%
{\protect \APACyear {2018}}%
}]{%
delaney2017}
\APACinsertmetastar {%
delaney2017}%
\begin{APACrefauthors}%
Delaney, I.%
, Bauder, A.%
, Huss, M.%
\BCBL {}\ \BBA {} Weidmann, Y.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2018}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Proglacial erosion rates and processes in a glacierized
  catchment in the {S}wiss {A}lps} {Proglacial erosion rates and processes in a
  glacierized catchment in the {S}wiss {A}lps}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Earth Surface Processes and Landfroms}{43}{4}{765-778}.
\newblock
\begin{APACrefDOI} \doi{10.1002/esp.4239} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Delaney%
, Bauder%
, Werder%
\BCBL {}\ \BBA {} Farinotti%
}{%
Delaney%
, Bauder%
, Werder%
\BCBL {}\ \BBA {} Farinotti%
}{%
{\protect \APACyear {2018}}%
}]{%
delaney2018}
\APACinsertmetastar {%
delaney2018}%
\begin{APACrefauthors}%
Delaney, I.%
, Bauder, A.%
, Werder, M\BPBI A.%
\BCBL {}\ \BBA {} Farinotti, D.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2018}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Regional and annual variability in subglacial sediment
  transport by water for two glaciers in the {S}wiss {A}lps} {Regional and
  annual variability in subglacial sediment transport by water for two glaciers
  in the {S}wiss {A}lps}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Frontiers in Earth Science}{}{}{}.
\newblock
\begin{APACrefDOI} \doi{10.3389/feart.2018.00175} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Delaney%
, Werder%
\BCBL {}\ \BBA {} Farinotti%
}{%
Delaney%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2019}}%
}]{%
delaney2019}
\APACinsertmetastar {%
delaney2019}%
\begin{APACrefauthors}%
Delaney, I.%
, Werder, M.%
\BCBL {}\ \BBA {} Farinotti, D.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2019}{}{}.
\newblock
{\BBOQ}\APACrefatitle {{A} {N}umerical {M}odel for {F}luvial {T}ransport of
  {S}ubglacial {S}ediment} {{A} {N}umerical {M}odel for {F}luvial {T}ransport
  of {S}ubglacial {S}ediment}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Journal of Geophysical Research: Earth
  Surface}{124}{8}{2197--2223}.
\newblock
\begin{APACrefDOI} \doi{10.1029/2019JF005004} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Egli%
, Belotti%
, Ouvry%
, Irving%
\BCBL {}\ \BBA {} Lane%
}{%
Egli%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2021}}%
}]{%
egli2021b}
\APACinsertmetastar {%
egli2021b}%
\begin{APACrefauthors}%
Egli, P.%
, Belotti, B.%
, Ouvry, B.%
, Irving, J.%
\BCBL {}\ \BBA {} Lane, S.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2021}{}{}.
\newblock
{\BBOQ}\APACrefatitle {{Subglacial Channels, Climate Warming, and Increasing
  Frequency of Alpine Glacier Snout Collapse}} {{Subglacial Channels, Climate
  Warming, and Increasing Frequency of Alpine Glacier Snout Collapse}}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Geophysical Research Letters}{48}{21}{e2021GL096031}.
\newblock
\APACrefnote{e2021GL096031 2021GL096031}
\newblock
\begin{APACrefDOI} \doi{10.1029/2021GL096031} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Engelund%
\ \BBA {} Hansen%
}{%
Engelund%
\ \BBA {} Hansen%
}{%
{\protect \APACyear {1967}}%
}]{%
engelund1967}
\APACinsertmetastar {%
engelund1967}%
\begin{APACrefauthors}%
Engelund, F.%
\BCBT {}\ \BBA {} Hansen, E.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{1967}{}{}.
\newblock
\APACrefbtitle {A monograph on sediment transport in alluvial streams} {A
  monograph on sediment transport in alluvial streams}\
  \APACbVolEdTR{}{\BTR{}}.
\newblock
\APACaddressInstitution{Copenhagen, Denmark}{Technical University of Denmark}.
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Ganti%
\ \protect \BOthers {.}}{%
Ganti%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2016}}%
}]{%
ganti2016}
\APACinsertmetastar {%
ganti2016}%
\begin{APACrefauthors}%
Ganti, V.%
, von Hagke, C.%
, Scherler, D.%
, Lamb, M.%
, Fischer, W.%
\BCBL {}\ \BBA {} Avouac, J\BHBI P.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2016}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Time scale bias in erosion rates of glaciated
  landscapes} {Time scale bias in erosion rates of glaciated
  landscapes}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Science Advances}{2}{10}{}.
\newblock
\begin{APACrefDOI} \doi{10.1126/sciadv.1600204} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Gimbert%
, Tsai%
, Amundson%
, Bartholomaus%
\BCBL {}\ \BBA {} Walter%
}{%
Gimbert%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2016}}%
}]{%
gimbert2016}
\APACinsertmetastar {%
gimbert2016}%
\begin{APACrefauthors}%
Gimbert, F.%
, Tsai, V\BPBI C.%
, Amundson, J\BPBI M.%
, Bartholomaus, T\BPBI C.%
\BCBL {}\ \BBA {} Walter, J\BPBI I.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2016}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Subseasonal changes observed in subglacial channel
  pressure, size, and sediment transport} {Subseasonal changes observed in
  subglacial channel pressure, size, and sediment transport}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Geophysical Research Letters}{43}{8}{3786--3794}.
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Hallet%
}{%
Hallet%
}{%
{\protect \APACyear {1979}}%
}]{%
hallet1979}
\APACinsertmetastar {%
hallet1979}%
\begin{APACrefauthors}%
Hallet, B.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{1979}{}{}.
\newblock
{\BBOQ}\APACrefatitle {A theoretical model of glacial abrasion} {A theoretical
  model of glacial abrasion}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Journal of Glaciology}{23}{89}{39--50}.
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Hallet%
, Hunter%
\BCBL {}\ \BBA {} Bogen%
}{%
Hallet%
\ \protect \BOthers {.}}{%
{\protect \APACyear {1996}}%
}]{%
hallet1996}
\APACinsertmetastar {%
hallet1996}%
\begin{APACrefauthors}%
Hallet, B.%
, Hunter, L.%
\BCBL {}\ \BBA {} Bogen, J.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{1996}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Rates of erosion and sediment evacuation by glaciers:
  {A} review of field data and their implications} {Rates of erosion and
  sediment evacuation by glaciers: {A} review of field data and their
  implications}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Global and Planetary Change}{12}{1}{213--235}.
\newblock
\begin{APACrefDOI} \doi{10.1016/0921-8181(95)00021-6} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Hasholt%
, van As%
, Mikkelsen%
, Mernild%
\BCBL {}\ \BBA {} Yde%
}{%
Hasholt%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2018}}%
}]{%
hasholt2018}
\APACinsertmetastar {%
hasholt2018}%
\begin{APACrefauthors}%
Hasholt, B.%
, van As, D.%
, Mikkelsen, A.%
, Mernild, S.%
\BCBL {}\ \BBA {} Yde, J.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2018}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Observed sediment and solute transport from the
  {K}angerlussuaq sector of the {G}reenland {I}ce {S}heet (2006–2016)}
  {Observed sediment and solute transport from the {K}angerlussuaq sector of
  the {G}reenland {I}ce {S}heet (2006–2016)}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Arctic, Antarctic, and Alpine
  Research}{50}{1}{S100009}.
\newblock
\begin{APACrefDOI} \doi{10.1080/15230430.2018.1433789} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Herman%
\ \protect \BOthers {.}}{%
Herman%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2015}}%
}]{%
herman2015}
\APACinsertmetastar {%
herman2015}%
\begin{APACrefauthors}%
Herman, F.%
, Beyssac, O.%
, Brughelli, M.%
, Lane, S.%
, Leprince, S.%
, Adatte, T.%
\BDBL {}Cox, S\BPBI C.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2015}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Erosion by an alpine glacier} {Erosion by an alpine
  glacier}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Science}{350}{6257}{193--195}.
\newblock
\begin{APACrefDOI} \doi{10.1126/science.aab2386} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Hewitt%
\ \BBA {} Creyts%
}{%
Hewitt%
\ \BBA {} Creyts%
}{%
{\protect \APACyear {2019}}%
}]{%
hewitt2019}
\APACinsertmetastar {%
hewitt2019}%
\begin{APACrefauthors}%
Hewitt, I.%
\BCBT {}\ \BBA {} Creyts, T.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2019}{}{}.
\newblock
{\BBOQ}\APACrefatitle {A model for the formation of eskers} {A model for the
  formation of eskers}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Geophysical Research Letters}{46}{12}{6673--6680}.
\newblock
\begin{APACrefDOI} \doi{10.1029/2019GL082304} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Hooke%
, Laumann%
\BCBL {}\ \BBA {} Kohler%
}{%
Hooke%
\ \protect \BOthers {.}}{%
{\protect \APACyear {1990}}%
}]{%
hooke1990}
\APACinsertmetastar {%
hooke1990}%
\begin{APACrefauthors}%
Hooke, R\BPBI L.%
, Laumann, T.%
\BCBL {}\ \BBA {} Kohler, J.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{1990}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Subglacial Water Pressures and the Shape of Subglacial
  Conduits} {Subglacial water pressures and the shape of subglacial
  conduits}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Journal of Glaciology}{36}{122}{67--71}.
\newblock
\begin{APACrefDOI} \doi{10.3189/S0022143000005566} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Iken%
\ \BBA {} Bindschadler%
}{%
Iken%
\ \BBA {} Bindschadler%
}{%
{\protect \APACyear {1986}}%
}]{%
iken1986}
\APACinsertmetastar {%
iken1986}%
\begin{APACrefauthors}%
Iken, A.%
\BCBT {}\ \BBA {} Bindschadler, R\BPBI A.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{1986}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Combined measurements of subglacial water pressure and
  surface velocity of {F}indelengletscher, {S}witzerland: conclusions about
  drainage system and sliding mechanism} {Combined measurements of subglacial
  water pressure and surface velocity of {F}indelengletscher, {S}witzerland:
  conclusions about drainage system and sliding mechanism}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Journal of Glaciology}{32}{110}{101--119}.
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Iverson%
}{%
Iverson%
}{%
{\protect \APACyear {2012}}%
}]{%
iverson2012}
\APACinsertmetastar {%
iverson2012}%
\begin{APACrefauthors}%
Iverson, N\BPBI R.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2012}{}{}.
\newblock
{\BBOQ}\APACrefatitle {A theory of glacial quarrying for landscape evolution
  models} {A theory of glacial quarrying for landscape evolution
  models}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Geology}{40}{8}{679--682}.
\newblock
\begin{APACrefDOI} \doi{10.1130/G33079.1} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Jansson%
, Rosqvist%
\BCBL {}\ \BBA {} Schneider%
}{%
Jansson%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2005}}%
}]{%
jansson2005}
\APACinsertmetastar {%
jansson2005}%
\begin{APACrefauthors}%
Jansson, P.%
, Rosqvist, G.%
\BCBL {}\ \BBA {} Schneider, T.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2005}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Glacier fluctuations, suspended sediment flux and
  glacio‐lacustrine sediments} {Glacier fluctuations, suspended sediment flux
  and glacio‐lacustrine sediments}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Geografiska Annaler: Series A, Physical
  Geography}{87}{1}{37-50}.
\newblock
\begin{APACrefDOI} \doi{10.1111/j.0435-3676.2005.00243.x} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Jerolmack%
\ \BBA {} Paola%
}{%
Jerolmack%
\ \BBA {} Paola%
}{%
{\protect \APACyear {2010}}%
}]{%
jerolmack2010}
\APACinsertmetastar {%
jerolmack2010}%
\begin{APACrefauthors}%
Jerolmack, D.%
\BCBT {}\ \BBA {} Paola, C.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2010}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Shredding of environmental signals by sediment
  transport} {Shredding of environmental signals by sediment transport}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Geophysical Research Letters}{37}{19}{}.
\newblock
\begin{APACrefDOI} \doi{10.1029/2010GL044638} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Kasmalkar%
, Mantelli%
\BCBL {}\ \BBA {} Suckale%
}{%
Kasmalkar%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2019}}%
}]{%
kasmalkar2019}
\APACinsertmetastar {%
kasmalkar2019}%
\begin{APACrefauthors}%
Kasmalkar, I.%
, Mantelli, E.%
\BCBL {}\ \BBA {} Suckale, J.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2019}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Spatial heterogeneity in subglacial drainage driven by
  till erosion} {Spatial heterogeneity in subglacial drainage driven by till
  erosion}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Proceedings of the Royal Society A: Mathematical,
  Physical and Engineering Sciences}{475}{2228}{20190259}.
\newblock
\begin{APACrefDOI} \doi{10.1098/rspa.2019.0259} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Koppes%
\ \protect \BOthers {.}}{%
Koppes%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2015}}%
}]{%
koppes2015}
\APACinsertmetastar {%
koppes2015}%
\begin{APACrefauthors}%
Koppes, M.%
, Hallet, B.%
, Rignot, E.%
, Mouginot, J.%
, Wellner, J\BPBI S.%
\BCBL {}\ \BBA {} Boldt, K.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2015}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Observed latitudinal variations in erosion as a function
  of glacier dynamics} {Observed latitudinal variations in erosion as a
  function of glacier dynamics}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Nature}{526}{7571}{100--103}.
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Koppes%
\ \BBA {} Montgomery%
}{%
Koppes%
\ \BBA {} Montgomery%
}{%
{\protect \APACyear {2009}}%
}]{%
koppes2009a}
\APACinsertmetastar {%
koppes2009a}%
\begin{APACrefauthors}%
Koppes, M.%
\BCBT {}\ \BBA {} Montgomery, D\BPBI R.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2009}{}{}.
\newblock
{\BBOQ}\APACrefatitle {The relative efficacy of fluvial and glacial erosion
  over modern to orogenic timescales} {The relative efficacy of fluvial and
  glacial erosion over modern to orogenic timescales}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Nature Geoscience}{2}{9}{644--647}.
\newblock
\begin{APACrefDOI} \doi{10.1038/ngeo616} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Lane%
, Bakker%
, Gabbud%
, Micheletti%
\BCBL {}\ \BBA {} Saugy%
}{%
Lane%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2017}}%
}]{%
lane2016}
\APACinsertmetastar {%
lane2016}%
\begin{APACrefauthors}%
Lane, S.%
, Bakker, M.%
, Gabbud, C.%
, Micheletti, N.%
\BCBL {}\ \BBA {} Saugy, J.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2017}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Sediment export, transient landscape response and
  catchment-scale connectivity following rapid climate warming and alpine
  glacier recession} {Sediment export, transient landscape response and
  catchment-scale connectivity following rapid climate warming and alpine
  glacier recession}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Geomorphology}{277}{}{210 - 227}.
\newblock
\begin{APACrefDOI} \doi{10.1016/j.geomorph.2016.02.015} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Leopold%
\ \BBA {} Maddock%
}{%
Leopold%
\ \BBA {} Maddock%
}{%
{\protect \APACyear {1953}}%
}]{%
leopold1953}
\APACinsertmetastar {%
leopold1953}%
\begin{APACrefauthors}%
Leopold, L.%
\BCBT {}\ \BBA {} Maddock, T.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYear{1953}.
\newblock
\APACrefbtitle {The hydraulic geometry of stream channels and some
  physiographic implications} {The hydraulic geometry of stream channels and
  some physiographic implications}\ (\BVOL~252).
\newblock
\APACaddressPublisher{}{US Government Printing Office}.
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Li%
\ \protect \BOthers {.}}{%
Li%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2021}}%
}]{%
li2021}
\APACinsertmetastar {%
li2021}%
\begin{APACrefauthors}%
Li, D.%
, Lu, X.%
, Overeem, I.%
, Walling, D\BPBI E.%
, Syvitski, J.%
, Kettner, A\BPBI J.%
\BDBL {}Zhang, T.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2021}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Exceptional increases in fluvial sediment fluxes in a
  warmer and wetter {High Mountain Asia}} {Exceptional increases in fluvial
  sediment fluxes in a warmer and wetter {High Mountain Asia}}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Science}{374}{6567}{599--603}.
\newblock
\begin{APACrefDOI} \doi{10.1126/science.abi9649} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Mariotti%
\ \protect \BOthers {.}}{%
Mariotti%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2021}}%
}]{%
mariotti2021}
\APACinsertmetastar {%
mariotti2021}%
\begin{APACrefauthors}%
Mariotti, A.%
, Blard, P\BHBI H.%
, Charreau, J.%
, Toucanne, S.%
, Jorry, S.%
, Molliex, S.%
\BDBL {}Keddadouche, K.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2021}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Nonlinear forcing of climate on mountain denudation
  during glaciations} {Nonlinear forcing of climate on mountain denudation
  during glaciations}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Nature Geoscience}{}{}{1--7}.
\newblock
\begin{APACrefDOI} \doi{10.1038/s41561-020-00672-2} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Meyer-Peter%
\ \BBA {} M{\"u}ller%
}{%
Meyer-Peter%
\ \BBA {} M{\"u}ller%
}{%
{\protect \APACyear {1948}}%
}]{%
meyer1948}
\APACinsertmetastar {%
meyer1948}%
\begin{APACrefauthors}%
Meyer-Peter, E.%
\BCBT {}\ \BBA {} M{\"u}ller, R.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{1948}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Formulas for bedload transport} {Formulas for bedload
  transport}.{\BBCQ}
\newblock
\BIn{} \APACrefbtitle {Hydraulic Engineering Reports.} {Hydraulic engineering
  reports.}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
M{\"u}ller%
\ \BBA {} F{\"o}rstner%
}{%
M{\"u}ller%
\ \BBA {} F{\"o}rstner%
}{%
{\protect \APACyear {1968}}%
}]{%
muller1968}
\APACinsertmetastar {%
muller1968}%
\begin{APACrefauthors}%
M{\"u}ller, G.%
\BCBT {}\ \BBA {} F{\"o}rstner, U.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{1968}{}{}.
\newblock
{\BBOQ}\APACrefatitle {General relationship between suspended sediment
  concentration and water discharge in the {A}lpenrhein and some other rivers}
  {General relationship between suspended sediment concentration and water
  discharge in the {A}lpenrhein and some other rivers}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Nature}{217}{5125}{244--245}.
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Nanni%
\ \protect \BOthers {.}}{%
Nanni%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2020}}%
}]{%
nanni2020}
\APACinsertmetastar {%
nanni2020}%
\begin{APACrefauthors}%
Nanni, U.%
, Gimbert, F.%
, Vincent, C.%
, Gr\"aff, D.%
, Walter, F.%
, Piard, L.%
\BCBL {}\ \BBA {} Moreau, L.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2020}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Quantification of seasonal and diurnal dynamics of
  subglacial channels using seismic observations on an {A}lpine glacier}
  {Quantification of seasonal and diurnal dynamics of subglacial channels using
  seismic observations on an {A}lpine glacier}.{\BBCQ}
\newblock
\APACjournalVolNumPages{The Cryosphere}{14}{5}{1475--1496}.
\newblock
\begin{APACrefDOI} \doi{10.5194/tc-14-1475-2020} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Overeem%
\ \protect \BOthers {.}}{%
Overeem%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2017}}%
}]{%
overeem2017}
\APACinsertmetastar {%
overeem2017}%
\begin{APACrefauthors}%
Overeem, I.%
, Hudson, B\BPBI D.%
, Syvitski, J\BPBI P\BPBI M.%
, Mikkelsen, A\BPBI B.%
, Hasholt, B.%
, van~den Broeke, M\BPBI R.%
\BDBL {}Morlighem, M.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2017}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Substantial export of suspended sediment to the global
  oceans from glacial erosion in {G}reenland} {Substantial export of suspended
  sediment to the global oceans from glacial erosion in {G}reenland}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Nature Geoscience}{10}{}{859-863}.
\newblock
\begin{APACrefDOI} \doi{10.1038/NGEO3046} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Perolo%
\ \protect \BOthers {.}}{%
Perolo%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2018}}%
}]{%
perolo2018}
\APACinsertmetastar {%
perolo2018}%
\begin{APACrefauthors}%
Perolo, P.%
, Bakker, M.%
, Gabbud, C.%
, Moradi, G.%
, Rennie, C.%
\BCBL {}\ \BBA {} Lane, S.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2018}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Subglacial sediment production and snout marginal ice
  uplift during the late ablation season of a temperate valley glacier}
  {Subglacial sediment production and snout marginal ice uplift during the late
  ablation season of a temperate valley glacier}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Earth Surface Processes and Landforms}{0}{}{1-68}.
\newblock
\begin{APACrefDOI} \doi{10.1002/esp.4562} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Richards%
\ \BBA {} Moore%
}{%
Richards%
\ \BBA {} Moore%
}{%
{\protect \APACyear {2003}}%
}]{%
richards2003}
\APACinsertmetastar {%
richards2003}%
\begin{APACrefauthors}%
Richards, G.%
\BCBT {}\ \BBA {} Moore, R\BPBI D.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2003}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Suspended sediment dynamics in a steep, glacier-fed
  mountain stream, {P}lace {C}reek, {C}anada} {Suspended sediment dynamics in a
  steep, glacier-fed mountain stream, {P}lace {C}reek, {C}anada}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Hydrological Processes}{17}{9}{1733--1753}.
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Riihimaki%
, MacGregor%
, Anderson%
, Anderson%
\BCBL {}\ \BBA {} Loso%
}{%
Riihimaki%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2005}}%
}]{%
riihimaki2005}
\APACinsertmetastar {%
riihimaki2005}%
\begin{APACrefauthors}%
Riihimaki, C\BPBI A.%
, MacGregor, K\BPBI R.%
, Anderson, R\BPBI .%
, Anderson, S\BPBI P.%
\BCBL {}\ \BBA {} Loso, M\BPBI G.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2005}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Sediment evacuation and glacial erosion rates at a small
  alpine glacier} {Sediment evacuation and glacial erosion rates at a small
  alpine glacier}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Journal of Geophysical Research: Earth Surface
  (2003--2012)}{110}{F03003}{}.
\newblock
\begin{APACrefDOI} \doi{10.1029/2004JF000189} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Romans%
, Covault%
, Fildani%
\BCBL {}\ \BBA {} Walsh%
}{%
Romans%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2016}}%
}]{%
romans2016}
\APACinsertmetastar {%
romans2016}%
\begin{APACrefauthors}%
Romans, S., B.W.and~Castelltort%
, Covault, J.%
, Fildani, A.%
\BCBL {}\ \BBA {} Walsh, J.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2016}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Environmental signal propagation in sedimentary systems
  across timescales} {Environmental signal propagation in sedimentary systems
  across timescales}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Earth-Science Reviews}{153}{}{7-29}.
\newblock
\APACrefnote{Source-to-Sink Systems: Sediment \& Solute Transfer on the Earth
  Surface}
\newblock
\begin{APACrefDOI} \doi{10.1016/j.earscirev.2015.07.012} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
R{\"o}thlisberger%
}{%
R{\"o}thlisberger%
}{%
{\protect \APACyear {1972}}%
}]{%
rothlisberger1972}
\APACinsertmetastar {%
rothlisberger1972}%
\begin{APACrefauthors}%
R{\"o}thlisberger, H.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{1972}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Water pressure in intra-- and subglacial channels}
  {Water pressure in intra-- and subglacial channels}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Journal of Glaciology}{11}{62}{177--203}.
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Shields%
}{%
Shields%
}{%
{\protect \APACyear {1936}}%
}]{%
shields1936}
\APACinsertmetastar {%
shields1936}%
\begin{APACrefauthors}%
Shields, A.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{1936}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Anwendung der {A}ehnlichkeitsmechanik und der
  {T}urbulenzforschung auf die {G}eschiebebewegung} {Anwendung der
  {A}ehnlichkeitsmechanik und der {T}urbulenzforschung auf die
  {G}eschiebebewegung}.{\BBCQ}
\newblock
\APACjournalVolNumPages{PhD Thesis Technical University Berlin}{}{}{}.
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Shreve%
}{%
Shreve%
}{%
{\protect \APACyear {1972}}%
}]{%
shreve1972}
\APACinsertmetastar {%
shreve1972}%
\begin{APACrefauthors}%
Shreve, R\BPBI L.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{1972}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Movement of water in glaciers} {Movement of water in
  glaciers}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Journal of Glaciology}{11}{62}{205--214}.
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Stevens%
\ \protect \BOthers {.}}{%
Stevens%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2022}}%
}]{%
stevens2022}
\APACinsertmetastar {%
stevens2022}%
\begin{APACrefauthors}%
Stevens, D.%
, Ely, J.%
, Livingstone, S.%
, Clark, C.%
, Butcher, F.%
\BCBL {}\ \BBA {} Hewitt, I.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2022}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Effects of basal topography and ice-sheet surface slope
  in a subglacial glaciofluvial deposition model} {Effects of basal topography
  and ice-sheet surface slope in a subglacial glaciofluvial deposition
  model}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Journal of Glaciology}{}{}{1--13}.
\newblock
\begin{APACrefDOI} \doi{10.1017/jog.2022.71} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Swift%
, Nienow%
\BCBL {}\ \BBA {} Hoey%
}{%
Swift%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2005}}%
}]{%
swift2005}
\APACinsertmetastar {%
swift2005}%
\begin{APACrefauthors}%
Swift, D.%
, Nienow, P\BPBI W.%
\BCBL {}\ \BBA {} Hoey, T\BPBI B.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2005}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Basal sediment evacuation by subglacial meltwater:
  suspended sediment transport from {H}aut {G}lacier d'{A}rolla, {S}witzerland}
  {Basal sediment evacuation by subglacial meltwater: suspended sediment
  transport from {H}aut {G}lacier d'{A}rolla, {S}witzerland}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Earth Surface Processes and
  Landforms}{30}{7}{867--883}.
\newblock
\begin{APACrefDOI} \doi{10.1002/esp.1197} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Swift%
\ \protect \BOthers {.}}{%
Swift%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2021}}%
}]{%
swift2021}
\APACinsertmetastar {%
swift2021}%
\begin{APACrefauthors}%
Swift, D.%
, Tallentire, G.%
, Farinotti, D.%
, Cook, S.%
, Higson, W.%
\BCBL {}\ \BBA {} Bryant, R.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2021}{}{}.
\newblock
{\BBOQ}\APACrefatitle {The hydrology of glacier-bed overdeepenings: Sediment
  transport mechanics, drainage system morphology, and geomorphological
  implications} {The hydrology of glacier-bed overdeepenings: Sediment
  transport mechanics, drainage system morphology, and geomorphological
  implications}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Earth Surface Processes and
  Landforms}{46}{11}{2264-2278}.
\newblock
\begin{APACrefDOI} \doi{10.1002/esp.5173} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Tofelde%
, Bernhardt%
, Guerit%
\BCBL {}\ \BBA {} Romans%
}{%
Tofelde%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2021}}%
}]{%
tofelde2021}
\APACinsertmetastar {%
tofelde2021}%
\begin{APACrefauthors}%
Tofelde, S.%
, Bernhardt, A.%
, Guerit, L.%
\BCBL {}\ \BBA {} Romans, B.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2021}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Times Associated With Source-to-Sink Propagation of
  Environmental Signals During Landscape Transience} {Times associated with
  source-to-sink propagation of environmental signals during landscape
  transience}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Frontiers in Earth Science}{9}{}{}.
\newblock
\begin{APACrefDOI} \doi{10.3389/feart.2021.628315} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Tucker%
\ \BBA {} Slingerland%
}{%
Tucker%
\ \BBA {} Slingerland%
}{%
{\protect \APACyear {1997}}%
}]{%
tucker1997}
\APACinsertmetastar {%
tucker1997}%
\begin{APACrefauthors}%
Tucker, G\BPBI E.%
\BCBT {}\ \BBA {} Slingerland, R.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{1997}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Drainage basin responses to climate change} {Drainage
  basin responses to climate change}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Water Resources Research}{33}{8}{2031--2047}.
\newblock
\begin{APACrefDOI} \doi{10.1029/97WR00409} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Ugelvig%
, Egholm%
, Anderson%
\BCBL {}\ \BBA {} Iverson%
}{%
Ugelvig%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2018}}%
}]{%
ugelvig2018}
\APACinsertmetastar {%
ugelvig2018}%
\begin{APACrefauthors}%
Ugelvig, S\BPBI V.%
, Egholm, D\BPBI L.%
, Anderson, R\BPBI S.%
\BCBL {}\ \BBA {} Iverson, N\BPBI R.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2018}{}{}.
\newblock
{\BBOQ}\APACrefatitle {{Glacial Erosion Driven by Variations in Meltwater
  Drainage}} {{Glacial Erosion Driven by Variations in Meltwater
  Drainage}}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Journal of Geophysical Research: Earth
  Surface}{123}{0}{}.
\newblock
\begin{APACrefDOI} \doi{10.1029/2018JF004680} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
van As%
\ \protect \BOthers {.}}{%
van As%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2017}}%
}]{%
vanas2017}
\APACinsertmetastar {%
vanas2017}%
\begin{APACrefauthors}%
van As, D.%
, Mikkelsen, A.%
, Nielsen, M.%
, Box, J.%
, Liljedahl, L.%
, Lindb{\"a}ck, K.%
\BDBL {}Hasholt, B.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2017}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Hypsometric amplification and routing moderation of
  Greenland ice sheet meltwater release} {Hypsometric amplification and routing
  moderation of greenland ice sheet meltwater release}.{\BBCQ}
\newblock
\APACjournalVolNumPages{The Cryosphere}{11}{3}{1371--1386}.
\newblock
\begin{APACrefDOI} \doi{10.5194/tc-11-1371-2017} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Vergara%
, Garreaud%
\BCBL {}\ \BBA {} Ayala%
}{%
Vergara%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2022}}%
}]{%
vergara2022}
\APACinsertmetastar {%
vergara2022}%
\begin{APACrefauthors}%
Vergara, I.%
, Garreaud, R.%
\BCBL {}\ \BBA {} Ayala, {\'A}.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2022}{}{}.
\newblock
{\BBOQ}\APACrefatitle {{Sharp increase of extreme turbidity events due to
  deglaciation in the subtropical Andes}} {{Sharp increase of extreme turbidity
  events due to deglaciation in the subtropical Andes}}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Journal of Geophysical Research: Earth
  Surface}{127}{6}{e2021JF006584}.
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Walder%
\ \BBA {} Fowler%
}{%
Walder%
\ \BBA {} Fowler%
}{%
{\protect \APACyear {1994}}%
}]{%
walder1994}
\APACinsertmetastar {%
walder1994}%
\begin{APACrefauthors}%
Walder, J\BPBI S.%
\BCBT {}\ \BBA {} Fowler, A.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{1994}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Channelized subglacial drainage over a deformable bed}
  {Channelized subglacial drainage over a deformable bed}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Journal of Glaciology}{40}{134}{3--15}.
\newblock
\begin{APACrefDOI} \doi{10.3189/S0022143000003750} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Werder%
, Bauder%
, Funk%
\BCBL {}\ \BBA {} Keusen%
}{%
Werder%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2010}}%
}]{%
werder2010}
\APACinsertmetastar {%
werder2010}%
\begin{APACrefauthors}%
Werder, M\BPBI A.%
, Bauder, A.%
, Funk, M.%
\BCBL {}\ \BBA {} Keusen, H\BPBI R.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2010}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Hazard assessment investigations in connection with the
  formation of a lake on the tongue of {U}nterer {G}rindelwaldgletscher,
  {B}ernese {A}lps, {S}witzerland} {Hazard assessment investigations in
  connection with the formation of a lake on the tongue of {U}nterer
  {G}rindelwaldgletscher, {B}ernese {A}lps, {S}witzerland}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Natural Hazards and Earth System
  Science}{10}{2}{227--237}.
\newblock
\begin{APACrefDOI} \doi{10.5194/nhess-10-227-2010} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Werder%
, Hewitt%
, Schoof%
\BCBL {}\ \BBA {} Flowers%
}{%
Werder%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2013}}%
}]{%
werder2013}
\APACinsertmetastar {%
werder2013}%
\begin{APACrefauthors}%
Werder, M\BPBI A.%
, Hewitt, I\BPBI J.%
, Schoof, C\BPBI G.%
\BCBL {}\ \BBA {} Flowers, G\BPBI E.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2013}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Modeling channelized and distributed subglacial drainage
  in two dimensions} {Modeling channelized and distributed subglacial drainage
  in two dimensions}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Journal of Geophysical Research: Earth
  Surface}{118}{4}{2140--2158}.
\newblock
\begin{APACrefDOI} \doi{10.1002/jgrf.20146} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Whipple%
\ \BBA {} Tucker%
}{%
Whipple%
\ \BBA {} Tucker%
}{%
{\protect \APACyear {1999}}%
}]{%
whipple1999}
\APACinsertmetastar {%
whipple1999}%
\begin{APACrefauthors}%
Whipple, K.%
\BCBT {}\ \BBA {} Tucker, G.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{1999}{}{}.
\newblock
{\BBOQ}\APACrefatitle {{Dynamics of the stream-power river incision model:
  Implications for height limits of mountain ranges, landscape response
  timescales, and research needs}} {{Dynamics of the stream-power river
  incision model: Implications for height limits of mountain ranges, landscape
  response timescales, and research needs}}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Journal of Geophysical Research: Solid
  Earth}{104}{B8}{17661-17674}.
\newblock
\begin{APACrefDOI} \doi{10.1029/1999JB900120} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Wickert%
\ \BBA {} Schildgen%
}{%
Wickert%
\ \BBA {} Schildgen%
}{%
{\protect \APACyear {2019}}%
}]{%
wickert2019}
\APACinsertmetastar {%
wickert2019}%
\begin{APACrefauthors}%
Wickert, A\BPBI D.%
\BCBT {}\ \BBA {} Schildgen, T\BPBI F.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2019}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Long-profile evolution of transport-limited gravel-bed
  rivers} {Long-profile evolution of transport-limited gravel-bed
  rivers}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Earth Surface Dynamics}{7}{1}{17--43}.
\newblock
\begin{APACrefDOI} \doi{10.5194/esurf-7-17-2019} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Willenbring%
\ \BBA {} Jerolmack%
}{%
Willenbring%
\ \BBA {} Jerolmack%
}{%
{\protect \APACyear {2016}}%
}]{%
willenbring2016}
\APACinsertmetastar {%
willenbring2016}%
\begin{APACrefauthors}%
Willenbring, J.%
\BCBT {}\ \BBA {} Jerolmack, D.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2016}{}{}.
\newblock
{\BBOQ}\APACrefatitle {{The null hypothesis: globally steady rates of erosion,
  weathering fluxes and shelf sediment accumulation during Late Cenozoic
  mountain uplift and glaciation}} {{The null hypothesis: globally steady rates
  of erosion, weathering fluxes and shelf sediment accumulation during Late
  Cenozoic mountain uplift and glaciation}}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Terra Nova}{28}{1}{11-18}.
\newblock
\begin{APACrefDOI} \doi{10.1111/ter.12185} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Williams%
}{%
Williams%
}{%
{\protect \APACyear {1989}}%
}]{%
williams1989}
\APACinsertmetastar {%
williams1989}%
\begin{APACrefauthors}%
Williams, G\BPBI .%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{1989}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Sediment concentration versus water discharge during
  single hydrologic events in rivers} {Sediment concentration versus water
  discharge during single hydrologic events in rivers}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Journal of Hydrology}{111}{1}{89 -- 106}.
\newblock
\begin{APACrefURL}
  \url{http://www.sciencedirect.com/science/article/pii/0022169489902540}
  \end{APACrefURL}
\newblock
\begin{APACrefDOI} \doi{10.1016/0022-1694(89)90254-0} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Wolman%
\ \BBA {} Miller%
}{%
Wolman%
\ \BBA {} Miller%
}{%
{\protect \APACyear {1960}}%
}]{%
wolman1960}
\APACinsertmetastar {%
wolman1960}%
\begin{APACrefauthors}%
Wolman, M\BPBI G.%
\BCBT {}\ \BBA {} Miller, J\BPBI P.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{1960}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Magnitude and frequency of forces in geomorphic
  processes} {Magnitude and frequency of forces in geomorphic
  processes}.{\BBCQ}
\newblock
\APACjournalVolNumPages{The Journal of Geology}{68}{1}{54--74}.
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Wong%
\ \BBA {} Parker%
}{%
Wong%
\ \BBA {} Parker%
}{%
{\protect \APACyear {2006}}%
}]{%
wong2006}
\APACinsertmetastar {%
wong2006}%
\begin{APACrefauthors}%
Wong, M.%
\BCBT {}\ \BBA {} Parker, G.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2006}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Reanalysis and correction of bed-load relation of
  {M}eyer-{P}eter and {M}{\"u}ller using their own database} {Reanalysis and
  correction of bed-load relation of {M}eyer-{P}eter and {M}{\"u}ller using
  their own database}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Journal of Hydraulic Engineering}{132}{11}{1159--1168}.
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Yager%
, Venditti%
, Smith%
\BCBL {}\ \BBA {} Schmeeckle%
}{%
Yager%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2018}}%
}]{%
yager2018}
\APACinsertmetastar {%
yager2018}%
\begin{APACrefauthors}%
Yager, E.%
, Venditti, J.%
, Smith, H.%
\BCBL {}\ \BBA {} Schmeeckle, M.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2018}{}{}.
\newblock
{\BBOQ}\APACrefatitle {The trouble with shear stress} {The trouble with shear
  stress}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Geomorphology}{323}{}{41--50}.
\PrintBackRefs{\CurrentBib}

\bibitem [\protect \citeauthoryear {%
Zhang%
\ \protect \BOthers {.}}{%
Zhang%
\ \protect \BOthers {.}}{%
{\protect \APACyear {2022}}%
}]{%
zhang2022}
\APACinsertmetastar {%
zhang2022}%
\begin{APACrefauthors}%
Zhang, T.%
, Li, D.%
, East, A.%
, Walling, D.%
, Lane, S.%
, Overeem, I.%
\BDBL {}Lu, X.%
\end{APACrefauthors}%
\unskip\
\newblock
\APACrefYearMonthDay{2022}{}{}.
\newblock
{\BBOQ}\APACrefatitle {Warming-driven erosion and sediment transport in cold
  regions} {Warming-driven erosion and sediment transport in cold
  regions}.{\BBCQ}
\newblock
\APACjournalVolNumPages{Nature Reviews Earth \& Environment}{}{}{1--20}.
\newblock
\begin{APACrefDOI} \doi{10.1038/s43017-022-00362-0} \end{APACrefDOI}
\PrintBackRefs{\CurrentBib}

\end{thebibliography}

\section{Supplement}

\renewcommand{\figurename}{Figure S}
\setcounter{figure}{0}

\begin{center}
  \begin{figure}[H]
    \includegraphics[width=0.7\linewidth]{multi_run_vars.pdf}
    \caption{Variability $\chi$ in velocity ($v_g$, $v_f$; a) shear stress ($\tau_g$, $\tau_f$; b) and width-integrated shear stress ($\tau_{tg}$, $\tau_{tf}$; c)  in glacial and subaerial systems. Across the parameter space in Figure~\ref{fig:range}. }
    \label{fig:gammas}
  \end{figure}
\end{center}

\end{document}

### 2023-03-23 theory.tex
GRL template added

@@
   Additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to subaerial systems \citep{jansson2005,ganti2016}. 
-  For instance, sediment expelled from a glacier in a transport-limited regime reacts, in addition to water discharge, to sediment discharge capacity defined by the thickness of the ice, controlling the closure of the channel, and the surface slope of the glacier, controlling water velocity \citep[Section~\ref{sect:sub_mode}; ] []{rothlisberger1972,shreve1972,delaney2022,stevens2022}.
+  For instance, sediment transport capacity from a glacier reacts to the ice thickness, controlling the closure of the channel, and the surface slope of the glacier, controlling water velocity \citep[Section~\ref{sect:sub_mode}; ] []{rothlisberger1972,shreve1972,delaney2022,stevens2022}, in addition to water discharge.
   Conversely, sediment transport in most subaerial systems increases with water discharge and bed slope, with the bed slope, probably being more stable over yearly to century-scale time periods \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wong2006,wickert2019}.
@@
   \section{Conclusions}
-  Contrary to subaerial channels that can alter their channel width and water velocity in response to changing water discharge,
-   pressurized subglacial channels accommodate changing water discharge by altering water velocity and shear stress, upon which sediment transport depends.
+
+  Pressurized subglacial channels accommodate changing water discharge by altering water velocity and shear stress, upon which sediment transport depends.
   This occurs because the size of subglacial channels evolves slowly compared to variations in water discharge.
+  Conversely, subaerial channels can alter their channel width and water velocity in response to changing water discharge.
 

### 2023-03-23 Delaney2023.tex
more small changes

@@
 
-\item  Greater variability in water velocity in subglacial systems causes sediment transport capacity to vary more compared to subaerial systems.
+\item  Greater variability in water velocity in subglacial systems causes sediment transport capacity to vary more than subaerial systems.
 
@@
 Sediment transport capacity depends mainly on the shear stress or velocity of water flowing through a channel and the channel width over which to mobilize sediment.
-In subaerial channels, changing water discharge can be accommodated by rapid changes in the width of the channel over which water flows, the water depth, and the water velocity.
+In subaerial channels, changing water discharge can be accommodated by rapid changes in the channel width over which water flows, the water depth, and the water velocity.
 In subglacial channels, however, water flows through a conduit whose size evolves  slowly, because the water is pressurized by the glacier ice above.
@@
 \section*{Plain language summary} % 200 words
-In both subglacial and subaerial channels, sediment transport capacity depends mainly on 1) the channel width over which the sediment is mobilized  and 2) the shear stress applied by flowing water on sediment at the channel bottom; the shear stress depends largely on the water velocity.
+In both subglacial and subaerial channels, sediment transport capacity depends mainly on 1) the channel width over which the sediment is mobilized  and 2) the shear stress applied by flowing water on sediment at the channel bottom; the shear stress depends mainly on the water velocity.
 In subaerial channels, changing water discharge will  change the water depth,  the channel width (which dictates the distance  over which sediment transport  occurs), and the  water velocity (which controls the sediment transport capacity).
@@
   Increases in sediment transport  have been observed in Greenland \cite{bendixen2017}, the European Alps \cite{costa2017}, the Himalayas \cite{li2021}, and the Andes \cite{vergara2022}.
-  In some of these regions, increased water discharge and glacier melt have been attributed to the greater sediment transport \cite{bendixen2017,costa2017,li2021}.
+  In some of these regions, increased water discharge and glacier melt have been attributed to greater sediment transport \cite{bendixen2017,costa2017,li2021}.
   In other cases, enhanced access of meltwater to subglacial sediment at high elevations was attributed as the cause of increased sediment transport \cite{delaney2020,vergara2022}.
@@
   Across the hydrological forcing, variable water discharge results in a substantial increase in the range of water velocities in the glacial system compared to the subaerial system (Figure~\ref{fig:model_outs}).
-  Water velocity in the glacier system vary by  $0.5$ (Equation~\ref{eq:var}) over the discharge values, in the subaerial system the water velocity only varies by $0.11$ (Figure~\ref{fig:model_outs}\,e).
+  Water velocity in the glacier system varies by  $0.5$ (Equation~\ref{eq:var}) over the discharge values, in the subaerial system the water velocity only varies by $0.11$ (Figure~\ref{fig:model_outs}\,e).
   Shear stress at the glacier bed varies by $0.75$ across the range of water discharge values, whereas shear stress across the subaerial system varies by roughly $0.21$ (Figure~\ref{fig:model_outs}\,f).
@@
   For instance, sediment transport capacity from a glacier reacts to the ice thickness, controlling the closure of the channel, and the surface slope of the glacier, controlling water velocity \cite<Section~\ref{sect:sub_mode}; >{rothlisberger1972,shreve1972,delaney2022,stevens2022}, in addition to water discharge.
-  Conversely, sediment transport in most subaerial systems increases with water discharge and bed slope, with the bed slope probably being more stable over yearly to century-scale time periods \cite<Section~\ref{sect:fluv}; e.g.>{muller1968,whipple1999,wong2006,wickert2019}.
-
-  Records of sediment discharge have been used to establish the relationship, or lack thereof, between sediment discharge and climate in glacial systems \cite<e.g.>[]{koppes2009a,willenbring2016,mariotti2021}.
+  Conversely, sediment transport in most subaerial systems increases with water discharge and bed slope, and the bed slope probably remains relatively stable over yearly to century-scale time periods \cite<Section~\ref{sect:fluv}; e.g.>{muller1968,whipple1999,wong2006,wickert2019}.
+
+  Records of sediment discharge have been used to establish the relationship, or lack thereof, between sediment discharge and climate in glacial systems \, cite<e.g.>[]{koppes2009a,willenbring2016,mariotti2021}.
   The variability in subglacial sediment transport capacity presented here applies most generally to short-time scales responsible for the size of subglacial channels.
@@
   For instance, sediment discharge records from two glaciers in the Swiss Alps show that $40$ to $50$\% of a season's sediment discharge occurs when water discharge is below the $75^{\mathrm{th}}$ quantile of the season \cite{delaney2018}.
-  Such a large quantity of sediment transported at lower water discharges may occur in part because sediment transport capacity responds the size of the glacier conduit, in addition to water discharge (Section~\ref{sect:sub_mode}).
+  Such a large quantity of sediment transported at lower water discharges may occur in part because sediment transport capacity responds to the size of the glacier conduit, in addition to water discharge (Section~\ref{sect:sub_mode}).
 

### 2023-03-23 theory.tex
more small changes

@@
 \begin{enumerate}
-\item Water discharge variations in subglacial systems are accommodated by water velocity, not channel size, given the slow evolution of channels.
+\item Water discharge variations in subglacial systems are accommodated by water velocity, not channel size, given the channel's slow evolution.
 
@@
   % 
-  \noindent Below we examine variables, $\Gamma$, as water discharge ($Q_w$) water velocity ($v_g$, $v_f$), shear stress ($\tau_g$, $\tau_f$), and integrated shear stress ($\tau_{tg}$, $\tau_{tf}$) with the purpose of evaluating total shear stress across the glacier bed.
+  \noindent The results below examine variables $\Gamma$, as water discharge ($Q_w$) water velocity ($v_g$, $v_f$), shear stress ($\tau_g$, $\tau_f$), and integrated shear stress ($\tau_{tg}$, $\tau_{tf}$) with the purpose of evaluating total shear stress across the glacier bed.
 
@@
 Horizontal bars in d, e, and f represent the range of values in the  output.
-Gray space represents water discharge below the mean value.
-}
+Gray space represents water discharge below the mean value.}
       \label{fig:model_outs}
@@
   The subaerial parameterization assumes that sediment is transported across all channel widths, as opposed to large flows such as floods when much sediment transport occurs \citep{wolman1960}.
-  As a result, the variability in the subaerial system across the range of discharge values could be underestimated in the parameterization (Figure~\ref{fig:range}).
+  As a result, the variability in velocity, shear-stress, and width integrated shear stress in the subaerial system across the range of discharge values could be underestimated in the parameterization (Figure~\ref{fig:range}).
 
@@
   As a result, variations in sediment transport capacity could be responsible for the fluctuations in sediment discharge (``flushings'') from glaciers that are not attributed to variations in water discharge \citep[e.g.][]{richards2003,swift2021}.
-  Aside from the sporadic nature of these events, the observed flushings could result from increases in water velocity resulting from reduced channel size, not water discharge, which increase sediment transport capacity.
+  Aside from the sporadic nature of these events, the observed "flushing" could result from increases in water velocity resulting from reduced channel size, not water discharge, which increase sediment transport capacity.
 

### 2023-05-16 Delaney2023.tex
Rejected version by GRL

@@
 
-\title{Comparing variations in sediment transport capacity between subglacial and subaerial systems }
+\title{Comparing variations in sediment transport capacity between subglacial and subaerial channels }
 
@@
 
-\item The systems' divergence in transport capacity response may impact sediment dynamics at glacier margins and interpretation of sediment records.
+\item The systems' divergence in transport response may impact sediment dynamics at glacier margins and interpretation of sediment records.
 \end{keypoints}
@@
     \begin{figure}[H]
-      \includegraphics[width=0.65\linewidth]{Cartoon.pdf}
+      \includegraphics[width=0.8\linewidth]{Fig1.pdf}
       \caption{Cartoon of the situation with various parameters.
@@
     \begin{figure}[H]
-      \includegraphics[width=0.7\linewidth]{model_outputs.pdf}
+      \includegraphics[width=0.8\linewidth]{model_outputs.pdf}
       \caption{Response of channel width (a), hydraulic diameter (b),  Reynolds number  (c),  water velocity (d), shear-stress (e), and width-integrated shear-stress (f)  to different water discharge values.
@@
     \begin{figure}[H]
-      \includegraphics[width=0.7\linewidth]{multi_run.pdf}
+      \includegraphics[width=0.8\linewidth]{multi_run.pdf}
       \caption{Ratio of range of water velocity, shear-stress, and width integrated shear-stress for glacial and subaerial systems in response to mean water discharge (a, b, c), and  variability in water discharge (d, e, f).
@@
 
-  Records of sediment discharge have been used to establish the relationship, or lack thereof, between sediment discharge and climate in glacial systems \, cite<e.g.>[]{koppes2009a,willenbring2016,mariotti2021}.
+  Records of sediment discharge have been used to establish the relationship, or lack thereof, between sediment discharge and climate in glacial systems \cite<e.g.>[]{koppes2009a,willenbring2016,mariotti2021}.
   The variability in subglacial sediment transport capacity presented here applies most generally to short-time scales responsible for the size of subglacial channels.
@@
 
-  These different characteristics between glacial and subaerial channels show that records of sediment transport downstream of glaciers capture both subglacial and subaerial processes together.
-  Such processes might be particularly relevant at glacier margins, where water transitions from pressurized subglacial flow to open-channel subaerial flow.
+  These different characteristics between glacial and subaerial channels show that records of sediment transport downstream of glaciers represent both subglacial and subaerial processes together.
+  Divergent processes discussed here might be particularly relevant at glacier margins, where water transitions from pressurized subglacial flow to open-channel subaerial flow.
   The inconsistent response of sediment transport in subglacial and subaerial systems to changing water discharge could add uncertainty in evaluating  sediment transport signals in subglacial systems.
@@
 % 
-% don't specify bibliographystylei
+% don't specify bibliographystyleiz
 
@@
 
-%\bibliography{Paperlib.bib}
 \begin{thebibliography}{}
@@
 \newblock
-\begin{APACrefURL}
-  \url{https://www.sciencedirect.com/science/article/pii/S0037073803000666}
-  \end{APACrefURL}
-\newblock
 \begin{APACrefDOI} \doi{10.1016/S0037-0738(03)00066-6} \end{APACrefDOI}
@@
 \begin{APACrefauthors}%
-Williams, G\BPBI .%
+Williams, G\BPBI P.%
 \end{APACrefauthors}%
@@
 \APACjournalVolNumPages{Journal of Hydrology}{111}{1}{89 -- 106}.
-\newblock
-\begin{APACrefURL}
-  \url{http://www.sciencedirect.com/science/article/pii/0022169489902540}
-  \end{APACrefURL}
 \newblock
@@
   \begin{figure}[H]
-    \includegraphics[width=0.7\linewidth]{multi_run_vars.pdf}
+    \includegraphics[width=0.8\linewidth]{multi_run_vars.pdf}
     \caption{Variability $\chi$ in velocity ($v_g$, $v_f$; a) shear stress ($\tau_g$, $\tau_f$; b) and width-integrated shear stress ($\tau_{tg}$, $\tau_{tf}$; c)  in glacial and subaerial systems.
-Across the parameter space in Figure~\ref{fig:range}.
+Values are determined from examining the same parameter space in Figure~\ref{fig:range}.
 }

### 2023-07-10 Delaney2023.tex
more stumbling

@@
 
-\title{Comparing variations in sediment transport capacity between subglacial and subaerial channels }
+\title{Water discharge has a variable impact on sediment transport capacity in subglacial systems}
+
+\ian{Alpine glacier vs Ice sheet}
 
@@
 \begin{keypoints}
-\item Water discharge variations in subglacial channels are accommodated by water velocity, not channel size, given their slow evolution.
+\item Water discharge variations in subglacial channels are mainly accommodated by water velocity, not channel size, given their slow evolution.
 
@@
 \begin{abstract} % 150 words
-Sediment transport capacity depends mainly on the shear stress or velocity of water flowing through a channel and the channel width over which to mobilize sediment.
-In subaerial channels, changing water discharge can be accommodated by rapid changes in the channel width over which water flows, the water depth, and the water velocity.
-In subglacial channels, however, water flows through a conduit whose size evolves  slowly, because the water is pressurized by the glacier ice above.
-As a result, variations in water discharge may modify the water velocity, but not the channel's width.
-Here, parameterizations of subglacial and subaerial channels indicate that sediment transport capacity varies more in subglacial systems compared with subaerial systems, across a wide range of channel geometries and flow conditions.
-Large variations in subglacial water sediment transport capacity may impact sediment transport processes and the interpretation of records of sediment transport from glacierized catchments.
-
-\end{abstract}
-
-\section*{Plain language summary} % 200 words
-In both subglacial and subaerial channels, sediment transport capacity depends mainly on 1) the channel width over which the sediment is mobilized  and 2) the shear stress applied by flowing water on sediment at the channel bottom; the shear stress depends mainly on the water velocity.
+
+In both subglacial and subaerial channels, sediment transport capacity of a sediment size depends mainly on 1) the channel width over which the sediment is mobilized  and 2) the shear stress applied by flowing water on sediment at the channel bottom; the shear stress depends mainly on the water velocity.
 In subaerial channels, changing water discharge will  change the water depth,  the channel width (which dictates the distance  over which sediment transport  occurs), and the  water velocity (which controls the sediment transport capacity).
@@
 Parameterizations of these processes show that sediment transport capacity varies far more in subglacial systems compared to subaerial systems, even though changing channel width can accommodate some sediment transport variability in subaerial channels.
-The increased variability in sediment transport capacity in subglacial systems, compared to subaerial ones, has several important implications for sediment transport in glacierized catchments and the interpretation of sediment discharge records.
+The increased variability in sediment transport capacity in subglacial systems, compared to subaerial ones, has several important implications for sediment transport processes in glacierized catchments and the interpretation of sediment discharge records.
+
+\end{abstract}
 
@@
 %%% Suggested section heads:
+\ian{comparison with Alley 1997... mention his variability-- also deal with his $Qs \propto Qw^{\frac{9}{2}}$  treats like rivers?  does not consider channel size evolution}
+
 \section{Introduction}
@@
 Increases in sediment transport  have been observed in Greenland \cite{bendixen2017}, the European Alps \cite{costa2017}, the Himalayas \cite{li2021}, and the Andes \cite{vergara2022}.
-  In some of these regions, increased water discharge and glacier melt have been attributed to greater sediment transport \cite{bendixen2017,costa2017,li2021}.
+In some of these regions, increased water discharge and glacier melt have been attributed to greater sediment transport capacity \cite{bendixen2017,costa2017,li2021}.
 In other cases, enhanced access of meltwater to subglacial sediment at high elevations was attributed as the cause of increased sediment transport \cite{delaney2020,vergara2022}.
@@
 where $w_c$ is the width of the channel.
-  Following the conservation of mass, the velocity of the water flowing through a rectangular channel is given as 
+Following the conservation of mass, the velocity of the water flowing through a  channel is given as 
 \begin{linenomath*}
@@
 During a flood with greater water discharge $Q_w$, for instance, the cross-sectional area $S$ increases as the channel widens and deepens, along with water velocity $v$.
-As a result, the shear stress across the channel $\tau_t$ will increase in response to both the channel width $w_c$  and water velocity $v$.
+As a result, the shear stress across the channel $\tau_t$ will increase in response to both the channel width $w_c$  and water velocity $v$, which is squared.
 
 Yet, the response of water velocity to changing water discharge in subglacial channels is different than subaerial ones.
-  In subglacial systems, the water is pressurized by the ice above \cite{shreve1972}.
+In subglacial systems, the water is pressurized by the ice above \cite{rothlisberger1972}.
 The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by frictional heating of water flowing through the channel \cite{rothlisberger1972}.
-  As a result, the subglacial channel size evolves over days to months, whereas water discharge varies over hours \cite<e.g.>[]{iken1986,andrews2014,nanni2020}.
-  In turn, changes in water discharge $Q_w$ are mainly accommodated with water velocity $v$, because the size of the channel $S$ is largely fixed (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
-  It follows that greater water discharge $Q_w$ will cause the shear stress across the channel $\tau_t$ to respond to the water velocity $v$, as opposed to the subaerial system where width evolves as well.
+As a result, the subglacial channel size evolves over days to months, whereas water discharge can vary over hours \cite<e.g.>[]{iken1986,andrews2014,nanni2020}.
+In turn, changes in water discharge $Q_w$ are mainly accommodated with water velocity $v$, because the size of the channel $S$ changes on a much shorter timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
+It follows that an increase in water discharge $Q_w$ will cause the shear stress across the channel $\tau_t$ to respond to the water velocity $v$, as opposed to the subaerial system where channel width $w_c$ evolves as well.
 
@@
 These differences have been implicitly included in a wide range of available models that  quantify sediment transport from catchments in both subglacial and subaerial systems  \cite<e.g.>[]{walder1994,tucker1997,creyts2013,wickert2019,hewitt2019}.
-  The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \cite<e.g.>[]{lane2016}.
+The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \cite<e.g.>[]{lane2016,perolo2018}.
 Additionally, the variable response of sediment transport capacity to water discharge in the two systems may impact the interpretation of records of sediment transport in glacial compared to subaerial systems \cite<e.g.>[]{muller1968,richards2003,ganti2016}.
 Yet, the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial systems has been minimally discussed.
-
-  This manuscript leverages parameterizations of subglacial and subaerial hydraulics to explore differences in the relationships between  water discharge, channel geometry, and water velocity. 
-  With these parameterizations, the manuscript examines the impact of water discharge variability on sediment transport capacity in idealized subglacial and subaerial systems.
-  Lastly, the manuscript discusses the implications for sediment transport processes in subglacial systems, compared to subaerial ones.
+Furthermore, thorough observations of water velocity and channel width in subglacial systems are limited and not contrasted with subaerial conditions at the glacier margins.
+
+To better understand the relationship between sediment transport capacity and water discharge in subglacial systems, this manuscript leverages parameterizations of subglacial and subaerial hydraulics.
+These parameterizations are applied to hydrological records from the Fieschergletscher, an alpine glacier in Switzerland, and Leverett glacier,  a land-terminating glacier in Greenland.
+Outputs are used to explore differences in the relationships between  water discharge, channel geometry, and water velocity in the subglacial and subaerial. 
+With these parameterizations, the manuscript examines the impact of water discharge variability on sediment transport capacity in these subglacial and subaerial systems.
+The manuscript then uses the demonstrated differences between the two systems to discuss the implications for sediment transport processes in subglacial systems, compared to subaerial ones, and for interpreting records of sediment transport from glacierized catchments.
 
@@
     \includegraphics[width=0.8\linewidth]{Fig1.pdf}
-      \caption{Cartoon of the situation with various parameters.
+    \caption{Sketch of the situation with various parameters.
 Water velocity magnitude in the subglacial, $v_g$, and subaerial, $v_f$, systems are shown by arrow length.
@@
 
+\ian{update hooke angle }
 \begin{table}[H]
@@
     \textbf{Variables}  & & & \\
-      Channel hydraulic diameter (glacier, subaerial) &  $D_{hg},\,D_{hf}$&  & $\mathrm{m}$     \\
+    Water discharge  & $Q_w$& & $\mathrm{m^{3}\,s^{-1}}$ \\
+    Water velocity (glacier, subaerial)  & $v_g,\,v_{f}$& & $\mathrm{m\,s^{-1}}$ \\
+    Channel cross-sectional area (glacier, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
+    Hydraulic diameter &$D_h$&&$\mathrm{m}$\\
     Width of channel floor (glacial, subaerial) & $w_{cg},w_{cf}$&  & $\mathrm{m}$     \\
-      Channel cross-sectional area (glacier, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
-      Water discharge (instantaneous) & $Q_w$& & $\mathrm{m^{3}\,s^{-1}}$ \\
-      Representative water discharge & $Q_{w}^*$& & $\mathrm{m^{3}\,s^{-1}}$ \\
-      Representative hydraulic gradient  &$\Psi^*$ & & $\mathrm{Pa\, m^{-1}}$\\
-      Water velocity (glacier, subaerial)  & $v_g,\,v_{f}$& & $\mathrm{m\,s^{-1}}$ \\
+    Hydraulic head &$\Delta h$&& $\mathrm{m}$\\
     Shear stress (glacier, subaerial) & $\tau_g,\,\tau_f$&& $\mathrm{Pa \, m^{-1}}$ \\
     Width-integrated shear-stress (glacier, subaerial) & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$ \\
-      Reynolds number &$R_e$& & $\mathrm{(-)}$\\
-      Variable &$\Gamma$&&$\mathrm{(-)}$\\
-      Variability of a variable ($\Gamma$) &$\chi$& &$\mathrm{(-)}$\\
+
          &&&\\
@@
     Density of ice & $\rho_i$& $900$ & $\mathrm{kg\,m^{-3}}$ \\
-      Kinematic viscosity of water &$\nu$& $10^{-4}$& $\mathrm{m^2\,s^{-1}}$\\
     Hooke angle of channel & $\beta$ & $\frac{\pi}{2}$ & \unit{rad}\\
-      Flotation fraction & $f_f$&$0.7$& $\mathrm{(-)}$\\
+    Effective glacier thickness &$h_o$& $225^a$ or $700^b$  &\unit{m}\\
+    Effective glacier length &$l$&$7000^a$ or $20,000^b$&\unit{m}\\
+    Constant $1$ &$C_1$&$2.2\times10^{-5}$&\unit{m}$^{-1}$\\
+    Constant $2$ &$C_2$&$3.7\times10^{-13}$&\unit{m}$^{-n}\,s^{-1}$\\
+    Latent heat of fusion &$L$&$333.5 $&\unit{kJ\,kg}$^{-1}$\\
+    Pressure melting coefficient &$c_t$&$7.5\times 10^{-8}$&\unit{K\,Pa}$^{-1}$\\
+    Specific heat capacity of water &$c_p$&$4180$&\unit{J\,kg}$^{-1}$\unit{K}$^{-1}$\\
+
+    Ice flow constant &$A$& $5.3\times10^{-24}$ &\unit{Pa}$^{-n}$\,$s^{-1}$\\
+    Ice flow exponent &$n$& $3$ &$\mathrm{(-)}$\\
     Friction factor (glacier) & $f_r$ & $4$ & $\mathrm{(-)}$ \\
     Friction factor (subaerial) & $f_p$ & $3$ & $\mathrm{(-)}$\\
-      Gradient of glacier surface & $\frac{\partial z_{sg}}{\partial x}$ &$0.25$& $\mathrm{(-)}$\\
     Gradient of channel bed (subaerial) &$\frac{\partial z_c}{\partial x}$ &$0.05$& $\mathrm{(-)}$\\
@@
   \label{table:vpm}
+
 \end{table}
+  $^a$ for Fieschergletscher
+
+  $^b$ for Leverett glacier
 
@@
 \label{sect:sub_mode}
+
 To evaluate the shear stress of water flowing across sediments below a glacier, the subglacial channel parameterization evaluates the channel geometry and the velocity of the flowing water.
-  To accomplish this, the hydraulics model presented in \citeA{delaney2019} is used.
-
-  Here, it is assumed that the water is transported through subglacial channels \cite<Figure~\ref{fig:cartoon}; >{rothlisberger1972}, and that the channel  geometry will respond to a representative water discharge over a certain time period $Q_{w}^*$, using the Darcy-Weisbach formulation for water-flow through a pipe  \cite<e.g.>[]{rothlisberger1972,clarke2003,werder2013}.
-  Thus, the hydraulic diameter of the subglacial channel, $D_h$, responds to a representative water discharge, $Q_{w}^*$, and a representative gradient of the hydraulic potential, $\Psi^*$ such that:
+To accomplish this, a modified version of  the lumped hydraulics model presented in \citeA{clarke1996} and \citeA{werder2010} is used.
+
+Here, it is assumed that the water is transported through a subglacial channel flowing down the glacier; \cite<Figure~\ref{fig:cartoon}; >{rothlisberger1972}, and that the channel  size responds to frictional heating from water flow and creep closure by ice, using the Darcy-Weisbach formulation for water-flow through a pipe  \cite<e.g.>[]{rothlisberger1972,clarke2003}.
+In the formulation here does not consider englacial storage of water, and thus that changes in hydraulic head at the top of the glacier are negligable in the evolution of the channel size. 
+The evolution of subglacial channel size $S_g$ is given as
 \begin{linenomath*}
   \begin{equation}
-      \label{eq:DW}
-      D_h = \big(s\, f_r\,\rho_w\, \frac{Q_w^{*\,2}}{\Psi^*}\big)^{\frac{1}{5}}~~.
+    \label{eq:dS_dt}
+    \frac{\partial S_g}{\partial t} = C_2 \frac{Q \Delta h}{l} - C_2 (h_{o}-\frac{\Delta h}{2})^n\,S_g,
   \end{equation}
 \end{linenomath*}
-  % 
-  The representative hydraulic gradient is based upon the pressurized flow in Equation~\ref{eq:DW},
-  \begin{equation}
-    \label{eq:psi}
-    \Psi^*= f_f \,  \rho_i \, g (\frac{\partial  z_{sg}}{\partial x} - \frac{\partial z_c}{\partial x}) +  \rho_i \, g \, \frac{\partial z_c}{\partial x},
-  \end{equation}
-  % 
-  \noindent
-  where, $f_f$ is the flotation fraction, $\rho_i$ is density of ice $\frac{\partial z_{sg}}{\partial x}$ is the surface slope of the glacier and $\frac{\partial z_c}{\partial x}$ is the glacier bed slope.
-
-  In Equation~\ref{eq:DW}, $f_r$ is the Darcy-Weisbach friction factor and $\rho_w$ is the density of water (Table \ref{table:vpm}).
+\noindent $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (Table \ref{table:vpm}), $l$ is the length of subglacial channel, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change from the glacier terminus to its top, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the ice overburden pressure, and $n$ is Glen's n.
+The first term on the right side of the equation represents the opening of the channel through frictional heating, while the second term represents the creep closure of the channel from ice deformation.
+
+Following the Darcy-Weisbach, head drop $\Delta h$ is,
+\begin{linenomath*}
+  \begin{equation}
+    \label{eq:dh}
+    \Delta h \,  = l \,\frac{1}{g}\,s\,f_r\,\frac{Q^2}{D_h^5}
+  \end{equation}
+\end{linenomath*}
+\noindent where $f_r$ is a friction factor, $D_h$ is the hydraulic diameter, and $l$ is the length overwhich head change occurs.
 $s$ is a factor accounting for channel geometry \cite{hooke1990}, calculated as:
@@
 smaller values of $\beta$ result in shallow, wide channels.
-  The width of the flat channel floor $w_c$, given angle $beta$ is 
-  \begin{equation}
-    \label{eq:dh2wc}
-    w_{cg} = 2  \sin \frac{\beta}{2} \sqrt{\frac{2\, S}{\beta -\sin \beta}},
-  \end{equation}
-  % 
-  where $S$ is the cross-sectional area of the channel given by
-  \begin{equation}
-    \label{eq:dh2S}
-    S_g =  \frac{D_h^2}{2}~ \frac{\Big(\frac{\beta}{2} \,+ \, \sin \frac{\beta}{2}\Big)^2  }{\beta\,-\,\sin \beta}.
-  \end{equation}
-
-  The shear stress, $\tau$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
+
+With knowledge of cross sectional area $S$, the shear stress, $\tau$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
 \begin{equation}
@@
 % 
-  where  the water velocity $v_g$ is $v_g = \frac{Q_w}{S_g}$.
-  Here, $Q_w$ represents the instantaneous water discharge, rather than $Q_w^*$, which is used to evaluate the channel geometry.
-  With this formulation, the processes in an R-channel, through which water flows subglacially, can be represented algebraically \cite{rothlisberger1972,delaney2019}.
+where  the water velocity $v_g = \frac{Q_w}{S_g}$.
+
+The width of the flat channel floor $w_{cg}$, given angle $beta$ is 
+\begin{equation}
+  \label{eq:dh2wc}
+  w_{cg} = 2  \sin \frac{\beta}{2} \sqrt{\frac{2\, S}{\beta -\sin \beta}}.
+\end{equation}
 
@@
 where $k$ is a constant and $e$ is an exponent commonly equal to $\frac{1}{2}$ \cite{leopold1953}.
-  Cross-sectional area, $ S_f$, is 
-  \begin{equation}
-    \label{eq:Sf}
-    S_f = \frac{Q_w}{v_f},
-  \end{equation}
-  % 
-  where $v_f$ is water velocity, given by rearranging Equation \ref{eq:tau} as
+Water velocity, $v_f$, is given by rearranging Equation \ref{eq:tau} as
 \begin{equation}
@@
 As in Section~\ref{sect:sub_mode}, the width integrated shear stress is $\tau_{tf}=w_{cf}\,\tau_f$.
-  Hydraulic diameter of the channel is given by $D_{hf} = \frac{4\,w_{cf}\,d}{2\,d+w_{cf}}$, with flow depth $d$ determined from knowledge of $S_f$ and $w_{cf}$ in a rectangular channel.
 
@@
 
-  The formulation of the glacial system (Section~\ref{sect:sub_mode}) requires inputs of the hydraulic gradient (Equation \ref{eq:psi}; $\frac{\partial z_{sg}}{\partial x}$, $\frac{\partial z_{c}}{\partial x}$), water discharge, $Q_w$ representative water discharge, $Q_w^*$,  Hooke angle (Equation~\ref{eq:Hf}; $\beta$), and friction factor (Equation~\ref{eq:DW}; $f_r$).
+The parameterizations above are applied to hydrological records from the Fieschergletscher in the Swiss Alps (info) and the Leverett glacier on the Greenland Ice Sheet (info).
+The hydrographs applied to the subglacial and subaerial parameterizations represent the subglacial flow of water as it leaves the glacier and moves into a subaerial flow in front of the glacier, as water moves through the catchment.
+Outputs of the parameterizations are meant to represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
+
+The formulation of the glacial system (Section~\ref{sect:sub_mode}) requires inputs of the ice thickness ($h_o$, Equation\,\ref{eq:dS_dt}), water discharge, $Q_w$,  Hooke angle (Equation~\ref{eq:Hf}; $\beta$), and friction factor (Equations~\ref{eq:dh}~and~\ref{eq:tau}; $f_r$).
+Values of these parameters are given in Table~\ref{table:vpm}.
 
@@
 
-  The Reynolds number $R_e$, which gives an evaluation of the degree of turbulence of the water, is given as 
-  \begin{equation}
-    \label{eq:re}
-    R_e\,=\, v \,\frac{D_h}{\nu},
-  \end{equation}
-  % 
-  \noindent where $\nu$ is the kinematic viscosity of water, and  $v$ is velocity of water, $v_g$ or $v_f$.
-
-  The variance $ \chi$ of a variable $\Gamma$ is given by 
-  \begin{equation}
-    \label{eq:var}
-    \chi(\Gamma) \,=\, 1 - \frac{\mathrm{min}(\Gamma)}{\mathrm{max}(\Gamma)}\,\,.
-  \end{equation}
-  % 
-  \noindent Below we examine variables, $\Gamma$, as water discharge ($Q_w$) water velocity ($v_g$, $v_f$), shear stress ($\tau_g$, $\tau_f$), and integrated shear stress ($\tau_{tg}$, $\tau_{tf}$) with the purpose of evaluating total shear stress across the glacier bed.
+The parameterizations are first applied to a reference test case for each glacier that assumes a semi-circular subglacial channel ($\beta=\pi$) and friction factors $f_r$ and $f_p$ are tuned so that the model reproduces reasonable velocities for both  the Greenland Ice Sheet and alpine glaciers \cite<$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$>{werder2010b,chandler2013}.
+Water velocity, shear-stress and width-integrated shear-stress in the subglacial and subaerial systems are presented below.
+
+To characterize the variability in sediment discharge capacity in subglacial compared to subaerial systems further, the parameterizations are applied to range of different channel geometry factors ($\beta$ and $k$) and friction factors $f_r$ and $f_p$.
+Model runs with parameter combinations  are culled, if their maximum subglacial water velocity exceeds $2\,$\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity exceeds $1\,$\unit{m}$^3$\,\unit{s}$^{-1}$.
+The standard deviation of water velocity, shear-stress and width-integrated shear-stress across the season long hydrographs are presented below.
 
 \section{Results}
-  Initially, both parameterizations are forced with a water discharge ($Q_w$) varying between $8$ and $16$ \,\unit{m}$^{3}$\,\unit{s}$^{-1}$, to represent a range of sediment transport conditions, for instance over diurnal fluctuation in water discharge.
-The subaerial channel width factor $k$ is $3$\,\unit{m}$^{2}$\,\unit{s}$^{-1}$, and the Hooke angle $\beta$ is $\frac{\pi}{2}$.
-  Values of Hooke angle $\beta$ and subglacial friction factor $f_r$ are chosen so that they produce reasonable values of water velocity ($0.6- 1.2$\,\unit{m}\,\unit{s}$^{-1}$) in accordance with measurements made through dye-tracing experiments to examine subglacial hydraulic conditions \cite<Section~\ref{sect:sub_mode}, Figure~\ref{fig:model_outs}; e.g.>{werder2010}.
-  The subglacial representative water discharge ($Q_w^*$; Equation~\ref{eq:DW}) of $12$\,\unit{m}$^{3}$\,\unit{s}$^{-1}$ is applied to establish the hydraulic diameter and shape of the channel.
+
+% there is more variability across the season subglacial and subaerial cases. peaks in subaerial cases lag peaks in subglacial cases. (Figure 2)
+
+% For a fixed channel geometry, there is a single shear stress for a water discharge.
+In the subglacial case, peak water discharge that is 1.5 times the mean velocity.
+In Fieschercase, peak water discharge does not result in the greatest of any variable.
+In Leverett case the mean total shear stress can occur from 100 m3/s to 200 m3/s.
+The mean velocity and shear stress can occur across effectively all water discharges below 250 m3/s
+
+% across a wide range of channel geometries
+
+Following the parameterizations defined in Sections~\ref{sect:sub_mode}~and~\ref{sect:fluv}, the water velocity, shear-stress and total shear-stress exhibit vastly different seasonal evolutions and peaks of water velocity and shear stress (representing the potential for sediment mobilization) and total shear stress (representing the total sediment transport capacity; Figure~\ref{fig:model_outs}).
+Furthermore, these variables respond differently to water discharge in the subglacial and subaerial systems (Figure~\ref{fig:Qw_vari}).
+For each water discharge value in the subaerial system, there is a single water velocity, shear-stress and total shear-stress (Figure~\ref{fig:Qw_vari}).
+In the subglacial system, water velocity, shear-stress and total shear-stress vary substantially for a given water discharge(Figure~\ref{fig:Qw_vari}).
+For instance, over $15$ minute time periods at Fieschergletscher, the range of water velocities and shear-stress spans from those below the mean to those over $1.5$ times the mean.
+Substantially high values of water velocity and shear-stress can occur from minimal water discharge at $2.5$\,\unit{m}$^3$\,\unit{s}$^{-1}$ to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
+At Leverett glacier, reduced diurnal variability in the hydrograph  (Figure~\ref{fig:model_outs} e) results in smaller variations in water velocity and shear stress compared to Fieschergletscher (Figure~\ref{fig:Qw_vari}).
+However, mean values of water velocity can occur at water discharges between roughly $50$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $225$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, effectively spanning the observed water discharge.
+
+When accounting for the evolving channel width of the subglacial conduit, total shear-stress generally increases with water discharge.
+Yet the total shear-stress across the can very substantially, with the highest values in total shear-stress occurring at water discharge values ranging from roughly $11$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
+The variance is less at the Leverett glacier case, yet a range of total shear-stress can occur at a given discharge(Figure~\ref{fig:Qw_vari} c, e).
+
+Over the course of the season, large peaks in water velocity and shear stress, along with total shear stress, occur in response to increases in water discharge.
+In the subglacial system, these peaks generally occur over the period when water discharge increases at the fastest rate (Figure~\ref{fig:model_outs}).
+Conversely, in the subaerial system peaks in these variables occur when water discharge is at its highest value.
+As a result, peaks in the subglacial variables occur prior to the peaks in the subglacial variables, representing lags sediment mobilization and  transport capacity as sediment is transferred from on system to the other.
+
+The water velocity and shear stress remain higher across the different friction factors and channel geometries examined in the ensemble of runs across averaging periods.
+However, the Leverett case, some parameter combinations the result in greater variability in subaerial system compared to the subaerial channels.
+The effect of channel evolving width largely counteracts the variability in water velocity, especially in the alpine glacier case.
+Here the large diurnal cycles in the hydrograph can cause the maximum and minimum amounts of variability in subaerial total shear-stress to be greater than those in the subglacial system.
+At the Leverett glacier, with reduced diurnal variability in water discharge (Figure\,\ref{fig:model_outs}), the variability in total shear-stress remains higher in the subglacial system compared to the subaerial one.
+However, some variability in subaerial parameter combinations exceeds those subglacial ones.
+Note however, that this occurs across the total of xxx parameter combinations examined.
 
@@
     \includegraphics[width=0.8\linewidth]{model_outputs.pdf}
-      \caption{Response of channel width (a), hydraulic diameter (b),  Reynolds number  (c),  water velocity (d), shear-stress (e), and width-integrated shear-stress (f)  to different water discharge values.
-The glacial system is in blue and the subaerial system is in green.
-Horizontal bars in d, e, and f represent the range of values in the  output.
-Gray space represents water discharge below the mean value.
+    \caption{Model outputs as in response to hydrograph (a and e) from Fieschergletscher (a-d) and Leverett glacier (e-h).
+Gray line in a and e represent subglacial channel cross sectional area.
+Below this black lines represent model outputs of velocity, shear-stress and width-integrated shear-stress from the subglacial system, while grey lines represent the subaerial system.
 }
@@
 
-  Across the hydrological forcing, variable water discharge results in a substantial increase in the range of water velocities in the glacial system compared to the subaerial system (Figure~\ref{fig:model_outs}).
-  Water velocity in the glacier system varies by  $0.5$ (Equation~\ref{eq:var}) over the discharge values, in the subaerial system the water velocity only varies by $0.11$ (Figure~\ref{fig:model_outs}\,e).
-  Shear stress at the glacier bed varies by $0.75$ across the range of water discharge values, whereas shear stress across the subaerial system varies by roughly $0.21$ (Figure~\ref{fig:model_outs}\,f).
-The increased variability in shear stress relative to water velocity between the two systems occurs as the water velocity is raised to the power of $2$ in Equation~\ref{eq:tau}.
-
-  In the subaerial systems, changing channel geometry accommodates evolving water discharge  (Equation~\ref{eq:wcf}), whereas in the subglacial case, increased water flux can only result in changing water velocity (Equations~\ref{eq:v}~ and ~\ref{eq:DW}).
-  As a result, in the subaerial system, variability in the shear stress $\tau_f$ and width-integrated shear stress $\tau_{tf}$ increases from $0.21$ to $0.44$.
-Variability in width-integrated shear stress  in the glacial system $\tau_{tg}$ is $0.83$ (Figure~\ref{fig:range}).
-  The reduced variability in width-integrated shear stress results from the additional width across the subaerial channel that accommodates some variations in water discharge.
-  Indeed, the relative  difference in width-integrated shear stress is less between the subaerial and glacier cases, compared to the water velocity and shear stress outputs (Figure~\ref{fig:model_outs}\,e--g).
-
-  Next, the parameterizations are run $25,000$ times with randomly selected channel geometries ($k=(2,6)$, $\beta=(\frac{\pi}{10},\pi)$), friction factors ($f_i=(0.001,20)$, $f_p=(0.001,20)$) and water discharges (maximum and minimum values selected between $5$ to $500$ \,\mmms).
-$Q_w^*$ in Equation ~\ref{eq:DW} is determined from the mean in the water discharge range.
-Parameter combinations are only selected if their maximum and minimum velocities for both the subglacial and subaerial systems are between $0.2$ and $3$\,\unit{m}\,\unit{s}$^{-1}$, resulting in roughly $10,600$ adequate parameter combinations.
-
 \begin{center}
   \begin{figure}[H]
-      \includegraphics[width=0.8\linewidth]{multi_run.pdf}
-      \caption{Ratio of range of water velocity, shear-stress, and width integrated shear-stress for glacial and subaerial systems in response to mean water discharge (a, b, c), and  variability in water discharge (d, e, f).
-        Range  $\chi$ for each variable is given in Equation~\ref{eq:var}. Values greater than $1$ shown in white space are parameter combinations that vary more in a subglacial  system.
-        Histograms (g-i) show the frequency of relative variations across variables in the parameterization.}
-      \label{fig:range}
+    \includegraphics[width=0.8\linewidth]{Qw_vari.png}
+    \caption{Relationship between water discharge and velocity, shear stress and width-integrated shear stress for the Fieschergletscher hydrograph (a-c) and Leverett glacier (d-f).
+Variables have been normalized to by the mean values for better comparison amongst the model runs.}
+    \label{fig:Qw_vari}
   \end{figure}
@@
 
-  Across the accepted model runs, variability in water velocity and shear stress remains substantially higher across the subglacial model compared to subaerial ones (Figure~\ref{fig:range}).
-  Water velocity, shear stress, and width-integrated shear stress all have a variability ratio ($\frac{\chi(\Gamma_{g})}{\chi(\Gamma_f)}$) greater than $1$ in the subglacial  system compared to subaerial one (Figure~S \ref{fig:gammas}).
-  Runs with the smallest variability in discharge ($\chi(Q_w)$) have the greatest variability in water velocity, shear stress, and width-integrated shear stress in subglacial systems compared to subaerial ones (Figure~\ref{fig:range}\,d--f and S \ref{fig:gammas}).
-
-  The minimum ratio of  water velocity, $\frac{\chi(v_{g})}{\chi(v_{f})}$, and shear-stress, $\frac{\chi(\tau_{g})}{\chi(\tau_{f})}$, is greater than $1.5$.
-  These results show that increased variability occurs across a wide range of subaerial and subglacial channel configurations (Figure~\ref{fig:range} g, h).
-The width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, spans between $1.02$ and $2.4$, a smaller amount of relative variability compared to  water velocity $\frac{\chi(v_{g})}{\chi(v_{f})}$ or shear stress $\frac{\chi(\tau_{g})}{\chi(\tau_{f})}$ (Figure~\ref{fig:range}).
-Consistent with the individual case (Figure ~\ref{fig:model_outs}), the subaerial systems' reduce variability in width-integrated shear stress occurs  as the evolving channel width accommodates some of the reduced shear stress variability in subaerial systems (Figure~\ref{fig:range} i).
-  Importantly though, no runs experienced relative variability in width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, less than $1$, showing that the subglacial has consistently high variability in width integrated shear stress compared to subaerial systems across the range of tested situations.
+\begin{center}
+  \begin{figure}[H]
+    \includegraphics[width=0.8\linewidth]{multi_run.png}
+    \caption{} 
+    \label{fig:multi_run}
+  \end{figure}
+\end{center}
 
 \section{Discussion}
-
-  \subsection{Geomorphic Implications}
-  \label{sect:GI}
+\ian{add cautionary note as this is an example}
+
+\ian{Zwally and tedstone effect with increased discharge... non-linear response of increasing water discharge. }
+
 The relationship between water discharge variability and relative variability of the water velocity, shear stress, and width-integrated shear stress results from the fixed geometry in the glacier system and variable geometry in the subaerial system as they respond to changes in water discharge (Equations~\ref{eq:tau_t} and \ref{eq:v}).
@@
 In the total sediment transport relationship \citeA{engelund1967}, the shear stress is scaled to $\frac{5}{2}$ power.
-  In both cases, the exponent greater than $1$ of shear stress in the sediment transport relationship magnifies the variability in sediment transport beyond the highly variable shear stress and width-integrated shear stress described above in subglacial environments (Figure~\ref{fig:model_outs}\, b and d; Figure~S~\ref{fig:gammas}).
-
-  Both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress of water across the channel bed is homogeneous \cite<Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}; >{yager2018}.
-  The subaerial parameterization assumes that sediment is transported across all channel widths, as opposed to large flows such as floods when much sediment transport occurs \cite{wolman1960}.
-  As a result, the variability in the subaerial system across the range of discharge values could be underestimated in the parameterization (Figure~\ref{fig:range}).
-
-  \subsubsection{Sediment transport below glaciers and at their margins}
+In both cases, the exponent greater than $1$ of shear stress in the sediment transport relationship magnifies the variability in sediment transport beyond the highly variable shear stress and width-integrated shear stress described above in subglacial environments (Figure~\ian{xxx}).
+
+% Both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress of water across the channel bed is homogeneous \cite<Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}; e.g.
+>{yager2018}.
+% \ian{Phillips and Jeromack 2016... stabilization here.}
+% In the subaerial parameterization, this means that sediment is transported across all channel widths, as opposed to large flows such as floods when much sediment transport occurs \cite{wolman1960}.
+% As a result, the  the subaerial systems variability in sediment transport capacity across the range of discharge values could be underestimated in the parameterization here (Figure~~\ian{xxx}).
+
+\subsection{Sediment transport below glaciers and at their margins}
 
@@
 Yet, in the subglacial system reaching the threshold of motion for sediment frequently, across a range of discharges,  may result in additional sediment transport.
-  The variable sediment transport capacity may result in a tendency for glaciers to evacuate sediment and transition to a supply-limited regime whereby sediment discharge scales with glacier sliding  \cite<e.g.>[]{herman2015,koppes2015}.
+The variable sediment transport capacity may result in a tendency for glaciers to evacuate sediment and transition to a supply-limited regime whereby sediment discharge scales with glacier sliding  \cite<e.g.>[]{herman2015}.
 Sediment exhaustion by this process may also explain the dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, as opposed to glacier melt \cite{overeem2017}.
@@
 
-  Increased variability in sediment transport may also impact sediment export from the subglacial environment into proglacial areas \cite<e.g.>[]{delaney2017,perolo2018}.
+Increased variability in sediment transport may also impact sediment export from the subglacial environment into proglacial areas \cite<e.g.>[]{perolo2018}.
 Here, velocities at the highest water discharges could generally decrease (Figure~\ref{fig:model_outs}, blue and green horizontal bars), in the transition of pressurized subglacial flow to subaerial open-channel flow as the water leaves the glacier.
@@
 
-  In large catchments and further downstream of proglacial margins, water discharge is generally less variable than at the tops of catchments, where glaciers lie in alpine settings \cite<c.f.>{swift2005,riihimaki2005,costa2017,vanas2017}.
+In large catchments and further downstream of proglacial margins, water discharge is generally less variable than at the tops of catchments, where glaciers lie in
+alpine settings \cite<c.f.>{swift2005,riihimaki2005,costa2017,vanas2017}.
 Here, external hydrological factors further downstream of glaciers could result in smaller variations in sediment transport capacity in many subaerial systems compared to subglacial ones.
 
-  The parameterization here shows that the smallest variability in water discharge has the largest difference in relative variability in sediment transport metrics (Figure~\ref{fig:range}\, d,\,e,\,f).
+The parameterization here shows that the smallest variability in water discharge has the largest difference in relative variability in sediment transport metrics (Figure~~\ian{xxx}).
 At locations such as the Greenland Ice Sheet, smaller diurnal variations in water discharge compared to many alpine catchments \cite<c.f.>{swift2005,riihimaki2005,vanas2017,hasholt2018}.
 Therefore, the variability in sediment transport capacity in the transition from the ice sheet to the proglacial river could be more dramatic in Greenland than in alpine systems as water leaves the glacier and moves downstream.
-  Thus the disparity in sediment transport capacity between the subglacial and subaerial systems could be most pronounced in larger catchments, potentially resulting in temporally mismatched periods between when glaciers introduce sediment to these river systems  and when the river mobilizes it (Figures~\ref{fig:model_outs}~ and~\ref{fig:range}).
-
-  \subsubsection{Records of sediment discharge  from glaciers}
+Thus the disparity in sediment transport capacity between the subglacial and subaerial systems could be most pronounced in larger catchments, potentially resulting in temporally mismatched periods between when glaciers introduce sediment to these river systems  and when the river mobilizes it (Figures~\ref{fig:model_outs}).
+
+\subsection{Records of sediment discharge  from glaciers}
+
+\ian{include comments about the limits of hysterisis analysis from glacial systems}
 
@@
 As a result, variations in sediment transport capacity could be responsible for the fluctuations in sediment discharge (``flushings'') from glaciers that are not attributed to variations in water discharge \cite<e.g.>[]{richards2003,swift2021}.
-  Aside from the sporadic nature of these events, the observed flushings could result from increases in water velocity resulting from reduced channel size, not water discharge, which increase sediment transport capacity.
+Aside from the sporadic nature of these events, the observed ``flushings'' could result from increases in water velocity resulting from reduced channel size, not water discharge, which increase sediment transport capacity.
 
@@
 For instance, sediment transport capacity from a glacier reacts to the ice thickness, controlling the closure of the channel, and the surface slope of the glacier, controlling water velocity \cite<Section~\ref{sect:sub_mode}; >{rothlisberger1972,shreve1972,delaney2022,stevens2022}, in addition to water discharge.
-  Conversely, sediment transport in most subaerial systems increases with water discharge and bed slope, and the bed slope probably remains relatively stable over yearly to century-scale time periods \cite<Section~\ref{sect:fluv}; e.g.>{muller1968,whipple1999,wong2006,wickert2019}.
+Conversely, sediment transport in most subaerial systems increases with water discharge and hydraulic gradient, and the hydraulic gradient probably remains relatively stable over yearly to century-scale time periods \cite<Section~\ref{sect:fluv}; e.g.>{muller1968,whipple1999,wong2006,wickert2019}.
 
@@
 
-  This suggests that a larger  perturbation of water discharge might be needed to significantly alter the sediment discharge capacity from the glaciers compared to subaerial systems, to overcome the noise resulting from increased variability in sediment transport capacity.
-Furthermore, if no sediment is available and the glacier is in a supply-limited case, then the sediment discharge record will also represent additional processes of bedrock erosion from sliding and water pressure variations  \cite{iverson2012,herman2015}.
+This suggests that a larger climatic perturbations in water discharge might be needed to isolate significant changes sediment discharge from glacial systems compared to subaerial on, as the noise resulting from increased variability in sediment transport capacity must be overcome.
+Furthermore, if no sediment is available and the glacier is in a supply-limited case, then the sediment discharge record will also represent additional processes of bedrock erosion from  water pressure variations and sliding  \cite{iverson2012,herman2015}, which are also complex.
 
@@
 
+The differences in subglacial and subaerial channels' response to variations in water discharge should be considered when examining sediment transport processes in glacierized catchments.
+
+\section*{Open Research}
+\noindent
+The julia code used herein is included in the supplementary material.
+
+\acknowledgments
+I was funded by SNSF Project No. PZ00P2\_202024.
+
+%% ------------------------------------------------------------------------ %%
+%% References and Citations
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+% 
+% \bibliography{<name of your .bib file>} don't specify the file extension
+% 
+% don't specify bibliographystyleiz
+
+% In the References section, cite the data/software described in the Availability Statement (this includes primary and processed data used for your research).
+For details on data/software citation as well as examples, see the Data & Software Citation section of the Data & Software for Authors guidance
+% https://www.agu.org/Publish-with-AGU/Publish/Author-Resources/Data-and-Software-for-Authors#citation
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\bibliography{Paperlib.bib}
+
+% \section{Supplement}
+
+% \renewcommand{\figurename}{Figure S}
+% \setcounter{figure}{0}
+
+% \begin{center}
+%   \begin{figure}[H]
+%     \includegraphics[width=0.8\linewidth]{multi_run_vars.pdf}
+%     \caption{Variability $\chi$ in velocity ($v_g$, $v_f$; a) shear stress ($\tau_g$, $\tau_f$; b) and width-integrated shear stress ($\tau_{tg}$, $\tau_{tf}$; c)  in glacial and subaerial systems.
+Values are determined from examining the same parameter space in Figure~\ref{fig:range}.
+}
+%     \label{fig:gammas}
+%   \end{figure}
+% \end{center}
+
+\end{document}
+
 These different characteristics between glacial and subaerial channels show that records of sediment transport downstream of glaciers represent both subglacial and subaerial processes together.
@@
 In addition, the subglacial parameterization shows that sediment transport capacity in glacier systems responds to a large number of factors such as channel size, ice thickness, and glacier surface slope, which may react to climate forcing differently than water discharge alone.
-  The differences in subglacial and subaerial channels' response to changing hydrology should be considered when examining sediment transport processes in glacierized catchments.
-
-  \section*{Open Research}
-  \noindent
-  The julia code used herein is included in the supplementary material.
-
-\acknowledgments
-I was funded by SNSF Project No. PZ00P2\_202024.
-
-%% ------------------------------------------------------------------------ %%
-%% References and Citations
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-% 
-% \bibliography{<name of your .bib file>} don't specify the file extension
-% 
-% don't specify bibliographystyleiz
-
-% In the References section, cite the data/software described in the Availability Statement (this includes primary and processed data used for your research).
-For details on data/software citation as well as examples, see the Data & Software Citation section of the Data & Software for Authors guidance
-% https://www.agu.org/Publish-with-AGU/Publish/Author-Resources/Data-and-Software-for-Authors#citation
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{thebibliography}{}
-
-\bibitem [\protect \citeauthoryear {%
-Andrews%
-\ \protect \BOthers {.}}{%
-Andrews%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2014}}%
-}]{%
-andrews2014}
-\APACinsertmetastar {%
-andrews2014}%
-\begin{APACrefauthors}%
-Andrews, L\BPBI C.%
-, Catania, G\BPBI A.%
-, Hoffman, M\BPBI J.%
-, Gulley, J\BPBI D.%
-, L{\"u}thi, M\BPBI P.%
-, Ryser, C.%
-\BDBL {}Neumann, T\BPBI A.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2014}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Direct observations of evolving subglacial drainage
-  beneath the {G}reenland {I}ce {S}heet} {Direct observations of evolving
-  subglacial drainage beneath the {G}reenland {I}ce {S}heet}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Nature}{514}{7520}{80}.
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Beaud%
-, Flowers%
-\BCBL {}\ \BBA {} Venditti%
-}{%
-Beaud%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2018}}%
-}]{%
-beaud2018}
-\APACinsertmetastar {%
-beaud2018}%
-\begin{APACrefauthors}%
-Beaud, F.%
-, Flowers, G.%
-\BCBL {}\ \BBA {} Venditti, J\BPBI G.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2018}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Modeling sediment transport in ice-walled subglacial
-  channels and its implications for esker formation and pro-glacial sediment
-  yields} {Modeling sediment transport in ice-walled subglacial channels and
-  its implications for esker formation and pro-glacial sediment yields}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Journal of Geophysical Research: Earth
-  Surface}{123}{}{1--56}.
-\newblock
-\begin{APACrefDOI} \doi{10.1029/2018JF004779} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Bendixen%
-\ \protect \BOthers {.}}{%
-Bendixen%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2017}}%
-}]{%
-bendixen2017}
-\APACinsertmetastar {%
-bendixen2017}%
-\begin{APACrefauthors}%
-Bendixen, M.%
-, Iversen, L\BPBI L.%
-, Bj{\o}rk, A\BPBI A.%
-, Elberling, B.%
-, Westergaard-Nielsen, A.%
-, Overeem, I.%
-\BDBL {}others%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2017}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Delta progradation in {G}reenland driven by increasing
-  glacial mass loss} {Delta progradation in {G}reenland driven by increasing
-  glacial mass loss}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Nature}{550}{7674}{101}.
-\newblock
-\begin{APACrefDOI} \doi{10.1038/nature23873} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Castelltort%
-\ \BBA {} {Van Den Driessche}%
-}{%
-Castelltort%
-\ \BBA {} {Van Den Driessche}%
-}{%
-{\protect \APACyear {2003}}%
-}]{%
-castletort2003}
-\APACinsertmetastar {%
-castletort2003}%
-\begin{APACrefauthors}%
-Castelltort, S.%
-\BCBT {}\ \BBA {} {Van Den Driessche}, J.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2003}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {{How plausible are high-frequency sediment supply-driven
-  cycles in the stratigraphic record?}} {{How plausible are high-frequency
-  sediment supply-driven cycles in the stratigraphic record?}}{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Sedimentary Geology}{157}{1}{3-13}.
-\newblock
-\begin{APACrefDOI} \doi{10.1016/S0037-0738(03)00066-6} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Clarke%
-}{%
-Clarke%
-}{%
-{\protect \APACyear {2003}}%
-}]{%
-clarke2003}
-\APACinsertmetastar {%
-clarke2003}%
-\begin{APACrefauthors}%
-Clarke, G\BPBI K\BPBI C.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2003}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Hydraulics of subglacial outburst floods: new insights
-  from the {S}pring--{H}utter formulation} {Hydraulics of subglacial outburst
-  floods: new insights from the {S}pring--{H}utter formulation}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Journal of Glaciology}{49}{165}{299--313}.
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Costa%
-\ \protect \BOthers {.}}{%
-Costa%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2018}}%
-}]{%
-costa2017}
-\APACinsertmetastar {%
-costa2017}%
-\begin{APACrefauthors}%
-Costa, A.%
-, Molnar, P.%
-, Stutenbecker, L.%
-, Bakker, M.%
-, Silva, T\BPBI A.%
-, Schlunegger, F.%
-\BDBL {}Girardclos, S.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2018}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Temperature signal in suspended sediment export from an
-  {A}lpine catchment} {Temperature signal in suspended sediment export from an
-  {A}lpine catchment}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Hydrology and Earth System Sciences}{22}{1}{509-528}.
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Cowan%
-, Powell%
-\BCBL {}\ \BBA {} Smith%
-}{%
-Cowan%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {1988}}%
-}]{%
-cowan1988}
-\APACinsertmetastar {%
-cowan1988}%
-\begin{APACrefauthors}%
-Cowan, E.%
-, Powell, R.%
-\BCBL {}\ \BBA {} Smith, N.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{1988}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {{Rainstorm-induced event sedimentation at the tidewater
-  front of a temperate glacier}} {{Rainstorm-induced event sedimentation at the
-  tidewater front of a temperate glacier}}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Geology}{16}{5}{409--412}.
-\newblock
-\begin{APACrefDOI}
-  \doi{https://doi.org/10.1130/0091-7613(1988)016%3C0409:RIESAT%3E2.3.CO;2}
-  \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Creyts%
-, Clarke%
-\BCBL {}\ \BBA {} Church%
-}{%
-Creyts%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2013}}%
-}]{%
-creyts2013}
-\APACinsertmetastar {%
-creyts2013}%
-\begin{APACrefauthors}%
-Creyts, T\BPBI T.%
-, Clarke, G\BPBI K\BPBI C.%
-\BCBL {}\ \BBA {} Church, M.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2013}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Evolution of subglacial overdeepenings in response to
-  sediment redistribution and glaciohydraulic supercooling} {Evolution of
-  subglacial overdeepenings in response to sediment redistribution and
-  glaciohydraulic supercooling}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Journal of Geophysical Research: Earth
-  Surface}{118}{2}{423--446}.
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Delaney%
-\ \BBA {} Adhikari%
-}{%
-Delaney%
-\ \BBA {} Adhikari%
-}{%
-{\protect \APACyear {2020}}%
-}]{%
-delaney2020}
-\APACinsertmetastar {%
-delaney2020}%
-\begin{APACrefauthors}%
-Delaney, I.%
-\BCBT {}\ \BBA {} Adhikari, S.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2020}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Increased subglacial sediment discharge during century
-  scale glacier retreat: consideration of ice dynamics, glacial erosion and
-  fluvial sediment transport} {Increased subglacial sediment discharge during
-  century scale glacier retreat: consideration of ice dynamics, glacial erosion
-  and fluvial sediment transport}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Geophyiscal Research Letters}{}{}{e2019GL085672}.
-\newblock
-\begin{APACrefDOI} \doi{10.1029/2019GL085672} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Delaney%
-\ \BBA {} Anderson%
-}{%
-Delaney%
-\ \BBA {} Anderson%
-}{%
-{\protect \APACyear {2022}}%
-}]{%
-delaney2022}
-\APACinsertmetastar {%
-delaney2022}%
-\begin{APACrefauthors}%
-Delaney, I.%
-\BCBT {}\ \BBA {} Anderson, L\BPBI S.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2022}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Debris Cover Limits Subglacial Erosion and Promotes Till
-  Accumulation} {Debris cover limits subglacial erosion and promotes till
-  accumulation}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Geophysical Research Letters}{49}{16}{e2022GL099049}.
-\newblock
-\APACrefnote{e2022GL099049 2022GL099049}
-\newblock
-\begin{APACrefDOI} \doi{10.1029/2022GL099049} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Delaney%
-, Bauder%
-, Huss%
-\BCBL {}\ \BBA {} Weidmann%
-}{%
-Delaney%
-, Bauder%
-, Huss%
-\BCBL {}\ \BBA {} Weidmann%
-}{%
-{\protect \APACyear {2018}}%
-}]{%
-delaney2017}
-\APACinsertmetastar {%
-delaney2017}%
-\begin{APACrefauthors}%
-Delaney, I.%
-, Bauder, A.%
-, Huss, M.%
-\BCBL {}\ \BBA {} Weidmann, Y.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2018}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Proglacial erosion rates and processes in a glacierized
-  catchment in the {S}wiss {A}lps} {Proglacial erosion rates and processes in a
-  glacierized catchment in the {S}wiss {A}lps}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Earth Surface Processes and Landfroms}{43}{4}{765-778}.
-\newblock
-\begin{APACrefDOI} \doi{10.1002/esp.4239} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Delaney%
-, Bauder%
-, Werder%
-\BCBL {}\ \BBA {} Farinotti%
-}{%
-Delaney%
-, Bauder%
-, Werder%
-\BCBL {}\ \BBA {} Farinotti%
-}{%
-{\protect \APACyear {2018}}%
-}]{%
-delaney2018}
-\APACinsertmetastar {%
-delaney2018}%
-\begin{APACrefauthors}%
-Delaney, I.%
-, Bauder, A.%
-, Werder, M\BPBI A.%
-\BCBL {}\ \BBA {} Farinotti, D.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2018}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Regional and annual variability in subglacial sediment
-  transport by water for two glaciers in the {S}wiss {A}lps} {Regional and
-  annual variability in subglacial sediment transport by water for two glaciers
-  in the {S}wiss {A}lps}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Frontiers in Earth Science}{}{}{}.
-\newblock
-\begin{APACrefDOI} \doi{10.3389/feart.2018.00175} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Delaney%
-, Werder%
-\BCBL {}\ \BBA {} Farinotti%
-}{%
-Delaney%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2019}}%
-}]{%
-delaney2019}
-\APACinsertmetastar {%
-delaney2019}%
-\begin{APACrefauthors}%
-Delaney, I.%
-, Werder, M.%
-\BCBL {}\ \BBA {} Farinotti, D.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2019}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {{A} {N}umerical {M}odel for {F}luvial {T}ransport of
-  {S}ubglacial {S}ediment} {{A} {N}umerical {M}odel for {F}luvial {T}ransport
-  of {S}ubglacial {S}ediment}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Journal of Geophysical Research: Earth
-  Surface}{124}{8}{2197--2223}.
-\newblock
-\begin{APACrefDOI} \doi{10.1029/2019JF005004} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Egli%
-, Belotti%
-, Ouvry%
-, Irving%
-\BCBL {}\ \BBA {} Lane%
-}{%
-Egli%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2021}}%
-}]{%
-egli2021b}
-\APACinsertmetastar {%
-egli2021b}%
-\begin{APACrefauthors}%
-Egli, P.%
-, Belotti, B.%
-, Ouvry, B.%
-, Irving, J.%
-\BCBL {}\ \BBA {} Lane, S.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2021}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {{Subglacial Channels, Climate Warming, and Increasing
-  Frequency of Alpine Glacier Snout Collapse}} {{Subglacial Channels, Climate
-  Warming, and Increasing Frequency of Alpine Glacier Snout Collapse}}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Geophysical Research Letters}{48}{21}{e2021GL096031}.
-\newblock
-\APACrefnote{e2021GL096031 2021GL096031}
-\newblock
-\begin{APACrefDOI} \doi{10.1029/2021GL096031} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Engelund%
-\ \BBA {} Hansen%
-}{%
-Engelund%
-\ \BBA {} Hansen%
-}{%
-{\protect \APACyear {1967}}%
-}]{%
-engelund1967}
-\APACinsertmetastar {%
-engelund1967}%
-\begin{APACrefauthors}%
-Engelund, F.%
-\BCBT {}\ \BBA {} Hansen, E.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{1967}{}{}.
-\newblock
-\APACrefbtitle {A monograph on sediment transport in alluvial streams} {A
-  monograph on sediment transport in alluvial streams}\
-  \APACbVolEdTR{}{\BTR{}}.
-\newblock
-\APACaddressInstitution{Copenhagen, Denmark}{Technical University of Denmark}.
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Ganti%
-\ \protect \BOthers {.}}{%
-Ganti%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2016}}%
-}]{%
-ganti2016}
-\APACinsertmetastar {%
-ganti2016}%
-\begin{APACrefauthors}%
-Ganti, V.%
-, von Hagke, C.%
-, Scherler, D.%
-, Lamb, M.%
-, Fischer, W.%
-\BCBL {}\ \BBA {} Avouac, J\BHBI P.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2016}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Time scale bias in erosion rates of glaciated
-  landscapes} {Time scale bias in erosion rates of glaciated
-  landscapes}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Science Advances}{2}{10}{}.
-\newblock
-\begin{APACrefDOI} \doi{10.1126/sciadv.1600204} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Gimbert%
-, Tsai%
-, Amundson%
-, Bartholomaus%
-\BCBL {}\ \BBA {} Walter%
-}{%
-Gimbert%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2016}}%
-}]{%
-gimbert2016}
-\APACinsertmetastar {%
-gimbert2016}%
-\begin{APACrefauthors}%
-Gimbert, F.%
-, Tsai, V\BPBI C.%
-, Amundson, J\BPBI M.%
-, Bartholomaus, T\BPBI C.%
-\BCBL {}\ \BBA {} Walter, J\BPBI I.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2016}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Subseasonal changes observed in subglacial channel
-  pressure, size, and sediment transport} {Subseasonal changes observed in
-  subglacial channel pressure, size, and sediment transport}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Geophysical Research Letters}{43}{8}{3786--3794}.
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Hallet%
-}{%
-Hallet%
-}{%
-{\protect \APACyear {1979}}%
-}]{%
-hallet1979}
-\APACinsertmetastar {%
-hallet1979}%
-\begin{APACrefauthors}%
-Hallet, B.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{1979}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {A theoretical model of glacial abrasion} {A theoretical
-  model of glacial abrasion}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Journal of Glaciology}{23}{89}{39--50}.
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Hallet%
-, Hunter%
-\BCBL {}\ \BBA {} Bogen%
-}{%
-Hallet%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {1996}}%
-}]{%
-hallet1996}
-\APACinsertmetastar {%
-hallet1996}%
-\begin{APACrefauthors}%
-Hallet, B.%
-, Hunter, L.%
-\BCBL {}\ \BBA {} Bogen, J.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{1996}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Rates of erosion and sediment evacuation by glaciers:
-  {A} review of field data and their implications} {Rates of erosion and
-  sediment evacuation by glaciers: {A} review of field data and their
-  implications}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Global and Planetary Change}{12}{1}{213--235}.
-\newblock
-\begin{APACrefDOI} \doi{10.1016/0921-8181(95)00021-6} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Hasholt%
-, van As%
-, Mikkelsen%
-, Mernild%
-\BCBL {}\ \BBA {} Yde%
-}{%
-Hasholt%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2018}}%
-}]{%
-hasholt2018}
-\APACinsertmetastar {%
-hasholt2018}%
-\begin{APACrefauthors}%
-Hasholt, B.%
-, van As, D.%
-, Mikkelsen, A.%
-, Mernild, S.%
-\BCBL {}\ \BBA {} Yde, J.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2018}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Observed sediment and solute transport from the
-  {K}angerlussuaq sector of the {G}reenland {I}ce {S}heet (2006–2016)}
-  {Observed sediment and solute transport from the {K}angerlussuaq sector of
-  the {G}reenland {I}ce {S}heet (2006–2016)}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Arctic, Antarctic, and Alpine
-  Research}{50}{1}{S100009}.
-\newblock
-\begin{APACrefDOI} \doi{10.1080/15230430.2018.1433789} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Herman%
-\ \protect \BOthers {.}}{%
-Herman%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2015}}%
-}]{%
-herman2015}
-\APACinsertmetastar {%
-herman2015}%
-\begin{APACrefauthors}%
-Herman, F.%
-, Beyssac, O.%
-, Brughelli, M.%
-, Lane, S.%
-, Leprince, S.%
-, Adatte, T.%
-\BDBL {}Cox, S\BPBI C.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2015}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Erosion by an alpine glacier} {Erosion by an alpine
-  glacier}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Science}{350}{6257}{193--195}.
-\newblock
-\begin{APACrefDOI} \doi{10.1126/science.aab2386} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Hewitt%
-\ \BBA {} Creyts%
-}{%
-Hewitt%
-\ \BBA {} Creyts%
-}{%
-{\protect \APACyear {2019}}%
-}]{%
-hewitt2019}
-\APACinsertmetastar {%
-hewitt2019}%
-\begin{APACrefauthors}%
-Hewitt, I.%
-\BCBT {}\ \BBA {} Creyts, T.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2019}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {A model for the formation of eskers} {A model for the
-  formation of eskers}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Geophysical Research Letters}{46}{12}{6673--6680}.
-\newblock
-\begin{APACrefDOI} \doi{10.1029/2019GL082304} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Hooke%
-, Laumann%
-\BCBL {}\ \BBA {} Kohler%
-}{%
-Hooke%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {1990}}%
-}]{%
-hooke1990}
-\APACinsertmetastar {%
-hooke1990}%
-\begin{APACrefauthors}%
-Hooke, R\BPBI L.%
-, Laumann, T.%
-\BCBL {}\ \BBA {} Kohler, J.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{1990}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Subglacial Water Pressures and the Shape of Subglacial
-  Conduits} {Subglacial water pressures and the shape of subglacial
-  conduits}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Journal of Glaciology}{36}{122}{67--71}.
-\newblock
-\begin{APACrefDOI} \doi{10.3189/S0022143000005566} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Iken%
-\ \BBA {} Bindschadler%
-}{%
-Iken%
-\ \BBA {} Bindschadler%
-}{%
-{\protect \APACyear {1986}}%
-}]{%
-iken1986}
-\APACinsertmetastar {%
-iken1986}%
-\begin{APACrefauthors}%
-Iken, A.%
-\BCBT {}\ \BBA {} Bindschadler, R\BPBI A.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{1986}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Combined measurements of subglacial water pressure and
-  surface velocity of {F}indelengletscher, {S}witzerland: conclusions about
-  drainage system and sliding mechanism} {Combined measurements of subglacial
-  water pressure and surface velocity of {F}indelengletscher, {S}witzerland:
-  conclusions about drainage system and sliding mechanism}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Journal of Glaciology}{32}{110}{101--119}.
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Iverson%
-}{%
-Iverson%
-}{%
-{\protect \APACyear {2012}}%
-}]{%
-iverson2012}
-\APACinsertmetastar {%
-iverson2012}%
-\begin{APACrefauthors}%
-Iverson, N\BPBI R.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2012}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {A theory of glacial quarrying for landscape evolution
-  models} {A theory of glacial quarrying for landscape evolution
-  models}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Geology}{40}{8}{679--682}.
-\newblock
-\begin{APACrefDOI} \doi{10.1130/G33079.1} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Jansson%
-, Rosqvist%
-\BCBL {}\ \BBA {} Schneider%
-}{%
-Jansson%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2005}}%
-}]{%
-jansson2005}
-\APACinsertmetastar {%
-jansson2005}%
-\begin{APACrefauthors}%
-Jansson, P.%
-, Rosqvist, G.%
-\BCBL {}\ \BBA {} Schneider, T.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2005}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Glacier fluctuations, suspended sediment flux and
-  glacio‐lacustrine sediments} {Glacier fluctuations, suspended sediment flux
-  and glacio‐lacustrine sediments}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Geografiska Annaler: Series A, Physical
-  Geography}{87}{1}{37-50}.
-\newblock
-\begin{APACrefDOI} \doi{10.1111/j.0435-3676.2005.00243.x} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Jerolmack%
-\ \BBA {} Paola%
-}{%
-Jerolmack%
-\ \BBA {} Paola%
-}{%
-{\protect \APACyear {2010}}%
-}]{%
-jerolmack2010}
-\APACinsertmetastar {%
-jerolmack2010}%
-\begin{APACrefauthors}%
-Jerolmack, D.%
-\BCBT {}\ \BBA {} Paola, C.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2010}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Shredding of environmental signals by sediment
-  transport} {Shredding of environmental signals by sediment transport}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Geophysical Research Letters}{37}{19}{}.
-\newblock
-\begin{APACrefDOI} \doi{10.1029/2010GL044638} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Kasmalkar%
-, Mantelli%
-\BCBL {}\ \BBA {} Suckale%
-}{%
-Kasmalkar%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2019}}%
-}]{%
-kasmalkar2019}
-\APACinsertmetastar {%
-kasmalkar2019}%
-\begin{APACrefauthors}%
-Kasmalkar, I.%
-, Mantelli, E.%
-\BCBL {}\ \BBA {} Suckale, J.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2019}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Spatial heterogeneity in subglacial drainage driven by
-  till erosion} {Spatial heterogeneity in subglacial drainage driven by till
-  erosion}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Proceedings of the Royal Society A: Mathematical,
-  Physical and Engineering Sciences}{475}{2228}{20190259}.
-\newblock
-\begin{APACrefDOI} \doi{10.1098/rspa.2019.0259} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Koppes%
-\ \protect \BOthers {.}}{%
-Koppes%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2015}}%
-}]{%
-koppes2015}
-\APACinsertmetastar {%
-koppes2015}%
-\begin{APACrefauthors}%
-Koppes, M.%
-, Hallet, B.%
-, Rignot, E.%
-, Mouginot, J.%
-, Wellner, J\BPBI S.%
-\BCBL {}\ \BBA {} Boldt, K.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2015}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Observed latitudinal variations in erosion as a function
-  of glacier dynamics} {Observed latitudinal variations in erosion as a
-  function of glacier dynamics}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Nature}{526}{7571}{100--103}.
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Koppes%
-\ \BBA {} Montgomery%
-}{%
-Koppes%
-\ \BBA {} Montgomery%
-}{%
-{\protect \APACyear {2009}}%
-}]{%
-koppes2009a}
-\APACinsertmetastar {%
-koppes2009a}%
-\begin{APACrefauthors}%
-Koppes, M.%
-\BCBT {}\ \BBA {} Montgomery, D\BPBI R.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2009}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {The relative efficacy of fluvial and glacial erosion
-  over modern to orogenic timescales} {The relative efficacy of fluvial and
-  glacial erosion over modern to orogenic timescales}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Nature Geoscience}{2}{9}{644--647}.
-\newblock
-\begin{APACrefDOI} \doi{10.1038/ngeo616} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Lane%
-, Bakker%
-, Gabbud%
-, Micheletti%
-\BCBL {}\ \BBA {} Saugy%
-}{%
-Lane%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2017}}%
-}]{%
-lane2016}
-\APACinsertmetastar {%
-lane2016}%
-\begin{APACrefauthors}%
-Lane, S.%
-, Bakker, M.%
-, Gabbud, C.%
-, Micheletti, N.%
-\BCBL {}\ \BBA {} Saugy, J.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2017}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Sediment export, transient landscape response and
-  catchment-scale connectivity following rapid climate warming and alpine
-  glacier recession} {Sediment export, transient landscape response and
-  catchment-scale connectivity following rapid climate warming and alpine
-  glacier recession}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Geomorphology}{277}{}{210 - 227}.
-\newblock
-\begin{APACrefDOI} \doi{10.1016/j.geomorph.2016.02.015} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Leopold%
-\ \BBA {} Maddock%
-}{%
-Leopold%
-\ \BBA {} Maddock%
-}{%
-{\protect \APACyear {1953}}%
-}]{%
-leopold1953}
-\APACinsertmetastar {%
-leopold1953}%
-\begin{APACrefauthors}%
-Leopold, L.%
-\BCBT {}\ \BBA {} Maddock, T.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYear{1953}.
-\newblock
-\APACrefbtitle {The hydraulic geometry of stream channels and some
-  physiographic implications} {The hydraulic geometry of stream channels and
-  some physiographic implications}\ (\BVOL~252).
-\newblock
-\APACaddressPublisher{}{US Government Printing Office}.
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Li%
-\ \protect \BOthers {.}}{%
-Li%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2021}}%
-}]{%
-li2021}
-\APACinsertmetastar {%
-li2021}%
-\begin{APACrefauthors}%
-Li, D.%
-, Lu, X.%
-, Overeem, I.%
-, Walling, D\BPBI E.%
-, Syvitski, J.%
-, Kettner, A\BPBI J.%
-\BDBL {}Zhang, T.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2021}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Exceptional increases in fluvial sediment fluxes in a
-  warmer and wetter {High Mountain Asia}} {Exceptional increases in fluvial
-  sediment fluxes in a warmer and wetter {High Mountain Asia}}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Science}{374}{6567}{599--603}.
-\newblock
-\begin{APACrefDOI} \doi{10.1126/science.abi9649} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Mariotti%
-\ \protect \BOthers {.}}{%
-Mariotti%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2021}}%
-}]{%
-mariotti2021}
-\APACinsertmetastar {%
-mariotti2021}%
-\begin{APACrefauthors}%
-Mariotti, A.%
-, Blard, P\BHBI H.%
-, Charreau, J.%
-, Toucanne, S.%
-, Jorry, S.%
-, Molliex, S.%
-\BDBL {}Keddadouche, K.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2021}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Nonlinear forcing of climate on mountain denudation
-  during glaciations} {Nonlinear forcing of climate on mountain denudation
-  during glaciations}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Nature Geoscience}{}{}{1--7}.
-\newblock
-\begin{APACrefDOI} \doi{10.1038/s41561-020-00672-2} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Meyer-Peter%
-\ \BBA {} M{\"u}ller%
-}{%
-Meyer-Peter%
-\ \BBA {} M{\"u}ller%
-}{%
-{\protect \APACyear {1948}}%
-}]{%
-meyer1948}
-\APACinsertmetastar {%
-meyer1948}%
-\begin{APACrefauthors}%
-Meyer-Peter, E.%
-\BCBT {}\ \BBA {} M{\"u}ller, R.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{1948}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Formulas for bedload transport} {Formulas for bedload
-  transport}.{\BBCQ}
-\newblock
-\BIn{} \APACrefbtitle {Hydraulic Engineering Reports.} {Hydraulic engineering
-  reports.}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-M{\"u}ller%
-\ \BBA {} F{\"o}rstner%
-}{%
-M{\"u}ller%
-\ \BBA {} F{\"o}rstner%
-}{%
-{\protect \APACyear {1968}}%
-}]{%
-muller1968}
-\APACinsertmetastar {%
-muller1968}%
-\begin{APACrefauthors}%
-M{\"u}ller, G.%
-\BCBT {}\ \BBA {} F{\"o}rstner, U.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{1968}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {General relationship between suspended sediment
-  concentration and water discharge in the {A}lpenrhein and some other rivers}
-  {General relationship between suspended sediment concentration and water
-  discharge in the {A}lpenrhein and some other rivers}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Nature}{217}{5125}{244--245}.
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Nanni%
-\ \protect \BOthers {.}}{%
-Nanni%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2020}}%
-}]{%
-nanni2020}
-\APACinsertmetastar {%
-nanni2020}%
-\begin{APACrefauthors}%
-Nanni, U.%
-, Gimbert, F.%
-, Vincent, C.%
-, Gr\"aff, D.%
-, Walter, F.%
-, Piard, L.%
-\BCBL {}\ \BBA {} Moreau, L.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2020}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Quantification of seasonal and diurnal dynamics of
-  subglacial channels using seismic observations on an {A}lpine glacier}
-  {Quantification of seasonal and diurnal dynamics of subglacial channels using
-  seismic observations on an {A}lpine glacier}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{The Cryosphere}{14}{5}{1475--1496}.
-\newblock
-\begin{APACrefDOI} \doi{10.5194/tc-14-1475-2020} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Overeem%
-\ \protect \BOthers {.}}{%
-Overeem%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2017}}%
-}]{%
-overeem2017}
-\APACinsertmetastar {%
-overeem2017}%
-\begin{APACrefauthors}%
-Overeem, I.%
-, Hudson, B\BPBI D.%
-, Syvitski, J\BPBI P\BPBI M.%
-, Mikkelsen, A\BPBI B.%
-, Hasholt, B.%
-, van~den Broeke, M\BPBI R.%
-\BDBL {}Morlighem, M.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2017}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Substantial export of suspended sediment to the global
-  oceans from glacial erosion in {G}reenland} {Substantial export of suspended
-  sediment to the global oceans from glacial erosion in {G}reenland}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Nature Geoscience}{10}{}{859-863}.
-\newblock
-\begin{APACrefDOI} \doi{10.1038/NGEO3046} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Perolo%
-\ \protect \BOthers {.}}{%
-Perolo%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2018}}%
-}]{%
-perolo2018}
-\APACinsertmetastar {%
-perolo2018}%
-\begin{APACrefauthors}%
-Perolo, P.%
-, Bakker, M.%
-, Gabbud, C.%
-, Moradi, G.%
-, Rennie, C.%
-\BCBL {}\ \BBA {} Lane, S.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2018}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Subglacial sediment production and snout marginal ice
-  uplift during the late ablation season of a temperate valley glacier}
-  {Subglacial sediment production and snout marginal ice uplift during the late
-  ablation season of a temperate valley glacier}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Earth Surface Processes and Landforms}{0}{}{1-68}.
-\newblock
-\begin{APACrefDOI} \doi{10.1002/esp.4562} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Richards%
-\ \BBA {} Moore%
-}{%
-Richards%
-\ \BBA {} Moore%
-}{%
-{\protect \APACyear {2003}}%
-}]{%
-richards2003}
-\APACinsertmetastar {%
-richards2003}%
-\begin{APACrefauthors}%
-Richards, G.%
-\BCBT {}\ \BBA {} Moore, R\BPBI D.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2003}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Suspended sediment dynamics in a steep, glacier-fed
-  mountain stream, {P}lace {C}reek, {C}anada} {Suspended sediment dynamics in a
-  steep, glacier-fed mountain stream, {P}lace {C}reek, {C}anada}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Hydrological Processes}{17}{9}{1733--1753}.
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Riihimaki%
-, MacGregor%
-, Anderson%
-, Anderson%
-\BCBL {}\ \BBA {} Loso%
-}{%
-Riihimaki%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2005}}%
-}]{%
-riihimaki2005}
-\APACinsertmetastar {%
-riihimaki2005}%
-\begin{APACrefauthors}%
-Riihimaki, C\BPBI A.%
-, MacGregor, K\BPBI R.%
-, Anderson, R\BPBI .%
-, Anderson, S\BPBI P.%
-\BCBL {}\ \BBA {} Loso, M\BPBI G.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2005}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Sediment evacuation and glacial erosion rates at a small
-  alpine glacier} {Sediment evacuation and glacial erosion rates at a small
-  alpine glacier}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Journal of Geophysical Research: Earth Surface
-  (2003--2012)}{110}{F03003}{}.
-\newblock
-\begin{APACrefDOI} \doi{10.1029/2004JF000189} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Romans%
-, Covault%
-, Fildani%
-\BCBL {}\ \BBA {} Walsh%
-}{%
-Romans%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2016}}%
-}]{%
-romans2016}
-\APACinsertmetastar {%
-romans2016}%
-\begin{APACrefauthors}%
-Romans, S., B.W.and~Castelltort%
-, Covault, J.%
-, Fildani, A.%
-\BCBL {}\ \BBA {} Walsh, J.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2016}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Environmental signal propagation in sedimentary systems
-  across timescales} {Environmental signal propagation in sedimentary systems
-  across timescales}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Earth-Science Reviews}{153}{}{7-29}.
-\newblock
-\APACrefnote{Source-to-Sink Systems: Sediment \& Solute Transfer on the Earth
-  Surface}
-\newblock
-\begin{APACrefDOI} \doi{10.1016/j.earscirev.2015.07.012} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-R{\"o}thlisberger%
-}{%
-R{\"o}thlisberger%
-}{%
-{\protect \APACyear {1972}}%
-}]{%
-rothlisberger1972}
-\APACinsertmetastar {%
-rothlisberger1972}%
-\begin{APACrefauthors}%
-R{\"o}thlisberger, H.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{1972}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Water pressure in intra-- and subglacial channels}
-  {Water pressure in intra-- and subglacial channels}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Journal of Glaciology}{11}{62}{177--203}.
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Shields%
-}{%
-Shields%
-}{%
-{\protect \APACyear {1936}}%
-}]{%
-shields1936}
-\APACinsertmetastar {%
-shields1936}%
-\begin{APACrefauthors}%
-Shields, A.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{1936}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Anwendung der {A}ehnlichkeitsmechanik und der
-  {T}urbulenzforschung auf die {G}eschiebebewegung} {Anwendung der
-  {A}ehnlichkeitsmechanik und der {T}urbulenzforschung auf die
-  {G}eschiebebewegung}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{PhD Thesis Technical University Berlin}{}{}{}.
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Shreve%
-}{%
-Shreve%
-}{%
-{\protect \APACyear {1972}}%
-}]{%
-shreve1972}
-\APACinsertmetastar {%
-shreve1972}%
-\begin{APACrefauthors}%
-Shreve, R\BPBI L.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{1972}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Movement of water in glaciers} {Movement of water in
-  glaciers}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Journal of Glaciology}{11}{62}{205--214}.
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Stevens%
-\ \protect \BOthers {.}}{%
-Stevens%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2022}}%
-}]{%
-stevens2022}
-\APACinsertmetastar {%
-stevens2022}%
-\begin{APACrefauthors}%
-Stevens, D.%
-, Ely, J.%
-, Livingstone, S.%
-, Clark, C.%
-, Butcher, F.%
-\BCBL {}\ \BBA {} Hewitt, I.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2022}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Effects of basal topography and ice-sheet surface slope
-  in a subglacial glaciofluvial deposition model} {Effects of basal topography
-  and ice-sheet surface slope in a subglacial glaciofluvial deposition
-  model}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Journal of Glaciology}{}{}{1--13}.
-\newblock
-\begin{APACrefDOI} \doi{10.1017/jog.2022.71} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Swift%
-, Nienow%
-\BCBL {}\ \BBA {} Hoey%
-}{%
-Swift%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2005}}%
-}]{%
-swift2005}
-\APACinsertmetastar {%
-swift2005}%
-\begin{APACrefauthors}%
-Swift, D.%
-, Nienow, P\BPBI W.%
-\BCBL {}\ \BBA {} Hoey, T\BPBI B.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2005}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Basal sediment evacuation by subglacial meltwater:
-  suspended sediment transport from {H}aut {G}lacier d'{A}rolla, {S}witzerland}
-  {Basal sediment evacuation by subglacial meltwater: suspended sediment
-  transport from {H}aut {G}lacier d'{A}rolla, {S}witzerland}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Earth Surface Processes and
-  Landforms}{30}{7}{867--883}.
-\newblock
-\begin{APACrefDOI} \doi{10.1002/esp.1197} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Swift%
-\ \protect \BOthers {.}}{%
-Swift%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2021}}%
-}]{%
-swift2021}
-\APACinsertmetastar {%
-swift2021}%
-\begin{APACrefauthors}%
-Swift, D.%
-, Tallentire, G.%
-, Farinotti, D.%
-, Cook, S.%
-, Higson, W.%
-\BCBL {}\ \BBA {} Bryant, R.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2021}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {The hydrology of glacier-bed overdeepenings: Sediment
-  transport mechanics, drainage system morphology, and geomorphological
-  implications} {The hydrology of glacier-bed overdeepenings: Sediment
-  transport mechanics, drainage system morphology, and geomorphological
-  implications}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Earth Surface Processes and
-  Landforms}{46}{11}{2264-2278}.
-\newblock
-\begin{APACrefDOI} \doi{10.1002/esp.5173} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Tofelde%
-, Bernhardt%
-, Guerit%
-\BCBL {}\ \BBA {} Romans%
-}{%
-Tofelde%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2021}}%
-}]{%
-tofelde2021}
-\APACinsertmetastar {%
-tofelde2021}%
-\begin{APACrefauthors}%
-Tofelde, S.%
-, Bernhardt, A.%
-, Guerit, L.%
-\BCBL {}\ \BBA {} Romans, B.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2021}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Times Associated With Source-to-Sink Propagation of
-  Environmental Signals During Landscape Transience} {Times associated with
-  source-to-sink propagation of environmental signals during landscape
-  transience}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Frontiers in Earth Science}{9}{}{}.
-\newblock
-\begin{APACrefDOI} \doi{10.3389/feart.2021.628315} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Tucker%
-\ \BBA {} Slingerland%
-}{%
-Tucker%
-\ \BBA {} Slingerland%
-}{%
-{\protect \APACyear {1997}}%
-}]{%
-tucker1997}
-\APACinsertmetastar {%
-tucker1997}%
-\begin{APACrefauthors}%
-Tucker, G\BPBI E.%
-\BCBT {}\ \BBA {} Slingerland, R.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{1997}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Drainage basin responses to climate change} {Drainage
-  basin responses to climate change}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Water Resources Research}{33}{8}{2031--2047}.
-\newblock
-\begin{APACrefDOI} \doi{10.1029/97WR00409} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Ugelvig%
-, Egholm%
-, Anderson%
-\BCBL {}\ \BBA {} Iverson%
-}{%
-Ugelvig%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2018}}%
-}]{%
-ugelvig2018}
-\APACinsertmetastar {%
-ugelvig2018}%
-\begin{APACrefauthors}%
-Ugelvig, S\BPBI V.%
-, Egholm, D\BPBI L.%
-, Anderson, R\BPBI S.%
-\BCBL {}\ \BBA {} Iverson, N\BPBI R.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2018}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {{Glacial Erosion Driven by Variations in Meltwater
-  Drainage}} {{Glacial Erosion Driven by Variations in Meltwater
-  Drainage}}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Journal of Geophysical Research: Earth
-  Surface}{123}{0}{}.
-\newblock
-\begin{APACrefDOI} \doi{10.1029/2018JF004680} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-van As%
-\ \protect \BOthers {.}}{%
-van As%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2017}}%
-}]{%
-vanas2017}
-\APACinsertmetastar {%
-vanas2017}%
-\begin{APACrefauthors}%
-van As, D.%
-, Mikkelsen, A.%
-, Nielsen, M.%
-, Box, J.%
-, Liljedahl, L.%
-, Lindb{\"a}ck, K.%
-\BDBL {}Hasholt, B.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2017}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Hypsometric amplification and routing moderation of
-  Greenland ice sheet meltwater release} {Hypsometric amplification and routing
-  moderation of greenland ice sheet meltwater release}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{The Cryosphere}{11}{3}{1371--1386}.
-\newblock
-\begin{APACrefDOI} \doi{10.5194/tc-11-1371-2017} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Vergara%
-, Garreaud%
-\BCBL {}\ \BBA {} Ayala%
-}{%
-Vergara%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2022}}%
-}]{%
-vergara2022}
-\APACinsertmetastar {%
-vergara2022}%
-\begin{APACrefauthors}%
-Vergara, I.%
-, Garreaud, R.%
-\BCBL {}\ \BBA {} Ayala, {\'A}.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2022}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {{Sharp increase of extreme turbidity events due to
-  deglaciation in the subtropical Andes}} {{Sharp increase of extreme turbidity
-  events due to deglaciation in the subtropical Andes}}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Journal of Geophysical Research: Earth
-  Surface}{127}{6}{e2021JF006584}.
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Walder%
-\ \BBA {} Fowler%
-}{%
-Walder%
-\ \BBA {} Fowler%
-}{%
-{\protect \APACyear {1994}}%
-}]{%
-walder1994}
-\APACinsertmetastar {%
-walder1994}%
-\begin{APACrefauthors}%
-Walder, J\BPBI S.%
-\BCBT {}\ \BBA {} Fowler, A.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{1994}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Channelized subglacial drainage over a deformable bed}
-  {Channelized subglacial drainage over a deformable bed}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Journal of Glaciology}{40}{134}{3--15}.
-\newblock
-\begin{APACrefDOI} \doi{10.3189/S0022143000003750} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Werder%
-, Bauder%
-, Funk%
-\BCBL {}\ \BBA {} Keusen%
-}{%
-Werder%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2010}}%
-}]{%
-werder2010}
-\APACinsertmetastar {%
-werder2010}%
-\begin{APACrefauthors}%
-Werder, M\BPBI A.%
-, Bauder, A.%
-, Funk, M.%
-\BCBL {}\ \BBA {} Keusen, H\BPBI R.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2010}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Hazard assessment investigations in connection with the
-  formation of a lake on the tongue of {U}nterer {G}rindelwaldgletscher,
-  {B}ernese {A}lps, {S}witzerland} {Hazard assessment investigations in
-  connection with the formation of a lake on the tongue of {U}nterer
-  {G}rindelwaldgletscher, {B}ernese {A}lps, {S}witzerland}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Natural Hazards and Earth System
-  Science}{10}{2}{227--237}.
-\newblock
-\begin{APACrefDOI} \doi{10.5194/nhess-10-227-2010} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Werder%
-, Hewitt%
-, Schoof%
-\BCBL {}\ \BBA {} Flowers%
-}{%
-Werder%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2013}}%
-}]{%
-werder2013}
-\APACinsertmetastar {%
-werder2013}%
-\begin{APACrefauthors}%
-Werder, M\BPBI A.%
-, Hewitt, I\BPBI J.%
-, Schoof, C\BPBI G.%
-\BCBL {}\ \BBA {} Flowers, G\BPBI E.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2013}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Modeling channelized and distributed subglacial drainage
-  in two dimensions} {Modeling channelized and distributed subglacial drainage
-  in two dimensions}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Journal of Geophysical Research: Earth
-  Surface}{118}{4}{2140--2158}.
-\newblock
-\begin{APACrefDOI} \doi{10.1002/jgrf.20146} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Whipple%
-\ \BBA {} Tucker%
-}{%
-Whipple%
-\ \BBA {} Tucker%
-}{%
-{\protect \APACyear {1999}}%
-}]{%
-whipple1999}
-\APACinsertmetastar {%
-whipple1999}%
-\begin{APACrefauthors}%
-Whipple, K.%
-\BCBT {}\ \BBA {} Tucker, G.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{1999}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {{Dynamics of the stream-power river incision model:
-  Implications for height limits of mountain ranges, landscape response
-  timescales, and research needs}} {{Dynamics of the stream-power river
-  incision model: Implications for height limits of mountain ranges, landscape
-  response timescales, and research needs}}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Journal of Geophysical Research: Solid
-  Earth}{104}{B8}{17661-17674}.
-\newblock
-\begin{APACrefDOI} \doi{10.1029/1999JB900120} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Wickert%
-\ \BBA {} Schildgen%
-}{%
-Wickert%
-\ \BBA {} Schildgen%
-}{%
-{\protect \APACyear {2019}}%
-}]{%
-wickert2019}
-\APACinsertmetastar {%
-wickert2019}%
-\begin{APACrefauthors}%
-Wickert, A\BPBI D.%
-\BCBT {}\ \BBA {} Schildgen, T\BPBI F.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2019}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Long-profile evolution of transport-limited gravel-bed
-  rivers} {Long-profile evolution of transport-limited gravel-bed
-  rivers}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Earth Surface Dynamics}{7}{1}{17--43}.
-\newblock
-\begin{APACrefDOI} \doi{10.5194/esurf-7-17-2019} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Willenbring%
-\ \BBA {} Jerolmack%
-}{%
-Willenbring%
-\ \BBA {} Jerolmack%
-}{%
-{\protect \APACyear {2016}}%
-}]{%
-willenbring2016}
-\APACinsertmetastar {%
-willenbring2016}%
-\begin{APACrefauthors}%
-Willenbring, J.%
-\BCBT {}\ \BBA {} Jerolmack, D.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2016}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {{The null hypothesis: globally steady rates of erosion,
-  weathering fluxes and shelf sediment accumulation during Late Cenozoic
-  mountain uplift and glaciation}} {{The null hypothesis: globally steady rates
-  of erosion, weathering fluxes and shelf sediment accumulation during Late
-  Cenozoic mountain uplift and glaciation}}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Terra Nova}{28}{1}{11-18}.
-\newblock
-\begin{APACrefDOI} \doi{10.1111/ter.12185} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Williams%
-}{%
-Williams%
-}{%
-{\protect \APACyear {1989}}%
-}]{%
-williams1989}
-\APACinsertmetastar {%
-williams1989}%
-\begin{APACrefauthors}%
-Williams, G\BPBI P.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{1989}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Sediment concentration versus water discharge during
-  single hydrologic events in rivers} {Sediment concentration versus water
-  discharge during single hydrologic events in rivers}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Journal of Hydrology}{111}{1}{89 -- 106}.
-\newblock
-\begin{APACrefDOI} \doi{10.1016/0022-1694(89)90254-0} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Wolman%
-\ \BBA {} Miller%
-}{%
-Wolman%
-\ \BBA {} Miller%
-}{%
-{\protect \APACyear {1960}}%
-}]{%
-wolman1960}
-\APACinsertmetastar {%
-wolman1960}%
-\begin{APACrefauthors}%
-Wolman, M\BPBI G.%
-\BCBT {}\ \BBA {} Miller, J\BPBI P.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{1960}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Magnitude and frequency of forces in geomorphic
-  processes} {Magnitude and frequency of forces in geomorphic
-  processes}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{The Journal of Geology}{68}{1}{54--74}.
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Wong%
-\ \BBA {} Parker%
-}{%
-Wong%
-\ \BBA {} Parker%
-}{%
-{\protect \APACyear {2006}}%
-}]{%
-wong2006}
-\APACinsertmetastar {%
-wong2006}%
-\begin{APACrefauthors}%
-Wong, M.%
-\BCBT {}\ \BBA {} Parker, G.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2006}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Reanalysis and correction of bed-load relation of
-  {M}eyer-{P}eter and {M}{\"u}ller using their own database} {Reanalysis and
-  correction of bed-load relation of {M}eyer-{P}eter and {M}{\"u}ller using
-  their own database}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Journal of Hydraulic Engineering}{132}{11}{1159--1168}.
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Yager%
-, Venditti%
-, Smith%
-\BCBL {}\ \BBA {} Schmeeckle%
-}{%
-Yager%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2018}}%
-}]{%
-yager2018}
-\APACinsertmetastar {%
-yager2018}%
-\begin{APACrefauthors}%
-Yager, E.%
-, Venditti, J.%
-, Smith, H.%
-\BCBL {}\ \BBA {} Schmeeckle, M.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2018}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {The trouble with shear stress} {The trouble with shear
-  stress}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Geomorphology}{323}{}{41--50}.
-\PrintBackRefs{\CurrentBib}
-
-\bibitem [\protect \citeauthoryear {%
-Zhang%
-\ \protect \BOthers {.}}{%
-Zhang%
-\ \protect \BOthers {.}}{%
-{\protect \APACyear {2022}}%
-}]{%
-zhang2022}
-\APACinsertmetastar {%
-zhang2022}%
-\begin{APACrefauthors}%
-Zhang, T.%
-, Li, D.%
-, East, A.%
-, Walling, D.%
-, Lane, S.%
-, Overeem, I.%
-\BDBL {}Lu, X.%
-\end{APACrefauthors}%
-\unskip\
-\newblock
-\APACrefYearMonthDay{2022}{}{}.
-\newblock
-{\BBOQ}\APACrefatitle {Warming-driven erosion and sediment transport in cold
-  regions} {Warming-driven erosion and sediment transport in cold
-  regions}.{\BBCQ}
-\newblock
-\APACjournalVolNumPages{Nature Reviews Earth \& Environment}{}{}{1--20}.
-\newblock
-\begin{APACrefDOI} \doi{10.1038/s43017-022-00362-0} \end{APACrefDOI}
-\PrintBackRefs{\CurrentBib}
-
-\end{thebibliography}
-
-\section{Supplement}
-
-\renewcommand{\figurename}{Figure S}
-\setcounter{figure}{0}
-
-\begin{center}
-  \begin{figure}[H]
-    \includegraphics[width=0.8\linewidth]{multi_run_vars.pdf}
-    \caption{Variability $\chi$ in velocity ($v_g$, $v_f$; a) shear stress ($\tau_g$, $\tau_f$; b) and width-integrated shear stress ($\tau_{tg}$, $\tau_{tf}$; c)  in glacial and subaerial systems.
-Values are determined from examining the same parameter space in Figure~\ref{fig:range}.
-}
-    \label{fig:gammas}
-  \end{figure}
-\end{center}
-
-\end{document}

### 2023-07-10 Delaney2023.tex
to ol

@@
 \newcommand{\leif}[1]{{\textbf{\color{red}Leif says:} \color{red} #1} }
-\newcommand{\db}{\textit{DEBRIS}\,}
-\newcommand{\cl}{\textit{CLEAN}\,}
+\newcommand{\alpine}{\textit{ALPINE}\,}
+\newcommand{\icesheet}{\textit{ICESHEET}\,}
 \newcommand{\m}{$\,\mathrm{m}$\,}
@@
 \title{Water discharge has a variable impact on sediment transport capacity in subglacial systems}
-
-\ian{Alpine glacier vs Ice sheet}
 
@@
 These differences have been implicitly included in a wide range of available models that  quantify sediment transport from catchments in both subglacial and subaerial systems  \cite<e.g.>[]{walder1994,tucker1997,creyts2013,wickert2019,hewitt2019}.
+However, some suggest that subglacial sediment discharge  $Q_s \propto Q_w^{\frac{9}{2}}$, assuming that water discharge, changing channel width and depth are each proportionally responsible for the variance of water velocity and sediment transport capacity \cite{alley1997,swift2021}.
 The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \cite<e.g.>[]{lane2016,perolo2018}.
@@
 
-The parameterizations above are applied to hydrological records from the Fieschergletscher in the Swiss Alps (info) and the Leverett glacier on the Greenland Ice Sheet (info).
-The hydrographs applied to the subglacial and subaerial parameterizations represent the subglacial flow of water as it leaves the glacier and moves into a subaerial flow in front of the glacier, as water moves through the catchment.
-Outputs of the parameterizations are meant to represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
-
 The formulation of the glacial system (Section~\ref{sect:sub_mode}) requires inputs of the ice thickness ($h_o$, Equation\,\ref{eq:dS_dt}), water discharge, $Q_w$,  Hooke angle (Equation~\ref{eq:Hf}; $\beta$), and friction factor (Equations~\ref{eq:dh}~and~\ref{eq:tau}; $f_r$).
@@
 The friction factor, $f_f$, and channel shape factor, $k$, are assigned such that they produce reasonable values for water velocity and channel width  (Section~\ref{sect:fluv}).
+
+The parameterizations above are applied to hydrological records from the Fieschergletscher in the Swiss Alps (info) and the Leverett glacier on the Greenland Ice Sheet (info).
+The hydrographs applied to the subglacial and subaerial parameterizations represent the subglacial flow of water as it leaves the glacier and moves into a subaerial flow in front of the glacier, as water moves through the catchment.
+Outputs of the parameterizations are meant to represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
+To generalize these cases, the Fieschergletscher case is referred to as \alpine{} and is exemplified by relatively thin ice thickness ($h_o$: xxx), low discharge (< $20$\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in discharge.
+The Leverett glacier case is referred to as \icesheet{} and is exemplified by thick ice  ($h_o$: xxx), high water discharge (< $300$\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in discharge.
 
@@
 Note however, that this occurs across the total of xxx parameter combinations examined.
+In both glaciers and across all variables, the variability reduces from one day to five day aggradations (Figure\,\ref{fig:multi_run}).
 
@@
 \section{Discussion}
-\ian{add cautionary note as this is an example}
-
-\ian{Zwally and tedstone effect with increased discharge... non-linear response of increasing water discharge. }
 
@@
 In the total sediment transport relationship \citeA{engelund1967}, the shear stress is scaled to $\frac{5}{2}$ power.
-In both cases, the exponent greater than $1$ of shear stress in the sediment transport relationship magnifies the variability in sediment transport beyond the highly variable shear stress and width-integrated shear stress described above in subglacial environments (Figure~\ian{xxx}).
+In both cases, the exponent greater than $1$ of shear stress in the sediment transport relationship magnifies the variability in sediment transport beyond the highly variable shear stress and total shear stress described above in subglacial environments (Figure~\ref{fig:Qw_vari}.
 
-% Both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress of water across the channel bed is homogeneous \cite<Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}; e.g.
+Both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress of water across the channel bed is homogeneous \cite<Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}; e.g.
 >{yager2018}.
-% \ian{Phillips and Jeromack 2016... stabilization here.}
-% In the subaerial parameterization, this means that sediment is transported across all channel widths, as opposed to large flows such as floods when much sediment transport occurs \cite{wolman1960}.
-% As a result, the  the subaerial systems variability in sediment transport capacity across the range of discharge values could be underestimated in the parameterization here (Figure~~\ian{xxx}).
+In the subaerial parameterization, this means that sediment is transported across all channel widths, as opposed to large flows such as floods when much sediment transport occurs \cite{wolman1960}.
+As a result, the  the subaerial systems variability in sediment transport capacity across the range of discharge values could be underestimated in the parameterization here (Figure~\ref{fig:Qw_vari}).
+Furthermore, the similar variance of total shear stress shows (Figure~\ref{fig:multi_run}) that subglacial channels contain self-organizing feedbacks that limit the mobilization of sediment, as in fluvial systems \cite{phillips2016}.
 
@@
 
-The parameterization here shows that the smallest variability in water discharge has the largest difference in relative variability in sediment transport metrics (Figure~~\ian{xxx}).
-At locations such as the Greenland Ice Sheet, smaller diurnal variations in water discharge compared to many alpine catchments \cite<c.f.>{swift2005,riihimaki2005,vanas2017,hasholt2018}.
-Therefore, the variability in sediment transport capacity in the transition from the ice sheet to the proglacial river could be more dramatic in Greenland than in alpine systems as water leaves the glacier and moves downstream.
-Thus the disparity in sediment transport capacity between the subglacial and subaerial systems could be most pronounced in larger catchments, potentially resulting in temporally mismatched periods between when glaciers introduce sediment to these river systems  and when the river mobilizes it (Figures~\ref{fig:model_outs}).
-
 \subsection{Records of sediment discharge  from glaciers}
 
-\ian{include comments about the limits of hysterisis analysis from glacial systems}
-
+The pressurized nature of subglacial water means that the relationship between water discharge and suspended sediment concentration, or sediment discharge could be disparate, especially over the timescales examined here.
+The rapid increase and decreases in sediment transport capacity, for example over a diurnal cycle, in a subglacial system may compound the variability in sediment transport already present in subaerial systems \cite<Figure~\ref{fig:Qw_vari}; >{williams1989,jerolmack2010}.
 Large variations in sediment transport capacity below glaciers may result in sediment mobilization and deposition processes in close spatial and temporal proximity to each other below the glacier \cite{gimbert2016,perolo2018}.
-The rapid increase and decreases in sediment transport capacity, for example over a diurnal cycle, in a subglacial system may compound the variability in sediment transport already present in subaerial systems \cite{williams1989,jerolmack2010}.
 As a result, variations in sediment transport capacity could be responsible for the fluctuations in sediment discharge (``flushings'') from glaciers that are not attributed to variations in water discharge \cite<e.g.>[]{richards2003,swift2021}.
-Aside from the sporadic nature of these events, the observed ``flushings'' could result from increases in water velocity resulting from reduced channel size, not water discharge, which increase sediment transport capacity.
+Aside from the sporadic nature of these events, the observed ``flushings'' could result from increases in water velocity resulting from reduced channel size, not water discharge, which increase sediment transport capacity (Figure~\ref{fig:Qw_vari}).
 
@@
 Records of sediment discharge have been used to establish the relationship, or lack thereof, between sediment discharge and climate in glacial systems \cite<e.g.>[]{koppes2009a,willenbring2016,mariotti2021}.
-The variability in subglacial sediment transport capacity presented here applies most generally to short-time scales responsible for the size of subglacial channels.
+The variability in subglacial sediment transport capacity presented here applies most generally to short-time scales responsible for the size of subglacial channels (Figure~\ref{fig:multi_run}).
 Yet, identifying climatic signals in sediment transport from transport-limited glaciers may require higher thresholds of climatic perturbation compared to subaerial systems \cite{tofelde2021}.
 This higher threshold of climatic perturbation may persist in addition to the stochastic nature of erosion and deposition in fluvial environments \cite{castletort2003,jerolmack2010,romans2016}.
+Results here suggest that increased variability in sediment transport in subglacial systems  could be particularly pronouced on ice sheets with large catchments and reduced variability.
 For instance, sediment discharge records from two glaciers in the Swiss Alps show that $40$ to $50$\% of a season's sediment discharge occurs when water discharge is below the $75^{\mathrm{th}}$ quantile of the season \cite{delaney2018}.
-Such a large quantity of sediment transported at lower water discharges may occur in part because sediment transport capacity responds to the size of the glacier conduit, in addition to water discharge (Section~\ref{sect:sub_mode}).
+Such a large quantity of sediment transported at lower water discharges may occur in part because sediment transport capacity responds to the size of the glacier conduit, in addition to water discharge (Figure~\ref{fig:model_outs} and Section~\ref{sect:sub_mode}).
 

### 2023-07-13 Delaney2023.tex
more bits

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
-% 
-% This file includes commands and instructions
-% given in the order necessary to produce a final output that will
-% satisfy AGU requirements, including customized APA reference formatting.
-% 
-% You may copy this file and give it your
-% article name, and enter your text.
-% 
-% 
-% Step 1: Set the \documentclass
-% 
-%
+\documentclass[11pt]{article}
+\usepackage [colon,square]{natbib}
+\usepackage {comment}
+\usepackage{cite}
+\usepackage[mathlines]{lineno} %% <- mathlines turns on line numbering in equations
+\usepackage{amsmath} 
+\usepackage{printlen}
+\setlength{\parindent}{10pt}
+\setlength{\parskip}{5pt}
+\usepackage[explicit]{titlesec}
+\usepackage{helvet}
+\usepackage{array}
+\usepackage{printlen}
+\usepackage[a4paper, margin=1.25cm]{geometry}
+\usepackage{color,hyperref}
+\catcode`\_=11\relax
+\newcommand\email[1]{\_email #1\q_nil}
+\def\_email#1@#2\q_nil{%
+  \href{mailto:#1@#2}{{\emailfont #1\emailampersat #2}}
+}
+\newcommand\emailfont{\sffamily}
+\newcommand\emailampersat{{\color{red}\small@}}
+\catcode`\_=8\relax
 
-%% To submit your paper:
-\documentclass[draft]{agujournal2019}
-\usepackage{url} %this package should fix any errors with URLs in refs.
-\usepackage{lineno}
-\usepackage{hyphenat}
-\usepackage{float}
-\usepackage{multirow}
-\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
-\usepackage{soul}
-\usepackage{array}
+\usepackage{hyperref}
+\usepackage[export]{adjustbox}
+\usepackage{bibentry}
+\renewcommand{\bibfont}{\footnotesize}
+\usepackage{tikz}
+\usepackage{setspace}
+\usepackage[inline,shortlabels]{enumitem}
+\usepackage[all]{nowidow}
+\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
+\pagenumbering{gobble} 
+\hypersetup{ % Remove ugly boxes around URLs
+  colorlinks, linkcolor={black!100!black}, citecolor={black!50!black}, urlcolor={black!80!black}
+}
+
 \newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
@@
 \hbadness=99999
-%%%%%%% 
-% As of 2018 we recommend use of the TrackChanges package to mark revisions.
-% The trackchanges package adds five new LaTeX commands:
-% 
-% \note[editor]{The note}
-% \annote[editor]{Text to annotate}{The note}
-% \add[editor]{Text to add}
-% \remove[editor]{Text to remove}
-% \change[editor]{Text to remove}{Text to add}
-% 
-% complete documentation is here: http://trackchanges.sourceforge.net/
-%%%%%%%
-
-\draftfalse
-
-%% Enter journal name below.
-%% Choose from this list of Journals:
-% 
-% JGR: Atmospheres
-% JGR: Biogeosciences
-% JGR: Earth Surface
-% JGR: Oceans
-% JGR: Planets
-% JGR: Solid Earth
-% JGR: Space Physics
-% Global Biogeochemical Cycles
-% Geophysical Research Letters
-% Paleoceanography and Paleoclimatology
-% Radio Science
-% Reviews of Geophysics
-% Tectonics
-% Space Weather
-% Water Resources Research
-% Geochemistry, Geophysics, Geosystems
-% Journal of Advances in Modeling Earth Systems (JAMES)
-% Earth's Future
-% Earth and Space Science
-% Geohealth
-% 
-% ie, \journalname{Water Resources Research}
-
-\journalname{Geophysical Research Letters}
 
@@
 
-\newenvironment{laysummary}{
-  \chapter{\centering \textbf{Plain Language Summary}}
-}
-
 \begin{document}
@@
 
-\title{Water discharge has a variable impact on sediment transport capacity in subglacial systems}
+\begin{center}
+\Large{\textbf{Water discharge quantity has a variable impact on sediment transport capacity in subglacial systems}}
+\normalsize
 
-%% ------------------------------------------------------------------------ %%
-% 
-% AUTHORS AND AFFILIATIONS
-%
+Ian Delaney
 
-% Authors are individuals who have significantly contributed to the
-% research and preparation of the article. Group authors are allowed, if
-% each author in the group is separately identified in an appendix.)
+Institut des dynamiques de la surface terrestre (IDYST), Universit\'{e} de Lausanne, B\^{a}timent G\'{e}opolis, CH-1015 Lausanne
 
-% List authors by first name or initial followed by last name and
-% separated by commas. Use \affil{} to number affiliations, and
-% \thanks{} for author notes.
-% Additional author notes should be indicated with \thanks{} (for
-% example, for current addresses).
+ianarburua.delaney@unil.ch
 
-% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antarctica}, B. C. Author\affil{2,3}, and D. E.
-% Author\affil{3,4}\thanks{Also funded by Monsanto.}}
-
-\authors{Ian Delaney\affil{1}}
-
-\affiliation{1}{Institut des dynamiques de la surface terrestre (IDYST), Universit\'{e} de Lausanne, B\^{a}timent G\'{e}opolis, CH-1015 Lausanne }
-
-% \affiliation{3}{Third Affiliation}
-% \affiliation{4}{Fourth Affiliation}
-
-% \affiliation{=number=}{=Affiliation Address=}
-% (repeat as many times as is necessary)
-
-%% Corresponding Author:
-% Corresponding author mailing address and e-mail address:
-
-% (include name and email addresses of the corresponding author.  More
-% than one corresponding author is allowed in this LaTeX file and for
-% publication; but only one corresponding author is allowed in our
-% editorial system.)
-
-% Example: \correspondingauthor{First and Last Name}{email@address.edu}
-
-\correspondingauthor{Ian Delaney}{ianarburua.delaney@unil.ch}
+\end{center}
 
@@
 
-\begin{keypoints}
+\paragraph{Keypoints}
+\begin{itemize}
 \item Water discharge variations in subglacial channels are mainly accommodated by water velocity, not channel size, given their slow evolution.
@@
 \item The systems' divergence in transport response may impact sediment dynamics at glacier margins and interpretation of sediment records.
-\end{keypoints}
+\end{itemize}
 
@@
 %%% Suggested section heads:
-\ian{comparison with Alley 1997... mention his variability-- also deal with his $Qs \propto Qw^{\frac{9}{2}}$  treats like rivers?  does not consider channel size evolution}
+%TODO: 
+%Check with Daniel and Mauro about data
+
+Recheck results figure names.
 
@@
 
-Glaciers expel large quantities of sediment from their termini \cite{hallet1996}.
-Changes in glacier dynamics, geomorphology, and hydrology  have prompted numerous  recent studies of  sediment transport processes in cold regions \cite<e.g.>[]{zhang2022}.
-Increases in sediment transport  have been observed in Greenland \cite{bendixen2017}, the European Alps \cite{costa2017}, the Himalayas \cite{li2021}, and the Andes \cite{vergara2022}.
-In some of these regions, increased water discharge and glacier melt have been attributed to greater sediment transport capacity \cite{bendixen2017,costa2017,li2021}.
-In other cases, enhanced access of meltwater to subglacial sediment at high elevations was attributed as the cause of increased sediment transport \cite{delaney2020,vergara2022}.
+Glaciers expel large quantities of sediment from their termini \citep{hallet1996}.
+Changes in glacier dynamics, geomorphology, and hydrology  have prompted numerous  recent studies of  sediment transport processes in cold regions \citep[e.g.][]{zhang2022}.
+Increases in sediment transport  have been observed in Greenland \citep{bendixen2017}, the European Alps \citep{costa2017}, the Himalayas \citep{li2021}, and the Andes \citep{vergara2022}.
+In some of these regions, increased water discharge and glacier melt have been attributed to greater sediment transport capacity \citep{bendixen2017,costa2017,li2021}.
+In other cases, enhanced access of meltwater to subglacial sediment at high elevations was attributed as the cause of increased sediment transport \citep{delaney2020,vergara2022}.
 Observed changes to sediment transport in glacierized catchments create an imperative to examine the processes controlling sediment discharge here. 
@@
 
-Over long periods, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from below glaciers \cite<c.f.>{hallet1979,iverson2012,ugelvig2018}.
-On shorter time periods, pressurized water transports sediment from the below glacier \cite{walder1994,creyts2013,beaud2018}, should enough sediment be present below the glacier (i.e.
+Over long periods, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from below glaciers \citep[c.f.][]{hallet1979,iverson2012,ugelvig2018}.
+On shorter time periods, pressurized water transports sediment from the below glacier \citep{walder1994,creyts2013,beaud2018}, should enough sediment be present below the glacier (i.e.
 in a transport-limited regime).
@@
 In a transport-limited regime, sediment discharge responds to the sediment transport capacity or the amount of sediment that could be carried by the water.
-Sediment transport capacity depends on the shear stress between water and sediment it flows  over \cite{shields1936,meyer1948,engelund1967} and the width of the channel bottom $w_c$ over which to mobilize sediment.
+Sediment transport capacity depends on the shear stress between water and sediment it flows  over \citep{shields1936,meyer1948,engelund1967} and the width of the channel bottom $w_c$ over which to mobilize sediment.
 The shear stress  across the channel width $\tau_t$ responds to the velocity of water $v$ flowing through at channel so that
@@
 
-In subaerial channels, with open channel flow, the cross-sectional area $S$ of the channel evolves with changing water discharge $Q_w$  by changing the width of the channel along with the depth \cite{leopold1953}.
+In subaerial channels, with open channel flow, the cross-sectional area $S$ of the channel evolves with changing water discharge $Q_w$  by changing the width of the channel along with the depth \citep{leopold1953}.
 During a flood with greater water discharge $Q_w$, for instance, the cross-sectional area $S$ increases as the channel widens and deepens, along with water velocity $v$.
@@
 Yet, the response of water velocity to changing water discharge in subglacial channels is different than subaerial ones.
-In subglacial systems, the water is pressurized by the ice above \cite{rothlisberger1972}.
-The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by frictional heating of water flowing through the channel \cite{rothlisberger1972}.
-As a result, the subglacial channel size evolves over days to months, whereas water discharge can vary over hours \cite<e.g.>[]{iken1986,andrews2014,nanni2020}.
+In subglacial systems, the water is pressurized by the ice above \citep{rothlisberger1972}.
+The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by frictional heating of water flowing through the channel \citep{rothlisberger1972}.
+As a result, the subglacial channel size evolves over days to months, whereas water discharge can vary over hours \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
 In turn, changes in water discharge $Q_w$ are mainly accommodated with water velocity $v$, because the size of the channel $S$ changes on a much shorter timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
@@
 As a result, sediment mobilization in subaerial and subglacial systems responds differently to changing water discharge.
-These differences have been implicitly included in a wide range of available models that  quantify sediment transport from catchments in both subglacial and subaerial systems  \cite<e.g.>[]{walder1994,tucker1997,creyts2013,wickert2019,hewitt2019}.
-However, some suggest that subglacial sediment discharge  $Q_s \propto Q_w^{\frac{9}{2}}$, assuming that water discharge, changing channel width and depth are each proportionally responsible for the variance of water velocity and sediment transport capacity \cite{alley1997,swift2021}.
-The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \cite<e.g.>[]{lane2016,perolo2018}.
-Additionally, the variable response of sediment transport capacity to water discharge in the two systems may impact the interpretation of records of sediment transport in glacial compared to subaerial systems \cite<e.g.>[]{muller1968,richards2003,ganti2016}.
+These differences have been implicitly included in a wide range of available models that  quantify sediment transport from catchments in both subglacial and subaerial systems  \citep[e.g.][]{walder1994,tucker1997,creyts2013,wickert2019,hewitt2019}.
+However, some suggest that subglacial sediment discharge  $Q_s \propto Q_w^{\frac{9}{2}}$, assuming that water discharge, changing channel width and depth are each proportionally responsible for the variance of water velocity and sediment transport capacity \citep{alley1997,swift2021}.
+The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \citep[e.g.][]{lane2016,perolo2018}.
+Additionally, the variable response of sediment transport capacity to water discharge in the two systems may impact the interpretation of records of sediment transport in glacial compared to subaerial systems \citep[e.g.][]{muller1968,richards2003,ganti2016}.
 Yet, the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial systems has been minimally discussed.
@@
 To better understand the relationship between sediment transport capacity and water discharge in subglacial systems, this manuscript leverages parameterizations of subglacial and subaerial hydraulics.
-These parameterizations are applied to hydrological records from the Fieschergletscher, an alpine glacier in Switzerland, and Leverett glacier,  a land-terminating glacier in Greenland.
+These parameterizations are applied to hydrological records from an alpine glacier in Switzerland (Fieschergletscher), and  a land-terminating glacier in Greenland (Leverett glacier).
 Outputs are used to explore differences in the relationships between  water discharge, channel geometry, and water velocity in the subglacial and subaerial. 
@@
 \begin{center}
-  \begin{figure}[H]
+  \begin{figure}[h]
     \includegraphics[width=0.8\linewidth]{Fig1.pdf}
@@
 Parameterizations below represent relationships amongst water discharge, subglacial water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
-Because sediment transport largely depends on shear stress applied by water flow \cite{shields1936}, both parameterizations calculate this value and integrate  it across the channel bed (Figure \ref{fig:cartoon}).
-The evaluation of shear stress omits a grain size parameter and, thus, leaves open the selection of a sediment transport relationship  \cite<e.g.>[]{shields1936,meyer1948} to  calculate sediment transport capacity.
+Because sediment transport largely depends on shear stress applied by water flow \citep{shields1936}, both parameterizations calculate this value and integrate  it across the channel bed (Figure \ref{fig:cartoon}).
+The evaluation of shear stress omits a grain size parameter and, thus, leaves open the selection of a sediment transport relationship  \citep[e.g.][]{shields1936,meyer1948} to  calculate sediment transport capacity.
 As a result, relative relationships in shear stress, as opposed to sediment transport capacity, are examined.
 
-\ian{update hooke angle }
-\begin{table}[H]
+\begin{table}[ht]
   \centering
@@
 
+  $^a$ for \alpine{} (Fieschergletscher)
+
+  $^b$ for \icesheet{} (Leverett glacier)
+
 \end{table}
-  $^a$ for Fieschergletscher
-
-  $^b$ for Leverett glacier
 
@@
 To evaluate the shear stress of water flowing across sediments below a glacier, the subglacial channel parameterization evaluates the channel geometry and the velocity of the flowing water.
-To accomplish this, a modified version of  the lumped hydraulics model presented in \citeA{clarke1996} and \citeA{werder2010} is used.
+To accomplish this, a modified version of  the lumped hydraulics model presented in \citet{clarke1996} and \citet{werder2010} is used.
 
-Here, it is assumed that the water is transported through a subglacial channel flowing down the glacier; \cite<Figure~\ref{fig:cartoon}; >{rothlisberger1972}, and that the channel  size responds to frictional heating from water flow and creep closure by ice, using the Darcy-Weisbach formulation for water-flow through a pipe  \cite<e.g.>[]{rothlisberger1972,clarke2003}.
-In the formulation here does not consider englacial storage of water, and thus that changes in hydraulic head at the top of the glacier are negligable in the evolution of the channel size. 
+Here, it is assumed that the water is transported through a subglacial channel flowing down the glacier; \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972}, and that the channel  size responds to frictional heating from water flow and creep closure by ice, using the Darcy-Weisbach formulation for water-flow through a pipe  \citep[e.g.][]{rothlisberger1972,clarke2003}.
+The formulation here does not consider englacial storage of water, and thus that changes in hydraulic head at the top of the glacier are negligable in the evolution of the channel size. 
 The evolution of subglacial channel size $S_g$ is given as
@@
 \noindent where $f_r$ is a friction factor, $D_h$ is the hydraulic diameter, and $l$ is the length overwhich head change occurs.
-$s$ is a factor accounting for channel geometry \cite{hooke1990}, calculated as:
+$s$ is a factor accounting for channel geometry \citep{hooke1990}, calculated as:
 \begin{equation}
@@
 
-The width of the flat channel floor $w_{cg}$, given angle $beta$ is 
+The width of the flat channel floor $w_{cg}$, given angle $\beta$ is 
 \begin{equation}
@@
 
-To parameterize the shear stress of water flowing across sediments in the subaerial channel,  the hydraulics parameterization presented in \citeA{tucker1997} is implemented.
+To parameterize the shear stress of water flowing across sediments in the subaerial channel,  the hydraulics parameterization presented in \citet{tucker1997} is implemented.
 Here, it is assumed that there is a conservation of mass and a sufficiently wide channel so that the hydraulic radius is consistent with flow depth, uniform flow, and the Darcy-Weisbach relationship.
@@
 % 
-where $k$ is a constant and $e$ is an exponent commonly equal to $\frac{1}{2}$ \cite{leopold1953}.
+where $k$ is a constant and $e$ is an exponent commonly equal to $\frac{1}{2}$ \citep{leopold1953}.
 Water velocity, $v_f$, is given by rearranging Equation \ref{eq:tau} as
@@
 
-The parameterizations above are applied to hydrological records from the Fieschergletscher in the Swiss Alps (info) and the Leverett glacier on the Greenland Ice Sheet (info).
+The parameterizations above are applied to hydrological records from the Fieschergletscher in the Swiss Alps and the Leverett glacier on the Greenland Ice Sheet.
 The hydrographs applied to the subglacial and subaerial parameterizations represent the subglacial flow of water as it leaves the glacier and moves into a subaerial flow in front of the glacier, as water moves through the catchment.
 Outputs of the parameterizations are meant to represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
-To generalize these cases, the Fieschergletscher case is referred to as \alpine{} and is exemplified by relatively thin ice thickness ($h_o$: xxx), low discharge (< $20$\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in discharge.
-The Leverett glacier case is referred to as \icesheet{} and is exemplified by thick ice  ($h_o$: xxx), high water discharge (< $300$\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in discharge.
+To generalize these cases, the Fieschergletscher case is referred to as \alpine{} and is exemplified by relatively thin ice thickness ($h_o$: $225$\,\unit{m}), low discharge ( $\le\,20$\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in discharge.
+The Leverett glacier case is referred to as \icesheet{} and is exemplified by thick ice  ($h_o$: $700$\,\unit{m}), high water discharge ($\le\,300$\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in discharge.
 
-The parameterizations are first applied to a reference test case for each glacier that assumes a semi-circular subglacial channel ($\beta=\pi$) and friction factors $f_r$ and $f_p$ are tuned so that the model reproduces reasonable velocities for both  the Greenland Ice Sheet and alpine glaciers \cite<$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$>{werder2010b,chandler2013}.
+The parameterizations are first applied to a reference test case for each glacier that assumes a semi-circular subglacial channel ($\beta=\pi$) and friction factors $f_r$ and $f_p$ are tuned so that the model reproduces reasonable velocities for both  the Greenland Ice Sheet and alpine glaciers \citep[$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$][]{werder2010b,chandler2013}.
 Water velocity, shear-stress and width-integrated shear-stress in the subglacial and subaerial systems are presented below.
@@
 To characterize the variability in sediment discharge capacity in subglacial compared to subaerial systems further, the parameterizations are applied to range of different channel geometry factors ($\beta$ and $k$) and friction factors $f_r$ and $f_p$.
-Model runs with parameter combinations  are culled, if their maximum subglacial water velocity exceeds $2\,$\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity exceeds $1\,$\unit{m}$^3$\,\unit{s}$^{-1}$.
+Model runs with parameter combinations  are culled if their maximum subglacial water velocity exceeds $2\,$\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity exceeds $1\,$\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with some dye tracing measurements \citep[e.g.]{werder2010b,chandler2013}.
 The standard deviation of water velocity, shear-stress and width-integrated shear-stress across the season long hydrographs are presented below.
@@
 
-% there is more variability across the season subglacial and subaerial cases. peaks in subaerial cases lag peaks in subglacial cases. (Figure 2)
-
-% For a fixed channel geometry, there is a single shear stress for a water discharge.
-In the subglacial case, peak water discharge that is 1.5 times the mean velocity.
-In Fieschercase, peak water discharge does not result in the greatest of any variable.
-In Leverett case the mean total shear stress can occur from 100 m3/s to 200 m3/s.
-The mean velocity and shear stress can occur across effectively all water discharges below 250 m3/s
-
-% across a wide range of channel geometries
-
-Following the parameterizations defined in Sections~\ref{sect:sub_mode}~and~\ref{sect:fluv}, the water velocity, shear-stress and total shear-stress exhibit vastly different seasonal evolutions and peaks of water velocity and shear stress (representing the potential for sediment mobilization) and total shear stress (representing the total sediment transport capacity; Figure~\ref{fig:model_outs}).
-Furthermore, these variables respond differently to water discharge in the subglacial and subaerial systems (Figure~\ref{fig:Qw_vari}).
-For each water discharge value in the subaerial system, there is a single water velocity, shear-stress and total shear-stress (Figure~\ref{fig:Qw_vari}).
-In the subglacial system, water velocity, shear-stress and total shear-stress vary substantially for a given water discharge(Figure~\ref{fig:Qw_vari}).
-For instance, over $15$ minute time periods at Fieschergletscher, the range of water velocities and shear-stress spans from those below the mean to those over $1.5$ times the mean.
-Substantially high values of water velocity and shear-stress can occur from minimal water discharge at $2.5$\,\unit{m}$^3$\,\unit{s}$^{-1}$ to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
-At Leverett glacier, reduced diurnal variability in the hydrograph  (Figure~\ref{fig:model_outs} e) results in smaller variations in water velocity and shear stress compared to Fieschergletscher (Figure~\ref{fig:Qw_vari}).
-However, mean values of water velocity can occur at water discharges between roughly $50$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $225$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, effectively spanning the observed water discharge.
-
-When accounting for the evolving channel width of the subglacial conduit, total shear-stress generally increases with water discharge.
-Yet the total shear-stress across the can very substantially, with the highest values in total shear-stress occurring at water discharge values ranging from roughly $11$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
-The variance is less at the Leverett glacier case, yet a range of total shear-stress can occur at a given discharge(Figure~\ref{fig:Qw_vari} c, e).
-
-Over the course of the season, large peaks in water velocity and shear stress, along with total shear stress, occur in response to increases in water discharge.
-In the subglacial system, these peaks generally occur over the period when water discharge increases at the fastest rate (Figure~\ref{fig:model_outs}).
-Conversely, in the subaerial system peaks in these variables occur when water discharge is at its highest value.
-As a result, peaks in the subglacial variables occur prior to the peaks in the subglacial variables, representing lags sediment mobilization and  transport capacity as sediment is transferred from on system to the other.
-
-The water velocity and shear stress remain higher across the different friction factors and channel geometries examined in the ensemble of runs across averaging periods.
-However, the Leverett case, some parameter combinations the result in greater variability in subaerial system compared to the subaerial channels.
-The effect of channel evolving width largely counteracts the variability in water velocity, especially in the alpine glacier case.
-Here the large diurnal cycles in the hydrograph can cause the maximum and minimum amounts of variability in subaerial total shear-stress to be greater than those in the subglacial system.
-At the Leverett glacier, with reduced diurnal variability in water discharge (Figure\,\ref{fig:model_outs}), the variability in total shear-stress remains higher in the subglacial system compared to the subaerial one.
-However, some variability in subaerial parameter combinations exceeds those subglacial ones.
-Note however, that this occurs across the total of xxx parameter combinations examined.
-In both glaciers and across all variables, the variability reduces from one day to five day aggradations (Figure\,\ref{fig:multi_run}).
-
 \begin{center}
-  \begin{figure}[H]
-    \includegraphics[width=0.8\linewidth]{model_outputs.pdf}
-    \caption{Model outputs as in response to hydrograph (a and e) from Fieschergletscher (a-d) and Leverett glacier (e-h).
+  \begin{figure}[h]
+    \includegraphics[width=0.8\linewidth]{Fig2.pdf}
+    \caption{Model outputs as in response to hydrograph (a and e) from \alpine{} (a-d) and \icesheet{} (e-h).
 Gray line in a and e represent subglacial channel cross sectional area.
 Below this black lines represent model outputs of velocity, shear-stress and width-integrated shear-stress from the subglacial system, while grey lines represent the subaerial system.
-}
+      Plots are shown at $15$\unit{min} interval.
+    Arrows denote some peak events where peak the variable peaks in the subglacial system prior to the subaerial one.} 
     \label{fig:model_outs}
@@
 
+Following the parameterizations defined in Sections~\ref{sect:sub_mode}~and~\ref{sect:fluv}, the water velocity, shear-stress and total shear-stress exhibit different seasonal evolutions and peaks of water velocity and shear stress (representing the potential for sediment mobilization) and total shear stress (representing the total sediment transport capacity; Figure~\ref{fig:model_outs}).
+
 \begin{center}
-  \begin{figure}[H]
-    \includegraphics[width=0.8\linewidth]{Qw_vari.png}
-    \caption{Relationship between water discharge and velocity, shear stress and width-integrated shear stress for the Fieschergletscher hydrograph (a-c) and Leverett glacier (d-f).
-Variables have been normalized to by the mean values for better comparison amongst the model runs.}
+  \begin{figure}[h]
+    \includegraphics[width=0.8\linewidth]{Fig3.png}
+    \caption{Relationship between water discharge and velocity, shear stress and width-integrated shear stress for the \alpine{} hydrograph (a-c) and \icesheet (d-f).
+Variables have been normalized to by the mean values for better comparison amongst the model runs.
+      Plots are shown at $15$\unit{min} interval.} 
     \label{fig:Qw_vari}
@@
 
+For each water discharge value in the subaerial system, there is a single water velocity, shear-stress and total shear-stress (Figure~\ref{fig:Qw_vari}).
+Conversely in the subglacial system,  water velocity, shear-stress and total shear-stress vary substantially for a given water discharge (Figure~\ref{fig:Qw_vari}).
+For instance, over $15$ minute time periods at \alpine{}, the range of water velocities and shear-stress spans from those below the mean to those almost $2$ times the mean velocity.
+Substantially high values of water velocity and shear-stress can occur from minimal water discharge at $2.5$\,\unit{m}$^3$\,\unit{s}$^{-1}$ to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
+At \icesheet, reduced diurnal variability in the hydrograph  (Figure~\ref{fig:model_outs} e) results in smaller variations in water velocity and shear stress compared to \alpine{} (Figure~\ref{fig:Qw_vari}).
+However, mean values of water velocity can occur at water discharges between roughly $50$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $225$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, effectively spanning the observed range of water discharge.
+
+When accounting for the evolving channel width of the subglacial conduit, total shear-stress generally increases with water discharge.
+Yet the total shear-stress across the can very substantially, with the highest values in total shear-stress occurring at water discharge values ranging from roughly $11$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
+The variance is less at the \icesheet case, yet a range of total shear-stress can occur at a given discharge(Figure~\ref{fig:Qw_vari} c, e).
+
+Over the course of the season, large peaks in water velocity and shear stress, along with total shear stress, occur in response to increases in water discharge in both systems.
+In the subglacial system, these peaks generally occur over the period when water discharge increases at the fastest rate (Figure~\ref{fig:model_outs}).
+Conversely, in the subaerial system peaks in these variables occur when water discharge is at its highest value.
+As a result, peaks in the subglacial variables occur prior to the peaks in the subaerial variables, representing lags sediment mobilization and  transport capacity when sediment is transferred from on system to the other.
+
 \begin{center}
-  \begin{figure}[H]
-    \includegraphics[width=0.8\linewidth]{multi_run.png}
-    \caption{} 
+  \begin{figure}[h]
+    \includegraphics[width=0.8\linewidth]{Fig4.png}
+    \caption{Variance (standard deviation) in water velocity, shear stress and total shear stress against aggregation time for the subglacial (blue) and subaerial (orange) systems from the \alpine{} (a-c) and \icesheet{} (d-f).
+      Shaded areas denote the $25^{\mathrm{th}}$ and $75^{\mathrm{th}}$ quantiles of the $100$ parameter combinations for each the subglacial and subaerial systems.
+      Solid lines denote the mean values.
+      Time intervals show in markers denote $15$\,\unit{min}, $1$\,\unit{hr}, $12$\,\unit{hr}, $1$\,\unit{d}, $5$\,\unit{d}, $10$\,\unit{d}, $15$\,\unit{d}, and $30$\,\unit{d} aggregation periods.
+    } 
     \label{fig:multi_run}
@@
 
+Variability in water velocity and shear stress remain higher in the subglacial system compared to the subaerial one, across the range of parameter values examined for both systems.
+However, in the \alpine{} system accounting for channel width makes the variability in the systems comparable (Figure~\ref{fig:multi_run}).
+Yet, the variability over time periods less than one half day, the variability in the subglacial system is somewhat higher. 
+Conversely, in \icesheet{}, with lower water discharge variability, the variability in total shear stress remains higher in the subglacial system across the time aggregation periods.
+
+In all cases, the variability in velocity, shear stress and total shear stress decreases after $1-5$ day periods (Figure~\ref{fig:multi_run}).
+This suggests that the effects of the variable response to hydrologic forcing in these cases is reduced after these time periods.
+Yet, substantial differences persist in periods of up to $30$ days, as the variability might not converge given seasonal variations in water discharge.
+Note that lags and time of peak sediment transport conditions will remain offset (Figure~\ref{fig:model_outs}), while the variability in sediment transport parameters is reduced with increasing time aggregation (Figure~\ref{fig:multi_run}).
+
 \section{Discussion}
 
-The relationship between water discharge variability and relative variability of the water velocity, shear stress, and width-integrated shear stress results from the fixed geometry in the glacier system and variable geometry in the subaerial system as they respond to changes in water discharge (Equations~\ref{eq:tau_t} and \ref{eq:v}).
+The relationship between water discharge variability and relative variability of the water velocity, shear stress, and width-integrated shear stress results from the slowly evolving channel geometry in the glacier system and variable geometry in the subaerial system as they respond to changes in water discharge (Equations~\ref{eq:tau_t} and \ref{eq:v}).
 Increased variability in sediment transport capacity with respect to discharge has several implications for sediment transport in subglacial systems compared to subaerial ones.
-Shear stress is scaled to the $\frac{3}{2}$ power and relies on shear stress exceeding a threshold in sediment transport relationships  such as in \citeA{meyer1948}.
-In the total sediment transport relationship \citeA{engelund1967}, the shear stress is scaled to $\frac{5}{2}$ power.
-In both cases, the exponent greater than $1$ of shear stress in the sediment transport relationship magnifies the variability in sediment transport beyond the highly variable shear stress and total shear stress described above in subglacial environments (Figure~\ref{fig:Qw_vari}.
+Shear stress is scaled to the $\frac{3}{2}$ power and relies on shear stress exceeding a threshold in sediment transport relationships  such as in \citet{meyer1948}.
+In the total sediment transport relationship \citet{engelund1967}, the shear stress is scaled to $\frac{5}{2}$ power.
+In both cases, the exponent greater than $1$ of shear stress in the sediment transport relationship magnifies the variability in sediment transport beyond the highly variable shear stress and total shear stress described above in subglacial environments (Figure~\ref{fig:Qw_vari}).
 
-Both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress of water across the channel bed is homogeneous \cite<Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}; e.g.
->{yager2018}.
-In the subaerial parameterization, this means that sediment is transported across all channel widths, as opposed to large flows such as floods when much sediment transport occurs \cite{wolman1960}.
+Both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress of water across the channel bed is homogeneous \citep[Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}; e.g.
+]{yager2018}.
+In the subaerial parameterization, this means that sediment is transported across all channel widths, as opposed to large channel widths during floods when much sediment transport occurs \citep{wolman1960}.
 As a result, the  the subaerial systems variability in sediment transport capacity across the range of discharge values could be underestimated in the parameterization here (Figure~\ref{fig:Qw_vari}).
-Furthermore, the similar variance of total shear stress shows (Figure~\ref{fig:multi_run}) that subglacial channels contain self-organizing feedbacks that limit the mobilization of sediment, as in fluvial systems \cite{phillips2016}.
+Furthermore, the similar variance of total shear stress in some cases shows (Figure~\ref{fig:multi_run}) that subglacial channels contain self-organizing feedbacks that limit sediment transport capacity, as in fluvial systems \citep{phillips2016}.
+
+In natural systems, variable mobilization and deposition of sediment may accumulate as water and sediment accumulates through the two-dimensional channel network will compound the already substantial variability in sediment mobilization parameters accounted for here in the non-dimensional parameterization.
+Such variability could be especially strong in the subglacial system \citep{delaney2023}.
 
@@
 
-The response of sediment discharge caused by the greater variability in sediment transport in subglacial channels will be greatest in completely transport-limited regimes \cite<e.g.>[]{kasmalkar2019}.
-In supply-limited systems, the changing sediment transport capacity (Figure~\ref{fig:model_outs}) would have a minimal effect on sediment discharge due to the absence of sediment available for transport \cite{delaney2019}.
+The response of sediment discharge caused by the greater variability in sediment transport in subglacial channels will be greatest in completely transport-limited regimes \citep[e.g.][]{kasmalkar2019}.
+In supply-limited systems, the changing sediment transport capacity (Figure~\ref{fig:model_outs}) would have a minimal effect on sediment discharge due to the absence of sediment available for transport \citep{delaney2019}.
 Yet, in the subglacial system reaching the threshold of motion for sediment frequently, across a range of discharges,  may result in additional sediment transport.
-The variable sediment transport capacity may result in a tendency for glaciers to evacuate sediment and transition to a supply-limited regime whereby sediment discharge scales with glacier sliding  \cite<e.g.>[]{herman2015}.
-Sediment exhaustion by this process may also explain the dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, as opposed to glacier melt \cite{overeem2017}.
+The variable sediment transport capacity may result in a tendency for glaciers to evacuate sediment and transition to a supply-limited regime whereby sediment discharge scales with glacier sliding  \citep[e.g.][]{herman2015}.
+Sediment exhaustion by this process may also explain the dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, as opposed to glacier melt \citep{overeem2017}.
 
-The effect of pressurized channel flow could be most pronounced during precipitation or flood events when subglacial conduits have not adapted to the water discharge flowing through them, resulting in a large increase in sediment transport capacity and discharge \cite<e.g.>[]{cowan1988,delaney2019}.
+The effect of pressurized channel flow could be most pronounced during precipitation or flood events when subglacial conduits have not adapted to the water discharge flowing through them, resulting in a large increase in sediment transport capacity and discharge \citep[e.g.][]{cowan1988,delaney2019}.
 Exceptional sediment discharge during floods may result from changes to sediment access below the glacier, yet rapid increases in water velocity from water flowing through a small channel will also cause a large increase in sediment transport capacity, not possible in a subaerial system (Section~\ref{sect:sub_mode}).
 
-Increased variability in sediment transport may also impact sediment export from the subglacial environment into proglacial areas \cite<e.g.>[]{perolo2018}.
-Here, velocities at the highest water discharges could generally decrease (Figure~\ref{fig:model_outs}, blue and green horizontal bars), in the transition of pressurized subglacial flow to subaerial open-channel flow as the water leaves the glacier.
-The decrease in water velocity potentially results in sediment deposition once sediment enters the proglacial area.
-Periodic sediment deposition in sediment deposition may even result in the transition between pressurized and subaerial flow below the glacier terminus \cite{perolo2018}, especially when cavity closure rates of the ice are slow \cite{egli2021b}.
+Increased variability in sediment transport may also impact sediment export from the subglacial environment into proglacial areas \citep[e.g.][]{perolo2018}.
+The decrease in sediment transport capacity potentially results in sediment deposition once sediment enters the proglacial area.
+Periodic sediment deposition in sediment deposition may even result in the transition between pressurized and subaerial flow below the glacier terminus \citep{perolo2018}, especially when cavity closure rates of the ice are slow \citep{egli2021b}.
 
 In large catchments and further downstream of proglacial margins, water discharge is generally less variable than at the tops of catchments, where glaciers lie in
-alpine settings \cite<c.f.>{swift2005,riihimaki2005,costa2017,vanas2017}.
+alpine settings \citep[][]{swift2005,riihimaki2005,costa2017,vanas2017}.
 Here, external hydrological factors further downstream of glaciers could result in smaller variations in sediment transport capacity in many subaerial systems compared to subglacial ones.
@@
 The pressurized nature of subglacial water means that the relationship between water discharge and suspended sediment concentration, or sediment discharge could be disparate, especially over the timescales examined here.
-The rapid increase and decreases in sediment transport capacity, for example over a diurnal cycle, in a subglacial system may compound the variability in sediment transport already present in subaerial systems \cite<Figure~\ref{fig:Qw_vari}; >{williams1989,jerolmack2010}.
-Large variations in sediment transport capacity below glaciers may result in sediment mobilization and deposition processes in close spatial and temporal proximity to each other below the glacier \cite{gimbert2016,perolo2018}.
-As a result, variations in sediment transport capacity could be responsible for the fluctuations in sediment discharge (``flushings'') from glaciers that are not attributed to variations in water discharge \cite<e.g.>[]{richards2003,swift2021}.
+The rapid increase and decreases in sediment transport capacity, for example over a diurnal cycle, in a subglacial system may compound the variability in sediment transport already present in subaerial systems \citep[Figure~\ref{fig:Qw_vari}; ][]{williams1989,jerolmack2010}.
+Large variations in sediment transport capacity below glaciers may result in sediment mobilization and deposition processes in close spatial and temporal proximity to each other below the glacier \citep{gimbert2016,perolo2018}.
+As a result, variations in sediment transport capacity could be responsible for the fluctuations in sediment discharge (``flushings'') from glaciers that are not attributed to variations in water discharge \citep[e.g.][]{richards2003,swift2021}.
 Aside from the sporadic nature of these events, the observed ``flushings'' could result from increases in water velocity resulting from reduced channel size, not water discharge, which increase sediment transport capacity (Figure~\ref{fig:Qw_vari}).
 
-Additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to subaerial systems \cite{jansson2005,ganti2016}. 
-For instance, sediment transport capacity from a glacier reacts to the ice thickness, controlling the closure of the channel, and the surface slope of the glacier, controlling water velocity \cite<Section~\ref{sect:sub_mode}; >{rothlisberger1972,shreve1972,delaney2022,stevens2022}, in addition to water discharge.
-Conversely, sediment transport in most subaerial systems increases with water discharge and hydraulic gradient, and the hydraulic gradient probably remains relatively stable over yearly to century-scale time periods \cite<Section~\ref{sect:fluv}; e.g.>{muller1968,whipple1999,wong2006,wickert2019}.
+Additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to subaerial systems \citep{jansson2005,ganti2016}. 
+For instance, sediment transport capacity from a glacier reacts to the ice thickness, controlling the closure of the channel, and the surface slope of the glacier, controlling water velocity \citep[Section~\ref{sect:sub_mode}; ][]{rothlisberger1972,shreve1972,delaney2022,stevens2022}, in addition to water discharge.
+Conversely, sediment transport in most subaerial systems increases with water discharge and hydraulic gradient, and the hydraulic gradient probably remains relatively stable over yearly to century-scale time periods \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wong2006,wickert2019}.
 
-Records of sediment discharge have been used to establish the relationship, or lack thereof, between sediment discharge and climate in glacial systems \cite<e.g.>[]{koppes2009a,willenbring2016,mariotti2021}.
+Records of sediment discharge have been used to establish the relationship, or lack thereof, between sediment discharge and climate in glacial systems \citep[e.g.][]{koppes2009a,willenbring2016,mariotti2021}.
 The variability in subglacial sediment transport capacity presented here applies most generally to short-time scales responsible for the size of subglacial channels (Figure~\ref{fig:multi_run}).
-Yet, identifying climatic signals in sediment transport from transport-limited glaciers may require higher thresholds of climatic perturbation compared to subaerial systems \cite{tofelde2021}.
-This higher threshold of climatic perturbation may persist in addition to the stochastic nature of erosion and deposition in fluvial environments \cite{castletort2003,jerolmack2010,romans2016}.
+Yet, identifying climatic signals in sediment transport from transport-limited glaciers may require higher thresholds of climatic perturbation compared to subaerial systems \citep{tofelde2021}.
+This higher threshold of climatic perturbation may persist in addition to the stochastic nature of erosion and deposition in fluvial environments \citep{castletort2003,jerolmack2010,romans2016}.
+
 Results here suggest that increased variability in sediment transport in subglacial systems  could be particularly pronouced on ice sheets with large catchments and reduced variability.
-For instance, sediment discharge records from two glaciers in the Swiss Alps show that $40$ to $50$\% of a season's sediment discharge occurs when water discharge is below the $75^{\mathrm{th}}$ quantile of the season \cite{delaney2018}.
+Furthermore, total shear stress can remain high across a range of water discharges (Figures~\ref{fig:model_outs} and \ref{fig:Qw_vari}).
+In fact, sediment discharge records from two glaciers in the Swiss Alps show that $40$ to $50$\% of a season's sediment discharge occurs when water discharge is below the $75^{\mathrm{th}}$ quantile of the season \citep{delaney2018}.
 Such a large quantity of sediment transported at lower water discharges may occur in part because sediment transport capacity responds to the size of the glacier conduit, in addition to water discharge (Figure~\ref{fig:model_outs} and Section~\ref{sect:sub_mode}).
 
-This suggests that a larger climatic perturbations in water discharge might be needed to isolate significant changes sediment discharge from glacial systems compared to subaerial on, as the noise resulting from increased variability in sediment transport capacity must be overcome.
-Furthermore, if no sediment is available and the glacier is in a supply-limited case, then the sediment discharge record will also represent additional processes of bedrock erosion from  water pressure variations and sliding  \cite{iverson2012,herman2015}, which are also complex.
+This suggests that larger climatic perturbations that isolate  water discharge might be needed to isolate significant changes sediment discharge from glacial systems compared to subaerial on, as the noise resulting from increased variability in sediment transport capacity must be overcome.
+Furthermore, if no sediment is available and the glacier is in a supply-limited case, then the sediment discharge record will also represent additional processes of bedrock erosion from  water pressure variations and sliding  \citep{iverson2012,herman2015}, which are also complex.
 
@@
 Subaerial channels can alter their channel width and water velocity in response to changing water discharge.
-Conversely, pressurized subglacial channels accommodate changing water discharge by altering water velocity and shear stress, upon which sediment transport depends.
+Conversely, pressurized subglacial channels mostly accommodate changing water discharge by altering water velocity and shear stress, upon which sediment transport depends.
 This occurs because the size of subglacial channels evolves slowly compared to variations in water discharge.
@@
 Parameterizations of subglacial and subaerial water flow show that subglacial systems exhibit increased variability in sediment transport capacity as they respond  to changes in water discharge.
-The variability in sediment transport capacity is reduced when the evolving channel width is accounted for in subaerial systems.
-Even so, variability in sediment transport capacity across a channel's width is consistently higher in subglacial channels compared to subaerial ones.
+Variability in sediment transport capacity across a channel's width can be higher in subglacial channels compared to subaerial ones, but also the variability might not be substantially different.
+Furthermore, the timing of peak sediment discharge conditions with respect to water discharge occurs on the rising limb of the hydrograph in subglacial systems, as opposed to the peak in subaerial systems.
 
@@
 
-\section*{Open Research}
+\section*{Acknowledgments}
 \noindent
+I was funded by SNSF Project No. PZ00P2\_202024.
 The julia code used herein is included in the supplementary material.
-
-\acknowledgments
-I was funded by SNSF Project No. PZ00P2\_202024.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-
-\bibliography{Paperlib.bib}
+\bibliographystyle{apa} 
+\bibliography{PaperLib.bib}
 

### 2023-07-13 Delaney2023.tex
Updates from Overleaf

--- 2023-07-10 Delaney2023.tex
@@
 \usepackage{multirow}
-\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
+\usepackage[finalnew]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
 \usepackage{soul}
@@
 
-%%% Suggested section heads:
-\ian{comparison with Alley 1997... mention his variability-- also deal with his $Qs \propto Qw^{\frac{9}{2}}$  treats like rivers?  does not consider channel size evolution}
-
 \section{Introduction}
@@
 Over long periods, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from below glaciers \cite<c.f.>{hallet1979,iverson2012,ugelvig2018}.
-On shorter time periods, pressurized water transports sediment from the below glacier \cite{walder1994,creyts2013,beaud2018}, should enough sediment be present below the glacier (i.e.
-in a transport-limited regime).
+On shorter time periods, pressurized water transports sediment from the below glacier \cite{walder1994,creyts2013,beaud2018}, should enough sediment be present below the glacier.
+In an end-member case, enough sediment is available below the glacier so that its export is limited only by hydraulic conditions below the glacier (i.e.
+transport-limited regime).
 
-In a transport-limited regime, sediment discharge responds to the sediment transport capacity or the amount of sediment that could be carried by the water.
+In a transport-limited regime, sediment discharge responds to the sediment transport capacity or the amount of sediment that the water could carry.
 Sediment transport capacity depends on the shear stress between water and sediment it flows  over \cite{shields1936,meyer1948,engelund1967} and the width of the channel bottom $w_c$ over which to mobilize sediment.
@@
 These differences have been implicitly included in a wide range of available models that  quantify sediment transport from catchments in both subglacial and subaerial systems  \cite<e.g.>[]{walder1994,tucker1997,creyts2013,wickert2019,hewitt2019}.
-However, some suggest that subglacial sediment discharge  $Q_s \propto Q_w^{\frac{9}{2}}$, assuming that water discharge, changing channel width and depth are each proportionally responsible for the variance of water velocity and sediment transport capacity \cite{alley1997,swift2021}.
+However, some suggest that subglacial sediment discharge  $Q_s \propto Q_w^{\frac{9}{2}}$, assuming that water discharge, changing channel width, and depth are each proportionally responsible for the variance of water velocity and sediment transport capacity \cite{alley1997,swift2021}.
 The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \cite<e.g.>[]{lane2016,perolo2018}.
@@
 \end{table}
-  $^a$ for Fieschergletscher
+  $^a$ for Fieschergletscher (\alpine case)
 
-  $^b$ for Leverett glacier
+  $^b$ for Leverett glacier (\icesheet case)
 
@@
 Outputs of the parameterizations are meant to represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
-To generalize these cases, the Fieschergletscher case is referred to as \alpine{} and is exemplified by relatively thin ice thickness ($h_o$: xxx), low discharge (< $20$\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in discharge.
-The Leverett glacier case is referred to as \icesheet{} and is exemplified by thick ice  ($h_o$: xxx), high water discharge (< $300$\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in discharge.
+To generalize these cases, the Fieschergletscher case is referred to as \alpine{} and is exemplified by relatively thin ice thickness ($h_o$: $225$\unit{m}), low discharge (< $20$\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in discharge.
+The Leverett glacier case is referred to as \icesheet{} and is exemplified by thick ice  ($h_o$: $700$\unit{m}), high water discharge (< $300$\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in discharge.
 
@@
     \includegraphics[width=0.8\linewidth]{model_outputs.pdf}
-    \caption{Model outputs as in response to hydrograph (a and e) from Fieschergletscher (a-d) and Leverett glacier (e-h).
+    \caption{Model outputs as in response to hydrograph (a and e) from \alpine (a-d) and \icesheet (e-h).
 Gray line in a and e represent subglacial channel cross sectional area.
@@
     \includegraphics[width=0.8\linewidth]{Qw_vari.png}
-    \caption{Relationship between water discharge and velocity, shear stress and width-integrated shear stress for the Fieschergletscher hydrograph (a-c) and Leverett glacier (d-f).
+    \caption{Relationship between water discharge and velocity, shear stress and width-integrated shear stress for the \alpine hydrograph (a-c) and \icesheet (d-f).
 Variables have been normalized to by the mean values for better comparison amongst the model runs.}

### 2023-07-13 Delaney2023.tex
small grammer bits

--- 2023-07-13 Delaney2023.tex
@@
 
-\item Greater variability in water velocity in subglacial systems causes sediment transport capacity to vary more than subaerial systems.
+\item Greater variability in water velocity in subglacial systems causes sediment transport capacity to vary more than in subaerial systems.
 
-\item The systems' divergence in transport response may impact sediment dynamics at glacier margins and interpretation of sediment records.
+\item The systems' divergence in transport response may impact sediment dynamics at glacier margins and the interpretation of sediment records.
 \end{itemize}
@@
 
-Recheck results figure names.
+%Recheck results figure names.
 
@@
 
-In a transport-limited regime, sediment discharge responds to the sediment transport capacity or the amount of sediment that could be carried by the water.
+In a transport-limited regime, sediment discharge responds to the sediment transport capacity or the amount of sediment that the water could carry.
 Sediment transport capacity depends on the shear stress between water and sediment it flows  over \citep{shields1936,meyer1948,engelund1967} and the width of the channel bottom $w_c$ over which to mobilize sediment.
@@
 These differences have been implicitly included in a wide range of available models that  quantify sediment transport from catchments in both subglacial and subaerial systems  \citep[e.g.][]{walder1994,tucker1997,creyts2013,wickert2019,hewitt2019}.
-However, some suggest that subglacial sediment discharge  $Q_s \propto Q_w^{\frac{9}{2}}$, assuming that water discharge, changing channel width and depth are each proportionally responsible for the variance of water velocity and sediment transport capacity \citep{alley1997,swift2021}.
+However, some suggest that subglacial sediment discharge  $Q_s \propto Q_w^{\frac{9}{2}}$, assuming that water discharge, changing channel width, and depth are each proportionally responsible for the variance of water velocity and sediment transport capacity \citep{alley1997,swift2021}.
 The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \citep[e.g.][]{lane2016,perolo2018}.
@@
     Shear stress (glacier, subaerial) & $\tau_g,\,\tau_f$&& $\mathrm{Pa \, m^{-1}}$ \\
-    Width-integrated shear-stress (glacier, subaerial) & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$ \\
+    Width-integrated shear stress (glacier, subaerial) & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$ \\
 
@@
 Here, it is assumed that the water is transported through a subglacial channel flowing down the glacier; \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972}, and that the channel  size responds to frictional heating from water flow and creep closure by ice, using the Darcy-Weisbach formulation for water-flow through a pipe  \citep[e.g.][]{rothlisberger1972,clarke2003}.
-The formulation here does not consider englacial storage of water, and thus that changes in hydraulic head at the top of the glacier are negligable in the evolution of the channel size. 
+The formulation here does not consider the englacial storage of water, and thus that changes in the hydraulic head at the top of the glacier are negligible in the evolution of the channel size. 
 The evolution of subglacial channel size $S_g$ is given as
@@
 
-With knowledge of cross sectional area $S$, the shear stress, $\tau$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
+With knowledge of cross-sectional area $S$, the shear stress, $\tau$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
 \begin{equation}
@@
 The parameterizations are first applied to a reference test case for each glacier that assumes a semi-circular subglacial channel ($\beta=\pi$) and friction factors $f_r$ and $f_p$ are tuned so that the model reproduces reasonable velocities for both  the Greenland Ice Sheet and alpine glaciers \citep[$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$][]{werder2010b,chandler2013}.
-Water velocity, shear-stress and width-integrated shear-stress in the subglacial and subaerial systems are presented below.
+Water velocity, shear stress, and width-integrated shear stress in the subglacial and subaerial systems are presented below.
 
-To characterize the variability in sediment discharge capacity in subglacial compared to subaerial systems further, the parameterizations are applied to range of different channel geometry factors ($\beta$ and $k$) and friction factors $f_r$ and $f_p$.
+To characterize the variability in sediment discharge capacity in subglacial compared to subaerial systems further, the parameterizations are applied to a range of different channel geometry factors ($\beta$ and $k$) and friction factors $f_r$ and $f_p$.
 Model runs with parameter combinations  are culled if their maximum subglacial water velocity exceeds $2\,$\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity exceeds $1\,$\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with some dye tracing measurements \citep[e.g.]{werder2010b,chandler2013}.
-The standard deviation of water velocity, shear-stress and width-integrated shear-stress across the season long hydrographs are presented below.
+The standard deviation of water velocity, shear stress, and width-integrated shear stress across the season-long hydrographs are presented below.
 
@@
     \caption{Model outputs as in response to hydrograph (a and e) from \alpine{} (a-d) and \icesheet{} (e-h).
-Gray line in a and e represent subglacial channel cross sectional area.
-Below this black lines represent model outputs of velocity, shear-stress and width-integrated shear-stress from the subglacial system, while grey lines represent the subaerial system.
+Gray lines in a and e represent the subglacial channel cross-sectional area.
+Below this blue lines represent model outputs of velocity, shear stress, and width-integrated shear stress from the subglacial system, while orange lines represent the subaerial system.
       Plots are shown at $15$\unit{min} interval.
@@
 
-Following the parameterizations defined in Sections~\ref{sect:sub_mode}~and~\ref{sect:fluv}, the water velocity, shear-stress and total shear-stress exhibit different seasonal evolutions and peaks of water velocity and shear stress (representing the potential for sediment mobilization) and total shear stress (representing the total sediment transport capacity; Figure~\ref{fig:model_outs}).
+Following the parameterizations defined in Sections~\ref{sect:sub_mode}~and~\ref{sect:fluv}, the water velocity, shear stress, and total shear stress exhibit different seasonal evolutions and peaks of water velocity and shear stress (representing the potential for sediment mobilization) and total shear stress (representing the total sediment transport capacity; Figure~\ref{fig:model_outs}).
 
@@
     \includegraphics[width=0.8\linewidth]{Fig3.png}
-    \caption{Relationship between water discharge and velocity, shear stress and width-integrated shear stress for the \alpine{} hydrograph (a-c) and \icesheet (d-f).
-Variables have been normalized to by the mean values for better comparison amongst the model runs.
+    \caption{Relationship between water discharge and velocity, shear stress, and width-integrated shear stress for the \alpine{} hydrograph (a-c) and \icesheet (d-f).
+Variables have been normalized to mean values for better comparison amongst the model runs.
       Plots are shown at $15$\unit{min} interval.} 
@@
 
-For each water discharge value in the subaerial system, there is a single water velocity, shear-stress and total shear-stress (Figure~\ref{fig:Qw_vari}).
-Conversely in the subglacial system,  water velocity, shear-stress and total shear-stress vary substantially for a given water discharge (Figure~\ref{fig:Qw_vari}).
-For instance, over $15$ minute time periods at \alpine{}, the range of water velocities and shear-stress spans from those below the mean to those almost $2$ times the mean velocity.
-Substantially high values of water velocity and shear-stress can occur from minimal water discharge at $2.5$\,\unit{m}$^3$\,\unit{s}$^{-1}$ to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
+For each water discharge value in the subaerial system, there is a single water velocity, shear stress and total shear stress (Figure~\ref{fig:Qw_vari}).
+Conversely in the subglacial system,  water velocity, shear stress and total shear stress vary substantially for a given water discharge (Figure~\ref{fig:Qw_vari}).
+For instance, over $15$ minute time periods at \alpine{}, the range of water velocities and shear stress spans from those below the mean to those almost $2$ times the mean velocity.
+Substantially high values of water velocity and shear stress can occur from minimal water discharge at $2.5$\,\unit{m}$^3$\,\unit{s}$^{-1}$ to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
 At \icesheet, reduced diurnal variability in the hydrograph  (Figure~\ref{fig:model_outs} e) results in smaller variations in water velocity and shear stress compared to \alpine{} (Figure~\ref{fig:Qw_vari}).
@@
 
-When accounting for the evolving channel width of the subglacial conduit, total shear-stress generally increases with water discharge.
-Yet the total shear-stress across the can very substantially, with the highest values in total shear-stress occurring at water discharge values ranging from roughly $11$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
-The variance is less at the \icesheet case, yet a range of total shear-stress can occur at a given discharge(Figure~\ref{fig:Qw_vari} c, e).
+When accounting for the evolving channel width of the subglacial conduit, total shear stress generally increases with water discharge.
+Yet the total shear stress across the can vary substantially, with the highest values in total shear stress occurring at water discharge values ranging from roughly $11$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
+The variance is less at the \icesheet case, yet a range of total shear stress can occur at a given discharge(Figure~\ref{fig:Qw_vari} c, e).
 
@@
 Conversely, in the subaerial system peaks in these variables occur when water discharge is at its highest value.
-As a result, peaks in the subglacial variables occur prior to the peaks in the subaerial variables, representing lags sediment mobilization and  transport capacity when sediment is transferred from on system to the other.
+As a result, peaks in the subglacial variables occur prior to the peaks in the subaerial variables, representing lags in sediment mobilization and  transport capacity when sediment is transferred from one system to the other.
 
@@
     \caption{Variance (standard deviation) in water velocity, shear stress and total shear stress against aggregation time for the subglacial (blue) and subaerial (orange) systems from the \alpine{} (a-c) and \icesheet{} (d-f).
-      Shaded areas denote the $25^{\mathrm{th}}$ and $75^{\mathrm{th}}$ quantiles of the $100$ parameter combinations for each the subglacial and subaerial systems.
+      Shaded areas denote the $25^{\mathrm{th}}$ and $75^{\mathrm{th}}$ quantiles of the $100$ parameter combinations for each of the subglacial and subaerial systems.
       Solid lines denote the mean values.
@@
 However, in the \alpine{} system accounting for channel width makes the variability in the systems comparable (Figure~\ref{fig:multi_run}).
-Yet, the variability over time periods less than one half day, the variability in the subglacial system is somewhat higher. 
+Yet, the variability over time periods is less than one-half day, and the variability in the subglacial system is somewhat higher. 
 Conversely, in \icesheet{}, with lower water discharge variability, the variability in total shear stress remains higher in the subglacial system across the time aggregation periods.
 
-In all cases, the variability in velocity, shear stress and total shear stress decreases after $1-5$ day periods (Figure~\ref{fig:multi_run}).
-This suggests that the effects of the variable response to hydrologic forcing in these cases is reduced after these time periods.
+In all cases, the variability in velocity, shear stress, and total shear stress decrease after $1-5$ day periods (Figure~\ref{fig:multi_run}).
+This suggests that the effects of the variable response to hydrologic forcing in these cases are reduced after these time periods.
 Yet, substantial differences persist in periods of up to $30$ days, as the variability might not converge given seasonal variations in water discharge.
@@
 In the subaerial parameterization, this means that sediment is transported across all channel widths, as opposed to large channel widths during floods when much sediment transport occurs \citep{wolman1960}.
-As a result, the  the subaerial systems variability in sediment transport capacity across the range of discharge values could be underestimated in the parameterization here (Figure~\ref{fig:Qw_vari}).
+As a result, the  subaerial systems variability in sediment transport capacity across the range of discharge values could be underestimated in the parameterization here (Figure~\ref{fig:Qw_vari}).
 Furthermore, the similar variance of total shear stress in some cases shows (Figure~\ref{fig:multi_run}) that subglacial channels contain self-organizing feedbacks that limit sediment transport capacity, as in fluvial systems \citep{phillips2016}.
 
-In natural systems, variable mobilization and deposition of sediment may accumulate as water and sediment accumulates through the two-dimensional channel network will compound the already substantial variability in sediment mobilization parameters accounted for here in the non-dimensional parameterization.
+In natural systems, variable mobilization and deposition of sediment may accumulate as water and sediment accumulate through the two-dimensional channel network will compound the already substantial variability in sediment mobilization parameters accounted for here in the non-dimensional parameterization.
 Such variability could be especially strong in the subglacial system \citep{delaney2023}.
@@
 
-Results here suggest that increased variability in sediment transport in subglacial systems  could be particularly pronouced on ice sheets with large catchments and reduced variability.
+Results here suggest that increased variability in sediment transport in subglacial systems  could be particularly pronounced on ice sheets with large catchments and reduced variability.
 Furthermore, total shear stress can remain high across a range of water discharges (Figures~\ref{fig:model_outs} and \ref{fig:Qw_vari}).
@@
 
-This suggests that larger climatic perturbations that isolate  water discharge might be needed to isolate significant changes sediment discharge from glacial systems compared to subaerial on, as the noise resulting from increased variability in sediment transport capacity must be overcome.
+This suggests that larger climatic perturbations that isolate  water discharge might be needed to isolate significant changes in sediment discharge from glacial systems compared to subaerial on, as the noise resulting from increased variability in sediment transport capacity must be overcome.
 Furthermore, if no sediment is available and the glacier is in a supply-limited case, then the sediment discharge record will also represent additional processes of bedrock erosion from  water pressure variations and sliding  \citep{iverson2012,herman2015}, which are also complex.

### 2023-08-07 Delaneyetal2023b.tex
working on edits

\documentclass[11pt]{article}
\usepackage [colon,square]{natbib}
\usepackage {comment}
\usepackage{cite}
\usepackage[mathlines]{lineno} %% <- mathlines turns on line numbering in equations
\usepackage{amsmath} 
\usepackage{printlen}
\setlength{\parindent}{10pt}
\setlength{\parskip}{5pt}
\usepackage[explicit]{titlesec}
\usepackage{helvet}
\usepackage{array}
\usepackage{printlen}
\usepackage[a4paper, margin=1.25cm]{geometry}
\usepackage{color,hyperref}
\catcode`\_=11\relax
\newcommand\email[1]{\_email #1\q_nil}
\def\_email#1@#2\q_nil{%
  \href{mailto:#1@#2}{{\emailfont #1\emailampersat #2}}
}
\newcommand\emailfont{\sffamily}
\newcommand\emailampersat{{\color{red}\small@}}
\catcode`\_=8\relax

\usepackage{hyperref}
\usepackage[export]{adjustbox}
\usepackage{bibentry}
\renewcommand{\bibfont}{\footnotesize}
\usepackage{tikz}
\usepackage{setspace}
\usepackage[inline,shortlabels]{enumitem}
\usepackage[all]{nowidow}
\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
\pagenumbering{gobble} 
\hypersetup{ % Remove ugly boxes around URLs
  colorlinks, linkcolor={black!100!black}, citecolor={black!50!black}, urlcolor={black!80!black}
}

\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\newcolumntype{C}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}

\linenumbers
\sloppy
\hbadness=99999

\newcommand{\aref}[1]{\textbf{Reference #1}}
\newcommand{\TODO}[1]{\textbf{TODO: \color{red}#1}}
\newcommand{\ian}[1]{{\textbf{\color{blue}Ian says:} \color{blue} #1} }
\newcommand{\alpine}{\textit{ALPINE}\,}
\newcommand{\icesheet}{\textit{ICESHEET}\,}
\newcommand{\m}{$\,\mathrm{m}$\,}
\newcommand{\cm}{$\,\mathrm{cm}$\,}
\newcommand{\mma}{$\,\mathrm{mm  \, a^{-1}}$\,}
\newcommand{\mmma}{$\,\mathrm{m^3\, a^{-1}}$\,}
\newcommand{\mmms}{$\,\mathrm{m^3\, s^{-1}}$\,}
\newcommand{\unit}[1]{$\mathrm{#1}$}

\begin{document}
%% ------------------------------------------------------------------------ %%
% Title
% 
% (A title should be specific, informative, and brief. Use
% abbreviations only if they are defined in the abstract. Titles that
% start with general keywords then specific terms are optimized in
% searches)
% 
%% ------------------------------------------------------------------------ %%

% Example: \title{This is a test title}

\begin{center}
\Large{\textbf{Water discharge quantity has a variable impact on sediment transport capacity in subglacial channels}}
\normalsize

Ian Delaney\footnote{Institut des dynamiques de la surface terrestre (IDYST), Universit\'{e} de Lausanne, B\^{a}timent G\'{e}opolis, 1015 Lausanne, Switzerland

ianarburua.delaney@unil.ch}, Daniel Farinotti\footnote{Laboratory of Hydraulics, Hydrology and Glaciology (VAW), ETH-Z\"urich, H\"onggerbergring 26, 8093 89 Z\"urich, Switzerland}$^{,}$\footnote{Swiss Federal Institute for Forest, Snow and Landscape Research (WSL) Z\"uricherstrasse 111, 8903 1011 Birmensdorf, Switzerland}, Andrew J. Tedstone\footnote{Department of Geosciences, University of Fribourg, Ch. du Musée 1700, Fribourg, Switzerland}

\end{center}
\paragraph{Keypoints}
\begin{itemize}
\item Water discharge variations in subglacial channels are mainly accommodated by water velocity, not channel size, given their subglacial channel's slow evolution.

\item Greater variability in water velocity in subglacial channels causes sediment transport capacity to vary more than in subaerial channels.

\item The channels' divergence in transport response may impact sediment dynamics at glacier margins and the interpretation of sediment records.
\end{itemize}

\TODO{Study site description}

\TODO{Consider removing ``case'' ``system'', around \alpine and \icesheet to reduce words}

\begin{abstract} % 150 words

\noindent
In both subglacial and subaerial channels, sediment transport capacity of a sediment size depends mainly on 1) the channel width over which the sediment is mobilized  and 2) water velocity, as this controls the shear stress that the flowing water applies to the channel bottom.
In subaerial channels, changing water discharge  changes the water depth,  the channel width, and the  water velocity.
In subglacial environments and on time scales of hours to weeks, however, changing water discharge  mainly alters the  water velocity, not channel area or width, because the subglacial conduit maintains a largely fixed geometry.
Here, parameterizations of sediment transport capacity for pressurized flow under glaciers and subaerial open channel flow are applied to hydrographs from an Alpine glacier and the Greenland Ice Sheet.
Results of these parameterizations show that sediment transport capacity varies far more in subglacial channels compared to subaerial channels both over time and with respect to water discharge.
However, changing channel width can accommodate some   variability in sediment transport capacity in subaerial channels.
The different response of water discharge on sediment transport capacity in subglacial channels, compared to subaerial ones, may impact the occurence of sediment transport capacity reaching the threshold of mobilization, thus should be considered in the interpretation of sediment discharge records with respect to water discharge.

%\ian{last sentence is a bit rough}

\end{abstract}

\section{Introduction}
\label{sect:intro}

Changes in glacier dynamics, geomorphology, and hydrology  have prompted numerous  recent studies of  sediment transport processes in cold regions \citep[e.g.][]{zhang2022}.
Increases in sediment transport  have been observed in Greenland \citep{bendixen2017}, the European Alps \citep{costa2017}, the Himalayas \citep{li2021}, and the Andes \citep{vergara2022}.
In some of these regions, increased water discharge and glacier melt have been interpreted to be greater sediment transport capacity \citep{bendixen2017,costa2017,li2021}.
Observed changes to sediment transport in glacierized catchments call for examining the processes controlling sediment discharge in these catchments. 
Yet, the relationship between water discharge and sediment transport in subglacial environments remains minimally discussed and not compared with the subaerial river systems that these glaciers feed.

\TODO{Add Riihimaki and Swift}

Over long periods, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from under glaciers \citep[c.f.][]{hallet1979,iverson2012,ugelvig2018}. 
On shorter time periods, pressurized water transports sediment from the under glacier \citep{walder1994,creyts2013,beaud2018}, should enough sediment be present subglacially (i.e. in a transport-limited regime).

In a transport-limited regime, sediment discharge responds to the sediment transport capacity or the amount of sediment that the water could carry. Sediment transport capacity depends on the shear stress between water and sediment it flows  over \citep{shields1936,meyer1948,engelund1967} and the width of the channel bottom $w_c$ over which to mobilize sediment. The shear stress  across the channel width $\tau_t$ responds to the velocity of water $v$ flowing through the channel so that 
\begin{linenomath*}
  \begin{equation}
    \label{eq:tau_t}
    \tau_t \propto w_c\, v^2,
  \end{equation}
\end{linenomath*}
% 
where $w_c$ is the width of the channel.
Following the conservation of mass, the velocity of the water flowing through a  channel is given as 
\begin{linenomath*}
  \begin{equation}
    \label{eq:v}
    v = \frac{Q_w}{S},
  \end{equation}
\end{linenomath*}
% 
where $Q_w$ is water discharge,  and $S$ is the channel's cross-sectional area.
\TODO{comment on the squared?}

In subaerial channels, with open channel flow, the cross-sectional area $S$ of the channel evolves with changing water discharge $Q_w$  by changing both the channel's width and the water depth \citep{leopold1953}. During a flood with greater water discharge $Q_w$, for instance, the cross-sectional area $S$ increases as the channel widens and deepens, along with water velocity $v$, which is squared. As a result, the shear stress across the channel $\tau_t$ will increase according to Equation~\ref{eq:tau_t}.

Yet, the response of water velocity to changing water discharge in subglacial channels is different from subaerial ones.
The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by frictional heating of water flowing through the channel \citep{rothlisberger1972}.
As a result, in subglacial channels, the water is pressurized by the ice above \citep{rothlisberger1972}.
As a result, the subglacial channel size only evolves over days to months, whereas water discharge can vary over hours \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
In turn, changes in water discharge $Q_w$ are mainly accommodated with water velocity $v$, because the size of the channel $S$ changes on a much shorter timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
It follows that an increase in water discharge $Q_w$ will cause the shear stress across the channel $\tau_t$ to respond to the water velocity $v$, as opposed to the subaerial channel where channel width $w_c$ evolves as well. 
\begin{center}
  \begin{figure}[h]
    \includegraphics[width=0.8\linewidth]{Fig1.pdf}
    \caption{Sketch for the different response that subglacial and subaerial channels have to an increase in water discharge.  Water velocity magnitude in the subglacial, $v_g$, and subaerial, $v_f$, channels are shown by arrow length. $S_g$ and $S_f$ represent the cross-sectional area in subglacial and subaerial channels. The subglacial channel widths $w_{cg}$ remain unchanged, while the subaerial channel widths evolve $w_{cf}$. Shear stress $\tau$ is responsible for the mobilization of sediment and increases with the number of makers at the channel bed. 
      Note that in the subaerial channel parameterization, channels have rectangular shapes (Section~\ref{sect:fluv}).} 
    \label{fig:cartoon}
  \end{figure}
\end{center}
\ian{Come back to Daniel's comments}

As a result, sediment mobilization in subaerial and subglacial channels responds differently to changing water discharge.
These differences have been implicitly included in a wide range of available models that  quantify sediment transport from catchments in both subglacial and subaerial channels  \citep[e.g.][]{walder1994,tucker1997,creyts2013,wickert2019,hewitt2019}.
%However, some modelss suggest that subglacial sediment discharge  $Q_s \propto Q_w^{\frac{9}{2}}$, assuming that water discharge, changing channel width, and depth are each proportionally responsible for the variance of water velocity and sediment transport capacity \citep{alley1997,swift2021}.
The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \citep[e.g.][]{lane2016,perolo2018}.
Additionally, the variable response of sediment transport capacity to water discharge in the two systems may impact the interpretation of records of sediment transport in glacial compared to subaerial systems \citep[e.g.][]{muller1968,richards2003,ganti2016}.
Yet, the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial channels, and implications for sediment discharge records, has been minimally discussed.
Furthermore, thorough observations of water velocity and channel width in subglacial channels are limited and, to these authors' knowledge, not contrasted with subaerial conditions at the glacier margins.

This manuscript aims to evaluate relationship between sediment transport capacity and water discharge in subglacial channels, compared to subaerial ones.
Numerical parameterizations of subglacial and subaerial hydraulics are untilized, in the absence of observations of continuous water velocity measurements and the evolution of subglacial and subaerial channels.
Parameterizations are applied to hydrological records from an Alpine glacier in Switzerland (Fieschergletscher), and  a land-terminating glacier in Greenland (Leverett glacier).
The parameterizations' lumped and simplistic nature is intended to show the disparity between water discharge and sediment transport in subglacial systems  without considering effects such as the upstream drainage network and the role of sediment supply.
Outputs are used to explore differences in the relationships between  water discharge, channel geometry, and water velocity in the subglacial and subaerial channels. 
With these parameterizations, the manuscript examines the impact of water discharge variability on sediment transport capacity in subglacial and subaerial channels.
The manuscript then uses the demonstrated differences between the two channel types to discuss the implications for sediment transport processes in subglacial channels, compared to subaerial ones, and for interpreting records of sediment transport from glacierized catchments.

\section{Methods}
\label{sect:meth}
Parameterizations below (Sections ~\ref{sect:sub_mode}~and~\ref{sect:fluv}) represent relationships amongst water discharge, subglacial water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
Because sediment transport largely depends on shear stress applied by water flow \citep{shields1936}, both parameterizations calculate this value and integrate  it across the channel bed (Figure \ref{fig:cartoon}).
The evaluation of shear stress omits a grain size parameter and, thus, leaves open the selection of a sediment transport relationship  \citep[e.g.][]{shields1936,meyer1948} to  calculate sediment transport capacity.
As a result, relative relationships in shear stress, as opposed to sediment transport capacity, are examined.

\begin{table}[ht]
  \centering
  \caption{Variables, parameters, and constants used in this work. Where two values are given, the first refers to an \alpine{} case (Fieschergletscher) and the second to a glacier marginal to the \icesheet{}case (Leverett Glacier) }
  \begin{tabular}{ l  c  c c }
    Name &Symbol&  Value&Units \\ \hline
    \textbf{Variables}  & & & \\
    Water discharge  & $Q_w$& & $\mathrm{m^{3}\,s^{-1}}$ \\
    Water velocity (glacier, subaerial)  & $v_g,\,v_{f}$& & $\mathrm{m\,s^{-1}}$ \\
    Channel cross-sectional area (glacier, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
    Hydraulic diameter &$D_h$&&$\mathrm{m}$\\
    Width of channel floor (glacial, subaerial) & $w_{cg},w_{cf}$&  & $\mathrm{m}$     \\
    Hydraulic head &$\Delta h$&& $\mathrm{m}$\\
    Shear stress (glacier, subaerial) & $\tau_g,\,\tau_f$&& $\mathrm{Pa \, m^{-1}}$ \\
    Width-integrated shear stress (glacier, subaerial) & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$ \\

         &&&\\

    \textbf{Parameters and Constants}  & & &\\
    Gravitational constant&$g$& $9.81$&$\mathrm{m\,s^{-2}}$\\
    Density of water & $\rho_w$& $1000$ & $\mathrm{kg\,m^{-3}}$ \\
    Density of ice & $\rho_i$& $900$ & $\mathrm{kg\,m^{-3}}$ \\
    Hooke angle of channel & $\beta$ & $\frac{\pi}{2}$ & \unit{rad}\\

    Glacier thickness &$h_{ice}$& $225^a$ or $700^b$  &\unit{m}\\
    Effective glacier thickness &$h_o$&$\frac{\rho_i}{\rho_w} h_{ice}$  &\unit{m}\\
    Effective glacier length &$l$&$7000^a$ or $20,000^b$&\unit{m}\\
    Constant $1$ in Equation~\ref{eq:dS_dt} &$C_1$&$2.2\times10^{-5}$&\unit{m}$^{-1}$\\
    Constant $2$ in Equation~\ref{eq:dS_dt} &$C_2$&$3.7\times10^{-13}$&\unit{m}$^{-n}\,s^{-1}$\\
    Latent heat of fusion &$L$&$333.5 $&\unit{kJ\,kg}$^{-1}$\\
    Pressure melting coefficient &$c_t$&$7.5\times 10^{-8}$&\unit{K\,Pa}$^{-1}$\\
    Specific heat capacity of water &$c_p$&$4180$&\unit{J\,kg}$^{-1}$\unit{K}$^{-1}$\\

    Ice flow constant &$A$& $5.3\times10^{-24}$ &\unit{Pa}$^{-n}$\,$s^{-1}$\\
    Ice flow exponent &$n$& $3$ &$\mathrm{(-)}$\\
    Friction factor (glacier) & $f_r$ & $4$ & $\mathrm{(-)}$ \\
    Friction factor (subaerial) & $f_p$ & $3$ & $\mathrm{(-)}$\\
    Gradient of channel bed (subaerial) &$\frac{\partial z_c}{\partial x}$ &$0.05$& $\mathrm{(-)}$\\
    Subaerial channel factor & $k$ &$3$ & $\mathrm{s\,m^{-2}}$\\
    Channel geometry exponent &$e$& $\frac{1}{2}$&$\mathrm{(-)}$ \\
    \hline
  \end{tabular}
  \label{table:vpm}

%  $^a$ for \alpine{} (Fieschergletscher)

 % $^b$ for \icesheet{} (Leverett glacier)

\end{table}

\subsection{Subglacial channel  parameterization}
\label{sect:sub_mode}

To evaluate the shear stress of water flowing across sediments underneath a glacier, the subglacial channel parameterization takes into account the channel geometry and the velocity of the flowing water.
To do this, a modified version of  the lumped hydraulics model presented in \citet{clarke1996} and \citet{werder2010} is used.

Here, it is assumed that the water is transported through a subglacial channel flowing down the glacier; \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972}, and that the channel  size responds to frictional heating from water flow and creep closure by ice, using the Darcy-Weisbach formulation for water-flow through a pipe  \citep[e.g.][]{rothlisberger1972,clarke2003}.
The formulation here does not consider the englacial storage of water.
Thus, it is assumed that the changing hydraulic head at the top of the glacier is negligible in the evolution of the channel size. 
The evolution of subglacial channel size $S_g$ is given as
\begin{linenomath*}
  \begin{equation}
    \label{eq:dS_dt}
    \frac{\partial S_g}{\partial t} = C_2 \frac{Q \Delta h}{l} - C_2 (h_{o}-\frac{\Delta h}{2})^n\,S_g,
  \end{equation}
\end{linenomath*}
\noindent where $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (Table \ref{table:vpm}), $l$ is the length of subglacial channel, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change from the glacier terminus to its top, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the ice overburden pressure, and $n$ is Glen's n \citep{glen1955}.
The first term on the right side of the equation represents the opening of the channel through frictional heating, while the second term represents the creep closure of the channel from ice deformation.

Following the Darcy-Weisbach, head drop $\Delta h$ is,
\begin{linenomath*}
  \begin{equation}
    \label{eq:dh}
    \Delta h \,  = l \,\frac{1}{g}\,s\,f_r\,\frac{Q^2}{D_h^5}
  \end{equation}
\end{linenomath*}
\noindent where $f_r$ is a friction factor, $D_h$ is the hydraulic diameter, and $l$ is the length overwhich head change occurs. $s$ is a factor accounting for channel geometry \citep{hooke1990}, calculated as:
\begin{equation}
  \label{eq:Hf}
  s = \frac{2\,(\beta -\sin \beta)^2}{(\frac{\beta}{2}\,+\,\sin \frac{\beta}{2})^4},
\end{equation}
where $\beta$ is the central angle of the circular segment that comprises the channel (the so-called Hooke angle). Note that $\beta =\pi$ corresponds to a semi-circle and
smaller values of $\beta$ result in shallow, wide channels.

With knowledge of cross-sectional area $S_g$, the shear stress, $\tau_g$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
\begin{equation}
  \label{eq:tau}
  \tau_g=\frac{1}{8}\,f_r\,\rho_w\,v_g^2,
\end{equation}
% 
where  the water velocity $v_g = \frac{Q_w}{S_g}$.

The width of the flat channel floor $w_{cg}$, given angle $\beta$ is 
\begin{equation}
  \label{eq:dh2wc}
  w_{cg} = 2  \sin \frac{\beta}{2} \sqrt{\frac{2\, S_g}{\beta -\sin \beta}}.
\end{equation}

The width-integrated shear stress is represented as $\tau_{tg}=w_{cg}\,\tau_g $.

\subsection{Subaerial channel  parameterization}
\label{sect:fluv}

To parameterize the shear stress of water flowing across sediments in the subaerial channel,  the hydraulics parameterization presented in \citet{tucker1997} is implemented.
Here, given  that mass is conserved mass, there is a sufficiently wide channel so that the hydraulic radius is consistent with flow depth, and the Darcy-Weisbach relationship,
the shear stress $\tau_f$ at the river bed is represented as
\begin{linenomath*}
  \begin{equation}
    \label{eq:DW_tau}
    \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q_w}{w_{cf}} \Big)^{\frac{2}{3}} \,\frac{\partial z_c}{\partial x}^{\frac{2}{3}},
  \end{equation}
\end{linenomath*}
where $\frac{\partial z_c}{\partial x}$ is the channel slope, and $f_p$ is the friction factor for subaerial channels.
Channel width $w_{cf}$ is 
\begin{equation}
  \label{eq:wcf}
  w_{cf} = k \, Q_w^\frac{1}{2},
\end{equation}
% 
where $k$ is a constant and the exponent is commonly given as $\frac{1}{2}$ \citep{leopold1953}.
Water velocity, $v_f$, is given by rearranging Equation \ref{eq:tau} as
\begin{equation}
  \label{eq:vf}
  v_f = \sqrt{\frac{8\,\tau_f}{f_p\,\rho_w}}.
\end{equation}
% 
As in Section~\ref{sect:sub_mode}, the width integrated shear stress is $\tau_{tf}=w_{cf}\,\tau_f$.

\subsection{Implementation}

The formulation of the glacial channel (Section~\ref{sect:sub_mode}) requires inputs of the ice thickness ($h_{ice}$, Equation\,\ref{eq:dS_dt}), water discharge, $Q_w$,  Hooke angle ($\beta$; Equation~\ref{eq:Hf}), and friction factor (Equations~\ref{eq:dh}~and~\ref{eq:tau}; $f_r$). Values of these parameters are given in Table~\ref{table:vpm}.

In the subaerial channel, the bed slope, $\frac{\partial z_c}{\partial x}$, is needed in Equation~\ref{eq:DW_tau}. The friction factor, $f_f$, and channel shape factor, $k$, are assigned such that they produce reasonable values for water velocity and channel width  (Section~\ref{sect:fluv}).
\TODO{Check what is the ``Reasonable values'' for water velocity. see below also.}

\TODO{Address Daniel's comment: I think that these records need a few more words: where are they measured exactly? How long are the time series? What temporal resolution have they got? Do they cover the entire year, or only the summer months? }

The parameterizations above are applied to hydrological records from the Fieschergletscher in the Swiss Alps and the Leverett glacier on the Greenland Ice Sheet.
The hydrographs applied to the subglacial and subaerial parameterizations represent the subglacial flow of water as it leaves the glacier and transitions to subaerial flow in front of the glacier, as water moves through the catchment.
Outputs of the parameterizations are meant to represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
To generalize these cases, the Fieschergletscher case is referred to as \alpine{} and is exemplified by relatively thin ice thickness ($h_{ice}$: $225$\,\unit{m}), low water discharge ( $\le\,20$\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
The Leverett glacier case is referred to as \icesheet{} and is exemplified by thick ice  ($h_{ice}$: $700$\,\unit{m}), high water discharge ($\le\,300$\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).

The parameterizations are first applied to a reference test case for each glacier that assumes a semi-circular subglacial channel ($\beta=\pi$) and friction factors $f_r$ and $f_p$ are tuned so that the model reproduces reasonable velocities for both  the Greenland Ice Sheet and Alpine glaciers \citep[$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$][]{werder2010b,chandler2013}.
Water velocity, shear stress, and width-integrated shear stress in the subglacial and subaerial channels are presented below.

Second, to characterize the variability in sediment discharge capacity in subglacial compared to subaerial channels further, $100$  parameterizations are applied to a range of different channel geometry factors ($\beta$ and $k$) and friction factors ($f_r$ and $f_p$).
Model runs with parameter combinations  are culled if their maximum subglacial water velocity exceeds $2\,$\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity exceeds $1\,$\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with some dye tracing measurements \citep[e.g.]{werder2010b,chandler2013}.
Variability in a variable is established from the standard deviation that variable over an aggregation time. 
\ian{Check this again... last sentence needs works (see Daniel's comments)}

\section{Results}
\TODO{needs clarification}

\begin{center}
  \begin{figure}[h]
    \includegraphics[width=0.8\linewidth]{Fig2.pdf}
    \caption{Model outputs resulting from the hydrographs shown in panels a and e for the cases called \alpine (a-d) and \icesheet{} (e-h).". Gray (black) lines in a and e represent the subglacial channel size (water discharge).  In other panels, blue lines represent model outputs from the subglacial channel, while orange lines represent the subaerial channel.
      Plots are shown at $15$\unit{min} intervals.
      Arrows denote some events where the variable peaks in the subglacial channel prior to the subaerial one.
      Insets in e-h show peak events at the onset of melt in the \icesheet{} case.
      Note that the x and y axis are different for the \alpine{} and \icesheet{} cases such that they each cover a summer.} 
    \label{fig:model_outs}
  \end{figure}
\end{center}

Following the parameterizations defined in Sections~\ref{sect:sub_mode}~and~\ref{sect:fluv}, the water velocity, shear stress, and width-integrated shear stress exhibit different seasonal evolutions and peaks. Water velocity and shear stress represent the potential for sediment mobilization, and width-integrated shear stress represents the total sediment transport capacity across the channel bed (Figure~\ref{fig:model_outs}).

\begin{center}
  \begin{figure}[h]
    \includegraphics[width=0.8\linewidth]{Fig3.png}
    \caption{Relationship between water discharge and velocity, shear stress, and width-integrated shear stress for the \alpine{}  (a-c) and \icesheet (d-f) cases. Variables have been normalized to mean values for better comparison amongst the model runs.
      Plots are shown at $15$\unit{min} interval.} 
    \label{fig:Qw_vari}
  \end{figure}
\end{center}

Each water discharge value in the subaerial channel is associated with a single water velocity, shear stress and width-integrated shear stress (Figure~\ref{fig:Qw_vari}).
Conversely in the subglacial channel,  water velocity, shear stress and width-integrated shear stress cab vary substantially for a given water discharge (Figure~\ref{fig:Qw_vari}).
For instance, over $15$ minute time periods at \alpine{}, the range of water velocities and shear stress spans from those below the mean to those almost $2$ times the mean velocity.
Very high values of water velocity and shear stress can occur from minimal water discharge at $2.5$\,\unit{m}$^3$\,\unit{s}$^{-1}$ to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
At \icesheet, reduced diurnal variability in the hydrograph  (Figure~\ref{fig:model_outs} e) results in smaller variations in water velocity and shear stress compared to \alpine{} (Figure~\ref{fig:Qw_vari}).
However in \icesheet, mean values of water velocity can occur at water discharges between roughly $50$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $225$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, effectively spanning the observed range of water discharge. 
\TODO{add sentence about why the two cases were different}

When accounting for the evolving channel width of the subglacial conduit, width-integrated shear stress integrated across the channel width generally increases with water discharge.
Yet even the width-integrated shear stress  can vary substantially, with the highest values occurring at water discharge values ranging from roughly $11$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, to over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
The variability is less pronounced in the \icesheet case, yet a range of width-integrated shear stress can occur at a given discharge(Figure~\ref{fig:Qw_vari} c, e).

Over the course of the season, large peaks in water velocity and shear stress, along with width-integrated shear stress, occur in response to increases in water discharge in both channels.
In the subglacial channel, these peaks generally occur over the period when water discharge increases at the fastest rate (Figure~\ref{fig:model_outs}).
Conversely, in the subaerial channel, peaks in these variables occur when water discharge is at its highest value.
As a result, peaks in the subglacial sediment transport variables occur  when the change in water discharge as highest, as opposed to when the maximum amount of water discharge occurs.
The timing of the fastest rate of change in water discharge occurs before the peaks in water discharge. 
Here, sediment could be deposited in the proglacial area as the water depressurizes and remobilized when sediment transport capacity peaks in the subaerial system.

\begin{center}
  \begin{figure}[h]
    \includegraphics[width=0.8\linewidth]{Fig4.png}
    \caption{Variance (measured by the standard deviation) in water velocity, shear stress and total shear stress for different time aggregations. The subglacial (blue) and subaerial (orange) systems are distinguished as are the \alpine{} (a-c) and \icesheet{} (d-f) case.
      Shaded areas denote the $25^{\mathrm{th}}$ and $75^{\mathrm{th}}$ quantiles of the $100$ parameter combinations tested for each of the systems.
      Solid lines denote  mean values.
      Time intervals shown with markers denote $15$\,\unit{min}, $1$\,\unit{hr}, $12$\,\unit{hr}, $1$\,\unit{d}, $5$\,\unit{d}, $10$\,\unit{d}, and $15$\,\unit{d} aggregation periods.
      \TODO{Introduce and define this experiment better!!!!}
    } 
    \label{fig:multi_run}
  \end{figure}
\end{center}

\TODO{Make sense of these paragraphs}

\TODO{Introduce this paragraph better}
Across the range of parameter values examined for both systems, variability in water velocity and shear stress remain higher in the subglacial system compared to the subaerial one.
However, in the \alpine{} system accounting for channel width makes the variability between the two systems comparable (Figure~\ref{fig:multi_run}).
The variability over time periods is less than one-half day, and the variability in the subglacial system is somewhat higher. 
Conversely, in the \icesheet{} case, with lower water discharge variability, the variability in width-integrated shear stress remains higher in the subglacial system across the time aggregation periods.

In all cases, the variability in velocity, shear stress, and width-integrated shear stress decrease after $1-5$ day periods (Figure~\ref{fig:multi_run}).
This suggests that the effects of the variable response to hydrologic forcing are reduced over longer time periods than these.
Yet, substantial differences persist in periods of up to $15$ days, as the variability might in time periods longer than this can be conflated with seasonal variations in water discharge.
Note that lags and time of peak sediment transport conditions still occur over these periods (Figure~\ref{fig:model_outs}), while the variability in sediment transport parameters is reduced with increasing time aggregation (Figure~\ref{fig:multi_run}).

\section{Discussion}

Quantities that control sediment transport capacity, such as water velocity, shear stress, and width-integrated shear stress, vary differently in response to water discharge in subglacial channels compared to subaerial ones, with several implications for sediment transport in subglacial systems compared to subaerial ones.
In sediment transport relationships, such as in \citet{meyer1948}, that evaluate sediment discharge, shear stress is scaled to the power of $\frac{3}{2}$ and relies on shear stress exceeding a threshold.
In the total sediment transport relationship of \citet{engelund1967}, the shear stress is scaled to $\frac{5}{2}$ power.
In both cases, the exponent greater than $1$ in the sediment transport relationship magnifies the variability in sediment discharge beyond the highly variable shear stress and width-integrated shear stress described above (Figure~\ref{fig:Qw_vari}).

Both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress is homogeneous across the channel bed, despite the impacts of this assumption \citep[e.g.][]{yager2018}.
In the subaerial parameterization, this means that sediment is transported across all channel widths and water discharges, as opposed to large channel widths that occur during floods when much sediment transport occurs \citep{wolman1960}.
As a result, the parameterization could underestimate the variability in sediment transport caused by different discharge values in the case of subaerial systems.  (Figure~\ref{fig:Qw_vari}).

In subglacial channels, peaks occur when water discharge increases at the fastest rate compared with growth of the channel (Figure~\ref{fig:multi_run}).
In turn, the parameterization of subglacial channels here contains self-organizing feedbacks that limits sediment transport capacity as the subglacial channel grows with increasing water discharge capacity, effectively reducing water velocity.
A similar feedback occurs in subaerial systems \citep{phillips2016}, albeit at longer timescales of centuries or more, as opposed to days here.

\subsection{Sediment transport underneath glaciers and at their margins}

The response of sediment discharge caused by the greater variability in sediment transport in subglacial channels will be greatest in strongly transport-limited regimes \citep[e.g.][]{kasmalkar2019}.
In supply-limited systems, the changing sediment transport capacity (Figure~\ref{fig:model_outs}) would have a minimal effect on sediment discharge due to the absence of sediment available for transport \citep{delaney2019}.
Yet, in subglacial systems, the threshold of motion for sediment can be reached more frequently, across a range of water discharges,  could result in additional sediment transport.
The variable sediment transport capacity may result in a tendency for glaciers to evacuate sediment and transition to a supply-limited regime whereby sediment discharge scales sediment production from glacier sliding  \citep[e.g.][]{herman2015}.
Sediment exhaustion by this process may also explain the dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, as opposed to glacier melt \citep{overeem2017}.

The effect of pressurized channel flow could be most pronounced during precipitation or flood events, when subglacial conduits have not adapted to the water discharge flowing through them, resulting in a large increase in sediment transport capacity and discharge \citep[e.g.][]{cowan1988,delaney2019}.

\TODO{COMEBACK HERE with peak event analysis}
Exceptional sediment discharge during floods may result from changes to sediment access underneath the glacier, yet rapid increases in water velocity from water flowing through a small channel will also cause a large increase in sediment transport capacity, not possible in a subaerial system (Section~\ref{sect:sub_mode}).

Increased variability in sediment transport may also impact sediment export from the subglacial environment into proglacial areas \citep[e.g.][]{perolo2018}.
The decrease in sediment transport capacity potentially results in sediment deposition once sediment enters the proglacial area, and thus transitions to subaerial conditions.
Periodic sediment deposition may even occur in the transition between pressurized and subaerial flow under glacier terminus \citep{perolo2018}, especially when cavity closure rates of the ice are slow \citep{egli2021b}.

In large catchments and further downstream of proglacial margins, water discharge is generally less variable than at catchments' headwaters, where glaciers lie in
alpine settings \citep[][]{swift2005,riihimaki2005,costa2017,vanas2017}.
Here, reduced hydrological variability far  downstream of glaciers could result in smaller variations in sediment transport capacity in many subaerial systems compared to subglacial ones.
Smaller variations here would occur as result of hydrological forcings, as opposed the roll the slowly evolving channel size below glaciers.

\subsection{Records of sediment discharge  from glaciers}

The pressurized nature of subglacial systems means that the relationship between water discharge and  sediment concentration, or sediment discharge, could well be disparate, especially over the sub-monthly timescales examined here.
The rapid increase and decreases in sediment transport capacity may, for example, compound the variability in sediment transport already present in subaerial systems \citep[Figure~\ref{fig:Qw_vari}; ][]{williams1989,jerolmack2010}.
Large variations in sediment transport capacity underneath glaciers may result in sediment mobilization and deposition processes in close spatial and temporal proximity to each other \citep{gimbert2016,perolo2018}.
As a result, variations in sediment transport capacity could be responsible for the fluctuations in sediment discharge (``flushings'') from glaciers that could not be attributed to variations in water discharge \citep[e.g.][]{richards2003,swift2021}.
In addition to resulting in sporadic changes to sediment access, the observed ``flushings'' could result from increases in water velocity resulting from reduced channel size, not water discharge, which increase sediment transport capacity (Figure~\ref{fig:Qw_vari}).
Furthermore, correlations between water discharge and sediment transport may occur in response to increased access to subglacial sediment as opposed to sediment transport capacity.

Additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to subaerial systems \citep{jansson2005,ganti2016}. 
In addition to water discharge, for instance sediment transport capacity from a glacier reacts to the ice thickness, controlling the closure of the channel, and the surface slope of the glacier, controlling water velocity \citep[Section~\ref{sect:sub_mode}; ][]{rothlisberger1972,shreve1972,stevens2022}.
Conversely, sediment transport in most subaerial systems increases with water discharge and hydraulic gradient, and the hydraulic gradient of most subaerial systems most likely  remains stable over yearly to century-scale time periods \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wong2006,wickert2019}.

In natural systems, variable mobilization and deposition of sediment occur as water and sediment discharge accumulate as they move through the two-dimensional channel network will compound the already substantial variability in sediment mobilization parameters accounted for here in the lumped parameterization.
Such variability could be especially strong in the subglacial system where melt water pathways change in response to the evolving hydraulic gradient \citep{delaney2023}.

Results here suggest that increased variability in sediment transport in subglacial systems  could be particularly pronounced on ice sheets with large catchments and reduced water discharge variability over sub-monthly timescales.
Furthermore, width-integrated shear stress across the channel bed can remain relatively high across a range of water discharges (Figures~\ref{fig:model_outs} and \ref{fig:Qw_vari}).
\TODO{is this needed?... does if conflict with observations of the Leverett and not testable} In fact, sediment discharge records from two glaciers in the Swiss Alps show that $40$ to $50$\% of a season's sediment discharge occurs when water discharge is below the $75^{\mathrm{th}}$ quantile of the season \citep{delaney2018}.
Such a large quantity of sediment transported at lower water discharges may occur in part because sediment transport capacity responds to the size of the glacier conduit, in addition to water discharge (Figure~\ref{fig:model_outs} and Section~\ref{sect:sub_mode}).

Records of sediment discharge have been used to establish the relationship, or lack thereof, between sediment discharge and climate in glacial systems \citep[e.g.][]{koppes2009a,willenbring2016,mariotti2021}.
The variability in subglacial sediment transport capacity presented here mostly applies to short-time scales controlling the size of subglacial channels (Figure~\ref{fig:multi_run}).
Yet, identifying climatic signals in sediment transport from transport-limited glaciers may require higher thresholds of water discharge compared to subaerial systems \citep[Figure~\ref{fig:Qw_vari}][]{tofelde2021}.
This higher threshold of climatic perturbation, impacting sediment transport capacity, may persist in addition to the stochastic nature of erosion and deposition in fluvial environments \citep{castletort2003,jerolmack2010,romans2016}.
Furthermore, if no sediment is available and the glacier is in a supply-limited state, often assumed in landscape evolution models, then the sediment discharge record will also represent additional processes of bedrock erosion from  water pressure variations and sliding  \citep{iverson2012,herman2015}, which are also complex.

\ian{stopped with daniel's comments here.}

\section{Conclusions}
Subaerial channels  alter their channel width and water velocity immediately in response to changing water discharge.
Conversely, pressurized subglacial channels mostly accommodate changing water discharge by altering water velocity and shear stress, upon which sediment transport depends.
This occurs because the size of subglacial channels evolves slowly compared to variations in water discharge.

Parameterizations of subglacial and subaerial water flow show that subglacial systems exhibit increased variability in sediment transport capacity as they respond  to changes in water discharge.
Variability in sediment transport capacity across a channel's width can be higher in subglacial channels compared to subaerial ones, but the overall variability might be similar if the variations in water discharge are already large.
Furthermore, in subglacial system the timing of peak sediment discharge conditions with respect to water discharge occurs on the rising limb of the hydrograph.
This is in contrast to subaerial systems, where the peak in sediment discharge generally occurs at the same time as the peak in discharge.
The parameterizations and differences presented here do not consider the supply of sediment or evolution of a channel morphology following the evacuation or deposition of sediment.

The differences between subglacial and subaerial channels' response to variations in water discharge should be considered when examining and interpreting sediment transport processes and records in glacierized catchments.

\begin{itemize}
\item questioning if a record of sediment discharge represents subglacial or subaerial sediment transport processes.
\item recognizing that high sediment transport can occur at low water discharges.
\item 
\end{itemize}

\TODO{1-3 concrete recommendations for what to consider would make this conclusion much stronger.}

\section*{Acknowledgments}
\noindent
SNSF Project No. PZ00P2\_202024 provided  funding for I. Delaney.
%The julia code used herein is included in the supplementary material.
%Two reviewers provided constructive and valuable comments on previous version of this manuscript.

%% ------------------------------------------------------------------------ %%
%% References and Citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
% 
% \bibliography{<name of your .bib file>} don't specify the file extension
% 
% don't specify bibliographystyleiz

% In the References section, cite the data/software described in the Availability Statement (this includes primary and processed data used for your research). For details on data/software citation as well as examples, see the Data & Software Citation section of the Data & Software for Authors guidance
% https://www.agu.org/Publish-with-AGU/Publish/Author-Resources/Data-and-Software-for-Authors#citation

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\bibliographystyle{apa} 
\bibliography{PaperLib.bib}

 \section{Supplement}

 \renewcommand{\figurename}{Figure S}
 \setcounter{figure}{0}
\begin{center}
  \begin{figure}[h]
    \includegraphics[width=0.5\linewidth]{Fig2_hr.png}
    \caption{As Figure \ref{fig:model_outs}, with $1$ \,\unit{hr} aggregation.} 
    \label{fig:model_outs_1hr}
  \end{figure}
\end{center}

\begin{center}
  \begin{figure}[h]
    \includegraphics[width=0.5\linewidth]{Fig2_12hr.png}
    \caption{As Figure \ref{fig:model_outs}, with $12$ \,\unit{hr} aggregation.} 
    \label{fig:model_outs_12hr}
  \end{figure}
\end{center}

\begin{center}
  \begin{figure}[h]
    \includegraphics[width=0.5\linewidth]{Fig2_1day.png}
    \caption{As Figure \ref{fig:model_outs}, with $1$ \,\unit{day} aggregation.} 
    \label{fig:model_outs_1day}
  \end{figure}
\end{center}

\begin{center}
  \begin{figure}[h]
    \includegraphics[width=0.5\linewidth]{Fig2_5day.png}
    \caption{As Figure \ref{fig:model_outs}, with $5$ \,\unit{day} aggregation.} 
    \label{fig:model_outs_5day}
  \end{figure}
\end{center}

\begin{center}
  \begin{figure}[h]
    \includegraphics[width=0.5\linewidth]{Fig2_10day.png}
    \caption{As Figure \ref{fig:model_outs}, with $10$ \,\unit{day} aggregation.} 
    \label{fig:model_outs_10day}
  \end{figure}
\end{center}
\begin{center}
  \begin{figure}[h]
    \includegraphics[width=0.5\linewidth]{Fig2_15day.png}
    \caption{As Figure \ref{fig:model_outs}, with $15$ \,\unit{day} aggregation.} 
    \label{fig:model_outs_10day}
  \end{figure}
\end{center}

\end{document}

These different characteristics between glacial and subaerial channels show that records of sediment transport downstream of glaciers represent both subglacial and subaerial processes together.
Divergent processes discussed here might be particularly relevant at glacier margins, where water transitions from pressurized subglacial flow to open-channel subaerial flow.
The inconsistent response of sediment transport in subglacial and subaerial systems to changing water discharge could add uncertainty in evaluating  sediment transport signals in subglacial systems.
In addition, the subglacial parameterization shows that sediment transport capacity in glacier systems responds to a large number of factors such as channel size, ice thickness, and glacier surface slope, which may react to climate forcing differently than water discharge alone.

### 2023-08-11 Delaneyetal2023b.tex
updates

@@
 \usepackage [colon,square]{natbib}
-\usepackage {comment}
-\usepackage{cite}
 \usepackage[mathlines]{lineno} %% <- mathlines turns on line numbering in equations
@@
 \setlength{\parindent}{10pt}
-\setlength{\parskip}{5pt}
+\setlength{\parskip}{10pt}
 \usepackage[explicit]{titlesec}
@@
 \usepackage{printlen}
-\usepackage[a4paper, margin=1.25cm]{geometry}
+\usepackage[a4paper, margin=1.3cm]{geometry}
 \usepackage{color,hyperref}
@@
 
-ianarburua.delaney@unil.ch}, Daniel Farinotti\footnote{Laboratory of Hydraulics, Hydrology and Glaciology (VAW), ETH-Z\"urich, H\"onggerbergring 26, 8093 89 Z\"urich, Switzerland}$^{,}$\footnote{Swiss Federal Institute for Forest, Snow and Landscape Research (WSL) Z\"uricherstrasse 111, 8903 1011 Birmensdorf, Switzerland}, Andrew J.
-Tedstone\footnote{Department of Geosciences, University of Fribourg, Ch.
-du Musée 1700, Fribourg, Switzerland}
+    ianarburua.delaney@unil.ch},
+  Daniel Farinotti\footnote{Laboratory of Hydraulics, Hydrology and Glaciology (VAW), ETH-Z\"urich, H\"onggerbergring 26, 8093 89 Z\"urich, Switzerland}$^{,}$\footnote{Swiss Federal Institute for Forest, Snow and Landscape Research (WSL) Z\"uricherstrasse 111, 8903 1011 Birmensdorf, Switzerland},
+  Andrew J. Tedstone\footnote{Department of Geosciences, University of Fribourg, Ch. du Musée 1700, Fribourg, Switzerland},
 
@@
 \begin{itemize}
-\item Water discharge variations in subglacial channels are mainly accommodated by water velocity, not channel size, given their subglacial channel's slow evolution.
+\item Water discharge variations in subglacial channels are mainly accommodated by water velocity, not channel size, given their subglacial channel's evolution.
 
@@
 
-\item The channels' divergence in transport response may impact sediment dynamics at glacier margins and the interpretation of sediment records.
+\item The channels' divergence in sediment transport response represents multiple processes represented by water discharge in sediment transport.
 \end{itemize}
-
-\TODO{Study site description}
-
-\TODO{Consider removing ``case'' ``system'', around \alpine and \icesheet to reduce words}
 
@@
 
-%\ian{last sentence is a bit rough}
-
 \end{abstract}
@@
 In some of these regions, increased water discharge and glacier melt have been interpreted to be greater sediment transport capacity \citep{bendixen2017,costa2017,li2021}.
-Observed changes to sediment transport in glacierized catchments call for examining the processes controlling sediment discharge in these catchments. 
-Yet, the relationship between water discharge and sediment transport in subglacial environments remains minimally discussed and not compared with the subaerial river systems that these glaciers feed.
-
-\TODO{Add Riihimaki and Swift}
+Observed changes to sediment transport in glacierized catchments require examining the processes controlling sediment discharge in these catchments and its variations with water discharge  \citep[e.g.][]{riihimaki2005,swift2005}.
 
@@
 
-In a transport-limited regime, sediment discharge responds to the sediment transport capacity or the amount of sediment that the water could carry.
+In a transport-limited regime, sediment discharge responds to the sediment transport capacity or the amount of sediment that the water can carry.
 Sediment transport capacity depends on the shear stress between water and sediment it flows  over \citep{shields1936,meyer1948,engelund1967} and the width of the channel bottom $w_c$ over which to mobilize sediment.
@@
 where $Q_w$ is water discharge,  and $S$ is the channel's cross-sectional area.
-\TODO{comment on the squared?}
 
@@
 
-Yet, the response of water velocity to changing water discharge in subglacial channels is different from subaerial ones.
+Yet, the response of water velocity to changing water discharge in subglacial channels differs from subaerial ones.
 The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by frictional heating of water flowing through the channel \citep{rothlisberger1972}.
-As a result, in subglacial channels, the water is pressurized by the ice above \citep{rothlisberger1972}.
 As a result, the subglacial channel size only evolves over days to months, whereas water discharge can vary over hours \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
-In turn, changes in water discharge $Q_w$ are mainly accommodated with water velocity $v$, because the size of the channel $S$ changes on a much shorter timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
+In turn, changes in water discharge $Q_w$ are mainly accommodated with water velocity $v$ \citep[e.g.][]{swift2005}, because the size of the channel $S$ changes on a much shorter timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
 It follows that an increase in water discharge $Q_w$ will cause the shear stress across the channel $\tau_t$ to respond to the water velocity $v$, as opposed to the subaerial channel where channel width $w_c$ evolves as well.
@@
 \end{center}
-\ian{Come back to Daniel's comments}
+% \ian{Come back to Daniel's comments}
 
@@
 These differences have been implicitly included in a wide range of available models that  quantify sediment transport from catchments in both subglacial and subaerial channels  \citep[e.g.][]{walder1994,tucker1997,creyts2013,wickert2019,hewitt2019}.
-%However, some modelss suggest that subglacial sediment discharge  $Q_s \propto Q_w^{\frac{9}{2}}$, assuming that water discharge, changing channel width, and depth are each proportionally responsible for the variance of water velocity and sediment transport capacity \citep{alley1997,swift2021}.
 The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \citep[e.g.][]{lane2016,perolo2018}.
-Additionally, the variable response of sediment transport capacity to water discharge in the two systems may impact the interpretation of records of sediment transport in glacial compared to subaerial systems \citep[e.g.][]{muller1968,richards2003,ganti2016}.
+Additionally, the variable response of sediment transport capacity to water discharge in the two systems may impact the interpretation of records of sediment transport in glacial systems \citep[e.g.][]{muller1968,richards2003,swift2005,ganti2016}.
 Yet, the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial channels, and implications for sediment discharge records, has been minimally discussed.
@@
 Outputs are used to explore differences in the relationships between  water discharge, channel geometry, and water velocity in the subglacial and subaerial channels. 
-With these parameterizations, the manuscript examines the impact of water discharge variability on sediment transport capacity in subglacial and subaerial channels.
-The manuscript then uses the demonstrated differences between the two channel types to discuss the implications for sediment transport processes in subglacial channels, compared to subaerial ones, and for interpreting records of sediment transport from glacierized catchments.
+Furthermore, outputs are compared to sediment discharge data from these two catchments.
+The manuscript then uses the demonstrated differences between the two channel types to discuss the implications for interpreting the processes driving variability in records of sediment transport from glacierized catchments.
+
+\section{Study sites and data}
+\label{sect:ss_data}
+
+% \TODO{Address Daniel's comment: I think that these records need a few more words: where are they measured exactly?
+How long are the time series?
+What temporal resolution have they got?
+Do they cover the entire year, or only the summer months?
+}
+
+Water and suspended sediment data are leveraged from two sites in this study.
+Water discharge and suspended sediment concentration were collected at a $1$\unit{min} interval from the Fieschergletscher in the Swiss Alps  continuously over the period 2014--2021 ($46^\circ\,29'\,07"$ N, $8^\circ\,08'\,34"$ E).
+A description of the data and its collection is available in \citet{felix2022} and \citet{felix2021}.
+
+Additionally, water discharge and suspended sediment concentration were collected roughly $2$\unit{km} downstream from Leverett Glacier in Greenland from over  periods from June to September from 2009 to 2012 ($67^\circ\,03'\,50"$ N, $50^\circ\,12'\,59"$ W).
+This data was presented in part in \citet{cowton2012} and is available from \citet{tedstone2017} at a $5$\unit{min} time interval.
+Note that in some case the removal and relocation of sensors created variations in the sediment concentration that do not result from natural processes.
 
@@
   \caption{Variables, parameters, and constants used in this work.
-Where two values are given, the first refers to an \alpine{} case (Fieschergletscher) and the second to a glacier marginal to the \icesheet{}case (Leverett Glacier) }
+Where two values are given, the first refers to  \alpine{}, a case from Fieschergletscher and the second to a glacier marginal to the \icesheet{} a case from Leverett Glacier }
   \begin{tabular}{ l  c  c c }
@@
     Friction factor (subaerial) & $f_p$ & $3$ & $\mathrm{(-)}$\\
-    Gradient of channel bed (subaerial) &$\frac{\partial z_c}{\partial x}$ &$0.05$& $\mathrm{(-)}$\\
+    Gradient of channel bed (subaerial) &$\nabla z_c$ &$0.05$& $\mathrm{(-)}$\\
     Subaerial channel factor & $k$ &$3$ & $\mathrm{s\,m^{-2}}$\\
@@
     \label{eq:dS_dt}
-    \frac{\partial S_g}{\partial t} = C_2 \frac{Q \Delta h}{l} - C_2 (h_{o}-\frac{\Delta h}{2})^n\,S_g,
+    \frac{\partial S_g}{\partial t} = C_1 \frac{Q \Delta h}{l} - C_2 (h_{o}-\frac{\Delta h}{2})^n\,S_g,
   \end{equation}
@@
 \noindent where $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (Table \ref{table:vpm}), $l$ is the length of subglacial channel, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change from the glacier terminus to its top, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the ice overburden pressure, and $n$ is Glen's n \citep{glen1955}.
-The first term on the right side of the equation represents the opening of the channel through frictional heating, while the second term represents the creep closure of the channel from ice deformation.
+The first term on the right side of the equation represents the opening of the channel through frictional heating, while the following term represents the creep closure of the channel from ice deformation.
 
@@
     \label{eq:DW_tau}
-    \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q_w}{w_{cf}} \Big)^{\frac{2}{3}} \,\frac{\partial z_c}{\partial x}^{\frac{2}{3}},
+    \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q_w}{w_{cf}} \Big)^{\frac{2}{3}} \,\nabla z_c^{\frac{2}{3}},
   \end{equation}
 \end{linenomath*}
-where $\frac{\partial z_c}{\partial x}$ is the channel slope, and $f_p$ is the friction factor for subaerial channels.
+where $\nabla z_c$ is the channel slope, and $f_p$ is the friction factor for subaerial channels.
 Channel width $w_{cf}$ is 
@@
 % 
-As in Section~\ref{sect:sub_mode}, the width integrated shear stress is $\tau_{tf}=w_{cf}\,\tau_f$.
+As in Section~\ref{sect:sub_mode}, the width-integrated shear stress is $\tau_{tf}=w_{cf}\,\tau_f$.
 
 \subsection{Implementation}
-
-The formulation of the glacial channel (Section~\ref{sect:sub_mode}) requires inputs of the ice thickness ($h_{ice}$, Equation\,\ref{eq:dS_dt}), water discharge, $Q_w$,  Hooke angle ($\beta$; Equation~\ref{eq:Hf}), and friction factor (Equations~\ref{eq:dh}~and~\ref{eq:tau}; $f_r$).
-Values of these parameters are given in Table~\ref{table:vpm}.
-
-In the subaerial channel, the bed slope, $\frac{\partial z_c}{\partial x}$, is needed in Equation~\ref{eq:DW_tau}.
-The friction factor, $f_f$, and channel shape factor, $k$, are assigned such that they produce reasonable values for water velocity and channel width  (Section~\ref{sect:fluv}).
-\TODO{Check what is the ``Reasonable values'' for water velocity. see below also.}
-
-\TODO{Address Daniel's comment: I think that these records need a few more words: where are they measured exactly?
-How long are the time series?
-What temporal resolution have they got?
-Do they cover the entire year, or only the summer months?
-}
+\label{sect:imp}
 
@@
 
-The parameterizations are first applied to a reference test case for each glacier that assumes a semi-circular subglacial channel ($\beta=\pi$) and friction factors $f_r$ and $f_p$ are tuned so that the model reproduces reasonable velocities for both  the Greenland Ice Sheet and Alpine glaciers \citep[$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$][]{werder2010b,chandler2013}.
-Water velocity, shear stress, and width-integrated shear stress in the subglacial and subaerial channels are presented below.
+The parameterizations are first applied to a reference test case for each glacier that assumes  subglacial channel with $\beta=\frac{\pi}{2}$ and friction factors $f_r$ and $f_p$ are tuned so that the model reproduces reasonable velocities for both  the Greenland Ice Sheet and Alpine glaciers \citep[$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$][]{werder2010b,chandler2013}.
+In both subaerial channels, the bed slope, $\nabla z_c$ is $0.05$.
+Water velocity, shear stress, and width-integrated shear stress in the subglacial and subaerial channels from this formulation are presented below.
 
 Second, to characterize the variability in sediment discharge capacity in subglacial compared to subaerial channels further, $100$  parameterizations are applied to a range of different channel geometry factors ($\beta$ and $k$) and friction factors ($f_r$ and $f_p$).
-Model runs with parameter combinations  are culled if their maximum subglacial water velocity exceeds $2\,$\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity exceeds $1\,$\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with some dye tracing measurements \citep[e.g.]{werder2010b,chandler2013}.
-Variability in a variable is established from the standard deviation that variable over an aggregation time. 
-\ian{Check this again... last sentence needs works (see Daniel's comments)}
+Model runs with parameter combinations  are culled if their mean subglacial water velocity over the season lies outside of the range between $0.75$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $2$\,\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity lies outside of the range $0.3$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with some measurements \citep[e.g.]{werder2010b,magnusson2012,chandler2013}.
+The standard deviation of a model output over the course of the season is used to quantify the variability of the model output.
+
+Lastly, model outputs are compared with record of sediment discharge from the two glaciers using Spearman rank correlation to reduce the impact of non-linear response of sediment discharge to the output.
 
 \section{Results}
-\TODO{needs clarification}
 
-\begin{center}
+\subsection{Role of water discharge in sediment transport capacity}
 \begin{figure}[h]
+  \centering
   \includegraphics[width=0.8\linewidth]{Fig2.pdf}
@@
 In other panels, blue lines represent model outputs from the subglacial channel, while orange lines represent the subaerial channel.
-      Plots are shown at $15$\unit{min} intervals.
+    Data in plots are at $15$\unit{min} intervals.
     Arrows denote some events where the variable peaks in the subglacial channel prior to the subaerial one.
-      Insets in e-h show peak events at the onset of melt in the \icesheet{} case.
-      Note that the x and y axis are different for the \alpine{} and \icesheet{} cases such that they each cover a summer.} 
+    Insets in e-h show peak events at the onset of melt in \icesheet{}.
+    Note that the x and y axis are different for \alpine{} and \icesheet{} such that they each cover a summer.} 
   \label{fig:model_outs}
 \end{figure}
-\end{center}
-
 Following the parameterizations defined in Sections~\ref{sect:sub_mode}~and~\ref{sect:fluv}, the water velocity, shear stress, and width-integrated shear stress exhibit different seasonal evolutions and peaks.
@@
 
-\begin{center}
+Over the course of the season, large peaks in water velocity and shear stress, along with width-integrated shear stress, occur in response to increases in water discharge in both cases.
+In the subglacial channel, these peaks generally occur over the period when water discharge increases at the fastest rate (Figure~\ref{fig:model_outs}).
+Conversely, in the subaerial channel, peaks in these variables occur when water discharge is at its highest value.
+As a result, peaks in the subglacial sediment transport variables occur  when water discharge change is greatest, as opposed to when the maximum amount of water discharge occurs.
+The timing of the greatest rate of change in water discharge occurs before the peaks in water discharge. 
+Here, sediment could be deposited in the proglacial area as the water depressurizes and remobilized when sediment transport capacity peaks in the subaerial system.
+
 \begin{figure}[h]
+  \centering
   \includegraphics[width=0.8\linewidth]{Fig3.png}
-    \caption{Relationship between water discharge and velocity, shear stress, and width-integrated shear stress for the \alpine{}  (a-c) and \icesheet (d-f) cases.
+  \caption{Relationship between water discharge and velocity, shear stress, and width-integrated shear stress for \alpine{}  (a-c) and \icesheet (d-f).
 Variables have been normalized to mean values for better comparison amongst the model runs.
+    $R_g$ ($R_f$) show the Spearman rank correlation values for the subglacial (subaerial) model outputs.
     Plots are shown at $15$\unit{min} interval.} 
@@
 \end{figure}
-\end{center}
 
-Each water discharge value in the subaerial channel is associated with a single water velocity, shear stress and width-integrated shear stress (Figure~\ref{fig:Qw_vari}).
-Conversely in the subglacial channel,  water velocity, shear stress and width-integrated shear stress cab vary substantially for a given water discharge (Figure~\ref{fig:Qw_vari}).
-For instance, over $15$ minute time periods at \alpine{}, the range of water velocities and shear stress spans from those below the mean to those almost $2$ times the mean velocity.
-Very high values of water velocity and shear stress can occur from minimal water discharge at $2.5$\,\unit{m}$^3$\,\unit{s}$^{-1}$ to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
-At \icesheet, reduced diurnal variability in the hydrograph  (Figure~\ref{fig:model_outs} e) results in smaller variations in water velocity and shear stress compared to \alpine{} (Figure~\ref{fig:Qw_vari}).
-However in \icesheet, mean values of water velocity can occur at water discharges between roughly $50$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $225$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, effectively spanning the observed range of water discharge.
-\TODO{add sentence about why the two cases were different}
+Each water discharge value in the subaerial channel is associated with a single water velocity, shear stress and width-integrated shear stress, resulting in a perfect rank correlation between water discharge and the model output (Figure~\ref{fig:Qw_vari} a--f).
+Conversely in the subglacial channel,  water velocity, shear stress and width-integrated shear stress can vary substantially for a given water discharge and width-integrated shear stress varies with respect to the other two model outputs as well (Figure~\ref{fig:Qw_vari} a--f).
+For instance, very high values of water velocity and shear stress can occur from minimal water discharge at $2.5$\,\unit{m}$^3$\,\unit{s}$^{-1}$ to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
+In \icesheet, mean values of water velocity can occur at water discharges between roughly $145$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $215$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, effectively spanning the much of the water discharge range.
 
-When accounting for the evolving channel width of the subglacial conduit, width-integrated shear stress integrated across the channel width generally increases with water discharge.
+When accounting for the evolving channel width of the subglacial conduit, width-integrated shear stress  across the channel generally increases with water discharge, with improved rank correlation compared to water discharge and water velocity or shear stress (Figure~\ref{fig:Qw_vari} a--f).
 Yet even the width-integrated shear stress  can vary substantially, with the highest values occurring at water discharge values ranging from roughly $11$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, to over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
-The variability is less pronounced in the \icesheet case, yet a range of width-integrated shear stress can occur at a given discharge(Figure~\ref{fig:Qw_vari} c, e).
+The variability in width-integrated shear stress is less pronounced in the \icesheet case, in part due to the reduced variability in water discharge.
+Yet, a range of width-integrated shear stress can occur at a given discharge (Figure~\ref{fig:Qw_vari} c, f).
 
-Over the course of the season, large peaks in water velocity and shear stress, along with width-integrated shear stress, occur in response to increases in water discharge in both channels.
-In the subglacial channel, these peaks generally occur over the period when water discharge increases at the fastest rate (Figure~\ref{fig:model_outs}).
-Conversely, in the subaerial channel, peaks in these variables occur when water discharge is at its highest value.
-As a result, peaks in the subglacial sediment transport variables occur  when the change in water discharge as highest, as opposed to when the maximum amount of water discharge occurs.
-The timing of the fastest rate of change in water discharge occurs before the peaks in water discharge. 
-Here, sediment could be deposited in the proglacial area as the water depressurizes and remobilized when sediment transport capacity peaks in the subaerial system.
-
-\begin{center}
 \begin{figure}[h]
+  \centering
     \includegraphics[width=0.8\linewidth]{Fig4.png}
-    \caption{Variance (measured by the standard deviation) in water velocity, shear stress and total shear stress for different time aggregations.
-The subglacial (blue) and subaerial (orange) systems are distinguished as are the \alpine{} (a-c) and \icesheet{} (d-f) case.
-      Shaded areas denote the $25^{\mathrm{th}}$ and $75^{\mathrm{th}}$ quantiles of the $100$ parameter combinations tested for each of the systems.
+    \caption{Variance (measured by the standard deviation) in water velocity, shear stress and total shear stress for different time aggregations across a large number of subglacial and subaerial channel shapes and friction factors The subglacial (blue) and subaerial (orange) systems are distinguished as are \alpine{} (a-c) and \icesheet{} (d-f).
+      Shaded areas denote the $25^{\mathrm{th}}$ and $75^{\mathrm{th}}$ quantiles of the $100$ parameter combinations tested for each of the systems (Section~\ref{sect:imp}).
       Solid lines denote  mean values.
       Time intervals shown with markers denote $15$\,\unit{min}, $1$\,\unit{hr}, $12$\,\unit{hr}, $1$\,\unit{d}, $5$\,\unit{d}, $10$\,\unit{d}, and $15$\,\unit{d} aggregation periods.
-      \TODO{Introduce and define this experiment better!!!!}
     } 
@@
   \end{figure}
-\end{center}
 
-\TODO{Make sense of these paragraphs}
+The second model application examines the impact of a variety channel shapes and friction factors on the relationship between water discharge and sediment transport capacity, in both systems.
+Across the range of parameter values examined, variability in water velocity and shear stress remain higher in the subglacial system compared to the subaerial one.
+However, in \alpine{}, the width-integrated shear stress makes the variability between the two systems comparable (Figure~\ref{fig:multi_run}).
+Conversely, in \icesheet{}, with lower water discharge variability, the variability in width-integrated shear stress remains higher in the subglacial system across the time aggregation periods.
 
-\TODO{Introduce this paragraph better}
-Across the range of parameter values examined for both systems, variability in water velocity and shear stress remain higher in the subglacial system compared to the subaerial one.
-However, in the \alpine{} system accounting for channel width makes the variability between the two systems comparable (Figure~\ref{fig:multi_run}).
-The variability over time periods is less than one-half day, and the variability in the subglacial system is somewhat higher. 
-Conversely, in the \icesheet{} case, with lower water discharge variability, the variability in width-integrated shear stress remains higher in the subglacial system across the time aggregation periods.
-
-In all cases, the variability in velocity, shear stress, and width-integrated shear stress decrease after $1-5$ day periods (Figure~\ref{fig:multi_run}).
-This suggests that the effects of the variable response to hydrologic forcing are reduced over longer time periods than these.
-Yet, substantial differences persist in periods of up to $15$ days, as the variability might in time periods longer than this can be conflated with seasonal variations in water discharge.
+It is prescient to evaluate the time scales overwhich the subglacial channels may respond to changing water discharge to understand the impact on variability in sediment transport capacity.
+In both cases, the variability in velocity, shear stress, and width-integrated shear stress decrease after $1-5$ day aggregations (Figure~\ref{fig:multi_run}).
+This suggests that the effects of the variable response to hydrologic forcing are reduced over longer time aggregations than these.
+Yet, substantial differences in variability persist in periods of up to $15$ days, as the variability in aggregations periods longer than this might be conflated with seasonal variations in water discharge.
 Note that lags and time of peak sediment transport conditions still occur over these periods (Figure~\ref{fig:model_outs}), while the variability in sediment transport parameters is reduced with increasing time aggregation (Figure~\ref{fig:multi_run}).
 
+\subsection{Relationship between model outputs and observed sediment discharge}
+
+In comparing model outputs the with suspended sediment discharge data, velocity, shear stress and width-integrated shear stress from the subaerial system rank correlate better with sediment discharge compared to subglacial outputs (Figure~\ref{fig:Qw_vari} g--l).
+The rank correlation among the three model outputs for the subaerial parameterization remains constant due to the absence of hystersis in channel shape, as shown above (see Section~\ref{sect:fluv}).
+In the subglacial system, velocity, and shear stress in the subglacial system perform substantially worse compared to width-integrated shear stress in rank correlating with sediment discharge.
+In many both \alpine{} and \icesheet{}, the rank correlation of the subglacial system approaches that of the subaerial system.
+Over longer time aggregation periods in some cases, model outputs from the subglacial model out perform those from the subaerial model (see Figures S1-S6).
+
 \section{Discussion}
-
-Quantities that control sediment transport capacity, such as water velocity, shear stress, and width-integrated shear stress, vary differently in response to water discharge in subglacial channels compared to subaerial ones, with several implications for sediment transport in subglacial systems compared to subaerial ones.
+\subsection{Context of the sediment transport parameterizations}
+Quantities that control sediment transport capacity, such as water velocity, shear stress, and width-integrated shear stress, vary differently in response to water discharge in subglacial channels compared to subaerial ones (Figure~\ref{fig:Qw_vari}).
 In sediment transport relationships, such as in \citet{meyer1948}, that evaluate sediment discharge, shear stress is scaled to the power of $\frac{3}{2}$ and relies on shear stress exceeding a threshold.
 In the total sediment transport relationship of \citet{engelund1967}, the shear stress is scaled to $\frac{5}{2}$ power.
-In both cases, the exponent greater than $1$ in the sediment transport relationship magnifies the variability in sediment discharge beyond the highly variable shear stress and width-integrated shear stress described above (Figure~\ref{fig:Qw_vari}).
+In both cases, the exponent greater than $1$ in the sediment transport relationship magnifies the variability in sediment discharge beyond the highly variable shear stress and width-integrated shear stress described above (Figure~\ref{fig:multi_run}).
 
-Both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress is homogeneous across the channel bed, despite the impacts of this assumption \citep[e.g.][]{yager2018}.
+Both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress is homogeneous across the channel bed, despite the limitations of this assumption \citep[e.g.][]{yager2018}.
 In the subaerial parameterization, this means that sediment is transported across all channel widths and water discharges, as opposed to large channel widths that occur during floods when much sediment transport occurs \citep{wolman1960}.
-As a result, the parameterization could underestimate the variability in sediment transport caused by different discharge values in the case of subaerial systems.  (Figure~\ref{fig:Qw_vari}).
 
 In subglacial channels, peaks occur when water discharge increases at the fastest rate compared with growth of the channel (Figure~\ref{fig:multi_run}).
-In turn, the parameterization of subglacial channels here contains self-organizing feedbacks that limits sediment transport capacity as the subglacial channel grows with increasing water discharge capacity, effectively reducing water velocity.
+In turn, the parameterization of subglacial channels here contains feedbacks that limits sediment transport capacity as the subglacial channel grows with increasing water discharge capacity, effectively reducing water velocity.
 A similar feedback occurs in subaerial systems \citep{phillips2016}, albeit at longer timescales of centuries or more, as opposed to days here.
-
-\subsection{Sediment transport underneath glaciers and at their margins}
-
-The response of sediment discharge caused by the greater variability in sediment transport in subglacial channels will be greatest in strongly transport-limited regimes \citep[e.g.][]{kasmalkar2019}.
-In supply-limited systems, the changing sediment transport capacity (Figure~\ref{fig:model_outs}) would have a minimal effect on sediment discharge due to the absence of sediment available for transport \citep{delaney2019}.
-Yet, in subglacial systems, the threshold of motion for sediment can be reached more frequently, across a range of water discharges,  could result in additional sediment transport.
-The variable sediment transport capacity may result in a tendency for glaciers to evacuate sediment and transition to a supply-limited regime whereby sediment discharge scales sediment production from glacier sliding  \citep[e.g.][]{herman2015}.
-Sediment exhaustion by this process may also explain the dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, as opposed to glacier melt \citep{overeem2017}.
-
-The effect of pressurized channel flow could be most pronounced during precipitation or flood events, when subglacial conduits have not adapted to the water discharge flowing through them, resulting in a large increase in sediment transport capacity and discharge \citep[e.g.][]{cowan1988,delaney2019}.
-
-\TODO{COMEBACK HERE with peak event analysis}
-Exceptional sediment discharge during floods may result from changes to sediment access underneath the glacier, yet rapid increases in water velocity from water flowing through a small channel will also cause a large increase in sediment transport capacity, not possible in a subaerial system (Section~\ref{sect:sub_mode}).
-
-Increased variability in sediment transport may also impact sediment export from the subglacial environment into proglacial areas \citep[e.g.][]{perolo2018}.
-The decrease in sediment transport capacity potentially results in sediment deposition once sediment enters the proglacial area, and thus transitions to subaerial conditions.
-Periodic sediment deposition may even occur in the transition between pressurized and subaerial flow under glacier terminus \citep{perolo2018}, especially when cavity closure rates of the ice are slow \citep{egli2021b}.
 
@@
 Here, reduced hydrological variability far  downstream of glaciers could result in smaller variations in sediment transport capacity in many subaerial systems compared to subglacial ones.
-Smaller variations here would occur as result of hydrological forcings, as opposed the roll the slowly evolving channel size below glaciers.
+Smaller variations here would occur as result of hydrological forcings, as opposed the roll the slowly evolving subglacial channel size.
 
@@
 
-The pressurized nature of subglacial systems means that the relationship between water discharge and  sediment concentration, or sediment discharge, could well be disparate, especially over the sub-monthly timescales examined here.
-The rapid increase and decreases in sediment transport capacity may, for example, compound the variability in sediment transport already present in subaerial systems \citep[Figure~\ref{fig:Qw_vari}; ][]{williams1989,jerolmack2010}.
-Large variations in sediment transport capacity underneath glaciers may result in sediment mobilization and deposition processes in close spatial and temporal proximity to each other \citep{gimbert2016,perolo2018}.
-As a result, variations in sediment transport capacity could be responsible for the fluctuations in sediment discharge (``flushings'') from glaciers that could not be attributed to variations in water discharge \citep[e.g.][]{richards2003,swift2021}.
-In addition to resulting in sporadic changes to sediment access, the observed ``flushings'' could result from increases in water velocity resulting from reduced channel size, not water discharge, which increase sediment transport capacity (Figure~\ref{fig:Qw_vari}).
-Furthermore, correlations between water discharge and sediment transport may occur in response to increased access to subglacial sediment as opposed to sediment transport capacity.
+The model outputs from the subaerial system rank correlate higher with sediment discharge in the \alpine and \icesheet systems here,
+despite model outputs showing the disparate relationship between sediment transport capacity and water discharge in subglacial systems.
+This could be attributed to several reasons.
 
-Additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to subaerial systems \citep{jansson2005,ganti2016}. 
-In addition to water discharge, for instance sediment transport capacity from a glacier reacts to the ice thickness, controlling the closure of the channel, and the surface slope of the glacier, controlling water velocity \citep[Section~\ref{sect:sub_mode}; ][]{rothlisberger1972,shreve1972,stevens2022}.
-Conversely, sediment transport in most subaerial systems increases with water discharge and hydraulic gradient, and the hydraulic gradient of most subaerial systems most likely  remains stable over yearly to century-scale time periods \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wong2006,wickert2019}.
+First, large variations in sediment transport capacity underneath glaciers may result in sediment mobilization and deposition processes in close spatial and temporal proximity to each other \citep{gimbert2016,perolo2018,delaney2023}.
+Rapid increase and decreases in sediment transport capacity may, for example, compound the variability in sediment transport already present in subaerial systems \citep[Figure~\ref{fig:Qw_vari}; ][]{williams1989,jerolmack2010}.
+Due to these chaotic nature of these sediment transport processes, differentiating the characteristics of sediment transport capacity in subglacial and subaerial systems could be difficult.
+Indeed, the model performance metrics of the subglacial width-integrated shear stress and the subaerial metrics are similar (Figures~\ref{fig:Qw_vari}\,i,l and S1-S6).
 
-In natural systems, variable mobilization and deposition of sediment occur as water and sediment discharge accumulate as they move through the two-dimensional channel network will compound the already substantial variability in sediment mobilization parameters accounted for here in the lumped parameterization.
-Such variability could be especially strong in the subglacial system where melt water pathways change in response to the evolving hydraulic gradient \citep{delaney2023}.
+Second, these measurement stations are $0.5$ to $2$\,\unit{km} downstream of the glacier terminus \citep{cowton2012,felix2022}, respectively.
+This could potentially be a long enough distance for the hydraulic characteristics to change from the pressurized subglacial regime to the open channel subaerial regime.
+The transition of subglacial to subaerial conditions may occur in the transition between pressurized and subaerial flow under glacier terminus \citep{perolo2018}, especially when cavity closure rates of the ice are slow here \citep{egli2021b}.
+Furthermore, sediment transport in the glacier margin may persist in a transport limited regime where sediment is deposited and mobilized with respect to sediment transport capacity, resulting in a stronger correlation between sediment transport and water discharge just a short distance downstream.
 
-Results here suggest that increased variability in sediment transport in subglacial systems  could be particularly pronounced on ice sheets with large catchments and reduced water discharge variability over sub-monthly timescales.
-Furthermore, width-integrated shear stress across the channel bed can remain relatively high across a range of water discharges (Figures~\ref{fig:model_outs} and \ref{fig:Qw_vari}).
-\TODO{is this needed?...
-does if conflict with observations of the Leverett and not testable} In fact, sediment discharge records from two glaciers in the Swiss Alps show that $40$ to $50$\% of a season's sediment discharge occurs when water discharge is below the $75^{\mathrm{th}}$ quantile of the season \citep{delaney2018}.
-Such a large quantity of sediment transported at lower water discharges may occur in part because sediment transport capacity responds to the size of the glacier conduit, in addition to water discharge (Figure~\ref{fig:model_outs} and Section~\ref{sect:sub_mode}).
+Third, the parameterizations here only consider sediment transport capacity, not the availability or routing of sediment below the glacier and through the proglacial area \citep{delaney2023}.
+The sediment discharge records could represent factors such as increased sediment accessibility when melt extends up glacier, thereby concurrently increasing sediment discharge with water discharge \citep[e.g.][]{delaney2020}.
+As a result, the sediment transport records presented here may represent a supply-limited sediment transport, less dependent on sediment transport capacity.
+Furthermore, in subglacial systems, the threshold of motion for sediment can be reached more frequently, and across a range of water discharges, potentially resulting in additional sediment transport (Figure~\ref{fig:Qw_vari}\,a-f).
+For instance, sediment discharge records from two glaciers in the Swiss Alps show that $40$ to $50$\% of a season's sediment discharge occurs when water discharge is below the $75^{\mathrm{th}}$ quantile of the season \citep{delaney2018}.
+This process may lead to glaciers evacuating sediment across a range of water discharges and transitioning to a supply-limited regime, where sediment discharge also scales with sediment production from glacier quarrying and abrasion \citep[e.g.][]{herman2015,ugelvig2018}.
+Sediment exhaustion through this process may also explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \citep{overeem2017}.
 
+Regardless of the cause of the increased performance of water discharge compared to subglacial sediment transport capacity in the observations, model outputs suggest that additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to subaerial systems \citep{jansson2005,ganti2016}.
 Records of sediment discharge have been used to establish the relationship, or lack thereof, between sediment discharge and climate in glacial systems \citep[e.g.][]{koppes2009a,willenbring2016,mariotti2021}.
@@
 Yet, identifying climatic signals in sediment transport from transport-limited glaciers may require higher thresholds of water discharge compared to subaerial systems \citep[Figure~\ref{fig:Qw_vari}][]{tofelde2021}.
-This higher threshold of climatic perturbation, impacting sediment transport capacity, may persist in addition to the stochastic nature of erosion and deposition in fluvial environments \citep{castletort2003,jerolmack2010,romans2016}.
-Furthermore, if no sediment is available and the glacier is in a supply-limited state, often assumed in landscape evolution models, then the sediment discharge record will also represent additional processes of bedrock erosion from  water pressure variations and sliding  \citep{iverson2012,herman2015}, which are also complex.
+This higher threshold of perturbation by water discharge, which impacts sediment transport capacity, may persist in addition to the stochastic nature of erosion and deposition in fluvial environments \citep{castletort2003,jerolmack2010,romans2016}.
 
-\ian{stopped with daniel's comments here.}
+In addition to water discharge, the glacier's sediment transport capacity reacts to the ice thickness, controlling the closure of the channel, and the surface slope of the glacier, controlling water velocity \citep[Section~\ref{sect:sub_mode}; ][]{rothlisberger1972,shreve1972,stevens2022}.
+Conversely, sediment transport in most subaerial systems typically increases with water discharge and hydraulic gradient, and the hydraulic gradient of most subaerial systems likely remains stable over time scales ranging from years to centuries \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wong2006,wickert2019}.
+Furthermore, if no sediment is available and the glacier is in a supply-limited state, often assumed in glacier landscape evolution models, then the sediment discharge record also represents additional processes of bedrock erosion from  water pressure variations and sliding  \citep{iverson2012,herman2015}, which are also complex.
 
 \section{Conclusions}
+Sediment transport capacity is largely driven by the velocity of water flowing through a channel and the channel's width.
 Subaerial channels  alter their channel width and water velocity immediately in response to changing water discharge.
-Conversely, pressurized subglacial channels mostly accommodate changing water discharge by altering water velocity and shear stress, upon which sediment transport depends.
-This occurs because the size of subglacial channels evolves slowly compared to variations in water discharge.
+In contrast, pressurized subglacial channels mostly accommodate changing water discharge by altering water velocity and shear stress, upon which sediment transport depends.
+The response of water velocity occurs because the size of subglacial channels evolves slowly compared to variations in water discharge.
 
-Parameterizations of subglacial and subaerial water flow show that subglacial systems exhibit increased variability in sediment transport capacity as they respond  to changes in water discharge.
-Variability in sediment transport capacity across a channel's width can be higher in subglacial channels compared to subaerial ones, but the overall variability might be similar if the variations in water discharge are already large.
-Furthermore, in subglacial system the timing of peak sediment discharge conditions with respect to water discharge occurs on the rising limb of the hydrograph.
-This is in contrast to subaerial systems, where the peak in sediment discharge generally occurs at the same time as the peak in discharge.
-The parameterizations and differences presented here do not consider the supply of sediment or evolution of a channel morphology following the evacuation or deposition of sediment.
+Model outputs show substantial differences in subglacial and subaerial channels' response to changing water discharge.
+However, subaerial hydraulics demonstrate a better representation of sediment discharge variations compared to subglacial hydraulics.
+This could be attributed to the role of open channel flow beneath the glacier terminus and in the proglacial area, or the influence of water discharge in driving sediment availability, rather than subglacial sediment discharge capacity.
 
-The differences between subglacial and subaerial channels' response to variations in water discharge should be considered when examining and interpreting sediment transport processes and records in glacierized catchments.
-
-\begin{itemize}
-\item questioning if a record of sediment discharge represents subglacial or subaerial sediment transport processes.
-\item recognizing that high sediment transport can occur at low water discharges.
-\item 
-\end{itemize}
-
-\TODO{1-3 concrete recommendations for what to consider would make this conclusion much stronger.}
+Considerations in interpretations of sediment discharge should be made given the findings presented here.
+These considerations include:
+\begin{enumerate}
+\item the inconsistent relationship between water discharge and sediment transport capacity underneath glaciers;
+\item the timing of peak sediment transport capacity below glaciers can occur when the rate of change in water discharge is large, as opposed to the maximum quantity of water discharge;
+\item the potentially increased variability in sediment discharge from underneath glaciers than in rivers they feed;
+\item the multiple roles of water discharge to represent processes such as subglacial sediment transport capacity, subglacial sediment access, or open channel processes near the glacier margin in sediment records collected near glaciers.
+\end{enumerate}
 
@@
 SNSF Project No. PZ00P2\_202024 provided  funding for I. Delaney.
-%The julia code used herein is included in the supplementary material.
-%Two reviewers provided constructive and valuable comments on previous version of this manuscript.
+% The julia code used herein is included in the supplementary material.
+% Two reviewers provided constructive and valuable comments on previous version of this manuscript.
 
@@
   \begin{figure}[h]
-    \includegraphics[width=0.5\linewidth]{Fig2_hr.png}
+    \includegraphics[width=0.7\linewidth]{Fig2_hr.png}
     \caption{As Figure \ref{fig:model_outs}, with $1$ \,\unit{hr} aggregation.} 
@@
   \begin{figure}[h]
-    \includegraphics[width=0.5\linewidth]{Fig2_12hr.png}
+    \includegraphics[width=0.7\linewidth]{Fig2_12hr.png}
     \caption{As Figure \ref{fig:model_outs}, with $12$ \,\unit{hr} aggregation.} 
@@
   \begin{figure}[h]
-    \includegraphics[width=0.5\linewidth]{Fig2_1day.png}
+    \includegraphics[width=0.7\linewidth]{Fig2_1day.png}
     \caption{As Figure \ref{fig:model_outs}, with $1$ \,\unit{day} aggregation.} 
@@
   \begin{figure}[h]
-    \includegraphics[width=0.5\linewidth]{Fig2_5day.png}
+    \includegraphics[width=0.7\linewidth]{Fig2_5day.png}
     \caption{As Figure \ref{fig:model_outs}, with $5$ \,\unit{day} aggregation.} 
@@
   \begin{figure}[h]
-    \includegraphics[width=0.5\linewidth]{Fig2_10day.png}
+    \includegraphics[width=0.7\linewidth]{Fig2_10day.png}
     \caption{As Figure \ref{fig:model_outs}, with $10$ \,\unit{day} aggregation.} 
@@
   \begin{figure}[h]
-    \includegraphics[width=0.5\linewidth]{Fig2_15day.png}
+    \includegraphics[width=0.7\linewidth]{Fig2_15day.png}
     \caption{As Figure \ref{fig:model_outs}, with $15$ \,\unit{day} aggregation.} 
-    \label{fig:model_outs_10day}
+    \label{fig:model_outs_15day}
   \end{figure}
@@
 In addition, the subglacial parameterization shows that sediment transport capacity in glacier systems responds to a large number of factors such as channel size, ice thickness, and glacier surface slope, which may react to climate forcing differently than water discharge alone.
+
+In natural systems, variable mobilization and deposition of sediment occur as water and sediment discharge accumulate as they move through the two-dimensional channel network will compound the already substantial variability in sediment mobilization parameters accounted for here in the lumped parameterization.
+Such variability could be especially strong in the subglacial system where melt water pathways change in response to the evolving hydraulic gradient \citep{delaney2023}.
+
+Results here suggest that increased variability in sediment transport in subglacial systems  could be particularly pronounced on ice sheets with large catchments and reduced water discharge variability over sub-monthly timescales.
+Furthermore, width-integrated shear stress across the channel bed can remain relatively high across a range of water discharges (Figures~\ref{fig:model_outs} and \ref{fig:Qw_vari}).
+\TODO{is this needed?... does if conflict with observations of the Leverett and not testable}

### 2023-08-11 Delaneyetal2023b.tex
Fig2 updated

@@
   Daniel Farinotti\footnote{Laboratory of Hydraulics, Hydrology and Glaciology (VAW), ETH-Z\"urich, H\"onggerbergring 26, 8093 89 Z\"urich, Switzerland}$^{,}$\footnote{Swiss Federal Institute for Forest, Snow and Landscape Research (WSL) Z\"uricherstrasse 111, 8903 1011 Birmensdorf, Switzerland},
-  Andrew J. Tedstone\footnote{Department of Geosciences, University of Fribourg, Ch. du Musée 1700, Fribourg, Switzerland},
+  Andrew J. Tedstone\footnote{Department of Geosciences, University of Fribourg, Ch. du Musée 1700, Fribourg, Switzerland}
 
@@
     Arrows denote some events where the variable peaks in the subglacial channel prior to the subaerial one.
-    Insets in e-h show peak events at the onset of melt in \icesheet{}.
+    Insets in e-h show the  peak melt event \icesheet{} denoted by shaded area.
     Note that the x and y axis are different for \alpine{} and \icesheet{} such that they each cover a summer.} 
@@
 \item the potentially increased variability in sediment discharge from underneath glaciers than in rivers they feed;
-\item the multiple roles of water discharge to represent processes such as subglacial sediment transport capacity, subglacial sediment access, or open channel processes near the glacier margin in sediment records collected near glaciers.
+\item the representations of water discharge such as subglacial sediment transport capacity, subglacial sediment access, or open channel processes near the glacier margin in sediment records collected near glaciers.
 \end{enumerate}

### 2023-08-11 Delaneyetal2023b.tex
merged and updated

@@
   \centering
-  \includegraphics[width=0.8\linewidth]{Fig2.pdf}
+  \includegraphics[width=0.9\linewidth]{Fig2.pdf}
   \caption{Model outputs resulting from the hydrographs shown in panels a and e for the cases called \alpine (a-d) and \icesheet{} (e-h).".
@@
   \centering
-  \includegraphics[width=0.8\linewidth]{Fig3.png}
+  \includegraphics[width=0.9\linewidth]{Fig3.png}
   \caption{Relationship between water discharge and velocity, shear stress, and width-integrated shear stress for \alpine{}  (a-c) and \icesheet (d-f).
@@
   \centering
-    \includegraphics[width=0.8\linewidth]{Fig4.png}
+    \includegraphics[width=0.9\linewidth]{Fig4.png}
     \caption{Variance (measured by the standard deviation) in water velocity, shear stress and total shear stress for different time aggregations across a large number of subglacial and subaerial channel shapes and friction factors The subglacial (blue) and subaerial (orange) systems are distinguished as are \alpine{} (a-c) and \icesheet{} (d-f).
@@
 \item the inconsistent relationship between water discharge and sediment transport capacity underneath glaciers;
-\item the timing of peak sediment transport capacity below glaciers can occur when the rate of change in water discharge is large, as opposed to the maximum quantity of water discharge;
-\item the potentially increased variability in sediment discharge from underneath glaciers than in rivers they feed;
+\item the timing of peak sediment transport capacity below glaciers generally coincides with the large rates of change in water discharge, as opposed to the maximum quantity of water discharge;
+\item the potentially increased variability in sediment discharge capacity from underneath glaciers than in rivers they feed;
 \item the representations of water discharge such as subglacial sediment transport capacity, subglacial sediment access, or open channel processes near the glacier margin in sediment records collected near glaciers.
@@
   \begin{figure}[h]
-    \includegraphics[width=0.7\linewidth]{Fig2_hr.png}
+    \includegraphics[width=0.7\linewidth]{Fig3_hr.png}
     \caption{As Figure \ref{fig:model_outs}, with $1$ \,\unit{hr} aggregation.} 
@@
   \begin{figure}[h]
-    \includegraphics[width=0.7\linewidth]{Fig2_12hr.png}
+    \includegraphics[width=0.7\linewidth]{Fig3_12hr.png}
     \caption{As Figure \ref{fig:model_outs}, with $12$ \,\unit{hr} aggregation.} 
@@
   \begin{figure}[h]
-    \includegraphics[width=0.7\linewidth]{Fig2_1day.png}
+    \includegraphics[width=0.7\linewidth]{Fig3_1day.png}
     \caption{As Figure \ref{fig:model_outs}, with $1$ \,\unit{day} aggregation.} 
@@
   \begin{figure}[h]
-    \includegraphics[width=0.7\linewidth]{Fig2_5day.png}
+    \includegraphics[width=0.7\linewidth]{Fig3_5day.png}
     \caption{As Figure \ref{fig:model_outs}, with $5$ \,\unit{day} aggregation.} 
@@
   \begin{figure}[h]
-    \includegraphics[width=0.7\linewidth]{Fig2_10day.png}
+    \includegraphics[width=0.7\linewidth]{Fig3_10day.png}
     \caption{As Figure \ref{fig:model_outs}, with $10$ \,\unit{day} aggregation.} 
@@
   \begin{figure}[h]
-    \includegraphics[width=0.7\linewidth]{Fig2_15day.png}
+    \includegraphics[width=0.7\linewidth]{Fig3_15day.png}
     \caption{As Figure \ref{fig:model_outs}, with $15$ \,\unit{day} aggregation.} 

### 2023-08-15 Delaneyetal2023b.tex
Small bits

@@
   \noindent
-  In both subglacial and subaerial channels, sediment transport capacity of a sediment size depends mainly on 1) the channel width over which the sediment is mobilized  and 2) water velocity, as this controls the shear stress that the flowing water applies to the channel bottom.
-  In subaerial channels, changing water discharge  changes the water depth,  the channel width, and the  water velocity.
-  In subglacial environments and on time scales of hours to weeks, however, changing water discharge  mainly alters the  water velocity, not channel area or width, because the subglacial conduit maintains a largely fixed geometry.
+  In both subglacial and subaerial channels, the sediment transport capacity of a sediment size depends mainly on 1) the channel width over which the sediment is mobilized  and 2) water velocity, as this controls the shear stress that the flowing water applies to the channel bottom.
+  In subaerial channels, changing water discharge  changes the water depth,  channel width, and velocity.
+  However, in subglacial environments and on time scales of hours to weeks, changing water discharge  mainly alters the  water velocity, not channel area or width, because the subglacial conduit maintains a largely fixed geometry.
   Here, parameterizations of sediment transport capacity for pressurized flow under glaciers and subaerial open channel flow are applied to hydrographs from an Alpine glacier and the Greenland Ice Sheet.
   Results of these parameterizations show that sediment transport capacity varies far more in subglacial channels compared to subaerial channels both over time and with respect to water discharge.
+  In the subglacial case, high sediment transport capacity can persist across a wide range of water discharges.
   However, changing channel width can accommodate some  variability in sediment transport capacity in subaerial channels.
-  The different response of water discharge on sediment transport capacity in subglacial channels, compared to subaerial ones, may impact the occurence of sediment transport capacity reaching the threshold of mobilization, thus should be considered in the interpretation of sediment discharge records with respect to water discharge.
+  The different responses of water discharge on sediment transport capacity in subglacial channels, compared to subaerial ones, may impact the occurrence of sediment transport capacity reaching mobilization thresholds and thus should be considered in the interpretation of sediment discharge records from glaciers.
 
@@
     \includegraphics[width=0.8\linewidth]{Fig1.pdf}
-    \caption{Sketch for the different response that subglacial and subaerial channels have to an increase in water discharge.
+    \caption{Sketch for the different responses that subglacial and subaerial channels have to an increase in water discharge.
 Water velocity magnitude in the subglacial, $v_g$, and subaerial, $v_f$, channels are shown by arrow length.
@@
 
-This manuscript aims to evaluate relationship between sediment transport capacity and water discharge in subglacial channels, compared to subaerial ones.
-Numerical parameterizations of subglacial and subaerial hydraulics are untilized, in the absence of observations of continuous water velocity measurements and the evolution of subglacial and subaerial channels.
+This manuscript aims to evaluate the relationship between sediment transport capacity and water discharge in subglacial channels, compared to subaerial ones.
+Numerical parameterizations of subglacial and subaerial hydraulics are utilized, in the absence of observations of continuous water velocity measurements and the evolution of subglacial and subaerial channels.
 Parameterizations are applied to hydrological records from an Alpine glacier in Switzerland (Fieschergletscher), and  a land-terminating glacier in Greenland (Leverett glacier).
@@
 
-Additionally, water discharge and suspended sediment concentration were collected roughly $2$\unit{km} downstream from Leverett Glacier in Greenland from over  periods from June to September from 2009 to 2012 ($67^\circ\,03'\,50"$ N, $50^\circ\,12'\,59"$ W).
+Additionally, water discharge and suspended sediment concentration were collected roughly $2$\unit{km} downstream from Leverett Glacier in Greenland over  periods from June to September from 2009 to 2012 ($67^\circ\,03'\,50"$ N, $50^\circ\,12'\,59"$ W).
 This data was presented in part in \citet{cowton2012} and is available from \citet{tedstone2017} at a $5$\unit{min} time interval.
-Note that in some case the removal and relocation of sensors created variations in the sediment concentration that do not result from natural processes.
+Note that in some cases the removal and relocation of sensors created variations in the sediment concentration that do not result from natural processes.
 
@@
   \caption{Variables, parameters, and constants used in this work.
-Where two values are given, the first refers to  \alpine{}, a case from Fieschergletscher and the second to a glacier marginal to the \icesheet{} a case from Leverett Glacier }
+Where two values are given, the first refers to  \alpine{}, a case from Fieschergletscher, and the second to a glacier marginal to the \icesheet{} a case from Leverett Glacier }
   \begin{tabular}{ l  c  c c }
@@
 
-Lastly, model outputs are compared with record of sediment discharge from the two glaciers using Spearman rank correlation to reduce the impact of non-linear response of sediment discharge to the output.
+Lastly, model outputs are compared with the record of sediment discharge from the two glaciers using Spearman rank correlation to reduce the impact of the non-linear response of sediment discharge to the output.
 
@@
     Arrows denote some events where the variable peaks in the subglacial channel prior to the subaerial one.
-    Insets in e-h show the  peak melt event \icesheet{} denoted by shaded area.
+    Insets in e-h show the  peak melt event \icesheet{} denoted by the shaded area.
     Note that the x and y axis are different for \alpine{} and \icesheet{} such that they each cover a summer.} 
@@
 The timing of the greatest rate of change in water discharge occurs before the peaks in water discharge. 
-Here, sediment could be deposited in the proglacial area as the water depressurizes and remobilized when sediment transport capacity peaks in the subaerial system.
+Here, sediment could be deposited in the proglacial area as the water depressurizes and is remobilized when sediment transport capacity peaks in the subaerial system.
 
@@
 
-Each water discharge value in the subaerial channel is associated with a single water velocity, shear stress and width-integrated shear stress, resulting in a perfect rank correlation between water discharge and the model output (Figure~\ref{fig:Qw_vari} a--f).
-Conversely in the subglacial channel,  water velocity, shear stress and width-integrated shear stress can vary substantially for a given water discharge and width-integrated shear stress varies with respect to the other two model outputs as well (Figure~\ref{fig:Qw_vari} a--f).
+Each water discharge value in the subaerial channel is associated with a single water velocity, shear stress, and width-integrated shear stress, resulting in a perfect rank correlation between water discharge and the model output (Figure~\ref{fig:Qw_vari} a--f).
+Conversely in the subglacial channel,  water velocity, shear stress, and width-integrated shear stress can vary substantially for a given water discharge and width-integrated shear stress varies with respect to the other two model outputs as well (Figure~\ref{fig:Qw_vari} a--f).
 For instance, very high values of water velocity and shear stress can occur from minimal water discharge at $2.5$\,\unit{m}$^3$\,\unit{s}$^{-1}$ to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
-In \icesheet, mean values of water velocity can occur at water discharges between roughly $145$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $215$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, effectively spanning the much of the water discharge range.
+In \icesheet, mean values of water velocity can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, effectively spanning the much of the water discharge range.
 
@@
     \includegraphics[width=0.9\linewidth]{Fig4.png}
-    \caption{Variance (measured by the standard deviation) in water velocity, shear stress and total shear stress for different time aggregations across a large number of subglacial and subaerial channel shapes and friction factors The subglacial (blue) and subaerial (orange) systems are distinguished as are \alpine{} (a-c) and \icesheet{} (d-f).
+    \caption{Variance (measured by the standard deviation) in water velocity, shear stress, and total shear stress for different time aggregations across a large number of subglacial and subaerial channel shapes and friction factors The subglacial (blue) and subaerial (orange) systems are distinguished as are \alpine{} (a-c) and \icesheet{} (d-f).
       Shaded areas denote the $25^{\mathrm{th}}$ and $75^{\mathrm{th}}$ quantiles of the $100$ parameter combinations tested for each of the systems (Section~\ref{sect:imp}).
@@
 
-The second model application examines the impact of a variety channel shapes and friction factors on the relationship between water discharge and sediment transport capacity, in both systems.
-Across the range of parameter values examined, variability in water velocity and shear stress remain higher in the subglacial system compared to the subaerial one.
+The second model application examines the impact of a variety of channel shapes and friction factors on the relationship between water discharge and sediment transport capacity, in both systems.
+Across the range of parameter values examined, variability in water velocity and shear stress remains higher in the subglacial system compared to the subaerial one.
 However, in \alpine{}, the width-integrated shear stress makes the variability between the two systems comparable (Figure~\ref{fig:multi_run}).
@@
 
-It is prescient to evaluate the time scales overwhich the subglacial channels may respond to changing water discharge to understand the impact on variability in sediment transport capacity.
+It is prescient to evaluate the time scales over which the subglacial channels may respond to changing water discharge to understand the impact on variability in sediment transport capacity.
 In both cases, the variability in velocity, shear stress, and width-integrated shear stress decrease after $1-5$ day aggregations (Figure~\ref{fig:multi_run}).
@@
 
-In comparing model outputs the with suspended sediment discharge data, velocity, shear stress and width-integrated shear stress from the subaerial system rank correlate better with sediment discharge compared to subglacial outputs (Figure~\ref{fig:Qw_vari} g--l).
-The rank correlation among the three model outputs for the subaerial parameterization remains constant due to the absence of hystersis in channel shape, as shown above (see Section~\ref{sect:fluv}).
+In comparing model outputs with suspended sediment discharge data, velocity, shear stress and width-integrated shear stress from the subaerial system rank correlate better with sediment discharge compared to subglacial outputs (Figure~\ref{fig:Qw_vari} g--l).
+The rank correlation among the three model outputs for the subaerial parameterization remains constant due to the absence of hysteresis in channel shape, as shown above (see Section~\ref{sect:fluv}).
 In the subglacial system, velocity, and shear stress in the subglacial system perform substantially worse compared to width-integrated shear stress in rank correlating with sediment discharge.
 In many both \alpine{} and \icesheet{}, the rank correlation of the subglacial system approaches that of the subaerial system.
-Over longer time aggregation periods in some cases, model outputs from the subglacial model out perform those from the subaerial model (see Figures S1-S6).
+Over longer time aggregation periods in some cases, model outputs from the subglacial model outperform those from the subaerial model (see Figures S1-S6).
 
@@
 
-In subglacial channels, peaks occur when water discharge increases at the fastest rate compared with growth of the channel (Figure~\ref{fig:multi_run}).
-In turn, the parameterization of subglacial channels here contains feedbacks that limits sediment transport capacity as the subglacial channel grows with increasing water discharge capacity, effectively reducing water velocity.
-A similar feedback occurs in subaerial systems \citep{phillips2016}, albeit at longer timescales of centuries or more, as opposed to days here.
+In subglacial channels, peaks occur when water discharge increases at the fastest rate compared with the growth of the channel (Figure~\ref{fig:multi_run}).
+In turn, the parameterization of subglacial channels here contains feedback that limits sediment transport capacity as the subglacial channel grows with increasing water discharge capacity, effectively reducing water velocity.
+Similar feedback occurs in subaerial systems \citep{phillips2016}, albeit at longer timescales of centuries or more, as opposed to days here.
 
@@
 Here, reduced hydrological variability far  downstream of glaciers could result in smaller variations in sediment transport capacity in many subaerial systems compared to subglacial ones.
-Smaller variations here would occur as result of hydrological forcings, as opposed the roll the slowly evolving subglacial channel size.
+Smaller variations here would occur as a result of hydrological forcings, as opposed to slowly evolving subglacial channel size.
 
@@
 First, large variations in sediment transport capacity underneath glaciers may result in sediment mobilization and deposition processes in close spatial and temporal proximity to each other \citep{gimbert2016,perolo2018,delaney2023}.
-Rapid increase and decreases in sediment transport capacity may, for example, compound the variability in sediment transport already present in subaerial systems \citep[Figure~\ref{fig:Qw_vari}; ][]{williams1989,jerolmack2010}.
-Due to these chaotic nature of these sediment transport processes, differentiating the characteristics of sediment transport capacity in subglacial and subaerial systems could be difficult.
-Indeed, the model performance metrics of the subglacial width-integrated shear stress and the subaerial metrics are similar (Figures~\ref{fig:Qw_vari}\,i,l and S1-S6).
+The rapid increase and decreases in sediment transport capacity may, for example, compound the variability in sediment transport already present in subaerial systems \citep[Figure~\ref{fig:Qw_vari}; ][]{williams1989,jerolmack2010}.
+Due to the chaotic nature of these sediment transport processes, differentiating the characteristics of sediment transport capacity in subglacial and subaerial systems could be difficult.
+Indeed, the model performance metrics of the subglacial width-integrated shear stress and the subaerial metrics are similar (Figures~\ref{fig:Qw_vari}\, i, l and S1-S6).
 
@@
 The transition of subglacial to subaerial conditions may occur in the transition between pressurized and subaerial flow under glacier terminus \citep{perolo2018}, especially when cavity closure rates of the ice are slow here \citep{egli2021b}.
-Furthermore, sediment transport in the glacier margin may persist in a transport limited regime where sediment is deposited and mobilized with respect to sediment transport capacity, resulting in a stronger correlation between sediment transport and water discharge just a short distance downstream.
+Furthermore, sediment transport in the glacier margin may persist in a transport-limited regime where sediment is deposited and mobilized with respect to sediment transport capacity, resulting in a stronger correlation between sediment transport and water discharge just a short distance downstream.
 
@@
 SNSF Project No. PZ00P2\_202024 provided  funding for I. Delaney.
-% The julia code used herein is included in the supplementary material.
-% Two reviewers provided constructive and valuable comments on previous version of this manuscript.
 
-%% ------------------------------------------------------------------------ %%
-%% References and Citations
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-% 
-% \bibliography{<name of your .bib file>} don't specify the file extension
-% 
-% don't specify bibliographystyleiz
-
-% In the References section, cite the data/software described in the Availability Statement (this includes primary and processed data used for your research).
-For details on data/software citation as well as examples, see the Data & Software Citation section of the Data & Software for Authors guidance
-% https://www.agu.org/Publish-with-AGU/Publish/Author-Resources/Data-and-Software-for-Authors#citation
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-\bibliographystyle{apa} 
+\bibliographystyle{apalike}
 \bibliography{PaperLib.bib}

### 2023-08-17 Delaneyetal2023b.tex
Updates from Overleaf

@@
 In natural systems, variable mobilization and deposition of sediment occur as water and sediment discharge accumulate as they move through the two-dimensional channel network will compound the already substantial variability in sediment mobilization parameters accounted for here in the lumped parameterization.
-Such variability could be especially strong in the subglacial system where melt water pathways change in response to the evolving hydraulic gradient \citep{delaney2023}.
+Such variability could be especially strong in the subglacial system where meltwater pathways change in response to the evolving hydraulic gradient \citep{delaney2023}.
 

### 2023-08-17 Delaneyetal2023b.tex
more editing...

@@
 
-The parameterizations are first applied to a reference test case for each glacier that assumes  subglacial channel with $\beta=\frac{\pi}{2}$ and friction factors $f_r$ and $f_p$ are tuned so that the model reproduces reasonable velocities for both  the Greenland Ice Sheet and Alpine glaciers \citep[$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$][]{werder2010b,chandler2013}.
+The parameterizations are first applied to a reference test case for each glacier that assumes  subglacial channel with $\beta=\frac{\pi}{2}$ and friction factors $f_r$ and $f_p$ are tuned so that the parameterizations reproduce reasonable velocities for both  the Greenland Ice Sheet and Alpine glaciers \citep[$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$][]{werder2010b,chandler2013}.
 In both subaerial channels, the bed slope, $\nabla z_c$ is $0.05$.
@@
 Second, to characterize the variability in sediment discharge capacity in subglacial compared to subaerial channels further, $100$  parameterizations are applied to a range of different channel geometry factors ($\beta$ and $k$) and friction factors ($f_r$ and $f_p$).
-Model runs with parameter combinations  are culled if their mean subglacial water velocity over the season lies outside of the range between $0.75$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $2$\,\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity lies outside of the range $0.3$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with some measurements \citep[e.g.]{werder2010b,magnusson2012,chandler2013}.
-The standard deviation of a model output over the course of the season is used to quantify the variability of the model output.
+Runs with parameter combinations  are culled if their mean subglacial water velocity over the season lies outside of the range between $0.75$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $2$\,\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity lies outside of the range $0.3$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with some measurements \citep[e.g.]{werder2010b,magnusson2012,chandler2013}.
+The standard deviation of a output variable over the course of the season is used to quantify the variability of thats output.
 
-Lastly, model outputs are compared with the record of sediment discharge from the two glaciers using Spearman rank correlation to reduce the impact of the non-linear response of sediment discharge to the output.
+Lastly,  outputs are compared with the record of sediment discharge from the two glaciers using Spearman rank correlation to reduce the impact of the non-linear response of sediment discharge to the hydrological forcing.
+Rank correlation also reduces the impact uncertainties in the sediment transport record form \icesheet{}.
 
@@
   \includegraphics[width=0.9\linewidth]{Fig2.pdf}
-  \caption{Model outputs resulting from the hydrographs shown in panels a and e for the cases called \alpine (a-d) and \icesheet{} (e-h).".
+  \caption{Parameterization outputs resulting from the hydrographs shown in panels a and e for the cases called \alpine (a-d) and \icesheet{} (e-h).
 Gray (black) lines in a and e represent the subglacial channel size (water discharge).
-In other panels, blue lines represent model outputs from the subglacial channel, while orange lines represent the subaerial channel.
+In other panels, blue lines represent outputs from the subglacial channel, while orange lines represent the subaerial channel.
     Data in plots are at $15$\unit{min} intervals.
@@
   \caption{Relationship between water discharge and velocity, shear stress, and width-integrated shear stress for \alpine{}  (a-c) and \icesheet (d-f).
-Variables have been normalized to mean values for better comparison amongst the model runs.
-    $R_g$ ($R_f$) show the Spearman rank correlation values for the subglacial (subaerial) model outputs.
+Variables have been normalized to mean values for better comparison amongst the runs.
+    $R_g$ ($R_f$) show the Spearman rank correlation values for the subglacial (subaerial) outputs.
     Plots are shown at $15$\unit{min} interval.} 
@@
 
-Each water discharge value in the subaerial channel is associated with a single water velocity, shear stress, and width-integrated shear stress, resulting in a perfect rank correlation between water discharge and the model output (Figure~\ref{fig:Qw_vari} a--f).
-Conversely in the subglacial channel,  water velocity, shear stress, and width-integrated shear stress can vary substantially for a given water discharge and width-integrated shear stress varies with respect to the other two model outputs as well (Figure~\ref{fig:Qw_vari} a--f).
+Each water discharge value in the subaerial channel is associated with a single water velocity, shear stress, and width-integrated shear stress, resulting in a perfect rank correlation between water discharge and the output (Figure~\ref{fig:Qw_vari} a--f).
+Conversely in the subglacial channel,  water velocity, shear stress, and width-integrated shear stress can vary substantially for a given water discharge and width-integrated shear stress varies with respect to the other two  outputs as well (Figure~\ref{fig:Qw_vari} a--f).
 For instance, very high values of water velocity and shear stress can occur from minimal water discharge at $2.5$\,\unit{m}$^3$\,\unit{s}$^{-1}$ to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
@@
 
-The second model application examines the impact of a variety of channel shapes and friction factors on the relationship between water discharge and sediment transport capacity, in both systems.
+The second application examines the impact of a variety of channel shapes and friction factors on the relationship between water discharge and sediment transport capacity, in both systems.
 Across the range of parameter values examined, variability in water velocity and shear stress remains higher in the subglacial system compared to the subaerial one.
@@
 
-\subsection{Relationship between model outputs and observed sediment discharge}
+\subsection{Relationship between parameterization  outputs and observed sediment discharge}
 
-In comparing model outputs with suspended sediment discharge data, velocity, shear stress and width-integrated shear stress from the subaerial system rank correlate better with sediment discharge compared to subglacial outputs (Figure~\ref{fig:Qw_vari} g--l).
-The rank correlation among the three model outputs for the subaerial parameterization remains constant due to the absence of hysteresis in channel shape, as shown above (see Section~\ref{sect:fluv}).
+In comparing output variable with suspended sediment discharge data, velocity, shear stress and width-integrated shear stress from the subaerial system rank correlate better with sediment discharge compared to subglacial outputs (Figure~\ref{fig:Qw_vari} g--l).
+The rank correlation among the three output variables for the subaerial parameterization remains constant due to the absence of hysteresis in channel shape, as shown above (see Section~\ref{sect:fluv}).
 In the subglacial system, velocity, and shear stress in the subglacial system perform substantially worse compared to width-integrated shear stress in rank correlating with sediment discharge.
 In many both \alpine{} and \icesheet{}, the rank correlation of the subglacial system approaches that of the subaerial system.
-Over longer time aggregation periods in some cases, model outputs from the subglacial model outperform those from the subaerial model (see Figures S1-S6).
+Over longer time aggregation periods in some cases, outputs from the subglacial parameterization outperform those from the subaerial on (see Figures S1-S6).
 
@@
 In both cases, the exponent greater than $1$ in the sediment transport relationship magnifies the variability in sediment discharge beyond the highly variable shear stress and width-integrated shear stress described above (Figure~\ref{fig:multi_run}).
+The rapid increase and decreases in subglacial sediment transport capacity may, for example, compound the variability in sediment transport already present in subaerial systems \citep[Figure~\ref{fig:Qw_vari}; ][]{williams1989,jerolmack2010}.
+First, large variations in sediment transport capacity underneath glaciers may result in sediment mobilization and deposition processes in close spatial and temporal proximity to each other \citep{gimbert2016,perolo2018}.
 
@@
 In subglacial channels, peaks occur when water discharge increases at the fastest rate compared with the growth of the channel (Figure~\ref{fig:multi_run}).
-In turn, the parameterization of subglacial channels here contains feedback that limits sediment transport capacity as the subglacial channel grows with increasing water discharge capacity, effectively reducing water velocity.
+In turn, the parameterization of subglacial channels here contains a feedback that limits sediment transport capacity as the subglacial channel grows with increasing water discharge capacity, effectively reducing water velocity.
 Similar feedback occurs in subaerial systems \citep{phillips2016}, albeit at longer timescales of centuries or more, as opposed to days here.
 
-In large catchments and further downstream of proglacial margins, water discharge is generally less variable than at catchments' headwaters, where glaciers lie in
-alpine settings \citep[][]{swift2005,riihimaki2005,costa2017,vanas2017}.
+In large catchments and further downstream of proglacial margins, water discharge is generally less variable than at catchments' headwaters, where glaciers lie \citep[][]{swift2005,riihimaki2005,costa2017,vanas2017}.
 Here, reduced hydrological variability far  downstream of glaciers could result in smaller variations in sediment transport capacity in many subaerial systems compared to subglacial ones.
-Smaller variations here would occur as a result of hydrological forcings, as opposed to slowly evolving subglacial channel size.
+Smaller variations here would occur as a result of the hydrological forcing, as opposed to slowly evolving subglacial channel size.
 
@@
 
-The model outputs from the subaerial system rank correlate higher with sediment discharge in the \alpine and \icesheet systems here,
-despite model outputs showing the disparate relationship between sediment transport capacity and water discharge in subglacial systems.
+The  outputs from the subaerial system rank correlate higher with sediment discharge in the \alpine and \icesheet systems here.
+This increased performance occurs despite  output variables showing the disparate relationship between sediment transport capacity and water discharge in subglacial systems (Figure~\ref{fig:Qw_vari}).
 This could be attributed to several reasons.
 
-First, large variations in sediment transport capacity underneath glaciers may result in sediment mobilization and deposition processes in close spatial and temporal proximity to each other \citep{gimbert2016,perolo2018,delaney2023}.
-The rapid increase and decreases in sediment transport capacity may, for example, compound the variability in sediment transport already present in subaerial systems \citep[Figure~\ref{fig:Qw_vari}; ][]{williams1989,jerolmack2010}.
-Due to the chaotic nature of these sediment transport processes, differentiating the characteristics of sediment transport capacity in subglacial and subaerial systems could be difficult.
-Indeed, the model performance metrics of the subglacial width-integrated shear stress and the subaerial metrics are similar (Figures~\ref{fig:Qw_vari}\, i, l and S1-S6).
-
-Second, these measurement stations are $0.5$ to $2$\,\unit{km} downstream of the glacier terminus \citep{cowton2012,felix2022}, respectively.
+First, these measurement stations are $0.5$ to $2$\,\unit{km} downstream of the glacier terminus \citep{cowton2012,felix2022}, respectively.
 This could potentially be a long enough distance for the hydraulic characteristics to change from the pressurized subglacial regime to the open channel subaerial regime.
 The transition of subglacial to subaerial conditions may occur in the transition between pressurized and subaerial flow under glacier terminus \citep{perolo2018}, especially when cavity closure rates of the ice are slow here \citep{egli2021b}.
-Furthermore, sediment transport in the glacier margin may persist in a transport-limited regime where sediment is deposited and mobilized with respect to sediment transport capacity, resulting in a stronger correlation between sediment transport and water discharge just a short distance downstream.
+Furthermore,  if a transport-limited regime may persists at the glacier margin, then the representation of subaerial transport processes could be represented just a short distance downstream, as sediment transport could be quickly adjusted to these variations.
+Indeed, a strong correlation persists between sediment plume size and water discharge in the Kangerlussuaq fjord, ~ $20$\,\unit{km} downstream of the Leverett measurement site \citep{mcgrath2010}, possibly as a result of open channel flow conditions.
+However, at other marine-terminating glacier fjords in Greenland, a less consistent relationships persists between water discharge, or melt extend, and sediment plume size \citep{tedstone2012}.
 
-Third, the parameterizations here only consider sediment transport capacity, not the availability or routing of sediment below the glacier and through the proglacial area \citep{delaney2023}.
+Second, the parameterizations here consider sediment transport capacity, not the availability or routing of sediment below the glacier and through the proglacial area \citep{delaney2023}.
 The sediment discharge records could represent factors such as increased sediment accessibility when melt extends up glacier, thereby concurrently increasing sediment discharge with water discharge \citep[e.g.][]{delaney2020}.
-As a result, the sediment transport records presented here may represent a supply-limited sediment transport, less dependent on sediment transport capacity.
+As a result of sediment availability, the sediment transport records presented here may represent a supply-limited sediment transport, less dependent on sediment transport capacity.
 Furthermore, in subglacial systems, the threshold of motion for sediment can be reached more frequently, and across a range of water discharges, potentially resulting in additional sediment transport (Figure~\ref{fig:Qw_vari}\,a-f).
@@
 
-Regardless of the cause of the increased performance of water discharge compared to subglacial sediment transport capacity in the observations, model outputs suggest that additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to subaerial systems \citep{jansson2005,ganti2016}.
+%Performance metrics of the subglacial width-integrated shear stress and the subaerial metrics are similar (Figures~\ref{fig:Qw_vari}\, i, l and S1-S6).
+%Yet, further analysis would be needed to understand the roles of water discharge representing sediment availability, as opposed to  width-integrated shear stress, in this correlation.
+
+Regardless of the cause of the increased performance of water discharge compared to subglacial sediment transport capacity in the observations, parameterization outputs suggest that additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to subaerial systems \citep{jansson2005,ganti2016}.
 Records of sediment discharge have been used to establish the relationship, or lack thereof, between sediment discharge and climate in glacial systems \citep[e.g.][]{koppes2009a,willenbring2016,mariotti2021}.
-The variability in subglacial sediment transport capacity presented here mostly applies to short-time scales controlling the size of subglacial channels (Figure~\ref{fig:multi_run}).
+The variable relationship between water discharge and subglacial sediment transport capacity presented here mostly applies to the daily- to weekly- time scales controlling the size of subglacial channels (Figure~\ref{fig:multi_run}).
 Yet, identifying climatic signals in sediment transport from transport-limited glaciers may require higher thresholds of water discharge compared to subaerial systems \citep[Figure~\ref{fig:Qw_vari}][]{tofelde2021}.
@@
 In addition to water discharge, the glacier's sediment transport capacity reacts to the ice thickness, controlling the closure of the channel, and the surface slope of the glacier, controlling water velocity \citep[Section~\ref{sect:sub_mode}; ][]{rothlisberger1972,shreve1972,stevens2022}.
-Conversely, sediment transport in most subaerial systems typically increases with water discharge and hydraulic gradient, and the hydraulic gradient of most subaerial systems likely remains stable over time scales ranging from years to centuries \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wong2006,wickert2019}.
-Furthermore, if no sediment is available and the glacier is in a supply-limited state, often assumed in glacier landscape evolution models, then the sediment discharge record also represents additional processes of bedrock erosion from  water pressure variations and sliding  \citep{iverson2012,herman2015}, which are also complex.
+Conversely, sediment transport in most subaerial systems typically increases with water discharge and hydraulic gradient, and the hydraulic gradient of most subaerial systems likely remains stable over time scales ranging from years to millenia \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wong2006,wickert2019}.
+%Furthermore, if no sediment is available and the glacier is in a supply-limited state, often assumed in glacier landscape evolution models, then the sediment discharge record also represents additional processes of bedrock erosion from  water pressure variations and sliding  \citep{iverson2012,herman2015}, which are also complex.
 
@@
 Subaerial channels  alter their channel width and water velocity immediately in response to changing water discharge.
-In contrast, pressurized subglacial channels mostly accommodate changing water discharge by altering water velocity and shear stress, upon which sediment transport depends.
-The response of water velocity occurs because the size of subglacial channels evolves slowly compared to variations in water discharge.
+In contrast, pressurized subglacial channels largely accommodate changing water discharge by altering water velocity and shear stress, upon which sediment transport depends.
+Water velocity responds quickly to changes in water discharge because the size of subglacial channels evolves slowly compared to variations in water discharge.
 
-Model outputs show substantial differences in subglacial and subaerial channels' response to changing water discharge.
-However, subaerial hydraulics demonstrate a better representation of sediment discharge variations compared to subglacial hydraulics.
-This could be attributed to the role of open channel flow beneath the glacier terminus and in the proglacial area, or the influence of water discharge in driving sediment availability, rather than subglacial sediment discharge capacity.
+The parameterizations of these processes show substantial differences in subglacial and subaerial channels' response to changing water discharge.
+However, subaerial hydraulics demonstrate a better representation of observed sediment discharge variations.
+This could be attributed to the role of open channel flow at the glacier margin and in the proglacial area, or the influence of water discharge in driving sediment availability, rather than subglacial sediment discharge capacity.
 
-Considerations in interpretations of sediment discharge should be made given the findings presented here.
-These considerations include:
+Implications of the findings here include:
 \begin{enumerate}
-\item the inconsistent relationship between water discharge and sediment transport capacity underneath glaciers;
-\item the timing of peak sediment transport capacity below glaciers generally coincides with the large rates of change in water discharge, as opposed to the maximum quantity of water discharge;
-\item the potentially increased variability in sediment discharge capacity from underneath glaciers than in rivers they feed;
-\item the representations of water discharge such as subglacial sediment transport capacity, subglacial sediment access, or open channel processes near the glacier margin in sediment records collected near glaciers.
+\item the inconsistent relationship between water discharge and sediment transport capacity underneath glaciers, causing increased variability in sediment transport capacity compared to river systems;
+\item the timing of peak sediment transport capacity below glaciers generally coincides with large rates of change in water discharge, as opposed to the maximum quantity of water discharge;
+\item the possibility for water discharge to represent processes such as subglacial sediment transport capacity, subglacial sediment access, or open channel processes near the glacier margin. 
 \end{enumerate}
@@
 SNSF Project No. PZ00P2\_202024 provided  funding for I. Delaney.
+Code, with links to data, can be found at \url{https://bitbucket.org/IanDelaney/xsection/src/master/}.
 

### 2023-08-21 Delaneyetal2023b.tex
local chages here

@@
       Time intervals shown with markers denote $15$\,\unit{min}, $1$\,\unit{hr}, $12$\,\unit{hr}, $1$\,\unit{d}, $5$\,\unit{d}, $10$\,\unit{d}, and $15$\,\unit{d} aggregation periods.
+      Text in insets shows rank correlation between the variable and sediment discharge for the corresponding time period for both the subglacial (right column) and subaerial (left column) systems.
+      \ian{problem with $\tau_g$}
     } 
@@
 \subsection{Relationship between parameterization  outputs and observed sediment discharge}
-
-In comparing output variable with suspended sediment discharge data, velocity, shear stress and width-integrated shear stress from the subaerial system rank correlate better with sediment discharge compared to subglacial outputs (Figure~\ref{fig:Qw_vari} g--l).
-The rank correlation among the three output variables for the subaerial parameterization remains constant due to the absence of hysteresis in channel shape, as shown above (see Section~\ref{sect:fluv}).
-In the subglacial system, velocity, and shear stress in the subglacial system perform substantially worse compared to width-integrated shear stress in rank correlating with sediment discharge.
-In many both \alpine{} and \icesheet{}, the rank correlation of the subglacial system approaches that of the subaerial system.
-Over longer time aggregation periods in some cases, outputs from the subglacial parameterization outperform those from the subaerial on (see Figures S1-S6).
+When comparing the output variables with suspended sediment discharge data, velocity, shear stress, and width-integrated shear stress from the subaerial system generally show better correlation with sediment discharge compared to subglacial outputs across the tested subglacial channel parameters (Figures~\ref{fig:Qw_vari} g--l and \ref{fig:multi_run}).
+The rank correlation among the three output variables for the subaerial parameterization remains more or less constant despite the absence of hysteresis in channel shape and consistent scaling with water discharge, as shown above (Figure \ref{fig:Qw_vari} and Section~\ref{sect:fluv}).
+In the subglacial system, velocity and shear stress perform substantially worse in terms of rank correlation with sediment discharge compared to width-integrated shear stress (Figures~ \ref{fig:multi_run}).
+In both \alpine{} and \icesheet{}, the rank correlation of the subglacial system approaches that of the subaerial system with longer time aggregation periods (Figures~ \ref{fig:multi_run}).
 
 \section{Discussion}
-\subsection{Context of the sediment transport parameterizations}
+\subsection{Context of sediment mobilization differences}
 Quantities that control sediment transport capacity, such as water velocity, shear stress, and width-integrated shear stress, vary differently in response to water discharge in subglacial channels compared to subaerial ones (Figure~\ref{fig:Qw_vari}).
@@
 
-The  outputs from the subaerial system rank correlate higher with sediment discharge in the \alpine and \icesheet systems here.
-This increased performance occurs despite  output variables showing the disparate relationship between sediment transport capacity and water discharge in subglacial systems (Figure~\ref{fig:Qw_vari}).
-This could be attributed to several reasons.
+Outputs from the subaerial system rank correlate higher with sediment discharge in the \alpine and \icesheet systems here.
+The increased performance occurs despite  output variables showing the disparate relationship between sediment transport capacity and water discharge in subglacial systems (Figure~\ref{fig:Qw_vari}), which could occur for two reasons.
 
@@
 Indeed, a strong correlation persists between sediment plume size and water discharge in the Kangerlussuaq fjord, ~ $20$\,\unit{km} downstream of the Leverett measurement site \citep{mcgrath2010}, possibly as a result of open channel flow conditions.
-However, at other marine-terminating glacier fjords in Greenland, a less consistent relationships persists between water discharge, or melt extend, and sediment plume size \citep{tedstone2012}.
+This lies in contrast to other marine-terminating glacier fjords in Greenland that could represent the subglacial signal where less consistent relationships persists between water discharge, or melt extend, and sediment plume size \citep{tedstone2012}.
 
@@
 \section{Conclusions}
+
 Sediment transport capacity is largely driven by the velocity of water flowing through a channel and the channel's width.
@@
 
-The parameterizations of these processes show substantial differences in subglacial and subaerial channels' response to changing water discharge.
-However, subaerial hydraulics demonstrate a better representation of observed sediment discharge variations.
+Despite these differences, parameterizations of subaerial hydraulics generally demonstrate a better representation of observed sediment discharge variations compared to subglacial channels.
 This could be attributed to the role of open channel flow at the glacier margin and in the proglacial area, or the influence of water discharge in driving sediment availability, rather than subglacial sediment discharge capacity.
 
-Implications of the findings here include:
+Findings here potentially impacting the interpretation of sediment transport records include:
 \begin{enumerate}

### 2023-08-24 Delaneyetal2023b.tex
from overleaf -> starting Andrew's comments

@@
 
-\begin{abstract} % 150 words
+\begin{abstract}
 
@@
   Here, parameterizations of sediment transport capacity for pressurized flow under glaciers and subaerial open channel flow are applied to hydrographs from an Alpine glacier and the Greenland Ice Sheet.
-  Results of these parameterizations show that sediment transport capacity varies far more in subglacial channels compared to subaerial channels both over time and with respect to water discharge.
-  In the subglacial case, high sediment transport capacity can persist across a wide range of water discharges.
+  Results of these parameterizations show that sediment transport capacity varies far more in subglacial channels compared to subaerial ones both over time and with respect to water discharge.
+  Furthermore, high subglacial sediment transport capacity can persist across a wide range of water discharges.
   However, changing channel width can accommodate some  variability in sediment transport capacity in subaerial channels.
-  The different responses of water discharge on sediment transport capacity in subglacial channels, compared to subaerial ones, may impact the occurrence of sediment transport capacity reaching mobilization thresholds and thus should be considered in the interpretation of sediment discharge records from glaciers.
+  Despite the different responses of sediment transport capacity to water discharge in the two-channel types, measurements of sediment discharge from the catchments generally correlate better with subaerial hydraulics.
+  This suggests that other processes such as sediment access or subaerial processes in the proglacial area are represented in the sediment discharge records.
+  The different responses of water discharge on sediment transport capacity in subglacial channels, compared to subaerial ones, increase the occurrence of sediment transport capacity reaching mobilization thresholds, not necessarily related to water discharge quantity.
 
@@
 It follows that an increase in water discharge $Q_w$ will cause the shear stress across the channel $\tau_t$ to respond to the water velocity $v$, as opposed to the subaerial channel where channel width $w_c$ evolves as well.
-\begin{center}
+
   \begin{figure}[h]
+  \centering
     \includegraphics[width=0.8\linewidth]{Fig1.pdf}
@@
   \end{figure}
-\end{center}
 % \ian{Come back to Daniel's comments}
@@
 Parameterizations are applied to hydrological records from an Alpine glacier in Switzerland (Fieschergletscher), and  a land-terminating glacier in Greenland (Leverett glacier).
-The parameterizations' lumped and simplistic nature is intended to show the disparity between water discharge and sediment transport in subglacial systems  without considering effects such as the upstream drainage network and the role of sediment supply.
+The parameterizations' lumped and simplistic nature is intended to show the disparity between water discharge and sediment transport capacity in subglacial systems  without considering effects such as the upstream drainage network and the role of sediment supply.
 Outputs are used to explore differences in the relationships between  water discharge, channel geometry, and water velocity in the subglacial and subaerial channels. 
@@
 \label{sect:meth}
-Parameterizations below (Sections ~\ref{sect:sub_mode}~and~\ref{sect:fluv}) represent relationships amongst water discharge, subglacial water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
+Parameterizations below (Sections ~\ref{sect:sub_mode}~and~\ref{sect:fluv}) represent relationships amongst water discharge, water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
 Because sediment transport largely depends on shear stress applied by water flow \citep{shields1936}, both parameterizations calculate this value and integrate  it across the channel bed (Figure \ref{fig:cartoon}).
@@
     \includegraphics[width=0.9\linewidth]{Fig4.png}
-    \caption{Variance (measured by the standard deviation) in water velocity, shear stress, and total shear stress for different time aggregations across a large number of subglacial and subaerial channel shapes and friction factors The subglacial (blue) and subaerial (orange) systems are distinguished as are \alpine{} (a-c) and \icesheet{} (d-f).
-      Shaded areas denote the $25^{\mathrm{th}}$ and $75^{\mathrm{th}}$ quantiles of the $100$ parameter combinations tested for each of the systems (Section~\ref{sect:imp}).
+    \caption{Variance (measured by the standard deviation) in water velocity, shear stress, and total shear stress for different time aggregations across the range of subglacial and subaerial channel shapes and friction factors.
+    %The subglacial (blue) and subaerial (orange) systems are distinguished as are \alpine{} (a-c) and \icesheet{} (d-f).
+      Shaded areas denote the $5^{\mathrm{th}}$ and $95^{\mathrm{th}}$ quantiles of the $100$ parameter combinations tested for each of the systems (Section~\ref{sect:imp}).
       Solid lines denote  mean values.
       Time intervals shown with markers denote $15$\,\unit{min}, $1$\,\unit{hr}, $12$\,\unit{hr}, $1$\,\unit{d}, $5$\,\unit{d}, $10$\,\unit{d}, and $15$\,\unit{d} aggregation periods.
-      Text in insets shows rank correlation between the variable and sediment discharge for the corresponding time period for both the subglacial (right column) and subaerial (left column) systems.
+      Text in columns shows rank correlation between the variable and sediment discharge for the corresponding time period for the subglacial (right column) and subaerial (left column) channels.
       \ian{problem with $\tau_g$}
@@
 In subglacial channels, peaks occur when water discharge increases at the fastest rate compared with the growth of the channel (Figure~\ref{fig:multi_run}).
-In turn, the parameterization of subglacial channels here contains a feedback that limits sediment transport capacity as the subglacial channel grows with increasing water discharge capacity, effectively reducing water velocity.
+In turn, the parameterization of subglacial channels here contains feedbacks that limits sediment transport capacity as the subglacial channel grows with increasing water discharge capacity, effectively reducing water velocity.
 Similar feedback occurs in subaerial systems \citep{phillips2016}, albeit at longer timescales of centuries or more, as opposed to days here.
@@
 The transition of subglacial to subaerial conditions may occur in the transition between pressurized and subaerial flow under glacier terminus \citep{perolo2018}, especially when cavity closure rates of the ice are slow here \citep{egli2021b}.
-Furthermore,  if a transport-limited regime may persists at the glacier margin, then the representation of subaerial transport processes could be represented just a short distance downstream, as sediment transport could be quickly adjusted to these variations.
+Furthermore,  if a transport-limited regime may persist at the glacier margin, then the representation of subaerial transport processes could be represented just a short distance downstream, as sediment transport could be quickly adjusted to these variations.
 Indeed, a strong correlation persists between sediment plume size and water discharge in the Kangerlussuaq fjord, ~ $20$\,\unit{km} downstream of the Leverett measurement site \citep{mcgrath2010}, possibly as a result of open channel flow conditions.
-This lies in contrast to other marine-terminating glacier fjords in Greenland that could represent the subglacial signal where less consistent relationships persists between water discharge, or melt extend, and sediment plume size \citep{tedstone2012}.
+This lies in contrast to other marine-terminating glacier fjords in Greenland that could represent the subglacial signal where less consistent relationships persist between water discharge, or melt extend, and sediment plume size \citep{tedstone2012}.
 
@@
 Records of sediment discharge have been used to establish the relationship, or lack thereof, between sediment discharge and climate in glacial systems \citep[e.g.][]{koppes2009a,willenbring2016,mariotti2021}.
-The variable relationship between water discharge and subglacial sediment transport capacity presented here mostly applies to the daily- to weekly- time scales controlling the size of subglacial channels (Figure~\ref{fig:multi_run}).
+The variable relationship between water discharge and subglacial sediment transport capacity presented here mostly applies to the daily- to weekly-time scales controlling the size of subglacial channels (Figure~\ref{fig:multi_run}).
 Yet, identifying climatic signals in sediment transport from transport-limited glaciers may require higher thresholds of water discharge compared to subaerial systems \citep[Figure~\ref{fig:Qw_vari}][]{tofelde2021}.
@@
 In addition to water discharge, the glacier's sediment transport capacity reacts to the ice thickness, controlling the closure of the channel, and the surface slope of the glacier, controlling water velocity \citep[Section~\ref{sect:sub_mode}; ][]{rothlisberger1972,shreve1972,stevens2022}.
-Conversely, sediment transport in most subaerial systems typically increases with water discharge and hydraulic gradient, and the hydraulic gradient of most subaerial systems likely remains stable over time scales ranging from years to millenia \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wong2006,wickert2019}.
+Conversely, sediment transport in most subaerial systems typically increases with water discharge and hydraulic gradient, and the hydraulic gradient of most subaerial systems likely remains stable over time scales ranging from years to millennia \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wong2006,wickert2019}.
 %Furthermore, if no sediment is available and the glacier is in a supply-limited state, often assumed in glacier landscape evolution models, then the sediment discharge record also represents additional processes of bedrock erosion from  water pressure variations and sliding  \citep{iverson2012,herman2015}, which are also complex.
@@
 
-Findings here potentially impacting the interpretation of sediment transport records include:
+Implications of the findings here include:
 \begin{enumerate}
@@
 \item the timing of peak sediment transport capacity below glaciers generally coincides with large rates of change in water discharge, as opposed to the maximum quantity of water discharge;
-\item the possibility for water discharge to represent processes such as subglacial sediment transport capacity, subglacial sediment access, or open channel processes near the glacier margin. 
+\item the possibility for water discharge to represent processes such as subglacial sediment transport capacity, subglacial sediment access, or open channel processes near the glacier margin
 \end{enumerate}
@@
 SNSF Project No. PZ00P2\_202024 provided  funding for I. Delaney.
-Code, with links to data, can be found at \url{https://bitbucket.org/IanDelaney/xsection/src/master/}.
+Code to make the figures, with links to data, can be found at \url{https://bitbucket.org/IanDelaney/xsection/src/master/}.
 

### 2023-08-24 Delaneyetal2023b.tex
few of Andrews comments

@@
 
-\item The channels' divergence in sediment transport response represents multiple processes represented by water discharge in sediment transport.
+\item Water discharge may represent processes including proglacial sediment mobilization and sediment access driving variations in sediment discharge.
 \end{itemize}
@@
 Increases in sediment transport  have been observed in Greenland \citep{bendixen2017}, the European Alps \citep{costa2017}, the Himalayas \citep{li2021}, and the Andes \citep{vergara2022}.
-In some of these regions, increased water discharge and glacier melt have been interpreted to be greater sediment transport capacity \citep{bendixen2017,costa2017,li2021}.
+In some of these regions, increased water discharge and glacier melt have been interpreted to yeild greater sediment transport capacity \citep{bendixen2017,costa2017,li2021}.
 Observed changes to sediment transport in glacierized catchments require examining the processes controlling sediment discharge in these catchments and its variations with water discharge  \citep[e.g.][]{riihimaki2005,swift2005}.
@@
 
+\ian{comeback to Andrew's comments here}
 Water and suspended sediment data are leveraged from two sites in this study.
@@
 This data was presented in part in \citet{cowton2012} and is available from \citet{tedstone2017} at a $5$\unit{min} time interval.
-Note that in some cases the removal and relocation of sensors created variations in the sediment concentration that do not result from natural processes.
+Note that in some cases the removal and relocation of sensors created variations in the sediment concentration not resulting from natural processes.
 
@@
 The parameterizations above are applied to hydrological records from the Fieschergletscher in the Swiss Alps and the Leverett glacier on the Greenland Ice Sheet.
-The hydrographs applied to the subglacial and subaerial parameterizations represent the subglacial flow of water as it leaves the glacier and transitions to subaerial flow in front of the glacier, as water moves through the catchment.
+The subglacial and subaerial parameterizations represent the subglacial flow of water exiting a glacier and transitioning to subaerial flow, as it moves through the catchment.
 Outputs of the parameterizations are meant to represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
@@
 Runs with parameter combinations  are culled if their mean subglacial water velocity over the season lies outside of the range between $0.75$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $2$\,\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity lies outside of the range $0.3$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with some measurements \citep[e.g.]{werder2010b,magnusson2012,chandler2013}.
-The standard deviation of a output variable over the course of the season is used to quantify the variability of thats output.
+The standard deviation of an output variable over the course of the season is used to quantify the variability of thats output.
 
 Lastly,  outputs are compared with the record of sediment discharge from the two glaciers using Spearman rank correlation to reduce the impact of the non-linear response of sediment discharge to the hydrological forcing.
-Rank correlation also reduces the impact uncertainties in the sediment transport record form \icesheet{}.
+Rank correlation also reduces the impact uncertainties from sensor disturbances in the sediment transport record from \icesheet{}.
 
@@
     Insets in e-h show the  peak melt event \icesheet{} denoted by the shaded area.
-    Note that the x and y axis are different for \alpine{} and \icesheet{} such that they each cover a summer.} 
+    Note that the x and y axis are different for \alpine{} and \icesheet{} such that they each cover a summer.
+    \ian{Andrew's comment: for a given channel width, subglacial channels are more effective at mobilising and transporting sediment that subaerial ones}
+  } 
   \label{fig:model_outs}
@@
 
-Over the course of the season, large peaks in water velocity and shear stress, along with width-integrated shear stress, occur in response to increases in water discharge in both cases.
+Over the course of the season, large peaks in water velocity and shear stress, along with width-integrated shear stress, occur in response to increases in water discharge in both \alpine{} and \icesheet{}.
 In the subglacial channel, these peaks generally occur over the period when water discharge increases at the fastest rate (Figure~\ref{fig:model_outs}).
 Conversely, in the subaerial channel, peaks in these variables occur when water discharge is at its highest value.
-As a result, peaks in the subglacial sediment transport variables occur  when water discharge change is greatest, as opposed to when the maximum amount of water discharge occurs.
+As a result, peaks in the subglacial sediment transport variables occur  when the rate of water discharge increase is greatest
+is greatest, as opposed to when the maximum amount of water discharge occurs.
 The timing of the greatest rate of change in water discharge occurs before the peaks in water discharge. 
-Here, sediment could be deposited in the proglacial area as the water depressurizes and is remobilized when sediment transport capacity peaks in the subaerial system.
+Here, sediment could be deposited in the proglacial area as the water depressurizes, then remobilized when sediment transport capacity peaks in the subaerial system.
 
@@
   \caption{Relationship between water discharge and velocity, shear stress, and width-integrated shear stress for \alpine{} (a-c) and \icesheet (d-f).
+    These outputs' relationship with sediment discharge is given for \alpine in (g-i) and \icesheet (j-l).
     Variables have been normalized to mean values for better comparison amongst the runs.
     $R_g$ ($R_f$) show the Spearman rank correlation values for the subglacial (subaerial) outputs.
-    Plots are shown at $15$\unit{min} interval.} 
+    Plots are shown at $15$\unit{min} interval.
+  } 
   \label{fig:Qw_vari}
@@
 For instance, very high values of water velocity and shear stress can occur from minimal water discharge at $2.5$\,\unit{m}$^3$\,\unit{s}$^{-1}$ to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
-In \icesheet, mean values of water velocity can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, effectively spanning the much of the water discharge range.
+In \icesheet, mean values of water velocity can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, effectively spanning much of the water discharge range.
 
@@
       Text in columns shows rank correlation between the variable and sediment discharge for the corresponding time period for the subglacial (right column) and subaerial (left column) channels.
-      \ian{problem with $\tau_g$}
       } 
@@
 The rapid increase and decreases in subglacial sediment transport capacity may, for example, compound the variability in sediment transport already present in subaerial systems \citep[Figure~\ref{fig:Qw_vari}; ][]{williams1989,jerolmack2010}.
-First, large variations in sediment transport capacity underneath glaciers may result in sediment mobilization and deposition processes in close spatial and temporal proximity to each other \citep{gimbert2016,perolo2018}.
+Large variations in sediment transport capacity underneath glaciers may result in sediment mobilization and deposition processes in close spatial and temporal proximity to each other \citep{gimbert2016,perolo2018}.
+Note also that both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress is homogeneous across the channel bed, despite the limitations of this assumption \citep[e.g.][]{yager2018}.
 
-Both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress is homogeneous across the channel bed, despite the limitations of this assumption \citep[e.g.][]{yager2018}.
-In the subaerial parameterization, this means that sediment is transported across all channel widths and water discharges, as opposed to large channel widths that occur during floods when much sediment transport occurs \citep{wolman1960}.
+% In the subaerial parameterization, this means that sediment is transported across all channel widths and water discharges, as opposed to large channel widths that occur during floods when much sediment transport occurs \citep{wolman1960}.
 
@@
 
-In large catchments and further downstream of proglacial margins, water discharge is generally less variable than at catchments' headwaters, where glaciers lie \citep[][]{swift2005,riihimaki2005,costa2017,vanas2017}.
-Here, reduced hydrological variability far  downstream of glaciers could result in smaller variations in sediment transport capacity in many subaerial systems compared to subglacial ones.
-Smaller variations here would occur as a result of the hydrological forcing, as opposed to slowly evolving subglacial channel size.
+%In large catchments and further downstream of proglacial margins, water discharge is generally less variable than at catchments' headwaters, where glaciers lie \citep[][]{swift2005,riihimaki2005,costa2017,vanas2017}.
+%Here, reduced hydrological variability far  downstream of glaciers could result in smaller variations in sediment transport capacity in many subaerial systems compared to subglacial ones.
+%Smaller variations here would occur as a result of the hydrological forcing, as opposed to slowly evolving subglacial channel size.
 
@@
 
-First, these measurement stations are $0.5$ to $2$\,\unit{km} downstream of the glacier terminus \citep{cowton2012,felix2022}, respectively.
-This could potentially be a long enough distance for the hydraulic characteristics to change from the pressurized subglacial regime to the open channel subaerial regime.
-The transition of subglacial to subaerial conditions may occur in the transition between pressurized and subaerial flow under glacier terminus \citep{perolo2018}, especially when cavity closure rates of the ice are slow here \citep{egli2021b}.
-Furthermore,  if a transport-limited regime may persist at the glacier margin, then the representation of subaerial transport processes could be represented just a short distance downstream, as sediment transport could be quickly adjusted to these variations.
-Indeed, a strong correlation persists between sediment plume size and water discharge in the Kangerlussuaq fjord, ~ $20$\,\unit{km} downstream of the Leverett measurement site \citep{mcgrath2010}, possibly as a result of open channel flow conditions.
-This lies in contrast to other marine-terminating glacier fjords in Greenland that could represent the subglacial signal where less consistent relationships persist between water discharge, or melt extend, and sediment plume size \citep{tedstone2012}.
+First, these measurement stations are $0.5$ to $2$\,\unit{km} downstream of the glacier terminus \citep{cowton2012,felix2022}, respectively, likely a long enough distance for the pressurized subglacial regime to transition to the open channel subaerial regime.
+The transition of subglacial to subaerial conditions may occur in the transition between pressurized and subaerial flow under glacier terminus \citep{perolo2018}, especially if cavity closure rates of the ice are slow  \citep{egli2021b}.
+Furthermore,  if a transport-limited regime persists at the glacier margin, then sediment can be easily mobilized and subaerial processes could be represented just a short distance downstream.
 
-Second, the parameterizations here consider sediment transport capacity, not the availability or routing of sediment below the glacier and through the proglacial area \citep{delaney2023}.
-The sediment discharge records could represent factors such as increased sediment accessibility when melt extends up glacier, thereby concurrently increasing sediment discharge with water discharge \citep[e.g.][]{delaney2020}.
-As a result of sediment availability, the sediment transport records presented here may represent a supply-limited sediment transport, less dependent on sediment transport capacity.
-Furthermore, in subglacial systems, the threshold of motion for sediment can be reached more frequently, and across a range of water discharges, potentially resulting in additional sediment transport (Figure~\ref{fig:Qw_vari}\,a-f).
-For instance, sediment discharge records from two glaciers in the Swiss Alps show that $40$ to $50$\% of a season's sediment discharge occurs when water discharge is below the $75^{\mathrm{th}}$ quantile of the season \citep{delaney2018}.
-This process may lead to glaciers evacuating sediment across a range of water discharges and transitioning to a supply-limited regime, where sediment discharge also scales with sediment production from glacier quarrying and abrasion \citep[e.g.][]{herman2015,ugelvig2018}.
+The strong link between sediment discharge capacity and open channel flow conditions could be manifested ~ $20$\,\unit{km} downstream of the Leverett measurement site, where a strong correlation persists between sediment plume size and water discharge in to the Kangerlussuaq fjord \citep{mcgrath2010}.
+This lies in contrast to other marine-terminating glacier fjords in Greenland where the subglacial channels are pressurized by ocean water at the ice front \citep[e.g.][]{how2017}. 
+In these catchments, a less consistent relationships may persist between water discharge, or melt extend, and sediment plume size \citep{tedstone2012}, possibly due to the variable link between sediment transport capacity and water discharge in subglacial conditions.
+
+Second, the sediment discharge records could represent factors such as increased sediment accessibility when melt extends up glacier, thereby concurrently increasing sediment discharge with water discharge \citep[e.g.][]{delaney2020}.
+In this case, the sediment transport records presented here may represent supply-limited sediment transport, less dependent on sediment transport capacity.
+The threshold of motion for sediment can be reached more frequently, and across a range of water discharges, in subglacial systems.
+Therefore, glaciers could well  persist in a supply-limited regime, where sediment discharge also scales with sediment production from glacier quarrying and abrasion, along with melt water access \citep[e.g.][]{herman2015,ugelvig2018,delaney2023}.
 Sediment exhaustion through this process may also explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \citep{overeem2017}.
+Note, however, that the absences of lags between sediment and water discharge may suggest that sediment is mobilized close to the measurement stations, as opposed to distal areas of the glacier bed  \citep[Figure~\ref{fig:model_outs}\,a and e][]{williams1989}.
+\ian{might need to tighten this better}
 
-%Performance metrics of the subglacial width-integrated shear stress and the subaerial metrics are similar (Figures~\ref{fig:Qw_vari}\, i, l and S1-S6).
-%Yet, further analysis would be needed to understand the roles of water discharge representing sediment availability, as opposed to  width-integrated shear stress, in this correlation.
-
-Regardless of the cause of the increased performance of water discharge compared to subglacial sediment transport capacity in the observations, parameterization outputs suggest that additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to subaerial systems \citep{jansson2005,ganti2016}.
+Regardless of the increased performance of water discharge compared to subglacial sediment transport capacity in the observations, parameterization outputs suggest that additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to subaerial systems \citep{jansson2005,ganti2016}.
 Records of sediment discharge have been used to establish the relationship, or lack thereof, between sediment discharge and climate in glacial systems \citep[e.g.][]{koppes2009a,willenbring2016,mariotti2021}.
@@
 Yet, identifying climatic signals in sediment transport from transport-limited glaciers may require higher thresholds of water discharge compared to subaerial systems \citep[Figure~\ref{fig:Qw_vari}][]{tofelde2021}.
-This higher threshold of perturbation by water discharge, which impacts sediment transport capacity, may persist in addition to the stochastic nature of erosion and deposition in fluvial environments \citep{castletort2003,jerolmack2010,romans2016}.
+This higher threshold of perturbation by water discharge, which impacts sediment transport capacity, might compound with the stochastic nature of erosion and deposition in fluvial environments \citep{castletort2003,jerolmack2010,romans2016}.
 
@@
 Conversely, sediment transport in most subaerial systems typically increases with water discharge and hydraulic gradient, and the hydraulic gradient of most subaerial systems likely remains stable over time scales ranging from years to millennia \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wong2006,wickert2019}.
-%Furthermore, if no sediment is available and the glacier is in a supply-limited state, often assumed in glacier landscape evolution models, then the sediment discharge record also represents additional processes of bedrock erosion from  water pressure variations and sliding  \citep{iverson2012,herman2015}, which are also complex.
+% Furthermore, if no sediment is available and the glacier is in a supply-limited state, often assumed in glacier landscape evolution models, then the sediment discharge record also represents additional processes of bedrock erosion from  water pressure variations and sliding  \citep{iverson2012,herman2015}, which are also complex.
+
+\ian{Comeback here and link with paragraph above better.}
 
@@
 \item the inconsistent relationship between water discharge and sediment transport capacity underneath glaciers, causing increased variability in sediment transport capacity compared to river systems;
-\item the timing of peak sediment transport capacity below glaciers generally coincides with large rates of change in water discharge, as opposed to the maximum quantity of water discharge;
-\item the possibility for water discharge to represent processes such as subglacial sediment transport capacity, subglacial sediment access, or open channel processes near the glacier margin
+\item the timing of peak sediment transport capacity below glaciers generally coincides with large rates of increasing in water discharge, as opposed to the maximum quantity of water discharge;
+\item the possibility for a consistent relationship between water and sediment discharge to represent processes such as subglacial sediment access, or open channel processes near the glacier margin, instead of subglacial sediment transport capacity.
+
 \end{enumerate}
@@
 SNSF Project No. PZ00P2\_202024 provided  funding for I. Delaney.
-Code to make the figures, with links to data, can be found at \url{https://bitbucket.org/IanDelaney/xsection/src/master/}.
+Code to make the figures, with links to the data, can be found at \url{https://bitbucket.org/IanDelaney/xsection/src/master/}.
 

### 2023-08-24 Delaneyetal2023b.tex
Small bits in discussion

@@
     Insets in e-h show the  peak melt event \icesheet{} denoted by the shaded area.
-    Note that the x and y axis are different for \alpine{} and \icesheet{} such that they each cover a summer.
-    \ian{Andrew's comment: for a given channel width, subglacial channels are more effective at mobilising and transporting sediment that subaerial ones}
+    Note that the x and y axes are different for \alpine{} and \icesheet{} such that they each cover a summer.
+    \ian{Andrew's comment: for a given channel width, subglacial channels are more effective at mobilizing and transporting sediment than subaerial ones}
   } 
@@
 
-The strong link between sediment discharge capacity and open channel flow conditions could be manifested ~ $20$\,\unit{km} downstream of the Leverett measurement site, where a strong correlation persists between sediment plume size and water discharge in to the Kangerlussuaq fjord \citep{mcgrath2010}.
+The strong link between sediment discharge capacity and open channel flow conditions could be manifested ~ $20$\,\unit{km} downstream of the Leverett measurement site, where a strong correlation persists between sediment plume size and water discharge into the Kangerlussuaq fjord \citep{mcgrath2010}.
 This lies in contrast to other marine-terminating glacier fjords in Greenland where the subglacial channels are pressurized by ocean water at the ice front \citep[e.g.][]{how2017}. 
@@
 In this case, the sediment transport records presented here may represent supply-limited sediment transport, less dependent on sediment transport capacity.
-The threshold of motion for sediment can be reached more frequently, and across a range of water discharges, in subglacial systems.
-Therefore, glaciers could well  persist in a supply-limited regime, where sediment discharge also scales with sediment production from glacier quarrying and abrasion, along with melt water access \citep[e.g.][]{herman2015,ugelvig2018,delaney2023}.
+Glaciers could well trend toward a supply-limited regime, because the threshold of motion for sediment can be reached more frequently, and across a range of water discharges, compared to subaerial systems.
+In a supply-limited system, sediment discharge would also scale with sediment production from glacier quarrying and abrasion, along with meltwater access \citep[e.g.][]{herman2015,ugelvig2018,delaney2023}.
  Sediment exhaustion through this process may also explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \citep{overeem2017}.
-Note, however, that the absences of lags between sediment and water discharge may suggest that sediment is mobilized close to the measurement stations, as opposed to distal areas of the glacier bed  \citep[Figure~\ref{fig:model_outs}\,a and e][]{williams1989}.
-\ian{might need to tighten this better}
+Note, however, that the absence of lags between sediment and water discharge may suggest that sediment is mobilized close to the measurement stations, as opposed to distal areas of the glacier bed as in the supply-limited regime \citep[Figure~\ref{fig:model_outs}\,a and e][]{williams1989}.
 

### 2023-08-25 Delaneyetal2023b.tex
small changes

@@
 
-\item Water discharge may represent processes including proglacial sediment mobilization and sediment access driving variations in sediment discharge.
+\item In interpreting sediment transport records, water discharge may represent processes including proglacial sediment mobilization and sediment access driving variations in sediment discharge.
 \end{itemize}
@@
 Runs with parameter combinations  are culled if their mean subglacial water velocity over the season lies outside of the range between $0.75$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $2$\,\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity lies outside of the range $0.3$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with some measurements \citep[e.g.]{werder2010b,magnusson2012,chandler2013}.
-The standard deviation of an output variable over the course of the season is used to quantify the variability of thats output.
+The standard deviation of an output variable over the course of the season is used to quantify the variability of that output.
 
@@
 This lies in contrast to other marine-terminating glacier fjords in Greenland where the subglacial channels are pressurized by ocean water at the ice front \citep[e.g.][]{how2017}. 
-In these catchments, a less consistent relationships may persist between water discharge, or melt extend, and sediment plume size \citep{tedstone2012}, possibly due to the variable link between sediment transport capacity and water discharge in subglacial conditions.
+In these catchments, a less consistent relationships may persist between water discharge or melt extend and sediment plume size \citep{tedstone2012}, possibly due to the variable relationship between subglacial sediment transport capacity and water discharge.
 
@@
 In this case, the sediment transport records presented here may represent supply-limited sediment transport, less dependent on sediment transport capacity.
-Glaciers could well trend toward a supply-limited regime, because the threshold of motion for sediment can be reached more frequently, and across a range of water discharges, compared to subaerial systems.
+Glaciers could well trend toward a supply-limited regime, because the threshold of motion for sediment can be reached more frequently, and across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
 In a supply-limited system, sediment discharge would also scale with sediment production from glacier quarrying and abrasion, along with meltwater access \citep[e.g.][]{herman2015,ugelvig2018,delaney2023}.
  Sediment exhaustion through this process may also explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \citep{overeem2017}.
-Note, however, that the absence of lags between sediment and water discharge may suggest that sediment is mobilized close to the measurement stations, as opposed to distal areas of the glacier bed as in the supply-limited regime \citep[Figure~\ref{fig:model_outs}\,a and e][]{williams1989}.
+Note, however, that the absence of lags between sediment and water discharge may suggest that sediment is mobilized close to the measurement stations, as opposed to distal areas of the glacier bed as in the supply-limited regime \citep[Figure~\ref{fig:model_outs}\, a and e; ][]{williams1989}.
 
@@
 The variable relationship between water discharge and subglacial sediment transport capacity presented here mostly applies to the daily- to weekly-time scales controlling the size of subglacial channels (Figure~\ref{fig:multi_run}).
-Yet, identifying climatic signals in sediment transport from transport-limited glaciers may require higher thresholds of water discharge compared to subaerial systems \citep[Figure~\ref{fig:Qw_vari}][]{tofelde2021}.
-This higher threshold of perturbation by water discharge, which impacts sediment transport capacity, might compound with the stochastic nature of erosion and deposition in fluvial environments \citep{castletort2003,jerolmack2010,romans2016}.
+Yet, identifying climatic signals in sediment transport from the transport-limited component of glacier systems may require higher thresholds of water discharge compared to subaerial systems \citep[Figure~\ref{fig:Qw_vari}; ][]{tofelde2021}.
+This occurs as the variable relationship between water discharge and sediment transport capacity
+may compound  the already stochastic nature of erosion and deposition in fluvial environments \citep{castletort2003,jerolmack2010}.
 
-In addition to water discharge, the glacier's sediment transport capacity reacts to the ice thickness, controlling the closure of the channel, and the surface slope of the glacier, controlling water velocity \citep[Section~\ref{sect:sub_mode}; ][]{rothlisberger1972,shreve1972,stevens2022}.
-Conversely, sediment transport in most subaerial systems typically increases with water discharge and hydraulic gradient, and the hydraulic gradient of most subaerial systems likely remains stable over time scales ranging from years to millennia \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wong2006,wickert2019}.
-% Furthermore, if no sediment is available and the glacier is in a supply-limited state, often assumed in glacier landscape evolution models, then the sediment discharge record also represents additional processes of bedrock erosion from  water pressure variations and sliding  \citep{iverson2012,herman2015}, which are also complex.
-
-\ian{Comeback here and link with paragraph above better.}
+Over time scales longer than days to weeks, the glacier's sediment transport capacity reacts to the ice thickness, controlling the channel closure rate, and the surface slope of the glacier \citep[Figure~\ref{fig:multi_run}, ~ Section~\ref{sect:sub_mode}; ][]{rothlisberger1972,shreve1972,stevens2022}.
+If the glacier is in a supply-limited state, often assumed in landscape evolution models, then the sediment discharge record also represents additional processes of sediment availability and bedrock erosion from  water pressure variations and sliding  \citep[e.g.][]{iverson2012,herman2015,delaney2023}.
+This multitude of processes lies in contrast to most subaerial systems, where transport capacity typically increases with water discharge and hydraulic gradient, and the hydraulic gradient generally remains stable over time scales ranging from years to millennia \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wickert2019}.
+The stronger link between sediment export and hydrological forcing in subaerial channels means that variations in sediment transport records can be more easily attributed to climatic conditions than in  subglacial systems
 
@@
 \item the timing of peak sediment transport capacity below glaciers generally coincides with large rates of increasing in water discharge, as opposed to the maximum quantity of water discharge;
-\item the possibility for a consistent relationship between water and sediment discharge to represent processes such as subglacial sediment access, or open channel processes near the glacier margin, instead of subglacial sediment transport capacity.
+\item the possibility for a consistent relationship between water and sediment discharge to represent subglacial sediment access or open channel hydraulics near the glacier margin, instead of subglacial sediment transport capacity.
 

### 2023-08-25 Delaneyetal2023b.tex
more small changes

@@
     \caption{Variance (measured by the standard deviation) in water velocity, shear stress, and total shear stress for different time aggregations across the range of subglacial and subaerial channel shapes and friction factors.
-    %The subglacial (blue) and subaerial (orange) systems are distinguished as are \alpine{} (a-c) and \icesheet{} (d-f).
-      Shaded areas denote the $5^{\mathrm{th}}$ and $95^{\mathrm{th}}$ quantiles of the $100$ parameter combinations tested for each of the systems (Section~\ref{sect:imp}).
+      Shaded areas denote the minimum and maximum of the $100$ parameter combinations tested for each of the systems (Section~\ref{sect:imp}).
       Solid lines denote  mean values.
@@
 This multitude of processes lies in contrast to most subaerial systems, where transport capacity typically increases with water discharge and hydraulic gradient, and the hydraulic gradient generally remains stable over time scales ranging from years to millennia \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wickert2019}.
-The stronger link between sediment export and hydrological forcing in subaerial channels means that variations in sediment transport records can be more easily attributed to climatic conditions than in  subglacial systems
+The stronger link between sediment transport capacity and hydrological forcing in subaerial channels means that variations in sediment transport records can be more easily attributed to climatic conditions than in  subglacial systems.
 

### 2023-08-25 Delaneyetal2023b.tex
Updates from Overleaf

--- 2023-08-25 Delaneyetal2023b.tex
@@
     ianarburua.delaney@unil.ch},
-  Daniel Farinotti\footnote{Laboratory of Hydraulics, Hydrology and Glaciology (VAW), ETH-Z\"urich, H\"onggerbergring 26, 8093 89 Z\"urich, Switzerland}$^{,}$\footnote{Swiss Federal Institute for Forest, Snow and Landscape Research (WSL) Z\"uricherstrasse 111, 8903 1011 Birmensdorf, Switzerland},
-  Andrew J. Tedstone\footnote{Department of Geosciences, University of Fribourg, Ch. du Musée 1700, Fribourg, Switzerland}
+  Andrew J. Tedstone\footnote{Department of Geosciences, University of Fribourg, Ch. du Musée 1700, Fribourg, Switzerland},
+  Daniel Farinotti\footnote{Laboratory of Hydraulics, Hydrology and Glaciology (VAW), ETH-Z\"urich, H\"onggerbergring 26, 8093 89 Z\"urich, Switzerland}$^{,}$\footnote{Swiss Federal Institute for Forest, Snow and Landscape Research (WSL) Z\"uricherstrasse 111, 8903 1011 Birmensdorf, Switzerland}
 
@@
 Increases in sediment transport  have been observed in Greenland \citep{bendixen2017}, the European Alps \citep{costa2017}, the Himalayas \citep{li2021}, and the Andes \citep{vergara2022}.
-In some of these regions, increased water discharge and glacier melt have been interpreted to yeild greater sediment transport capacity \citep{bendixen2017,costa2017,li2021}.
+In some of these regions, increased water discharge and glacier melt have been interpreted to yield greater sediment transport capacity \citep{bendixen2017,costa2017,li2021}.
 Observed changes to sediment transport in glacierized catchments require examining the processes controlling sediment discharge in these catchments and its variations with water discharge  \citep[e.g.][]{riihimaki2005,swift2005}.
@@
 
-\ian{comeback to Andrew's comments here}
-Water and suspended sediment data are leveraged from two sites in this study.
-Water discharge and suspended sediment concentration were collected at a $1$\unit{min} interval from the Fieschergletscher in the Swiss Alps  continuously over the period 2014--2021 ($46^\circ\,29'\,07"$ N, $8^\circ\,08'\,34"$ E).
+Water and suspended sediment data are leveraged from an Alpine and an ice sheet setting.
+The Alpine site is  Fieschergletscher in the Swiss Alps ($46^\circ\,29'\,07"$ N, $8^\circ\,08'\,34"$ E).
+Water discharge and suspended sediment concentration were collected here at a $1$\unit{min} interval continuously over the period 2014--2021. 
 A description of the data and its collection is available in \citet{felix2022} and \citet{felix2021}.
 
-Additionally, water discharge and suspended sediment concentration were collected roughly $2$\unit{km} downstream from Leverett Glacier in Greenland over  periods from June to September from 2009 to 2012 ($67^\circ\,03'\,50"$ N, $50^\circ\,12'\,59"$ W).
+The Leverett Glacier in Greenland served as the ice sheet setting.
+Water discharge and suspended sediment concentration were collected roughly $2$\unit{km} downstream from the terminus over  periods from June to September from 2009 to 2012 ($67^\circ\,03'\,50"$ N, $50^\circ\,12'\,59"$ W).
 This data was presented in part in \citet{cowton2012} and is available from \citet{tedstone2017} at a $5$\unit{min} time interval.
@@
 
-Over time scales longer than days to weeks, the glacier's sediment transport capacity reacts to the ice thickness, controlling the channel closure rate, and the surface slope of the glacier \citep[Figure~\ref{fig:multi_run}, ~ Section~\ref{sect:sub_mode}; ][]{rothlisberger1972,shreve1972,stevens2022}.
+Over time scales longer than days to weeks, the glacier's sediment transport capacity reacts to the ice thickness, controlling the channel closure rate, and the surface slope of the glacier, in addition to water discharge \citep[Figure~\ref{fig:multi_run}, ~ Section~\ref{sect:sub_mode}; ][]{rothlisberger1972,shreve1972,stevens2022}.
 If the glacier is in a supply-limited state, often assumed in landscape evolution models, then the sediment discharge record also represents additional processes of sediment availability and bedrock erosion from  water pressure variations and sliding  \citep[e.g.][]{iverson2012,herman2015,delaney2023}.
@@
 \item the timing of peak sediment transport capacity below glaciers generally coincides with large rates of increasing in water discharge, as opposed to the maximum quantity of water discharge;
-\item the possibility for a consistent relationship between water and sediment discharge to represent subglacial sediment access or open channel hydraulics near the glacier margin, instead of subglacial sediment transport capacity.
+\item a potentially consistent relationship between water and sediment discharge may represent subglacial sediment access or open channel hydraulics near the glacier margin, instead of subglacial sediment transport capacity.
 

### 2023-08-28 Delaneyetal2023b.tex
polishing

@@
 Increases in sediment transport  have been observed in Greenland \citep{bendixen2017}, the European Alps \citep{costa2017}, the Himalayas \citep{li2021}, and the Andes \citep{vergara2022}.
-In some of these regions, increased water discharge and glacier melt have been interpreted to yield greater sediment transport capacity \citep{bendixen2017,costa2017,li2021}.
+In some regions, increased water discharge and glacier melt have been interpreted to yield greater sediment transport capacity \citep{bendixen2017,costa2017,li2021}.
 Observed changes to sediment transport in glacierized catchments require examining the processes controlling sediment discharge in these catchments and its variations with water discharge  \citep[e.g.][]{riihimaki2005,swift2005}.
@@
 
-In a transport-limited regime, sediment discharge responds to the sediment transport capacity or the amount of sediment that the water can carry.
+In a transport-limited regime, sediment discharge responds to the sediment transport capacity or the amount of sediment the water can carry.
 Sediment transport capacity depends on the shear stress between water and sediment it flows  over \citep{shields1936,meyer1948,engelund1967} and the width of the channel bottom $w_c$ over which to mobilize sediment.
@@
 Parameterizations are applied to hydrological records from an Alpine glacier in Switzerland (Fieschergletscher), and  a land-terminating glacier in Greenland (Leverett glacier).
-The parameterizations' lumped and simplistic nature is intended to show the disparity between water discharge and sediment transport capacity in subglacial systems  without considering effects such as the upstream drainage network and the role of sediment supply.
-Outputs are used to explore differences in the relationships between  water discharge, channel geometry, and water velocity in the subglacial and subaerial channels. 
+The parameterizations' lumped nature is intended to isolate the disparity between water discharge and sediment transport capacity in subglacial systems, independent of the upstream drainage network and sediment access.
+Outputs demonstrate differences in the relationships between  water discharge, channel geometry, and water velocity in the subglacial and subaerial channels. 
 Furthermore, outputs are compared to sediment discharge data from these two catchments.
@@
 Water and suspended sediment data are leveraged from an Alpine and an ice sheet setting.
-The Alpine site is  Fieschergletscher in the Swiss Alps ($46^\circ\,29'\,07"$ N, $8^\circ\,08'\,34"$ E).
+The Alpine site (\alpine) is  Fieschergletscher in the Swiss Alps ($46^\circ\,29'\,07"$ N, $8^\circ\,08'\,34"$ E).
 Water discharge and suspended sediment concentration were collected here at a $1$\unit{min} interval continuously over the period 2014--2021. 
-A description of the data and its collection is available in \citet{felix2022} and \citet{felix2021}.
+See \citet{felix2022} and \citet{felix2021} for more details.
 
-The Leverett Glacier in Greenland served as the ice sheet setting.
-Water discharge and suspended sediment concentration were collected roughly $2$\unit{km} downstream from the terminus over  periods from June to September from 2009 to 2012 ($67^\circ\,03'\,50"$ N, $50^\circ\,12'\,59"$ W).
-This data was presented in part in \citet{cowton2012} and is available from \citet{tedstone2017} at a $5$\unit{min} time interval.
+The Leverett Glacier in Greenland (\icesheet) serves as the ice sheet setting.
+Water discharge and suspended sediment concentration data were collected roughly $2$\unit{km} downstream from the terminus from June to September 2009 to 2012 ($67^\circ\,03'\,50"$ N, $50^\circ\,12'\,59"$ W).
+This data is available from \citet{tedstone2017} at a $5$\unit{min} time interval.
 Note that in some cases the removal and relocation of sensors created variations in the sediment concentration not resulting from natural processes.
+% was presented in part in \citet{cowton2012} and
 
@@
 Parameterizations below (Sections ~\ref{sect:sub_mode}~and~\ref{sect:fluv}) represent relationships amongst water discharge, water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
-Because sediment transport largely depends on shear stress applied by water flow \citep{shields1936}, both parameterizations calculate this value and integrate  it across the channel bed (Figure \ref{fig:cartoon}).
-The evaluation of shear stress omits a grain size parameter and, thus, leaves open the selection of a sediment transport relationship  \citep[e.g.][]{shields1936,meyer1948} to  calculate sediment transport capacity.
-As a result, relative relationships in shear stress, as opposed to sediment transport capacity, are examined.
+Both parameterizations calculate   water velocity, shear stress and width-integrated shear stress, upon which sediment transport depends \citep[Figure \ref{fig:cartoon}; ][]{shields1936}.
+The evaluation of these variables omits the selection of a sediment transport relationship and grain-size parameter \citep[e.g.][]{shields1936,meyer1948} needed to calculate sediment transport capacity from shear stress.
 
@@
   \centering
-  \caption{Variables, parameters, and constants used in this work.
-Where two values are given, the first refers to  \alpine{}, a case from Fieschergletscher, and the second to a glacier marginal to the \icesheet{} a case from Leverett Glacier }
+  \caption{Variables, parameters, and constants used in this work.} 
+  %Where two values are given, the first refers to  \alpine{}, a case from Fieschergletscher, and the second to a glacier marginal to the \icesheet{} a case from Leverett Glacier }
   \begin{tabular}{ l  c  c c }
@@
 Outputs of the parameterizations are meant to represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
-To generalize these cases, the Fieschergletscher case is referred to as \alpine{} and is exemplified by relatively thin ice thickness ($h_{ice}$: $225$\,\unit{m}), low water discharge ( $\le\,20$\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
-The Leverett glacier case is referred to as \icesheet{} and is exemplified by thick ice  ($h_{ice}$: $700$\,\unit{m}), high water discharge ($\le\,300$\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).
+To generalize these cases, the Fieschergletscher case is referred to as \alpine{} and is exemplified by relatively thin ice thickness ($h_{ice}$: $225$\,\unit{m}), low water discharge ( $<\,20$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
+The Leverett glacier case is referred to as \icesheet{} and is exemplified by thick ice  ($h_{ice}$: $700$\,\unit{m}), high water discharge (generally $<\,500$\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).
 
@@
 Second, to characterize the variability in sediment discharge capacity in subglacial compared to subaerial channels further, $100$  parameterizations are applied to a range of different channel geometry factors ($\beta$ and $k$) and friction factors ($f_r$ and $f_p$).
-Runs with parameter combinations  are culled if their mean subglacial water velocity over the season lies outside of the range between $0.75$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $2$\,\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity lies outside of the range $0.3$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with some measurements \citep[e.g.]{werder2010b,magnusson2012,chandler2013}.
+Runs with parameter combinations are culled if their mean subglacial water velocity over the season lies outside of the range between $0.75$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $2$\,\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity lies outside of the range $0.3$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with some measurements \citep[e.g.][]{werder2010b,magnusson2012,chandler2013}.
 The standard deviation of an output variable over the course of the season is used to quantify the variability of that output.
@@
     Note that the x and y axes are different for \alpine{} and \icesheet{} such that they each cover a summer.
-    \ian{Andrew's comment: for a given channel width, subglacial channels are more effective at mobilizing and transporting sediment than subaerial ones}
+    %    \ian{Andrew's comment: for a given channel width, subglacial channels are more effective at mobilizing and transporting sediment than subaerial ones}
   } 
@@
 
-The strong link between sediment discharge capacity and open channel flow conditions could be manifested ~ $20$\,\unit{km} downstream of the Leverett measurement site, where a strong correlation persists between sediment plume size and water discharge into the Kangerlussuaq fjord \citep{mcgrath2010}.
-This lies in contrast to other marine-terminating glacier fjords in Greenland where the subglacial channels are pressurized by ocean water at the ice front \citep[e.g.][]{how2017}. 
-In these catchments, a less consistent relationships may persist between water discharge or melt extend and sediment plume size \citep{tedstone2012}, possibly due to the variable relationship between subglacial sediment transport capacity and water discharge.
+The consistent relationship between sediment discharge capacity and open channel flow conditions could be pronounced ~ $20$\,\unit{km} downstream of the Leverett measurement site, where a strong correlation persists between sediment plume size and water discharge into the Kangerlussuaq fjord \citep{mcgrath2010}.
+In marine-terminating glacier catchments, a less consistent relationship may persist between water discharge or melt extend and sediment plume size \citep{tedstone2012}.
+The lack of correlation could be due to the variable relationship between subglacial sediment transport capacity and water discharge at marine-terminating glaciers where subglacial channels are pressurized by ocean water at the ice front \citep[e.g.][]{how2017}.
 
-Second, the sediment discharge records could represent factors such as increased sediment accessibility when melt extends up glacier, thereby concurrently increasing sediment discharge with water discharge \citep[e.g.][]{delaney2020}.
-In this case, the sediment transport records presented here may represent supply-limited sediment transport, less dependent on sediment transport capacity.
-Glaciers could well trend toward a supply-limited regime, because the threshold of motion for sediment can be reached more frequently, and across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
+Second, the sediment discharge records could represent processes such as increased sediment accessibility when melt extends up glacier, thereby concurrently increasing sediment and water discharge \citep[e.g.][]{delaney2020}.
+A supply-limited regime could well persist at many glaciers, because the threshold of motion for sediment can be reached more frequently, and across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
 In a supply-limited system, sediment discharge would also scale with sediment production from glacier quarrying and abrasion, along with meltwater access \citep[e.g.][]{herman2015,ugelvig2018,delaney2023}.
 Sediment exhaustion through this process may also explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \citep{overeem2017}.
-Note, however, that the absence of lags between sediment and water discharge may suggest that sediment is mobilized close to the measurement stations, as opposed to distal areas of the glacier bed as in the supply-limited regime \citep[Figure~\ref{fig:model_outs}\, a and e; ][]{williams1989}.
+Note that in the data here, however, the absence of lags between the observed sediment and water discharge may suggest that sediment is mobilized close to the measurement stations, as opposed to distal areas of the glacier bed as would be expected in a supply-limited regime \citep[Figure~\ref{fig:model_outs}\, a and e; ][]{williams1989}.
 
@@
 \item the timing of peak sediment transport capacity below glaciers generally coincides with large rates of increasing in water discharge, as opposed to the maximum quantity of water discharge;
-\item a potentially consistent relationship between water and sediment discharge may represent subglacial sediment access or open channel hydraulics near the glacier margin, instead of subglacial sediment transport capacity.
+\item a consistent relationship between water and sediment discharge may represent subglacial sediment access or open channel hydraulics near the glacier margin, instead of subglacial sediment transport capacity.
 
@@
 Code to make the figures, with links to the data, can be found at \url{https://bitbucket.org/IanDelaney/xsection/src/master/}.
+Code and data will be uploaded to a permanent repository pending acceptance of the paper.
 

### 2023-09-01 Delaneyetal2023b.tex
Latest changes to mauro

@@
 \end{center}
+
 \paragraph{Keypoints}
@@
   Here, parameterizations of sediment transport capacity for pressurized flow under glaciers and subaerial open channel flow are applied to hydrographs from an Alpine glacier and the Greenland Ice Sheet.
-  Results of these parameterizations show that sediment transport capacity varies far more in subglacial channels compared to subaerial ones both over time and with respect to water discharge.
+  Results of these parameterizations show that sediment transport capacity varies far more in subglacial channels than subaerial ones over time and with respect to water discharge.
   Furthermore, high subglacial sediment transport capacity can persist across a wide range of water discharges.
@@
   Despite the different responses of sediment transport capacity to water discharge in the two-channel types, measurements of sediment discharge from the catchments generally correlate better with subaerial hydraulics.
-  This suggests that other processes such as sediment access or subaerial processes in the proglacial area are represented in the sediment discharge records.
+  This suggests that the sediment discharge records represent other processes such as sediment access or subaerial processes in the proglacial area.
   The different responses of water discharge on sediment transport capacity in subglacial channels, compared to subaerial ones, increase the occurrence of sediment transport capacity reaching mobilization thresholds, not necessarily related to water discharge quantity.
@@
 In turn, changes in water discharge $Q_w$ are mainly accommodated with water velocity $v$ \citep[e.g.][]{swift2005}, because the size of the channel $S$ changes on a much shorter timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
-It follows that an increase in water discharge $Q_w$ will cause the shear stress across the channel $\tau_t$ to respond to the water velocity $v$, as opposed to the subaerial channel where channel width $w_c$ evolves as well.
+It follows that an increase in water discharge $Q_w$ will cause the shear stress across the channel $\tau_t$ to respond with water velocity $v$, as opposed to subaerial channels where channel width $w_c$ evolves as well.
 
@@
 These differences have been implicitly included in a wide range of available models that  quantify sediment transport from catchments in both subglacial and subaerial channels  \citep[e.g.][]{walder1994,tucker1997,creyts2013,wickert2019,hewitt2019}.
-The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \citep[e.g.][]{lane2016,perolo2018}.
+The divergent response of changing water discharge on sediment mobilization capacity may well alter sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \citep[e.g.][]{lane2016,perolo2018}.
 Additionally, the variable response of sediment transport capacity to water discharge in the two systems may impact the interpretation of records of sediment transport in glacial systems \citep[e.g.][]{muller1968,richards2003,swift2005,ganti2016}.
@@
 This manuscript aims to evaluate the relationship between sediment transport capacity and water discharge in subglacial channels, compared to subaerial ones.
-Numerical parameterizations of subglacial and subaerial hydraulics are utilized, in the absence of observations of continuous water velocity measurements and the evolution of subglacial and subaerial channels.
+In the absence of continuous measurements of subglacial and subaerial water velocity and channel morphology, numerical parameterizations of subglacial and subaerial hydraulics are utilized.
 Parameterizations are applied to hydrological records from an Alpine glacier in Switzerland (Fieschergletscher), and  a land-terminating glacier in Greenland (Leverett glacier).
@@
 
-The parameterizations above are applied to hydrological records from the Fieschergletscher in the Swiss Alps and the Leverett glacier on the Greenland Ice Sheet.
+The parameterizations above are applied to hydrological records from the Fieschergletscher in the Swiss Alps (\alpine) and the Leverett glacier on the Greenland Ice Sheet (\icesheet).
 The subglacial and subaerial parameterizations represent the subglacial flow of water exiting a glacier and transitioning to subaerial flow, as it moves through the catchment.
 Outputs of the parameterizations are meant to represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
-To generalize these cases, the Fieschergletscher case is referred to as \alpine{} and is exemplified by relatively thin ice thickness ($h_{ice}$: $225$\,\unit{m}), low water discharge ( $<\,20$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
-The Leverett glacier case is referred to as \icesheet{} and is exemplified by thick ice  ($h_{ice}$: $700$\,\unit{m}), high water discharge (generally $<\,500$\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).
+To generalize these cases, \alpine{}  is exemplified by relatively thin ice thickness ($h_{ice}$: $225$\,\unit{m}), low water discharge ($<\,20$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
+ \icesheet{}  is exemplified by thick ice  ($h_{ice}$: $700$\,\unit{m}), high water discharge (generally $<\,500$\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).
 
@@
     Variables have been normalized to mean values for better comparison amongst the runs.
-    $R_g$ ($R_f$) show the Spearman rank correlation values for the subglacial (subaerial) outputs.
-    Plots are shown at $15$\unit{min} interval.
+    $R_g$ ($R_f$) shows the Spearman rank correlation values for the subglacial (subaerial) outputs.
+    Plots are shown with data and outputs at $15$\unit{min} interval.
+    \ian{TOtal vs Width integrated}
   } 
@@
       Text in columns shows rank correlation between the variable and sediment discharge for the corresponding time period for the subglacial (right column) and subaerial (left column) channels.
+
       } 
@@
 
-It is prescient to evaluate the time scales over which the subglacial channels may respond to changing water discharge to understand the impact on variability in sediment transport capacity.
+It is prescient to evaluate the time scales over which the subglacial channels may respond to changing water discharge to understand the influence of variability in sediment transport capacity.
 In both cases, the variability in velocity, shear stress, and width-integrated shear stress decrease after $1-5$ day aggregations (Figure~\ref{fig:multi_run}).
@@
 The consistent relationship between sediment discharge capacity and open channel flow conditions could be pronounced ~ $20$\,\unit{km} downstream of the Leverett measurement site, where a strong correlation persists between sediment plume size and water discharge into the Kangerlussuaq fjord \citep{mcgrath2010}.
-In marine-terminating glacier catchments, a less consistent relationship may persist between water discharge or melt extend and sediment plume size \citep{tedstone2012}.
+In marine-terminating glacier catchments, for contrast, a less consistent relationship may persist between water discharge or melt extend and sediment plume size \citep{tedstone2012}.
 The lack of correlation could be due to the variable relationship between subglacial sediment transport capacity and water discharge at marine-terminating glaciers where subglacial channels are pressurized by ocean water at the ice front \citep[e.g.][]{how2017}.
 
-Second, the sediment discharge records could represent processes such as increased sediment accessibility when melt extends up glacier, thereby concurrently increasing sediment and water discharge \citep[e.g.][]{delaney2020}.
+Second, the sediment discharge records could represent processes such as increased sediment accessibility when melt extends up glacier, thereby concurrently increasing sediment and water discharge \citep[e.g.][]{vergara2022}.
 A supply-limited regime could well persist at many glaciers, because the threshold of motion for sediment can be reached more frequently, and across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
@@
 The variable relationship between water discharge and subglacial sediment transport capacity presented here mostly applies to the daily- to weekly-time scales controlling the size of subglacial channels (Figure~\ref{fig:multi_run}).
-Yet, identifying climatic signals in sediment transport from the transport-limited component of glacier systems may require higher thresholds of water discharge compared to subaerial systems \citep[Figure~\ref{fig:Qw_vari}; ][]{tofelde2021}.
-This occurs as the variable relationship between water discharge and sediment transport capacity
-may compound  the already stochastic nature of erosion and deposition in fluvial environments \citep{castletort2003,jerolmack2010}.
+Yet, identifying climatic signals in sediment transport from the transport-limited component of glacier-sedimentary systems may require higher thresholds of water discharge compared to subaerial systems \citep[Figure~\ref{fig:Qw_vari}; ][]{tofelde2021}.
 
@@
 Code to make the figures, with links to the data, can be found at \url{https://bitbucket.org/IanDelaney/xsection/src/master/}.
-Code and data will be uploaded to a permanent repository pending acceptance of the paper.
+Code and data will be uploaded to a permanent repository, with FAIR principles, pending acceptance of the paper.
+We thank G. King for commenting on a previous version of this manuscript.
 

### 2023-09-01 Delaneyetal2023b.tex
a few line-by-line changes

@@
 \newcommand{\ian}[1]{{\textbf{\color{blue}Ian says:} \color{blue} #1} }
+\newcommand{\mauro}[1]{{\textbf{\color{green}Mauro says:} \color{green} #1} }
 \newcommand{\alpine}{\textit{ALPINE}\,}
@@
   \noindent
-  In both subglacial and subaerial channels, the sediment transport capacity of a sediment size depends mainly on 1) the channel width over which the sediment is mobilized  and 2) water velocity, as this controls the shear stress that the flowing water applies to the channel bottom.
+  In both subglacial and subaerial channels, the sediment transport capacity for sediment size depends mainly on 1) the channel width over which the sediment is mobilized  and 2) water velocity, as this controls the shear stress that the flowing water applies to the channel bottom.
   In subaerial channels, changing water discharge  changes the water depth,  channel width, and velocity.
   However, in subglacial environments and on time scales of hours to weeks, changing water discharge  mainly alters the  water velocity, not channel area or width, because the subglacial conduit maintains a largely fixed geometry.
-  Here, parameterizations of sediment transport capacity for pressurized flow under glaciers and subaerial open channel flow are applied to hydrographs from an Alpine glacier and the Greenland Ice Sheet.
+  Here, parameterizations of sediment transport capacity for pressurized flow under glaciers and for subaerial open channel flow are applied to hydrographs from an Alpine glacier and the Greenland Ice Sheet.
   Results of these parameterizations show that sediment transport capacity varies far more in subglacial channels than subaerial ones over time and with respect to water discharge.
@@
 
-In a transport-limited regime, sediment discharge responds to the sediment transport capacity or the amount of sediment the water can carry.
+In a transport-limited regime, sediment discharge responds to the sediment transport capacity defined as the amount of sediment the water can carry.
 Sediment transport capacity depends on the shear stress between water and sediment it flows  over \citep{shields1936,meyer1948,engelund1967} and the width of the channel bottom $w_c$ over which to mobilize sediment.
@@
 \end{linenomath*}
-% 
-where $Q_w$ is water discharge,  and $S$ is the channel's cross-sectional area.
+\mauro{later $Q$ is used.  drop $_w$?}
+where $Q_w$ is water discharge,  and $S$ is the channel's wetted area. \mauro{I think ``wetted area'' is better and less ambiguous, changed throughout.}
 
-In subaerial channels, with open channel flow, the cross-sectional area $S$ of the channel evolves with changing water discharge $Q_w$  by changing both the channel's width and the water depth \citep{leopold1953}.
-During a flood with greater water discharge $Q_w$, for instance, the cross-sectional area $S$ increases as the channel widens and deepens, along with water velocity $v$, which is squared.
+In subaerial channels, operating with open channel flow, the wetted area $S$ of the channel evolves with changing water discharge $Q_w$, by changing both the channel's width and the water depth \citep{leopold1953}.
+The change in water depth results in a commensurate increase in water velocity.
+Thus higher discharge is accommodated by both an increase in wetted area and flow velocity such that Equation~\eqref{eq:v} remains fulfilled.
 As a result, the shear stress across the channel $\tau_t$ will increase according to Equation~\ref{eq:tau_t}.
 
-Yet, the response of water velocity to changing water discharge in subglacial channels differs from subaerial ones.
+The response of water velocity to changing water discharge in subglacial channels differs from subaerial ones.
 The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by frictional heating of water flowing through the channel \citep{rothlisberger1972}.
-As a result, the subglacial channel size only evolves over days to months, whereas water discharge can vary over hours \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
-In turn, changes in water discharge $Q_w$ are mainly accommodated with water velocity $v$ \citep[e.g.][]{swift2005}, because the size of the channel $S$ changes on a much shorter timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
+As a result, the subglacial channel size only evolves relatively slowly over days to months, whereas water discharge can vary over hours \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
+Therefore, subglacial water flow behaves more like pipe-flow and changes in water discharge $Q_w$ are mainly accommodated with increased water velocity $v$ \citep[e.g.][]{swift2005}, because the size of the channel $S$ changes on a much longer timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
 It follows that an increase in water discharge $Q_w$ will cause the shear stress across the channel $\tau_t$ to respond with water velocity $v$, as opposed to subaerial channels where channel width $w_c$ evolves as well.
@@
 Water velocity magnitude in the subglacial, $v_g$, and subaerial, $v_f$, channels are shown by arrow length.
-$S_g$ and $S_f$ represent the cross-sectional area in subglacial and subaerial channels.
+$S_g$ and $S_f$ represent the wetted area in subglacial and subaerial channels.
 The subglacial channel widths $w_{cg}$ remain unchanged, while the subaerial channel widths evolve $w_{cf}$.
 Shear stress $\tau$ is responsible for the mobilization of sediment and increases with the number of makers at the channel bed.
-      Note that in the subaerial channel parameterization, channels have rectangular shapes (Section~\ref{sect:fluv}).} 
+      Note that in the subaerial channel parameterization, channels have rectangular shapes with their width much larger than the depth (Section~\ref{sect:fluv}).}
     \label{fig:cartoon}
@@
 Furthermore, thorough observations of water velocity and channel width in subglacial channels are limited and, to these authors' knowledge, not contrasted with subaerial conditions at the glacier margins.
+\mauro{self-cite here?}
 
@@
 \label{sect:meth}
-Parameterizations below (Sections ~\ref{sect:sub_mode}~and~\ref{sect:fluv}) represent relationships amongst water discharge, water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
+The parameterizations below (Sections ~\ref{sect:sub_mode}~and~\ref{sect:fluv}) represent relationships amongst water discharge, water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
 Both parameterizations calculate   water velocity, shear stress and width-integrated shear stress, upon which sediment transport depends \citep[Figure \ref{fig:cartoon}; ][]{shields1936}.
 The evaluation of these variables omits the selection of a sediment transport relationship and grain-size parameter \citep[e.g.][]{shields1936,meyer1948} needed to calculate sediment transport capacity from shear stress.
+
+\mauro{Yeah, I wonder whether width-integrated should be scaled by width to 2/3.}
 
@@
   \caption{Variables, parameters, and constants used in this work.}
-  %Where two values are given, the first refers to  \alpine{}, a case from Fieschergletscher, and the second to a glacier marginal to the \icesheet{} a case from Leverett Glacier }
+  %Where two values are given, the first refers to  \alpine{}, a case from Fieschergletscher, and the second to a glacier marginal to the \icesheet{} a case from Leverett Glacier.
+\mauro{those f are high}
   \begin{tabular}{ l  c  c c }
@@
     Water velocity (glacier, subaerial)  & $v_g,\,v_{f}$& & $\mathrm{m\,s^{-1}}$ \\
-    Channel cross-sectional area (glacier, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
+    Channel wetted area (glacier, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
     Hydraulic diameter &$D_h$&&$\mathrm{m}$\\
@@
     \label{eq:dS_dt}
-    \frac{\partial S_g}{\partial t} = C_1 \frac{Q \Delta h}{l} - C_2 (h_{o}-\frac{\Delta h}{2})^n\,S_g,
+    \frac{\partial S_g}{\partial t} = C_1 \frac{Q \Delta h}{l} - C_2 \left(h_{o}-\frac{\Delta h}{2}\right)^n\,S_g,
   \end{equation}
 \end{linenomath*}
-\noindent where $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (Table \ref{table:vpm}), $l$ is the length of subglacial channel, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change from the glacier terminus to its top, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the ice overburden pressure, and $n$ is Glen's n \citep{glen1955}.
+\noindent where $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (Table \ref{table:vpm}), $l$ is the length of subglacial channel, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change from the glacier terminus to its top, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the ice overburden pressure, and $n$ is Glen's n \citep{glen1955} (usually $n=3$).
 The first term on the right side of the equation represents the opening of the channel through frictional heating, while the following term represents the creep closure of the channel from ice deformation.
@@
 \end{linenomath*}
-\noindent where $f_r$ is a friction factor, $D_h$ is the hydraulic diameter, and $l$ is the length overwhich head change occurs.
+\mauro{the $s$ is iffy}
+\noindent where $f_r$ is a friction factor, $D_h$ is the hydraulic diameter, and $l$ is the length over which head change occurs.
 $s$ is a factor accounting for channel geometry \citep{hooke1990}, calculated as:
@@
 
-With knowledge of cross-sectional area $S_g$, the shear stress, $\tau_g$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
+With knowledge of wetted area $S_g$, the shear stress, $\tau_g$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
 \begin{equation}
@@
 To parameterize the shear stress of water flowing across sediments in the subaerial channel,  the hydraulics parameterization presented in \citet{tucker1997} is implemented.
-Here, given  that mass is conserved mass, there is a sufficiently wide channel so that the hydraulic radius is consistent with flow depth, and the Darcy-Weisbach relationship,
-the shear stress $\tau_f$ at the river bed is represented as
+Using mass conservation and the Darcy-Weisbach relationship as well as assuming that the channel is wide so that the hydraulic radius is well approximated by the flow depth, then
+the shear stress $\tau_f$ at the river bed is given by
 \begin{linenomath*}
@@
 where $k$ is a constant and the exponent is commonly given as $\frac{1}{2}$ \citep{leopold1953}.
+\mauro{This is for so called ``regime channels''}
 Water velocity, $v_f$, is given by rearranging Equation \ref{eq:tau} as
@@
 Outputs of the parameterizations are meant to represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
-To generalize these cases, \alpine{}  is exemplified by relatively thin ice thickness ($h_{ice}$: $225$\,\unit{m}), low water discharge ($<\,20$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
- \icesheet{}  is exemplified by thick ice  ($h_{ice}$: $700$\,\unit{m}), high water discharge (generally $<\,500$\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).
+To generalize these cases, \alpine{}  is exemplified by relatively thin ice thickness ($h_{ice}$: $225$\,\unit{m}), low water discharge ($\sim\,10$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
+ \icesheet{}  is exemplified by thick ice  ($h_{ice}$: $700$\,\unit{m}), high water discharge ($\sim\,300$\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).
 
@@
 Rank correlation also reduces the impact uncertainties from sensor disturbances in the sediment transport record from \icesheet{}.
+\mauro{Here a brief explanation of the rank-stuff would be good.  Also describe the time aggregation.}
 
@@
 Following the parameterizations defined in Sections~\ref{sect:sub_mode}~and~\ref{sect:fluv}, the water velocity, shear stress, and width-integrated shear stress exhibit different seasonal evolutions and peaks.
-Water velocity and shear stress represent the potential for sediment mobilization, and width-integrated shear stress represents the total sediment transport capacity across the channel bed (Figure~\ref{fig:model_outs}).
+Water velocity and shear stress represent the potential for sediment mobilization, and width-integrated shear stress represents a proxy for the total sediment transport capacity across the channel bed (Figure~\ref{fig:model_outs}).
 
@@
   \caption{Relationship between water discharge and velocity, shear stress, and width-integrated shear stress for \alpine{} (a-c) and \icesheet (d-f).
-    These outputs' relationship with sediment discharge is given for \alpine in (g-i) and \icesheet (j-l).
+    These outputs' relationship with measured sediment discharge is given for \alpine in (g-i) and \icesheet (j-l).
     Variables have been normalized to mean values for better comparison amongst the runs.
@@
     Plots are shown with data and outputs at $15$\unit{min} interval.
-    \ian{TOtal vs Width integrated}
+    \ian{Total vs Width integrated} \mauro{Why slightly different colours between ALPINE and ICESHEET?}
   }
@@
     \includegraphics[width=0.9\linewidth]{Fig4.png}
-    \caption{Variance (measured by the standard deviation) in water velocity, shear stress, and total shear stress for different time aggregations across the range of subglacial and subaerial channel shapes and friction factors.
+    \caption{Variance (measured by the standard deviation) in water velocity, shear stress, and width integrated shear stress for different time aggregations across the range of subglacial and subaerial channel shapes and friction factors.
       Shaded areas denote the minimum and maximum of the $100$ parameter combinations tested for each of the systems (Section~\ref{sect:imp}).
@@
       Text in columns shows rank correlation between the variable and sediment discharge for the corresponding time period for the subglacial (right column) and subaerial (left column) channels.
-
       }
@@
 \subsection{Relationship between parameterization  outputs and observed sediment discharge}
-When comparing the output variables with suspended sediment discharge data, velocity, shear stress, and width-integrated shear stress from the subaerial system generally show better correlation with sediment discharge compared to subglacial outputs across the tested subglacial channel parameters (Figures~\ref{fig:Qw_vari} g--l and \ref{fig:multi_run}).
+When comparing the output variables with suspended sediment discharge measurements, then velocity, shear stress, and width-integrated shear stress from the subaerial system generally show better correlation with sediment discharge compared to subglacial outputs across the tested subglacial channel parameters (Figures~\ref{fig:Qw_vari} g--l and \ref{fig:multi_run}).
 The rank correlation among the three output variables for the subaerial parameterization remains more or less constant despite the absence of hysteresis in channel shape and consistent scaling with water discharge, as shown above (Figure \ref{fig:Qw_vari} and Section~\ref{sect:fluv}).
@@
 First, these measurement stations are $0.5$ to $2$\,\unit{km} downstream of the glacier terminus \citep{cowton2012,felix2022}, respectively, likely a long enough distance for the pressurized subglacial regime to transition to the open channel subaerial regime.
-The transition of subglacial to subaerial conditions may occur in the transition between pressurized and subaerial flow under glacier terminus \citep{perolo2018}, especially if cavity closure rates of the ice are slow  \citep{egli2021b}.
+The transition of subglacial to subaerial conditions may occur in the transition between pressurized and subaerial flow under glacier terminus \citep{perolo2018}, especially if closure rates of the ice channels are slow  \citep{egli2021b} \mauro{this would not be the citation I use here.
+Maybe Röthlisberge1972?}.
 Furthermore,  if a transport-limited regime persists at the glacier margin, then sediment can be easily mobilized and subaerial processes could be represented just a short distance downstream.
 
-The consistent relationship between sediment discharge capacity and open channel flow conditions could be pronounced ~ $20$\,\unit{km} downstream of the Leverett measurement site, where a strong correlation persists between sediment plume size and water discharge into the Kangerlussuaq fjord \citep{mcgrath2010}.
-In marine-terminating glacier catchments, for contrast, a less consistent relationship may persist between water discharge or melt extend and sediment plume size \citep{tedstone2012}.
+The consistent relationship between sediment discharge capacity and open channel flow conditions could be pronounced $\sim20$\,\unit{km} downstream of the Leverett measurement site, where a strong correlation persists between sediment plume size and water discharge into the Kangerlussuaq fjord \citep{mcgrath2010}.
+In marine-terminating glacier catchments, for contrast, a less consistent relationship may persist between water discharge or melt extent and sediment plume size \citep{tedstone2012}.
 The lack of correlation could be due to the variable relationship between subglacial sediment transport capacity and water discharge at marine-terminating glaciers where subglacial channels are pressurized by ocean water at the ice front \citep[e.g.][]{how2017}.
@@
 Regardless of the increased performance of water discharge compared to subglacial sediment transport capacity in the observations, parameterization outputs suggest that additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to subaerial systems \citep{jansson2005,ganti2016}.
-Records of sediment discharge have been used to establish the relationship, or lack thereof, between sediment discharge and climate in glacial systems \citep[e.g.][]{koppes2009a,willenbring2016,mariotti2021}.
+Records of sediment discharge have been used to establish the relationship, or lack thereof, with climate-warming induced glacial changes \citep[e.g.][]{koppes2009a,willenbring2016,mariotti2021}.
 The variable relationship between water discharge and subglacial sediment transport capacity presented here mostly applies to the daily- to weekly-time scales controlling the size of subglacial channels (Figure~\ref{fig:multi_run}).
@@
 
-Sediment transport capacity is largely driven by the velocity of water flowing through a channel and the channel's width.
+Sediment transport capacity is largely driven by the shear stress of water flowing through a channel and the channel's width.
 Subaerial channels  alter their channel width and water velocity immediately in response to changing water discharge.

### 2023-09-06 Delaneyetal2023b.tex
starting to mess with D-W

@@
   Andrew J. Tedstone\footnote{Department of Geosciences, University of Fribourg, Ch. du Musée 1700, Fribourg, Switzerland},
-  Daniel Farinotti\footnote{Laboratory of Hydraulics, Hydrology and Glaciology (VAW), ETH-Z\"urich, H\"onggerbergring 26, 8093 89 Z\"urich, Switzerland}$^{,}$\footnote{Swiss Federal Institute for Forest, Snow and Landscape Research (WSL) Z\"uricherstrasse 111, 8903 1011 Birmensdorf, Switzerland}
+  Daniel Farinotti\footnote{Laboratory of Hydraulics, Hydrology and Glaciology (VAW), ETH-Z\"urich, H\"onggerbergring 26, 8093 89 Z\"urich, Switzerland}$^{,}$\footnote{Swiss Federal Institute for Forest, Snow and Landscape Research (WSL) Z\"uricherstrasse 111, 8903 1011 Birmensdorf, Switzerland},
+  Mauro A. Werder$^{3,4}$
 
@@
 \end{itemize}
+
+\ian{TODO!!! 2019 paper.... and scale $\tau$ with 3 div 2 as in MPM???... friction factor f}
 
@@
     \label{eq:v}
-    v = \frac{Q_w}{S},
+    v = \frac{Q}{S},
   \end{equation}
 \end{linenomath*}
-\mauro{later $Q$ is used.  drop $_w$?}
-where $Q_w$ is water discharge,  and $S$ is the channel's wetted area. \mauro{I think ``wetted area'' is better and less ambiguous, changed throughout.}
+where $Q$ is water discharge,  and $S$ is the channel's wetted area. %\mauro{I think ``wetted area'' is better and less ambiguous, changed throughout.}
+% \mauro{later $Q$ is used.  drop $_w$?}
 
-In subaerial channels, operating with open channel flow, the wetted area $S$ of the channel evolves with changing water discharge $Q_w$, by changing both the channel's width and the water depth \citep{leopold1953}.
+In subaerial channels, operating with open channel flow, the wetted area $S$ of the channel evolves with changing water discharge $Q$, by changing both the channel's width and the water depth \citep{leopold1953}.
 The change in water depth results in a commensurate increase in water velocity.
@@
 As a result, the subglacial channel size only evolves relatively slowly over days to months, whereas water discharge can vary over hours \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
-Therefore, subglacial water flow behaves more like pipe-flow and changes in water discharge $Q_w$ are mainly accommodated with increased water velocity $v$ \citep[e.g.][]{swift2005}, because the size of the channel $S$ changes on a much longer timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
-It follows that an increase in water discharge $Q_w$ will cause the shear stress across the channel $\tau_t$ to respond with water velocity $v$, as opposed to subaerial channels where channel width $w_c$ evolves as well.
+Therefore, subglacial water flow behaves more like pipe-flow and changes in water discharge $Q$ are mainly accommodated with increased water velocity $v$ \citep[e.g.][]{swift2005}, because the size of the channel $S$ changes on a much longer timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
+It follows that an increase in water discharge $Q$ will cause the shear stress across the channel $\tau_t$ to respond with water velocity $v$, as opposed to subaerial channels where channel width $w_c$ evolves as well.
 
@@
 Furthermore, thorough observations of water velocity and channel width in subglacial channels are limited and, to these authors' knowledge, not contrasted with subaerial conditions at the glacier margins.
-\mauro{self-cite here?}
+%\mauro{self-cite here?}
 
@@
     \textbf{Variables}  & & & \\
-    Water discharge  & $Q_w$& & $\mathrm{m^{3}\,s^{-1}}$ \\
+    Water discharge  & $Q$& & $\mathrm{m^{3}\,s^{-1}}$ \\
     Water velocity (glacier, subaerial)  & $v_g,\,v_{f}$& & $\mathrm{m\,s^{-1}}$ \\
@@
 \end{linenomath*}
-\mauro{the $s$ is iffy}
+\mauro{the $s$ is iffy}\ian{... not sure of the exact problem here... I have given the relationship between $D_h$ and $S_g$ below}
 \noindent where $f_r$ is a friction factor, $D_h$ is the hydraulic diameter, and $l$ is the length over which head change occurs.
@@
 smaller values of $\beta$ result in shallow, wide channels.
+The hydraulic diameter $D_h$ is converted to wetted area $S_g$ given
+\begin{equation}
+  \label{eq:Dh2S}
+  S_g= \frac{D_h^2}{2}\,\frac{(\frac{\beta}{2}+\sin \frac{\beta}{2})^2}{\beta - \sin \beta}.
+\end{equation}
 
@@
 %
-where  the water velocity $v_g = \frac{Q_w}{S_g}$.
+where  the water velocity $v_g = \frac{Q}{S_g}$.
 
@@
     \label{eq:DW_tau}
-    \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q_w}{w_{cf}} \Big)^{\frac{2}{3}} \,\nabla z_c^{\frac{2}{3}},
+    \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q}{w_{cf}} \Big)^{\frac{2}{3}} \,\nabla z_c^{\frac{2}{3}},
   \end{equation}
@@
   \label{eq:wcf}
-  w_{cf} = k \, Q_w^\frac{1}{2},
+  w_{cf} = k \, Q^\frac{1}{2},
 \end{equation}
@@
 where $k$ is a constant and the exponent is commonly given as $\frac{1}{2}$ \citep{leopold1953}.
-\mauro{This is for so called ``regime channels''}
+%\mauro{This is for so called ``regime channels''}
 Water velocity, $v_f$, is given by rearranging Equation \ref{eq:tau} as
@@
 
-The parameterizations above are applied to hydrological records from the Fieschergletscher in the Swiss Alps (\alpine) and the Leverett glacier on the Greenland Ice Sheet (\icesheet).
+The parameterizations above are applied to hydrological records from the Fieschergletscher (\alpine) and the Leverett glacier (\icesheet).
 The subglacial and subaerial parameterizations represent the subglacial flow of water exiting a glacier and transitioning to subaerial flow, as it moves through the catchment.
@@
 
-Lastly,  outputs are compared with the record of sediment discharge from the two glaciers using Spearman rank correlation to reduce the impact of the non-linear response of sediment discharge to the hydrological forcing.
-Rank correlation also reduces the impact uncertainties from sensor disturbances in the sediment transport record from \icesheet{}.
-\mauro{Here a brief explanation of the rank-stuff would be good.  Also describe the time aggregation.}
+Lastly,  outputs are compared with the record of sediment discharge from the two glaciers using Spearman rank correlation, or the correlation with respect to ordering of values from the outputs and observations.
+Rank correlation reduces the impact of the non-linear response of sediment discharge to the hydrological forcing and the uncertainties from sensor disturbances in the sediment transport record from \icesheet{}.
+Outputs and observations are aggregated across timescales from $1$\,\unit{hr} to $15$\,\unit{days} from the $15$\,\unit{min} outputs, by averaging values one half of the timescale before and after a point time.
+%\mauro{Here a brief explanation of the rank-stuff would be good.  Also describe the time aggregation.}
 
@@
     Note that the x and y axes are different for \alpine{} and \icesheet{} such that they each cover a summer.
-    %    \ian{Andrew's comment: for a given channel width, subglacial channels are more effective at mobilizing and transporting sediment than subaerial ones}
+        \ian{Andrew's comment: for a given channel width, subglacial channels are more effective at mobilizing and transporting sediment than subaerial ones}
+        \ian{also, might need to go back in Timemaching to pull out old version}
       }
@@
     Plots are shown with data and outputs at $15$\unit{min} interval.
-    \ian{Total vs Width integrated} \mauro{Why slightly different colours between ALPINE and ICESHEET?}
+    \mauro{Why slightly different colours between ALPINE and ICESHEET?}
   }
@@
 In both cases, the exponent greater than $1$ in the sediment transport relationship magnifies the variability in sediment discharge beyond the highly variable shear stress and width-integrated shear stress described above (Figure~\ref{fig:multi_run}).
+Note that sediment transport relationships describe sediment flux per unit width so that the evaluation across channel occurs after application of the exponent, unlike the width-integrated shear stress evaluation here \citep[c.f.][]{tucker1997}.
+Note also that both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress is homogeneous across the channel bed, despite the limitations of this assumption \citep[e.g.][]{yager2018}.
+
 The rapid increase and decreases in subglacial sediment transport capacity may, for example, compound the variability in sediment transport already present in subaerial systems \citep[Figure~\ref{fig:Qw_vari}; ][]{williams1989,jerolmack2010}.
-Large variations in sediment transport capacity underneath glaciers may result in sediment mobilization and deposition processes in close spatial and temporal proximity to each other \citep{gimbert2016,perolo2018}.
-Note also that both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress is homogeneous across the channel bed, despite the limitations of this assumption \citep[e.g.][]{yager2018}.
+For example, by applying the discharge - width scaling in Equation~\ref{eq:wcf} to Equation \ref{eq:tau_t} and to the $S \propto \,Q^{\frac{1}{2}}$ in Equation \ref{eq:v}, the response of the subaerial channel's width-integrated shear stress to changing water discharge is $\frac{\partial \tau_{tf}}{\partial t} \propto \frac{\partial Q^{\frac{3}{2}}}{\partial t}$.
+For fixed channel width and wetter area over sufficiently short time periods in the subglacial system, the response of width-integrated shear stress is $\frac{\partial \tau_{tg}}{\partial t} \propto \frac{\partial Q^{2}}{\partial t}$, with a larger exponent on $Q$.
 
@@
 
-Outputs from the subaerial system rank correlate higher with sediment discharge in the \alpine and \icesheet systems here.
-The increased performance occurs despite  output variables showing the disparate relationship between sediment transport capacity and water discharge in subglacial systems (Figure~\ref{fig:Qw_vari}), which could occur for two reasons.
+The increased performance of subaerial hydraulics, in representing observations occurs despite  output variables showing the disparate relationship between sediment transport capacity and water discharge in subglacial systems (Figure~\ref{fig:Qw_vari}).
+This could occur for two reasons.
 
-First, these measurement stations are $0.5$ to $2$\,\unit{km} downstream of the glacier terminus \citep{cowton2012,felix2022}, respectively, likely a long enough distance for the pressurized subglacial regime to transition to the open channel subaerial regime.
-The transition of subglacial to subaerial conditions may occur in the transition between pressurized and subaerial flow under glacier terminus \citep{perolo2018}, especially if closure rates of the ice channels are slow  \citep{egli2021b} \mauro{this would not be the citation I use here.
-Maybe Röthlisberge1972?}.
+First, these measurement stations are $0.5$ to $2$\,\unit{km} downstream of the glacier terminus \citep{cowton2012,felix2022}, likely a long enough distance for the pressurized subglacial regime to transition to the open channel subaerial regime.
+The transition of subglacial to subaerial conditions may occur in the transition between pressurized and subaerial flow under glacier terminus \citep{perolo2018}, especially if closure rates of the ice channels are slow  \citep{rothlisberger1972}.
 Furthermore,  if a transport-limited regime persists at the glacier margin, then sediment can be easily mobilized and subaerial processes could be represented just a short distance downstream.
@@
 The consistent relationship between sediment discharge capacity and open channel flow conditions could be pronounced $\sim20$\,\unit{km} downstream of the Leverett measurement site, where a strong correlation persists between sediment plume size and water discharge into the Kangerlussuaq fjord \citep{mcgrath2010}.
-In marine-terminating glacier catchments, for contrast, a less consistent relationship may persist between water discharge or melt extent and sediment plume size \citep{tedstone2012}.
+In marine-terminating glacier catchments, for contrast, a less consistent relationship may occur between water discharge or melt extent and sediment plume size \citep{tedstone2012}.
 The lack of correlation could be due to the variable relationship between subglacial sediment transport capacity and water discharge at marine-terminating glaciers where subglacial channels are pressurized by ocean water at the ice front \citep[e.g.][]{how2017}.
@@
 
-Regardless of the increased performance of water discharge compared to subglacial sediment transport capacity in the observations, parameterization outputs suggest that additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to subaerial systems \citep{jansson2005,ganti2016}.
-Records of sediment discharge have been used to establish the relationship, or lack thereof, with climate-warming induced glacial changes \citep[e.g.][]{koppes2009a,willenbring2016,mariotti2021}.
+%Regardless of the increased performance of water discharge compared to subglacial sediment transport capacity in the observations, parameterization outputs suggest that additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to subaerial systems \citep{jansson2005,ganti2016}.
+
+Records of sediment discharge have been used to establish the relationship, or lack thereof, with climate forcing and glacial erosion \citep[e.g.][]{koppes2009a,willenbring2016,mariotti2021}.
 The variable relationship between water discharge and subglacial sediment transport capacity presented here mostly applies to the daily- to weekly-time scales controlling the size of subglacial channels (Figure~\ref{fig:multi_run}).

### 2023-09-12 Delaneyetal2023b.tex
most edits are included

@@
   Andrew J. Tedstone\footnote{Department of Geosciences, University of Fribourg, Ch. du Musée 1700, Fribourg, Switzerland},
-  Daniel Farinotti\footnote{Laboratory of Hydraulics, Hydrology and Glaciology (VAW), ETH-Z\"urich, H\"onggerbergring 26, 8093 89 Z\"urich, Switzerland}$^{,}$\footnote{Swiss Federal Institute for Forest, Snow and Landscape Research (WSL) Z\"uricherstrasse 111, 8903 1011 Birmensdorf, Switzerland},
-  Mauro A. Werder$^{3,4}$
+   Mauro A. Werder$^{3,4}$,
+  Daniel Farinotti\footnote{Laboratory of Hydraulics, Hydrology and Glaciology (VAW), ETH-Z\"urich, H\"onggerbergring 26, 8093 89 Z\"urich, Switzerland}$^{,}$\footnote{Swiss Federal Institute for Forest, Snow and Landscape Research (WSL) Z\"uricherstrasse 111, 8903 1011 Birmensdorf, Switzerland}
 
@@
 
-\ian{TODO!!! 2019 paper.... and scale $\tau$ with 3 div 2 as in MPM???... friction factor f}
-
 \begin{abstract}
@@
   Furthermore, high subglacial sediment transport capacity can persist across a wide range of water discharges.
-  However, changing channel width can accommodate some  variability in sediment transport capacity in subaerial channels.
+  However, slowly evolving  channel width can accommodate some  variability in sediment transport capacity in subglacial channels.
   Despite the different responses of sediment transport capacity to water discharge in the two-channel types, measurements of sediment discharge from the catchments generally correlate better with subaerial hydraulics.
@@
 %
-where $w_c$ is the width of the channel.
-Following the conservation of mass, the velocity of the water flowing through a  channel is given as
+where $w_c$ is the width of the channel. Following mass conservation, the velocity of the water flowing through a  channel is 
 \begin{linenomath*}
@@
 The change in water depth results in a commensurate increase in water velocity.
-Thus higher discharge is accommodated by both an increase in wetted area and flow velocity such that Equation~\eqref{eq:v} remains fulfilled.
-As a result, the shear stress across the channel $\tau_t$ will increase according to Equation~\ref{eq:tau_t}.
+Thus,  fulfilling Equation~\eqref{eq:v}, the higher discharge is accommodated by both an increase in wetted area and flow velocity.
+As a result, the shear stress across the channel $\tau_t$ increases according to Equation~\ref{eq:tau_t}.
 
@@
 These differences have been implicitly included in a wide range of available models that  quantify sediment transport from catchments in both subglacial and subaerial channels  \citep[e.g.][]{walder1994,tucker1997,creyts2013,wickert2019,hewitt2019}.
-The divergent response of changing water discharge on sediment mobilization capacity may well alter sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \citep[e.g.][]{lane2016,perolo2018}.
-Additionally, the variable response of sediment transport capacity to water discharge in the two systems may impact the interpretation of records of sediment transport in glacial systems \citep[e.g.][]{muller1968,richards2003,swift2005,ganti2016}.
+The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \citep[e.g.][]{lane2016,perolo2018}.
+Additionally, the variable response of sediment transport capacity to water discharge in the two systems may affect the interpretation of sediment transport records in glacial systems \citep[e.g.][]{muller1968,richards2003,swift2005,ganti2016}.
 Yet, the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial channels, and implications for sediment discharge records, has been minimally discussed.
@@
 In the absence of continuous measurements of subglacial and subaerial water velocity and channel morphology, numerical parameterizations of subglacial and subaerial hydraulics are utilized.
-Parameterizations are applied to hydrological records from an Alpine glacier in Switzerland (Fieschergletscher), and  a land-terminating glacier in Greenland (Leverett glacier).
-The parameterizations' lumped nature is intended to isolate the disparity between water discharge and sediment transport capacity in subglacial systems, independent of the upstream drainage network and sediment access.
+Parameterizations are applied to hydrological records from an Alpine glacier in Switzerland (Fieschergletscher) and  a land-terminating glacier in Greenland (Leverett glacier).
+The parameterizations' lumped nature isolates the disparity between water discharge and sediment transport capacity in subglacial systems, independent of the upstream drainage network and sediment access.
 Outputs demonstrate differences in the relationships between  water discharge, channel geometry, and water velocity in the subglacial and subaerial channels.
@@
 \label{sect:ss_data}
-
-% \TODO{Address Daniel's comment: I think that these records need a few more words: where are they measured exactly?
-How long are the time series?
-What temporal resolution have they got?
-Do they cover the entire year, or only the summer months?
-}
 
@@
 
-\mauro{Yeah, I wonder whether width-integrated should be scaled by width to 2/3.}
-
+\ian{values of f are still high}
 \begin{table}[ht]
   \centering
-  \caption{Variables, parameters, and constants used in this work.}
-  %Where two values are given, the first refers to  \alpine{}, a case from Fieschergletscher, and the second to a glacier marginal to the \icesheet{} a case from Leverett Glacier.
-\mauro{those f are high}
+  \caption{Variables, parameters, and constants used in this work. 
+    Where two values are given, the first refers to  \alpine{}, a case from Fieschergletscher, and the second to a glacier marginal to the \icesheet{} a case from Leverett Glacier}
   \begin{tabular}{ l  c  c c }
@@
 
-    Glacier thickness &$h_{ice}$& $225^a$ or $700^b$  &\unit{m}\\
+    Glacier thickness &$h_{ice}$& $225$ or $700$  &\unit{m}\\
     Effective glacier thickness &$h_o$&$\frac{\rho_i}{\rho_w} h_{ice}$  &\unit{m}\\
-    Effective glacier length &$l$&$7000^a$ or $20,000^b$&\unit{m}\\
+    Effective glacier length &$l$&$7000$ or $26,000$&\unit{m}\\
     Constant $1$ in Equation~\ref{eq:dS_dt} &$C_1$&$2.2\times10^{-5}$&\unit{m}$^{-1}$\\
@@
     Ice flow exponent &$n$& $3$ &$\mathrm{(-)}$\\
-    Friction factor (glacier) & $f_r$ & $4$ & $\mathrm{(-)}$ \\
+    Friction factor (subglacial) & $f_r$ &$10$ or $20$ & $\mathrm{(-)}$ \\
     Friction factor (subaerial) & $f_p$ & $3$ & $\mathrm{(-)}$\\
     Gradient of channel bed (subaerial) &$\nabla z_c$ &$0.05$& $\mathrm{(-)}$\\
-    Subaerial channel factor & $k$ &$3$ & $\mathrm{s\,m^{-2}}$\\
+    Subaerial channel factor & $k$ &$3$ or $6.5$ & $\mathrm{s\,m^{-2}}$\\
     Channel geometry exponent &$e$& $\frac{1}{2}$&$\mathrm{(-)}$ \\
@@
   \label{table:vpm}
-
-  % $^a$ for \alpine{} (Fieschergletscher)
-
-  % $^b$ for \icesheet{} (Leverett glacier)
 
@@
 The formulation here does not consider the englacial storage of water.
-Thus, it is assumed that the changing hydraulic head at the top of the glacier is negligible in the evolution of the channel size.
+Thus, it is assumed that the changing hydraulic head at the top of the glacier is negligible in the channel size evolution.
 The evolution of subglacial channel size $S_g$ is given as
@@
 \end{linenomath*}
-\mauro{the $s$ is iffy}\ian{... not sure of the exact problem here... I have given the relationship between $D_h$ and $S_g$ below}
-\noindent where $f_r$ is a friction factor, $D_h$ is the hydraulic diameter, and $l$ is the length over which head change occurs.
+\noindent where $f_r$ is a friction factor, $D_h$ is the hydraulic diameter, and $l$ is the length overwhich head change occurs.
 $s$ is a factor accounting for channel geometry \citep{hooke1990}, calculated as:
@@
 smaller values of $\beta$ result in shallow, wide channels.
-The hydraulic diameter $D_h$ is converted to wetted area $S_g$ given
+The hydraulic diameter $D_h$ is converted to wetted area $S_g$ given a channel geometry following \citep{hooke1990}
 \begin{equation}
   \label{eq:Dh2S}
-  S_g= \frac{D_h^2}{2}\,\frac{(\frac{\beta}{2}+\sin \frac{\beta}{2})^2}{\beta - \sin \beta}.
+  S_g= \frac{D_h^2}{2}\,\frac{(\frac{\beta}{2}+\sin \frac{\beta}{2})^2}{\beta - \sin \beta},
 \end{equation}
+where $\beta$ is the central angle of the circular segment that comprises the channel (the so-called Hooke angle). Note that $\beta =\pi$ corresponds to a semi-circle and
+smaller values of $\beta$ result in shallow, wide channels.
 
@@
 Using mass conservation and the Darcy-Weisbach relationship as well as assuming that the channel is wide so that the hydraulic radius is well approximated by the flow depth, then
-the shear stress $\tau_f$ at the river bed is given by
+the shear stress $\tau_f$ at the river bed is
 \begin{linenomath*}
@@
 To generalize these cases, \alpine{}  is exemplified by relatively thin ice thickness ($h_{ice}$: $225$\,\unit{m}), low water discharge ($\sim\,10$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
- \icesheet{}  is exemplified by thick ice  ($h_{ice}$: $700$\,\unit{m}), high water discharge ($\sim\,300$\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).
+\icesheet{}  is exemplified by thick ice  ($h_{ice}$: $700$\,\unit{m}), high water discharge ($\sim\,300$\,\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).
 
@@
 
-Lastly,  outputs are compared with the record of sediment discharge from the two glaciers using Spearman rank correlation, or the correlation with respect to ordering of values from the outputs and observations.
-Rank correlation reduces the impact of the non-linear response of sediment discharge to the hydrological forcing and the uncertainties from sensor disturbances in the sediment transport record from \icesheet{}.
+Lastly,  outputs are compared with the sediment discharge records the glaciers using Spearman rank correlation, or the correlation with respect to the ordering of values from the outputs and observations.
+Rank correlation reduces the impact of the non-linear response of sediment transport capacity to the hydrological forcing.
+Periods of uncertainty in the data in \icesheet are removed from the analysis.
 Outputs and observations are aggregated across timescales from $1$\,\unit{hr} to $15$\,\unit{days} from the $15$\,\unit{min} outputs, by averaging values one half of the timescale before and after a point time.
@@
     Insets in e-h show the  peak melt event \icesheet{} denoted by the shaded area.
-    Note that the x and y axes are different for \alpine{} and \icesheet{} such that they each cover a summer.
-        \ian{Andrew's comment: for a given channel width, subglacial channels are more effective at mobilizing and transporting sediment than subaerial ones}
-        \ian{also, might need to go back in Timemaching to pull out old version}
+    Note that the x and y axes are different for \alpine{} and \icesheet{} such that they each cover a 
+        \ian{NaN values will be removed in final version}
       }
@@
 Over the course of the season, large peaks in water velocity and shear stress, along with width-integrated shear stress, occur in response to increases in water discharge in both \alpine{} and \icesheet{}.
+This occurs as the subglacial values are generally larger than the subaerial values, thus more efficient at transporting sediment.
 In the subglacial channel, these peaks generally occur over the period when water discharge increases at the fastest rate (Figure~\ref{fig:model_outs}).
@@
     Plots are shown with data and outputs at $15$\unit{min} interval.
-    \mauro{Why slightly different colours between ALPINE and ICESHEET?}
   }
@@
 
-Each water discharge value in the subaerial channel is associated with a single water velocity, shear stress, and width-integrated shear stress, resulting in a perfect rank correlation between water discharge and the output (Figure~\ref{fig:Qw_vari} a--f).
+Each water discharge in the subaerial channel is associated with a single water velocity, shear stress, and width-integrated shear stress, resulting in a perfect rank correlation between the variables (Figure~\ref{fig:Qw_vari} a--f).
 Conversely in the subglacial channel,  water velocity, shear stress, and width-integrated shear stress can vary substantially for a given water discharge and width-integrated shear stress varies with respect to the other two  outputs as well (Figure~\ref{fig:Qw_vari} a--f).
@@
       Solid lines denote  mean values.
+      Top row shows results from \alpine{} case, while bottom row show outputs from \icesheet{} case
       Time intervals shown with markers denote $15$\,\unit{min}, $1$\,\unit{hr}, $12$\,\unit{hr}, $1$\,\unit{d}, $5$\,\unit{d}, $10$\,\unit{d}, and $15$\,\unit{d} aggregation periods.
@@
 \subsection{Relationship between parameterization  outputs and observed sediment discharge}
-When comparing the output variables with suspended sediment discharge measurements, then velocity, shear stress, and width-integrated shear stress from the subaerial system generally show better correlation with sediment discharge compared to subglacial outputs across the tested subglacial channel parameters (Figures~\ref{fig:Qw_vari} g--l and \ref{fig:multi_run}).
-The rank correlation among the three output variables for the subaerial parameterization remains more or less constant despite the absence of hysteresis in channel shape and consistent scaling with water discharge, as shown above (Figure \ref{fig:Qw_vari} and Section~\ref{sect:fluv}).
-In the subglacial system, velocity and shear stress perform substantially worse in terms of rank correlation with sediment discharge compared to width-integrated shear stress (Figures~ \ref{fig:multi_run}).
-In both \alpine{} and \icesheet{}, the rank correlation of the subglacial system approaches that of the subaerial system with longer time aggregation periods (Figures~ \ref{fig:multi_run}).
+When comparing the output variables with suspended sediment discharge measurements, velocity, shear stress, and width-integrated shear stress from the subaerial system generally show better correlation with sediment discharge compared to subglacial outputs across the tested subglacial channel parameters (Figures~\ref{fig:Qw_vari} g--l and \ref{fig:multi_run}).
+The rank correlation amongst the  subaerial parameterization's three output variables remains fairly constant despite the variable impact of channel shape and consistent scaling with water discharge (Figure \ref{fig:Qw_vari} and Section~\ref{sect:fluv}).
+In subglacial channels, velocity and shear stress perform substantially worse in terms of rank correlation with sediment discharge compared to width-integrated shear stress (Figures~ \ref{fig:multi_run}).
+In both \alpine{} and \icesheet{}, the  subglacial rank correlation approaches that of subaerial channels over longer time aggregation periods (Figures~ \ref{fig:multi_run}).
 
 \section{Discussion}
-\subsection{Context of sediment mobilization differences}
+\subsection{Context of sediment capacity differences between the channel types}
 Quantities that control sediment transport capacity, such as water velocity, shear stress, and width-integrated shear stress, vary differently in response to water discharge in subglacial channels compared to subaerial ones (Figure~\ref{fig:Qw_vari}).
-In sediment transport relationships, such as in \citet{meyer1948}, that evaluate sediment discharge, shear stress is scaled to the power of $\frac{3}{2}$ and relies on shear stress exceeding a threshold.
-In the total sediment transport relationship of \citet{engelund1967}, the shear stress is scaled to $\frac{5}{2}$ power.
-In both cases, the exponent greater than $1$ in the sediment transport relationship magnifies the variability in sediment discharge beyond the highly variable shear stress and width-integrated shear stress described above (Figure~\ref{fig:multi_run}).
-Note that sediment transport relationships describe sediment flux per unit width so that the evaluation across channel occurs after application of the exponent, unlike the width-integrated shear stress evaluation here \citep[c.f.][]{tucker1997}.
-Note also that both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress is homogeneous across the channel bed, despite the limitations of this assumption \citep[e.g.][]{yager2018}.
+In sediment transport relationships, such as in \citet{meyer1948} or \citet{engelund1967}, that evaluate sediment discharge, shear stress is scaled to the $\frac{3}{2}$ or $\frac{5}{2}$ power, respectively.
+The shear stress exponent greater than $1$ magnifies sediment discharge variability beyond the highly variable sediment transport parameters described above (Figure~\ref{fig:multi_run}).
+Note that sediment transport relationships describe sediment flux per unit width so that the integrated capacity across channel occurs after application of the exponent, unlike the width-integrated shear stress evaluation here \citep[c.f.][]{tucker1997}.
+%Note also that both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress is homogeneous across the channel bed, despite the limitations of this assumption \citep[e.g.][]{yager2018}.
 
-The rapid increase and decreases in subglacial sediment transport capacity may, for example, compound the variability in sediment transport already present in subaerial systems \citep[Figure~\ref{fig:Qw_vari}; ][]{williams1989,jerolmack2010}.
-For example, by applying the discharge - width scaling in Equation~\ref{eq:wcf} to Equation \ref{eq:tau_t} and to the $S \propto \,Q^{\frac{1}{2}}$ in Equation \ref{eq:v}, the response of the subaerial channel's width-integrated shear stress to changing water discharge is $\frac{\partial \tau_{tf}}{\partial t} \propto \frac{\partial Q^{\frac{3}{2}}}{\partial t}$.
-For fixed channel width and wetter area over sufficiently short time periods in the subglacial system, the response of width-integrated shear stress is $\frac{\partial \tau_{tg}}{\partial t} \propto \frac{\partial Q^{2}}{\partial t}$, with a larger exponent on $Q$.
+The rapid increases and decreases in subglacial sediment transport capacity may compound the variability in sediment transport already present in subaerial systems \citep[Figure~\ref{fig:Qw_vari}; ][]{williams1989,jerolmack2010}.
+By applying the discharge - width scaling in Equation~\ref{eq:wcf} to Equation~\ref{eq:v}, by setting $S\, \propto \,Q^{\frac{1}{2}}$, the subaerial channel's width-integrated shear stress response is $\tau_{tf}\, \propto\,  Q$.
+For sufficiently short time periods,  channel width and wetted area are fixed in the subglacial system so that the response of width-integrated shear stress is $ \tau_{tg}\, \propto\,  Q^{2}$, with a larger exponent on $Q$ than in the subaerial case.
+As a result, when water discharge fluctuates at shorter timescale than the subglacial channel evolution \citep{gimbert2016}, the width-integrated shear stress varies more with water discharge than in subaerial channels.
+Note also that the subglacial shear-stress response to water discharge will be greater than the subaerial's provided that the exponent is less than one in Equation~\ref{eq:wcf}.
+However, when subglacial channel size lies in equilibrium with water discharge, the  width-integrated shear stress response is $\tau_{tg}\, \propto \, Q^{\frac{4}{5}}$, similar to subaerial channels.
+Hydraulics could trend toward this equilibrium regime after $1-5$ days, when variability in subglacial channels begins to approach that of subaerial channels (Figure~\ref{fig:multi_run}). 
+\ian{mauro: should a supplement be made for this? if so, could you send me the one for R-channels? is this essentially equation 14 in werder et al., 2010}
 
-% In the subaerial parameterization, this means that sediment is transported across all channel widths and water discharges, as opposed to large channel widths that occur during floods when much sediment transport occurs \citep{wolman1960}.
-
-In subglacial channels, peaks occur when water discharge increases at the fastest rate compared with the growth of the channel (Figure~\ref{fig:multi_run}).
-In turn, the parameterization of subglacial channels here contains feedbacks that limits sediment transport capacity as the subglacial channel grows with increasing water discharge capacity, effectively reducing water velocity.
-Similar feedback occurs in subaerial systems \citep{phillips2016}, albeit at longer timescales of centuries or more, as opposed to days here.
-
-%In large catchments and further downstream of proglacial margins, water discharge is generally less variable than at catchments' headwaters, where glaciers lie \citep[][]{swift2005,riihimaki2005,costa2017,vanas2017}.
-%Here, reduced hydrological variability far  downstream of glaciers could result in smaller variations in sediment transport capacity in many subaerial systems compared to subglacial ones.
-%Smaller variations here would occur as a result of the hydrological forcing, as opposed to slowly evolving subglacial channel size.
+In subglacial channels, peaks occur with the fastest rate of  water discharge increase relative to the channel's growth (Figure~\ref{fig:model_outs}).
+In turn, the subglacial channels here contain feedbacks that limit sediment transport capacity as the subglacial channel grows with increasing water discharge capacity, effectively reducing sediment transport capacity with the stabilization of water discharge.
+Similar feedback occurs in subaerial systems \citep{phillips2016}, albeit at longer timescales of centuries or more, as opposed to days here \citep{gimbert2016}.
 
@@
 
-The increased performance of subaerial hydraulics, in representing observations occurs despite  output variables showing the disparate relationship between sediment transport capacity and water discharge in subglacial systems (Figure~\ref{fig:Qw_vari}).
+The increased performance of subaerial hydraulics in representing observations occurs despite the proximity of glaciers to these measurements (Figure~\ref{fig:Qw_vari}).
 This could occur for two reasons.
 
-First, these measurement stations are $0.5$ to $2$\,\unit{km} downstream of the glacier terminus \citep{cowton2012,felix2022}, likely a long enough distance for the pressurized subglacial regime to transition to the open channel subaerial regime.
-The transition of subglacial to subaerial conditions may occur in the transition between pressurized and subaerial flow under glacier terminus \citep{perolo2018}, especially if closure rates of the ice channels are slow  \citep{rothlisberger1972}.
-Furthermore,  if a transport-limited regime persists at the glacier margin, then sediment can be easily mobilized and subaerial processes could be represented just a short distance downstream.
+First, these measurement stations are $0.5$ to $2$\,\unit{km} downstream of the glacier terminus \citep{cowton2012,felix2022}, likely a long enough distance for the pressurized subglacial regime to transition to the open channel subaerial one.
+The transition of subglacial to subaerial conditions may occur underneath glacier terminus \citep{perolo2018}, especially if closure rates of the ice channels are slow  \citep{rothlisberger1972}.
+Furthermore,  if a transport-limited regime persists at the glacier margin, then sediment can be easily mobilized, and subaerial processes could be represented just a short distance downstream.
 
-The consistent relationship between sediment discharge capacity and open channel flow conditions could be pronounced $\sim20$\,\unit{km} downstream of the Leverett measurement site, where a strong correlation persists between sediment plume size and water discharge into the Kangerlussuaq fjord \citep{mcgrath2010}.
+The consistent relationship between sediment discharge capacity and open channel flow conditions could be pronounced $\sim20$\,\unit{km} downstream of the Leverett site, where a strong correlation persists between sediment plume size and the river's water discharge into the Kangerlussuaq fjord \citep{mcgrath2010}.
 In marine-terminating glacier catchments, for contrast, a less consistent relationship may occur between water discharge or melt extent and sediment plume size \citep{tedstone2012}.
-The lack of correlation could be due to the variable relationship between subglacial sediment transport capacity and water discharge at marine-terminating glaciers where subglacial channels are pressurized by ocean water at the ice front \citep[e.g.][]{how2017}.
+Here, ocean water pressurizes subglacial channels at the ice front \citep[e.g.][]{how2017}, so the minimal correlation could result from the inconsistent relationship between subglacial sediment transport capacity and water discharge (Figure~\ref{fig:Qw_vari}).
 
@@
 A supply-limited regime could well persist at many glaciers, because the threshold of motion for sediment can be reached more frequently, and across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
-In a supply-limited system, sediment discharge would also scale with sediment production from glacier quarrying and abrasion, along with meltwater access \citep[e.g.][]{herman2015,ugelvig2018,delaney2023}.
 Sediment exhaustion through this process may also explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \citep{overeem2017}.
@@
 
-Records of sediment discharge have been used to establish the relationship, or lack thereof, with climate forcing and glacial erosion \citep[e.g.][]{koppes2009a,willenbring2016,mariotti2021}.
+Sediment discharge records have been used to establish the relationship, or lack thereof, with climate forcing and glacial erosion \citep[e.g.][]{koppes2009a,ganti2016,willenbring2016,mariotti2021}.
 The variable relationship between water discharge and subglacial sediment transport capacity presented here mostly applies to the daily- to weekly-time scales controlling the size of subglacial channels (Figure~\ref{fig:multi_run}).
-Yet, identifying climatic signals in sediment transport from the transport-limited component of glacier-sedimentary systems may require higher thresholds of water discharge compared to subaerial systems \citep[Figure~\ref{fig:Qw_vari}; ][]{tofelde2021}.
+Yet, identifying climatic signals in sediment transport from the transport-limited component of glacier-sedimentary systems may require higher thresholds of water discharge compared to subaerial systems when water discharge is variable  \citep[Figure~\ref{fig:Qw_vari}; ][]{tofelde2021}.
 
-Over time scales longer than days to weeks, the glacier's sediment transport capacity reacts to the ice thickness, controlling the channel closure rate, and the surface slope of the glacier, in addition to water discharge \citep[Figure~\ref{fig:multi_run}, ~ Section~\ref{sect:sub_mode}; ][]{rothlisberger1972,shreve1972,stevens2022}.
-If the glacier is in a supply-limited state, often assumed in landscape evolution models, then the sediment discharge record also represents additional processes of sediment availability and bedrock erosion from  water pressure variations and sliding  \citep[e.g.][]{iverson2012,herman2015,delaney2023}.
-This multitude of processes lies in contrast to most subaerial systems, where transport capacity typically increases with water discharge and hydraulic gradient, and the hydraulic gradient generally remains stable over time scales ranging from years to millennia \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wickert2019}.
-The stronger link between sediment transport capacity and hydrological forcing in subaerial channels means that variations in sediment transport records can be more easily attributed to climatic conditions than in  subglacial systems.
+Over time scales longer than days to weeks, the glacier's sediment transport capacity reacts to the ice thickness, controlling the channel closure rate, and the glacier's surface slope, in addition to water discharge \citep[Figure~\ref{fig:multi_run}, ~ Section~\ref{sect:sub_mode}; ][]{rothlisberger1972,shreve1972,stevens2022}.
+In a supply-limited state, often assumed in landscape evolution models, then glaciers' sediment discharge record also represents additional processes of sediment availability and bedrock erosion from  water pressure variations and sliding  \citep[e.g.][]{iverson2012,herman2015,delaney2023}.
+This multitude of processes lies in contrast to most subaerial systems, where transport capacity typically responds to sediment size, water discharge and hydraulic gradient, and the hydraulic gradient generally remains stable over time scales ranging from years to millennia \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wickert2019}.
+The stronger link between sediment transport capacity and hydrological forcing in subaerial channels may mean that variations in sediment transport records can be more easily attributed to climatic conditions than in subglacial systems.
 
@@
 In contrast, pressurized subglacial channels largely accommodate changing water discharge by altering water velocity and shear stress, upon which sediment transport depends.
-Water velocity responds quickly to changes in water discharge because the size of subglacial channels evolves slowly compared to variations in water discharge.
+Water velocity responds quickly to changes in water discharge because the size of subglacial channels generally evolves slowly compared to variations in water discharge.
 
@@
 \begin{enumerate}
-\item the inconsistent relationship between water discharge and sediment transport capacity underneath glaciers, causing increased variability in sediment transport capacity compared to river systems;
+\item the inconsistent relationship between water discharge and sediment transport capacity underneath glaciers creates increased variability in sediment transport capacity compared to river systems;
 \item the timing of peak sediment transport capacity below glaciers generally coincides with large rates of increasing in water discharge, as opposed to the maximum quantity of water discharge;

### 2023-09-13 Delaneyetal2023b.tex
Mauro's changes largely integrated and polishing of text

@@
   In subaerial channels, changing water discharge  changes the water depth,  channel width, and velocity.
-  However, in subglacial environments and on time scales of hours to weeks, changing water discharge  mainly alters the  water velocity, not channel area or width, because the subglacial conduit maintains a largely fixed geometry.
+  However, in subglacial environments and on time scales of hours to weeks, changing water discharge  mainly alters the  water velocity, not channel area or width, because the subglacial channel maintains a largely fixed geometry.
   Here, parameterizations of sediment transport capacity for pressurized flow under glaciers and for subaerial open channel flow are applied to hydrographs from an Alpine glacier and the Greenland Ice Sheet.
@@
 Over long periods, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from under glaciers \citep[c.f.][]{hallet1979,iverson2012,ugelvig2018}.
-On shorter time periods, pressurized water transports sediment from the under glacier \citep{walder1994,creyts2013,beaud2018}, should enough sediment be present subglacially (i.e.
+On shorter time periods, pressurized water transports sediment from the under glacier \citep{walder1994,creyts2013,beaud2018,delaney2019}, should enough sediment be present subglacially (i.e.
 in a transport-limited regime).
@@
 Water and suspended sediment data are leveraged from an Alpine and an ice sheet setting.
-The Alpine site (\alpine) is  Fieschergletscher in the Swiss Alps ($46^\circ\,29'\,07"$ N, $8^\circ\,08'\,34"$ E).
+The Alpine site (\alpine) is  Fieschergletscher in the Swiss Alps ($46^\circ\,29'\,07''$ N, $8^\circ\,08'\,3''$ E).
 Water discharge and suspended sediment concentration were collected here at a $1$\unit{min} interval continuously over the period 2014--2021.
@@
 The Leverett Glacier in Greenland (\icesheet) serves as the ice sheet setting.
-Water discharge and suspended sediment concentration data were collected roughly $2$\unit{km} downstream from the terminus from June to September 2009 to 2012 ($67^\circ\,03'\,50"$ N, $50^\circ\,12'\,59"$ W).
-This data is available from \citet{tedstone2017} at a $5$\unit{min} time interval.
+Water discharge and suspended sediment concentration data were collected roughly $2$\unit{km} downstream from the terminus from June to August/September 2009 to 2012 ($67^\circ\,03'\,5''$ N, $50^\circ\,12'\,59''$ W).
+This data is available from \citet{tedstone2017} at a $5$\,\unit{min} time interval of the 2012 season.
 Note that in some cases the removal and relocation of sensors created variations in the sediment concentration not resulting from natural processes.
@@
 \end{equation}
-where $\beta$ is the central angle of the circular segment that comprises the channel (the so-called Hooke angle). Note that $\beta =\pi$ corresponds to a semi-circle and
-smaller values of $\beta$ result in shallow, wide channels.
+where $\beta$ is the central angle of the circular segment that comprises the channel (the so-called Hooke angle).
 
@@
 
-The width of the flat channel floor $w_{cg}$, given angle $\beta$ is
+The flat channel floor width $w_{cg}$, given angle $\beta$, is
 \begin{equation}
@@
 
-The width-integrated shear stress is represented as $\tau_{tg}=w_{cg}\,\tau_g $.
+The width-integrated shear stress is represented as
+\begin{equation}
+  \label{eq:tautg}
+  \tau_{tg}=w_{cg}\,\tau_g.
+\end{equation}
 
@@
 %
-As in Section~\ref{sect:sub_mode}, the width-integrated shear stress is $\tau_{tf}=w_{cf}\,\tau_f$.
-
+As above, the width-integrated shear stress is
+\begin{equation}
+    \label{eq:tautf}
+    \tau_{tf}=w_{cf}\,\tau_f.
+\end{equation}
 \subsection{Implementation}
@@
 This occurs as the subglacial values are generally larger than the subaerial values, thus more efficient at transporting sediment.
-In the subglacial channel, these peaks generally occur over the period when water discharge increases at the fastest rate (Figure~\ref{fig:model_outs}).
-Conversely, in the subaerial channel, peaks in these variables occur when water discharge is at its highest value.
-As a result, peaks in the subglacial sediment transport variables occur  when the rate of water discharge increase is greatest
-is greatest, as opposed to when the maximum amount of water discharge occurs.
+In the subglacial channel, peaks generally occur over the period when water discharge increases at the fastest rate (Figure~\ref{fig:model_outs}).
+Conversely, in the subaerial channel, peaks in these sediment transport variables occur when water discharge is highest.
 The timing of the greatest rate of change in water discharge occurs before the peaks in water discharge.
@@
 
-Each water discharge in the subaerial channel is associated with a single water velocity, shear stress, and width-integrated shear stress, resulting in a perfect rank correlation between the variables (Figure~\ref{fig:Qw_vari} a--f).
+Each water discharge value in the subaerial channel results in a unique water velocity, shear stress, and width-integrated shear stress, creating a perfect rank correlation between the variables (Figure~\ref{fig:Qw_vari} a--f).
 Conversely in the subglacial channel,  water velocity, shear stress, and width-integrated shear stress can vary substantially for a given water discharge and width-integrated shear stress varies with respect to the other two  outputs as well (Figure~\ref{fig:Qw_vari} a--f).
-For instance, very high values of water velocity and shear stress can occur from minimal water discharge at $2.5$\,\unit{m}$^3$\,\unit{s}$^{-1}$ to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
-In \icesheet, mean values of water velocity can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, effectively spanning much of the water discharge range.
+For instance, very high  water velocity values and shear stresses can occur across minimal water discharge at $2.5$\,\unit{m}$^3$\,\unit{s}$^{-1}$ to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
+In \icesheet, mean values of water velocity can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, effectively spanning much of the water discharge range (Figure~\ref{fig:Qw_vari} d--f).
 
-When accounting for the evolving channel width of the subglacial conduit, width-integrated shear stress  across the channel generally increases with water discharge, with improved rank correlation compared to water discharge and water velocity or shear stress (Figure~\ref{fig:Qw_vari} a--f).
+When considering the subglacial channel's evolving width, width-integrated shear stress  across the channel generally increases with water discharge, with improved rank correlation compared to water discharge and water velocity or shear stress (Figure~\ref{fig:Qw_vari} a--f).
 Yet even the width-integrated shear stress  can vary substantially, with the highest values occurring at water discharge values ranging from roughly $11$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, to over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
-The variability in width-integrated shear stress is less pronounced in the \icesheet case, in part due to the reduced variability in water discharge.
-Yet, a range of width-integrated shear stress can occur at a given discharge (Figure~\ref{fig:Qw_vari} c, f).
+The variability in width-integrated shear stress is less pronounced in the \icesheet case, with reduced variability in water discharge (Figure~\ref{fig:Qw_vari} c, f).
+Yet, a range of width-integrated shear stress can occur at a given discharge.
 
@@
       Text in columns shows rank correlation between the variable and sediment discharge for the corresponding time period for the subglacial (right column) and subaerial (left column) channels.
+      \ian{time aggregation ``smooths'' between data gaps here....}
     }
@@
 Across the range of parameter values examined, variability in water velocity and shear stress remains higher in the subglacial system compared to the subaerial one.
-However, in \alpine{}, the width-integrated shear stress makes the variability between the two systems comparable (Figure~\ref{fig:multi_run}).
+However, in \alpine{}, the variability in width-integrated shear stress is comparable between the two systems, due to the variable water discharge (Figure~\ref{fig:multi_run}).
 Conversely, in \icesheet{}, with lower water discharge variability, the variability in width-integrated shear stress remains higher in the subglacial system across the time aggregation periods.
@@
 In both cases, the variability in velocity, shear stress, and width-integrated shear stress decrease after $1-5$ day aggregations (Figure~\ref{fig:multi_run}).
-This suggests that the effects of the variable response to hydrologic forcing are reduced over longer time aggregations than these.
 Yet, substantial differences in variability persist in periods of up to $15$ days, as the variability in aggregations periods longer than this might be conflated with seasonal variations in water discharge.
-Note that lags and time of peak sediment transport conditions still occur over these periods (Figure~\ref{fig:model_outs}), while the variability in sediment transport parameters is reduced with increasing time aggregation (Figure~\ref{fig:multi_run}).
+Note that lags and time of peak sediment transport conditions still occur over these periods (Figure~\ref{fig:model_outs}), while the variability in sediment transport parameters decreases with larger time aggregation (Figure~\ref{fig:multi_run}).
 
@@
 Quantities that control sediment transport capacity, such as water velocity, shear stress, and width-integrated shear stress, vary differently in response to water discharge in subglacial channels compared to subaerial ones (Figure~\ref{fig:Qw_vari}).
-In sediment transport relationships, such as in \citet{meyer1948} or \citet{engelund1967}, that evaluate sediment discharge, shear stress is scaled to the $\frac{3}{2}$ or $\frac{5}{2}$ power, respectively.
-The shear stress exponent greater than $1$ magnifies sediment discharge variability beyond the highly variable sediment transport parameters described above (Figure~\ref{fig:multi_run}).
-Note that sediment transport relationships describe sediment flux per unit width so that the integrated capacity across channel occurs after application of the exponent, unlike the width-integrated shear stress evaluation here \citep[c.f.][]{tucker1997}.
+In sediment transport relationships, such as in \citet{meyer1948} or \citet{engelund1967}, that evaluate sediment discharge capacity, shear stress is scaled to the $\frac{3}{2}$ or $\frac{5}{2}$ power, respectively.
+The exponent greater than $1$ magnifies sediment discharge variability beyond the variable sediment transport parameters described above (Figure~\ref{fig:multi_run}).
+Note that sediment transport relationships describe sediment flux per unit width so that the integrated capacity across channel occurs after application of the exponent, unlike the width-integrated shear stress evaluation here \citep[Equations\,\ref{eq:tautg} and \ref{eq:tautf}; c.f.][]{tucker1997}.
 %Note also that both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress is homogeneous across the channel bed, despite the limitations of this assumption \citep[e.g.][]{yager2018}.
@@
 For sufficiently short time periods,  channel width and wetted area are fixed in the subglacial system so that the response of width-integrated shear stress is $ \tau_{tg}\, \propto\,  Q^{2}$, with a larger exponent on $Q$ than in the subaerial case.
-As a result, when water discharge fluctuates at shorter timescale than the subglacial channel evolution \citep{gimbert2016}, the width-integrated shear stress varies more with water discharge than in subaerial channels.
-Note also that the subglacial shear-stress response to water discharge will be greater than the subaerial's provided that the exponent is less than one in Equation~\ref{eq:wcf}.
+In this case, when water discharge fluctuates at shorter timescale than the subglacial channel evolution \citep{gimbert2016}, the width-integrated shear stress varies more with water discharge than in subaerial channels.
+Note also that the subglacial shear-stress response to water discharge will be greater than the subaerial's provided that the exponent is below one in Equation~\ref{eq:wcf}.
 However, when subglacial channel size lies in equilibrium with water discharge, the  width-integrated shear stress response is $\tau_{tg}\, \propto \, Q^{\frac{4}{5}}$, similar to subaerial channels.
-Hydraulics could trend toward this equilibrium regime after $1-5$ days, when variability in subglacial channels begins to approach that of subaerial channels (Figure~\ref{fig:multi_run}). 
-\ian{mauro: should a supplement be made for this? if so, could you send me the one for R-channels? is this essentially equation 14 in werder et al., 2010}
+Hydraulics could trend toward this equilibrium regime with average discharge over $1-5$ days, when variability in subglacial channels  approaches that of subaerial channels (Figure~\ref{fig:multi_run}).
+Furthermore, the stronger correlation between hydraulics and water discharge in \icesheet{} suggests that this system is closer to this equilibrium compared to \alpine{}, given the reduced diurnal discharge amplitude in \icesheet{}.
+
+\ian{mauro: should a supplement be made for this simplification? if so, could you send me the one for R-channels? is this essentially equation 14 in werder et al., 2010?}
 
@@
 
-First, these measurement stations are $0.5$ to $2$\,\unit{km} downstream of the glacier terminus \citep{cowton2012,felix2022}, likely a long enough distance for the pressurized subglacial regime to transition to the open channel subaerial one.
+First, these measurement stations lie $0.5$ to $2$\,\unit{km} downstream of the glacier terminus \citep{cowton2012,felix2022}, likely a long enough distance for the pressurized subglacial regime to transition to the open channel subaerial one.
 The transition of subglacial to subaerial conditions may occur underneath glacier terminus \citep{perolo2018}, especially if closure rates of the ice channels are slow  \citep{rothlisberger1972}.
@@
 Second, the sediment discharge records could represent processes such as increased sediment accessibility when melt extends up glacier, thereby concurrently increasing sediment and water discharge \citep[e.g.][]{vergara2022}.
-A supply-limited regime could well persist at many glaciers, because the threshold of motion for sediment can be reached more frequently, and across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
-Sediment exhaustion through this process may also explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \citep{overeem2017}.
+A supply-limited regime could well persist at many glaciers, because the sediment's threshold of motion can be reached more frequently, across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
+Sediment exhaustion occurring through this process may also explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \citep{overeem2017}.
 Note that in the data here, however, the absence of lags between the observed sediment and water discharge may suggest that sediment is mobilized close to the measurement stations, as opposed to distal areas of the glacier bed as would be expected in a supply-limited regime \citep[Figure~\ref{fig:model_outs}\, a and e; ][]{williams1989}.
@@
 
-Sediment discharge records have been used to establish the relationship, or lack thereof, with climate forcing and glacial erosion \citep[e.g.][]{koppes2009a,ganti2016,willenbring2016,mariotti2021}.
+Sediment discharge records have been used to establish the relationship, or lack thereof, between climate forcing and glacial erosion \citep[e.g.][]{koppes2009a,ganti2016,willenbring2016,mariotti2021}.
 The variable relationship between water discharge and subglacial sediment transport capacity presented here mostly applies to the daily- to weekly-time scales controlling the size of subglacial channels (Figure~\ref{fig:multi_run}).
-Yet, identifying climatic signals in sediment transport from the transport-limited component of glacier-sedimentary systems may require higher thresholds of water discharge compared to subaerial systems when water discharge is variable  \citep[Figure~\ref{fig:Qw_vari}; ][]{tofelde2021}.
+Yet, identifying climatic signals in sediment transport from the transport-limited component of glacier-sedimentary systems may require higher thresholds of water discharge compared to subaerial systems when water discharge varies substantial over short timescales \citep[Figure~\ref{fig:Qw_vari}; ][]{tofelde2021}.
 
-Over time scales longer than days to weeks, the glacier's sediment transport capacity reacts to the ice thickness, controlling the channel closure rate, and the glacier's surface slope, in addition to water discharge \citep[Figure~\ref{fig:multi_run}, ~ Section~\ref{sect:sub_mode}; ][]{rothlisberger1972,shreve1972,stevens2022}.
+Over time scales of weeks or longer, the glacier's sediment transport capacity responds to the ice thickness, controlling the channel closure rate, and the glacier's surface slope, in addition to water discharge \citep[Figure~\ref{fig:multi_run}, ~ Section~\ref{sect:sub_mode}; ][]{rothlisberger1972,shreve1972,gimbert2016,stevens2022}.
 In a supply-limited state, often assumed in landscape evolution models, then glaciers' sediment discharge record also represents additional processes of sediment availability and bedrock erosion from  water pressure variations and sliding  \citep[e.g.][]{iverson2012,herman2015,delaney2023}.
-This multitude of processes lies in contrast to most subaerial systems, where transport capacity typically responds to sediment size, water discharge and hydraulic gradient, and the hydraulic gradient generally remains stable over time scales ranging from years to millennia \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wickert2019}.
+This multitude of processes lies in contrast to most subaerial systems, where transport capacity typically responds to sediment size, water discharge and hydraulic gradient, and the hydraulic gradient generally remains stable over time scales ranging from years to millennia \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,tucker1997,wickert2019}.
 The stronger link between sediment transport capacity and hydrological forcing in subaerial channels may mean that variations in sediment transport records can be more easily attributed to climatic conditions than in subglacial systems.
@@
 Water velocity responds quickly to changes in water discharge because the size of subglacial channels generally evolves slowly compared to variations in water discharge.
-
-Despite these differences, parameterizations of subaerial hydraulics generally demonstrate a better representation of observed sediment discharge variations compared to subglacial channels.
-This could be attributed to the role of open channel flow at the glacier margin and in the proglacial area, or the influence of water discharge in driving sediment availability, rather than subglacial sediment discharge capacity.
+Despite these differences, parameterizations of subaerial hydraulics generally represent observed sediment discharge variations better than subglacial hydraulics.
+%This could be attributed to the role of open channel flow at the glacier margin and in the proglacial area, or the influence of water discharge in driving sediment availability, rather than subglacial sediment discharge capacity.
 

### 2023-09-13 Delaneyetal2023b.tex
Updates from Overleaf

--- 2023-09-06 Delaneyetal2023b.tex
@@
 \newcommand{\ian}[1]{{\textbf{\color{blue}Ian says:} \color{blue} #1} }
-\newcommand{\mauro}[1]{{\textbf{\color{green}Mauro says:} \color{green} #1} }
 \newcommand{\alpine}{\textit{ALPINE}\,}
@@
   Andrew J. Tedstone\footnote{Department of Geosciences, University of Fribourg, Ch. du Musée 1700, Fribourg, Switzerland},
-  Daniel Farinotti\footnote{Laboratory of Hydraulics, Hydrology and Glaciology (VAW), ETH-Z\"urich, H\"onggerbergring 26, 8093 89 Z\"urich, Switzerland}$^{,}$\footnote{Swiss Federal Institute for Forest, Snow and Landscape Research (WSL) Z\"uricherstrasse 111, 8903 1011 Birmensdorf, Switzerland},
-  Mauro A. Werder$^{3,4}$
+  Daniel Farinotti\footnote{Laboratory of Hydraulics, Hydrology and Glaciology (VAW), ETH-Z\"urich, H\"onggerbergring 26, 8093 89 Z\"urich, Switzerland}$^{,}$\footnote{Swiss Federal Institute for Forest, Snow and Landscape Research (WSL) Z\"uricherstrasse 111, 8903 1011 Birmensdorf, Switzerland}
 
@@
   \noindent
-  In both subglacial and subaerial channels, the sediment transport capacity for sediment size depends mainly on 1) the channel width over which the sediment is mobilized  and 2) water velocity, as this controls the shear stress that the flowing water applies to the channel bottom.
+  In both subglacial and subaerial channels, the sediment transport capacity of a sediment size depends mainly on 1) the channel width over which the sediment is mobilized  and 2) water velocity, as this controls the shear stress that the flowing water applies to the channel bottom.
   In subaerial channels, changing water discharge  changes the water depth,  channel width, and velocity.
   However, in subglacial environments and on time scales of hours to weeks, changing water discharge  mainly alters the  water velocity, not channel area or width, because the subglacial conduit maintains a largely fixed geometry.
-  Here, parameterizations of sediment transport capacity for pressurized flow under glaciers and for subaerial open channel flow are applied to hydrographs from an Alpine glacier and the Greenland Ice Sheet.
+  Here, parameterizations of sediment transport capacity for pressurized flow under glaciers and subaerial open channel flow are applied to hydrographs from an Alpine glacier and the Greenland Ice Sheet.
   Results of these parameterizations show that sediment transport capacity varies far more in subglacial channels than subaerial ones over time and with respect to water discharge.
@@
 
-In a transport-limited regime, sediment discharge responds to the sediment transport capacity defined as the amount of sediment the water can carry.
+In a transport-limited regime, sediment discharge responds to the sediment transport capacity or the amount of sediment the water can carry.
 Sediment transport capacity depends on the shear stress between water and sediment it flows  over \citep{shields1936,meyer1948,engelund1967} and the width of the channel bottom $w_c$ over which to mobilize sediment.
@@
     \label{eq:v}
-    v = \frac{Q}{S},
+    v = \frac{Q_w}{S},
   \end{equation}
 \end{linenomath*}
-where $Q$ is water discharge,  and $S$ is the channel's wetted area. %\mauro{I think ``wetted area'' is better and less ambiguous, changed throughout.}
-% \mauro{later $Q$ is used.  drop $_w$?}
+% 
+where $Q_w$ is water discharge,  and $S$ is the channel's cross-sectional area.
 
-In subaerial channels, operating with open channel flow, the wetted area $S$ of the channel evolves with changing water discharge $Q$, by changing both the channel's width and the water depth \citep{leopold1953}.
-The change in water depth results in a commensurate increase in water velocity.
-Thus higher discharge is accommodated by both an increase in wetted area and flow velocity such that Equation~\eqref{eq:v} remains fulfilled.
+In subaerial channels, with open channel flow, the cross-sectional area $S$ of the channel evolves with changing water discharge $Q_w$  by changing both the channel's width and the water depth \citep{leopold1953}.
+During a flood with greater water discharge $Q_w$, for instance, the cross-sectional area $S$ increases as the channel widens and deepens, along with water velocity $v$, which is squared.
 As a result, the shear stress across the channel $\tau_t$ will increase according to Equation~\ref{eq:tau_t}.
 
-The response of water velocity to changing water discharge in subglacial channels differs from subaerial ones.
+Yet, the response of water velocity to changing water discharge in subglacial channels differs from subaerial ones.
 The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by frictional heating of water flowing through the channel \citep{rothlisberger1972}.
-As a result, the subglacial channel size only evolves relatively slowly over days to months, whereas water discharge can vary over hours \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
-Therefore, subglacial water flow behaves more like pipe-flow and changes in water discharge $Q$ are mainly accommodated with increased water velocity $v$ \citep[e.g.][]{swift2005}, because the size of the channel $S$ changes on a much longer timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
-It follows that an increase in water discharge $Q$ will cause the shear stress across the channel $\tau_t$ to respond with water velocity $v$, as opposed to subaerial channels where channel width $w_c$ evolves as well.
+As a result, the subglacial channel size only evolves over days to months, whereas water discharge can vary over hours \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
+In turn, changes in water discharge $Q_w$ are mainly accommodated with water velocity $v$ \citep[e.g.][]{swift2005}, because the size of the channel $S$ changes on a much shorter timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
+It follows that an increase in water discharge $Q_w$ will cause the shear stress across the channel $\tau_t$ to respond with water velocity $v$, as opposed to subaerial channels where channel width $w_c$ evolves as well.
 
@@
 Water velocity magnitude in the subglacial, $v_g$, and subaerial, $v_f$, channels are shown by arrow length.
-$S_g$ and $S_f$ represent the wetted area in subglacial and subaerial channels.
+$S_g$ and $S_f$ represent the cross-sectional area in subglacial and subaerial channels.
 The subglacial channel widths $w_{cg}$ remain unchanged, while the subaerial channel widths evolve $w_{cf}$.
 Shear stress $\tau$ is responsible for the mobilization of sediment and increases with the number of makers at the channel bed.
-      Note that in the subaerial channel parameterization, channels have rectangular shapes with their width much larger than the depth (Section~\ref{sect:fluv}).}
+      Note that in the subaerial channel parameterization, channels have rectangular shapes (Section~\ref{sect:fluv}).} 
     \label{fig:cartoon}
@@
 Furthermore, thorough observations of water velocity and channel width in subglacial channels are limited and, to these authors' knowledge, not contrasted with subaerial conditions at the glacier margins.
-%\mauro{self-cite here?}
 
 This manuscript aims to evaluate the relationship between sediment transport capacity and water discharge in subglacial channels, compared to subaerial ones.
-In the absence of continuous measurements of subglacial and subaerial water velocity and channel morphology, numerical parameterizations of subglacial and subaerial hydraulics are utilized.
+Without continuous measurements of subglacial and subaerial water velocity and channel morphology, numerical parameterizations of subglacial and subaerial hydraulics are utilized.
 Parameterizations are applied to hydrological records from an Alpine glacier in Switzerland (Fieschergletscher), and  a land-terminating glacier in Greenland (Leverett glacier).
@@
 
-% \TODO{Address Daniel's comment: I think that these records need a few more words: where are they measured exactly?
+% \TODO{Address Daniel's comment: I think these records need a few more words: where are they measured exactly?
 How long are the time series?
@@
 \label{sect:meth}
-The parameterizations below (Sections ~\ref{sect:sub_mode}~and~\ref{sect:fluv}) represent relationships amongst water discharge, water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
+Parameterizations below (Sections ~\ref{sect:sub_mode}~and~\ref{sect:fluv}) represent relationships amongst water discharge, water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
 Both parameterizations calculate   water velocity, shear stress and width-integrated shear stress, upon which sediment transport depends \citep[Figure \ref{fig:cartoon}; ][]{shields1936}.
 The evaluation of these variables omits the selection of a sediment transport relationship and grain-size parameter \citep[e.g.][]{shields1936,meyer1948} needed to calculate sediment transport capacity from shear stress.
-
-\mauro{Yeah, I wonder whether width-integrated should be scaled by width to 2/3.}
 
@@
   \caption{Variables, parameters, and constants used in this work.} 
-  %Where two values are given, the first refers to  \alpine{}, a case from Fieschergletscher, and the second to a glacier marginal to the \icesheet{} a case from Leverett Glacier.
-\mauro{those f are high}
+  %Where two values are given, the first refers to  \alpine{}, a case from Fieschergletscher, and the second to a glacier marginal to the \icesheet{} a case from Leverett Glacier }
   \begin{tabular}{ l  c  c c }
@@
     \textbf{Variables}  & & & \\
-    Water discharge  & $Q$& & $\mathrm{m^{3}\,s^{-1}}$ \\
+    Water discharge  & $Q_w$& & $\mathrm{m^{3}\,s^{-1}}$ \\
     Water velocity (glacier, subaerial)  & $v_g,\,v_{f}$& & $\mathrm{m\,s^{-1}}$ \\
-    Channel wetted area (glacier, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
+    Channel cross-sectional area (glacier, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
     Hydraulic diameter &$D_h$&&$\mathrm{m}$\\
@@
     \label{eq:dS_dt}
-    \frac{\partial S_g}{\partial t} = C_1 \frac{Q \Delta h}{l} - C_2 \left(h_{o}-\frac{\Delta h}{2}\right)^n\,S_g,
+    \frac{\partial S_g}{\partial t} = C_1 \frac{Q \Delta h}{l} - C_2 (h_{o}-\frac{\Delta h}{2})^n\,S_g,
   \end{equation}
 \end{linenomath*}
-\noindent where $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (Table \ref{table:vpm}), $l$ is the length of subglacial channel, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change from the glacier terminus to its top, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the ice overburden pressure, and $n$ is Glen's n \citep{glen1955} (usually $n=3$).
+\noindent where $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (Table \ref{table:vpm}), $l$ is the length of subglacial channel, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change from the glacier terminus to its top, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the ice overburden pressure, and $n$ is Glen's n \citep{glen1955}.
 The first term on the right side of the equation represents the opening of the channel through frictional heating, while the following term represents the creep closure of the channel from ice deformation.
@@
 smaller values of $\beta$ result in shallow, wide channels.
-The hydraulic diameter $D_h$ is converted to wetted area $S_g$ given
-\begin{equation}
-  \label{eq:Dh2S}
-  S_g= \frac{D_h^2}{2}\,\frac{(\frac{\beta}{2}+\sin \frac{\beta}{2})^2}{\beta - \sin \beta}.
-\end{equation}
 
-With knowledge of wetted area $S_g$, the shear stress, $\tau_g$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
+With knowledge of cross-sectional area $S_g$, the shear stress, $\tau_g$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
 \begin{equation}
@@
 % 
-where  the water velocity $v_g = \frac{Q}{S_g}$.
+where  the water velocity $v_g = \frac{Q_w}{S_g}$.
 
@@
 To parameterize the shear stress of water flowing across sediments in the subaerial channel,  the hydraulics parameterization presented in \citet{tucker1997} is implemented.
-Using mass conservation and the Darcy-Weisbach relationship as well as assuming that the channel is wide so that the hydraulic radius is well approximated by the flow depth, then
-the shear stress $\tau_f$ at the river bed is given by
+Here, given  that mass is conserved mass, there is a sufficiently wide channel so that the hydraulic radius is consistent with flow depth, and the Darcy-Weisbach relationship,
+the shear stress $\tau_f$ at the river bed is represented as
 \begin{linenomath*}
@@
     \label{eq:DW_tau}
-    \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q}{w_{cf}} \Big)^{\frac{2}{3}} \,\nabla z_c^{\frac{2}{3}},
+    \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q_w}{w_{cf}} \Big)^{\frac{2}{3}} \,\nabla z_c^{\frac{2}{3}},
   \end{equation}
@@
   \label{eq:wcf}
-  w_{cf} = k \, Q^\frac{1}{2},
+  w_{cf} = k \, Q_w^\frac{1}{2},
 \end{equation}
@@
 where $k$ is a constant and the exponent is commonly given as $\frac{1}{2}$ \citep{leopold1953}.
-%\mauro{This is for so called ``regime channels''}
 Water velocity, $v_f$, is given by rearranging Equation \ref{eq:tau} as
@@
 
-The parameterizations above are applied to hydrological records from the Fieschergletscher (\alpine) and the Leverett glacier (\icesheet).
+The parameterizations above are applied to hydrological records from the Fieschergletscher in the Swiss Alps (\alpine) and the Leverett glacier on the Greenland Ice Sheet (\icesheet).
 The subglacial and subaerial parameterizations represent the subglacial flow of water exiting a glacier and transitioning to subaerial flow, as it moves through the catchment.
 Outputs of the parameterizations are meant to represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
-To generalize these cases, \alpine{}  is exemplified by relatively thin ice thickness ($h_{ice}$: $225$\,\unit{m}), low water discharge ($\sim\,10$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
- \icesheet{}  is exemplified by thick ice  ($h_{ice}$: $700$\,\unit{m}), high water discharge ($\sim\,300$\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).
+To generalize these cases, \alpine{}  is exemplified by relatively thin ice thickness ($h_{ice}$: $225$\,\unit{m}), low water discharge ($<\,20$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
+ \icesheet{}  is exemplified by thick ice  ($h_{ice}$: $700$\,\unit{m}), high water discharge (generally $<\,500$\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).
 
@@
 
-Lastly,  outputs are compared with the record of sediment discharge from the two glaciers using Spearman rank correlation, or the correlation with respect to ordering of values from the outputs and observations.
-Rank correlation reduces the impact of the non-linear response of sediment discharge to the hydrological forcing and the uncertainties from sensor disturbances in the sediment transport record from \icesheet{}.
-Outputs and observations are aggregated across timescales from $1$\,\unit{hr} to $15$\,\unit{days} from the $15$\,\unit{min} outputs, by averaging values one half of the timescale before and after a point time.
-%\mauro{Here a brief explanation of the rank-stuff would be good.  Also describe the time aggregation.}
+Lastly,  outputs are compared with the record of sediment discharge from the two glaciers using Spearman rank correlation to reduce the impact of the non-linear response of sediment discharge to the hydrological forcing.
+Rank correlation also reduces the impact uncertainties from sensor disturbances in the sediment transport record from \icesheet{}.
 
@@
     Note that the x and y axes are different for \alpine{} and \icesheet{} such that they each cover a summer.
-        \ian{Andrew's comment: for a given channel width, subglacial channels are more effective at mobilizing and transporting sediment than subaerial ones}
-        \ian{also, might need to go back in Timemaching to pull out old version}
+    %    \ian{Andrew's comment: for a given channel width, subglacial channels are more effective at mobilizing and transporting sediment than subaerial ones}
   } 
@@
 Following the parameterizations defined in Sections~\ref{sect:sub_mode}~and~\ref{sect:fluv}, the water velocity, shear stress, and width-integrated shear stress exhibit different seasonal evolutions and peaks.
-Water velocity and shear stress represent the potential for sediment mobilization, and width-integrated shear stress represents a proxy for the total sediment transport capacity across the channel bed (Figure~\ref{fig:model_outs}).
+Water velocity and shear stress represent the potential for sediment mobilization, and width-integrated shear stress represents the total sediment transport capacity across the channel bed (Figure~\ref{fig:model_outs}).
 
@@
   \caption{Relationship between water discharge and velocity, shear stress, and width-integrated shear stress for \alpine{} (a-c) and \icesheet (d-f).
-    These outputs' relationship with measured sediment discharge is given for \alpine in (g-i) and \icesheet (j-l).
+    These outputs' relationship with sediment discharge is given for \alpine in (g-i) and \icesheet (j-l).
     Variables have been normalized to mean values for better comparison amongst the runs.
@@
     Plots are shown with data and outputs at $15$\unit{min} interval.
-    \mauro{Why slightly different colours between ALPINE and ICESHEET?}
+    \ian{TOtal vs Width integrated}
   } 
@@
     \includegraphics[width=0.9\linewidth]{Fig4.png}
-    \caption{Variance (measured by the standard deviation) in water velocity, shear stress, and width integrated shear stress for different time aggregations across the range of subglacial and subaerial channel shapes and friction factors.
+    \caption{Variance (measured by the standard deviation) in water velocity, shear stress, and total shear stress for different time aggregations across the range of subglacial and subaerial channel shapes and friction factors.
       Shaded areas denote the minimum and maximum of the $100$ parameter combinations tested for each of the systems (Section~\ref{sect:imp}).
@@
 \subsection{Relationship between parameterization  outputs and observed sediment discharge}
-When comparing the output variables with suspended sediment discharge measurements, then velocity, shear stress, and width-integrated shear stress from the subaerial system generally show better correlation with sediment discharge compared to subglacial outputs across the tested subglacial channel parameters (Figures~\ref{fig:Qw_vari} g--l and \ref{fig:multi_run}).
+When comparing the output variables with suspended sediment discharge data, velocity, shear stress, and width-integrated shear stress from the subaerial system generally show better correlation with sediment discharge compared to subglacial outputs across the tested subglacial channel parameters (Figures~\ref{fig:Qw_vari} g--l and \ref{fig:multi_run}).
 The rank correlation among the three output variables for the subaerial parameterization remains more or less constant despite the absence of hysteresis in channel shape and consistent scaling with water discharge, as shown above (Figure \ref{fig:Qw_vari} and Section~\ref{sect:fluv}).
@@
 In both cases, the exponent greater than $1$ in the sediment transport relationship magnifies the variability in sediment discharge beyond the highly variable shear stress and width-integrated shear stress described above (Figure~\ref{fig:multi_run}).
-Note that sediment transport relationships describe sediment flux per unit width so that the evaluation across channel occurs after application of the exponent, unlike the width-integrated shear stress evaluation here \citep[c.f.][]{tucker1997}.
+The rapid increase and decreases in subglacial sediment transport capacity may, for example, compound the variability in sediment transport already present in subaerial systems \citep[Figure~\ref{fig:Qw_vari}; ][]{williams1989,jerolmack2010}.
+Large variations in sediment transport capacity underneath glaciers may result in sediment mobilization and deposition processes in close spatial and temporal proximity to each other \citep{gimbert2016,perolo2018}.
 Note also that both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress is homogeneous across the channel bed, despite the limitations of this assumption \citep[e.g.][]{yager2018}.
-
-The rapid increase and decreases in subglacial sediment transport capacity may, for example, compound the variability in sediment transport already present in subaerial systems \citep[Figure~\ref{fig:Qw_vari}; ][]{williams1989,jerolmack2010}.
-For example, by applying the discharge - width scaling in Equation~\ref{eq:wcf} to Equation \ref{eq:tau_t} and to the $S \propto \,Q^{\frac{1}{2}}$ in Equation \ref{eq:v}, the response of the subaerial channel's width-integrated shear stress to changing water discharge is $\frac{\partial \tau_{tf}}{\partial t} \propto \frac{\partial Q^{\frac{3}{2}}}{\partial t}$.
-For fixed channel width and wetter area over sufficiently short time periods in the subglacial system, the response of width-integrated shear stress is $\frac{\partial \tau_{tg}}{\partial t} \propto \frac{\partial Q^{2}}{\partial t}$, with a larger exponent on $Q$.
 
@@
 
-The increased performance of subaerial hydraulics, in representing observations occurs despite  output variables showing the disparate relationship between sediment transport capacity and water discharge in subglacial systems (Figure~\ref{fig:Qw_vari}).
-This could occur for two reasons.
+Outputs from the subaerial system rank correlate higher with sediment discharge in the \alpine and \icesheet systems here.
+The increased performance occurs despite  output variables showing the disparate relationship between sediment transport capacity and water discharge in subglacial systems (Figure~\ref{fig:Qw_vari}), which could occur for two reasons.
 
-First, these measurement stations are $0.5$ to $2$\,\unit{km} downstream of the glacier terminus \citep{cowton2012,felix2022}, likely a long enough distance for the pressurized subglacial regime to transition to the open channel subaerial regime.
-The transition of subglacial to subaerial conditions may occur in the transition between pressurized and subaerial flow under glacier terminus \citep{perolo2018}, especially if closure rates of the ice channels are slow  \citep{rothlisberger1972}.
+First, these measurement stations are $0.5$ to $2$\,\unit{km} downstream of the glacier terminus \citep{cowton2012,felix2022}, respectively, likely a long enough distance for the pressurized subglacial regime to transition to the open channel subaerial regime.
+The transition of subglacial to subaerial conditions may occur in the transition between pressurized and subaerial flow under glacier terminus \citep{perolo2018}, especially if cavity closure rates of the ice are slow  \citep{egli2021b}.
 Furthermore,  if a transport-limited regime persists at the glacier margin, then sediment can be easily mobilized and subaerial processes could be represented just a short distance downstream.
 
-The consistent relationship between sediment discharge capacity and open channel flow conditions could be pronounced $\sim20$\,\unit{km} downstream of the Leverett measurement site, where a strong correlation persists between sediment plume size and water discharge into the Kangerlussuaq fjord \citep{mcgrath2010}.
-In marine-terminating glacier catchments, for contrast, a less consistent relationship may occur between water discharge or melt extent and sediment plume size \citep{tedstone2012}.
+The consistent relationship between sediment discharge capacity and open channel flow conditions could be pronounced ~ $20$\,\unit{km} downstream of the Leverett measurement site, where a strong correlation persists between sediment plume size and water discharge into the Kangerlussuaq fjord \citep{mcgrath2010}.
+In marine-terminating glacier catchments, for contrast, a less consistent relationship may persist between water discharge or melt extend and sediment plume size \citep{tedstone2012}.
 The lack of correlation could be due to the variable relationship between subglacial sediment transport capacity and water discharge at marine-terminating glaciers where subglacial channels are pressurized by ocean water at the ice front \citep[e.g.][]{how2017}.
@@
 
-%Regardless of the increased performance of water discharge compared to subglacial sediment transport capacity in the observations, parameterization outputs suggest that additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to subaerial systems \citep{jansson2005,ganti2016}.
-
-Records of sediment discharge have been used to establish the relationship, or lack thereof, with climate forcing and glacial erosion \citep[e.g.][]{koppes2009a,willenbring2016,mariotti2021}.
+Regardless of the increased performance of water discharge compared to subglacial sediment transport capacity in the observations, parameterization outputs suggest that additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to subaerial systems \citep{jansson2005,ganti2016}.
+Records of sediment discharge have been used to establish the relationship, or lack thereof, between sediment discharge and climate in glacial systems \citep[e.g.][]{koppes2009a,willenbring2016,mariotti2021}.
 The variable relationship between water discharge and subglacial sediment transport capacity presented here mostly applies to the daily- to weekly-time scales controlling the size of subglacial channels (Figure~\ref{fig:multi_run}).
@@
 
-Sediment transport capacity is largely driven by the shear stress of water flowing through a channel and the channel's width.
+Sediment transport capacity is largely driven by the velocity of water flowing through a channel and the channel's width.
 Subaerial channels  alter their channel width and water velocity immediately in response to changing water discharge.

### 2023-09-13 Delaneyetal2023b.tex
Following small changes on overleaf

--- 2023-09-13 Delaneyetal2023b.tex
@@
 The change in water depth results in a commensurate increase in water velocity.
-Thus,  fulfilling Equation~\eqref{eq:v}, the higher discharge is accommodated by both an increase in wetted area and flow velocity.
+Thus,  fulfilling Equation~\ref {eq:v}, the higher discharge is accommodated by both an increase in wetted area and flow velocity.
 As a result, the shear stress across the channel $\tau_t$ increases according to Equation~\ref{eq:tau_t}.
@@
 As a result, the subglacial channel size only evolves relatively slowly over days to months, whereas water discharge can vary over hours \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
-Therefore, subglacial water flow behaves more like pipe-flow and changes in water discharge $Q$ are mainly accommodated with increased water velocity $v$ \citep[e.g.][]{swift2005}, because the size of the channel $S$ changes on a much longer timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
+Therefore, subglacial water flow behaves more like pipe flow, and changes in water discharge $Q$ are mainly accommodated with increased water velocity $v$ \citep[e.g.][]{swift2005}, because the size of the channel $S$ changes on a much longer timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
 It follows that an increase in water discharge $Q$ will cause the shear stress across the channel $\tau_t$ to respond with water velocity $v$, as opposed to subaerial channels where channel width $w_c$ evolves as well.
@@
 The parameterizations below (Sections ~\ref{sect:sub_mode}~and~\ref{sect:fluv}) represent relationships amongst water discharge, water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
-Both parameterizations calculate   water velocity, shear stress and width-integrated shear stress, upon which sediment transport depends \citep[Figure \ref{fig:cartoon}; ][]{shields1936}.
+Both parameterizations calculate   water velocity, shear stress, and width-integrated shear stress, upon which sediment transport depends \citep[Figure \ref{fig:cartoon}; ][]{shields1936}.
 The evaluation of these variables omits the selection of a sediment transport relationship and grain-size parameter \citep[e.g.][]{shields1936,meyer1948} needed to calculate sediment transport capacity from shear stress.
@@
     Note that the x and y axes are different for \alpine{} and \icesheet{} such that they each cover a 
-        \ian{NaN values will be removed in final version}
+        \ian{NaN values will be removed in the final version}
       }
@@
 For sufficiently short time periods,  channel width and wetted area are fixed in the subglacial system so that the response of width-integrated shear stress is $ \tau_{tg}\, \propto\,  Q^{2}$, with a larger exponent on $Q$ than in the subaerial case.
-In this case, when water discharge fluctuates at shorter timescale than the subglacial channel evolution \citep{gimbert2016}, the width-integrated shear stress varies more with water discharge than in subaerial channels.
+In this case, when water discharge fluctuates at a shorter timescale than the subglacial channel evolution \citep{gimbert2016}, the width-integrated shear stress varies more with water discharge than in subaerial channels.
 Note also that the subglacial shear-stress response to water discharge will be greater than the subaerial's provided that the exponent is below one in Equation~\ref{eq:wcf}.
 However, when subglacial channel size lies in equilibrium with water discharge, the  width-integrated shear stress response is $\tau_{tg}\, \propto \, Q^{\frac{4}{5}}$, similar to subaerial channels.
-Hydraulics could trend toward this equilibrium regime with average discharge over $1-5$ days, when variability in subglacial channels  approaches that of subaerial channels (Figure~\ref{fig:multi_run}).
+Hydraulics could trend toward this equilibrium regime with average discharge over $1-5$ days when variability in subglacial channels  approaches that of subaerial channels (Figure~\ref{fig:multi_run}).
 Furthermore, the stronger correlation between hydraulics and water discharge in \icesheet{} suggests that this system is closer to this equilibrium compared to \alpine{}, given the reduced diurnal discharge amplitude in \icesheet{}.
 
-\ian{mauro: should a supplement be made for this simplification? if so, could you send me the one for R-channels? is this essentially equation 14 in werder et al., 2010?}
+\ian{mauro: should a supplement be made for this simplification? if so, could you send me the one for R-channels? Is this essentially equation 14 in Werder et al., 2010?}
 

### 2023-09-14 Delaneyetal2023b.tex
not as many changes as I would like... still need some work for Alley 1997

@@
 As a result, the subglacial channel size only evolves relatively slowly over days to months, whereas water discharge can vary over hours \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
-Therefore, subglacial water flow behaves more like pipe flow, and changes in water discharge $Q$ are mainly accommodated with increased water velocity $v$ \citep[e.g.][]{swift2005}, because the size of the channel $S$ changes on a much longer timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
+Therefore, subglacial water flow behaves more like pipe flow, and changes in water discharge $Q$ are mainly accommodated with increased water velocity $v$ \citep[][]{alley1997}, because the size of the channel $S$ changes on a much longer timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
 It follows that an increase in water discharge $Q$ will cause the shear stress across the channel $\tau_t$ to respond with water velocity $v$, as opposed to subaerial channels where channel width $w_c$ evolves as well.
@@
 As a result, sediment mobilization in subaerial and subglacial channels responds differently to changing water discharge.
-These differences have been implicitly included in a wide range of available models that  quantify sediment transport from catchments in both subglacial and subaerial channels  \citep[e.g.][]{walder1994,tucker1997,creyts2013,wickert2019,hewitt2019}.
+These differences have been implicitly included in a wide range of available models that  quantify sediment transport from catchments in both subglacial and subaerial channels  \citep[e.g.][]{walder1994,alley1997,tucker1997,creyts2013,wickert2019,hewitt2019}.
 The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \citep[e.g.][]{lane2016,perolo2018}.
 Additionally, the variable response of sediment transport capacity to water discharge in the two systems may affect the interpretation of sediment transport records in glacial systems \citep[e.g.][]{muller1968,richards2003,swift2005,ganti2016}.
-Yet, the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial channels, and implications for sediment discharge records, has been minimally discussed.
-Furthermore, thorough observations of water velocity and channel width in subglacial channels are limited and, to these authors' knowledge, not contrasted with subaerial conditions at the glacier margins.
+Yet, little work has examined the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial channel \citep{alley1997}.
+%Furthermore, thorough observations of water velocity and channel width in subglacial channels are limited and, to these authors' knowledge, not contrasted with subaerial conditions at the glacier margins.
 %\mauro{self-cite here?}
@@
 This data is available from \citet{tedstone2017} at a $5$\,\unit{min} time interval of the 2012 season.
-Note that in some cases the removal and relocation of sensors created variations in the sediment concentration not resulting from natural processes.
+Note that in some cases the removal and relocation of sensors created variations in the sediment concentration not resulting from natural processes (see Supplement xxx).
 % was presented in part in \citet{cowton2012} and
@@
 Outputs and observations are aggregated across timescales from $1$\,\unit{hr} to $15$\,\unit{days} from the $15$\,\unit{min} outputs, by averaging values one half of the timescale before and after a point time.
-%\mauro{Here a brief explanation of the rank-stuff would be good.  Also describe the time aggregation.}
 
@@
 By applying the discharge - width scaling in Equation~\ref{eq:wcf} to Equation~\ref{eq:v}, by setting $S\, \propto \,Q^{\frac{1}{2}}$, the subaerial channel's width-integrated shear stress response is $\tau_{tf}\, \propto\,  Q$.
-For sufficiently short time periods,  channel width and wetted area are fixed in the subglacial system so that the response of width-integrated shear stress is $ \tau_{tg}\, \propto\,  Q^{2}$, with a larger exponent on $Q$ than in the subaerial case.
-In this case, when water discharge fluctuates at a shorter timescale than the subglacial channel evolution \citep{gimbert2016}, the width-integrated shear stress varies more with water discharge than in subaerial channels.
+For sufficiently short time periods,  channel width and wetted area are fixed in the subglacial system so that the response of width-integrated shear stress is $ \tau_{tg}\, \propto\,  Q^{2}$, with a larger exponent on $Q$ than in the subaerial case \citep{alley1997}.
+In this case, when water discharge fluctuates at a shorter timescale than the subglacial channel evolution \citep{gimbert2016}
 Note also that the subglacial shear-stress response to water discharge will be greater than the subaerial's provided that the exponent is below one in Equation~\ref{eq:wcf}.
@@
 
-\ian{mauro: should a supplement be made for this simplification? if so, could you send me the one for R-channels? Is this essentially equation 14 in Werder et al., 2010?}
+\ian{mauro: I think a supplement be made for this.  could you send me the one for R-channels? Is this essentially equation 14 in Werder et al., 2010?}
 

### 2023-09-21 Delaneyetal2023b.tex
figure update and formating for GRL

@@
-\documentclass[11pt]{article}
-\usepackage [colon,square]{natbib}
-\usepackage[mathlines]{lineno} %% <- mathlines turns on line numbering in equations
-\usepackage{amsmath}
-\usepackage{printlen}
-\setlength{\parindent}{10pt}
-\setlength{\parskip}{10pt}
-\usepackage[explicit]{titlesec}
-\usepackage{helvet}
-\usepackage{array}
-\usepackage{printlen}
-\usepackage[a4paper, margin=1.3cm]{geometry}
-\usepackage{color,hyperref}
-\catcode`\_=11\relax
-\newcommand\email[1]{\_email #1\q_nil}
-\def\_email#1@#2\q_nil{%
-  \href{mailto:#1@#2}{{\emailfont #1\emailampersat #2}}
-}
-\newcommand\emailfont{\sffamily}
-\newcommand\emailampersat{{\color{red}\small@}}
-\catcode`\_=8\relax
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
+%
+% This file includes commands and instructions
+% given in the order necessary to produce a final output that will
+% satisfy AGU requirements, including customized APA reference formatting.
+%
+% You may copy this file and give it your
+% article name, and enter your text.
+%
+% guidelines and troubleshooting are here:
 
-\usepackage{hyperref}
-\usepackage[export]{adjustbox}
-\usepackage{bibentry}
-\renewcommand{\bibfont}{\footnotesize}
-\usepackage{tikz}
-\usepackage{setspace}
-\usepackage[inline,shortlabels]{enumitem}
-\usepackage[all]{nowidow}
-\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
-\pagenumbering{gobble}
-\hypersetup{ % Remove ugly boxes around URLs
-  colorlinks, linkcolor={black!100!black}, citecolor={black!50!black}, urlcolor={black!80!black}
-}
+%% To submit your paper:
+\documentclass[draft]{agujournal2019}
+\usepackage{url} %this package should fix any errors with URLs in refs.
+\usepackage{lineno}
+\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
+\usepackage{soul}
+\linenumbers
+%%%%%%%
+% As of 2018 we recommend use of the TrackChanges package to mark revisions.
+% The trackchanges package adds five new LaTeX commands:
+%
+%  \note[editor]{The note}
+%  \annote[editor]{Text to annotate}{The note}
+%  \add[editor]{Text to add}
+%  \remove[editor]{Text to remove}
+%  \change[editor]{Text to remove}{Text to add}
+%
+% complete documentation is here: http://trackchanges.sourceforge.net/
+%%%%%%%
 
-\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
-\newcolumntype{C}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}
-
-\linenumbers
-\sloppy
-\hbadness=99999
+\draftfalse
 
@@
 
+\journalname{Geophysical Research Letters}
+
 \begin{document}
-%% ------------------------------------------------------------------------ %%
-% Title
-%
-% (A title should be specific, informative, and brief. Use
-% abbreviations only if they are defined in the abstract. Titles that
-% start with general keywords then specific terms are optimized in
-% searches)
-%
-%% ------------------------------------------------------------------------ %%
 
-% Example: \title{This is a test title}
+\title{Water discharge quantity has a variable impact on sediment transport capacity in subglacial channels}
 
-\begin{center}
-  \Large{\textbf{Water discharge quantity has a variable impact on sediment transport capacity in subglacial channels}}
-  \normalsize
+\authors{Ian Delaney\affil{1},  Daniel Farinotti\affil{2,3},  Andrew J. Tedstone\affil{4},   Mauro A. Werder\affil{2,3}}
 
-  Ian Delaney\footnote{Institut des dynamiques de la surface terrestre (IDYST), Universit\'{e} de Lausanne, B\^{a}timent G\'{e}opolis, 1015 Lausanne, Switzerland
+\affiliation{1}{Institut des dynamiques de la surface terrestre (IDYST), Universit\'{e} de Lausanne, B\^{a}timent G\'{e}opolis, 1015 Lausanne, Switzerland}
+\affiliation{2}{Laboratory of Hydraulics, Hydrology and Glaciology (VAW), ETH-Z\"urich, H\"onggerbergring 26, 8093 89 Z\"urich, Switzerland}
+\affiliation{3}{Swiss Federal Institute for Forest, Snow and Landscape Research (WSL) Z\"uricherstrasse 111, 8903 1011 Birmensdorf, Switzerland}
+\affiliation{4}{Department of Geosciences, University of Fribourg, Ch. du Musée 1700, Fribourg, Switzerland}
 
-    ianarburua.delaney@unil.ch},
-  Andrew J. Tedstone\footnote{Department of Geosciences, University of Fribourg, Ch. du Musée 1700, Fribourg, Switzerland},
-   Mauro A. Werder$^{3,4}$,
-  Daniel Farinotti\footnote{Laboratory of Hydraulics, Hydrology and Glaciology (VAW), ETH-Z\"urich, H\"onggerbergring 26, 8093 89 Z\"urich, Switzerland}$^{,}$\footnote{Swiss Federal Institute for Forest, Snow and Landscape Research (WSL) Z\"uricherstrasse 111, 8903 1011 Birmensdorf, Switzerland}
+\correspondingauthor{Ian Delaney}{ianarburua.delaney@unil.ch}
 
-\end{center}
-
-\paragraph{Keypoints}
-\begin{itemize}
+\begin{keypoints}
 \item Water discharge variations in subglacial channels are mainly accommodated by water velocity, not channel size, given their subglacial channel's evolution.
-
 \item Greater variability in water velocity in subglacial channels causes sediment transport capacity to vary more than in subaerial channels.
-
 \item In interpreting sediment transport records, water discharge may represent processes including proglacial sediment mobilization and sediment access driving variations in sediment discharge.
-\end{itemize}
+\end{keypoints}
 
 \begin{abstract}
-
-  \noindent
-  In both subglacial and subaerial channels, the sediment transport capacity for sediment size depends mainly on 1) the channel width over which the sediment is mobilized  and 2) water velocity, as this controls the shear stress that the flowing water applies to the channel bottom.
-  In subaerial channels, changing water discharge  changes the water depth,  channel width, and velocity.
-  However, in subglacial environments and on time scales of hours to weeks, changing water discharge  mainly alters the  water velocity, not channel area or width, because the subglacial channel maintains a largely fixed geometry.
-  Here, parameterizations of sediment transport capacity for pressurized flow under glaciers and for subaerial open channel flow are applied to hydrographs from an Alpine glacier and the Greenland Ice Sheet.
-  Results of these parameterizations show that sediment transport capacity varies far more in subglacial channels than subaerial ones over time and with respect to water discharge.
-  Furthermore, high subglacial sediment transport capacity can persist across a wide range of water discharges.
-  However, slowly evolving  channel width can accommodate some  variability in sediment transport capacity in subglacial channels.
-  Despite the different responses of sediment transport capacity to water discharge in the two-channel types, measurements of sediment discharge from the catchments generally correlate better with subaerial hydraulics.
-  This suggests that the sediment discharge records represent other processes such as sediment access or subaerial processes in the proglacial area.
-  The different responses of water discharge on sediment transport capacity in subglacial channels, compared to subaerial ones, increase the occurrence of sediment transport capacity reaching mobilization thresholds, not necessarily related to water discharge quantity.
-
+  Sediment transport in subglacial and subaerial channels depends on  water velocity and channel width, which controls the shear stress exerted on the channel bottom and the space over which to integrate it.
+  In subaerial channels, depth, width, and velocity evolve quickly with water discharge, while in subglacial channels, water discharge changes mainly affect water velocity due to slow-evolving channel geometry.
+  Parameterizations applied to hydrographs from an Alpine glacier and the Greenland Ice Sheet reveal greater variability in sediment transport capacity in subglacial channels.
+  Furthermore, results show that high subglacial sediment transport rates can persist across various water discharges.
+  Slowly evolving channel width accommodates some variability in subglacial channels.
+  Despite differences in response to water discharge, sediment discharge measurements generally correlate better with subaerial hydraulics.
+  This implies that other factors like sediment access or proglacial processes might be at play, affecting sediment transport capacity differently in subglacial and subaerial channels.
 \end{abstract}
@@
 
-Changes in glacier dynamics, geomorphology, and hydrology  have prompted numerous  recent studies of  sediment transport processes in cold regions \citep[e.g.][]{zhang2022}.
-Increases in sediment transport  have been observed in Greenland \citep{bendixen2017}, the European Alps \citep{costa2017}, the Himalayas \citep{li2021}, and the Andes \citep{vergara2022}.
-In some regions, increased water discharge and glacier melt have been interpreted to yield greater sediment transport capacity \citep{bendixen2017,costa2017,li2021}.
-Observed changes to sediment transport in glacierized catchments require examining the processes controlling sediment discharge in these catchments and its variations with water discharge  \citep[e.g.][]{riihimaki2005,swift2005}.
+Changes in glacier dynamics, geomorphology, and hydrology  have prompted numerous  recent studies of  sediment transport processes in cold regions \cite<e.g.>{zhang2022}.
+Increases in sediment transport  have been observed in Greenland \cite{bendixen2017}, the European Alps \cite{costa2017}, the Himalayas \cite{li2021}, and the Andes \cite{vergara2022}.
+In some regions, increased water discharge and glacier melt have been interpreted to yield greater sediment transport capacity \cite{bendixen2017,costa2017,li2021}.
+Observed changes to sediment transport in glacierized catchments require examining the processes controlling sediment discharge in these catchments and its variations with water discharge  \cite<e.g.>{riihimaki2005,swift2005}.
 
-Over long periods, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from under glaciers \citep[c.f.][]{hallet1979,iverson2012,ugelvig2018}.
-On shorter time periods, pressurized water transports sediment from the under glacier \citep{walder1994,creyts2013,beaud2018,delaney2019}, should enough sediment be present subglacially (i.e.
+Over long periods, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from under glaciers \cite<c.f.>{hallet1979,iverson2012,ugelvig2018}.
+On shorter time periods, pressurized water transports sediment from the under glacier \cite{walder1994,creyts2013,beaud2018,delaney2019}, should enough sediment be present subglacially (i.e.
 in a transport-limited regime).
@@
 In a transport-limited regime, sediment discharge responds to the sediment transport capacity defined as the amount of sediment the water can carry.
-Sediment transport capacity depends on the shear stress between water and sediment it flows  over \citep{shields1936,meyer1948,engelund1967} and the width of the channel bottom $w_c$ over which to mobilize sediment.
+Sediment transport capacity depends on the shear stress between water and sediment it flows  over \cite{shields1936,meyer1948,engelund1967} and the width of the channel bottom $w_c$ over which to mobilize sediment.
 The shear stress  across the channel width $\tau_t$ responds to the velocity of water $v$ flowing through the channel so that
@@
 \end{linenomath*}
-where $Q$ is water discharge,  and $S$ is the channel's wetted area. %\mauro{I think ``wetted area'' is better and less ambiguous, changed throughout.}
-% \mauro{later $Q$ is used.  drop $_w$?}
+where $Q$ is water discharge,  and $S$ is the channel's wetted area.
 
-In subaerial channels, operating with open channel flow, the wetted area $S$ of the channel evolves with changing water discharge $Q$, by changing both the channel's width and the water depth \citep{leopold1953}.
+In subaerial channels, operating with open channel flow, the wetted area $S$ of the channel evolves with changing water discharge $Q$, by changing both the channel's width and the water depth \cite{leopold1953}.
 The change in water depth results in a commensurate increase in water velocity.
@@
 The response of water velocity to changing water discharge in subglacial channels differs from subaerial ones.
-The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by frictional heating of water flowing through the channel \citep{rothlisberger1972}.
-As a result, the subglacial channel size only evolves relatively slowly over days to months, whereas water discharge can vary over hours \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
-Therefore, subglacial water flow behaves more like pipe flow, and changes in water discharge $Q$ are mainly accommodated with increased water velocity $v$ \citep[][]{alley1997}, because the size of the channel $S$ changes on a much longer timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
+The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by frictional heating of water flowing through the channel \cite{rothlisberger1972}.
+As a result, the subglacial channel size only evolves relatively slowly over days to months, whereas water discharge can vary over hours \cite<e.g.>{iken1986,andrews2014,nanni2020}.
+Therefore, subglacial water flow behaves more like pipe flow, and changes in water discharge $Q$ are mainly accommodated with increased water velocity $v$ \cite{alley1997}, because the size of the channel $S$ changes on a much longer timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
 It follows that an increase in water discharge $Q$ will cause the shear stress across the channel $\tau_t$ to respond with water velocity $v$, as opposed to subaerial channels where channel width $w_c$ evolves as well.
@@
 As a result, sediment mobilization in subaerial and subglacial channels responds differently to changing water discharge.
-These differences have been implicitly included in a wide range of available models that  quantify sediment transport from catchments in both subglacial and subaerial channels  \citep[e.g.][]{walder1994,alley1997,tucker1997,creyts2013,wickert2019,hewitt2019}.
-The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \citep[e.g.][]{lane2016,perolo2018}.
-Additionally, the variable response of sediment transport capacity to water discharge in the two systems may affect the interpretation of sediment transport records in glacial systems \citep[e.g.][]{muller1968,richards2003,swift2005,ganti2016}.
-Yet, little work has examined the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial channel \citep{alley1997}.
+These differences have been implicitly included in a wide range of available models that  quantify sediment transport from catchments in both subglacial and subaerial channels  \cite<e.g.>{walder1994,alley1997,tucker1997,creyts2013,delaney2019,hewitt2019,wickert2019}.
+The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \cite<e.g.>{lane2016,perolo2018}.
+Additionally, the variable response of sediment transport capacity to water discharge in the two systems may affect the interpretation of sediment transport records in glacial systems \cite<e.g.>{muller1968,richards2003,swift2005,ganti2016}.
+Yet, little work has examined the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial channel \cite{alley1997}.
 %Furthermore, thorough observations of water velocity and channel width in subglacial channels are limited and, to these authors' knowledge, not contrasted with subaerial conditions at the glacier margins.
@@
 Water discharge and suspended sediment concentration were collected here at a $1$\unit{min} interval continuously over the period 2014--2021.
-See \citet{felix2022} and \citet{felix2021} for more details.
+See \citeA{felix2022} and \citeA{felix2021} for more details.
 
@@
 Water discharge and suspended sediment concentration data were collected roughly $2$\unit{km} downstream from the terminus from June to August/September 2009 to 2012 ($67^\circ\,03'\,5''$ N, $50^\circ\,12'\,59''$ W).
-This data is available from \citet{tedstone2017} at a $5$\,\unit{min} time interval of the 2012 season.
-Note that in some cases the removal and relocation of sensors created variations in the sediment concentration not resulting from natural processes (see Supplement xxx).
-% was presented in part in \citet{cowton2012} and
+Data is used from \citeA{tedstone2017} and \citeA{hawkings2014} at a $5$\,\unit{min} time interval from May 28, 2012 to August 8, 2012.
+Two gaps in suspended sediment data of roughly day in length are interpolated across with a linear spline.
 
@@
 The parameterizations below (Sections ~\ref{sect:sub_mode}~and~\ref{sect:fluv}) represent relationships amongst water discharge, water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
-Both parameterizations calculate   water velocity, shear stress, and width-integrated shear stress, upon which sediment transport depends \citep[Figure \ref{fig:cartoon}; ][]{shields1936}.
-The evaluation of these variables omits the selection of a sediment transport relationship and grain-size parameter \citep[e.g.][]{shields1936,meyer1948} needed to calculate sediment transport capacity from shear stress.
+Both parameterizations calculate   water velocity, shear stress, and width-integrated shear stress, upon which sediment transport depends \cite<Figure \ref{fig:cartoon}; >{shields1936}.
+The evaluation of these variables omits the selection of a sediment transport relationship and grain-size parameter \cite<e.g.>{shields1936,meyer1948} needed to calculate sediment transport capacity from shear stress.
 
-\ian{values of f are still high}
-\begin{table}[ht]
+\subsection{Subglacial channel  parameterization}
+\label{sect:sub_mode}
+
+To evaluate the shear stress of water flowing across sediments underneath a glacier, the subglacial channel parameterization takes into account the channel geometry and the velocity of the flowing water.
+To do this, a modified version of the lumped hydraulics model presented in \citeA{clarke1996} and \citeA{werder2010} is used.
+
+Here, it is assumed that the water is transported through a subglacial channel flowing down the glacier; \cite<Figure~\ref{fig:cartoon}; >{rothlisberger1972}, and that the channel  size responds to frictional heating from water flow and creep closure by ice, using the Darcy-Weisbach formulation for water-flow through a pipe  \cite<e.g.>{rothlisberger1972,clarke2003}.
+The formulation here does not consider the englacial storage of water.
+Thus, it is assumed that the changing hydraulic head at the top of the glacier is negligible in the channel size evolution.
+The evolution of subglacial channel size $S_g$ is given as
+\begin{linenomath*}
+  \begin{equation}
+    \label{eq:dS_dt}
+    \frac{\partial S_g}{\partial t} = C_1 \frac{Q \Delta h}{l} - C_2 \left(h_{o}-\frac{\Delta h}{2}\right)^n\,S_g,
+  \end{equation}
+\end{linenomath*}
+\noindent where $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (Table \ref{table:vpm}), $l$ is the length of subglacial channel, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change from the glacier terminus to its top, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the ice overburden pressure, and $n$ is Glen's n \cite{glen1955} (usually $n=3$).
+The first term on the right side of the equation represents the opening of the channel through frictional heating, while the following term represents the creep closure of the channel from ice deformation.
+
+Following the Darcy-Weisbach, head drop $\Delta h$ is,
+\begin{linenomath*}
+  \begin{equation}
+    \label{eq:dh}
+    \Delta h \,  = l \,\frac{1}{g}\,s\,f_r\,\frac{Q^2}{D_h^5}
+  \end{equation}
+\end{linenomath*}
+\noindent where $f_r$ is a friction factor, $D_h$ is the hydraulic diameter, and $l$ is the length overwhich head change occurs.
+$s$ is a factor accounting for channel geometry \cite{hooke1990}, calculated as
+\begin{equation}
+  \label{eq:Hf}
+  s = \frac{2\,(\beta -\sin \beta)^2}{(\frac{\beta}{2}\,+\,\sin \frac{\beta}{2})^4},
+\end{equation}
+where $\beta$ is the central angle of the circular segment that comprises the channel (the so-called Hooke angle). Note that $\beta =\pi$ corresponds to a semi-circle and
+smaller values of $\beta$ result in shallow, wide channels.
+The hydraulic diameter $D_h$ is converted to wetted area $S_g$ given a channel geometry following \cite{hooke1990}
+\begin{equation}
+  \label{eq:Dh2S}
+  S_g= \frac{D_h^2}{2}\,\frac{(\frac{\beta}{2}+\sin \frac{\beta}{2})^2}{\beta - \sin \beta},
+\end{equation}
+where $\beta$ is the central angle of the circular segment that comprises the channel (the so-called Hooke angle).
+
+With knowledge of wetted area $S_g$, the shear stress, $\tau_g$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
+\begin{equation}
+  \label{eq:tau}
+  \tau_g=\frac{1}{8}\,f_r\,\rho_w\,v_g^2,
+\end{equation}
+%
+where  the water velocity $v_g = \frac{Q}{S_g}$.
+
+The flat channel floor width $w_{cg}$, given angle $\beta$, is
+\begin{equation}
+  \label{eq:dh2wc}
+  w_{cg} = 2  \sin \frac{\beta}{2} \sqrt{\frac{2\, S_g}{\beta -\sin \beta}}.
+\end{equation}
+
+The width-integrated shear stress is represented as
+\begin{equation}
+  \label{eq:tautg}
+  \tau_{tg}=w_{cg}\,\tau_g.
+\end{equation}
+
+\subsection{Subaerial channel  parameterization}
+\label{sect:fluv}
+
+To parameterize the shear stress of water flowing across sediments in the subaerial channel,  the hydraulics parameterization presented in \citeA{tucker1997} is implemented.
+Using mass conservation and the Darcy-Weisbach relationship as well as assuming that the channel is wide so that the hydraulic radius is well approximated by the flow depth, then
+the shear stress $\tau_f$ at the river bed is
+\begin{linenomath*}
+  \begin{equation}
+    \label{eq:DW_tau}
+    \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q}{w_{cf}} \Big)^{\frac{2}{3}} \,\nabla z_c^{\frac{2}{3}},
+  \end{equation}
+\end{linenomath*}
+where $\nabla z_c$ is the channel slope, and $f_p$ is the friction factor for subaerial channels.
+Channel width $w_{cf}$ is
+\begin{equation}
+  \label{eq:wcf}
+  w_{cf} = k \, Q^\frac{1}{2},
+\end{equation}
+%
+where $k$ is a constant and the exponent is commonly given as $\frac{1}{2}$ \cite{leopold1953}.
+%\mauro{This is for so called ``regime channels''}
+Water velocity, $v_f$, is given by rearranging Equation \ref{eq:tau} as
+\begin{equation}
+  \label{eq:vf}
+  v_f = \sqrt{\frac{8\,\tau_f}{f_p\,\rho_w}}.
+\end{equation}
+%
+As above, the width-integrated shear stress is
+\begin{equation}
+    \label{eq:tautf}
+    \tau_{tf}=w_{cf}\,\tau_f.
+  \end{equation}
+
+\subsection{Implementation}
+\label{sect:imp}
+
+The parameterizations above are applied to hydrological records from the Fieschergletscher (\alpine) and the Leverett glacier (\icesheet).
+The subglacial and subaerial parameterizations represent the subglacial flow of water exiting a glacier and transitioning to subaerial flow, as it moves through the catchment.
+Outputs of the parameterizations are meant to represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
+To generalize these cases, \alpine{}  is exemplified by relatively thin ice thickness ($h_{ice}$: $225$\,\unit{m}), low water discharge ($\sim\,10$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
+\icesheet{}  is exemplified by thick ice  ($h_{ice}$: $700$\,\unit{m}), high water discharge ($\sim\,300$\,\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).
+
+The parameterizations are first applied to a reference test case for each glacier that assumes  subglacial channel with $\beta=\frac{\pi}{2}$ and friction factors $f_r$ and $f_p$ are tuned so that the parameterizations reproduce reasonable velocities for both  the Greenland Ice Sheet and Alpine glaciers \cite<$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$>{werder2010b,chandler2013}.
+In both subaerial channels, the bed slope, $\nabla z_c$ is $0.05$.
+Water velocity, shear stress, and width-integrated shear stress in the subglacial and subaerial channels from this formulation are presented below.
+
+Second, to characterize the variability in sediment discharge capacity in subglacial compared to subaerial channels further, $100$  parameterizations are applied to a range of different channel geometry factors ($\beta$ and $k$) and friction factors ($f_r$ and $f_p$).
+Runs with parameter combinations are culled if their mean subglacial water velocity over the season lies outside of the range between $0.75$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $2$\,\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity lies outside of the range $0.3$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with some measurements \cite<e.g.>{werder2010b,magnusson2012,chandler2013}.
+The standard deviation of an output variable over the course of the season is used to quantify the variability of that output.
+
+Lastly,  outputs are compared with the sediment discharge records the glaciers using Spearman rank correlation, or the correlation with respect to the ordering of values from the outputs and observations.
+Rank correlation reduces the impact of the non-linear response of sediment transport capacity to the hydrological forcing.
+Periods of uncertainty in the data in \icesheet are removed from the analysis.
+Outputs and observations are aggregated across timescales from $1$\,\unit{hr} to $15$\,\unit{days} from the $15$\,\unit{min} outputs, by averaging values one half of the timescale before and after a point time.
+
+\section{Results}
+
+\subsection{Role of water discharge in sediment transport capacity}
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.9\linewidth]{Fig2.pdf}
+  \caption{Parameterization outputs resulting from the hydrographs shown in panels a and e for the cases called \alpine (a-d) and \icesheet{} (e-h).
+Gray (black) lines in a and e represent the subglacial channel size (water discharge).
+In other panels, blue lines represent outputs from the subglacial channel, while orange lines represent the subaerial channel.
+    Data in plots are at $15$\unit{min} intervals.
+    Arrows denote some events where the variable peaks in the subglacial channel prior to the subaerial one.
+    Insets in e-h show the  peak melt event \icesheet{} denoted by the shaded area.
+    Note that the x and y axes are different for \alpine{} and \icesheet{}.
+      }
+  \label{fig:model_outs}
+\end{figure}
+Following the parameterizations defined in Sections~\ref{sect:sub_mode}~and~\ref{sect:fluv}, the water velocity, shear stress, and width-integrated shear stress exhibit different seasonal evolutions and peaks.
+Water velocity and shear stress represent the potential for sediment mobilization, and width-integrated shear stress represents a proxy for the total sediment transport capacity across the channel bed (Figure~\ref{fig:model_outs}).
+
+Over the course of the season, large peaks in water velocity and shear stress, along with width-integrated shear stress, occur in response to increases in water discharge in both \alpine{} and \icesheet{}.
+This occurs as the subglacial values are generally larger than the subaerial values, thus more efficient at transporting sediment.
+In the subglacial channel, peaks generally occur over the period when water discharge increases at the fastest rate (Figure~\ref{fig:model_outs}).
+Conversely, in the subaerial channel, peaks in these sediment transport variables occur when water discharge is highest.
+The timing of the greatest rate of change in water discharge occurs before the peaks in water discharge.
+Here, sediment could be deposited in the proglacial area as the water depressurizes, then remobilized when sediment transport capacity peaks in the subaerial system.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.9\linewidth]{Fig3.png}
+  \caption{Relationship between water discharge and velocity, shear stress, and width-integrated shear stress for \alpine{} (a-c) and \icesheet (d-f).
+    These outputs' relationship with measured sediment discharge is given for \alpine in (g-i) and \icesheet (j-l).
+    Variables have been normalized to mean values for better comparison amongst the runs.
+    $R_g$ ($R_f$) shows the Spearman rank correlation values for the subglacial (subaerial) outputs.
+    Plots are shown with data and outputs at $15$\unit{min} interval.
+  }
+  \label{fig:Qw_vari}
+\end{figure}
+
+Each water discharge value in the subaerial channel results in a unique water velocity, shear stress, and width-integrated shear stress, creating a perfect rank correlation between the variables (Figure~\ref{fig:Qw_vari} a--f).
+Conversely in the subglacial channel,  water velocity, shear stress, and width-integrated shear stress can vary substantially for a given water discharge and width-integrated shear stress varies with respect to the other two  outputs as well (Figure~\ref{fig:Qw_vari} a--f).
+For instance, very high  water velocity values and shear stresses can occur across minimal water discharge at $2.5$\,\unit{m}$^3$\,\unit{s}$^{-1}$ to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
+In \icesheet, mean values of water velocity can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, effectively spanning much of the water discharge range (Figure~\ref{fig:Qw_vari} d--f).
+
+When considering the subglacial channel's evolving width, width-integrated shear stress  across the channel generally increases with water discharge, with improved rank correlation compared to water discharge and water velocity or shear stress (Figure~\ref{fig:Qw_vari} a--f).
+Yet even the width-integrated shear stress  can vary substantially, with the highest values occurring at water discharge values ranging from roughly $11$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, to over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
+The variability in width-integrated shear stress is less pronounced in the \icesheet case, with reduced variability in water discharge (Figure~\ref{fig:Qw_vari} c, f).
+Yet, a range of width-integrated shear stress can occur at a given discharge.
+
+\begin{figure}[h]
+  \centering
+    \includegraphics[width=0.9\linewidth]{Fig4.png}
+    \caption{Variance (measured by the standard deviation) in water velocity, shear stress, and width integrated shear stress for different time aggregations across the range of subglacial and subaerial channel shapes and friction factors.
+      Shaded areas denote the minimum and maximum of the $100$ parameter combinations tested for each of the systems (Section~\ref{sect:imp}).
+      Solid lines denote  mean values.
+      Top row shows results from \alpine{} case, while bottom row show outputs from \icesheet{} case
+      Time intervals shown with markers denote $15$\,\unit{min}, $1$\,\unit{hr}, $12$\,\unit{hr}, $1$\,\unit{d}, $5$\,\unit{d}, $10$\,\unit{d}, and $15$\,\unit{d} aggregation periods.
+      Text in columns shows rank correlation between the variable and sediment discharge for the corresponding time period for the subglacial (right column) and subaerial (left column) channels.
+    }
+    \label{fig:multi_run}
+  \end{figure}
+
+The second application examines the impact of a variety of channel shapes and friction factors on the relationship between water discharge and sediment transport capacity, in both systems.
+Across the range of parameter values examined, variability in water velocity and shear stress remains higher in the subglacial system compared to the subaerial one.
+However, in \alpine{}, the variability in width-integrated shear stress is comparable between the two systems, due to the variable water discharge (Figure~\ref{fig:multi_run}).
+Conversely, in \icesheet{}, with lower water discharge variability, the variability in width-integrated shear stress remains higher in the subglacial system across the time aggregation periods.
+
+It is prescient to evaluate the time scales over which the subglacial channels may respond to changing water discharge to understand the influence of variability in sediment transport capacity.
+In both cases, the variability in velocity, shear stress, and width-integrated shear stress decrease after $1-5$ day aggregations (Figure~\ref{fig:multi_run}).
+Yet, substantial differences in variability persist in periods of up to $15$ days, as the variability in aggregations periods longer than this might be conflated with seasonal variations in water discharge.
+Note that lags and time of peak sediment transport conditions still occur over these periods (Figure~\ref{fig:model_outs}), while the variability in sediment transport parameters decreases with larger time aggregation (Figure~\ref{fig:multi_run}).
+
+\subsection{Relationship between parameterization  outputs and observed sediment discharge}
+When comparing the output variables with suspended sediment discharge measurements, velocity, shear stress, and width-integrated shear stress from the subaerial system generally show better correlation with sediment discharge compared to subglacial outputs across the tested subglacial channel parameters (Figures~\ref{fig:Qw_vari} g--l and \ref{fig:multi_run}).
+The rank correlation amongst the  subaerial parameterization's three output variables remains fairly constant despite the variable impact of channel shape and consistent scaling with water discharge (Figure \ref{fig:Qw_vari} and Section~\ref{sect:fluv}).
+In subglacial channels, velocity and shear stress perform substantially worse in terms of rank correlation with sediment discharge compared to width-integrated shear stress (Figures~ \ref{fig:multi_run}).
+In both \alpine{} and \icesheet{}, the  subglacial rank correlation approaches that of subaerial channels over longer time aggregation periods (Figures~ \ref{fig:multi_run} and \ref{fig:model_outs_1hr} to \ref{fig:model_outs_15day}).
+
+\section{Discussion}
+\subsection{Context of sediment capacity differences between the channel types}
+\label{sect:dis_qsc}
+Quantities that control sediment transport capacity, such as water velocity, shear stress, and width-integrated shear stress, vary differently in response to water discharge in subglacial channels compared to subaerial ones (Figure~\ref{fig:Qw_vari}).
+In sediment transport relationships, such as in \citeA{meyer1948} or \citeA{engelund1967}, that evaluate sediment discharge capacity, shear stress is scaled to the $\frac{3}{2}$ or $\frac{5}{2}$ power, respectively.
+The exponent greater than $1$ magnifies sediment discharge variability beyond the variable sediment transport parameters described above (Figure~\ref{fig:multi_run}).
+Note that sediment transport relationships describe sediment flux per unit width so that the integrated capacity across channel occurs after application of the exponent, unlike the width-integrated shear stress evaluation here \cite<Equations\,\ref{eq:tautg} and \ref{eq:tautf}; c.f.>{tucker1997}.
+Note also that both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress is homogeneous across the channel bed, despite the limitations of this assumption \cite<e.g.>{yager2018}.
+
+The rapid increases and decreases in subglacial sediment transport capacity may compound the variability in sediment transport already present in subaerial systems \cite<Figure~\ref{fig:Qw_vari}; >{williams1989,jerolmack2010}.
+By applying the discharge - width scaling in Equation~\ref{eq:wcf} to Equation~\ref{eq:v}, by setting $S\, \propto \,Q^{\frac{1}{2}}$, the subaerial channel's width-integrated shear stress response is $\tau_{tf}\, \propto\,  Q$.
+For sufficiently short time periods,  channel width and wetted area are fixed in the subglacial system so that the response of width-integrated shear stress is $ \tau_{tg}\, \propto\,  Q^{2}$, with a larger exponent on $Q$ than in the subaerial case following \citeA{alley1997}.
+In this case, water discharge fluctuates at a shorter timescale than the subglacial channel evolution \cite{gimbert2016}.
+Note also that the subglacial width integrated shear stress response to water discharge will be greater than the subaerial's provided that  Equation~\ref{eq:wcf}'s exponent is below one.
+However, when subglacial channel size lies in equilibrium with water discharge, the  width-integrated shear stress response is $\tau_{tg}\, \propto \, Q^{\frac{4}{5}}$, similar to subaerial channels.
+It follows that, the stronger correlation between hydraulics and water discharge in \icesheet{} could result from being closer to this equilibrium between variations in the channels wetter area and water discharge compared to \alpine{}.
+This state could result from the reduced diurnal discharge amplitude in \icesheet{} (Figures~\ref{fig:model_outs} and ~\ref{fig:Qw_vari}).
+In both cases, hydraulics could trend toward this equilibrium regime when average discharge over $1-5$ days when variability in subglacial channels  approaches that of subaerial channels (Figure~\ref{fig:multi_run}).
+
+In subglacial channels, peaks occur with the fastest rate of  water discharge increase relative to the channel's growth (Figure~\ref{fig:model_outs}).
+In turn, the subglacial channels here contain feedbacks that limit sediment transport capacity as the subglacial channel grows with increasing water discharge capacity, effectively reducing sediment transport capacity with the stabilization of water discharge.
+Similar feedback occurs in subaerial systems \cite{phillips2016}, albeit at longer timescales, as opposed to days here \cite{gimbert2016}.
+
+\subsection{Records of sediment discharge  from glaciers}
+
+The increased performance of subaerial hydraulics in representing observations occurs despite the proximity of glaciers to these measurements (Figure~\ref{fig:Qw_vari}).
+This could occur for two reasons.
+
+First, these measurement stations lie $0.5$ to $2$\,\unit{km} downstream of the glacier terminus \cite{cowton2012,felix2022}, likely a long enough distance for the pressurized subglacial regime to transition to the open channel subaerial one.
+The transition of subglacial to subaerial conditions may occur underneath glacier terminus \cite{perolo2018}, especially if closure rates of the ice channels are slow  \cite{rothlisberger1972}.
+Furthermore,  if a transport-limited regime persists at the glacier margin, then sediment can be easily mobilized, and subaerial processes could be represented just a short distance downstream.
+
+The consistent relationship between sediment discharge capacity and open channel flow conditions could be pronounced $\sim20$\,\unit{km} downstream of the Leverett site, where a strong correlation persists between sediment plume size and the river's water discharge into the Kangerlussuaq fjord \cite{mcgrath2010}.
+In marine-terminating glacier catchments, for contrast, a less consistent relationship may occur between water discharge or melt extent and sediment plume size \cite{tedstone2012}.
+Here, ocean water pressurizes subglacial channels at the ice front \cite<e.g.>{how2017}, so the minimal correlation could result from the inconsistent relationship between subglacial sediment transport capacity and water discharge (Figure~\ref{fig:Qw_vari}).
+
+Second, the sediment discharge records could represent processes such as increased sediment accessibility when melt extends up glacier, thereby concurrently increasing sediment and water discharge \cite<e.g.>{vergara2022}.
+A supply-limited regime could well persist at many glaciers, because the sediment's threshold of motion can be reached more frequently, across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
+Sediment exhaustion occurring through this process may also explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \cite{overeem2017}.
+Note that in the data here, however, the absence of lags between the observed sediment and water discharge may suggest that sediment is mobilized close to the measurement stations, as opposed to distal areas of the glacier bed as would be expected in a supply-limited regime \cite<Figure~\ref{fig:model_outs}\, a and e; >{williams1989}.
+
+Regardless of the increased performance of water discharge compared to subglacial sediment transport capacity in the observations, parameterization outputs suggest that additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to subaerial systems \cite{jansson2005,ganti2016}.
+Sediment discharge records have been used to establish the relationship, or lack thereof, between climate forcing and glacial erosion \cite<e.g.>{koppes2009a,ganti2016,willenbring2016,mariotti2021}.
+The variable relationship between water discharge and subglacial sediment transport capacity presented here mostly applies to the daily- to weekly-time scales controlling the size of subglacial channels (Figure~\ref{fig:multi_run}).
+Yet, identifying climatic signals in sediment transport from the transport-limited component of glacier-sedimentary systems may require higher thresholds of water discharge compared to subaerial systems when water discharge varies substantial over short timescales \cite<Figure~\ref{fig:Qw_vari}; >{tofelde2021}.
+
+Over time scales of weeks or longer, the glacier's sediment transport capacity responds to the ice thickness, controlling the channel closure rate, and the glacier's surface slope, in addition to water discharge \cite<Figure~\ref{fig:multi_run}, ~ Section~\ref{sect:sub_mode}; >{rothlisberger1972,shreve1972,gimbert2016,stevens2022}.
+In a supply-limited state, often assumed in landscape evolution models, then glaciers' sediment discharge record also represents additional processes of sediment availability and bedrock erosion from  water pressure variations and sliding  \cite<e.g.>{iverson2012,herman2015,delaney2023}.
+This multitude of processes lies in contrast to most subaerial systems, where transport capacity typically responds to sediment size, water discharge and hydraulic gradient, and the hydraulic gradient generally remains stable over time scales ranging from years to millennia \cite<Section~\ref{sect:fluv}; e.g.>{muller1968,tucker1997,wickert2019}.
+The stronger link between sediment transport capacity and hydrological forcing in subaerial channels may mean that variations in sediment transport records can be more easily attributed to climatic conditions than in subglacial systems.
+
+\section{Conclusions}
+
+Sediment transport capacity is largely driven by the shear stress of water flowing through a channel and the channel's width.
+Subaerial channels  alter their channel width and water velocity immediately in response to changing water discharge.
+In contrast, pressurized subglacial channels largely accommodate changing water discharge by altering water velocity and shear stress, upon which sediment transport depends.
+Water velocity responds quickly to changes in water discharge because the size of subglacial channels generally evolves slowly compared to variations in water discharge.
+This makes the response of sediment transport capacity much more sensitive to changes in water discharge in subglacial systems compared to subaerial ones. 
+Despite these differences, parameterizations of subaerial hydraulics in two glacierized catchments generally represent observed sediment discharge variations better than subglacial hydraulics.
+
+Implications of the findings here include:
+\begin{enumerate}
+\item the inconsistent relationship between water discharge and sediment transport capacity underneath glaciers creates increased variability in sediment transport capacity compared to river systems;
+\item the timing of peak sediment transport capacity below glaciers generally coincides with large rates of increasing in water discharge, as opposed to the maximum quantity of water discharge;
+\item a consistent relationship between water and sediment discharge may represent subglacial sediment access or open channel hydraulics near the glacier margin, instead of subglacial sediment transport capacity.
+
+\end{enumerate}
+
+\section*{Author Contributions}
+\ian{let me know if these accurately describe your contributions... I hope I am not leaving anything out}
+I.D. designed the study, developed and implement the model and experiments and lead the writing.
+D.F. provided data from Fieschergletscher.
+A.J.T. helped interpret data from Leverett Glacier and provided guidance on experiment design.
+M.A.W. provided data from Fieschergletscher, added the relationship between shear stress and water discharge (see Section~\ref{sect:dis_qsc} and Supplement).
+All authors provided  key inputs in writing and editing the manuscript.
+
+\section*{Open Research Section}
+Code to make the figures, with links to the data, can be found at \url{https://bitbucket.org/IanDelaney/xsection/src/master/}.
+Code and data will be uploaded to a permanent repository, with FAIR principles, pending acceptance of the paper.
+
+\acknowledgments
+
+SNSF Project No. PZ00P2\_202024 provided  funding for I. Delaney.
+We thank G. King for commenting on a previous version of this manuscript.
+
+\bibliography{PaperLib}
+
+\appendix
+\section{Table of variables, parameters and constants}
+\begin{table}[h]
   \centering
@@
 
-\subsection{Subglacial channel  parameterization}
-\label{sect:sub_mode}
+\newpage
 
-To evaluate the shear stress of water flowing across sediments underneath a glacier, the subglacial channel parameterization takes into account the channel geometry and the velocity of the flowing water.
-To do this, a modified version of the lumped hydraulics model presented in \citet{clarke1996} and \citet{werder2010} is used.
+\section{Figure 3 over various time aggregations}
 
-Here, it is assumed that the water is transported through a subglacial channel flowing down the glacier; \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972}, and that the channel  size responds to frictional heating from water flow and creep closure by ice, using the Darcy-Weisbach formulation for water-flow through a pipe  \citep[e.g.][]{rothlisberger1972,clarke2003}.
-The formulation here does not consider the englacial storage of water.
-Thus, it is assumed that the changing hydraulic head at the top of the glacier is negligible in the channel size evolution.
-The evolution of subglacial channel size $S_g$ is given as
-\begin{linenomath*}
-  \begin{equation}
-    \label{eq:dS_dt}
-    \frac{\partial S_g}{\partial t} = C_1 \frac{Q \Delta h}{l} - C_2 \left(h_{o}-\frac{\Delta h}{2}\right)^n\,S_g,
-  \end{equation}
-\end{linenomath*}
-\noindent where $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (Table \ref{table:vpm}), $l$ is the length of subglacial channel, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change from the glacier terminus to its top, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the ice overburden pressure, and $n$ is Glen's n \citep{glen1955} (usually $n=3$).
-The first term on the right side of the equation represents the opening of the channel through frictional heating, while the following term represents the creep closure of the channel from ice deformation.
-
-Following the Darcy-Weisbach, head drop $\Delta h$ is,
-\begin{linenomath*}
-  \begin{equation}
-    \label{eq:dh}
-    \Delta h \,  = l \,\frac{1}{g}\,s\,f_r\,\frac{Q^2}{D_h^5}
-  \end{equation}
-\end{linenomath*}
-\noindent where $f_r$ is a friction factor, $D_h$ is the hydraulic diameter, and $l$ is the length overwhich head change occurs.
-$s$ is a factor accounting for channel geometry \citep{hooke1990}, calculated as:
-\begin{equation}
-  \label{eq:Hf}
-  s = \frac{2\,(\beta -\sin \beta)^2}{(\frac{\beta}{2}\,+\,\sin \frac{\beta}{2})^4},
-\end{equation}
-where $\beta$ is the central angle of the circular segment that comprises the channel (the so-called Hooke angle). Note that $\beta =\pi$ corresponds to a semi-circle and
-smaller values of $\beta$ result in shallow, wide channels.
-The hydraulic diameter $D_h$ is converted to wetted area $S_g$ given a channel geometry following \citep{hooke1990}
-\begin{equation}
-  \label{eq:Dh2S}
-  S_g= \frac{D_h^2}{2}\,\frac{(\frac{\beta}{2}+\sin \frac{\beta}{2})^2}{\beta - \sin \beta},
-\end{equation}
-where $\beta$ is the central angle of the circular segment that comprises the channel (the so-called Hooke angle).
-
-With knowledge of wetted area $S_g$, the shear stress, $\tau_g$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
-\begin{equation}
-  \label{eq:tau}
-  \tau_g=\frac{1}{8}\,f_r\,\rho_w\,v_g^2,
-\end{equation}
-%
-where  the water velocity $v_g = \frac{Q}{S_g}$.
-
-The flat channel floor width $w_{cg}$, given angle $\beta$, is
-\begin{equation}
-  \label{eq:dh2wc}
-  w_{cg} = 2  \sin \frac{\beta}{2} \sqrt{\frac{2\, S_g}{\beta -\sin \beta}}.
-\end{equation}
-
-The width-integrated shear stress is represented as
-\begin{equation}
-  \label{eq:tautg}
-  \tau_{tg}=w_{cg}\,\tau_g.
-\end{equation}
-
-\subsection{Subaerial channel  parameterization}
-\label{sect:fluv}
-
-To parameterize the shear stress of water flowing across sediments in the subaerial channel,  the hydraulics parameterization presented in \citet{tucker1997} is implemented.
-Using mass conservation and the Darcy-Weisbach relationship as well as assuming that the channel is wide so that the hydraulic radius is well approximated by the flow depth, then
-the shear stress $\tau_f$ at the river bed is
-\begin{linenomath*}
-  \begin{equation}
-    \label{eq:DW_tau}
-    \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q}{w_{cf}} \Big)^{\frac{2}{3}} \,\nabla z_c^{\frac{2}{3}},
-  \end{equation}
-\end{linenomath*}
-where $\nabla z_c$ is the channel slope, and $f_p$ is the friction factor for subaerial channels.
-Channel width $w_{cf}$ is
-\begin{equation}
-  \label{eq:wcf}
-  w_{cf} = k \, Q^\frac{1}{2},
-\end{equation}
-%
-where $k$ is a constant and the exponent is commonly given as $\frac{1}{2}$ \citep{leopold1953}.
-%\mauro{This is for so called ``regime channels''}
-Water velocity, $v_f$, is given by rearranging Equation \ref{eq:tau} as
-\begin{equation}
-  \label{eq:vf}
-  v_f = \sqrt{\frac{8\,\tau_f}{f_p\,\rho_w}}.
-\end{equation}
-%
-As above, the width-integrated shear stress is
-\begin{equation}
-    \label{eq:tautf}
-    \tau_{tf}=w_{cf}\,\tau_f.
-\end{equation}
-\subsection{Implementation}
-\label{sect:imp}
-
-The parameterizations above are applied to hydrological records from the Fieschergletscher (\alpine) and the Leverett glacier (\icesheet).
-The subglacial and subaerial parameterizations represent the subglacial flow of water exiting a glacier and transitioning to subaerial flow, as it moves through the catchment.
-Outputs of the parameterizations are meant to represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
-To generalize these cases, \alpine{}  is exemplified by relatively thin ice thickness ($h_{ice}$: $225$\,\unit{m}), low water discharge ($\sim\,10$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
-\icesheet{}  is exemplified by thick ice  ($h_{ice}$: $700$\,\unit{m}), high water discharge ($\sim\,300$\,\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).
-
-The parameterizations are first applied to a reference test case for each glacier that assumes  subglacial channel with $\beta=\frac{\pi}{2}$ and friction factors $f_r$ and $f_p$ are tuned so that the parameterizations reproduce reasonable velocities for both  the Greenland Ice Sheet and Alpine glaciers \citep[$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$][]{werder2010b,chandler2013}.
-In both subaerial channels, the bed slope, $\nabla z_c$ is $0.05$.
-Water velocity, shear stress, and width-integrated shear stress in the subglacial and subaerial channels from this formulation are presented below.
-
-Second, to characterize the variability in sediment discharge capacity in subglacial compared to subaerial channels further, $100$  parameterizations are applied to a range of different channel geometry factors ($\beta$ and $k$) and friction factors ($f_r$ and $f_p$).
-Runs with parameter combinations are culled if their mean subglacial water velocity over the season lies outside of the range between $0.75$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $2$\,\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity lies outside of the range $0.3$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with some measurements \citep[e.g.][]{werder2010b,magnusson2012,chandler2013}.
-The standard deviation of an output variable over the course of the season is used to quantify the variability of that output.
-
-Lastly,  outputs are compared with the sediment discharge records the glaciers using Spearman rank correlation, or the correlation with respect to the ordering of values from the outputs and observations.
-Rank correlation reduces the impact of the non-linear response of sediment transport capacity to the hydrological forcing.
-Periods of uncertainty in the data in \icesheet are removed from the analysis.
-Outputs and observations are aggregated across timescales from $1$\,\unit{hr} to $15$\,\unit{days} from the $15$\,\unit{min} outputs, by averaging values one half of the timescale before and after a point time.
-
-\section{Results}
-
-\subsection{Role of water discharge in sediment transport capacity}
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=0.9\linewidth]{Fig2.pdf}
-  \caption{Parameterization outputs resulting from the hydrographs shown in panels a and e for the cases called \alpine (a-d) and \icesheet{} (e-h).
-Gray (black) lines in a and e represent the subglacial channel size (water discharge).
-In other panels, blue lines represent outputs from the subglacial channel, while orange lines represent the subaerial channel.
-    Data in plots are at $15$\unit{min} intervals.
-    Arrows denote some events where the variable peaks in the subglacial channel prior to the subaerial one.
-    Insets in e-h show the  peak melt event \icesheet{} denoted by the shaded area.
-    Note that the x and y axes are different for \alpine{} and \icesheet{} such that they each cover a 
-        \ian{NaN values will be removed in the final version}
-      }
-  \label{fig:model_outs}
-\end{figure}
-Following the parameterizations defined in Sections~\ref{sect:sub_mode}~and~\ref{sect:fluv}, the water velocity, shear stress, and width-integrated shear stress exhibit different seasonal evolutions and peaks.
-Water velocity and shear stress represent the potential for sediment mobilization, and width-integrated shear stress represents a proxy for the total sediment transport capacity across the channel bed (Figure~\ref{fig:model_outs}).
-
-Over the course of the season, large peaks in water velocity and shear stress, along with width-integrated shear stress, occur in response to increases in water discharge in both \alpine{} and \icesheet{}.
-This occurs as the subglacial values are generally larger than the subaerial values, thus more efficient at transporting sediment.
-In the subglacial channel, peaks generally occur over the period when water discharge increases at the fastest rate (Figure~\ref{fig:model_outs}).
-Conversely, in the subaerial channel, peaks in these sediment transport variables occur when water discharge is highest.
-The timing of the greatest rate of change in water discharge occurs before the peaks in water discharge.
-Here, sediment could be deposited in the proglacial area as the water depressurizes, then remobilized when sediment transport capacity peaks in the subaerial system.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=0.9\linewidth]{Fig3.png}
-  \caption{Relationship between water discharge and velocity, shear stress, and width-integrated shear stress for \alpine{} (a-c) and \icesheet (d-f).
-    These outputs' relationship with measured sediment discharge is given for \alpine in (g-i) and \icesheet (j-l).
-    Variables have been normalized to mean values for better comparison amongst the runs.
-    $R_g$ ($R_f$) shows the Spearman rank correlation values for the subglacial (subaerial) outputs.
-    Plots are shown with data and outputs at $15$\unit{min} interval.
-  }
-  \label{fig:Qw_vari}
-\end{figure}
-
-Each water discharge value in the subaerial channel results in a unique water velocity, shear stress, and width-integrated shear stress, creating a perfect rank correlation between the variables (Figure~\ref{fig:Qw_vari} a--f).
-Conversely in the subglacial channel,  water velocity, shear stress, and width-integrated shear stress can vary substantially for a given water discharge and width-integrated shear stress varies with respect to the other two  outputs as well (Figure~\ref{fig:Qw_vari} a--f).
-For instance, very high  water velocity values and shear stresses can occur across minimal water discharge at $2.5$\,\unit{m}$^3$\,\unit{s}$^{-1}$ to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
-In \icesheet, mean values of water velocity can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, effectively spanning much of the water discharge range (Figure~\ref{fig:Qw_vari} d--f).
-
-When considering the subglacial channel's evolving width, width-integrated shear stress  across the channel generally increases with water discharge, with improved rank correlation compared to water discharge and water velocity or shear stress (Figure~\ref{fig:Qw_vari} a--f).
-Yet even the width-integrated shear stress  can vary substantially, with the highest values occurring at water discharge values ranging from roughly $11$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, to over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
-The variability in width-integrated shear stress is less pronounced in the \icesheet case, with reduced variability in water discharge (Figure~\ref{fig:Qw_vari} c, f).
-Yet, a range of width-integrated shear stress can occur at a given discharge.
-
-\begin{figure}[h]
-  \centering
-    \includegraphics[width=0.9\linewidth]{Fig4.png}
-    \caption{Variance (measured by the standard deviation) in water velocity, shear stress, and width integrated shear stress for different time aggregations across the range of subglacial and subaerial channel shapes and friction factors.
-      Shaded areas denote the minimum and maximum of the $100$ parameter combinations tested for each of the systems (Section~\ref{sect:imp}).
-      Solid lines denote  mean values.
-      Top row shows results from \alpine{} case, while bottom row show outputs from \icesheet{} case
-      Time intervals shown with markers denote $15$\,\unit{min}, $1$\,\unit{hr}, $12$\,\unit{hr}, $1$\,\unit{d}, $5$\,\unit{d}, $10$\,\unit{d}, and $15$\,\unit{d} aggregation periods.
-      Text in columns shows rank correlation between the variable and sediment discharge for the corresponding time period for the subglacial (right column) and subaerial (left column) channels.
-      \ian{time aggregation ``smooths'' between data gaps here....}
-    }
-    \label{fig:multi_run}
-  \end{figure}
-
-The second application examines the impact of a variety of channel shapes and friction factors on the relationship between water discharge and sediment transport capacity, in both systems.
-Across the range of parameter values examined, variability in water velocity and shear stress remains higher in the subglacial system compared to the subaerial one.
-However, in \alpine{}, the variability in width-integrated shear stress is comparable between the two systems, due to the variable water discharge (Figure~\ref{fig:multi_run}).
-Conversely, in \icesheet{}, with lower water discharge variability, the variability in width-integrated shear stress remains higher in the subglacial system across the time aggregation periods.
-
-It is prescient to evaluate the time scales over which the subglacial channels may respond to changing water discharge to understand the influence of variability in sediment transport capacity.
-In both cases, the variability in velocity, shear stress, and width-integrated shear stress decrease after $1-5$ day aggregations (Figure~\ref{fig:multi_run}).
-Yet, substantial differences in variability persist in periods of up to $15$ days, as the variability in aggregations periods longer than this might be conflated with seasonal variations in water discharge.
-Note that lags and time of peak sediment transport conditions still occur over these periods (Figure~\ref{fig:model_outs}), while the variability in sediment transport parameters decreases with larger time aggregation (Figure~\ref{fig:multi_run}).
-
-\subsection{Relationship between parameterization  outputs and observed sediment discharge}
-When comparing the output variables with suspended sediment discharge measurements, velocity, shear stress, and width-integrated shear stress from the subaerial system generally show better correlation with sediment discharge compared to subglacial outputs across the tested subglacial channel parameters (Figures~\ref{fig:Qw_vari} g--l and \ref{fig:multi_run}).
-The rank correlation amongst the  subaerial parameterization's three output variables remains fairly constant despite the variable impact of channel shape and consistent scaling with water discharge (Figure \ref{fig:Qw_vari} and Section~\ref{sect:fluv}).
-In subglacial channels, velocity and shear stress perform substantially worse in terms of rank correlation with sediment discharge compared to width-integrated shear stress (Figures~ \ref{fig:multi_run}).
-In both \alpine{} and \icesheet{}, the  subglacial rank correlation approaches that of subaerial channels over longer time aggregation periods (Figures~ \ref{fig:multi_run}).
-
-\section{Discussion}
-\subsection{Context of sediment capacity differences between the channel types}
-Quantities that control sediment transport capacity, such as water velocity, shear stress, and width-integrated shear stress, vary differently in response to water discharge in subglacial channels compared to subaerial ones (Figure~\ref{fig:Qw_vari}).
-In sediment transport relationships, such as in \citet{meyer1948} or \citet{engelund1967}, that evaluate sediment discharge capacity, shear stress is scaled to the $\frac{3}{2}$ or $\frac{5}{2}$ power, respectively.
-The exponent greater than $1$ magnifies sediment discharge variability beyond the variable sediment transport parameters described above (Figure~\ref{fig:multi_run}).
-Note that sediment transport relationships describe sediment flux per unit width so that the integrated capacity across channel occurs after application of the exponent, unlike the width-integrated shear stress evaluation here \citep[Equations\,\ref{eq:tautg} and \ref{eq:tautf}; c.f.][]{tucker1997}.
-%Note also that both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress is homogeneous across the channel bed, despite the limitations of this assumption \citep[e.g.][]{yager2018}.
-
-The rapid increases and decreases in subglacial sediment transport capacity may compound the variability in sediment transport already present in subaerial systems \citep[Figure~\ref{fig:Qw_vari}; ][]{williams1989,jerolmack2010}.
-By applying the discharge - width scaling in Equation~\ref{eq:wcf} to Equation~\ref{eq:v}, by setting $S\, \propto \,Q^{\frac{1}{2}}$, the subaerial channel's width-integrated shear stress response is $\tau_{tf}\, \propto\,  Q$.
-For sufficiently short time periods,  channel width and wetted area are fixed in the subglacial system so that the response of width-integrated shear stress is $ \tau_{tg}\, \propto\,  Q^{2}$, with a larger exponent on $Q$ than in the subaerial case \citep{alley1997}.
-In this case, when water discharge fluctuates at a shorter timescale than the subglacial channel evolution \citep{gimbert2016}
-Note also that the subglacial shear-stress response to water discharge will be greater than the subaerial's provided that the exponent is below one in Equation~\ref{eq:wcf}.
-However, when subglacial channel size lies in equilibrium with water discharge, the  width-integrated shear stress response is $\tau_{tg}\, \propto \, Q^{\frac{4}{5}}$, similar to subaerial channels.
-Hydraulics could trend toward this equilibrium regime with average discharge over $1-5$ days when variability in subglacial channels  approaches that of subaerial channels (Figure~\ref{fig:multi_run}).
-Furthermore, the stronger correlation between hydraulics and water discharge in \icesheet{} suggests that this system is closer to this equilibrium compared to \alpine{}, given the reduced diurnal discharge amplitude in \icesheet{}.
-
-\ian{mauro: I think a supplement be made for this.  could you send me the one for R-channels? Is this essentially equation 14 in Werder et al., 2010?}
-
-In subglacial channels, peaks occur with the fastest rate of  water discharge increase relative to the channel's growth (Figure~\ref{fig:model_outs}).
-In turn, the subglacial channels here contain feedbacks that limit sediment transport capacity as the subglacial channel grows with increasing water discharge capacity, effectively reducing sediment transport capacity with the stabilization of water discharge.
-Similar feedback occurs in subaerial systems \citep{phillips2016}, albeit at longer timescales of centuries or more, as opposed to days here \citep{gimbert2016}.
-
-\subsection{Records of sediment discharge  from glaciers}
-
-The increased performance of subaerial hydraulics in representing observations occurs despite the proximity of glaciers to these measurements (Figure~\ref{fig:Qw_vari}).
-This could occur for two reasons.
-
-First, these measurement stations lie $0.5$ to $2$\,\unit{km} downstream of the glacier terminus \citep{cowton2012,felix2022}, likely a long enough distance for the pressurized subglacial regime to transition to the open channel subaerial one.
-The transition of subglacial to subaerial conditions may occur underneath glacier terminus \citep{perolo2018}, especially if closure rates of the ice channels are slow  \citep{rothlisberger1972}.
-Furthermore,  if a transport-limited regime persists at the glacier margin, then sediment can be easily mobilized, and subaerial processes could be represented just a short distance downstream.
-
-The consistent relationship between sediment discharge capacity and open channel flow conditions could be pronounced $\sim20$\,\unit{km} downstream of the Leverett site, where a strong correlation persists between sediment plume size and the river's water discharge into the Kangerlussuaq fjord \citep{mcgrath2010}.
-In marine-terminating glacier catchments, for contrast, a less consistent relationship may occur between water discharge or melt extent and sediment plume size \citep{tedstone2012}.
-Here, ocean water pressurizes subglacial channels at the ice front \citep[e.g.][]{how2017}, so the minimal correlation could result from the inconsistent relationship between subglacial sediment transport capacity and water discharge (Figure~\ref{fig:Qw_vari}).
-
-Second, the sediment discharge records could represent processes such as increased sediment accessibility when melt extends up glacier, thereby concurrently increasing sediment and water discharge \citep[e.g.][]{vergara2022}.
-A supply-limited regime could well persist at many glaciers, because the sediment's threshold of motion can be reached more frequently, across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
-Sediment exhaustion occurring through this process may also explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \citep{overeem2017}.
-Note that in the data here, however, the absence of lags between the observed sediment and water discharge may suggest that sediment is mobilized close to the measurement stations, as opposed to distal areas of the glacier bed as would be expected in a supply-limited regime \citep[Figure~\ref{fig:model_outs}\, a and e; ][]{williams1989}.
-
-%Regardless of the increased performance of water discharge compared to subglacial sediment transport capacity in the observations, parameterization outputs suggest that additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to subaerial systems \citep{jansson2005,ganti2016}.
-
-Sediment discharge records have been used to establish the relationship, or lack thereof, between climate forcing and glacial erosion \citep[e.g.][]{koppes2009a,ganti2016,willenbring2016,mariotti2021}.
-The variable relationship between water discharge and subglacial sediment transport capacity presented here mostly applies to the daily- to weekly-time scales controlling the size of subglacial channels (Figure~\ref{fig:multi_run}).
-Yet, identifying climatic signals in sediment transport from the transport-limited component of glacier-sedimentary systems may require higher thresholds of water discharge compared to subaerial systems when water discharge varies substantial over short timescales \citep[Figure~\ref{fig:Qw_vari}; ][]{tofelde2021}.
-
-Over time scales of weeks or longer, the glacier's sediment transport capacity responds to the ice thickness, controlling the channel closure rate, and the glacier's surface slope, in addition to water discharge \citep[Figure~\ref{fig:multi_run}, ~ Section~\ref{sect:sub_mode}; ][]{rothlisberger1972,shreve1972,gimbert2016,stevens2022}.
-In a supply-limited state, often assumed in landscape evolution models, then glaciers' sediment discharge record also represents additional processes of sediment availability and bedrock erosion from  water pressure variations and sliding  \citep[e.g.][]{iverson2012,herman2015,delaney2023}.
-This multitude of processes lies in contrast to most subaerial systems, where transport capacity typically responds to sediment size, water discharge and hydraulic gradient, and the hydraulic gradient generally remains stable over time scales ranging from years to millennia \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,tucker1997,wickert2019}.
-The stronger link between sediment transport capacity and hydrological forcing in subaerial channels may mean that variations in sediment transport records can be more easily attributed to climatic conditions than in subglacial systems.
-
-\section{Conclusions}
-
-Sediment transport capacity is largely driven by the shear stress of water flowing through a channel and the channel's width.
-Subaerial channels  alter their channel width and water velocity immediately in response to changing water discharge.
-In contrast, pressurized subglacial channels largely accommodate changing water discharge by altering water velocity and shear stress, upon which sediment transport depends.
-Water velocity responds quickly to changes in water discharge because the size of subglacial channels generally evolves slowly compared to variations in water discharge.
-Despite these differences, parameterizations of subaerial hydraulics generally represent observed sediment discharge variations better than subglacial hydraulics.
-%This could be attributed to the role of open channel flow at the glacier margin and in the proglacial area, or the influence of water discharge in driving sediment availability, rather than subglacial sediment discharge capacity.
-
-Implications of the findings here include:
-\begin{enumerate}
-\item the inconsistent relationship between water discharge and sediment transport capacity underneath glaciers creates increased variability in sediment transport capacity compared to river systems;
-\item the timing of peak sediment transport capacity below glaciers generally coincides with large rates of increasing in water discharge, as opposed to the maximum quantity of water discharge;
-\item a consistent relationship between water and sediment discharge may represent subglacial sediment access or open channel hydraulics near the glacier margin, instead of subglacial sediment transport capacity.
-
-\end{enumerate}
-
-\section*{Acknowledgments}
-\noindent
-SNSF Project No. PZ00P2\_202024 provided  funding for I. Delaney.
-Code to make the figures, with links to the data, can be found at \url{https://bitbucket.org/IanDelaney/xsection/src/master/}.
-Code and data will be uploaded to a permanent repository, with FAIR principles, pending acceptance of the paper.
-We thank G. King for commenting on a previous version of this manuscript.
-
-\bibliographystyle{apalike}
-\bibliography{PaperLib.bib}
-
-\section{Supplement}
-
-\renewcommand{\figurename}{Figure S}
-\setcounter{figure}{0}
 \begin{center}
@@
 
+\newpage
+
+\section{Mauro's bit about shear stress}
+% If you get an error about an unknown bounding box, try specifying the width and height of the figure with the natwidth and natheight options.
+This is common when trying to add a PDF figure without pdflatex.
+% \begin{figure}
+% \noindent\includegraphics[natwidth=800px,natheight=600px]{samplefigure.pdf}
+%\caption{caption}
+%\label{pdffiguresample}
+%\end{figure}
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Optional Appendices go here
+%
+% The \appendix command resets counters and redefines section heads
+%
+% After typing \appendix
+%
+%\section{Here Is Appendix Title}
+% will show
+% A: Here Is Appendix Title
+%
+%\appendix
+%\section{Here is a sample appendix}
 \end{document}
-
-These different characteristics between glacial and subaerial channels show that records of sediment transport downstream of glaciers represent both subglacial and subaerial processes together.
-Divergent processes discussed here might be particularly relevant at glacier margins, where water transitions from pressurized subglacial flow to open-channel subaerial flow.
-The inconsistent response of sediment transport in subglacial and subaerial systems to changing water discharge could add uncertainty in evaluating  sediment transport signals in subglacial systems.
-In addition, the subglacial parameterization shows that sediment transport capacity in glacier systems responds to a large number of factors such as channel size, ice thickness, and glacier surface slope, which may react to climate forcing differently than water discharge alone.
-
-In natural systems, variable mobilization and deposition of sediment occur as water and sediment discharge accumulate as they move through the two-dimensional channel network will compound the already substantial variability in sediment mobilization parameters accounted for here in the lumped parameterization.
-Such variability could be especially strong in the subglacial system where meltwater pathways change in response to the evolving hydraulic gradient \citep{delaney2023}.
-
-Results here suggest that increased variability in sediment transport in subglacial systems  could be particularly pronounced on ice sheets with large catchments and reduced water discharge variability over sub-monthly timescales.
-Furthermore, width-integrated shear stress across the channel bed can remain relatively high across a range of water discharges (Figures~\ref{fig:model_outs} and \ref{fig:Qw_vari}).
-\TODO{is this needed?... does if conflict with observations of the Leverett and not testable}

### 2023-09-21 Delaneyetal2023b.tex
getting ready to pass around
Updates from Overleaf
small changes to be passed around

@@
 \usepackage{lineno}
+\usepackage{placeins}
 \usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
@@
 In the subglacial channel, peaks generally occur over the period when water discharge increases at the fastest rate (Figure~\ref{fig:model_outs}).
-Conversely, in the subaerial channel, peaks in these sediment transport variables occur when water discharge is highest.
+Alternatively, in the subaerial channel, peaks in these sediment transport variables occur when water discharge is highest.
 The timing of the greatest rate of change in water discharge occurs before the peaks in water discharge.
@@
 Each water discharge value in the subaerial channel results in a unique water velocity, shear stress, and width-integrated shear stress, creating a perfect rank correlation between the variables (Figure~\ref{fig:Qw_vari} a--f).
-Conversely in the subglacial channel,  water velocity, shear stress, and width-integrated shear stress can vary substantially for a given water discharge and width-integrated shear stress varies with respect to the other two  outputs as well (Figure~\ref{fig:Qw_vari} a--f).
+In the subglacial channel, however,  water velocity, shear stress, and width-integrated shear stress can vary substantially for a given water discharge and width-integrated shear stress varies with respect to the other two  outputs as well (Figure~\ref{fig:Qw_vari} a--f).
 For instance, very high  water velocity values and shear stresses can occur across minimal water discharge at $2.5$\,\unit{m}$^3$\,\unit{s}$^{-1}$ to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
@@
 However, in \alpine{}, the variability in width-integrated shear stress is comparable between the two systems, due to the variable water discharge (Figure~\ref{fig:multi_run}).
-Conversely, in \icesheet{}, with lower water discharge variability, the variability in width-integrated shear stress remains higher in the subglacial system across the time aggregation periods.
+ However, in \icesheet{}, with lower water discharge variability, the variability in width-integrated shear stress remains higher in the subglacial system across the time aggregation periods.
 
@@
 The rapid increases and decreases in subglacial sediment transport capacity may compound the variability in sediment transport already present in subaerial systems \cite<Figure~\ref{fig:Qw_vari}; >{williams1989,jerolmack2010}.
-By applying the discharge - width scaling in Equation~\ref{eq:wcf} to Equation~\ref{eq:v}, by setting $S\, \propto \,Q^{\frac{1}{2}}$, the subaerial channel's width-integrated shear stress response is $\tau_{tf}\, \propto\,  Q$.
+By applying the discharge - width scaling in Equation~\ref{eq:wcf} to Equation~\ref{eq:v}, by setting $S\, \propto \,Q^{\frac{1}{2}}$, the subaerial channel's width-integrated shear stress response is $\tau_{tf}\, \propto\,  Q$ (see \ref{sect:scaling}).
 For sufficiently short time periods,  channel width and wetted area are fixed in the subglacial system so that the response of width-integrated shear stress is $ \tau_{tg}\, \propto\,  Q^{2}$, with a larger exponent on $Q$ than in the subaerial case following \citeA{alley1997}.
@@
 However, when subglacial channel size lies in equilibrium with water discharge, the  width-integrated shear stress response is $\tau_{tg}\, \propto \, Q^{\frac{4}{5}}$, similar to subaerial channels.
-It follows that, the stronger correlation between hydraulics and water discharge in \icesheet{} could result from being closer to this equilibrium between variations in the channels wetter area and water discharge compared to \alpine{}.
+It follows that the stronger correlation between hydraulics and water discharge in \icesheet{} could result from being closer to this equilibrium between variations in the channels' wetter area and water discharge compared to \alpine{}.
 This state could result from the reduced diurnal discharge amplitude in \icesheet{} (Figures~\ref{fig:model_outs} and ~\ref{fig:Qw_vari}).
@@
 SNSF Project No. PZ00P2\_202024 provided  funding for I. Delaney.
-We thank G. King for commenting on a previous version of this manuscript.
+G. King provided useful comments on a previous version of this manuscript.
 
 \bibliography{PaperLib}
+
+\newpage
 
@@
 \end{table}
-
-\newpage
+\FloatBarrier
 
@@
 \begin{center}
-  \begin{figure}[h]
+  \begin{figure}[!h]
     \includegraphics[width=0.7\linewidth]{Fig3_hr.png}
@@
 \begin{center}
-  \begin{figure}[h]
+  \begin{figure}[!h]
     \includegraphics[width=0.7\linewidth]{Fig3_12hr.png}
@@
 \begin{center}
-  \begin{figure}[h]
+  \begin{figure}[!h]
     \includegraphics[width=0.7\linewidth]{Fig3_1day.png}
@@
 \begin{center}
-  \begin{figure}[h]
+  \begin{figure}[!h]
     \includegraphics[width=0.7\linewidth]{Fig3_5day.png}
@@
 \begin{center}
-  \begin{figure}[h]
+  \begin{figure}[!h]
     \includegraphics[width=0.7\linewidth]{Fig3_10day.png}
@@
 \end{center}
+
 \begin{center}
-  \begin{figure}[h]
+  \begin{figure}[!h]
     \includegraphics[width=0.7\linewidth]{Fig3_15day.png}
@@
 \end{center}
+\FloatBarrier
 
 \newpage
-
 \section{Mauro's bit about shear stress}
+\label{sect:scaling}
 % If you get an error about an unknown bounding box, try specifying the width and height of the figure with the natwidth and natheight options.

### 2023-09-21 Delaneyetal2023b.tex
small bits

@@
 The variable relationship between water discharge and subglacial sediment transport capacity presented here mostly applies to the daily- to weekly-time scales controlling the size of subglacial channels (Figure~\ref{fig:multi_run}).
-Yet, identifying climatic signals in sediment transport from the transport-limited component of glacier-sedimentary systems may require higher thresholds of water discharge compared to subaerial systems when water discharge varies substantial over short timescales \cite<Figure~\ref{fig:Qw_vari}; >{tofelde2021}.
+Yet, identifying climatic signals in sediment transport from the transport-limited component of glacier-sedimentary systems may require higher thresholds of water discharge compared to subaerial systems when water discharge varies substantially over short timescales \cite<Figure~\ref{fig:Qw_vari}; >{tofelde2021}.
 
@@
 \ian{let me know if these accurately describe your contributions... I hope I am not leaving anything out}
-I.D. designed the study, developed and implement the model and experiments and lead the writing.
+I.D. designed the study, developed, designed, and implemented the model and experiments, and wrote the manuscript.
 D.F. provided data from Fieschergletscher.
 A.J.T. helped interpret data from Leverett Glacier and provided guidance on experiment design.
-M.A.W. provided data from Fieschergletscher, added the relationship between shear stress and water discharge (see Section~\ref{sect:dis_qsc} and Supplement).
+M.A.W. provided data from Fieschergletscher and added the relationship between shear stress and water discharge (see Section~\ref{sect:dis_qsc} and \ref{sect:scaling}).
 All authors provided  key inputs in writing and editing the manuscript.

### 2023-09-22 Delaneyetal2023b.tex
Draft to pass around

@@
 
-\authors{Ian Delaney\affil{1},  Daniel Farinotti\affil{2,3},  Andrew J. Tedstone\affil{4},   Mauro A. Werder\affil{2,3}}
+\authors{Ian Delaney\affil{1},  Andrew J. Tedstone\affil{2},   Mauro A. Werder\affil{3,4},  Daniel Farinotti\affil{3,4}, }
 
 \affiliation{1}{Institut des dynamiques de la surface terrestre (IDYST), Universit\'{e} de Lausanne, B\^{a}timent G\'{e}opolis, 1015 Lausanne, Switzerland}
-\affiliation{2}{Laboratory of Hydraulics, Hydrology and Glaciology (VAW), ETH-Z\"urich, H\"onggerbergring 26, 8093 89 Z\"urich, Switzerland}
-\affiliation{3}{Swiss Federal Institute for Forest, Snow and Landscape Research (WSL) Z\"uricherstrasse 111, 8903 1011 Birmensdorf, Switzerland}
-\affiliation{4}{Department of Geosciences, University of Fribourg, Ch. du Musée 1700, Fribourg, Switzerland}
+\affiliation{2}{Department of Geosciences, University of Fribourg, Ch. du Musée 1700, Fribourg, Switzerland}
+\affiliation{3}{Laboratory of Hydraulics, Hydrology and Glaciology (VAW), ETH-Z\"urich, H\"onggerbergring 26, 8093 89 Z\"urich, Switzerland}
+\affiliation{4}{Swiss Federal Institute for Forest, Snow and Landscape Research (WSL) Z\"uricherstrasse 111, 8903 1011 Birmensdorf, Switzerland}
 

### 2023-09-25 Delaneyetal2023b.tex
Daniel's comments

@@
 \begin{keypoints}
-\item Water discharge variations in subglacial channels are mainly accommodated by water velocity, not channel size, given their subglacial channel's evolution.
-\item Greater variability in water velocity in subglacial channels causes sediment transport capacity to vary more than in subaerial channels.
+\item Water discharge variations in subglacial channels are mainly accommodated by water velocity, not channel size.
+\item Greater variability in water velocity causes sediment transport capacity to vary more in subglacial channels than in subaerial channels.
 \item In interpreting sediment transport records, water discharge may represent processes including proglacial sediment mobilization and sediment access driving variations in sediment discharge.
@@
 Over long periods, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from under glaciers \cite<c.f.>{hallet1979,iverson2012,ugelvig2018}.
-On shorter time periods, pressurized water transports sediment from the under glacier \cite{walder1994,creyts2013,beaud2018,delaney2019}, should enough sediment be present subglacially (i.e.
-in a transport-limited regime).
+On shorter time periods, pressurized water transports sediment from underneath glaciers \cite{walder1994,creyts2013,beaud2018,delaney2019}, should enough sediment be present subglacially.
 
-In a transport-limited regime, sediment discharge responds to the sediment transport capacity defined as the amount of sediment the water can carry.
-Sediment transport capacity depends on the shear stress between water and sediment it flows  over \cite{shields1936,meyer1948,engelund1967} and the width of the channel bottom $w_c$ over which to mobilize sediment.
+In a transport-limited regime, sediment discharge responds to the sediment transport capacity, defined as the amount of sediment the water can carry.
+Sediment transport capacity depends on the shear stress between water and the sediment it flows  over \cite{shields1936,meyer1948,engelund1967} and the width of the channel bottom $w_c$ over which to mobilize sediment.
 The shear stress  across the channel width $\tau_t$ responds to the velocity of water $v$ flowing through the channel so that
@@
 As a result, the subglacial channel size only evolves relatively slowly over days to months, whereas water discharge can vary over hours \cite<e.g.>{iken1986,andrews2014,nanni2020}.
-Therefore, subglacial water flow behaves more like pipe flow, and changes in water discharge $Q$ are mainly accommodated with increased water velocity $v$ \cite{alley1997}, because the size of the channel $S$ changes on a much longer timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
+Therefore, subglacial water flow behaves more like pipe flow, and changes in water discharge $Q$ are mainly accommodated with increased or decreased water velocity $v$ \cite{alley1997}, because the size of the channel $S$ changes on a much longer timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
 It follows that an increase in water discharge $Q$ will cause the shear stress across the channel $\tau_t$ to respond with water velocity $v$, as opposed to subaerial channels where channel width $w_c$ evolves as well.
@@
     \caption{Sketch for the different responses that subglacial and subaerial channels have to an increase in water discharge.
-Water velocity magnitude in the subglacial, $v_g$, and subaerial, $v_f$, channels are shown by arrow length.
-$S_g$ and $S_f$ represent the wetted area in subglacial and subaerial channels.
-The subglacial channel widths $w_{cg}$ remain unchanged, while the subaerial channel widths evolve $w_{cf}$.
+Water velocity magnitudes in the subglacial, $v_g$, and subaerial, $v_f$, channels are shown by arrow length.
+$S_g$ and $S_f$ represent the wetted area in subglacial and subaerial channels, respectively.
+The subglacial channel widths $w_{cg}$ remain unchanged, while the subaerial channel widths $w_{cf}$ evolve.
 Shear stress $\tau$ is responsible for the mobilization of sediment and increases with the number of makers at the channel bed.
-      Note that in the subaerial channel parameterization, channels have rectangular shapes with their width much larger than the depth (Section~\ref{sect:fluv}).}
+      Note that in the subaerial channel parameterization, channels have rectangular shapes with their width much larger than their depth (Section~\ref{sect:fluv}).}
     \label{fig:cartoon}
@@
 
-This manuscript aims to evaluate the relationship between sediment transport capacity and water discharge in subglacial channels, compared to subaerial ones.
+This manuscript evaluates the relationship between sediment transport capacity and water discharge in subglacial channels, and compares it to subaerial ones.
 In the absence of continuous measurements of subglacial and subaerial water velocity and channel morphology, numerical parameterizations of subglacial and subaerial hydraulics are utilized.
-Parameterizations are applied to hydrological records from an Alpine glacier in Switzerland (Fieschergletscher) and  a land-terminating glacier in Greenland (Leverett glacier).
+Parameterizations are applied to hydrological records from an Alpine glacier in Switzerland (Fieschergletscher) and  a land-terminating glacier in Greenland (Leverett Glacier).
 The parameterizations' lumped nature isolates the disparity between water discharge and sediment transport capacity in subglacial systems, independent of the upstream drainage network and sediment access.
@@
 Data is used from \citeA{tedstone2017} and \citeA{hawkings2014} at a $5$\,\unit{min} time interval from May 28, 2012 to August 8, 2012.
-Two gaps in suspended sediment data of roughly day in length are interpolated across with a linear spline.
+Two gaps in suspended sediment data of roughly one day in length are interpolated across with a linear spline.
 
@@
 Both parameterizations calculate   water velocity, shear stress, and width-integrated shear stress, upon which sediment transport depends \cite<Figure \ref{fig:cartoon}; >{shields1936}.
-The evaluation of these variables omits the selection of a sediment transport relationship and grain-size parameter \cite<e.g.>{shields1936,meyer1948} needed to calculate sediment transport capacity from shear stress.
+The evaluation of these variables omits the selection of a sediment transport relationship and a grain-size parameter \cite<e.g.>{shields1936,meyer1948} needed to calculate sediment transport capacity from shear stress.
 
@@
 Thus, it is assumed that the changing hydraulic head at the top of the glacier is negligible in the channel size evolution.
-The evolution of subglacial channel size $S_g$ is given as
+The evolution of subglacial channel size $S_g$ is then given as
 \begin{linenomath*}
@@
 \end{linenomath*}
-\noindent where $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (Table \ref{table:vpm}), $l$ is the length of subglacial channel, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change from the glacier terminus to its top, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the ice overburden pressure, and $n$ is Glen's n \cite{glen1955} (usually $n=3$).
+\noindent where $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (individual quantities defined in Table \ref{table:vpm}), $l$ is the length of the subglacial channel, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change from the glacier terminus to its top, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the ice overburden pressure, and $n$ is the exponent in Glen's law for ice flow \cite{glen1955} (usually $n=3$).
 The first term on the right side of the equation represents the opening of the channel through frictional heating, while the following term represents the creep closure of the channel from ice deformation.
 
-Following the Darcy-Weisbach, head drop $\Delta h$ is,
+Following the Darcy-Weisbach equation, head drop $\Delta h$ is,
 \begin{linenomath*}
@@
 \end{linenomath*}
-\noindent where $f_r$ is a friction factor, $D_h$ is the hydraulic diameter, and $l$ is the length overwhich head change occurs.
+\noindent where $f_r$ is a friction factor, $D_h$ is the hydraulic diameter, and $l$ is the length over which head change occurs.
 $s$ is a factor accounting for channel geometry \cite{hooke1990}, calculated as
@@
   \label{eq:Dh2S}
-  S_g= \frac{D_h^2}{2}\,\frac{(\frac{\beta}{2}+\sin \frac{\beta}{2})^2}{\beta - \sin \beta},
+  S_g= \frac{D_h^2}{2}\,\frac{(\frac{\beta}{2}+\sin \frac{\beta}{2})^2}{\beta - \sin \beta}.
 \end{equation}
-where $\beta$ is the central angle of the circular segment that comprises the channel (the so-called Hooke angle).
 
@@
 %
-where  the water velocity $v_g = \frac{Q}{S_g}$.
+where $v_g = \frac{Q}{S_g}$ is the water velocity.
 
-The flat channel floor width $w_{cg}$, given angle $\beta$, is
+The width-integrated shear stress can then be represented by
+\begin{equation}
+  \label{eq:tautg}
+  \tau_{tg}=w_{cg}\,\tau_g,
+\end{equation}
+
+where $w_{cg}$ is the flat channel floor width:
 \begin{equation}
@@
   w_{cg} = 2  \sin \frac{\beta}{2} \sqrt{\frac{2\, S_g}{\beta -\sin \beta}}.
-\end{equation}
-
-The width-integrated shear stress is represented as
-\begin{equation}
-  \label{eq:tautg}
-  \tau_{tg}=w_{cg}\,\tau_g.
 \end{equation}
@@
 %
-where $k$ is a constant and the exponent is commonly given as $\frac{1}{2}$ \cite{leopold1953}.
+where $k$ is a constant and $\frac{1}{2}$ is a commonly chosen exponent \cite{leopold1953}.
 %\mauro{This is for so called ``regime channels''}
@@
 %
-As above, the width-integrated shear stress is
+As above, the width-integrated shear stress is then
 \begin{equation}
@@
 
-The parameterizations above are applied to hydrological records from the Fieschergletscher (\alpine) and the Leverett glacier (\icesheet).
+The parameterizations above are applied to hydrological records from the Fieschergletscher (\alpine) and the Leverett Glacier (\icesheet).
 The subglacial and subaerial parameterizations represent the subglacial flow of water exiting a glacier and transitioning to subaerial flow, as it moves through the catchment.
@@
 
-The parameterizations are first applied to a reference test case for each glacier that assumes  subglacial channel with $\beta=\frac{\pi}{2}$ and friction factors $f_r$ and $f_p$ are tuned so that the parameterizations reproduce reasonable velocities for both  the Greenland Ice Sheet and Alpine glaciers \cite<$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$>{werder2010b,chandler2013}.
-In both subaerial channels, the bed slope, $\nabla z_c$ is $0.05$.
+For each glacier, the parameterizations are first applied to a reference test case that assumes  subglacial channel with $\beta=\frac{\pi}{2}$.
+In this case the friction factors $f_r$ and $f_p$ are tuned so that the parameterizations reproduce reasonable velocities for both  the Greenland Ice Sheet and the Alpine glacier \cite<$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$>{werder2010b,chandler2013}.
+In both subaerial channels, the bed slope is $\nabla z_c = 0.05$.
 Water velocity, shear stress, and width-integrated shear stress in the subglacial and subaerial channels from this formulation are presented below.
@@
 Second, to characterize the variability in sediment discharge capacity in subglacial compared to subaerial channels further, $100$  parameterizations are applied to a range of different channel geometry factors ($\beta$ and $k$) and friction factors ($f_r$ and $f_p$).
-Runs with parameter combinations are culled if their mean subglacial water velocity over the season lies outside of the range between $0.75$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $2$\,\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity lies outside of the range $0.3$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with some measurements \cite<e.g.>{werder2010b,magnusson2012,chandler2013}.
-The standard deviation of an output variable over the course of the season is used to quantify the variability of that output.
+Runs with parameter combinations are culled if their mean subglacial water velocity over the season lies outside of the range between $0.75$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $2$\,\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity lies outside of the range $0.3$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with measurements \cite<e.g.>{werder2010b,magnusson2012,chandler2013}.
+For any given variable, the standard deviation of that variable over the course of the season is used to quantify its variability.
 
-Lastly,  outputs are compared with the sediment discharge records the glaciers using Spearman rank correlation, or the correlation with respect to the ordering of values from the outputs and observations.
+Lastly,  outputs are compared with the sediment discharge records of the glaciers using Spearman rank correlation, that is the correlation with respect to the ordering of values from the outputs and the observations.
 Rank correlation reduces the impact of the non-linear response of sediment transport capacity to the hydrological forcing.
 Periods of uncertainty in the data in \icesheet are removed from the analysis.
-Outputs and observations are aggregated across timescales from $1$\,\unit{hr} to $15$\,\unit{days} from the $15$\,\unit{min} outputs, by averaging values one half of the timescale before and after a point time.
+Outputs and observations, which have a native time resolution of $15$\,\unit{min}, are aggregated across timescales from $1$\,\unit{hr} to $15$\,\unit{days} by averaging values one half of the timescale before and after a given point time.
 
@@
 In other panels, blue lines represent outputs from the subglacial channel, while orange lines represent the subaerial channel.
-    Data in plots are at $15$\unit{min} intervals.
-    Arrows denote some events where the variable peaks in the subglacial channel prior to the subaerial one.
+    Data are shown at $15$\unit{min} intervals.
+    Arrows denote some events where the variables peaks in the subglacial channel prior to the subaerial one.
     Insets in e-h show the  peak melt event \icesheet{} denoted by the shaded area.
@@
     Variables have been normalized to mean values for better comparison amongst the runs.
-    $R_g$ ($R_f$) shows the Spearman rank correlation values for the subglacial (subaerial) outputs.
-    Plots are shown with data and outputs at $15$\unit{min} interval.
+    $R_g$ and $R_f$ show the Spearman rank correlation values for the subglacial and the subaerial outputs, respectively.
+    Plots are shown with data and outputs at $15$\unit{min} intervals.
   }
@@
     \includegraphics[width=0.9\linewidth]{Fig4.png}
-    \caption{Variance (measured by the standard deviation) in water velocity, shear stress, and width integrated shear stress for different time aggregations across the range of subglacial and subaerial channel shapes and friction factors.
+    \caption{Variability (measured by the standard deviation) in water velocity, shear stress, and width integrated shear stress for different time aggregations across the range of subglacial and subaerial channel shapes and friction factors.
       Shaded areas denote the minimum and maximum of the $100$ parameter combinations tested for each of the systems (Section~\ref{sect:imp}).
       Solid lines denote  mean values.
-      Top row shows results from \alpine{} case, while bottom row show outputs from \icesheet{} case
+      Top row shows results from the \alpine{} case, while bottom row show outputs from the \icesheet{} case
       Time intervals shown with markers denote $15$\,\unit{min}, $1$\,\unit{hr}, $12$\,\unit{hr}, $1$\,\unit{d}, $5$\,\unit{d}, $10$\,\unit{d}, and $15$\,\unit{d} aggregation periods.
@@
 
-The second application examines the impact of a variety of channel shapes and friction factors on the relationship between water discharge and sediment transport capacity, in both systems.
+The second application examines the impact of a variety of channel shapes and friction factors on the relationship between water discharge and sediment transport capacity in both systems.
 Across the range of parameter values examined, variability in water velocity and shear stress remains higher in the subglacial system compared to the subaerial one.
@@
 \subsection{Relationship between parameterization  outputs and observed sediment discharge}
-When comparing the output variables with suspended sediment discharge measurements, velocity, shear stress, and width-integrated shear stress from the subaerial system generally show better correlation with sediment discharge compared to subglacial outputs across the tested subglacial channel parameters (Figures~\ref{fig:Qw_vari} g--l and \ref{fig:multi_run}).
+When comparing the output variables with suspended sediment discharge measurements, the variables velocity, shear stress, and width-integrated shear stress from the subaerial system generally show better correlation with sediment discharge compared to subglacial outputs across the tested subglacial channel parameters (Figures~\ref{fig:Qw_vari} g--l and \ref{fig:multi_run}).
 The rank correlation amongst the  subaerial parameterization's three output variables remains fairly constant despite the variable impact of channel shape and consistent scaling with water discharge (Figure \ref{fig:Qw_vari} and Section~\ref{sect:fluv}).
@@
 Quantities that control sediment transport capacity, such as water velocity, shear stress, and width-integrated shear stress, vary differently in response to water discharge in subglacial channels compared to subaerial ones (Figure~\ref{fig:Qw_vari}).
-In sediment transport relationships, such as in \citeA{meyer1948} or \citeA{engelund1967}, that evaluate sediment discharge capacity, shear stress is scaled to the $\frac{3}{2}$ or $\frac{5}{2}$ power, respectively.
+In sediment transport relationships that evaluate sediment discharge capacity, such as in \citeA{meyer1948} or \citeA{engelund1967}, shear stress is scaled to the power of $\frac{3}{2}$ or $\frac{5}{2}$, respectively.
 The exponent greater than $1$ magnifies sediment discharge variability beyond the variable sediment transport parameters described above (Figure~\ref{fig:multi_run}).
-Note that sediment transport relationships describe sediment flux per unit width so that the integrated capacity across channel occurs after application of the exponent, unlike the width-integrated shear stress evaluation here \cite<Equations\,\ref{eq:tautg} and \ref{eq:tautf}; c.f.>{tucker1997}.
+Note that sediment transport relationships describe sediment flux per unit width, so that the integrated capacity across channel occurs after application of the exponent, unlike the width-integrated shear stress evaluation here \cite<Equations\,\ref{eq:tautg} and \ref{eq:tautf}; c.f.>{tucker1997}.
 Note also that both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress is homogeneous across the channel bed, despite the limitations of this assumption \cite<e.g.>{yager2018}.
 
+\ian{Mauro: Can you double check this?}
 The rapid increases and decreases in subglacial sediment transport capacity may compound the variability in sediment transport already present in subaerial systems \cite<Figure~\ref{fig:Qw_vari}; >{williams1989,jerolmack2010}.
@@
 In subglacial channels, peaks occur with the fastest rate of  water discharge increase relative to the channel's growth (Figure~\ref{fig:model_outs}).
-In turn, the subglacial channels here contain feedbacks that limit sediment transport capacity as the subglacial channel grows with increasing water discharge capacity, effectively reducing sediment transport capacity with the stabilization of water discharge.
+In turn, the subglacial channels contain feedbacks that limit sediment transport capacity as the subglacial channel grows with increasing water discharge capacity.
+This effectively reduces sediment transport capacity with the stabilization of water discharge.
 Similar feedback occurs in subaerial systems \cite{phillips2016}, albeit at longer timescales, as opposed to days here \cite{gimbert2016}.
@@
 
-First, these measurement stations lie $0.5$ to $2$\,\unit{km} downstream of the glacier terminus \cite{cowton2012,felix2022}, likely a long enough distance for the pressurized subglacial regime to transition to the open channel subaerial one.
-The transition of subglacial to subaerial conditions may occur underneath glacier terminus \cite{perolo2018}, especially if closure rates of the ice channels are slow  \cite{rothlisberger1972}.
+First, these measurement stations lie $0.5$ to $2$\,\unit{km} downstream of the glacier terminus \cite{cowton2012,felix2022}, a long enough distance for the pressurized subglacial regime to transition to the open channel subaerial one.
+Actually, the transition of subglacial to subaerial conditions may occur underneath glacier terminus \cite{perolo2018}, especially if closure rates of the ice channels are slow  \cite{rothlisberger1972}.
 Furthermore,  if a transport-limited regime persists at the glacier margin, then sediment can be easily mobilized, and subaerial processes could be represented just a short distance downstream.
@@
 
-Second, the sediment discharge records could represent processes such as increased sediment accessibility when melt extends up glacier, thereby concurrently increasing sediment and water discharge \cite<e.g.>{vergara2022}.
-A supply-limited regime could well persist at many glaciers, because the sediment's threshold of motion can be reached more frequently, across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
+Second, the sediment discharge records could represent processes such as increased sediment accessibility when melt extends up glacier, thereby  increasing sediment and water discharge \cite<e.g.>{vergara2022}.
+A supply-limited regime could well persist at many glaciers, because the sediment's threshold of motion can be reached more frequently and across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
 Sediment exhaustion occurring through this process may also explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \cite{overeem2017}.
@@
 
-Regardless of the increased performance of water discharge compared to subglacial sediment transport capacity in the observations, parameterization outputs suggest that additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to subaerial systems \cite{jansson2005,ganti2016}.
+Regardless of the increased performance of water discharge compared to subglacial sediment transport capacity in the observations, parameterization outputs suggest that additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate \cite{jansson2005,ganti2016}.
 Sediment discharge records have been used to establish the relationship, or lack thereof, between climate forcing and glacial erosion \cite<e.g.>{koppes2009a,ganti2016,willenbring2016,mariotti2021}.
@@
 In a supply-limited state, often assumed in landscape evolution models, then glaciers' sediment discharge record also represents additional processes of sediment availability and bedrock erosion from  water pressure variations and sliding  \cite<e.g.>{iverson2012,herman2015,delaney2023}.
-This multitude of processes lies in contrast to most subaerial systems, where transport capacity typically responds to sediment size, water discharge and hydraulic gradient, and the hydraulic gradient generally remains stable over time scales ranging from years to millennia \cite<Section~\ref{sect:fluv}; e.g.>{muller1968,tucker1997,wickert2019}.
+This multitude of processes lies in contrast to most subaerial systems, where transport capacity typically responds to sediment size, water discharge, and hydraulic gradient, and the hydraulic gradient generally remains stable over time scales ranging from years to millennia \cite<Section~\ref{sect:fluv}; e.g.>{muller1968,tucker1997,wickert2019}.
 The stronger link between sediment transport capacity and hydrological forcing in subaerial channels may mean that variations in sediment transport records can be more easily attributed to climatic conditions than in subglacial systems.
@@
 In contrast, pressurized subglacial channels largely accommodate changing water discharge by altering water velocity and shear stress, upon which sediment transport depends.
-Water velocity responds quickly to changes in water discharge because the size of subglacial channels generally evolves slowly compared to variations in water discharge.
-This makes the response of sediment transport capacity much more sensitive to changes in water discharge in subglacial systems compared to subaerial ones. 
+In subglacial channels, water velocity responds quickly to changes in water discharge because their size generally evolves slowly compared to variations in water discharge.
+This makes the response of their sediment transport capacity much more sensitive to changes in water discharge when compared to subaerial channels. 
 Despite these differences, parameterizations of subaerial hydraulics in two glacierized catchments generally represent observed sediment discharge variations better than subglacial hydraulics.
 
-Implications of the findings here include:
+Implications of the presented findings include:
 \begin{enumerate}

### 2023-09-26 Delaneyetal2023b.tex
daniel and andrews comments

@@
 \item Greater variability in water velocity causes sediment transport capacity to vary more in subglacial channels than in subaerial channels.
-\item In interpreting sediment transport records, water discharge may represent processes including proglacial sediment mobilization and sediment access driving variations in sediment discharge.
+\item Water discharge may represent processes including proglacial sediment mobilization and sediment access driving variations in sediment discharge, as opposed to subglacial sediment transport capacity in glacierized catchments.
 \end{keypoints}
@@
 \begin{abstract}
-  Sediment transport  in subglacial and subaerial channels depends on  water velocity and channel width, which controls the shear stress exerted on the channel bottom and the space over which to integrate it.
-  In subaerial channels, depth, width, and velocity evolve quickly with water discharge, while in subglacial channels, water discharge changes mainly affect water velocity due to slow-evolving channel geometry.
-  Parameterizations applied to hydrographs from an Alpine glacier and the Greenland Ice Sheet reveal greater variability in sediment transport capacity in subglacial channels.
+  Sediment transport capacity in subglacial and subaerial channels depends on  water velocity and channel width, which together control the shear stress exerted on the channel bottom.
+  In subaerial channels, width, and velocity evolve quickly with water discharge, while in subglacial channels, water discharge changes mainly affect water velocity due to slow-evolving channel geometry.
+  Parameterizations of these subglacial and subaerial hydraulics are applied to hydrographs from an Alpine glacier and the Greenland Ice Sheet revealing greater variability in sediment transport capacity in subglacial channels.
   Furthermore, results show that high subglacial sediment transport rates can persist across various water discharges.
   Slowly evolving channel width accommodates some variability in subglacial channels.
-  Despite differences in response to water discharge, sediment discharge measurements generally correlate better with subaerial hydraulics.
+  Despite differences in response to water discharge, sediment discharge measurements generally correlate better with subaerial hydraulic parameters.
   This implies that other factors like sediment access or proglacial processes might be at play, affecting sediment transport capacity differently in subglacial and subaerial channels.
@@
 Over long periods, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from under glaciers \cite<c.f.>{hallet1979,iverson2012,ugelvig2018}.
-On shorter time periods, pressurized water transports sediment from underneath glaciers \cite{walder1994,creyts2013,beaud2018,delaney2019}, should enough sediment be present subglacially.
+On shorter time periods, pressurized water transports sediment from underneath glaciers \cite{walder1994,creyts2013,beaud2018,delaney2019}, should enough sediment be present subglacially (i.e.
+in a transport-limited regime).
 
 In a transport-limited regime, sediment discharge responds to the sediment transport capacity, defined as the amount of sediment the water can carry.
-Sediment transport capacity depends on the shear stress between water and the sediment it flows  over \cite{shields1936,meyer1948,engelund1967} and the width of the channel bottom $w_c$ over which to mobilize sediment.
+Sediment transport capacity depends on the shear stress between water and the sediment it flows  over \cite{shields1936,meyer1948,engelund1967} along with the width of the channel bottom $w_c$ over which to mobilize sediment.
 The shear stress  across the channel width $\tau_t$ responds to the velocity of water $v$ flowing through the channel so that
@@
 As a result, the subglacial channel size only evolves relatively slowly over days to months, whereas water discharge can vary over hours \cite<e.g.>{iken1986,andrews2014,nanni2020}.
-Therefore, subglacial water flow behaves more like pipe flow, and changes in water discharge $Q$ are mainly accommodated with increased or decreased water velocity $v$ \cite{alley1997}, because the size of the channel $S$ changes on a much longer timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
+Therefore, subglacial water flow behaves more like pipe flow, and changes in water discharge $Q$ are mainly accommodated with increased or decreased water velocity $v$ \cite{alley1997} because the size of the channel $S$ changes on a much longer timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
 It follows that an increase in water discharge $Q$ will cause the shear stress across the channel $\tau_t$ to respond with water velocity $v$, as opposed to subaerial channels where channel width $w_c$ evolves as well.
@@
 Additionally, the variable response of sediment transport capacity to water discharge in the two systems may affect the interpretation of sediment transport records in glacial systems \cite<e.g.>{muller1968,richards2003,swift2005,ganti2016}.
-Yet, little work has examined the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial channel \cite{alley1997}.
+Yet, little work has examined the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial channels \cite{alley1997}.
 %Furthermore, thorough observations of water velocity and channel width in subglacial channels are limited and, to these authors' knowledge, not contrasted with subaerial conditions at the glacier margins.
@@
 
-Water and suspended sediment data are leveraged from an Alpine and an ice sheet setting.
+Water and suspended sediment data are leveraged from Alpine and ice sheet settings.
 The Alpine site (\alpine) is  Fieschergletscher in the Swiss Alps ($46^\circ\,29'\,07''$ N, $8^\circ\,08'\,3''$ E).
@@
 The Leverett Glacier in Greenland (\icesheet) serves as the ice sheet setting.
-Water discharge and suspended sediment concentration data were collected roughly $2$\unit{km} downstream from the terminus from June to August/September 2009 to 2012 ($67^\circ\,03'\,5''$ N, $50^\circ\,12'\,59''$ W).
+Water discharge and suspended sediment concentration data were collected roughly $2$\unit{km} downstream from the terminus between June and August/September 2009 to 2012 ($67^\circ\,03'\,5''$ N, $50^\circ\,12'\,59''$ W).
 Data is used from \citeA{tedstone2017} and \citeA{hawkings2014} at a $5$\,\unit{min} time interval from May 28, 2012 to August 8, 2012.
@@
 
-Here, it is assumed that the water is transported through a subglacial channel flowing down the glacier; \cite<Figure~\ref{fig:cartoon}; >{rothlisberger1972}, and that the channel  size responds to frictional heating from water flow and creep closure by ice, using the Darcy-Weisbach formulation for water-flow through a pipe  \cite<e.g.>{rothlisberger1972,clarke2003}.
+Here, it is assumed that the water is transported through a subglacial channel \cite<Figure~\ref{fig:cartoon}; >{rothlisberger1972}, and that the channel  size responds to frictional heating from water flow and creep closure by ice, using the Darcy-Weisbach formulation for water-flow through a pipe  \cite<e.g.>{rothlisberger1972,clarke2003}.
 The formulation here does not consider the englacial storage of water.
 Thus, it is assumed that the changing hydraulic head at the top of the glacier is negligible in the channel size evolution.
-The evolution of subglacial channel size $S_g$ is then given as
+The evolution of subglacial channel size $S_g$ is given as
 \begin{linenomath*}
@@
 \end{linenomath*}
-\noindent where $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (individual quantities defined in Table \ref{table:vpm}), $l$ is the length of the subglacial channel, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change from the glacier terminus to its top, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the ice overburden pressure, and $n$ is the exponent in Glen's law for ice flow \cite{glen1955} (usually $n=3$).
+\noindent where $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (Table \ref{table:vpm}), $l$ is the length of the subglacial channel, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change from the glacier terminus to its top, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the ice overburden pressure, and $n$ is Glen's n \cite{glen1955} (usually $n=3$).
 The first term on the right side of the equation represents the opening of the channel through frictional heating, while the following term represents the creep closure of the channel from ice deformation.
@@
 
-The width-integrated shear stress can then be represented by
+The width-integrated shear stress is represented as
 \begin{equation}
@@
 %
-As above, the width-integrated shear stress is then
+As above, the width-integrated shear stress is
 \begin{equation}
@@
 
-For each glacier, the parameterizations are first applied to a reference test case that assumes  subglacial channel with $\beta=\frac{\pi}{2}$.
-In this case the friction factors $f_r$ and $f_p$ are tuned so that the parameterizations reproduce reasonable velocities for both  the Greenland Ice Sheet and the Alpine glacier \cite<$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$>{werder2010b,chandler2013}.
+For each glacier, the parameterizations are first applied to a reference test case for each glacier that assumes  subglacial channel with $\beta=\frac{\pi}{2}$.
+Friction factors $f_r$ and $f_p$ are tuned so that the parameterizations reproduce reasonable water velocities for both \alpine and \icesheet{}\cite<$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$>{werder2010b,chandler2013}.
 In both subaerial channels, the bed slope is $\nabla z_c = 0.05$.
@@
 Runs with parameter combinations are culled if their mean subglacial water velocity over the season lies outside of the range between $0.75$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $2$\,\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity lies outside of the range $0.3$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with measurements \cite<e.g.>{werder2010b,magnusson2012,chandler2013}.
-For any given variable, the standard deviation of that variable over the course of the season is used to quantify its variability.
+The standard deviation of an output over the course of the season is used to quantify its variability.
 
@@
 Rank correlation reduces the impact of the non-linear response of sediment transport capacity to the hydrological forcing.
-Periods of uncertainty in the data in \icesheet are removed from the analysis.
-Outputs and observations, which have a native time resolution of $15$\,\unit{min}, are aggregated across timescales from $1$\,\unit{hr} to $15$\,\unit{days} by averaging values one half of the timescale before and after a given point time.
+Outputs and observations, which have a native time resolution of $5$\,\unit{min}, are aggregated across timescales from $15$\,\unit{min} to $15$\,\unit{days} by averaging values one half of the timescale before and after a given point time.
 
@@
 Over the course of the season, large peaks in water velocity and shear stress, along with width-integrated shear stress, occur in response to increases in water discharge in both \alpine{} and \icesheet{}.
-This occurs as the subglacial values are generally larger than the subaerial values, thus more efficient at transporting sediment.
+This occurs as the subglacial outputs are generally larger than the subaerial values, thus more efficient at transporting sediment.
 In the subglacial channel, peaks generally occur over the period when water discharge increases at the fastest rate (Figure~\ref{fig:model_outs}).
@@
 Each water discharge value in the subaerial channel results in a unique water velocity, shear stress, and width-integrated shear stress, creating a perfect rank correlation between the variables (Figure~\ref{fig:Qw_vari} a--f).
-In the subglacial channel, however,  water velocity, shear stress, and width-integrated shear stress can vary substantially for a given water discharge and width-integrated shear stress varies with respect to the other two  outputs as well (Figure~\ref{fig:Qw_vari} a--f).
+In the subglacial channel, however,  the outputs can vary substantially for a given water discharge and width-integrated shear stress varies with respect to the other two  outputs as well (Figure~\ref{fig:Qw_vari} a--f).
 For instance, very high  water velocity values and shear stresses can occur across minimal water discharge at $2.5$\,\unit{m}$^3$\,\unit{s}$^{-1}$ to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
-In \icesheet, mean values of water velocity can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, effectively spanning much of the water discharge range (Figure~\ref{fig:Qw_vari} d--f).
+In \icesheet, mean values of water velocity can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
 
-When considering the subglacial channel's evolving width, width-integrated shear stress  across the channel generally increases with water discharge, with improved rank correlation compared to water discharge and water velocity or shear stress (Figure~\ref{fig:Qw_vari} a--f).
-Yet even the width-integrated shear stress  can vary substantially, with the highest values occurring at water discharge values ranging from roughly $11$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, to over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
+When considering the subglacial channel's evolving width, width-integrated shear stress  across the channel generally increases with water discharge, with improved rank correlation compared to water velocity or shear stress (Figure~\ref{fig:Qw_vari} a--f).
+Yet even the width-integrated shear stress  can vary substantially, with the highest values occurring at water discharge values ranging from roughly $11$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ to over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
 The variability in width-integrated shear stress is less pronounced in the \icesheet case, with reduced variability in water discharge (Figure~\ref{fig:Qw_vari} c, f).
-Yet, a range of width-integrated shear stress can occur at a given discharge.
+The persistence of high water velocity, shear stress, and width-integrated shear stress suggests that substantial sediment transport may occur across much of the water discharge range in the subglacial system (Figure~\ref{fig:Qw_vari} a--f).
 
@@
 Across the range of parameter values examined, variability in water velocity and shear stress remains higher in the subglacial system compared to the subaerial one.
-However, in \alpine{}, the variability in width-integrated shear stress is comparable between the two systems, due to the variable water discharge (Figure~\ref{fig:multi_run}).
+n \alpine{}, the variability in width-integrated shear stress is comparable between the two systems, due to the variable water discharge (Figure~\ref{fig:multi_run}).
  However, in \icesheet{}, with lower water discharge variability, the variability in width-integrated shear stress remains higher in the subglacial system across the time aggregation periods.
@@
 Note that lags and time of peak sediment transport conditions still occur over these periods (Figure~\ref{fig:model_outs}), while the variability in sediment transport parameters decreases with larger time aggregation (Figure~\ref{fig:multi_run}).
+\ian{add sentence here}
 
 \subsection{Relationship between parameterization  outputs and observed sediment discharge}
-When comparing the output variables with suspended sediment discharge measurements, the variables velocity, shear stress, and width-integrated shear stress from the subaerial system generally show better correlation with sediment discharge compared to subglacial outputs across the tested subglacial channel parameters (Figures~\ref{fig:Qw_vari} g--l and \ref{fig:multi_run}).
+
+When comparing the outputs with suspended sediment discharge measurements, the subaerial parameterizations generally show a better correlation with sediment discharge compared to subglacial outputs across the tested parameters (Figures~\ref{fig:Qw_vari} g--l and \ref{fig:multi_run}).
 The rank correlation amongst the  subaerial parameterization's three output variables remains fairly constant despite the variable impact of channel shape and consistent scaling with water discharge (Figure \ref{fig:Qw_vari} and Section~\ref{sect:fluv}).
@@
 \section{Discussion}
-\subsection{Context of sediment capacity differences between the channel types}
+\subsection{Sediment capacity differences between the channel types}
 \label{sect:dis_qsc}
@@
 
-\ian{Mauro: Can you double check this?}
+\ian{Mauro: Can you double-check this?}
 The rapid increases and decreases in subglacial sediment transport capacity may compound the variability in sediment transport already present in subaerial systems \cite<Figure~\ref{fig:Qw_vari}; >{williams1989,jerolmack2010}.
@@
 This state could result from the reduced diurnal discharge amplitude in \icesheet{} (Figures~\ref{fig:model_outs} and ~\ref{fig:Qw_vari}).
-In both cases, hydraulics could trend toward this equilibrium regime when average discharge over $1-5$ days when variability in subglacial channels  approaches that of subaerial channels (Figure~\ref{fig:multi_run}).
+In both cases, hydraulics could trend toward this equilibrium regime if the discharge is averaged discharge over $1-5$ days.
+In this case, variability in subglacial channels  approaches that of subaerial channels (Figure~\ref{fig:multi_run}).
 
-In subglacial channels, peaks occur with the fastest rate of  water discharge increase relative to the channel's growth (Figure~\ref{fig:model_outs}).
-In turn, the subglacial channels contain feedbacks that limit sediment transport capacity as the subglacial channel grows with increasing water discharge capacity.
-This effectively reduces sediment transport capacity with the stabilization of water discharge.
-Similar feedback occurs in subaerial systems \cite{phillips2016}, albeit at longer timescales, as opposed to days here \cite{gimbert2016}.
+In subglacial channels, peaks in water velocity, shear stress and width-integrated shear stress occur with the fastest rate of  water discharge increase relative to the channel's growth (Figure~\ref{fig:model_outs}).
+In turn, the subglacial channels contain feedbacks limiting sediment transport capacity as the subglacial channel grows with increasing water discharge capacity.
+This effectively reduces sediment transport capacity with the stabilization of water discharge at its peak.
+Similar feedback occurs in subaerial systems \cite{phillips2016}, albeit at timescales longer than days \cite{gimbert2016}.
 
@@
 
-The increased performance of subaerial hydraulics in representing observations occurs despite the proximity of glaciers to these measurements (Figure~\ref{fig:Qw_vari}).
+The increased performance of subaerial hydraulics outputs, compared to subglacial ones, in representing observations occurs despite the proximity of measurement locations to these glaciers.(Figure~\ref{fig:Qw_vari}).
 This could occur for two reasons.
@@
 The consistent relationship between sediment discharge capacity and open channel flow conditions could be pronounced $\sim20$\,\unit{km} downstream of the Leverett site, where a strong correlation persists between sediment plume size and the river's water discharge into the Kangerlussuaq fjord \cite{mcgrath2010}.
-In marine-terminating glacier catchments, for contrast, a less consistent relationship may occur between water discharge or melt extent and sediment plume size \cite{tedstone2012}.
+In contrast, in marine-terminating glacier catchments a less consistent relationship may occur between water discharge or melt extent and sediment plume size \cite{tedstone2012}.
 Here, ocean water pressurizes subglacial channels at the ice front \cite<e.g.>{how2017}, so the minimal correlation could result from the inconsistent relationship between subglacial sediment transport capacity and water discharge (Figure~\ref{fig:Qw_vari}).
 
-Second, the sediment discharge records could represent processes such as increased sediment accessibility when melt extends up glacier, thereby  increasing sediment and water discharge \cite<e.g.>{vergara2022}.
+Second, the sediment discharge records could represent processes such as increased sediment accessibility when melt extends up the glacier, thereby increasing sediment and water discharge \cite<e.g.>{vergara2022}.
 A supply-limited regime could well persist at many glaciers, because the sediment's threshold of motion can be reached more frequently and across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
@@
 In subglacial channels, water velocity responds quickly to changes in water discharge because their size generally evolves slowly compared to variations in water discharge.
-This makes the response of their sediment transport capacity much more sensitive to changes in water discharge when compared to subaerial channels. 
+This makes the response of sediment transport capacity much more sensitive to changes in water discharge when compared to subaerial channels. 
 Despite these differences, parameterizations of subaerial hydraulics in two glacierized catchments generally represent observed sediment discharge variations better than subglacial hydraulics.
@@
 I.D. designed the study, developed, designed, and implemented the model and experiments, and wrote the manuscript.
-D.F. provided data from Fieschergletscher.
 A.J.T. helped interpret data from Leverett Glacier and provided guidance on experiment design.
 M.A.W. provided data from Fieschergletscher and added the relationship between shear stress and water discharge (see Section~\ref{sect:dis_qsc} and \ref{sect:scaling}).
+D.F. provided data from Fieschergletscher.
 All authors provided  key inputs in writing and editing the manuscript.
@@
 SNSF Project No. PZ00P2\_202024 provided  funding for I. Delaney.
+A. Tedstone acknowledges funding from the European Research Council, award 818994 -- CASSANDRA.
 G. King provided useful comments on a previous version of this manuscript.

### 2023-09-29 Delaneyetal2023b.tex
Working on the algebra

@@
 \usepackage{placeins}
+\usepackage{amsmath}
+
 \usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
@@
     Hydraulic head &$\Delta h$&& $\mathrm{m}$\\
-    Shear stress (glacier, subaerial) & $\tau_g,\,\tau_f$&& $\mathrm{Pa \, m^{-1}}$ \\
+    Shear stress (glacier, subaerial) & $\tau_g,\,\tau_f$&& $\mathrm{Pa \, m^{-2}}$ \\
     Width-integrated shear stress (glacier, subaerial) & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$ \\
@@
 \label{sect:scaling}
-% If you get an error about an unknown bounding box, try specifying the width and height of the figure with the natwidth and natheight options.
-This is common when trying to add a PDF figure without pdflatex.
-% \begin{figure}
-% \noindent\includegraphics[natwidth=800px,natheight=600px]{samplefigure.pdf}
-%\caption{caption}
-%\label{pdffiguresample}
-%\end{figure}
-%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Optional Appendices go here
-%
-% The \appendix command resets counters and redefines section heads
-%
-% After typing \appendix
-%
-%\section{Here Is Appendix Title}
-% will show
-% A: Here Is Appendix Title
-%
-%\appendix
-%\section{Here is a sample appendix}
+
+Here we look at three possibilities, namely, rivers, steady-state
+R-channels, and pipe-flow (i.e. R-channels which do not have time to
+adjust their size to discharge conditions).
+
+Basics of hydraulics
+\begin{equation}
+  \Psi \propto f\frac{v^2}{D}
+\end{equation}
+or
+\begin{equation}
+  \Psi \propto f\frac{Q^2}{D^5}.
+\end{equation}
+Then
+\begin{equation}
+  \tau \propto v^2 = (Q/S)^2.
+\end{equation}
+
+For all flow types we assume that the discharge $Q$ is given.
+A flow in a wide river has $\Psi$ constant (given by topography), $w \gg h$, and we assume the width can be approximated by a relation $w \propto Q^\alpha$ with $\alpha\in [0,1]$ and typically $\alpha \approx 1/2$.
+A R-channel flow has $\Psi$ constant (approximated by the Shreve potential) and $S$ adjusts such that it is in steady state.
+Pipe flow has $S$ fixed and $\Psi$ adjusts.
+
+Shear stresses are given by
+\begin{align}
+  \text{wide river} \quad\quad &  \tau \propto \left(\frac{Q}{w}\right)^{2/3} \Psi^{2/3}\\
+  \text{R-channel} \quad\quad & \tau =  \\
+  \text{pipe} \quad\quad &  \tau \propto v^2 \propto\\
+\end{align}
+
+Stream power are
+
+Basics on sediment transport
+\begin{align}
+  \text{MPM} \quad\quad & Q_s \propto w \tau^{3/2}\\
+  \text{EH}  \quad\quad& Q_s \propto w \tau^{5/2}\\
+  \text{Bagnold}  \quad\quad & Q_s \propto w (\Psi Q)^{3/2}
+\end{align}
+for conditions well above transport onset threshold.
+
+% \begin{tabular}{lll}
+%  & Shear stress & Stream power\\[0pt]
+%  & \(\tau \propto\) & \(\Omega \propto\)\\[0pt]
+% \hline
+% River & \(\big(\tfrac{Q}{w}\big)^{2/3} \Psi^{2/3}\) & \(Q \Psi\)\\[3pt]
+% R-channel & \(Q^{2/5} \, \Psi^{4/5}\) & \(Q \Psi\)\\
+% Pipe & \(Q^2\) & \(Q^3\)\\
+% \end{tabular}
+
+\noindent
+\begin{tabular}{llllll}
+Kind & Fixed & Determined via D-W & Additional relations & Shear stress & Stream p.\\[0pt]
+ &  &  &  & \(\tau \propto\) & \(\Omega \propto\)\\[0pt]
+\hline
+River & \(\Psi\) & \(H \propto (f \Psi^{-1} Q^{2-2\alpha})^{1/3}\) & \(w\propto Q^\alpha\), \(S=WH\), \(D_h=H\) & \(f^{1/3} Q^{2/3-2\alpha/3}  \Psi^{2/3}\) & \(Q \Psi\)\\[0pt]
+R-ch. & \(\Psi\) & \(S \propto (f \Psi^{-1} Q^2)^{2/5}\) & \(D_h\propto S^{1/2}\) & \(f^{1/5} Q^{2/5} \, \Psi^{4/5}\) & \(Q \Psi\)\\[0pt]
+Pipe & \(S\) & \(\Psi \propto f Q^2 S^{-5/2}\) & \(D_h\propto S^{1/2}\) & \(f Q^2 S^{-2}\) & \(f Q^3 S^{-5/2}\)\\[0pt]
+\end{tabular}
+
+\vspace{10mm}
+
+\noindent
+\begin{tabular}{llll}
+ & MPM & EH & Bagnold\\[0pt]
+\(Q_s \propto\) & \$W$\backslash$, \(\tau\)\textsuperscript{3/2} & \(W\, \tau^{5/2}\) & \(W\, \Omega^{3/2} H^{-2/3}\)\\[0pt]
+\hline
+River & \(f^{1/2}\, Q \, \Psi\) & \(f^{5/6}\, Q^{5/3 - 2\alpha/3} \, \Psi^{5/3}\) & \(f^{-2/9}\, Q^{13\alpha/9 - 19/18} \, \Psi^{31/18}\)\\[0pt]
+R-ch. & \(f^{1/2}\, Q \, \Psi\) & \(f^{7/10}\, Q^{7/5}\, \Psi^{9/5}\) & \(f^{1/15}\, Q^{49/30}\, \Psi^{43/30}\)\\[0pt]
+Pipe & \(f^{3/2}\, Q^3 \, S^{-5/2}\) & \(f^{5/2}\, Q^5\, S^{-9/2}\) & \(f^{3/2} \, Q^{9/2} \, S^{-43/12}\)\\[0pt]
+\end{tabular}\\
+\mauro{That negative exponent of $f$ for Bagnold-River is pretty crazy! Not sure what that is about.}
+
 \end{document}

### 2023-09-30 Delaneyetal2023b.tex
Now added w*tau to table too
Scaling-text slowly getting there

--- 2023-09-26 Delaneyetal2023b.tex
@@
 \newpage
-\section{Mauro's bit about shear stress}
+\section{Scalings for sediment transport in different drainage regimes}
 \label{sect:scaling}
-% If you get an error about an unknown bounding box, try specifying the width and height of the figure with the natwidth and natheight options.
-This is common when trying to add a PDF figure without pdflatex.
-% \begin{figure}
-% \noindent\includegraphics[natwidth=800px,natheight=600px]{samplefigure.pdf}
-%\caption{caption}
-%\label{pdffiguresample}
-%\end{figure}
+
+\mauro{Notation: either use $H$ and $W$ for the river, or otherwise, somehow free $h$. Below I use $H$ and $W$ for now.}
+
+Here we look at three kind of drainage regimes, namely, rivers, steady-state R-channels, and pipe-flow (i.e.
+R-channels which do not have time to adjust their size to discharge conditions).
+We calculate how in these regimes shear stress and sediment flux scale with respect to a given discharge for three different transport formulas: MPM \cite{meyer1948}, EH \cite{engelund1967}, and Bagnold \cite{Bagnold1980}.
+
+A few basic hydraulic relations are used in the following.  The Darcy-Weisbach equation can be stated
+\begin{equation}
+  \label{eq:DW}
+  \Psi \propto f\frac{Q^2}{D_h S^2},
+\end{equation}
+with $\Psi$ the head gradient and friction factor $f$.  The shear stress and stream power are, respectively,
+\begin{equation}
+    \label{eq:tau-omega}
+  \tau \propto f v^2 = f \left(\frac{Q}{S}\right)^2, \quad  \Omega \propto \Psi Q.
+\end{equation}
 %
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Optional Appendices go here
-%
-% The \appendix command resets counters and redefines section heads
-%
-% After typing \appendix
-%
-%\section{Here Is Appendix Title}
-% will show
-% A: Here Is Appendix Title
-%
-%\appendix
-%\section{Here is a sample appendix}
+Sediment discharge is given by the MPM, EH and Bagnold formulations as, respectively,
+\begin{equation}
+  Q_s \propto w\, \tau^{3/2}, \quad Q_s \propto w\, \tau^{5/2}, \quad Q_s \propto w\, \Omega^{3/2}
+\end{equation}
+for conditions well above transport onset threshold.
+
+We assume a wide river with depth $H$, such that $W \gg H$ and $D_h\approx 4H$, that has a constant head gradient ($\Psi$) given by the topography and we assume that its width can be approximated by a relation $w \propto Q^\alpha$ with $\alpha\in [0,1]$; typically $\alpha \approx 1/2$, $\alpha=0$ or $\alpha=1$ correspond a river of constant width or depth, respectively.
+
+We assume that a steady state R-channel has a constant $\Psi$ (approximated by the gradient of the Shreve \citeyear{shreve1972} potential) and that $S$ adjusts such that it is in steady state.
+Conversely, pipe flow has $S$ fixed and $\Psi$ adjusts.
+We assume that for pipe flow, which occurs when a R-channel is subjected to rapid discharge variations such that it cannot adjust its size significantly, the cross-sectional area $S$ is fixed and that $\Psi$ adjusts.
+For both R-channel and pipe flow, we assume that $D_h \propto S^{1/2}$.
+
+With these assumptions the Darcy-Weisbach equation~\eqref{eq:DW} can be solved for the not-fixed quantity: $H$ for river, $S$ for R-channel, and $\Psi$ for pipe-flow.
+Then, using equations~\eqref{eq:tau-omega}, the shear stress and stream power can be calculated.
+These resulted are summarised in Table~\ref{tab:eqs1}.
+Now sediment transport can be calculated for all regimes and all transport formulas.
+The results of the nine combinations are presented in Table~\ref{tab:Qs} and also in Table~\ref{tab:Qs2} where the fractions in the exponents are approximately given by decimal numbers for ease of comparison.
+
+\begin{table}
+  \caption{}
+  \label{tab:eqs1}
+  \small
+\begin{tabular}{llllll}
+Regime & Fixed & Determined via D-W & Additional relations & Shear stress & Stream p.\\[0pt]
+ &  &  &  & \(\tau \propto\) & \(\Omega \propto\)\\[0pt]
+\hline
+River & \(\Psi\) & \(H \propto f^{1/3}\, Q^{2/3-2\alpha/3} \, \Psi^{-1/3}\) & \(W\,\propto Q^\alpha\), \(S=WH\), \(D_h=H\) & \(f^{1/3} Q^{2/3-2\alpha/3}  \Psi^{2/3}\) & \(Q \Psi\)\\[0pt]
+R-ch. & \(\Psi\) & \(S\, \propto f^{2/5}\, Q^{4/5} \, \Psi^{-2/5}\) & \(D_h\propto S^{1/2}\) & \(f^{1/5} Q^{2/5} \, \Psi^{4/5}\) & \(Q \Psi\)\\[0pt]
+Pipe & \(S\) & \(\Psi \propto f \, Q^2\, S^{-5/2}\) & \(D_h\propto S^{1/2}\) & \(f Q^2 S^{-2}\) & \(f Q^3 S^{-5/2}\)\\[0pt]
+\end{tabular}
+\end{table}
+
+\begin{table}
+  \caption{
+    \mauro{That negative exponent of $f$ for Bagnold-River is pretty crazy! Not sure what that is about.}
+    }
+  \label{tab:Qs}
+\begin{tabular}{lllll}
+ & Width*\(\tau\) & MPM & EH & Bagnold\\[0pt]
+ & \(W\, \tau\) & \(Q_s \propto W\, \tau^{3/2}\) & \(Q_s \propto W\, \tau^{5/2}\) & \(Q_s \propto W\, \Omega^{3/2} H^{-2/3}\)\\[0pt]
+\hline
+River & \(f^{1/3}\, Q^{2/3+\alpha/3}\,  \Psi^{2/3}\) & \(f^{1/2}\, Q \, \Psi\) & \(f^{5/6}\, Q^{5/3 - 2\alpha/3} \, \Psi^{5/3}\) & \(f^{-2/9}\, Q^{13\alpha/9 - 19/18} \, \Psi^{31/18}\)\\[0pt]
+R-ch. & \(f^{2/5}\, Q^{4/5} \, \Psi^{3/5}\) & \(f^{1/2}\, Q \, \Psi\) & \(f^{7/10}\, Q^{7/5}\, \Psi^{9/5}\) & \(f^{1/15}\, Q^{49/30}\, \Psi^{43/30}\)\\[0pt]
+Pipe & \(f \, Q^2 \, S^{-1}\) & \(f^{3/2}\, Q^3 \, S^{-5/2}\) & \(f^{5/2}\, Q^5\, S^{-9/2}\) & \(f^{3/2} \, Q^{9/2} \, S^{-43/12}\)\\[0pt]
+\end{tabular}
+\end{table}
+
+\mauro{TODO: make third table C3.  Discuss results briefly.}
+
 \end{document}

### 2023-09-30 Delaneyetal2023b.tex
Scaling-text now done.  Except need to recheck that Bagnold is calcualted right as those formulas are way off

--- 2023-09-26 Delaneyetal2023b.tex
@@
 \newpage
-\section{Mauro's bit about shear stress}
+\section{Formulas for sediment transport in different drainage regimes}
 \label{sect:scaling}
-% If you get an error about an unknown bounding box, try specifying the width and height of the figure with the natwidth and natheight options.
-This is common when trying to add a PDF figure without pdflatex.
-% \begin{figure}
-% \noindent\includegraphics[natwidth=800px,natheight=600px]{samplefigure.pdf}
-%\caption{caption}
-%\label{pdffiguresample}
-%\end{figure}
+
+\mauro{Notation: either use $H$ and $W$ for the river, or otherwise, somehow free $h$. Below I use $H$ and $W$ for now.}
+
+Here we look at three kind of drainage regimes, namely, rivers, steady-state R-channels, and pipe-flow (i.e.
+R-channels which do not have time to adjust their size to discharge conditions).
+We calculate how in these regimes sediment flux scales with respect to a given discharge for three different transport formulas: MPM \cite{meyer1948}, EH \cite{engelund1967}, and Bagnold \cite{Bagnold1980}; additionally we state the width integrated shear stress used as proxy for sediment transport throughout the manuscript.
+
+A few basic hydraulic relations are used in the following.  The Darcy-Weisbach equation can be stated as
+\begin{equation}
+  \label{eq:DW}
+  \Psi \propto f\frac{Q^2}{D_h S^2},
+\end{equation}
+with $\Psi$ the head gradient and friction factor $f$.  The shear stress and stream power are, respectively,
+\begin{equation}
+    \label{eq:tau-omega}
+  \tau \propto f v^2 = f \left(\frac{Q}{S}\right)^2, \quad  \Omega \propto \Psi Q.
+\end{equation}
 %
+Sediment discharge is given by the MPM, EH and Bagnold formulations as, respectively,
+\begin{equation}
+  Q_s \propto w\, \tau^{3/2}, \quad Q_s \propto w\, \tau^{5/2}, \quad Q_s \propto w\, \Omega^{3/2}
+\end{equation}
+for conditions well above the transport onset threshold.
+
+The river is assumed to have a width $W$ much greater than its depth $H$, such that $D_h\approx 4H$, and to have a constant head gradient ($\Psi$) given by the topography.
+Further, we assume that its width can be approximated by a relation $w \propto Q^\alpha$ with $\alpha\in [0,1]$; typically $\alpha \approx 1/2$, $\alpha=0$ or $\alpha=1$ correspond a river of constant width or depth, respectively.
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Optional Appendices go here
+For a steady state R-channel, we assume that it has a constant $\Psi$ (approximated by the gradient of the Shreve \citeyear{shreve1972} potential) and that $S$ adjusts such that it is in steady state corresponding to $\Psi$ and $Q$.
+%Conversely, pipe flow has $S$ fixed and $\Psi$ adjusts to satisfy the given $Q$.
 %
-% The \appendix command resets counters and redefines section heads
+For pipe flow, which occurs when a R-channel is subjected to rapid discharge variations such that it cannot adjust its size significantly,  we assume that the cross-sectional area $S$ is fixed and that $\Psi$ adjusts to meet the specified $Q$.
+For both R-channel and pipe flow, we assume that $D_h \propto S^{1/2}$.
+
+With these assumptions the Darcy-Weisbach equation~\eqref{eq:DW} can be solved for the not-fixed quantity: $H$ for a river, $S$ for a R-channel, and $\Psi$ for pipe-flow.
+Then, using equations~\eqref{eq:tau-omega}, the shear stress and stream power can be calculated.
+These results are summarised in Table~\ref{tab:eqs1}.
+Now width integrated shear stress and sediment transport can be calculated for all regimes and for all transport formulas.
+The results of the 12 combinations are presented in Table~\ref{tab:Qs} and also in Table~\ref{tab:Qs2} where the fractions in the exponents are approximately given by decimal numbers for ease of comparison.
+
+\mauro{Bagnold gives pretty crazy results. Check again!}
+
+Table~\ref{tab:Qs2} shows that the proxy $W\tau$ as well as the MPM and EH sediment transport formulas give roughly a similar scaling of $Q_s$ on $\Psi$ or $S$ and on $Q$.
+The total transport formula EH giving a stronger dependence on all variables as should be expected duto to the larger exponent on $\tau$ of 2.5 versus 1.5 for the others.
+The exponent for the width scaling $\alpha$ only impacts the EH relation, with its value almost certainly around or bigger than 0.5 for a braided proglacial stream.
 %
-% After typing \appendix
+The sediment discharge in the steady state R-channel scales very similarly to the river case, and almost identically for the $\alpha=0.5$ case.
+However, note that the head gradients $\Psi$ are likely much higher for comparable $Q$ in a glacial setting than a river setting, as is well described in \citeA{alley1997}, and thus transport rates may still be much higher in a steady state R-channel.
 %
-%\section{Here Is Appendix Title}
-% will show
-% A: Here Is Appendix Title
-%
-%\appendix
-%\section{Here is a sample appendix}
+Furthermore, R-channels will rarely operate in steady state as variations in discharge, in particular on the diurnal timescale or the timescale of severe rain or melting events, are too fast for such a channel to reach a steady state.
+In such cases they operate more like a pipe of fixed cross-section.
+Table~\ref{tab:Qs2} shows that then the sediment transport scales much more severely with discharge, with the exponent of $Q$ being between 2 and 5, than in the other two regimes when that exponent is at most 1.3.
+Thus fluctuations of discharge on short timescales (on the order of a day) have the potential to cause very high sediment transport conditions.
+
+\begin{table}
+  \caption{Relations for hydraulic variables for the three drainage regimes: river, R-channel and pipe-flow.
+Darcy-Weisbach equation is abbreviated with ``D-W'', and stream power with ``Stream p.''.
+}
+  \label{tab:eqs1}
+  \small
+\begin{tabular}{llllll}
+Regime & Fixed & Determined via D-W & Additional relations & Shear stress & Stream p.\\
+ &  &  &  & \(\tau \propto\) & \(\Omega \propto\)\\
+\hline
+River & \(\Psi\) & \(H \propto f^{1/3}\, Q^{2/3-2\alpha/3} \, \Psi^{-1/3}\) & \(W\,\propto Q^\alpha\), \(S=WH\), \(D_h=H\) & \(f^{1/3} Q^{2/3-2\alpha/3}  \Psi^{2/3}\) & \(Q \Psi\)\\
+R-ch. & \(\Psi\) & \(S\, \propto f^{2/5}\, Q^{4/5} \, \Psi^{-2/5}\) & \(D_h\propto S^{1/2}\) & \(f^{1/5} Q^{2/5} \, \Psi^{4/5}\) & \(Q \Psi\)\\
+Pipe & \(S\) & \(\Psi \propto f \, Q^2\, S^{-5/2}\) & \(D_h\propto S^{1/2}\) & \(f Q^2 S^{-2}\) & \(f Q^3 S^{-5/2}\)\\
+\end{tabular}
+\end{table}
+
+\begin{table}
+  \caption{Sediment transport proxy ($W\tau$) and rates for the three considered different transport formulas: MPM \cite{meyer1948}, EH \cite{engelund1967}, and Bagnold \cite{Bagnold1980}.
+    }
+  \label{tab:Qs}
+\begin{tabular}{lllll}
+ & Width*\(\tau\) & MPM & EH & Bagnold\\
+ & \(W\, \tau\) & \(Q_s \propto W\, \tau^{3/2}\) & \(Q_s \propto W\, \tau^{5/2}\) & \(Q_s \propto W\, \Omega^{3/2} H^{-2/3}\)\\
+\hline
+River & \(f^{1/3}\, Q^{2/3+\alpha/3}\,  \Psi^{2/3}\) & \(f^{1/2}\, Q \, \Psi\) & \(f^{5/6}\, Q^{5/3 - 2\alpha/3} \, \Psi^{5/3}\) & \(f^{-2/9}\, Q^{13\alpha/9 - 19/18} \, \Psi^{31/18}\)\\
+R-ch. & \(f^{2/5}\, Q^{4/5} \, \Psi^{3/5}\) & \(f^{1/2}\, Q \, \Psi\) & \(f^{7/10}\, Q^{7/5}\, \Psi^{9/5}\) & \(f^{1/15}\, Q^{49/30}\, \Psi^{43/30}\)\\
+Pipe & \(f \, Q^2 \, S^{-1}\) & \(f^{3/2}\, Q^3 \, S^{-5/2}\) & \(f^{5/2}\, Q^5\, S^{-9/2}\) & \(f^{3/2} \, Q^{9/2} \, S^{-43/12}\)\\
+\end{tabular}
+\end{table}
+
+\begin{table}
+  \caption{Sediment transport proxy ($W\tau$) and rates for the three considered different transport formulas: MPM \cite{meyer1948}, EH \cite{engelund1967}, and Bagnold \cite{Bagnold1980}.
+    }
+  \label{tab:Qs2}
+\begin{tabular}{lllll}
+ & Width*\(\tau\) & MPM & EH & Bagnold\\
+ & \(W\, \tau\) & \(Q_s \propto W\, \tau^{3/2}\) & \(Q_s \propto W\, \tau^{5/2}\) & \(Q_s \propto W\, \Omega^{3/2} H^{-2/3}\)\\
+\hline
+River (\(\alpha=0\)) & \(f^{0.7}\, Q^{0.7}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{1.7} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{-1.1} \, \Psi^{1.7}\)\\
+River (\(\alpha=0.5\)) & \(f^{0.7}\, Q^{0.8}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\,\Psi\) & \(f^{0.8}\, Q^{1.3} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{-0.3} \, \Psi^{1.7}\)\\
+River (\(\alpha=1\)) & \(f^{0.7}\, Q^{1.0}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{0.7} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{0.4} \, \Psi^{1.7}\)\\[3pt]
+R-ch. & \(f^{0.4}\, Q^{0.8} \, \Psi^{0.6}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.7}\, Q^{1.4}\, \Psi^{1.8}\) & \(f^{0.07}\, Q^{1.6}\, \Psi^{1.4}\)\\
+Pipe & \(f \quad\, Q^{2\phantom{.0}} \, S^{-1}\) & \(f^{1.5}\, Q^3 \, S^{-2.5}\) & \(f^{2.5}\, Q^{5\phantom{.0}}\, S^{-4.5}\) & \(f^{1.5} \, Q^{4.5} \, S^{-3.6}\)\\
+\end{tabular}
+\end{table}
+
 \end{document}

### 2023-10-02 Delaneyetal2023b.tex

@@
 \begin{equation}
-  Q_s \propto w\, \tau^{3/2}, \quad Q_s \propto w\, \tau^{5/2}, \quad Q_s \propto w\, \Omega^{3/2}
+  Q_s \propto W\, \tau^{3/2}, \quad Q_s \propto W\, \tau^{5/2}, \quad Q_s \propto W \left(\frac{\Omega}{W}\right)^{3/2} H^{-2/3}
 \end{equation}
@@
 The river is assumed to have a width $W$ much greater than its depth $H$, such that $D_h\approx 4H$, and to have a constant head gradient ($\Psi$) given by the topography.
-Further, we assume that its width can be approximated by a relation $w \propto Q^\alpha$ with $\alpha\in [0,1]$; typically $\alpha \approx 1/2$, $\alpha=0$ or $\alpha=1$ correspond a river of constant width or depth, respectively.
+Further, we assume that its width can be approximated by a relation $W \propto Q^\alpha$ with $\alpha\in [0,1]$; typically $\alpha \approx 1/2$ (a so-called regime channel), $\alpha=0$ or $\alpha=1$ correspond a river of constant width (a slot canyon) or depth (no natural equivalent), respectively, i.e.
+the two end members.
 %
@@
 
-\mauro{Bagnold gives pretty crazy results. Check again!}
+Table~\ref{tab:Qs2} shows that the proxy $W\tau$ as well as the MPM, EH and Bagnold sediment transport formulas give roughly a similar scaling of $Q_s$ on $\Psi$ or $S$ and on $Q$.
+Remarkably, the Bagnold formula has a negative exponent for $f$ in all but the pipe-flow regime, which seems rather unexpected.
+The total transport formula EH gives a slightly stronger dependence on all variables as should be expected due to to the larger exponent on $\tau$ of 2.5 versus 1.5 for the others.
+Albeit, for pipe-flow Bagnold is close to EH.
+Conversely, the proxy used within this publication scales only as $Q^2$ for pipe-flow, whereas the others scale at least with $Q^3$.
+The exponent for the width scaling $\alpha$ only impacts the EH relation in any meaningful way; but for the value of $\alpha$ around 0.5, as is suitable for a braided proglacial stream and indeed most rivers, the exponent on $Q$ is only slightly above what the other relations give.
 
-Table~\ref{tab:Qs2} shows that the proxy $W\tau$ as well as the MPM and EH sediment transport formulas give roughly a similar scaling of $Q_s$ on $\Psi$ or $S$ and on $Q$.
-The total transport formula EH giving a stronger dependence on all variables as should be expected duto to the larger exponent on $\tau$ of 2.5 versus 1.5 for the others.
-The exponent for the width scaling $\alpha$ only impacts the EH relation, with its value almost certainly around or bigger than 0.5 for a braided proglacial stream.
-%
-The sediment discharge in the steady state R-channel scales very similarly to the river case, and almost identically for the $\alpha=0.5$ case.
-However, note that the head gradients $\Psi$ are likely much higher for comparable $Q$ in a glacial setting than a river setting, as is well described in \citeA{alley1997}, and thus transport rates may still be much higher in a steady state R-channel.
-%
+The sediment discharge in the steady state R-channel scales very similarly to the river case for all relations and virtually identically for the $\alpha=0.5$ case.
+However, note that the head gradients $\Psi$ are likely higher for comparable $Q$ in a ice-sheet marginal or alpine glacier setting than in a river setting, as is well described in \citeA{alley1997}, and thus transport rates may still be much higher in a steady state R-channel.
+
 Furthermore, R-channels will rarely operate in steady state as variations in discharge, in particular on the diurnal timescale or the timescale of severe rain or melting events, are too fast for such a channel to reach a steady state.
 In such cases they operate more like a pipe of fixed cross-section.
-Table~\ref{tab:Qs2} shows that then the sediment transport scales much more severely with discharge, with the exponent of $Q$ being between 2 and 5, than in the other two regimes when that exponent is at most 1.3.
+Table~\ref{tab:Qs2} shows that in such a situation, the sediment transport scales much more severely with discharge, with the exponent of $Q$ being likely between 3 and 5, compared to other two regimes when that exponent is at most 1.3.
 Thus fluctuations of discharge on short timescales (on the order of a day) have the potential to cause very high sediment transport conditions.
@@
 River & \(\Psi\) & \(H \propto f^{1/3}\, Q^{2/3-2\alpha/3} \, \Psi^{-1/3}\) & \(W\,\propto Q^\alpha\), \(S=WH\), \(D_h=H\) & \(f^{1/3} Q^{2/3-2\alpha/3}  \Psi^{2/3}\) & \(Q \Psi\)\\
-R-ch. & \(\Psi\) & \(S\, \propto f^{2/5}\, Q^{4/5} \, \Psi^{-2/5}\) & \(D_h\propto S^{1/2}\) & \(f^{1/5} Q^{2/5} \, \Psi^{4/5}\) & \(Q \Psi\)\\
-Pipe & \(S\) & \(\Psi \propto f \, Q^2\, S^{-5/2}\) & \(D_h\propto S^{1/2}\) & \(f Q^2 S^{-2}\) & \(f Q^3 S^{-5/2}\)\\
+R-ch. & \(\Psi\) & \(S\, \propto f^{2/5}\, Q^{4/5} \, \Psi^{-2/5}\) & \(D_h\propto W \propto H \propto S^{1/2}\) & \(f^{1/5} Q^{2/5} \, \Psi^{4/5}\) & \(Q \Psi\)\\
+Pipe & \(S\) & \(\Psi \propto f \, Q^2\, S^{-5/2}\) & \(D_h\propto W \propto H \propto S^{1/2}\) & \(f Q^2 S^{-2}\) & \(f Q^3 S^{-5/2}\)\\
 \end{tabular}
@@
  & Width*\(\tau\) & MPM & EH & Bagnold\\
- & \(W\, \tau\) & \(Q_s \propto W\, \tau^{3/2}\) & \(Q_s \propto W\, \tau^{5/2}\) & \(Q_s \propto W\, \Omega^{3/2} H^{-2/3}\)\\
+ & \(W\, \tau\) & \(Q_s \propto W\, \tau^{3/2}\) & \(Q_s \propto W\, \tau^{5/2}\) & \(Q_s \propto W^{-1/2}\, \Omega^{3/2} H^{-2/3}\)\\
 \hline
-River & \(f^{1/3}\, Q^{2/3+\alpha/3}\,  \Psi^{2/3}\) & \(f^{1/2}\, Q \, \Psi\) & \(f^{5/6}\, Q^{5/3 - 2\alpha/3} \, \Psi^{5/3}\) & \(f^{-2/9}\, Q^{13\alpha/9 - 19/18} \, \Psi^{31/18}\)\\
-R-ch. & \(f^{2/5}\, Q^{4/5} \, \Psi^{3/5}\) & \(f^{1/2}\, Q \, \Psi\) & \(f^{7/10}\, Q^{7/5}\, \Psi^{9/5}\) & \(f^{1/15}\, Q^{49/30}\, \Psi^{43/30}\)\\
-Pipe & \(f \, Q^2 \, S^{-1}\) & \(f^{3/2}\, Q^3 \, S^{-5/2}\) & \(f^{5/2}\, Q^5\, S^{-9/2}\) & \(f^{3/2} \, Q^{9/2} \, S^{-43/12}\)\\
+River & \(f^{1/3}\, Q^{2/3+\alpha/3}\,  \Psi^{2/3}\) & \(f^{1/2}\, Q \, \Psi\) & \(f^{5/6}\, Q^{5/3 - 2\alpha/3} \, \Psi^{5/3}\) & \(f^{-2/9}\, Q^{19/18-\alpha/18} \, \Psi^{31/18}\)\\
+R-ch. & \(f^{2/5}\, Q^{4/5} \, \Psi^{3/5}\) & \(f^{1/2}\, Q \, \Psi\) & \(f^{7/10}\, Q^{7/5}\, \Psi^{9/5}\) & \(f^{-7/30}\, Q^{31/30}\, \Psi^{26/15}\)\\
+Pipe & \(f \, Q^2 \, S^{-1}\) & \(f^{3/2}\, Q^3 \, S^{-5/2}\) & \(f^{5/2}\, Q^5\, S^{-9/2}\) & \(f^{3/2} \, Q^{9/2} \, S^{-14/3}\)\\
 \end{tabular}
@@
 \begin{table}
-  \caption{Sediment transport proxy ($W\tau$) and rates for the three considered different transport formulas: MPM \cite{meyer1948}, EH \cite{engelund1967}, and Bagnold \cite{Bagnold1980}.
+  \caption{As Table~\ref{tab:Qs} but with the fractional exponents stated as rounded decimal numbers.
+For the River regime the exponents are for displayed for the likely width-exponent $\alpha=1$, as well as its end members 0 (slot canyon) and 1 (only width increases and not depth).
     }
@@
  & Width*\(\tau\) & MPM & EH & Bagnold\\
- & \(W\, \tau\) & \(Q_s \propto W\, \tau^{3/2}\) & \(Q_s \propto W\, \tau^{5/2}\) & \(Q_s \propto W\, \Omega^{3/2} H^{-2/3}\)\\
+ & \(W\, \tau\) & \(Q_s \propto W\, \tau^{3/2}\) & \(Q_s \propto W\, \tau^{5/2}\) & \(Q_s \propto W^{-1/2}\, \Omega^{3/2} H^{-2/3}\)\\
 \hline
-River (\(\alpha=0\)) & \(f^{0.7}\, Q^{0.7}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{1.7} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{-1.1} \, \Psi^{1.7}\)\\
-River (\(\alpha=0.5\)) & \(f^{0.7}\, Q^{0.8}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\,\Psi\) & \(f^{0.8}\, Q^{1.3} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{-0.3} \, \Psi^{1.7}\)\\
-River (\(\alpha=1\)) & \(f^{0.7}\, Q^{1.0}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{0.7} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{0.4} \, \Psi^{1.7}\)\\[3pt]
-R-ch. & \(f^{0.4}\, Q^{0.8} \, \Psi^{0.6}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.7}\, Q^{1.4}\, \Psi^{1.8}\) & \(f^{0.07}\, Q^{1.6}\, \Psi^{1.4}\)\\
-Pipe & \(f \quad\, Q^{2\phantom{.0}} \, S^{-1}\) & \(f^{1.5}\, Q^3 \, S^{-2.5}\) & \(f^{2.5}\, Q^{5\phantom{.0}}\, S^{-4.5}\) & \(f^{1.5} \, Q^{4.5} \, S^{-3.6}\)\\
+River (\(\alpha=0\)) & \(f^{0.7}\, Q^{0.7}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{1.7} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.1} \, \Psi^{1.7}\)\\
+River (\(\alpha=0.5\)) & \(f^{0.7}\, Q^{0.8}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{1.3} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\
+River (\(\alpha=1\)) & \(f^{0.7}\, Q^{1.0}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{0.7} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\[3pt]
+R-ch. & \(f^{0.4}\, Q^{0.8} \, \Psi^{0.6}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.7}\, Q^{1.4}\, \Psi^{1.8}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\
+Pipe & \(f \,\quad Q^{2\phantom{.0}} \, S^{-1}\) & \(f^{1.5}\, Q^3 \, S^{-2.5}\) & \(f^{2.5}\, Q^{5\phantom{.0}}\, S^{-4.5}\) & \(f^{1.5} \,\,\,\,\, Q^{4.5} \, S^{-4.7}\)\\
 \end{tabular}

### 2023-10-02 Delaneyetal2023b.tex
integrating appendix and updating text

@@
 \usepackage{amsmath}
+\usepackage{makecell}
 
@@
 
-\title{Water discharge quantity has a variable impact on sediment transport capacity in subglacial channels}
+\title{Natural variations in water discharge quantity has an inconsistent impact on sediment transport capacity in subglacial channels}
 
-\authors{Ian Delaney\affil{1},  Andrew J. Tedstone\affil{2},   Mauro A. Werder\affil{3,4},  Daniel Farinotti\affil{3,4}, }
+\authors{Ian Delaney\affil{1},  Andrew J. Tedstone\affil{2},   Mauro A. Werder\affil{3,4},  Daniel Farinotti\affil{3,4} }
 
@@
 Therefore, subglacial water flow behaves more like pipe flow, and changes in water discharge $Q$ are mainly accommodated with increased or decreased water velocity $v$ \cite{alley1997} because the size of the channel $S$ changes on a much longer timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
-It follows that an increase in water discharge $Q$ will cause the shear stress across the channel $\tau_t$ to respond with water velocity $v$, as opposed to subaerial channels where channel width $w_c$ evolves as well.
+%It follows that an increase in water discharge $Q$ will cause the shear stress across the channel $\tau_t$ to respond with water velocity $v$, as opposed to subaerial channels where channel width $w_c$ evolves as well.
 
@@
     \includegraphics[width=0.8\linewidth]{Fig1.pdf}
-    \caption{Sketch for the different responses that subglacial and subaerial channels have to an increase in water discharge.
+    \caption{Sketch for the different responses that subglacial and subaerial channels to increased water discharge.
       Water velocity magnitudes in the subglacial, $v_g$, and subaerial, $v_f$, channels are shown by arrow length.
       $S_g$ and $S_f$ represent the wetted area in subglacial and subaerial channels, respectively.
-The subglacial channel widths $w_{cg}$ remain unchanged, while the subaerial channel widths $w_{cf}$ evolve.
-Shear stress $\tau$ is responsible for the mobilization of sediment and increases with the number of makers at the channel bed.
-      Note that in the subaerial channel parameterization, channels have rectangular shapes with their width much larger than their depth (Section~\ref{sect:fluv}).}
+      The subglacial channel widths $w_{cg}$ remain unchanged, while the subaerial channel widths $w_{cf}$ evolve. Shear stress $\tau$ is responsible for the mobilization of sediment.
+      %Note that in the subaerial channel parameterization, channels have rectangular shapes with their width much larger than their depth (Section~\ref{sect:fluv}).
+    }
     \label{fig:cartoon}
@@
 As a result, sediment mobilization in subaerial and subglacial channels responds differently to changing water discharge.
-These differences have been implicitly included in a wide range of available models that  quantify sediment transport from catchments in both subglacial and subaerial channels  \cite<e.g.>{walder1994,alley1997,tucker1997,creyts2013,delaney2019,hewitt2019,wickert2019}.
-The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \cite<e.g.>{lane2016,perolo2018}.
+These differences are implicitly included in a wide range of available models that  quantify sediment transport in both subglacial and subaerial channels  \cite<e.g.>{walder1994,alley1997,tucker1997,creyts2013,delaney2019,hewitt2019,wickert2019}.
+The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \cite<e.g.>{perolo2018}.
 Additionally, the variable response of sediment transport capacity to water discharge in the two systems may affect the interpretation of sediment transport records in glacial systems \cite<e.g.>{muller1968,richards2003,swift2005,ganti2016}.
@@
 Data is used from \citeA{tedstone2017} and \citeA{hawkings2014} at a $5$\,\unit{min} time interval from May 28, 2012 to August 8, 2012.
-Two gaps in suspended sediment data of roughly one day in length are interpolated across with a linear spline.
+Two gaps (14.07.2012 to 15.07.2012 and 16.07.2012 to 17.07.2021) in suspended sediment data interpolated using a linear spline.
 
@@
 
-With knowledge of wetted area $S_g$, the shear stress, $\tau_g$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
+With wetted area $S_g$, the shear stress, $\tau_g$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
 \begin{equation}
@@
 
-where $w_{cg}$ is the flat channel floor width:
+where  channel floor width $w_{cg}$ is
 \begin{equation}
@@
   \label{eq:wcf}
-  w_{cf} = k \, Q^\frac{1}{2},
+  w_{cf} = k \, Q^{\alpha},
 \end{equation}
 %
-where $k$ is a constant and $\frac{1}{2}$ is a commonly chosen exponent \cite{leopold1953}.
-%\mauro{This is for so called ``regime channels''}
+where $k$ is a constant and $\alpha=\frac{1}{2}$ is a commonly chosen exponent \cite{leopold1953}.
 Water velocity, $v_f$, is given by rearranging Equation \ref{eq:tau} as
@@
 The parameterizations above are applied to hydrological records from the Fieschergletscher (\alpine) and the Leverett Glacier (\icesheet).
-The subglacial and subaerial parameterizations represent the subglacial flow of water exiting a glacier and transitioning to subaerial flow, as it moves through the catchment.
+The subglacial and subaerial parameterizations represent the subglacial flow of water exiting a glacier and transitioning to subaerial flow.
 Outputs of the parameterizations are meant to represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
@@
 In both subaerial channels, the bed slope is $\nabla z_c = 0.05$.
-Water velocity, shear stress, and width-integrated shear stress in the subglacial and subaerial channels from this formulation are presented below.
+% Water velocity, shear stress, and width-integrated shear stress in the subglacial and subaerial channels from this formulation are presented below.
 
-Second, to characterize the variability in sediment discharge capacity in subglacial compared to subaerial channels further, $100$  parameterizations are applied to a range of different channel geometry factors ($\beta$ and $k$) and friction factors ($f_r$ and $f_p$).
+Second, to further characterize the variability in sediment discharge capacity in subglacial compared to subaerial channels, $100$  parameterizations are applied to a range of different channel geometry factors ($\beta$ and $k$) and friction factors ($f_r$ and $f_p$).
 Runs with parameter combinations are culled if their mean subglacial water velocity over the season lies outside of the range between $0.75$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $2$\,\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity lies outside of the range $0.3$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with measurements \cite<e.g.>{werder2010b,magnusson2012,chandler2013}.
-The standard deviation of an output over the course of the season is used to quantify its variability.
+Variability of an output is represented by the  standard deviation of an output over the course of the season.
 
-Lastly,  outputs are compared with the sediment discharge records of the glaciers using Spearman rank correlation, that is the correlation with respect to the ordering of values from the outputs and the observations.
+Lastly, outputs are compared with the sediment discharge records of the glaciers using Spearman rank correlation (i.e.
+correlation depends on the ordering of values from the outputs and the observations).
 Rank correlation reduces the impact of the non-linear response of sediment transport capacity to the hydrological forcing.
-Outputs and observations, which have a native time resolution of $5$\,\unit{min}, are aggregated across timescales from $15$\,\unit{min} to $15$\,\unit{days} by averaging values one half of the timescale before and after a given point time.
+Outputs and observations at $15$\,\unit{min} intervals are aggregated across timescales of up to $15$\,\unit{days} by averaging values one half of the timescale before and after a given point time.
 
@@
   \includegraphics[width=0.9\linewidth]{Fig2.pdf}
-  \caption{Parameterization outputs resulting from the hydrographs shown in panels a and e for the cases called \alpine (a-d) and \icesheet{} (e-h).
+  \caption{Parameterization outputs resulting from the hydrographs shown in panels a and e for \alpine cases (a-d) and \icesheet{} (e-h).
     Gray (black) lines in a and e represent the subglacial channel size (water discharge).
@@
     Data are shown at $15$\unit{min} intervals.
-    Arrows denote some events where the variables peaks in the subglacial channel prior to the subaerial one.
+    Arrows denote some events where  variables peak in the subglacial channel prior to the subaerial one.
     Insets in e-h show the  peak melt event \icesheet{} denoted by the shaded area.
@@
 \end{figure}
+
 Following the parameterizations defined in Sections~\ref{sect:sub_mode}~and~\ref{sect:fluv}, the water velocity, shear stress, and width-integrated shear stress exhibit different seasonal evolutions and peaks.
@@
 
-Over the course of the season, large peaks in water velocity and shear stress, along with width-integrated shear stress, occur in response to increases in water discharge in both \alpine{} and \icesheet{}.
-This occurs as the subglacial outputs are generally larger than the subaerial values, thus more efficient at transporting sediment.
+Over the course of the season, large peaks in water velocity and shear stress, along with width-integrated shear stress, occur in response to water discharge variations in both \alpine{} and \icesheet{}.
 In the subglacial channel, peaks generally occur over the period when water discharge increases at the fastest rate (Figure~\ref{fig:model_outs}).
@@
 The timing of the greatest rate of change in water discharge occurs before the peaks in water discharge.
+Furthermore, the subglacial outputs are generally larger than the subaerial values, thus more efficient at transporting sediment.
 Here, sediment could be deposited in the proglacial area as the water depressurizes, then remobilized when sediment transport capacity peaks in the subaerial system.
@@
   \includegraphics[width=0.9\linewidth]{Fig3.png}
-  \caption{Relationship between water discharge and velocity, shear stress, and width-integrated shear stress for \alpine{} (a-c) and \icesheet (d-f).
+  \caption{
+    Relationship between water discharge and velocity, shear stress, and width-integrated shear stress for \alpine{} (a-c) and \icesheet (d-f).
     These outputs' relationship with measured sediment discharge is given for \alpine in (g-i) and \icesheet (j-l).
-    Variables have been normalized to mean values for better comparison amongst the runs.
-    $R_g$ and $R_f$ show the Spearman rank correlation values for the subglacial and the subaerial outputs, respectively.
+    Variables have been normalized to mean values.
+    $R_g$ and $R_f$ show the Spearman rank correlations for the subglacial and the subaerial outputs, respectively.
     Plots are shown with data and outputs at $15$\unit{min} intervals.
@@
 Each water discharge value in the subaerial channel results in a unique water velocity, shear stress, and width-integrated shear stress, creating a perfect rank correlation between the variables (Figure~\ref{fig:Qw_vari} a--f).
-In the subglacial channel, however,  the outputs can vary substantially for a given water discharge and width-integrated shear stress varies with respect to the other two  outputs as well (Figure~\ref{fig:Qw_vari} a--f).
+In the subglacial channel, however, the persistence of high water velocity, shear stress, and width-integrated shear stress across a range of discharges suggests that substantial sediment transport may occur across much of the season (Figure~\ref{fig:Qw_vari} a--f).
 For instance, very high  water velocity values and shear stresses can occur across minimal water discharge at $2.5$\,\unit{m}$^3$\,\unit{s}$^{-1}$ to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
 In \icesheet, mean values of water velocity can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
-
 When considering the subglacial channel's evolving width, width-integrated shear stress  across the channel generally increases with water discharge, with improved rank correlation compared to water velocity or shear stress (Figure~\ref{fig:Qw_vari} a--f).
@@
 The variability in width-integrated shear stress is less pronounced in the \icesheet case, with reduced variability in water discharge (Figure~\ref{fig:Qw_vari} c, f).
-The persistence of high water velocity, shear stress, and width-integrated shear stress suggests that substantial sediment transport may occur across much of the water discharge range in the subglacial system (Figure~\ref{fig:Qw_vari} a--f).
 
@@
     \includegraphics[width=0.9\linewidth]{Fig4.png}
-    \caption{Variability (measured by the standard deviation) in water velocity, shear stress, and width integrated shear stress for different time aggregations across the range of subglacial and subaerial channel shapes and friction factors.
-      Shaded areas denote the minimum and maximum of the $100$ parameter combinations tested for each of the systems (Section~\ref{sect:imp}).
+    \caption{Variability (standard deviation) in water velocity, shear stress, and width integrated shear stress for different time aggregations across the range of subglacial and subaerial channel shapes and friction factors.
+      Shaded areas denote the minimum and maximum of the $100$ parameter combinations tested (Section~\ref{sect:imp}).
       Solid lines denote  mean values.
-      Top row shows results from the \alpine{} case, while bottom row show outputs from the \icesheet{} case
+      %Top row shows results from the \alpine{} case, while bottom row show outputs from the \icesheet{} case
       Time intervals shown with markers denote $15$\,\unit{min}, $1$\,\unit{hr}, $12$\,\unit{hr}, $1$\,\unit{d}, $5$\,\unit{d}, $10$\,\unit{d}, and $15$\,\unit{d} aggregation periods.
@@
 
-The second application examines the impact of a variety of channel shapes and friction factors on the relationship between water discharge and sediment transport capacity in both systems.
+The second application examines the impact of a variety of channel shapes and friction factors on the relationship between water discharge and sediment transport capacity.
 Across the range of parameter values examined, variability in water velocity and shear stress remains higher in the subglacial system compared to the subaerial one.
-n \alpine{}, the variability in width-integrated shear stress is comparable between the two systems, due to the variable water discharge (Figure~\ref{fig:multi_run}).
- However, in \icesheet{}, with lower water discharge variability, the variability in width-integrated shear stress remains higher in the subglacial system across the time aggregation periods.
+In \alpine{}, the variability in width-integrated shear stress is comparable between the two systems, due to the variable water discharge (Figure~\ref{fig:multi_run}).
+However, in \icesheet{}, with lower water discharge fluctuations, the variability in width-integrated shear stress remains higher in the subglacial system across the time aggregation periods.
 
@@
 Note that lags and time of peak sediment transport conditions still occur over these periods (Figure~\ref{fig:model_outs}), while the variability in sediment transport parameters decreases with larger time aggregation (Figure~\ref{fig:multi_run}).
-\ian{add sentence here}
 
@@
 
-When comparing the outputs with suspended sediment discharge measurements, the subaerial parameterizations generally show a better correlation with sediment discharge compared to subglacial outputs across the tested parameters (Figures~\ref{fig:Qw_vari} g--l and \ref{fig:multi_run}).
-The rank correlation amongst the  subaerial parameterization's three output variables remains fairly constant despite the variable impact of channel shape and consistent scaling with water discharge (Figure \ref{fig:Qw_vari} and Section~\ref{sect:fluv}).
+When comparing the outputs with suspended sediment discharge measurements, the subaerial parameterizations generally correlate better with sediment discharge compared to subglacial outputs across the tested parameters (Figures~\ref{fig:Qw_vari} g--l and \ref{fig:multi_run}).
+The rank correlation amongst the subaerial parameterization's outputs remains fairly constant despite the variable impact of channel shape and consistent scaling with water discharge (Figure \ref{fig:Qw_vari} and Section~\ref{sect:fluv}).
 In subglacial channels, velocity and shear stress perform substantially worse in terms of rank correlation with sediment discharge compared to width-integrated shear stress (Figures~ \ref{fig:multi_run}).
@@
 \label{sect:dis_qsc}
+
 Quantities that control sediment transport capacity, such as water velocity, shear stress, and width-integrated shear stress, vary differently in response to water discharge in subglacial channels compared to subaerial ones (Figure~\ref{fig:Qw_vari}).
 In sediment transport relationships that evaluate sediment discharge capacity, such as in \citeA{meyer1948} or \citeA{engelund1967}, shear stress is scaled to the power of $\frac{3}{2}$ or $\frac{5}{2}$, respectively.
-The exponent greater than $1$ magnifies sediment discharge variability beyond the variable sediment transport parameters described above (Figure~\ref{fig:multi_run}).
+The exponent greater than $1$ magnifies sediment discharge variability beyond the variable sediment transport parameters described above (Figure~\ref{fig:multi_run}; Table~\ref{tab:Qs}).
 Note that sediment transport relationships describe sediment flux per unit width, so that the integrated capacity across channel occurs after application of the exponent, unlike the width-integrated shear stress evaluation here \cite<Equations\,\ref{eq:tautg} and \ref{eq:tautf}; c.f.>{tucker1997}.
-Note also that both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress is homogeneous across the channel bed, despite the limitations of this assumption \cite<e.g.>{yager2018}.
+%Note also that both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress is homogeneous across the channel bed, despite the limitations of this assumption \cite<e.g.>{yager2018}.
 
-\ian{Mauro: Can you double-check this?}
-The rapid increases and decreases in subglacial sediment transport capacity may compound the variability in sediment transport already present in subaerial systems \cite<Figure~\ref{fig:Qw_vari}; >{williams1989,jerolmack2010}.
-By applying the discharge - width scaling in Equation~\ref{eq:wcf} to Equation~\ref{eq:v}, by setting $S\, \propto \,Q^{\frac{1}{2}}$, the subaerial channel's width-integrated shear stress response is $\tau_{tf}\, \propto\,  Q$ (see \ref{sect:scaling}).
-For sufficiently short time periods,  channel width and wetted area are fixed in the subglacial system so that the response of width-integrated shear stress is $ \tau_{tg}\, \propto\,  Q^{2}$, with a larger exponent on $Q$ than in the subaerial case following \citeA{alley1997}.
-In this case, water discharge fluctuates at a shorter timescale than the subglacial channel evolution \cite{gimbert2016}.
-Note also that the subglacial width integrated shear stress response to water discharge will be greater than the subaerial's provided that  Equation~\ref{eq:wcf}'s exponent is below one.
-However, when subglacial channel size lies in equilibrium with water discharge, the  width-integrated shear stress response is $\tau_{tg}\, \propto \, Q^{\frac{4}{5}}$, similar to subaerial channels.
-It follows that the stronger correlation between hydraulics and water discharge in \icesheet{} could result from being closer to this equilibrium between variations in the channels' wetter area and water discharge compared to \alpine{}.
-This state could result from the reduced diurnal discharge amplitude in \icesheet{} (Figures~\ref{fig:model_outs} and ~\ref{fig:Qw_vari}).
-In both cases, hydraulics could trend toward this equilibrium regime if the discharge is averaged discharge over $1-5$ days.
-In this case, variability in subglacial channels  approaches that of subaerial channels (Figure~\ref{fig:multi_run}).
+% The rapid increases and decreases in subglacial sediment transport capacity compounds the variability in sediment transport already present in subaerial systems \cite<Figure~\ref{fig:Qw_vari}; >{williams1989,jerolmack2010}.
+Sediment transport capacity in subglacial channels rapidly responds to changes in water discharge compared to their subaerial counter parts \cite{alley1997}.
+In subglacial R\"othlisberger channels where water discharge evolves more slowly than channel size or subaerial channels the width-integrated shear stress response to water discharge is $ \tau_{tg}\, \propto\,  Q^{\sim\frac{4}{5}}$ (Table\,\ref{tab:eqs1}).
+However, subglacial channels rarely operate in a steady state \cite<e.g.>{gimbert2016}.
+For sufficiently short time periods, channel width and wetted area are fixed in the subglacial system (assuming pipe flow) so that the response of width-integrated shear stress is $ \tau_{tg}\, \propto\,  Q^{2}$ (Table\,\ref{tab:eqs1}).
+Appendix~\ref{sect:scaling} provides a complete description of sediment transport capacity in different drainage conditions.
+
+The strong correlation between hydraulics and water discharge in \icesheet{} could result from being closer to an equilibrium between variations in the channels' wetter area and water discharge with smaller diurnal variations in amplitude (Figures~\ref{fig:model_outs} and ~\ref{fig:Qw_vari}).
+Conversely, the correlation between hydraulics and water discharge in \alpine is smaller due to the larger discharge variations. 
+In both cases, hydraulics could trend toward equilibrium by averaging the discharge over $1-5$ days, when the variability in subglacial channels approaches that of subaerial one (Figure~\ref{fig:multi_run}).
 
 In subglacial channels, peaks in water velocity, shear stress and width-integrated shear stress occur with the fastest rate of water discharge increase relative to the channel's growth (Figure~\ref{fig:model_outs}).
-In turn, the subglacial channels contain feedbacks limiting sediment transport capacity as the subglacial channel grows with increasing water discharge capacity.
-This effectively reduces sediment transport capacity with the stabilization of water discharge at its peak.
+In turn, subglacial channels grow with increasing water discharge, effectively reduces sediment transport capacity with the stabilization of water discharge at its peak.
 Similar feedback occurs in subaerial systems \cite{phillips2016}, albeit at timescales longer than days \cite{gimbert2016}.
@@
 
-The increased performance of subaerial hydraulics outputs, compared to subglacial ones, in representing observations occurs despite the proximity of measurement locations to these glaciers.(Figure~\ref{fig:Qw_vari}).
+The increased performance of subaerial hydraulics outputs, compared to subglacial ones, in representing observations occurs despite the proximity of measurement locations to these glaciers (Figure~\ref{fig:Qw_vari}).
 This could occur for two reasons.
@@
 First, these measurement stations lie $0.5$ to $2$\,\unit{km} downstream of the glacier terminus \cite{cowton2012,felix2022}, a long enough distance for the pressurized subglacial regime to transition to the open channel subaerial one.
-Actually, the transition of subglacial to subaerial conditions may occur underneath glacier terminus \cite{perolo2018}, especially if closure rates of the ice channels are slow  \cite{rothlisberger1972}.
+Actually, the transition of subglacial to subaerial conditions may occur underneath glacier terminus \cite{perolo2018}, especially if closure rates of the subglacial channels are slow  \cite{rothlisberger1972}.
 Furthermore, if a transport-limited regime persists at the glacier margin, then sediment can be easily mobilized, and subaerial processes could be represented just a short distance downstream.
@@
 A supply-limited regime could well persist at many glaciers, because the sediment's threshold of motion can be reached more frequently and across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
-Sediment exhaustion occurring through this process may also explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \cite{overeem2017}.
-Note that in the data here, however, the absence of lags between the observed sediment and water discharge may suggest that sediment is mobilized close to the measurement stations, as opposed to distal areas of the glacier bed as would be expected in a supply-limited regime \cite<Figure~\ref{fig:model_outs}\, a and e; >{williams1989}.
+Sediment exhaustion occurring through increased variability may also explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \cite{overeem2017}.
+Note, however, the absence of lags between the observed sediment and water discharge may suggest that sediment is mobilized close to the measurement stations, as opposed to distal areas of the glacier bed as would be expected in a supply-limited regime \cite<Figure~\ref{fig:model_outs}\, a and e; >{williams1989}.
 
-Regardless of the increased performance of water discharge compared to subglacial sediment transport capacity in the observations, parameterization outputs suggest that additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate \cite{jansson2005,ganti2016}.
+Regardless of the increased performance of water discharge compared to subglacial sediment transport capacity in the observations, outputs suggest that additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate \cite{jansson2005,ganti2016}.
 Sediment discharge records have been used to establish the relationship, or lack thereof, between climate forcing and glacial erosion \cite<e.g.>{koppes2009a,ganti2016,willenbring2016,mariotti2021}.
@@
 \item a consistent relationship between water and sediment discharge may represent subglacial sediment access or open channel hydraulics near the glacier margin, instead of subglacial sediment transport capacity.
-
 \end{enumerate}
@@
 \section*{Author Contributions}
-\ian{let me know if these accurately describe your contributions... I hope I am not leaving anything out}
-I.D. designed the study, developed, designed, and implemented the model and experiments, and wrote the manuscript.
+
+I.D. designed the study, developed, and implemented the model and experiments, and wrote the manuscript.
 A.J.T. helped interpret data from Leverett Glacier and provided guidance on experiment design.
-M.A.W. provided data from Fieschergletscher and added the relationship between shear stress and water discharge (see Section~\ref{sect:dis_qsc} and \ref{sect:scaling}).
+M.A.W. provided data from Fieschergletscher and added the relationship between shear stress and water discharge (see Sections~\ref{sect:dis_qsc} and \ref{sect:scaling}).
 D.F. provided data from Fieschergletscher.
@@
 \section*{Open Research Section}
+
 Code to make the figures, with links to the data, can be found at \url{https://bitbucket.org/IanDelaney/xsection/src/master/}.
@@
     Channel wetted area (glacier, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
+    Channel depth (subaerial) & &$H$& $\mathrm{m}$\\
     Hydraulic diameter &$D_h$&&$\mathrm{m}$\\
@@
     Hydraulic head &$\Delta h$&& $\mathrm{m}$\\
+    Hydraulic gradient &$\Psi=\frac{\Delta h}{l}$&& $\mathrm{m\, m^{-1}}$\\
+
     Shear stress (glacier, subaerial) & $\tau_g,\,\tau_f$&& $\mathrm{Pa \, m^{-2}}$ \\
     Width-integrated shear stress (glacier, subaerial) & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$ \\
+    Stream power & $\Omega$ && $\mathrm{ kg \, m\, s^{-3}}$ \\
 
@@
     Hooke angle of channel & $\beta$ & $\frac{\pi}{2}$ & \unit{rad}\\
-
     Glacier thickness &$h_{ice}$& $225$ or $700$  &\unit{m}\\
@@
     Specific heat capacity of water &$c_p$&$4180$&\unit{J\,kg}$^{-1}$\unit{K}$^{-1}$\\
-
     Ice flow constant &$A$& $5.3\times10^{-24}$ &\unit{Pa}$^{-n}$\,$s^{-1}$\\
@@
     Subaerial channel factor & $k$ &$3$ or $6.5$ & $\mathrm{s\,m^{-2}}$\\ 
-    Channel geometry exponent &$e$& $\frac{1}{2}$&$\mathrm{(-)}$ \\
+    Channel geometry exponent &$\alpha$& $\frac{1}{2}$&$\mathrm{(-)}$ \\
     \hline
@@
   \label{table:vpm}
-
 \end{table}
@@
 
-\mauro{Notation: either use $H$ and $W$ for the river, or otherwise, somehow free $h$. Below I use $H$ and $W$ for now.}
+\mauro{Notation: either use $h$ and $w_c$ for the river, or otherwise, somehow free $h$. Below I use $H$ and $W$ for now.}
 
-Here we look at three kind of drainage regimes, namely, rivers, steady-state R-channels, and pipe-flow (i.e.
+Here three kinds of drainage regimes are examined, namely, subaerial channels, steady-state R-channels \cite{rothlisberger1972}, and pipe-flow (i.e.
 R-channels which do not have time to adjust their size to discharge conditions).
-We calculate how in these regimes sediment flux scales with respect to a given discharge for three different transport formulas: MPM \cite{meyer1948}, EH \cite{engelund1967}, and Bagnold \cite{Bagnold1980}; additionally we state the width integrated shear stress used as proxy for sediment transport throughout the manuscript.
+Sediment transport capacity is scaled with respect to a given water discharge for three different transport formulas: MPM \cite{meyer1948}, EH \cite{engelund1967}, and Bagnold \cite{bagnold1980}; additionally width-integrated shear stress is assessed, as it is  used as proxy for sediment transport throughout the manuscript.
 
-A few basic hydraulic relations are used in the following.  The Darcy-Weisbach equation can be stated as
+A few basic hydraulic relations are used in the following.
+The Darcy-Weisbach equation can be stated as
 \begin{equation}
@@
 \end{equation}
-with $\Psi$ the head gradient and friction factor $f$.  The shear stress and stream power are, respectively,
+with $\Psi$ the head gradient and friction factor $f$.
+The shear stress and stream power are, respectively,
 \begin{equation}
@@
 %
+
 Sediment discharge is given by the MPM, EH and Bagnold formulations as, respectively,
 \begin{equation}
-  Q_s \propto W\, \tau^{3/2}, \quad Q_s \propto W\, \tau^{5/2}, \quad Q_s \propto W \left(\frac{\Omega}{W}\right)^{3/2} H^{-2/3}
+  Q_s \propto w_c\, \tau^{3/2}, \quad Q_s \propto w_c\, \tau^{5/2}, \quad Q_s \propto w_c \left(\frac{\Omega}{w_c}\right)^{3/2} H^{-2/3}
 \end{equation}
@@
 
-The river is assumed to have a width $W$ much greater than its depth $H$, such that $D_h\approx 4H$, and to have a constant head gradient ($\Psi$) given by the topography.
-Further, we assume that its width can be approximated by a relation $W \propto Q^\alpha$ with $\alpha\in [0,1]$; typically $\alpha \approx 1/2$ (a so-called regime channel), $\alpha=0$ or $\alpha=1$ correspond a river of constant width (a slot canyon) or depth (no natural equivalent), respectively, i.e.
-the two end members.
+The  subaerial channels is assumed to have a width $w_c$ much greater than its depth $H$, such that $D_h\approx 4h$, and to have a constant head gradient ($\Psi$) given by the topography.
+Further, it is  assumed that its width can be approximated by a relation $w_c \propto Q^\alpha$ (Equation~\ref{eq:wcf})with $\alpha\in [0,1]$; typically $\alpha \approx 1/2$ (a so-called regime channel).
+End members $\alpha=0$ or $\alpha=1$ correspond a  subaerial channels of constant width (a slot canyon) or depth (no natural equivalent), respectively.
 %
-For a steady state R-channel, we assume that it has a constant $\Psi$ (approximated by the gradient of the Shreve \citeyear{shreve1972} potential) and that $S$ adjusts such that it is in steady state corresponding to $\Psi$ and $Q$.
+For a steady state R-channel, it is assumed that  $\Psi$ is constant (approximated by the gradient of the Shreve \citeyear{shreve1972} potential) and that $S$ adjusts such that it is in steady state with $\Psi$ and $Q$.
 %Conversely, pipe flow has $S$ fixed and $\Psi$ adjusts to satisfy the given $Q$.
 %
-For pipe flow, which occurs when a R-channel is subjected to rapid discharge variations such that it cannot adjust its size significantly,  we assume that the cross-sectional area $S$ is fixed and that $\Psi$ adjusts to meet the specified $Q$.
-For both R-channel and pipe flow, we assume that $D_h \propto S^{1/2}$.
+Pipe flow is approximated when a R-channel is subjected to rapid discharge variations such that the channel does not adjust its size significantly on the time scale of the discharge variations.
+In this case, it is assumed that the cross-sectional area $S$ is fixed, and $\Psi$ adjusts to the specified $Q$.
+For both R-channel and pipe flow, it is assumed assume that $D_h \propto S^{1/2}$.
 
-With these assumptions the Darcy-Weisbach equation~\eqref{eq:DW} can be solved for the not-fixed quantity: $H$ for a river, $S$ for a R-channel, and $\Psi$ for pipe-flow.
+With these assumptions the Darcy-Weisbach equation~\eqref{eq:DW} can be solved for the not-fixed quantity: $H$ for a  subaerial channel, $S$ for a R-channel, and $\Psi$ for pipe-flow.
 Then, using equations~\eqref{eq:tau-omega}, the shear stress and stream power can be calculated.
@@
 
-Table~\ref{tab:Qs2} shows that the proxy $W\tau$ as well as the MPM, EH and Bagnold sediment transport formulas give roughly a similar scaling of $Q_s$ on $\Psi$ or $S$ and on $Q$.
+Table~\ref{tab:Qs2} shows that the proxy $w_c\tau$ as well as the MPM, EH and Bagnold sediment transport formulas give roughly a similar scaling of $Q_s$ on hydraulic gradient $\Psi$ or $S$ and on water discharge $Q$.
 Remarkably, the Bagnold formula has a negative exponent for $f$ in all but the pipe-flow regime, which seems rather unexpected.
-The total transport formula EH gives a slightly stronger dependence on all variables as should be expected due to to the larger exponent on $\tau$ of 2.5 versus 1.5 for the others.
-Albeit, for pipe-flow Bagnold is close to EH.
-Conversely, the proxy used within this publication scales only as $Q^2$ for pipe-flow, whereas the others scale at least with $Q^3$.
-The exponent for the width scaling $\alpha$ only impacts the EH relation in any meaningful way; but for the value of $\alpha$ around 0.5, as is suitable for a braided proglacial stream and indeed most rivers, the exponent on $Q$ is only slightly above what the other relations give.
+The total transport formula EH gives a slightly stronger dependence on all variables as should be expected due to the larger exponent on $\tau$ of $\frac{5}{2}$ versus  $\frac{3}{2}$  for the others.
+Albeit, the sediment transport response in pipe-flow for the Bagnold case is close to EH.
+Conversely, the sediment transport proxy $t_{tg},t_{tf}$ used within this publication scales only as $Q^2$ for pipe-flow, whereas the others scale at least with $Q^3$.
+The exponent for the width scaling $\alpha$ only impacts relationship between sediment transport and water discharge in the EH relation in any meaningful way.
+However, for the value of $\alpha$ around $\frac{1}{2}$,  suits braided proglacial stream and indeed most  subaerial channels.
+Furthermore, when $\alpha=1$, the exponent on $Q$ is only slightly above the other values of $\alpha$.
 
-The sediment discharge in the steady state R-channel scales very similarly to the river case for all relations and virtually identically for the $\alpha=0.5$ case.
-However, note that the head gradients $\Psi$ are likely higher for comparable $Q$ in a ice-sheet marginal or alpine glacier setting than in a river setting, as is well described in \citeA{alley1997}, and thus transport rates may still be much higher in a steady state R-channel.
+The sediment discharge in the steady state R-channel scales very similarly to the  subaerial channel case for all relations and virtually identically for the $\alpha=0.5$ case.
+However, note that the head gradients $\Psi$ are likely higher for comparable $Q$ in a ice-sheet marginal or alpine glacier setting than in a  subaerial channel, as is well described in \citeA{alley1997}, and thus transport rates may still be much higher in a steady state R-channel.
 
 Furthermore, R-channels will rarely operate in steady state as variations in discharge, in particular on the diurnal timescale or the timescale of severe rain or melting events, are too fast for such a channel to reach a steady state.
-In such cases they operate more like a pipe of fixed cross-section.
-Table~\ref{tab:Qs2} shows that in such a situation, the sediment transport scales much more severely with discharge, with the exponent of $Q$ being likely between 3 and 5, compared to other two regimes when that exponent is at most 1.3.
+In such cases they operate more like a pipe of fixed cross-section \cite<e.g.>{gimbert2016}.
+Table~\ref{tab:Qs2} shows that in such a situation, the sediment transport scales much more severely with discharge, with the exponent on $Q$ being likely between $3$ and $5$, compared to other two regimes when that exponent is at most $1.3$ (Table~\ref{tab:Qs2}).
 Thus fluctuations of discharge on short timescales (on the order of a day) have the potential to cause very high sediment transport conditions.
@@
 \begin{table}
-  \caption{Relations for hydraulic variables for the three drainage regimes: river, R-channel and pipe-flow.
+  \caption{Relations for hydraulic variables for the three drainage regimes:  subaerial channels, R-channel and pipe-flow.
 Darcy-Weisbach equation is abbreviated with ``D-W'', and stream power with ``Stream p.''.
@@
 \begin{tabular}{llllll}
-Regime & Fixed & Determined via D-W & Additional relations & Shear stress & Stream p.\\
+  Regime & Fixed & Determined via D-W
+  \makecell{Equation~\ref{eq:DW}} & Additional relations & Shear stress & Stream p.\\
  &  &  &  & \(\tau \propto\) & \(\Omega \propto\)\\
 \hline
-River & \(\Psi\) & \(H \propto f^{1/3}\, Q^{2/3-2\alpha/3} \, \Psi^{-1/3}\) & \(W\,\propto Q^\alpha\), \(S=WH\), \(D_h=H\) & \(f^{1/3} Q^{2/3-2\alpha/3}  \Psi^{2/3}\) & \(Q \Psi\)\\
-R-ch. & \(\Psi\) & \(S\, \propto f^{2/5}\, Q^{4/5} \, \Psi^{-2/5}\) & \(D_h\propto W \propto H \propto S^{1/2}\) & \(f^{1/5} Q^{2/5} \, \Psi^{4/5}\) & \(Q \Psi\)\\
-Pipe & \(S\) & \(\Psi \propto f \, Q^2\, S^{-5/2}\) & \(D_h\propto W \propto H \propto S^{1/2}\) & \(f Q^2 S^{-2}\) & \(f Q^3 S^{-5/2}\)\\
+  Subaerial & \(\Psi\) & \(H \propto f^{1/3}\, Q^{2/3-2\alpha/3} \, \Psi^{-1/3}\) & 
+\makecell{\(w_c\,\propto Q^\alpha\) \\ \(S=w_cH\) \\ \(D_h=H\)} & 
+\(f^{1/3} Q^{2/3-2\alpha/3}  \Psi^{2/3}\) & \(Q \Psi\)\\
+R-ch. & \(\Psi\) & \(S\, \propto f^{2/5}\, Q^{4/5} \, \Psi^{-2/5}\) & \(D_h\propto w_c \propto H \propto S^{1/2}\) & \(f^{1/5} Q^{2/5} \, \Psi^{4/5}\) & \(Q \Psi\)\\
+Pipe & \(S\) & \(\Psi \propto f \, Q^2\, S^{-5/2}\) & \(D_h\propto w_c \propto H \propto S^{1/2}\) & \(f Q^2 S^{-2}\) & \(f Q^3 S^{-5/2}\)\\
 \end{tabular}
@@
 \begin{table}
-  \caption{Sediment transport proxy ($W\tau$) and rates for the three considered different transport formulas: MPM \cite{meyer1948}, EH \cite{engelund1967}, and Bagnold \cite{Bagnold1980}.
+  \caption{Sediment transport proxy ($w_c\tau$) and rates for the three considered different transport formulas: MPM \cite{meyer1948}, EH \cite{engelund1967}, and Bagnold \cite{bagnold1980}.
     }
@@
  & Width*\(\tau\) & MPM & EH & Bagnold\\
- & \(W\, \tau\) & \(Q_s \propto W\, \tau^{3/2}\) & \(Q_s \propto W\, \tau^{5/2}\) & \(Q_s \propto W^{-1/2}\, \Omega^{3/2} H^{-2/3}\)\\
+ & \(w_c\, \tau\) & \(Q_s \propto w_c\, \tau^{3/2}\) & \(Q_s \propto w_c\, \tau^{5/2}\) & \(Q_s \propto w_c^{-1/2}\, \Omega^{3/2} H^{-2/3}\)\\
 \hline
-River & \(f^{1/3}\, Q^{2/3+\alpha/3}\,  \Psi^{2/3}\) & \(f^{1/2}\, Q \, \Psi\) & \(f^{5/6}\, Q^{5/3 - 2\alpha/3} \, \Psi^{5/3}\) & \(f^{-2/9}\, Q^{19/18-\alpha/18} \, \Psi^{31/18}\)\\
+Subaerial  & \(f^{1/3}\, Q^{2/3+\alpha/3}\,  \Psi^{2/3}\) & \(f^{1/2}\, Q \, \Psi\) & \(f^{5/6}\, Q^{5/3 - 2\alpha/3} \, \Psi^{5/3}\) & \(f^{-2/9}\, Q^{19/18-\alpha/18} \, \Psi^{31/18}\)\\
 R-ch. & \(f^{2/5}\, Q^{4/5} \, \Psi^{3/5}\) & \(f^{1/2}\, Q \, \Psi\) & \(f^{7/10}\, Q^{7/5}\, \Psi^{9/5}\) & \(f^{-7/30}\, Q^{31/30}\, \Psi^{26/15}\)\\
@@
   \caption{As Table~\ref{tab:Qs} but with the fractional exponents stated as rounded decimal numbers.
-For the River regime the exponents are for displayed for the likely width-exponent $\alpha=1$, as well as its end members 0 (slot canyon) and 1 (only width increases and not depth).
+For the subaerial regime the exponents are for displayed for the likely width-exponent $\alpha=1$, as well as its end members 0 (slot canyon) and 1 (only width increases and not depth).
     }
@@
  & Width*\(\tau\) & MPM & EH & Bagnold\\
- & \(W\, \tau\) & \(Q_s \propto W\, \tau^{3/2}\) & \(Q_s \propto W\, \tau^{5/2}\) & \(Q_s \propto W^{-1/2}\, \Omega^{3/2} H^{-2/3}\)\\
+ & \(w_c\, \tau\) & \(Q_s \propto w_c\, \tau^{3/2}\) & \(Q_s \propto w_c\, \tau^{5/2}\) & \(Q_s \propto w_c^{-1/2}\, \Omega^{3/2} H^{-2/3}\)\\
 \hline
-River (\(\alpha=0\)) & \(f^{0.7}\, Q^{0.7}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{1.7} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.1} \, \Psi^{1.7}\)\\
-River (\(\alpha=0.5\)) & \(f^{0.7}\, Q^{0.8}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{1.3} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\
-River (\(\alpha=1\)) & \(f^{0.7}\, Q^{1.0}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{0.7} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\[3pt]
+Subaerial (\(\alpha=0\)) & \(f^{0.7}\, Q^{0.7}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{1.7} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.1} \, \Psi^{1.7}\)\\
+Subaerial (\(\alpha=0.5\)) & \(f^{0.7}\, Q^{0.8}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{1.3} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\
+Subaerial (\(\alpha=1\)) & \(f^{0.7}\, Q^{1.0}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{0.7} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\[3pt]
 R-ch. & \(f^{0.4}\, Q^{0.8} \, \Psi^{0.6}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.7}\, Q^{1.4}\, \Psi^{1.8}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\

### 2023-10-05 Delaneyetal2023b.tex
more small bits

@@
 
-\title{Natural variations in water discharge quantity has an inconsistent impact on sediment transport capacity in subglacial channels}
+\title{Natural variations in water discharge quantity have an inconsistent impact on sediment transport capacity in subglacial channels}
 
@@
 Second, to further characterize the variability in sediment discharge capacity in subglacial compared to subaerial channels, $100$  parameterizations are applied to a range of different channel geometry factors ($\beta$ and $k$) and friction factors ($f_r$ and $f_p$).
-Runs with parameter combinations are culled if their mean subglacial water velocity over the season lies outside of the range between $0.75$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $2$\,\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity lies outside of the range $0.3$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with measurements \cite<e.g.>{werder2010b,magnusson2012,chandler2013}.
+Runs with parameter combinations are culled if their mean subglacial water velocity over the season lies between $0.75$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $2$\,\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity lies between $0.3$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with measurements \cite<e.g.>{werder2010b,magnusson2012,chandler2013}.
 Variability of an output is represented by the  standard deviation of an output over the course of the season.
 
-Lastly, outputs are compared with the sediment discharge records of the glaciers using Spearman rank correlation (i.e.
-correlation depends on the ordering of values from the outputs and the observations).
+Lastly, outputs are compared with the sediment discharge records of the glaciers using Spearman rank correlation, dependent on the ordering of values.
 Rank correlation reduces the impact of the non-linear response of sediment transport capacity to the hydrological forcing.
-Outputs and observations at $15$\,\unit{min} intervals are aggregated across timescales of up to $15$\,\unit{days} by averaging values one half of the timescale before and after a given point time.
+Outputs and observations at $15$\,\unit{min} intervals are aggregated across timescales of up to $15$\,\unit{days} by averaging values one half of the timescale before and after a point time.
 
@@
   \includegraphics[width=0.9\linewidth]{Fig2.pdf}
-  \caption{Parameterization outputs resulting from the hydrographs shown in panels a and e for \alpine cases (a-d) and \icesheet{} (e-h).
+  \caption{Parameterization outputs from the hydrographs n panels a and e for \alpine cases (a-d) and \icesheet{} (e-h).
     Gray (black) lines in a and e represent the subglacial channel size (water discharge).
-    In other panels, blue lines represent outputs from the subglacial channel, while orange lines represent the subaerial channel.
+    Blue lines represent outputs from the subglacial channel, while orange lines represent the subaerial channel.
     Data are shown at $15$\unit{min} intervals.
-    Arrows denote some events where  variables peak in the subglacial channel prior to the subaerial one.
+    Arrows denote where variables peaking in the subglacial channel prior to the subaerial one.
     Insets in e-h show the  peak melt event \icesheet{} denoted by the shaded area.
-    Note that the x and y axes are different for \alpine{} and \icesheet{}.
+    %Note that the x and y axes are different for \alpine{} and \icesheet{}.
   }
@@
 
-Over the course of the season, large peaks in water velocity and shear stress, along with width-integrated shear stress, occur in response to water discharge variations in both \alpine{} and \icesheet{}.
-In the subglacial channel, peaks generally occur over the period when water discharge increases at the fastest rate (Figure~\ref{fig:model_outs}).
-Alternatively, in the subaerial channel, peaks in these sediment transport variables occur when water discharge is highest.
+In the subglacial channel, peaks generally occur when water discharge increases at the fastest rate (Figure~\ref{fig:model_outs}).
+Alternatively, in the subaerial channel,  sediment transport variables peak  hen water discharge is highest.
 The timing of the greatest rate of change in water discharge occurs before the peaks in water discharge.
@@
 Each water discharge value in the subaerial channel results in a unique water velocity, shear stress, and width-integrated shear stress, creating a perfect rank correlation between the variables (Figure~\ref{fig:Qw_vari} a--f).
-In the subglacial channel, however, the persistence of high water velocity, shear stress, and width-integrated shear stress across a range of discharges suggests that substantial sediment transport may occur across much of the season (Figure~\ref{fig:Qw_vari} a--f).
+
+In the subglacial channel, however, the persistence of high water velocity, shear stress, and width-integrated shear stress across a range of discharges suggests that substantial sediment transport may occur across  (Figure~\ref{fig:Qw_vari} a--f).
+
 For instance, very high  water velocity values and shear stresses can occur across minimal water discharge at $2.5$\,\unit{m}$^3$\,\unit{s}$^{-1}$ to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
@@
     \includegraphics[width=0.9\linewidth]{Fig4.png}
-    \caption{Variability (standard deviation) in water velocity, shear stress, and width integrated shear stress for different time aggregations across the range of subglacial and subaerial channel shapes and friction factors.
-      Shaded areas denote the minimum and maximum of the $100$ parameter combinations tested (Section~\ref{sect:imp}).
+    \caption{Variability (standard deviation) of output variables for different time aggregations across the range of subglacial and subaerial channel shapes and friction factors.
+      Shaded areas denote the range of the $100$ parameter combinations tested (Section~\ref{sect:imp}).
       Solid lines denote  mean values.
@@
 
-The second application examines the impact of a variety of channel shapes and friction factors on the relationship between water discharge and sediment transport capacity.
-Across the range of parameter values examined, variability in water velocity and shear stress remains higher in the subglacial system compared to the subaerial one.
+Across the range of channel shape parameters and friction factors  examined, variability in water velocity and shear stress remains higher in the subglacial system compared to the subaerial one.
 In \alpine{}, the variability in width-integrated shear stress is comparable between the two systems, due to the variable water discharge (Figure~\ref{fig:multi_run}).
@@
 It is prescient to evaluate the time scales over which the subglacial channels may respond to changing water discharge to understand the influence of variability in sediment transport capacity.
-In both cases, the variability in velocity, shear stress, and width-integrated shear stress decrease after $1-5$ day aggregations (Figure~\ref{fig:multi_run}).
-Yet, substantial differences in variability persist in periods of up to $15$ days, as the variability in aggregations periods longer than this might be conflated with seasonal variations in water discharge.
+In both cases, the variability in velocity, shear stress, and width-integrated shear stress decreases after $1-5$ day aggregations (Figure~\ref{fig:multi_run}).
+Yet, substantial differences in variability persist in periods of up to $15$ days.
+Aggregations periods longer than this might be conflated with seasonal variations in water discharge.
 Note that lags and time of peak sediment transport conditions still occur over these periods (Figure~\ref{fig:model_outs}), while the variability in sediment transport parameters decreases with larger time aggregation (Figure~\ref{fig:multi_run}).
@@
 
-Quantities that control sediment transport capacity, such as water velocity, shear stress, and width-integrated shear stress, vary differently in response to water discharge in subglacial channels compared to subaerial ones (Figure~\ref{fig:Qw_vari}).
+Quantities that control sediment transport capacity (water velocity, shear stress, and width-integrated shear stress) vary differently in response to water discharge in subglacial channels compared to subaerial ones (Figure~\ref{fig:Qw_vari}).
 In sediment transport relationships that evaluate sediment discharge capacity, such as in \citeA{meyer1948} or \citeA{engelund1967}, shear stress is scaled to the power of $\frac{3}{2}$ or $\frac{5}{2}$, respectively.
@@
 % The rapid increases and decreases in subglacial sediment transport capacity compounds the variability in sediment transport already present in subaerial systems \cite<Figure~\ref{fig:Qw_vari}; >{williams1989,jerolmack2010}.
-Sediment transport capacity in subglacial channels rapidly responds to changes in water discharge compared to their subaerial counter parts \cite{alley1997}.
+Sediment transport capacity in subglacial channels responds more rapidly to changes in water discharge compared to their subaerial counter parts \cite{alley1997}.
 In subglacial R\"othlisberger channels where water discharge evolves more slowly than channel size or subaerial channels the width-integrated shear stress response to water discharge is $ \tau_{tg}\, \propto\,  Q^{\sim\frac{4}{5}}$ (Table\,\ref{tab:eqs1}).
@@
 
-The strong correlation between hydraulics and water discharge in \icesheet{} could result from being closer to an equilibrium between variations in the channels' wetter area and water discharge with smaller diurnal variations in amplitude (Figures~\ref{fig:model_outs} and ~\ref{fig:Qw_vari}).
+The stronger correlation between subglacial hydraulics and water discharge in \icesheet{} could result from the channel being closer to equilibrium between variations in the channels' wetter area and water discharge with smaller diurnal variations in amplitude (Figures~\ref{fig:model_outs} and ~\ref{fig:Qw_vari}).
 Conversely, the correlation between hydraulics and water discharge in \alpine is smaller due to the larger discharge variations. 
@@
 
-In subglacial channels, peaks in water velocity, shear stress and width-integrated shear stress occur with the fastest rate of water discharge increase relative to the channel's growth (Figure~\ref{fig:model_outs}).
+In subglacial channels, peaks in water velocity, shear stress and width-integrated shear stress occur with  fast rates of water discharge increase relative to the channel's growth (Figure~\ref{fig:model_outs}).
 In turn, subglacial channels grow with increasing water discharge, effectively reduces sediment transport capacity with the stabilization of water discharge at its peak.
@@
 First, these measurement stations lie $0.5$ to $2$\,\unit{km} downstream of the glacier terminus \cite{cowton2012,felix2022}, a long enough distance for the pressurized subglacial regime to transition to the open channel subaerial one.
-Actually, the transition of subglacial to subaerial conditions may occur underneath glacier terminus \cite{perolo2018}, especially if closure rates of the subglacial channels are slow  \cite{rothlisberger1972}.
+The transition of subglacial to subaerial conditions may occur underneath glacier terminus \cite{perolo2018}, especially if closure rates of the subglacial channels are slow  \cite{rothlisberger1972}.
 Furthermore, if a transport-limited regime persists at the glacier margin, then sediment can be easily mobilized, and subaerial processes could be represented just a short distance downstream.

### 2023-10-05 Delaneyetal2023b.tex
more edits

@@
 
-% The rapid increases and decreases in subglacial sediment transport capacity compounds the variability in sediment transport already present in subaerial systems \cite<Figure~\ref{fig:Qw_vari}; >{williams1989,jerolmack2010}.
+The rapid increases and decreases in subglacial sediment transport capacity compounds the variability in sediment transport already present in subaerial systems \cite<Figure~\ref{fig:Qw_vari}; >{williams1989,jerolmack2010}.
 Sediment transport capacity in subglacial channels responds more rapidly to changes in water discharge compared to their subaerial counter parts \cite{alley1997}.
@@
 For sufficiently short time periods, channel width and wetted area are fixed in the subglacial system (assuming pipe flow) so that the response of width-integrated shear stress is $ \tau_{tg}\, \propto\,  Q^{2}$ (Table\,\ref{tab:eqs1}).
-Appendix~\ref{sect:scaling} provides a complete description of sediment transport capacity in different drainage conditions.
+\ref{sect:scaling} provides a complete description of sediment transport capacity in different drainage conditions.
 

### 2023-10-09 Delaneyetal2023b.tex
Updates from Overleaf

@@
 
-\begin{abstract}
-  Sediment transport capacity in subglacial and subaerial channels depends on  water velocity and channel width, which together control the shear stress exerted on the channel bottom.
+\begin{abstract}  Sediment transport capacity in subglacial and subaerial channels depends on  water velocity and channel width, which together control the shear stress exerted on the channel bottom.
   In subaerial channels, width, and velocity evolve quickly with water discharge, while in subglacial channels, water discharge changes mainly affect water velocity due to slow-evolving channel geometry.
@@
 
-This manuscript evaluates the relationship between sediment transport capacity and water discharge in subglacial channels, and compares it to subaerial ones.
+This manuscript evaluates the relationship between sediment transport capacity and water discharge in subglacial channels and compares it to subaerial ones.
 In the absence of continuous measurements of subglacial and subaerial water velocity and channel morphology, numerical parameterizations of subglacial and subaerial hydraulics are utilized.
@@
 Data is used from \citeA{tedstone2017} and \citeA{hawkings2014} at a $5$\,\unit{min} time interval from May 28, 2012 to August 8, 2012.
-Two gaps (14.07.2012 to 15.07.2012 and 16.07.2012 to 17.07.2021) in suspended sediment data interpolated using a linear spline.
+Two gaps (14.07.2012 to 15.07.2012 and 16.07.2012 to 17.07.2021) in suspended sediment data were interpolated using a linear spline.
 
@@
 The subglacial and subaerial parameterizations represent the subglacial flow of water exiting a glacier and transitioning to subaerial flow.
-Outputs of the parameterizations are meant to represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
+Outputs of the parameterizations represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
 To generalize these cases, \alpine{}  is exemplified by relatively thin ice thickness ($h_{ice}$: $225$\,\unit{m}), low water discharge ($\sim\,10$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
@@
 
-For each glacier, the parameterizations are first applied to a reference test case for each glacier that assumes  subglacial channel with $\beta=\frac{\pi}{2}$.
+For each glacier, the parameterizations are first applied to a reference test case for each glacier that assumes  a subglacial channel with $\beta=\frac{\pi}{2}$.
 Friction factors $f_r$ and $f_p$ are tuned so that the parameterizations reproduce reasonable water velocities for both \alpine and \icesheet{}\cite<$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$>{werder2010b,chandler2013}.
@@
 Runs with parameter combinations are culled if their mean subglacial water velocity over the season lies between $0.75$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $2$\,\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity lies between $0.3$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with measurements \cite<e.g.>{werder2010b,magnusson2012,chandler2013}.
-Variability of an output is represented by the  standard deviation of an output over the course of the season.
+The variability of output is represented by the  standard deviation of that output over the course of the season.
 
@@
 Rank correlation reduces the impact of the non-linear response of sediment transport capacity to the hydrological forcing.
-Outputs and observations at $15$\,\unit{min} intervals are aggregated across timescales of up to $15$\,\unit{days} by averaging values one half of the timescale before and after a point time.
+Outputs and observations at $15$\,\unit{min} intervals are aggregated across timescales of up to $15$\,\unit{days} by averaging values one-half of the timescale before and after a point time.
 
@@
 
-Over time scales of weeks or longer, the glacier's sediment transport capacity responds to the ice thickness, controlling the channel closure rate, and the glacier's surface slope, in addition to water discharge \cite<Figure~\ref{fig:multi_run}, ~ Section~\ref{sect:sub_mode}; >{rothlisberger1972,shreve1972,gimbert2016,stevens2022}.
+Over time scales of weeks or longer, the glacier's sediment transport capacity responds to the ice thickness, controlling the channel closure rate, and the glacier's surface slope, in addition to water discharge \cite<Figure~\ref{fig:multi_run}, Section~\ref{sect:sub_mode}; >{rothlisberger1972,shreve1972,gimbert2016,stevens2022}.
 In a supply-limited state, often assumed in landscape evolution models, then glaciers' sediment discharge record also represents additional processes of sediment availability and bedrock erosion from  water pressure variations and sliding  \cite<e.g.>{iverson2012,herman2015,delaney2023}.
@@
 \begin{tabular}{llllll}
-  Regime & Fixed & Determined via D-W
-  \makecell{Equation~\ref{eq:DW}} & Additional relations & Shear stress & Stream p.\\
+  Regime & Fixed & \makecell{Determined via D-W
+  \\Equation~\ref{eq:DW}} & Additional relations & Shear stress & Stream p.\\
  &  &  &  & \(\tau \propto\) & \(\Omega \propto\)\\

### 2023-10-09 Delaneyetal2023b.tex
Appendix done

@@
 R-channels which do not have time to adjust their size to discharge conditions).
-Sediment transport capacity is scaled with respect to a given water discharge for three different transport formulas: MPM \cite{meyer1948}, EH \cite{engelund1967}, and Bagnold \cite{bagnold1980}; additionally width-integrated shear stress is assessed, as it is  used as proxy for sediment transport throughout the manuscript.
+Here we show how sediment transport capacity scales with respect to a given water discharge and hydraulic gradient for three different transport formulas: MPM \cite{meyer1948}, EH \cite{engelund1967}, and Bagnold \cite{bagnold1980}; additionally width-integrated shear stress is assessed, as used as proxy for sediment transport throughout the manuscript.
 
@@
 \end{equation}
-with $\Psi$ the head gradient and friction factor $f$.
+with $\Psi = \Delta h / l$ the head gradient and friction factor $f$.
 The shear stress and stream power are, respectively,
@@
 
-Sediment discharge is given by the MPM, EH and Bagnold formulations as, respectively,
+Sediment discharge is given by the MPM, EH and Bagnold formulations as
 \begin{equation}
-  Q_s \propto w_c\, \tau^{3/2}, \quad Q_s \propto w_c\, \tau^{5/2}, \quad Q_s \propto w_c \left(\frac{\Omega}{w_c}\right)^{3/2} H^{-2/3}
+  Q_s \propto w_c\, \tau^{3/2}, \quad Q_s \propto w_c\, \tau^{5/2}, \quad Q_s \propto w_c \left(\frac{\Omega}{w_c}\right)^{3/2} H^{-2/3},
 \end{equation}
-for conditions well above the transport onset threshold.
+respectively, for conditions well above the transport onset threshold.
 
-The  subaerial channels is assumed to have a width $w_c$ much greater than its depth $H$, such that $D_h\approx 4h$, and to have a constant head gradient ($\Psi$) given by the topography.
+The  subaerial channel is assumed to have a width $w_c$ much greater than its depth $H$, such that $D_h\approx 4h$, and to have a constant head gradient ($\Psi$) given by the topography.
 Further, it is  assumed that its width can be approximated by a relation $w_c \propto Q^\alpha$ (Equation~\ref{eq:wcf})with $\alpha\in [0,1]$; typically $\alpha \approx 1/2$ (a so-called regime channel).
@@
 %
-For a steady state R-channel, it is assumed that  $\Psi$ is constant (approximated by the gradient of the Shreve \citeyear{shreve1972} potential) and that $S$ adjusts such that it is in steady state with $\Psi$ and $Q$.
-%Conversely, pipe flow has $S$ fixed and $\Psi$ adjusts to satisfy the given $Q$.
+For a steady state R-channel, it is assumed that  $\Psi$ is constant (approximated by the gradient of the Shreve \citeyear{shreve1972} potential \footnote{Note that the R-channel model presented above (Eq.~\eqref{eq:dS_dt}--\eqref{eq:dh2wc}), calculates $\Psi$ from the time evolving $S_g$ (Eq.~\eqref{eq:dS_dt}) via the Darcy-Weisbach equation~\eqref{eq:dh} and thus no Shreve approximation is then needed.}) and that $S$ adjusts such that it is in steady state with $\Psi$ and $Q$.
+% Conversely, pipe flow has $S$ fixed and $\Psi$ adjusts to satisfy the given $Q$.
 %
-Pipe flow is approximated when a R-channel is subjected to rapid discharge variations such that the channel does not adjust its size significantly on the time scale of the discharge variations.
+Pipe flow like conditions occur when a R-channel is subjected to rapid discharge variations such that the channel cannot adjust its size significantly on the time scale of the discharge variations.
 In this case, it is assumed that the cross-sectional area $S$ is fixed, and $\Psi$ adjusts to the specified $Q$.
@@
 
-Table~\ref{tab:Qs2} shows that the proxy $w_c\tau$ as well as the MPM, EH and Bagnold sediment transport formulas give roughly a similar scaling of $Q_s$ on hydraulic gradient $\Psi$ or $S$ and on water discharge $Q$.
+Table~\ref{tab:Qs2} shows how the proxy $w_c\tau$ as well as $Q_s$ of the MPM, EH and Bagnold sediment transport formulas scale with respect to $Q$, $\Psi$ or $S$ for the three different regimes and for different $\alpha$.
 Remarkably, the Bagnold formula has a negative exponent for $f$ in all but the pipe-flow regime, which seems rather unexpected.
@@
 Albeit, the sediment transport response in pipe-flow for the Bagnold case is close to EH.
-Conversely, the sediment transport proxy $t_{tg},t_{tf}$ used within this publication scales only as $Q^2$ for pipe-flow, whereas the others scale at least with $Q^3$.
+Conversely, the sediment transport proxy $w_g\,\tau_{g}, w_f\,\tau_{f}$ used within this publication scales only as $Q^2$ for pipe-flow, whereas the others scale at least with $Q^3$.
 The exponent for the width scaling $\alpha$ only impacts relationship between sediment transport and water discharge in the EH relation in any meaningful way.
-However, for the value of $\alpha$ around $\frac{1}{2}$,  suits braided proglacial stream and indeed most  subaerial channels.
-Furthermore, when $\alpha=1$, the exponent on $Q$ is only slightly above the other values of $\alpha$.
+However, for the value of $\alpha$ around $\frac{1}{2}$ (a value appropriate for most streams), the exponent on $Q$ for EH is only slightly above what the other transport relations give.
 
 The sediment discharge in the steady state R-channel scales very similarly to the  subaerial channel case for all relations and virtually identically for the $\alpha=0.5$ case.
-However, note that the head gradients $\Psi$ are likely higher for comparable $Q$ in a ice-sheet marginal or alpine glacier setting than in a  subaerial channel, as is well described in \citeA{alley1997}, and thus transport rates may still be much higher in a steady state R-channel.
+However, note that the head gradients $\Psi$ are likely higher for comparable $Q$ in an ice-sheet marginal or alpine glacier setting than in a  subaerial channel, as is well described in \citeA{alley1997}, and thus transport rates may still be much higher in a steady state R-channel.
 
@@
 Table~\ref{tab:Qs2} shows that in such a situation, the sediment transport scales much more severely with discharge, with the exponent on $Q$ being likely between $3$ and $5$, compared to other two regimes when that exponent is at most $1.3$ (Table~\ref{tab:Qs2}).
-Thus fluctuations of discharge on short timescales (on the order of a day) have the potential to cause very high sediment transport conditions.
+Thus fluctuations of discharge on short timescales (on the order of a day) have the potential to cause conditions with very high sediment transport capacities.
 
@@
   \label{tab:eqs1}
-  \small
 \begin{tabular}{llllll}
-  Regime & Fixed & \makecell{Determined via D-W
-  \\Equation~\ref{eq:DW}} & Additional relations & Shear stress & Stream p.\\
- &  &  &  & \(\tau \propto\) & \(\Omega \propto\)\\
+  Regime & Fixed & Determined via D-W
+   & Additional relations & Shear stress & Stream p.\\
+& & (Equation~\ref{eq:DW})  &  &  \(\tau \propto\) & \(\Omega \propto\)\\
 \hline
-  Subaerial & \(\Psi\) & \(H \propto f^{1/3}\, Q^{2/3-2\alpha/3} \, \Psi^{-1/3}\) & 
-\makecell{\(w_c\,\propto Q^\alpha\) \\ \(S=w_cH\) \\ \(D_h=H\)} & 
-\(f^{1/3} Q^{2/3-2\alpha/3}  \Psi^{2/3}\) & \(Q \Psi\)\\
-R-ch. & \(\Psi\) & \(S\, \propto f^{2/5}\, Q^{4/5} \, \Psi^{-2/5}\) & \(D_h\propto w_c \propto H \propto S^{1/2}\) & \(f^{1/5} Q^{2/5} \, \Psi^{4/5}\) & \(Q \Psi\)\\
-Pipe & \(S\) & \(\Psi \propto f \, Q^2\, S^{-5/2}\) & \(D_h\propto w_c \propto H \propto S^{1/2}\) & \(f Q^2 S^{-2}\) & \(f Q^3 S^{-5/2}\)\\
+  Subaerial & \(\Psi\) & \(H \propto f^{1/3}\, Q^{2/3-2\alpha/3} \, \Psi^{-1/3}\) & \makecell{\(w_c\,\propto Q^\alpha\) \\ \(S=w_cH\) \\ \(D_h\propto H\)} & \(f^{1/3} Q^{2/3-2\alpha/3}  \Psi^{2/3}\) & \(Q\, \Psi\)\\
+  R-channel & \(\Psi\) & \(S\, \propto f^{2/5}\, Q^{4/5} \, \Psi^{-2/5}\) & \(D_h\propto w_c \propto H \propto S^{1/2}\) & \(f^{1/5} Q^{2/5} \, \Psi^{4/5}\) & \(Q\, \Psi\)\\
+Pipe & \(S\) & \(\Psi \propto f \, Q^2\, S^{-5/2}\) & \(D_h\propto w_c \propto H \propto S^{1/2}\) & \(f Q^2 S^{-2}\) & \(f\, Q^3 S^{-5/2}\)\\
 \end{tabular}
@@
 \begin{tabular}{lllll}
- & Width*\(\tau\) & MPM & EH & Bagnold\\
+ & Width \(\times \,\, \tau\) & MPM & EH & Bagnold\\
  & \(w_c\, \tau\) & \(Q_s \propto w_c\, \tau^{3/2}\) & \(Q_s \propto w_c\, \tau^{5/2}\) & \(Q_s \propto w_c^{-1/2}\, \Omega^{3/2} H^{-2/3}\)\\
@@
 Subaerial  & \(f^{1/3}\, Q^{2/3+\alpha/3}\,  \Psi^{2/3}\) & \(f^{1/2}\, Q \, \Psi\) & \(f^{5/6}\, Q^{5/3 - 2\alpha/3} \, \Psi^{5/3}\) & \(f^{-2/9}\, Q^{19/18-\alpha/18} \, \Psi^{31/18}\)\\
-R-ch. & \(f^{2/5}\, Q^{4/5} \, \Psi^{3/5}\) & \(f^{1/2}\, Q \, \Psi\) & \(f^{7/10}\, Q^{7/5}\, \Psi^{9/5}\) & \(f^{-7/30}\, Q^{31/30}\, \Psi^{26/15}\)\\
+R-channel & \(f^{2/5}\, Q^{4/5} \, \Psi^{3/5}\) & \(f^{1/2}\, Q \, \Psi\) & \(f^{7/10}\, Q^{7/5}\, \Psi^{9/5}\) & \(f^{-7/30}\, Q^{31/30}\, \Psi^{26/15}\)\\
 Pipe & \(f \, Q^2 \, S^{-1}\) & \(f^{3/2}\, Q^3 \, S^{-5/2}\) & \(f^{5/2}\, Q^5\, S^{-9/2}\) & \(f^{3/2} \, Q^{9/2} \, S^{-14/3}\)\\
@@
 Subaerial (\(\alpha=1\)) & \(f^{0.7}\, Q^{1.0}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{0.7} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\[3pt]
-R-ch. & \(f^{0.4}\, Q^{0.8} \, \Psi^{0.6}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.7}\, Q^{1.4}\, \Psi^{1.8}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\
+R-channel & \(f^{0.4}\, Q^{0.8} \, \Psi^{0.6}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.7}\, Q^{1.4}\, \Psi^{1.8}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\
 Pipe & \(f \,\quad Q^{2\phantom{.0}} \, S^{-1}\) & \(f^{1.5}\, Q^3 \, S^{-2.5}\) & \(f^{2.5}\, Q^{5\phantom{.0}}\, S^{-4.5}\) & \(f^{1.5} \,\,\,\,\, Q^{4.5} \, S^{-4.7}\)\\

### 2023-10-09 Delaneyetal2023b.tex
dropped subscript c of variable w

@@
 In a transport-limited regime, sediment discharge responds to the sediment transport capacity, defined as the amount of sediment the water can carry.
-Sediment transport capacity depends on the shear stress between water and the sediment it flows  over \cite{shields1936,meyer1948,engelund1967} along with the width of the channel bottom $w_c$ over which to mobilize sediment.
+Sediment transport capacity depends on the shear stress between water and the sediment it flows  over \cite{shields1936,meyer1948,engelund1967} along with the width of the channel bottom $w$ over which to mobilize sediment.
 The shear stress  across the channel width $\tau_t$ responds to the velocity of water $v$ flowing through the channel so that
@@
     \label{eq:tau_t}
-    \tau_t \propto w_c\, v^2,
+    \tau_t \propto w\, v^2,
   \end{equation}
@@
 %
-where $w_c$ is the width of the channel. Following mass conservation, the velocity of the water flowing through a  channel is
+where $w$ is the width of the channel. Following mass conservation, the velocity of the water flowing through a  channel is
 \begin{linenomath*}
@@
 Therefore, subglacial water flow behaves more like pipe flow, and changes in water discharge $Q$ are mainly accommodated with increased or decreased water velocity $v$ \cite{alley1997} because the size of the channel $S$ changes on a much longer timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
-%It follows that an increase in water discharge $Q$ will cause the shear stress across the channel $\tau_t$ to respond with water velocity $v$, as opposed to subaerial channels where channel width $w_c$ evolves as well.
+%It follows that an increase in water discharge $Q$ will cause the shear stress across the channel $\tau_t$ to respond with water velocity $v$, as opposed to subaerial channels where channel width $w$ evolves as well.
 
@@
       $S_g$ and $S_f$ represent the wetted area in subglacial and subaerial channels, respectively.
-      The subglacial channel widths $w_{cg}$ remain unchanged, while the subaerial channel widths $w_{cf}$ evolve. Shear stress $\tau$ is responsible for the mobilization of sediment.
+      The subglacial channel widths $w_g$ remain unchanged, while the subaerial channel widths $w_f$ evolve. Shear stress $\tau$ is responsible for the mobilization of sediment.
       %Note that in the subaerial channel parameterization, channels have rectangular shapes with their width much larger than their depth (Section~\ref{sect:fluv}).
@@
   \label{eq:tautg}
-  \tau_{tg}=w_{cg}\,\tau_g,
+  \tau_{tg}=w_g\,\tau_g,
 \end{equation}
 
-where  channel floor width $w_{cg}$ is
+where  channel floor width $w_g$ is
 \begin{equation}
   \label{eq:dh2wc}
-  w_{cg} = 2  \sin \frac{\beta}{2} \sqrt{\frac{2\, S_g}{\beta -\sin \beta}}.
+  w_g = 2  \sin \frac{\beta}{2} \sqrt{\frac{2\, S_g}{\beta -\sin \beta}}.
 \end{equation}
@@
     \label{eq:DW_tau}
-    \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q}{w_{cf}} \Big)^{\frac{2}{3}} \,\nabla z_c^{\frac{2}{3}},
+    \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q}{w_f} \Big)^{\frac{2}{3}} \,\nabla z_c^{\frac{2}{3}},
   \end{equation}
@@
 where $\nabla z_c$ is the channel slope, and $f_p$ is the friction factor for subaerial channels.
-Channel width $w_{cf}$ is
+Channel width $w_f$ is
 \begin{equation}
   \label{eq:wcf}
-  w_{cf} = k \, Q^{\alpha},
+  w_f = k \, Q^{\alpha},
 \end{equation}
@@
     \label{eq:tautf}
-    \tau_{tf}=w_{cf}\,\tau_f.
+    \tau_{tf}=w_f\,\tau_f.
   \end{equation}
@@
     Hydraulic diameter &$D_h$&&$\mathrm{m}$\\
-    Width of channel floor (glacial, subaerial) & $w_{cg},w_{cf}$&  & $\mathrm{m}$     \\
+    Width of channel floor (glacial, subaerial) & $w_g,w_f$&  & $\mathrm{m}$     \\
     Hydraulic head &$\Delta h$&& $\mathrm{m}$\\
@@
 
-\mauro{Notation: either use $h$ and $w_c$ for the river, or otherwise, somehow free $h$. Below I use $H$ and $W$ for now.}
+\mauro{Notation: either use $h$ and $w$ for the river, or otherwise, somehow free $h$. Below I use $H$ and $W$ for now.}
 
@@
 \begin{equation}
-  Q_s \propto w_c\, \tau^{3/2}, \quad Q_s \propto w_c\, \tau^{5/2}, \quad Q_s \propto w_c \left(\frac{\Omega}{w_c}\right)^{3/2} H^{-2/3},
+  Q_s \propto w\, \tau^{3/2}, \quad Q_s \propto w\, \tau^{5/2}, \quad Q_s \propto w \left(\frac{\Omega}{w}\right)^{3/2} H^{-2/3},
 \end{equation}
@@
 
-The  subaerial channel is assumed to have a width $w_c$ much greater than its depth $H$, such that $D_h\approx 4h$, and to have a constant head gradient ($\Psi$) given by the topography.
-Further, it is  assumed that its width can be approximated by a relation $w_c \propto Q^\alpha$ (Equation~\ref{eq:wcf})with $\alpha\in [0,1]$; typically $\alpha \approx 1/2$ (a so-called regime channel).
+The  subaerial channel is assumed to have a width $w$ much greater than its depth $H$, such that $D_h\approx 4h$, and to have a constant head gradient ($\Psi$) given by the topography.
+Further, it is  assumed that its width can be approximated by a relation $w \propto Q^\alpha$ (Equation~\ref{eq:wcf})with $\alpha\in [0,1]$; typically $\alpha \approx 1/2$ (a so-called regime channel).
 End members $\alpha=0$ or $\alpha=1$ correspond a  subaerial channels of constant width (a slot canyon) or depth (no natural equivalent), respectively.
@@
 
-Table~\ref{tab:Qs2} shows how the proxy $w_c\tau$ as well as $Q_s$ of the MPM, EH and Bagnold sediment transport formulas scale with respect to $Q$, $\Psi$ or $S$ for the three different regimes and for different $\alpha$.
+Table~\ref{tab:Qs2} shows how the proxy $w\tau$ as well as $Q_s$ of the MPM, EH and Bagnold sediment transport formulas scale with respect to $Q$, $\Psi$ or $S$ for the three different regimes and for different $\alpha$.
 Remarkably, the Bagnold formula has a negative exponent for $f$ in all but the pipe-flow regime, which seems rather unexpected.
@@
 \hline
-  Subaerial & \(\Psi\) & \(H \propto f^{1/3}\, Q^{2/3-2\alpha/3} \, \Psi^{-1/3}\) & \makecell{\(w_c\,\propto Q^\alpha\) \\ \(S=w_cH\) \\ \(D_h\propto H\)} & \(f^{1/3} Q^{2/3-2\alpha/3}  \Psi^{2/3}\) & \(Q\, \Psi\)\\
-  R-channel & \(\Psi\) & \(S\, \propto f^{2/5}\, Q^{4/5} \, \Psi^{-2/5}\) & \(D_h\propto w_c \propto H \propto S^{1/2}\) & \(f^{1/5} Q^{2/5} \, \Psi^{4/5}\) & \(Q\, \Psi\)\\
-Pipe & \(S\) & \(\Psi \propto f \, Q^2\, S^{-5/2}\) & \(D_h\propto w_c \propto H \propto S^{1/2}\) & \(f Q^2 S^{-2}\) & \(f\, Q^3 S^{-5/2}\)\\
+  Subaerial & \(\Psi\) & \(H \propto f^{1/3}\, Q^{2/3-2\alpha/3} \, \Psi^{-1/3}\) & \makecell{\(w\,\propto Q^\alpha\) \\ \(S=wH\) \\ \(D_h\propto H\)} & \(f^{1/3} Q^{2/3-2\alpha/3}  \Psi^{2/3}\) & \(Q\, \Psi\)\\
+  R-channel & \(\Psi\) & \(S\, \propto f^{2/5}\, Q^{4/5} \, \Psi^{-2/5}\) & \(D_h\propto w \propto H \propto S^{1/2}\) & \(f^{1/5} Q^{2/5} \, \Psi^{4/5}\) & \(Q\, \Psi\)\\
+Pipe & \(S\) & \(\Psi \propto f \, Q^2\, S^{-5/2}\) & \(D_h\propto w \propto H \propto S^{1/2}\) & \(f Q^2 S^{-2}\) & \(f\, Q^3 S^{-5/2}\)\\
 \end{tabular}
@@
 \begin{table}
-  \caption{Sediment transport proxy ($w_c\tau$) and rates for the three considered different transport formulas: MPM \cite{meyer1948}, EH \cite{engelund1967}, and Bagnold \cite{bagnold1980}.
+  \caption{Sediment transport proxy ($w\tau$) and rates for the three considered different transport formulas: MPM \cite{meyer1948}, EH \cite{engelund1967}, and Bagnold \cite{bagnold1980}.
     }
@@
  & Width \(\times \,\, \tau\) & MPM & EH & Bagnold\\
- & \(w_c\, \tau\) & \(Q_s \propto w_c\, \tau^{3/2}\) & \(Q_s \propto w_c\, \tau^{5/2}\) & \(Q_s \propto w_c^{-1/2}\, \Omega^{3/2} H^{-2/3}\)\\
+ & \(w\, \tau\) & \(Q_s \propto w\, \tau^{3/2}\) & \(Q_s \propto w\, \tau^{5/2}\) & \(Q_s \propto w^{-1/2}\, \Omega^{3/2} H^{-2/3}\)\\
 \hline

### 2023-10-09 Delaneyetal2023b.tex
table A1

@@
     Water discharge  & $Q$& & $\mathrm{m^{3}\,s^{-1}}$ \\
-    Water velocity (glacier, subaerial)  & $v_g,\,v_{f}$& & $\mathrm{m\,s^{-1}}$ \\
+    Water velocity (glacier, subaerial)  & $v$, ($v_g,\,v_{f}$)& & $\mathrm{m\,s^{-1}}$ \\
     Channel wetted area (glacier, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
-    Channel depth (subaerial) & &$H$& $\mathrm{m}$\\
+    Channel depth (subaerial) & $H$&& $\mathrm{m}$\\
     Hydraulic diameter &$D_h$&&$\mathrm{m}$\\
-    Width of channel floor (glacial, subaerial) & $w_g,w_f$&  & $\mathrm{m}$     \\
+    Width of channel floor (glacial, subaerial) & $w$, ($w_g,w_f$)&  & $\mathrm{m}$     \\
     Hydraulic head &$\Delta h$&& $\mathrm{m}$\\
@@
 
-    Shear stress (glacier, subaerial) & $\tau_g,\,\tau_f$&& $\mathrm{Pa \, m^{-2}}$ \\
-    Width-integrated shear stress (glacier, subaerial) & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$ \\
+    Shear stress (glacier, subaerial) & $\tau$, ($\tau_g,\,\tau_f$) && $\mathrm{Pa \, m^{-2}}$ \\
+    Width-integrated shear stress (glacier, subaerial) \mauro{get rid of this} & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$\\
     Stream power & $\Omega$ && $\mathrm{ kg \, m\, s^{-3}}$ \\
@@
 Albeit, the sediment transport response in pipe-flow for the Bagnold case is close to EH.
-Conversely, the sediment transport proxy $w_g\,\tau_{g}, w_f\,\tau_{f}$ used within this publication scales only as $Q^2$ for pipe-flow, whereas the others scale at least with $Q^3$.
+Conversely, the sediment transport proxies $w_g\,\tau_g$ \& $w_f\,\tau_f$ used within this publication scale only as $Q^2$ for pipe-flow, whereas the transport scales at least with $Q^3$.
 The exponent for the width scaling $\alpha$ only impacts relationship between sediment transport and water discharge in the EH relation in any meaningful way.

### 2023-10-09 Delaneyetal2023b.tex
General changes and comments

@@
 
-\title{Natural variations in water discharge quantity have an inconsistent impact on sediment transport capacity in subglacial channels}
+\title{Natural variations in water discharge quantity have an inconsistent impact on sediment transport capacity in subglacial channels\mauro{\\Title does not work.
+I don't think ``inconsistent'' can be used like this.
+Maybe:}\\
+Subglacial and subaerial fluvial sediment transport respond differently to discharge variations
+}
 
@@
 \begin{keypoints}
-\item Water discharge variations in subglacial channels are mainly accommodated by water velocity, not channel size.
+\item Short term water discharge variations in subglacial channels are mainly accommodated by water velocity, not channel size.
 \item Greater variability in water velocity causes sediment transport capacity to vary more in subglacial channels than in subaerial channels.
 \item Water discharge may represent processes including proglacial sediment mobilization and sediment access driving variations in sediment discharge, as opposed to subglacial sediment transport capacity in glacierized catchments.
+\mauro{I don't understand this sentence.}
 \end{keypoints}
 
-\begin{abstract}  Sediment transport capacity in subglacial and subaerial channels depends on  water velocity and channel width, which together control the shear stress exerted on the channel bottom.
-  In subaerial channels, width, and velocity evolve quickly with water discharge, while in subglacial channels, water discharge changes mainly affect water velocity due to slow-evolving channel geometry.
-  Parameterizations of these subglacial and subaerial hydraulics are applied to hydrographs from an Alpine glacier and the Greenland Ice Sheet revealing greater variability in sediment transport capacity in subglacial channels.
-  Furthermore, results show that high subglacial sediment transport rates can persist across various water discharges.
-  Slowly evolving channel width accommodates some variability in subglacial channels.
+\begin{abstract}  Sediment transport capacity in subglacial and subaerial channels depends on  water velocity and channel width, which together control the total shear stress exerted on the channel bottom.
+  In subaerial channels water discharge variations are accommodated largely by changes in flow depth and width.
+  While in subglacial channels, water discharge changes primarily affect water velocity due to slow-evolving channel geometry.
+  Parameterizations of these subglacial and subaerial hydraulics are applied to hydrographs from an Alpine glacier and the Greenland Ice Sheet revealing greater variability in the inferred sediment transport capacity in subglacial than in subaerial channels.
+  Results show that high subglacial sediment transport rates can occur across a wide range of water discharges.
+%  Slowly evolving channel width accommodates some variability in subglacial channels. \mauro{not sure what that sentence should do}
   Despite differences in response to water discharge, sediment discharge measurements generally correlate better with subaerial hydraulic parameters.
-  This implies that other factors like sediment access or proglacial processes might be at play, affecting sediment transport capacity differently in subglacial and subaerial channels.
+  This implies that other factors like sediment access or proglacial processes might be at play that affect sediment transport capacity differently in subglacial and subaerial channels.
 \end{abstract}
@@
 
-In a transport-limited regime, sediment discharge responds to the sediment transport capacity, defined as the amount of sediment the water can carry.
+In a transport-limited regime, sediment discharge tracks sediment transport capacity, defined as the amount of sediment the water can carry.
 Sediment transport capacity depends on the shear stress between water and the sediment it flows  over \cite{shields1936,meyer1948,engelund1967} along with the width of the channel bottom $w$ over which to mobilize sediment.
-The shear stress  across the channel width $\tau_t$ responds to the velocity of water $v$ flowing through the channel so that
+The shear stress $\tau$ responds to the velocity of water $v$ flowing through the channel so that
 \begin{linenomath*}
   \begin{equation}
-    \label{eq:tau_t}
-    \tau_t \propto w\, v^2,
+    \label{eq:tau}
+    \tau_t \propto v^2.
   \end{equation}
 \end{linenomath*}
+\mauro{I removed the width from here as I found it odd to introduce that here.}
 %
-where $w$ is the width of the channel. Following mass conservation, the velocity of the water flowing through a  channel is
+Following mass conservation, the velocity of the water flowing through a  channel is
 \begin{linenomath*}
@@
 \end{linenomath*}
-where $Q$ is water discharge,  and $S$ is the channel's wetted area.
+where $Q$ is water discharge,  and $S$ is the channel's wetted area. \mauro{I find it odd to have formulas in the Intro}
 
@@
 The change in water depth results in a commensurate increase in water velocity.
-Thus,  fulfilling Equation~\ref {eq:v}, the higher discharge is accommodated by both an increase in wetted area and flow velocity.
-As a result, the shear stress across the channel $\tau_t$ increases according to Equation~\ref{eq:tau_t}.
+%Thus,  fulfilling Equation~\ref {eq:v}, the higher discharge is accommodated by both an increase in wetted area and flow velocity. \mauro{not needed}
+As a result, the shear stress $\tau$ increases according to Equation~\ref{eq:tau}.
 
 The response of water velocity to changing water discharge in subglacial channels differs from subaerial ones.
-The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by frictional heating of water flowing through the channel \cite{rothlisberger1972}.
+The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by melting due to frictional heating of water flowing through the channel \cite{rothlisberger1972}.
 As a result, the subglacial channel size only evolves relatively slowly over days to months, whereas water discharge can vary over hours \cite<e.g.>{iken1986,andrews2014,nanni2020}.
-Therefore, subglacial water flow behaves more like pipe flow, and changes in water discharge $Q$ are mainly accommodated with increased or decreased water velocity $v$ \cite{alley1997} because the size of the channel $S$ changes on a much longer timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
+Therefore, on short time scales, subglacial water flow behaves more like pipe flow and changes in water discharge $Q$ are mainly accommodated with increased or decreased water velocity $v$ \cite{alley1997} because the size of the channel $S$ changes on a much longer timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
 %It follows that an increase in water discharge $Q$ will cause the shear stress across the channel $\tau_t$ to respond with water velocity $v$, as opposed to subaerial channels where channel width $w$ evolves as well.
@@
       %Note that in the subaerial channel parameterization, channels have rectangular shapes with their width much larger than their depth (Section~\ref{sect:fluv}).
+\mauro{Needs renaming $w_{cg/f}$ to $w_{g/f}$.
+Then, I find the downward arrow confusing.
+Maybe just leave away?
+Also, I would write $\tau_f$ and $\tau_g$, it's ok to not have the width integrated tau here (and maybe we remove $\tau_{tf}$ from the paper anyway...}
     }
@@
 
-As a result, sediment mobilization in subaerial and subglacial channels responds differently to changing water discharge.
+Consequently, sediment mobilization in subaerial and subglacial channels responds differently to changing water discharge.
 These differences are implicitly included in a wide range of available models that  quantify sediment transport in both subglacial and subaerial channels  \cite<e.g.>{walder1994,alley1997,tucker1997,creyts2013,delaney2019,hewitt2019,wickert2019}.
@@
 This manuscript evaluates the relationship between sediment transport capacity and water discharge in subglacial channels and compares it to subaerial ones.
-In the absence of continuous measurements of subglacial and subaerial water velocity and channel morphology, numerical parameterizations of subglacial and subaerial hydraulics are utilized.
+Due to the absence of continuous measurements of subglacial and subaerial water velocity and channel morphology, numerical parameterizations of subglacial and subaerial hydraulics are utilized.
+\mauro{Aren't those things always parameterised, irrespective of whether there are measurements or not?}
 Parameterizations are applied to hydrological records from an Alpine glacier in Switzerland (Fieschergletscher) and  a land-terminating glacier in Greenland (Leverett Glacier).
-The parameterizations' lumped nature isolates the disparity between water discharge and sediment transport capacity in subglacial systems, independent of the upstream drainage network and sediment access.
+Our parameterizations' lumped nature isolates the disparity between water discharge and sediment transport capacity in subglacial systems, independent of the upstream drainage network and sediment access.
 Outputs demonstrate differences in the relationships between  water discharge, channel geometry, and water velocity in the subglacial and subaerial channels.
@@
 The manuscript then uses the demonstrated differences between the two channel types to discuss the implications for interpreting the processes driving variability in records of sediment transport from glacierized catchments.
+\mauro{As stated in the general comments.  I'd rather use ``model'' than ``parameterization'' as the latter is, for me, something which is used in a model and not a model itself.}
 
@@
 To evaluate the shear stress of water flowing across sediments underneath a glacier, the subglacial channel parameterization takes into account the channel geometry and the velocity of the flowing water.
-To do this, a modified version of the lumped hydraulics model presented in \citeA{clarke1996} and \citeA{werder2010} is used.
+To do this, a lumped hydraulics model \cite{clarke1996} modified from the one presented by \citeA{werder2010b} is used.
 
-Here, it is assumed that the water is transported through a subglacial channel \cite<Figure~\ref{fig:cartoon}; >{rothlisberger1972}, and that the channel  size responds to frictional heating from water flow and creep closure by ice, using the Darcy-Weisbach formulation for water-flow through a pipe  \cite<e.g.>{rothlisberger1972,clarke2003}.
+Here, it is assumed that the water is transported through a subglacial channel \cite<Figure~\ref{fig:cartoon}; >{rothlisberger1972}, and that the channel  size responds to melt due to frictional heating from water flow and to creep closure of the ice, using the Darcy-Weisbach formulation for water-flow through a pipe  \cite<e.g.>{rothlisberger1972,clarke2003}.
 The formulation here does not consider the englacial storage of water.
-Thus, it is assumed that the changing hydraulic head at the top of the glacier is negligible in the channel size evolution.
+Thus, it is assumed that the changing hydraulic head at the top of the glacier is negligible in the channel size evolution \mauro{I don't understand this sentence}.
 The evolution of subglacial channel size $S_g$ is given as
@@
 \end{linenomath*}
-\noindent where $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (Table \ref{table:vpm}), $l$ is the length of the subglacial channel, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change from the glacier terminus to its top, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the ice overburden pressure, and $n$ is Glen's n \cite{glen1955} (usually $n=3$).
+\noindent where $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (Table \ref{table:vpm}), $l$ is the length of the subglacial channel, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change from the glacier terminus to its top, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the ice overburden pressure expressed in meter water equivalent, and $n$ is Glen's n \cite{glen1955} (usually $n=3$).
 The first term on the right side of the equation represents the opening of the channel through frictional heating, while the following term represents the creep closure of the channel from ice deformation.
 
-Following the Darcy-Weisbach equation, head drop $\Delta h$ is,
+Following the Darcy-Weisbach equation, the head drop $\Delta h$ is
 \begin{linenomath*}
@@
     \label{eq:dh}
-    \Delta h \,  = l \,\frac{1}{g}\,s\,f_r\,\frac{Q^2}{D_h^5}
+    \Delta h \,  = l \,\frac{1}{2g} \,f_r\,\frac{v_g^2}{D_h},
   \end{equation}
 \end{linenomath*}
-\noindent where $f_r$ is a friction factor, $D_h$ is the hydraulic diameter, and $l$ is the length over which head change occurs.
-$s$ is a factor accounting for channel geometry \cite{hooke1990}, calculated as
-\begin{equation}
-  \label{eq:Hf}
-  s = \frac{2\,(\beta -\sin \beta)^2}{(\frac{\beta}{2}\,+\,\sin \frac{\beta}{2})^4},
-\end{equation}
-where $\beta$ is the central angle of the circular segment that comprises the channel (the so-called Hooke angle). Note that $\beta =\pi$ corresponds to a semi-circle and
-smaller values of $\beta$ result in shallow, wide channels.
-The hydraulic diameter $D_h$ is converted to wetted area $S_g$ given a channel geometry following \cite{hooke1990}
+\noindent where $f_r$ is a friction factor, $D_h$ is the hydraulic diameter, $l$ is the length over which head change occurs, and $v_g=Q/S_g$ is the water flow speed.
+%
+The hydraulic diameter $D_h$ is converted to wetted area $S_g$ with
 \begin{equation}
   \label{eq:Dh2S}
-  S_g= \frac{D_h^2}{2}\,\frac{(\frac{\beta}{2}+\sin \frac{\beta}{2})^2}{\beta - \sin \beta}.
+  S_g= \frac{D_h^2}{2}\,\frac{(\frac{\beta}{2}+\sin \frac{\beta}{2})^2}{\beta - \sin \beta},
 \end{equation}
-
-With wetted area $S_g$, the shear stress, $\tau_g$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
+where $\beta$ is the central angle of the circular segment that comprises the channel (the so-called Hooke angle, \citeA{hooke1990}).
+Note that $\beta =\pi$ corresponds to a semi-circular channel and smaller values of $\beta$ result in shallow, wide channels.
+%
+The shear stress, $\tau_g$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
 \begin{equation}
@@
 where $v_g = \frac{Q}{S_g}$ is the water velocity.
-
+%
 The width-integrated shear stress is represented as
@@
 \end{equation}
-
 where  channel floor width $w_g$ is
@@
 \end{equation}
+\mauro{If we get rid of $\tau_{tf}$, above needs a bit of massaging.  I'd then move the $w_g$ eq to just after the $S_g$ eq.}
 
@@
 
-To parameterize the shear stress of water flowing across sediments in the subaerial channel, the hydraulics parameterization presented in \citeA{tucker1997} is implemented.
+% To parameterize the shear stress of water flowing across sediments in the subaerial channel, the hydraulics parameterization presented in \citeA{tucker1997} is implemented.
 Using mass conservation and the Darcy-Weisbach relationship as well as assuming that the channel is wide so that the hydraulic radius is well approximated by the flow depth, then
@@
 \end{linenomath*}
-where $\nabla z_c$ is the channel slope, and $f_p$ is the friction factor for subaerial channels.
+where $\nabla z_c$ is the channel slope, and $f_p$ is the friction factor for subaerial channels \cite{tucker1997}.
 Channel width $w_f$ is
@@
 
+  Note that the subaerial channel model \mauro{or ``parameterization''} is a purely algebraic model, whereas the subglacial model comprises a differential equation for the evolution of $S_g$ and thus will have a history dependence on the discharge.
+
 \subsection{Implementation}
@@
 
-The parameterizations above are applied to hydrological records from the Fieschergletscher (\alpine) and the Leverett Glacier (\icesheet).
+\mauro{I would call the ``cases'' ``scenarios'' instead.
+The word ``scenario'' does not appear in the manuscript so far, so it would be unambiguous.
+Also, ``cases'' has the disadvantage that ``In both cases,'' is a saying, and thus it is confusing.
+I implemented this change.}
+
+The parameterizations above are applied to hydrological records from the Fieschergletscher (scenario \alpine) and the Leverett Glacier (scenario \icesheet).
 The subglacial and subaerial parameterizations represent the subglacial flow of water exiting a glacier and transitioning to subaerial flow.
 Outputs of the parameterizations represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
-To generalize these cases, \alpine{}  is exemplified by relatively thin ice thickness ($h_{ice}$: $225$\,\unit{m}), low water discharge ($\sim\,10$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
+To generalize these scenarios, \alpine{}  is exemplified by relatively thin ice thickness ($h_{ice}$: $225$\,\unit{m}), low water discharge ($\sim\,10$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
 \icesheet{}  is exemplified by thick ice  ($h_{ice}$: $700$\,\unit{m}), high water discharge ($\sim\,300$\,\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).
+%
+However, both scenarios share a discharge with a similar, relative seasonal amplitude.
 
-For each glacier, the parameterizations are first applied to a reference test case for each glacier that assumes  a subglacial channel with $\beta=\frac{\pi}{2}$.
+For each glacier, the parameterizations are first applied to a reference test case for each glacier that assumes  a subglacial channel with $\beta=\frac{\pi}{2}$, i.e. a segment of a quarter-circle.
 Friction factors $f_r$ and $f_p$ are tuned so that the parameterizations reproduce reasonable water velocities for both \alpine and \icesheet{}\cite<$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$>{werder2010b,chandler2013}.
-In both subaerial channels, the bed slope is $\nabla z_c = 0.05$.
+\mauro{Those $f$ reported in the table are super large!  Maybe an issue down the line...}
+The subaerial channels of both scenarios have a bed slope of $\nabla z_c = 0.05$ \mauro{Why do you report a value here but most are hidden in the table?  Maybe just ref the table?}.
 % Water velocity, shear stress, and width-integrated shear stress in the subglacial and subaerial channels from this formulation are presented below.
 
-Second, to further characterize the variability in sediment discharge capacity in subglacial compared to subaerial channels, $100$  parameterizations are applied to a range of different channel geometry factors ($\beta$ and $k$) and friction factors ($f_r$ and $f_p$).
-Runs with parameter combinations are culled if their mean subglacial water velocity over the season lies between $0.75$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $2$\,\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity lies between $0.3$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with measurements \cite<e.g.>{werder2010b,magnusson2012,chandler2013}.
-The variability of output is represented by the  standard deviation of that output over the course of the season.
+Second, to further characterize the variability in sediment discharge capacity in subglacial compared to subaerial channels, $100$  parameter combinations are applied to a range of different channel geometry factors ($\beta$ and $k$) and friction factors ($f_r$ and $f_p$).
+\mauro{So how many runs are actually run?
+Not clear.}
+Runs with parameter combinations are culled if their mean subglacial water velocity over the season lies outwith $0.75$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $2$\,\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity lies outwith $0.3$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with measurements \cite<e.g.>{werder2010b,magnusson2012,chandler2013} \mauro{Awkward sentence.
+(also note that ``cull'' means to remove, thus I changed ``between'' to ``outwith''.}.
+The variability of output is represented by the  standard deviation of that output over the course of the season \mauro{This sentence needs context.  What is variability used for?}.
 
-Lastly, outputs are compared with the sediment discharge records of the glaciers using Spearman rank correlation, dependent on the ordering of values.
-Rank correlation reduces the impact of the non-linear response of sediment transport capacity to the hydrological forcing.
-Outputs and observations at $15$\,\unit{min} intervals are aggregated across timescales of up to $15$\,\unit{days} by averaging values one-half of the timescale before and after a point time.
+Lastly, outputs of $w\tau$ are compared with the sediment discharge records of the glaciers using Spearman rank correlation, dependent on the ordering of values.
+Rank correlation reduces the impact of the non-linear response of sediment transport capacity to the hydrological forcing, thus making our interpretation independent of the exact details of the sediment transport formula.
+Outputs and observations, which have a $15$\,\unit{min} resolution, are aggregated across time intervals of up to $15$\,\unit{days} by averaging values over a symmetric window of corresponding size.
 
@@
   \includegraphics[width=0.9\linewidth]{Fig2.pdf}
-  \caption{Parameterization outputs from the hydrographs n panels a and e for \alpine cases (a-d) and \icesheet{} (e-h).
+  \caption{Parameterization outputs from the hydrographs n panels a and e for \alpine scenarios (a-d) and \icesheet{} (e-h).
     Gray (black) lines in a and e represent the subglacial channel size (water discharge).
     Blue lines represent outputs from the subglacial channel, while orange lines represent the subaerial channel.
-    Data are shown at $15$\unit{min} intervals.
-    Arrows denote where variables peaking in the subglacial channel prior to the subaerial one.
-    Insets in e-h show the  peak melt event \icesheet{} denoted by the shaded area.
-    %Note that the x and y axes are different for \alpine{} and \icesheet{}.
+    Data are shown at $15$\unit{min} intervals, i.e. not aggregated.
+    Arrows denote where variables peaking in the subglacial channel prior to the subaerial one \mauro{wouldn't this need more zoomed-in plots?}.
+    Insets in e-h show the  peak melt event \icesheet{} denoted by the shaded area in panel e.
+    % Note that the x and y axes are different for \alpine{} and \icesheet{}.
+        \mauro{Wrong units! Shear stress is Pa\,m$^{-2}$ and width integrated Pa\,m$^{-1}$}
   }
@@
 
+\mauro{Good para}
 Following the parameterizations defined in Sections~\ref{sect:sub_mode}~and~\ref{sect:fluv}, the water velocity, shear stress, and width-integrated shear stress exhibit different seasonal evolutions and peaks.
@@
 
-In the subglacial channel, peaks generally occur when water discharge increases at the fastest rate (Figure~\ref{fig:model_outs}).
-Alternatively, in the subaerial channel,  sediment transport variables peak  hen water discharge is highest.
-The timing of the greatest rate of change in water discharge occurs before the peaks in water discharge.
-Furthermore, the subglacial outputs are generally larger than the subaerial values, thus more efficient at transporting sediment.
-Here, sediment could be deposited in the proglacial area as the water depressurizes, then remobilized when sediment transport capacity peaks in the subaerial system.
+\mauro{In principle good paragraph, the meat.  But suffers from sloppy writing.}
+In the subglacial channel, peaks in $w_g\tau_g$ generally occur when water discharge increases at the fastest rate (Figure~\ref{fig:model_outs}d,h).
+In contrast, in the subaerial channel,   $w_g\tau_g$  peaks when water discharge is highest.
+The timing of the greatest rate of change in water discharge occurs before the peaks in water discharge \mauro{what?}.
+Furthermore, the subglacial outputs \mauro{What are ``outputs''?}  are generally larger than the subaerial values, thus more efficient at transporting sediment.
+Here, sediment could be deposited in the proglacial area as the water depressurizes, then remobilized when sediment transport capacity peaks in the subaerial system \mauro{not sure this tangent is worth taking in this paragraph}.
 
@@
     These outputs' relationship with measured sediment discharge is given for \alpine in (g-i) and \icesheet (j-l).
-    Variables have been normalized to mean values.
+    Variables on the x-axis have been normalized to mean values.
     $R_g$ and $R_f$ show the Spearman rank correlations for the subglacial and the subaerial outputs, respectively.
     Plots are shown with data and outputs at $15$\unit{min} intervals.
+    \mauro{Pixely, needs more resolution, at least 2x. The two dots in the legend are not visible.  I would add ``Measured'' to the two y-labels.}
   }
@@
 
-Each water discharge value in the subaerial channel results in a unique water velocity, shear stress, and width-integrated shear stress, creating a perfect rank correlation between the variables (Figure~\ref{fig:Qw_vari} a--f).
+In the subaerial system there is no history dependence and thus each water discharge value in the subaerial channel results in a unique water velocity, shear stress, and width-integrated shear stress, creating a perfect rank correlation between the variables (Figure~\ref{fig:Qw_vari} a--f).
 
-In the subglacial channel, however, the persistence of high water velocity, shear stress, and width-integrated shear stress across a range of discharges suggests that substantial sediment transport may occur across  (Figure~\ref{fig:Qw_vari} a--f).
+\mauro{it feels this sentence should be in the same para as the previous one.  But not sure what it tries to say.}
+In the subglacial channel, however, the persistence of high water velocity, shear stress, and width-integrated shear stress across a range of discharges suggests that substantial sediment transport may occur across \mauro{across what?} (Figure~\ref{fig:Qw_vari} a--f).
 
 For instance, very high  water velocity values and shear stresses can occur across minimal water discharge at $2.5$\,\unit{m}$^3$\,\unit{s}$^{-1}$ to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
-In \icesheet, mean values of water velocity can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
-When considering the subglacial channel's evolving width, width-integrated shear stress  across the channel generally increases with water discharge, with improved rank correlation compared to water velocity or shear stress (Figure~\ref{fig:Qw_vari} a--f).
+In \icesheet, mean values \mauro{what ``mean''?} of water velocity can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
+When considering the subglacial channel's evolving width, width-integrated shear stress  across the channel generally increases with water discharge, with improved rank correlation compared to water velocity or shear stress (\alpine, Figure~\ref{fig:Qw_vari} a--f).
 Yet even the width-integrated shear stress  can vary substantially, with the highest values occurring at water discharge values ranging from roughly $11$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ to over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
-The variability in width-integrated shear stress is less pronounced in the \icesheet case, with reduced variability in water discharge (Figure~\ref{fig:Qw_vari} c, f).
+The variability in width-integrated shear stress is less pronounced in the \icesheet scenario, with reduced variability in water discharge (Figure~\ref{fig:Qw_vari} c, f).
 
@@
       Solid lines denote  mean values.
-      %Top row shows results from the \alpine{} case, while bottom row show outputs from the \icesheet{} case
+      %Top row shows results from the \alpine{} scenario, while bottom row show outputs from the \icesheet{} scenario
       Time intervals shown with markers denote $15$\,\unit{min}, $1$\,\unit{hr}, $12$\,\unit{hr}, $1$\,\unit{d}, $5$\,\unit{d}, $10$\,\unit{d}, and $15$\,\unit{d} aggregation periods.
-      Text in columns shows rank correlation between the variable and sediment discharge for the corresponding time period for the subglacial (right column) and subaerial (left column) channels.
-    }
+      Text in columns shows rank correlation between the variable and sediment discharge for the corresponding time period for the subglacial (left column) and subaerial (right column) channels.
+      \mauro{Again units are wrong as in Fig 2.
+        I would add \alpine and \icesheet labels to the plots; and for sure that needs to be stated in the caption.
+        It's a bit confusing to have the rank-correlations here as well as they are not on the variability but just on the variable themselves... not sure what to do though. } }
     \label{fig:multi_run}
@@
 
-Across the range of channel shape parameters and friction factors  examined, variability in water velocity and shear stress remains higher in the subglacial system compared to the subaerial one.
-In \alpine{}, the variability in width-integrated shear stress is comparable between the two systems, due to the variable water discharge (Figure~\ref{fig:multi_run}).
+Across the range of channel shape parameters and friction factors  examined, variability in water velocity and shear stress remains higher in the subglacial system compared to the subaerial one (Figure~\ref{fig:multi_run}).
+In \alpine{}, the variability in width-integrated shear stress is comparable between the two systems, due to the variable water discharge.
 However, in \icesheet{}, with lower water discharge fluctuations, the variability in width-integrated shear stress remains higher in the subglacial system across the time aggregation periods.
 
-It is prescient to evaluate the time scales over which the subglacial channels may respond to changing water discharge to understand the influence of variability in sediment transport capacity.
-In both cases, the variability in velocity, shear stress, and width-integrated shear stress decreases after $1-5$ day aggregations (Figure~\ref{fig:multi_run}).
+\mauro{this paragraph needs more motivation: why are we looking at this?}
+It is prescient \mauro{I don't think that works ``to have foresight''} to evaluate the time scales over which the subglacial channels may respond to changing water discharge to understand the influence of variability in sediment transport capacity.
+In both scenarios, the variability in velocity, shear stress, and width-integrated shear stress decreases with aggregations longer than approximately 1--5 days (Figure~\ref{fig:multi_run}).
 Yet, substantial differences in variability persist in periods of up to $15$ days.
-Aggregations periods longer than this might be conflated with seasonal variations in water discharge.
-Note that lags and time of peak sediment transport conditions still occur over these periods (Figure~\ref{fig:model_outs}), while the variability in sediment transport parameters decreases with larger time aggregation (Figure~\ref{fig:multi_run}).
+Aggregations periods longer than this might be conflated with seasonal variations in water discharge and are thus omitted.
+Note that lags and time of peak sediment transport conditions still occur over these periods (Figure~\ref{fig:model_outs}), while the variability in sediment transport parameters decreases with larger time aggregation (Figure~\ref{fig:multi_run}) \mauro{not clear sentence}.
 
 \subsection{Relationship between parameterization  outputs and observed sediment discharge}
+\mauro{Not sure a separate section is needed for this one paragraph.}
 
@@
 
-Quantities that control sediment transport capacity (water velocity, shear stress, and width-integrated shear stress) vary differently in response to water discharge in subglacial channels compared to subaerial ones (Figure~\ref{fig:Qw_vari}).
+Quantities that relate to sediment transport capacity (water velocity, shear stress, and width-integrated shear stress) vary differently in response to water discharge in subglacial channels compared to subaerial ones (Figure~\ref{fig:Qw_vari}).
 In sediment transport relationships that evaluate sediment discharge capacity, such as in \citeA{meyer1948} or \citeA{engelund1967}, shear stress is scaled to the power of $\frac{3}{2}$ or $\frac{5}{2}$, respectively.
 The exponent greater than $1$ magnifies sediment discharge variability beyond the variable sediment transport parameters described above (Figure~\ref{fig:multi_run}; Table~\ref{tab:Qs}).
-Note that sediment transport relationships describe sediment flux per unit width, so that the integrated capacity across channel occurs after application of the exponent, unlike the width-integrated shear stress evaluation here \cite<Equations\,\ref{eq:tautg} and \ref{eq:tautf}; c.f.>{tucker1997}.
+Note that sediment transport relationships describe sediment flux per unit width, so that the integrated capacity across channel occurs after application of the exponent, unlike the width-integrated shear stress evaluation here \cite<Equations\,\ref{eq:tautg} and \ref{eq:tautf}; c.f.>{tucker1997} \mauro{sentence does not parse; I think you should state limitations of width integrated shear stress and then be done with it.}.
 %Note also that both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress is homogeneous across the channel bed, despite the limitations of this assumption \cite<e.g.>{yager2018}.
 
+\mauro{This paragraph needs rewriting.  I guess I should do that.}
 The rapid increases and decreases in subglacial sediment transport capacity compounds the variability in sediment transport already present in subaerial systems \cite<Figure~\ref{fig:Qw_vari}; >{williams1989,jerolmack2010}.
 Sediment transport capacity in subglacial channels responds more rapidly to changes in water discharge compared to their subaerial counter parts \cite{alley1997}.
-In subglacial R\"othlisberger channels where water discharge evolves more slowly than channel size or subaerial channels the width-integrated shear stress response to water discharge is $ \tau_{tg}\, \propto\,  Q^{\sim\frac{4}{5}}$ (Table\,\ref{tab:eqs1}).
+In subglacial R\"othlisberger channels in steady state with respect to water discharge the width-integrated shear stress response to water discharge is $ \tau_{tg}\, \propto\,  Q^{\sim\frac{4}{5}}$ (Table\,\ref{tab:eqs1}).
 However, subglacial channels rarely operate in a steady state \cite<e.g.>{gimbert2016}.
-For sufficiently short time periods, channel width and wetted area are fixed in the subglacial system (assuming pipe flow) so that the response of width-integrated shear stress is $ \tau_{tg}\, \propto\,  Q^{2}$ (Table\,\ref{tab:eqs1}).
+For sufficiently short time periods, channel width and wetted area are fixed in the subglacial system (akin to pipe flow) so that the response of width-integrated shear stress is $ \tau_{tg}\, \propto\,  Q^{2}$ (Table\,\ref{tab:eqs1}).
 \ref{sect:scaling} provides a complete description of sediment transport capacity in different drainage conditions.
 
-The stronger correlation between subglacial hydraulics and water discharge in \icesheet{} could result from the channel being closer to equilibrium between variations in the channels' wetter area and water discharge with smaller diurnal variations in amplitude (Figures~\ref{fig:model_outs} and ~\ref{fig:Qw_vari}).
-Conversely, the correlation between hydraulics and water discharge in \alpine is smaller due to the larger discharge variations.
-In both cases, hydraulics could trend toward equilibrium by averaging the discharge over $1-5$ days, when the variability in subglacial channels approaches that of subaerial one (Figure~\ref{fig:multi_run}).
+\mauro{You often use ``subglacial hydraulics''.
+I'm not sure what that means.
+For instance in the next sentence.
+Isn't discharge part of hydraulics too?
+Maybe you could define a bunch of equations as the subglacial hydraulic model and subaerial hydraulic model and then talk of that?}
+%
+The stronger correlation between subglacial hydraulics and water discharge in \icesheet{} (Figure~\ref{fig:Qw_vari}a-f) could result from the channel being closer to equilibrium between variations in the channels' wetter area and water discharge with smaller diurnal variations in amplitude (Figures~\ref{fig:model_outs} and ~\ref{fig:Qw_vari}) \mauro{needs rewriting}.
+Conversely, the correlation between hydraulics and water discharge in \alpine is smaller due to the relatively larger discharge variations on short time scales.
+In both scenarios, hydraulics could \mauro{``could''?
+Does it or does it not?} trend toward equilibrium by averaging the discharge over $1-5$ days, when the variability in subglacial channels approaches that of subaerial ones (Figure~\ref{fig:multi_run}).
 
-In subglacial channels, peaks in water velocity, shear stress and width-integrated shear stress occur with  fast rates of water discharge increase relative to the channel's growth (Figure~\ref{fig:model_outs}).
-In turn, subglacial channels grow with increasing water discharge, effectively reduces sediment transport capacity with the stabilization of water discharge at its peak.
-Similar feedback occurs in subaerial systems \cite{phillips2016}, albeit at timescales longer than days \cite{gimbert2016}.
+In subglacial channels, peaks in water velocity, shear stress and width-integrated shear stress occur with  fast rates of water discharge increase relative to the channel's growth rate (Figure~\ref{fig:model_outs}).
+In turn, subglacial channels grow with increasing water discharge, effectively reducing sediment transport capacity with the stabilization of water discharge at its peak.
+Similar feedback occurs in subaerial systems \cite{phillips2016}, albeit at timescales longer than days.  \mauro{\citeA{gimbert2016}, which you cited here, says nothing of subaerial flow, or do they?}
 
 \subsection{Records of sediment discharge  from glaciers}
+\mauro{This section header sounds like this will be about measurements only.}
 
+\mauro{I find this confusing.
+The measurements are sediment discharge, so how can they be represent by hydraulic outputs?
+Maybe talk of sediment discharge proxy ($w\tau$) or some such?
+Or use something establishing a less direct link than ``representing''.}
 The increased performance of subaerial hydraulics outputs, compared to subglacial ones, in representing observations occurs despite the proximity of measurement locations to these glaciers (Figure~\ref{fig:Qw_vari}).
@@
 First, these measurement stations lie $0.5$ to $2$\,\unit{km} downstream of the glacier terminus \cite{cowton2012,felix2022}, a long enough distance for the pressurized subglacial regime to transition to the open channel subaerial one.
-The transition of subglacial to subaerial conditions may occur underneath glacier terminus \cite{perolo2018}, especially if closure rates of the subglacial channels are slow  \cite{rothlisberger1972}.
-Furthermore, if a transport-limited regime persists at the glacier margin, then sediment can be easily mobilized, and subaerial processes could be represented just a short distance downstream.
+The transition of subglacial to subaerial conditions typically occur still underneath glacier but close to the terminus \cite{perolo2018}, where closure rates of the subglacial channels are slow and open channel flow condition prevail \cite{rothlisberger1972}.
+Furthermore, if a transport-limited regime persists at the glacier margin, then sediment can be easily mobilized, and subaerial processes could be represented just a short distance downstream \mauro{Not sure what that sentence means}.
 
-The consistent relationship between sediment discharge capacity and open channel flow conditions could be pronounced $\sim20$\,\unit{km} downstream of the Leverett site, where a strong correlation persists between sediment plume size and the river's water discharge into the Kangerlussuaq fjord \cite{mcgrath2010}.
+\mauro{I perceived this paragraph pretty out of the blue.}
+The consistent relationship between sediment discharge capacity and open channel flow conditions \mauro{What?
+How do you compare those two things?} could be pronounced $\sim20$\,\unit{km} downstream of the Leverett site, where a strong correlation persists between sediment plume size and the river's water discharge into the Kangerlussuaq fjord \cite{mcgrath2010}.
 In contrast, in marine-terminating glacier catchments a less consistent relationship may occur between water discharge or melt extent and sediment plume size \cite{tedstone2012}.
@@
 Sediment exhaustion occurring through increased variability may also explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \cite{overeem2017}.
-Note, however, the absence of lags between the observed sediment and water discharge may suggest that sediment is mobilized close to the measurement stations, as opposed to distal areas of the glacier bed as would be expected in a supply-limited regime \cite<Figure~\ref{fig:model_outs}\, a and e; >{williams1989}.
+Note, however, the absence of lags between the observed sediment and water discharge \mauro{reference a figure here?} may suggest that sediment is mobilized close to the measurement stations, as opposed to distal areas of the glacier bed as would be expected in a supply-limited regime \cite<Figure~\ref{fig:model_outs}\, a and e; >{williams1989}.
 
-Regardless of the increased performance of water discharge compared to subglacial sediment transport capacity in the observations, outputs suggest that additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate \cite{jansson2005,ganti2016}.
+Regardless of the increased performance of water discharge compared to subglacial sediment transport capacity in the observations \mauro{does not parse}, outputs suggest that additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate \cite{jansson2005,ganti2016}.
 Sediment discharge records have been used to establish the relationship, or lack thereof, between climate forcing and glacial erosion \cite<e.g.>{koppes2009a,ganti2016,willenbring2016,mariotti2021}.
@@
 Yet, identifying climatic signals in sediment transport from the transport-limited component of glacier-sedimentary systems may require higher thresholds of water discharge compared to subaerial systems when water discharge varies substantially over short timescales \cite<Figure~\ref{fig:Qw_vari}; >{tofelde2021}.
+\mauro{I don't understand the last two sentences.}
 
-Over time scales of weeks or longer, the glacier's sediment transport capacity responds to the ice thickness, controlling the channel closure rate, and the glacier's surface slope, in addition to water discharge \cite<Figure~\ref{fig:multi_run}, Section~\ref{sect:sub_mode}; >{rothlisberger1972,shreve1972,gimbert2016,stevens2022}.
-In a supply-limited state, often assumed in landscape evolution models, then glaciers' sediment discharge record also represents additional processes of sediment availability and bedrock erosion from  water pressure variations and sliding  \cite<e.g.>{iverson2012,herman2015,delaney2023}.
-This multitude of processes lies in contrast to most subaerial systems, where transport capacity typically responds to sediment size, water discharge, and hydraulic gradient, and the hydraulic gradient generally remains stable over time scales ranging from years to millennia \cite<Section~\ref{sect:fluv}; e.g.>{muller1968,tucker1997,wickert2019}.
+Over time scales of weeks or longer, the glacier's sediment transport capacity is impacted by the ice thickness, controlling the channel closure rate, and the glacier's surface slope, in addition to water discharge \cite<Figure~\ref{fig:multi_run}, Section~\ref{sect:sub_mode}; >{rothlisberger1972,shreve1972,gimbert2016,stevens2022}.
+In a supply-limited state, often assumed in landscape evolution models, then glaciers' sediment discharge record are also impacted by the additional processes of sediment availability and bedrock erosion from  water pressure variations and sliding  \cite<e.g.>{iverson2012,herman2015,delaney2023}.
+This multitude of processes lies in contrast to processes occurring in most subaerial systems, where transport capacity typically responds to sediment size, water discharge, and hydraulic gradient, and the hydraulic gradient generally remains stable over time scales ranging from years to millennia \cite<Section~\ref{sect:fluv}; e.g.>{muller1968,tucker1997,wickert2019}.
 The stronger link between sediment transport capacity and hydrological forcing in subaerial channels may mean that variations in sediment transport records can be more easily attributed to climatic conditions than in subglacial systems.
@@
 
-Sediment transport capacity is largely driven by the shear stress of water flowing through a channel and the channel's width.
-Subaerial channels  alter their channel width and water velocity immediately in response to changing water discharge.
-In contrast, pressurized subglacial channels largely accommodate changing water discharge by altering water velocity and shear stress, upon which sediment transport depends.
-In subglacial channels, water velocity responds quickly to changes in water discharge because their size generally evolves slowly compared to variations in water discharge.
-This makes the response of sediment transport capacity much more sensitive to changes in water discharge when compared to subaerial channels.
-Despite these differences, parameterizations of subaerial hydraulics in two glacierized catchments generally represent observed sediment discharge variations better than subglacial hydraulics.
+The sediment transport capacity of a channel is driven by its width and the shear stress exerted by the flowing water, which, in turn, is proportional to the square of the flow velocity.
+Subaerial channels alter both their channel width and water velocity immediately in response to changing water discharge.
+In contrast, pressurized subglacial channels largely accommodate rapidly changing water discharge by altering water velocity but not their size; the latter, in contrast, only responds over the course of several days to changes in discharge.
+Thus the response of sediment transport capacity is much more sensitive to changes in water discharge in subglacial channels compared to subaerial ones.
+Despite these differences, parameterizations of subaerial hydraulics in two glacierized catchments generally represent observed sediment discharge variations better than subglacial hydraulics \mauro{This sentence does not quite parse.
+Why ``despite''?
+Also state that we found this.}
 
@@
 \begin{enumerate}
-\item the inconsistent relationship between water discharge and sediment transport capacity underneath glaciers creates increased variability in sediment transport capacity compared to river systems;
-\item the timing of peak sediment transport capacity below glaciers generally coincides with large rates of increasing in water discharge, as opposed to the maximum quantity of water discharge;
-\item a consistent relationship between water and sediment discharge may represent subglacial sediment access or open channel hydraulics near the glacier margin, instead of subglacial sediment transport capacity.
+\item the inconsistent \mauro{``inconsistent'' does not work for me here.
+What do you mean?
+More non-linear?} relationship between water discharge and sediment transport capacity underneath glaciers creates larger variability in sediment transport capacity compared to river systems;
+\item the timing of peak sediment transport capacity below glaciers generally coincides with large rates of increase in water discharge, as opposed to the maximum water discharge;
+\item the timing of peak sediment transport capacity in subaerial channels coincides with the maximum water discharge;
+\item a consistent \mauro{and what does ``consistent'' mean?} relationship between water and sediment discharge may represent subglacial sediment access or open channel hydraulics near the glacier margin, instead of subglacial sediment transport capacity.
 \end{enumerate}
@@
   \caption{Variables, parameters, and constants used in this work.
-    Where two values are given, the first refers to  \alpine{}, a case from Fieschergletscher, and the second to a glacier marginal to the \icesheet{} a case from Leverett Glacier}
+    Where two values are given, the first refers to  \alpine{}, a scenario from Fieschergletscher, and the second to a glacier marginal to the \icesheet{} a scenario from Leverett Glacier}
+  \tiny % TODO remove again
   \begin{tabular}{ l  c  c c }
-    Name &Symbol&  Value&Units \\ \hline
+    Name &Symbol&  Value&Units \\
+    && (\alpine or \icesheet)\\
+    \hline
     \textbf{Variables}  & & & \\
     Water discharge  & $Q$& & $\mathrm{m^{3}\,s^{-1}}$ \\
-    Water velocity (glacier, subaerial)  & $v$, ($v_g,\,v_{f}$)& & $\mathrm{m\,s^{-1}}$ \\
-    Channel wetted area (glacier, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
+    Water velocity (subglacial, subaerial)  & $v$, ($v_g,\,v_{f}$)& & $\mathrm{m\,s^{-1}}$ \\
+    Channel wetted area (subglacial, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
     Channel depth (subaerial) & $H$&& $\mathrm{m}$\\
     Hydraulic diameter &$D_h$&&$\mathrm{m}$\\
-    Width of channel floor (glacial, subaerial) & $w$, ($w_g,w_f$)&  & $\mathrm{m}$     \\
+    Width of channel floor (subglacial, subaerial) & $w$, ($w_g,w_f$)&  & $\mathrm{m}$     \\
     Hydraulic head &$\Delta h$&& $\mathrm{m}$\\
@@
 
-    Shear stress (glacier, subaerial) & $\tau$, ($\tau_g,\,\tau_f$) && $\mathrm{Pa \, m^{-2}}$ \\
-    Width-integrated shear stress (glacier, subaerial) \mauro{get rid of this} & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$\\
+    Shear stress (subglacial, subaerial) & $\tau$, ($\tau_g,\,\tau_f$) && $\mathrm{Pa \, m^{-2}}$ \\
+    Width-integrated shear stress (subglacial, subaerial) \mauro{get rid of this} & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$\\
     Stream power & $\Omega$ && $\mathrm{ kg \, m\, s^{-3}}$ \\
@@
     Effective glacier thickness &$h_o$&$\frac{\rho_i}{\rho_w} h_{ice}$  &\unit{m}\\
-    Effective glacier length &$l$&$7000$ or $26,000$&\unit{m}\\
+    Effective glacier length &$l$& 7,000 or 26,000&\unit{m}\\
     Constant $1$ in Equation~\ref{eq:dS_dt} &$C_1$&$2.2\times10^{-5}$&\unit{m}$^{-1}$\\
@@
     Ice flow exponent &$n$& $3$ &$\mathrm{(-)}$\\
-    Friction factor (subglacial) & $f_r$ &$10$ or $20$ & $\mathrm{(-)}$ \\
+    Friction factor (subglacial, subaerial)  \mauro{why not $f_g$ and $f_f$?} & $f_r$ &$10$ or $20$ & $\mathrm{(-)}$ \\
     Friction factor (subaerial) & $f_p$ & $3$ & $\mathrm{(-)}$\\

### 2023-10-10 Delaneyetal2023b.tex
small bit before mauro's comments

--- 2023-10-09 Delaneyetal2023b.tex
@@
 
-\begin{abstract}  Sediment transport capacity in subglacial and subaerial channels depends on  water velocity and channel width, which together control the shear stress exerted on the channel bottom.
+\begin{abstract}
+  Sediment transport capacity in subglacial and subaerial channels depends on  water velocity and channel width, which together control the shear stress exerted on the channel bottom.
   In subaerial channels, width, and velocity evolve quickly with water discharge, while in subglacial channels, water discharge changes mainly affect water velocity due to slow-evolving channel geometry.
@@
 In a transport-limited regime, sediment discharge responds to the sediment transport capacity, defined as the amount of sediment the water can carry.
-Sediment transport capacity depends on the shear stress between water and the sediment it flows  over \cite{shields1936,meyer1948,engelund1967} along with the width of the channel bottom $w$ over which to mobilize sediment.
+Sediment transport capacity depends on the shear stress between water and the sediment it flows  over \cite{shields1936,meyer1948,engelund1967} along with the width of the channel bottom $w_c$ over which to mobilize sediment.
 The shear stress  across the channel width $\tau_t$ responds to the velocity of water $v$ flowing through the channel so that
@@
     \label{eq:tau_t}
-    \tau_t \propto w\, v^2,
+    \tau_t \propto w_c\, v^2,
   \end{equation}
@@
 %
-where $w$ is the width of the channel. Following mass conservation, the velocity of the water flowing through a  channel is
+where $w_c$ is the width of the channel. Following mass conservation, the velocity of the water flowing through a  channel is
 \begin{linenomath*}
@@
 Therefore, subglacial water flow behaves more like pipe flow, and changes in water discharge $Q$ are mainly accommodated with increased or decreased water velocity $v$ \cite{alley1997} because the size of the channel $S$ changes on a much longer timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
-%It follows that an increase in water discharge $Q$ will cause the shear stress across the channel $\tau_t$ to respond with water velocity $v$, as opposed to subaerial channels where channel width $w$ evolves as well.
+%It follows that an increase in water discharge $Q$ will cause the shear stress across the channel $\tau_t$ to respond with water velocity $v$, as opposed to subaerial channels where channel width $w_c$ evolves as well.
 
@@
       $S_g$ and $S_f$ represent the wetted area in subglacial and subaerial channels, respectively.
-      The subglacial channel widths $w_g$ remain unchanged, while the subaerial channel widths $w_f$ evolve. Shear stress $\tau$ is responsible for the mobilization of sediment.
+      The subglacial channel widths $w_{cg}$ remain unchanged, while the subaerial channel widths $w_{cf}$ evolve. Shear stress $\tau$ is responsible for the mobilization of sediment.
       %Note that in the subaerial channel parameterization, channels have rectangular shapes with their width much larger than their depth (Section~\ref{sect:fluv}).
@@
 
-This manuscript evaluates the relationship between sediment transport capacity and water discharge in subglacial channels and compares it to subaerial ones.
+This manuscript evaluates the relationship between sediment transport capacity and water discharge in subglacial channels, and compares it to subaerial ones.
 In the absence of continuous measurements of subglacial and subaerial water velocity and channel morphology, numerical parameterizations of subglacial and subaerial hydraulics are utilized.
@@
 Data is used from \citeA{tedstone2017} and \citeA{hawkings2014} at a $5$\,\unit{min} time interval from May 28, 2012 to August 8, 2012.
-Two gaps (14.07.2012 to 15.07.2012 and 16.07.2012 to 17.07.2021) in suspended sediment data were interpolated using a linear spline.
+Two gaps (14.07.2012 to 15.07.2012 and 16.07.2012 to 17.07.2021) in suspended sediment data interpolated using a linear spline.
 
@@
   \label{eq:tautg}
-  \tau_{tg}=w_g\,\tau_g,
+  \tau_{tg}=w_{cg}\,\tau_g,
 \end{equation}
 
-where  channel floor width $w_g$ is
+where  channel floor width $w_{cg}$ is
 \begin{equation}
   \label{eq:dh2wc}
-  w_g = 2  \sin \frac{\beta}{2} \sqrt{\frac{2\, S_g}{\beta -\sin \beta}}.
+  w_{cg} = 2  \sin \frac{\beta}{2} \sqrt{\frac{2\, S_g}{\beta -\sin \beta}}.
 \end{equation}
@@
     \label{eq:DW_tau}
-    \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q}{w_f} \Big)^{\frac{2}{3}} \,\nabla z_c^{\frac{2}{3}},
+    \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q}{w_{cf}} \Big)^{\frac{2}{3}} \,\nabla z_c^{\frac{2}{3}},
   \end{equation}
@@
 where $\nabla z_c$ is the channel slope, and $f_p$ is the friction factor for subaerial channels.
-Channel width $w_f$ is
+Channel width $w_{cf}$ is
 \begin{equation}
   \label{eq:wcf}
-  w_f = k \, Q^{\alpha},
+  w_{cf} = k \, Q^{\alpha},
 \end{equation}
@@
     \label{eq:tautf}
-    \tau_{tf}=w_f\,\tau_f.
+    \tau_{tf}=w_{cf}\,\tau_f.
   \end{equation}
@@
 The subglacial and subaerial parameterizations represent the subglacial flow of water exiting a glacier and transitioning to subaerial flow.
-Outputs of the parameterizations represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
+Outputs of the parameterizations are meant to represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
 To generalize these cases, \alpine{}  is exemplified by relatively thin ice thickness ($h_{ice}$: $225$\,\unit{m}), low water discharge ($\sim\,10$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
@@
 
-For each glacier, the parameterizations are first applied to a reference test case for each glacier that assumes  a subglacial channel with $\beta=\frac{\pi}{2}$.
+For each glacier, the parameterizations are first applied to a reference test case for each glacier that assumes  subglacial channel with $\beta=\frac{\pi}{2}$.
 Friction factors $f_r$ and $f_p$ are tuned so that the parameterizations reproduce reasonable water velocities for both \alpine and \icesheet{}\cite<$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$>{werder2010b,chandler2013}.
@@
 Runs with parameter combinations are culled if their mean subglacial water velocity over the season lies between $0.75$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $2$\,\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity lies between $0.3$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with measurements \cite<e.g.>{werder2010b,magnusson2012,chandler2013}.
-The variability of output is represented by the  standard deviation of that output over the course of the season.
+Variability of an output is represented by the  standard deviation of an output over the course of the season.
 
@@
 Rank correlation reduces the impact of the non-linear response of sediment transport capacity to the hydrological forcing.
-Outputs and observations at $15$\,\unit{min} intervals are aggregated across timescales of up to $15$\,\unit{days} by averaging values one-half of the timescale before and after a point time.
+Outputs and observations at $15$\,\unit{min} intervals are aggregated across timescales of up to $15$\,\unit{days} by averaging values one half of the timescale before and after a point time.
 
@@
 
-Across the range of channel shape parameters and friction factors  examined, variability in water velocity and shear stress remains higher in the subglacial system compared to the subaerial one.
+Across the range of channel shape parameters and friction factors  examined, variability in water velocity and shear stress remains higher in the subglacial system compared to the subaerial ones.
 In \alpine{}, the variability in width-integrated shear stress is comparable between the two systems, due to the variable water discharge (Figure~\ref{fig:multi_run}).
@@
 First, these measurement stations lie $0.5$ to $2$\,\unit{km} downstream of the glacier terminus \cite{cowton2012,felix2022}, a long enough distance for the pressurized subglacial regime to transition to the open channel subaerial one.
-The transition of subglacial to subaerial conditions may occur underneath glacier terminus \cite{perolo2018}, especially if closure rates of the subglacial channels are slow  \cite{rothlisberger1972}.
+The transition of subglacial to subaerial conditions may even occur underneath glacier terminus \cite{perolo2018}, especially if closure rates of the subglacial channels are slow  \cite{rothlisberger1972}.
 Furthermore, if a transport-limited regime persists at the glacier margin, then sediment can be easily mobilized, and subaerial processes could be represented just a short distance downstream.
@@
 
-Over time scales of weeks or longer, the glacier's sediment transport capacity responds to the ice thickness, controlling the channel closure rate, and the glacier's surface slope, in addition to water discharge \cite<Figure~\ref{fig:multi_run}, Section~\ref{sect:sub_mode}; >{rothlisberger1972,shreve1972,gimbert2016,stevens2022}.
+Over time scales of weeks or longer, the glacier's sediment transport capacity responds to the ice thickness, controlling the channel closure rate, and the glacier's surface slope, in addition to water discharge \cite<Figure~\ref{fig:multi_run}, ~ Section~\ref{sect:sub_mode}; >{rothlisberger1972,shreve1972,gimbert2016,stevens2022}.
 In a supply-limited state, often assumed in landscape evolution models, then glaciers' sediment discharge record also represents additional processes of sediment availability and bedrock erosion from  water pressure variations and sliding  \cite<e.g.>{iverson2012,herman2015,delaney2023}.
@@
     Hydraulic diameter &$D_h$&&$\mathrm{m}$\\
-    Width of channel floor (glacial, subaerial) & $w_g,w_f$&  & $\mathrm{m}$     \\
+    Width of channel floor (glacial, subaerial) & $w_{cg},w_{cf}$&  & $\mathrm{m}$     \\
     Hydraulic head &$\Delta h$&& $\mathrm{m}$\\
@@
 
-\mauro{Notation: either use $h$ and $w$ for the river, or otherwise, somehow free $h$. Below I use $H$ and $W$ for now.}
+\mauro{Notation: either use $h$ and $w_c$ for the river, or otherwise, somehow free $h$. Below I use $H$ and $W$ for now.}
 
@@
 R-channels which do not have time to adjust their size to discharge conditions).
-Here we show how sediment transport capacity scales with respect to a given water discharge and hydraulic gradient for three different transport formulas: MPM \cite{meyer1948}, EH \cite{engelund1967}, and Bagnold \cite{bagnold1980}; additionally width-integrated shear stress is assessed, as used as proxy for sediment transport throughout the manuscript.
+Sediment transport capacity is scaled with respect to a given water discharge for three different transport formulas: MPM \cite{meyer1948}, EH \cite{engelund1967}, and Bagnold \cite{bagnold1980}; additionally width-integrated shear stress is assessed, as it is  used as proxy for sediment transport throughout the manuscript.
 
@@
 \end{equation}
-with $\Psi = \Delta h / l$ the head gradient and friction factor $f$.
+with $\Psi$ the head gradient and friction factor $f$.
 The shear stress and stream power are, respectively,
@@
 
-Sediment discharge is given by the MPM, EH and Bagnold formulations as
+Sediment discharge is given by the MPM, EH and Bagnold formulations as, respectively,
 \begin{equation}
-  Q_s \propto w\, \tau^{3/2}, \quad Q_s \propto w\, \tau^{5/2}, \quad Q_s \propto w \left(\frac{\Omega}{w}\right)^{3/2} H^{-2/3},
+  Q_s \propto w_c\, \tau^{3/2}, \quad Q_s \propto w_c\, \tau^{5/2}, \quad Q_s \propto w_c \left(\frac{\Omega}{w_c}\right)^{3/2} H^{-2/3}
 \end{equation}
-respectively, for conditions well above the transport onset threshold.
+for conditions well above the transport onset threshold.
 
-The  subaerial channel is assumed to have a width $w$ much greater than its depth $H$, such that $D_h\approx 4h$, and to have a constant head gradient ($\Psi$) given by the topography.
-Further, it is  assumed that its width can be approximated by a relation $w \propto Q^\alpha$ (Equation~\ref{eq:wcf})with $\alpha\in [0,1]$; typically $\alpha \approx 1/2$ (a so-called regime channel).
+The  subaerial channels is assumed to have a width $w_c$ much greater than its depth $H$, such that $D_h\approx 4h$, and to have a constant head gradient ($\Psi$) given by the topography.
+Further, it is  assumed that its width can be approximated by a relation $w_c \propto Q^\alpha$ (Equation~\ref{eq:wcf})with $\alpha\in [0,1]$; typically $\alpha \approx 1/2$ (a so-called regime channel).
 End members $\alpha=0$ or $\alpha=1$ correspond a  subaerial channels of constant width (a slot canyon) or depth (no natural equivalent), respectively.
 %
-For a steady state R-channel, it is assumed that  $\Psi$ is constant (approximated by the gradient of the Shreve \citeyear{shreve1972} potential \footnote{Note that the R-channel model presented above (Eq.~\eqref{eq:dS_dt}--\eqref{eq:dh2wc}), calculates $\Psi$ from the time evolving $S_g$ (Eq.~\eqref{eq:dS_dt}) via the Darcy-Weisbach equation~\eqref{eq:dh} and thus no Shreve approximation is then needed.}) and that $S$ adjusts such that it is in steady state with $\Psi$ and $Q$.
-% Conversely, pipe flow has $S$ fixed and $\Psi$ adjusts to satisfy the given $Q$.
+For a steady state R-channel, it is assumed that  $\Psi$ is constant (approximated by the gradient of the Shreve \citeyear{shreve1972} potential) and that $S$ adjusts such that it is in steady state with $\Psi$ and $Q$.
+%Conversely, pipe flow has $S$ fixed and $\Psi$ adjusts to satisfy the given $Q$.
 %
-Pipe flow like conditions occur when a R-channel is subjected to rapid discharge variations such that the channel cannot adjust its size significantly on the time scale of the discharge variations.
+Pipe flow is approximated when a R-channel is subjected to rapid discharge variations such that the channel does not adjust its size significantly on the time scale of the discharge variations.
 In this case, it is assumed that the cross-sectional area $S$ is fixed, and $\Psi$ adjusts to the specified $Q$.
@@
 
-Table~\ref{tab:Qs2} shows how the proxy $w\tau$ as well as $Q_s$ of the MPM, EH and Bagnold sediment transport formulas scale with respect to $Q$, $\Psi$ or $S$ for the three different regimes and for different $\alpha$.
+Table~\ref{tab:Qs2} shows that the proxy $w_c\tau$ as well as the MPM, EH and Bagnold sediment transport formulas give roughly a similar scaling of $Q_s$ on hydraulic gradient $\Psi$ or $S$ and on water discharge $Q$.
 Remarkably, the Bagnold formula has a negative exponent for $f$ in all but the pipe-flow regime, which seems rather unexpected.
@@
 Albeit, the sediment transport response in pipe-flow for the Bagnold case is close to EH.
-Conversely, the sediment transport proxy $w_g\,\tau_{g}, w_f\,\tau_{f}$ used within this publication scales only as $Q^2$ for pipe-flow, whereas the others scale at least with $Q^3$.
+Conversely, the sediment transport proxy $t_{tg},t_{tf}$ used within this publication scales only as $Q^2$ for pipe-flow, whereas the others scale at least with $Q^3$.
 The exponent for the width scaling $\alpha$ only impacts relationship between sediment transport and water discharge in the EH relation in any meaningful way.
-However, for the value of $\alpha$ around $\frac{1}{2}$ (a value appropriate for most streams), the exponent on $Q$ for EH is only slightly above what the other transport relations give.
+However, for the value of $\alpha$ around $\frac{1}{2}$,  suits braided proglacial stream and indeed most  subaerial channels.
+Furthermore, when $\alpha=1$, the exponent on $Q$ is only slightly above the other values of $\alpha$.
 
 The sediment discharge in the steady state R-channel scales very similarly to the  subaerial channel case for all relations and virtually identically for the $\alpha=0.5$ case.
-However, note that the head gradients $\Psi$ are likely higher for comparable $Q$ in an ice-sheet marginal or alpine glacier setting than in a  subaerial channel, as is well described in \citeA{alley1997}, and thus transport rates may still be much higher in a steady state R-channel.
+However, note that the head gradients $\Psi$ are likely higher for comparable $Q$ in a ice-sheet marginal or alpine glacier setting than in a  subaerial channel, as is well described in \citeA{alley1997}, and thus transport rates may still be much higher in a steady state R-channel.
 
@@
 Table~\ref{tab:Qs2} shows that in such a situation, the sediment transport scales much more severely with discharge, with the exponent on $Q$ being likely between $3$ and $5$, compared to other two regimes when that exponent is at most $1.3$ (Table~\ref{tab:Qs2}).
-Thus fluctuations of discharge on short timescales (on the order of a day) have the potential to cause conditions with very high sediment transport capacities.
+Thus fluctuations of discharge on short timescales (on the order of a day) have the potential to cause very high sediment transport conditions.
 
@@
   \label{tab:eqs1}
+  \small
 \begin{tabular}{llllll}
   Regime & Fixed & Determined via D-W
-   & Additional relations & Shear stress & Stream p.\\
-& & (Equation~\ref{eq:DW})  &  &  \(\tau \propto\) & \(\Omega \propto\)\\
+  \makecell{Equation~\ref{eq:DW}} & Additional relations & Shear stress & Stream p.\\
+ &  &  &  & \(\tau \propto\) & \(\Omega \propto\)\\
 \hline
-  Subaerial & \(\Psi\) & \(H \propto f^{1/3}\, Q^{2/3-2\alpha/3} \, \Psi^{-1/3}\) & \makecell{\(w\,\propto Q^\alpha\) \\ \(S=wH\) \\ \(D_h\propto H\)} & \(f^{1/3} Q^{2/3-2\alpha/3}  \Psi^{2/3}\) & \(Q\, \Psi\)\\
-  R-channel & \(\Psi\) & \(S\, \propto f^{2/5}\, Q^{4/5} \, \Psi^{-2/5}\) & \(D_h\propto w \propto H \propto S^{1/2}\) & \(f^{1/5} Q^{2/5} \, \Psi^{4/5}\) & \(Q\, \Psi\)\\
-Pipe & \(S\) & \(\Psi \propto f \, Q^2\, S^{-5/2}\) & \(D_h\propto w \propto H \propto S^{1/2}\) & \(f Q^2 S^{-2}\) & \(f\, Q^3 S^{-5/2}\)\\
+  Subaerial & \(\Psi\) & \(H \propto f^{1/3}\, Q^{2/3-2\alpha/3} \, \Psi^{-1/3}\) & 
+\makecell{\(w_c\,\propto Q^\alpha\) \\ \(S=w_cH\) \\ \(D_h=H\)} & 
+\(f^{1/3} Q^{2/3-2\alpha/3}  \Psi^{2/3}\) & \(Q \Psi\)\\
+R-ch. & \(\Psi\) & \(S\, \propto f^{2/5}\, Q^{4/5} \, \Psi^{-2/5}\) & \(D_h\propto w_c \propto H \propto S^{1/2}\) & \(f^{1/5} Q^{2/5} \, \Psi^{4/5}\) & \(Q \Psi\)\\
+Pipe & \(S\) & \(\Psi \propto f \, Q^2\, S^{-5/2}\) & \(D_h\propto w_c \propto H \propto S^{1/2}\) & \(f Q^2 S^{-2}\) & \(f Q^3 S^{-5/2}\)\\
 \end{tabular}
@@
 \begin{table}
-  \caption{Sediment transport proxy ($w\tau$) and rates for the three considered different transport formulas: MPM \cite{meyer1948}, EH \cite{engelund1967}, and Bagnold \cite{bagnold1980}.
+  \caption{Sediment transport proxy ($w_c\tau$) and rates for the three considered different transport formulas: MPM \cite{meyer1948}, EH \cite{engelund1967}, and Bagnold \cite{bagnold1980}.
     }
@@
 \begin{tabular}{lllll}
- & Width \(\times \,\, \tau\) & MPM & EH & Bagnold\\
- & \(w\, \tau\) & \(Q_s \propto w\, \tau^{3/2}\) & \(Q_s \propto w\, \tau^{5/2}\) & \(Q_s \propto w^{-1/2}\, \Omega^{3/2} H^{-2/3}\)\\
+ & Width*\(\tau\) & MPM & EH & Bagnold\\
+ & \(w_c\, \tau\) & \(Q_s \propto w_c\, \tau^{3/2}\) & \(Q_s \propto w_c\, \tau^{5/2}\) & \(Q_s \propto w_c^{-1/2}\, \Omega^{3/2} H^{-2/3}\)\\
 \hline
 Subaerial  & \(f^{1/3}\, Q^{2/3+\alpha/3}\,  \Psi^{2/3}\) & \(f^{1/2}\, Q \, \Psi\) & \(f^{5/6}\, Q^{5/3 - 2\alpha/3} \, \Psi^{5/3}\) & \(f^{-2/9}\, Q^{19/18-\alpha/18} \, \Psi^{31/18}\)\\
-R-channel & \(f^{2/5}\, Q^{4/5} \, \Psi^{3/5}\) & \(f^{1/2}\, Q \, \Psi\) & \(f^{7/10}\, Q^{7/5}\, \Psi^{9/5}\) & \(f^{-7/30}\, Q^{31/30}\, \Psi^{26/15}\)\\
+R-ch. & \(f^{2/5}\, Q^{4/5} \, \Psi^{3/5}\) & \(f^{1/2}\, Q \, \Psi\) & \(f^{7/10}\, Q^{7/5}\, \Psi^{9/5}\) & \(f^{-7/30}\, Q^{31/30}\, \Psi^{26/15}\)\\
 Pipe & \(f \, Q^2 \, S^{-1}\) & \(f^{3/2}\, Q^3 \, S^{-5/2}\) & \(f^{5/2}\, Q^5\, S^{-9/2}\) & \(f^{3/2} \, Q^{9/2} \, S^{-14/3}\)\\
@@
 \begin{tabular}{lllll}
- & Width \(\times \,\, \tau\) & MPM & EH & Bagnold\\
- & \(w\, \tau\) & \(Q_s \propto w\, \tau^{3/2}\) & \(Q_s \propto w\, \tau^{5/2}\) & \(Q_s \propto w^{-1/2}\, \Omega^{3/2} H^{-2/3}\)\\
+ & Width*\(\tau\) & MPM & EH & Bagnold\\
+ & \(w_c\, \tau\) & \(Q_s \propto w_c\, \tau^{3/2}\) & \(Q_s \propto w_c\, \tau^{5/2}\) & \(Q_s \propto w_c^{-1/2}\, \Omega^{3/2} H^{-2/3}\)\\
 \hline
@@
 Subaerial (\(\alpha=1\)) & \(f^{0.7}\, Q^{1.0}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{0.7} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\[3pt]
-R-channel & \(f^{0.4}\, Q^{0.8} \, \Psi^{0.6}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.7}\, Q^{1.4}\, \Psi^{1.8}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\
+R-ch. & \(f^{0.4}\, Q^{0.8} \, \Psi^{0.6}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.7}\, Q^{1.4}\, \Psi^{1.8}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\
 Pipe & \(f \,\quad Q^{2\phantom{.0}} \, S^{-1}\) & \(f^{1.5}\, Q^3 \, S^{-2.5}\) & \(f^{2.5}\, Q^{5\phantom{.0}}\, S^{-4.5}\) & \(f^{1.5} \,\,\,\,\, Q^{4.5} \, S^{-4.7}\)\\

### 2023-11-14 Delaneyetal2023b.tex

--- 2023-10-09 Delaneyetal2023b.tex
@@
 
-\title{Natural variations in water discharge quantity have an inconsistent impact on sediment transport capacity in subglacial channels\mauro{\\Title does not work.
-I don't think ``inconsistent'' can be used like this.
-Maybe:}\\
-Subglacial and subaerial fluvial sediment transport respond differently to discharge variations
-}
+\title{Subglacial and subaerial fluvial sediment transport capacity respond differently to water discharge variations}
 
@@
 \correspondingauthor{Ian Delaney}{ianarburua.delaney@unil.ch}
+
+change subglacial hydraulics to subglacial model outputs??
 
@@
   Results show that high subglacial sediment transport rates can occur across a wide range of water discharges.
-%  Slowly evolving channel width accommodates some variability in subglacial channels. \mauro{not sure what that sentence should do}
+  Slowly evolving channel width accommodates some variability in subglacial channels. \mauro{not sure what that sentence should do}
   Despite differences in response to water discharge, sediment discharge measurements generally correlate better with subaerial hydraulic parameters.
@@
     \label{eq:tau}
-    \tau_t \propto v^2.
+    \tau_t \propto w\,v^2.
   \end{equation}
@@
 
-Across the range of channel shape parameters and friction factors  examined, variability in water velocity and shear stress remains higher in the subglacial system compared to the subaerial one (Figure~\ref{fig:multi_run}).
-In \alpine{}, the variability in width-integrated shear stress is comparable between the two systems, due to the variable water discharge.
+Across the range of channel shape parameters and friction factors  examined, variability in water velocity and shear stress remains higher in the subglacial system compared to the subaerial one.
+In \alpine{}, the variability in width-integrated shear stress is comparable between the two systems, due to the variable water discharge (Figure~\ref{fig:multi_run}).
+
 However, in \icesheet{}, with lower water discharge fluctuations, the variability in width-integrated shear stress remains higher in the subglacial system across the time aggregation periods.
@@
 %
-The stronger correlation between subglacial hydraulics and water discharge in \icesheet{} (Figure~\ref{fig:Qw_vari}a-f) could result from the channel being closer to equilibrium between variations in the channels' wetter area and water discharge with smaller diurnal variations in amplitude (Figures~\ref{fig:model_outs} and ~\ref{fig:Qw_vari}) \mauro{needs rewriting}.
+The stronger correlation between subglacial hydraulics and water discharge in \icesheet{} (Figure~\ref{fig:Qw_vari}\,a-f) could result from the channel being closer to equilibrium between variations in the channels' wetter area and water discharge with smaller diurnal variations in amplitude (Figures~\ref{fig:model_outs} and ~\ref{fig:Qw_vari}) \mauro{needs rewriting}.
 Conversely, the correlation between hydraulics and water discharge in \alpine is smaller due to the relatively larger discharge variations on short time scales.
-In both scenarios, hydraulics could \mauro{``could''?
-Does it or does it not?} trend toward equilibrium by averaging the discharge over $1-5$ days, when the variability in subglacial channels approaches that of subaerial ones (Figure~\ref{fig:multi_run}).
+In both scenarios, the variability in subglacial channels approaches that of subaerial ones  over $1-5$ days (Figure~\ref{fig:multi_run}), suggesting a parity between the two channel types might be approached over this period.
 
@@
 In turn, subglacial channels grow with increasing water discharge, effectively reducing sediment transport capacity with the stabilization of water discharge at its peak.
-Similar feedback occurs in subaerial systems \cite{phillips2016}, albeit at timescales longer than days.  \mauro{\citeA{gimbert2016}, which you cited here, says nothing of subaerial flow, or do they?}
+Similar feedback occurs in subaerial systems \cite{phillips2016}, albeit at timescales longer than days in the subglacial system.
 
-\subsection{Records of sediment discharge  from glaciers}
+\subsection{Effect of variable sediment transport capacity on sediment discharge  from glaciers}
 \mauro{This section header sounds like this will be about measurements only.}
@@
 The transition of subglacial to subaerial conditions typically occur still underneath glacier but close to the terminus \cite{perolo2018}, where closure rates of the subglacial channels are slow and open channel flow condition prevail \cite{rothlisberger1972}.
-Furthermore, if a transport-limited regime persists at the glacier margin, then sediment can be easily mobilized, and subaerial processes could be represented just a short distance downstream \mauro{Not sure what that sentence means}.
+Furthermore, if abundant sediment persists in the proglacial area, then it can be easily mobilized, and subaerial processes could be represented just a short distance downstream \mauro{Not sure what that sentence means}.
 
 \mauro{I perceived this paragraph pretty out of the blue.}
-The consistent relationship between sediment discharge capacity and open channel flow conditions \mauro{What?
-How do you compare those two things?} could be pronounced $\sim20$\,\unit{km} downstream of the Leverett site, where a strong correlation persists between sediment plume size and the river's water discharge into the Kangerlussuaq fjord \cite{mcgrath2010}.
+The consistent relationship between sediment discharge capacity in  open channel flow conditions could be pronounced $\sim20$\,\unit{km} downstream of the Leverett site, where a strong correlation persists between sediment plume size and the river's water discharge into the Kangerlussuaq fjord \cite{mcgrath2010}.
 In contrast, in marine-terminating glacier catchments a less consistent relationship may occur between water discharge or melt extent and sediment plume size \cite{tedstone2012}.
@@
 Sediment exhaustion occurring through increased variability may also explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \cite{overeem2017}.
-Note, however, the absence of lags between the observed sediment and water discharge \mauro{reference a figure here?} may suggest that sediment is mobilized close to the measurement stations, as opposed to distal areas of the glacier bed as would be expected in a supply-limited regime \cite<Figure~\ref{fig:model_outs}\, a and e; >{williams1989}.
+Note, however, the absence of lags between the observed sediment and water discharge (Figure~\ref{fig:model_outs}\, a and e) may suggest that sediment is mobilized close to the measurement stations, as opposed to distal areas of the glacier bed as would be expected in a supply-limited regime \cite{williams1989}.
 
-Regardless of the increased performance of water discharge compared to subglacial sediment transport capacity in the observations \mauro{does not parse}, outputs suggest that additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate \cite{jansson2005,ganti2016}.
+Regardless of the increased performance of water discharge compared to subglacial sediment transport capacity in representing the observations,  suggest that additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate \cite{jansson2005,ganti2016}.
 Sediment discharge records have been used to establish the relationship, or lack thereof, between climate forcing and glacial erosion \cite<e.g.>{koppes2009a,ganti2016,willenbring2016,mariotti2021}.
 The variable relationship between water discharge and subglacial sediment transport capacity presented here mostly applies to the daily- to weekly-time scales controlling the size of subglacial channels (Figure~\ref{fig:multi_run}).
-Yet, identifying climatic signals in sediment transport from the transport-limited component of glacier-sedimentary systems may require higher thresholds of water discharge compared to subaerial systems when water discharge varies substantially over short timescales \cite<Figure~\ref{fig:Qw_vari}; >{tofelde2021}.
+Yet, identifying climatic effects in sediment transport from the transport-limited component of glacier-sedimentary systems may require higher thresholds of water discharge compared to subaerial systems when water discharge varies substantially over short timescales \cite<Figure~\ref{fig:Qw_vari}; >{tofelde2021}.
 \mauro{I don't understand the last two sentences.}
@@
 
+\ian{move this shit up or delete}
+\subsection{Comparison of Alpine and Arctic hydrological forcing}
+
 \section{Conclusions}
@@
 Subaerial channels alter both their channel width and water velocity immediately in response to changing water discharge.
-In contrast, pressurized subglacial channels largely accommodate rapidly changing water discharge by altering water velocity but not their size; the latter, in contrast, only responds over the course of several days to changes in discharge.
+In contrast, pressurized subglacial channels largely accommodate rapidly changing water discharge by altering water velocity.
+Their size, in contrast, responds over the course of several days to changes in discharge.
 Thus the response of sediment transport capacity is much more sensitive to changes in water discharge in subglacial channels compared to subaerial ones.
-Despite these differences, parameterizations of subaerial hydraulics in two glacierized catchments generally represent observed sediment discharge variations better than subglacial hydraulics \mauro{This sentence does not quite parse.
+%Despite these differences, parameterizations of subaerial hydraulics in two glacierized catchments generally represent observed sediment discharge variations better than subglacial hydraulics \mauro{This sentence does not quite parse.
 Why ``despite''?
@@
 \begin{enumerate}
-\item the inconsistent \mauro{``inconsistent'' does not work for me here.
-What do you mean?
-More non-linear?} relationship between water discharge and sediment transport capacity underneath glaciers creates larger variability in sediment transport capacity compared to river systems;
-\item the timing of peak sediment transport capacity below glaciers generally coincides with large rates of increase in water discharge, as opposed to the maximum water discharge;
+\item the changing relationship between water discharge and sediment transport capacity underneath glaciers creates larger variability in sediment transport capacity compared to river systems;
+\item the timing of peak sediment transport capacity below glaciers generally coincides with large rates of increase in water discharge, as opposed to the maximum water discharge, as would occur in pipe flow of a fixed size;
 \item the timing of peak sediment transport capacity in subaerial channels coincides with the maximum water discharge;
-\item a consistent \mauro{and what does ``consistent'' mean?} relationship between water and sediment discharge may represent subglacial sediment access or open channel hydraulics near the glacier margin, instead of subglacial sediment transport capacity.
+\item a steady relationship between water and sediment discharge may represent subglacial sediment access or open channel hydraulics near the glacier margin, instead of subglacial sediment transport capacity.
 \end{enumerate}
@@
 R-channels which do not have time to adjust their size to discharge conditions).
-Here we show how sediment transport capacity scales with respect to a given water discharge and hydraulic gradient for three different transport formulas: MPM \cite{meyer1948}, EH \cite{engelund1967}, and Bagnold \cite{bagnold1980}; additionally width-integrated shear stress is assessed, as used as proxy for sediment transport throughout the manuscript.
+Here we show how sediment transport capacity scales with respect to a given water discharge and hydraulic gradient for three different transport formulas: Meyer-Peter M\"uller  \cite<MPM;>{meyer1948}, Engelund and Hansen \cite<EH;>{engelund1967}, and Bagnold \cite{bagnold1980}; additionally width-integrated shear stress is assessed, as used as proxy for sediment transport throughout the manuscript.
 

### 2024-03-07 Delaneyetal2024.tex

--- 2023-11-14 Delaneyetal2023b.tex
@@
 
-change subglacial hydraulics to subglacial model outputs??
+%\ian{only glaciers with subglacial channels... }
 
@@
 \item Greater variability in water velocity causes sediment transport capacity to vary more in subglacial channels than in subaerial channels.
-\item Water discharge may represent processes including proglacial sediment mobilization and sediment access driving variations in sediment discharge, as opposed to subglacial sediment transport capacity in glacierized catchments.
-\mauro{I don't understand this sentence.}
+\item Results suggest that significant sediment transport capacity can persist across a range of water discharge conditions, and timing of peak events occurs under different hydrological conditions.
 \end{keypoints}
 
-\begin{abstract}  Sediment transport capacity in subglacial and subaerial channels depends on  water velocity and channel width, which together control the total shear stress exerted on the channel bottom.
-  In subaerial channels water discharge variations are accommodated largely by changes in flow depth and width.
+\begin{abstract}
+  Sediment transport capacity in both subaerial and subglacial channels depends on the shear stress exerted across the channel bottom, which vary with the  water velocity and channel width.
+  In subaerial channels, water discharge variations are accommodated largely by changes in flow depth and width.
   While in subglacial channels, water discharge changes primarily affect water velocity due to slow-evolving channel geometry.
-  Parameterizations of these subglacial and subaerial hydraulics are applied to hydrographs from an Alpine glacier and the Greenland Ice Sheet revealing greater variability in the inferred sediment transport capacity in subglacial than in subaerial channels.
-  Results show that high subglacial sediment transport rates can occur across a wide range of water discharges.
-  Slowly evolving channel width accommodates some variability in subglacial channels. \mauro{not sure what that sentence should do}
-  Despite differences in response to water discharge, sediment discharge measurements generally correlate better with subaerial hydraulic parameters.
-  This implies that other factors like sediment access or proglacial processes might be at play that affect sediment transport capacity differently in subglacial and subaerial channels.
+  Hydrographs from an Alpine glacier and the Greenland Ice Sheet are applied to models of subglacial and subaerial hydraulics, revealing greater variability in the inferred sediment transport capacity in subglacial than in subaerial channels.
+  Results also show that high subglacial sediment transport rates can occur across a wide range of water discharges.
+  However, slowly evolving channel width accommodates some variability in subglacial channels, compared to subaerial channels.
+  These findings are discussed in the context of understanding sediment discharge from glaciers with different hydrological forcings and assumptions of channel type.
 \end{abstract}
+
+\section*{Plain Language Summary} % 200 words
+Sediment export from glaciers impacts downstream ecosystems and communities and it has been documented to change in response to climate warming and glacier retreat.
+In these catchments, sediment export depends in part on the sediment transport capacity of water flowing below the glacier, or the amount of sediment that could be transported given the abundance of sediment.
+This capacity largely depends on the width of the channel and the water's velocity. 
+Unlike subaerial channels, like rivers, where the size of the river changes rapidly with respect to water discharge, however, subglacial channels are pressurized by the ice above.
+This pressurized flow means that channel size evolves slowly compared to the variations in water discharge, therefore variations in water discharge are largely accommodated by variations in velocity.
+Models of these processes show that sediment transport capacity is highly variable in subglacial channels, and a wide range of water discharges can result in comparable sediment transport capacity.
+In subaerial channels, by contrast, sediment transport capacity is less variable and depends on a single water discharge quantity.
+The channel's different response to water discharge may impact the interpretation of sediment transport records and responds to the type of hydrological forcing applied.
 
@@
 
-Changes in glacier dynamics, geomorphology, and hydrology  have prompted numerous  recent studies of  sediment transport processes in cold regions \cite<e.g.>{zhang2022}.
+Changes in glacier dynamics and hydrology  have prompted numerous  recent studies of  sediment transport processes in cold regions \cite<e.g.>{zhang2022}.
 Increases in sediment transport  have been observed in Greenland \cite{bendixen2017}, the European Alps \cite{costa2017}, the Himalayas \cite{li2021}, and the Andes \cite{vergara2022}.
@@
     \label{eq:tau}
-    \tau_t \propto w\,v^2.
+    \tau \propto v^2.
   \end{equation}
 \end{linenomath*}
-\mauro{I removed the width from here as I found it odd to introduce that here.}
 %
-Following mass conservation, the velocity of the water flowing through a  channel is
+Following mass conservation, the mean velocity of the water flowing through a  channel is
 \begin{linenomath*}
@@
 \end{linenomath*}
-where $Q$ is water discharge,  and $S$ is the channel's wetted area. \mauro{I find it odd to have formulas in the Intro}
+where $Q$ is water discharge,  and $S$ is the channel's wetted area.
 
@@
 The change in water depth results in a commensurate increase in water velocity.
-%Thus,  fulfilling Equation~\ref {eq:v}, the higher discharge is accommodated by both an increase in wetted area and flow velocity. \mauro{not needed}
 As a result, the shear stress $\tau$ increases according to Equation~\ref{eq:tau}.
@@
 Therefore, on short time scales, subglacial water flow behaves more like pipe flow and changes in water discharge $Q$ are mainly accommodated with increased or decreased water velocity $v$ \cite{alley1997} because the size of the channel $S$ changes on a much longer timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
-%It follows that an increase in water discharge $Q$ will cause the shear stress across the channel $\tau_t$ to respond with water velocity $v$, as opposed to subaerial channels where channel width $w$ evolves as well.
 
@@
       The subglacial channel widths $w_g$ remain unchanged, while the subaerial channel widths $w_f$ evolve. Shear stress $\tau$ is responsible for the mobilization of sediment.
-      %Note that in the subaerial channel parameterization, channels have rectangular shapes with their width much larger than their depth (Section~\ref{sect:fluv}).
-\mauro{Needs renaming $w_{cg/f}$ to $w_{g/f}$.
-Then, I find the downward arrow confusing.
-Maybe just leave away?
-Also, I would write $\tau_f$ and $\tau_g$, it's ok to not have the width integrated tau here (and maybe we remove $\tau_{tf}$ from the paper anyway...}
     }
@@
   \end{figure}
-% \ian{Come back to Daniel's comments}
 
@@
 Yet, little work has examined the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial channels \cite{alley1997}.
-%Furthermore, thorough observations of water velocity and channel width in subglacial channels are limited and, to these authors' knowledge, not contrasted with subaerial conditions at the glacier margins.
-%\mauro{self-cite here?}
 
 This manuscript evaluates the relationship between sediment transport capacity and water discharge in subglacial channels and compares it to subaerial ones.
-Due to the absence of continuous measurements of subglacial and subaerial water velocity and channel morphology, numerical parameterizations of subglacial and subaerial hydraulics are utilized.
-\mauro{Aren't those things always parameterised, irrespective of whether there are measurements or not?}
-Parameterizations are applied to hydrological records from an Alpine glacier in Switzerland (Fieschergletscher) and  a land-terminating glacier in Greenland (Leverett Glacier).
-Our parameterizations' lumped nature isolates the disparity between water discharge and sediment transport capacity in subglacial systems, independent of the upstream drainage network and sediment access.
+Models of subglacial and subaerial hydraulics are used to quantify subglacial and subaerial water velocity and channel morphology.
+The models are applied to hydrological records from an Alpine glacier in Switzerland (Fieschergletscher) and  a land-terminating glacier in Greenland (Leverett Glacier).
+The  lumped nature of the models isolates the disparity between water discharge and sediment transport capacity in subglacial systems, independent of the upstream drainage network and sediment access.
 Outputs demonstrate differences in the relationships between  water discharge, channel geometry, and water velocity in the subglacial and subaerial channels.
-Furthermore, outputs are compared to sediment discharge data from these two catchments.
-The manuscript then uses the demonstrated differences between the two channel types to discuss the implications for interpreting the processes driving variability in records of sediment transport from glacierized catchments.
-\mauro{As stated in the general comments.  I'd rather use ``model'' than ``parameterization'' as the latter is, for me, something which is used in a model and not a model itself.}
+The manuscript then uses the demonstrated differences between the two channel types to discuss the implications for interpreting the processes driving variability in records of sediment transport, and their relationship with hydrology from glacierized catchments.
 
@@
 
-Water and suspended sediment data are leveraged from Alpine and ice sheet settings.
+Water discharge data is leveraged from Alpine and ice sheet settings collected downstream of the glacier, assuming that no water storage occurred in the proglacial area.
 The Alpine site (\alpine) is  Fieschergletscher in the Swiss Alps ($46^\circ\,29'\,07''$ N, $8^\circ\,08'\,3''$ E).
-Water discharge and suspended sediment concentration were collected here at a $1$\unit{min} interval continuously over the period 2014--2021.
+Water discharge was collected here at a $1$\unit{min} interval continuously over the period 2014--2021.
 See \citeA{felix2022} and \citeA{felix2021} for more details.
@@
 The Leverett Glacier in Greenland (\icesheet) serves as the ice sheet setting.
-Water discharge and suspended sediment concentration data were collected roughly $2$\unit{km} downstream from the terminus between June and August/September 2009 to 2012 ($67^\circ\,03'\,5''$ N, $50^\circ\,12'\,59''$ W).
-Data is used from \citeA{tedstone2017} and \citeA{hawkings2014} at a $5$\,\unit{min} time interval from May 28, 2012 to August 8, 2012.
-Two gaps (14.07.2012 to 15.07.2012 and 16.07.2012 to 17.07.2021) in suspended sediment data were interpolated using a linear spline.
+Water discharge was collected roughly $2$\unit{km} downstream from the terminus between June and August/September 2009 to 2012 ($67^\circ\,03'\,5''$ N, $50^\circ\,12'\,59''$ W).
+Data is used in \citeA{tedstone2017} and \citeA{hawkings2014} at a $5$\,\unit{min} time interval from May 28, 2012 to August 8, 2012.
 
@@
 \label{sect:meth}
-The parameterizations below (Sections ~\ref{sect:sub_mode}~and~\ref{sect:fluv}) represent relationships amongst water discharge, water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
-Both parameterizations calculate   water velocity, shear stress, and width-integrated shear stress, upon which sediment transport depends \cite<Figure \ref{fig:cartoon}; >{shields1936}.
+The models described below (Sections ~\ref{sect:sub_mode}~and~\ref{sect:fluv}) represent relationships amongst water discharge, water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
+Both models calculate   water velocity, shear stress, and width-integrated shear stress, upon which sediment transport depends \cite<Figure \ref{fig:cartoon}; >{shields1936}.
 The evaluation of these variables omits the selection of a sediment transport relationship and a grain-size parameter \cite<e.g.>{shields1936,meyer1948} needed to calculate sediment transport capacity from shear stress.
 
-\subsection{Subglacial channel  parameterization}
+\subsection{Subglacial channel  model}
 \label{sect:sub_mode}
 
-To evaluate the shear stress of water flowing across sediments underneath a glacier, the subglacial channel parameterization takes into account the channel geometry and the velocity of the flowing water.
+To evaluate the shear stress of water flowing across sediments underneath a glacier, the subglacial channel model accounts for the channel geometry and the velocity of the flowing water.
 To do this, a lumped hydraulics model \cite{clarke1996} modified from the one presented by \citeA{werder2010b} is used.
@@
 The formulation here does not consider the englacial storage of water.
-Thus, it is assumed that the changing hydraulic head at the top of the glacier is negligible in the channel size evolution \mauro{I don't understand this sentence}.
 The evolution of subglacial channel size $S_g$ is given as
@@
 \end{linenomath*}
-\noindent where $f_r$ is a friction factor, $D_h$ is the hydraulic diameter, $l$ is the length over which head change occurs, and $v_g=Q/S_g$ is the water flow speed.
+\noindent where $f_r$ is a friction factor, $D_h$ is the hydraulic diameter, $l$ is the length over which head change occurs, and $v_g=\frac{Q}{S_g}$ is the water flow speed.
 %
@@
 \begin{equation}
-  \label{eq:tau}
+  \label{eq:tau_g}
   \tau_g=\frac{1}{8}\,f_r\,\rho_w\,v_g^2,
@@
 %
-The width-integrated shear stress is represented as
-\begin{equation}
-  \label{eq:tautg}
-  \tau_{tg}=w_g\,\tau_g,
-\end{equation}
-where  channel floor width $w_g$ is
+The width of the channel floor $w_g$ is represented as
 \begin{equation}
@@
 \end{equation}
-\mauro{If we get rid of $\tau_{tf}$, above needs a bit of massaging.  I'd then move the $w_g$ eq to just after the $S_g$ eq.}
+%
+This value is used to establish the integrated shear stress across the channel, $w_g\tau_g$.
 
-\subsection{Subaerial channel  parameterization}
+\subsection{Subaerial channel  model}
 \label{sect:fluv}
@@
     \label{eq:DW_tau}
-    \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q}{w_f} \Big)^{\frac{2}{3}} \,\nabla z_c^{\frac{2}{3}},
+    \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_f^{\frac{1}{3}}}{2}\, \Big(\frac{Q}{w_f} \Big)^{\frac{2}{3}} \,\nabla z_c^{\frac{2}{3}},
   \end{equation}
 \end{linenomath*}
-where $\nabla z_c$ is the channel slope, and $f_p$ is the friction factor for subaerial channels \cite{tucker1997}.
+where $\nabla z_c$ is the channel slope, and $f_f$ is the friction factor for subaerial channels \cite{tucker1997}.
 Channel width $w_f$ is
@@
 where $k$ is a constant and $\alpha=\frac{1}{2}$ is a commonly chosen exponent \cite{leopold1953}.
-Water velocity, $v_f$, is given by rearranging Equation \ref{eq:tau} as
+Water velocity, $v_f$, is given by rearranging Equation \ref{eq:tau_g} as
 \begin{equation}
   \label{eq:vf}
-  v_f = \sqrt{\frac{8\,\tau_f}{f_p\,\rho_w}}.
+  v_f = \sqrt{\frac{8\,\tau_f}{f_f\,\rho_w}}.
 \end{equation}
 %
-As above, the width-integrated shear stress is
-\begin{equation}
-    \label{eq:tautf}
-    \tau_{tf}=w_f\,\tau_f.
-  \end{equation}
+As above, the width-integrated shear stress is $w_f\tau_f$.
 
-  Note that the subaerial channel model \mauro{or ``parameterization''} is a purely algebraic model, whereas the subglacial model comprises a differential equation for the evolution of $S_g$ and thus will have a history dependence on the discharge.
+Note that this subaerial channel model is purely algebraic, whereas the subglacial model comprises a differential equation for the evolution of $S_g$.
+Thus the models have a history dependence on the discharge $Q_w$.
 
@@
 
-\mauro{I would call the ``cases'' ``scenarios'' instead.
-The word ``scenario'' does not appear in the manuscript so far, so it would be unambiguous.
-Also, ``cases'' has the disadvantage that ``In both cases,'' is a saying, and thus it is confusing.
-I implemented this change.}
-
-The parameterizations above are applied to hydrological records from the Fieschergletscher (scenario \alpine) and the Leverett Glacier (scenario \icesheet).
-The subglacial and subaerial parameterizations represent the subglacial flow of water exiting a glacier and transitioning to subaerial flow.
-Outputs of the parameterizations represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
+The models above are applied to hydrological records from the Fieschergletscher (scenario \alpine) and the Leverett Glacier (scenario \icesheet).
+The subglacial and subaerial models represent the subglacial flow of water exiting a glacier and transitioning to subaerial flow.
+Outputs of the models represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
 To generalize these scenarios, \alpine{}  is exemplified by relatively thin ice thickness ($h_{ice}$: $225$\,\unit{m}), low water discharge ($\sim\,10$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
@@
 
-For each glacier, the parameterizations are first applied to a reference test case for each glacier that assumes  a subglacial channel with $\beta=\frac{\pi}{2}$, i.e. a segment of a quarter-circle.
-Friction factors $f_r$ and $f_p$ are tuned so that the parameterizations reproduce reasonable water velocities for both \alpine and \icesheet{}\cite<$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$>{werder2010b,chandler2013}.
-\mauro{Those $f$ reported in the table are super large!  Maybe an issue down the line...}
-The subaerial channels of both scenarios have a bed slope of $\nabla z_c = 0.05$ \mauro{Why do you report a value here but most are hidden in the table?  Maybe just ref the table?}.
-% Water velocity, shear stress, and width-integrated shear stress in the subglacial and subaerial channels from this formulation are presented below.
+For each glacier, the models are first applied to a reference test case for each glacier that assumes  a subglacial channel with $\beta=\frac{\pi}{6}$, i.e.
+a segment of a quarter-circle, (Table~\ref{table:vpm}).
+Friction factors $f_r$ and $f_f$ are tuned so that reasonable water velocities occur for both \alpine and \icesheet{}\cite<$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$>{werder2010b,chandler2013}.
+The subaerial channels of both scenarios have a bed slope of $\nabla z_c = 0.05$ (Table~\ref{table:vpm}).
 
-Second, to further characterize the variability in sediment discharge capacity in subglacial compared to subaerial channels, $100$  parameter combinations are applied to a range of different channel geometry factors ($\beta$ and $k$) and friction factors ($f_r$ and $f_p$).
-\mauro{So how many runs are actually run?
-Not clear.}
-Runs with parameter combinations are culled if their mean subglacial water velocity over the season lies outwith $0.75$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $2$\,\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity lies outwith $0.3$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with measurements \cite<e.g.>{werder2010b,magnusson2012,chandler2013} \mauro{Awkward sentence.
-(also note that ``cull'' means to remove, thus I changed ``between'' to ``outwith''.}.
-The variability of output is represented by the  standard deviation of that output over the course of the season \mauro{This sentence needs context.  What is variability used for?}.
+A second experiment aims to further characterize the variability in sediment discharge capacity in subglacial compared to subaerial channels across a range channel shapes, friction factors and water discharge variability.
+To accomplish this, the water discharge records are aggregated across a select time period in the past ranging from $15$\,\unit{min} up to $15$\,\unit{days}, and the model is run with random  parameter combinations of channel geometry factors ($\beta$ and $k$) and friction factors ($f_r$ and $f_f$).
+Runs with parameter combinations are accepted if their mean subglacial water velocity over the season lies between $0.75$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $2$\,\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity lies between $0.3$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with measurements \cite<e.g.>{werder2010b,magnusson2012,chandler2013}.
+The routine runs until $100$ different parameter combinations are accepted from each water discharge aggregation period.
+The variability of an output, a key difference between the subaerial and subglacial systems, is represented by  its  standard deviation over the course of the season.
 
-Lastly, outputs of $w\tau$ are compared with the sediment discharge records of the glaciers using Spearman rank correlation, dependent on the ordering of values.
+The results below present water velocity ($v_g$, $v_f$), shear stress ($\tau_g$, $\tau_f$), and width-integrated shear stress ($w_g\tau_g$, $w_f\tau_f$) from the subglacial and subaerial models.
+For brevity in the text these values together are referred to as ``model outputs''.
+Lastly, outputs  are compared with the water discharge from glaciers using Spearman rank correlation, dependent on the ordering of values.
 Rank correlation reduces the impact of the non-linear response of sediment transport capacity to the hydrological forcing, thus making our interpretation independent of the exact details of the sediment transport formula.
-Outputs and observations, which have a $15$\,\unit{min} resolution, are aggregated across time intervals of up to $15$\,\unit{days} by averaging values over a symmetric window of corresponding size.
 
@@
   \includegraphics[width=0.9\linewidth]{Fig2.pdf}
-  \caption{Parameterization outputs from the hydrographs n panels a and e for \alpine scenarios (a-d) and \icesheet{} (e-h).
+  \caption{Model outputs from the hydrographs n panels a and e for \alpine scenarios (a-d) and \icesheet{} (e-h).
     Gray (black) lines in a and e represent the subglacial channel size (water discharge).
@@
     Data are shown at $15$\unit{min} intervals, i.e. not aggregated.
-    Arrows denote where variables peaking in the subglacial channel prior to the subaerial one \mauro{wouldn't this need more zoomed-in plots?}.
+    Arrows denote where variables peaking in the subglacial channel prior to the subaerial one.
     Insets in e-h show the  peak melt event \icesheet{} denoted by the shaded area in panel e.
     % Note that the x and y axes are different for \alpine{} and \icesheet{}.
-        \mauro{Wrong units! Shear stress is Pa\,m$^{-2}$ and width integrated Pa\,m$^{-1}$}
   }
@@
 
-\mauro{Good para}
-Following the parameterizations defined in Sections~\ref{sect:sub_mode}~and~\ref{sect:fluv}, the water velocity, shear stress, and width-integrated shear stress exhibit different seasonal evolutions and peaks.
+Following the models defined in Sections~\ref{sect:sub_mode}~and~\ref{sect:fluv}, the water velocity, shear stress, and width-integrated shear stress exhibit different seasonal evolutions and peaks.
 Water velocity and shear stress represent the potential for sediment mobilization, and width-integrated shear stress represents a proxy for the total sediment transport capacity across the channel bed (Figure~\ref{fig:model_outs}).
-
-\mauro{In principle good paragraph, the meat.  But suffers from sloppy writing.}
-In the subglacial channel, peaks in $w_g\tau_g$ generally occur when water discharge increases at the fastest rate (Figure~\ref{fig:model_outs}d,h).
-In contrast, in the subaerial channel,   $w_g\tau_g$  peaks when water discharge is highest.
-The timing of the greatest rate of change in water discharge occurs before the peaks in water discharge \mauro{what?}.
-Furthermore, the subglacial outputs \mauro{What are ``outputs''?}  are generally larger than the subaerial values, thus more efficient at transporting sediment.
-Here, sediment could be deposited in the proglacial area as the water depressurizes, then remobilized when sediment transport capacity peaks in the subaerial system \mauro{not sure this tangent is worth taking in this paragraph}.
+Furthermore, the subglacial hydraulic outputs are generally larger than their subaerial equivalents , thus greater sediment transport capacity would be expected in subglacial channels compared to subaerial ones.
 
@@
     Relationship between water discharge and velocity, shear stress, and width-integrated shear stress for \alpine{} (a-c) and \icesheet (d-f).
-    These outputs' relationship with measured sediment discharge is given for \alpine in (g-i) and \icesheet (j-l).
+    %These outputs' relationship with measured sediment discharge is given for \alpine in (g-i) and \icesheet (j-l).
     Variables on the x-axis have been normalized to mean values.
@@
     Plots are shown with data and outputs at $15$\unit{min} intervals.
-    \mauro{Pixely, needs more resolution, at least 2x. The two dots in the legend are not visible.  I would add ``Measured'' to the two y-labels.}
+    %\mauro{Pixely, needs more resolution, at least 2x. The two dots in the legend are not visible.  I would add ``Measured'' to the two y-labels.}
   }
@@
 
-In the subaerial system there is no history dependence and thus each water discharge value in the subaerial channel results in a unique water velocity, shear stress, and width-integrated shear stress, creating a perfect rank correlation between the variables (Figure~\ref{fig:Qw_vari} a--f).
+In examining the relationship between water discharge and the model outputs, different relationship emerges between the subaerial and subglacial cases.
+Because there is no history dependence in  subaerial channels, each water discharge value in the subaerial channel results in a unique water velocity, shear stress, and width-integrated shear stress (Section~\ref{sect:sub_mode}).
+This characteristic results in a perfect rank correlation between the variables (Figure~\ref{fig:Qw_vari} a--f).
+In the subglacial channel, however, the hysteresis in channel size evolution (Equation~\ref{eq:dS_dt}) means that subglacial model outputs respond differently to water discharge than subaerial ones.
+Here, the persistence of high water velocity, shear stress, and width-integrated shear stress across a range of water discharges suggests that substantial sediment transport may occur under a wide range of hydrological forcings (Figure~\ref{fig:Qw_vari} a--f).
 
-\mauro{it feels this sentence should be in the same para as the previous one.  But not sure what it tries to say.}
-In the subglacial channel, however, the persistence of high water velocity, shear stress, and width-integrated shear stress across a range of discharges suggests that substantial sediment transport may occur across \mauro{across what?} (Figure~\ref{fig:Qw_vari} a--f).
-
-For instance, very high  water velocity values and shear stresses can occur across minimal water discharge at $2.5$\,\unit{m}$^3$\,\unit{s}$^{-1}$ to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
-In \icesheet, mean values \mauro{what ``mean''?} of water velocity can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
+For instance in \alpine, very high water velocity values and shear stresses can occur at a minimal water discharge  ($\sim\,4$\,\unit{m}$^3$\,\unit{s}$^{-1}$) to the maximum water discharge at over ($17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, Figure~\ref{fig:Qw_vari}\,a).
+In \icesheet, mean values water velocity over the season can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
 When considering the subglacial channel's evolving width, width-integrated shear stress  across the channel generally increases with water discharge, with improved rank correlation compared to water velocity or shear stress (\alpine, Figure~\ref{fig:Qw_vari} a--f).
 Yet even the width-integrated shear stress  can vary substantially, with the highest values occurring at water discharge values ranging from roughly $11$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ to over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
-The variability in width-integrated shear stress is less pronounced in the \icesheet scenario, with reduced variability in water discharge (Figure~\ref{fig:Qw_vari} c, f).
+The variability in width-integrated shear stress is less pronounced in the \icesheet scenario, where the hydrograph has less diurnal variability in water discharge (Figure~\ref{fig:Qw_vari} c, f).
+
+The different response of subaerial and subglacial model outputs to water discharge impacts the timing of peak events.
+In the subglacial channel, peaks in $w_g\tau_g$ generally occur when water discharge increases at the fastest rate, but before the maximum quantity of water discharge (Figure~\ref{fig:model_outs}\, d and h).
+By contrast, in the subaerial channel,   $w_f\tau_f$  peaks when water discharge is highest.
 
@@
       Time intervals shown with markers denote $15$\,\unit{min}, $1$\,\unit{hr}, $12$\,\unit{hr}, $1$\,\unit{d}, $5$\,\unit{d}, $10$\,\unit{d}, and $15$\,\unit{d} aggregation periods.
-      Text in columns shows rank correlation between the variable and sediment discharge for the corresponding time period for the subglacial (left column) and subaerial (right column) channels.
-      \mauro{Again units are wrong as in Fig 2.
-        I would add \alpine and \icesheet labels to the plots; and for sure that needs to be stated in the caption.
-        It's a bit confusing to have the rank-correlations here as well as they are not on the variability but just on the variable themselves... not sure what to do though. } }
+       }
     \label{fig:multi_run}
@@
 
+The second numerical experiment aims to evaluate the role of channel shape and friction factors on the findings above.
 Across the range of channel shape parameters and friction factors  examined, variability in water velocity and shear stress remains higher in the subglacial system compared to the subaerial one.
-In \alpine{}, the variability in width-integrated shear stress is comparable between the two systems, due to the variable water discharge (Figure~\ref{fig:multi_run}).
+In \alpine{}, the variability in width-integrated shear stress is comparable between the subaerial and subglacial systems, due to the highly variable water discharge, particularly over diurnal time periods (Figure~\ref{fig:multi_run}).
+However, in \icesheet{}, with lower water discharge fluctuations, the variability in width-integrated shear stress remains higher in the subglacial system across the time aggregation periods.
+Together these findings suggest that sediment transport capacity variations over a melt season can be comparable in subaerial and subglacial systems if substantial variability persists in the hydrological forcing.
 
-However, in \icesheet{}, with lower water discharge fluctuations, the variability in width-integrated shear stress remains higher in the subglacial system across the time aggregation periods.
-
-\mauro{this paragraph needs more motivation: why are we looking at this?}
-It is prescient \mauro{I don't think that works ``to have foresight''} to evaluate the time scales over which the subglacial channels may respond to changing water discharge to understand the influence of variability in sediment transport capacity.
-In both scenarios, the variability in velocity, shear stress, and width-integrated shear stress decreases with aggregations longer than approximately 1--5 days (Figure~\ref{fig:multi_run}).
+Evaluating the different variability in sediment transport characteristics across timescales can help inform the sediment transport processes omitted from daily or week values of water discharge.
+Results from  both \alpine and \icesheet scenarios show that the variability in velocity, shear stress, and width-integrated shear stress decreases with aggregations longer than approximately 1--5 days (Figure~\ref{fig:multi_run}).
 Yet, substantial differences in variability persist in periods of up to $15$ days.
 Aggregations periods longer than this might be conflated with seasonal variations in water discharge and are thus omitted.
-Note that lags and time of peak sediment transport conditions still occur over these periods (Figure~\ref{fig:model_outs}), while the variability in sediment transport parameters decreases with larger time aggregation (Figure~\ref{fig:multi_run}) \mauro{not clear sentence}.
-
-\subsection{Relationship between parameterization  outputs and observed sediment discharge}
-\mauro{Not sure a separate section is needed for this one paragraph.}
-
-When comparing the outputs with suspended sediment discharge measurements, the subaerial parameterizations generally correlate better with sediment discharge compared to subglacial outputs across the tested parameters (Figures~\ref{fig:Qw_vari} g--l and \ref{fig:multi_run}).
-The rank correlation amongst the subaerial parameterization's outputs remains fairly constant despite the variable impact of channel shape and consistent scaling with water discharge (Figure \ref{fig:Qw_vari} and Section~\ref{sect:fluv}).
-In subglacial channels, velocity and shear stress perform substantially worse in terms of rank correlation with sediment discharge compared to width-integrated shear stress (Figures~ \ref{fig:multi_run}).
-In both \alpine{} and \icesheet{}, the subglacial rank correlation approaches that of subaerial channels over longer time aggregation periods (Figures~ \ref{fig:multi_run} and \ref{fig:model_outs_1hr} to \ref{fig:model_outs_15day}).
+Note that lags and time of peak sediment transport conditions still occur over these periods, and these characteristics are thus note considered in aggregating outputs and water discharge over longer time periods (Figure~\ref{fig:model_outs}).
 
@@
 The exponent greater than $1$ magnifies sediment discharge variability beyond the variable sediment transport parameters described above (Figure~\ref{fig:multi_run}; Table~\ref{tab:Qs}).
-Note that sediment transport relationships describe sediment flux per unit width, so that the integrated capacity across channel occurs after application of the exponent, unlike the width-integrated shear stress evaluation here \cite<Equations\,\ref{eq:tautg} and \ref{eq:tautf}; c.f.>{tucker1997} \mauro{sentence does not parse; I think you should state limitations of width integrated shear stress and then be done with it.}.
-%Note also that both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress is homogeneous across the channel bed, despite the limitations of this assumption \cite<e.g.>{yager2018}.
 
-\mauro{This paragraph needs rewriting.  I guess I should do that.}
-The rapid increases and decreases in subglacial sediment transport capacity compounds the variability in sediment transport already present in subaerial systems \cite<Figure~\ref{fig:Qw_vari}; >{williams1989,jerolmack2010}.
+In R-channel models, viable water pressure regimes can be achieved by changing the shape and friction factor of the channels, enabling assumptions of semi-circular channels in most applications \cite{werder2010}.
+However, the role of channel width in the findings here suggests that more thorough parameterization of channel shape and channel width is needed in assessing subglacial sediment transport capacity (Section\ref{sect:sub_mode}, \ref{sect:scaling}).
+Note as well that sediment transport relationships describe sediment flux per unit width, so that the integrated sediment transport capacity across channel occurs after application of the exponent, unlike the width-integrated shear stress evaluation here \cite<c.f.>{tucker1997}.
+
 Sediment transport capacity in subglacial channels responds more rapidly to changes in water discharge compared to their subaerial counter parts \cite{alley1997}.
@@
 
-\mauro{You often use ``subglacial hydraulics''.
+The consistent relationship between sediment discharge capacity in  open channel flow conditions could be pronounced $\sim20$\,\unit{km} downstream of the Leverett site, where a strong correlation persists between sediment plume size and the river's water discharge into the Kangerlussuaq fjord \cite{mcgrath2010}.
+In contrast, in marine-terminating glacier catchments a less consistent relationship may occur between water discharge or melt extent and sediment plume size \cite{tedstone2012}.
+Here, ocean water pressurizes subglacial channels at the ice front \cite<e.g.>{how2017}, so the minimal correlation could result from the inconsistent relationship between subglacial sediment transport capacity and water discharge (Figure~\ref{fig:Qw_vari}).
+
+Over time scales of weeks or longer, the glacier's sediment transport capacity is impacted by the ice thickness, controlling the channel closure rate, and the glacier's surface slope, in addition to water discharge \cite<Figure~\ref{fig:multi_run}, Section~\ref{sect:sub_mode}; >{rothlisberger1972,shreve1972,gimbert2016,stevens2022}.
+In a supply-limited state, often assumed in landscape evolution models, then glaciers' sediment discharge record are also impacted by the additional processes of sediment availability and bedrock erosion from  water pressure variations and sliding  \cite<e.g.>{iverson2012,herman2015,delaney2023}.
+This multitude of processes lies in contrast to processes occurring in most subaerial channels, where transport capacity typically responds to sediment size, water discharge, and hydraulic gradient, and the hydraulic gradient generally remains stable over time scales ranging from years to millennia \cite<Section~\ref{sect:fluv}; e.g.>{muller1968,tucker1997,wickert2019}.
+The stronger link between sediment transport capacity and hydrological forcing in subaerial channels may mean that variations in sediment transport records can be more easily attributed to climatic conditions than in subglacial systems.
+
+In subglacial channels, peaks in water velocity, shear stress and width-integrated shear stress occur with  fast rates of water discharge increase relative to the channel's growth rate (Figure~\ref{fig:model_outs}).
+In turn, subglacial channels grow with increasing water discharge, effectively reducing sediment transport capacity with the stabilization of water discharge at its peak.
+Similar feedback occurs in subaerial channels \cite{phillips2016}, albeit at  timescales apparently longer than days in subglacial channels.
+
+\subsection{Role of hydrological forcing}
+
+%\mauro{You often use ``subglacial hydraulics''.
 I'm not sure what that means.
@@
 Maybe you could define a bunch of equations as the subglacial hydraulic model and subaerial hydraulic model and then talk of that?}
-%
-The stronger correlation between subglacial hydraulics and water discharge in \icesheet{} (Figure~\ref{fig:Qw_vari}\,a-f) could result from the channel being closer to equilibrium between variations in the channels' wetter area and water discharge with smaller diurnal variations in amplitude (Figures~\ref{fig:model_outs} and ~\ref{fig:Qw_vari}) \mauro{needs rewriting}.
-Conversely, the correlation between hydraulics and water discharge in \alpine is smaller due to the relatively larger discharge variations on short time scales.
-In both scenarios, the variability in subglacial channels approaches that of subaerial ones  over $1-5$ days (Figure~\ref{fig:multi_run}), suggesting a parity between the two channel types might be approached over this period.
 
-In subglacial channels, peaks in water velocity, shear stress and width-integrated shear stress occur with  fast rates of water discharge increase relative to the channel's growth rate (Figure~\ref{fig:model_outs}).
-In turn, subglacial channels grow with increasing water discharge, effectively reducing sediment transport capacity with the stabilization of water discharge at its peak.
-Similar feedback occurs in subaerial systems \cite{phillips2016}, albeit at timescales longer than days in the subglacial system.
+%\ian{make comment about lumped nature of model}
 
-\subsection{Effect of variable sediment transport capacity on sediment discharge  from glaciers}
-\mauro{This section header sounds like this will be about measurements only.}
+Results point to the highly variable response of water discharge quantity to proxies for sediment transport capacity, yet substantial differences emerge between the \alpine{} and \icesheet{}.
+Indeed, the correlation between hydraulics and water discharge in \alpine is smaller compared to \icesheet due to the relatively larger discharge variations on daily to weekly timescales.
+This could occur as the evolution of the R-channel's size is small compared to the variations in water discharge.
+Such variability in alpine catchments may cause shear stress to be proportional to the water discharge squared \cite< \ref{sect:scaling}; c.f.>{alley1997}.
+In turn, a closer approximation of the hydraulics could be represented by pipe flow, where the channel size remains fixed.
+Conversely, the reduced variability in water discharge in \icesheet may result in the R-channel's size being closer to equilibrium with respect to the variations in water discharge.
+In this case, the exponent on water discharge with respect to shear stress is likely significantly less than $2$, but greater than $\frac{4}{5}$ that occurs when the R-channel's size evolves in equilibrium with water discharge (see \ref{sect:scaling}).
+As a result in the stronger correlation persists between subglacial model outputs and water discharge in \icesheet{} compared to \alpine (Figure~\ref{fig:Qw_vari}\,a-f).
 
-\mauro{I find this confusing.
-The measurements are sediment discharge, so how can they be represent by hydraulic outputs?
-Maybe talk of sediment discharge proxy ($w\tau$) or some such?
-Or use something establishing a less direct link than ``representing''.}
-The increased performance of subaerial hydraulics outputs, compared to subglacial ones, in representing observations occurs despite the proximity of measurement locations to these glaciers (Figure~\ref{fig:Qw_vari}).
-This could occur for two reasons.
+The hydrological forcing that could cause the greatest increases in subglacial sediment transport capacity may occur during glacier lake drainages or intense precipitation events \cite<e.g.>{cowan1988,delaney2018}.
+In these cases, massive amounts of sediment can be evacuated from the subglacial environment, due to increased sediment access, but also the large increases in sediment transport capacity.
+If water input prior to these events has been minimal, the channel size would be small and sediment transport capacity would very rapidly increase at the onset of the event (see \ref{sect:scaling}).
 
-First, these measurement stations lie $0.5$ to $2$\,\unit{km} downstream of the glacier terminus \cite{cowton2012,felix2022}, a long enough distance for the pressurized subglacial regime to transition to the open channel subaerial one.
-The transition of subglacial to subaerial conditions typically occur still underneath glacier but close to the terminus \cite{perolo2018}, where closure rates of the subglacial channels are slow and open channel flow condition prevail \cite{rothlisberger1972}.
-Furthermore, if abundant sediment persists in the proglacial area, then it can be easily mobilized, and subaerial processes could be represented just a short distance downstream \mauro{Not sure what that sentence means}.
+The greater variations in subglacial sediment transport capacity in both catchment types compared to subaerial ones could cause a supply-limited regime to persist at many glaciers.
+This regime may persist as the sediment's threshold of motion can be reached more frequently and across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
+Sediment exhaustion occurring through increased variability may also explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \cite{overeem2017}.
 
-\mauro{I perceived this paragraph pretty out of the blue.}
-The consistent relationship between sediment discharge capacity in  open channel flow conditions could be pronounced $\sim20$\,\unit{km} downstream of the Leverett site, where a strong correlation persists between sediment plume size and the river's water discharge into the Kangerlussuaq fjord \cite{mcgrath2010}.
-In contrast, in marine-terminating glacier catchments a less consistent relationship may occur between water discharge or melt extent and sediment plume size \cite{tedstone2012}.
-Here, ocean water pressurizes subglacial channels at the ice front \cite<e.g.>{how2017}, so the minimal correlation could result from the inconsistent relationship between subglacial sediment transport capacity and water discharge (Figure~\ref{fig:Qw_vari}).
-
-Second, the sediment discharge records could represent processes such as increased sediment accessibility when melt extends up the glacier, thereby increasing sediment and water discharge \cite<e.g.>{vergara2022}.
-A supply-limited regime could well persist at many glaciers, because the sediment's threshold of motion can be reached more frequently and across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
-Sediment exhaustion occurring through increased variability may also explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \cite{overeem2017}.
-Note, however, the absence of lags between the observed sediment and water discharge (Figure~\ref{fig:model_outs}\, a and e) may suggest that sediment is mobilized close to the measurement stations, as opposed to distal areas of the glacier bed as would be expected in a supply-limited regime \cite{williams1989}.
-
-Regardless of the increased performance of water discharge compared to subglacial sediment transport capacity in representing the observations,  suggest that additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate \cite{jansson2005,ganti2016}.
-Sediment discharge records have been used to establish the relationship, or lack thereof, between climate forcing and glacial erosion \cite<e.g.>{koppes2009a,ganti2016,willenbring2016,mariotti2021}.
-The variable relationship between water discharge and subglacial sediment transport capacity presented here mostly applies to the daily- to weekly-time scales controlling the size of subglacial channels (Figure~\ref{fig:multi_run}).
-Yet, identifying climatic effects in sediment transport from the transport-limited component of glacier-sedimentary systems may require higher thresholds of water discharge compared to subaerial systems when water discharge varies substantially over short timescales \cite<Figure~\ref{fig:Qw_vari}; >{tofelde2021}.
-\mauro{I don't understand the last two sentences.}
-
-Over time scales of weeks or longer, the glacier's sediment transport capacity is impacted by the ice thickness, controlling the channel closure rate, and the glacier's surface slope, in addition to water discharge \cite<Figure~\ref{fig:multi_run}, Section~\ref{sect:sub_mode}; >{rothlisberger1972,shreve1972,gimbert2016,stevens2022}.
-In a supply-limited state, often assumed in landscape evolution models, then glaciers' sediment discharge record are also impacted by the additional processes of sediment availability and bedrock erosion from  water pressure variations and sliding  \cite<e.g.>{iverson2012,herman2015,delaney2023}.
-This multitude of processes lies in contrast to processes occurring in most subaerial systems, where transport capacity typically responds to sediment size, water discharge, and hydraulic gradient, and the hydraulic gradient generally remains stable over time scales ranging from years to millennia \cite<Section~\ref{sect:fluv}; e.g.>{muller1968,tucker1997,wickert2019}.
-The stronger link between sediment transport capacity and hydrological forcing in subaerial channels may mean that variations in sediment transport records can be more easily attributed to climatic conditions than in subglacial systems.
-
-\ian{move this shit up or delete}
-\subsection{Comparison of Alpine and Arctic hydrological forcing}
+Water discharge measurement at the sub-hour timescale remain unavailable in many catchments and daily values might be used, for instance.
+The variability in subglacial channels approaches that of subaerial ones  over $1-5$ days (Figure~\ref{fig:multi_run}), possibly resulting in a parity between subaerial and subglacial channels over these periods.
+This suggests that aggregating water discharge over daily to weekly time periods could yield similar behavior between subglacial and subaerial systems. 
+At the same time, water discharge over such time periods, may omit key process, such as water velocity reaching sediments' threshold of motion more frequently or the timing of peak events.
 
@@
 In contrast, pressurized subglacial channels largely accommodate rapidly changing water discharge by altering water velocity.
-Their size, in contrast, responds over the course of several days to changes in discharge.
+Their size responds over the course of several days to changes in discharge.
 Thus the response of sediment transport capacity is much more sensitive to changes in water discharge in subglacial channels compared to subaerial ones.
-%Despite these differences, parameterizations of subaerial hydraulics in two glacierized catchments generally represent observed sediment discharge variations better than subglacial hydraulics \mauro{This sentence does not quite parse.
-Why ``despite''?
-Also state that we found this.}
 
@@
 \item the changing relationship between water discharge and sediment transport capacity underneath glaciers creates larger variability in sediment transport capacity compared to river systems;
-\item the timing of peak sediment transport capacity below glaciers generally coincides with large rates of increase in water discharge, as opposed to the maximum water discharge, as would occur in pipe flow of a fixed size;
-\item the timing of peak sediment transport capacity in subaerial channels coincides with the maximum water discharge;
-\item a steady relationship between water and sediment discharge may represent subglacial sediment access or open channel hydraulics near the glacier margin, instead of subglacial sediment transport capacity.
+  \vspace{.1cm}
+
+\item the timing of peak sediment transport capacity below glaciers generally coincides with large rates of increase in water discharge, as opposed to the maximum water discharge in subaerial channels;
+  \vspace{.1cm}
+
+\item the degree of difference in sediment transport capacity variations between the two channels types depends on the water discharge variability in the hydrograph, which varies with geography and across timescales.
+  However, representing subglacial sediment transport capacity with either open-channel flow or pipe flow omits key subglacial sediment transport characteristics.
+  \vspace{.1cm}
+
+  \item the uneven response of subglacial sediment transport capacity to water discharge variations suggests different magnitude-frequency relationships occur in most subglacial channels compared to subaerial ones with similar hydrological forcings.
 \end{enumerate}
@@
 A.J.T. helped interpret data from Leverett Glacier and provided guidance on experiment design.
-M.A.W. provided data from Fieschergletscher and added the relationship between shear stress and water discharge (see Sections~\ref{sect:dis_qsc} and \ref{sect:scaling}).
+M.A.W. provided data from Fieschergletscher and added the relationship between shear stress and water discharge (see \ref{sect:scaling}).
 D.F. provided data from Fieschergletscher.
@@
 
-SNSF Project No. PZ00P2\_202024 provided  funding for I. Delaney.
-A. Tedstone acknowledges funding from the European Research Council, award 818994 -- CASSANDRA.
+SNSF Project No. $\mathrm{PZ00P2}$\_$202024$ provided  funding for I. Delaney.
+A. Tedstone acknowledges funding from the European Research Council, award $818994$ -- CASSANDRA.
 G. King provided useful comments on a previous version of this manuscript.
@@
     Where two values are given, the first refers to  \alpine{}, a scenario from Fieschergletscher, and the second to a glacier marginal to the \icesheet{} a scenario from Leverett Glacier}
-  \tiny % TODO remove again
+  \small % TODO remove again
   \begin{tabular}{ l  c  c c }
@@
     Shear stress (subglacial, subaerial) & $\tau$, ($\tau_g,\,\tau_f$) && $\mathrm{Pa \, m^{-2}}$ \\
-    Width-integrated shear stress (subglacial, subaerial) \mauro{get rid of this} & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$\\
     Stream power & $\Omega$ && $\mathrm{ kg \, m\, s^{-3}}$ \\
@@
     Density of ice & $\rho_i$& $900$ & $\mathrm{kg\,m^{-3}}$ \\
-    Hooke angle of channel & $\beta$ & $\frac{\pi}{2}$ & \unit{rad}\\
-    Glacier thickness &$h_{ice}$& $225$ or $700$  &\unit{m}\\
+    Hooke angle of channel & $\beta$ & $\frac{\pi}{6}$ & \unit{rad}\\
+    Friction factor (subglacial, subaerial) & ($f_r$, $f_f$) &$5,\,16$ or $3$ & $\mathrm{(-)}$ \\
+    Glacier thickness &$h_{ice}$& $225$ or $740$  &\unit{m}\\
     Effective glacier thickness &$h_o$&$\frac{\rho_i}{\rho_w} h_{ice}$  &\unit{m}\\
-    Effective glacier length &$l$& 7,000 or 26,000&\unit{m}\\
+    Effective glacier length &$l$& $7,000$ or $26,000$&\unit{m}\\
     Constant $1$ in Equation~\ref{eq:dS_dt} &$C_1$&$2.2\times10^{-5}$&\unit{m}$^{-1}$\\
@@
     Ice flow exponent &$n$& $3$ &$\mathrm{(-)}$\\
-    Friction factor (subglacial, subaerial)  \mauro{why not $f_g$ and $f_f$?} & $f_r$ &$10$ or $20$ & $\mathrm{(-)}$ \\
-    Friction factor (subaerial) & $f_p$ & $3$ & $\mathrm{(-)}$\\
+
     Gradient of channel bed (subaerial) &$\nabla z_c$ &$0.05$& $\mathrm{(-)}$\\
@@
 
-\mauro{Notation: either use $h$ and $w$ for the river, or otherwise, somehow free $h$. Below I use $H$ and $W$ for now.}
+%\mauro{Notation: either use $h$ and $w$ for the river, or otherwise, somehow free $h$. Below I use $H$ and $W$ for now.}
 
@@
 }
+  \small
   \label{tab:eqs1}
@@
   }
+  \small
   \label{tab:Qs}
@@
   }
+  \small
   \label{tab:Qs2}
@@
 \end{document}
+
+% Graveyard
+% Additionally uggest that additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate \cite{jansson2005,ganti2016}.
+% Sediment discharge records have been used to establish the relationship, or lack thereof, between climate forcing and glacial erosion \cite<e.g.>{koppes2009a,ganti2016,willenbring2016,mariotti2021}.
+% The variable relationship between water discharge and subglacial sediment transport capacity presented here mostly applies to the daily- to weekly-time scales controlling the size of subglacial channels (Figure~\ref{fig:multi_run}).
+% Yet, identifying climatic effects in sediment transport from the transport-limited component of glacier-sedimentary systems may require higher thresholds of water discharge compared to subaerial systems when water discharge varies substantially over short timescales \cite<Figure~\ref{fig:Qw_vari}; >{tofelde2021}.
+\mauro{I don't understand the last two sentences.}
+%Furthermore evaluating the different variability in sediment transport characteristics across timescales can help inform the sediment transport processes omitted from daily or week values of water discharge..

### 2024-03-18 Delaneyetal2024.tex
mauros edits moving forward

@@
 \linenumbers
+%% make line numbers work with amsmath
+% https://tex.stackexchange.com/questions/461186/how-to-use-lineno-with-amsmath-align
+\usepackage{lineno}   %% <- no mathlines option
+\usepackage{amsmath}  %% <- after lineno
+\usepackage{etoolbox} %% <- for \cspreto, \csappto
+
+%% Patch 'normal' math environments:
+\newcommand*\linenomathpatch[1]{%
+  \cspreto{#1}{\linenomath}%
+  \cspreto{#1*}{\linenomath}%
+  \csappto{end#1}{\endlinenomath}%
+  \csappto{end#1*}{\endlinenomath}%
+}
+
+\linenomathpatch{equation}
+\linenomathpatch{gather}
+\linenomathpatch{multline}
+\linenomathpatch{align}
+\linenomathpatch{alignat}
+\linenomathpatch{flalign}
+
 %%%%%%%
@@
 \newcommand{\mauro}[1]{{\textbf{\color{green}Mauro says:} \color{green} #1} }
-\newcommand{\alpine}{\textit{ALPINE}\,}
-\newcommand{\icesheet}{\textit{ICESHEET}\,}
+\newcommand{\alpine}{\textit{ALPINE}}
+\newcommand{\icesheet}{\textit{ICESHEET}}
 \newcommand{\m}{$\,\mathrm{m}$\,}
@@
 \affiliation{2}{Department of Geosciences, University of Fribourg, Ch. du Musée 1700, Fribourg, Switzerland}
-\affiliation{3}{Laboratory of Hydraulics, Hydrology and Glaciology (VAW), ETH-Z\"urich, H\"onggerbergring 26, 8093 89 Z\"urich, Switzerland}
-\affiliation{4}{Swiss Federal Institute for Forest, Snow and Landscape Research (WSL) Z\"uricherstrasse 111, 8903 1011 Birmensdorf, Switzerland}
+\affiliation{3}{Laboratory of Hydraulics, Hydrology and Glaciology (VAW), ETH-Z\"urich, H\"onggerbergring 26, 8093 Z\"urich, Switzerland}
+\affiliation{4}{Swiss Federal Institute for Forest, Snow and Landscape Research (WSL) Z\"uricherstrasse 111, 8903 Birmensdorf, Switzerland}
 
@@
 \begin{keypoints}
-\item Short term water discharge variations in subglacial channels are mainly accommodated by water velocity, not channel size.
+\item Short term water discharge variations in subglacial channels are mainly accommodated by water velocity, not channel size, as is the case in subaerial systems.
 \item Greater variability in water velocity causes sediment transport capacity to vary more in subglacial channels than in subaerial channels.
-\item Results suggest that significant sediment transport capacity can persist across a range of water discharge conditions, and timing of peak events occurs under different hydrological conditions.
+\item Results suggest that significant subglacial sediment transport capacity can persist across a range of water discharge conditions, and the timing of peak events occurs under different hydrological conditions.
 \end{keypoints}
@@
 \begin{abstract}
-  Sediment transport capacity in both subaerial and subglacial channels depends on the shear stress exerted across the channel bottom, which vary with the  water velocity and channel width.
+  Sediment transport capacity in both subaerial and subglacial channels depends on the shear stress exerted across the channel bottom, which varies with the  water velocity and channel width.
   In subaerial channels, water discharge variations are accommodated largely by changes in flow depth and width.
-  While in subglacial channels, water discharge changes primarily affect water velocity due to slow-evolving channel geometry.
-  Hydrographs from an Alpine glacier and the Greenland Ice Sheet are applied to models of subglacial and subaerial hydraulics, revealing greater variability in the inferred sediment transport capacity in subglacial than in subaerial channels.
-  Results also show that high subglacial sediment transport rates can occur across a wide range of water discharges.
-  However, slowly evolving channel width accommodates some variability in subglacial channels, compared to subaerial channels.
+  While in subglacial channels, water discharge changes primarily affect water velocity due to evolving channel geometry over hours to weeks.
+  Hydrographs from an Alpine glacier and the Greenland Ice Sheet are used to drive models of subglacial and subaerial hydraulics, revealing greater variability in the inferred sediment transport capacity in subglacial than in subaerial channels.
+  Results also show that high subglacial sediment transport rates and water velocities can occur across a wide range of water discharges.
   These findings are discussed in the context of understanding sediment discharge from glaciers with different hydrological forcings and assumptions of channel type.
@@
 Sediment export from glaciers impacts downstream ecosystems and communities and it has been documented to change in response to climate warming and glacier retreat.
-In these catchments, sediment export depends in part on the sediment transport capacity of water flowing below the glacier, or the amount of sediment that could be transported given the abundance of sediment.
+In these catchments, sediment export depends in part on the sediment transport capacity, the amount of sediment that could be transported for the hydraulic conditions given the abundance of sediment.
 This capacity largely depends on the width of the channel and the water's velocity. 
-Unlike subaerial channels, like rivers, where the size of the river changes rapidly with respect to water discharge, however, subglacial channels are pressurized by the ice above.
-This pressurized flow means that channel size evolves slowly compared to the variations in water discharge, therefore variations in water discharge are largely accommodated by variations in velocity.
-Models of these processes show that sediment transport capacity is highly variable in subglacial channels, and a wide range of water discharges can result in comparable sediment transport capacity.
-In subaerial channels, by contrast, sediment transport capacity is less variable and depends on a single water discharge quantity.
-The channel's different response to water discharge may impact the interpretation of sediment transport records and responds to the type of hydrological forcing applied.
+Unlike subaerial channels, like rivers, where the size of the river changes rapidly with water discharge, however, subglacial channels are pressurized by the ice above.
+This pressurized flow means that channel size evolves slowly compared to the variations in water discharge, therefore variations in water discharge are largely accommodated by variations in velocity as opposed to channel size.
+Models of these processes show that sediment transport capacity is highly variable in subglacial channels, and a wide range of water discharges can result in comparable sediment transport capacities.
+In subaerial channels, by contrast, sediment transport capacity is less variable and depends solely on water discharge.
+The channel's different responses to water discharge may impact the interpretation of sediment transport records and respond to the type of hydrological forcing applied.
 
@@
 In some regions, increased water discharge and glacier melt have been interpreted to yield greater sediment transport capacity \cite{bendixen2017,costa2017,li2021}.
-Observed changes to sediment transport in glacierized catchments require examining the processes controlling sediment discharge in these catchments and its variations with water discharge  \cite<e.g.>{riihimaki2005,swift2005}.
+To explain observed changes to sediment transport in glacierized catchments require examining the processes controlling sediment discharge in these catchments and its variations with water discharge  \cite<e.g.>{riihimaki2005,swift2005}.
 
 Over long periods, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from under glaciers \cite<c.f.>{hallet1979,iverson2012,ugelvig2018}.
-On shorter time periods, pressurized water transports sediment from underneath glaciers \cite{walder1994,creyts2013,beaud2018,delaney2019}, should enough sediment be present subglacially (i.e.
+Pressurized subglacial water can transport this sediment from underneath glaciers \cite{walder1994,creyts2013,beaud2018,delaney2019}, should enough sediment be present subglacially (i.e.
 in a transport-limited regime).
@@
 In a transport-limited regime, sediment discharge tracks sediment transport capacity, defined as the amount of sediment the water can carry.
-Sediment transport capacity depends on the shear stress between water and the sediment it flows  over \cite{shields1936,meyer1948,engelund1967} along with the width of the channel bottom $w$ over which to mobilize sediment.
+Sediment transport capacity depends on the shear stress between water and the sediment it flows  over \cite{shields1936,meyer1948,engelund1967} along with the width of the channel bottom $w$ over which sediment mobilizes.
 The shear stress $\tau$ responds to the velocity of water $v$ flowing through the channel so that
@@
 The response of water velocity to changing water discharge in subglacial channels differs from subaerial ones.
-The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by melting due to frictional heating of water flowing through the channel \cite{rothlisberger1972}.
+The size of subglacial channels responds to the creep closure of the ice and the opening of the channel by melting due to frictional heating of water flowing through the channel \cite{rothlisberger1972}.
 As a result, the subglacial channel size only evolves relatively slowly over days to months, whereas water discharge can vary over hours \cite<e.g.>{iken1986,andrews2014,nanni2020}.
@@
     \includegraphics[width=0.8\linewidth]{Fig1.pdf}
-    \caption{Sketch for the different responses that subglacial and subaerial channels to increased water discharge.
+    \caption{Sketch for the different short time-scale responses of subglacial and subaerial channels to increased water discharge.
       Water velocity magnitudes in the subglacial, $v_g$, and subaerial, $v_f$, channels are shown by arrow length.
       $S_g$ and $S_f$ represent the wetted area in subglacial and subaerial channels, respectively.
-      The subglacial channel widths $w_g$ remain unchanged, while the subaerial channel widths $w_f$ evolve. Shear stress $\tau$ is responsible for the mobilization of sediment.
+      The subglacial channel width $w_g$ remain unchanged, while the subaerial channel width $w_f$ evolves. Shear stress $\tau$ is responsible for the mobilization of sediment.
     }
@@
 These differences are implicitly included in a wide range of available models that  quantify sediment transport in both subglacial and subaerial channels  \cite<e.g.>{walder1994,alley1997,tucker1997,creyts2013,delaney2019,hewitt2019,wickert2019}.
-The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \cite<e.g.>{perolo2018}.
+The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics near glacier margins where flow transitions from pressurized to open channel flow \cite<e.g.>{perolo2018}.
 Additionally, the variable response of sediment transport capacity to water discharge in the two systems may affect the interpretation of sediment transport records in glacial systems \cite<e.g.>{muller1968,richards2003,swift2005,ganti2016}.
@@
 This manuscript evaluates the relationship between sediment transport capacity and water discharge in subglacial channels and compares it to subaerial ones.
-Models of subglacial and subaerial hydraulics are used to quantify subglacial and subaerial water velocity and channel morphology.
+Models of subglacial and subaerial hydraulics are used to quantify the respective water velocity and channel morphology.
 The models are applied to hydrological records from an Alpine glacier in Switzerland (Fieschergletscher) and  a land-terminating glacier in Greenland (Leverett Glacier).
@@
 
-Water discharge data is leveraged from Alpine and ice sheet settings collected downstream of the glacier, assuming that no water storage occurred in the proglacial area.
-The Alpine site (\alpine) is  Fieschergletscher in the Swiss Alps ($46^\circ\,29'\,07''$ N, $8^\circ\,08'\,3''$ E).
-Water discharge was collected here at a $1$\unit{min} interval continuously over the period 2014--2021.
-See \citeA{felix2022} and \citeA{felix2021} for more details.
+Water discharge data is used from Alpine and ice sheet settings collected downstream of the glacier, assuming that no water storage occurred in the proglacial area.
+The Alpine site (\alpine{}) is  Fieschergletscher in the Swiss Alps ($46^\circ\,29'\,07''$ N, $8^\circ\,08'\,3''$ E).
+Water discharge was collected here at a $1$\unit{min} interval continuously over the period 2014--2021;
+see \citeA{felix2022} and \citeA{felix2021} for more details.
 
-The Leverett Glacier in Greenland (\icesheet) serves as the ice sheet setting.
+The Leverett Glacier in Greenland (\icesheet{}) serves as the ice sheet setting.
 Water discharge was collected roughly $2$\unit{km} downstream from the terminus between June and August/September 2009 to 2012 ($67^\circ\,03'\,5''$ N, $50^\circ\,12'\,59''$ W).
@@
 The models described below (Sections ~\ref{sect:sub_mode}~and~\ref{sect:fluv}) represent relationships amongst water discharge, water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
-Both models calculate   water velocity, shear stress, and width-integrated shear stress, upon which sediment transport depends \cite<Figure \ref{fig:cartoon}; >{shields1936}.
+Both models use the measured discharge to calculate water velocity, shear stress, and width-integrated shear stress, upon which sediment transport depends \cite<Figure \ref{fig:cartoon}; >{shields1936}.
 The evaluation of these variables omits the selection of a sediment transport relationship and a grain-size parameter \cite<e.g.>{shields1936,meyer1948} needed to calculate sediment transport capacity from shear stress.
@@
 \end{linenomath*}
-\noindent where $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (Table \ref{table:vpm}), $l$ is the length of the subglacial channel, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change from the glacier terminus to its top, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the ice overburden pressure expressed in meter water equivalent, and $n$ is Glen's n \cite{glen1955} (usually $n=3$).
+\noindent where $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (Table \ref{table:vpm}), $g$ is the acceleration due to gravity, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change over $l$, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the mean ice overburden pressure expressed in meter water equivalent, and $n$ is Glen's n \cite{glen1955} (usually $n=3$).
 The first term on the right side of the equation represents the opening of the channel through frictional heating, while the following term represents the creep closure of the channel from ice deformation.
@@
 Note that $\beta =\pi$ corresponds to a semi-circular channel and smaller values of $\beta$ result in shallow, wide channels.
-%
+This completes the subglacial hydraulic model which is described by the state variables $S_g$ and $\Delta h$.
+
 The shear stress, $\tau_g$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
@@
 %
-where $v_g = \frac{Q}{S_g}$ is the water velocity.
+where $v_g = \frac{Q}{S_g}$ is the water velocity and $\rho_w$ the density of water.
 %
@@
 %
-This value is used to establish the integrated shear stress across the channel, $w_g\tau_g$.
+This value is used to establish the integrated shear stress across the channel $w_g\tau_g$.
 
@@
 Note that this subaerial channel model is purely algebraic, whereas the subglacial model comprises a differential equation for the evolution of $S_g$.
-Thus the models have a history dependence on the discharge $Q_w$.
+Thus the subglacial model has a history dependence on the discharge $Q_w$.
 
@@
 
-The models above are applied to hydrological records from the Fieschergletscher (scenario \alpine) and the Leverett Glacier (scenario \icesheet).
+The models above are applied to proglacial discharge records from the Fieschergletscher (scenario \alpine{}) and the Leverett Glacier (scenario \icesheet{}).
 The subglacial and subaerial models represent the subglacial flow of water exiting a glacier and transitioning to subaerial flow.
@@
 a segment of a quarter-circle, (Table~\ref{table:vpm}).
-Friction factors $f_r$ and $f_f$ are tuned so that reasonable water velocities occur for both \alpine and \icesheet{}\cite<$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$>{werder2010b,chandler2013}.
+Both friction factors $f_r$ and $f_f$ are tuned so that reasonable water velocities occur for both \alpine{} and \icesheet{} \cite<$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$>{werder2010b,chandler2013}.
+The initial condition of cross-sectional area $S_g$ is evaluated by applying a model spinup with the maximum water discharge for the first $3$ days of the study until the change in channel area is negligible.
 The subaerial channels of both scenarios have a bed slope of $\nabla z_c = 0.05$ (Table~\ref{table:vpm}).
-
-A second experiment aims to further characterize the variability in sediment discharge capacity in subglacial compared to subaerial channels across a range channel shapes, friction factors and water discharge variability.
-To accomplish this, the water discharge records are aggregated across a select time period in the past ranging from $15$\,\unit{min} up to $15$\,\unit{days}, and the model is run with random  parameter combinations of channel geometry factors ($\beta$ and $k$) and friction factors ($f_r$ and $f_f$).
-Runs with parameter combinations are accepted if their mean subglacial water velocity over the season lies between $0.75$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $2$\,\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity lies between $0.3$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with measurements \cite<e.g.>{werder2010b,magnusson2012,chandler2013}.
-The routine runs until $100$ different parameter combinations are accepted from each water discharge aggregation period.
-The variability of an output, a key difference between the subaerial and subglacial systems, is represented by  its  standard deviation over the course of the season.
 
@@
 Rank correlation reduces the impact of the non-linear response of sediment transport capacity to the hydrological forcing, thus making our interpretation independent of the exact details of the sediment transport formula.
+
+A second experiment aims to further characterize the variability in sediment discharge capacity in subglacial compared to subaerial channels across a range of channel shapes, friction factors, and water discharge variability.
+The variability of one model output (i.e water velocity), a key difference between the subaerial and subglacial systems, is represented by its standard deviation throughout the season.
+
+The effect of channel shape and roughness is established by running the model with random parameter values of channel geometry factors ($\beta$ and $k$) and friction factors ($f_r$ and $f_f$).
+Runs with parameter combinations are accepted if their mean subglacial water velocity over the season lies between $0.75$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $2$\,\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity lies between $0.3$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}$^3$\,\unit{s}$^{-1}$, following measurements \cite<e.g.>{werder2010b,magnusson2012,chandler2013}.
+Additionally, to be accepted, model runs can only experience a flotation fraction of above $1.2$ for $2.5$\,\% of the run.
+
+To establish the role of water discharge variability in driving variations in sediment transport capacity, the water discharge records are aggregated across a select period ranging from $15$\,\unit{min} up to $15$\,\unit{days}.
+Aggregation periods longer than this might be conflated with seasonal variations in water discharge and are thus omitted.
+In this case, the water discharge input at a given time is the average over the aggregation period before that time.  
+The routine runs until $100$ different parameter combinations are accepted for each water discharge aggregation period.
 
@@
   \includegraphics[width=0.9\linewidth]{Fig2.pdf}
-  \caption{Model outputs from the hydrographs n panels a and e for \alpine scenarios (a-d) and \icesheet{} (e-h).
+  \caption{Model outputs from simulations using the hydrographs in panels a and e for the scenarios \alpine{} (a-d) and \icesheet{} (e-h).
     Gray (black) lines in a and e represent the subglacial channel size (water discharge).
@@
     Data are shown at $15$\unit{min} intervals, i.e. not aggregated.
-    Arrows denote where variables peaking in the subglacial channel prior to the subaerial one.
-    Insets in e-h show the  peak melt event \icesheet{} denoted by the shaded area in panel e.
     % Note that the x and y axes are different for \alpine{} and \icesheet{}.
@@
 
-Following the models defined in Sections~\ref{sect:sub_mode}~and~\ref{sect:fluv}, the water velocity, shear stress, and width-integrated shear stress exhibit different seasonal evolutions and peaks.
+Following the first experiment defined in Sections~\ref{sect:imp}, model outputs exhibit different seasonal evolutions and peaks.
 Water velocity and shear stress represent the potential for sediment mobilization, and width-integrated shear stress represents a proxy for the total sediment transport capacity across the channel bed (Figure~\ref{fig:model_outs}).
-Furthermore, the subglacial hydraulic outputs are generally larger than their subaerial equivalents , thus greater sediment transport capacity would be expected in subglacial channels compared to subaerial ones.
+Furthermore, the subglacial model outputs are generally larger than their subaerial equivalents, thus greater sediment transport capacity would be expected in subglacial channels compared to subaerial ones.
 
@@
   \caption{
-    Relationship between water discharge and velocity, shear stress, and width-integrated shear stress for \alpine{} (a-c) and \icesheet (d-f).
-    %These outputs' relationship with measured sediment discharge is given for \alpine in (g-i) and \icesheet (j-l).
+    Relationship between water discharge and normalized velocity, shear stress, and width-integrated shear stress for \alpine{} (a-c) and \icesheet{} (d-f).
+    %These outputs' relationship with measured sediment discharge is given for \alpine{} in (g-i) and \icesheet{} (j-l).
     Variables on the x-axis have been normalized to mean values.
@@
 
-In examining the relationship between water discharge and the model outputs, different relationship emerges between the subaerial and subglacial cases.
+Examining the relationship between water discharge and the model outputs, different relationship emerges between the subaerial and subglacial cases.
 Because there is no history dependence in  subaerial channels, each water discharge value in the subaerial channel results in a unique water velocity, shear stress, and width-integrated shear stress (Section~\ref{sect:sub_mode}).
-This characteristic results in a perfect rank correlation between the variables (Figure~\ref{fig:Qw_vari} a--f).
+This characteristic results in a perfect rank correlation between the variables (Figure~\ref{fig:Qw_vari}).
 In the subglacial channel, however, the hysteresis in channel size evolution (Equation~\ref{eq:dS_dt}) means that subglacial model outputs respond differently to water discharge than subaerial ones.
-Here, the persistence of high water velocity, shear stress, and width-integrated shear stress across a range of water discharges suggests that substantial sediment transport may occur under a wide range of hydrological forcings (Figure~\ref{fig:Qw_vari} a--f).
+Here, the persistence of high water velocity, shear stress, and width-integrated shear stress across a range of water discharges suggests that substantial sediment transport may occur under a wide range of hydrological forcings (Figure~\ref{fig:Qw_vari}.
 
-For instance in \alpine, very high water velocity values and shear stresses can occur at a minimal water discharge  ($\sim\,4$\,\unit{m}$^3$\,\unit{s}$^{-1}$) to the maximum water discharge at over ($17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, Figure~\ref{fig:Qw_vari}\,a).
-In \icesheet, mean values water velocity over the season can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
-When considering the subglacial channel's evolving width, width-integrated shear stress  across the channel generally increases with water discharge, with improved rank correlation compared to water velocity or shear stress (\alpine, Figure~\ref{fig:Qw_vari} a--f).
+For instance, in \alpine{}, very high water velocity values and shear stresses can occur from a low water discharge  ($\sim\,4$\,\unit{m}$^3$\,\unit{s}$^{-1}$) to the maximum water discharge at over ($17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, Figure~\ref{fig:Qw_vari}\,a).
+In \icesheet{}, mean values water velocity over the season can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
+When considering the subglacial channel's evolving width, width-integrated shear stress  across the channel generally increases with water discharge, with improved rank correlation compared to water velocity or shear stress (\alpine{}, Figure~\ref{fig:Qw_vari} a--c).
 Yet even the width-integrated shear stress  can vary substantially, with the highest values occurring at water discharge values ranging from roughly $11$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ to over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
-The variability in width-integrated shear stress is less pronounced in the \icesheet scenario, where the hydrograph has less diurnal variability in water discharge (Figure~\ref{fig:Qw_vari} c, f).
+The variability in width-integrated shear stress is less pronounced in the \icesheet{} scenario, where the hydrograph has less diurnal variability in water discharge (Figure~\ref{fig:Qw_vari} c, f).
 
@@
     \includegraphics[width=0.9\linewidth]{Fig4.png}
-    \caption{Variability (standard deviation) of output variables for different time aggregations across the range of subglacial and subaerial channel shapes and friction factors.
-      Shaded areas denote the range of the $100$ parameter combinations tested (Section~\ref{sect:imp}).
+    \caption{Variability (standard deviation) of model outputs for different time aggregations across the range of subglacial and subaerial channel shapes and friction factors.
+      Shaded areas denote the range of results from the accepted $100$ parameter combinations tested (Section~\ref{sect:imp}).
       Solid lines denote  mean values.
@@
 The second numerical experiment aims to evaluate the role of channel shape and friction factors on the findings above.
-Across the range of channel shape parameters and friction factors  examined, variability in water velocity and shear stress remains higher in the subglacial system compared to the subaerial one.
-In \alpine{}, the variability in width-integrated shear stress is comparable between the subaerial and subglacial systems, due to the highly variable water discharge, particularly over diurnal time periods (Figure~\ref{fig:multi_run}).
+Across the range of channel shape parameters and friction factors  examined (Section\,\ref{sect:imp}), variability in water velocity and shear stress remains higher in the subglacial system compared to the subaerial one.
+In contrast, in \alpine{}, the variability in width-integrated shear stress is comparable between the subaerial and subglacial systems, due to the highly variable water discharge, largely due to diurnal variations (Figure~\ref{fig:multi_run}).
 However, in \icesheet{}, with lower water discharge fluctuations, the variability in width-integrated shear stress remains higher in the subglacial system across the time aggregation periods.
-Together these findings suggest that sediment transport capacity variations over a melt season can be comparable in subaerial and subglacial systems if substantial variability persists in the hydrological forcing.
+Together these findings suggest that sediment transport capacity variations over a melt season can approach in subaerial and subglacial systems if substantial variability persists in the hydrological forcing.
+\ian{state that these are generally higher, is t-test needed? }
 
-Evaluating the different variability in sediment transport characteristics across timescales can help inform the sediment transport processes omitted from daily or week values of water discharge.
-Results from  both \alpine and \icesheet scenarios show that the variability in velocity, shear stress, and width-integrated shear stress decreases with aggregations longer than approximately 1--5 days (Figure~\ref{fig:multi_run}).
-Yet, substantial differences in variability persist in periods of up to $15$ days.
-Aggregations periods longer than this might be conflated with seasonal variations in water discharge and are thus omitted.
-Note that lags and time of peak sediment transport conditions still occur over these periods, and these characteristics are thus note considered in aggregating outputs and water discharge over longer time periods (Figure~\ref{fig:model_outs}).
+Evaluating the different variability in sediment transport characteristics across timescales can help inform the sediment transport processes omitted from applying daily or weekly values of water discharge to understand sediment transport.
+Results from  both \alpine{} and \icesheet{} scenarios show that the variability in subglacial  decreases with aggregations longer than approximately 1--5 days (Figure~\ref{fig:multi_run}).
+Yet, substantial differences in the variability of model outputs for subglacial and subaerial cases persist in periods of up to $15$ days.
+Note that lags and time of peak sediment transport conditions still occur over these periods, and these characteristics are thus not considered in aggregating outputs and water discharge over longer periods (Figure~\ref{fig:model_outs}).
 
@@
 
-In R-channel models, viable water pressure regimes can be achieved by changing the shape and friction factor of the channels, enabling assumptions of semi-circular channels in most applications \cite{werder2010}.
+In R-channel models, viable water pressure regimes can be achieved by changing the shape and friction factor of the channels, enabling assumptions of semi-circular channels in most applications \cite{werder2010b}.
 However, the role of channel width in the findings here suggests that more thorough parameterization of channel shape and channel width is needed in assessing subglacial sediment transport capacity (Section\ref{sect:sub_mode}, \ref{sect:scaling}).
-Note as well that sediment transport relationships describe sediment flux per unit width, so that the integrated sediment transport capacity across channel occurs after application of the exponent, unlike the width-integrated shear stress evaluation here \cite<c.f.>{tucker1997}.
+\ian{Similar feedback occurs in subaerial channels \cite{phillips2016}, albeit at  timescales longer than days in subglacial channels. }
 
-Sediment transport capacity in subglacial channels responds more rapidly to changes in water discharge compared to their subaerial counter parts \cite{alley1997}.
-In subglacial R\"othlisberger channels in steady state with respect to water discharge the width-integrated shear stress response to water discharge is $ \tau_{tg}\, \propto\,  Q^{\sim\frac{4}{5}}$ (Table\,\ref{tab:eqs1}).
+Sediment transport capacity in subglacial channels responds more rapidly to changes in water discharge compared to their subaerial counterparts \cite{alley1997}.
+In subglacial R\"othlisberger channels in steady state with respect to water discharge, the width-integrated shear stress response to water discharge is $ w\tau_{tg}\, \propto\,  Q^{\sim\frac{4}{5}}$ (Table\,\ref{tab:eqs1}).
 However, subglacial channels rarely operate in a steady state \cite<e.g.>{gimbert2016}.
-For sufficiently short time periods, channel width and wetted area are fixed in the subglacial system (akin to pipe flow) so that the response of width-integrated shear stress is $ \tau_{tg}\, \propto\,  Q^{2}$ (Table\,\ref{tab:eqs1}).
+For sufficiently short periods, channel width and wetted area are fixed in the subglacial system (akin to pipe flow) so that the response of width-integrated shear stress is $ \tau_{tg}\, \propto\,  Q^{2}$ (Table\,\ref{tab:eqs1}).
 \ref{sect:scaling} provides a complete description of sediment transport capacity in different drainage conditions.
 
-The consistent relationship between sediment discharge capacity in  open channel flow conditions could be pronounced $\sim20$\,\unit{km} downstream of the Leverett site, where a strong correlation persists between sediment plume size and the river's water discharge into the Kangerlussuaq fjord \cite{mcgrath2010}.
+The co-varying relationship between sediment discharge capacity in  subaerial conditions could be pronounced $\sim20$\,\unit{km} downstream of the Leverett site, where a strong correlation persists between sediment plume size and the river's water discharge into the Kangerlussuaq fjord \cite{mcgrath2010}.
 In contrast, in marine-terminating glacier catchments a less consistent relationship may occur between water discharge or melt extent and sediment plume size \cite{tedstone2012}.
@@
 
+The stronger link between sediment transport capacity and hydrological forcing in subaerial channels may mean that variations in sediment transport records can be more easily attributed to climatic conditions than in subglacial systems.
 Over time scales of weeks or longer, the glacier's sediment transport capacity is impacted by the ice thickness, controlling the channel closure rate, and the glacier's surface slope, in addition to water discharge \cite<Figure~\ref{fig:multi_run}, Section~\ref{sect:sub_mode}; >{rothlisberger1972,shreve1972,gimbert2016,stevens2022}.
@@
 This multitude of processes lies in contrast to processes occurring in most subaerial channels, where transport capacity typically responds to sediment size, water discharge, and hydraulic gradient, and the hydraulic gradient generally remains stable over time scales ranging from years to millennia \cite<Section~\ref{sect:fluv}; e.g.>{muller1968,tucker1997,wickert2019}.
-The stronger link between sediment transport capacity and hydrological forcing in subaerial channels may mean that variations in sediment transport records can be more easily attributed to climatic conditions than in subglacial systems.
-
-In subglacial channels, peaks in water velocity, shear stress and width-integrated shear stress occur with  fast rates of water discharge increase relative to the channel's growth rate (Figure~\ref{fig:model_outs}).
-In turn, subglacial channels grow with increasing water discharge, effectively reducing sediment transport capacity with the stabilization of water discharge at its peak.
-Similar feedback occurs in subaerial channels \cite{phillips2016}, albeit at  timescales apparently longer than days in subglacial channels.
 
@@
 
-Results point to the highly variable response of water discharge quantity to proxies for sediment transport capacity, yet substantial differences emerge between the \alpine{} and \icesheet{}.
-Indeed, the correlation between hydraulics and water discharge in \alpine is smaller compared to \icesheet due to the relatively larger discharge variations on daily to weekly timescales.
+Our results point to the highly variable response of proxies for sediment transport capacity to  water discharge quantity, yet substantial differences emerge between the \alpine{} and \icesheet{} experiments.
+Indeed, the more independent relationship between model outputs and water discharge in \alpine{} is smaller compared to \icesheet{} due to the relatively larger discharge variations on daily to weekly timescales.
 This could occur as the evolution of the R-channel's size is small compared to the variations in water discharge.
@@
 In turn, a closer approximation of the hydraulics could be represented by pipe flow, where the channel size remains fixed.
-Conversely, the reduced variability in water discharge in \icesheet may result in the R-channel's size being closer to equilibrium with respect to the variations in water discharge.
-In this case, the exponent on water discharge with respect to shear stress is likely significantly less than $2$, but greater than $\frac{4}{5}$ that occurs when the R-channel's size evolves in equilibrium with water discharge (see \ref{sect:scaling}).
-As a result in the stronger correlation persists between subglacial model outputs and water discharge in \icesheet{} compared to \alpine (Figure~\ref{fig:Qw_vari}\,a-f).
+Conversely, the reduced variability in water discharge in \icesheet{} may result in the R-channel's size being closer to equilibrium concerning the variations in water discharge.
+In this case, the exponent on water discharge for shear stress is likely significantly less than $2$, but greater than $\frac{4}{5}$ that occurs when the R-channel's size evolves in equilibrium with water discharge (see \ref{sect:scaling}).
+As a result, a stronger correlation persists between subglacial model outputs and water discharge in \icesheet{} compared to \alpine{} (Figure~\ref{fig:Qw_vari}\,a-f).
 
@@
 In these cases, massive amounts of sediment can be evacuated from the subglacial environment, due to increased sediment access, but also the large increases in sediment transport capacity.
-If water input prior to these events has been minimal, the channel size would be small and sediment transport capacity would very rapidly increase at the onset of the event (see \ref{sect:scaling}).
+If water input before these events have been minimal, the channel size would be small and sediment transport capacity would very rapidly increase at the onset of the event (see \ref{sect:scaling}).
 
@@
 
-Water discharge measurement at the sub-hour timescale remain unavailable in many catchments and daily values might be used, for instance.
-The variability in subglacial channels approaches that of subaerial ones  over $1-5$ days (Figure~\ref{fig:multi_run}), possibly resulting in a parity between subaerial and subglacial channels over these periods.
-This suggests that aggregating water discharge over daily to weekly time periods could yield similar behavior between subglacial and subaerial systems. 
-At the same time, water discharge over such time periods, may omit key process, such as water velocity reaching sediments' threshold of motion more frequently or the timing of peak events.
+Water discharge measurements at the sub-hour timescale remain unavailable in many catchments and daily values might be used, for instance.
+The variability in subglacial channels approaches that of subaerial ones  over $1-5$ days (Figure~\ref{fig:multi_run}), possibly resulting in parity between subaerial and subglacial channels over these periods.
+This suggests that aggregating water discharge over daily to weekly periods could yield similar behavior between subglacial and subaerial systems. 
+At the same time, water discharge over such periods may omit key processes, such as water velocity reaching sediments' threshold of motion more frequently or the timing of peak events.
 
@@
 In contrast, pressurized subglacial channels largely accommodate rapidly changing water discharge by altering water velocity.
-Their size responds over the course of several days to changes in discharge.
+Their size responds over several days to changes in discharge.
 Thus the response of sediment transport capacity is much more sensitive to changes in water discharge in subglacial channels compared to subaerial ones.
@@
 
-\item the degree of difference in sediment transport capacity variations between the two channels types depends on the water discharge variability in the hydrograph, which varies with geography and across timescales.
-  However, representing subglacial sediment transport capacity with either open-channel flow or pipe flow omits key subglacial sediment transport characteristics.
+\item the degree of difference in sediment transport capacity variations between the two channel types depends on the water discharge variability in the hydrograph, which varies with geography and across timescales.
+  However, representing subglacial sediment transport capacity with either open-channel flow or pipe flow omits key subglacial sediment transport characteristics;
   \vspace{.1cm}
 
-  \item the uneven response of subglacial sediment transport capacity to water discharge variations suggests different magnitude-frequency relationships occur in most subglacial channels compared to subaerial ones with similar hydrological forcings.
+  \item the uneven response of subglacial sediment transport capacity to water discharge variations shows a different response to periods of elevated water discharge in subglacial channels compared to subaerial ones.
 \end{enumerate}
@@
 A. Tedstone acknowledges funding from the European Research Council, award $818994$ -- CASSANDRA.
-G. King provided useful comments on a previous version of this manuscript.
+G. King and M.J. Gevers provided useful comments on a previous version of this manuscript.
 
@@
     Name &Symbol&  Value&Units \\
-    && (\alpine or \icesheet)\\
+    && (\alpine{} or \icesheet{})\\
     \hline
@@
 The  subaerial channel is assumed to have a width $w$ much greater than its depth $H$, such that $D_h\approx 4h$, and to have a constant head gradient ($\Psi$) given by the topography.
-Further, it is  assumed that its width can be approximated by a relation $w \propto Q^\alpha$ (Equation~\ref{eq:wcf})with $\alpha\in [0,1]$; typically $\alpha \approx 1/2$ (a so-called regime channel).
+Further, it is  assumed that its width can be approximated by a relation $w \propto Q^\alpha$ (Equation~\ref{eq:wcf}) with $\alpha\in [0,1]$; typically $\alpha \approx 1/2$ (a so-called regime channel).
 End members $\alpha=0$ or $\alpha=1$ correspond a  subaerial channels of constant width (a slot canyon) or depth (no natural equivalent), respectively.
@@
 In this case, it is assumed that the cross-sectional area $S$ is fixed, and $\Psi$ adjusts to the specified $Q$.
-For both R-channel and pipe flow, it is assumed assume that $D_h \propto S^{1/2}$.
+For both R-channel and pipe flow, it is assumed that $D_h \propto S^{1/2}$.
 
@@
 In such cases they operate more like a pipe of fixed cross-section \cite<e.g.>{gimbert2016}.
-Table~\ref{tab:Qs2} shows that in such a situation, the sediment transport scales much more severely with discharge, with the exponent on $Q$ being likely between $3$ and $5$, compared to other two regimes when that exponent is at most $1.3$ (Table~\ref{tab:Qs2}).
+Table~\ref{tab:Qs2} shows that in such a situation, the sediment transport scales much more severely with discharge, with the exponent on $Q$ being between $3$ and $5$, compared to the other two regimes when that exponent is at most $1.7$ (Table~\ref{tab:Qs2}).
 Thus fluctuations of discharge on short timescales (on the order of a day) have the potential to cause conditions with very high sediment transport capacities.

### 2024-03-20 Delaneyetal2024.tex
progress on discussion

@@
 
-%\ian{only glaciers with subglacial channels... }
-
 \begin{keypoints}
@@
   Sediment transport capacity in both subaerial and subglacial channels depends on the shear stress exerted across the channel bottom, which varies with the water velocity and channel width.
-  In subaerial channels, water discharge variations are accommodated largely by changes in flow depth and width.
-  While in subglacial channels, water discharge changes primarily affect water velocity due to evolving channel geometry over hours to weeks.
+  In subaerial channels, water discharge variations are accommodated largely by flow depth and width changes.
+  In subglacial channels, water discharge changes primarily affect water velocity due to evolving channel geometry over hours to weeks.
   Hydrographs from an Alpine glacier and the Greenland Ice Sheet are used to drive models of subglacial and subaerial hydraulics, revealing greater variability in the inferred sediment transport capacity in subglacial than in subaerial channels.
@@
 \section*{Plain Language Summary} % 200 words
-Sediment export from glaciers impacts downstream ecosystems and communities and it has been documented to change in response to climate warming and glacier retreat.
-In these catchments, sediment export depends in part on the sediment transport capacity, the amount of sediment that could be transported for the hydraulic conditions given the abundance of sediment.
+Sediment export from glaciers impacts downstream ecosystems and communities and has changed in response to climate warming and glacier retreat.
+Sediment export depends in part on the sediment transport capacity, the amount of sediment that could be transported for the hydraulic conditions with abundant sediment availability.
 This capacity largely depends on the width of the channel and the water's velocity.
-Unlike subaerial channels, like rivers, where the size of the river changes rapidly with water discharge, however, subglacial channels are pressurized by the ice above.
-This pressurized flow means that channel size evolves slowly compared to the variations in water discharge, therefore variations in water discharge are largely accommodated by variations in velocity as opposed to channel size.
+In subaerial channels, like rivers, water velocity and channel width both respond rapidly to water discharge.
+However, subglacial channels are pressurized by the ice above, so that the water flows through a subglacial pipe that evolves in size.
+This pressurized flow means that channel size evolves slowly compared to variations in water discharge, therefore these variations are largely accommodated by changes in water velocity as opposed to channel size.
 Models of these processes show that sediment transport capacity is highly variable in subglacial channels, and a wide range of water discharges can result in comparable sediment transport capacities.
-In subaerial channels, by contrast, sediment transport capacity is less variable and depends solely on water discharge.
-The channel's different responses to water discharge may impact the interpretation of sediment transport records and respond to the type of hydrological forcing applied.
+In subaerial channels, by contrast, sediment transport capacity is less variable, and variations depends solely on water discharge, not the channel size.
+The channel's different responses to water discharge may impact the interpretation of sediment transport records from glacierized catchments, complicating the link between glacier melt and sediment discharge.
 
@@
 The Alpine site (\alpine{}) is  Fieschergletscher in the Swiss Alps ($46^\circ\,29'\,07''$ N, $8^\circ\,08'\,3''$ E).
-Water discharge was collected here at a $1$\unit{min} interval continuously over the period 2014--2021;
+The water discharge data used here was collected at a $1$\unit{min} interval from May 24, 2014 to October 10, 2014
 see \citeA{felix2022} and \citeA{felix2021} for more details.
@@
 The models described below (Sections ~\ref{sect:sub_mode}~and~\ref{sect:fluv}) represent relationships amongst water discharge, water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
-Both models use the measured discharge to calculate water velocity, shear stress, and width-integrated shear stress, upon which sediment transport depends \cite<Figure \ref{fig:cartoon}; >{shields1936}.
+Both models use the measured discharge to calculate water velocity, shear stress, and width-integrated shear stress, upon which sediment transport both suspended load and bedload depend \cite<Figure \ref{fig:cartoon}; >{shields1936}.
 The evaluation of these variables omits the selection of a sediment transport relationship and a grain-size parameter \cite<e.g.>{shields1936,meyer1948} needed to calculate sediment transport capacity from shear stress.
@@
 To evaluate the shear stress of water flowing across sediments underneath a glacier, the subglacial channel model accounts for the channel geometry and the velocity of the flowing water.
-To do this, a lumped hydraulics model \cite{clarke1996} modified from the one presented by \citeA{werder2010b} is used.
+To do this, we use a lumped hydraulics model \cite{clarke1996} modified from the one presented by \citeA{werder2010b}.
 
@@
 The results below present water velocity ($v_g$, $v_f$), shear stress ($\tau_g$, $\tau_f$), and width-integrated shear stress ($w_g\tau_g$, $w_f\tau_f$) from the subglacial and subaerial models.
-For brevity in the text these values together are referred to as ``model outputs''.
-Lastly, outputs  are compared with the water discharge from glaciers using Spearman rank correlation, dependent on the ordering of values.
+For brevity, these values together are referred to as ``model outputs'' in the text.
+Outputs  are compared to water discharge from the glaciers using Spearman rank correlation, which varies with the ordering of values.
 Rank correlation reduces the impact of the non-linear response of sediment transport capacity to the hydrological forcing, thus making our interpretation independent of the exact details of the sediment transport formula.
@@
 A second experiment aims to further characterize the variability in sediment discharge capacity in subglacial compared to subaerial channels across a range of channel shapes, friction factors, and water discharge variability.
-The variability of one model output (i.e water velocity), a key difference between the subaerial and subglacial systems, is represented by its standard deviation throughout the season.
+To accomplish this, we examine the variability of one model output (i.e water velocity) by evaluating its standard deviation throughout the season.
 
@@
 Runs with parameter combinations are accepted if their mean subglacial water velocity over the season lies between $0.75$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $2$\,\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity lies between $0.3$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}$^3$\,\unit{s}$^{-1}$, following measurements \cite<e.g.>{werder2010b,magnusson2012,chandler2013}.
-Additionally, to be accepted, model runs can only experience a flotation fraction of above $1.2$ for $2.5$\,\% of the run.
+Additionally, to be accepted, model runs can only experience a flotation fraction of above $1.2$ for less than $2.5$\,\% of the run.
 
 To establish the role of water discharge variability in driving variations in sediment transport capacity, the water discharge records are aggregated across a select period ranging from $15$\,\unit{min} up to $15$\,\unit{days}.
-Aggregation periods longer than this might be conflated with seasonal variations in water discharge and are thus omitted.
-In this case, the water discharge input at a given time is the average over the aggregation period before that time.  
+Aggregation periods longer than this  are omitted as they might be conflated with seasonal variations in water discharge.
+Here, the water discharge input at a given time is the average over the aggregation period before that time.  
 The routine runs until $100$ different parameter combinations are accepted for each water discharge aggregation period.
@@
     Relationship between water discharge and normalized velocity, shear stress, and width-integrated shear stress for \alpine{} (a-c) and \icesheet{} (d-f).
-    %These outputs' relationship with measured sediment discharge is given for \alpine{} in (g-i) and \icesheet{} (j-l).
     Variables on the x-axis have been normalized to mean values.
@@
     Plots are shown with data and outputs at $15$\unit{min} intervals.
-    %\mauro{Pixely, needs more resolution, at least 2x. The two dots in the legend are not visible.  I would add ``Measured'' to the two y-labels.}
   }
@@
 
-The second numerical experiment aims to evaluate the role of channel shape and friction factors on the findings above.
-Across the range of channel shape parameters and friction factors  examined (Section\,\ref{sect:imp}), variability in water velocity and shear stress remains higher in the subglacial system compared to the subaerial one.
+The second numerical experiment aims to evaluate the role of channel shape and friction factors on the findings above and to examine the effect of variability in water discharge.
+Across the range of channel shapes and friction factors  examined (Section\,\ref{sect:imp}), variability in water velocity and shear stress remains higher in the subglacial system compared to the subaerial one in both \alpine{} and \icesheet{}.
+This suggests that there is likely greater intermittency in sediment transport in subglacial systems compared to subaerial ones, especially in supply-limited systems where sediment could be exhausted or mobilized over diurnal discharge cycles.
 In contrast, in \alpine{}, the variability in width-integrated shear stress is comparable between the subaerial and subglacial systems, due to the highly variable water discharge, largely due to diurnal variations (Figure~\ref{fig:multi_run}).
-However, in \icesheet{}, with lower water discharge fluctuations, the variability in width-integrated shear stress remains higher in the subglacial system across the time aggregation periods.
-Together these findings suggest that sediment transport capacity variations over a melt season can approach in subaerial and subglacial systems if substantial variability persists in the hydrological forcing.
-\ian{state that these are generally higher, is t-test needed? }
+While the seasonal variability is comparable, substantial variations in sediment discharge likely persist between the two channel types from the different timing of peak discharge events.
+However, in \icesheet{}, with lower water discharge fluctuations, the variability in width-integrated shear stress remains consistently higher in the subglacial system across the time aggregation periods.
 
@@
 Results from  both \alpine{} and \icesheet{} scenarios show that the variability in subglacial  decreases with aggregations longer than approximately 1--5 days (Figure~\ref{fig:multi_run}).
-Yet, substantial differences in the variability of model outputs for subglacial and subaerial cases persist in periods of up to $15$ days.
-Note that lags and time of peak sediment transport conditions still occur over these periods, and these characteristics are thus not considered in aggregating outputs and water discharge over longer periods (Figure~\ref{fig:model_outs}).
+Yet, substantial differences in the variability between the model outputs for subglacial and subaerial cases persist in periods of up to $15$ days.
+Note again that the timing of peak sediment transport conditions still occur across these periods, but  are thus not captured in aggregating water discharge over longer periods (Figure~\ref{fig:model_outs}).
 
 \section{Discussion}
-\subsection{Sediment capacity differences between the channel types}
+\subsection{Hydrology and evolving channel shape in driving sediment transport variability}
 \label{sect:dis_qsc}
 
-Quantities that relate to sediment transport capacity (water velocity, shear stress, and width-integrated shear stress) vary differently in response to water discharge in subglacial channels compared to subaerial ones (Figure~\ref{fig:Qw_vari}).
+Model outputs (water velocity, shear stress, and width-integrated shear stress) vary differently in response to water discharge in subglacial channels compared to subaerial ones (Figure~\ref{fig:Qw_vari}).
 In sediment transport relationships that evaluate sediment discharge capacity, such as in \citeA{meyer1948} or \citeA{engelund1967}, shear stress is scaled to the power of $\frac{3}{2}$ or $\frac{5}{2}$, respectively.
@@
 
-In R-channel models, viable water pressure regimes can be achieved by changing the shape and friction factor of the channels, enabling assumptions of semi-circular channels in most applications \cite{werder2010b}.
-However, the role of channel width in the findings here suggests that more thorough parameterization of channel shape and channel width is needed in assessing subglacial sediment transport capacity (Section\ref{sect:sub_mode}, \ref{sect:scaling}).
-\ian{Similar feedback occurs in subaerial channels \cite{phillips2016}, albeit at  timescales longer than days in subglacial channels. }
-
 Sediment transport capacity in subglacial channels responds more rapidly to changes in water discharge compared to their subaerial counterparts \cite{alley1997}.
-In subglacial R\"othlisberger channels in steady state with respect to water discharge, the width-integrated shear stress response to water discharge is $ w\tau_{tg}\, \propto\,  Q^{\sim\frac{4}{5}}$ (Table\,\ref{tab:eqs1}).
+In subglacial R\"othlisberger channels in steady state with respect to water discharge, the width-integrated shear stress response to water discharge is $ w_g\,\tau_{g}\, \propto\,  Q^{\sim\frac{4}{5}}$ (Table\,\ref{tab:eqs1}).
 However, subglacial channels rarely operate in a steady state \cite<e.g.>{gimbert2016}.
-For sufficiently short periods, channel width and wetted area are fixed in the subglacial system (akin to pipe flow) so that the response of width-integrated shear stress is $ \tau_{tg}\, \propto\,  Q^{2}$ (Table\,\ref{tab:eqs1}).
+For sufficiently short periods, channel width and wetted area are fixed in the subglacial system (akin to pipe flow) so that the response of width-integrated shear stress is $w_g\, \tau_{g}\, \propto\,  Q^{2}$ (Table\,\ref{tab:eqs1}).
 \ref{sect:scaling} provides a complete description of sediment transport capacity in different drainage conditions.
-
-The co-varying relationship between sediment discharge capacity in  subaerial conditions could be pronounced $\sim20$\,\unit{km} downstream of the Leverett site, where a strong correlation persists between sediment plume size and the river's water discharge into the Kangerlussuaq fjord \cite{mcgrath2010}.
-In contrast, in marine-terminating glacier catchments a less consistent relationship may occur between water discharge or melt extent and sediment plume size \cite{tedstone2012}.
-Here, ocean water pressurizes subglacial channels at the ice front \cite<e.g.>{how2017}, so the minimal correlation could result from the inconsistent relationship between subglacial sediment transport capacity and water discharge (Figure~\ref{fig:Qw_vari}).
-
-The stronger link between sediment transport capacity and hydrological forcing in subaerial channels may mean that variations in sediment transport records can be more easily attributed to climatic conditions than in subglacial systems.
-Over time scales of weeks or longer, the glacier's sediment transport capacity is impacted by the ice thickness, controlling the channel closure rate, and the glacier's surface slope, in addition to water discharge \cite<Figure~\ref{fig:multi_run}, Section~\ref{sect:sub_mode}; >{rothlisberger1972,shreve1972,gimbert2016,stevens2022}.
-In a supply-limited state, often assumed in landscape evolution models, then glaciers' sediment discharge record are also impacted by the additional processes of sediment availability and bedrock erosion from  water pressure variations and sliding  \cite<e.g.>{iverson2012,herman2015,delaney2023}.
-This multitude of processes lies in contrast to processes occurring in most subaerial channels, where transport capacity typically responds to sediment size, water discharge, and hydraulic gradient, and the hydraulic gradient generally remains stable over time scales ranging from years to millennia \cite<Section~\ref{sect:fluv}; e.g.>{muller1968,tucker1997,wickert2019}.
-
-\subsection{Role of hydrological forcing}
-
-%\mauro{You often use ``subglacial hydraulics''.
-I'm not sure what that means.
-For instance in the next sentence.
-Isn't discharge part of hydraulics too?
-Maybe you could define a bunch of equations as the subglacial hydraulic model and subaerial hydraulic model and then talk of that?}
-
-%\ian{make comment about lumped nature of model}
+Note that in pipe flow conditions, maximum sediment transport would coincide with peak water discharge and a co-varying relationship with water discharge would occur, unlike the case here with evolving channel size (Figure \ref{fig:model_outs} and \ref{fig:Qw_vari}).
+However, similar adjustments in channel shape occur in subaerial channels \cite{phillips2016}, albeit at  timescales longer than hours to days in subglacial channels.
 
@@
 This could occur as the evolution of the R-channel's size is small compared to the variations in water discharge.
-Such variability in alpine catchments may cause shear stress to be proportional to the water discharge squared \cite< \ref{sect:scaling}; c.f.>{alley1997}.
+Such water discharge variability in alpine catchments may cause shear stress to approach the water discharge squared \cite< \ref{sect:scaling}; c.f.>{alley1997}.
 In turn, a closer approximation of the hydraulics could be represented by pipe flow, where the channel size remains fixed.
@@
 In this case, the exponent on water discharge for shear stress is likely significantly less than $2$, but greater than $\frac{4}{5}$ that occurs when the R-channel's size evolves in equilibrium with water discharge (see \ref{sect:scaling}).
-As a result, a stronger correlation persists between subglacial model outputs and water discharge in \icesheet{} compared to \alpine{} (Figure~\ref{fig:Qw_vari}\,a-f).
+As a result, a stronger correlation persists between subglacial model outputs and water discharge in \icesheet{} compared to \alpine{} (Figure~\ref{fig:Qw_vari}).
 
@@
 
-The greater variations in subglacial sediment transport capacity in both catchment types compared to subaerial ones could cause a supply-limited regime to persist at many glaciers.
+\subsection{Implications for parameterizing  sediment transport in glacierized catchments}
+
+\ian{comeback here}
+
+Numerical models of subglacial sediment transport generally apply classical parameterizations of subglacial hydrology to evaluating sediment transport conditions \cite<e.g>{hewitt2019,beaud2018}.
+In these R-channel models, viable water pressure regimes can be achieved by changing the shape and friction factor of the channels, enabling assumptions of semi-circular channels in most applications \cite{werder2010b}.
+However, the role of channel width in the findings here suggests that more thorough parameterization of channel shape and channel width is needed in assessing subglacial sediment transport capacity as compared to evaluating water pressure (Section\,\ref{sect:sub_mode}, \ref{sect:scaling}).
+
+In applying these models to different climate scenarios, water discharge measurements at the sub-hour timescale remain unavailable in many catchments and daily values might be used, for instance.
+The variability in subglacial channels approaches that of subaerial ones  over $1-5$ days (Figure~\ref{fig:multi_run}), possibly resulting in parity in variability between subaerial and subglacial channels over these periods.
+This suggests that aggregating water discharge over daily to weekly periods could yield similar variations between subglacial and subaerial systems, although generally steeper hydraulic gradients of glaciers may cause elevated sediment transport capacity there (Figure\,\ref{fig:model_outs}).
+At the same time, water discharge over periods longer than days may omit key processes, such as the maximum sediment transport capacity occurring with increased water discharge, as opposed to the maximum quantity (Figure~\ref{fig:model_outs}).
+
+\subsection{Effect sediment transport and landscape evolution}
+
+The stronger link between sediment transport capacity and water discharge in subaerial channels may mean that variations in sediment transport records can be more easily attributed to hydro-climatic conditions than in subglacial systems.
+Over time scales of weeks or longer, the glacier's sediment transport capacity is impacted by the ice thickness, controlling the channel closure rate, and the glacier's surface slope, in addition to water discharge \cite<Figure~\ref{fig:multi_run}, Section~\ref{sect:sub_mode}; >{rothlisberger1972,shreve1972,gimbert2016,stevens2022}.
+In a supply-limited state, then glaciers' sediment discharge record also responds to  sediment availability and bedrock erosion from  water pressure variations and sliding  \cite<e.g.>{iverson2012,herman2015,delaney2023}.
+This multitude of processes lies in contrast to processes occurring in most subaerial channels, where transport capacity typically responds to sediment size, channel shape, water discharge, and hydraulic gradient.
+The hydraulic gradient and channel shape can remain stable over years to millennia \cite<Section~\ref{sect:fluv}; e.g.>{tucker1997,wickert2019}.
+
+This co-varying relationship between sediment discharge capacity in  subaerial conditions could be pronounced $\sim20$\,\unit{km} downstream of the Leverett site, where a strong correlation persists between sediment plume size and the river's water discharge into the Kangerlussuaq fjord \cite{mcgrath2010}.
+In contrast, in marine-terminating glacier catchments, a less consistent relationship may occur between water discharge or melt extent and sediment plume size \cite{tedstone2012}.
+Here, ocean water pressurizes subglacial channels at the ice front \cite<e.g.>{how2017}, so the reduced correlation could result from the inconsistent relationship between subglacial sediment transport capacity and water discharge (e.g.
+Figure~\ref{fig:Qw_vari}).
+
+Furthermore, the greater variations in subglacial sediment transport capacity in both catchment types compared to subaerial ones could cause a supply-limited regime to persist at many glaciers.
 This regime may persist as the sediment's threshold of motion can be reached more frequently and across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
 Sediment exhaustion occurring through increased variability may also explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \cite{overeem2017}.
-
-Water discharge measurements at the sub-hour timescale remain unavailable in many catchments and daily values might be used, for instance.
-The variability in subglacial channels approaches that of subaerial ones  over $1-5$ days (Figure~\ref{fig:multi_run}), possibly resulting in parity between subaerial and subglacial channels over these periods.
-This suggests that aggregating water discharge over daily to weekly periods could yield similar behavior between subglacial and subaerial systems. 
-At the same time, water discharge over such periods may omit key processes, such as water velocity reaching sediments' threshold of motion more frequently or the timing of peak events.
 
@@
 
-\item the timing of peak sediment transport capacity below glaciers generally coincides with large rates of increase in water discharge, as opposed to the maximum water discharge in subaerial channels;
+\item the timing of peak sediment transport capacity below glaciers generally coincides with large rates of increase in water discharge, as opposed to the maximum water discharge in subaerial channels.
+  However, representing subglacial sediment transport capacity with either open-channel flow or pipe flow results in coincident variations; 
   \vspace{.1cm}
 
-\item the degree of difference in sediment transport capacity variations between the two channel types depends on the water discharge variability in the hydrograph, which varies with geography and across timescales.
-  However, representing subglacial sediment transport capacity with either open-channel flow or pipe flow omits key subglacial sediment transport characteristics;
+\item the degree of difference in sediment transport capacity variations between the two channel types depends on the water discharge variability in the hydrograph, which varies with catchment characteristics and water discharge evolution;
   \vspace{.1cm}
@@
 A. Tedstone acknowledges funding from the European Research Council, award $818994$ -- CASSANDRA.
-G. King and M.J. Gevers provided useful comments on a previous version of this manuscript.
+G. King and M. J. Gevers provided useful comments on a previous version of this manuscript.
 
@@
 \appendix
+\TODO{match Friction factor in appendix}
 \section{Table of variables, parameters and constants}
@@
     Hooke angle of channel & $\beta$ & $\frac{\pi}{6}$ & \unit{rad}\\
-    Friction factor (subglacial, subaerial) & ($f_r$, $f_f$) &$5,\,16$ or $3$ & $\mathrm{(-)}$ \\
+    Friction factor (subglacial, subaerial) & $f$, ($f_r$, $f_f$) &$5,\,16$ or $3$ & $\mathrm{(-)}$ \\
     Glacier thickness &$h_{ice}$& $225$ or $740$  &\unit{m}\\

### 2024-03-21 Delaneyetal2024.tex
edits here.. push to overleaf

@@
 
+\ian{add comment about glacier geometry}
 Here, it is assumed that the water is transported through a subglacial channel \cite<Figure~\ref{fig:cartoon}; >{rothlisberger1972}, and that the channel  size responds to melt due to frictional heating from water flow and to creep closure of the ice, using the Darcy-Weisbach formulation for water-flow through a pipe  \cite<e.g.>{rothlisberger1972,clarke2003}.

### 2024-03-21 Delaneyetal2024.tex
back to mac

--- 2024-03-20 Delaneyetal2024.tex
@@
 \begin{keypoints}
-\item Short term water discharge variations in subglacial channels are mainly accommodated by water velocity, not channel size, as is the case in subaerial systems.
+\item Short-term water discharge variations in subglacial channels are mainly accommodated by water velocity, not channel size, as is the case in subaerial systems.
 \item Greater variability in water velocity causes sediment transport capacity to vary more in subglacial channels than in subaerial channels.
@@
   In subglacial channels, water discharge changes primarily affect water velocity due to evolving channel geometry over hours to weeks.
-  Hydrographs from an Alpine glacier and the Greenland Ice Sheet are used to drive models of subglacial and subaerial hydraulics, revealing greater variability in the inferred sediment transport capacity in subglacial than in subaerial channels.
-  Results also show that high subglacial sediment transport rates and water velocities can occur across a wide range of water discharges.
+  Hydrographs from an Alpine glacier and the Greenland Ice Sheet are used to drive subglacial and subaerial hydraulics models, revealing greater variability in the inferred sediment transport capacity in subglacial than in subaerial channels.
+  Results also show that high subglacial sediment transport rates and water velocities can occur across wide ranging water discharges.
   These findings are discussed in the context of understanding sediment discharge from glaciers with different hydrological forcings and assumptions of channel type.
@@
 Sediment export from glaciers impacts downstream ecosystems and communities and has changed in response to climate warming and glacier retreat.
-Sediment export depends in part on the sediment transport capacity, the amount of sediment that could be transported for the hydraulic conditions with abundant sediment availability.
+Sediment export depends in part on the sediment transport capacity, and the amount of sediment that could be transported for the hydraulic conditions with abundant sediment availability.
 This capacity largely depends on the width of the channel and the water's velocity.
-In subaerial channels, like rivers, water velocity and channel width both respond rapidly to water discharge.
-However, subglacial channels are pressurized by the ice above, so that the water flows through a subglacial pipe that evolves in size.
+In subaerial channels, like rivers, water velocity, and channel width both respond rapidly to water discharge.
+However, subglacial channels are pressurized by the ice above, so water flows through a subglacial pipe that evolves in size.
 This pressurized flow means that channel size evolves slowly compared to variations in water discharge, therefore these variations are largely accommodated by changes in water velocity as opposed to channel size.
 Models of these processes show that sediment transport capacity is highly variable in subglacial channels, and a wide range of water discharges can result in comparable sediment transport capacities.
-In subaerial channels, by contrast, sediment transport capacity is less variable, and variations depends solely on water discharge, not the channel size.
+In subaerial channels, by contrast, sediment transport capacity is less variable, and variations depend solely on water discharge, not the channel size.
 The channel's different responses to water discharge may impact the interpretation of sediment transport records from glacierized catchments, complicating the link between glacier melt and sediment discharge.
@@
 
-Here, it is assumed that the water is transported through a subglacial channel \cite<Figure~\ref{fig:cartoon}; >{rothlisberger1972}, and that the channel  size responds to melt due to frictional heating from water flow and to creep closure of the ice, using the Darcy-Weisbach formulation for water-flow through a pipe  \cite<e.g.>{rothlisberger1972,clarke2003}.
+Here, it is assumed that the water is transported through a subglacial channel \cite<Figure~\ref{fig:cartoon}; >{rothlisberger1972} beneath a glacier with length $l$, a flat bed and thickness $h_{ice}$.
+The channel  size responds to melt due to frictional heating from water flow and to creep closure of the ice, using the Darcy-Weisbach formulation for water-flow through a pipe  \cite<e.g.>{rothlisberger1972,clarke2003}.
 The formulation here does not consider the englacial storage of water.
@@
     Data are shown at $15$\unit{min} intervals, i.e. not aggregated.
+    Arrows denote where variables peak in the subglacial channel before the subaerial one.
+    Insets in e-h show the  peak melt event \icesheet{} denoted by the shaded area in panel e.
     % Note that the x and y axes are different for \alpine{} and \icesheet{}.
@@
 
-Examining the relationship between water discharge and the model outputs, different relationship emerges between the subaerial and subglacial cases.
+Examining the relationship between water discharge and the model outputs, a different relationship emerges between the subaerial and subglacial cases.
 Because there is no history dependence in  subaerial channels, each water discharge value in the subaerial channel results in a unique water velocity, shear stress, and width-integrated shear stress (Section~\ref{sect:sub_mode}).
@@
 Yet, substantial differences in the variability between the model outputs for subglacial and subaerial cases persist in periods of up to $15$ days.
-Note again that the timing of peak sediment transport conditions still occur across these periods, but  are thus not captured in aggregating water discharge over longer periods (Figure~\ref{fig:model_outs}).
+Note again that the timing of peak sediment transport conditions still occurs across these periods, and they are are thus not captured in aggregating water discharge over longer periods (Figure~\ref{fig:model_outs}).
 
@@
 Sediment transport capacity in subglacial channels responds more rapidly to changes in water discharge compared to their subaerial counterparts \cite{alley1997}.
-In subglacial R\"othlisberger channels in steady state with respect to water discharge, the width-integrated shear stress response to water discharge is $ w_g\,\tau_{g}\, \propto\,  Q^{\sim\frac{4}{5}}$ (Table\,\ref{tab:eqs1}).
+In subglacial R\"othlisberger channels in steady state with water discharge, the width-integrated shear stress response to water discharge is $ w_g\,\tau_{g}\, \propto\,  Q^{\sim\frac{4}{5}}$ (Table\,\ref{tab:eqs1}).
 However, subglacial channels rarely operate in a steady state \cite<e.g.>{gimbert2016}.
@@
 
-Our results point to the highly variable response of proxies for sediment transport capacity to  water discharge quantity, yet substantial differences emerge between the \alpine{} and \icesheet{} experiments.
-Indeed, the more independent relationship between model outputs and water discharge in \alpine{} is smaller compared to \icesheet{} due to the relatively larger discharge variations on daily to weekly timescales.
-This could occur as the evolution of the R-channel's size is small compared to the variations in water discharge.
+ Substantial differences emerge in model outputs between the different hydrological forcings of the \alpine{} and \icesheet{} cases.
+Indeed, the more independent relationship between model outputs and water discharge in \alpine{} results from the
+relatively larger discharge variations on daily to weekly timescales compared to \icesheet{}.
+This occurs as the evolution of the R-channel's size is small compared to the variations in water discharge.
 Such water discharge variability in alpine catchments may cause shear stress to approach the water discharge squared \cite< \ref{sect:scaling}; c.f.>{alley1997}.
-In turn, a closer approximation of the hydraulics could be represented by pipe flow, where the channel size remains fixed.
+With respect to \icesheet{}, a closer approximation of the hydraulics could be represented by pipe flow here, where the channel size remains fixed.
 Conversely, the reduced variability in water discharge in \icesheet{} may result in the R-channel's size being closer to equilibrium concerning the variations in water discharge.
-In this case, the exponent on water discharge for shear stress is likely significantly less than $2$, but greater than $\frac{4}{5}$ that occurs when the R-channel's size evolves in equilibrium with water discharge (see \ref{sect:scaling}).
+In this case, the exponent on water discharge for shear stress is likely less than $2$, but greater than $\frac{4}{5}$ that occurs when the R-channel's size evolves in equilibrium with water discharge (see \ref{sect:scaling}).
 As a result, a stronger correlation persists between subglacial model outputs and water discharge in \icesheet{} compared to \alpine{} (Figure~\ref{fig:Qw_vari}).
 
-The hydrological forcing that could cause the greatest increases in subglacial sediment transport capacity may occur during glacier lake drainages or intense precipitation events \cite<e.g.>{cowan1988,delaney2018}.
+Very large quantities of sediment are known to be evacuated during glacier lake drainages or intense precipitation events \cite<e.g.>{cowan1988,delaney2018}.
 In these cases, massive amounts of sediment can be evacuated from the subglacial environment, due to increased sediment access, but also the large increases in sediment transport capacity.
 If water input before these events have been minimal, the channel size would be small and sediment transport capacity would very rapidly increase at the onset of the event (see \ref{sect:scaling}).
-
-\subsection{Implications for parameterizing  sediment transport in glacierized catchments}
-
-\ian{comeback here}
-
-Numerical models of subglacial sediment transport generally apply classical parameterizations of subglacial hydrology to evaluating sediment transport conditions \cite<e.g>{hewitt2019,beaud2018}.
-In these R-channel models, viable water pressure regimes can be achieved by changing the shape and friction factor of the channels, enabling assumptions of semi-circular channels in most applications \cite{werder2010b}.
-However, the role of channel width in the findings here suggests that more thorough parameterization of channel shape and channel width is needed in assessing subglacial sediment transport capacity as compared to evaluating water pressure (Section\,\ref{sect:sub_mode}, \ref{sect:scaling}).
-
-In applying these models to different climate scenarios, water discharge measurements at the sub-hour timescale remain unavailable in many catchments and daily values might be used, for instance.
-The variability in subglacial channels approaches that of subaerial ones  over $1-5$ days (Figure~\ref{fig:multi_run}), possibly resulting in parity in variability between subaerial and subglacial channels over these periods.
-This suggests that aggregating water discharge over daily to weekly periods could yield similar variations between subglacial and subaerial systems, although generally steeper hydraulic gradients of glaciers may cause elevated sediment transport capacity there (Figure\,\ref{fig:model_outs}).
-At the same time, water discharge over periods longer than days may omit key processes, such as the maximum sediment transport capacity occurring with increased water discharge, as opposed to the maximum quantity (Figure~\ref{fig:model_outs}).
 
@@
 Sediment exhaustion occurring through increased variability may also explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \cite{overeem2017}.
+
+\subsection{Implications for parameterizing  sediment transport in glacierized catchments}
+
+Numerical models of subglacial sediment transport generally apply classical parameterizations of subglacial hydrology to evaluating sediment transport conditions \cite<e.g>{hewitt2019,beaud2018}.
+In these R-channel models, viable water pressure regimes can be achieved by changing the shape and friction factor of the channels, enabling assumptions of semi-circular channels in most applications \cite{werder2010b}.
+However, the role of channel width in the findings here suggests that more thorough parameterization of channel shape and channel width is needed in assessing subglacial sediment transport capacity as compared to evaluating water pressure (Section\,\ref{sect:sub_mode}, \ref{sect:scaling}).
+
+In applying these models to different climate scenarios, water discharge measurements at the sub-hour timescale remain unavailable in many catchments and daily values might be used, for instance.
+The variability in subglacial channels approaches that of subaerial ones  over $1-5$ days (Figure~\ref{fig:multi_run}), possibly resulting in parity in variability between subaerial and subglacial channels over these periods.
+This suggests that aggregating water discharge over daily to weekly periods could yield similar variations between subglacial and subaerial systems, although generally steeper hydraulic gradients of glaciers may cause elevated sediment transport capacity there (Figure\,\ref{fig:model_outs}).
+At the same time, water discharge over periods longer than days may omit key processes, such as the maximum sediment transport capacity occurring with increased water discharge, as opposed to the maximum quantity (Figure~\ref{fig:model_outs}).
 

### 2024-03-21 Delaneyetal2024.tex
small changes

@@
 \section*{Plain Language Summary} % 200 words
-Sediment export from glaciers impacts downstream ecosystems and communities and has changed in response to climate warming and glacier retreat.
-Sediment export depends in part on the sediment transport capacity, and the amount of sediment that could be transported for the hydraulic conditions with abundant sediment availability.
+Sediment export from glaciers impacts downstream ecosystems and communities and has changed in response to climate warming.
+Sediment export depends partly on the sediment transport capacity, and the amount of sediment that could be transported for the hydraulic conditions with abundant sediment availability.
 This capacity largely depends on the width of the channel and the water's velocity.
@@
     \caption{Sketch for the different short time-scale responses of subglacial and subaerial channels to increased water discharge.
-      Water velocity magnitudes in the subglacial, $v_g$, and subaerial, $v_f$, channels are shown by arrow length.
-      $S_g$ and $S_f$ represent the wetted area in subglacial and subaerial channels, respectively.
-      The subglacial channel width $w_g$ remain unchanged, while the subaerial channel width $w_f$ evolves. Shear stress $\tau$ is responsible for the mobilization of sediment.
+      Arrow length denotes Water velocity magnitudes in the subglacial, $v_g$, and subaerial, $v_f$, channels.
+      $S_g$ ($S_f$) represent the wetted area in subglacial (subaerial) channels.
+      The subglacial channel width $w_g$ remain unchanged, while the subaerial channel width $w_f$ evolves.
+      Shear stress $\tau$ is responsible for the mobilization of sediment.
     }
@@
 
-Water discharge data is used from Alpine and ice sheet settings collected downstream of the glacier, assuming that no water storage occurred in the proglacial area.
+Water discharge data is used from Alpine and ice sheet settings collected downstream of the glacier, assuming no water storage in the proglacial area.
 The Alpine site (\alpine{}) is  Fieschergletscher in the Swiss Alps ($46^\circ\,29'\,07''$ N, $8^\circ\,08'\,3''$ E).
@@
 The models described below (Sections ~\ref{sect:sub_mode}~and~\ref{sect:fluv}) represent relationships amongst water discharge, water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
-Both models use the measured discharge to calculate water velocity, shear stress, and width-integrated shear stress, upon which sediment transport both suspended load and bedload depend \cite<Figure \ref{fig:cartoon}; >{shields1936}.
+Both models use the measured discharge to calculate water velocity, shear stress, and width-integrated shear stress, upon which both suspended sediment and bedload transport depend \cite<Figure \ref{fig:cartoon}; >{shields1936}.
 The evaluation of these variables omits the selection of a sediment transport relationship and a grain-size parameter \cite<e.g.>{shields1936,meyer1948} needed to calculate sediment transport capacity from shear stress.
@@
 \icesheet{}  is exemplified by thick ice  ($h_{ice}$: $700$\,\unit{m}), high water discharge ($\sim\,300$\,\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).
-%
-However, both scenarios share a discharge with a similar, relative seasonal amplitude.
+%However, both scenarios share a discharge with a similar, relative seasonal amplitude.
 
-For each glacier, the models are first applied to a reference test case for each glacier that assumes  a subglacial channel with $\beta=\frac{\pi}{6}$, i.e.
-a segment of a quarter-circle, (Table~\ref{table:vpm}).
+For each glacier, the models are first applied to a reference test case for each glacier that assumes  a subglacial channel with $\beta=\frac{\pi}{6}$%, i.e. a segment of a quarter-circle,
+(Table~\ref{table:vpm}).
 Both friction factors $f_r$ and $f_f$ are tuned so that reasonable water velocities occur for both \alpine{} and \icesheet{} \cite<$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$>{werder2010b,chandler2013}.
@@
 Outputs  are compared to water discharge from the glaciers using Spearman rank correlation, which varies with the ordering of values.
-Rank correlation reduces the impact of the non-linear response of sediment transport capacity to the hydrological forcing, thus making our interpretation independent of the exact details of the sediment transport formula.
+Rank correlation reduces the impact of the non-linear response of sediment transport capacity to the hydrological forcing.
 
@@
 To establish the role of water discharge variability in driving variations in sediment transport capacity, the water discharge records are aggregated across a select period ranging from $15$\,\unit{min} up to $15$\,\unit{days}.
-Aggregation periods longer than this  are omitted as they might be conflated with seasonal variations in water discharge.
 Here, the water discharge input at a given time is the average over the aggregation period before that time.  
@@
     Gray (black) lines in a and e represent the subglacial channel size (water discharge).
-    Blue lines represent outputs from the subglacial channel, while orange lines represent the subaerial channel.
-    Data are shown at $15$\unit{min} intervals, i.e. not aggregated.
+    Blue (orange) lines represent outputs from the subglacial (subaerial) channel.
+    Data are shown at $15$\unit{min} intervals.
     Arrows denote where variables peak in the subglacial channel before the subaerial one.
-    Insets in e-h show the  peak melt event \icesheet{} denoted by the shaded area in panel e.
+    Insets in f-h show the  peak melt event \icesheet{} denoted by the shaded area in panel e.
     % Note that the x and y axes are different for \alpine{} and \icesheet{}.
@@
     Variables on the x-axis have been normalized to mean values.
-    $R_g$ and $R_f$ show the Spearman rank correlations for the subglacial and the subaerial outputs, respectively.
-    Plots are shown with data and outputs at $15$\unit{min} intervals.
+    $R_g$ ($R_f$) show the Spearman rank correlations for the subglacial (subaerial) outputs.
+    Plots are shown with discharge and outputs at $15$\unit{min} intervals.
   }
@@
     \includegraphics[width=0.9\linewidth]{Fig4.png}
-    \caption{Variability (standard deviation) of model outputs for different time aggregations across the range of subglacial and subaerial channel shapes and friction factors.
-      Shaded areas denote the range of results from the accepted $100$ parameter combinations tested (Section~\ref{sect:imp}).
+    \caption{Standard deviation of model outputs for different time aggregations with different subglacial and subaerial channel shapes and friction factors.
+      Shaded areas denote the range of results from the accepted $100$ parameter combinations (Section~\ref{sect:imp}).
       Solid lines denote  mean values.
-      %Top row shows results from the \alpine{} scenario, while bottom row show outputs from the \icesheet{} scenario
-      Time intervals shown with markers denote $15$\,\unit{min}, $1$\,\unit{hr}, $12$\,\unit{hr}, $1$\,\unit{d}, $5$\,\unit{d}, $10$\,\unit{d}, and $15$\,\unit{d} aggregation periods.
+      Markers show aggregation periods ($15$\,\unit{min}, $1$\,\unit{hr}, $12$\,\unit{hr}, $1$\,\unit{d}, $5$\,\unit{d}, $10$\,\unit{d}, and $15$\,\unit{d}).
        }
@@
 In contrast, in \alpine{}, the variability in width-integrated shear stress is comparable between the subaerial and subglacial systems, due to the highly variable water discharge, largely due to diurnal variations (Figure~\ref{fig:multi_run}).
-While the seasonal variability is comparable, substantial variations in sediment discharge likely persist between the two channel types from the different timing of peak discharge events.
 However, in \icesheet{}, with lower water discharge fluctuations, the variability in width-integrated shear stress remains consistently higher in the subglacial system across the time aggregation periods.
 
-Evaluating the different variability in sediment transport characteristics across timescales can help inform the sediment transport processes omitted from applying daily or weekly values of water discharge to understand sediment transport.
 Results from  both \alpine{} and \icesheet{} scenarios show that the variability in subglacial  decreases with aggregations longer than approximately 1--5 days (Figure~\ref{fig:multi_run}).
-Yet, substantial differences in the variability between the model outputs for subglacial and subaerial cases persist in periods of up to $15$ days.
-Note again that the timing of peak sediment transport conditions still occurs across these periods, and they are are thus not captured in aggregating water discharge over longer periods (Figure~\ref{fig:model_outs}).
+Yet, substantial differences in the variability between the model outputs for subglacial and subaerial cases persist in periods of up to $15$ days, after which seasonal variations might be conflated.
+Note that the timing of peak sediment transport conditions still occurs across these periods, and they are are thus not captured in aggregating water discharge over longer periods (Figure~\ref{fig:model_outs}).
 
 \section{Discussion}
+
 \subsection{Hydrology and evolving channel shape in driving sediment transport variability}
@@
 \ref{sect:scaling} provides a complete description of sediment transport capacity in different drainage conditions.
-Note that in pipe flow conditions, maximum sediment transport would coincide with peak water discharge and a co-varying relationship with water discharge would occur, unlike the case here with evolving channel size (Figure \ref{fig:model_outs} and \ref{fig:Qw_vari}).
-However, similar adjustments in channel shape occur in subaerial channels \cite{phillips2016}, albeit at  timescales longer than hours to days in subglacial channels.
+Note that in pipe flow conditions, maximum sediment transport would coincide with peak water discharge and a co-varying relationship with water discharge would occur, unlike with the evolving channel size here (Section~\ref{sect:sub_mode}; Figure \ref{fig:model_outs} and \ref{fig:Qw_vari}).
+Similar adjustments in channel shape occur in subaerial channels \cite{phillips2016}, albeit at timescales longer than hours to days in subglacial channels.
 
@@
 Conversely, the reduced variability in water discharge in \icesheet{} may result in the R-channel's size being closer to equilibrium concerning the variations in water discharge.
-In this case, the exponent on water discharge for shear stress is likely less than $2$, but greater than $\frac{4}{5}$ that occurs when the R-channel's size evolves in equilibrium with water discharge (see \ref{sect:scaling}).
+In this case, the exponent on water discharge for shear stress is likely less than $2$, but greater than $\frac{4}{5}$ occurring when the R-channel's size evolves in equilibrium with water discharge (see \ref{sect:scaling}).
 As a result, a stronger correlation persists between subglacial model outputs and water discharge in \icesheet{} compared to \alpine{} (Figure~\ref{fig:Qw_vari}).
-
-Very large quantities of sediment are known to be evacuated during glacier lake drainages or intense precipitation events \cite<e.g.>{cowan1988,delaney2018}.
-In these cases, massive amounts of sediment can be evacuated from the subglacial environment, due to increased sediment access, but also the large increases in sediment transport capacity.
-If water input before these events have been minimal, the channel size would be small and sediment transport capacity would very rapidly increase at the onset of the event (see \ref{sect:scaling}).
 
@@
 The stronger link between sediment transport capacity and water discharge in subaerial channels may mean that variations in sediment transport records can be more easily attributed to hydro-climatic conditions than in subglacial systems.
-Over time scales of weeks or longer, the glacier's sediment transport capacity is impacted by the ice thickness, controlling the channel closure rate, and the glacier's surface slope, in addition to water discharge \cite<Figure~\ref{fig:multi_run}, Section~\ref{sect:sub_mode}; >{rothlisberger1972,shreve1972,gimbert2016,stevens2022}.
+A glacier's sediment transport capacity is impacted by the ice thickness controlling the channel closure rate, and the glacier's surface slope, in addition to water discharge and sediment size \cite<Figure~\ref{fig:multi_run}, Section~\ref{sect:sub_mode}; >{rothlisberger1972,shreve1972,gimbert2016,stevens2022,walder1994}.
 In a supply-limited state, then glaciers' sediment discharge record also responds to  sediment availability and bedrock erosion from  water pressure variations and sliding  \cite<e.g.>{iverson2012,herman2015,delaney2023}.
-This multitude of processes lies in contrast to processes occurring in most subaerial channels, where transport capacity typically responds to sediment size, channel shape, water discharge, and hydraulic gradient.
+This multitude of processes lies in contrast to many subaerial channels, where transport capacity typically responds to sediment size, channel shape, water discharge, and hydraulic gradient.
 The hydraulic gradient and channel shape can remain stable over years to millennia \cite<Section~\ref{sect:fluv}; e.g.>{tucker1997,wickert2019}.
@@
 Furthermore, the greater variations in subglacial sediment transport capacity in both catchment types compared to subaerial ones could cause a supply-limited regime to persist at many glaciers.
-This regime may persist as the sediment's threshold of motion can be reached more frequently and across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
-Sediment exhaustion occurring through increased variability may also explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \cite{overeem2017}.
+In subglacial systems, sediment's threshold of motion can be reached more frequently and across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
+Sediment exhaustion through this process may  explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \cite{overeem2017}.
 
-\subsection{Implications for parameterizing  sediment transport in glacierized catchments}
+\subsection{Parameterizing  sediment transport in glacierized catchments}
 
@@
 In these R-channel models, viable water pressure regimes can be achieved by changing the shape and friction factor of the channels, enabling assumptions of semi-circular channels in most applications \cite{werder2010b}.
-However, the role of channel width in the findings here suggests that more thorough parameterization of channel shape and channel width is needed in assessing subglacial sediment transport capacity as compared to evaluating water pressure (Section\,\ref{sect:sub_mode}, \ref{sect:scaling}).
+However, the role of channel width in the findings here suggests that more thorough parameterization of channel shape and channel width is needed in assessing subglacial sediment transport capacity as compared to water pressure (Section\,\ref{sect:sub_mode}, \ref{sect:scaling}).
 
-In applying these models to different climate scenarios, water discharge measurements at the sub-hour timescale remain unavailable in many catchments and daily values might be used, for instance.
-The variability in subglacial channels approaches that of subaerial ones  over $1-5$ days (Figure~\ref{fig:multi_run}), possibly resulting in parity in variability between subaerial and subglacial channels over these periods.
+In applying these models to different climate scenarios, water discharge measurements at the sub-hour timescale remain unavailable in many catchments.
+The model outputs' variability in subglacial channels approaches that of subaerial ones  over $1-5$ days (Figure~\ref{fig:multi_run}), possibly resulting in parity between subaerial and subglacial channels over these periods.
 This suggests that aggregating water discharge over daily to weekly periods could yield similar variations between subglacial and subaerial systems, although generally steeper hydraulic gradients of glaciers may cause elevated sediment transport capacity there (Figure\,\ref{fig:model_outs}).
-At the same time, water discharge over periods longer than days may omit key processes, such as the maximum sediment transport capacity occurring with increased water discharge, as opposed to the maximum quantity (Figure~\ref{fig:model_outs}).
+Yet, water discharge over periods longer than days may omit key processes, such as the time of peaks in sediment transport capacity (Figure~\ref{fig:model_outs}).
 
@@
 
-The sediment transport capacity of a channel is driven by its width and the shear stress exerted by the flowing water, which, in turn, is proportional to the square of the flow velocity.
+The sediment transport capacity of a channel is driven by its width and the shear stress exerted by the flowing water, which is proportional to the square of the flow velocity.
 Subaerial channels alter both their channel width and water velocity immediately in response to changing water discharge.
@@
 \begin{enumerate}
-\item the changing relationship between water discharge and sediment transport capacity underneath glaciers creates larger variability in sediment transport capacity compared to river systems;
+\item the changing relationship between water discharge and sediment transport capacity below glaciers creates larger variability in sediment transport capacity compared to river systems;
   \vspace{.1cm}
 
-\item the timing of peak sediment transport capacity below glaciers generally coincides with large rates of increase in water discharge, as opposed to the maximum water discharge in subaerial channels.
-  However, representing subglacial sediment transport capacity with either open-channel flow or pipe flow results in coincident variations; 
+\item  water discharge variability in the two hydrographs impacts subglacial sediment transport capacity variations;
   \vspace{.1cm}
 
-\item the degree of difference in sediment transport capacity variations between the two channel types depends on the water discharge variability in the hydrograph, which varies with catchment characteristics and water discharge evolution;
-  \vspace{.1cm}
-
-  \item the uneven response of subglacial sediment transport capacity to water discharge variations shows a different response to periods of elevated water discharge in subglacial channels compared to subaerial ones.
+\item the uneven response of sediment transport capacity to water discharge variations shows a different response and timing to periods of elevated water discharge in subglacial channels compared to subaerial ones.
+  However, representing subglacial sediment transport capacity with either open-channel flow or pipe flow neglects variations in timing.
 \end{enumerate}

### 2024-03-21 Delaneyetal2024.tex
more small changes

@@
 \begin{keypoints}
-\item Short-term water discharge variations in subglacial channels are mainly accommodated by water velocity, not channel size, as is the case in subaerial systems.
+\item Short water discharge fluctuations in subglacial channels are accommodated by water velocity, not channel size, as in subaerial systems.
 \item Greater variability in water velocity causes sediment transport capacity to vary more in subglacial channels than in subaerial channels.
-\item Results suggest that significant subglacial sediment transport capacity can persist across a range of water discharge conditions, and the timing of peak events occurs under different hydrological conditions.
+\item Significant subglacial sediment transport capacity can persist across a range of water discharges, impacting the  timing of peak events.
 \end{keypoints}
@@
 \section{Results}
-
-\subsection{Role of water discharge in sediment transport capacity}
 \begin{figure}[h]
@@
 
-In applying these models to different climate scenarios, water discharge measurements at the sub-hour timescale remain unavailable in many catchments.
-The model outputs' variability in subglacial channels approaches that of subaerial ones  over $1-5$ days (Figure~\ref{fig:multi_run}), possibly resulting in parity between subaerial and subglacial channels over these periods.
-This suggests that aggregating water discharge over daily to weekly periods could yield similar variations between subglacial and subaerial systems, although generally steeper hydraulic gradients of glaciers may cause elevated sediment transport capacity there (Figure\,\ref{fig:model_outs}).
-Yet, water discharge over periods longer than days may omit key processes, such as the time of peaks in sediment transport capacity (Figure~\ref{fig:model_outs}).
+Water discharge measurements at the sub-hour timescale remain unavailable in many catchments, further limiting the ability of subglacial  sediment transport models to capture certain events.
+The model outputs' variability in subglacial channels approaches that of subaerial ones over $1-5$ days of water aggregation time (Figure~\ref{fig:multi_run}), possibly resulting in parity in sediment transport capacity variations between subaerial and subglacial channels over these periods.
+However, generally steeper hydraulic gradients of glaciers may cause elevated sediment transport capacity in subglacial channels (Figure\,\ref{fig:model_outs}). 
+Even so, water discharge forcings aggregated over periods longer than days may omit key sediment discharge characteristics, such as the timing of peaks in sediment transport capacity (Figure~\ref{fig:model_outs}).
 
@@
 
-\item  water discharge variability in the two hydrographs impacts subglacial sediment transport capacity variations;
+\item
+  water discharge variability in the two hydrographs impacts subglacial sediment transport capacity fluctuations;
   \vspace{.1cm}
@@
 \item the uneven response of sediment transport capacity to water discharge variations shows a different response and timing to periods of elevated water discharge in subglacial channels compared to subaerial ones.
-  However, representing subglacial sediment transport capacity with either open-channel flow or pipe flow neglects variations in timing.
+  However, representing subglacial sediment transport capacity with either open-channel flow or pipe flow neglects differences in timing.
 \end{enumerate}
@@
 \appendix
-\TODO{match Friction factor in appendix}
+
 \section{Table of variables, parameters and constants}
@@
     Hydraulic gradient &$\Psi=\frac{\Delta h}{l}$&& $\mathrm{m\, m^{-1}}$\\
-
     Shear stress (subglacial, subaerial) & $\tau$, ($\tau_g,\,\tau_f$) && $\mathrm{Pa \, m^{-2}}$ \\
@@
     Hooke angle of channel & $\beta$ & $\frac{\pi}{6}$ & \unit{rad}\\
-    Friction factor (subglacial, subaerial) & $f$, ($f_r$, $f_f$) &$5,\,16$ or $3$ & $\mathrm{(-)}$ \\
+    Friction factor (subglacial, subaerial) & $f$, ($f_r$, $f_f$) &$5$ or $16$, $3$ & $\mathrm{(-)}$ \\
     Glacier thickness &$h_{ice}$& $225$ or $740$  &\unit{m}\\

### 2024-03-21 Delaneyetal2024.tex
small changes

@@
 The model outputs' variability in subglacial channels approaches that of subaerial ones over $1-5$ days of water aggregation time (Figure~\ref{fig:multi_run}), possibly resulting in parity in sediment transport capacity variations between subaerial and subglacial channels over these periods.
-However, generally steeper hydraulic gradients of glaciers may cause elevated sediment transport capacity in subglacial channels (Figure\,\ref{fig:model_outs}). 
-Even so, water discharge forcings aggregated over periods longer than days may omit key sediment discharge characteristics, such as the timing of peaks in sediment transport capacity (Figure~\ref{fig:model_outs}).
+Note that although variability between the events could remain comparable, water discharge forcings aggregated over periods longer than days may omit key sediment discharge characteristics, such as the timing of peaks in sediment transport capacity (Figure~\ref{fig:model_outs}).
 

### 2024-03-21 Delaneyetal2024.tex
Updates from Overleaf

--- 2024-03-21 Delaneyetal2024.tex
@@
   Hydrographs from an Alpine glacier and the Greenland Ice Sheet are used to drive subglacial and subaerial hydraulics models, revealing greater variability in the inferred sediment transport capacity in subglacial than in subaerial channels.
-  Results also show that high subglacial sediment transport rates and water velocities can occur across wide ranging water discharges.
+  Results also show that high subglacial sediment transport rates and water velocities can occur across wide-ranging water discharges.
   These findings are discussed in the context of understanding sediment discharge from glaciers with different hydrological forcings and assumptions of channel type.
@@
 As a result, the subglacial channel size only evolves relatively slowly over days to months, whereas water discharge can vary over hours \cite<e.g.>{iken1986,andrews2014,nanni2020}.
-Therefore, on short time scales, subglacial water flow behaves more like pipe flow and changes in water discharge $Q$ are mainly accommodated with increased or decreased water velocity $v$ \cite{alley1997} because the size of the channel $S$ changes on a much longer timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
+Therefore, on short time scales, subglacial water flow behaves more like pipe flow, and changes in water discharge $Q$ are mainly accommodated with increased or decreased water velocity $v$ \cite{alley1997} because the size of the channel $S$ changes on a much longer timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
 
@@
       Arrow length denotes Water velocity magnitudes in the subglacial, $v_g$, and subaerial, $v_f$, channels.
-      $S_g$ ($S_f$) represent the wetted area in subglacial (subaerial) channels.
-      The subglacial channel width $w_g$ remain unchanged, while the subaerial channel width $w_f$ evolves.
+      $S_g$ ($S_f$) represents the wetted area in subglacial (subaerial) channels.
+      The subglacial channel width $w_g$ remains unchanged, while the subaerial channel width $w_f$ evolves.
       Shear stress $\tau$ is responsible for the mobilization of sediment.
@@
 
-Here, it is assumed that the water is transported through a subglacial channel \cite<Figure~\ref{fig:cartoon}; >{rothlisberger1972} beneath a glacier with length $l$, a flat bed and thickness $h_{ice}$.
-The channel  size responds to melt due to frictional heating from water flow and to creep closure of the ice, using the Darcy-Weisbach formulation for water-flow through a pipe  \cite<e.g.>{rothlisberger1972,clarke2003}.
+Here, it is assumed that the water is transported through a subglacial channel \cite<Figure~\ref{fig:cartoon}; >{rothlisberger1972} beneath a glacier with length $l$, with a flatbed and a mean thickness of $h_{ice}$.
+The channel  size grows from melt due to frictional heating from water flow and closes due to the creep  of the ice.
 The formulation here does not consider the englacial storage of water.
@@
 Note that this subaerial channel model is purely algebraic, whereas the subglacial model comprises a differential equation for the evolution of $S_g$.
-Thus the subglacial model has a history dependence on the discharge $Q_w$.
+Thus the channel size in the subglacial model has a history dependence on the discharge $Q_w$.
 
@@
     Variables on the x-axis have been normalized to mean values.
-    $R_g$ ($R_f$) show the Spearman rank correlations for the subglacial (subaerial) outputs.
+    $R_g$ ($R_f$) shows the Spearman rank correlations for the subglacial (subaerial) outputs.
     Plots are shown with discharge and outputs at $15$\unit{min} intervals.
@@
 A glacier's sediment transport capacity is impacted by the ice thickness controlling the channel closure rate, and the glacier's surface slope, in addition to water discharge and sediment size \cite<Figure~\ref{fig:multi_run}, Section~\ref{sect:sub_mode}; >{rothlisberger1972,shreve1972,gimbert2016,stevens2022,walder1994}.
-In a supply-limited state, then glaciers' sediment discharge record also responds to  sediment availability and bedrock erosion from  water pressure variations and sliding  \cite<e.g.>{iverson2012,herman2015,delaney2023}.
+In a supply-limited state, the glaciers' sediment discharge record also responds to  sediment availability and bedrock erosion from  water pressure variations and sliding  \cite<e.g.>{iverson2012,herman2015,delaney2023}.
 This multitude of processes lies in contrast to many subaerial channels, where transport capacity typically responds to sediment size, channel shape, water discharge, and hydraulic gradient.

### 2024-03-22 Delaneyetal2024.tex
to daniel

@@
 Results from  both \alpine{} and \icesheet{} scenarios show that the variability in subglacial  decreases with aggregations longer than approximately 1--5 days (Figure~\ref{fig:multi_run}).
-Yet, substantial differences in the variability between the model outputs for subglacial and subaerial cases persist in periods of up to $15$ days, after which seasonal variations might be conflated.
+Yet, substantial differences in the variability between the model outputs for subglacial and subaerial cases persist in periods of up to $15$ days, longer than which seasonal variations might be conflated.
 Note that the timing of peak sediment transport conditions still occurs across these periods, and they are are thus not captured in aggregating water discharge over longer periods (Figure~\ref{fig:model_outs}).
@@
 \ref{sect:scaling} provides a complete description of sediment transport capacity in different drainage conditions.
-Note that in pipe flow conditions, maximum sediment transport would coincide with peak water discharge and a co-varying relationship with water discharge would occur, unlike with the evolving channel size here (Section~\ref{sect:sub_mode}; Figure \ref{fig:model_outs} and \ref{fig:Qw_vari}).
+Note that in pipe flow conditions, maximum sediment transport would coincide with peak water discharge and a co-varying relationship with water discharge would occur, unlike with the evolving channel size here where a disparate relationship with water discharge occurs (Section~\ref{sect:sub_mode}; Figure \ref{fig:model_outs} and \ref{fig:Qw_vari}).
 Similar adjustments in channel shape occur in subaerial channels \cite{phillips2016}, albeit at timescales longer than hours to days in subglacial channels.

### 2024-03-27 Delaneyetal2024.tex
first pass at Daniel's comments done

@@
   In subaerial channels, water discharge variations are accommodated largely by flow depth and width changes.
-  In subglacial channels, water discharge changes primarily affect water velocity due to evolving channel geometry over hours to weeks.
-  Hydrographs from an Alpine glacier and the Greenland Ice Sheet are used to drive subglacial and subaerial hydraulics models, revealing greater variability in the inferred sediment transport capacity in subglacial than in subaerial channels.
+  In subglacial channels rather, water discharge changes primarily affect water velocity due to evolving channel geometry over hours to weeks.
+  Here, hydrographs from an Alpine glacier and the Greenland Ice Sheet are used to drive subglacial and subaerial hydraulics models, revealing greater variability in the inferred sediment transport capacity in subglacial than in subaerial channels.
   Results also show that high subglacial sediment transport rates and water velocities can occur across wide-ranging water discharges.
@@
 
-Changes in glacier dynamics and hydrology  have prompted numerous  recent studies of  sediment transport processes in cold regions \cite<e.g.>{zhang2022}.
+Changes in glacier dynamics and hydrology  have prompted numerous  recent studies of  sediment transport processes in cold regions \cite<e.g.>{li2022,vergara2022,zhang2022}.
 Increases in sediment transport  have been observed in Greenland \cite{bendixen2017}, the European Alps \cite{costa2017}, the Himalayas \cite{li2021}, and the Andes \cite{vergara2022}.
 In some regions, increased water discharge and glacier melt have been interpreted to yield greater sediment transport capacity \cite{bendixen2017,costa2017,li2021}.
-To explain observed changes to sediment transport in glacierized catchments require examining the processes controlling sediment discharge in these catchments and its variations with water discharge  \cite<e.g.>{riihimaki2005,swift2005}.
+To explain observed changes to sediment transport in glacierized catchments requires examining the processes controlling sediment discharge and its variations with water discharge  \cite<e.g.>{riihimaki2005,swift2005}.
 
-Over long periods, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from under glaciers \cite<c.f.>{hallet1979,iverson2012,ugelvig2018}.
-Pressurized subglacial water can transport this sediment from underneath glaciers \cite{walder1994,creyts2013,beaud2018,delaney2019}, should enough sediment be present subglacially (i.e.
-in a transport-limited regime).
+Over long periods, such as millenia, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from under glaciers \cite<c.f.>{hallet1979,iverson2012,ugelvig2018}.
+Pressurized subglacial water can transport this sediment from underneath glaciers \cite{walder1994,creyts2013,beaud2018,delaney2019}, should enough sediment be present subglacially.
 
-In a transport-limited regime, sediment discharge tracks sediment transport capacity, defined as the amount of sediment the water can carry.
-Sediment transport capacity depends on the shear stress between water and the sediment it flows  over \cite{shields1936,meyer1948,engelund1967} along with the width of the channel bottom $w$ over which sediment mobilizes.
+In a transport-limited regime, sediment discharge is controlled by sediment transport capacity, defined as the amount of sediment the water can carry.
+In both subglacial and subaerial channels, sediment transport capacity depends on the shear stress between water and the sediment it flows  over \cite{shields1936,meyer1948,engelund1967} along with the width of the channel bottom $w$ over which sediment mobilizes.
 The shear stress $\tau$ responds to the velocity of water $v$ flowing through the channel so that
@@
 As a result, the subglacial channel size only evolves relatively slowly over days to months, whereas water discharge can vary over hours \cite<e.g.>{iken1986,andrews2014,nanni2020}.
-Therefore, on short time scales, subglacial water flow behaves more like pipe flow, and changes in water discharge $Q$ are mainly accommodated with increased or decreased water velocity $v$ \cite{alley1997} because the size of the channel $S$ changes on a much longer timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
+Therefore, on short time scales, subglacial water flow behaves more like pipe flow, and changes in water discharge $Q$ are mainly accommodated by increased or decreased water velocity $v$ \cite<Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}; >{alley1997}.
 
@@
     \includegraphics[width=0.8\linewidth]{Fig1.pdf}
-    \caption{Sketch for the different short time-scale responses of subglacial and subaerial channels to increased water discharge.
-      Arrow length denotes Water velocity magnitudes in the subglacial, $v_g$, and subaerial, $v_f$, channels.
+    \caption{Sketch for the different responses of subglacial and subaerial channels to increased water discharge over short time-scales.
+      Arrow length denotes water velocity magnitudes in the subglacial (subaerial) $v_g$ ($v_f$) channels.
       $S_g$ ($S_f$) represents the wetted area in subglacial (subaerial) channels.
-      The subglacial channel width $w_g$ remains unchanged, while the subaerial channel width $w_f$ evolves.
-      Shear stress $\tau$ is responsible for the mobilization of sediment.
+      The subglacial channel width $w_g$ remains unchanged, while the subaerial channel width $w_f$ evolves with water discharge.
+      Subglacial (subaerial) shear stress $\tau_g$ ($\tau_f$) is responsible for the mobilization of sediment.
     }
@@
 
-Consequently, sediment mobilization in subaerial and subglacial channels responds differently to changing water discharge.
+Because of the above, sediment mobilization in subaerial and subglacial channels responds differently to changing water discharge.
 These differences are implicitly included in a wide range of available models that  quantify sediment transport in both subglacial and subaerial channels  \cite<e.g.>{walder1994,alley1997,tucker1997,creyts2013,delaney2019,hewitt2019,wickert2019}.
@@
 Water discharge data is used from Alpine and ice sheet settings collected downstream of the glacier, assuming no water storage in the proglacial area.
-The Alpine site (\alpine{}) is  Fieschergletscher in the Swiss Alps ($46^\circ\,29'\,07''$ N, $8^\circ\,08'\,3''$ E).
-The water discharge data used here was collected at a $1$\unit{min} interval from May 24, 2014 to October 10, 2014
-see \citeA{felix2022} and \citeA{felix2021} for more details.
+The Alpine site (labeled \alpine{}) is  Fieschergletscher in the Swiss Alps ($46^\circ\,29'\,07''$ N, $8^\circ\,08'\,3''$ E).
+The water discharge data used here was collected at a $1$\unit{min} interval from May 24, 2014 to October 10, 2014 \citeA{felix2021,felix2022}.
 
-The Leverett Glacier in Greenland (\icesheet{}) serves as the ice sheet setting.
+The Leverett Glacier in Greenland (labeled \icesheet{}) serves as the ice sheet setting.
 Water discharge was collected roughly $2$\unit{km} downstream from the terminus between June and August/September 2009 to 2012 ($67^\circ\,03'\,5''$ N, $50^\circ\,12'\,59''$ W).
@@
 To evaluate the shear stress of water flowing across sediments underneath a glacier, the subglacial channel model accounts for the channel geometry and the velocity of the flowing water.
-To do this, we use a lumped hydraulics model \cite{clarke1996} modified from the one presented by \citeA{werder2010b}.
+To do this, we use a lumped hydraulics model from \citeA{werder2010b}, itself based upon  \cite{clarke1996}.
 
-Here, it is assumed that the water is transported through a subglacial channel \cite<Figure~\ref{fig:cartoon}; >{rothlisberger1972} beneath a glacier with length $l$, with a flatbed and a mean thickness of $h_{ice}$.
+Here, it is assumed that the water is transported through a subglacial channel \cite<Figure~\ref{fig:cartoon}; >{rothlisberger1972} beneath a glacier with channel length $l$, with a flatbed and a mean thickness of $h_{ice}$.
 The channel  size grows from melt due to frictional heating from water flow and closes due to the creep  of the ice.
@@
 \end{linenomath*}
-\noindent where $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (Table \ref{table:vpm}), $g$ is the acceleration due to gravity, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change over $l$, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the mean ice overburden pressure expressed in meter water equivalent, and $n$ is Glen's n \cite{glen1955} (usually $n=3$).
+\noindent where $t$ is time, $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (complete values given in Table \ref{table:vpm}), $g$ is the acceleration due to gravity, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change over $l$, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the mean ice overburden pressure expressed in meter water equivalent ($\rho_w$ is density of water; $\rho_i$ is density of ice), and $n$ is Glen's n \cite{glen1955} (usually $n=3$).
 The first term on the right side of the equation represents the opening of the channel through frictional heating, while the following term represents the creep closure of the channel from ice deformation.
@@
 \end{linenomath*}
-\noindent where $f_r$ is a friction factor, $D_h$ is the hydraulic diameter, $l$ is the length over which head change occurs, and $v_g=\frac{Q}{S_g}$ is the water flow speed.
+\noindent where $f_r$ is a friction factor, $D_h$ is the hydraulic diameter, $l$ is the channel length, and $v_g=\frac{Q}{S_g}$ is the water flow speed.
 %
@@
 %
-where $v_g = \frac{Q}{S_g}$ is the water velocity and $\rho_w$ the density of water.
+where $v_g = \frac{Q}{S_g}$ is the water velocity.
 %
@@
 % To parameterize the shear stress of water flowing across sediments in the subaerial channel, the hydraulics parameterization presented in \citeA{tucker1997} is implemented.
-Using mass conservation and the Darcy-Weisbach relationship as well as assuming that the channel is wide so that the hydraulic radius is well approximated by the flow depth, then
+Using mass conservation and the Darcy-Weisbach relationship as well as assuming that the channel is sufficiently wide compared to its depth so that the hydraulic radius is well approximated by the flow depth,
 the shear stress $\tau_f$ at the river bed is
@@
 where $k$ is a constant and $\alpha=\frac{1}{2}$ is a commonly chosen exponent \cite{leopold1953}.
-Water velocity, $v_f$, is given by rearranging Equation \ref{eq:tau_g} as
+Water velocity, $v_f$, is given as
 \begin{equation}
@@
 Note that this subaerial channel model is purely algebraic, whereas the subglacial model comprises a differential equation for the evolution of $S_g$.
-Thus the channel size in the subglacial model has a history dependence on the discharge $Q_w$.
+Thus the channel size in the subaerial model has no history dependence on the discharge $Q_w$, where as the subglacial one does.
 
@@
 Outputs of the models represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
-To generalize these scenarios, \alpine{}  is exemplified by relatively thin ice thickness ($h_{ice}$: $225$\,\unit{m}), low water discharge ($\sim\,10$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
-\icesheet{}  is exemplified by thick ice  ($h_{ice}$: $700$\,\unit{m}), high water discharge ($\sim\,300$\,\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).
+To generalize these scenarios, \alpine{}  is exemplified by Fieschergletscher's relatively thin ice thickness ($h_{ice}$: $225$\,\unit{m}), low water discharge ($\sim\,10$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
+\icesheet{}  is exemplified by Leverett glacier's  thick ice  ($h_{ice}$: $700$\,\unit{m}), high water discharge ($\sim\,300$\,\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).
 %However, both scenarios share a discharge with a similar, relative seasonal amplitude.
 
-For each glacier, the models are first applied to a reference test case for each glacier that assumes  a subglacial channel with $\beta=\frac{\pi}{6}$%, i.e. a segment of a quarter-circle,
+In the first experiment, for each glacier, the models are  applied to a reference test case  that assumes  a subglacial channel with $\beta=\frac{\pi}{6}$%, i.e. a segment of a quarter-circle,
 (Table~\ref{table:vpm}).
-Both friction factors $f_r$ and $f_f$ are tuned so that reasonable water velocities occur for both \alpine{} and \icesheet{} \cite<$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$>{werder2010b,chandler2013}.
-The initial condition of cross-sectional area $S_g$ is evaluated by applying a model spinup with the maximum water discharge for the first $3$ days of the study until the change in channel area is negligible.
+Both friction factors $f_r$ and $f_f$ are tuned so that reasonable water velocities \cite<$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$>{werder2010b,chandler2013} occur for both \alpine{} and \icesheet{}.
+The initial condition of cross-sectional area $S_g$ is evaluated by applying a model spinup with the maximum observed water discharge for the first $4$ days of the study until there is no change in channel area.
 The subaerial channels of both scenarios have a bed slope of $\nabla z_c = 0.05$ (Table~\ref{table:vpm}).
@@
 For brevity, these values together are referred to as ``model outputs'' in the text.
-Outputs  are compared to water discharge from the glaciers using Spearman rank correlation, which varies with the ordering of values.
+Model outputs are compared to water discharge from the glaciers using Spearman rank correlation, which varies with the ordering of values.
 Rank correlation reduces the impact of the non-linear response of sediment transport capacity to the hydrological forcing.
@@
 
-The effect of channel shape and roughness is established by running the model with random parameter values of channel geometry factors ($\beta$ and $k$) and friction factors ($f_r$ and $f_f$).
-Runs with parameter combinations are accepted if their mean subglacial water velocity over the season lies between $0.75$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $2$\,\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity lies between $0.3$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}$^3$\,\unit{s}$^{-1}$, following measurements \cite<e.g.>{werder2010b,magnusson2012,chandler2013}.
-Additionally, to be accepted, model runs can only experience a flotation fraction of above $1.2$ for less than $2.5$\,\% of the run.
+The effect of channel shape and roughness is established by running the model with random parameter values of channel geometry factors ($\beta$ and $k$) and friction factors ($f_r$ and $f_f$; see Table~\ref{table:vpm} for values).
+Runs with parameter combinations are accepted if their mean subglacial water velocity over the season lies between $0.75$\,\unit{m}\,\unit{s}$^{-1}$ and $2$\,\unit{m}\,\unit{s}$^{-1}$ or if subaerial water velocity lies between $0.3$\,\unit{m}\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}\,\unit{s}$^{-1}$, following measurements \cite<e.g.>{werder2010b,magnusson2012,chandler2013}.
+Additionally, to be accepted, model runs can only experience a flotation fraction ($\frac{\Delta h}{2\,h_o}$) of above $1.2$ for less than $2.5$\,\% of the run.
 
-To establish the role of water discharge variability in driving variations in sediment transport capacity, the water discharge records are aggregated across a select period ranging from $15$\,\unit{min} up to $15$\,\unit{days}.
-Here, the water discharge input at a given time is the average over the aggregation period before that time.  
-The routine runs until $100$ different parameter combinations are accepted for each water discharge aggregation period.
+The routine runs until $100$ different parameter combinations for each water discharge aggregation period are accepted with the conditions described above.
+Here, the water discharge input at a given time is the average over the aggregation period before that time, ranging from $15$\,\unit{min} up to $15$\,\unit{days}.
 
@@
     Arrows denote where variables peak in the subglacial channel before the subaerial one.
-    Insets in f-h show the  peak melt event \icesheet{} denoted by the shaded area in panel e.
+    Insets in f-h show the  peak melt event denoted by the shaded area in panels e--h, but have arbitrary y-axis.
     % Note that the x and y axes are different for \alpine{} and \icesheet{}.
@@
 
-Examining the relationship between water discharge and the model outputs, a different relationship emerges between the subaerial and subglacial cases.
+Examining the relationship between water discharge and the model outputs, a different relationship emerges between the subaerial and subglacial cases, due to the hysteresis in channel size in the subglacial model (Equation~\ref{eq:dS_dt}).
 Because there is no history dependence in  subaerial channels, each water discharge value in the subaerial channel results in a unique water velocity, shear stress, and width-integrated shear stress (Section~\ref{sect:sub_mode}).
 This characteristic results in a perfect rank correlation between the variables (Figure~\ref{fig:Qw_vari}).
-In the subglacial channel, however, the hysteresis in channel size evolution (Equation~\ref{eq:dS_dt}) means that subglacial model outputs respond differently to water discharge than subaerial ones.
-Here, the persistence of high water velocity, shear stress, and width-integrated shear stress across a range of water discharges suggests that substantial sediment transport may occur under a wide range of hydrological forcings (Figure~\ref{fig:Qw_vari}.
-
-For instance, in \alpine{}, very high water velocity values and shear stresses can occur from a low water discharge  ($\sim\,4$\,\unit{m}$^3$\,\unit{s}$^{-1}$) to the maximum water discharge at over ($17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, Figure~\ref{fig:Qw_vari}\,a).
-In \icesheet{}, mean values water velocity over the season can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
+For instance, in subglacial channels in \alpine{}, very high seasonal water velocity values and shear stresses can occur from a low water discharge  ($\sim\,4$\,\unit{m}$^3$\,\unit{s}$^{-1}$) to the maximum water discharge at over ($17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, Figure~\ref{fig:Qw_vari}\,a).
+In \icesheet{}, water velocities close to the seasonal mean value over the season can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
 When considering the subglacial channel's evolving width, width-integrated shear stress  across the channel generally increases with water discharge, with improved rank correlation compared to water velocity or shear stress (\alpine{}, Figure~\ref{fig:Qw_vari} a--c).
 Yet even the width-integrated shear stress  can vary substantially, with the highest values occurring at water discharge values ranging from roughly $11$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ to over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
-The variability in width-integrated shear stress is less pronounced in the \icesheet{} scenario, where the hydrograph has less diurnal variability in water discharge (Figure~\ref{fig:Qw_vari} c, f).
+The variability in width-integrated shear stress is less pronounced in the \icesheet{} scenario, where the hydrograph has less diurnal variability (Figure~\ref{fig:Qw_vari} c, f).
 
-The different response of subaerial and subglacial model outputs to water discharge impacts the timing of peak events.
-In the subglacial channel, peaks in $w_g\tau_g$ generally occur when water discharge increases at the fastest rate, but before the maximum quantity of water discharge (Figure~\ref{fig:model_outs}\, d and h).
+The different response of subaerial and subglacial model outputs to water discharge impacts the timing of peaks in sediment transport capacity.
+In the subglacial channel, peaks in $w_g\tau_g$ generally occur when water discharge increases at the fastest rate, but before the maximum water discharge (Figure~\ref{fig:model_outs}\, d and h).
 By contrast, in the subaerial channel,   $w_f\tau_f$  peaks when water discharge is highest.
@@
 Across the range of channel shapes and friction factors  examined (Section\,\ref{sect:imp}), variability in water velocity and shear stress remains higher in the subglacial system compared to the subaerial one in both \alpine{} and \icesheet{}.
-This suggests that there is likely greater intermittency in sediment transport in subglacial systems compared to subaerial ones, especially in supply-limited systems where sediment could be exhausted or mobilized over diurnal discharge cycles.
+This suggests that there is greater intermittency and variability in sediment transport in subglacial systems compared to subaerial ones, especially in supply-limited systems where sediment could be exhausted or mobilized over diurnal discharge cycles.
 In contrast, in \alpine{}, the variability in width-integrated shear stress is comparable between the subaerial and subglacial systems, due to the highly variable water discharge, largely due to diurnal variations (Figure~\ref{fig:multi_run}).
@@
 
-Sediment transport capacity in subglacial channels responds more rapidly to changes in water discharge compared to their subaerial counterparts \cite{alley1997}.
+Sediment transport capacity in subglacial channels responds more rapidly to changes in water discharge compared to their subaerial counterparts \cite<Figures~\ref{fig:Qw_vari}~and~\ref{fig:multi_run}>{alley1997}.
 In subglacial R\"othlisberger channels in steady state with water discharge, the width-integrated shear stress response to water discharge is $ w_g\,\tau_{g}\, \propto\,  Q^{\sim\frac{4}{5}}$ (Table\,\ref{tab:eqs1}).
@@
 Note that in pipe flow conditions, maximum sediment transport would coincide with peak water discharge and a co-varying relationship with water discharge would occur, unlike with the evolving channel size here where a disparate relationship with water discharge occurs (Section~\ref{sect:sub_mode}; Figure \ref{fig:model_outs} and \ref{fig:Qw_vari}).
-Similar adjustments in channel shape occur in subaerial channels \cite{phillips2016}, albeit at timescales longer than hours to days in subglacial channels.
+Similar adjustments in channel shape occur in subaerial channels \cite{phillips2016}, albeit at timescales longer than the subglacial channel's response of hours to days.
 
 Substantial differences emerge in model outputs between the different hydrological forcings of the \alpine{} and \icesheet{} cases.
-Indeed, the more independent relationship between model outputs and water discharge in \alpine{} results from the
+Indeed, the more decoupled and sporadic relationship between model outputs and water discharge in \alpine{} results from the
 relatively larger discharge variations on daily to weekly timescales compared to \icesheet{}.
@@
 
-This co-varying relationship between sediment discharge capacity in  subaerial conditions could be pronounced $\sim20$\,\unit{km} downstream of the Leverett site, where a strong correlation persists between sediment plume size and the river's water discharge into the Kangerlussuaq fjord \cite{mcgrath2010}.
+Previous work suggests that this co-varying relationship between sediment discharge capacity in  subaerial conditions could be noticeable $\sim20$\,\unit{km} downstream of the Leverett site, where a strong correlation persists between sediment plume size and the river's water discharge into the Kangerlussuaq fjord \cite{mcgrath2010}.
 In contrast, in marine-terminating glacier catchments, a less consistent relationship may occur between water discharge or melt extent and sediment plume size \cite{tedstone2012}.
-Here, ocean water pressurizes subglacial channels at the ice front \cite<e.g.>{how2017}, so the reduced correlation could result from the inconsistent relationship between subglacial sediment transport capacity and water discharge (e.g.
-Figure~\ref{fig:Qw_vari}).
+Here, ocean water pressurizes subglacial channels at the ice front \cite<e.g.>{how2017}, so the reduced correlation could result from the inconsistent relationship between subglacial sediment transport capacity and water discharge.
 
-Furthermore, the greater variations in subglacial sediment transport capacity in both catchment types compared to subaerial ones could cause a supply-limited regime to persist at many glaciers.
-In subglacial systems, sediment's threshold of motion can be reached more frequently and across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
+Furthermore, the greater variations in subglacial sediment transport capacity compared to subaerial ones could cause a supply-limited regime to persist at many glaciers.
+In subglacial systems, sediment's critical shear stress, or threshold at which sediment mobilization begins, can be reached more frequently and across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
 Sediment exhaustion through this process may  explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \cite{overeem2017}.
@@
 Numerical models of subglacial sediment transport generally apply classical parameterizations of subglacial hydrology to evaluating sediment transport conditions \cite<e.g>{hewitt2019,beaud2018}.
-In these R-channel models, viable water pressure regimes can be achieved by changing the shape and friction factor of the channels, enabling assumptions of semi-circular channels in most applications \cite{werder2010b}.
+In these R-channel models, realistic water pressure regimes can be achieved by changing the channels shape and friction factor, therefore  model tuning is done with the friction factor and leaving the channel shape fixed as a semi-circle \cite<$\beta =\pi$; >{werder2010b}.
 However, the role of channel width in the findings here suggests that more thorough parameterization of channel shape and channel width is needed in assessing subglacial sediment transport capacity as compared to water pressure (Section\,\ref{sect:sub_mode}, \ref{sect:scaling}).
 
-Water discharge measurements at the sub-hour timescale remain unavailable in many catchments, further limiting the ability of subglacial  sediment transport models to capture certain events.
-The model outputs' variability in subglacial channels approaches that of subaerial ones over $1-5$ days of water aggregation time (Figure~\ref{fig:multi_run}), possibly resulting in parity in sediment transport capacity variations between subaerial and subglacial channels over these periods.
+Water discharge measurements at the hourly timescale remain unavailable in many catchments, further limiting the ability of subglacial  sediment transport models to capture certain events.
+The model outputs' variability in subglacial channels approaches that of subaerial ones over $1-5$ days of  aggregation (Figure~\ref{fig:multi_run}), possibly resulting in parity in sediment transport capacity variations between subaerial and subglacial channels over these periods.
 Note that although variability between the events could remain comparable, water discharge forcings aggregated over periods longer than days may omit key sediment discharge characteristics, such as the timing of peaks in sediment transport capacity (Figure~\ref{fig:model_outs}).
@@
 \item
-  water discharge variability in the two hydrographs impacts subglacial sediment transport capacity fluctuations;
+  water discharge variability impacts subglacial sediment transport capacity fluctuations to a larger degree than subaerial channels;
   \vspace{.1cm}
@@
 \end{enumerate}
+
+This study calls for the careful analysis of the relationship between sediment transport capacity and water discharge when examining sediment discharge from glacierized catchments.
 
@@
   \caption{Variables, parameters, and constants used in this work.
-    Where two values are given, the first refers to  \alpine{}, a scenario from Fieschergletscher, and the second to a glacier marginal to the \icesheet{} a scenario from Leverett Glacier}
+    Where two values are given, the first refers to  \alpine{}, a scenario from Fieschergletscher, and the second to a glacier marginal to the \icesheet{} a scenario from Leverett Glacier.
+If a second line is given, it refers to the range of values examined in the parameter search}
   \small % TODO remove again
@@
     Hooke angle of channel & $\beta$ & $\frac{\pi}{6}$ & \unit{rad}\\
+         && $\beta$:($\frac{\pi}{10}$, $\pi$) & \\
     Friction factor (subglacial, subaerial) & $f$, ($f_r$, $f_f$) &$5$ or $16$, $3$ & $\mathrm{(-)}$ \\
+         && $f_g,f_f$:($0.01$, $21$) & \\
     Glacier thickness &$h_{ice}$& $225$ or $740$  &\unit{m}\\
@@
     Subaerial channel factor & $k$ &$3$ or $6.5$ & $\mathrm{s\,m^{-2}}$\\
+     && $k$:($5$, $20$) & \\
     Channel geometry exponent &$\alpha$& $\frac{1}{2}$&$\mathrm{(-)}$ \\

### 2024-03-27 Delaneyetal2024.tex
Updates from Overleaf
small changes with grammer/formating

@@
 \correspondingauthor{Ian Delaney}{ianarburua.delaney@unil.ch}
-
+\ian{quick check on shear stress units: seems like Pa, not Pa/m}
 \begin{keypoints}
@@
 
-Over long periods, such as millenia, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from under glaciers \cite<c.f.>{hallet1979,iverson2012,ugelvig2018}.
+Over long periods, such as millennia, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from under glaciers \cite<c.f.>{hallet1979,iverson2012,ugelvig2018}.
 Pressurized subglacial water can transport this sediment from underneath glaciers \cite{walder1994,creyts2013,beaud2018,delaney2019}, should enough sediment be present subglacially.
@@
 In subaerial channels, operating with open channel flow, the wetted area $S$ of the channel evolves with changing water discharge $Q$, by changing both the channel's width and the water depth \cite{leopold1953}.
-The change in water depth results in a commensurate increase in water velocity.
+The change in water depth results in a proportional increase in water velocity.
 As a result, the shear stress $\tau$ increases according to Equation~\ref{eq:tau}.
@@
 Outputs demonstrate differences in the relationships between  water discharge, channel geometry, and water velocity in the subglacial and subaerial channels.
-The manuscript then uses the demonstrated differences between the two channel types to discuss the implications for interpreting the processes driving variability in records of sediment transport, and their relationship with hydrology from glacierized catchments.
+The manuscript then uses the demonstrated differences between the two channel types to discuss the implications for interpreting the processes driving variability in sediment transport records and their relationship with hydrology from glacierized catchments.
 
@@
 The Alpine site (labeled \alpine{}) is  Fieschergletscher in the Swiss Alps ($46^\circ\,29'\,07''$ N, $8^\circ\,08'\,3''$ E).
-The water discharge data used here was collected at a $1$\unit{min} interval from May 24, 2014 to October 10, 2014 \citeA{felix2021,felix2022}.
+The water discharge data used here was collected at a $1$\unit{min} interval from May 24, 2014, to October 10, 2014 \cite{felix2021,felix2022}.
 

### 2024-03-28 Delaneyetal2024.tex
push to overleaf

@@
 \correspondingauthor{Ian Delaney}{ianarburua.delaney@unil.ch}
-\ian{quick check on shear stress units: seems like Pa, not Pa/m}
+
 \begin{keypoints}
@@
   Sediment transport capacity in both subaerial and subglacial channels depends on the shear stress exerted across the channel bottom, which varies with the water velocity and channel width.
-  In subaerial channels, water discharge variations are accommodated largely by flow depth and width changes.
-  In subglacial channels rather, water discharge changes primarily affect water velocity due to evolving channel geometry over hours to weeks.
+  In subaerial channels, water discharge variations are accommodated  by flow depth and width changes, along with water velocity.
+  In subglacial channels rather, water discharge changes primarily affect solely water velocity due to evolving channel geometry over hours to weeks.
   Here, hydrographs from an Alpine glacier and the Greenland Ice Sheet are used to drive subglacial and subaerial hydraulics models, revealing greater variability in the inferred sediment transport capacity in subglacial than in subaerial channels.
   Results also show that high subglacial sediment transport rates and water velocities can occur across wide-ranging water discharges.
-  These findings are discussed in the context of understanding sediment discharge from glaciers with different hydrological forcings and assumptions of channel type.
+  These findings are discussed in the context of understanding sediment discharge from glaciers with different hydrological forcings and assumptions of channel shape.
 \end{abstract}
@@
 
+Note that sediment discharge records are omitted from the analysis here given the role of sediment supply, in addition to capacity, on capturing variations in sediment discharge \cite<e.g.>{delaney2024}.
+
 \section{Methods}
@@
 Both models use the measured discharge to calculate water velocity, shear stress, and width-integrated shear stress, upon which both suspended sediment and bedload transport depend \cite<Figure \ref{fig:cartoon}; >{shields1936}.
-The evaluation of these variables omits the selection of a sediment transport relationship and a grain-size parameter \cite<e.g.>{shields1936,meyer1948} needed to calculate sediment transport capacity from shear stress.
+The evaluation of these variables omits the selection of a sediment transport relationship and a grain-size parameter \cite<e.g.>{shields1936,meyer1948}.
 
@@
 To evaluate the shear stress of water flowing across sediments underneath a glacier, the subglacial channel model accounts for the channel geometry and the velocity of the flowing water.
-To do this, we use a lumped hydraulics model from \citeA{werder2010b}, itself based upon  \cite{clarke1996}.
+To do this, we use a lumped hydraulics model from \citeA{werder2010b}, itself based upon  \citeA{clarke1996}.
 
@@
 where $k$ is a constant and $\alpha=\frac{1}{2}$ is a commonly chosen exponent \cite{leopold1953}.
-Water velocity, $v_f$, is given as
+Also following the Darcy-Weisbach, subaerial water velocity, $v_f$, is given as
 \begin{equation}
@@
 Note that this subaerial channel model is purely algebraic, whereas the subglacial model comprises a differential equation for the evolution of $S_g$.
-Thus the channel size in the subaerial model has no history dependence on the discharge $Q_w$, where as the subglacial one does.
+Thus the channel size in the subaerial model has no history dependence on the discharge $Q_w$, where as the subglacial one does (Equation~\ref{eq:dS_dt}).
 
@@
 
-The models above are applied to proglacial discharge records from the Fieschergletscher (scenario \alpine{}) and the Leverett Glacier (scenario \icesheet{}).
-The subglacial and subaerial models represent the subglacial flow of water exiting a glacier and transitioning to subaerial flow.
+The models above are applied to proglacial discharge records from the Fieschergletscher (scenario \alpine{}) and the Leverett Glacier (scenario \icesheet{}), using subglacial model outputs that are contrasted with subaerial ones.
 Outputs of the models represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
-To generalize these scenarios, \alpine{}  is exemplified by Fieschergletscher's relatively thin ice thickness ($h_{ice}$: $225$\,\unit{m}), low water discharge ($\sim\,10$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
-\icesheet{}  is exemplified by Leverett glacier's  thick ice  ($h_{ice}$: $700$\,\unit{m}), high water discharge ($\sim\,300$\,\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).
+To generalize these scenarios, \alpine{}  is exemplified by Fieschergletscher's relatively thin ice thickness \cite<$h_{ice}$: $225$\,\unit{m}>{grab2021}, low water discharge ($\sim\,10$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
+\icesheet{}  is exemplified by Leverett glacier's  thick ice  \cite<$h_{ice}$: $700$\,\unit{m}; >{morlighem2017}, high water discharge \cite<$\sim\,300$\,\unit{m}$^3$\,\unit{s}$^{-1}$>{tedstone2017}  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).
 %However, both scenarios share a discharge with a similar, relative seasonal amplitude.
 
-In the first experiment, for each glacier, the models are  applied to a reference test case  that assumes  a subglacial channel with $\beta=\frac{\pi}{6}$%, i.e. a segment of a quarter-circle,
+In the first experiment, the models are  applied to a reference test case for each glacier that assumes  a subglacial channel with $\beta=\frac{\pi}{6}$%, i.e. a segment of a quarter-circle,
 (Table~\ref{table:vpm}).
-Both friction factors $f_r$ and $f_f$ are tuned so that reasonable water velocities \cite<$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$>{werder2010b,chandler2013} occur for both \alpine{} and \icesheet{}.
+Both friction factors $f_r$ and $f_f$ are tuned so that reasonable water velocities \cite<$\sim\,1.6\,$\unit{m}\,\unit{s}$^{-1}$>{werder2010b,chandler2013} occur for both \alpine{} and \icesheet{}.
 The initial condition of cross-sectional area $S_g$ is evaluated by applying a model spinup with the maximum observed water discharge for the first $4$ days of the study until there is no change in channel area.
@@
 For brevity, these values together are referred to as ``model outputs'' in the text.
-Model outputs are compared to water discharge from the glaciers using Spearman rank correlation, which varies with the ordering of values.
+To quantify the link between sediment transport capacity and water discharge, model outputs are compared to water discharge from the glaciers using Spearman rank correlation, which varies with the ordering of values.
 Rank correlation reduces the impact of the non-linear response of sediment transport capacity to the hydrological forcing.
@@
 
-The effect of channel shape and roughness is established by running the model with random parameter values of channel geometry factors ($\beta$ and $k$) and friction factors ($f_r$ and $f_f$; see Table~\ref{table:vpm} for values).
+The effect of channel shape and roughness is established by running the model with random parameter values of channel geometry factors ($\beta$ and $k$) and friction factors ($f_r$ and $f_f$; see Table~\ref{table:vpm} for value range).
 Runs with parameter combinations are accepted if their mean subglacial water velocity over the season lies between $0.75$\,\unit{m}\,\unit{s}$^{-1}$ and $2$\,\unit{m}\,\unit{s}$^{-1}$ or if subaerial water velocity lies between $0.3$\,\unit{m}\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}\,\unit{s}$^{-1}$, following measurements \cite<e.g.>{werder2010b,magnusson2012,chandler2013}.
@@
   \centering
-  \includegraphics[width=0.9\linewidth]{Fig3.png}
+  \includegraphics[width=0.9\linewidth]{Fig3.pdf}
   \caption{
@@
 
-Examining the relationship between water discharge and the model outputs, a different relationship emerges between the subaerial and subglacial cases, due to the hysteresis in channel size in the subglacial model (Equation~\ref{eq:dS_dt}).
+Examining the relationship between water discharge and the model outputs, a different relationships emerge between the subaerial and subglacial cases, due to the hysteresis in channel size in the subglacial model (Equation~\ref{eq:dS_dt}).
 Because there is no history dependence in  subaerial channels, each water discharge value in the subaerial channel results in a unique water velocity, shear stress, and width-integrated shear stress (Section~\ref{sect:sub_mode}).
 This characteristic results in a perfect rank correlation between the variables (Figure~\ref{fig:Qw_vari}).
+
+Conversely, the history dependence on channel size in subglacial channels means different outputs occur across a range of discharges.
 For instance, in subglacial channels in \alpine{}, very high seasonal water velocity values and shear stresses can occur from a low water discharge  ($\sim\,4$\,\unit{m}$^3$\,\unit{s}$^{-1}$) to the maximum water discharge at over ($17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, Figure~\ref{fig:Qw_vari}\,a).
-In \icesheet{}, water velocities close to the seasonal mean value over the season can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
+In \icesheet{}, water velocities close to the seasonal mean value can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
 When considering the subglacial channel's evolving width, width-integrated shear stress  across the channel generally increases with water discharge, with improved rank correlation compared to water velocity or shear stress (\alpine{}, Figure~\ref{fig:Qw_vari} a--c).
-Yet even the width-integrated shear stress  can vary substantially, with the highest values occurring at water discharge values ranging from roughly $11$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ to over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
+Yet even the width-integrated shear stress  can vary substantially, with the highest values occurring at water discharge values ranging from roughly $11$ \,\unit{m}\,\unit{s}$^{-1}$ to over $17$ \,\unit{m}\,\unit{s}$^{-1}$.
 The variability in width-integrated shear stress is less pronounced in the \icesheet{} scenario, where the hydrograph has less diurnal variability (Figure~\ref{fig:Qw_vari} c, f).
@@
   \centering
-    \includegraphics[width=0.9\linewidth]{Fig4.png}
+    \includegraphics[width=0.9\linewidth]{Fig4.pdf}
     \caption{Standard deviation of model outputs for different time aggregations with different subglacial and subaerial channel shapes and friction factors.
@@
 Yet, substantial differences in the variability between the model outputs for subglacial and subaerial cases persist in periods of up to $15$ days, longer than which seasonal variations might be conflated.
-Note that the timing of peak sediment transport conditions still occurs across these periods, and they are are thus not captured in aggregating water discharge over longer periods (Figure~\ref{fig:model_outs}).
+Note that the timing of peak sediment transport conditions still occurs, across these periods, and they are are thus not captured in aggregating water discharge over longer periods (Figure~\ref{fig:model_outs}).
 
@@
 Note that in pipe flow conditions, maximum sediment transport would coincide with peak water discharge and a co-varying relationship with water discharge would occur, unlike with the evolving channel size here where a disparate relationship with water discharge occurs (Section~\ref{sect:sub_mode}; Figure \ref{fig:model_outs} and \ref{fig:Qw_vari}).
-Similar adjustments in channel shape occur in subaerial channels \cite{phillips2016}, albeit at timescales longer than the subglacial channel's response of hours to days.
+Similar adjustments in channel shape occur in subaerial channels \cite{phillips2016}, albeit at timescales longer than the subglacial channel's response time of hours to days.
 
@@
 In contrast, in marine-terminating glacier catchments, a less consistent relationship may occur between water discharge or melt extent and sediment plume size \cite{tedstone2012}.
-Here, ocean water pressurizes subglacial channels at the ice front \cite<e.g.>{how2017}, so the reduced correlation could result from the inconsistent relationship between subglacial sediment transport capacity and water discharge.
+Here, ocean water pressurizes subglacial channels at the ice front \cite<e.g.>{how2017}, so the reduced correlation could result from the inconsistent relationship between subglacial sediment transport capacity and water discharge presented here.
 
 Furthermore, the greater variations in subglacial sediment transport capacity compared to subaerial ones could cause a supply-limited regime to persist at many glaciers.
-In subglacial systems, sediment's critical shear stress, or threshold at which sediment mobilization begins, can be reached more frequently and across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
+In subglacial systems, sediment's critical shear stress, or threshold at which sediment mobilization occurs, can be reached more frequently and across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
 Sediment exhaustion through this process may  explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \cite{overeem2017}.
@@
 The model outputs' variability in subglacial channels approaches that of subaerial ones over $1-5$ days of  aggregation (Figure~\ref{fig:multi_run}), possibly resulting in parity in sediment transport capacity variations between subaerial and subglacial channels over these periods.
-Note that although variability between the events could remain comparable, water discharge forcings aggregated over periods longer than days may omit key sediment discharge characteristics, such as the timing of peaks in sediment transport capacity (Figure~\ref{fig:model_outs}).
+Note that although variability between the events could remain comparable, water discharge forcings aggregated over periods longer than days may omit key sediment discharge characteristics, such as the timing of peaks in sediment transport capacity from glaciers (Figure~\ref{fig:model_outs}).
 
@@
     Hooke angle of channel & $\beta$ & $\frac{\pi}{6}$ & \unit{rad}\\
-         && $\beta$:($\frac{\pi}{10}$, $\pi$) & \\
+         && ($\frac{\pi}{10}$, $\pi$) & \\
     Friction factor (subglacial, subaerial) & $f$, ($f_r$, $f_f$) &$5$ or $16$, $3$ & $\mathrm{(-)}$ \\
-         && $f_g,f_f$:($0.01$, $21$) & \\
+         && ($0.01$, $21$) & \\
     Glacier thickness &$h_{ice}$& $225$ or $740$  &\unit{m}\\
@@
     Subaerial channel factor & $k$ &$3$ or $6.5$ & $\mathrm{s\,m^{-2}}$\\
-     && $k$:($5$, $20$) & \\
+     && ($5$, $20$) & \\
     Channel geometry exponent &$\alpha$& $\frac{1}{2}$&$\mathrm{(-)}$ \\

### 2024-04-05 Delaneyetal2024.tex
Delaneyetal2024.tex

@@
 \begin{keypoints}
-\item Short water discharge fluctuations in subglacial channels are accommodated by water velocity, not channel size, as in subaerial systems.
+\item Short water discharge fluctuations in subglacial channels are accommodated by water velocity, not channel size as in subaerial systems.
 \item Greater variability in water velocity causes sediment transport capacity to vary more in subglacial channels than in subaerial channels.
@@
 \begin{abstract}
-  Sediment transport capacity in both subaerial and subglacial channels depends on the shear stress exerted across the channel bottom, which varies with the water velocity and channel width.
+  Sediment transport capacity in both subaerial and subglacial channels depends on the shear stress exerted across the channel bottom, which varies with water velocity and channel width.
   In subaerial channels, water discharge variations are accommodated  by flow depth and width changes, along with water velocity.
-  In subglacial channels rather, water discharge changes primarily affect solely water velocity due to evolving channel geometry over hours to weeks.
-  Here, hydrographs from an Alpine glacier and the Greenland Ice Sheet are used to drive subglacial and subaerial hydraulics models, revealing greater variability in the inferred sediment transport capacity in subglacial than in subaerial channels.
+  However, in subglacial channels  water discharge changes mainly result in velocity variations, as the channel geometry evolves slowly (over days).
+  Here, hydrographs from an Alpine glacier and the Greenland Ice Sheet are applied to subglacial and subaerial hydraulics models, revealing greater variability in the inferred sediment transport capacity in subglacial than in subaerial channels.
   Results also show that high subglacial sediment transport rates and water velocities can occur across wide-ranging water discharges.
-  These findings are discussed in the context of understanding sediment discharge from glaciers with different hydrological forcings and assumptions of channel shape.
+  These findings are discussed in the context of understanding sediment discharge from glaciers with different hydro-climatic forcings and assumptions of channel shape.
 \end{abstract}
@@
 Sediment export from glaciers impacts downstream ecosystems and communities and has changed in response to climate warming.
-Sediment export depends partly on the sediment transport capacity, and the amount of sediment that could be transported for the hydraulic conditions with abundant sediment availability.
+Sediment export depends on two main factors: sediment availability, and the capacity of the hydrological system to transport that sediment.
 This capacity largely depends on the width of the channel and the water's velocity.
 In subaerial channels, like rivers, water velocity, and channel width both respond rapidly to water discharge.
-However, subglacial channels are pressurized by the ice above, so water flows through a subglacial pipe that evolves in size.
-This pressurized flow means that channel size evolves slowly compared to variations in water discharge, therefore these variations are largely accommodated by changes in water velocity as opposed to channel size.
-Models of these processes show that sediment transport capacity is highly variable in subglacial channels, and a wide range of water discharges can result in comparable sediment transport capacities.
+However, subglacial channels are pressurized by the ice above, so water flows through a subglacial pipe that evolves in size only slowly.
+Variations in water discharge and thus also sediment transport capacity are therefore mainly accommodated by changes in the water velocity.
+Models of these processes show that sediment transport capacity is highly variable in subglacial channels, and a wide range of water discharges can result in similar sediment transport capacities.
 In subaerial channels, by contrast, sediment transport capacity is less variable, and variations depend solely on water discharge, not the channel size.
-The channel's different responses to water discharge may impact the interpretation of sediment transport records from glacierized catchments, complicating the link between glacier melt and sediment discharge.
+These findings may impact the interpretation of sediment transport records from glacierized catchments, complicating the link between glacier melt and sediment discharge.
 
@@
 Over long periods, such as millennia, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from under glaciers \cite<c.f.>{hallet1979,iverson2012,ugelvig2018}.
-Pressurized subglacial water can transport this sediment from underneath glaciers \cite{walder1994,creyts2013,beaud2018,delaney2019}, should enough sediment be present subglacially.
+Pressurized subglacial water can transport this sediment from underneath glaciers \cite{walder1994,creyts2013,beaud2018,delaney2019}, should adequate sediment be present subglacially.
 
@@
 The response of water velocity to changing water discharge in subglacial channels differs from subaerial ones.
-The size of subglacial channels responds to the creep closure of the ice and the opening of the channel by melting due to frictional heating of water flowing through the channel \cite{rothlisberger1972}.
-As a result, the subglacial channel size only evolves relatively slowly over days to months, whereas water discharge can vary over hours \cite<e.g.>{iken1986,andrews2014,nanni2020}.
+The size of subglacial channels is controlled on the opposing processes of channel opening by frictional heating of water flow on the one hand versus creep closure by ice flow on the other \cite{rothlisberger1972}.
+As a result, the subglacial channel size only evolves relatively slowly over days or longer, whereas water discharge can vary over hours \cite<e.g.>{iken1986,andrews2014,nanni2020}.
 Therefore, on short time scales, subglacial water flow behaves more like pipe flow, and changes in water discharge $Q$ are mainly accommodated by increased or decreased water velocity $v$ \cite<Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}; >{alley1997}.
@@
 
-Because of the above, sediment mobilization in subaerial and subglacial channels responds differently to changing water discharge.
+Sediment mobilization in subaerial and subglacial channels, therefore, responds differently to changing water discharge.
 These differences are implicitly included in a wide range of available models that  quantify sediment transport in both subglacial and subaerial channels  \cite<e.g.>{walder1994,alley1997,tucker1997,creyts2013,delaney2019,hewitt2019,wickert2019}.
-The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics near glacier margins where flow transitions from pressurized to open channel flow \cite<e.g.>{perolo2018}.
+The different response of  sediment transport capacity to changing water discharge  may well impact sediment dynamics near glacier margins where flow transitions from pressurized to open channel flow \cite<e.g.>{perolo2018}.
 Additionally, the variable response of sediment transport capacity to water discharge in the two systems may affect the interpretation of sediment transport records in glacial systems \cite<e.g.>{muller1968,richards2003,swift2005,ganti2016}.
@@
 Models of subglacial and subaerial hydraulics are used to quantify the respective water velocity and channel morphology.
-The models are applied to hydrological records from an Alpine glacier in Switzerland (Fieschergletscher) and  a land-terminating glacier in Greenland (Leverett Glacier).
-The  lumped nature of the models isolates the disparity between water discharge and sediment transport capacity in subglacial systems, independent of the upstream drainage network and sediment access.
+The models are applied to proglacial hydrological records from an Alpine glacier in Switzerland (Fieschergletscher) and  a land-terminating glacier in Greenland (Leverett Glacier).
+The  lumped nature of the models isolates the relationship between water discharge and sediment transport capacity in subglacial and subaerial systems, independent of the upstream drainage network and sediment access.
 Outputs demonstrate differences in the relationships between  water discharge, channel geometry, and water velocity in the subglacial and subaerial channels.
@@
 The Leverett Glacier in Greenland (labeled \icesheet{}) serves as the ice sheet setting.
-Water discharge was collected roughly $2$\unit{km} downstream from the terminus between June and August/September 2009 to 2012 ($67^\circ\,03'\,5''$ N, $50^\circ\,12'\,59''$ W).
-Data is used in \citeA{tedstone2017} and \citeA{hawkings2014} at a $5$\,\unit{min} time interval from May 28, 2012 to August 8, 2012.
+Water discharge was measured roughly $2$\unit{km} downstream from the terminus between June and August/September 2009 to 2012 ($67^\circ\,03'\,5''$ N, $50^\circ\,12'\,59''$ W).
+Data is used in \citeA{tedstone2013} at a $5$\,\unit{min} time interval from May 28, 2012 to August 8, 2012.
 
-Note that sediment discharge records are omitted from the analysis here given the role of sediment supply, in addition to capacity, on capturing variations in sediment discharge \cite<e.g.>{delaney2024}.
+Note that we omit analysis suspended sediment discharge records due the role of sediment supply on capturing variations in sediment discharge, in addition to sediment transport capacity \cite<e.g.>{delaney2024}.
 
@@
 Both models use the measured discharge to calculate water velocity, shear stress, and width-integrated shear stress, upon which both suspended sediment and bedload transport depend \cite<Figure \ref{fig:cartoon}; >{shields1936}.
-The evaluation of these variables omits the selection of a sediment transport relationship and a grain-size parameter \cite<e.g.>{shields1936,meyer1948}.
+The evaluation of shear stress omits the selection of a sediment transport relationship and a grain-size parameter \cite<e.g.>{shields1936,meyer1948}.
 
@@
     \label{eq:dh}
-    \Delta h \,  = l \,\frac{1}{2g} \,f_r\,\frac{v_g^2}{D_h},
+    \Delta h \,  = l \,\frac{1}{2g} \,f_r\,\frac{v_{g}^{2}}{D_h},
   \end{equation}
@@
 
-The models above are applied to proglacial discharge records from the Fieschergletscher (scenario \alpine{}) and the Leverett Glacier (scenario \icesheet{}), using subglacial model outputs that are contrasted with subaerial ones.
+The models above are applied to proglacial discharge records from the Fieschergletscher (scenario \alpine{}) and the Leverett Glacier (scenario \icesheet{}).
 Outputs of the models represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
 To generalize these scenarios, \alpine{}  is exemplified by Fieschergletscher's relatively thin ice thickness \cite<$h_{ice}$: $225$\,\unit{m}>{grab2021}, low water discharge ($\sim\,10$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
-\icesheet{}  is exemplified by Leverett glacier's  thick ice  \cite<$h_{ice}$: $700$\,\unit{m}; >{morlighem2017}, high water discharge \cite<$\sim\,300$\,\unit{m}$^3$\,\unit{s}$^{-1}$>{tedstone2017}  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).
+\icesheet{}  is exemplified by Leverett glacier's  thick ice  \cite<$h_{ice}$: $700$\,\unit{m}; >{morlighem2017}, high water discharge ($\sim\,300$\,\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).
 %However, both scenarios share a discharge with a similar, relative seasonal amplitude.
@@
   \caption{Model outputs from simulations using the hydrographs in panels a and e for the scenarios \alpine{} (a-d) and \icesheet{} (e-h).
-    Gray (black) lines in a and e represent the subglacial channel size (water discharge).
+    Black (gray) lines in a and e represent the subglacial channel size (water discharge).
     Blue (orange) lines represent outputs from the subglacial (subaerial) channel.
@@
     Arrows denote where variables peak in the subglacial channel before the subaerial one.
-    Insets in f-h show the  peak melt event denoted by the shaded area in panels e--h, but have arbitrary y-axis.
+    Insets in f-h show the  peak melt event denoted by the shaded area in panels e--h, with an arbitrary y-axis.
     % Note that the x and y axes are different for \alpine{} and \icesheet{}.
@@
 
-Following the first experiment defined in Sections~\ref{sect:imp}, model outputs exhibit different seasonal evolutions and peaks.
+In the first experiment defined in Section~\ref{sect:imp}, model outputs exhibit different seasonal evolutions and peaks.
 Water velocity and shear stress represent the potential for sediment mobilization, and width-integrated shear stress represents a proxy for the total sediment transport capacity across the channel bed (Figure~\ref{fig:model_outs}).
-Furthermore, the subglacial model outputs are generally larger than their subaerial equivalents, thus greater sediment transport capacity would be expected in subglacial channels compared to subaerial ones.
+Additionally, the subglacial model outputs are generally larger than their subaerial equivalents, thus greater sediment transport capacity would be expected in subglacial channels compared to subaerial ones.
 
@@
 
-Examining the relationship between water discharge and the model outputs, a different relationships emerge between the subaerial and subglacial cases, due to the hysteresis in channel size in the subglacial model (Equation~\ref{eq:dS_dt}).
+Examining the relationship between water discharge and the model outputs, different relationships emerge between the subaerial and subglacial cases due to the hysteresis in channel size in the subglacial model (Equation~\ref{eq:dS_dt}).
 Because there is no history dependence in  subaerial channels, each water discharge value in the subaerial channel results in a unique water velocity, shear stress, and width-integrated shear stress (Section~\ref{sect:sub_mode}).
@@
 This suggests that there is greater intermittency and variability in sediment transport in subglacial systems compared to subaerial ones, especially in supply-limited systems where sediment could be exhausted or mobilized over diurnal discharge cycles.
-In contrast, in \alpine{}, the variability in width-integrated shear stress is comparable between the subaerial and subglacial systems, due to the highly variable water discharge, largely due to diurnal variations (Figure~\ref{fig:multi_run}).
+Considering the width integrated shear stress, in \alpine{} the variability is comparable between the subaerial and subglacial systems, due to the highly variable water discharge, largely due to diurnal variations (Figure~\ref{fig:multi_run}).
 However, in \icesheet{}, with lower water discharge fluctuations, the variability in width-integrated shear stress remains consistently higher in the subglacial system across the time aggregation periods.
 
-Results from  both \alpine{} and \icesheet{} scenarios show that the variability in subglacial  decreases with aggregations longer than approximately 1--5 days (Figure~\ref{fig:multi_run}).
+Results from  both \alpine{} and \icesheet{} scenarios show that the variability in subglacial outputs decreases with aggregations longer than approximately 1--5 days (Figure~\ref{fig:multi_run}).
 Yet, substantial differences in the variability between the model outputs for subglacial and subaerial cases persist in periods of up to $15$ days, longer than which seasonal variations might be conflated.
-Note that the timing of peak sediment transport conditions still occurs, across these periods, and they are are thus not captured in aggregating water discharge over longer periods (Figure~\ref{fig:model_outs}).
+Note that across longer aggregation times, the timing of peak sediment transport conditions still occur in the natural environment.
+Yet they are not captured in model outputs(Figure~\ref{fig:model_outs}), despite their potential geomorphic importance.
 
@@
 
-Model outputs (water velocity, shear stress, and width-integrated shear stress) vary differently in response to water discharge in subglacial channels compared to subaerial ones (Figure~\ref{fig:Qw_vari}).
+The different response to water discharge in the model outputs subglacial channels compared to subaerial ones (Figure~\ref{fig:Qw_vari}) may in fact cause even greater variations in sediment transport capacity in subglacial channels than presented above.
 In sediment transport relationships that evaluate sediment discharge capacity, such as in \citeA{meyer1948} or \citeA{engelund1967}, shear stress is scaled to the power of $\frac{3}{2}$ or $\frac{5}{2}$, respectively.
 The exponent greater than $1$ magnifies sediment discharge variability beyond the variable sediment transport parameters described above (Figure~\ref{fig:multi_run}; Table~\ref{tab:Qs}).
+Variability increases further due to the application of the exponent occurs before integrating the sediment transport capacity across the channel width, as was done here in evaluating $w\tau$.
 
@@
 \ref{sect:scaling} provides a complete description of sediment transport capacity in different drainage conditions.
-Note that in pipe flow conditions, maximum sediment transport would coincide with peak water discharge and a co-varying relationship with water discharge would occur, unlike with the evolving channel size here where a disparate relationship with water discharge occurs (Section~\ref{sect:sub_mode}; Figure \ref{fig:model_outs} and \ref{fig:Qw_vari}).
+Note that in pipe flow conditions, maximum sediment transport would coincide with peak water discharge, unlike with the evolving channel size here where a disparate relationship with water discharge occurs (Section~\ref{sect:sub_mode}; Figure \ref{fig:model_outs} and \ref{fig:Qw_vari}).
 Similar adjustments in channel shape occur in subaerial channels \cite{phillips2016}, albeit at timescales longer than the subglacial channel's response time of hours to days.
 
-Substantial differences emerge in model outputs between the different hydrological forcings of the \alpine{} and \icesheet{} cases.
-Indeed, the more decoupled and sporadic relationship between model outputs and water discharge in \alpine{} results from the
-relatively larger discharge variations on daily to weekly timescales compared to \icesheet{}.
+Large catchments, or those with reduced water discharge variability, may experience yet less variability in sediment transport capacity, as the two hydrological forcings from \alpine{} and \icesheet{} show different variations in  sediment transport capacity.
+Indeed, the more decoupled and sporadic relationship between model outputs and water discharge in \alpine{} results from the relatively larger discharge variations on daily to weekly timescales compared to \icesheet{}.
 This occurs as the evolution of the R-channel's size is small compared to the variations in water discharge.
-Such water discharge variability in alpine catchments may cause shear stress to approach the water discharge squared \cite< \ref{sect:scaling}; c.f.>{alley1997}.
-With respect to \icesheet{}, a closer approximation of the hydraulics could be represented by pipe flow here, where the channel size remains fixed.
+Such water discharge variability in alpine catchments may cause shear stress to approach the water discharge squared in assuming pipe-flow conditions \cite< \ref{sect:scaling}; c.f.>{alley1997}.
 Conversely, the reduced variability in water discharge in \icesheet{} may result in the R-channel's size being closer to equilibrium concerning the variations in water discharge.
+Due the smaller water discharge variability, a closer approximation of the hydraulics could be represented by pipe flow here, where the channel size remains fixed.
 In this case, the exponent on water discharge for shear stress is likely less than $2$, but greater than $\frac{4}{5}$ occurring when the R-channel's size evolves in equilibrium with water discharge (see \ref{sect:scaling}).
-As a result, a stronger correlation persists between subglacial model outputs and water discharge in \icesheet{} compared to \alpine{} (Figure~\ref{fig:Qw_vari}).
+As a result, there is a stronger between subglacial model outputs and water discharge in \icesheet{} compared to \alpine{} (Figure~\ref{fig:Qw_vari}) due to the channel size being in closer equilibrium with water discharge in the \icesheet{} case.
 
@@
 
-The stronger link between sediment transport capacity and water discharge in subaerial channels may mean that variations in sediment transport records can be more easily attributed to hydro-climatic conditions than in subglacial systems.
+Subglacial channels exhibit far weaker relationship between sediment transport capacity and water discharge than is the case in subaerial channels (Figure~\ref{fig:Qw_vari}).
+In subaerial channels sediment transport variations in subaerial channels can often be attributed to hydro climatic conditions \cite{wolman1960}, which could be less straightforward in subglacial environments \cite{lane2019}.
 A glacier's sediment transport capacity is impacted by the ice thickness controlling the channel closure rate, and the glacier's surface slope, in addition to water discharge and sediment size \cite<Figure~\ref{fig:multi_run}, Section~\ref{sect:sub_mode}; >{rothlisberger1972,shreve1972,gimbert2016,stevens2022,walder1994}.
-In a supply-limited state, the glaciers' sediment discharge record also responds to  sediment availability and bedrock erosion from  water pressure variations and sliding  \cite<e.g.>{iverson2012,herman2015,delaney2023}.
-This multitude of processes lies in contrast to many subaerial channels, where transport capacity typically responds to sediment size, channel shape, water discharge, and hydraulic gradient.
-The hydraulic gradient and channel shape can remain stable over years to millennia \cite<Section~\ref{sect:fluv}; e.g.>{tucker1997,wickert2019}.
+In a supply-limited state, a glacier's sediment discharge record also responds to  sediment availability and bedrock erosion from  water pressure variations and sliding  \cite<e.g.>{iverson2012,herman2015,delaney2023}.
+This multitude of processes lies in contrast to many subaerial channels, where transport capacity typically responds to sediment size, channel shape, water discharge, and hydraulic gradient,
+while the hydraulic gradient and channel shape can remain stable over years to millennia \cite<Section~\ref{sect:fluv}; e.g.>{tucker1997}.
 
@@
 
-Furthermore, the greater variations in subglacial sediment transport capacity compared to subaerial ones could cause a supply-limited regime to persist at many glaciers.
+Greater variations in subglacial sediment transport capacity compared to subaerial ones could cause a supply-limited regime at many glaciers.
 In subglacial systems, sediment's critical shear stress, or threshold at which sediment mobilization occurs, can be reached more frequently and across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
-Sediment exhaustion through this process may  explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \cite{overeem2017}.
+This result suggests sediment export in these catchments depend on water discharge variability, in addition to quantity \cite{lane2019}.
+Sediment exhaustion through repeated crossing of the mobilization threshold may explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \cite{overeem2017}.
+The repeated crossing of the mobilization threshold could also result in sediment discharge's strong dependence on sediment production from sliding in many cases \cite{herman2015,koppes2015}.
 
@@
 Numerical models of subglacial sediment transport generally apply classical parameterizations of subglacial hydrology to evaluating sediment transport conditions \cite<e.g>{hewitt2019,beaud2018}.
-In these R-channel models, realistic water pressure regimes can be achieved by changing the channels shape and friction factor, therefore  model tuning is done with the friction factor and leaving the channel shape fixed as a semi-circle \cite<$\beta =\pi$; >{werder2010b}.
-However, the role of channel width in the findings here suggests that more thorough parameterization of channel shape and channel width is needed in assessing subglacial sediment transport capacity as compared to water pressure (Section\,\ref{sect:sub_mode}, \ref{sect:scaling}).
+In these R-channel models, realistic water pressure regimes can be achieved by changing the channels shape and friction factor, therefore  models are tuned with the friction factor and channel shape remains fixed as a semi-circle \cite<$\beta =\pi$; >{werder2010b}.
+However, the role of channel width in the findings here suggests that more thorough parameterization of channel shape and channel width is needed in assessing subglacial sediment transport capacity compared to water pressure (Section\,\ref{sect:sub_mode}, \ref{sect:scaling}).
 
-Water discharge measurements at the hourly timescale remain unavailable in many catchments, further limiting the ability of subglacial  sediment transport models to capture certain events.
-The model outputs' variability in subglacial channels approaches that of subaerial ones over $1-5$ days of  aggregation (Figure~\ref{fig:multi_run}), possibly resulting in parity in sediment transport capacity variations between subaerial and subglacial channels over these periods.
+Water discharge measurements at the hourly timescale remain unavailable in many catchments, further limiting the ability of subglacial  sediment transport models to capture specific events.
+The model output variability in subglacial channels approaches that of subaerial ones over $1-5$ days of  aggregation (Figure~\ref{fig:multi_run}), possibly resulting in parity in sediment transport capacity variations between subaerial and subglacial channels over these periods.
 Note that although variability between the events could remain comparable, water discharge forcings aggregated over periods longer than days may omit key sediment discharge characteristics, such as the timing of peaks in sediment transport capacity from glaciers (Figure~\ref{fig:model_outs}).
@@
 \begin{enumerate}
-\item the changing relationship between water discharge and sediment transport capacity below glaciers creates larger variability in sediment transport capacity compared to river systems;
+\item  evolving subglacial channel size causes  water discharge to have a changing impact on sediment transport capacity, complicating the relationship between sediment export and hydro-climatic conditions below glaciers;
   \vspace{.1cm}
 
-\item
-  water discharge variability impacts subglacial sediment transport capacity fluctuations to a larger degree than subaerial channels;
+\item the  variable relationship between water discharge and sediment transport capacity below glaciers creates larger variability in sediment transport capacity compared to river systems;
  \vspace{.1cm}
 
-\item the uneven response of sediment transport capacity to water discharge variations shows a different response and timing to periods of elevated water discharge in subglacial channels compared to subaerial ones.
+\item the uneven response of sediment transport capacity to water discharge variations shows a different response and timing to periods of peak discharge in subglacial channels compared to subaerial ones.
 However, representing subglacial sediment transport capacity with either open-channel flow or pipe flow neglects differences in timing.
@@
 
-This study calls for the careful analysis of the relationship between sediment transport capacity and water discharge when examining sediment discharge from glacierized catchments.
+This study calls for the careful analysis when examining the relationship between sediment transport capacity and hydro climatic conditions from glacierized catchments.
 
@@
 If a second line is given, it refers to the range of values examined in the parameter search}
-  \small % TODO remove again
+  \small 
   \begin{tabular}{ l  c  c c }

### 2024-04-16 Delaneyetal2024.tex
push to overleaf

@@
 \begin{keypoints}
-\item Short water discharge fluctuations in subglacial channels are accommodated by water velocity, not channel size as in subaerial systems.
+\item Rapid water discharge fluctuations in subglacial channels are accommodated by water velocity, not channel size as in subaerial systems.
 \item Greater variability in water velocity causes sediment transport capacity to vary more in subglacial channels than in subaerial channels.
-\item Significant subglacial sediment transport capacity can persist across a range of water discharges, impacting the  timing of peak events.
+\item Significant subglacial sediment transport capacity can persist across a large range of water discharges, impacting  timing of peak events.
 \end{keypoints}
@@
   Here, hydrographs from an Alpine glacier and the Greenland Ice Sheet are applied to subglacial and subaerial hydraulics models, revealing greater variability in the inferred sediment transport capacity in subglacial than in subaerial channels.
-  Results also show that high subglacial sediment transport rates and water velocities can occur across wide-ranging water discharges.
+  Results also show that high subglacial sediment transport rates and water velocities can occur across a wide range of water discharges.
   These findings are discussed in the context of understanding sediment discharge from glaciers with different hydro-climatic forcings and assumptions of channel shape.
@@
 
-Over long periods, such as millennia, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from under glaciers \cite<c.f.>{hallet1979,iverson2012,ugelvig2018}.
-Pressurized subglacial water can transport this sediment from underneath glaciers \cite{walder1994,creyts2013,beaud2018,delaney2019}, should adequate sediment be present subglacially.
+Periods of millennia or longer, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from under glaciers \cite<c.f.>{hallet1979,iverson2012,ugelvig2018}.
+Pressurized subglacial water can transport this sediment from underneath glaciers \cite{walder1994,creyts2013,beaud2018,delaney2019}, if it is reachable by the water.
 
@@
 In subaerial channels, operating with open channel flow, the wetted area $S$ of the channel evolves with changing water discharge $Q$, by changing both the channel's width and the water depth \cite{leopold1953}.
-The change in water depth results in a proportional increase in water velocity.
-As a result, the shear stress $\tau$ increases according to Equation~\ref{eq:tau}.
+The change in water depth results in a proportional increase in water velocity, and  the shear stress $\tau$ increases according to Equation~\ref{eq:tau}.
 
@@
 The size of subglacial channels is controlled on the opposing processes of channel opening by frictional heating of water flow on the one hand versus creep closure by ice flow on the other \cite{rothlisberger1972}.
-As a result, the subglacial channel size only evolves relatively slowly over days or longer, whereas water discharge can vary over hours \cite<e.g.>{iken1986,andrews2014,nanni2020}.
-Therefore, on short time scales, subglacial water flow behaves more like pipe flow, and changes in water discharge $Q$ are mainly accommodated by increased or decreased water velocity $v$ \cite<Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}; >{alley1997}.
+As a result, the subglacial channel size only evolves relatively slowly over days, whereas water discharge can vary more quickly over hours \cite<e.g.>{iken1986,andrews2014,nanni2020}.
+Therefore, on short periods, subglacial water flow behaves more like pipe flow, and changes in water discharge $Q$ are mainly accommodated by changing water velocity $v$ \cite<Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}; >{alley1997}.
 
@@
 Sediment mobilization in subaerial and subglacial channels, therefore, responds differently to changing water discharge.
-These differences are implicitly included in a wide range of available models that  quantify sediment transport in both subglacial and subaerial channels  \cite<e.g.>{walder1994,alley1997,tucker1997,creyts2013,delaney2019,hewitt2019,wickert2019}.
-The different response of  sediment transport capacity to changing water discharge  may well impact sediment dynamics near glacier margins where flow transitions from pressurized to open channel flow \cite<e.g.>{perolo2018}.
-Additionally, the variable response of sediment transport capacity to water discharge in the two systems may affect the interpretation of sediment transport records in glacial systems \cite<e.g.>{muller1968,richards2003,swift2005,ganti2016}.
+These differences are implicitly included in a range of models quantifying sediment transport in both subglacial and subaerial channels  \cite<e.g.>{walder1994,alley1997,tucker1997,creyts2013,delaney2019,hewitt2019,wickert2019}.
+Additionally, the variable response of sediment transport capacity to water discharge in the two systems may affect the interpretation of sediment transport records in glacial systems \cite<e.g.>{muller1968,richards2003,swift2005,ganti2016,pitlick2021}.
 Yet, little work has examined the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial channels \cite{alley1997}.
@@
 The Alpine site (labeled \alpine{}) is  Fieschergletscher in the Swiss Alps ($46^\circ\,29'\,07''$ N, $8^\circ\,08'\,3''$ E).
-The water discharge data used here was collected at a $1$\unit{min} interval from May 24, 2014, to October 10, 2014 \cite{felix2021,felix2022}.
+The water discharge data used here was collected at a $1$\unit{min} interval from May 24, 2014, to October 10, 2014 \cite{felix2022}.
 
 The Leverett Glacier in Greenland (labeled \icesheet{}) serves as the ice sheet setting.
-Water discharge was measured roughly $2$\unit{km} downstream from the terminus between June and August/September 2009 to 2012 ($67^\circ\,03'\,5''$ N, $50^\circ\,12'\,59''$ W).
-Data is used in \citeA{tedstone2013} at a $5$\,\unit{min} time interval from May 28, 2012 to August 8, 2012.
+Water discharge was measured roughly $2$\unit{km} downstream from the terminus ($67^\circ\,03'\,5''$ N, $50^\circ\,12'\,59''$ W), at a $5$\,\unit{min} time interval from May 28, 2012 to August 8, 2012 \citeA{tedstone2013}.
 
@@
 \label{sect:meth}
-The models described below (Sections ~\ref{sect:sub_mode}~and~\ref{sect:fluv}) represent relationships amongst water discharge, water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
+The two models described below (Sections ~\ref{sect:sub_mode}~and~\ref{sect:fluv}) represent relationships amongst water discharge, water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
 Both models use the measured discharge to calculate water velocity, shear stress, and width-integrated shear stress, upon which both suspended sediment and bedload transport depend \cite<Figure \ref{fig:cartoon}; >{shields1936}.
-The evaluation of shear stress omits the selection of a sediment transport relationship and a grain-size parameter \cite<e.g.>{shields1936,meyer1948}.
+Our evaluation in terms of shear stress omits the selection of a sediment transport relationship and a grain-size parameter \cite<e.g.>{shields1936,meyer1948}.
 
@@
 
-Here, it is assumed that the water is transported through a subglacial channel \cite<Figure~\ref{fig:cartoon}; >{rothlisberger1972} beneath a glacier with channel length $l$, with a flatbed and a mean thickness of $h_{ice}$.
+Here, it is assumed that the water is transported through a subglacial channel \cite<Figure~\ref{fig:cartoon}; >{rothlisberger1972} beneath a glacier with channel length $l$, with a flat bed and a mean thickness of $h_{ice}$.
 The channel  size grows from melt due to frictional heating from water flow and closes due to the creep  of the ice.
@@
     \label{eq:dh}
-    \Delta h \,  = l \,\frac{1}{2g} \,f_r\,\frac{v_{g}^{2}}{D_h},
+    \Delta h \,  = l \,\frac{1}{2g} \,f_i\,\frac{v_{g}^{2}}{D_h},
   \end{equation}
@@
 Outputs of the models represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
-To generalize these scenarios, \alpine{}  is exemplified by Fieschergletscher's relatively thin ice thickness \cite<$h_{ice}$: $225$\,\unit{m}>{grab2021}, low water discharge ($\sim\,10$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
-\icesheet{}  is exemplified by Leverett glacier's  thick ice  \cite<$h_{ice}$: $700$\,\unit{m}; >{morlighem2017}, high water discharge ($\sim\,300$\,\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).
-%However, both scenarios share a discharge with a similar, relative seasonal amplitude.
+To generalize these scenarios, \alpine{}  is exemplified by Fieschergletscher's relatively thin ice thickness \cite<$h_{ice}$= $225$\,\unit{m}>{grab2021}, low water discharge ($\sim\,10$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
+\icesheet{}  is exemplified by Leverett glacier's  thick ice  \cite<$h_{ice}$= $700$\,\unit{m}; >{morlighem2017}, high water discharge ($\sim\,300$\,\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).
+% However, both scenarios share a discharge with a similar, relative seasonal amplitude.
 
-In the first experiment, the models are  applied to a reference test case for each glacier that assumes  a subglacial channel with $\beta=\frac{\pi}{6}$%, i.e. a segment of a quarter-circle,
+In the first experiment, the models are  applied to a reference test case for each glacier that assumes  a subglacial channel with $\beta=\frac{\pi}{6}$
 (Table~\ref{table:vpm}).
 Both friction factors $f_r$ and $f_f$ are tuned so that reasonable water velocities \cite<$\sim\,1.6\,$\unit{m}\,\unit{s}$^{-1}$>{werder2010b,chandler2013} occur for both \alpine{} and \icesheet{}.
-The initial condition of cross-sectional area $S_g$ is evaluated by applying a model spinup with the maximum observed water discharge for the first $4$ days of the study until there is no change in channel area.
+The initial condition of cross-sectional area $S_g$ is evaluated by applying a model spinup with repeatedly applying the maximum observed water discharge of the first $4$ days of the study until there is no change in channel area.
 The subaerial channels of both scenarios have a bed slope of $\nabla z_c = 0.05$ (Table~\ref{table:vpm}).
@@
 For brevity, these values together are referred to as ``model outputs'' in the text.
-To quantify the link between sediment transport capacity and water discharge, model outputs are compared to water discharge from the glaciers using Spearman rank correlation, which varies with the ordering of values.
+To quantify the link between sediment transport capacity and water discharge, model outputs are compared to water discharge from the glaciers using Spearman rank correlation, which accounts the ordering of values, but not their magnitude.
 Rank correlation reduces the impact of the non-linear response of sediment transport capacity to the hydrological forcing.
 
-A second experiment aims to further characterize the variability in sediment discharge capacity in subglacial compared to subaerial channels across a range of channel shapes, friction factors, and water discharge variability.
-To accomplish this, we examine the variability of one model output (i.e water velocity) by evaluating its standard deviation throughout the season.
+The second experiment aims to further characterize the variability in sediment discharge capacity in subglacial compared to subaerial channels across a range of channel shapes and friction factors.
+Additionally, we examine the effects of different water discharge variability by smoothing the water discharge over different periods. 
+To accomplish this, we examine the variability of each model output by evaluating its standard deviation throughout the season.
 
@@
 Runs with parameter combinations are accepted if their mean subglacial water velocity over the season lies between $0.75$\,\unit{m}\,\unit{s}$^{-1}$ and $2$\,\unit{m}\,\unit{s}$^{-1}$ or if subaerial water velocity lies between $0.3$\,\unit{m}\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}\,\unit{s}$^{-1}$, following measurements \cite<e.g.>{werder2010b,magnusson2012,chandler2013}.
-Additionally, to be accepted, model runs can only experience a flotation fraction ($\frac{\Delta h}{2\,h_o}$) of above $1.2$ for less than $2.5$\,\% of the run.
+Additionally, to be accepted, subglacial model runs can only experience a flotation fraction ($\frac{\Delta h}{2\,h_o}$) of above $1.2$ for less than $2.5$\,\% of the run.
 
-The routine runs until $100$ different parameter combinations for each water discharge aggregation period are accepted with the conditions described above.
-Here, the water discharge input at a given time is the average over the aggregation period before that time, ranging from $15$\,\unit{min} up to $15$\,\unit{days}.
+The routine runs until $100$ different parameter combinations for each water discharge smoothing period are accepted with the conditions described above.
+Here, the water discharge input at a given time is the average over the smoothing period before that time, ranging from $15$\,\unit{min} up to $15$\,\unit{days}.
 
@@
   \centering
-  \includegraphics[width=0.9\linewidth]{Fig2.pdf}
+  \includegraphics[width=0.9\linewidth]{Fig2.png}
   \caption{Model outputs from simulations using the hydrographs in panels a and e for the scenarios \alpine{} (a-d) and \icesheet{} (e-h).
@@
     Data are shown at $15$\unit{min} intervals.
-    Arrows denote where variables peak in the subglacial channel before the subaerial one.
+    Arrows show examples where variables peak in the subglacial channel before the subaerial one.
     Insets in f-h show the  peak melt event denoted by the shaded area in panels e--h, with an arbitrary y-axis.
@@
   \centering
-  \includegraphics[width=0.9\linewidth]{Fig3.pdf}
+  \includegraphics[width=0.9\linewidth]{Fig3.png}
   \caption{
@@
 
-Examining the relationship between water discharge and the model outputs, different relationships emerge between the subaerial and subglacial cases due to the hysteresis in channel size in the subglacial model (Equation~\ref{eq:dS_dt}).
+Examining the relationship between water discharge and the model outputs, different relationships emerge for the subaerial and subglacial cases due to the hysteresis in channel size in the subglacial model (Equation~\ref{eq:dS_dt}).
 Because there is no history dependence in  subaerial channels, each water discharge value in the subaerial channel results in a unique water velocity, shear stress, and width-integrated shear stress (Section~\ref{sect:sub_mode}).
@@
 Conversely, the history dependence on channel size in subglacial channels means different outputs occur across a range of discharges.
-For instance, in subglacial channels in \alpine{}, very high seasonal water velocity values and shear stresses can occur from a low water discharge  ($\sim\,4$\,\unit{m}$^3$\,\unit{s}$^{-1}$) to the maximum water discharge at over ($17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, Figure~\ref{fig:Qw_vari}\,a).
+For instance, in subglacial channels in \alpine{}, very high seasonal water velocity values and shear stresses can occur from a low water discharge  ($\sim\,4$\,\unit{m}$^3$\,\unit{s}$^{-1}$) to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ (Figure~\ref{fig:Qw_vari}\,a).
 In \icesheet{}, water velocities close to the seasonal mean value can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
@@
 
-The different response of subaerial and subglacial model outputs to water discharge impacts the timing of peaks in sediment transport capacity.
+The different response of subaerial and subglacial model outputs to water discharge impacts the timing of peaks in sediment transport capacity, and thus its  relationship with water discharge.
 In the subglacial channel, peaks in $w_g\tau_g$ generally occur when water discharge increases at the fastest rate, but before the maximum water discharge (Figure~\ref{fig:model_outs}\, d and h).
@@
   \includegraphics[width=0.9\linewidth]{Fig4.pdf}
-    \caption{Standard deviation of model outputs for different time aggregations with different subglacial and subaerial channel shapes and friction factors.
-      Shaded areas denote the range of results from the accepted $100$ parameter combinations (Section~\ref{sect:imp}).
-      Solid lines denote  mean values.
-      Markers show aggregation periods ($15$\,\unit{min}, $1$\,\unit{hr}, $12$\,\unit{hr}, $1$\,\unit{d}, $5$\,\unit{d}, $10$\,\unit{d}, and $15$\,\unit{d}).
+  \caption{Standard deviation of model outputs for different smoothing times with different subglacial and subaerial channel shapes and friction factors.
+    Shaded areas denote the range of standard deviations from the accepted $100$ parameter combinations (Section~\ref{sect:imp}).
+    Solid lines denote  mean value of standard deviations.
+    Markers show smoothing periods ($15$\,\unit{min}, $1$\,\unit{hr}, $12$\,\unit{hr}, $1$\,\unit{d}, $5$\,\unit{d}, $10$\,\unit{d}, and $15$\,\unit{d}).
   }
@@
 
-The second numerical experiment aims to evaluate the role of channel shape and friction factors on the findings above and to examine the effect of variability in water discharge.
-Across the range of channel shapes and friction factors  examined (Section\,\ref{sect:imp}), variability in water velocity and shear stress remains higher in the subglacial system compared to the subaerial one in both \alpine{} and \icesheet{}.
+The second numerical experiment aims to evaluate the response of variability in model outputs to a range of channel shapes, friction factors, and water discharge variability.
+Across the range of parameters  examined (Section\,\ref{sect:imp}), variability in water velocity and shear stress remains higher in the subglacial system compared to the subaerial one in both \alpine{} and \icesheet{}.
 This suggests that there is greater intermittency and variability in sediment transport in subglacial systems compared to subaerial ones, especially in supply-limited systems where sediment could be exhausted or mobilized over diurnal discharge cycles.
-Considering the width integrated shear stress, in \alpine{} the variability is comparable between the subaerial and subglacial systems, due to the highly variable water discharge, largely due to diurnal variations (Figure~\ref{fig:multi_run}).
-However, in \icesheet{}, with lower water discharge fluctuations, the variability in width-integrated shear stress remains consistently higher in the subglacial system across the time aggregation periods.
+Considering the width integrated shear stress, in \alpine{} the variability is comparable between the subaerial and subglacial systems (shown by overlap of plots), due to the highly variable water discharge, largely due to diurnal variations (Figure~\ref{fig:multi_run}).
+In contrast, in \icesheet{}, with lower water discharge fluctuations, the variability in width-integrated shear stress remains consistently higher in the subglacial system across the time smoothing periods.
 
-Results from  both \alpine{} and \icesheet{} scenarios show that the variability in subglacial outputs decreases with aggregations longer than approximately 1--5 days (Figure~\ref{fig:multi_run}).
+Results from  both \alpine{} and \icesheet{} scenarios show that the variability in subglacial outputs decreases with smoothing time longer than approximately 1--5 days (Figure~\ref{fig:multi_run}).
 Yet, substantial differences in the variability between the model outputs for subglacial and subaerial cases persist in periods of up to $15$ days, longer than which seasonal variations might be conflated.
-Note that across longer aggregation times, the timing of peak sediment transport conditions still occur in the natural environment.
+Note that across longer smoothing times, the timing of peak sediment transport conditions still occur in the natural environment.
 Yet they are not captured in model outputs(Figure~\ref{fig:model_outs}), despite their potential geomorphic importance.
@@
 
-The different response to water discharge in the model outputs subglacial channels compared to subaerial ones (Figure~\ref{fig:Qw_vari}) may in fact cause even greater variations in sediment transport capacity in subglacial channels than presented above.
+Increased variability in subglacial sediment discharge capacity makes it more difficult to differentiate the role of long term climatic controls on sediment transport with shorter term hydrological variations.
+The different response to water discharge variations in the model outputs subglacial channels compared to subaerial ones (Figure~\ref{fig:Qw_vari}) may in fact cause even greater variations in sediment transport capacity in subglacial channels than presented above.
 In sediment transport relationships that evaluate sediment discharge capacity, such as in \citeA{meyer1948} or \citeA{engelund1967}, shear stress is scaled to the power of $\frac{3}{2}$ or $\frac{5}{2}$, respectively.
 The exponent greater than $1$ magnifies sediment discharge variability beyond the variable sediment transport parameters described above (Figure~\ref{fig:multi_run}; Table~\ref{tab:Qs}).
-Variability increases further due to the application of the exponent occurs before integrating the sediment transport capacity across the channel width, as was done here in evaluating $w\tau$.
 
-Sediment transport capacity in subglacial channels responds more rapidly to changes in water discharge compared to their subaerial counterparts \cite<Figures~\ref{fig:Qw_vari}~and~\ref{fig:multi_run}>{alley1997}.
-In subglacial R\"othlisberger channels in steady state with water discharge, the width-integrated shear stress response to water discharge is $ w_g\,\tau_{g}\, \propto\,  Q^{\sim\frac{4}{5}}$ (Table\,\ref{tab:eqs1}).
-However, subglacial channels rarely operate in a steady state \cite<e.g.>{gimbert2016}.
-For sufficiently short periods, channel width and wetted area are fixed in the subglacial system (akin to pipe flow) so that the response of width-integrated shear stress is $w_g\, \tau_{g}\, \propto\,  Q^{2}$ (Table\,\ref{tab:eqs1}).
+For sufficiently short periods, channel width and wetted area are fixed in the subglacial system (akin to pipe flow) so that the response of width-integrated shear stress is $w_g\, \tau_{g}\, \propto\,  Q^{2}$ \cite<Table\,\ref{tab:eqs1}; >{alley1997}.
 \ref{sect:scaling} provides a complete description of sediment transport capacity in different drainage conditions.
-Note that in pipe flow conditions, maximum sediment transport would coincide with peak water discharge, unlike with the evolving channel size here where a disparate relationship with water discharge occurs (Section~\ref{sect:sub_mode}; Figure \ref{fig:model_outs} and \ref{fig:Qw_vari}).
-Similar adjustments in channel shape occur in subaerial channels \cite{phillips2016}, albeit at timescales longer than the subglacial channel's response time of hours to days.
+Note that in pipe flow conditions, maximum sediment transport would coincide with peak water discharge, unlike here where evolving channel size here alters the timing (Figure \ref{fig:model_outs} and \ref{fig:Qw_vari}).
+R\"othlisberger channels in steady state with water discharge, the width-integrated shear stress response to water discharge is $ w_g\,\tau_{g}\, \propto\,  Q^{\sim\frac{4}{5}}$ (Table\,\ref{tab:eqs1}).
+This relationship suggests that when discharge varies more slowly than channel shape subglacial channels in fact have less variable sediment transport capacity than subaerial ones.
+Yet, subglacial channels rarely operate in steady state \cite{gimbert2016}.
 
-Large catchments, or those with reduced water discharge variability, may experience yet less variability in sediment transport capacity, as the two hydrological forcings from \alpine{} and \icesheet{} show different variations in  sediment transport capacity.
-Indeed, the more decoupled and sporadic relationship between model outputs and water discharge in \alpine{} results from the relatively larger discharge variations on daily to weekly timescales compared to \icesheet{}.
-This occurs as the evolution of the R-channel's size is small compared to the variations in water discharge.
+The channel width and size variations in subglacial channels presented here also occur in subaerial channels in response to water discharge \cite{phillips2016}, likely over decadal timescales or in response to individual extreme events \cite<>{slater2013,dean2013}.
+These timescales are considerably longer than the continuous changes to subglacial channel's response time of days.
+Furthermore, some observations from subaerial systems show a co-varying relationship between water discharge and sediment transport in time \cite<e.g. >{schmidt2006,pitlick2021}.
+Even so, if channel width changes occur in subaerial channels, then the values of $\alpha$ in Equation~\ref{eq:wcf} could change in time.
+$\alpha$ must be less than $1$, and even this would cause the exponent on $Q$ in the  $w_g\,\tau_{g}$ relationships to remain substantially below that of pipeflow (\ref{sect:scaling}, Table\,\ref{tab:Qs} and \ref{tab:Qs2})).
+
+Catchments with reduced water discharge variability may experience yet less variability in sediment transport capacity, as the two hydrological forcings from \alpine{} and \icesheet{} show different variations in  sediment transport capacity.
+Indeed, the more decoupled and sporadic relationship between model outputs and water discharge in \alpine{} results from the relatively larger discharge variations on daily to weekly timescales compared to \icesheet{} (Figure~\ref{fig:Qw_vari}).
+This occurs as the evolution of the R-channel's size is slow compared to the variations in water discharge.
 Such water discharge variability in alpine catchments may cause shear stress to approach the water discharge squared in assuming pipe-flow conditions \cite< \ref{sect:scaling}; c.f.>{alley1997}.
-Conversely, the reduced variability in water discharge in \icesheet{} may result in the R-channel's size being closer to equilibrium concerning the variations in water discharge.
-Due the smaller water discharge variability, a closer approximation of the hydraulics could be represented by pipe flow here, where the channel size remains fixed.
+Conversely, the reduced variability in water discharge in \icesheet{} may result in the R-channel's size being closer to equilibrium from the smaller variations in water discharge.
+Due the smaller water discharge variability, pipe flow with fixed channel size could more closely approximate hydraulics here.
 In this case, the exponent on water discharge for shear stress is likely less than $2$, but greater than $\frac{4}{5}$ occurring when the R-channel's size evolves in equilibrium with water discharge (see \ref{sect:scaling}).
@@
 
+Water discharge measurements at the hourly timescale remain unavailable in many catchments, further limiting subglacial  sediment transport models in capturing specific events.
+The model output variability in subglacial channels approaches that of subaerial ones over $1-5$ days of smoothing (Figure~\ref{fig:multi_run}), possibly resulting in parity in sediment transport capacity variations between subaerial and subglacial channels.
+However, this smoothing omits key sediment discharge characteristics, such as the timing of peaks in sediment transport capacity from glaciers (Figure~\ref{fig:model_outs}).
+
 \subsection{Effect sediment transport and landscape evolution}
 
-Subglacial channels exhibit far weaker relationship between sediment transport capacity and water discharge than is the case in subaerial channels (Figure~\ref{fig:Qw_vari}).
-In subaerial channels sediment transport variations in subaerial channels can often be attributed to hydro climatic conditions \cite{wolman1960}, which could be less straightforward in subglacial environments \cite{lane2019}.
-A glacier's sediment transport capacity is impacted by the ice thickness controlling the channel closure rate, and the glacier's surface slope, in addition to water discharge and sediment size \cite<Figure~\ref{fig:multi_run}, Section~\ref{sect:sub_mode}; >{rothlisberger1972,shreve1972,gimbert2016,stevens2022,walder1994}.
+Subglacial channels exhibit far weaker relationship between sediment transport capacity and water discharge than in subaerial channels (Figure~\ref{fig:Qw_vari}).
+In subaerial channels, sediment transport variations in subaerial channels can be attributed to hydro climatic conditions \cite{wolman1960}
+A glacier's sediment transport capacity is impacted by the ice thickness controlling the channel closure rate, and the glacier's surface slope, in addition to water discharge and sediment size \cite<Figure~\ref{fig:multi_run}, Section~\ref{sect:sub_mode}; >{rothlisberger1972,gimbert2016,stevens2022,walder1994}.
 In a supply-limited state, a glacier's sediment discharge record also responds to  sediment availability and bedrock erosion from  water pressure variations and sliding  \cite<e.g.>{iverson2012,herman2015,delaney2023}.
-This multitude of processes lies in contrast to many subaerial channels, where transport capacity typically responds to sediment size, channel shape, water discharge, and hydraulic gradient,
-while the hydraulic gradient and channel shape can remain stable over years to millennia \cite<Section~\ref{sect:fluv}; e.g.>{tucker1997}.
+This multitude of processes lies in contrast to many subaerial channels, where transport capacity typically responds to water discharge in addition to sediment size, channel shape,  and hydraulic gradient, that can remain stable over years to millennia \cite<Section~\ref{sect:fluv}; e.g.>{tucker1997}.
+As a result, characteristics such as bankfull or effective water discharge could have limited meaning in evaluating subglacial sediment transport \cite{wolman1960,lenzi2006}.
 
-Previous work suggests that this co-varying relationship between sediment discharge capacity in  subaerial conditions could be noticeable $\sim20$\,\unit{km} downstream of the Leverett site, where a strong correlation persists between sediment plume size and the river's water discharge into the Kangerlussuaq fjord \cite{mcgrath2010}.
+Previous work suggests that this co-varying relationship between sediment discharge capacity and water discharge in  subaerial conditions could be noticeable $\sim20$\,\unit{km} downstream of the Leverett site, where a strong correlation persists between sediment plume size and the river's water discharge into the Kangerlussuaq fjord \cite{mcgrath2010}.
 In contrast, in marine-terminating glacier catchments, a less consistent relationship may occur between water discharge or melt extent and sediment plume size \cite{tedstone2012}.
@@
 In subglacial systems, sediment's critical shear stress, or threshold at which sediment mobilization occurs, can be reached more frequently and across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
-This result suggests sediment export in these catchments depend on water discharge variability, in addition to quantity \cite{lane2019}.
+This result suggests sediment export here is especially sensitive to water discharge variability, in addition to quantity \cite{lane2019b}.
 Sediment exhaustion through repeated crossing of the mobilization threshold may explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \cite{overeem2017}.
-The repeated crossing of the mobilization threshold could also result in sediment discharge's strong dependence on sediment production from sliding in many cases \cite{herman2015,koppes2015}.
-
-\subsection{Parameterizing  sediment transport in glacierized catchments}
-
-Numerical models of subglacial sediment transport generally apply classical parameterizations of subglacial hydrology to evaluating sediment transport conditions \cite<e.g>{hewitt2019,beaud2018}.
-In these R-channel models, realistic water pressure regimes can be achieved by changing the channels shape and friction factor, therefore  models are tuned with the friction factor and channel shape remains fixed as a semi-circle \cite<$\beta =\pi$; >{werder2010b}.
-However, the role of channel width in the findings here suggests that more thorough parameterization of channel shape and channel width is needed in assessing subglacial sediment transport capacity compared to water pressure (Section\,\ref{sect:sub_mode}, \ref{sect:scaling}).
-
-Water discharge measurements at the hourly timescale remain unavailable in many catchments, further limiting the ability of subglacial  sediment transport models to capture specific events.
-The model output variability in subglacial channels approaches that of subaerial ones over $1-5$ days of  aggregation (Figure~\ref{fig:multi_run}), possibly resulting in parity in sediment transport capacity variations between subaerial and subglacial channels over these periods.
-Note that although variability between the events could remain comparable, water discharge forcings aggregated over periods longer than days may omit key sediment discharge characteristics, such as the timing of peaks in sediment transport capacity from glaciers (Figure~\ref{fig:model_outs}).
+The repeated crossing of the mobilization threshold across a range of discharges could also result in sediment discharge's strong dependence on sediment production from sliding in many cases \cite{herman2015,koppes2015}.
 
@@
 
-The sediment transport capacity of a channel is driven by its width and the shear stress exerted by the flowing water, which is proportional to the square of the flow velocity.
+The sediment transport capacity of a channel is driven by its width and the shear stress exerted by the flowing water, which is proportional to  flow velocity squared.
 Subaerial channels alter both their channel width and water velocity immediately in response to changing water discharge.
@@
 
-\item the  variable relationship between water discharge and sediment transport capacity below glaciers creates larger variability in sediment transport capacity compared to river systems;
+\item the  variable relationship between water discharge and sediment transport capacity below glaciers creates larger variability in sediment transport capacity compared to subaerial systems;
   \vspace{.1cm}
 
-\item the uneven response of sediment transport capacity to water discharge variations shows a different response and timing to periods of peak discharge in subglacial channels compared to subaerial ones.
-However, representing subglacial sediment transport capacity with either open-channel flow or pipe flow neglects differences in timing.
+\item the timing of peak sediment discharge capacity in subglacial channels occurs prior to peak  water discharge, whereas it is coincident in subaerial ones, limiting the meaning of certain water discharge characteristics in subglacial systems.
+  Representing subglacial sediment transport capacity with either open-channel flow or pipe flow neglects differences in timing.
 \end{enumerate}
 
-This study calls for the careful analysis when examining the relationship between sediment transport capacity and hydro climatic conditions from glacierized catchments.
+This study calls for the careful analysis when examining the relationship between sediment transport capacity and hydro-climatic conditions from glacierized catchments.
 
@@
 
-\appendix
-
-\section{Table of variables, parameters and constants}
+\section{Supplement}
+\subsection{Table of variables, parameters and constants}
 \begin{table}[h]
@@
 
-\section{Figure 3 over various time aggregations}
+\subsection{Figure 3 over various smoothing times}
 
@@
     \includegraphics[width=0.7\linewidth]{Fig3_hr.png}
-    \caption{As Figure \ref{fig:model_outs}, with $1$ \,\unit{hr} aggregation.}
+    \caption{As Figure \ref{fig:model_outs}, with $1$ \,\unit{hr} smoothing.}
     \label{fig:model_outs_1hr}
@@
     \includegraphics[width=0.7\linewidth]{Fig3_12hr.png}
-    \caption{As Figure \ref{fig:model_outs}, with $12$ \,\unit{hr} aggregation.}
+    \caption{As Figure \ref{fig:model_outs}, with $12$ \,\unit{hr} smoothing.}
     \label{fig:model_outs_12hr}
@@
     \includegraphics[width=0.7\linewidth]{Fig3_1day.png}
-    \caption{As Figure \ref{fig:model_outs}, with $1$ \,\unit{day} aggregation.}
+    \caption{As Figure \ref{fig:model_outs}, with $1$ \,\unit{day} smoothing.}
     \label{fig:model_outs_1day}
@@
     \includegraphics[width=0.7\linewidth]{Fig3_5day.png}
-    \caption{As Figure \ref{fig:model_outs}, with $5$ \,\unit{day} aggregation.}
+    \caption{As Figure \ref{fig:model_outs}, with $5$ \,\unit{day} smoothing.}
     \label{fig:model_outs_5day}
@@
     \includegraphics[width=0.7\linewidth]{Fig3_10day.png}
-    \caption{As Figure \ref{fig:model_outs}, with $10$ \,\unit{day} aggregation.}
+    \caption{As Figure \ref{fig:model_outs}, with $10$ \,\unit{day} smoothing.}
     \label{fig:model_outs_10day}
@@
     \includegraphics[width=0.7\linewidth]{Fig3_15day.png}
-    \caption{As Figure \ref{fig:model_outs}, with $15$ \,\unit{day} aggregation.}
+    \caption{As Figure \ref{fig:model_outs}, with $15$ \,\unit{day} smoothing.}
     \label{fig:model_outs_15day}
@@
 \newpage
-\section{Formulas for sediment transport in different drainage regimes}
+\subsection{Formulas for sediment transport in different drainage regimes}
 \label{sect:scaling}
 
-%\mauro{Notation: either use $h$ and $w$ for the river, or otherwise, somehow free $h$. Below I use $H$ and $W$ for now.}
+% \mauro{Notation: either use $h$ and $w$ for the river, or otherwise, somehow free $h$. Below I use $H$ and $W$ for now.}
 
@@
 \mauro{I don't understand the last two sentences.}
-%Furthermore evaluating the different variability in sediment transport characteristics across timescales can help inform the sediment transport processes omitted from daily or week values of water discharge..
+% Furthermore evaluating the different variability in sediment transport characteristics across timescales can help inform the sediment transport processes omitted from daily or week values of water discharge..

### 2024-04-16 Delaneyetal2024.tex
changes from overleaf

@@
 \item Greater variability in water velocity causes sediment transport capacity to vary more in subglacial channels than in subaerial channels.
-\item Significant subglacial sediment transport capacity can persist across a large range of water discharges, impacting  timing of peak events.
+\item Significant subglacial sediment transport capacity can persist across a large range of water discharges, impacting  the timing of peak events.
 \end{keypoints}
@@
   In subaerial channels, water discharge variations are accommodated  by flow depth and width changes, along with water velocity.
-  However, in subglacial channels  water discharge changes mainly result in velocity variations, as the channel geometry evolves slowly (over days).
+  However, in subglacial channels,  water discharge changes mainly result in velocity variations, as the channel geometry evolves slowly (over days).
   Here, hydrographs from an Alpine glacier and the Greenland Ice Sheet are applied to subglacial and subaerial hydraulics models, revealing greater variability in the inferred sediment transport capacity in subglacial than in subaerial channels.
@@
 The response of water velocity to changing water discharge in subglacial channels differs from subaerial ones.
-The size of subglacial channels is controlled on the opposing processes of channel opening by frictional heating of water flow on the one hand versus creep closure by ice flow on the other \cite{rothlisberger1972}.
+The size of subglacial channels is controlled by the opposing processes of channel opening by frictional heating of water flow on the one hand versus creep closure by ice flow on the other \cite{rothlisberger1972}.
 As a result, the subglacial channel size only evolves relatively slowly over days, whereas water discharge can vary more quickly over hours \cite<e.g.>{iken1986,andrews2014,nanni2020}.
@@
 
-Note that we omit analysis suspended sediment discharge records due the role of sediment supply on capturing variations in sediment discharge, in addition to sediment transport capacity \cite<e.g.>{delaney2024}.
+Note that we omit analysis of suspended sediment discharge records due to the role of sediment supply in capturing variations in sediment discharge, in addition to sediment transport capacity \cite<e.g.>{delaney2024}.
 
@@
 Note that this subaerial channel model is purely algebraic, whereas the subglacial model comprises a differential equation for the evolution of $S_g$.
-Thus the channel size in the subaerial model has no history dependence on the discharge $Q_w$, where as the subglacial one does (Equation~\ref{eq:dS_dt}).
+Thus the channel size in the subaerial model has no history dependence on the discharge $Q_w$, whereas the subglacial one does (Equation~\ref{eq:dS_dt}).
 
@@
 Both friction factors $f_r$ and $f_f$ are tuned so that reasonable water velocities \cite<$\sim\,1.6\,$\unit{m}\,\unit{s}$^{-1}$>{werder2010b,chandler2013} occur for both \alpine{} and \icesheet{}.
-The initial condition of cross-sectional area $S_g$ is evaluated by applying a model spinup with repeatedly applying the maximum observed water discharge of the first $4$ days of the study until there is no change in channel area.
+The initial condition of cross-sectional area $S_g$ is evaluated by applying a model spinup with repeatedly applying the maximum observed water discharge of the first $4$ days of the study until there is no change in the channel area.
 The subaerial channels of both scenarios have a bed slope of $\nabla z_c = 0.05$ (Table~\ref{table:vpm}).
@@
 For brevity, these values together are referred to as ``model outputs'' in the text.
-To quantify the link between sediment transport capacity and water discharge, model outputs are compared to water discharge from the glaciers using Spearman rank correlation, which accounts the ordering of values, but not their magnitude.
+To quantify the link between sediment transport capacity and water discharge, model outputs are compared to water discharge from the glaciers using Spearman rank correlation, which accounts for the ordering of values, but not their magnitude.
 Rank correlation reduces the impact of the non-linear response of sediment transport capacity to the hydrological forcing.
@@
     Shaded areas denote the range of standard deviations from the accepted $100$ parameter combinations (Section~\ref{sect:imp}).
-    Solid lines denote  mean value of standard deviations.
+    Solid lines denote  the mean value of standard deviations.
     Markers show smoothing periods ($15$\,\unit{min}, $1$\,\unit{hr}, $12$\,\unit{hr}, $1$\,\unit{d}, $5$\,\unit{d}, $10$\,\unit{d}, and $15$\,\unit{d}).
@@
 This suggests that there is greater intermittency and variability in sediment transport in subglacial systems compared to subaerial ones, especially in supply-limited systems where sediment could be exhausted or mobilized over diurnal discharge cycles.
-Considering the width integrated shear stress, in \alpine{} the variability is comparable between the subaerial and subglacial systems (shown by overlap of plots), due to the highly variable water discharge, largely due to diurnal variations (Figure~\ref{fig:multi_run}).
+Considering the width integrated shear stress, in \alpine{} the variability is comparable between the subaerial and subglacial systems (shown by the overlap of plots), due to the highly variable water discharge, largely due to diurnal variations (Figure~\ref{fig:multi_run}).
 In contrast, in \icesheet{}, with lower water discharge fluctuations, the variability in width-integrated shear stress remains consistently higher in the subglacial system across the time smoothing periods.
@@
 Yet, substantial differences in the variability between the model outputs for subglacial and subaerial cases persist in periods of up to $15$ days, longer than which seasonal variations might be conflated.
-Note that across longer smoothing times, the timing of peak sediment transport conditions still occur in the natural environment.
+Note that across longer smoothing times, the timing of peak sediment transport conditions still occurs in the natural environment.
 Yet they are not captured in model outputs(Figure~\ref{fig:model_outs}), despite their potential geomorphic importance.
@@
 
-Increased variability in subglacial sediment discharge capacity makes it more difficult to differentiate the role of long term climatic controls on sediment transport with shorter term hydrological variations.
-The different response to water discharge variations in the model outputs subglacial channels compared to subaerial ones (Figure~\ref{fig:Qw_vari}) may in fact cause even greater variations in sediment transport capacity in subglacial channels than presented above.
+Increased variability in subglacial sediment discharge capacity makes it more difficult to differentiate the role of long-term climatic controls on sediment transport with shorter-term hydrological variations.
+The different response to water discharge variations in the model outputs subglacial channels compared to subaerial ones (Figure~\ref{fig:Qw_vari}) may actually cause even greater variations in sediment transport capacity in subglacial channels than presented above.
 In sediment transport relationships that evaluate sediment discharge capacity, such as in \citeA{meyer1948} or \citeA{engelund1967}, shear stress is scaled to the power of $\frac{3}{2}$ or $\frac{5}{2}$, respectively.
@@
 R\"othlisberger channels in steady state with water discharge, the width-integrated shear stress response to water discharge is $ w_g\,\tau_{g}\, \propto\,  Q^{\sim\frac{4}{5}}$ (Table\,\ref{tab:eqs1}).
-This relationship suggests that when discharge varies more slowly than channel shape subglacial channels in fact have less variable sediment transport capacity than subaerial ones.
-Yet, subglacial channels rarely operate in steady state \cite{gimbert2016}.
+This relationship suggests that when discharge varies more slowly than channel shape subglacial channels have less variable sediment transport capacity than subaerial ones.
+Yet, subglacial channels rarely operate in a steady state \cite{gimbert2016}.
 
 The channel width and size variations in subglacial channels presented here also occur in subaerial channels in response to water discharge \cite{phillips2016}, likely over decadal timescales or in response to individual extreme events \cite<>{slater2013,dean2013}.
-These timescales are considerably longer than the continuous changes to subglacial channel's response time of days.
+These timescales are considerably longer than the continuous changes to the subglacial channel's response time of days.
 Furthermore, some observations from subaerial systems show a co-varying relationship between water discharge and sediment transport in time \cite<e.g. >{schmidt2006,pitlick2021}.
@@
 Conversely, the reduced variability in water discharge in \icesheet{} may result in the R-channel's size being closer to equilibrium from the smaller variations in water discharge.
-Due the smaller water discharge variability, pipe flow with fixed channel size could more closely approximate hydraulics here.
+Due to the smaller water discharge variability, pipe flow with fixed channel size could more closely approximate hydraulics here.
 In this case, the exponent on water discharge for shear stress is likely less than $2$, but greater than $\frac{4}{5}$ occurring when the R-channel's size evolves in equilibrium with water discharge (see \ref{sect:scaling}).
@@
 
-Subglacial channels exhibit far weaker relationship between sediment transport capacity and water discharge than in subaerial channels (Figure~\ref{fig:Qw_vari}).
-In subaerial channels, sediment transport variations in subaerial channels can be attributed to hydro climatic conditions \cite{wolman1960}
+Subglacial channels exhibit a far weaker relationship between sediment transport capacity and water discharge than in subaerial channels (Figure~\ref{fig:Qw_vari}).
+In subaerial channels, sediment transport variations in subaerial channels can be attributed to hydro-climatic conditions \cite{wolman1960}
 A glacier's sediment transport capacity is impacted by the ice thickness controlling the channel closure rate, and the glacier's surface slope, in addition to water discharge and sediment size \cite<Figure~\ref{fig:multi_run}, Section~\ref{sect:sub_mode}; >{rothlisberger1972,gimbert2016,stevens2022,walder1994}.
@@
 
-\item the timing of peak sediment discharge capacity in subglacial channels occurs prior to peak  water discharge, whereas it is coincident in subaerial ones, limiting the meaning of certain water discharge characteristics in subglacial systems.
+\item the timing of peak sediment discharge capacity in subglacial channels occurs before peak  water discharge, whereas it is coincident in subaerial ones, limiting the meaning of certain water discharge characteristics in subglacial systems.
   Representing subglacial sediment transport capacity with either open-channel flow or pipe flow neglects differences in timing.
@@
 
-This study calls for the careful analysis when examining the relationship between sediment transport capacity and hydro-climatic conditions from glacierized catchments.
+This study calls for careful analysis when examining the relationship between sediment transport capacity and hydro-climatic conditions from glacierized catchments.
 

### 2024-06-27 Delaneyetal2024.tex
reworking

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
-% 
-% This file includes commands and instructions
-% given in the order necessary to produce a final output that will
-% satisfy AGU requirements, including customized APA reference formatting.
-% 
-% You may copy this file and give it your
-% article name, and enter your text.
-% 
-% guidelines and troubleshooting are here:
+%% Copernicus Publications Manuscript Preparation Template for LaTeX Submissions
+%% ---------------------------------
+%% This template should be used for copernicus.cls
+%% The class file and some style files are bundled in the Copernicus Latex Package, which can be downloaded from the different journal webpages.
+%% For further assistance please contact Copernicus Publications at: production@copernicus.org
+%% https://publications.copernicus.org/for_authors/manuscript_preparation.html
 
-%% To submit your paper:
-\documentclass[draft]{agujournal2019}
-\usepackage{url} %this package should fix any errors with URLs in refs.
-\usepackage{lineno}
-\usepackage{placeins}
-\usepackage{amsmath}
+%% Please use the following documentclass and journal abbreviations for preprints and final revised papers.
+
+%% 2-column papers and preprints
+\documentclass[esurf, manuscript]{copernicus}
+
 \usepackage{makecell}
-
-\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
-\usepackage{soul}
-\linenumbers
-%% make line numbers work with amsmath
-% https://tex.stackexchange.com/questions/461186/how-to-use-lineno-with-amsmath-align
-\usepackage{lineno}   %% <- no mathlines option
-\usepackage{amsmath}  %% <- after lineno
-\usepackage{etoolbox} %% <- for \cspreto, \csappto
-
-%% Patch 'normal' math environments:
-\newcommand*\linenomathpatch[1]{%
-  \cspreto{#1}{\linenomath}%
-  \cspreto{#1*}{\linenomath}%
-  \csappto{end#1}{\endlinenomath}%
-  \csappto{end#1*}{\endlinenomath}%
-}
-
-\linenomathpatch{equation}
-\linenomathpatch{gather}
-\linenomathpatch{multline}
-\linenomathpatch{align}
-\linenomathpatch{alignat}
-\linenomathpatch{flalign}
-
-%%%%%%% 
-% As of 2018 we recommend use of the TrackChanges package to mark revisions.
-% The trackchanges package adds five new LaTeX commands:
-% 
-% \note[editor]{The note}
-% \annote[editor]{Text to annotate}{The note}
-% \add[editor]{Text to add}
-% \remove[editor]{Text to remove}
-% \change[editor]{Text to remove}{Text to add}
-% 
-% complete documentation is here: http://trackchanges.sourceforge.net/
-%%%%%%%
-
-\draftfalse
+\usepackage{float}
+\usepackage[section]{placeins}
 
 \newcommand{\aref}[1]{\textbf{Reference #1}}
+\newcommand{\afig}[1]{\textbf{Figure #1}}
 \newcommand{\TODO}[1]{\textbf{TODO: \color{red}#1}}
-\newcommand{\ian}[1]{{\textbf{\color{blue}Ian says:} \color{blue} #1} }
+\newcommand{\ian}[1]{{\textbf{\color{blue}Ian sayz:} \color{blue} #1} }
 \newcommand{\mauro}[1]{{\textbf{\color{green}Mauro says:} \color{green} #1} }
@@
 \newcommand{\mmms}{$\,\mathrm{m^3\, s^{-1}}$\,}
-\newcommand{\unit}[1]{$\mathrm{#1}$}
-
-\journalname{Geophysical Research Letters}
+%\newcommand{\unit}[1]{$\mathrm{#1}$}
 
@@
 
-\authors{Ian Delaney\affil{1},  Andrew J. Tedstone\affil{2},   Mauro A. Werder\affil{3,4},  Daniel Farinotti\affil{3,4} }
+% \Author[affil]{given_name}{surname}
 
-\affiliation{1}{Institut des dynamiques de la surface terrestre (IDYST), Universit\'{e} de Lausanne, B\^{a}timent G\'{e}opolis, 1015 Lausanne, Switzerland}
-\affiliation{2}{Department of Geosciences, University of Fribourg, Ch. du Musée 1700, Fribourg, Switzerland}
-\affiliation{3}{Laboratory of Hydraulics, Hydrology and Glaciology (VAW), ETH-Z\"urich, H\"onggerbergring 26, 8093 Z\"urich, Switzerland}
-\affiliation{4}{Swiss Federal Institute for Forest, Snow and Landscape Research (WSL) Z\"uricherstrasse 111, 8903 Birmensdorf, Switzerland}
+\Author[1][IanArburua.Delaney@unil.ch]{Ian}{Delaney} %% correspondence author
+\Author[2]{Andrew J.}{Tedstone}
+\Author[3,4]{Mauro A.}{Werder}
+\Author[3,4]{Daniel}{Farinotti}
 
-\correspondingauthor{Ian Delaney}{ianarburua.delaney@unil.ch}
+\affil[1]{Institut des dynamiques de la surface terrestre (IDYST), Universit\'{e} de Lausanne, B\^{a}timent G\'{e}opolis, 1015 Lausanne, Switzerland}
+\affil[2]{Department of Geosciences, University of Fribourg, Ch. du Musée 1700, Fribourg, Switzerland}
+\affil[3]{Laboratory of Hydraulics, Hydrology and Glaciology (VAW), ETH-Z\"urich, H\"onggerbergring 26, 8093 Z\"urich, Switzerland}
+\affil[4]{Swiss Federal Institute for Forest, Snow and Landscape Research (WSL) Z\"uricherstrasse 111, 8903 Birmensdorf, Switzerland}
+%% The [] brackets identify the author with the corresponding affiliation. 1, 2, 3, etc. should be inserted.
 
-\begin{keypoints}
-\item Rapid water discharge fluctuations in subglacial channels are accommodated by water velocity, not channel size as in subaerial systems.
-\item Greater variability in water velocity causes sediment transport capacity to vary more in subglacial channels than in subaerial channels.
-\item Significant subglacial sediment transport capacity can persist across a large range of water discharges, impacting  the timing of peak events.
-\end{keypoints}
+%% If an author is deceased, please mark the respective author name(s) with a dagger, e.g. "\Author[2,$\dag$]{Anton}{Smith}", and add a further "\affil[$\dag$]{deceased, 1 July 2019}".
+
+%% If authors contributed equally, please mark the respective author names with an asterisk, e.g.
+"\Author[2,*]{Anton}{Smith}" and "\Author[3,*]{Bradley}{Miller}" and add a further affiliation: "\affil[*]{These authors contributed equally to this work.}".
+
+\runningtitle{Subglacial sediment transport capacity}
+
+\runningauthor{Delaney et al.}
+\received{}
+\pubdiscuss{} %% only important for two-stage journals
+\revised{}
+\accepted{}
+\published{}
+
+%% These dates will be inserted by Copernicus Publications during the typesetting process.
+
+\firstpage{1}
+
+\maketitle
 
@@
   However, in subglacial channels,  water discharge changes mainly result in velocity variations, as the channel geometry evolves slowly (over days).
-  Here, hydrographs from an Alpine glacier and the Greenland Ice Sheet are applied to subglacial and subaerial hydraulics models, revealing greater variability in the inferred sediment transport capacity in subglacial than in subaerial channels.
-  Results also show that high subglacial sediment transport rates and water velocities can occur across a wide range of water discharges.
+  Here, we present relationships between water discharge and sediment transport under several channel types.
+  They demonstrate that sediment transport is highly non-linear with respect to water discharge with fixed subglacial channel size, and in fact can very less that subaerial flow if water discharge changes slowly enough to be in equilibrium with subglacial channel size.
+  We then apply hydrographs from an Alpine glacier and the Greenland Ice Sheet to subglacial and subaerial hydraulics models, revealing greater variability in the inferred sediment transport capacity in subglacial than in subaerial channels, as subglacial channels evolve in size.
+  This variability stems from the slowly evolving channel size that forces water velocity to accommodate variations in water discharge.
+  Furthermore, the changing channel size means that sediment transport capacity peaks prior to the maximum water discharge.
+  Results also show that high subglacial sediment transport rates and water velocities can occur across a wide range of water discharges, meaning that in a transport-limited system there could be a variable relationship between sediment and water discharge.
   These findings are discussed in the context of understanding sediment discharge from glaciers with different hydro-climatic forcings and assumptions of channel shape.
+  We also examine the implications for interpreting records of sediment discharge in glacierized catchments
 \end{abstract}
 
-\section*{Plain Language Summary} % 200 words
-Sediment export from glaciers impacts downstream ecosystems and communities and has changed in response to climate warming.
-Sediment export depends on two main factors: sediment availability, and the capacity of the hydrological system to transport that sediment.
-This capacity largely depends on the width of the channel and the water's velocity.
-In subaerial channels, like rivers, water velocity, and channel width both respond rapidly to water discharge.
-However, subglacial channels are pressurized by the ice above, so water flows through a subglacial pipe that evolves in size only slowly.
-Variations in water discharge and thus also sediment transport capacity are therefore mainly accommodated by changes in the water velocity.
-Models of these processes show that sediment transport capacity is highly variable in subglacial channels, and a wide range of water discharges can result in similar sediment transport capacities.
-In subaerial channels, by contrast, sediment transport capacity is less variable, and variations depend solely on water discharge, not the channel size.
-These findings may impact the interpretation of sediment transport records from glacierized catchments, complicating the link between glacier melt and sediment discharge.
+\introduction  %% \introduction[modified heading if necessary]
 
-\section{Introduction}
-\label{sect:intro}
+Changes in glacier dynamics and hydrology  have prompted numerous  recent studies of  sediment transport processes in cold regions \citep[e.g.][]{li2022,vergara2022,zhang2022}.
+Increases in sediment transport  have been observed in Greenland \citep{bendixen2017}, the European Alps \citep{costa2017}, the Himalayas \citep{li2021}, and the Andes \citep{vergara2022}.
+To explain observed changes to sediment transport in glacierized catchments requires examining the processes controlling sediment discharge and its variations with water discharge  \citep[e.g.][]{riihimaki2005,swift2005}.
 
-Changes in glacier dynamics and hydrology  have prompted numerous  recent studies of  sediment transport processes in cold regions \cite<e.g.>{li2022,vergara2022,zhang2022}.
-Increases in sediment transport  have been observed in Greenland \cite{bendixen2017}, the European Alps \cite{costa2017}, the Himalayas \cite{li2021}, and the Andes \cite{vergara2022}.
-In some regions, increased water discharge and glacier melt have been interpreted to yield greater sediment transport capacity \cite{bendixen2017,costa2017,li2021}.
-To explain observed changes to sediment transport in glacierized catchments requires examining the processes controlling sediment discharge and its variations with water discharge  \cite<e.g.>{riihimaki2005,swift2005}.
-
-Periods of millennia or longer, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from under glaciers \cite<c.f.>{hallet1979,iverson2012,ugelvig2018}.
-Pressurized subglacial water can transport this sediment from underneath glaciers \cite{walder1994,creyts2013,beaud2018,delaney2019}, if it is reachable by the water.
+Periods of millennia or longer, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from under glaciers \citep[c.f.][]{hallet1979,iverson2012,ugelvig2018}.
+Pressurized subglacial water can transport this sediment from underneath glaciers \citep{walder1994,creyts2013,beaud2018,delaney2019}, if it is reachable by the water.
 
 In a transport-limited regime, sediment discharge is controlled by sediment transport capacity, defined as the amount of sediment the water can carry.
-In both subglacial and subaerial channels, sediment transport capacity depends on the shear stress between water and the sediment it flows  over \cite{shields1936,meyer1948,engelund1967} along with the width of the channel bottom $w$ over which sediment mobilizes.
+In both subglacial and subaerial channels, sediment transport capacity depends on the shear stress between water and the sediment it flows  over \citep{shields1936,meyer1948,engelund1967} along with the width of the channel bottom $w$ over which sediment mobilizes.
 The shear stress $\tau$ responds to the velocity of water $v$ flowing through the channel so that
@@
 
-In subaerial channels, operating with open channel flow, the wetted area $S$ of the channel evolves with changing water discharge $Q$, by changing both the channel's width and the water depth \cite{leopold1953}.
+In subaerial channels, operating with open channel flow, the wetted area $S$ of the channel evolves with changing water discharge $Q$, by changing both the channel's width and the water depth \citep{leopold1953}.
 The change in water depth results in a proportional increase in water velocity, and  the shear stress $\tau$ increases according to Equation~\ref{eq:tau}.
@@
 The response of water velocity to changing water discharge in subglacial channels differs from subaerial ones.
-The size of subglacial channels is controlled by the opposing processes of channel opening by frictional heating of water flow on the one hand versus creep closure by ice flow on the other \cite{rothlisberger1972}.
-As a result, the subglacial channel size only evolves relatively slowly over days, whereas water discharge can vary more quickly over hours \cite<e.g.>{iken1986,andrews2014,nanni2020}.
-Therefore, on short periods, subglacial water flow behaves more like pipe flow, and changes in water discharge $Q$ are mainly accommodated by changing water velocity $v$ \cite<Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}; >{alley1997}.
+The size of subglacial channels is controlled by the opposing processes of channel opening by frictional heating of water flow on the one hand versus creep closure by ice flow on the other \citep{rothlisberger1972}.
+As a result, the subglacial channel size only evolves relatively slowly over days, whereas water discharge can vary more quickly over hours \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
+Therefore, on short periods, subglacial water flow behaves more like pipe flow, and changes in water discharge $Q$ are mainly accommodated by changing water velocity $v$ \citep[Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}; ][]{alley1997}.
 
-\begin{figure}[h]
+\begin{figure}[hbt!]
   \centering
-  \includegraphics[width=0.8\linewidth]{Fig1.pdf}
+  \includegraphics[width=0.6\linewidth]{Fig1.pdf}
   \caption{Sketch for the different responses of subglacial and subaerial channels to increased water discharge over short time-scales.
@@
 Sediment mobilization in subaerial and subglacial channels, therefore, responds differently to changing water discharge.
-These differences are implicitly included in a range of models quantifying sediment transport in both subglacial and subaerial channels  \cite<e.g.>{walder1994,alley1997,tucker1997,creyts2013,delaney2019,hewitt2019,wickert2019}.
-Additionally, the variable response of sediment transport capacity to water discharge in the two systems may affect the interpretation of sediment transport records in glacial systems \cite<e.g.>{muller1968,richards2003,swift2005,ganti2016,pitlick2021}.
-Yet, little work has examined the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial channels \cite{alley1997}.
+These differences are implicitly included in a range of models quantifying sediment transport in both subglacial and subaerial channels  \citep[e.g.][]{walder1994,alley1997,tucker1997,creyts2013,beaud2018,delaney2019,hewitt2019,wickert2019}.
+To date, explicit parameterizations of subglacial sediment transport capacity with respect to water discharge has assumed fixed channel size and demonstrates a very non-linear response in sediment transport capacity to water discharge \citep{alley1997}.
+The modeling frameworks examining subglacial sediment transport with evolving channel size minimally discuss the impact of different hydrological forcing on sediment transport capacity with the aim of understanding the impact of hydro-climatic conditions on subglacial sediment dynamics \citep{creyts2013,beaud2018,delaney2019,hewitt2019}.
+Yet, sediment discharge capacity controls the mobilization and deposition of sediment, making it one of the most fundamental aspects of subglacial sediment evacuation, especially over short timescales.
 
-This manuscript evaluates the relationship between sediment transport capacity and water discharge in subglacial channels and compares it to subaerial ones.
-Models of subglacial and subaerial hydraulics are used to quantify the respective water velocity and channel morphology.
-The models are applied to proglacial hydrological records from an Alpine glacier in Switzerland (Fieschergletscher) and  a land-terminating glacier in Greenland (Leverett Glacier).
-The  lumped nature of the models isolates the relationship between water discharge and sediment transport capacity in subglacial and subaerial systems, independent of the upstream drainage network and sediment access.
-Outputs demonstrate differences in the relationships between  water discharge, channel geometry, and water velocity in the subglacial and subaerial channels.
-The manuscript then uses the demonstrated differences between the two channel types to discuss the implications for interpreting the processes driving variability in sediment transport records and their relationship with hydrology from glacierized catchments.
+Despite the physical differences between the subaerial and subglacial systems, water discharge and sediment export are often compared in contemporary glacial environments \citep[e.g.][]{willis1996,pearce2003,richards2003,swift2005,chu2009,tedstone2012,chu2012,overeem2017,delaney2018,swift2021,lu2022,andresen2024}.
+These comparisons are natural given that the majority of sediment evacuated from glaciers over contemporary seasonal timescales is done so fluvially \citep[e.g.][]{collins1989, riihimaki2005,delaney2024}.
+In subaerial systems, water discharge and sediment transport capacity for a grain size are closely linked \citep[e.g. ][]{schmidt2006,pitlick2021} \ian{check citiation}.
+However, the potential for a highly variable response in sediment transport capacity in subglacial sediment transport may affect the interpretation of sediment transport records in glacial systems \citep[e.g.][]{ganti2016,mancini2023}, if properties associate with subaerial systems are assumed.
+This relationship should be clarified as sediment dynamics change in glacierized regions along with changing hydrology, increased variability of water discharge \citep{lane2019} and extreme glacier melt  \citep[e.g.][]{cremona2023,overeem2015} that may drive high magnitude sediment transport events.
+
+This manuscript as two objectives: 1) evaluate the variability of subglacial sediment transport capacity with evolving subglacial channels size and different hydrographs and 2) establish if subseasonal variations in water discharge can be covary with sediment transport capacity in subglacial systems.
+First, to reenforce the highly variable behavior of subglacial sediment transport capacity,  we present algebric formulations of the sediment transport capacity response to water discharge in subaerial, pipe-flow and equilibrium R-channel conditions, extending the relationships presented in \citet{alley1997}.
+We then use numerical models to examine  sediment transport capacity in subglacial channels that evolve in size and subaerial channels.
+They are then applied to proglacial hydrological records from an Alpine glacier in Switzerland (Fieschergletscher) and  a land-terminating glacier in Greenland (Leverett Glacier).
+We run models with an ensemble of parameters to evaluate the sediment transport capacity variability in subglacial channels compared to subaerial ones.
+Lastly, model outputs demonstrate the particular processes that can drive variability in sediment discharge capacity in subglacial systems.
+Together, these findings indicate differences in the relationships between  water discharge, channel geometry, and water velocity in the subglacial and subaerial channels.
+The manuscript then uses the identified differences between the subaerial and subglacial channel types to discuss the implications for interpreting the processes driving variability in sediment transport records and the impact of these differences on sediment export from glaciers.
 
@@
 The Alpine site (labeled \alpine{}) is  Fieschergletscher in the Swiss Alps ($46^\circ\,29'\,07''$ N, $8^\circ\,08'\,3''$ E).
-The water discharge data used here was collected at a $1$\unit{min} interval from May 24, 2014, to October 10, 2014 \cite{felix2022}.
+The water discharge data used here was collected at a $1$\unit{min} interval from May 24, 2014, to October 10, 2014 \citep[Figure~\ref{fig:Qw}\,a][]{felix2022}.
 
 The Leverett Glacier in Greenland (labeled \icesheet{}) serves as the ice sheet setting.
-Water discharge was measured roughly $2$\unit{km} downstream from the terminus ($67^\circ\,03'\,5''$ N, $50^\circ\,12'\,59''$ W), at a $5$\,\unit{min} time interval from May 28, 2012 to August 8, 2012 \citeA{tedstone2013}.
+Water discharge was measured roughly $2$\unit{km} downstream from the terminus ($67^\circ\,03'\,5''$ N, $50^\circ\,12'\,59''$ W), at a $5$\,\unit{min} time interval from May 28, 2012 to August 8, 2012 \citet[][Figure~\ref{fig:Qw}\,b]{tedstone2013}.
 
-Note that we omit analysis of suspended sediment discharge records due to the role of sediment supply in capturing variations in sediment discharge, in addition to sediment transport capacity \cite<e.g.>{delaney2024}.
+Note that we omit analysis of suspended sediment discharge records due to the role of sediment supply in capturing variations in sediment discharge, in addition to sediment transport capacity \citep[e.g.][]{delaney2019}.
+
+\begin{figure}[hbt!]
+  \centering
+  \includegraphics[width=0.8\linewidth]{Qw_plot.pdf}
+  \caption{Water discharge from Fieschergletcher (\alpine) and Leverett Glacier (\icesheet).
+    Note differences in variability and quantity of water discharge over the season.
+  }
+  \label{fig:Qw}
+\end{figure}
 
@@
 \label{sect:meth}
-The two models described below (Sections ~\ref{sect:sub_mode}~and~\ref{sect:fluv}) represent relationships amongst water discharge, water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
-Both models use the measured discharge to calculate water velocity, shear stress, and width-integrated shear stress, upon which both suspended sediment and bedload transport depend \cite<Figure \ref{fig:cartoon}; >{shields1936}.
-Our evaluation in terms of shear stress omits the selection of a sediment transport relationship and a grain-size parameter \cite<e.g.>{shields1936,meyer1948}.
+The two models described below (Sections ~\ref{sect:sub_mode}~and~\ref{sect:fluv}) represent relationships amongst water discharge, water velocity, and channel geometry in both subaerial and subglacial channels (Table~\ref{table:vpm}).
+Both models use the measured discharge to calculate water velocity, shear stress, and width-integrated shear stress, upon which both suspended sediment and bedload transport depend \citep[Figure \ref{fig:cartoon}; ][]{shields1936}.
+Our evaluation in terms of shear stress omits the selection of a sediment transport relationship and a grain-size parameter \citep[e.g.][]{shields1936,meyer1948}.
+Through this choice we also negate sorting processes
 
@@
 
-To evaluate the shear stress of water flowing across sediments underneath a glacier, the subglacial channel model accounts for the channel geometry and the velocity of the flowing water.
-To do this, we use a lumped hydraulics model from \citeA{werder2010b}, itself based upon  \citeA{clarke1996}.
+To evaluate the shear stress of water flowing across sediments underneath a glacier, the subglacial channel model accounts for the channel geometry and the water's velocity.
+To do this, we use a lumped hydraulics model from \citet{werder2010b}, itself based upon  \citet{clarke1996}.
 
-Here, it is assumed that the water is transported through a subglacial channel \cite<Figure~\ref{fig:cartoon}; >{rothlisberger1972} beneath a glacier with channel length $l$, with a flat bed and a mean thickness of $h_{ice}$.
+Here, it is assumed that the water is transported through a subglacial channel \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972} beneath a glacier with channel length $l$, with a flat bed and a mean thickness of $h_{ice}$.
 The channel  size grows from melt due to frictional heating from water flow and closes due to the creep  of the ice.
@@
 \end{linenomath*}
-\noindent where $t$ is time, $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (complete values given in Table \ref{table:vpm}), $g$ is the acceleration due to gravity, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change over $l$, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the mean ice overburden pressure expressed in meter water equivalent ($\rho_w$ is density of water; $\rho_i$ is density of ice), and $n$ is Glen's n \cite{glen1955} (usually $n=3$).
-The first term on the right side of the equation represents the opening of the channel through frictional heating, while the following term represents the creep closure of the channel from ice deformation.
+\noindent where $t$ is time, $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (complete values given in Table~\ref{table:vpm}), $g$ is the acceleration due to gravity, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change over $l$, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the mean ice overburden pressure expressed in meter water equivalent ($\rho_w$ is density of water; $\rho_i$ is density of ice), and $n$ is Glen's n \citep{glen1955} (usually $n=3$).
+The first term on the equation's right side represents the opening of the channel through frictional heating, while the following term represents the creep closure of the channel from ice deformation.
 
@@
 \end{equation}
-where $\beta$ is the central angle of the circular segment that comprises the channel (the so-called Hooke angle, \citeA{hooke1990}).
-Note that $\beta =\pi$ corresponds to a semi-circular channel and smaller values of $\beta$ result in shallow, wide channels.
+where $\beta$ is the central angle of the circular segment that comprises the channel (the Hooke angle, \citet{hooke1990}).
+$\beta =\pi$ corresponds to a semi-circular channel and smaller values of $\beta$ result in shallow, wide channels.
 This completes the subglacial hydraulic model which is described by the state variables $S_g$ and $\Delta h$.
 
-The shear stress, $\tau_g$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
+The shear stress, $\tau_g$, between the water and the channel bed results from the Darcy-Weisbach formulation
 \begin{equation}
@@
 
-% To parameterize the shear stress of water flowing across sediments in the subaerial channel, the hydraulics parameterization presented in \citeA{tucker1997} is implemented.
+The hydraulics parameterization presented in \citet{tucker1997} is implemented to represent the subaerial channel.
 Using mass conservation and the Darcy-Weisbach relationship as well as assuming that the channel is sufficiently wide compared to its depth so that the hydraulic radius is well approximated by the flow depth,
@@
 \end{linenomath*}
-where $\nabla z_c$ is the channel slope, and $f_f$ is the friction factor for subaerial channels \cite{tucker1997}.
+where $\nabla z_c$ is the channel slope, and $f_f$ is the friction factor for subaerial channels \citep{tucker1997}.
 Channel width $w_f$ is
@@
 % 
-where $k$ is a constant and $\alpha=\frac{1}{2}$ is a commonly chosen exponent \cite{leopold1953}.
+where $k$ is a constant and $\alpha=\frac{1}{3}$ is a commonly chosen exponent \citep{leopold1953}.
 Also following the Darcy-Weisbach, subaerial water velocity, $v_f$, is given as
@@
 
-\subsection{Implementation}
-\label{sect:imp}
-
-The models above are applied to proglacial discharge records from the Fieschergletscher (scenario \alpine{}) and the Leverett Glacier (scenario \icesheet{}).
-Outputs of the models represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
-To generalize these scenarios, \alpine{}  is exemplified by Fieschergletscher's relatively thin ice thickness \cite<$h_{ice}$= $225$\,\unit{m}>{grab2021}, low water discharge ($\sim\,10$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
-\icesheet{}  is exemplified by Leverett glacier's  thick ice  \cite<$h_{ice}$= $700$\,\unit{m}; >{morlighem2017}, high water discharge ($\sim\,300$\,\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).
-% However, both scenarios share a discharge with a similar, relative seasonal amplitude.
-
-In the first experiment, the models are  applied to a reference test case for each glacier that assumes  a subglacial channel with $\beta=\frac{\pi}{6}$
-(Table~\ref{table:vpm}).
-Both friction factors $f_r$ and $f_f$ are tuned so that reasonable water velocities \cite<$\sim\,1.6\,$\unit{m}\,\unit{s}$^{-1}$>{werder2010b,chandler2013} occur for both \alpine{} and \icesheet{}.
-The initial condition of cross-sectional area $S_g$ is evaluated by applying a model spinup with repeatedly applying the maximum observed water discharge of the first $4$ days of the study until there is no change in the channel area.
-The subaerial channels of both scenarios have a bed slope of $\nabla z_c = 0.05$ (Table~\ref{table:vpm}).
-
-The results below present water velocity ($v_g$, $v_f$), shear stress ($\tau_g$, $\tau_f$), and width-integrated shear stress ($w_g\tau_g$, $w_f\tau_f$) from the subglacial and subaerial models.
-For brevity, these values together are referred to as ``model outputs'' in the text.
-To quantify the link between sediment transport capacity and water discharge, model outputs are compared to water discharge from the glaciers using Spearman rank correlation, which accounts for the ordering of values, but not their magnitude.
-Rank correlation reduces the impact of the non-linear response of sediment transport capacity to the hydrological forcing.
-
-The second experiment aims to further characterize the variability in sediment discharge capacity in subglacial compared to subaerial channels across a range of channel shapes and friction factors.
-Additionally, we examine the effects of different water discharge variability by smoothing the water discharge over different periods. 
-To accomplish this, we examine the variability of each model output by evaluating its standard deviation throughout the season.
-
-The effect of channel shape and roughness is established by running the model with random parameter values of channel geometry factors ($\beta$ and $k$) and friction factors ($f_r$ and $f_f$; see Table~\ref{table:vpm} for value range).
-Runs with parameter combinations are accepted if their mean subglacial water velocity over the season lies between $0.75$\,\unit{m}\,\unit{s}$^{-1}$ and $2$\,\unit{m}\,\unit{s}$^{-1}$ or if subaerial water velocity lies between $0.3$\,\unit{m}\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}\,\unit{s}$^{-1}$, following measurements \cite<e.g.>{werder2010b,magnusson2012,chandler2013}.
-Additionally, to be accepted, subglacial model runs can only experience a flotation fraction ($\frac{\Delta h}{2\,h_o}$) of above $1.2$ for less than $2.5$\,\% of the run.
-
-The routine runs until $100$ different parameter combinations for each water discharge smoothing period are accepted with the conditions described above.
-Here, the water discharge input at a given time is the average over the smoothing period before that time, ranging from $15$\,\unit{min} up to $15$\,\unit{days}.
-
-\section{Results}
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=0.9\linewidth]{Fig2.png}
-  \caption{Model outputs from simulations using the hydrographs in panels a and e for the scenarios \alpine{} (a-d) and \icesheet{} (e-h).
-    Black (gray) lines in a and e represent the subglacial channel size (water discharge).
-    Blue (orange) lines represent outputs from the subglacial (subaerial) channel.
-    Data are shown at $15$\unit{min} intervals.
-    Arrows show examples where variables peak in the subglacial channel before the subaerial one.
-    Insets in f-h show the  peak melt event denoted by the shaded area in panels e--h, with an arbitrary y-axis.
-    % Note that the x and y axes are different for \alpine{} and \icesheet{}.
-  }
-  \label{fig:model_outs}
-\end{figure}
-
-In the first experiment defined in Section~\ref{sect:imp}, model outputs exhibit different seasonal evolutions and peaks.
-Water velocity and shear stress represent the potential for sediment mobilization, and width-integrated shear stress represents a proxy for the total sediment transport capacity across the channel bed (Figure~\ref{fig:model_outs}).
-Additionally, the subglacial model outputs are generally larger than their subaerial equivalents, thus greater sediment transport capacity would be expected in subglacial channels compared to subaerial ones.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=0.9\linewidth]{Fig3.png}
-  \caption{
-    Relationship between water discharge and normalized velocity, shear stress, and width-integrated shear stress for \alpine{} (a-c) and \icesheet{} (d-f).
-    Variables on the x-axis have been normalized to mean values.
-    $R_g$ ($R_f$) shows the Spearman rank correlations for the subglacial (subaerial) outputs.
-    Plots are shown with discharge and outputs at $15$\unit{min} intervals.
-  }
-  \label{fig:Qw_vari}
-\end{figure}
-
-Examining the relationship between water discharge and the model outputs, different relationships emerge for the subaerial and subglacial cases due to the hysteresis in channel size in the subglacial model (Equation~\ref{eq:dS_dt}).
-Because there is no history dependence in  subaerial channels, each water discharge value in the subaerial channel results in a unique water velocity, shear stress, and width-integrated shear stress (Section~\ref{sect:sub_mode}).
-This characteristic results in a perfect rank correlation between the variables (Figure~\ref{fig:Qw_vari}).
-
-Conversely, the history dependence on channel size in subglacial channels means different outputs occur across a range of discharges.
-For instance, in subglacial channels in \alpine{}, very high seasonal water velocity values and shear stresses can occur from a low water discharge  ($\sim\,4$\,\unit{m}$^3$\,\unit{s}$^{-1}$) to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ (Figure~\ref{fig:Qw_vari}\,a).
-In \icesheet{}, water velocities close to the seasonal mean value can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
-When considering the subglacial channel's evolving width, width-integrated shear stress  across the channel generally increases with water discharge, with improved rank correlation compared to water velocity or shear stress (\alpine{}, Figure~\ref{fig:Qw_vari} a--c).
-Yet even the width-integrated shear stress  can vary substantially, with the highest values occurring at water discharge values ranging from roughly $11$ \,\unit{m}\,\unit{s}$^{-1}$ to over $17$ \,\unit{m}\,\unit{s}$^{-1}$.
-The variability in width-integrated shear stress is less pronounced in the \icesheet{} scenario, where the hydrograph has less diurnal variability (Figure~\ref{fig:Qw_vari} c, f).
-
-The different response of subaerial and subglacial model outputs to water discharge impacts the timing of peaks in sediment transport capacity, and thus its  relationship with water discharge.
-In the subglacial channel, peaks in $w_g\tau_g$ generally occur when water discharge increases at the fastest rate, but before the maximum water discharge (Figure~\ref{fig:model_outs}\, d and h).
-By contrast, in the subaerial channel,   $w_f\tau_f$  peaks when water discharge is highest.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=0.9\linewidth]{Fig4.pdf}
-  \caption{Standard deviation of model outputs for different smoothing times with different subglacial and subaerial channel shapes and friction factors.
-    Shaded areas denote the range of standard deviations from the accepted $100$ parameter combinations (Section~\ref{sect:imp}).
-    Solid lines denote  the mean value of standard deviations.
-    Markers show smoothing periods ($15$\,\unit{min}, $1$\,\unit{hr}, $12$\,\unit{hr}, $1$\,\unit{d}, $5$\,\unit{d}, $10$\,\unit{d}, and $15$\,\unit{d}).
-  }
-  \label{fig:multi_run}
-\end{figure}
-
-The second numerical experiment aims to evaluate the response of variability in model outputs to a range of channel shapes, friction factors, and water discharge variability.
-Across the range of parameters  examined (Section\,\ref{sect:imp}), variability in water velocity and shear stress remains higher in the subglacial system compared to the subaerial one in both \alpine{} and \icesheet{}.
-This suggests that there is greater intermittency and variability in sediment transport in subglacial systems compared to subaerial ones, especially in supply-limited systems where sediment could be exhausted or mobilized over diurnal discharge cycles.
-Considering the width integrated shear stress, in \alpine{} the variability is comparable between the subaerial and subglacial systems (shown by the overlap of plots), due to the highly variable water discharge, largely due to diurnal variations (Figure~\ref{fig:multi_run}).
-In contrast, in \icesheet{}, with lower water discharge fluctuations, the variability in width-integrated shear stress remains consistently higher in the subglacial system across the time smoothing periods.
-
-Results from  both \alpine{} and \icesheet{} scenarios show that the variability in subglacial outputs decreases with smoothing time longer than approximately 1--5 days (Figure~\ref{fig:multi_run}).
-Yet, substantial differences in the variability between the model outputs for subglacial and subaerial cases persist in periods of up to $15$ days, longer than which seasonal variations might be conflated.
-Note that across longer smoothing times, the timing of peak sediment transport conditions still occurs in the natural environment.
-Yet they are not captured in model outputs(Figure~\ref{fig:model_outs}), despite their potential geomorphic importance.
-
-\section{Discussion}
-
-\subsection{Hydrology and evolving channel shape in driving sediment transport variability}
-\label{sect:dis_qsc}
-
-Increased variability in subglacial sediment discharge capacity makes it more difficult to differentiate the role of long-term climatic controls on sediment transport with shorter-term hydrological variations.
-The different response to water discharge variations in the model outputs subglacial channels compared to subaerial ones (Figure~\ref{fig:Qw_vari}) may actually cause even greater variations in sediment transport capacity in subglacial channels than presented above.
-In sediment transport relationships that evaluate sediment discharge capacity, such as in \citeA{meyer1948} or \citeA{engelund1967}, shear stress is scaled to the power of $\frac{3}{2}$ or $\frac{5}{2}$, respectively.
-The exponent greater than $1$ magnifies sediment discharge variability beyond the variable sediment transport parameters described above (Figure~\ref{fig:multi_run}; Table~\ref{tab:Qs}).
-
-For sufficiently short periods, channel width and wetted area are fixed in the subglacial system (akin to pipe flow) so that the response of width-integrated shear stress is $w_g\, \tau_{g}\, \propto\,  Q^{2}$ \cite<Table\,\ref{tab:eqs1}; >{alley1997}.
-\ref{sect:scaling} provides a complete description of sediment transport capacity in different drainage conditions.
-Note that in pipe flow conditions, maximum sediment transport would coincide with peak water discharge, unlike here where evolving channel size here alters the timing (Figure \ref{fig:model_outs} and \ref{fig:Qw_vari}).
-R\"othlisberger channels in steady state with water discharge, the width-integrated shear stress response to water discharge is $ w_g\,\tau_{g}\, \propto\,  Q^{\sim\frac{4}{5}}$ (Table\,\ref{tab:eqs1}).
-This relationship suggests that when discharge varies more slowly than channel shape subglacial channels have less variable sediment transport capacity than subaerial ones.
-Yet, subglacial channels rarely operate in a steady state \cite{gimbert2016}.
-
-The channel width and size variations in subglacial channels presented here also occur in subaerial channels in response to water discharge \cite{phillips2016}, likely over decadal timescales or in response to individual extreme events \cite<>{slater2013,dean2013}.
-These timescales are considerably longer than the continuous changes to the subglacial channel's response time of days.
-Furthermore, some observations from subaerial systems show a co-varying relationship between water discharge and sediment transport in time \cite<e.g. >{schmidt2006,pitlick2021}.
-Even so, if channel width changes occur in subaerial channels, then the values of $\alpha$ in Equation~\ref{eq:wcf} could change in time.
-$\alpha$ must be less than $1$, and even this would cause the exponent on $Q$ in the  $w_g\,\tau_{g}$ relationships to remain substantially below that of pipeflow (\ref{sect:scaling}, Table\,\ref{tab:Qs} and \ref{tab:Qs2})).
-
-Catchments with reduced water discharge variability may experience yet less variability in sediment transport capacity, as the two hydrological forcings from \alpine{} and \icesheet{} show different variations in  sediment transport capacity.
-Indeed, the more decoupled and sporadic relationship between model outputs and water discharge in \alpine{} results from the relatively larger discharge variations on daily to weekly timescales compared to \icesheet{} (Figure~\ref{fig:Qw_vari}).
-This occurs as the evolution of the R-channel's size is slow compared to the variations in water discharge.
-Such water discharge variability in alpine catchments may cause shear stress to approach the water discharge squared in assuming pipe-flow conditions \cite< \ref{sect:scaling}; c.f.>{alley1997}.
-Conversely, the reduced variability in water discharge in \icesheet{} may result in the R-channel's size being closer to equilibrium from the smaller variations in water discharge.
-Due to the smaller water discharge variability, pipe flow with fixed channel size could more closely approximate hydraulics here.
-In this case, the exponent on water discharge for shear stress is likely less than $2$, but greater than $\frac{4}{5}$ occurring when the R-channel's size evolves in equilibrium with water discharge (see \ref{sect:scaling}).
-As a result, there is a stronger between subglacial model outputs and water discharge in \icesheet{} compared to \alpine{} (Figure~\ref{fig:Qw_vari}) due to the channel size being in closer equilibrium with water discharge in the \icesheet{} case.
-
-Water discharge measurements at the hourly timescale remain unavailable in many catchments, further limiting subglacial  sediment transport models in capturing specific events.
-The model output variability in subglacial channels approaches that of subaerial ones over $1-5$ days of smoothing (Figure~\ref{fig:multi_run}), possibly resulting in parity in sediment transport capacity variations between subaerial and subglacial channels.
-However, this smoothing omits key sediment discharge characteristics, such as the timing of peaks in sediment transport capacity from glaciers (Figure~\ref{fig:model_outs}).
-
-\subsection{Effect sediment transport and landscape evolution}
-
-Subglacial channels exhibit a far weaker relationship between sediment transport capacity and water discharge than in subaerial channels (Figure~\ref{fig:Qw_vari}).
-In subaerial channels, sediment transport variations in subaerial channels can be attributed to hydro-climatic conditions \cite{wolman1960}
-A glacier's sediment transport capacity is impacted by the ice thickness controlling the channel closure rate, and the glacier's surface slope, in addition to water discharge and sediment size \cite<Figure~\ref{fig:multi_run}, Section~\ref{sect:sub_mode}; >{rothlisberger1972,gimbert2016,stevens2022,walder1994}.
-In a supply-limited state, a glacier's sediment discharge record also responds to  sediment availability and bedrock erosion from  water pressure variations and sliding  \cite<e.g.>{iverson2012,herman2015,delaney2023}.
-This multitude of processes lies in contrast to many subaerial channels, where transport capacity typically responds to water discharge in addition to sediment size, channel shape,  and hydraulic gradient, that can remain stable over years to millennia \cite<Section~\ref{sect:fluv}; e.g.>{tucker1997}.
-As a result, characteristics such as bankfull or effective water discharge could have limited meaning in evaluating subglacial sediment transport \cite{wolman1960,lenzi2006}.
-
-Previous work suggests that this co-varying relationship between sediment discharge capacity and water discharge in  subaerial conditions could be noticeable $\sim20$\,\unit{km} downstream of the Leverett site, where a strong correlation persists between sediment plume size and the river's water discharge into the Kangerlussuaq fjord \cite{mcgrath2010}.
-In contrast, in marine-terminating glacier catchments, a less consistent relationship may occur between water discharge or melt extent and sediment plume size \cite{tedstone2012}.
-Here, ocean water pressurizes subglacial channels at the ice front \cite<e.g.>{how2017}, so the reduced correlation could result from the inconsistent relationship between subglacial sediment transport capacity and water discharge presented here.
-
-Greater variations in subglacial sediment transport capacity compared to subaerial ones could cause a supply-limited regime at many glaciers.
-In subglacial systems, sediment's critical shear stress, or threshold at which sediment mobilization occurs, can be reached more frequently and across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
-This result suggests sediment export here is especially sensitive to water discharge variability, in addition to quantity \cite{lane2019b}.
-Sediment exhaustion through repeated crossing of the mobilization threshold may explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \cite{overeem2017}.
-The repeated crossing of the mobilization threshold across a range of discharges could also result in sediment discharge's strong dependence on sediment production from sliding in many cases \cite{herman2015,koppes2015}.
-
-\section{Conclusions}
-
-The sediment transport capacity of a channel is driven by its width and the shear stress exerted by the flowing water, which is proportional to  flow velocity squared.
-Subaerial channels alter both their channel width and water velocity immediately in response to changing water discharge.
-In contrast, pressurized subglacial channels largely accommodate rapidly changing water discharge by altering water velocity.
-Their size responds over several days to changes in discharge.
-Thus the response of sediment transport capacity is much more sensitive to changes in water discharge in subglacial channels compared to subaerial ones.
-
-Implications of the presented findings include:
-\begin{enumerate}
-\item  evolving subglacial channel size causes  water discharge to have a changing impact on sediment transport capacity, complicating the relationship between sediment export and hydro-climatic conditions below glaciers;
-  \vspace{.1cm}
-
-\item the  variable relationship between water discharge and sediment transport capacity below glaciers creates larger variability in sediment transport capacity compared to subaerial systems;
-  \vspace{.1cm}
-
-\item the timing of peak sediment discharge capacity in subglacial channels occurs before peak  water discharge, whereas it is coincident in subaerial ones, limiting the meaning of certain water discharge characteristics in subglacial systems.
-  Representing subglacial sediment transport capacity with either open-channel flow or pipe flow neglects differences in timing.
-\end{enumerate}
-
-This study calls for careful analysis when examining the relationship between sediment transport capacity and hydro-climatic conditions from glacierized catchments.
-
-\section*{Author Contributions}
-
-I.D. designed the study, developed, and implemented the model and experiments, and wrote the manuscript.
-A.J.T. helped interpret data from Leverett Glacier and provided guidance on experiment design.
-M.A.W. provided data from Fieschergletscher and added the relationship between shear stress and water discharge (see \ref{sect:scaling}).
-D.F. provided data from Fieschergletscher.
-All authors provided key inputs in writing and editing the manuscript.
-
-\section*{Open Research Section}
-
-Code to make the figures, with links to the data, can be found at \url{https://bitbucket.org/IanDelaney/xsection/src/master/}.
-Code and data will be uploaded to a permanent repository, with FAIR principles, pending acceptance of the paper.
-
-\acknowledgments
-
-SNSF Project No. $\mathrm{PZ00P2}$\_$202024$ provided  funding for I. Delaney.
-A. Tedstone acknowledges funding from the European Research Council, award $818994$ -- CASSANDRA.
-G. King and M. J. Gevers provided useful comments on a previous version of this manuscript.
-
-\bibliography{PaperLib}
-
-\newpage
-
-\section{Supplement}
-\subsection{Table of variables, parameters and constants}
-\begin{table}[h]
+\begin{table}[hbt!]
   \centering
@@
 
-         &&&\\
-
     \textbf{Parameters and Constants}  & & &\\
@@
     Effective glacier length &$l$& $7,000$ or $26,000$&\unit{m}\\
-    Constant $1$ in Equation~\ref{eq:dS_dt} &$C_1$&$2.2\times10^{-5}$&\unit{m}$^{-1}$\\
-    Constant $2$ in Equation~\ref{eq:dS_dt} &$C_2$&$3.7\times10^{-13}$&\unit{m}$^{-n}\,s^{-1}$\\
+    Constant $1$ in Equation~3 &$C_1$&$2.2\times10^{-5}$&\unit{m}$^{-1}$\\
+    Constant $2$ in Equation~3 &$C_2$&$3.7\times10^{-13}$&\unit{m}$^{-n}\,s^{-1}$\\
     Latent heat of fusion &$L$&$333.5 $&\unit{kJ\,kg}$^{-1}$\\
@@
 
-    Gradient of channel bed (subaerial) &$\nabla z_c$ &$0.05$& $\mathrm{(-)}$\\
-    Subaerial channel factor & $k$ &$3$ or $6.5$ & $\mathrm{s\,m^{-2}}$\\
-         && ($5$, $20$) & \\
-    Channel geometry exponent &$\alpha$& $\frac{1}{2}$&$\mathrm{(-)}$ \\
+    Gradient of channel bed (subaerial) &$\nabla z_c$ &$0.02$& $\mathrm{(-)}$\\
+    &&($.01$, $0.05$) & \\
+    Subaerial channel factor & $k$ &$8$ & $\mathrm{s\,m^{-2}}$\\
+    Channel geometry exponent &$\alpha$&$\frac{1}{3}$ &$\mathrm{(-)}$ \\
+         &&($\frac{1}{3}$, $\frac{1}{2}$) & \\
     \hline
@@
 \end{table}
+
 \FloatBarrier
+\subsection{Implementation}
+\label{sect:imp}
 
-\subsection{Figure 3 over various smoothing times}
+The models above are applied to proglacial discharge records from the Fieschergletscher (scenario \alpine{}) and the Leverett Glacier (scenario \icesheet{}).
+Outputs of the models represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
+To generalize these scenarios, \alpine{}  is exemplified by Fieschergletscher's relatively thin ice thickness \citep[$h_{ice}$= $225$\,\unit{m}][]{grab2021}, low water discharge ($\sim\,10$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:Qw}\, b).
+\icesheet{}  is exemplified by Leverett glacier's  thick ice  \citep[$h_{ice}$= $700$\,\unit{m}; ][]{morlighem2017}, high water discharge ($\sim\,300$\,\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:Qw}\, b).
 
-\begin{center}
-  \begin{figure}[!h]
-    \includegraphics[width=0.7\linewidth]{Fig3_hr.png}
-    \caption{As Figure \ref{fig:model_outs}, with $1$ \,\unit{hr} smoothing.}
-    \label{fig:model_outs_1hr}
-  \end{figure}
-\end{center}
+The first experiment aims to characterize the variability in sediment discharge capacity in subglacial compared to subaerial channels across a range of channels slopes and  shapes, along with friction factors.
+Additionally, we examine the effects of different water discharge variability by smoothing the water discharge over different periods.
+The results below present water velocity ($v_g$, $v_f$), shear stress ($\tau_g$, $\tau_f$), and width-integrated shear stress ($w_g\tau_g$, $w_f\tau_f$) from the subglacial and subaerial models.
+For brevity, these values together are referred to as ``model outputs'' in the text.
+The subglacial model outputs are generally larger than their subaerial equivalents due to their large hydraulic gradients, thus greater sediment transport capacity would be expected in subglacial channels compared to subaerial ones \citep{alley1997}.
+To evaluate variability, we subtract the model outputs from their values smoothed using daily averages,  detrending the seasonal signal (Figure~\ref{fig:Qw}).
+We then calculate the variability in the model output by taking the standard deviation of this detrended time series.
 
-\begin{center}
-  \begin{figure}[!h]
-    \includegraphics[width=0.7\linewidth]{Fig3_12hr.png}
-    \caption{As Figure \ref{fig:model_outs}, with $12$ \,\unit{hr} smoothing.}
-    \label{fig:model_outs_12hr}
-  \end{figure}
-\end{center}
+The effect of channel shape, gradient and roughness is established by running the model with random parameter values of channel slope and geometry factors ($\beta$,$\nabla z_c$ and $\alpha$) and friction factors ($f_r$ and $f_f$; see Table~\ref{table:vpm} for value range).
+Runs with parameter combinations are accepted if their mean subglacial water velocity over the season lies between $0.5$\,\unit{m}\,\unit{s}$^{-1}$ and $2$\,\unit{m}\,\unit{s}$^{-1}$ or if subaerial water velocity lies between $0.3$\,\unit{m}\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}\,\unit{s}$^{-1}$, following measurements \citep[e.g.][]{werder2010b,magnusson2012,chandler2013}.
+Additionally, to be accepted, subglacial model runs can only experience a flotation fraction ($\frac{\Delta h}{2\,h_o}$) of above $1.2$ for less than $2.5$\,\% of the run.
+The initial condition of cross-sectional area $S_g$ is evaluated by applying a model spinup with repeatedly applying the maximum observed water discharge of the first $4$ days of the study until there is no change in the channel area.
 
-\begin{center}
-  \begin{figure}[!h]
-    \includegraphics[width=0.7\linewidth]{Fig3_1day.png}
-    \caption{As Figure \ref{fig:model_outs}, with $1$ \,\unit{day} smoothing.}
-    \label{fig:model_outs_1day}
-  \end{figure}
-\end{center}
+The routine runs until $100$ different parameter combinations for each water discharge smoothing period are accepted with the conditions described above.
+Here, the water discharge input at a given time is the average over the smoothing period before that time, ranging from $15$\,\unit{min} up to $15$\,\unit{days}.
 
-\begin{center}
-  \begin{figure}[!h]
-    \includegraphics[width=0.7\linewidth]{Fig3_5day.png}
-    \caption{As Figure \ref{fig:model_outs}, with $5$ \,\unit{day} smoothing.}
-    \label{fig:model_outs_5day}
-  \end{figure}
-\end{center}
+In the second experiment, the models are  applied to a reference test case for each glacier.
+These experiments assumes  a subglacial channel with $\beta=\frac{\pi}{6}$ and a subaerial channel with $\alpha = \frac{1}{3}$ and slope of $0.02$ (Table~\ref{table:vpm}).
+Both friction factors $f_r$ and $f_f$ are tuned so that reasonable water velocities \citep[$\sim\,1.6\,$\unit{m}\,\unit{s}$^{-1}$][]{werder2010b,chandler2013} occur for both \alpine{} and \icesheet{}.
+%
+To quantify the link between sediment transport capacity and water discharge, model outputs are compared to water discharge from the glaciers using Spearman rank correlation, which accounts for the ordering of values, but not their magnitude.
+Rank correlation reduces the impact of the non-linear response of sediment transport capacity to the hydrological forcing.
 
-\begin{center}
-  \begin{figure}[!h]
-    \includegraphics[width=0.7\linewidth]{Fig3_10day.png}
-    \caption{As Figure \ref{fig:model_outs}, with $10$ \,\unit{day} smoothing.}
-    \label{fig:model_outs_10day}
-  \end{figure}
-\end{center}
+\section{Results}
+We first present formulations of sediment transport capacity in several channel types to build upon the analysis in \citet{alley1997}.
+We then apply the models above (Section~\ref{sect:meth}) in two experiments by examining three model outputs.
+Water velocity and shear stress represent the potential for sediment mobilization, and width-integrated shear stress represents a proxy for the total sediment transport capacity across the channel bed (Figure~\ref{fig:model_outs}).
 
-\begin{center}
-  \begin{figure}[!h]
-    \includegraphics[width=0.7\linewidth]{Fig3_15day.png}
-    \caption{As Figure \ref{fig:model_outs}, with $15$ \,\unit{day} smoothing.}
-    \label{fig:model_outs_15day}
-  \end{figure}
-\end{center}
-\FloatBarrier
-
-\newpage
-\subsection{Formulas for sediment transport in different drainage regimes}
+\subsection{Sensitivity of sediment transport in different drainage regimes}
 \label{sect:scaling}
 
-% \mauro{Notation: either use $h$ and $w$ for the river, or otherwise, somehow free $h$. Below I use $H$ and $W$ for now.}
+Here three kinds of drainage regimes are examined, namely, subaerial channels, steady-state R-channels \citep{rothlisberger1972}, and pipe-flow (i.e.
+R-channels which do not have time to adjust their size to discharge conditions).
+The evolving subglacial channel geometry presented in Section \ref{sect:sub_mode} is not considered.
+Here we show the sediment transport capacity scaling with  a given water discharge and hydraulic gradient for three different transport formulas: Meyer-Peter M\"uller  \citep[MPM; ][]{meyer1948}, Engelund and Hansen \citep[EH; ][]{engelund1967}, and Bagnold \citep{bagnold1980}; additionally width-integrated shear stress is assessed, as used as proxy for sediment transport throughout the manuscript.
 
-Here three kinds of drainage regimes are examined, namely, subaerial channels, steady-state R-channels \cite{rothlisberger1972}, and pipe-flow (i.e.
-R-channels which do not have time to adjust their size to discharge conditions).
-Here we show how sediment transport capacity scales with respect to a given water discharge and hydraulic gradient for three different transport formulas: Meyer-Peter M\"uller  \cite<MPM;>{meyer1948}, Engelund and Hansen \cite<EH;>{engelund1967}, and Bagnold \cite{bagnold1980}; additionally width-integrated shear stress is assessed, as used as proxy for sediment transport throughout the manuscript.
-
-A few basic hydraulic relations are used in the following.
+The following hydraulic relations are used.
 The Darcy-Weisbach equation can be stated as
@@
 
-The  subaerial channel is assumed to have a width $w$ much greater than its depth $H$, such that $D_h\approx 4h$, and to have a constant head gradient ($\Psi$) given by the topography.
+As in Section~\ref{sect:fluv}, the  subaerial channel is assumed to have a width $w$ much greater than its depth $H$, such that $D_h\approx 4h$, and to have a constant head gradient ($\Psi$) given by the topography.
 Further, it is  assumed that its width can be approximated by a relation $w \propto Q^\alpha$ (Equation~\ref{eq:wcf}) with $\alpha\in [0,1]$; typically $\alpha \approx 1/2$ (a so-called regime channel).
@@
 % 
-For a steady state R-channel, it is assumed that  $\Psi$ is constant (approximated by the gradient of the Shreve \citeyear{shreve1972} potential \footnote{Note that the R-channel model presented above (Eq.~\eqref{eq:dS_dt}--\eqref{eq:dh2wc}), calculates $\Psi$ from the time evolving $S_g$ (Eq.~\eqref{eq:dS_dt}) via the Darcy-Weisbach equation~\eqref{eq:dh} and thus no Shreve approximation is then needed.}) and that $S$ adjusts such that it is in steady state with $\Psi$ and $Q$.
+For a steady state R-channel, it is assumed that  $\Psi$ is constant (approximated by the gradient of the Shreve \citeyear{shreve1972} potential and that $S$ adjusts such that it is in steady state with $\Psi$ and $Q$.
+Note that the R-channel model presented above in Section~\ref{sect:sub_mode} Equation~\ref{eq:dS_dt} calculates $\Psi$ from the time evolving $S_g$ via the Darcy-Weisbach Equation~\ref{eq:dh} and thus no Shreve approximation is then needed.
 % Conversely, pipe flow has $S$ fixed and $\Psi$ adjusts to satisfy the given $Q$.
@@
 Now width integrated shear stress and sediment transport can be calculated for all regimes and for all transport formulas.
-The results of the 12 combinations are presented in Table~\ref{tab:Qs} and also in Table~\ref{tab:Qs2} where the fractions in the exponents are approximately given by decimal numbers for ease of comparison.
+The results of the 12 combinations are presented in Table~\ref{tab:Qs} and also in Table~\ref{tab:Qs} where the fractions in the exponents are approximately given by decimal numbers for ease of comparison.
 
-Table~\ref{tab:Qs2} shows how the proxy $w\tau$ as well as $Q_s$ of the MPM, EH and Bagnold sediment transport formulas scale with respect to $Q$, $\Psi$ or $S$ for the three different regimes and for different $\alpha$.
-Remarkably, the Bagnold formula has a negative exponent for $f$ in all but the pipe-flow regime, which seems rather unexpected.
-The total transport formula EH gives a slightly stronger dependence on all variables as should be expected due to the larger exponent on $\tau$ of $\frac{5}{2}$ versus  $\frac{3}{2}$  for the others.
-Albeit, the sediment transport response in pipe-flow for the Bagnold case is close to EH.
-Conversely, the sediment transport proxies $w_g\,\tau_g$ \& $w_f\,\tau_f$ used within this publication scale only as $Q^2$ for pipe-flow, whereas the transport scales at least with $Q^3$.
-The exponent for the width scaling $\alpha$ only impacts relationship between sediment transport and water discharge in the EH relation in any meaningful way.
-However, for the value of $\alpha$ around $\frac{1}{2}$ (a value appropriate for most streams), the exponent on $Q$ for EH is only slightly above what the other transport relations give.
-
-The sediment discharge in the steady state R-channel scales very similarly to the  subaerial channel case for all relations and virtually identically for the $\alpha=0.5$ case.
-However, note that the head gradients $\Psi$ are likely higher for comparable $Q$ in an ice-sheet marginal or alpine glacier setting than in a  subaerial channel, as is well described in \citeA{alley1997}, and thus transport rates may still be much higher in a steady state R-channel.
-
-Furthermore, R-channels will rarely operate in steady state as variations in discharge, in particular on the diurnal timescale or the timescale of severe rain or melting events, are too fast for such a channel to reach a steady state.
-In such cases they operate more like a pipe of fixed cross-section \cite<e.g.>{gimbert2016}.
-Table~\ref{tab:Qs2} shows that in such a situation, the sediment transport scales much more severely with discharge, with the exponent on $Q$ being between $3$ and $5$, compared to the other two regimes when that exponent is at most $1.7$ (Table~\ref{tab:Qs2}).
-Thus fluctuations of discharge on short timescales (on the order of a day) have the potential to cause conditions with very high sediment transport capacities.
-
-\begin{table}
+\begin{table}[hbt!]
   \caption{Relations for hydraulic variables for the three drainage regimes:  subaerial channels, R-channel and pipe-flow.
@@
     R-channel & \(\Psi\) & \(S\, \propto f^{2/5}\, Q^{4/5} \, \Psi^{-2/5}\) & \(D_h\propto w \propto H \propto S^{1/2}\) & \(f^{1/5} Q^{2/5} \, \Psi^{4/5}\) & \(Q\, \Psi\)\\
-    Pipe & \(S\) & \(\Psi \propto f \, Q^2\, S^{-5/2}\) & \(D_h\propto w \propto H \propto S^{1/2}\) & \(f Q^2 S^{-2}\) & \(f\, Q^3 S^{-5/2}\)\\
+    Pipe & \(S\) & \(\Psi \propto f \, Q^2\, S^{-5/2}\) & \(D_h\propto w \propto H \propto S^{1/2}\) & \(fx Q^2 S^{-2}\) & \(f\, Q^3 S^{-5/2}\)\\
   \end{tabular}
@@
 
-\begin{table}
-  \caption{Sediment transport proxy ($w\tau$) and rates for the three considered different transport formulas: MPM \cite{meyer1948}, EH \cite{engelund1967}, and Bagnold \cite{bagnold1980}.
+Table~\ref{tab:Qs} shows how the proxy $w\tau$ as well as $Q_s$ of the MPM, EH and Bagnold sediment transport formulas scale with respect to $Q$, $\Psi$ or $S$ for the three different regimes and for different $\alpha$.
+Remarkably, the Bagnold formula has a negative exponent for $f$ in all but the pipe-flow regime.
+The total transport formula EH gives a slightly stronger dependence on all variables as should be expected due to the larger exponent on $\tau$ of $\frac{5}{2}$ versus  $\frac{3}{2}$  for the others.
+Albeit, the sediment transport response in pipe-flow for the Bagnold case is close to EH.
+Conversely, the sediment transport proxies $w_g\,\tau_g$ \& $w_f\,\tau_f$ used below scale only as $Q^2$ for pipe-flow, whereas the transport scales at least with $Q^3$.
+The exponent for the width scaling $\alpha$ only impacts relationship between sediment transport and water discharge in the EH relation in any meaningful way.
+However, for the value of $\alpha$ around $\frac{1}{2}$ (a value appropriate for most streams), the exponent on $Q$ for EH is only slightly above what the other transport relations give.
+
+The sediment discharge in the steady state R-channel scales very similarly to the  subaerial channel case for all relations and virtually identically for the $\alpha=0.5$ case.
+However, note that the head gradients $\Psi$ are likely higher for comparable $Q$ in an ice-sheet marginal or alpine glacier setting than in a  subaerial channel, as is well described in \citet{alley1997}, and thus transport rates may still be much higher in a steady state R-channel.
+
+Furthermore, R-channels will rarely operate in steady state as variations in discharge, in particular on the diurnal timescale or the timescale of severe rain or melting events, are too fast for such a channel to reach a steady state.
+In such cases they operate more like a pipe of fixed cross-section \citep[e.g.][]{gimbert2016}.
+Table~\ref{tab:Qs} shows that in such a situation, the sediment transport scales much more severely with discharge, with the exponent on $Q$ being between $3$ and $5$, compared to the other two regimes when that exponent is at most $1.7$ (Table~\ref{tab:Qs}).
+Thus fluctuations of discharge on short timescales (on the order of a day) have the potential to cause conditions with very high sediment transport capacities, that are of far high magnitude than the subaerial channels.
+
+\begin{table}[hbt!]
+  \caption{Sediment transport proxy ($w\tau$) and rates for the three considered different transport formulas: MPM \citep{meyer1948}, EH \citep{engelund1967}, and Bagnold \citep{bagnold1980}.
   }
@@
 
+\FloatBarrier
+\subsection{Increased variability in evolving subglacial channels occurs across a range of channel shapes, slopes  and friction values}
+\label{sect:ensemble}
+
+The first numerical experiment aims to compare the variability between the subglacial and subaerial  model outputs to a range of channel shapes, friction factors, and water discharge variability, as the subglacial channel's cross-sectional area is allowed to evolve.
+Following the findings in Section~\ref{sect:scaling}, variability in subglacial sediment discharge capacity could approach that of subaerial systems, should water discharge variations be sufficienlty small or subglacial channels respond in size quickly.
+For simplicity, we only examine $w\tau$, not the sediment transport formulas mentioned above.
+%We have chosen to exclude the pipe flow case (Section~\ref{sect:scaling}) from analysis as it depends greatly on the choice of the subglacial cross sectional area $S_g$.
+
+Across the range of parameters  examined (Section\,\ref{sect:imp}), variability in all model outputs remains higher in the subglacial system compared to the subaerial one for both \alpine{} and \icesheet{} (Figure~\ref{fig:multi_run}).
+In some cases, the variability in subglacial model outputs is a  magnitude larger than their subaerial counterparts.
+Results from  both \alpine{} and \icesheet{} scenarios show that variability in both subglacial and subaerial outputs decreases with smoothing time longer than approximately 1--5 days (Figure~\ref{fig:multi_run}), when the diurnal variations in water discharge are removed from smoothing.
+In the \icesheet case, the smaller variability in model outputs compared to \alpine may in fact mean that the R-channel is in closer equilibrium between channel size and water discharge (Section~\ref{sect:scaling}).
+In this case, water discharge variations would create smaller disparities in sediment transport capacity than even in subaerial cases as the subglacial channel size evolves to accommodate them (Table~\ref{tab:Qs}).
+
+\begin{figure}[hbt!]
+  \centering
+  \includegraphics[width=0.9\linewidth]{multi_run_tmp.pdf}
+  \caption{Standard deviation of detrended model outputs for different smoothing times with different subglacial and subaerial channel shapes and friction factors.
+    Shaded areas denote the range of standard deviations from the accepted $100$ parameter combinations (Section~\ref{sect:imp}).
+    Solid lines denote  the mean value of standard deviations.
+    Markers show smoothing periods ($15$\,\unit{min}, $1$\,\unit{hr}, $12$\,\unit{hr}, $1$\,\unit{d}, $5$\,\unit{d}, $10$\,\unit{d}, and $15$\,\unit{d}).
+  }
+  \label{fig:multi_run}
+\end{figure}
+
+Across the range of parameters examined in both \alpine and \icesheet, variability in a subaerial model output did not exceed the corresponding subglacial one (Figures~\ref{fig:multi_run} and \ref{fig:params}).
+Analysis of parameter values shows that in both cases, smaller values of $\alpha$, or closer to a slot-canyon, result in slightly greater variability in velocity and shear stress (Figure~\ref{fig:params}).
+Yet, larger values of $\alpha$ can result in greater variability in width-integrated shear stress, as greater channel width changes increase the width-integrated shear stress, but not the shear stress.
+Additionally, steeper subaearial channel slopes result in greater variability in both shear stress and width-integrated shear stress, but only approach the variability of the subglacial channel in the \alpine case (Figure~\ref{fig:params}).
+%At steeper slopes, the alluvial  processes represented in the model (Section ~\ref{sect:fluv}) probably do not apply, as the.
+We note that these parameters span a commonly accepted range of values, therefore do not anticipate a scenario where variability in the subaerial system would exceed the subglacial system with this hydrology.
+
+Greater values of the subglacial and subaerial friction factor $f_i$ and $f_p$ result in greater variability in the  shear stress and width-integrate shear stress both cases (Figure~\ref{fig:params}), as expected given Equations~\ref{eq:tau_g} and \ref{eq:vf}.
+However, more velocity variability is present in low values of $f_i$ and $f_p$. 
+Low values of $f_i$ result in lower growth rates in subglacial channels (Equations~\ref{eq:dS_dt} and \ref{eq:dh}).
+As a result, water velocity, as opposed to channel growth, accommodates increases in water discharge, and velocity variability increases.
+Note as well that smaller values of channel factor $\beta$, creating low and broad channels, result in more variability in width-integrated shear stress.
+Here, the channel width can grow more quickly in response to water discharge increases as compared to a semi-circular channel with $\beta = \pi$ (Equation~ \ref{eq:dh}).
+
+\begin{figure}[hbt!]
+  \centering
+  \includegraphics[width=0.9\linewidth]{Parameter.pdf}
+  \caption{Parameter values compared to variability in model outputs.
+    Plots on the right (orange, light blue) correspond to the \alpine{} case. Plots on the left (brown, dark blue) correspond to the \icesheet case.
+    Outputs are shown using $15$ minute smoothing time. 
+  }
+  \label{fig:params}
+\end{figure}
+
+\FloatBarrier
+\subsection{Changing subglacial channel size drives different timing and variability in sediment transport capacity}
+
+In the second numerical experiment defined in Section~\ref{sect:imp}, we aim to quantify the sources of increased variability in the subglacial model outputs as they exhibit different seasonal evolutions and peaks.
+Examining the relationship between water discharge and the model outputs, different relationships emerge for the subaerial and subglacial cases due to the hysteresis in channel size in the subglacial model (Equation~\ref{eq:dS_dt}).
+Because there is no history dependence in  subaerial channels, each water discharge value in the subaerial channel results in a unique water velocity, shear stress, and width-integrated shear stress (Section~\ref{sect:sub_mode}).
+This characteristic results in a perfect rank correlation between the variables (Figure~\ref{fig:Qw_vari}).
+\begin{figure}[hbt!]
+  \centering
+  \includegraphics[width=0.9\linewidth]{Fig2.png}
+  \caption{Model outputs from simulations using the hydrographs in panels a and e for the scenarios \alpine{} (a-d) and \icesheet{} (e-h).
+    Black (gray) lines in a and e represent the subglacial channel size (water discharge).
+    Blue (orange) lines represent outputs from the subglacial (subaerial) channel.
+    Data are shown at $15$\unit{min} intervals.
+    Arrows show examples where variables peak in the subglacial channel before the subaerial one.
+    Insets in f-h show the  peak melt event denoted by the shaded area in panels e--h, with an arbitrary y-axis.
+  }
+  \label{fig:model_outs}
+\end{figure}
+
+Conversely, the history dependence on channel size in subglacial channels means that different sediment transport characteristics, such as velocity, occur across a large range of water discharges.
+For instance, in subglacial channels in \alpine{},  high water velocity values and shear stresses can occur from a low water discharge  ($\sim\,4$\,\unit{m}$^3$\,\unit{s}$^{-1}$) to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ (Figure~\ref{fig:Qw_vari}\,a).
+In \icesheet{}, water velocities close to the seasonal mean value can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
+When considering the subglacial channel's evolving width, width-integrated shear stress across the channel generally increases with water discharge, with improved rank correlation compared to water velocity or shear stress (\alpine{}, Figure~\ref{fig:Qw_vari} a--c).
+Yet even the width-integrated shear stress  can vary substantially, with the highest values occurring at water discharge values ranging from roughly $11$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ to over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
+The variability in width-integrated shear stress is less pronounced in the \icesheet{} scenario, where the hydrograph has less diurnal variability (Figure~\ref{fig:Qw_vari} c, f).
+
+\begin{figure}[hbt!]
+  \centering
+  \includegraphics[width=0.9\linewidth]{Fig3.png}
+  \caption{
+    Relationship between water discharge and normalized velocity, shear stress, and width-integrated shear stress for \alpine{} (a-c) and \icesheet{} (d-f).
+    Variables on the x-axis have been normalized to mean values.
+    $R_g$ ($R_f$) shows the Spearman rank correlations for the subglacial (subaerial) outputs.
+    Plots are shown with discharge and outputs at $15$\unit{min} intervals.
+  }
+  \label{fig:Qw_vari}
+\end{figure}
+
+The different response of subaerial and subglacial model outputs to water discharge impacts the timing of peaks in sediment transport capacity, and thus its  relationship with water discharge.
+In the subglacial channel, peaks in $w_g\tau_g$ generally occur when water discharge increases at the fastest rate, but before the maximum water discharge (Figure~\ref{fig:model_outs}\, d and h).
+By contrast, in the subaerial channel,   $w_f\tau_f$  peaks when water discharge is highest.
+
+\FloatBarrier
+\section{Discussion}
+
+\subsection{Experiment limitations}
+
+Several aspects of the models and  experiment design limit the comparison of the model outputs to field data and make comparison of the two systems difficult.
+The lumped nature of the models mean that they operate independently of the upstream drainage network, and sediment access is not considered here.
+In reality, processes such as sediment access are highly important in controlling sediment export in glacierized catchments \citep[e.g.][]{herman2015,vergara2022}.
+Furthermore, under different glacier configurations and subglacial conditions, new channels can emerge quickly and water can be distributed to flow through several adjacent conduits impacting the sediment transport in each \citep[e.g.][]{werder2013,hewitt2019,delaney2023}.
+As a result, these models isolate the relationship between water discharge and sediment transport capacity in subglacial and subaerial systems, making comparison more direct, but neglecting more complex, yet important spatially distributed processes.
+
+Comparison of model outputs between the subaerial and subglacial case is difficult, given that while the water velocities are comparable in both channel types, the shear stresses and width-integrated shear stresses are much smaller in subaerial channels (Figure~\ref{fig:model_outs}).
+This occurs due to the different gradients in the two systems, with the subglacial one being much steeper \citep{alley1997}.
+We note, however, that in the ensemble runs in Section~\ref{sect:ensemble}, the range of parameters tested cover the range of viable shear stresses in both subglacial and subaerial channels.
+
+Width-integrated shear stress is examined in the two numerical experiments as it does not have a dependence on grain-size, unlike sediment transport relationships, such as \citet[][Section \ref{sect:scaling}]{meyer1948}.
+This makes comparison between the two systems simpler. 
+Preferential transport of smaller sediment clasts and the input of upstream sediment impacts sediment transport capacity as grain-size evolves \citep[e.g.][]{gomez1983}.
+These processes are only beginning to be evaluated below glaciers \citep{aitken2024}, but can be important in subaerial systems.
+Subglacial sorting processes could be an additional source of variability in sediment export from glacier with respect to water discharge, especially with respect to bedload.
+
+\subsection{Increased variability in  sediment transport capacity in subglacial systems}
+\label{sect:dis_qsc}
+
+Increased variability in subglacial sediment discharge capacity could make it more difficult to differentiate the role of long-term climatic controls on sediment transport with shorter-term hydrological variations.
+In sediment transport relationships that evaluate sediment discharge capacity, such as in \citet{meyer1948} or \citet{engelund1967}, shear stress is scaled to the power of $\frac{3}{2}$ or $\frac{5}{2}$, respectively (Section~\ref{sect:scaling}).
+The exponent greater than $1$ magnifies sediment discharge variability beyond the variable sediment transport parameters described above (Figure~\ref{fig:multi_run}; Table~\ref{tab:Qs}).
+As a result, the different response to water discharge variations in  shear stress in subglacial channels compared to subaerial ones (Figure~\ref{fig:Qw_vari}) may actually cause even greater variations in sediment transport capacity in subglacial channels than presented above.
+
+Greater variations in subglacial sediment transport capacity compared to subaerial ones could cause a supply-limited regime at many glaciers \citep{alley1997}.
+In subglacial systems, sediment's critical shear stress, or threshold at which sediment mobilization occurs, can be reached more frequently and across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
+This result suggests sediment export here is especially sensitive to observed changes in water discharge variability, in addition to quantity \citep{lane2019b}.
+Sediment exhaustion through repeated crossing of the mobilization threshold may explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \citep{overeem2017}.
+The repeated crossing of the mobilization threshold across a range of discharges could also result in sediment discharge's strong dependence on sediment production from sliding in many cases \citep{herman2015,koppes2015}.
+
+While transport-limited states below glaciers are likely rare \citep[e.g.][]{alley1997}, they can be possible \citep[e.g.][]{swift2005} and abundant sediment can persist below glaciers \citep[e.g.][]{walter2014,stevens2022,delaney2022}.
+At these glaciers, the great variability in subglacial sediment transport capacity may make it difficult to link sediment discharge to hydrology, especially when peak events occur \citep{cowan1988,delaney2018}.
+As a result, it may prove challenging to establish the impact on sediment transport capacity  of individual extreme events, such as extreme precipitation or melt \citep{lu2022}, given the potential for different channel sizes with the same hydrology forcing.
+However, in the case of bedload, the signal of this variable subglacial sediment transport relationship may be destroyed only a short distance downstream as subaerial hydraulics controls these sediment transport \citep{mancini2023}.
+
+Catchments with reduced water discharge variability may experience yet less variability in sediment transport capacity, as the two hydrological forcings from \alpine{} and \icesheet{} show different variations in model outputs.
+Indeed, the more decoupled and sporadic relationship between model outputs and water discharge in \alpine{} results from the relatively larger discharge variations on daily to weekly timescales (Figure~\ref{fig:Qw_vari}).
+This occurs as the evolution of the R-channel's size is slow compared to the variations in water discharge.
+Such water discharge variability in \alpine{} may cause shear stress to approach $Q^{2}$in assuming pipe-flow conditions \citep[Section \ref{sect:scaling}; c.f.][]{alley1997}.
+Conversely, the reduced variability in water discharge in \icesheet{} may result in the R-channel's size being closer to equilibrium from the smaller variations in water discharge.
+In this case, the exponent on water discharge for shear stress is likely less than $2$, but greater than $\frac{4}{5}$ that occurs when the R-channel's size evolves in equilibrium with water discharge (see Section \ref{sect:scaling}).
+As a result, there is a stronger relationship between subglacial model outputs and water discharge in \icesheet{} compared to \alpine{} (Figure~\ref{fig:Qw_vari}) and less variability model outputs in the \icesheet case (Figure~\ref{fig:multi_run}).
+
+\subsection{Interpreting sediment transport records from glacierized catchments with respect to water discharge}
+
+A sporadic relationship between subglacial sediment transport capacities and water discharge can occur depending on the subglacial channel size (Figure~\ref{fig:Qw_vari}), limiting the use of water discharge as indicator of sediment discharge capacity in these systems.
+As a result, characteristics such as bankfull or effective water discharge  that link geomorphic work to hydro-climatic conditions could have limited meaning in evaluating subglacial sediment transport \citep{wolman1960,lenzi2006}.
+A glacier's sediment transport capacity is impacted by the ice thickness controlling the channel closure rate, and the glacier's surface slope, in addition to water discharge and sediment size \citep[Figure~\ref{fig:multi_run}, Section~\ref{sect:sub_mode}; ][]{rothlisberger1972,gimbert2016,stevens2022,walder1994}.
+In a supply-limited state, a glacier's sediment discharge record also responds to  sediment availability and bedrock erosion from  water pressure variations and sliding  \citep[e.g.
+][]{iverson2012,herman2015,delaney2023}.
+As a result, the greater amount of in variability the sediment transport evacuation can make it  difficult to isolate the drives of sediment export with respect to hydrological and glaciological conditions \citep{delaney2024}.
+This multitude of processes lies in contrast to many subaerial channels, where transport capacity typically responds to water discharge in addition to sediment size, channel shape,  and hydraulic gradient, that can remain stable over years or longer \citep[Section~\ref{sect:fluv}; e.g.][]{tucker1997}.
+Strong correlations between water discharge and sediment export in glacier systems could be indicative of other processes, such as increased sediment access \citep{zhang2022}.
+Furthermore observed correlation between water discharge and sediment discharge far downstream of the glacier likely represents subaerial transport processes, especially with respect to bedload \citep{mancini2023}.
+
+The co-varying relationship between sediment discharge capacity and water discharge in  subaerial conditions could be noticeable $\sim20$\,\unit{km} downstream of the Leverett site, where a strong correlation persists between sediment plume size and the river's water discharge into the Kangerlussuaq fjord \citep{chu2009,mcgrath2010}.
+In contrast, in marine-terminating glacier catchments, a less consistent relationship may occur between water discharge or melt extent and sediment plume size \citep{chu2012,tedstone2012}.
+Here, ocean water pressurizes subglacial channels at the ice front \citep[e.g.][]{how2017}, so the reduced correlation could result from the inconsistent relationship between subglacial sediment transport capacity and water discharge presented here.
+
+Water discharge measurements at the hourly timescale remain unavailable in many catchments, further limiting subglacial  sediment transport models in capturing specific events.
+The model output variability in subglacial channels decreases substantially beyond $1-5$ days of smoothing (Figure~\ref{fig:multi_run}).
+Over this time period, subglacial channels could be in closer equilibrium conditions with water discharge variations leading to a stronger relationship with sediment discharge capacity (Section ~\ref{sect:scaling}).
+Applying water discharge averaged over these time periods omits key sediment discharge characteristics, such as the timing of peaks in sediment transport capacity from glaciers (Figure~\ref{fig:model_outs}), that could play a key role in sediment export from glaciers.
+Yet, hydrographs with minimal variability may result in a strong correlation between sediment transport capacity and water discharge, if the water discharge evolves slowly with respect the subglacial channel size.
+
+\conclusions
+
+The sediment transport capacity of a channel is driven by its width and the shear stress exerted by the flowing water, which is proportional to  flow velocity squared.
+Subaerial channels alter both their channel width and water velocity immediately in response to changing water discharge.
+In contrast, pressurized subglacial channels largely accommodate rapidly changing water discharge by altering water velocity.
+Their size responds over several days to changes in discharge.
+Thus the response of sediment transport capacity is much more sensitive to changes in water discharge in subglacial channels compared to subaerial ones, although not as variable as if the size of the subglacial conduit was fixed.
+
+With respect to the  manuscript's first objective,
+results demonstrate that sediment transport capacity variability is higher in subglacial channels with evolving size across a range of channel shapes and friction factors.
+Increased variability even occurs in environments where the water discharge has relatively small diurnal variations, as in the \icesheet case.
+The role of this increased noise in sediment transport capacity in evaluating hydro-climatic signals from sediment records in glacierized catchments needs further evaluation.
+Furthermore, this processes may lead to sediment exhaustion by subglacial water repeatedly crossing the mobilization threshold across a range of water discharges.
+
+The manuscript's second objective is establish if water discharge varies with subglacial sediment transport capacity.
+Results here suggest that, even in a transport-limited subglacial system, an incoherent relationship between water and subglacial sediment discharge could be expected.
+This presents a challenge in linking hydro-climatic conditions to sediment export from glaciers.
+Due to evolving subglacial channel size, the timing of peak sediment discharge capacity in subglacial channels occurs before peak  water discharge, whereas it is coincident in subaerial ones.
+Our analysis shows that high correlation between water discharge and sediment transport capacity could be possible when water discharge varies at a slower rate than subglacial channel size.
+In natural environments, this is appears to often not be the case, but might be conceivable in very large catchments or ones at high latitudes with muted diurnal variations in melt.
+
+Lack of observation of the subglacial environment hindered our ability to quantify processes such as
+the shape of subglacial channels, the response time of subglacial channels to water discharge variations, and
+subglacial sediment size and sorting processes, linking shear stress to  sediment discharge.
+Further quantifying these processes will help to better inform the response of sediment discharge capacity to water discharge forcing in subglacial environments.
+
+This study calls for careful analysis when examining the relationship between sediment transport capacity and hydro-climatic conditions from glacierized catchments, especially ones with high water discharge variability.
+
+\codedataavailability{
+Code, with links to the data, can be found at \url{https://bitbucket.org/IanDelaney/xsection/src/master/}.
+Data from Leverett glacier have been previous published in \citet{tedstone2013}.
+Data from Fiescher glacier have been previous published in \citet{delaney2024}.
+Code will be uploaded to a permanent repository, with FAIR principles pending acceptance of the paper.}
+
+\authorcontribution{
+I.D. designed the study, developed, and implemented the model and experiments, and wrote the manuscript.
+A.J.T. helped interpret data from Leverett Glacier and provided guidance on experiment design.
+M.A.W. provided data from Fieschergletscher and added the relationship between shear stress and water discharge (see Section \ref{sect:scaling}).
+D.F. provided data from Fieschergletscher.
+All authors provided key inputs in writing and editing the manuscript.} %% this section is mandatory
+
+\competinginterests{The authors declare no competing interests.} %% this section is mandatory even if you declare that no competing interests are present
+
+\begin{acknowledgements}
+SNSF Project No. $\mathrm{PZ00P2}$\_$202024$ provided  funding for I. Delaney.
+A. Tedstone acknowledges funding from the European Research Council, award $818994$ -- CASSANDRA.
+G. King and M. J. Gevers provided useful comments on a previous version of this manuscript.
+We benefited from fruitful discussions with J. Irving.
+We thank three reviewers for constructive and critical reviews on a previous version of the manuscript.
+\end{acknowledgements}
+\newpage
+
+\appendix
+\appendixtables   %% needs to be added in front of appendix tables
 \begin{table}
@@
     Subaerial (\(\alpha=0\)) & \(f^{0.7}\, Q^{0.7}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{1.7} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.1} \, \Psi^{1.7}\)\\
+    Subaerial (\(\alpha=0.3\)) & \(f^{0.7}\, Q^{0.8}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{1.3} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\
     Subaerial (\(\alpha=0.5\)) & \(f^{0.7}\, Q^{0.8}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{1.3} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\
@@
 
+%\appendixfigures  %% needs to be added in front of appendix figures
+
+%
+
+%% Please add \clearpage between each table and/or figure. Further guidelines on figures and tables can be found below.
+
+\bibliographystyle{copernicus}
+\bibliography{Paperlib.bib}
+
+%% REFERENCES
+
+%% The reference list is compiled as follows:
+
+% \begin{thebibliography}{}
+
+% \bibitem[AUTHOR(YEAR)]{LABEL1}
+% REFERENCE 1
+
+% \bibitem[AUTHOR(YEAR)]{LABEL2}
+% REFERENCE 2
+
+% \end{thebibliography}
+
+%% Since the Copernicus LaTeX package includes the BibTeX style file copernicus.bst,
+%% authors experienced with BibTeX only have to include the following two lines:
+%%
+%% 
+%% \bibliography{example.bib}
+%%
+%% URLs and DOIs can be entered in your BibTeX file as:
+%%
+%% URL = {http://www.xyz.org/~jones/idx_g.htm}
+%% DOI = {10.5194/xyz}
+
+%% LITERATURE CITATIONS
+%%
+%% command                        & example result
+%% \citet{jones90}|               & Jones et al. (1990)
+%% \citep{jones90}|               & (Jones et al., 1990)
+%% \citep{jones90,jones93}|       & (Jones et al., 1990, 1993)
+%% \citep[p.~32]{jones90}|        & (Jones et al., 1990, p.~32)
+%% \citep[e.g.,][]{jones90}|      & (e.g., Jones et al., 1990)
+%% \citep[e.g.,][p.~32]{jones90}| & (e.g., Jones et al., 1990, p.~32)
+%% \citeauthor{jones90}|          & Jones et al.
+%% \citeyear{jones90}|            & 1990
+
+%% FIGURES
+
+%% When figures and tables are placed at the end of the MS (article in one-column style), please add \clearpage
+%% between bibliography and first table and/or figure as well as between each table and/or figure.
+
+% The figure files should be labelled correctly with Arabic numerals (e.g. fig01.jpg, fig02.png).
+
+%% ONE-COLUMN FIGURES
+
+%%f
+%\begin{figure}[t]
+%\includegraphics[width=8.3cm]{FILE NAME}
+%\caption{TEXT}
+%\end{figure}
+%
+%%% TWO-COLUMN FIGURES
+%
+%%f
+%\begin{figure*}[t]
+%\includegraphics[width=12cm]{FILE NAME}
+%\caption{TEXT}
+%\end{figure*}
+%
+%
+%%% TABLES
+%%%
+%%% The different columns must be seperated with a & command and should
+%%% end with \\ to identify the column brake.
+%
+%%% ONE-COLUMN TABLE
+%
+%%t
+%\begin{table}[t]
+%\caption{TEXT}
+%\begin{tabular}{column = lcr}
+%\tophline
+%
+%\middlehline
+%
+%\bottomhline
+%\end{tabular}
+%\belowtable{} % Table Footnotes
+%\end{table}
+%
+%%% TWO-COLUMN TABLE
+%
+%%t
+%\begin{table*}[t]
+%\caption{TEXT}
+%\begin{tabular}{column = lcr}
+%\tophline
+%
+%\middlehline
+%
+%\bottomhline
+%\end{tabular}
+%\belowtable{} % Table Footnotes
+%\end{table*}
+%
+%%% LANDSCAPE TABLE
+%
+%%t
+%\begin{sidewaystable*}[t]
+%\caption{TEXT}
+%\begin{tabular}{column = lcr}
+%\tophline
+%
+%\middlehline
+%
+%\bottomhline
+%\end{tabular}
+%\belowtable{} % Table Footnotes
+%\end{sidewaystable*}
+%
+%
+%%% MATHEMATICAL EXPRESSIONS
+%
+%%% All papers typeset by Copernicus Publications follow the math typesetting regulations
+%%% given by the IUPAC Green Book (IUPAC: Quantities, Units and Symbols in Physical Chemistry,
+%%% 2nd Edn., Blackwell Science, available at: http://old.iupac.org/publications/books/gbook/green_book_2ed.pdf, 1993).
+%%%
+%%% Physical quantities/variables are typeset in italic font (t for time, T for Temperature)
+%%% Indices which are not defined are typeset in italic font (x, y, z, a, b, c)
+%%% Items/objects which are defined are typeset in roman font (Car A, Car B)
+%%% Descriptions/specifications which are defined by itself are typeset in roman font (abs, rel, ref, tot, net, ice)
+%%% Abbreviations from 2 letters are typeset in roman font (RH, LAI)
+%%% Vectors are identified in bold italic font using \vec{x}
+%%% Matrices are identified in bold roman font
+%%% Multiplication signs are typeset using the LaTeX commands \times (for vector products, grids, and exponential notations) or \cdot
+%%% The character * should not be applied as mutliplication sign
+%
+%
+%%% EQUATIONS
+%
+%%% Single-row equation
+%
+%\begin{equation}
+%
+%\end{equation}
+%
+%%% Multiline equation
+%
+%\begin{align}
+%& 3 + 5 = 8\\
+%& 3 + 5 = 8\\
+%& 3 + 5 = 8
+%\end{align}
+%
+%
+%%% MATRICES
+%
+%\begin{matrix}
+%x & y & z\\
+%x & y & z\\
+%x & y & z\\
+%\end{matrix}
+%
+%
+%%% ALGORITHM
+%
+%\begin{algorithm}
+%\caption{...}
+%\label{a1}
+%\begin{algorithmic}
+%...
+%\end{algorithmic}
+%\end{algorithm}
+%
+%
+%%% CHEMICAL FORMULAS AND REACTIONS
+%
+%%% For formulas embedded in the text, please use \chem{}
+%
+%%% The reaction environment creates labels including the letter R, i.e. (R1), (R2), etc.
+%
+%\begin{reaction}
+%%% \rightarrow should be used for normal (one-way) chemical reactions
+%%% \rightleftharpoons should be used for equilibria
+%%% \leftrightarrow should be used for resonance structures
+%\end{reaction}
+%
+%
+%%% PHYSICAL UNITS
+%%%
+%%% Please use \unit{} and apply the exponential notation
+
 \end{document}
 
-% Graveyard
-% Additionally uggest that additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate \cite{jansson2005,ganti2016}.
-% Sediment discharge records have been used to establish the relationship, or lack thereof, between climate forcing and glacial erosion \cite<e.g.>{koppes2009a,ganti2016,willenbring2016,mariotti2021}.
-% The variable relationship between water discharge and subglacial sediment transport capacity presented here mostly applies to the daily- to weekly-time scales controlling the size of subglacial channels (Figure~\ref{fig:multi_run}).
-% Yet, identifying climatic effects in sediment transport from the transport-limited component of glacier-sedimentary systems may require higher thresholds of water discharge compared to subaerial systems when water discharge varies substantially over short timescales \cite<Figure~\ref{fig:Qw_vari}; >{tofelde2021}.
-\mauro{I don't understand the last two sentences.}
-% Furthermore evaluating the different variability in sediment transport characteristics across timescales can help inform the sediment transport processes omitted from daily or week values of water discharge..
+% \ian{comeback here}
+% For sufficiently short periods, channel width and wetted area are fixed in the subglacial system (akin to pipe flow) so that the response of width-integrated shear stress is $w_g\, \tau_{g}\, \propto\,  Q^{2}$ \citep[Table\,~\ref{tab:Qs}; ][]{alley1997}.
+
+% R\"othlisberger channels in steady state with water discharge (i.e.
+channel sizes evolves at a faster rate than water discharge), the width-integrated shear stress response to water discharge is $ w_g\,\tau_{g}\, \propto\,  Q^{\sim\frac{4}{5}}$ (Table\,~\ref{tab:Qs}).
+% This relationship suggests that when discharge varies more slowly than channel shape subglacial channels have less variable sediment transport capacity than subaerial ones.
+% Yet, subglacial channels rarely operate in a steady state \citep{gimbert2016}.
+% However, the reduced variability in the  \icesheet{} case compared to the \alpine{} one suggests that it is closer to an equilibrium state (Figure~\ref{fig:Qw_vari}).
+% \ian{comeback here}
+% The channel width and size variations in subglacial channels presented here also occur in subaerial channels in response to water discharge \citep{phillips2016}, likely over decadal timescales or in response to individual extreme events \citep{slater2013,dean2013}.
+% These timescales are considerably longer than the continuous changes to the subglacial channel's response time of days.
+% Even so, if channel width changes occur in subaerial channels, then the values of $\alpha$ in Equation~\ref{eq:wcf} could change in time.
+% $\alpha$ must be less than $1$, and even this would cause the exponent on $Q$ in the  $w_g\,\tau_{g}$ relationships to remain substantially below that of pipeflow (see Sections \ref{sect:scaling}, Table~\ref{tab:Qs}, Figure~\ref{fig:multi_run}).
+
+With even smaller water discharge variations, the subglacial sediment transport capacity may behave similarly to the subaerial system.
+Note that in pipe flow conditions, maximum sediment transport would coincide with peak water discharge, unlike the evolving channel size alters the timing (Figure \ref{fig:model_outs} and \ref{fig:Qw_vari}).

### 2024-06-27 Delaneyetal2024.tex
online edit

@@
   Here, we present relationships between water discharge and sediment transport under several channel types.
-  They demonstrate that sediment transport is highly non-linear with respect to water discharge with fixed subglacial channel size, and in fact can very less that subaerial flow if water discharge changes slowly enough to be in equilibrium with subglacial channel size.
+  They demonstrate that sediment transport is highly non-linear with respect to water discharge with fixed subglacial channel size, and in fact, can very less than subaerial flow if water discharge changes slowly enough to be in equilibrium with subglacial channel size.
   We then apply hydrographs from an Alpine glacier and the Greenland Ice Sheet to subglacial and subaerial hydraulics models, revealing greater variability in the inferred sediment transport capacity in subglacial than in subaerial channels, as subglacial channels evolve in size.
   This variability stems from the slowly evolving channel size that forces water velocity to accommodate variations in water discharge.
-  Furthermore, the changing channel size means that sediment transport capacity peaks prior to the maximum water discharge.
-  Results also show that high subglacial sediment transport rates and water velocities can occur across a wide range of water discharges, meaning that in a transport-limited system there could be a variable relationship between sediment and water discharge.
+  Furthermore, the changing channel size means that sediment transport capacity peaks before the maximum water discharge.
+  Results also show that high subglacial sediment transport rates and water velocities can occur across a wide range of water discharges, meaning that in a transport-limited system, there could be a variable relationship between sediment and water discharge.
   These findings are discussed in the context of understanding sediment discharge from glaciers with different hydro-climatic forcings and assumptions of channel shape.
@@
 Periods of millennia or longer, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from under glaciers \citep[c.f.][]{hallet1979,iverson2012,ugelvig2018}.
-Pressurized subglacial water can transport this sediment from underneath glaciers \citep{walder1994,creyts2013,beaud2018,delaney2019}, if it is reachable by the water.
+Pressurized subglacial water can transport this sediment from underneath glaciers \citep{walder1994,creyts2013,beaud2018,delaney2019} if it is reachable by the water.
 
@@
   \includegraphics[width=0.6\linewidth]{Fig1.pdf}
-  \caption{Sketch for the different responses of subglacial and subaerial channels to increased water discharge over short time-scales.
+  \caption{Sketch for the different responses of subglacial and subaerial channels to increased water discharge over short time scales.
     Arrow length denotes water velocity magnitudes in the subglacial (subaerial) $v_g$ ($v_f$) channels.
@@
 These differences are implicitly included in a range of models quantifying sediment transport in both subglacial and subaerial channels  \citep[e.g.][]{walder1994,alley1997,tucker1997,creyts2013,beaud2018,delaney2019,hewitt2019,wickert2019}.
-To date, explicit parameterizations of subglacial sediment transport capacity with respect to water discharge has assumed fixed channel size and demonstrates a very non-linear response in sediment transport capacity to water discharge \citep{alley1997}.
-The modeling frameworks examining subglacial sediment transport with evolving channel size minimally discuss the impact of different hydrological forcing on sediment transport capacity with the aim of understanding the impact of hydro-climatic conditions on subglacial sediment dynamics \citep{creyts2013,beaud2018,delaney2019,hewitt2019}.
+To date, explicit parameterizations of subglacial sediment transport capacity with respect to water discharge have assumed fixed channel size and demonstrate a very non-linear response in sediment transport capacity to water discharge \citep{alley1997}.
+The modeling frameworks examining subglacial sediment transport with evolving channel size minimally discuss the impact of different hydrological forcing on sediment transport capacity to understand the result of hydro-climatic conditions on subglacial sediment dynamics \citep{creyts2013,beaud2018,delaney2019,hewitt2019}.
 Yet, sediment discharge capacity controls the mobilization and deposition of sediment, making it one of the most fundamental aspects of subglacial sediment evacuation, especially over short timescales.
@@
 In subaerial systems, water discharge and sediment transport capacity for a grain size are closely linked \citep[e.g. ][]{schmidt2006,pitlick2021} \ian{check citiation}.
-However, the potential for a highly variable response in sediment transport capacity in subglacial sediment transport may affect the interpretation of sediment transport records in glacial systems \citep[e.g.][]{ganti2016,mancini2023}, if properties associate with subaerial systems are assumed.
+However, the potential for a highly variable response in sediment transport capacity in subglacial sediment transport may affect the interpretation of sediment transport records in glacial systems \citep[e.g.][]{ganti2016,mancini2023}, if properties associated with subaerial systems are assumed.
 This relationship should be clarified as sediment dynamics change in glacierized regions along with changing hydrology, increased variability of water discharge \citep{lane2019} and extreme glacier melt  \citep[e.g.][]{cremona2023,overeem2015} that may drive high magnitude sediment transport events.
 
-This manuscript as two objectives: 1) evaluate the variability of subglacial sediment transport capacity with evolving subglacial channels size and different hydrographs and 2) establish if subseasonal variations in water discharge can be covary with sediment transport capacity in subglacial systems.
-First, to reenforce the highly variable behavior of subglacial sediment transport capacity,  we present algebric formulations of the sediment transport capacity response to water discharge in subaerial, pipe-flow and equilibrium R-channel conditions, extending the relationships presented in \citet{alley1997}.
+This manuscript has two objectives: 1) evaluate the variability of subglacial sediment transport capacity with evolving subglacial channel size and different hydrographs and 2) establish if sub-seasonal variations in water discharge can co-vary with sediment transport capacity in subglacial systems.
+First, to reinforce the highly variable behavior of subglacial sediment transport capacity,  we present algebraic formulations of the sediment transport capacity response to water discharge in subaerial, pipe-flow, and equilibrium R-channel conditions, extending the relationships presented in \citet{alley1997}.
 We then use numerical models to examine sediment transport capacity in subglacial channels that evolve in size and subaerial channels.
@@
 Our evaluation in terms of shear stress omits the selection of a sediment transport relationship and a grain-size parameter \citep[e.g.][]{shields1936,meyer1948}.
-Through this choice we also negate sorting processes
 
@@
 
-Here, it is assumed that the water is transported through a subglacial channel \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972} beneath a glacier with channel length $l$, with a flat bed and a mean thickness of $h_{ice}$.
+Here, it is assumed that the water is transported through a subglacial channel \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972} beneath a glacier with channel length $l$, with a flatbed and a mean thickness of $h_{ice}$.
 The channel size grows from melt due to frictional heating from water flow and closes due to the creep of the ice.
@@
 
-The first experiment aims to characterize the variability in sediment discharge capacity in subglacial compared to subaerial channels across a range of channels slopes and  shapes, along with friction factors.
+The first experiment aims to characterize the variability in sediment discharge capacity in subglacial compared to subaerial channels across a range of channel slopes and shapes, along with friction factors.
 Additionally, we examine the effects of different water discharge variability by smoothing the water discharge over different periods.
@@
 
-The effect of channel shape, gradient and roughness is established by running the model with random parameter values of channel slope and geometry factors ($\beta$,$\nabla z_c$ and $\alpha$) and friction factors ($f_r$ and $f_f$; see Table~\ref{table:vpm} for value range).
+The effect of channel shape, gradient, and roughness is established by running the model with random parameter values of channel slope and geometry factors ($\beta$,$\nabla z_c$, and $\alpha$) and friction factors ($f_r$ and $f_f$; see Table~\ref{table:vpm} for value range).
 Runs with parameter combinations are accepted if their mean subglacial water velocity over the season lies between $0.5$\,\unit{m}\,\unit{s}$^{-1}$ and $2$\,\unit{m}\,\unit{s}$^{-1}$ or if subaerial water velocity lies between $0.3$\,\unit{m}\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}\,\unit{s}$^{-1}$, following measurements \citep[e.g.][]{werder2010b,magnusson2012,chandler2013}.
@@
 
-Sediment discharge is given by the MPM, EH and Bagnold formulations as
+Sediment discharge is given by the MPM, EH, and Bagnold formulations as
 \begin{equation}
@@
 For a steady state R-channel, it is assumed that  $\Psi$ is constant (approximated by the gradient of the Shreve \citeyear{shreve1972} potential and that $S$ adjusts such that it is in steady state with $\Psi$ and $Q$.
-Note that the R-channel model presented above in Section~\ref{sect:sub_mode} Equation~\ref{eq:dS_dt} calculates $\Psi$ from the time evolving $S_g$ via the Darcy-Weisbach Equation~\ref{eq:dh} and thus no Shreve approximation is then needed.
+Note that the R-channel model presented above in Section~\ref{sect:sub_mode} Equation~\ref{eq:dS_dt} calculates $\Psi$ from the time-evolving $S_g$ via the Darcy-Weisbach Equation~\ref{eq:dh} and thus no Shreve approximation is then needed.
 % Conversely, pipe flow has $S$ fixed and $\Psi$ adjusts to satisfy the given $Q$.
 % 
-Pipe flow like conditions occur when a R-channel is subjected to rapid discharge variations such that the channel cannot adjust its size significantly on the time scale of the discharge variations.
+Pipe flow-like conditions occur when an R-channel is subjected to rapid discharge variations such that the channel cannot adjust its size significantly on the time scale of the discharge variations.
 In this case, it is assumed that the cross-sectional area $S$ is fixed, and $\Psi$ adjusts to the specified $Q$.
@@
 
-With these assumptions the Darcy-Weisbach equation~\eqref{eq:DW} can be solved for the not-fixed quantity: $H$ for a  subaerial channel, $S$ for a R-channel, and $\Psi$ for pipe-flow.
+With these assumptions, the Darcy-Weisbach equation~\eqref{eq:DW} can be solved for the not-fixed quantity: $H$ for a subaerial channel, $S$ for a R-channel, and $\Psi$ for pipe-flow.
 Then, using equations~\eqref{eq:tau-omega}, the shear stress and stream power can be calculated.
 These results are summarised in Table~\ref{tab:eqs1}.
-Now width integrated shear stress and sediment transport can be calculated for all regimes and for all transport formulas.
+Now width integrated shear stress and sediment transport can be calculated for all regimes and all transport formulas.
 The results of the 12 combinations are presented in Table~\ref{tab:Qs} and also in Table~\ref{tab:Qs} where the fractions in the exponents are approximately given by decimal numbers for ease of comparison.
@@
 \begin{table}[hbt!]
-  \caption{Relations for hydraulic variables for the three drainage regimes:  subaerial channels, R-channel and pipe-flow.
+  \caption{Relations for hydraulic variables for the three drainage regimes:  subaerial channels, R-channel, and pipe-flow.
 Darcy-Weisbach equation is abbreviated with ``D-W'', and stream power with ``Stream p.''.
@@
 
-Table~\ref{tab:Qs} shows how the proxy $w\tau$ as well as $Q_s$ of the MPM, EH and Bagnold sediment transport formulas scale with respect to $Q$, $\Psi$ or $S$ for the three different regimes and for different $\alpha$.
+Table~\ref{tab:Qs} shows how the proxy $w\tau$ as well as $Q_s$ of the MPM, EH, and Bagnold sediment transport formulas scale with respect to $Q$, $\Psi$ or $S$ for the three different regimes and for different $\alpha$.
 Remarkably, the Bagnold formula has a negative exponent for $f$ in all but the pipe-flow regime.
 The total transport formula EH gives a slightly stronger dependence on all variables as should be expected due to the larger exponent on $\tau$ of $\frac{5}{2}$ versus  $\frac{3}{2}$  for the others.
-Albeit, the sediment transport response in pipe-flow for the Bagnold case is close to EH.
+Albeit, the sediment transport response in pipe flow for the Bagnold case is close to EH.
 Conversely, the sediment transport proxies $w_g\,\tau_g$ \& $w_f\,\tau_f$ used below scale only as $Q^2$ for pipe-flow, whereas the transport scales at least with $Q^3$.
-The exponent for the width scaling $\alpha$ only impacts relationship between sediment transport and water discharge in the EH relation in any meaningful way.
+The exponent for the width scaling $\alpha$ only impacts the relationship between sediment transport and water discharge in the EH relation in any meaningful way.
 However, for the value of $\alpha$ around $\frac{1}{2}$ (a value appropriate for most streams), the exponent on $Q$ for EH is only slightly above what the other transport relations give.
@@
 
-Furthermore, R-channels will rarely operate in steady state as variations in discharge, in particular on the diurnal timescale or the timescale of severe rain or melting events, are too fast for such a channel to reach a steady state.
+Furthermore, R-channels will rarely operate in a steady state as variations in discharge, in particular on the diurnal timescale or the timescale of severe rain or melting events, are too fast for such a channel to reach a steady state.
 In such cases they operate more like a pipe of fixed cross-section \citep[e.g.][]{gimbert2016}.
 Table~\ref{tab:Qs} shows that in such a situation, the sediment transport scales much more severely with discharge, with the exponent on $Q$ being between $3$ and $5$, compared to the other two regimes when that exponent is at most $1.7$ (Table~\ref{tab:Qs}).
-Thus fluctuations of discharge on short timescales (on the order of a day) have the potential to cause conditions with very high sediment transport capacities, that are of far high magnitude than the subaerial channels.
+Thus fluctuations of discharge on short timescales (on the order of a day) have the potential to cause conditions with very high sediment transport capacities, that are of far higher magnitude than the subaerial channels.
 
@@
 \FloatBarrier
-\subsection{Increased variability in evolving subglacial channels occurs across a range of channel shapes, slopes  and friction values}
+\subsection{Increased variability in evolving subglacial channels occurs across a range of channel shapes, slopes,  and friction values}
 \label{sect:ensemble}
@@
 The first numerical experiment aims to compare the variability between the subglacial and subaerial model outputs to a range of channel shapes, friction factors, and water discharge variability, as the subglacial channel's cross-sectional area is allowed to evolve.
-Following the findings in Section~\ref{sect:scaling}, variability in subglacial sediment discharge capacity could approach that of subaerial systems, should water discharge variations be sufficienlty small or subglacial channels respond in size quickly.
+Following the findings in Section~\ref{sect:scaling}, variability in subglacial sediment discharge capacity could approach that of subaerial systems, should water discharge variations be sufficiently small or subglacial channels respond in size quickly.
 For simplicity, we only examine $w\tau$, not the sediment transport formulas mentioned above.
@@
 Results from both \alpine{} and \icesheet{} scenarios show that variability in both subglacial and subaerial outputs decreases with smoothing time longer than approximately 1--5 days (Figure~\ref{fig:multi_run}), when the diurnal variations in water discharge are removed from smoothing.
-In the \icesheet case, the smaller variability in model outputs compared to \alpine may in fact mean that the R-channel is in closer equilibrium between channel size and water discharge (Section~\ref{sect:scaling}).
+In the \icesheet case, the smaller variability in model outputs compared to \alpine may mean that the R-channel is in closer equilibrium between channel size and water discharge (Section~\ref{sect:scaling}).
 In this case, water discharge variations would create smaller disparities in sediment transport capacity than even in subaerial cases as the subglacial channel size evolves to accommodate them (Table~\ref{tab:Qs}).
@@
 Yet, larger values of $\alpha$ can result in greater variability in width-integrated shear stress, as greater channel width changes increase the width-integrated shear stress, but not the shear stress.
-Additionally, steeper subaearial channel slopes result in greater variability in both shear stress and width-integrated shear stress, but only approach the variability of the subglacial channel in the \alpine case (Figure~\ref{fig:params}).
+Additionally, steeper subaerial channel slopes result in greater variability in both shear stress and width-integrated shear stress but only approach the variability of the subglacial channel in the \alpine case (Figure~\ref{fig:params}).
 %At steeper slopes, the alluvial processes represented in the model (Section ~\ref{sect:fluv}) probably do not apply, as the.
@@
 
-Greater values of the subglacial and subaerial friction factor $f_i$ and $f_p$ result in greater variability in the  shear stress and width-integrate shear stress both cases (Figure~\ref{fig:params}), as expected given Equations~\ref{eq:tau_g} and \ref{eq:vf}.
+Greater values of the subglacial and subaerial friction factor $f_i$ and $f_p$ result in greater variability in the shear stress and width-integrate shear stress in both cases (Figure~\ref{fig:params}), as expected given Equations~\ref{eq:tau_g} and \ref{eq:vf}.
 However, more velocity variability is present in low values of $f_i$ and $f_p$. 
@@
 Several aspects of the models and experiment design limit the comparison of the model outputs to field data and make comparison of the two systems difficult.
-The lumped nature of the models mean that they operate independently of the upstream drainage network, and sediment access is not considered here.
+The lumped nature of the models means that they operate independently of the upstream drainage network, and sediment access is not considered here.
 In reality, processes such as sediment access are highly important in controlling sediment export in glacierized catchments \citep[e.g.][]{herman2015,vergara2022}.
@@
 This occurs due to the different gradients in the two systems, with the subglacial one being much steeper \citep{alley1997}.
-We note, however, that in the ensemble runs in Section~\ref{sect:ensemble}, the range of parameters tested cover the range of viable shear stresses in both subglacial and subaerial channels.
+We note, however, that in the ensemble runs in Section~\ref{sect:ensemble}, the range of parameters tested covers the range of viable shear stresses in both subglacial and subaerial channels.
 
-Width-integrated shear stress is examined in the two numerical experiments as it does not have a dependence on grain-size, unlike sediment transport relationships, such as \citet[][Section \ref{sect:scaling}]{meyer1948}.
+Width-integrated shear stress is examined in the two numerical experiments as it does not have a dependence on grain size, unlike sediment transport relationships, such as \citet[][Section \ref{sect:scaling}]{meyer1948}.
 This makes comparison between the two systems simpler. 
-Preferential transport of smaller sediment clasts and the input of upstream sediment impacts sediment transport capacity as grain-size evolves \citep[e.g.][]{gomez1983}.
+Preferential transport of smaller sediment clasts and the input of upstream sediment impact sediment transport capacity as grain size evolves \citep[e.g.][]{gomez1983}.
 These processes are only beginning to be evaluated below glaciers \citep{aitken2024}, but can be important in subaerial systems.
-Subglacial sorting processes could be an additional source of variability in sediment export from glacier with respect to water discharge, especially with respect to bedload.
+Subglacial sorting processes could be an additional source of variability in sediment export from glaciers with respect to water discharge, especially for bedload transport.
 
@@
 
-A sporadic relationship between subglacial sediment transport capacities and water discharge can occur depending on the subglacial channel size (Figure~\ref{fig:Qw_vari}), limiting the use of water discharge as indicator of sediment discharge capacity in these systems.
+A sporadic relationship between subglacial sediment transport capacities and water discharge can occur depending on the subglacial channel size (Figure~\ref{fig:Qw_vari}), limiting the use of water discharge as an indicator of sediment discharge capacity in these systems.
 As a result, characteristics such as bankfull or effective water discharge that link geomorphic work to hydro-climatic conditions could have limited meaning in evaluating subglacial sediment transport \citep{wolman1960,lenzi2006}.
@@
 As a result, the greater amount of in variability the sediment transport evacuation can make it difficult to isolate the drives of sediment export with respect to hydrological and glaciological conditions \citep{delaney2024}.
-This multitude of processes lies in contrast to many subaerial channels, where transport capacity typically responds to water discharge in addition to sediment size, channel shape,  and hydraulic gradient, that can remain stable over years or longer \citep[Section~\ref{sect:fluv}; e.g.][]{tucker1997}.
+This multitude of processes lies in contrast to many subaerial channels, where transport capacity typically responds to water discharge in addition to sediment size, channel shape,  and hydraulic gradient, that can remain stable over the years or longer \citep[Section~\ref{sect:fluv}; e.g.][]{tucker1997}.
 Strong correlations between water discharge and sediment export in glacier systems could be indicative of other processes, such as increased sediment access \citep{zhang2022}.
-Furthermore observed correlation between water discharge and sediment discharge far downstream of the glacier likely represents subaerial transport processes, especially with respect to bedload \citep{mancini2023}.
+Furthermore, the observed correlation between water discharge and sediment discharge far downstream of the glacier likely represents subaerial transport processes, especially with respect to bedload \citep{mancini2023}.
 
@@
 Water discharge measurements at the hourly timescale remain unavailable in many catchments, further limiting subglacial sediment transport models in capturing specific events.
-The model output variability in subglacial channels decreases substantially beyond $1-5$ days of smoothing (Figure~\ref{fig:multi_run}).
-Over this time period, subglacial channels could be in closer equilibrium conditions with water discharge variations leading to a stronger relationship with sediment discharge capacity (Section ~\ref{sect:scaling}).
-Applying water discharge averaged over these time periods omits key sediment discharge characteristics, such as the timing of peaks in sediment transport capacity from glaciers (Figure~\ref{fig:model_outs}), that could play a key role in sediment export from glaciers.
-Yet, hydrographs with minimal variability may result in a strong correlation between sediment transport capacity and water discharge, if the water discharge evolves slowly with respect the subglacial channel size.
+The model output variability in subglacial channels decreases substantially beyond $1--5$ days of smoothing (Figure~\ref{fig:multi_run}).
+Over this period, subglacial channels could be in closer equilibrium conditions with water discharge variations leading to a stronger relationship with sediment discharge capacity (Section ~\ref{sect:scaling}).
+Applying water discharge averaged over these periods omits key sediment discharge characteristics, such as the timing of peaks in sediment transport capacity from glaciers (Figure~\ref{fig:model_outs}), that could play a key role in sediment export from glaciers.
+Yet, hydrographs with minimal variability may result in a strong correlation between sediment transport capacity and water discharge, if the water discharge evolves slowly with respect to the subglacial channel size.
 
@@
 
-With respect to the  manuscript's first objective,
+Concerning the manuscript's first objective,
 results demonstrate that sediment transport capacity variability is higher in subglacial channels with evolving size across a range of channel shapes and friction factors.
-Increased variability even occurs in environments where the water discharge has relatively small diurnal variations, as in the \icesheet case.
+Increased variability occurs in environments where the water discharge has relatively small diurnal variations, as in the \icesheet case.
 The role of this increased noise in sediment transport capacity in evaluating hydro-climatic signals from sediment records in glacierized catchments needs further evaluation.
-Furthermore, this processes may lead to sediment exhaustion by subglacial water repeatedly crossing the mobilization threshold across a range of water discharges.
+Furthermore, this process may lead to sediment exhaustion by subglacial water repeatedly crossing the mobilization threshold across a range of water discharges.
 
-The manuscript's second objective is establish if water discharge varies with subglacial sediment transport capacity.
+The manuscript's second objective is to establish if water discharge varies with subglacial sediment transport capacity.
 Results here suggest that, even in a transport-limited subglacial system, an incoherent relationship between water and subglacial sediment discharge could be expected.
 This presents a challenge in linking hydro-climatic conditions to sediment export from glaciers.
-Due to evolving subglacial channel size, the timing of peak sediment discharge capacity in subglacial channels occurs before peak  water discharge, whereas it is coincident in subaerial ones.
-Our analysis shows that high correlation between water discharge and sediment transport capacity could be possible when water discharge varies at a slower rate than subglacial channel size.
-In natural environments, this is appears to often not be the case, but might be conceivable in very large catchments or ones at high latitudes with muted diurnal variations in melt.
+Due to evolving subglacial channel size, the timing of peak sediment discharge capacity in subglacial channels occurs before peak water discharge during an event, whereas it is coincident in subaerial ones.
+Our analysis shows that a high correlation between water discharge and sediment transport capacity could be possible when water discharge varies at a slower rate than subglacial channel size.
+In natural environments, this appears to often not be the case but might be conceivable in very large catchments or ones at high latitudes with muted diurnal variations in melt.
 

### 2024-07-01 Delaneyetal2024.tex
ESurf update

@@
-%% Copernicus Publications Manuscript Preparation Template for LaTeX Submissions
+%%Copernicus Publications Manuscript Preparation Template for LaTeX Submissions
 %% ---------------------------------
@@
 \newcommand{\mauro}[1]{{\textbf{\color{green}Mauro says:} \color{green} #1} }
-\newcommand{\alpine}{\textit{ALPINE}}
-\newcommand{\icesheet}{\textit{ICESHEET}}
+\newcommand{\alpine}{\textit{ALPINE}\,}
+\newcommand{\icesheet}{\textit{ICESHEET}\,}
 \newcommand{\m}{$\,\mathrm{m}$\,}
@@
 \newcommand{\mmms}{$\,\mathrm{m^3\, s^{-1}}$\,}
-%\newcommand{\unit}[1]{$\mathrm{#1}$}
+% \newcommand{\unit}[1]{$\mathrm{#1}$}
 
@@
 \title{Subglacial and subaerial fluvial sediment transport capacity respond differently to water discharge variations}
-
-% \Author[affil]{given_name}{surname}
 
@@
 \affil[4]{Swiss Federal Institute for Forest, Snow and Landscape Research (WSL) Z\"uricherstrasse 111, 8903 Birmensdorf, Switzerland}
-%% The [] brackets identify the author with the corresponding affiliation. 1, 2, 3, etc. should be inserted.
-
-%% If an author is deceased, please mark the respective author name(s) with a dagger, e.g. "\Author[2,$\dag$]{Anton}{Smith}", and add a further "\affil[$\dag$]{deceased, 1 July 2019}".
-
-%% If authors contributed equally, please mark the respective author names with an asterisk, e.g.
-"\Author[2,*]{Anton}{Smith}" and "\Author[3,*]{Bradley}{Miller}" and add a further affiliation: "\affil[*]{These authors contributed equally to this work.}".
 
@@
 
-%% These dates will be inserted by Copernicus Publications during the typesetting process.
-
 \firstpage{1}
@@
 \begin{abstract}
-  Sediment transport capacity in both subaerial and subglacial channels depends on the shear stress exerted across the channel bottom, which varies with water velocity and channel width.
-  In subaerial channels, water discharge variations are accommodated by flow depth and width changes, along with water velocity.
-  However, in subglacial channels,  water discharge changes mainly result in velocity variations, as the channel geometry evolves slowly (over days).
-  Here, we present relationships between water discharge and sediment transport under several channel types.
-  They demonstrate that sediment transport is highly non-linear with respect to water discharge with fixed subglacial channel size, and in fact, can very less than subaerial flow if water discharge changes slowly enough to be in equilibrium with subglacial channel size.
-  We then apply hydrographs from an Alpine glacier and the Greenland Ice Sheet to subglacial and subaerial hydraulics models, revealing greater variability in the inferred sediment transport capacity in subglacial than in subaerial channels, as subglacial channels evolve in size.
-  This variability stems from the slowly evolving channel size that forces water velocity to accommodate variations in water discharge.
-  Furthermore, the changing channel size means that sediment transport capacity peaks before the maximum water discharge.
-  Results also show that high subglacial sediment transport rates and water velocities can occur across a wide range of water discharges, meaning that in a transport-limited system, there could be a variable relationship between sediment and water discharge.
-  These findings are discussed in the context of understanding sediment discharge from glaciers with different hydro-climatic forcings and assumptions of channel shape.
-  We also examine the implications for interpreting records of sediment discharge in glacierized catchments
+  Sediment transport capacity in both subaerial and subglacial channels depends on the shear stress exerted across the channel bottom.
+  This quantity varies with water velocity and channel width.
+  In subaerial channels,  sediment transport capacity can evolve consistently because water discharge variations are accommodated by changes in flow depth, width, and water velocity.
+  By contrast, subglacial channels grow at a slower rate than most water discharge variations, causing a variable relationship between channel shape, water velocity, and discharge.
+  Using numerical models applied to two observed hydrographs and formulations, we evaluate the differences in sediment transport capacity variability between subaerial and subglacial channels.
+  We also establish the conditions under which subglacial sediment transport capacity could co-vary with water discharge.
+  The formulations of hydraulics and numerical models confirm that subglacial sediment transport is highly non-linear with respect to water discharge.
+  Yet, subglacial sediment transport capacity  can vary less if water discharge variations are closer to equilibrium with subglacial channel size.
+  In another numerical experiment, we show that the changing channel size results in sediment transport capacity peaking before the maximum water discharge.
+  Results also indicate that high subglacial sediment transport capacities can occur across a wide range of water discharges.
+  This implies that  hysteresis in channel size causes a variable relationship between sediment and water discharge in a transport-limited subglacial system.
+  The implications are discussed for sediment discharge from glaciers with different hydro-climatic forcings and assumptions of channel shape.
+  These findings can also improve interpretations of sediment discharge records in glacierized catchments.
 \end{abstract}
@@
 
-Changes in glacier dynamics and hydrology have prompted numerous recent studies of sediment transport processes in cold regions \citep[e.g.][]{li2022,vergara2022,zhang2022}.
+Changes in glacier dynamics and hydrology have motivated numerous recent studies of sediment transport processes in cold regions \citep[e.g.][]{li2022,vergara2022,zhang2022}.
 Increases in sediment transport have been observed in Greenland \citep{bendixen2017}, the European Alps \citep{costa2017}, the Himalayas \citep{li2021}, and the Andes \citep{vergara2022}.
-To explain observed changes to sediment transport in glacierized catchments requires examining the processes controlling sediment discharge and its variations with water discharge  \citep[e.g.][]{riihimaki2005,swift2005}.
+To accurately explain observed changes to sediment transport in glacierized catchments requires examining the processes controlling sediment discharge and its variations with water discharge  \citep[e.g.][]{riihimaki2005,swift2005}.
 
-Periods of millennia or longer, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from under glaciers \citep[c.f.][]{hallet1979,iverson2012,ugelvig2018}.
+Glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from under glaciers over periods of millennia or longer \citep[c.f.][]{hallet1979,iverson2012,ugelvig2018}.
 Pressurized subglacial water can transport this sediment from underneath glaciers \citep{walder1994,creyts2013,beaud2018,delaney2019} if it is reachable by the water.
@@
 In a transport-limited regime, sediment discharge is controlled by sediment transport capacity, defined as the amount of sediment the water can carry.
-In both subglacial and subaerial channels, sediment transport capacity depends on the shear stress between water and the sediment it flows over \citep{shields1936,meyer1948,engelund1967} along with the width of the channel bottom $w$ over which sediment mobilizes.
+In both subglacial and subaerial channels, sediment transport capacity depends on the shear stress between water and the sediment it flows over \citep{shields1936,meyer1948,engelund1967}, along with the width of the channel bottom $w$ over which sediment mobilizes.
 The shear stress $\tau$ responds to the velocity of water $v$ flowing through the channel so that
@@
 
-In subaerial channels, operating with open channel flow, the wetted area $S$ of the channel evolves with changing water discharge $Q$, by changing both the channel's width and the water depth \citep{leopold1953}.
+In subaerial channels operating with open channel flow, the wetted area $S$ of the channel evolves with changing water discharge $Q$, by changing both the channel width and the water depth \citep{leopold1953}.
 The change in water depth results in a proportional increase in water velocity, and the shear stress $\tau$ increases according to Equation~\ref{eq:tau}.
 
-The response of water velocity to changing water discharge in subglacial channels differs from subaerial ones.
+The response of water velocity to changing water discharge in subglacial channels differs from subaerial ones, however.
 The size of subglacial channels is controlled by the opposing processes of channel opening by frictional heating of water flow on the one hand versus creep closure by ice flow on the other \citep{rothlisberger1972}.
-As a result, the subglacial channel size only evolves relatively slowly over days, whereas water discharge can vary more quickly over hours \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
-Therefore, on short periods, subglacial water flow behaves more like pipe flow, and changes in water discharge $Q$ are mainly accommodated by changing water velocity $v$ \citep[Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}; ][]{alley1997}.
+As a result  the subglacial channel size only evolves relatively slowly over days, whereas water discharge can vary more quickly over hours \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
+Therefore, subglacial water flow behaves more like pipe flow over short periods (hours, days).
+Changes in water discharge $Q$ are mainly accommodated by changing water velocity $v$ \citep[Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}; ][]{alley1997}.
 
@@
 These differences are implicitly included in a range of models quantifying sediment transport in both subglacial and subaerial channels  \citep[e.g.][]{walder1994,alley1997,tucker1997,creyts2013,beaud2018,delaney2019,hewitt2019,wickert2019}.
-To date, explicit parameterizations of subglacial sediment transport capacity with respect to water discharge have assumed fixed channel size and demonstrate a very non-linear response in sediment transport capacity to water discharge \citep{alley1997}.
-The modeling frameworks examining subglacial sediment transport with evolving channel size minimally discuss the impact of different hydrological forcing on sediment transport capacity to understand the result of hydro-climatic conditions on subglacial sediment dynamics \citep{creyts2013,beaud2018,delaney2019,hewitt2019}.
-Yet, sediment discharge capacity controls the mobilization and deposition of sediment, making it one of the most fundamental aspects of subglacial sediment evacuation, especially over short timescales.
+To date, explicit parameterizations of subglacial sediment transport capacity with respect to water discharge have assumed fixed channel size  \citep{alley1997}. and
+This result demonstrates a strongly non-linear response in subglacial sediment transport capacity to water discharge.
+Several modeling frameworks examining subglacial sediment transport with evolving channel size \citep{creyts2013,beaud2018,delaney2019,hewitt2019}.
+However, these works minimally discuss the impact of different hydrologies on sediment transport capacity underneath glaciers for interpreting sediment transport records.
+Understanding these processes is imperative for establishing the effect of hydro-climatic conditions on subglacial sediment dynamics.
+Particularly because sediment discharge capacity controls the mobilization and deposition of sediment, making it a fundamental aspect of subglacial sediment evacuation, especially over short timescales.
 
-Despite the physical differences between the subaerial and subglacial systems, water discharge and sediment export are often compared in contemporary glacial environments \citep[e.g.][]{willis1996,pearce2003,richards2003,swift2005,chu2009,tedstone2012,chu2012,overeem2017,delaney2018,swift2021,lu2022,andresen2024}.
-These comparisons are natural given that the majority of sediment evacuated from glaciers over contemporary seasonal timescales is done so fluvially \citep[e.g.][]{collins1989, riihimaki2005,delaney2024}.
-In subaerial systems, water discharge and sediment transport capacity for a grain size are closely linked \citep[e.g. ][]{schmidt2006,pitlick2021} \ian{check citiation}.
-However, the potential for a highly variable response in sediment transport capacity in subglacial sediment transport may affect the interpretation of sediment transport records in glacial systems \citep[e.g.][]{ganti2016,mancini2023}, if properties associated with subaerial systems are assumed.
-This relationship should be clarified as sediment dynamics change in glacierized regions along with changing hydrology, increased variability of water discharge \citep{lane2019} and extreme glacier melt  \citep[e.g.][]{cremona2023,overeem2015} that may drive high magnitude sediment transport events.
+Despite the physical differences between the subaerial and subglacial systems, water discharge and sediment export are often compared in contemporary glacierized catchments \citep[e.g.][]{willis1996,pearce2003,richards2003,swift2005,chu2009,tedstone2012,chu2012,overeem2017,delaney2018,swift2021,lu2022,andresen2024}.
+Many of theses studies discuss the disparity between water discharge and sediment export, highlighting the non-linear relationship between the two, along with the role of sediment access or production.
+Yet, a variable relationship between water discharge and sediment export could be expected from glaciers with evolving conduit size in a transport-limited, or hydraulically controlled, regime.
+The response of sediment transport capacity in subglacial sediment transport may affect the interpretation of sediment transport records in these glacierized catchments systems \citep[e.g.][]{ganti2016,mancini2023}.
+It is important that this relationship is clarified as sediment dynamics change in glacierized regions along with changing hydrology \citep[e.g.][]{brunner2019}, increasing variability of water discharge \citep{lane2019} and occurrence of extreme glacier melt  \citep[e.g.][]{overeem2015,cremona2023} that may drive high magnitude sediment transport events.
 
-This manuscript has two objectives: 1) evaluate the variability of subglacial sediment transport capacity with evolving subglacial channel size and different hydrographs and 2) establish if sub-seasonal variations in water discharge can co-vary with sediment transport capacity in subglacial systems.
-First, to reinforce the highly variable behavior of subglacial sediment transport capacity,  we present algebraic formulations of the sediment transport capacity response to water discharge in subaerial, pipe-flow, and equilibrium R-channel conditions, extending the relationships presented in \citet{alley1997}.
+This manuscript has two objectives: 1) to evaluate the variability of subglacial sediment transport capacity with evolving subglacial channel size and different hydrographs and 2) to establish whether sub-seasonal water discharge can co-vary with sediment transport capacity in subglacial systems.
+First, to demonstrate subglacial sediment transport capacity's highly variable nature,  we present algebraic formulations of the sediment transport capacity response to water discharge in subaerial, pipe-flow, and equilibrium R-channel conditions.
+These formulations extend the relationships presented in \citet{alley1997}.
 We then use numerical models to examine sediment transport capacity in subglacial channels that evolve in size and subaerial channels.
-They are then applied to proglacial hydrological records from an Alpine glacier in Switzerland (Fieschergletscher) and a land-terminating glacier in Greenland (Leverett Glacier).
-We run models with an ensemble of parameters to evaluate the sediment transport capacity variability in subglacial channels compared to subaerial ones.
-Lastly, model outputs demonstrate the particular processes that can drive variability in sediment discharge capacity in subglacial systems.
-Together, these findings indicate differences in the relationships between water discharge, channel geometry, and water velocity in the subglacial and subaerial channels.
-The manuscript then uses the identified differences between the subaerial and subglacial channel types to discuss the implications for interpreting the processes driving variability in sediment transport records and the impact of these differences on sediment export from glaciers.
+They are applied to proglacial hydrological records from an Alpine glacier in Switzerland (Fieschergletscher) and a land-terminating glacier in Greenland (Leverett Glacier).
+We run a model ensemble to evaluate the sediment transport capacity variability in subglacial channels compared to subaerial channels.
+Model outputs demonstrate the specific processes that can drive variability in sediment discharge capacity in subglacial systems.
+These findings indicate differences in the relationships amongst water discharge, channel geometry, and water velocity in subglacial and subaerial channels.
+The manuscript then discusses the implications of variability in sediment transport capacity from glaciers and the interpretation of sediment transport records.
 
@@
 
-Water discharge data is used from Alpine and ice sheet settings collected downstream of the glacier, assuming no water storage in the proglacial area.
+Water discharge data is  from Alpine and ice sheet settings collected downstream of the glacier, and assumes no water storage in the proglacial area.
 The Alpine site (labeled \alpine{}) is  Fieschergletscher in the Swiss Alps ($46^\circ\,29'\,07''$ N, $8^\circ\,08'\,3''$ E).
@@
 
-Note that we omit analysis of suspended sediment discharge records due to the role of sediment supply in capturing variations in sediment discharge, in addition to sediment transport capacity \citep[e.g.][]{delaney2019}.
+Note that we omit analysis of suspended sediment discharge records due to their dependence on sediment supply in capturing variations in sediment discharge, in addition to sediment transport capacity \citep[e.g.][]{delaney2019}.
 
@@
   \centering
-  \includegraphics[width=0.8\linewidth]{Qw_plot.pdf}
+  \includegraphics[width=0.8\linewidth]{Fig2.pdf}
   \caption{Water discharge from Fieschergletcher (\alpine) and Leverett Glacier (\icesheet).
@@
 Both models use the measured discharge to calculate water velocity, shear stress, and width-integrated shear stress, upon which both suspended sediment and bedload transport depend \citep[Figure \ref{fig:cartoon}; ][]{shields1936}.
-Our evaluation in terms of shear stress omits the selection of a sediment transport relationship and a grain-size parameter \citep[e.g.][]{shields1936,meyer1948}.
+Our choice to evaluate shear stress omits selecting a sediment transport relationship and a grain-size parameter \citep[e.g.][]{shields1936,meyer1948}.
 
@@
 To evaluate the shear stress of water flowing across sediments underneath a glacier, the subglacial channel model accounts for the channel geometry and the water's velocity.
-To do this, we use a lumped hydraulics model from \citet{werder2010b}, itself based upon  \citet{clarke1996}.
+We use a lumped hydraulics model from \citet{werder2010b}, itself based upon  \citet{clarke1996}.
 
 Here, it is assumed that the water is transported through a subglacial channel \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972} beneath a glacier with channel length $l$, with a flatbed and a mean thickness of $h_{ice}$.
-The channel size grows from melt due to frictional heating from water flow and closes due to the creep of the ice.
+The channel size grows from melt due to frictional heating from water flow and closes due to ice creep.
 The formulation here does not consider the englacial storage of water.
@@
 \end{linenomath*}
-\noindent where $t$ is time, $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (complete values given in Table~\ref{table:vpm}), $g$ is the acceleration due to gravity, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change over $l$, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the mean ice overburden pressure expressed in meter water equivalent ($\rho_w$ is density of water; $\rho_i$ is density of ice), and $n$ is Glen's n \citep{glen1955} (usually $n=3$).
-The first term on the equation's right side represents the opening of the channel through frictional heating, while the following term represents the creep closure of the channel from ice deformation.
+\noindent where $t$ is time, $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (values in Table~\ref{table:vpm}), $g$ is the acceleration due to gravity, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change over $l$, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the mean ice overburden pressure expressed in meter water equivalent ($\rho_w$ is density of water; $\rho_i$ is density of ice), and $n$ is Glen's n \citep{glen1955} (usually $n=3$).
+The first term on the equation's right side represents the channel opening by frictional heating, while the following term represents channel closure from ice deformation.
 
@@
 The hydraulics parameterization presented in \citet{tucker1997} is implemented to represent the subaerial channel.
-Using mass conservation and the Darcy-Weisbach relationship as well as assuming that the channel is sufficiently wide compared to its depth so that the hydraulic radius is well approximated by the flow depth,
-the shear stress $\tau_f$ at the river bed is
+This model uses mass conservation and the Darcy-Weisbach relationship and assumes that the channel is sufficiently wide compared to its depth.
+This latter assumption means that the hydraulic radius is well approximated by the flow depth.
+Therefore, the resulting shear stress $\tau_f$ at the river bed is
 \begin{linenomath*}
@@
     Where two values are given, the first refers to  \alpine{}, a scenario from Fieschergletscher, and the second to a glacier marginal to the \icesheet{} a scenario from Leverett Glacier.
-If a second line is given, it refers to the range of values examined in the parameter search}
+    A second line is given refers to the range of values examined in the parameter search.}
   \small 
@@
 The models above are applied to proglacial discharge records from the Fieschergletscher (scenario \alpine{}) and the Leverett Glacier (scenario \icesheet{}).
-Outputs of the models represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
-To generalize these scenarios, \alpine{}  is exemplified by Fieschergletscher's relatively thin ice thickness \citep[$h_{ice}$= $225$\,\unit{m}][]{grab2021}, low water discharge ($\sim\,10$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:Qw}\, b).
-\icesheet{}  is exemplified by Leverett glacier's  thick ice  \citep[$h_{ice}$= $700$\,\unit{m}; ][]{morlighem2017}, high water discharge ($\sim\,300$\,\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:Qw}\, b).
+The model outputs represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
+To generalize these scenarios, \alpine{}  is exemplified by relatively thin ice thickness \citep[$h_{ice}$= $225$\,\unit{m}][]{grab2021}, low water discharge ($\sim\,10$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge at Fieschergletscher (Figure~\ref{fig:Qw}\, b).
+\icesheet{}  is exemplified by   thick ice  \citep[$h_{ice}$= $700$\,\unit{m}; ][]{morlighem2017}, high water discharge ($\sim\,300$\,\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge at Leverett glacier (Figure~\ref{fig:Qw}\, b).
 
-The first experiment aims to characterize the variability in sediment discharge capacity in subglacial compared to subaerial channels across a range of channel slopes and shapes, along with friction factors.
-Additionally, we examine the effects of different water discharge variability by smoothing the water discharge over different periods.
+The first experiment aims to characterize the variability in sediment discharge capacity in subglacial relative to subaerial channels across a range of channel slopes and shapes, along with friction factors.
+Additionally, we examine differing water discharge variability by smoothing the water discharge over different period from $15$\,\unit{min} up to $15$\,\unit{days}.
 The results below present water velocity ($v_g$, $v_f$), shear stress ($\tau_g$, $\tau_f$), and width-integrated shear stress ($w_g\tau_g$, $w_f\tau_f$) from the subglacial and subaerial models.
 For brevity, these values together are referred to as ``model outputs'' in the text.
-The subglacial model outputs are generally larger than their subaerial equivalents due to their large hydraulic gradients, thus greater sediment transport capacity would be expected in subglacial channels compared to subaerial ones \citep{alley1997}.
-To evaluate variability, we subtract the model outputs from their values smoothed using daily averages,  detrending the seasonal signal (Figure~\ref{fig:Qw}).
+To evaluate variability, we subtract the model outputs from their values smoothed using daily averages,  detrending the seasonal signal.
 We then calculate the variability in the model output by taking the standard deviation of this detrended time series.
 
-The effect of channel shape, gradient, and roughness is established by running the model with random parameter values of channel slope and geometry factors ($\beta$,$\nabla z_c$, and $\alpha$) and friction factors ($f_r$ and $f_f$; see Table~\ref{table:vpm} for value range).
-Runs with parameter combinations are accepted if their mean subglacial water velocity over the season lies between $0.5$\,\unit{m}\,\unit{s}$^{-1}$ and $2$\,\unit{m}\,\unit{s}$^{-1}$ or if subaerial water velocity lies between $0.3$\,\unit{m}\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}\,\unit{s}$^{-1}$, following measurements \citep[e.g.][]{werder2010b,magnusson2012,chandler2013}.
-Additionally, to be accepted, subglacial model runs can only experience a flotation fraction ($\frac{\Delta h}{2\,h_o}$) of above $1.2$ for less than $2.5$\,\% of the run.
+The effects of channel shape, gradient, and roughness is established by running the model with random parameter values of channel slope and geometry factors ($\beta$,$\nabla z_c$, and $\alpha$) and friction factors ($f_r$ and $f_f$; see Table~\ref{table:vpm} for value range).
+Runs with parameter combinations are accepted if their mean subglacial water velocity over the season lies between $0.5$\,\unit{m}\,\unit{s}$^{-1}$ and $2$\,\unit{m}\,\unit{s}$^{-1}$ or if subaerial water velocity lies between $0.3$\,\unit{m}\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}\,\unit{s}$^{-1}$ \citep[e.g.][]{werder2010b,magnusson2012,chandler2013}.
+To be accepted, subglacial model runs must experience a flotation fraction ($\frac{\Delta h}{2\,h_o}$) of above $1.2$ for less than $2.5$\,\% of the run.
 The initial condition of cross-sectional area $S_g$ is evaluated by applying a model spinup with repeatedly applying the maximum observed water discharge of the first $4$ days of the study until there is no change in the channel area.
@@
 The routine runs until $100$ different parameter combinations for each water discharge smoothing period are accepted with the conditions described above.
-Here, the water discharge input at a given time is the average over the smoothing period before that time, ranging from $15$\,\unit{min} up to $15$\,\unit{days}.
+To test the effects of variability in water discharge, it is averaged over the smoothing period ranging
 
@@
 % 
-To quantify the link between sediment transport capacity and water discharge, model outputs are compared to water discharge from the glaciers using Spearman rank correlation, which accounts for the ordering of values, but not their magnitude.
-Rank correlation reduces the impact of the non-linear response of sediment transport capacity to the hydrological forcing.
+To test covariance between sediment transport capacity and water discharge, model outputs are compared to water discharge from the glaciers using Spearman rank correlation.
+This metric accounts for the ordering of values, but not their magnitude.
+Rank correlation reduces the impact of the non-linear relationship between sediment transport capacity and hydrology.
 
@@
 % 
-For a steady state R-channel, it is assumed that  $\Psi$ is constant (approximated by the gradient of the Shreve \citeyear{shreve1972} potential and that $S$ adjusts such that it is in steady state with $\Psi$ and $Q$.
+For a steady state R-channel, it is assumed that  $\Psi$ is constant (approximated by the gradient of the Shreve \citeyear{shreve1972} potential and that $S$ adjusts in steady state with $\Psi$ and $Q$.
 Note that the R-channel model presented above in Section~\ref{sect:sub_mode} Equation~\ref{eq:dS_dt} calculates $\Psi$ from the time-evolving $S_g$ via the Darcy-Weisbach Equation~\ref{eq:dh} and thus no Shreve approximation is then needed.
@@
 Now width integrated shear stress and sediment transport can be calculated for all regimes and all transport formulas.
-The results of the 12 combinations are presented in Table~\ref{tab:Qs} and also in Table~\ref{tab:Qs} where the fractions in the exponents are approximately given by decimal numbers for ease of comparison.
+The results of the 12 combinations are presented in Table~\ref{tab:Qs} and also in Table~S1 where the fractions in the exponents are approximately given by decimal numbers for ease of comparison.
 
@@
 The exponent for the width scaling $\alpha$ only impacts the relationship between sediment transport and water discharge in the EH relation in any meaningful way.
-However, for the value of $\alpha$ around $\frac{1}{2}$ (a value appropriate for most streams), the exponent on $Q$ for EH is only slightly above what the other transport relations give.
+However, for the value of $\alpha$ around $\frac{1}{2}$ (a value appropriate for most streams), the exponent on $Q$ for EH is only slightly greater than th other transport relations.
 
 The sediment discharge in the steady state R-channel scales very similarly to the subaerial channel case for all relations and virtually identically for the $\alpha=0.5$ case.
-However, note that the head gradients $\Psi$ are likely higher for comparable $Q$ in an ice-sheet marginal or alpine glacier setting than in a subaerial channel, as is well described in \citet{alley1997}, and thus transport rates may still be much higher in a steady state R-channel.
+However, note that the head gradients $\Psi$ are likely higher for comparable $Q$ in an ice-sheet marginal or alpine glacier setting than in a subaerial channel \citep{alley1997}.
+Thus, sediment transport capacity is likely  much higher in a steady state R-channel.
 
-Furthermore, R-channels will rarely operate in a steady state as variations in discharge, in particular on the diurnal timescale or the timescale of severe rain or melting events, are too fast for such a channel to reach a steady state.
-In such cases they operate more like a pipe of fixed cross-section \citep[e.g.][]{gimbert2016}.
-Table~\ref{tab:Qs} shows that in such a situation, the sediment transport scales much more severely with discharge, with the exponent on $Q$ being between $3$ and $5$, compared to the other two regimes when that exponent is at most $1.7$ (Table~\ref{tab:Qs}).
-Thus fluctuations of discharge on short timescales (on the order of a day) have the potential to cause conditions with very high sediment transport capacities, that are of far higher magnitude than the subaerial channels.
+Furthermore, R-channels will rarely operate in steady state as variations in water discharge.
+The timescale of  diurnal melt or severe rain or melt  are too short  for the channel to reach a steady state.
+In these cases with high water discharge variability, channels operate more like a pipe of fixed cross-section \citep[e.g.][]{gimbert2016}.
+Table~\ref{tab:Qs} shows that  sediment transport in pipe-flow scales much more severely with discharge; the exponent on $Q$ being between $3$ and $5$, compared to the other two regimes when that exponent is at most $1.7$ (Table~\ref{tab:Qs}).
+Thus fluctuations of discharge on short timescales (on the order of a day) have the potential to cause conditions with very high sediment transport capacities.
+These sediment transport capacities variations are of far higher magnitude than those of subaerial channels \citep{alley1997}.
 
@@
 The first numerical experiment aims to compare the variability between the subglacial and subaerial model outputs to a range of channel shapes, friction factors, and water discharge variability, as the subglacial channel's cross-sectional area is allowed to evolve.
-Following the findings in Section~\ref{sect:scaling}, variability in subglacial sediment discharge capacity could approach that of subaerial systems, should water discharge variations be sufficiently small or subglacial channels respond in size quickly.
-For simplicity, we only examine $w\tau$, not the sediment transport formulas mentioned above.
-%We have chosen to exclude the pipe flow case (Section~\ref{sect:scaling}) from analysis as it depends greatly on the choice of the subglacial cross sectional area $S_g$.
+Following the findings in Section~\ref{sect:scaling}, variability in subglacial sediment discharge capacity could approach that of subaerial systems, should water discharge variations be sufficiently small.
+For simplicity, we only examine $w\tau$.
 
 Across the range of parameters examined (Section\,\ref{sect:imp}), variability in all model outputs remains higher in the subglacial system compared to the subaerial one for both \alpine{} and \icesheet{} (Figure~\ref{fig:multi_run}).
-In some cases, the variability in subglacial model outputs is a magnitude larger than their subaerial counterparts.
-Results from both \alpine{} and \icesheet{} scenarios show that variability in both subglacial and subaerial outputs decreases with smoothing time longer than approximately 1--5 days (Figure~\ref{fig:multi_run}), when the diurnal variations in water discharge are removed from smoothing.
-In the \icesheet case, the smaller variability in model outputs compared to \alpine may mean that the R-channel is in closer equilibrium between channel size and water discharge (Section~\ref{sect:scaling}).
-In this case, water discharge variations would create smaller disparities in sediment transport capacity than even in subaerial cases as the subglacial channel size evolves to accommodate them (Table~\ref{tab:Qs}).
+In some cases, subglacial model outputs' variability is a magnitude larger than their subaerial counterparts.
+Results show that variability in both subglacial and subaerial outputs decreases with smoothing time longer than approximately $1$--$5$ days (Figure~\ref{fig:multi_run}).
+These smoothing timescales remove the diurnal variations in water discharge.
+In the \icesheet case, the smaller variability in model outputs compared to \alpine means that the R-channel is in closer equilibrium between channel size and water discharge (Section~\ref{sect:scaling}).
+In this case, water discharge variations would result in  smaller  sediment transport capacity variations than even in subaerial cases as the subglacial channel size evolves to accommodate them (Table~\ref{tab:Qs}).
 
@@
   \centering
-  \includegraphics[width=0.9\linewidth]{multi_run_tmp.pdf}
+  \includegraphics[width=0.9\linewidth]{Fig3.pdf}
   \caption{Standard deviation of detrended model outputs for different smoothing times with different subglacial and subaerial channel shapes and friction factors.
@@
     Solid lines denote the mean value of standard deviations.
-    Markers show smoothing periods ($15$\,\unit{min}, $1$\,\unit{hr}, $12$\,\unit{hr}, $1$\,\unit{d}, $5$\,\unit{d}, $10$\,\unit{d}, and $15$\,\unit{d}).
+    Markers show smoothing periods ($15$\,\unit{min}, $1$\,\unit{hr}, $6$\,\unit{hr} $12$\,\unit{hr}, $1$\,\unit{d}, $5$\,\unit{d}, $10$\,\unit{d}, and $15$\,\unit{d}).
   }
@@
 
-Across the range of parameters examined in both \alpine and \icesheet, variability in a subaerial model output did not exceed the corresponding subglacial one (Figures~\ref{fig:multi_run} and \ref{fig:params}).
-Analysis of parameter values shows that in both cases, smaller values of $\alpha$, or closer to a slot-canyon, result in slightly greater variability in velocity and shear stress (Figure~\ref{fig:params}).
-Yet, larger values of $\alpha$ can result in greater variability in width-integrated shear stress, as greater channel width changes increase the width-integrated shear stress, but not the shear stress.
-Additionally, steeper subaerial channel slopes result in greater variability in both shear stress and width-integrated shear stress but only approach the variability of the subglacial channel in the \alpine case (Figure~\ref{fig:params}).
-%At steeper slopes, the alluvial processes represented in the model (Section ~\ref{sect:fluv}) probably do not apply, as the.
-We note that these parameters span a commonly accepted range of values, therefore do not anticipate a scenario where variability in the subaerial system would exceed the subglacial system with this hydrology.
+Smaller values of $\alpha$, or a channel cross section closer to a slot-canyon, result in greater variability in velocity and shear stress (Figure~\ref{fig:params}).
+In both cases, the variability of small values of $\alpha$ does not approach that of the subglacial channels.
+Yet, larger values of $\alpha$ can result in greater variability in width-integrated shear stress, as greater channel width changes increase the width-integrated shear stress.
+Additionally, steeper subaerial channel slopes result in greater variability in both shear stress and width-integrated shear stress but only approach the  subglacial channel's variability in the \alpine case (Figure~\ref{fig:params}).
+We note that these parameter values span a commonly accepted range.
+Therefore , we do not anticipate a scenario where variability in the subaerial system would exceed the subglacial system with this two hydrologic forcings.
 
-Greater values of the subglacial and subaerial friction factor $f_i$ and $f_p$ result in greater variability in the shear stress and width-integrate shear stress in both cases (Figure~\ref{fig:params}), as expected given Equations~\ref{eq:tau_g} and \ref{eq:vf}.
-However, more velocity variability is present in low values of $f_i$ and $f_p$. 
-Low values of $f_i$ result in lower growth rates in subglacial channels (Equations~\ref{eq:dS_dt} and \ref{eq:dh}).
+Greater  subglacial and subaerial friction factors $f_i$ and $f_p$ result in greater variability in the shear stress and width-integrate shear stress in both cases (Figure~\ref{fig:params}).
+This result is expected given Equations~\ref{eq:tau_g} and \ref{eq:vf}.
+However, more velocity variability is present in low  $f_i$ and $f_p$ values. 
+Low values of $f_i$ result in slower growth rates in subglacial channels (Equations~\ref{eq:dS_dt} and \ref{eq:dh}).
 As a result, water velocity, as opposed to channel growth, accommodates increases in water discharge, and velocity variability increases.
-Note as well that smaller values of channel factor $\beta$, creating low and broad channels, result in more variability in width-integrated shear stress.
+Smaller values of channel factor $\beta$, creating low and broad channels, result in more variability in width-integrated shear stress.
 Here, the channel width can grow more quickly in response to water discharge increases as compared to a semi-circular channel with $\beta = \pi$ (Equation~ \ref{eq:dh}).
@@
   \centering
-  \includegraphics[width=0.9\linewidth]{Parameter.pdf}
+  \includegraphics[width=0.9\linewidth]{Fig4.pdf}
   \caption{Parameter values compared to variability in model outputs.
@@
 
-In the second numerical experiment defined in Section~\ref{sect:imp}, we aim to quantify the sources of increased variability in the subglacial model outputs as they exhibit different seasonal evolutions and peaks.
-Examining the relationship between water discharge and the model outputs, different relationships emerge for the subaerial and subglacial cases due to the hysteresis in channel size in the subglacial model (Equation~\ref{eq:dS_dt}).
-Because there is no history dependence in subaerial channels, each water discharge value in the subaerial channel results in a unique water velocity, shear stress, and width-integrated shear stress (Section~\ref{sect:sub_mode}).
-This characteristic results in a perfect rank correlation between the variables (Figure~\ref{fig:Qw_vari}).
+The second numerical experiment aims to quantify the sources of increased variability in the subglacial model outputs as they exhibit different seasonal evolutions and peaks (Figure~\ref{fig:model_outs}).
+Variable relationships between model outputs and water discharge emerge for the subaerial and subglacial cases due to the hysteresis in channel size in the subglacial model (Equation~\ref{eq:dS_dt}).
+Because there subaerial channels have no history dependence, each water discharge value in the subaerial channel produces a unique water velocity, shear stress, and width-integrated shear stress (Section~\ref{sect:sub_mode}).
+This characteristic results in a perfect rank correlation between the variables and water discharge (Figure~\ref{fig:Qw_vari}).
 \begin{figure}[hbt!]
   \centering
-  \includegraphics[width=0.9\linewidth]{Fig2.png}
+  \includegraphics[width=0.9\linewidth]{Fig5.png}
   \caption{Model outputs from simulations using the hydrographs in panels a and e for the scenarios \alpine{} (a-d) and \icesheet{} (e-h).
@@
 
-Conversely, the history dependence on channel size in subglacial channels means that different sediment transport characteristics, such as velocity, occur across a large range of water discharges.
+Conversely, the history dependence on channel size in subglacial channels means that model outputs occur across a large range of water discharges.
 For instance, in subglacial channels in \alpine{},  high water velocity values and shear stresses can occur from a low water discharge  ($\sim\,4$\,\unit{m}$^3$\,\unit{s}$^{-1}$) to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ (Figure~\ref{fig:Qw_vari}\,a).
 In \icesheet{}, water velocities close to the seasonal mean value can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
-When considering the subglacial channel's evolving width, width-integrated shear stress across the channel generally increases with water discharge, with improved rank correlation compared to water velocity or shear stress (\alpine{}, Figure~\ref{fig:Qw_vari} a--c).
-Yet even the width-integrated shear stress can vary substantially, with the highest values occurring at water discharge values ranging from roughly $11$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ to over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
+The subglacial channel's evolving width can counteract some of these effects.
+Width-integrated shear stress generally increases with water discharge, with greater rank correlation compared to water velocity or shear stress (\alpine{}, Figure~\ref{fig:Qw_vari} a--c).
+Yet, even the width-integrated shear stress can vary substantially relative to water discharge.
+The highest values of width-integrated shear stress occur at water discharge values ranging from roughly $11$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ to over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
 The variability in width-integrated shear stress is less pronounced in the \icesheet{} scenario, where the hydrograph has less diurnal variability (Figure~\ref{fig:Qw_vari} c, f).
+This results from the discharge variations being in closer equilibrium with subglacial channels compared to \alpine{} (Sections~\ref{sect:scaling} and~\ref{sect:ensemble}).
 
@@
   \centering
-  \includegraphics[width=0.9\linewidth]{Fig3.png}
+  \includegraphics[width=0.9\linewidth]{Fig6.png}
   \caption{
@@
     $R_g$ ($R_f$) shows the Spearman rank correlations for the subglacial (subaerial) outputs.
-    Plots are shown with discharge and outputs at $15$\unit{min} intervals.
+    Plots are shown with discharge and model outputs at $15$\unit{min} intervals.
   }
@@
 
-The different response of subaerial and subglacial model outputs to water discharge impacts the timing of peaks in sediment transport capacity, and thus its relationship with water discharge.
-In the subglacial channel, peaks in $w_g\tau_g$ generally occur when water discharge increases at the fastest rate, but before the maximum water discharge (Figure~\ref{fig:model_outs}\, d and h).
-By contrast, in the subaerial channel,   $w_f\tau_f$  peaks when water discharge is highest.
+Peaks in subaerial model outputs occur coincident with peaks in water discharge (Figure~\ref{fig:model_outs}).
+In the subglacial channel, peaks in model outputs generally occur when water discharge increases at the fastest rate, but before the maximum water discharge.
+Reduced water velocity takes place when water discharge stabilizes at its peak and channel growth continues.
+As a result, subglacial sediment transport capacity is greatest on the hydrograph's rising limb, relative to the falling limb.
 
@@
 
-Several aspects of the models and experiment design limit the comparison of the model outputs to field data and make comparison of the two systems difficult.
-The lumped nature of the models means that they operate independently of the upstream drainage network, and sediment access is not considered here.
+Several aspects of the models and experiment design limit make comparing the model outputs to field observations difficult and make comparison of the two systems difficult.
+The lumped nature of the models means that they operate independently of the upstream drainage network.
+Additionally, sediment access is not considered here.
 In reality, processes such as sediment access are highly important in controlling sediment export in glacierized catchments \citep[e.g.][]{herman2015,vergara2022}.
-Furthermore, under different glacier configurations and subglacial conditions, new channels can emerge quickly and water can be distributed to flow through several adjacent conduits impacting the sediment transport in each \citep[e.g.][]{werder2013,hewitt2019,delaney2023}.
-As a result, these models isolate the relationship between water discharge and sediment transport capacity in subglacial and subaerial systems, making comparison more direct, but neglecting more complex, yet important spatially distributed processes.
+Furthermore, meltwater can be distributed to flow through several adjacent conduits impacting the sediment transport capacity in each \citep[e.g.][]{werder2013,hewitt2019,delaney2023}.
+The lumped models used here making comparison more direct by isolating the relationship between water discharge and sediment transport capacity in subglacial and subaerial systems. 
+However, they neglect more complex, yet important, spatially distributed processes.
 
-Comparison of model outputs between the subaerial and subglacial case is difficult, given that while the water velocities are comparable in both channel types, the shear stresses and width-integrated shear stresses are much smaller in subaerial channels (Figure~\ref{fig:model_outs}).
-This occurs due to the different gradients in the two systems, with the subglacial one being much steeper \citep{alley1997}.
-We note, however, that in the ensemble runs in Section~\ref{sect:ensemble}, the range of parameters tested covers the range of viable shear stresses in both subglacial and subaerial channels.
+Comparing model outputs between the subaerial and subglacial case is difficult due to the different gradients in the two systems, with the subglacial one being much steeper \citep{alley1997}.
+The water velocities are similar in both channel types (Figure~\ref{fig:model_outs}).
+However, the shear stresses and width-integrated shear stresses are much smaller in subaerial channels.
+We note that in the ensemble runs in Section~\ref{sect:ensemble}, the parameters' range tested covers a likely range of viable shear stresses in both subglacial and subaerial channels.
+The greater sediment transport capacity in the subglacial channels implies that sediment grain-size underneath subglacial channels is likely larger that subaerial counterparts.
 
-Width-integrated shear stress is examined in the two numerical experiments as it does not have a dependence on grain size, unlike sediment transport relationships, such as \citet[][Section \ref{sect:scaling}]{meyer1948}.
+Width-integrated shear stress is examined in the two numerical experiments as it does not have a dependence on grain size, unlike sediment transport relationships \citep[][Section \ref{sect:scaling}]{meyer1948}.
 This makes comparison between the two systems simpler. 
 Preferential transport of smaller sediment clasts and the input of upstream sediment impact sediment transport capacity as grain size evolves \citep[e.g.][]{gomez1983}.
-These processes are only beginning to be evaluated below glaciers \citep{aitken2024}, but can be important in subaerial systems.
+These processes are only beginning to be evaluated underneath glaciers \citep{aitken2024}, but can be important in subaerial systems.
 Subglacial sorting processes could be an additional source of variability in sediment export from glaciers with respect to water discharge, especially for bedload transport.
@@
 
-Increased variability in subglacial sediment discharge capacity could make it more difficult to differentiate the role of long-term climatic controls on sediment transport with shorter-term hydrological variations.
-In sediment transport relationships that evaluate sediment discharge capacity, such as in \citet{meyer1948} or \citet{engelund1967}, shear stress is scaled to the power of $\frac{3}{2}$ or $\frac{5}{2}$, respectively (Section~\ref{sect:scaling}).
+The greater variations in shear stress in subglacial channels presented here  may actually cause even greater variations in sediment transport capacity in subglacial channels than presented above  (Figure~\ref{fig:multi_run}).
+In sediment transport capacity relationships, such as in \citet{meyer1948} or \citet{engelund1967}, shear stress is scaled to the power of $\frac{3}{2}$ or $\frac{5}{2}$, respectively (Section~\ref{sect:scaling}).
 The exponent greater than $1$ magnifies sediment discharge variability beyond the variable sediment transport parameters described above (Figure~\ref{fig:multi_run}; Table~\ref{tab:Qs}).
-As a result, the different response to water discharge variations in shear stress in subglacial channels compared to subaerial ones (Figure~\ref{fig:Qw_vari}) may actually cause even greater variations in sediment transport capacity in subglacial channels than presented above.
 
-Greater variations in subglacial sediment transport capacity compared to subaerial ones could cause a supply-limited regime at many glaciers \citep{alley1997}.
-In subglacial systems, sediment's critical shear stress, or threshold at which sediment mobilization occurs, can be reached more frequently and across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
-This result suggests sediment export here is especially sensitive to observed changes in water discharge variability, in addition to quantity \citep{lane2019b}.
-Sediment exhaustion through repeated crossing of the mobilization threshold may explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \citep{overeem2017}.
-The repeated crossing of the mobilization threshold across a range of discharges could also result in sediment discharge's strong dependence on sediment production from sliding in many cases \citep{herman2015,koppes2015}.
+Greater variations in subglacial sediment transport capacity could cause a supply-limited regime at many glaciers \citep{alley1997}.
+In subglacial systems, sediment's critical shear stress, or threshold at which sediment mobilization occurs, can be reached more frequently and across many water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
+This result suggests sediment export here is especially sensitive to observed changes in water discharge variability \citep{lane2019b}.
+Sediment exhaustion through repeatedly crossing the mobilization threshold may explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \citep{overeem2017}.
+The same process could also result in sediment discharge's strong dependence on sediment production from sliding in many cases \citep{herman2015,koppes2015}.
 
-While transport-limited states below glaciers are likely rare \citep[e.g.][]{alley1997}, they can be possible \citep[e.g.][]{swift2005} and abundant sediment can persist below glaciers \citep[e.g.][]{walter2014,stevens2022,delaney2022}.
-At these glaciers, the great variability in subglacial sediment transport capacity may make it difficult to link sediment discharge to hydrology, especially when peak events occur \citep{cowan1988,delaney2018}.
-As a result, it may prove challenging to establish the impact on sediment transport capacity of individual extreme events, such as extreme precipitation or melt \citep{lu2022}, given the potential for different channel sizes with the same hydrology forcing.
-However, in the case of bedload, the signal of this variable subglacial sediment transport relationship may be destroyed only a short distance downstream as subaerial hydraulics controls these sediment transport \citep{mancini2023}.
+While transport-limited states at glaciers are likely rare \citep[e.g.][]{alley1997}, abundant sediment can persist underneath glaciers \citep[e.g.][]{walter2014,stevens2022,delaney2022}.
+As a result, in some cases sediment export can be hydrologically controlled and efficient \citep[e.g.][]{swift2005,andresen2024}.
+At these glaciers, the great variability in subglacial sediment transport capacity may make it difficult to link sediment discharge to hydrology, especially when peak events occur \citep{cowan1988,delaney2018,lu2022}.
+As a result, the sediment transport capacity response to individual extreme precipitation or melt events could be difficult to establish.
+The results here suggest that different channel sizes with the same hydrology forcing can result in very different sediment transport capacities.
 
-Catchments with reduced water discharge variability may experience yet less variability in sediment transport capacity, as the two hydrological forcings from \alpine{} and \icesheet{} show different variations in model outputs.
-Indeed, the more decoupled and sporadic relationship between model outputs and water discharge in \alpine{} results from the relatively larger discharge variations on daily to weekly timescales (Figure~\ref{fig:Qw_vari}).
-This occurs as the evolution of the R-channel's size is slow compared to the variations in water discharge.
-Such water discharge variability in \alpine{} may cause shear stress to approach $Q^{2}$in assuming pipe-flow conditions \citep[Section \ref{sect:scaling}; c.f.][]{alley1997}.
-Conversely, the reduced variability in water discharge in \icesheet{} may result in the R-channel's size being closer to equilibrium from the smaller variations in water discharge.
-In this case, the exponent on water discharge for shear stress is likely less than $2$, but greater than $\frac{4}{5}$ that occurs when the R-channel's size evolves in equilibrium with water discharge (see Section \ref{sect:scaling}).
-As a result, there is a stronger relationship between subglacial model outputs and water discharge in \icesheet{} compared to \alpine{} (Figure~\ref{fig:Qw_vari}) and less variability model outputs in the \icesheet case (Figure~\ref{fig:multi_run}).
+Catchments with reduced water discharge variability may experience yet less variability in sediment transport capacity, shown by the \alpine{} and \icesheet{} model outputs.
+Indeed, the more decoupled and sporadic relationship between model outputs and water discharge in \alpine{} results from the relatively larger discharge variations on sub-daily to weekly timescales (Figure~\ref{fig:Qw_vari}).
+This occurs as the subglacial channel's size evolves slowly compared to the variations in water discharge.
+High water discharge variability in \alpine{} may cause shear stress to approach $Q^{2}$in assuming pipe-flow conditions where the water discharge varies largely compared to channel size \citep[Section \ref{sect:scaling}; c.f.][]{alley1997}.
+Conversely, the reduced variability in water discharge in \icesheet{} may result in the subglacial channel's size being closer to equilibrium from the smaller variations in water discharge.
+In this case, the exponent on water discharge for shear stress is likely substantially less than $2$, but greater than $\frac{4}{5}$ that occurs in a steady-state R-channel (see Section \ref{sect:scaling}).
+As a result, there is a stronger relationship between subglacial model outputs and water discharge in \icesheet{} (Figure~\ref{fig:Qw_vari}) and less variability model outputs in the \icesheet case (Figure~\ref{fig:multi_run}).
 
@@
 
-A sporadic relationship between subglacial sediment transport capacities and water discharge can occur depending on the subglacial channel size (Figure~\ref{fig:Qw_vari}), limiting the use of water discharge as an indicator of sediment discharge capacity in these systems.
+A sporadic relationship between subglacial sediment transport capacities and water discharge occurs due to hysterisis in subglacial channel size (Figure~\ref{fig:Qw_vari}).
+This characteristic limits the use of water discharge as an indicator of sediment discharge capacity in these systems.
 As a result, characteristics such as bankfull or effective water discharge that link geomorphic work to hydro-climatic conditions could have limited meaning in evaluating subglacial sediment transport \citep{wolman1960,lenzi2006}.
-A glacier's sediment transport capacity is impacted by the ice thickness controlling the channel closure rate, and the glacier's surface slope, in addition to water discharge and sediment size \citep[Figure~\ref{fig:multi_run}, Section~\ref{sect:sub_mode}; ][]{rothlisberger1972,gimbert2016,stevens2022,walder1994}.
-In a supply-limited state, a glacier's sediment discharge record also responds to sediment availability and bedrock erosion from water pressure variations and sliding  \citep[e.g.
-][]{iverson2012,herman2015,delaney2023}.
-As a result, the greater amount of in variability the sediment transport evacuation can make it difficult to isolate the drives of sediment export with respect to hydrological and glaciological conditions \citep{delaney2024}.
-This multitude of processes lies in contrast to many subaerial channels, where transport capacity typically responds to water discharge in addition to sediment size, channel shape,  and hydraulic gradient, that can remain stable over the years or longer \citep[Section~\ref{sect:fluv}; e.g.][]{tucker1997}.
-Strong correlations between water discharge and sediment export in glacier systems could be indicative of other processes, such as increased sediment access \citep{zhang2022}.
-Furthermore, the observed correlation between water discharge and sediment discharge far downstream of the glacier likely represents subaerial transport processes, especially with respect to bedload \citep{mancini2023}.
+A glacier's sediment transport capacity is impacted by the ice thickness controlling the channel closure rate and the glacier's surface slope, in addition to water discharge and sediment size \citep[Figure~\ref{fig:multi_run}, Section~\ref{sect:sub_mode}; ][]{rothlisberger1972,gimbert2016,stevens2022,walder1994}.
+This multitude of processes lies in contrast to many subaerial channels, where transport capacity typically responds to water discharge, sediment size, channel shape,  and hydraulic gradient.
+The latter of these parameters can remain relatively stable over the years or longer \citep[Section~\ref{sect:fluv}; e.g.][]{tucker1997}.
+Strong correlations between water discharge and sediment export in glacier systems could indicate  other processes, such as increased sediment access \citep{zhang2022}.
+Furthermore, the observed correlation between water discharge and sediment discharge far downstream likely represents subaerial transport processes, especially with respect to bedload \citep{mancini2023}.
 
-The co-varying relationship between sediment discharge capacity and water discharge in subaerial conditions could be noticeable $\sim20$\,\unit{km} downstream of the Leverett site, where a strong correlation persists between sediment plume size and the river's water discharge into the Kangerlussuaq fjord \citep{chu2009,mcgrath2010}.
+The co-varying relationship between sediment discharge capacity and water discharge in subaerial conditions could be noticeable $\sim20$\,\unit{km} downstream of the Leverett site.
+A strong correlation persists between sediment plume size and the Watson river's water discharge into the Kangerlussuaq fjord \citep{chu2009,mcgrath2010}.
 In contrast, in marine-terminating glacier catchments, a less consistent relationship may occur between water discharge or melt extent and sediment plume size \citep{chu2012,tedstone2012}.
-Here, ocean water pressurizes subglacial channels at the ice front \citep[e.g.][]{how2017}, so the reduced correlation could result from the inconsistent relationship between subglacial sediment transport capacity and water discharge presented here.
+Here, ocean water pressurizes subglacial channels at the ice front \citep[e.g.][]{how2017}, so the observed reduced correlation could result from the inconsistent relationship between subglacial sediment transport capacity and water discharge (Figure~\ref{fig:Qw_vari}).
 
-Water discharge measurements at the hourly timescale remain unavailable in many catchments, further limiting subglacial sediment transport models in capturing specific events.
-The model output variability in subglacial channels decreases substantially beyond $1--5$ days of smoothing (Figure~\ref{fig:multi_run}).
-Over this period, subglacial channels could be in closer equilibrium conditions with water discharge variations leading to a stronger relationship with sediment discharge capacity (Section ~\ref{sect:scaling}).
-Applying water discharge averaged over these periods omits key sediment discharge characteristics, such as the timing of peaks in sediment transport capacity from glaciers (Figure~\ref{fig:model_outs}), that could play a key role in sediment export from glaciers.
-Yet, hydrographs with minimal variability may result in a strong correlation between sediment transport capacity and water discharge, if the water discharge evolves slowly with respect to the subglacial channel size.
+Water discharge measurements at the less than daily timescales could severely limit subglacial sediment transport models in capturing specific events.
+The decrease model output variability beyond $1$--$5$ days of smoothing could result in water discharge being in equilibrium with subglacial channel, when they are in fact not (Figure~\ref{fig:multi_run}).
+Over this period, water discharge variations could lead to a stronger relationship with sediment discharge capacity (Section ~\ref{sect:scaling}).
+Yet, such strong relationship would not represent the impact of shorter term fluctuations in water discharge.
+Note, that if basal melt largely dominates, then water discharge would evolve slowly with respect to the subglacial channel size.
+
+Clockwise hysterisis loops between sediment concentration and water discharge suggest that sediment availability is reduced over the event scale \citep[][]{williams1989}.
+These loops are often observed in glacierized catchments and used to suggest that access to subglacial sediment has been limited \citep[e.g.][]{collins1979,willis1996,richards2003,stott2007,delaney2018}.
+Results here suggest that clockwise hysterisis could also be expected in transport-limited regimes in many subglacial environments.
+Here, the smaller channel size on the rising limb would result in greater sediment transport capacity and sediment uptakes.
+Larger channel size would reduce sediment transport capacity for an equivalent water discharge on the falling limb, creating clockwise hysterisis.
 
@@
 Their size responds over several days to changes in discharge.
-Thus the response of sediment transport capacity is much more sensitive to changes in water discharge in subglacial channels compared to subaerial ones, although not as variable as if the size of the subglacial conduit was fixed.
+Thus, sediment transport capacity is  more sensitive to changes in water discharge in subglacial channels compared to subaerial ones.
 
@@
 results demonstrate that sediment transport capacity variability is higher in subglacial channels with evolving size across a range of channel shapes and friction factors.
-Increased variability occurs in environments where the water discharge has relatively small diurnal variations, as in the \icesheet case.
-The role of this increased noise in sediment transport capacity in evaluating hydro-climatic signals from sediment records in glacierized catchments needs further evaluation.
+Increased variability occurs even in environments where the water discharge has relatively small diurnal variations in the \icesheet case.
+Yet, elevated variability is especially strong with highly variable water discharge in the \alpine case.
+Increased noise in sediment transport capacity in evaluating hydro-climatic signals from sediment records in glacierized catchments needs further evaluation.
 Furthermore, this process may lead to sediment exhaustion by subglacial water repeatedly crossing the mobilization threshold across a range of water discharges.
@@
 The manuscript's second objective is to establish if water discharge varies with subglacial sediment transport capacity.
-Results here suggest that, even in a transport-limited subglacial system, an incoherent relationship between water and subglacial sediment discharge could be expected.
-This presents a challenge in linking hydro-climatic conditions to sediment export from glaciers.
+Results here suggest that, even in a transport-limited subglacial system, in most systems an incoherent relationship between water and subglacial sediment discharge could be expected.
+This presents a challenge in linking hydro-climatic conditions or events to sediment export from glaciers.
 Due to evolving subglacial channel size, the timing of peak sediment discharge capacity in subglacial channels occurs before peak water discharge during an event, whereas it is coincident in subaerial ones.
-Our analysis shows that a high correlation between water discharge and sediment transport capacity could be possible when water discharge varies at a slower rate than subglacial channel size.
-In natural environments, this appears to often not be the case but might be conceivable in very large catchments or ones at high latitudes with muted diurnal variations in melt.
+Our analysis shows, however, that covariance between water discharge and sediment transport capacity could be possible when water discharge varies at a slower rate than subglacial channel size.
+In natural environments, this appears to often not be the case, except where basal melt dominates.
 
-Lack of observation of the subglacial environment hindered our ability to quantify processes such as
-the shape of subglacial channels, the response time of subglacial channels to water discharge variations, and
-subglacial sediment size and sorting processes, linking shear stress to sediment discharge.
+The subglacial environment's few observations hindered our ability to quantify processes such as
+the shape of subglacial channels and response time of subglacial channels to water discharge variations.
+The poor constrains on subglacial sediment size and sorting processes make it difficult to link shear stress to sediment discharge capacity.
 Further quantifying these processes will help to better inform the response of sediment discharge capacity to water discharge forcing in subglacial environments.
 
-This study calls for careful analysis when examining the relationship between sediment transport capacity and hydro-climatic conditions from glacierized catchments, especially ones with high water discharge variability.
+This study calls for careful analysis when examining the relationship between sediment transport and hydro-climatic conditions from glacierized catchments, especially ones with high water discharge variability.
 
@@
   A.J.T. helped interpret data from Leverett Glacier and provided guidance on experiment design.
-M.A.W. provided data from Fieschergletscher and added the relationship between shear stress and water discharge (see Section \ref{sect:scaling}).
+  M.A.W. provided data from Fieschergletscher and added Section \ref{sect:scaling}.
   D.F. provided data from Fieschergletscher.
@@
   We benefited from fruitful discussions with J. Irving.
-We thank three reviewers for constructive and critical reviews on a previous version of the manuscript.
+  We thank two reviewers for constructive and critical reviews on a previous version of the manuscript.
 \end{acknowledgements}
@@
 
-\appendix
-\appendixtables   %% needs to be added in front of appendix tables
-\begin{table}
-  \caption{As Table~\ref{tab:Qs} but with the fractional exponents stated as rounded decimal numbers.
-For the subaerial regime the exponents are for displayed for the likely width-exponent $\alpha=1$, as well as its end members 0 (slot canyon) and 1 (only width increases and not depth).
-  }
-  \small
-  \label{tab:Qs2}
-  \begin{tabular}{lllll}
-    & Width \(\times \,\, \tau\) & MPM & EH & Bagnold\\
-    & \(w\, \tau\) & \(Q_s \propto w\, \tau^{3/2}\) & \(Q_s \propto w\, \tau^{5/2}\) & \(Q_s \propto w^{-1/2}\, \Omega^{3/2} H^{-2/3}\)\\
-    \hline
-    Subaerial (\(\alpha=0\)) & \(f^{0.7}\, Q^{0.7}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{1.7} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.1} \, \Psi^{1.7}\)\\
-    Subaerial (\(\alpha=0.3\)) & \(f^{0.7}\, Q^{0.8}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{1.3} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\
-    Subaerial (\(\alpha=0.5\)) & \(f^{0.7}\, Q^{0.8}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{1.3} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\
-    Subaerial (\(\alpha=1\)) & \(f^{0.7}\, Q^{1.0}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{0.7} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\[3pt]
-    R-channel & \(f^{0.4}\, Q^{0.8} \, \Psi^{0.6}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.7}\, Q^{1.4}\, \Psi^{1.8}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\
-    Pipe & \(f \,\quad Q^{2\phantom{.0}} \, S^{-1}\) & \(f^{1.5}\, Q^3 \, S^{-2.5}\) & \(f^{2.5}\, Q^{5\phantom{.0}}\, S^{-4.5}\) & \(f^{1.5} \,\,\,\,\, Q^{4.5} \, S^{-4.7}\)\\
-  \end{tabular}
-\end{table}
+% \appendix
+% \appendixtables   %% needs to be added in front of appendix tables
 
-%\appendixfigures  %% needs to be added in front of appendix figures
+% \appendixfigures  %% needs to be added in front of appendix figures
 
@@
 
-%%f
-%\begin{figure}[t]
-%\includegraphics[width=8.3cm]{FILE NAME}
-%\caption{TEXT}
-%\end{figure}
+%% f
+% \begin{figure}[t]
+%   \includegraphics[width=8.3cm]{FILE NAME}
+%   \caption{TEXT}
+% \end{figure}
 % 
@@
 % 
-%%f
-%\begin{figure*}[t]
-%\includegraphics[width=12cm]{FILE NAME}
-%\caption{TEXT}
-%\end{figure*}
+%% f
+% \begin{figure*}[t]
+%   \includegraphics[width=12cm]{FILE NAME}
+%   \caption{TEXT}
+% \end{figure*}
 % 
@@
 % 
-%%t
-%\begin{table}[t]
-%\caption{TEXT}
-%\begin{tabular}{column = lcr}
-%\tophline
+%% t
+% \begin{table}[t]
+%   \caption{TEXT}
+%   \begin{tabular}{column = lcr}
+      %       \tophline
       %       
-%\middlehline
+      %       \middlehline
       %       
-%\bottomhline
-%\end{tabular}
-%\belowtable{} % Table Footnotes
-%\end{table}
+      %       \bottomhline
+      %     \end{tabular}
+      %       \belowtable{} % Table Footnotes
+      %       \end{table}
       %       
@@
       %       
-%%t
-%\begin{table*}[t]
-%\caption{TEXT}
-%\begin{tabular}{column = lcr}
-%\tophline
+      %%       t
+      %       \begin{table*}[t]
+      %       \caption{TEXT}
+      %       \begin{tabular}{column = lcr}
+      %       \tophline
       %       
-%\middlehline
+      %       \middlehline
       %       
-%\bottomhline
-%\end{tabular}
-%\belowtable{} % Table Footnotes
-%\end{table*}
+      %       \bottomhline
+      %     \end{tabular}
+      %       \belowtable{} % Table Footnotes
+      %       \end{table*}
       %       
@@
       %       
-%%t
-%\begin{sidewaystable*}[t]
-%\caption{TEXT}
-%\begin{tabular}{column = lcr}
-%\tophline
+      %%       t
+      %       \begin{sidewaystable*}[t]
+      %       \caption{TEXT}
+      %       \begin{tabular}{column = lcr}
+      %       \tophline
       %       
-%\middlehline
+      %       \middlehline
       %       
-%\bottomhline
-%\end{tabular}
-%\belowtable{} % Table Footnotes
-%\end{sidewaystable*}
+      %       \bottomhline
+      %     \end{tabular}
+      %       \belowtable{} % Table Footnotes
+      %       \end{sidewaystable*}
       %       
@@
       %       
-%\begin{equation}
+      %       \begin{equation}
       %       
-%\end{equation}
+      %       \end{equation}
       %       
@@
       %       
-%\begin{align}
-%& 3 + 5 = 8\\
-%& 3 + 5 = 8\\
-%& 3 + 5 = 8
-%\end{align}
+      %       \begin{align}
+      %       & 3 + 5 = 8\\
+      %               & 3 + 5 = 8\\
+      %               & 3 + 5 = 8
+                        %     \end{align}
                         %                         
@@
                         %                         
-%\begin{matrix}
-%x & y & z\\
-%x & y & z\\
-%x & y & z\\
-%\end{matrix}
+                        %                         \begin{matrix}
+                        %                         x & y & z\\
+      %       x & y & z\\
+      %       x & y & z\\
+      %       \end{matrix}
       %       
@@
       %       
-%\begin{algorithm}
-%\caption{...}
-%\label{a1}
-%\begin{algorithmic}
-%...
-%\end{algorithmic}
-%\end{algorithm}
+      %       \begin{algorithm}
+      %       \caption{...}
+      %       \label{a1}
+      %       \begin{algorithmic}
+      %       ...
+      %       \end{algorithmic}
+      %       \end{algorithm}
       %       
@@
       %       
-%\begin{reaction}
+      %       \begin{reaction}
       %%%         \rightarrow should be used for normal (one-way) chemical reactions
@@
       %%%         \leftrightarrow should be used for resonance structures
-%\end{reaction}
+      %       \end{reaction}
       %       
@@
 Note that in pipe flow conditions, maximum sediment transport would coincide with peak water discharge, unlike the evolving channel size alters the timing (Figure \ref{fig:model_outs} and \ref{fig:Qw_vari}).
+
+When sediment availability processes are considered, a glacier's sediment discharge record also represent sediment availability and bedrock erosion from water pressure variations and sliding  \citep[e.g.
+][]{iverson2012,herman2015,delaney2023}.
+The greater amount of in variability of subglacial sediment evacuation can make it difficult to isolate the drivers of sediment export with respect to hydrological and glaciological conditions \citep{delaney2024}.

### 2024-07-01 Delaneyetal2024.tex
overleaf grammer

@@
 The size of subglacial channels is controlled by the opposing processes of channel opening by frictional heating of water flow on the one hand versus creep closure by ice flow on the other \citep{rothlisberger1972}.
-As a result  the subglacial channel size only evolves relatively slowly over days, whereas water discharge can vary more quickly over hours \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
+As a result, the subglacial channel size only evolves relatively slowly over days, whereas water discharge can vary more quickly over hours \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
 Therefore, subglacial water flow behaves more like pipe flow over short periods (hours, days).
@@
 Despite the physical differences between the subaerial and subglacial systems, water discharge and sediment export are often compared in contemporary glacierized catchments \citep[e.g.][]{willis1996,pearce2003,richards2003,swift2005,chu2009,tedstone2012,chu2012,overeem2017,delaney2018,swift2021,lu2022,andresen2024}.
-Many of theses studies discuss the disparity between water discharge and sediment export, highlighting the non-linear relationship between the two, along with the role of sediment access or production.
+Many of these studies discuss the disparity between water discharge and sediment export, highlighting the non-linear relationship between the two, along with the role of sediment access or production.
 Yet, a variable relationship between water discharge and sediment export could be expected from glaciers with evolving conduit size in a transport-limited, or hydraulically controlled, regime.
-The response of sediment transport capacity in subglacial sediment transport may affect the interpretation of sediment transport records in these glacierized catchments systems \citep[e.g.][]{ganti2016,mancini2023}.
-It is important that this relationship is clarified as sediment dynamics change in glacierized regions along with changing hydrology \citep[e.g.][]{brunner2019}, increasing variability of water discharge \citep{lane2019} and occurrence of extreme glacier melt  \citep[e.g.][]{overeem2015,cremona2023} that may drive high magnitude sediment transport events.
+The response of sediment transport capacity in subglacial sediment transport may affect the interpretation of sediment transport records in these glacierized catchment systems \citep[e.g.][]{ganti2016,mancini2023}.
+This relationship must be clarified as sediment dynamics change in glacierized regions along with changing hydrology \citep[e.g.][]{brunner2019}, increasing variability of water discharge \citep{lane2019} and occurrence of extreme glacier melt  \citep[e.g.][]{overeem2015,cremona2023} that may drive high magnitude sediment transport events.
 
@@
 Both models use the measured discharge to calculate water velocity, shear stress, and width-integrated shear stress, upon which both suspended sediment and bedload transport depend \citep[Figure \ref{fig:cartoon}; ][]{shields1936}.
-Our choice to evaluate shear stress omits selecting a sediment transport relationship and a grain-size parameter \citep[e.g.][]{shields1936,meyer1948}.
+Our choice to evaluate shear stress omits the selection of a sediment transport relationship and a grain-size parameter \citep[e.g.][]{shields1936,meyer1948}.
 
@@
 % 
-This value is used to establish the integrated shear stress across the channel $w_g\tau_g$.
+This value establishes the integrated shear stress across the channel $w_g\tau_g$.
 
@@
 % 
-As above, the width-integrated shear stress is $w_f\tau_f$.
+As mentioned above, the width-integrated shear stress is $w_f\tau_f$.
 
@@
     Where two values are given, the first refers to  \alpine{}, a scenario from Fieschergletscher, and the second to a glacier marginal to the \icesheet{} a scenario from Leverett Glacier.
-    A second line is given refers to the range of values examined in the parameter search.}
+    A second line refers to the range of values examined in the parameter search.}
   \small 
@@
 The first experiment aims to characterize the variability in sediment discharge capacity in subglacial relative to subaerial channels across a range of channel slopes and shapes, along with friction factors.
-Additionally, we examine differing water discharge variability by smoothing the water discharge over different period from $15$\,\unit{min} up to $15$\,\unit{days}.
+Additionally, we examine differing water discharge variability by smoothing the water discharge over different periods from $15$\,\unit{min} up to $15$\,\unit{days}.
 The results below present water velocity ($v_g$, $v_f$), shear stress ($\tau_g$, $\tau_f$), and width-integrated shear stress ($w_g\tau_g$, $w_f\tau_f$) from the subglacial and subaerial models.
@@
 
-The effects of channel shape, gradient, and roughness is established by running the model with random parameter values of channel slope and geometry factors ($\beta$,$\nabla z_c$, and $\alpha$) and friction factors ($f_r$ and $f_f$; see Table~\ref{table:vpm} for value range).
+The effects of channel shape, gradient, and roughness are established by running the model with random parameter values of channel slope and geometry factors ($\beta$,$\nabla z_c$, and $\alpha$) and friction factors ($f_r$ and $f_f$; see Table~\ref{table:vpm} for value range).
 Runs with parameter combinations are accepted if their mean subglacial water velocity over the season lies between $0.5$\,\unit{m}\,\unit{s}$^{-1}$ and $2$\,\unit{m}\,\unit{s}$^{-1}$ or if subaerial water velocity lies between $0.3$\,\unit{m}\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}\,\unit{s}$^{-1}$ \citep[e.g.][]{werder2010b,magnusson2012,chandler2013}.
@@
 % 
-To test covariance between sediment transport capacity and water discharge, model outputs are compared to water discharge from the glaciers using Spearman rank correlation.
+To test the covariance between sediment transport capacity and water discharge, model outputs are compared to water discharge from the glaciers using Spearman rank correlation.
 This metric accounts for the ordering of values, but not their magnitude.
@@
 The exponent for the width scaling $\alpha$ only impacts the relationship between sediment transport and water discharge in the EH relation in any meaningful way.
-However, for the value of $\alpha$ around $\frac{1}{2}$ (a value appropriate for most streams), the exponent on $Q$ for EH is only slightly greater than th other transport relations.
+However, for the value of $\alpha$ around $\frac{1}{2}$ (a value appropriate for most streams), the exponent on $Q$ for EH is only slightly greater than the other transport relations.
 
 The sediment discharge in the steady state R-channel scales very similarly to the subaerial channel case for all relations and virtually identically for the $\alpha=0.5$ case.
-However, note that the head gradients $\Psi$ are likely higher for comparable $Q$ in an ice-sheet marginal or alpine glacier setting than in a subaerial channel \citep{alley1997}.
-Thus, sediment transport capacity is likely  much higher in a steady state R-channel.
+However, the head gradients $\Psi$ are likely higher for comparable $Q$ in an ice-sheet marginal or alpine glacier setting than in a subaerial channel \citep{alley1997}.
+Thus, sediment transport capacity is likely much higher in a steady-state R-channel.
 
-Furthermore, R-channels will rarely operate in steady state as variations in water discharge.
-The timescale of  diurnal melt or severe rain or melt  are too short  for the channel to reach a steady state.
+Furthermore, R-channels will rarely operate in a steady state as variations in water discharge.
+The timescale of diurnal melt or severe rain or melt is too short for the channel to reach a steady state.
 In these cases with high water discharge variability, channels operate more like a pipe of fixed cross-section \citep[e.g.][]{gimbert2016}.
@@
 
-Smaller values of $\alpha$, or a channel cross section closer to a slot-canyon, result in greater variability in velocity and shear stress (Figure~\ref{fig:params}).
+Smaller values of $\alpha$, or a channel cross-section closer to a slot canyon, result in greater variability in velocity and shear stress (Figure~\ref{fig:params}).
 In both cases, the variability of small values of $\alpha$ does not approach that of the subglacial channels.
@@
 We note that these parameter values span a commonly accepted range.
-Therefore , we do not anticipate a scenario where variability in the subaerial system would exceed the subglacial system with this two hydrologic forcings.
+Therefore, we do not anticipate a scenario where variability in the subaerial system would exceed the subglacial system with these two hydrologic forcings.
 
@@
 Variable relationships between model outputs and water discharge emerge for the subaerial and subglacial cases due to the hysteresis in channel size in the subglacial model (Equation~\ref{eq:dS_dt}).
-Because there subaerial channels have no history dependence, each water discharge value in the subaerial channel produces a unique water velocity, shear stress, and width-integrated shear stress (Section~\ref{sect:sub_mode}).
+Because subaerial channels have no history dependence, each water discharge value in the subaerial channel produces a unique water velocity, shear stress, and width-integrated shear stress (Section~\ref{sect:sub_mode}).
 This characteristic results in a perfect rank correlation between the variables and water discharge (Figure~\ref{fig:Qw_vari}).
@@
 Furthermore, meltwater can be distributed to flow through several adjacent conduits impacting the sediment transport capacity in each \citep[e.g.][]{werder2013,hewitt2019,delaney2023}.
-The lumped models used here making comparison more direct by isolating the relationship between water discharge and sediment transport capacity in subglacial and subaerial systems. 
+The lumped models used here make the comparison more direct by isolating the relationship between water discharge and sediment transport capacity in subglacial and subaerial systems. 
 However, they neglect more complex, yet important, spatially distributed processes.
@@
 We note that in the ensemble runs in Section~\ref{sect:ensemble}, the parameters' range tested covers a likely range of viable shear stresses in both subglacial and subaerial channels.
-The greater sediment transport capacity in the subglacial channels implies that sediment grain-size underneath subglacial channels is likely larger that subaerial counterparts.
+The greater sediment transport capacity in the subglacial channels implies that sediment grain size underneath subglacial channels is likely larger than subaerial counterparts.
 
@@
 
-A sporadic relationship between subglacial sediment transport capacities and water discharge occurs due to hysterisis in subglacial channel size (Figure~\ref{fig:Qw_vari}).
+A sporadic relationship between subglacial sediment transport capacities and water discharge occurs due to hysteresis in subglacial channel size (Figure~\ref{fig:Qw_vari}).
 This characteristic limits the use of water discharge as an indicator of sediment discharge capacity in these systems.
@@
 The co-varying relationship between sediment discharge capacity and water discharge in subaerial conditions could be noticeable $\sim20$\,\unit{km} downstream of the Leverett site.
-A strong correlation persists between sediment plume size and the Watson river's water discharge into the Kangerlussuaq fjord \citep{chu2009,mcgrath2010}.
+A strong correlation persists between sediment plume size and the Watson River's water discharge into the Kangerlussuaq fjord \citep{chu2009,mcgrath2010}.
 In contrast, in marine-terminating glacier catchments, a less consistent relationship may occur between water discharge or melt extent and sediment plume size \citep{chu2012,tedstone2012}.
@@
 
-Water discharge measurements at the less than daily timescales could severely limit subglacial sediment transport models in capturing specific events.
-The decrease model output variability beyond $1$--$5$ days of smoothing could result in water discharge being in equilibrium with subglacial channel, when they are in fact not (Figure~\ref{fig:multi_run}).
+Water discharge measurements at less than daily timescales could severely limit subglacial sediment transport models in capturing specific events.
+The decreased model output variability beyond $1$--$5$ days of smoothing could result in water discharge being in equilibrium with the subglacial channel when they are in fact not (Figure~\ref{fig:multi_run}).
 Over this period, water discharge variations could lead to a stronger relationship with sediment discharge capacity (Section ~\ref{sect:scaling}).
-Yet, such strong relationship would not represent the impact of shorter term fluctuations in water discharge.
+Yet, such a strong relationship would not represent the impact of shorter-term fluctuations in water discharge.
 Note, that if basal melt largely dominates, then water discharge would evolve slowly with respect to the subglacial channel size.
 
-Clockwise hysterisis loops between sediment concentration and water discharge suggest that sediment availability is reduced over the event scale \citep[][]{williams1989}.
+Clockwise hysteresis loops between sediment concentration and water discharge suggest that sediment availability is reduced over the event scale \citep[][]{williams1989}.
 These loops are often observed in glacierized catchments and used to suggest that access to subglacial sediment has been limited \citep[e.g.][]{collins1979,willis1996,richards2003,stott2007,delaney2018}.
-Results here suggest that clockwise hysterisis could also be expected in transport-limited regimes in many subglacial environments.
+Results here suggest that clockwise hysteresis could also be expected in transport-limited regimes in many subglacial environments.
 Here, the smaller channel size on the rising limb would result in greater sediment transport capacity and sediment uptakes.
-Larger channel size would reduce sediment transport capacity for an equivalent water discharge on the falling limb, creating clockwise hysterisis.
+A larger channel size would reduce sediment transport capacity for an equivalent water discharge on the falling limb, creating clockwise hysteresis.
 
@@
 The manuscript's second objective is to establish if water discharge varies with subglacial sediment transport capacity.
-Results here suggest that, even in a transport-limited subglacial system, in most systems an incoherent relationship between water and subglacial sediment discharge could be expected.
+Results here suggest that, even in a transport-limited subglacial system, in most systems, an incoherent relationship between water and subglacial sediment discharge could be expected.
 This presents a challenge in linking hydro-climatic conditions or events to sediment export from glaciers.
@@
 The subglacial environment's few observations hindered our ability to quantify processes such as
-the shape of subglacial channels and response time of subglacial channels to water discharge variations.
-The poor constrains on subglacial sediment size and sorting processes make it difficult to link shear stress to sediment discharge capacity.
+the shape of subglacial channels and the response time of subglacial channels to water discharge variations.
+The poor constraints on subglacial sediment size and sorting processes make it difficult to link shear stress to sediment discharge capacity.
 Further quantifying these processes will help to better inform the response of sediment discharge capacity to water discharge forcing in subglacial environments.
@@
   Code, with links to the data, can be found at \url{https://bitbucket.org/IanDelaney/xsection/src/master/}.
-  Data from Leverett glacier have been previous published in \citet{tedstone2013}.
-  Data from Fiescher glacier have been previous published in \citet{delaney2024}.
-  Code will be uploaded to a permanent repository, with FAIR principles pending acceptance of the paper.}
+  Data from Leverett glacier have been previously published in \citet{tedstone2013}.
+  Data from Fieschergletscher have been previously published in \citet{delaney2024}.
+  The code will be uploaded to a permanent repository, with FAIR principles pending acceptance of the paper.}
 
@@
   I.D. designed the study, developed, and implemented the model and experiments, and wrote the manuscript.
-  A.J.T. helped interpret data from Leverett Glacier and provided guidance on experiment design.
+  A.J.T. helped interpret findings from Leverett Glacier and guided experiment design.
   M.A.W. provided data from Fieschergletscher and added Section \ref{sect:scaling}.

### 2024-07-01 Delaneyetal2024.tex
version to pass around

@@
 This characteristic results in a perfect rank correlation between the variables and water discharge (Figure~\ref{fig:Qw_vari}).
+An inconsistent relationship between subglacial model outputs and water discharge persists across the range of parameters examined in the ensemble runs (Figure S15).
 \begin{figure}[hbt!]
@@
 The lumped models used here make the comparison more direct by isolating the relationship between water discharge and sediment transport capacity in subglacial and subaerial systems. 
-However, they neglect more complex, yet important, spatially distributed processes.
+However, they neglect more complex, yet important, spatially distributed processes. a
 
@@
 Conversely, the reduced variability in water discharge in \icesheet{} may result in the subglacial channel's size being closer to equilibrium from the smaller variations in water discharge.
-In this case, the exponent on water discharge for shear stress is likely substantially less than $2$, but greater than $\frac{4}{5}$ that occurs in a steady-state R-channel (see Section \ref{sect:scaling}).
+In this case, the exponent on water discharge for shear stress is likely substantially less than $w\tau \propto Q^2$, but greater than $w \tau \propto Q^{\frac{4}{5}}$ that occurs in a steady-state R-channel (see Section \ref{sect:scaling}).
 As a result, there is a stronger relationship between subglacial model outputs and water discharge in \icesheet{} (Figure~\ref{fig:Qw_vari}) and less variability model outputs in the \icesheet case (Figure~\ref{fig:multi_run}).
@@
 Concerning the manuscript's first objective,
-results demonstrate that sediment transport capacity variability is higher in subglacial channels with evolving size across a range of channel shapes and friction factors.
+results demonstrate that sediment transport capacity variability is high in subglacial channels with evolving size across a range of channel shapes and friction factors.
 Increased variability occurs even in environments where the water discharge has relatively small diurnal variations in the \icesheet case.
@@
 Due to evolving subglacial channel size, the timing of peak sediment discharge capacity in subglacial channels occurs before peak water discharge during an event, whereas it is coincident in subaerial ones.
-Our analysis shows, however, that covariance between water discharge and sediment transport capacity could be possible when water discharge varies at a slower rate than subglacial channel size.
+Covariance between water discharge and sediment transport capacity could be possible when water discharge varies at a slower rate than subglacial channel size.
 In natural environments, this appears to often not be the case, except where basal melt dominates.

### 2024-07-04 Delaneyetal2024.tex
Updates from Overleaf
change on machine

@@
 
-Despite the physical differences between the subaerial and subglacial systems, water discharge and sediment export are often compared in contemporary glacierized catchments \citep[e.g.][]{willis1996,pearce2003,richards2003,swift2005,chu2009,tedstone2012,chu2012,overeem2017,delaney2018,swift2021,lu2022,andresen2024}.
-Many of these studies discuss the disparity between water discharge and sediment export, highlighting the non-linear relationship between the two, along with the role of sediment access or production.
+Despite the physical differences between the subaerial and subglacial systems, water discharge and sediment export are often compared in contemporary glacierized catchments \citep[e.g.][]{willis1996,hodson1998,pearce2003,richards2003,swift2005,chu2009,tedstone2012,chu2012,overeem2017,delaney2018,swift2021,lu2022,andresen2024}.
+Many of these studies discuss the disparity between water discharge and sediment export, particularly the non-linear relationship between the two, along with the role of sediment access or production.
 Yet, a variable relationship between water discharge and sediment export could be expected from glaciers with evolving conduit size in a transport-limited, or hydraulically controlled, regime.
@@
 These smoothing timescales remove the diurnal variations in water discharge.
-In the \icesheet case, the smaller variability in model outputs compared to \alpine means that the R-channel is in closer equilibrium between channel size and water discharge (Section~\ref{sect:scaling}).
-In this case, water discharge variations would result in smaller sediment transport capacity variations than even in subaerial cases as the subglacial channel size evolves to accommodate them (Table~\ref{tab:Qs}).
 

### 2024-07-08 Delaneyetal2024.tex

@@
 \Author[1][IanArburua.Delaney@unil.ch]{Ian}{Delaney} %% correspondence author
-\Author[2]{Andrew J.}{Tedstone}
+\Author[1,2]{Andrew J.}{Tedstone}
 \Author[3,4]{Mauro A.}{Werder}
@@
   Yet, subglacial sediment transport capacity can vary less if water discharge variations are closer to equilibrium with subglacial channel size.
-  In another numerical experiment, we show that the changing channel size results in sediment transport capacity peaking before the maximum water discharge.
+  In a dedicated numerical experiment, we show that the changing channel size results in sediment transport capacity peaking before the maximum water discharge.
   Results also indicate that high subglacial sediment transport capacities can occur across a wide range of water discharges.
@@
 Increases in sediment transport have been observed in Greenland \citep{bendixen2017}, the European Alps \citep{costa2017}, the Himalayas \citep{li2021}, and the Andes \citep{vergara2022}.
-To accurately explain observed changes to sediment transport in glacierized catchments requires examining the processes controlling sediment discharge and its variations with water discharge  \citep[e.g.][]{riihimaki2005,swift2005}.
+To accurately explain observed changes to sediment transport in glacierized catchments, the processes controlling sediment discharge and its variations with water discharge  need to be examined \citep[e.g.][]{riihimaki2005,swift2005}.
 
-Glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from under glaciers over periods of millennia or longer \citep[c.f.][]{hallet1979,iverson2012,ugelvig2018}.
+Glacier abrasion and quarrying sculpt landscapes, and create sediment that is transported fluvially over periods of millennia or longer \citep[c.f.][]{hallet1979,iverson2012,ugelvig2018}.
 Pressurized subglacial water can transport this sediment from underneath glaciers \citep{walder1994,creyts2013,beaud2018,delaney2019} if it is reachable by the water.
@@
 
-In subaerial channels operating with open channel flow, the wetted area $S$ of the channel evolves with changing water discharge $Q$, by changing both the channel width and the water depth \citep{leopold1953}.
+In subaerial channels operating with open channel flow, $S$  evolves with changing water discharge $Q$, by changing both the channel width and the water depth \citep{leopold1953}.
 The change in water depth results in a proportional increase in water velocity, and the shear stress $\tau$ increases according to Equation~\ref{eq:tau}.
@@
 The response of water velocity to changing water discharge in subglacial channels differs from subaerial ones, however.
-The size of subglacial channels is controlled by the opposing processes of channel opening by frictional heating of water flow on the one hand versus creep closure by ice flow on the other \citep{rothlisberger1972}.
+The size of subglacial channels is controlled by the opposing processes of channel opening by frictional heating of water flow on the one hand, versus creep closure by ice flow on the other \citep{rothlisberger1972}.
 As a result, the subglacial channel size only evolves relatively slowly over days, whereas water discharge can vary more quickly over hours \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
@@
   \centering
-  \includegraphics[width=0.6\linewidth]{Fig1.pdf}
+  \includegraphics[width=0.8\linewidth]{Fig1.pdf}
   \caption{Sketch for the different responses of subglacial and subaerial channels to increased water discharge over short time scales.
@@
 
-Sediment mobilization in subaerial and subglacial channels, therefore, responds differently to changing water discharge.
+Because of the above, sediment mobilization in subaerial and subglacial channels responds differently to changing water discharge.
 These differences are implicitly included in a range of models quantifying sediment transport in both subglacial and subaerial channels  \citep[e.g.][]{walder1994,alley1997,tucker1997,creyts2013,beaud2018,delaney2019,hewitt2019,wickert2019}.
-To date, explicit parameterizations of subglacial sediment transport capacity with respect to water discharge have assumed fixed channel size  \citep{alley1997}. and
-This result demonstrates a strongly non-linear response in subglacial sediment transport capacity to water discharge.
-Several modeling frameworks examining subglacial sediment transport with evolving channel size \citep{creyts2013,beaud2018,delaney2019,hewitt2019}.
-However, these works minimally discuss the impact of different hydrologies on sediment transport capacity underneath glaciers for interpreting sediment transport records.
+To date, several modeling frameworks examine subglacial sediment transport with evolving channel size \citep{creyts2013,beaud2018,delaney2019,hewitt2019}.
+However, these works minimally discuss the impact of different hydrological regimes on sediment transport capacity underneath glaciers for interpreting sediment transport records.
+The few explicit parameterizations of subglacial sediment transport capacity with respect to water discharge have assumed fixed channel size  \citep{alley1997}. 
+These formulations demonstrate a strongly non-linear response in subglacial sediment transport capacity to water discharge.
 Understanding these processes is imperative for establishing the effect of hydro-climatic conditions on subglacial sediment dynamics.
-Particularly because sediment discharge capacity controls the mobilization and deposition of sediment, making it a fundamental aspect of subglacial sediment evacuation, especially over short timescales.
+This is especially the case because sediment discharge capacity controls the mobilization and deposition of sediment, making it a fundamental aspect of subglacial sediment evacuation, especially over short timescales.
 
@@
 Yet, a variable relationship between water discharge and sediment export could be expected from glaciers with evolving conduit size in a transport-limited, or hydraulically controlled, regime.
-The response of sediment transport capacity in subglacial sediment transport may affect the interpretation of sediment transport records in these glacierized catchment systems \citep[e.g.][]{ganti2016,mancini2023}.
+The response of sediment transport capacity to water discharge variations may affect the interpretation of sediment transport records in these glacierized catchment systems \citep[e.g.][]{ganti2016,mancini2023}.
 This relationship must be clarified as sediment dynamics change in glacierized regions along with changing hydrology \citep[e.g.][]{brunner2019}, increasing variability of water discharge \citep{lane2019} and occurrence of extreme glacier melt  \citep[e.g.][]{overeem2015,cremona2023} that may drive high magnitude sediment transport events.
@@
 These formulations extend the relationships presented in \citet{alley1997}.
-We then use numerical models to examine sediment transport capacity in subglacial channels that evolve in size and subaerial channels.
+We then use numerical models to examine sediment transport capacity in both subglacial channels that evolve in size and subaerial channels that remain fixed.
 They are applied to proglacial hydrological records from an Alpine glacier in Switzerland (Fieschergletscher) and a land-terminating glacier in Greenland (Leverett Glacier).
@@
 
-Note that we omit analysis of suspended sediment discharge records due to their dependence on sediment supply in capturing variations in sediment discharge, in addition to sediment transport capacity \citep[e.g.][]{delaney2019}.
-
 \begin{figure}[hbt!]
   \centering
-  \includegraphics[width=0.8\linewidth]{Fig2.pdf}
+  \includegraphics[width=0.9\linewidth]{Fig2.pdf}
   \caption{Water discharge from Fieschergletcher (\alpine) and Leverett Glacier (\icesheet).
@@
 
-\subsection{Sensitivity of sediment transport in different drainage regimes}
+\FloatBarrier
+\subsection{Increased variability in evolving subglacial channels occurs across a range of channel shapes, slopes,  and friction values}
+\label{sect:ensemble}
+
+The first numerical experiment aims to compare the variability between the subglacial and subaerial model outputs to a range of channel shapes, friction factors, and water discharge variability, as the subglacial channel's cross-sectional area is allowed to evolve.
+Following the findings in Section~\ref{sect:scaling}, variability in subglacial sediment discharge capacity could approach that of subaerial systems, should water discharge variations be sufficiently small.
+For simplicity, we only examine $w\tau$.
+
+Across the range of parameters examined (Section\,\ref{sect:imp}), variability in all model outputs remains higher in the subglacial system compared to the subaerial one for both \alpine{} and \icesheet{} (Figure~\ref{fig:multi_run}).
+In some cases, subglacial model outputs' variability is a magnitude larger than their subaerial counterparts.
+Results show that variability in both subglacial and subaerial outputs decreases with smoothing time longer than approximately $1$--$5$ days (Figure~\ref{fig:multi_run}).
+These smoothing timescales remove the diurnal variations in water discharge, thereby reducing variability in model outputs. 
+Velocity variability in \icesheet{} is substantially smaller than \alpine{} (Figure~\ref{fig:multi_run}\,a and \, d).
+This could result from the subglacial conduit being in closer equilibrium with water discharge.
+Variations in \icesheet{} shear stress are comparable to \alpine due to the effects the evolution of shear stress with subglacial conduit size (Equations ~\ref{eq:dS_dt} and \ref{eq:DW};). 
+The much greater water flux in \icesheet{} drives the larger width integrated shear stress variations (Figure~\ref{fig:Qw}\,c and f).
+
+\begin{figure}[hbt!]
+  \centering
+  \includegraphics[width=0.9\linewidth]{Fig3.pdf}
+  \caption{Standard deviation of detrended model outputs for different smoothing times with different subglacial and subaerial channel shapes and friction factors.
+    Shaded areas denote the range of standard deviations from the accepted $100$ parameter combinations (Section~\ref{sect:imp}).
+    Solid lines denote the mean value of standard deviations.
+    Markers show smoothing periods ($15$\,\unit{min}, $1$\,\unit{hr}, $6$\,\unit{hr} $12$\,\unit{hr}, $1$\,\unit{d}, $5$\,\unit{d}, $10$\,\unit{d}, and $15$\,\unit{d}).
+  }
+  \label{fig:multi_run}
+\end{figure}
+
+Smaller values of $\alpha$, or a channel cross-section closer to a slot canyon, result in greater variability in velocity and shear stress (Figure~\ref{fig:params}).
+In both cases, the variability of small values of $\alpha$ does not approach that of the subglacial channels.
+Yet, larger values of $\alpha$ can result in greater variability in width-integrated shear stress, as greater channel width changes increase the width-integrated shear stress.
+Additionally, steeper subaerial channel slopes result in greater variability in both shear stress and width-integrated shear stress but only approach the subglacial channel's variability in the \alpine case (Figure~\ref{fig:params}).
+We note that these parameter values span a commonly accepted range.
+Therefore, we do not anticipate a scenario where variability in the subaerial system would exceed the subglacial system with these two hydrologic forcings.
+
+Greater subglacial and subaerial friction factors $f_i$ and $f_p$ result in greater variability in the shear stress and width-integrate shear stress in both cases (Figure~\ref{fig:params}).
+This result is expected given Equations~\ref{eq:tau_g} and \ref{eq:vf}.
+However, more velocity variability is present in low  $f_i$ and $f_p$ values. 
+Low values of $f_i$ result in slower growth rates in subglacial channels (Equations~\ref{eq:dS_dt} and \ref{eq:dh}).
+As a result, water velocity, as opposed to channel growth, accommodates increases in water discharge, and velocity variability increases.
+Smaller values of channel factor $\beta$, creating low and broad channels, result in more variability in width-integrated shear stress.
+Here, the channel width can grow more quickly in response to water discharge increases as compared to a semi-circular channel with $\beta = \pi$ (Equation~ \ref{eq:dh}).
+
+\begin{figure}[hbt!]
+  \centering
+  \includegraphics[width=0.9\linewidth]{Fig4.pdf}
+  \caption{Parameter values compared to variability in model outputs.
+    Plots on the right (orange, light blue) correspond to the \alpine{} case. Plots on the left (brown, dark blue) correspond to the \icesheet case.
+    Outputs are shown using $15$ minute smoothing time. 
+  }
+  \label{fig:params}
+\end{figure}
+
+\FloatBarrier
+\subsection{Changing subglacial channel size drives different timing and variability in sediment transport capacity}
+
+The second numerical experiment aims to quantify the sources of increased variability in the subglacial model outputs as they exhibit different seasonal evolutions and peaks (Figure~\ref{fig:model_outs}).
+Variable relationships between model outputs and water discharge emerge for the subaerial and subglacial cases due to the hysteresis in channel size in the subglacial model (Equation~\ref{eq:dS_dt}).
+Because subaerial channels have no history dependence, each water discharge value in the subaerial channel produces a unique water velocity, shear stress, and width-integrated shear stress (Section~\ref{sect:sub_mode}).
+This characteristic results in a perfect rank correlation between the variables and water discharge (Figure~\ref{fig:Qw_vari}).
+An inconsistent relationship between subglacial model outputs and water discharge persists across the range of parameters examined in the ensemble runs (Figure S15).
+\begin{figure}[hbt!]
+  \centering
+  \includegraphics[width=\linewidth]{Fig5.png}
+  \caption{Model outputs from simulations using the hydrographs in panels a and e for the scenarios \alpine{} (a-d) and \icesheet{} (e-h).
+    Black (gray) lines in a and e represent the subglacial channel size (water discharge).
+    Blue (orange) lines represent outputs from the subglacial (subaerial) channel.
+    Data are shown at $15$\unit{min} intervals.
+    Arrows show examples where variables peak in the subglacial channel before the subaerial one.
+    Insets in f-h show the peak melt event denoted by the shaded area in panels e--h, with an arbitrary y-axis.
+  }
+  \label{fig:model_outs}
+\end{figure}
+
+Conversely, the history dependence on channel size in subglacial channels means that model outputs occur across a large range of water discharges.
+For instance, in subglacial channels in \alpine{},  high water velocity values and shear stresses can occur from a low water discharge  ($\sim\,4$\,\unit{m}$^3$\,\unit{s}$^{-1}$) to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ (Figure~\ref{fig:Qw_vari}\,a).
+In \icesheet{}, water velocities close to the seasonal mean value can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
+The subglacial channel's evolving width can counteract some of these effects.
+Width-integrated shear stress generally increases with water discharge, with greater rank correlation compared to water velocity or shear stress (\alpine{}, Figure~\ref{fig:Qw_vari} a--c).
+Yet, even the width-integrated shear stress can vary substantially relative to water discharge.
+The highest values of width-integrated shear stress occur at water discharge values ranging from roughly $11$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ to over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
+The variability in width-integrated shear stress is less pronounced in the \icesheet{} scenario, where the hydrograph has less diurnal variability (Figure~\ref{fig:Qw_vari} c, f).
+This results from the discharge variations being in closer equilibrium with subglacial channels compared to \alpine{} (Sections~\ref{sect:scaling} and~\ref{sect:ensemble}).
+
+\begin{figure}[hbt!]
+  \centering
+  \includegraphics[width=\linewidth]{Fig6.png}
+  \caption{
+    Relationship between water discharge and normalized velocity, shear stress, and width-integrated shear stress for \alpine{} (a-c) and \icesheet{} (d-f).
+    Variables on the x-axis have been normalized to mean values.
+    $R_g$ ($R_f$) shows the Spearman rank correlations for the subglacial (subaerial) outputs.
+    Plots are shown with discharge and model outputs at $15$\unit{min} intervals.
+  }
+  \label{fig:Qw_vari}
+\end{figure}
+
+Peaks in subaerial model outputs occur coincident with peaks in water discharge (Figure~\ref{fig:model_outs}).
+In the subglacial channel, peaks in model outputs generally occur when water discharge increases at the fastest rate, but before the maximum water discharge.
+Reduced water velocity takes place when water discharge stabilizes at its peak and channel growth continues.
+As a result, subglacial sediment transport capacity is greatest on the hydrograph's rising limb, relative to the falling limb.
+
+\FloatBarrier
+\section{Discussion}
+
+\subsection{Increased variability in  sediment transport capacity in subglacial systems}
+\label{sect:dis_qsc}
+
+The greater variations in shear stress in subglacial channels presented here may actually cause even greater variations in sediment transport capacity in subglacial channels than presented above  (Figure~\ref{fig:multi_run}).
+In sediment transport capacity relationships, such as in \citet{meyer1948} or \citet{engelund1967}, shear stress is scaled to the power of $\frac{3}{2}$ or $\frac{5}{2}$, respectively (Section~\ref{sect:scaling}).
+The exponent greater than $1$ magnifies sediment discharge variability beyond the variable sediment transport parameters described above (Figure~\ref{fig:multi_run}; Table~\ref{tab:Qs}).
+
+Greater variations in subglacial sediment transport capacity could cause a supply-limited regime at many glaciers \citep{alley1997}.
+In subglacial systems, sediment's critical shear stress, or threshold at which sediment mobilization occurs, can be reached more frequently and across many water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
+This result suggests sediment export here is especially sensitive to observed changes in water discharge variability \citep{lane2019b}.
+Sediment exhaustion through repeatedly crossing the mobilization threshold may explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \citep{overeem2017}.
+The same process could also result in sediment discharge's strong dependence on sediment production from sliding in many cases \citep{herman2015,koppes2015}.
+
+While transport-limited states at glaciers are likely rare \citep[e.g.][]{alley1997}, abundant sediment can persist underneath glaciers \citep[e.g.][]{walter2014,stevens2022,delaney2022}.
+As a result, in some cases sediment export can be hydrologically controlled and efficient \citep[e.g.][]{swift2005,andresen2024}.
+At these glaciers, the great variability in subglacial sediment transport capacity may make it difficult to link sediment discharge to hydrology, especially when peak events occur \citep{cowan1988,delaney2018,lu2022}.
+As a result, the sediment transport capacity response to individual extreme precipitation or melt events could be difficult to establish.
+The results here suggest that different channel sizes with the same hydrology forcing can result in very different sediment transport capacities.
+
+Catchments with reduced water discharge variability may experience yet less variability in sediment transport capacity, shown by the \alpine{} and \icesheet{} model outputs.
+Indeed, the more decoupled and sporadic relationship between model outputs and water discharge in \alpine{} results from the relatively larger discharge variations on sub-daily to weekly timescales (Figure~\ref{fig:Qw_vari}).
+This occurs as the subglacial channel's size evolves slowly compared to the variations in water discharge.
+High water discharge variability in \alpine{} may cause shear stress to approach $Q^{2}$in assuming pipe-flow conditions where the water discharge varies largely compared to channel size \citep[Section \ref{sect:scaling}; c.f.][]{alley1997}.
+Conversely, the reduced relative variability in water discharge in \icesheet{} and stronger correlation between width-integrated shear stress may result in the subglacial channel's size being closer to equilibrium from the smaller variations in water discharge.
+In this case, the exponent on water discharge for shear stress is likely substantially less than $w\tau \propto Q^2$, but greater than $w \tau \propto Q^{\frac{4}{5}}$ that occurs in a steady-state R-channel (see Section \ref{sect:scaling}).
+As a result, there is a stronger relationship between subglacial model outputs and water discharge in \icesheet{} (Figure~\ref{fig:Qw_vari}) and less variability model outputs in the \icesheet case (Figure~\ref{fig:multi_run}).
+
+\subsection{Interpreting sediment transport records from glacierized catchments with respect to water discharge}
+
+A sporadic relationship between subglacial sediment transport capacities and water discharge occurs due to hysteresis in subglacial channel size (Figure~\ref{fig:Qw_vari}).
+This characteristic limits the use of water discharge as an indicator of sediment discharge capacity in these systems.
+As a result, characteristics such as bankfull or effective water discharge that link geomorphic work to hydro-climatic conditions could have limited meaning in evaluating subglacial sediment transport \citep{wolman1960,lenzi2006}.
+A glacier's sediment transport capacity is impacted by the ice thickness controlling the channel closure rate and the glacier's surface slope, in addition to water discharge and sediment size \citep[Figure~\ref{fig:multi_run}, Section~\ref{sect:sub_mode}; ][]{rothlisberger1972,gimbert2016,stevens2022,walder1994}.
+This multitude of processes lies in contrast to many subaerial channels, where transport capacity typically responds to water discharge, sediment size, channel shape,  and hydraulic gradient.
+The latter of these parameters can remain relatively stable over the years or longer \citep[Section~\ref{sect:fluv}; e.g.][]{tucker1997}.
+Strong correlations between water discharge and sediment export in glacier systems could indicate other processes, such as increased sediment access \citep{zhang2022}.
+Furthermore, the observed correlation between water discharge and sediment discharge far downstream likely represents subaerial transport processes, especially with respect to bedload \citep{mancini2023}.
+
+The co-varying relationship between sediment discharge capacity and water discharge in subaerial conditions could be noticeable $\sim20$\,\unit{km} downstream of the Leverett site.
+A strong correlation persists between sediment plume size and the Watson River's water discharge into the Kangerlussuaq fjord \citep{chu2009,mcgrath2010}.
+In contrast, in marine-terminating glacier catchments, a less consistent relationship may occur between water discharge or melt extent and sediment plume size \citep{chu2012,tedstone2012}.
+Here, ocean water pressurizes subglacial channels at the ice front \citep[e.g.][]{how2017}, so the observed reduced correlation could result from the inconsistent relationship between subglacial sediment transport capacity and water discharge (Figure~\ref{fig:Qw_vari}).
+
+Water discharge measurements at less than daily timescales could severely limit subglacial sediment transport models in capturing specific events.
+The decreased model output variability beyond $1$--$5$ days of smoothing could result in water discharge being in equilibrium with the subglacial channel when they are in fact not (Figure~\ref{fig:multi_run}).
+Over this period, water discharge variations could lead to a stronger relationship with sediment discharge capacity (Section ~\ref{sect:scaling}).
+Yet, such a strong relationship would not represent the impact of shorter-term fluctuations in water discharge.
+Note, that if basal melt largely dominates, then water discharge would evolve slowly with respect to the subglacial channel size.
+
+Clockwise hysteresis loops between sediment concentration and water discharge suggest that sediment availability is reduced over the event scale \citep[][]{williams1989}.
+These loops are often observed in glacierized catchments and used to suggest that access to subglacial sediment has been limited \citep[e.g.][]{collins1979,willis1996,richards2003,stott2007,delaney2018}.
+Results here suggest that clockwise hysteresis could also be expected in transport-limited regimes in many subglacial environments.
+Here, the smaller channel size on the rising limb would result in greater sediment transport capacity and sediment uptakes.
+A larger channel size would reduce sediment transport capacity for an equivalent water discharge on the falling limb, creating clockwise hysteresis.
+
+\subsection{Experiment limitations}
+
+Several aspects of the models and experiment design make comparison of the two systems difficult. 
+The lumped nature of the models means that they operate independently of the upstream drainage network.
+Additionally, we omit analysis of suspended sediment discharge records due to their dependence on sediment supply in capturing variations in sediment discharge, in addition to sediment transport capacity \citep[e.g.][]{delaney2019}.
+In reality, processes such as sediment access are highly important in controlling sediment export in glacierized catchments \citep[e.g.][]{herman2015,vergara2022}.
+Furthermore, meltwater can be distributed to flow through several adjacent conduits impacting the sediment transport capacity in each \citep[e.g.][]{werder2013,hewitt2019,delaney2023}.
+The lumped models used here make the comparison isolate the relationship between water discharge and sediment transport capacity in subglacial and subaerial systems. 
+However, they neglect more complex, yet important spatially distributed processes.
+
+Comparing model outputs between the subaerial and subglacial case is difficult due to the different hydraulic gradients in the two systems, with the subglacial one generally being much steeper \citep{alley1997}.
+The water velocities are similar in both channel types (Figure~\ref{fig:model_outs}).
+However, the shear stresses and width-integrated shear stresses across the channel bed that control sediment transport capacity are much smaller in subaerial channels.
+We note that in the ensemble runs in Section~\ref{sect:ensemble}, the parameters' range tested covers a likely range of viable shear stresses in both subglacial and subaerial channels.
+The greater sediment transport capacity in the subglacial channels implies that sediment grain size underneath subglacial channels is likely larger than subaerial counterparts.
+
+Width-integrated shear stress is examined in the two numerical experiments as it does not have a dependence on grain size, unlike sediment transport relationships \citep[][Section \ref{sect:scaling}]{meyer1948}.
+This makes comparison between the two systems simpler. 
+Preferential transport of smaller sediment clasts and the input of upstream sediment impact sediment transport capacity as grain size evolves \citep[e.g.][]{gomez1983}.
+These processes are only beginning to be evaluated underneath glaciers \citep{aitken2024}, but can be important in subaerial systems.
+Subglacial sorting processes could be an additional source of variability in sediment export from glaciers with respect to water discharge, especially for bedload transport.
+
+\conclusions
+
+The sediment transport capacity of both subglacial and subaerial channels are driven by its width and the shear stress exerted by the flowing water, which is proportional to flow velocity squared.
+Subaerial channels immediately alter both their width and water velocity in response to changing water discharge.
+In contrast, pressurized subglacial channels largely accommodate rapidly changing water discharge by altering water velocity, while their size only responds to changes in discharge over longer timescales, for instance several days.
+Thus, sediment transport capacity is more sensitive to changes in water discharge in subglacial channels compared to subaerial ones.
+
+The manuscript's first objective aims to evaluate the relative variability in sediment transport capacity between the two channel types.
+Results demonstrate that sediment transport capacity variability is higher in subglacial channels with evolving size across a range of channel conditions.
+Increased variability occurs even in environments where the water discharge has relatively small diurnal variations in the \icesheet case.
+Yet, elevated variability is especially strong with highly variable water discharge in the \alpine case.
+Increased variability in sediment transport capacity in evaluating hydro-climatic signals from sediment records in glacierized catchments needs further evaluation.
+Furthermore, greater variability in sediment transport capacity may lead to sediment exhaustion by subglacial water repeatedly crossing the mobilization threshold across a range of water discharges.
+
+The manuscript's second objective is to establish if water discharge varies with subglacial sediment transport capacity.
+Results here suggest that, even in a transport-limited subglacial system, an incoherent relationship between water and subglacial sediment discharge could be expected in most systems.
+This presents a challenge in linking hydro-climatic conditions or events to sediment export from glaciers.
+In subglacial channels, the timing of peak sediment discharge capacity occurs before peak water discharge during an event, due to evolving channel size.
+In subaerial channels, the timing of peak sediment discharge capacity and water discharge coincide.
+whereas it is coincident in subaerial ones.
+Covariance between water discharge and sediment transport capacity could be possible when water discharge varies at a slower rate than subglacial channel size.
+In natural environments, this appears to rarely be the case, except where basal melt dominates.
+
+The subglacial environment's few observations hinder our ability to quantify processes such as
+the shape of subglacial channels and the response time of subglacial channels to water discharge variations.
+The poor constraints on subglacial sediment size and sorting processes make it difficult to link shear stress to sediment discharge capacity.
+Further quantifying these processes will help to better inform the response of sediment discharge capacity to water discharge forcing in subglacial environments.
+
+This study calls for the explicit consideration of rapidly evolving channel size when examining the relationship between sediment transport and hydro-climatic conditions from glacierized catchments, especially ones with high water discharge variability.
+
+\codedataavailability{
+  Code, with links to the data, can be found at \url{https://bitbucket.org/IanDelaney/xsection/src/master/}.
+  Data from Leverett glacier have been previously published in \citet{tedstone2013}.
+  Data from Fieschergletscher have been previously published in \citet{delaney2024}.
+  The code will be uploaded to a permanent repository, with FAIR principles pending acceptance of the paper.}
+
+\authorcontribution{
+  I.D. designed the study, developed, and implemented the model and experiments, and wrote the manuscript.
+  A.J.T. helped interpret findings from Leverett Glacier and guided experiment design.
+  M.A.W. provided data from Fieschergletscher and added Section \ref{sect:scaling}.
+  D.F. provided data from Fieschergletscher.
+  All authors provided key inputs in writing and editing the manuscript.} %% this section is mandatory
+
+\competinginterests{The authors declare no competing interests.} %% this section is mandatory even if you declare that no competing interests are present
+
+\begin{acknowledgements}
+  SNSF Project No. $\mathrm{PZ00P2}$\_$202024$ provided  funding for I. Delaney.
+  A. Tedstone acknowledges funding from the European Research Council, award $818994$ -- CASSANDRA.
+  G. King and M. J. Gevers provided useful comments on a previous version of this manuscript.
+  We benefited from fruitful discussions with J. Irving.
+  We thank two reviewers for constructive and critical reviews on a previous version of the manuscript.
+\end{acknowledgements}
+\newpage
+\appendix
+\section{Sensitivity of sediment transport in different drainage regimes}
 \label{sect:scaling}
@@
 % 
-Pipe flow-like conditions occur when an R-channel is subjected to rapid discharge variations such that the channel cannot adjust its size significantly on the time scale of the discharge variations.
+Pipe flow-like conditions occur when an R-channel is subjected to rapid discharge variations such that the channel cannot adjust its size.
 In this case, it is assumed that the cross-sectional area $S$ is fixed, and $\Psi$ adjusts to the specified $Q$.
@@
 
-\FloatBarrier
-\subsection{Increased variability in evolving subglacial channels occurs across a range of channel shapes, slopes,  and friction values}
-\label{sect:ensemble}
-
-The first numerical experiment aims to compare the variability between the subglacial and subaerial model outputs to a range of channel shapes, friction factors, and water discharge variability, as the subglacial channel's cross-sectional area is allowed to evolve.
-Following the findings in Section~\ref{sect:scaling}, variability in subglacial sediment discharge capacity could approach that of subaerial systems, should water discharge variations be sufficiently small.
-For simplicity, we only examine $w\tau$.
-
-Across the range of parameters examined (Section\,\ref{sect:imp}), variability in all model outputs remains higher in the subglacial system compared to the subaerial one for both \alpine{} and \icesheet{} (Figure~\ref{fig:multi_run}).
-In some cases, subglacial model outputs' variability is a magnitude larger than their subaerial counterparts.
-Results show that variability in both subglacial and subaerial outputs decreases with smoothing time longer than approximately $1$--$5$ days (Figure~\ref{fig:multi_run}).
-These smoothing timescales remove the diurnal variations in water discharge.
-
-\begin{figure}[hbt!]
-  \centering
-  \includegraphics[width=0.9\linewidth]{Fig3.pdf}
-  \caption{Standard deviation of detrended model outputs for different smoothing times with different subglacial and subaerial channel shapes and friction factors.
-    Shaded areas denote the range of standard deviations from the accepted $100$ parameter combinations (Section~\ref{sect:imp}).
-    Solid lines denote the mean value of standard deviations.
-    Markers show smoothing periods ($15$\,\unit{min}, $1$\,\unit{hr}, $6$\,\unit{hr} $12$\,\unit{hr}, $1$\,\unit{d}, $5$\,\unit{d}, $10$\,\unit{d}, and $15$\,\unit{d}).
-  }
-  \label{fig:multi_run}
-\end{figure}
-
-Smaller values of $\alpha$, or a channel cross-section closer to a slot canyon, result in greater variability in velocity and shear stress (Figure~\ref{fig:params}).
-In both cases, the variability of small values of $\alpha$ does not approach that of the subglacial channels.
-Yet, larger values of $\alpha$ can result in greater variability in width-integrated shear stress, as greater channel width changes increase the width-integrated shear stress.
-Additionally, steeper subaerial channel slopes result in greater variability in both shear stress and width-integrated shear stress but only approach the subglacial channel's variability in the \alpine case (Figure~\ref{fig:params}).
-We note that these parameter values span a commonly accepted range.
-Therefore, we do not anticipate a scenario where variability in the subaerial system would exceed the subglacial system with these two hydrologic forcings.
-
-Greater subglacial and subaerial friction factors $f_i$ and $f_p$ result in greater variability in the shear stress and width-integrate shear stress in both cases (Figure~\ref{fig:params}).
-This result is expected given Equations~\ref{eq:tau_g} and \ref{eq:vf}.
-However, more velocity variability is present in low  $f_i$ and $f_p$ values. 
-Low values of $f_i$ result in slower growth rates in subglacial channels (Equations~\ref{eq:dS_dt} and \ref{eq:dh}).
-As a result, water velocity, as opposed to channel growth, accommodates increases in water discharge, and velocity variability increases.
-Smaller values of channel factor $\beta$, creating low and broad channels, result in more variability in width-integrated shear stress.
-Here, the channel width can grow more quickly in response to water discharge increases as compared to a semi-circular channel with $\beta = \pi$ (Equation~ \ref{eq:dh}).
-
-\begin{figure}[hbt!]
-  \centering
-  \includegraphics[width=0.9\linewidth]{Fig4.pdf}
-  \caption{Parameter values compared to variability in model outputs.
-    Plots on the right (orange, light blue) correspond to the \alpine{} case. Plots on the left (brown, dark blue) correspond to the \icesheet case.
-    Outputs are shown using $15$ minute smoothing time. 
-  }
-  \label{fig:params}
-\end{figure}
-
-\FloatBarrier
-\subsection{Changing subglacial channel size drives different timing and variability in sediment transport capacity}
-
-The second numerical experiment aims to quantify the sources of increased variability in the subglacial model outputs as they exhibit different seasonal evolutions and peaks (Figure~\ref{fig:model_outs}).
-Variable relationships between model outputs and water discharge emerge for the subaerial and subglacial cases due to the hysteresis in channel size in the subglacial model (Equation~\ref{eq:dS_dt}).
-Because subaerial channels have no history dependence, each water discharge value in the subaerial channel produces a unique water velocity, shear stress, and width-integrated shear stress (Section~\ref{sect:sub_mode}).
-This characteristic results in a perfect rank correlation between the variables and water discharge (Figure~\ref{fig:Qw_vari}).
-An inconsistent relationship between subglacial model outputs and water discharge persists across the range of parameters examined in the ensemble runs (Figure S15).
-\begin{figure}[hbt!]
-  \centering
-  \includegraphics[width=0.9\linewidth]{Fig5.png}
-  \caption{Model outputs from simulations using the hydrographs in panels a and e for the scenarios \alpine{} (a-d) and \icesheet{} (e-h).
-    Black (gray) lines in a and e represent the subglacial channel size (water discharge).
-    Blue (orange) lines represent outputs from the subglacial (subaerial) channel.
-    Data are shown at $15$\unit{min} intervals.
-    Arrows show examples where variables peak in the subglacial channel before the subaerial one.
-    Insets in f-h show the peak melt event denoted by the shaded area in panels e--h, with an arbitrary y-axis.
-  }
-  \label{fig:model_outs}
-\end{figure}
-
-Conversely, the history dependence on channel size in subglacial channels means that model outputs occur across a large range of water discharges.
-For instance, in subglacial channels in \alpine{},  high water velocity values and shear stresses can occur from a low water discharge  ($\sim\,4$\,\unit{m}$^3$\,\unit{s}$^{-1}$) to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ (Figure~\ref{fig:Qw_vari}\,a).
-In \icesheet{}, water velocities close to the seasonal mean value can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
-The subglacial channel's evolving width can counteract some of these effects.
-Width-integrated shear stress generally increases with water discharge, with greater rank correlation compared to water velocity or shear stress (\alpine{}, Figure~\ref{fig:Qw_vari} a--c).
-Yet, even the width-integrated shear stress can vary substantially relative to water discharge.
-The highest values of width-integrated shear stress occur at water discharge values ranging from roughly $11$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ to over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
-The variability in width-integrated shear stress is less pronounced in the \icesheet{} scenario, where the hydrograph has less diurnal variability (Figure~\ref{fig:Qw_vari} c, f).
-This results from the discharge variations being in closer equilibrium with subglacial channels compared to \alpine{} (Sections~\ref{sect:scaling} and~\ref{sect:ensemble}).
-
-\begin{figure}[hbt!]
-  \centering
-  \includegraphics[width=0.9\linewidth]{Fig6.png}
-  \caption{
-    Relationship between water discharge and normalized velocity, shear stress, and width-integrated shear stress for \alpine{} (a-c) and \icesheet{} (d-f).
-    Variables on the x-axis have been normalized to mean values.
-    $R_g$ ($R_f$) shows the Spearman rank correlations for the subglacial (subaerial) outputs.
-    Plots are shown with discharge and model outputs at $15$\unit{min} intervals.
-  }
-  \label{fig:Qw_vari}
-\end{figure}
-
-Peaks in subaerial model outputs occur coincident with peaks in water discharge (Figure~\ref{fig:model_outs}).
-In the subglacial channel, peaks in model outputs generally occur when water discharge increases at the fastest rate, but before the maximum water discharge.
-Reduced water velocity takes place when water discharge stabilizes at its peak and channel growth continues.
-As a result, subglacial sediment transport capacity is greatest on the hydrograph's rising limb, relative to the falling limb.
-
-\FloatBarrier
-\section{Discussion}
-
-\subsection{Experiment limitations}
-
-Several aspects of the models and experiment design limit make comparing the model outputs to field observations difficult and make comparison of the two systems difficult.
-The lumped nature of the models means that they operate independently of the upstream drainage network.
-Additionally, sediment access is not considered here.
-In reality, processes such as sediment access are highly important in controlling sediment export in glacierized catchments \citep[e.g.][]{herman2015,vergara2022}.
-Furthermore, meltwater can be distributed to flow through several adjacent conduits impacting the sediment transport capacity in each \citep[e.g.][]{werder2013,hewitt2019,delaney2023}.
-The lumped models used here make the comparison more direct by isolating the relationship between water discharge and sediment transport capacity in subglacial and subaerial systems. 
-However, they neglect more complex, yet important, spatially distributed processes. a
-
-Comparing model outputs between the subaerial and subglacial case is difficult due to the different gradients in the two systems, with the subglacial one being much steeper \citep{alley1997}.
-The water velocities are similar in both channel types (Figure~\ref{fig:model_outs}).
-However, the shear stresses and width-integrated shear stresses are much smaller in subaerial channels.
-We note that in the ensemble runs in Section~\ref{sect:ensemble}, the parameters' range tested covers a likely range of viable shear stresses in both subglacial and subaerial channels.
-The greater sediment transport capacity in the subglacial channels implies that sediment grain size underneath subglacial channels is likely larger than subaerial counterparts.
-
-Width-integrated shear stress is examined in the two numerical experiments as it does not have a dependence on grain size, unlike sediment transport relationships \citep[][Section \ref{sect:scaling}]{meyer1948}.
-This makes comparison between the two systems simpler. 
-Preferential transport of smaller sediment clasts and the input of upstream sediment impact sediment transport capacity as grain size evolves \citep[e.g.][]{gomez1983}.
-These processes are only beginning to be evaluated underneath glaciers \citep{aitken2024}, but can be important in subaerial systems.
-Subglacial sorting processes could be an additional source of variability in sediment export from glaciers with respect to water discharge, especially for bedload transport.
-
-\subsection{Increased variability in  sediment transport capacity in subglacial systems}
-\label{sect:dis_qsc}
-
-The greater variations in shear stress in subglacial channels presented here may actually cause even greater variations in sediment transport capacity in subglacial channels than presented above  (Figure~\ref{fig:multi_run}).
-In sediment transport capacity relationships, such as in \citet{meyer1948} or \citet{engelund1967}, shear stress is scaled to the power of $\frac{3}{2}$ or $\frac{5}{2}$, respectively (Section~\ref{sect:scaling}).
-The exponent greater than $1$ magnifies sediment discharge variability beyond the variable sediment transport parameters described above (Figure~\ref{fig:multi_run}; Table~\ref{tab:Qs}).
-
-Greater variations in subglacial sediment transport capacity could cause a supply-limited regime at many glaciers \citep{alley1997}.
-In subglacial systems, sediment's critical shear stress, or threshold at which sediment mobilization occurs, can be reached more frequently and across many water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
-This result suggests sediment export here is especially sensitive to observed changes in water discharge variability \citep{lane2019b}.
-Sediment exhaustion through repeatedly crossing the mobilization threshold may explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \citep{overeem2017}.
-The same process could also result in sediment discharge's strong dependence on sediment production from sliding in many cases \citep{herman2015,koppes2015}.
-
-While transport-limited states at glaciers are likely rare \citep[e.g.][]{alley1997}, abundant sediment can persist underneath glaciers \citep[e.g.][]{walter2014,stevens2022,delaney2022}.
-As a result, in some cases sediment export can be hydrologically controlled and efficient \citep[e.g.][]{swift2005,andresen2024}.
-At these glaciers, the great variability in subglacial sediment transport capacity may make it difficult to link sediment discharge to hydrology, especially when peak events occur \citep{cowan1988,delaney2018,lu2022}.
-As a result, the sediment transport capacity response to individual extreme precipitation or melt events could be difficult to establish.
-The results here suggest that different channel sizes with the same hydrology forcing can result in very different sediment transport capacities.
-
-Catchments with reduced water discharge variability may experience yet less variability in sediment transport capacity, shown by the \alpine{} and \icesheet{} model outputs.
-Indeed, the more decoupled and sporadic relationship between model outputs and water discharge in \alpine{} results from the relatively larger discharge variations on sub-daily to weekly timescales (Figure~\ref{fig:Qw_vari}).
-This occurs as the subglacial channel's size evolves slowly compared to the variations in water discharge.
-High water discharge variability in \alpine{} may cause shear stress to approach $Q^{2}$in assuming pipe-flow conditions where the water discharge varies largely compared to channel size \citep[Section \ref{sect:scaling}; c.f.][]{alley1997}.
-Conversely, the reduced variability in water discharge in \icesheet{} may result in the subglacial channel's size being closer to equilibrium from the smaller variations in water discharge.
-In this case, the exponent on water discharge for shear stress is likely substantially less than $w\tau \propto Q^2$, but greater than $w \tau \propto Q^{\frac{4}{5}}$ that occurs in a steady-state R-channel (see Section \ref{sect:scaling}).
-As a result, there is a stronger relationship between subglacial model outputs and water discharge in \icesheet{} (Figure~\ref{fig:Qw_vari}) and less variability model outputs in the \icesheet case (Figure~\ref{fig:multi_run}).
-
-\subsection{Interpreting sediment transport records from glacierized catchments with respect to water discharge}
-
-A sporadic relationship between subglacial sediment transport capacities and water discharge occurs due to hysteresis in subglacial channel size (Figure~\ref{fig:Qw_vari}).
-This characteristic limits the use of water discharge as an indicator of sediment discharge capacity in these systems.
-As a result, characteristics such as bankfull or effective water discharge that link geomorphic work to hydro-climatic conditions could have limited meaning in evaluating subglacial sediment transport \citep{wolman1960,lenzi2006}.
-A glacier's sediment transport capacity is impacted by the ice thickness controlling the channel closure rate and the glacier's surface slope, in addition to water discharge and sediment size \citep[Figure~\ref{fig:multi_run}, Section~\ref{sect:sub_mode}; ][]{rothlisberger1972,gimbert2016,stevens2022,walder1994}.
-This multitude of processes lies in contrast to many subaerial channels, where transport capacity typically responds to water discharge, sediment size, channel shape,  and hydraulic gradient.
-The latter of these parameters can remain relatively stable over the years or longer \citep[Section~\ref{sect:fluv}; e.g.][]{tucker1997}.
-Strong correlations between water discharge and sediment export in glacier systems could indicate other processes, such as increased sediment access \citep{zhang2022}.
-Furthermore, the observed correlation between water discharge and sediment discharge far downstream likely represents subaerial transport processes, especially with respect to bedload \citep{mancini2023}.
-
-The co-varying relationship between sediment discharge capacity and water discharge in subaerial conditions could be noticeable $\sim20$\,\unit{km} downstream of the Leverett site.
-A strong correlation persists between sediment plume size and the Watson River's water discharge into the Kangerlussuaq fjord \citep{chu2009,mcgrath2010}.
-In contrast, in marine-terminating glacier catchments, a less consistent relationship may occur between water discharge or melt extent and sediment plume size \citep{chu2012,tedstone2012}.
-Here, ocean water pressurizes subglacial channels at the ice front \citep[e.g.][]{how2017}, so the observed reduced correlation could result from the inconsistent relationship between subglacial sediment transport capacity and water discharge (Figure~\ref{fig:Qw_vari}).
-
-Water discharge measurements at less than daily timescales could severely limit subglacial sediment transport models in capturing specific events.
-The decreased model output variability beyond $1$--$5$ days of smoothing could result in water discharge being in equilibrium with the subglacial channel when they are in fact not (Figure~\ref{fig:multi_run}).
-Over this period, water discharge variations could lead to a stronger relationship with sediment discharge capacity (Section ~\ref{sect:scaling}).
-Yet, such a strong relationship would not represent the impact of shorter-term fluctuations in water discharge.
-Note, that if basal melt largely dominates, then water discharge would evolve slowly with respect to the subglacial channel size.
-
-Clockwise hysteresis loops between sediment concentration and water discharge suggest that sediment availability is reduced over the event scale \citep[][]{williams1989}.
-These loops are often observed in glacierized catchments and used to suggest that access to subglacial sediment has been limited \citep[e.g.][]{collins1979,willis1996,richards2003,stott2007,delaney2018}.
-Results here suggest that clockwise hysteresis could also be expected in transport-limited regimes in many subglacial environments.
-Here, the smaller channel size on the rising limb would result in greater sediment transport capacity and sediment uptakes.
-A larger channel size would reduce sediment transport capacity for an equivalent water discharge on the falling limb, creating clockwise hysteresis.
-
-\conclusions
-
-The sediment transport capacity of a channel is driven by its width and the shear stress exerted by the flowing water, which is proportional to flow velocity squared.
-Subaerial channels alter both their channel width and water velocity immediately in response to changing water discharge.
-In contrast, pressurized subglacial channels largely accommodate rapidly changing water discharge by altering water velocity.
-Their size responds over several days to changes in discharge.
-Thus, sediment transport capacity is more sensitive to changes in water discharge in subglacial channels compared to subaerial ones.
-
-Concerning the manuscript's first objective,
-results demonstrate that sediment transport capacity variability is high in subglacial channels with evolving size across a range of channel shapes and friction factors.
-Increased variability occurs even in environments where the water discharge has relatively small diurnal variations in the \icesheet case.
-Yet, elevated variability is especially strong with highly variable water discharge in the \alpine case.
-Increased noise in sediment transport capacity in evaluating hydro-climatic signals from sediment records in glacierized catchments needs further evaluation.
-Furthermore, this process may lead to sediment exhaustion by subglacial water repeatedly crossing the mobilization threshold across a range of water discharges.
-
-The manuscript's second objective is to establish if water discharge varies with subglacial sediment transport capacity.
-Results here suggest that, even in a transport-limited subglacial system, in most systems, an incoherent relationship between water and subglacial sediment discharge could be expected.
-This presents a challenge in linking hydro-climatic conditions or events to sediment export from glaciers.
-Due to evolving subglacial channel size, the timing of peak sediment discharge capacity in subglacial channels occurs before peak water discharge during an event, whereas it is coincident in subaerial ones.
-Covariance between water discharge and sediment transport capacity could be possible when water discharge varies at a slower rate than subglacial channel size.
-In natural environments, this appears to often not be the case, except where basal melt dominates.
-
-The subglacial environment's few observations hindered our ability to quantify processes such as
-the shape of subglacial channels and the response time of subglacial channels to water discharge variations.
-The poor constraints on subglacial sediment size and sorting processes make it difficult to link shear stress to sediment discharge capacity.
-Further quantifying these processes will help to better inform the response of sediment discharge capacity to water discharge forcing in subglacial environments.
-
-This study calls for careful analysis when examining the relationship between sediment transport and hydro-climatic conditions from glacierized catchments, especially ones with high water discharge variability.
-
-\codedataavailability{
-  Code, with links to the data, can be found at \url{https://bitbucket.org/IanDelaney/xsection/src/master/}.
-  Data from Leverett glacier have been previously published in \citet{tedstone2013}.
-  Data from Fieschergletscher have been previously published in \citet{delaney2024}.
-  The code will be uploaded to a permanent repository, with FAIR principles pending acceptance of the paper.}
-
-\authorcontribution{
-  I.D. designed the study, developed, and implemented the model and experiments, and wrote the manuscript.
-  A.J.T. helped interpret findings from Leverett Glacier and guided experiment design.
-  M.A.W. provided data from Fieschergletscher and added Section \ref{sect:scaling}.
-  D.F. provided data from Fieschergletscher.
-  All authors provided key inputs in writing and editing the manuscript.} %% this section is mandatory
-
-\competinginterests{The authors declare no competing interests.} %% this section is mandatory even if you declare that no competing interests are present
-
-\begin{acknowledgements}
-  SNSF Project No. $\mathrm{PZ00P2}$\_$202024$ provided  funding for I. Delaney.
-  A. Tedstone acknowledges funding from the European Research Council, award $818994$ -- CASSANDRA.
-  G. King and M. J. Gevers provided useful comments on a previous version of this manuscript.
-  We benefited from fruitful discussions with J. Irving.
-  We thank two reviewers for constructive and critical reviews on a previous version of the manuscript.
-\end{acknowledgements}
-\newpage
-
-% \appendix
 % \appendixtables   %% needs to be added in front of appendix tables

### 2024-07-08 Delaneyetal2024.tex
working here... might be conflicts of OL

@@
 \end{linenomath*}
-\noindent where $t$ is time, $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (values in Table~\ref{table:vpm}), $g$ is the acceleration due to gravity, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change over $l$, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the mean ice overburden pressure expressed in meter water equivalent ($\rho_w$ is density of water; $\rho_i$ is density of ice), and $n$ is Glen's n \citep{glen1955} (usually $n=3$).
+\noindent where $t$ is time, $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (values in Table~\ref{table:vpm}), $g$ is the acceleration due to gravity, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change over $l$, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the mean ice overburden pressure expressed in meter water equivalent ($\rho_w$ is density of water; $\rho_i$ is density of ice), and $n$ is Glen's n \citep[usually $n=3$][]{glen1955}.
 The first term on the equation's right side represents the channel opening by frictional heating, while the following term represents channel closure from ice deformation.
@@
 \section{Results}
-We first present formulations of sediment transport capacity in several channel types to build upon the analysis in \citet{alley1997}.
-We then apply the models above (Section~\ref{sect:meth}) in two experiments by examining three model outputs.
-Water velocity and shear stress represent the potential for sediment mobilization, and width-integrated shear stress represents a proxy for the total sediment transport capacity across the channel bed (Figure~\ref{fig:model_outs}).
+%We first present formulations of sediment transport capacity in several channel types to build upon the analysis in \citet{alley1997}.
+%We then apply the models above (Section~\ref{sect:meth}) in two experiments by examining three model outputs.
+%Water velocity and shear stress represent the potential for sediment mobilization, and width-integrated shear stress represents a proxy for the total sediment transport capacity across the channel bed (Figure~\ref{fig:model_outs}).
 
@@
 
+\begin{figure}[hbt!]
+  \centering
+  \includegraphics[width=\linewidth]{Fig7.pdf}
+  \caption{Response of sediment transport capacity (orange line) and channel size (bottom row) to water discharge variations under different channel conditions.
+    a) Pipe flow with fixed channel size.
+    b) R-channel described in Section \ref{sect:sub_mode}.
+    c) Subaerial channel described in Section~\ref{sect:fluv}.
+    d) Steady state R-channel described in Section~\ref{sect:scaling}.
+    Note axes not to scale.
+    Gyres on bottom panel represent shear stress on channel walls.
+    $t_1$ ($t_2$) represents low (high) water discharge. 
+  }
+  \label{fig:chan_types}
+\end{figure}
+
 Catchments with reduced water discharge variability may experience yet less variability in sediment transport capacity, shown by the \alpine{} and \icesheet{} model outputs.
 Indeed, the more decoupled and sporadic relationship between model outputs and water discharge in \alpine{} results from the relatively larger discharge variations on sub-daily to weekly timescales (Figure~\ref{fig:Qw_vari}).
-This occurs as the subglacial channel's size evolves slowly compared to the variations in water discharge.
-High water discharge variability in \alpine{} may cause shear stress to approach $Q^{2}$in assuming pipe-flow conditions where the water discharge varies largely compared to channel size \citep[Section \ref{sect:scaling}; c.f.][]{alley1997}.
-Conversely, the reduced relative variability in water discharge in \icesheet{} and stronger correlation between width-integrated shear stress may result in the subglacial channel's size being closer to equilibrium from the smaller variations in water discharge.
+This occurs as the subglacial channel's size evolves slowly compared to the variations in water discharge (Figure~\ref{fig:chan_types} b).
+High water discharge variability in \alpine{} may cause shear stress to approach $Q^{2}$ in assuming pipe-flow conditions where the water discharge varies largely compared to channel size \citep[Figure~\ref{fig:chan_types}; Section \ref{sect:scaling}; c.f.][]{alley1997}.
+Conversely, the reduced relative variability in water discharge in \icesheet{} and stronger correlation between width-integrated shear stress may result in the subglacial channel's size being closer to equilibrium from the smaller variations in water discharge (Figure~\ref{fig:chan_types} d).
 In this case, the exponent on water discharge for shear stress is likely substantially less than $w\tau \propto Q^2$, but greater than $w \tau \propto Q^{\frac{4}{5}}$ that occurs in a steady-state R-channel (see Section \ref{sect:scaling}).
@@
 The decreased model output variability beyond $1$--$5$ days of smoothing could result in water discharge being in equilibrium with the subglacial channel when they are in fact not (Figure~\ref{fig:multi_run}).
-Over this period, water discharge variations could lead to a stronger relationship with sediment discharge capacity (Section ~\ref{sect:scaling}).
-Yet, such a strong relationship would not represent the impact of shorter-term fluctuations in water discharge.
-Note, that if basal melt largely dominates, then water discharge would evolve slowly with respect to the subglacial channel size.
+Over this period, water discharge variations could lead to a stronger relationship with sediment discharge capacity (Figure~\ref{fig:chan_types} d and Section ~\ref{sect:scaling}).
+Particularly, if water discharge evolves slowly with respect to the subglacial channel size, then it will covary with sediment transport capacity (Figure~\ref{fig:chan_types} d).
+Such a strong relationship would not represent the impact of actual shorter-term fluctuations in water discharge.
+Note, in this case, that if  pipe flow is assumed, with pressurized flow, but no channel evolution, then  sediment discharge capacity would covary with water discharge (Figure~\ref{fig:chan_types} a).
 
@@
 Results here suggest that clockwise hysteresis could also be expected in transport-limited regimes in many subglacial environments.
-Here, the smaller channel size on the rising limb would result in greater sediment transport capacity and sediment uptakes.
+Here, the smaller channel size on the rising limb would result in greater sediment transport capacity and sediment uptakes (Figure~\ref{fig:chan_types} b).
 A larger channel size would reduce sediment transport capacity for an equivalent water discharge on the falling limb, creating clockwise hysteresis.
@@
 Furthermore, meltwater can be distributed to flow through several adjacent conduits impacting the sediment transport capacity in each \citep[e.g.][]{werder2013,hewitt2019,delaney2023}.
-The lumped models used here make the comparison isolate the relationship between water discharge and sediment transport capacity in subglacial and subaerial systems. 
+The lumped models used here isolate the relationship between water discharge and sediment transport capacity in subglacial and subaerial systems. 
 However, they neglect more complex, yet important spatially distributed processes.
@@
 Comparing model outputs between the subaerial and subglacial case is difficult due to the different hydraulic gradients in the two systems, with the subglacial one generally being much steeper \citep{alley1997}.
-The water velocities are similar in both channel types (Figure~\ref{fig:model_outs}).
-However, the shear stresses and width-integrated shear stresses across the channel bed that control sediment transport capacity are much smaller in subaerial channels.
-We note that in the ensemble runs in Section~\ref{sect:ensemble}, the parameters' range tested covers a likely range of viable shear stresses in both subglacial and subaerial channels.
+This results the shear stresses and width-integrated shear stresses across the channel bed that control sediment transport capacity are much smaller in subaerial channels (Figure~\ref{fig:model_outs}).
+The parameters' range tested in Section~\ref{sect:ensemble} covers a likely span of viable shear stresses in both subglacial and subaerial channels.
 The greater sediment transport capacity in the subglacial channels implies that sediment grain size underneath subglacial channels is likely larger than subaerial counterparts.
+Note however, that the water velocities are similar in both channel types.
 
@@
     Pipe & \(f \, Q^2 \, S^{-1}\) & \(f^{3/2}\, Q^3 \, S^{-5/2}\) & \(f^{5/2}\, Q^5\, S^{-9/2}\) & \(f^{3/2} \, Q^{9/2} \, S^{-14/3}\)\\
+  \end{tabular}
+\end{table}
+
+\begin{table}
+  \caption{As Table~3 but with the fractional exponents stated as rounded decimal numbers.
+For the subaerial regime the exponents are for displayed for the likely width-exponent $\alpha=1$, as well as its end members 0 (slot canyon) and 1 (only width increases and not depth).
+  }
+  \small
+  \label{tab:Qs2}
+  \begin{tabular}{lllll}
+    & Width \(\times \,\, \tau\) & MPM & EH & Bagnold\\
+    & \(w\, \tau\) & \(Q_s \propto w\, \tau^{3/2}\) & \(Q_s \propto w\, \tau^{5/2}\) & \(Q_s \propto w^{-1/2}\, \Omega^{3/2} H^{-2/3}\)\\
+    \hline
+    Subaerial (\(\alpha=0\)) & \(f^{0.7}\, Q^{0.7}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{1.7} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.1} \, \Psi^{1.7}\)\\
+    Subaerial (\(\alpha=0.3\)) & \(f^{0.7}\, Q^{0.8}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{1.3} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\
+    Subaerial (\(\alpha=0.5\)) & \(f^{0.7}\, Q^{0.8}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{1.3} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\
+    Subaerial (\(\alpha=1\)) & \(f^{0.7}\, Q^{1.0}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{0.7} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\[3pt]
+    R-channel & \(f^{0.4}\, Q^{0.8} \, \Psi^{0.6}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.7}\, Q^{1.4}\, \Psi^{1.8}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\
+    Pipe & \(f \,\quad Q^{2\phantom{.0}} \, S^{-1}\) & \(f^{1.5}\, Q^3 \, S^{-2.5}\) & \(f^{2.5}\, Q^{5\phantom{.0}}\, S^{-4.5}\) & \(f^{1.5} \,\,\,\,\, Q^{4.5} \, S^{-4.7}\)\\
   \end{tabular}

### 2024-07-09 Delaneyetal2024.tex
Overleaf work

--- 2024-07-08 Delaneyetal2024.tex
@@
 \section{Results}
-We first present formulations of sediment transport capacity in several channel types to build upon the analysis in \citet{alley1997}.
-We then apply the models above (Section~\ref{sect:meth}) in two experiments by examining three model outputs.
-Water velocity and shear stress represent the potential for sediment mobilization, and width-integrated shear stress represents a proxy for the total sediment transport capacity across the channel bed (Figure~\ref{fig:model_outs}).
+%We then apply the models above (Section~\ref{sect:meth}) in two experiments by examining three model outputs.
+%Water velocity and shear stress represent the potential for sediment mobilization, and width-integrated shear stress represents a proxy for the total sediment transport capacity across the channel bed (Figure~\ref{fig:model_outs}).
 
@@
 The first numerical experiment aims to compare the variability between the subglacial and subaerial model outputs to a range of channel shapes, friction factors, and water discharge variability, as the subglacial channel's cross-sectional area is allowed to evolve.
-Following the findings in Section~\ref{sect:scaling}, variability in subglacial sediment discharge capacity could approach that of subaerial systems, should water discharge variations be sufficiently small.
-For simplicity, we only examine $w\tau$.
+%Following the findings in Section~\ref{sect:scaling}, variability in subglacial sediment discharge capacity could approach that of subaerial systems, should water discharge variations be sufficiently small.
 
@@
 The greater variations in shear stress in subglacial channels presented here may actually cause even greater variations in sediment transport capacity in subglacial channels than presented above  (Figure~\ref{fig:multi_run}).
-In sediment transport capacity relationships, such as in \citet{meyer1948} or \citet{engelund1967}, shear stress is scaled to the power of $\frac{3}{2}$ or $\frac{5}{2}$, respectively (Section~\ref{sect:scaling}).
+In sediment transport capacity relationships, such as in \citet{meyer1948} or \citet{engelund1967}, shear stress is scaled to the power of $\frac{3}{2}$ or $\frac{5}{2}$, respectively (e.g.
+Section~\ref{sect:scaling}).
 The exponent greater than $1$ magnifies sediment discharge variability beyond the variable sediment transport parameters described above (Figure~\ref{fig:multi_run}; Table~\ref{tab:Qs}).
@@
 
-While transport-limited states at glaciers are likely rare \citep[e.g.][]{alley1997}, abundant sediment can persist underneath glaciers \citep[e.g.][]{walter2014,stevens2022,delaney2022}.
-As a result, in some cases sediment export can be hydrologically controlled and efficient \citep[e.g.][]{swift2005,andresen2024}.
+While transport-limited states at glaciers are likely rare \citep[e.g.][]{alley1997}, abundant sediment could persist underneath some glaciers \citep[e.g.][]{walter2014,stevens2022,delaney2022}.
+As a result, sediment export can be hydrologically controlled and efficient \citep[e.g.][]{swift2005,andresen2024}.
 At these glaciers, the great variability in subglacial sediment transport capacity may make it difficult to link sediment discharge to hydrology, especially when peak events occur \citep{cowan1988,delaney2018,lu2022}.
@@
 This occurs as the subglacial channel's size evolves slowly compared to the variations in water discharge.
-High water discharge variability in \alpine{} may cause shear stress to approach $Q^{2}$in assuming pipe-flow conditions where the water discharge varies largely compared to channel size \citep[Section \ref{sect:scaling}; c.f.][]{alley1997}.
+High water discharge variability in \alpine{} may cause shear stress to approach $Q^{2}$ in assuming pipe-flow conditions where the water discharge varies largely compared to channel size \citep[Section \ref{sect:scaling}; c.f.][]{alley1997}.
 Conversely, the reduced relative variability in water discharge in \icesheet{} and stronger correlation between width-integrated shear stress may result in the subglacial channel's size being closer to equilibrium from the smaller variations in water discharge.
@@
 
-\competinginterests{The authors declare no competing interests.} %% this section is mandatory even if you declare that no competing interests are present
+\competinginterests{The authors declare no competing interests.} %% This section is mandatory even if you declare that no competing interests are present
 
@@
 R-channels which do not have time to adjust their size to discharge conditions).
-The evolving subglacial channel geometry presented in Section \ref{sect:sub_mode} is not considered.
-Here we show the sediment transport capacity scaling with a given water discharge and hydraulic gradient for three different transport formulas: Meyer-Peter M\"uller  \citep[MPM; ][]{meyer1948}, Engelund and Hansen \citep[EH; ][]{engelund1967}, and Bagnold \citep{bagnold1980}; additionally width-integrated shear stress is assessed, as used as proxy for sediment transport throughout the manuscript.
+Here we show the sediment transport capacity scaling with a given water discharge and hydraulic gradient for three different sediment transport capacity formulas: Meyer-Peter M\"uller  \citep[MPM; ][]{meyer1948}, Engelund and Hansen \citep[EH; ][]{engelund1967}, and Bagnold \citep{bagnold1980}; additionally width-integrated shear stress is assessed, as used as proxy for sediment transport throughout the manuscript.
+
+Sediment discharge is given by the MPM, EH, and Bagnold formulations as
+\begin{equation}
+  Q_s \propto w\, \tau^{3/2}, \quad Q_s \propto w\, \tau^{5/2}, \quad Q_s \propto w \left(\frac{\Omega}{w}\right)^{3/2} H^{-2/3},
+\end{equation}
+respectively, for conditions well above the transport onset threshold.
 
@@
 
-Sediment discharge is given by the MPM, EH, and Bagnold formulations as
-\begin{equation}
-  Q_s \propto w\, \tau^{3/2}, \quad Q_s \propto w\, \tau^{5/2}, \quad Q_s \propto w \left(\frac{\Omega}{w}\right)^{3/2} H^{-2/3},
-\end{equation}
-respectively, for conditions well above the transport onset threshold.
-
 As in Section~\ref{sect:fluv}, the subaerial channel is assumed to have a width $w$ much greater than its depth $H$, such that $D_h\approx 4h$, and to have a constant head gradient ($\Psi$) given by the topography.
@@
 Then, using equations~\eqref{eq:tau-omega}, the shear stress and stream power can be calculated.
-These results are summarised in Table~\ref{tab:eqs1}.
+These results of are summarised in Table~\ref{tab:eqs1}.
 Now width integrated shear stress and sediment transport can be calculated for all regimes and all transport formulas.
@@
     R-channel & \(\Psi\) & \(S\, \propto f^{2/5}\, Q^{4/5} \, \Psi^{-2/5}\) & \(D_h\propto w \propto H \propto S^{1/2}\) & \(f^{1/5} Q^{2/5} \, \Psi^{4/5}\) & \(Q\, \Psi\)\\
-    Pipe & \(S\) & \(\Psi \propto f \, Q^2\, S^{-5/2}\) & \(D_h\propto w \propto H \propto S^{1/2}\) & \(fx Q^2 S^{-2}\) & \(f\, Q^3 S^{-5/2}\)\\
+    Pipe & \(S\) & \(\Psi \propto f \, Q^2\, S^{-5/2}\) & \(D_h\propto w \propto H \propto S^{1/2}\) & \(f Q^2 S^{-2}\) & \(f\, Q^3 S^{-5/2}\)\\
   \end{tabular}
@@
 
-The sediment discharge in the steady state R-channel scales very similarly to the subaerial channel case for all relations and virtually identically for the $\alpha=0.5$ case.
+The sediment discharge capacity in the steady state R-channel scales very similarly to the subaerial channel case for all relations and virtually identically for the $\alpha=0.5$ case.
 However, the head gradients $\Psi$ are likely higher for comparable $Q$ in an ice-sheet marginal or alpine glacier setting than in a subaerial channel \citep{alley1997}.
@@
 
-Furthermore, R-channels will rarely operate in a steady state as variations in water discharge.
-The timescale of diurnal melt or severe rain or melt is too short for the channel to reach a steady state.
-In these cases with high water discharge variability, channels operate more like a pipe of fixed cross-section \citep[e.g.][]{gimbert2016}.
-Table~\ref{tab:Qs} shows that sediment transport in pipe-flow scales much more severely with discharge; the exponent on $Q$ being between $3$ and $5$, compared to the other two regimes when that exponent is at most $1.7$ (Table~\ref{tab:Qs}).
-Thus fluctuations of discharge on short timescales (on the order of a day) have the potential to cause conditions with very high sediment transport capacities.
-These sediment transport capacities variations are of far higher magnitude than those of subaerial channels \citep{alley1997}.
+%Furthermore, R-channels will rarely operate in a steady state as variations in water discharge.
+%The timescale of diurnal melt or severe rain or melt is too short for the channel to reach a steady state.
+%In these cases with high water discharge variability, channels operate more like a pipe of fixed cross-section \citep[e.g.][]{gimbert2016}.
+%Table~\ref{tab:Qs} shows that sediment transport in pipe-flow scales much more severely with discharge; the exponent on $Q$ being between $3$ and $5$, compared to the other two regimes when that exponent is at most $1.7$ (Table~\ref{tab:Qs}).
+%Thus fluctuations of discharge on short timescales (on the order of a day) have the potential to cause conditions with very high sediment transport capacities.
+%These sediment transport capacities variations are of far higher magnitude than those of subaerial channels \citep{alley1997}.
 

### 2024-07-09 Delaneyetal2024.tex
small changes

--- 2024-07-08 Delaneyetal2024.tex
@@
 At these glaciers, the great variability in subglacial sediment transport capacity may make it difficult to link sediment discharge to hydrology, especially when peak events occur \citep{cowan1988,delaney2018,lu2022}.
-As a result, the sediment transport capacity response to individual extreme precipitation or melt events could be difficult to establish.
-The results here suggest that different channel sizes with the same hydrology forcing can result in very different sediment transport capacities.
+The results here suggest that different channel sizes with the same hydrology forcing can result in very different sediment transport capacities, making it challenging to establish the effects of individual extreme precipitation or melt events.
 
@@
   \caption{Response of sediment transport capacity (orange line) and channel size (bottom row) to water discharge variations under different channel conditions.
-    a) Pipe flow with fixed channel size.
+    a) Pipe flow with fixed channel size described in Section~\ref{sect:scaling}.
     b) R-channel described in Section \ref{sect:sub_mode}.

### 2024-07-09 Delaneyetal2024.tex

--- 2024-07-09 Delaneyetal2024.tex
@@
   By contrast, subglacial channels grow at a slower rate than most water discharge variations, causing a variable relationship between channel shape, water velocity, and discharge.
-  Using numerical models applied to two observed hydrographs and formulations, we evaluate the differences in sediment transport capacity variability between subaerial and subglacial channels.
-  We also establish the conditions under which subglacial sediment transport capacity could co-vary with water discharge.
-  The formulations of hydraulics and numerical models confirm that subglacial sediment transport is highly non-linear with respect to water discharge.
+  Using numerical models applied to two observed hydrographs, we evaluate the differences in sediment transport capacity variability between subaerial and subglacial channels.
+  The numerical experiments confirm that subglacial sediment transport is highly non-linear with respect to water discharge.
   Yet, subglacial sediment transport capacity can vary less if water discharge variations are closer to equilibrium with subglacial channel size.
@@
   This implies that hysteresis in channel size causes a variable relationship between sediment and water discharge in a transport-limited subglacial system.
-  The implications are discussed for sediment discharge from glaciers with different hydro-climatic forcings and assumptions of channel shape.
+  The implications are discussed for sediment discharge from glaciers with different hydro-climatic forcings.
+  We also evaluate the conditions under which subglacial sediment transport capacity could co-vary with water discharge.
   These findings can also improve interpretations of sediment discharge records in glacierized catchments.
@@
 This manuscript has two objectives: 1) to evaluate the variability of subglacial sediment transport capacity with evolving subglacial channel size and different hydrographs and 2) to establish whether sub-seasonal water discharge can co-vary with sediment transport capacity in subglacial systems.
-First, to demonstrate subglacial sediment transport capacity's highly variable nature,  we present algebraic formulations of the sediment transport capacity response to water discharge in subaerial, pipe-flow, and equilibrium R-channel conditions.
-These formulations extend the relationships presented in \citet{alley1997}.
-We then use numerical models to examine sediment transport capacity in both subglacial channels that evolve in size and subaerial channels that remain fixed.
+We use numerical models to examine sediment transport capacity in both subglacial channels that evolve in size and subaerial channels that remain fixed.
 They are applied to proglacial hydrological records from an Alpine glacier in Switzerland (Fieschergletscher) and a land-terminating glacier in Greenland (Leverett Glacier).
@@
     Shear stress (subglacial, subaerial) & $\tau$, ($\tau_g,\,\tau_f$) && $\mathrm{Pa \, m^{-2}}$ \\
-    Stream power & $\Omega$ && $\mathrm{ kg \, m\, s^{-3}}$ \\
+x    Stream power & $\Omega$ && $\mathrm{ kg \, m\, s^{-3}}$ \\
 
@@
 \section{Results}
-%We then apply the models above (Section~\ref{sect:meth}) in two experiments by examining three model outputs.
-%Water velocity and shear stress represent the potential for sediment mobilization, and width-integrated shear stress represents a proxy for the total sediment transport capacity across the channel bed (Figure~\ref{fig:model_outs}).
 
@@
 At these glaciers, the great variability in subglacial sediment transport capacity may make it difficult to link sediment discharge to hydrology, especially when peak events occur \citep{cowan1988,delaney2018,lu2022}.
-As a result, the sediment transport capacity response to individual extreme precipitation or melt events could be difficult to establish.
-The results here suggest that different channel sizes with the same hydrology forcing can result in very different sediment transport capacities.
+The results here suggest that different channel sizes with the same hydrology forcing can result in very different sediment transport capacities, making it challenging to establish the effects of individual extreme precipitation or melt events.
+
+\begin{figure}[hbt!]
+  \centering
+  \includegraphics[width=0.7\linewidth]{Fig7.pdf}
+  \caption{Response of sediment transport capacity (orange line) and channel size (bottom row) to a water discharge event under different channel conditions.
+    a) Pipe flow with fixed channel size described in Section~\ref{sect:scaling}.
+    b) R-channel described in Section \ref{sect:sub_mode}.
+    c) Subaerial channel described in Section~\ref{sect:fluv}.
+    d) Steady state R-channel (channel size evolves in equilibrium with water discharge) described in Section~\ref{sect:scaling}.
+    Note axes not to scale.
+    Gyres on bottom panels represent shear stress on channel walls.
+    $t_1$ ($t_2$) represents low (high) water discharge. 
+  }
+  \label{fig:chan_types}
+\end{figure}
 
@@
 Indeed, the more decoupled and sporadic relationship between model outputs and water discharge in \alpine{} results from the relatively larger discharge variations on sub-daily to weekly timescales (Figure~\ref{fig:Qw_vari}).
-This occurs as the subglacial channel's size evolves slowly compared to the variations in water discharge.
-High water discharge variability in \alpine{} may cause shear stress to approach $Q^{2}$ in assuming pipe-flow conditions where the water discharge varies largely compared to channel size \citep[Section \ref{sect:scaling}; c.f.][]{alley1997}.
-Conversely, the reduced relative variability in water discharge in \icesheet{} and stronger correlation between width-integrated shear stress may result in the subglacial channel's size being closer to equilibrium from the smaller variations in water discharge.
+This occurs as the subglacial channel's size evolves slowly compared to the variations in water discharge (Figure~\ref{fig:chan_types} b).
+High water discharge variability in \alpine{} may cause shear stress to approach $Q^{2}$ in assuming pipe-flow conditions where the water discharge varies largely compared to channel size \citep[Figure~\ref{fig:chan_types}; Section \ref{sect:scaling}; c.f.][]{alley1997}.
+Conversely, the reduced relative variability in water discharge in \icesheet{} and stronger correlation between width-integrated shear stress may result in the subglacial channel's size being closer to equilibrium from the smaller variations in water discharge (Figure~\ref{fig:chan_types} d).
 In this case, the exponent on water discharge for shear stress is likely substantially less than $w\tau \propto Q^2$, but greater than $w \tau \propto Q^{\frac{4}{5}}$ that occurs in a steady-state R-channel (see Section \ref{sect:scaling}).
@@
 The co-varying relationship between sediment discharge capacity and water discharge in subaerial conditions could be noticeable $\sim20$\,\unit{km} downstream of the Leverett site.
-A strong correlation persists between sediment plume size and the Watson River's water discharge into the Kangerlussuaq fjord \citep{chu2009,mcgrath2010}.
+A strong correlation persists between sediment plume size and the Watson River's water discharge into the Kangerlussuaq fjord \citep[Figure~\ref{fig:chan_types}\,c][]{chu2009,mcgrath2010}.
 In contrast, in marine-terminating glacier catchments, a less consistent relationship may occur between water discharge or melt extent and sediment plume size \citep{chu2012,tedstone2012}.
-Here, ocean water pressurizes subglacial channels at the ice front \citep[e.g.][]{how2017}, so the observed reduced correlation could result from the inconsistent relationship between subglacial sediment transport capacity and water discharge (Figure~\ref{fig:Qw_vari}).
+Here, ocean water pressurizes subglacial channels at the ice front \citep[e.g.][]{how2017}, so the observed reduced correlation could result from the inconsistent relationship between subglacial sediment transport capacity and water discharge (Figures~\ref{fig:Qw_vari} and \ref{fig:chan_types}\,b).
 
@@
 Yet, elevated variability is especially strong with highly variable water discharge in the \alpine case.
-Increased variability in sediment transport capacity in evaluating hydro-climatic signals from sediment records in glacierized catchments needs further evaluation.
-Furthermore, greater variability in sediment transport capacity may lead to sediment exhaustion by subglacial water repeatedly crossing the mobilization threshold across a range of water discharges.
+Further evaluation is needed to establish the role of highly variability in sediment transport capacity in evaluating hydro-climatic signals from sediment records.
+Greater variability in sediment transport capacity may also lead to sediment exhaustion by subglacial water repeatedly crossing the mobilization threshold across a range of water discharges.
 
@@
 In subaerial channels, the timing of peak sediment discharge capacity and water discharge coincide.
-whereas it is coincident in subaerial ones.
 Covariance between water discharge and sediment transport capacity could be possible when water discharge varies at a slower rate than subglacial channel size.
@@
   A.J.T. helped interpret findings from Leverett Glacier and guided experiment design.
-  M.A.W. provided data from Fieschergletscher and added Section \ref{sect:scaling}.
-  D.F. provided data from Fieschergletscher.
+  M.A.W. and D.F. provided data from Fieschergletscher and contributed to analysis.
   All authors provided key inputs in writing and editing the manuscript.} %% this section is mandatory
@@
 
-Here three kinds of drainage regimes are examined, namely, subaerial channels, steady-state R-channels \citep{rothlisberger1972}, and pipe-flow (i.e.
-R-channels which do not have time to adjust their size to discharge conditions).
+Here three kinds of drainage regimes are examined, namely, pipe-flow (i.e.
+R-channels which do not have time to adjust their size to discharge conditions; Figure~\ref{fig:chan_types} a), subaerial channels (Figure~\ref{fig:chan_types} c), steady-state R-channels \citep[Figure~\ref{fig:chan_types} d]{rothlisberger1972}.
 Here we show the sediment transport capacity scaling with a given water discharge and hydraulic gradient for three different sediment transport capacity formulas: Meyer-Peter M\"uller  \citep[MPM; ][]{meyer1948}, Engelund and Hansen \citep[EH; ][]{engelund1967}, and Bagnold \citep{bagnold1980}; additionally width-integrated shear stress is assessed, as used as proxy for sediment transport throughout the manuscript.
@@
 \begin{equation}
+  \label{eq:Qs_eq}
   Q_s \propto w\, \tau^{3/2}, \quad Q_s \propto w\, \tau^{5/2}, \quad Q_s \propto w \left(\frac{\Omega}{w}\right)^{3/2} H^{-2/3},
@@
 As in Section~\ref{sect:fluv}, the subaerial channel is assumed to have a width $w$ much greater than its depth $H$, such that $D_h\approx 4h$, and to have a constant head gradient ($\Psi$) given by the topography.
-Further, it is assumed that its width can be approximated by a relation $w \propto Q^\alpha$ (Equation~\ref{eq:wcf}) with $\alpha\in [0,1]$; typically $\alpha \approx 1/2$ (a so-called regime channel).
+Further, it is assumed that its width can be approximated by a relation $w \propto Q^\alpha$ (Equation~\ref{eq:wcf}) with $\alpha\in [0,1]$.
 End members $\alpha=0$ or $\alpha=1$ correspond a  subaerial channels of constant width (a slot canyon) or depth (no natural equivalent), respectively.
 % 
-For a steady state R-channel, it is assumed that  $\Psi$ is constant (approximated by the gradient of the Shreve \citeyear{shreve1972} potential and that $S$ adjusts in steady state with $\Psi$ and $Q$.
-Note that the R-channel model presented above in Section~\ref{sect:sub_mode} Equation~\ref{eq:dS_dt} calculates $\Psi$ from the time-evolving $S_g$ via the Darcy-Weisbach Equation~\ref{eq:dh} and thus no Shreve approximation is then needed.
+For a steady-state R-channel, it is assumed that  $\Psi$ is constant (approximated by the gradient of the Shreve \citeyear{shreve1972} potential and that $S$ adjusts in steady state with $\Psi$ and $Q$.
+Note that the R-channel model presented above in Section~\ref{sect:sub_mode}, Equation~\ref{eq:dS_dt} calculates $\Psi$ from the time-evolving $S_g$ via the Darcy-Weisbach Equation~\ref{eq:dh} and thus no Shreve approximation is then needed.
 % Conversely, pipe flow has $S$ fixed and $\Psi$ adjusts to satisfy the given $Q$.
@@
 In this case, it is assumed that the cross-sectional area $S$ is fixed, and $\Psi$ adjusts to the specified $Q$.
-For both R-channel and pipe flow, it is assumed that $D_h \propto S^{1/2}$.
+For both steady-state R-channel and pipe flow, it is assumed that $D_h \propto S^{1/2}$.
 
@@
 Now width integrated shear stress and sediment transport can be calculated for all regimes and all transport formulas.
-The results of the 12 combinations are presented in Table~\ref{tab:Qs} and also in Table~S1 where the fractions in the exponents are approximately given by decimal numbers for ease of comparison.
+The results of the 12 combinations are presented in Table~\ref{tab:Qs} and also in Table~\ref{tab:Qs2} where the fractions in the exponents are approximately given by decimal numbers for ease of comparison.
 
@@
 
-Table~\ref{tab:Qs} shows how the proxy $w\tau$ as well as $Q_s$ of the MPM, EH, and Bagnold sediment transport formulas scale with respect to $Q$, $\Psi$ or $S$ for the three different regimes and for different $\alpha$.
+Table~\ref{tab:Qs} shows the scaling of the proxy $w\tau$ as well as $Q_s$ of the MPM, EH, and Bagnold sediment transport formulas with respect to $Q$, $\Psi$ or $S$ for the three different regimes and for different $\alpha$.
 Remarkably, the Bagnold formula has a negative exponent for $f$ in all but the pipe-flow regime.
-The total transport formula EH gives a slightly stronger dependence on all variables as should be expected due to the larger exponent on $\tau$ of $\frac{5}{2}$ versus  $\frac{3}{2}$  for the others.
+The total transport formula EH gives a slightly stronger dependence on all variables due to the larger exponent on $\tau$ of $\frac{5}{2}$ versus  $\frac{3}{2}$  for the others (Equation~\ref{eq:Qs_eq}).
 Albeit, the sediment transport response in pipe flow for the Bagnold case is close to EH.
-Conversely, the sediment transport proxies $w_g\,\tau_g$ \& $w_f\,\tau_f$ used below scale only as $Q^2$ for pipe-flow, whereas the transport scales at least with $Q^3$.
+Conversely, the sediment transport proxies $w_g\,\tau_g$ and $w_f\,\tau_f$ used below scale only as $Q^2$ for pipe-flow, whereas the transport scales at least with $Q^3$.
 The exponent for the width scaling $\alpha$ only impacts the relationship between sediment transport and water discharge in the EH relation in any meaningful way.
-However, for the value of $\alpha$ around $\frac{1}{2}$ (a value appropriate for most streams), the exponent on $Q$ for EH is only slightly greater than the other transport relations.
+However, for the value of $\alpha$ around $\frac{1}{3}$ (a value appropriate for most streams), the exponent on $Q$ for EH is only slightly greater than the other transport relations.
 
-The sediment discharge capacity in the steady state R-channel scales very similarly to the subaerial channel case for all relations and virtually identically for the $\alpha=0.5$ case.
+The sediment discharge capacity in the steady state R-channel scales very similarly to the subaerial channel case for all relations and virtually identically for the $\alpha=0.5$ case (Table\ref{tab:eqs1}).
 However, the head gradients $\Psi$ are likely higher for comparable $Q$ in an ice-sheet marginal or alpine glacier setting than in a subaerial channel \citep{alley1997}.
-Thus, sediment transport capacity is likely much higher in a steady-state R-channel.
+Thus, sediment transport capacity is likely higher in a steady-state R-channel.
 
@@
     Pipe & \(f \, Q^2 \, S^{-1}\) & \(f^{3/2}\, Q^3 \, S^{-5/2}\) & \(f^{5/2}\, Q^5\, S^{-9/2}\) & \(f^{3/2} \, Q^{9/2} \, S^{-14/3}\)\\
+  \end{tabular}
+\end{table}
+
+\begin{table}
+  \caption{As Table~3 but with the fractional exponents stated as rounded decimal numbers.
+For the subaerial regime the exponents are for displayed for the likely width-exponent $\alpha=1$, as well as its end members 0 (slot canyon) and 1 (only width increases and not depth).
+  }
+  \small
+  \label{tab:Qs2}
+  \begin{tabular}{lllll}
+    & Width \(\times \,\, \tau\) & MPM & EH & Bagnold\\
+    & \(w\, \tau\) & \(Q_s \propto w\, \tau^{3/2}\) & \(Q_s \propto w\, \tau^{5/2}\) & \(Q_s \propto w^{-1/2}\, \Omega^{3/2} H^{-2/3}\)\\
+    \hline
+    Subaerial (\(\alpha=0\)) & \(f^{0.7}\, Q^{0.7}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{1.7} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.1} \, \Psi^{1.7}\)\\
+    Subaerial (\(\alpha=0.3\)) & \(f^{0.7}\, Q^{0.8}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{1.3} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\
+    %Subaerial (\(\alpha=0.5\)) & \(f^{0.7}\, Q^{0.8}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{1.3} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\
+    Subaerial (\(\alpha=1\)) & \(f^{0.7}\, Q^{1.0}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{0.7} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\[3pt]
+    R-channel & \(f^{0.4}\, Q^{0.8} \, \Psi^{0.6}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.7}\, Q^{1.4}\, \Psi^{1.8}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\
+    Pipe & \(f \,\quad Q^{2\phantom{.0}} \, S^{-1}\) & \(f^{1.5}\, Q^3 \, S^{-2.5}\) & \(f^{2.5}\, Q^{5\phantom{.0}}\, S^{-4.5}\) & \(f^{1.5} \,\,\,\,\, Q^{4.5} \, S^{-4.7}\)\\
   \end{tabular}

### 2024-07-09 Delaneyetal2024.tex
update to formulations section

@@
   By contrast, subglacial channels grow at a slower rate than most water discharge variations, causing a variable relationship between channel shape, water velocity, and discharge.
-  Using numerical models applied to two observed hydrographs, we evaluate the differences in sediment transport capacity variability between subaerial and subglacial channels.
+  Using numerical models applied to two observed hydrographs and sediment transport formulations, we evaluate the differences in sediment transport capacity variability between subaerial and subglacial channels.
   The numerical experiments confirm that subglacial sediment transport is highly non-linear with respect to water discharge.
@@
   In a dedicated numerical experiment, we show that the changing channel size results in sediment transport capacity peaking before the maximum water discharge.
-  Results also indicate that high subglacial sediment transport capacities can occur across a wide range of water discharges.
+  The results also indicate that high subglacial sediment transport capacities can occur across a wide range of water discharges.
   This implies that hysteresis in channel size causes a variable relationship between sediment and water discharge in a transport-limited subglacial system.
@@
   We also evaluate the conditions under which subglacial sediment transport capacity could co-vary with water discharge.
-  These findings can also improve interpretations of sediment discharge records in glacierized catchments.
+  These findings can also improve interpretation of sediment discharge records in glacierized catchments.
 \end{abstract}
@@
 
-Changes in glacier dynamics and hydrology have motivated numerous recent studies of sediment transport processes in cold regions \citep[e.g.][]{li2022,vergara2022,zhang2022}.
+Changes in glacier dynamics and hydrology have motivated numerous recent studies on sediment transport processes in cold regions \citep[e.g.][]{li2022,vergara2022,zhang2022}.
 Increases in sediment transport have been observed in Greenland \citep{bendixen2017}, the European Alps \citep{costa2017}, the Himalayas \citep{li2021}, and the Andes \citep{vergara2022}.
-To accurately explain observed changes to sediment transport in glacierized catchments, the processes controlling sediment discharge and its variations with water discharge  need to be examined \citep[e.g.][]{riihimaki2005,swift2005}.
+To accurately explain observed changes in sediment transport in glacierized catchments, the processes controlling sediment discharge and its variations with water discharge  need to be examined \citep[e.g.][]{riihimaki2005,swift2005}.
 
@@
 
-In a transport-limited regime, sediment discharge is controlled by sediment transport capacity, defined as the amount of sediment the water can carry.
+In a transport-limited regime, sediment discharge is controlled by sediment transport capacity, which is defined as the amount of sediment the water can carry.
 In both subglacial and subaerial channels, sediment transport capacity depends on the shear stress between water and the sediment it flows over \citep{shields1936,meyer1948,engelund1967}, along with the width of the channel bottom $w$ over which sediment mobilizes.
@@
 To date, several modeling frameworks examine subglacial sediment transport with evolving channel size \citep{creyts2013,beaud2018,delaney2019,hewitt2019}.
-However, these works minimally discuss the impact of different hydrological regimes on sediment transport capacity underneath glaciers for interpreting sediment transport records.
-The few explicit parameterizations of subglacial sediment transport capacity with respect to water discharge have assumed fixed channel size  \citep{alley1997}. 
+However, these works minimally discuss the impact of different hydrological regimes on sediment transport capacity underneath glaciers in the context of interpreting sediment transport records.
+The few explicit parameterizations of subglacial sediment transport capacity with respect to water discharge assume fixed channel size  \citep{alley1997}. 
 These formulations demonstrate a strongly non-linear response in subglacial sediment transport capacity to water discharge.
+Yet, the effects of evolving channels size could impact the sediment transport response. 
 Understanding these processes is imperative for establishing the effect of hydro-climatic conditions on subglacial sediment dynamics.
@@
 
-This manuscript has two objectives: 1) to evaluate the variability of subglacial sediment transport capacity with evolving subglacial channel size and different hydrographs and 2) to establish whether sub-seasonal water discharge can co-vary with sediment transport capacity in subglacial systems.
+This manuscript has two objectives: 1) to evaluate the variability of subglacial sediment transport capacity with evolving subglacial channel size and different hydrographs, and 2) to establish whether sub-seasonal water discharge can co-vary with sediment transport capacity in subglacial systems.
 We use numerical models to examine sediment transport capacity in both subglacial channels that evolve in size and subaerial channels that remain fixed.
@@
 Model outputs demonstrate the specific processes that can drive variability in sediment discharge capacity in subglacial systems.
-These findings indicate differences in the relationships amongst water discharge, channel geometry, and water velocity in subglacial and subaerial channels.
+Lastly,  we present algebraic formulations of the sediment transport capacity response to water discharge in subaerial, steady-state R-channel, and pipe flow conditions.
+These formulations extend the relationships presented in \citet{alley1997}, and illustrate the behavior of different channel types.
+Findings indicate differences in the relationships amongst water discharge, channel geometry, and water velocity in subglacial and subaerial channels.
 The manuscript then discusses the implications of variability in sediment transport capacity from glaciers and the interpretation of sediment transport records.
@@
 \end{linenomath*}
-\noindent where $f_r$ is a friction factor, $D_h$ is the hydraulic diameter, $l$ is the channel length, and $v_g=\frac{Q}{S_g}$ is the water flow speed.
+\noindent where $f_r$ is a friction factor, $D_h$ is the hydraulic diameter, $l$ is the channel length, and $v_g=\frac{Q}{S_g}$ is the water velocity.
 % 
@@
 Note that this subaerial channel model is purely algebraic, whereas the subglacial model comprises a differential equation for the evolution of $S_g$.
-Thus the channel size in the subaerial model has no history dependence on the discharge $Q_w$, whereas the subglacial one does (Equation~\ref{eq:dS_dt}).
+Thus, the channel size in the subaerial model has no history dependence on the discharge $Q_w$, whereas the subglacial one does (Equation~\ref{eq:dS_dt}).
 
@@
 
-The first numerical experiment aims to compare the variability between the subglacial and subaerial model outputs to a range of channel shapes, friction factors, and water discharge variability, as the subglacial channel's cross-sectional area is allowed to evolve.
-%Following the findings in Section~\ref{sect:scaling}, variability in subglacial sediment discharge capacity could approach that of subaerial systems, should water discharge variations be sufficiently small.
+The first numerical experiment aims to compare the variability between the subglacial and subaerial model outputs to a range of channel shapes, friction factors, and water discharge variability.
 
@@
 In some cases, subglacial model outputs' variability is a magnitude larger than their subaerial counterparts.
-Results show that variability in both subglacial and subaerial outputs decreases with smoothing time longer than approximately $1$--$5$ days (Figure~\ref{fig:multi_run}).
+Variability in both subglacial and subaerial outputs decreases with smoothing time longer than approximately $1$--$5$ days (Figure~\ref{fig:multi_run}).
 These smoothing timescales remove the diurnal variations in water discharge, thereby reducing variability in model outputs. 
@@
 Variations in \icesheet{} shear stress are comparable to \alpine due to the effects the evolution of shear stress with subglacial conduit size (Equations ~\ref{eq:dS_dt} and \ref{eq:DW};). 
-The much greater water flux in \icesheet{} drives the larger width integrated shear stress variations (Figure~\ref{fig:Qw}\,c and f).
+The much greater water flux and thus subglacial channel size in \icesheet{} drives the larger width integrated shear stress variations (Figure~\ref{fig:Qw}\,c and f).
 
@@
 
-Conversely, the history dependence on channel size in subglacial channels means that model outputs occur across a large range of water discharges.
+Conversely, the history dependence on channel size in subglacial channels means that different sediment transport characteristics, such as velocity, occur across a large range of water discharges.
 For instance, in subglacial channels in \alpine{},  high water velocity values and shear stresses can occur from a low water discharge  ($\sim\,4$\,\unit{m}$^3$\,\unit{s}$^{-1}$) to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ (Figure~\ref{fig:Qw_vari}\,a).
@@
 \FloatBarrier
+\section{Sensitivity of sediment transport in different drainage regimes}
+\label{sect:scaling}
+
+\begin{figure}[hbt!]
+  \centering
+  \includegraphics[width=0.7\linewidth]{Fig7.pdf}
+  \caption{Response of sediment transport capacity (orange line) and channel size (bottom row) to a water discharge event under different channel conditions.
+    a) Pipe flow with fixed channel size described in Section~\ref{sect:scaling}.
+    b) R-channel described in Section \ref{sect:sub_mode}.
+    c) Subaerial channel described in Section~\ref{sect:fluv}.
+    d) Steady state R-channel (channel size evolves in equilibrium with water discharge) described in Section~\ref{sect:scaling}.
+    Note axes not to scale.
+    Gyres on bottom panels represent shear stress on channel walls.
+    $t_1$ ($t_2$) represents low (high) water discharge. 
+  }
+  \label{fig:chan_types}
+\end{figure}
+
+Here we examine subaerial channels (Figure~\ref{fig:chan_types} c), steady-state R-channels \citep[Figure~\ref{fig:chan_types} d]{rothlisberger1972} and pipe-flow (i.e.
+R-channels of fixed size that do not adjust their size to discharge conditions; Figure~\ref{fig:chan_types} a).
+The sediment transport capacity is scaled with a given water discharge and hydraulic gradient for three different sediment transport capacity formulas: Meyer-Peter M\"uller  \citep[MPM; ][]{meyer1948}, Engelund and Hansen \citep[EH; ][]{engelund1967}, and Bagnold \citep{bagnold1980}; additionally width-integrated shear stress is assessed, as used as proxy for sediment transport throughout the manuscript.
+
+Sediment discharge is given by the MPM, EH, and Bagnold formulations as
+\begin{equation}
+  \label{eq:Qs_eq}
+  Q_s \propto w\, \tau^{3/2}, \quad Q_s \propto w\, \tau^{5/2}, \quad Q_s \propto w \left(\frac{\Omega}{w}\right)^{3/2} H^{-2/3},
+\end{equation}
+respectively, for conditions well above the transport onset threshold.
+
+We use the Darcy-Weisbach equation to evaluate shear stress
+\begin{equation}
+  \label{eq:DW}
+  \Psi \propto f\frac{Q^2}{D_h S^2},
+\end{equation}
+with $\Psi = \Delta h / l$ the head gradient and friction factor $f$.
+The shear stress and stream power are, respectively,
+\begin{equation}
+  \label{eq:tau-omega}
+  \tau \propto f v^2 = f \left(\frac{Q}{S}\right)^2, \quad  \Omega \propto \Psi Q.
+\end{equation}
+%
+
+As in Section~\ref{sect:fluv}, the subaerial channel is assumed to have a width $w$ much greater than its depth $H$, such that $D_h\approx 4h$, and to have a constant head gradient ($\Psi$) given by the topography.
+Further, it is assumed that its width can be approximated by a relation $w \propto Q^\alpha$ (Equation~\ref{eq:wcf}) with $\alpha\in [0,1]$.
+End members $\alpha=0$ or $\alpha=1$ correspond a  subaerial channels of constant width (a slot canyon) or depth (no natural equivalent), respectively.
+% 
+For a steady-state R-channel, it is assumed that  $\Psi$ is constant (approximated by the gradient of the Shreve \citeyear{shreve1972} potential and that $S$ adjusts in steady state with $\Psi$ and $Q$.
+Note that the R-channel model used above (Section~\ref{sect:sub_mode}), Equation~\ref{eq:dS_dt} calculates $\Psi$ from the time-evolving $S_g$ via the Darcy-Weisbach Equation~\ref{eq:dh} and thus no Shreve approximation is then needed.
+% Conversely, pipe flow has $S$ fixed and $\Psi$ adjusts to satisfy the given $Q$.
+% 
+Pipe flow-like conditions occur when an R-channel is subjected to rapid discharge variations such that the channel cannot adjust its size.
+In this case, it is assumed that the cross-sectional area $S$ is fixed, and $\Psi$ adjusts to the specified $Q$.
+For both steady-state R-channel and pipe flow, it is assumed that $D_h \propto S^{1/2}$.
+
+With these assumptions, the Darcy-Weisbach equation~\eqref{eq:DW} can be solved for the not-fixed quantity: $H$ for a subaerial channel, $S$ for a R-channel, and $\Psi$ for pipe-flow.
+Then, using equations~\eqref{eq:tau-omega}, the shear stress and stream power can be calculated.
+These results of are summarised in Table~\ref{tab:eqs1}.
+The results of the 12 combinations of shear stress and sediment transport capacity are presented in Table~\ref{tab:Qs} and also in Table~S1 where the fractions in the exponents are approximately given by decimal numbers for ease of comparison.
+
+\begin{table}[hbt!]
+  \caption{Relations for hydraulic variables for the three drainage regimes:  subaerial channels, R-channel, and pipe-flow.
+Darcy-Weisbach equation is abbreviated with ``D-W'', and stream power with ``Stream p.''.
+}
+  \small
+  \label{tab:eqs1}
+  \begin{tabular}{llllll}
+    Regime & Fixed & Determined via D-W
+    & Additional relations & Shear stress & Stream p.\\
+           & & (Equation~\ref{eq:DW})  &  &  \(\tau \propto\) & \(\Omega \propto\)\\
+    \hline
+    Subaerial & \(\Psi\) & \(H \propto f^{1/3}\, Q^{2/3-2\alpha/3} \, \Psi^{-1/3}\) & \makecell{\(w\,\propto Q^\alpha\) \\ \(S=wH\) \\ \(D_h\propto H\)} & \(f^{1/3} Q^{2/3-2\alpha/3}  \Psi^{2/3}\) & \(Q\, \Psi\)\\
+    R-channel & \(\Psi\) & \(S\, \propto f^{2/5}\, Q^{4/5} \, \Psi^{-2/5}\) & \(D_h\propto w \propto H \propto S^{1/2}\) & \(f^{1/5} Q^{2/5} \, \Psi^{4/5}\) & \(Q\, \Psi\)\\
+    Pipe & \(S\) & \(\Psi \propto f \, Q^2\, S^{-5/2}\) & \(D_h\propto w \propto H \propto S^{1/2}\) & \(f Q^2 S^{-2}\) & \(f\, Q^3 S^{-5/2}\)\\
+  \end{tabular}
+\end{table}
+
+Table~\ref{tab:Qs} shows the scaling of the proxy $w\tau$ as well as $Q_s$ of the MPM, EH, and Bagnold sediment transport formulas with respect to $Q$, $\Psi$ or $S$ for the three different regimes and for different $\alpha$.
+Remarkably, the Bagnold formula has a negative exponent for $f$ in all but the pipe-flow regime.
+The total transport formula EH gives a slightly stronger dependence on all variables due to the larger exponent on $\tau$ of $\frac{5}{2}$ versus  $\frac{3}{2}$  for the others (Equation~\ref{eq:Qs_eq}).
+Albeit, the sediment transport response in pipe flow for the Bagnold case is close to EH.
+Conversely, the sediment transport proxies $w_g\,\tau_g$ and $w_f\,\tau_f$ used below scale only as $Q^2$ for pipe-flow, whereas the sediment transport capacity scales at least with $Q^3$.
+The exponent for the width scaling $\alpha$ only impacts the relationship between sediment transport and water discharge in the EH relation in any meaningful way.
+However, for the value of $\alpha$ around $\frac{1}{3}$ (a value appropriate for most streams), the exponent on $Q$ for EH is only slightly greater than the other transport relations.
+
+The sediment discharge capacity in the steady state R-channel scales very similarly to the subaerial channel case for all relations and virtually identically for the $\alpha=\frac{1}{3}$ case (Table~\ref{tab:eqs1}).
+However, the head gradients $\Psi$ are likely higher for comparable $Q$ in an ice-sheet marginal or alpine glacier setting than in a subaerial channel \citep{alley1997}.
+Thus, sediment transport capacity is likely higher in a steady-state R-channel.
+
+\begin{table}[hbt!]
+  \caption{Sediment transport proxy ($w\tau$) and rates for the three considered different transport formulas: MPM \citep{meyer1948}, EH \citep{engelund1967}, and Bagnold \citep{bagnold1980}.
+  }
+  \small
+  \label{tab:Qs}
+  \begin{tabular}{lllll}
+    & Width \(\times \,\, \tau\) & MPM & EH & Bagnold\\
+    & \(w\, \tau\) & \(Q_s \propto w\, \tau^{3/2}\) & \(Q_s \propto w\, \tau^{5/2}\) & \(Q_s \propto w^{-1/2}\, \Omega^{3/2} H^{-2/3}\)\\
+    \hline
+    Subaerial  & \(f^{1/3}\, Q^{2/3+\alpha/3}\,  \Psi^{2/3}\) & \(f^{1/2}\, Q \, \Psi\) & \(f^{5/6}\, Q^{5/3 - 2\alpha/3} \, \Psi^{5/3}\) & \(f^{-2/9}\, Q^{19/18-\alpha/18} \, \Psi^{31/18}\)\\
+    R-channel & \(f^{2/5}\, Q^{4/5} \, \Psi^{3/5}\) & \(f^{1/2}\, Q \, \Psi\) & \(f^{7/10}\, Q^{7/5}\, \Psi^{9/5}\) & \(f^{-7/30}\, Q^{31/30}\, \Psi^{26/15}\)\\
+    Pipe & \(f \, Q^2 \, S^{-1}\) & \(f^{3/2}\, Q^3 \, S^{-5/2}\) & \(f^{5/2}\, Q^5\, S^{-9/2}\) & \(f^{3/2} \, Q^{9/2} \, S^{-14/3}\)\\
+  \end{tabular}
+\end{table}
+
+\FloatBarrier
 \section{Discussion}
@@
 
-\begin{figure}[hbt!]
-  \centering
-  \includegraphics[width=0.7\linewidth]{Fig7.pdf}
-  \caption{Response of sediment transport capacity (orange line) and channel size (bottom row) to a water discharge event under different channel conditions.
-    a) Pipe flow with fixed channel size described in Section~\ref{sect:scaling}.
-    b) R-channel described in Section \ref{sect:sub_mode}.
-    c) Subaerial channel described in Section~\ref{sect:fluv}.
-    d) Steady state R-channel (channel size evolves in equilibrium with water discharge) described in Section~\ref{sect:scaling}.
-    Note axes not to scale.
-    Gyres on bottom panels represent shear stress on channel walls.
-    $t_1$ ($t_2$) represents low (high) water discharge. 
-  }
-  \label{fig:chan_types}
-\end{figure}
+R-channels rarely operate in a steady state with variations in water discharge, especially during severe rain or melt is too short for the channel to reach a steady state (Figures~\ref{fig:model_outs} and \ref{fig:Qw_vari}).
+In these cases with high water discharge variability, channels can behave more like a pipe of fixed cross-section \citep[e.g.][]{gimbert2016}.
+Table~\ref{tab:Qs} shows that sediment transport in pipe-flow scales much more severely with discharge; the exponent on $Q$ being between $3$ and $5$, compared to the other two regimes when that exponent is at most $\frac{7}{5}$.
+Thus fluctuations of discharge on short timescales (on the order of a day) have the potential to cause conditions with very high sediment transport capacities.
+These sediment transport capacities variations are of far higher magnitude than those of subaerial channels \citep{alley1997}, when evolving subglacial channel size is not considered (Figure~\ref{fig:model_outs}).
+Note that pipe flow would cause sediment discharge capacity to covary with water discharge (Figure~\ref{fig:chan_types} a).
+However, this covariance does not occur with evolving channel size (Figures~\ref{fig:Qw_vari} and \ref{fig:chan_types} b).
 
@@
 Such a strong relationship would not represent the impact of actual shorter-term fluctuations in water discharge.
-Note, in this case, that if  pipe flow is assumed, with pressurized flow, but no channel evolution, then  sediment discharge capacity would covary with water discharge (Figure~\ref{fig:chan_types} a).
 
@@
 
+The channel width and size variations in subglacial channels presented here also occur in subaerial channels in response to water discharge \citep{phillips2016}, likely over decadal timescales or in response to individual extreme events \citep[e.g.]{slater2013,dean2013}.
+These timescales are likely  considerably longer than the continuous changes to the subglacial channel's response time of days.
+Furthermore, some observations from subaerial systems show a co-varying relationship between water discharge and sediment transport in time \citep[e.g.][]{schmidt2006,pitlick2021}.
+Even so, if channel width changes occur in subaerial channels, then the values of $\alpha$ in Equation~\ref{eq:wcf} could change in time.
+$\alpha$ must be less than $1$, and even this would cause the exponent on $Q$ in the  $w_g\,\tau_{g}$ relationships to remain substantially below that of pipeflow (Section ~\ref{sect:scaling}, Table\,\ref{tab:eqs1} and \ref{tab:Qs})).
+
 Width-integrated shear stress is examined in the two numerical experiments as it does not have a dependence on grain size, unlike sediment transport relationships \citep[][Section \ref{sect:scaling}]{meyer1948}.
@@
 The manuscript's second objective is to establish if water discharge varies with subglacial sediment transport capacity.
+In subglacial channels, the timing of peak sediment discharge capacity occurs before peak water discharge during an event, due to evolving channel size.
+In subaerial channels, the timing of peak sediment discharge capacity and water discharge coincide.
 Results here suggest that, even in a transport-limited subglacial system, an incoherent relationship between water and subglacial sediment discharge could be expected in most systems.
 This presents a challenge in linking hydro-climatic conditions or events to sediment export from glaciers.
-In subglacial channels, the timing of peak sediment discharge capacity occurs before peak water discharge during an event, due to evolving channel size.
-In subaerial channels, the timing of peak sediment discharge capacity and water discharge coincide.
 Covariance between water discharge and sediment transport capacity could be possible when water discharge varies at a slower rate than subglacial channel size.
@@
 \appendix
-\section{Sensitivity of sediment transport in different drainage regimes}
-\label{sect:scaling}
-
-Here three kinds of drainage regimes are examined, namely, pipe-flow (i.e.
-R-channels which do not have time to adjust their size to discharge conditions; Figure~\ref{fig:chan_types} a), subaerial channels (Figure~\ref{fig:chan_types} c), steady-state R-channels \citep[Figure~\ref{fig:chan_types} d]{rothlisberger1972}.
-Here we show the sediment transport capacity scaling with a given water discharge and hydraulic gradient for three different sediment transport capacity formulas: Meyer-Peter M\"uller  \citep[MPM; ][]{meyer1948}, Engelund and Hansen \citep[EH; ][]{engelund1967}, and Bagnold \citep{bagnold1980}; additionally width-integrated shear stress is assessed, as used as proxy for sediment transport throughout the manuscript.
-
-Sediment discharge is given by the MPM, EH, and Bagnold formulations as
-\begin{equation}
-  \label{eq:Qs_eq}
-  Q_s \propto w\, \tau^{3/2}, \quad Q_s \propto w\, \tau^{5/2}, \quad Q_s \propto w \left(\frac{\Omega}{w}\right)^{3/2} H^{-2/3},
-\end{equation}
-respectively, for conditions well above the transport onset threshold.
-
-The following hydraulic relations are used.
-The Darcy-Weisbach equation can be stated as
-\begin{equation}
-  \label{eq:DW}
-  \Psi \propto f\frac{Q^2}{D_h S^2},
-\end{equation}
-with $\Psi = \Delta h / l$ the head gradient and friction factor $f$.
-The shear stress and stream power are, respectively,
-\begin{equation}
-  \label{eq:tau-omega}
-  \tau \propto f v^2 = f \left(\frac{Q}{S}\right)^2, \quad  \Omega \propto \Psi Q.
-\end{equation}
-%
-
-As in Section~\ref{sect:fluv}, the subaerial channel is assumed to have a width $w$ much greater than its depth $H$, such that $D_h\approx 4h$, and to have a constant head gradient ($\Psi$) given by the topography.
-Further, it is assumed that its width can be approximated by a relation $w \propto Q^\alpha$ (Equation~\ref{eq:wcf}) with $\alpha\in [0,1]$.
-End members $\alpha=0$ or $\alpha=1$ correspond a  subaerial channels of constant width (a slot canyon) or depth (no natural equivalent), respectively.
-% 
-For a steady-state R-channel, it is assumed that  $\Psi$ is constant (approximated by the gradient of the Shreve \citeyear{shreve1972} potential and that $S$ adjusts in steady state with $\Psi$ and $Q$.
-Note that the R-channel model presented above in Section~\ref{sect:sub_mode}, Equation~\ref{eq:dS_dt} calculates $\Psi$ from the time-evolving $S_g$ via the Darcy-Weisbach Equation~\ref{eq:dh} and thus no Shreve approximation is then needed.
-% Conversely, pipe flow has $S$ fixed and $\Psi$ adjusts to satisfy the given $Q$.
-% 
-Pipe flow-like conditions occur when an R-channel is subjected to rapid discharge variations such that the channel cannot adjust its size.
-In this case, it is assumed that the cross-sectional area $S$ is fixed, and $\Psi$ adjusts to the specified $Q$.
-For both steady-state R-channel and pipe flow, it is assumed that $D_h \propto S^{1/2}$.
-
-With these assumptions, the Darcy-Weisbach equation~\eqref{eq:DW} can be solved for the not-fixed quantity: $H$ for a subaerial channel, $S$ for a R-channel, and $\Psi$ for pipe-flow.
-Then, using equations~\eqref{eq:tau-omega}, the shear stress and stream power can be calculated.
-These results of are summarised in Table~\ref{tab:eqs1}.
-Now width integrated shear stress and sediment transport can be calculated for all regimes and all transport formulas.
-The results of the 12 combinations are presented in Table~\ref{tab:Qs} and also in Table~\ref{tab:Qs2} where the fractions in the exponents are approximately given by decimal numbers for ease of comparison.
-
-\begin{table}[hbt!]
-  \caption{Relations for hydraulic variables for the three drainage regimes:  subaerial channels, R-channel, and pipe-flow.
-Darcy-Weisbach equation is abbreviated with ``D-W'', and stream power with ``Stream p.''.
-}
-  \small
-  \label{tab:eqs1}
-  \begin{tabular}{llllll}
-    Regime & Fixed & Determined via D-W
-    & Additional relations & Shear stress & Stream p.\\
-           & & (Equation~\ref{eq:DW})  &  &  \(\tau \propto\) & \(\Omega \propto\)\\
-    \hline
-    Subaerial & \(\Psi\) & \(H \propto f^{1/3}\, Q^{2/3-2\alpha/3} \, \Psi^{-1/3}\) & \makecell{\(w\,\propto Q^\alpha\) \\ \(S=wH\) \\ \(D_h\propto H\)} & \(f^{1/3} Q^{2/3-2\alpha/3}  \Psi^{2/3}\) & \(Q\, \Psi\)\\
-    R-channel & \(\Psi\) & \(S\, \propto f^{2/5}\, Q^{4/5} \, \Psi^{-2/5}\) & \(D_h\propto w \propto H \propto S^{1/2}\) & \(f^{1/5} Q^{2/5} \, \Psi^{4/5}\) & \(Q\, \Psi\)\\
-    Pipe & \(S\) & \(\Psi \propto f \, Q^2\, S^{-5/2}\) & \(D_h\propto w \propto H \propto S^{1/2}\) & \(f Q^2 S^{-2}\) & \(f\, Q^3 S^{-5/2}\)\\
-  \end{tabular}
-\end{table}
-
-Table~\ref{tab:Qs} shows the scaling of the proxy $w\tau$ as well as $Q_s$ of the MPM, EH, and Bagnold sediment transport formulas with respect to $Q$, $\Psi$ or $S$ for the three different regimes and for different $\alpha$.
-Remarkably, the Bagnold formula has a negative exponent for $f$ in all but the pipe-flow regime.
-The total transport formula EH gives a slightly stronger dependence on all variables due to the larger exponent on $\tau$ of $\frac{5}{2}$ versus  $\frac{3}{2}$  for the others (Equation~\ref{eq:Qs_eq}).
-Albeit, the sediment transport response in pipe flow for the Bagnold case is close to EH.
-Conversely, the sediment transport proxies $w_g\,\tau_g$ and $w_f\,\tau_f$ used below scale only as $Q^2$ for pipe-flow, whereas the transport scales at least with $Q^3$.
-The exponent for the width scaling $\alpha$ only impacts the relationship between sediment transport and water discharge in the EH relation in any meaningful way.
-However, for the value of $\alpha$ around $\frac{1}{3}$ (a value appropriate for most streams), the exponent on $Q$ for EH is only slightly greater than the other transport relations.
-
-The sediment discharge capacity in the steady state R-channel scales very similarly to the subaerial channel case for all relations and virtually identically for the $\alpha=0.5$ case (Table\ref{tab:eqs1}).
-However, the head gradients $\Psi$ are likely higher for comparable $Q$ in an ice-sheet marginal or alpine glacier setting than in a subaerial channel \citep{alley1997}.
-Thus, sediment transport capacity is likely higher in a steady-state R-channel.
-
-%Furthermore, R-channels will rarely operate in a steady state as variations in water discharge.
-%The timescale of diurnal melt or severe rain or melt is too short for the channel to reach a steady state.
-%In these cases with high water discharge variability, channels operate more like a pipe of fixed cross-section \citep[e.g.][]{gimbert2016}.
-%Table~\ref{tab:Qs} shows that sediment transport in pipe-flow scales much more severely with discharge; the exponent on $Q$ being between $3$ and $5$, compared to the other two regimes when that exponent is at most $1.7$ (Table~\ref{tab:Qs}).
-%Thus fluctuations of discharge on short timescales (on the order of a day) have the potential to cause conditions with very high sediment transport capacities.
-%These sediment transport capacities variations are of far higher magnitude than those of subaerial channels \citep{alley1997}.
-
-\begin{table}[hbt!]
-  \caption{Sediment transport proxy ($w\tau$) and rates for the three considered different transport formulas: MPM \citep{meyer1948}, EH \citep{engelund1967}, and Bagnold \citep{bagnold1980}.
-  }
-  \small
-  \label{tab:Qs}
-  \begin{tabular}{lllll}
-    & Width \(\times \,\, \tau\) & MPM & EH & Bagnold\\
-    & \(w\, \tau\) & \(Q_s \propto w\, \tau^{3/2}\) & \(Q_s \propto w\, \tau^{5/2}\) & \(Q_s \propto w^{-1/2}\, \Omega^{3/2} H^{-2/3}\)\\
-    \hline
-    Subaerial  & \(f^{1/3}\, Q^{2/3+\alpha/3}\,  \Psi^{2/3}\) & \(f^{1/2}\, Q \, \Psi\) & \(f^{5/6}\, Q^{5/3 - 2\alpha/3} \, \Psi^{5/3}\) & \(f^{-2/9}\, Q^{19/18-\alpha/18} \, \Psi^{31/18}\)\\
-    R-channel & \(f^{2/5}\, Q^{4/5} \, \Psi^{3/5}\) & \(f^{1/2}\, Q \, \Psi\) & \(f^{7/10}\, Q^{7/5}\, \Psi^{9/5}\) & \(f^{-7/30}\, Q^{31/30}\, \Psi^{26/15}\)\\
-    Pipe & \(f \, Q^2 \, S^{-1}\) & \(f^{3/2}\, Q^3 \, S^{-5/2}\) & \(f^{5/2}\, Q^5\, S^{-9/2}\) & \(f^{3/2} \, Q^{9/2} \, S^{-14/3}\)\\
-  \end{tabular}
-\end{table}
-
-\begin{table}
-  \caption{As Table~3 but with the fractional exponents stated as rounded decimal numbers.
-For the subaerial regime the exponents are for displayed for the likely width-exponent $\alpha=1$, as well as its end members 0 (slot canyon) and 1 (only width increases and not depth).
-  }
-  \small
-  \label{tab:Qs2}
-  \begin{tabular}{lllll}
-    & Width \(\times \,\, \tau\) & MPM & EH & Bagnold\\
-    & \(w\, \tau\) & \(Q_s \propto w\, \tau^{3/2}\) & \(Q_s \propto w\, \tau^{5/2}\) & \(Q_s \propto w^{-1/2}\, \Omega^{3/2} H^{-2/3}\)\\
-    \hline
-    Subaerial (\(\alpha=0\)) & \(f^{0.7}\, Q^{0.7}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{1.7} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.1} \, \Psi^{1.7}\)\\
-    Subaerial (\(\alpha=0.3\)) & \(f^{0.7}\, Q^{0.8}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{1.3} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\
-    %Subaerial (\(\alpha=0.5\)) & \(f^{0.7}\, Q^{0.8}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{1.3} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\
-    Subaerial (\(\alpha=1\)) & \(f^{0.7}\, Q^{1.0}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{0.7} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\[3pt]
-    R-channel & \(f^{0.4}\, Q^{0.8} \, \Psi^{0.6}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.7}\, Q^{1.4}\, \Psi^{1.8}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\
-    Pipe & \(f \,\quad Q^{2\phantom{.0}} \, S^{-1}\) & \(f^{1.5}\, Q^3 \, S^{-2.5}\) & \(f^{2.5}\, Q^{5\phantom{.0}}\, S^{-4.5}\) & \(f^{1.5} \,\,\,\,\, Q^{4.5} \, S^{-4.7}\)\\
-  \end{tabular}
-\end{table}
 

### 2024-07-12 Delaneyetal2024.tex
small edits
changes on overleaf

@@
   This implies that hysteresis in channel size causes a variable relationship between sediment and water discharge in a transport-limited subglacial system.
+  Lastly, we present formulations that demonstrate the impact of different assumptions of channel behavior on sediment transport capacity.
   The implications are discussed for sediment discharge from glaciers with different hydro-climatic forcings.
   We also evaluate the conditions under which subglacial sediment transport capacity could co-vary with water discharge.
-  These findings can also improve interpretation of sediment discharge records in glacierized catchments.
+  These findings can also improve the interpretation of sediment discharge records in glacierized catchments.
 \end{abstract}
@@
 These formulations demonstrate a strongly non-linear response in subglacial sediment transport capacity to water discharge.
-Yet, the effects of evolving channels size could impact the sediment transport response. 
-Understanding these processes is imperative for establishing the effect of hydro-climatic conditions on subglacial sediment dynamics.
-This is especially the case because sediment discharge capacity controls the mobilization and deposition of sediment, making it a fundamental aspect of subglacial sediment evacuation, especially over short timescales.
+Yet, the continually evolving channels' size can impact variations in sediment transport capacity.
+Understanding these processes is imperative for establishing the effect of hydro-climatic conditions on subglacial sediment dynamics,
+especially as sediment discharge capacity controls the mobilization and deposition of sediment.
+This makes it a fundamental aspect of subglacial sediment evacuation, especially over short timescales.
 
-Despite the physical differences between the subaerial and subglacial systems, water discharge and sediment export are often compared in contemporary glacierized catchments \citep[e.g.][]{willis1996,hodson1998,pearce2003,richards2003,swift2005,chu2009,tedstone2012,chu2012,overeem2017,delaney2018,swift2021,lu2022,andresen2024}.
+Despite the different physical processes between the subaerial and subglacial systems, observed water discharge and sediment export are often compared in contemporary glacierized catchments \citep[e.g.][]{willis1996,hodson1998,pearce2003,richards2003,swift2005,chu2009,tedstone2012,chu2012,overeem2017,delaney2018,swift2021,lu2022,andresen2024}.
 Many of these studies discuss the disparity between water discharge and sediment export, particularly the non-linear relationship between the two, along with the role of sediment access or production.
 Yet, a variable relationship between water discharge and sediment export could be expected from glaciers with evolving conduit size in a transport-limited, or hydraulically controlled, regime.
-The response of sediment transport capacity to water discharge variations may affect the interpretation of sediment transport records in these glacierized catchment systems \citep[e.g.][]{ganti2016,mancini2023}.
-This relationship must be clarified as sediment dynamics change in glacierized regions along with changing hydrology \citep[e.g.][]{brunner2019}, increasing variability of water discharge \citep{lane2019} and occurrence of extreme glacier melt  \citep[e.g.][]{overeem2015,cremona2023} that may drive high magnitude sediment transport events.
+The response of sediment transport capacity to water discharge variations may affect the interpretation of sediment transport records in these glacierized catchments \citep[e.g.][]{ganti2016,mancini2023}.
+This relationship must be clarified as sediment dynamics change in glacierized regions along with changing hydrology \citep[e.g.][]{brunner2019}, increasing variability of water discharge \citep{lane2019b} and occurrence of extreme glacier melt  \citep[e.g.][]{overeem2015,cremona2023} that may drive high magnitude sediment transport events.
 
@@
 Lastly,  we present algebraic formulations of the sediment transport capacity response to water discharge in subaerial, steady-state R-channel, and pipe flow conditions.
-These formulations extend the relationships presented in \citet{alley1997}, and illustrate the behavior of different channel types.
+These formulations extend the relationships presented in \citet{alley1997} and illustrate the sediment transport capacity behavior of different channel types.
 Findings indicate differences in the relationships amongst water discharge, channel geometry, and water velocity in subglacial and subaerial channels.
@@
 
-To evaluate the shear stress of water flowing across sediments underneath a glacier, the subglacial channel model accounts for the channel geometry and the water's velocity.
+The subglacial channel model accounts for the channel geometry and the water's velocity to evaluate the shear stress of water flowing across sediments underneath a glacier.
 We use a lumped hydraulics model from \citet{werder2010b}, itself based upon  \citet{clarke1996}.
@@
     Shear stress (subglacial, subaerial) & $\tau$, ($\tau_g,\,\tau_f$) && $\mathrm{Pa \, m^{-2}}$ \\
-x    Stream power & $\Omega$ && $\mathrm{ kg \, m\, s^{-3}}$ \\
+    Stream power & $\Omega$ && $\mathrm{ kg \, m\, s^{-3}}$ \\
 
@@
 The first numerical experiment aims to compare the variability between the subglacial and subaerial model outputs to a range of channel shapes, friction factors, and water discharge variability. 
+This is accomplished by applying a range of parameter values to models applied to the different hydrological regimes and examining the variability in model outputs (Section~\ref{sect:imp}).
 
-Across the range of parameters examined (Section\,\ref{sect:imp}), variability in all model outputs remains higher in the subglacial system compared to the subaerial one for both \alpine{} and \icesheet{} (Figure~\ref{fig:multi_run}).
+Across the range of parameters examined, variability in all model outputs (i.e.
+velocity, shear stress, and width-integrated shear stress) remains higher in the subglacial system compared to the subaerial one for both \alpine{} and \icesheet{} (Figure~\ref{fig:multi_run}).
 In some cases, subglacial model outputs' variability is a magnitude larger than their subaerial counterparts.
@@
 Variations in \icesheet{} shear stress are comparable to \alpine due to the effects the evolution of shear stress with subglacial conduit size (Equations ~\ref{eq:dS_dt} and \ref{eq:DW};). 
-The much greater water flux and thus subglacial channel size in \icesheet{} drives the larger width integrated shear stress variations (Figure~\ref{fig:Qw}\,c and f).
+The much greater water flux and thus subglacial channel size in \icesheet{} drives the larger width integrated shear stress variations (Figures~\ref{fig:Qw}~and~\ref{fig:Qw}\,c and f).
 
@@
   \centering
-  \includegraphics[width=0.9\linewidth]{Fig3.pdf}
+  \includegraphics[width=0.7\linewidth]{Fig3.pdf}
   \caption{Standard deviation of detrended model outputs for different smoothing times with different subglacial and subaerial channel shapes and friction factors.
@@
 Yet, larger values of $\alpha$ can result in greater variability in width-integrated shear stress, as greater channel width changes increase the width-integrated shear stress.
-Additionally, steeper subaerial channel slopes result in greater variability in both shear stress and width-integrated shear stress but only approach the subglacial channel's variability in the \alpine case (Figure~\ref{fig:params}).
+Additionally, steeper subaerial channel slopes $\nabla z_c$ result in greater variability in both shear stress and width-integrated shear stress but only approach the subglacial channel's variability in the \alpine case (Figure~\ref{fig:params}).
 We note that these parameter values span a commonly accepted range.
@@
   \centering
-  \includegraphics[width=0.9\linewidth]{Fig4.pdf}
+  \includegraphics[width=0.7\linewidth]{Fig4.pdf}
   \caption{Parameter values compared to variability in model outputs.
@@
   \centering
-  \includegraphics[width=\linewidth]{Fig6.png}
+  \includegraphics[width=0.8\linewidth]{Fig6.png}
   \caption{
@@
 \FloatBarrier
-\section{Sensitivity of sediment transport in different drainage regimes}
+\subsection{Sensitivity of sediment transport in different channel types}
 \label{sect:scaling}
@@
   \includegraphics[width=0.7\linewidth]{Fig7.pdf}
-  \caption{Response of sediment transport capacity (orange line) and channel size (bottom row) to a water discharge event under different channel conditions.
+  \caption{Response of sediment transport capacity (orange line) and channel size (bottom row) to a water discharge event (blue line) under different channel conditions.
     a) Pipe flow with fixed channel size described in Section~\ref{sect:scaling}.
@@
     c) Subaerial channel described in Section~\ref{sect:fluv}.
-    d) Steady state R-channel (channel size evolves in equilibrium with water discharge) described in Section~\ref{sect:scaling}.
+    d) Steady-state R-channel (channel size evolves in equilibrium with water discharge) described in Section~\ref{sect:scaling}.
     Note axes not to scale.
@@
 \begin{table}[hbt!]
-  \caption{Relations for hydraulic variables for the three drainage regimes:  subaerial channels, R-channel, and pipe-flow.
+  \caption{Relations for hydraulic variables for the three different channel types:  subaerial channels, R-channel, and pipe-flow.
 Darcy-Weisbach equation is abbreviated with ``D-W'', and stream power with ``Stream p.''.
@@
   \begin{tabular}{llllll}
-    Regime & Fixed & Determined via D-W
+    Channel type & Fixed & Determined via D-W
     & Additional relations & Shear stress & Stream p.\\
@@
 
-Table~\ref{tab:Qs} shows the scaling of the proxy $w\tau$ as well as $Q_s$ of the MPM, EH, and Bagnold sediment transport formulas with respect to $Q$, $\Psi$ or $S$ for the three different regimes and for different $\alpha$.
-Remarkably, the Bagnold formula has a negative exponent for $f$ in all but the pipe-flow regime.
+Table~\ref{tab:Qs} shows the scaling of the proxy $w\tau$ as well as $Q_s$ of the MPM, EH, and Bagnold sediment transport formulas with respect to $Q$, $\Psi$ or $S$ for the three different channel types and for different $\alpha$.
+Remarkably, the Bagnold formula has a negative exponent for $f$ in all but the pipe-flow channel type.
 The total transport formula EH gives a slightly stronger dependence on all variables due to the larger exponent on $\tau$ of $\frac{5}{2}$ versus  $\frac{3}{2}$  for the others (Equation~\ref{eq:Qs_eq}).
@@
 
+R-channels rarely operate in a steady state with variations in water discharge, especially during severe rain or melt is too short for the channel to reach a steady state (Figures~\ref{fig:model_outs} and \ref{fig:Qw_vari}).
+In these cases with high water discharge variability, channels can behave more like a pipe of fixed cross-section \citep[\alpine{} case in Figure \ref{fig:Qw_vari}; e.g.][]{gimbert2016}.
+Table~\ref{tab:Qs} shows that sediment transport in pipe-flow scales much more severely with discharge; the exponent on $Q$ being between $3$ and $5$, compared to the other two channel types when that exponent is at most $\frac{7}{5}$.
+Thus fluctuations of discharge on short timescales (on the order of a day; Figure~\ref{fig:multi_run}) have the potential to cause conditions with very high sediment transport capacities.
+These sediment transport capacities variations are of far higher magnitude than those of subaerial channels \citep[][]{alley1997}.
+Note that pipe flow would cause sediment discharge capacity to covary with water discharge (Figure~\ref{fig:chan_types} a).
+However, this covariance does not occur with evolving channel size (Figures~\ref{fig:Qw_vari} and \ref{fig:chan_types} b).
+
 \FloatBarrier
@@
 
-While transport-limited states at glaciers are likely rare \citep[e.g.][]{alley1997}, abundant sediment could persist underneath some glaciers \citep[e.g.][]{walter2014,stevens2022,delaney2022}.
-As a result, sediment export can be hydrologically controlled and efficient \citep[e.g.][]{swift2005,andresen2024}.
+While transport-limited states at glaciers are likely rare \citep[e.g.][]{alley1997}, abundant sediment could persist underneath some glaciers, potentially creating a transport-limited regime \citep[e.g.][]{walter2014,stevens2022,delaney2022}.
 At these glaciers, the great variability in subglacial sediment transport capacity may make it difficult to link sediment discharge to hydrology, especially when peak events occur \citep{cowan1988,delaney2018,lu2022}.
 The results here suggest that different channel sizes with the same hydrology forcing can result in very different sediment transport capacities, making it challenging to establish the effects of individual extreme precipitation or melt events.
-
-R-channels rarely operate in a steady state with variations in water discharge, especially during severe rain or melt is too short for the channel to reach a steady state (Figures~\ref{fig:model_outs} and \ref{fig:Qw_vari}).
-In these cases with high water discharge variability, channels can behave more like a pipe of fixed cross-section \citep[e.g.][]{gimbert2016}.
-Table~\ref{tab:Qs} shows that sediment transport in pipe-flow scales much more severely with discharge; the exponent on $Q$ being between $3$ and $5$, compared to the other two regimes when that exponent is at most $\frac{7}{5}$.
-Thus fluctuations of discharge on short timescales (on the order of a day) have the potential to cause conditions with very high sediment transport capacities.
-These sediment transport capacities variations are of far higher magnitude than those of subaerial channels \citep{alley1997}, when evolving subglacial channel size is not considered (Figure~\ref{fig:model_outs}).
-Note that pipe flow would cause sediment discharge capacity to covary with water discharge (Figure~\ref{fig:chan_types} a).
-However, this covariance does not occur with evolving channel size (Figures~\ref{fig:Qw_vari} and \ref{fig:chan_types} b).
 
@@
 This occurs as the subglacial channel's size evolves slowly compared to the variations in water discharge (Figure~\ref{fig:chan_types} b).
-High water discharge variability in \alpine{} may cause shear stress to approach $Q^{2}$ in assuming pipe-flow conditions where the water discharge varies largely compared to channel size \citep[Figure~\ref{fig:chan_types}; Section \ref{sect:scaling}; c.f.][]{alley1997}.
+High water discharge variability in \alpine{} may cause width integrated shear stress to approach $Q^{2}$ in assuming pipe-flow conditions where the water discharge varies largely compared to channel size \citep[Figure~\ref{fig:chan_types}; Section \ref{sect:scaling}; c.f.][]{alley1997}.
 Conversely, the reduced relative variability in water discharge in \icesheet{} and stronger correlation between width-integrated shear stress may result in the subglacial channel's size being closer to equilibrium from the smaller variations in water discharge (Figure~\ref{fig:chan_types} d).
@@
 The co-varying relationship between sediment discharge capacity and water discharge in subaerial conditions could be noticeable $\sim20$\,\unit{km} downstream of the Leverett site.
-A strong correlation persists between sediment plume size and the Watson River's water discharge into the Kangerlussuaq fjord \citep[Figure~\ref{fig:chan_types}\,c][]{chu2009,mcgrath2010}.
+A strong correlation persists between sediment plume size and the Watson River's water discharge into the Kangerlussuaq fjord \citep[Figure~\ref{fig:chan_types}\,c; ][]{chu2009,mcgrath2010}.
 In contrast, in marine-terminating glacier catchments, a less consistent relationship may occur between water discharge or melt extent and sediment plume size \citep{chu2012,tedstone2012}.
@@
 Results here suggest that clockwise hysteresis could also be expected in transport-limited regimes in many subglacial environments.
-Here, the smaller channel size on the rising limb would result in greater sediment transport capacity and sediment uptakes (Figure~\ref{fig:chan_types} b).
+Here, the smaller channel size on the rising limb would result in greater sediment transport capacity and sediment mobilization (Figure~\ref{fig:chan_types} b).
 A larger channel size would reduce sediment transport capacity for an equivalent water discharge on the falling limb, creating clockwise hysteresis.
@@
 
-Several aspects of the models and experiment design make comparison of the two systems difficult. 
+Several aspects of the models and experiment design make the comparison of the subaerial and subglaical systems difficult. 
 The lumped nature of the models means that they operate independently of the upstream drainage network.
@@
 
-Comparing model outputs between the subaerial and subglacial case is difficult due to the different hydraulic gradients in the two systems, with the subglacial one generally being much steeper \citep{alley1997}.
-This results the shear stresses and width-integrated shear stresses across the channel bed that control sediment transport capacity are much smaller in subaerial channels (Figure~\ref{fig:model_outs}).
+Different hydraulic gradients control the velocity and sediment transport capacity in the subglacial and subaerial cases (Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}). 
+The subglacial one is generally much steeper \citep{alley1997}.
+This results in the shear stresses and width-integrated shear stresses across the channel bed that controls sediment transport capacity being much greater in subglacial channels (Figure~\ref{fig:model_outs}).
 The parameters' range tested in Section~\ref{sect:ensemble} covers a likely span of viable shear stresses in both subglacial and subaerial channels.
 The greater sediment transport capacity in the subglacial channels implies that sediment grain size underneath subglacial channels is likely larger than subaerial counterparts.
-Note however, that the water velocities are similar in both channel types.
+Note, however, that the water velocities are similar in both channel types.
 
-The channel width and size variations in subglacial channels presented here also occur in subaerial channels in response to water discharge \citep{phillips2016}, likely over decadal timescales or in response to individual extreme events \citep[e.g.]{slater2013,dean2013}.
+The channel width and size variations in subglacial channels presented here also occur in subaerial channels in response to water discharge \citep{phillips2016}, likely over decadal timescales or in response to individual extreme events \citep[e.g.][]{slater2013,dean2013}.
 These timescales are likely considerably longer than the continuous changes to the subglacial channel's response time of days.
-Furthermore, some observations from subaerial systems show a co-varying relationship between water discharge and sediment transport in time \citep[e.g.][]{schmidt2006,pitlick2021}.
+Furthermore, observations from subaerial systems can show a co-varying relationship between water discharge and sediment transport in time \citep[e.g.][]{schmidt2006,pitlick2021}.
 Even so, if channel width changes occur in subaerial channels, then the values of $\alpha$ in Equation~\ref{eq:wcf} could change in time.
@@
 
-The sediment transport capacity of both subglacial and subaerial channels are driven by its width and the shear stress exerted by the flowing water, which is proportional to flow velocity squared.
+The sediment transport capacity of both subglacial and subaerial channels is driven by its width and the shear stress exerted by the flowing water, which is proportional to flow velocity squared.
 Subaerial channels immediately alter both their width and water velocity in response to changing water discharge.
-In contrast, pressurized subglacial channels largely accommodate rapidly changing water discharge by altering water velocity, while their size only responds to changes in discharge over longer timescales, for instance several days.
+In contrast, pressurized subglacial channels largely accommodate rapidly changing water discharge by altering water velocity, while their size only responds to changes in discharge over longer timescales, for instance, several days.
 Thus, sediment transport capacity is more sensitive to changes in water discharge in subglacial channels compared to subaerial ones.
 
-The manuscript's first objective aims to evaluate the relative variability in sediment transport capacity between the two channel types.
-Results demonstrate that sediment transport capacity variability is higher in subglacial channels with evolving size across a range of channel conditions.
+The manuscript's first objective aims to evaluate the relative variability in sediment transport capacity between subglacial and subaerial channels.
+Results demonstrate that sediment transport capacity variability is higher in pressurized subglacial channels independent of hydrological conditions and channel characteristics.
 Increased variability occurs even in environments where the water discharge has relatively small diurnal variations in the \icesheet case.
-Yet, elevated variability is especially strong with highly variable water discharge in the \alpine case.
-Further evaluation is needed to establish the role of highly variability in sediment transport capacity in evaluating hydro-climatic signals from sediment records.
+Yet, elevated variability is especially strong with greater water discharge variations in the \alpine case.
+Further evaluation is needed to establish the role of high variability in sediment transport capacity in evaluating hydro-climatic signals from sediment records.
 Greater variability in sediment transport capacity may also lead to sediment exhaustion by subglacial water repeatedly crossing the mobilization threshold across a range of water discharges.
@@
 Results here suggest that, even in a transport-limited subglacial system, an incoherent relationship between water and subglacial sediment discharge could be expected in most systems.
-This presents a challenge in linking hydro-climatic conditions or events to sediment export from glaciers.
+This incoherent relationship presents a challenge in linking hydro-climatic conditions or events to sediment export from glaciers.
 Covariance between water discharge and sediment transport capacity could be possible when water discharge varies at a slower rate than subglacial channel size.
-In natural environments, this appears to rarely be the case, except where basal melt dominates.
+In natural environments, this appears to rarely be the case.
 
@@
 
-This study calls for the explicit consideration of rapidly evolving channel size when examining the relationship between sediment transport and hydro-climatic conditions from glacierized catchments, especially ones with high water discharge variability.
+This study calls for the explicit consideration of evolving channel size when examining the relationship between sediment transport and hydro-climatic conditions from glacierized catchments, especially ones with high water discharge variability.
 
@@
 \end{acknowledgements}
-\newpage
-\appendix
+%\newpage
+%\appendix
 

### 2024-07-22 Delaneyetal2024.tex
paper updates

@@
   This quantity varies with water velocity and channel width.
-  In subaerial channels,  sediment transport capacity can evolve consistently because water discharge variations are accommodated by changes in flow depth, width, and water velocity.
-  By contrast, subglacial channels grow at a slower rate than most water discharge variations, causing a variable relationship between channel shape, water velocity, and discharge.
-  Using numerical models applied to two observed hydrographs and sediment transport formulations, we evaluate the differences in sediment transport capacity variability between subaerial and subglacial channels.
+  In subaerial channels,  sediment transport capacity can co-vary with water discharge because its variations are accommodated by changes in flow depth, width, and water velocity.
+  By contrast, subglacial channels evolve at a slower rate than most water discharge variations, causing a variable relationship between channel shape, water velocity, and discharge.
+  Using  sediment transport formulations and numerical models applied to two observed hydrographs, we evaluate the differences in sediment transport capacity variability between subaerial and subglacial channels.
   The numerical experiments confirm that subglacial sediment transport is highly non-linear with respect to water discharge.
@@
 Despite the different physical processes between the subaerial and subglacial systems, observed water discharge and sediment export are often compared in contemporary glacierized catchments \citep[e.g.][]{willis1996,hodson1998,pearce2003,richards2003,swift2005,chu2009,tedstone2012,chu2012,overeem2017,delaney2018,swift2021,lu2022,andresen2024}.
-Many of these studies discuss the disparity between water discharge and sediment export, particularly the non-linear relationship between the two, along with the role of sediment access or production.
-Yet, a variable relationship between water discharge and sediment export could be expected from glaciers with evolving conduit size in a transport-limited, or hydraulically controlled, regime.
+Many of these studies discuss the disparity between water discharge and sediment export given the role of sediment access or production, along with the sensitive and non-linear relationship between the two.
+Yet, a variable relationship between water discharge and sediment export could be expected from glaciers with evolving conduit size in a transport-limited regime, where sediment transport responds to the channel's hydraulic conditions, instead of sediment availability.
 The response of sediment transport capacity to water discharge variations may affect the interpretation of sediment transport records in these glacierized catchments \citep[e.g.][]{ganti2016,mancini2023}.
@@
 
-This manuscript has two objectives: 1) to evaluate the variability of subglacial sediment transport capacity with evolving subglacial channel size and different hydrographs, and 2) to establish whether sub-seasonal water discharge can co-vary with sediment transport capacity in subglacial systems.
+This manuscript has two objectives: 1) to evaluate the variability of  sediment transport capacity in subglacial channels with evolving channel size and different hydrographs compared to subaerial ones, and 2) to establish whether sub-seasonal water discharge can co-vary with sediment transport capacity in subglacial systems.
 We use numerical models to examine sediment transport capacity in both subglacial channels that evolve in size and subaerial channels that remain fixed.
@@
 Both models use the measured discharge to calculate water velocity, shear stress, and width-integrated shear stress, upon which both suspended sediment and bedload transport depend \citep[Figure \ref{fig:cartoon}; ][]{shields1936}.
-Our choice to evaluate shear stress omits the selection of a sediment transport relationship and a grain-size parameter \citep[e.g.][]{shields1936,meyer1948}.
+Our choice to evaluate our results in terms of shear stress omits the selection of a sediment transport relationship and a grain-size parameter \citep[e.g.][]{shields1936,meyer1948}.
 
@@
 
-Here, it is assumed that the water is transported through a subglacial channel \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972} beneath a glacier with channel length $l$, with a flatbed and a mean thickness of $h_{ice}$.
+Here, it is assumed that the water is transported through a subglacial channel \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972} beneath a glacier with channel length $l$, with a flat bed and a mean thickness of $h_{ice}$.
 The channel size grows from melt due to frictional heating from water flow and closes due to ice creep.
@@
 \end{linenomath*}
-\noindent where $t$ is time, $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (values in Table~\ref{table:vpm}), $g$ is the acceleration due to gravity, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change over $l$, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the mean ice overburden pressure expressed in meter water equivalent ($\rho_w$ is density of water; $\rho_i$ is density of ice), and $n$ is Glen's n \citep[usually $n=3$][]{glen1955}.
+\noindent where $t$ is time, $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (values in Table~\ref{table:vpm}), $g$ is the acceleration due to gravity, $Q$ is water discharge, $\Delta h$ is the hydraulic head drop change over $l$, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the mean ice overburden pressure expressed in meter water equivalent ($\rho_w$ is density of water; $\rho_i$ is density of ice), and $n$ is Glen's n \citep[usually $n=3$; ][]{glen1955}.
 The first term on the equation's right side represents the channel opening by frictional heating, while the following term represents channel closure from ice deformation.
@@
          && ($\frac{\pi}{10}$, $\pi$) & \\
-    Friction factor (subglacial, subaerial) & $f$, ($f_r$, $f_f$) &$5$ or $16$, $3$ & $\mathrm{(-)}$ \\
+    Friction factor (subglacial, subaerial) & $f$, ($f_r$, $f_f$) & -($5$, ($16$, $3$)) & $\mathrm{(-)}$ \\
          && ($0.01$, $21$) & \\
@@
 \label{sect:imp}
+
+\ian{redo some of this with Mauro's comments}
 
@@
 \FloatBarrier
+
+\subsection{Changing subglacial channel size drives different timing and variability in sediment transport capacity}
+
+The first numerical experiment aims to quantify the sources of increased variability in the subglacial model outputs as they exhibit different seasonal evolutions and peaks (Figure~\ref{fig:model_outs}).
+Variable relationships between model outputs and water discharge emerge for the subaerial and subglacial cases due to the hysteresis in channel size in the subglacial model (Equation~\ref{eq:dS_dt}).
+Because subaerial channels have no history dependence, each water discharge value in the subaerial channel produces a unique water velocity, shear stress, and width-integrated shear stress (Section~\ref{sect:sub_mode}).
+This characteristic results in a perfect rank correlation between the variables and water discharge (Figure~\ref{fig:Qw_vari}).
+An inconsistent relationship between subglacial model outputs and water discharge persists across the range of parameters examined in the ensemble runs (see Section~\ref{sect:ensemble}; Figure S15).
+\begin{figure}[hbt!]
+  \centering
+  \includegraphics[width=\linewidth]{Fig5.png}
+  \caption{Model outputs from simulations using the hydrographs in panels a and e for the scenarios \alpine{} (a-d) and \icesheet{} (e-h).
+    Black (gray) lines in a and e represent the subglacial channel size (water discharge).
+    Blue (orange) lines represent outputs from the subglacial (subaerial) channel.
+    Data are shown at $15$\unit{min} intervals.
+    Arrows show examples where variables peak in the subglacial channel before the subaerial one.
+    Insets in f-h show the peak melt event denoted by the shaded area in panels e--h, with an arbitrary y-axis.
+  }
+  \label{fig:model_outs}
+\end{figure}
+
+Peaks in subaerial model outputs occur coincident with peaks in water discharge (Figure~\ref{fig:model_outs}).
+In the subglacial channel, peaks in model outputs generally occur when water discharge increases at the fastest rate, but before the maximum water discharge.
+Reduced water velocity takes place when water discharge stabilizes at its peak and channel growth continues.
+As a result, subglacial sediment transport capacity is greatest on the hydrograph's rising limb, relative to the falling limb.
+
+The history dependence on channel size in subglacial channels means that different sediment transport characteristics, such as velocity, occur across a large range of water discharges.
+For instance, in subglacial channels in \alpine{},  high water velocity values and shear stresses can occur from a low water discharge  ($\sim\,4$\,\unit{m}$^3$\,\unit{s}$^{-1}$) to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ (Figure~\ref{fig:Qw_vari}\,a).
+In \icesheet{}, water velocities close to the seasonal mean value can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
+The subglacial channel's evolving width can counteract some of these effects.
+Width-integrated shear stress generally increases with water discharge, with greater rank correlation compared to water velocity or shear stress (\alpine{}, Figure~\ref{fig:Qw_vari} a--c).
+Yet, even the width-integrated shear stress can vary substantially relative to water discharge.
+The highest values of width-integrated shear stress occur at water discharge values ranging from roughly $11$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ to over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
+The variability in width-integrated shear stress is less pronounced in the \icesheet{} scenario, where the hydrograph has less diurnal variability (Figure~\ref{fig:Qw_vari} c, f).
+This results from the discharge variations being in closer equilibrium with subglacial channels compared to \alpine{} (Sections~\ref{sect:scaling} and~\ref{sect:ensemble}).
+
+\begin{figure}[hbt!]
+  \centering
+  \includegraphics[width=0.8\linewidth]{Fig6.png}
+  \caption{
+    Relationship between water discharge and normalized velocity, shear stress, and width-integrated shear stress for \alpine{} (a-c) and \icesheet{} (d-f).
+    Variables on the x-axis have been normalized to mean values.
+    $R_g$ ($R_f$) shows the Spearman rank correlations for the subglacial (subaerial) outputs.
+    Plots are shown with discharge and model outputs at $15$\unit{min} intervals.
+  }
+  \label{fig:Qw_vari}
+\end{figure}
+
+\FloatBarrier
 \subsection{Increased variability in evolving subglacial channels occurs across a range of channel shapes, slopes,  and friction values}
@@
 
-The first numerical experiment aims to compare the variability between the subglacial and subaerial model outputs to a range of channel shapes, friction factors, and water discharge variability. 
+The second numerical experiment aims to compare the variability between the subglacial and subaerial model outputs to a range of channel shapes, friction factors, and water discharge variability. 
 This is accomplished by applying a range of parameter values to models applied to the different hydrological regimes and examining the variability in model outputs (Section~\ref{sect:imp}).
@@
 \FloatBarrier
-\subsection{Changing subglacial channel size drives different timing and variability in sediment transport capacity}
-
-The second numerical experiment aims to quantify the sources of increased variability in the subglacial model outputs as they exhibit different seasonal evolutions and peaks (Figure~\ref{fig:model_outs}).
-Variable relationships between model outputs and water discharge emerge for the subaerial and subglacial cases due to the hysteresis in channel size in the subglacial model (Equation~\ref{eq:dS_dt}).
-Because subaerial channels have no history dependence, each water discharge value in the subaerial channel produces a unique water velocity, shear stress, and width-integrated shear stress (Section~\ref{sect:sub_mode}).
-This characteristic results in a perfect rank correlation between the variables and water discharge (Figure~\ref{fig:Qw_vari}).
-An inconsistent relationship between subglacial model outputs and water discharge persists across the range of parameters examined in the ensemble runs (Figure S15).
-\begin{figure}[hbt!]
-  \centering
-  \includegraphics[width=\linewidth]{Fig5.png}
-  \caption{Model outputs from simulations using the hydrographs in panels a and e for the scenarios \alpine{} (a-d) and \icesheet{} (e-h).
-    Black (gray) lines in a and e represent the subglacial channel size (water discharge).
-    Blue (orange) lines represent outputs from the subglacial (subaerial) channel.
-    Data are shown at $15$\unit{min} intervals.
-    Arrows show examples where variables peak in the subglacial channel before the subaerial one.
-    Insets in f-h show the peak melt event denoted by the shaded area in panels e--h, with an arbitrary y-axis.
-  }
-  \label{fig:model_outs}
-\end{figure}
-
-Conversely, the history dependence on channel size in subglacial channels means that different sediment transport characteristics, such as velocity, occur across a large range of water discharges.
-For instance, in subglacial channels in \alpine{},  high water velocity values and shear stresses can occur from a low water discharge  ($\sim\,4$\,\unit{m}$^3$\,\unit{s}$^{-1}$) to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ (Figure~\ref{fig:Qw_vari}\,a).
-In \icesheet{}, water velocities close to the seasonal mean value can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
-The subglacial channel's evolving width can counteract some of these effects.
-Width-integrated shear stress generally increases with water discharge, with greater rank correlation compared to water velocity or shear stress (\alpine{}, Figure~\ref{fig:Qw_vari} a--c).
-Yet, even the width-integrated shear stress can vary substantially relative to water discharge.
-The highest values of width-integrated shear stress occur at water discharge values ranging from roughly $11$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ to over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
-The variability in width-integrated shear stress is less pronounced in the \icesheet{} scenario, where the hydrograph has less diurnal variability (Figure~\ref{fig:Qw_vari} c, f).
-This results from the discharge variations being in closer equilibrium with subglacial channels compared to \alpine{} (Sections~\ref{sect:scaling} and~\ref{sect:ensemble}).
-
-\begin{figure}[hbt!]
-  \centering
-  \includegraphics[width=0.8\linewidth]{Fig6.png}
-  \caption{
-    Relationship between water discharge and normalized velocity, shear stress, and width-integrated shear stress for \alpine{} (a-c) and \icesheet{} (d-f).
-    Variables on the x-axis have been normalized to mean values.
-    $R_g$ ($R_f$) shows the Spearman rank correlations for the subglacial (subaerial) outputs.
-    Plots are shown with discharge and model outputs at $15$\unit{min} intervals.
-  }
-  \label{fig:Qw_vari}
-\end{figure}
-
-Peaks in subaerial model outputs occur coincident with peaks in water discharge (Figure~\ref{fig:model_outs}).
-In the subglacial channel, peaks in model outputs generally occur when water discharge increases at the fastest rate, but before the maximum water discharge.
-Reduced water velocity takes place when water discharge stabilizes at its peak and channel growth continues.
-As a result, subglacial sediment transport capacity is greatest on the hydrograph's rising limb, relative to the falling limb.
-
-\FloatBarrier
 \subsection{Sensitivity of sediment transport in different channel types}
@@
     d) Steady-state R-channel (channel size evolves in equilibrium with water discharge) described in Section~\ref{sect:scaling}.
-    Note axes not to scale.
-    Gyres on bottom panels represent shear stress on channel walls.
+    Note axes are not to scale.
     $t_1$ ($t_2$) represents low (high) water discharge. 
@@
 
-Here we examine subaerial channels (Figure~\ref{fig:chan_types} c), steady-state R-channels \citep[Figure~\ref{fig:chan_types} d]{rothlisberger1972} and pipe-flow (i.e.
+Here, we evaluate the relative sensitivity of different channel types as they scale with different water discharge, channel shape and hydraulic gradient.
+These include subaerial channels (Figure~\ref{fig:chan_types} c), steady-state R-channels \citep[Figure~\ref{fig:chan_types} d]{rothlisberger1972} and pipe-flow (i.e.
 R-channels of fixed size that do not adjust their size to discharge conditions; Figure~\ref{fig:chan_types} a).
-The sediment transport capacity is scaled with a given water discharge and hydraulic gradient for three different sediment transport capacity formulas: Meyer-Peter M\"uller  \citep[MPM; ][]{meyer1948}, Engelund and Hansen \citep[EH; ][]{engelund1967}, and Bagnold \citep{bagnold1980}; additionally width-integrated shear stress is assessed, as used as proxy for sediment transport throughout the manuscript.
+These formulations do not allow us to evaluate non-steady state R-channels (Section~\ref{sect:sub_mode}, Figure~\ref{fig:chan_types} b).
+The sediment transport capacity is calculated with a given water discharge and hydraulic gradient for three different sediment transport capacity formulas: Meyer-Peter M\"uller  \citep[MPM; ][]{meyer1948}, Engelund and Hansen \citep[EH; ][]{engelund1967}, and Bagnold \citep{bagnold1980}; additionally width-integrated shear stress is assessed, as used as proxy for sediment transport throughout the manuscript.
 
@@
 Remarkably, the Bagnold formula has a negative exponent for $f$ in all but the pipe-flow channel type.
-The total transport formula EH gives a slightly stronger dependence on all variables due to the larger exponent on $\tau$ of $\frac{5}{2}$ versus  $\frac{3}{2}$  for the others (Equation~\ref{eq:Qs_eq}).
+The total transport formula EH gives a slightly stronger dependence on all variables due to the larger exponent on $\tau$ of $\frac{5}{2}$ versus  $\frac{3}{2}$ for the MPM (Equation~\ref{eq:Qs_eq}).
 Albeit, the sediment transport response in pipe flow for the Bagnold case is close to EH.
@@
 
-The sediment discharge capacity in the steady state R-channel scales very similarly to the subaerial channel case for all relations and virtually identically for the $\alpha=\frac{1}{3}$ case (Table~\ref{tab:eqs1}).
+The sediment discharge capacity in the steady state R-channel scales very similarly to the subaerial channel case for all relations and virtually identically for the $\alpha=\frac{1}{3}$ case (Table~\ref{tab:Qs}).
 However, the head gradients $\Psi$ are likely higher for comparable $Q$ in an ice-sheet marginal or alpine glacier setting than in a subaerial channel \citep{alley1997}.
@@
 In these cases with high water discharge variability, channels can behave more like a pipe of fixed cross-section \citep[\alpine{} case in Figure \ref{fig:Qw_vari}; e.g.][]{gimbert2016}.
-Table~\ref{tab:Qs} shows that sediment transport in pipe-flow scales much more severely with discharge; the exponent on $Q$ being between $3$ and $5$, compared to the other two channel types when that exponent is at most $\frac{7}{5}$.
+Table~\ref{tab:Qs} shows that sediment transport in pipe-flow scales much more severely with discharge; the exponent on $Q$ being between $3$ and $5$, compared to the other two channel types when that exponent is at most $\frac{5}{3}$.
 Thus fluctuations of discharge on short timescales (on the order of a day; Figure~\ref{fig:multi_run}) have the potential to cause conditions with very high sediment transport capacities.
@@
 Note that pipe flow would cause sediment discharge capacity to covary with water discharge (Figure~\ref{fig:chan_types} a).
-However, this covariance does not occur with evolving channel size (Figures~\ref{fig:Qw_vari} and \ref{fig:chan_types} b).
+Covariance also occurs for steady state R-channels, but not for R-channels that evolve in time . (Figures~\ref{fig:Qw_vari} and \ref{fig:chan_types} b).
 
@@
 Subaerial channels immediately alter both their width and water velocity in response to changing water discharge.
-In contrast, pressurized subglacial channels largely accommodate rapidly changing water discharge by altering water velocity, while their size only responds to changes in discharge over longer timescales, for instance, several days.
+In contrast, pressurized subglacial channels largely accommodate rapidly changing water discharge by altering water velocity, while their size only responds to changes in discharge over longer timescales, typically, several days.
 Thus, sediment transport capacity is more sensitive to changes in water discharge in subglacial channels compared to subaerial ones.
@@
 
-The manuscript's second objective is to establish if water discharge varies with subglacial sediment transport capacity.
-In subglacial channels, the timing of peak sediment discharge capacity occurs before peak water discharge during an event, due to evolving channel size.
+The manuscript's second objective is to establish if water discharge covaries with subglacial sediment transport capacity.
+In subglacial channels, the timing of peak sediment discharge capacity occurs before peak water discharge during an discharge event, due to evolving channel size.
 In subaerial channels, the timing of peak sediment discharge capacity and water discharge coincide.
@@
 
-The subglacial environment's few observations hinder our ability to quantify processes such as
+Few observations of the subglacial environment hinder our ability to quantify processes such as
 the shape of subglacial channels and the response time of subglacial channels to water discharge variations.

### 2024-07-22 Delaneyetal2024.tex
Updates from Overleaf

--- 2024-07-12 Delaneyetal2024.tex
@@
 In subaerial channels, the timing of peak sediment discharge capacity and water discharge coincide.
-Results here suggest that, even in a transport-limited subglacial system, an incoherent relationship between water and subglacial sediment discharge could be expected in most systems.
+Results here suggest that, even in a transport-limited subglacial system, an incoherent relationship between water and subglacial sediment discharge could be expected in most channels.
 This incoherent relationship presents a challenge in linking hydro-climatic conditions or events to sediment export from glaciers.

### 2024-08-05 Delaneyetal2024.tex
mauro's edits

--- 2024-07-22 Delaneyetal2024.tex
@@
 Despite the different physical processes between the subaerial and subglacial systems, observed water discharge and sediment export are often compared in contemporary glacierized catchments \citep[e.g.][]{willis1996,hodson1998,pearce2003,richards2003,swift2005,chu2009,tedstone2012,chu2012,overeem2017,delaney2018,swift2021,lu2022,andresen2024}.
-Many of these studies discuss the disparity between water discharge and sediment export given the role of sediment access or production, along with the sensitive and non-linear relationship between the two.
+Many of these studies discuss the disparity between water discharge and sediment export, especially given the role of sediment access or production.
 Yet, a variable relationship between water discharge and sediment export could be expected from glaciers with evolving conduit size in a transport-limited regime, where sediment transport responds to the channel's hydraulic conditions, instead of sediment availability.
@@
 
-This manuscript has two objectives: 1) to evaluate the variability of  sediment transport capacity in subglacial channels with evolving channel size and different hydrographs compared to subaerial ones, and 2) to establish whether sub-seasonal water discharge can co-vary with sediment transport capacity in subglacial systems.
+This manuscript has two objectives: 1)  to establish whether sub-seasonal water discharge can co-vary with sediment transport capacity in subglacial systems, and 2) to evaluate the variability of  sediment transport capacity in subglacial channels with evolving channel size and different hydrographs compared to subaerial ones.
 We use numerical models to examine sediment transport capacity in both subglacial channels that evolve in size and subaerial channels that remain fixed.
 They are applied to proglacial hydrological records from an Alpine glacier in Switzerland (Fieschergletscher) and a land-terminating glacier in Greenland (Leverett Glacier).
-We run a model ensemble to evaluate the sediment transport capacity variability in subglacial channels compared to subaerial channels.
 Model outputs demonstrate the specific processes that can drive variability in sediment discharge capacity in subglacial systems.
+We also run a model ensemble to evaluate the sediment transport capacity variability in subglacial channels compared to subaerial channels.
 Lastly, we present algebraic formulations of the sediment transport capacity response to water discharge in subaerial, steady-state R-channel, and pipe flow conditions.
-These formulations extend the relationships presented in \citet{alley1997} and illustrate the sediment transport capacity behavior of different channel types.
-Findings indicate differences in the relationships amongst water discharge, channel geometry, and water velocity in subglacial and subaerial channels.
+These formulations extend the relationships presented in \citet{alley1997} and illustrate the sediment transport capacity behavior of different channel assumptions.
+Findings indicate differences in the relationships amongst water discharge, channel geometry, water velocity and sediment transport capacity in subglacial and subaerial channels.
 The manuscript then discusses the implications of variability in sediment transport capacity from glaciers and the interpretation of sediment transport records.
@@
 
-\ian{redo some of this with Mauro's comments}
-
 The models above are applied to proglacial discharge records from the Fieschergletscher (scenario \alpine{}) and the Leverett Glacier (scenario \icesheet{}).
@@
 
-The first experiment aims to characterize the variability in sediment discharge capacity in subglacial relative to subaerial channels across a range of channel slopes and shapes, along with friction factors.
-Additionally, we examine differing water discharge variability by smoothing the water discharge over different periods from $15$\,\unit{min} up to $15$\,\unit{days}.
+In the first experiment, the models are applied to a reference test case for each glacier.
+These experiments assumes  a subglacial channel with $\beta=\frac{\pi}{6}$ and a subaerial channel with $\alpha = \frac{1}{3}$ and slope of $0.02$ (Table~\ref{table:vpm}).
+Both friction factors $f_r$ and $f_f$ are tuned so that reasonable water velocities \citep[$\sim\,1- 1.5\,$\unit{m}\,\unit{s}$^{-1}$][]{werder2010b,chandler2013} occur for both \alpine{} and \icesheet{}.
+To test the covariance between sediment transport capacity and water discharge, model outputs are compared to water discharge from the glaciers using Spearman rank correlation.
+This metric accounts for the ordering of values, but not their magnitude.
+Rank correlation reduces the impact of the non-linear relationship between sediment transport capacity and hydrology.
+
+The second experiment aims to characterize the variability in sediment discharge capacity in subglacial relative to subaerial channels across a range of channel slopes and shapes, and friction factors.
+Additionally, we examine the effects differing water discharge smoothing periods from $15$\,\unit{min} up to $15$\,\unit{days}.
 The results below present water velocity ($v_g$, $v_f$), shear stress ($\tau_g$, $\tau_f$), and width-integrated shear stress ($w_g\tau_g$, $w_f\tau_f$) from the subglacial and subaerial models.
 For brevity, these values together are referred to as ``model outputs'' in the text.
-To evaluate variability, we subtract the model outputs from their values smoothed using daily averages,  detrending the seasonal signal.
-We then calculate the variability in the model output by taking the standard deviation of this detrended time series.
+To evaluate variability, we subtract the model outputs from their daily averages. 
+This creates a time series that is detrended from he seasonal variations in the model outputs and has an approximately normal distribution.
+We then present the variability in the model output by taking the standard deviation of this detrended time series.
 
@@
 To be accepted, subglacial model runs must experience a flotation fraction ($\frac{\Delta h}{2\,h_o}$) of above $1.2$ for less than $2.5$\,\% of the run.
-The initial condition of cross-sectional area $S_g$ is evaluated by applying a model spinup with repeatedly applying the maximum observed water discharge of the first $4$ days of the study until there is no change in the channel area.
+A model spinup dictates the initial condition of cross-sectional area $S_g$.
+The spinup consists of applying the maximum observed water discharge of the first $4$ days of the study period to the model until there is no change in the channel area $S_g$.
 
@@
 To test the effects of variability in water discharge, it is averaged over the smoothing period ranging
-
-In the second experiment, the models are applied to a reference test case for each glacier.
-These experiments assumes  a subglacial channel with $\beta=\frac{\pi}{6}$ and a subaerial channel with $\alpha = \frac{1}{3}$ and slope of $0.02$ (Table~\ref{table:vpm}).
-Both friction factors $f_r$ and $f_f$ are tuned so that reasonable water velocities \citep[$\sim\,1.6\,$\unit{m}\,\unit{s}$^{-1}$][]{werder2010b,chandler2013} occur for both \alpine{} and \icesheet{}.
-% 
-To test the covariance between sediment transport capacity and water discharge, model outputs are compared to water discharge from the glaciers using Spearman rank correlation.
-This metric accounts for the ordering of values, but not their magnitude.
-Rank correlation reduces the impact of the non-linear relationship between sediment transport capacity and hydrology.
 
@@
 
-As in Section~\ref{sect:fluv}, the subaerial channel is assumed to have a width $w$ much greater than its depth $H$, such that $D_h\approx 4h$, and to have a constant head gradient ($\Psi$) given by the topography.
+As in Section~\ref{sect:fluv}, the subaerial channel is assumed to have a width $w$ much greater than its depth $H$, such that $D_h\approx 4H$, and to have a constant head gradient ($\Psi$) given by the topography.
 Further, it is assumed that its width can be approximated by a relation $w \propto Q^\alpha$ (Equation~\ref{eq:wcf}) with $\alpha\in [0,1]$.
@@
 % 
-For a steady-state R-channel, it is assumed that  $\Psi$ is constant (approximated by the gradient of the Shreve \citeyear{shreve1972} potential and that $S$ adjusts in steady state with $\Psi$ and $Q$.
+For a steady-state R-channel, it is assumed that  $\Psi$ is constant (approximated by the gradient of the \citet{shreve1972} potential) and that $S$ adjusts in steady state with $\Psi$ and $Q$.
 Note that the R-channel model used above (Section~\ref{sect:sub_mode}), Equation~\ref{eq:dS_dt} calculates $\Psi$ from the time-evolving $S_g$ via the Darcy-Weisbach Equation~\ref{eq:dh} and thus no Shreve approximation is then needed.
@@
 Note that pipe flow would cause sediment discharge capacity to covary with water discharge (Figure~\ref{fig:chan_types} a).
-Covariance also occurs for steady state R-channels, but not for R-channels that evolve in time . (Figures~\ref{fig:Qw_vari} and \ref{fig:chan_types} b).
+Covariance also occurs for steady state R-channels, but not for R-channels where channel area evolves in time (Figures~\ref{fig:Qw_vari} and \ref{fig:chan_types} b).
 
@@
 Additionally, we omit analysis of suspended sediment discharge records due to their dependence on sediment supply in capturing variations in sediment discharge, in addition to sediment transport capacity \citep[e.g.][]{delaney2019}.
-In reality, processes such as sediment access are highly important in controlling sediment export in glacierized catchments \citep[e.g.][]{herman2015,vergara2022}.
+In reality, processes such as sediment access are usually important in controlling sediment export in glacierized catchments \citep[e.g.][]{herman2015,vergara2022}.
 Furthermore, meltwater can be distributed to flow through several adjacent conduits impacting the sediment transport capacity in each \citep[e.g.][]{werder2013,hewitt2019,delaney2023}.
 The lumped models used here isolate the relationship between water discharge and sediment transport capacity in subglacial and subaerial systems. 
-However, they neglect more complex, yet important spatially distributed processes.
+However, they neglect more complex, yet important, spatially distributed processes.
 
@@
 
-The manuscript's first objective aims to evaluate the relative variability in sediment transport capacity between subglacial and subaerial channels.
-Results demonstrate that sediment transport capacity variability is higher in pressurized subglacial channels independent of hydrological conditions and channel characteristics.
+The manuscript's first objective is to establish if water discharge covaries with subglacial sediment transport capacity.
+In subglacial channels, the timing of peak sediment discharge capacity occurs before peak water discharge during an discharge event, due to evolving channel size.
+In subaerial channels, the timing of peak sediment discharge capacity and water discharge coincide.
+Results here suggest that, even in a transport-limited subglacial system, an incoherent relationship between water and subglacial sediment discharge could be expected in most systems.
+This incoherent relationship presents a challenge in linking hydro-climatic conditions or events to sediment export from glaciers.
+Water discharge and sediment transport capacity covariance could be possible when water discharge varies at a slower rate than subglacial channel size.
+In natural environments, this appears to rarely be the case.
+
+The manuscript's second objective aims to evaluate the relative variability in sediment transport capacity between subglacial and subaerial channels.
+Results demonstrate that sediment transport capacity variability is higher in pressurized subglacial channels that are out of equilibrium with water discharge.
 Increased variability occurs even in environments where the water discharge has relatively small diurnal variations in the \icesheet case.
@@
 Greater variability in sediment transport capacity may also lead to sediment exhaustion by subglacial water repeatedly crossing the mobilization threshold across a range of water discharges.
-
-The manuscript's second objective is to establish if water discharge covaries with subglacial sediment transport capacity.
-In subglacial channels, the timing of peak sediment discharge capacity occurs before peak water discharge during an discharge event, due to evolving channel size.
-In subaerial channels, the timing of peak sediment discharge capacity and water discharge coincide.
-Results here suggest that, even in a transport-limited subglacial system, an incoherent relationship between water and subglacial sediment discharge could be expected in most systems.
-This incoherent relationship presents a challenge in linking hydro-climatic conditions or events to sediment export from glaciers.
-Covariance between water discharge and sediment transport capacity could be possible when water discharge varies at a slower rate than subglacial channel size.
-In natural environments, this appears to rarely be the case.
 

### 2024-08-05 Delaneyetal2024.tex
changes

@@
 %% 2-column papers and preprints
-\documentclass[esurf, manuscript]{copernicus}
+\documentclass[tc, manuscript]{copernicus}
 
@@
   By contrast, subglacial channels evolve at a slower rate than most water discharge variations, causing a variable relationship between channel shape, water velocity, and discharge.
-  Using  sediment transport formulations and numerical models applied to two observed hydrographs, we evaluate the differences in sediment transport capacity variability between subaerial and subglacial channels.
-  The numerical experiments confirm that subglacial sediment transport is highly non-linear with respect to water discharge.
-  Yet, subglacial sediment transport capacity can vary less if water discharge variations are closer to equilibrium with subglacial channel size.
-  In a dedicated numerical experiment, we show that the changing channel size results in sediment transport capacity peaking before the maximum water discharge.
+  Using  sediment transport formulations and numerical models applied to two observed hydrographs, we evaluate the conditions where sediment transport capacity varies with water discharge in subglacial channels.
+  Numerical experiments show that the changing channel size results in sediment transport capacity peaking before the maximum water discharge.
+  This hysteresis in channel size causes a highly variable relationship between sediment and water discharge in a transport-limited subglacial system.
   The results also indicate that high subglacial sediment transport capacities can occur across a wide range of water discharges.
-  This implies that hysteresis in channel size causes a variable relationship between sediment and water discharge in a transport-limited subglacial system.
-  Lastly, we present formulations that demonstrate the impact of different assumptions of channel behavior on sediment transport capacity.
-  The implications are discussed for sediment discharge from glaciers with different hydro-climatic forcings.
-  We also evaluate the conditions under which subglacial sediment transport capacity could co-vary with water discharge.
-  These findings can also improve the interpretation of sediment discharge records in glacierized catchments.
+  We  then evaluate the conditions where subglacial sediment transport capacity is more variable than its subaerial counterpart.
+  The numerical experiments of real hydrographs confirm that subglacial sediment transport is highly non-linear with respect to water discharge, creating more variability in sediment transport capacity.
+  Yet,  formulations of subglacial sediment transport capacity show that it can covary with water discharge variations when subglacial channel size is in equilibrium. 
+  The implications of these findings are discussed in the context of sediment discharge from glaciers with different hydro-climatic forcings.
+  We also discuss the impact of different assumptions of channel behavior on sediment transport capacity.
+  These findings can improve the interpretation of sediment discharge records in glacierized catchments.
 \end{abstract}

### 2024-08-05 Delaneyetal2024.tex
small changes

@@
 The second experiment aims to characterize the variability in sediment discharge capacity in subglacial relative to subaerial channels across a range of channel slopes and shapes, and friction factors.
-Additionally, we examine the effects differing water discharge smoothing periods from $15$\,\unit{min} up to $15$\,\unit{days}.
+Additionally, we examine the effects of differing water discharge smoothing periods from $15$\,\unit{min} up to $15$\,\unit{days}.
 The results below present water velocity ($v_g$, $v_f$), shear stress ($\tau_g$, $\tau_f$), and width-integrated shear stress ($w_g\tau_g$, $w_f\tau_f$) from the subglacial and subaerial models.
@@
 To evaluate variability, we subtract the model outputs from their daily averages. 
-This creates a time series that is detrended from he seasonal variations in the model outputs and has an approximately normal distribution.
+This creates a time series that is detrended from the seasonal variations in the model outputs and has an approximately normal distribution.
 We then present the variability in the model output by taking the standard deviation of this detrended time series.
@@
 This could result from the subglacial conduit being in closer equilibrium with water discharge.
-Variations in \icesheet{} shear stress are comparable to \alpine due to the effects the evolution of shear stress with subglacial conduit size (Equations ~\ref{eq:dS_dt} and \ref{eq:DW};). 
+Variations in \icesheet{} shear stress are comparable to \alpine due to the effects of the evolution of shear stress with subglacial conduit size (Equations ~\ref{eq:dS_dt} and \ref{eq:DW};). 
 The much greater water flux and thus subglacial channel size in \icesheet{} drives the larger width integrated shear stress variations (Figures~\ref{fig:Qw}~and~\ref{fig:Qw}\,c and f).
@@
 
-Here, we evaluate the relative sensitivity of different channel types as they scale with different water discharge, channel shape and hydraulic gradient.
+Here, we evaluate the relative sensitivity of different channel types as they scale with water discharge, channel shape, and hydraulic gradient.
 These include subaerial channels (Figure~\ref{fig:chan_types} c), steady-state R-channels \citep[Figure~\ref{fig:chan_types} d]{rothlisberger1972} and pipe-flow (i.e.
@@
 Then, using equations~\eqref{eq:tau-omega}, the shear stress and stream power can be calculated.
-These results of are summarised in Table~\ref{tab:eqs1}.
+These results are summarised in Table~\ref{tab:eqs1}.
 The results of the 12 combinations of shear stress and sediment transport capacity are presented in Table~\ref{tab:Qs} and also in Table~S1 where the fractions in the exponents are approximately given by decimal numbers for ease of comparison.
@@
 
-Table~\ref{tab:Qs} shows the scaling of the proxy $w\tau$ as well as $Q_s$ of the MPM, EH, and Bagnold sediment transport formulas with respect to $Q$, $\Psi$ or $S$ for the three different channel types and for different $\alpha$.
+Table~\ref{tab:Qs} shows the scaling of the proxy $w\tau$ as well as $Q_s$ of the MPM, EH, and Bagnold sediment transport formulas with respect to $Q$, $\Psi$ or $S$ for the three different channel types and  $\alpha$ values.
 Remarkably, the Bagnold formula has a negative exponent for $f$ in all but the pipe-flow channel type.
@@
 The manuscript's first objective is to establish if water discharge covaries with subglacial sediment transport capacity.
-In subglacial channels, the timing of peak sediment discharge capacity occurs before peak water discharge during an discharge event, due to evolving channel size.
+In subglacial channels, the timing of peak sediment discharge capacity occurs before peak water discharge during a discharge event, due to evolving channel size.
 In subaerial channels, the timing of peak sediment discharge capacity and water discharge coincide.

### 2024-08-08 Delaneyetal2024.tex
Working on Andrews comments

--- 2024-08-05 Delaneyetal2024.tex
@@
 \begin{abstract}
-  Sediment transport capacity in both subaerial and subglacial channels depends on the shear stress exerted across the channel bottom.
-  This quantity varies with water velocity and channel width.
-  In subaerial channels,  sediment transport capacity can co-vary with water discharge because its variations are accommodated by changes in flow depth, width, and water velocity.
-  By contrast, subglacial channels evolve at a slower rate than most water discharge variations, causing a variable relationship between channel shape, water velocity, and discharge.
-  Using  sediment transport formulations and numerical models applied to two observed hydrographs, we evaluate the conditions where sediment transport capacity varies with water discharge in subglacial channels.
+ Sediment transport capacity in both subaerial and subglacial channels depends on the shear stress exerted across the channel bottom, which varies with water velocity and channel width.
+  In subaerial channels, water discharge variations are accommodated  by flow depth and width changes, along with water velocity.
+  However, in subglacial channels, water is pressurized by the ice above and grow in response to frictional heating of water flowing through them. 
+  As a result, water discharge changes mainly result in velocity variations, as the channel geometry evolves slowly (over days).
+  Here, we present formulations of sediment transport capacity in different channel types and apply subglacial and subaerial hydraulics models to hydrographs from an Alpine glacier and the Greenland Ice.
   Numerical experiments show that the changing channel size results in sediment transport capacity peaking before the maximum water discharge.
@@
   The results also indicate that high subglacial sediment transport capacities can occur across a wide range of water discharges.
-  We  then evaluate the conditions where subglacial sediment transport capacity is more variable than its subaerial counterpart.
-  The numerical experiments of real hydrographs confirm that subglacial sediment transport is highly non-linear with respect to water discharge, creating more variability in sediment transport capacity.
-  Yet,  formulations of subglacial sediment transport capacity show that it can covary with water discharge variations when subglacial channel size is in equilibrium. 
+  A second set of numerical experiments show that subglacial sediment transport is highly non-linear with respect to water discharge, creating more variability in sediment transport capacity.
+  Yet, results and  formulations of subglacial sediment transport capacity show that its variability can approach that of subaerial systems when subglacial channel size is in equilibrium with water discharge.
   The implications of these findings are discussed in the context of sediment discharge from glaciers with different hydro-climatic forcings.
@@
 In the subglacial channel, peaks in model outputs generally occur when water discharge increases at the fastest rate, but before the maximum water discharge.
-Reduced water velocity takes place when water discharge stabilizes at its peak and channel growth continues.
-As a result, subglacial sediment transport capacity is greatest on the hydrograph's rising limb, relative to the falling limb.
+As the water discharge stabilizes at its peak, channel growth continues, causing water velocity and other model outputs to decrease from their peak values (Section~\ref{sect:sub_mode}).
+As a result, subglacial sediment transport capacity is greatest on the hydrograph's rising limb, relative to the falling limb, creating a hysteresis effect.
 
@@
 
-Smaller values of $\alpha$, or a channel cross-section closer to a slot canyon, result in greater variability in velocity and shear stress (Figure~\ref{fig:params}).
-In both cases, the variability of small values of $\alpha$ does not approach that of the subglacial channels.
-Yet, larger values of $\alpha$ can result in greater variability in width-integrated shear stress, as greater channel width changes increase the width-integrated shear stress.
-Additionally, steeper subaerial channel slopes $\nabla z_c$ result in greater variability in both shear stress and width-integrated shear stress but only approach the subglacial channel's variability in the \alpine case (Figure~\ref{fig:params}).
+Greater subglacial friction factors $f_i$ result in greater variability in the shear stress and width-integrate shear stress in the \alpine case (Figure~\ref{fig:params}).
+This result is expected given Equations~\ref{eq:tau_g}.
+Low values of $f_i$ result in slower growth rates in subglacial channels (Equations~\ref{eq:dS_dt} and \ref{eq:dh}).
+As a result, water velocity, as opposed to channel growth, accommodates increases in water discharge, and velocity variability increases.
+Higher values of $f_i$ allow faster channel growth that accommodates increase in water discharge, reducing velocity variability.
+Smaller values of channel factor $\beta$, creating low and broad channels, result in more variability in width-integrated shear stress.
+Here, the channel width can grow more quickly in response to water discharge increases as compared to a semi-circular channel with $\beta = \pi$ (Equation~ \ref{eq:dh}).
+
+Smaller values of subaerial channel shape factor $\alpha$, or a  channel cross-section shape closer to a slot canyon, result in greater variability in subaerial velocity and shear stress that can approach the variability of subglacial values (Figure~\ref{fig:params}).
+Steeper subaerial channel slopes $\nabla z_c$ result in greater variability in both shear stress and width-integrated shear stress but only approach the subglacial channel's variability in the \alpine case (Figure~\ref{fig:params}).
+Greater values of subaerial friction factors $f_p$ resulting in greater variability shear stress and width-integrated shear stress. 
+Yet, they to not exceed the variability of subglaical channels. 
 We note that these parameter values span a commonly accepted range.
 Therefore, we do not anticipate a scenario where variability in the subaerial system would exceed the subglacial system with these two hydrologic forcings.
-
-Greater subglacial and subaerial friction factors $f_i$ and $f_p$ result in greater variability in the shear stress and width-integrate shear stress in both cases (Figure~\ref{fig:params}).
-This result is expected given Equations~\ref{eq:tau_g} and \ref{eq:vf}.
-However, more velocity variability is present in low  $f_i$ and $f_p$ values. 
-Low values of $f_i$ result in slower growth rates in subglacial channels (Equations~\ref{eq:dS_dt} and \ref{eq:dh}).
-As a result, water velocity, as opposed to channel growth, accommodates increases in water discharge, and velocity variability increases.
-Smaller values of channel factor $\beta$, creating low and broad channels, result in more variability in width-integrated shear stress.
-Here, the channel width can grow more quickly in response to water discharge increases as compared to a semi-circular channel with $\beta = \pi$ (Equation~ \ref{eq:dh}).
 
@@
     Outputs are shown using $15$ minute smoothing time.
+    Linear trends are denoted in black if they are significant at  $p\ge0.01$ .
+    To compare the range of model outputs for the subaerial and subglacial cases, shaded areas show range of model outputs that are independent of the parameter values. 
   }
@@
 \FloatBarrier
-\subsection{Sensitivity of sediment transport in different channel types}
+\subsection{Sediment transport scaling in different channel types}
 \label{sect:scaling}
@@
 
-Here, we evaluate the relative sensitivity of different channel types as they scale with different water discharge, channel shape and hydraulic gradient.
-These include subaerial channels (Figure~\ref{fig:chan_types} c), steady-state R-channels \citep[Figure~\ref{fig:chan_types} d]{rothlisberger1972} and pipe-flow (i.e.
+The numerical experiments above consider the size evolution of subglacial channel and demonstrate that for these hydrographs subglacial sediment transport variability is greater than its subaerial counterpart (Section~\ref{sect:ensemble}).
+Here, we compare the sediment transport behavior of different channels types as they respond water discharge, channel shape, and hydraulic gradient. 
+This yields insights into the conditions where subglacial sediment transport capacity could exhibit similar variability as subaerial channels.
+
+Channel types examined include subaerial channels (Figure~\ref{fig:chan_types} c), steady-state R-channels \citep[Figure~\ref{fig:chan_types} d]{rothlisberger1972} and pipe-flow (i.e.
 R-channels of fixed size that do not adjust their size to discharge conditions; Figure~\ref{fig:chan_types} a).
-These formulations do not allow us to evaluate non-steady state R-channels (Section~\ref{sect:sub_mode}, Figure~\ref{fig:chan_types} b).
-The sediment transport capacity is calculated with a given water discharge and hydraulic gradient for three different sediment transport capacity formulas: Meyer-Peter M\"uller  \citep[MPM; ][]{meyer1948}, Engelund and Hansen \citep[EH; ][]{engelund1967}, and Bagnold \citep{bagnold1980}; additionally width-integrated shear stress is assessed, as used as proxy for sediment transport throughout the manuscript.
+These formulations do not allow us to evaluate non-steady state R-channels presented above (Section~\ref{sect:sub_mode}, Figure~\ref{fig:chan_types} b).
+However, the formulations establish the scaling of these different channel types  .
+The sediment transport capacity is calculated with a given water discharge and hydraulic gradient for three different sediment transport capacity formulas: Meyer-Peter M\"uller  \citep[MPM; ][]{meyer1948}, Engelund and Hansen \citep[EH; ][]{engelund1967}, and Bagnold \citep{bagnold1980}; additionally width-integrated shear stress is assessed, as used as proxy for sediment transport above.
 
@@
 \end{equation}
-respectively, for conditions well above the transport onset threshold.
+respectively, for sediment transport conditions well above the threshold of sediment motion.
 
@@
 Further, it is assumed that its width can be approximated by a relation $w \propto Q^\alpha$ (Equation~\ref{eq:wcf}) with $\alpha\in [0,1]$.
-End members $\alpha=0$ or $\alpha=1$ correspond a  subaerial channels of constant width (a slot canyon) or depth (no natural equivalent), respectively.
+End members $\alpha=0$ or $\alpha=1$ correspond to a subaerial channel of constant width (a slot canyon) or depth (no natural equivalent), respectively.
 % 
@@
 Albeit, the sediment transport response in pipe flow for the Bagnold case is close to EH.
-Conversely, the sediment transport proxies $w_g\,\tau_g$ and $w_f\,\tau_f$ used below scale only as $Q^2$ for pipe-flow, whereas the sediment transport capacity scales at least with $Q^3$.
+Conversely, the sediment transport proxies $w_g\,\tau_g$ and $w_f\,\tau_f$ used previously scale only as $Q^2$ for pipe-flow, whereas the sediment transport capacity scales at least with $Q^3$.
 The exponent for the width scaling $\alpha$ only impacts the relationship between sediment transport and water discharge in the EH relation in any meaningful way.
@@
 
+\ian{comeback here}
 R-channels rarely operate in a steady state with variations in water discharge, especially during severe rain or melt is too short for the channel to reach a steady state (Figures~\ref{fig:model_outs} and \ref{fig:Qw_vari}).
-In these cases with high water discharge variability, channels can behave more like a pipe of fixed cross-section \citep[\alpine{} case in Figure \ref{fig:Qw_vari}; e.g.][]{gimbert2016}.
+In these cases with high water discharge variability, channels can behave more like a pipe of fixed cross-section.
+Here the cross-section responds to a characteristic discharge, but variations in water discharge deviate substantially from the flow conditions responsible for the channel size \citep[i.e.
+diurnal water discharge variations in \alpine{} case Figure \ref{fig:Qw_vari}; e.g.][]{gimbert2016}.
 Table~\ref{tab:Qs} shows that sediment transport in pipe-flow scales much more severely with discharge; the exponent on $Q$ being between $3$ and $5$, compared to the other two channel types when that exponent is at most $\frac{5}{3}$.
@@
 
-The greater variations in shear stress in subglacial channels presented here may actually cause even greater variations in sediment transport capacity in subglacial channels than presented above  (Figure~\ref{fig:multi_run}).
+The greater variations in shear stress in subglacial channels compared with subaerial channels that we present here may cause even greater variations in sediment transport capacity in subglacial channels than our numerical results suggest  (Figure~\ref{fig:multi_run}).
 In sediment transport capacity relationships, such as in \citet{meyer1948} or \citet{engelund1967}, shear stress is scaled to the power of $\frac{3}{2}$ or $\frac{5}{2}$, respectively (e.g.
@@
 This result suggests sediment export here is especially sensitive to observed changes in water discharge variability \citep{lane2019b}.
-Sediment exhaustion through repeatedly crossing the mobilization threshold may explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \citep{overeem2017}.
-The same process could also result in sediment discharge's strong dependence on sediment production from sliding in many cases \citep{herman2015,koppes2015}.
+Sediment exhaustion through repeatedly crossing the mobilization threshold may explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on the glaciers' basal shear stress, a proxy for bedrock erosion, rather than glacier melt \citep{overeem2017}.
+More generally, the same process could also result in sediment discharge's strong dependence on sediment production from glacial sliding, itself dependent of basal shear stress of the glacier \citep{herman2015,koppes2015}.
 
-While transport-limited states at glaciers are likely rare \citep[e.g.][]{alley1997}, abundant sediment could persist underneath some glaciers, potentially creating a transport-limited regime \citep[e.g.][]{walter2014,stevens2022,delaney2022}.
+While transport-limited states likely do not occur at many glaciers \citep[e.g.][]{alley1997}, abundant sediment could persist underneath some glaciers, potentially creating a transport-limited regime \citep[e.g.][]{walter2014,stevens2022,delaney2022}.
 At these glaciers, the great variability in subglacial sediment transport capacity may make it difficult to link sediment discharge to hydrology, especially when peak events occur \citep{cowan1988,delaney2018,lu2022}.
-The results here suggest that different channel sizes with the same hydrology forcing can result in very different sediment transport capacities, making it challenging to establish the effects of individual extreme precipitation or melt events.
+The results here suggest that different channel sizes with the same hydrology forcing can result in very different sediment transport capacities, making it challenging to establish the effects of individual extreme precipitation or melt events without establishing the antecedent state of the subglacial drainage system.
 
-Catchments with reduced water discharge variability may experience yet less variability in sediment transport capacity, shown by the \alpine{} and \icesheet{} model outputs.
+Catchments with reduced water discharge variability may experience less variability in sediment transport capacity, shown by the \alpine{} and \icesheet{} model outputs.
 Indeed, the more decoupled and sporadic relationship between model outputs and water discharge in \alpine{} results from the relatively larger discharge variations on sub-daily to weekly timescales (Figure~\ref{fig:Qw_vari}).
@@
 High water discharge variability in \alpine{} may cause width integrated shear stress to approach $Q^{2}$ in assuming pipe-flow conditions where the water discharge varies largely compared to channel size \citep[Figure~\ref{fig:chan_types}; Section \ref{sect:scaling}; c.f.][]{alley1997}.
-Conversely, the reduced relative variability in water discharge in \icesheet{} and stronger correlation between width-integrated shear stress may result in the subglacial channel's size being closer to equilibrium from the smaller variations in water discharge (Figure~\ref{fig:chan_types} d).
+Conversely, the reduced relative variability in water discharge in \icesheet{} comes as a result of the larger catchment areas and longer travel times of the water \citep[e.g.][]{vanas2017}.
+Water discharge's stronger correlation with width-integrated shear stress in \icesheet{} may result in the subglacial channel's size being closer to equilibrium from the smaller variations in water discharge (Figure~\ref{fig:chan_types} d).
 In this case, the exponent on water discharge for shear stress is likely substantially less than $w\tau \propto Q^2$, but greater than $w \tau \propto Q^{\frac{4}{5}}$ that occurs in a steady-state R-channel (see Section \ref{sect:scaling}).
@@
 
-Water discharge measurements at less than daily timescales could severely limit subglacial sediment transport models in capturing specific events.
-The decreased model output variability beyond $1$--$5$ days of smoothing could result in water discharge being in equilibrium with the subglacial channel when they are in fact not (Figure~\ref{fig:multi_run}).
+Water discharge measurements at sub-daily timescales could severely limit the ability of a subglacial sediment transport model to capture specific events.
+The decreased model output variability beyond $1$--$5$ days of smoothing could result in water discharge appearing to be in equilibrium with the subglacial channel when they are in fact not (Figure~\ref{fig:multi_run}).
 Over this period, water discharge variations could lead to a stronger relationship with sediment discharge capacity (Figure~\ref{fig:chan_types} d and Section ~\ref{sect:scaling}).
@@
 Here, the smaller channel size on the rising limb would result in greater sediment transport capacity and sediment mobilization (Figure~\ref{fig:chan_types} b).
-A larger channel size would reduce sediment transport capacity for an equivalent water discharge on the falling limb, creating clockwise hysteresis.
+The subsequently larger channel size would reduce sediment transport capacity for an equivalent water discharge on the falling limb, creating clockwise hysteresis.
 
@@
 The manuscript's first objective is to establish if water discharge covaries with subglacial sediment transport capacity.
-In subglacial channels, the timing of peak sediment discharge capacity occurs before peak water discharge during an discharge event, due to evolving channel size.
+In subglacial channels, the timing of peak water velocity and sediment discharge capacity occurs before peak water discharge during a discharge event, due to evolving channel size.
 In subaerial channels, the timing of peak sediment discharge capacity and water discharge coincide.

### 2024-08-13 Delaneyetal2024.tex
pulling from OL

@@
 Water discharge was measured roughly $2$\unit{km} downstream from the terminus ($67^\circ\,03'\,5''$ N, $50^\circ\,12'\,59''$ W), at a $5$\,\unit{min} time interval from May 28, 2012 to August 8, 2012 \citet[][Figure~\ref{fig:Qw}\,b]{tedstone2013}.
+
+\ian{include figure}
 
@@
 The much greater water flux and thus subglacial channel size in \icesheet{} drives the larger width integrated shear stress variations (Figures~\ref{fig:Qw}~and~\ref{fig:Qw}\,c and f).
+
+\ian{is this figure better? ... text reorganized. What about a symbol for cross-referencing}
 

### 2024-08-15 Delaneyetal2024.tex
getting ready for submission

@@
   In subaerial channels, water discharge variations are accommodated  by flow depth and width changes, along with water velocity.
-  However, in subglacial channels, water is pressurized by the ice above and grow in response to frictional heating of water flowing through them. 
+  However, in subglacial channels, water is pressurized by the ice above, and they grow in response to frictional heating of water flowing through them. 
   As a result, water discharge changes mainly result in velocity variations, as the channel geometry evolves slowly (over days).
@@
   The results also indicate that high subglacial sediment transport capacities can occur across a wide range of water discharges.
-  A second set of numerical experiments show that subglacial sediment transport is highly non-linear with respect to water discharge, creating more variability in sediment transport capacity.
+  A second set of numerical experiments shows that subglacial sediment transport is highly non-linear with respect to water discharge, creating more variability in sediment transport capacity.
   Yet, results and  formulations of subglacial sediment transport capacity show that its variability can approach that of subaerial systems when subglacial channel size is in equilibrium with water discharge.
@@
 The Alpine site (labeled \alpine{}) is  Fieschergletscher in the Swiss Alps ($46^\circ\,29'\,07''$ N, $8^\circ\,08'\,3''$ E).
-The water discharge data used here was collected at a $1$\unit{min} interval from May 24, 2014, to October 10, 2014 \citep[Figure~\ref{fig:Qw}\,a][]{felix2022}.
+The water discharge data used here was collected at a $1$\unit{min} interval from May 24, 2014, to October 10, 2014 \citep[Figure~\ref{fig:model_outs}\,a][]{felix2022}.
 
 The Leverett Glacier in Greenland (labeled \icesheet{}) serves as the ice sheet setting.
-Water discharge was measured roughly $2$\unit{km} downstream from the terminus ($67^\circ\,03'\,5''$ N, $50^\circ\,12'\,59''$ W), at a $5$\,\unit{min} time interval from May 28, 2012 to August 8, 2012 \citet[][Figure~\ref{fig:Qw}\,b]{tedstone2013}.
-
-\ian{include figure}
-
-\begin{figure}[hbt!]
-  \centering
-  \includegraphics[width=0.9\linewidth]{Fig2.pdf}
-  \caption{Water discharge from Fieschergletcher (\alpine) and Leverett Glacier (\icesheet).
-    Note differences in variability and quantity of water discharge over the season.
-  }
-  \label{fig:Qw}
-\end{figure}
+Water discharge was measured roughly $2$\unit{km} downstream from the terminus ($67^\circ\,03'\,5''$ N, $50^\circ\,12'\,59''$ W), at a $5$\,\unit{min} time interval from May 28, 2012 to August 8, 2012 \citet[][Figure~\ref{fig:model_outs}\,b]{tedstone2013}.
 
@@
 The model outputs represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
-To generalize these scenarios, \alpine{}  is exemplified by relatively thin ice thickness \citep[$h_{ice}$= $225$\,\unit{m}][]{grab2021}, low water discharge ($\sim\,10$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge at Fieschergletscher (Figure~\ref{fig:Qw}\, b).
-\icesheet{}  is exemplified by thick ice  \citep[$h_{ice}$= $700$\,\unit{m}; ][]{morlighem2017}, high water discharge ($\sim\,300$\,\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge at Leverett glacier (Figure~\ref{fig:Qw}\, b).
+To generalize these scenarios, \alpine{}  is exemplified by relatively thin ice thickness \citep[$h_{ice}$= $225$\,\unit{m}][]{grab2021}, low water discharge ($\sim\,10$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge at Fieschergletscher (Figure~\ref{fig:model_outs}\, a).
+\icesheet{}  is exemplified by thick ice  \citep[$h_{ice}$= $700$\,\unit{m}; ][]{morlighem2017}, high water discharge ($\sim\,300$\,\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge at Leverett glacier (Figure~\ref{fig:model_outs}\, e).
 
@@
   \centering
-  \includegraphics[width=\linewidth]{Fig5.png}
+  \includegraphics[width=\linewidth]{Fig2.png}
   \caption{Model outputs from simulations using the hydrographs in panels a and e for the scenarios \alpine{} (a-d) and \icesheet{} (e-h).
@@
   \centering
-  \includegraphics[width=0.8\linewidth]{Fig6.png}
+  \includegraphics[width=0.8\linewidth]{Fig3.png}
   \caption{
@@
 Variations in \icesheet{} shear stress are comparable to \alpine due to the effects of the evolution of shear stress with subglacial conduit size (Equations ~\ref{eq:dS_dt} and \ref{eq:DW};). 
-The much greater water flux and thus subglacial channel size in \icesheet{} drives the larger width integrated shear stress variations (Figures~\ref{fig:Qw}~and~\ref{fig:Qw}\,c and f).
-
-\ian{is this figure better? ... text reorganized. What about a symbol for cross-referencing}
+The much greater water flux and thus subglacial channel size in \icesheet{} drives the larger width integrated shear stress variations (Figure \ref{fig:model_outs}\,c and f).
 
@@
   \centering
-  \includegraphics[width=0.7\linewidth]{Fig3.pdf}
+  \includegraphics[width=0.7\linewidth]{Fig4.pdf}
   \caption{Standard deviation of detrended model outputs for different smoothing times with different subglacial and subaerial channel shapes and friction factors.
@@
   \centering
-  \includegraphics[width=0.7\linewidth]{Fig4.pdf}
+  \includegraphics[width=0.7\linewidth]{Fig5.pdf}
   \caption{Parameter values compared to variability in model outputs.
@@
   \centering
-  \includegraphics[width=0.7\linewidth]{Fig7.pdf}
+  \includegraphics[width=0.7\linewidth]{Fig6.pdf}
   \caption{Response of sediment transport capacity (orange line) and channel size (bottom row) to a water discharge event (blue line) under different channel conditions.
@@
 
-The numerical experiments above consider the size evolution of subglacial channel and demonstrate that for these hydrographs subglacial sediment transport variability is greater than its subaerial counterpart (Section~\ref{sect:ensemble}).
+The numerical experiments above consider the size evolution of subglacial channels and demonstrate that for these hydrographs subglacial sediment transport variability is greater than its subaerial counterpart (Section~\ref{sect:ensemble}).
 Here, we compare the sediment transport behavior of different channels types as they respond water discharge, channel shape, and hydraulic gradient. 
@@
 
-\ian{comeback here}
 R-channels rarely operate in a steady state with variations in water discharge, especially during severe rain or melt is too short for the channel to reach a steady state (Figures~\ref{fig:model_outs} and \ref{fig:Qw_vari}).
@@
 Thus fluctuations of discharge on short timescales (on the order of a day; Figure~\ref{fig:multi_run}) have the potential to cause conditions with very high sediment transport capacities.
+Alternatively, at low water discharges these channels could become depressurized and transition to subearial flow \citep{perolo2018}.
 These sediment transport capacities variations are of far higher magnitude than those of subaerial channels \citep[][]{alley1997}.
-Note that pipe flow would cause sediment discharge capacity to covary with water discharge (Figure~\ref{fig:chan_types} a).
-Covariance also occurs for steady state R-channels, but not for R-channels where channel area evolves in time (Figures~\ref{fig:Qw_vari} and \ref{fig:chan_types} b).
+Note, however, that pipe flow assumptions would cause sediment discharge capacity to covary with water discharge, as is not the case in normal R-channels (Figures~\ref{fig:Qw_vari}~and~\ref{fig:chan_types}\,a\,and\,b).
+Covariance also occurs for steady state R-channels (Figure \ref{fig:chan_types} d).
 
@@
 
-Greater variations in subglacial sediment transport capacity could cause a supply-limited regime at many glaciers \citep{alley1997}.
+Greater variations in subglacial sediment transport capacity could cause a supply-limited regime at many glaciers due to their high sediment transport capacity under assumed pressurized flow conditions \citep{alley1997}.
 In subglacial systems, sediment's critical shear stress, or threshold at which sediment mobilization occurs, can be reached more frequently and across many water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
@@
 A sporadic relationship between subglacial sediment transport capacities and water discharge occurs due to hysteresis in subglacial channel size (Figure~\ref{fig:Qw_vari}).
-This characteristic limits the use of water discharge as an indicator of sediment discharge capacity in these systems.
+This hysteresis limits the use of water discharge as an indicator of sediment discharge capacity in these systems.
 As a result, characteristics such as bankfull or effective water discharge that link geomorphic work to hydro-climatic conditions could have limited meaning in evaluating subglacial sediment transport \citep{wolman1960,lenzi2006}.
@@
 
-The co-varying relationship between sediment discharge capacity and water discharge in subaerial conditions could be noticeable $\sim20$\,\unit{km} downstream of the Leverett site.
+Clockwise hysteresis loops between sediment concentration and water discharge in subaerial channels suggest that sediment availability is reduced over the event scale \citep[][]{williams1989}.
+These loops are often observed in glacierized catchments and used to suggest that access to subglacial sediment has been limited \citep[e.g.][]{collins1979,willis1996,richards2003,stott2007,delaney2018}.
+Results here suggest that clockwise hysteresis could also be expected in transport-limited regimes in many subglacial environments.
+Here, the smaller channel size on the rising limb would result in greater sediment transport capacity and sediment mobilization (Figure~\ref{fig:chan_types} b).
+The subsequently larger channel size would reduce sediment transport capacity for an equivalent water discharge on the falling limb, creating clockwise hysteresis.
+
+The co-varying relationship between sediment transport capacity and water discharge in subaerial conditions could be noticeable $\sim20$\,\unit{km} downstream of the Leverett site.
 A strong correlation persists between sediment plume size and the Watson River's water discharge into the Kangerlussuaq fjord \citep[Figure~\ref{fig:chan_types}\,c; ][]{chu2009,mcgrath2010}.
@@
 The decreased model output variability beyond $1$--$5$ days of smoothing could result in water discharge appearing to be in equilibrium with the subglacial channel when they are in fact not (Figure~\ref{fig:multi_run}).
-Over this period, water discharge variations could lead to a stronger relationship with sediment discharge capacity (Figure~\ref{fig:chan_types} d and Section ~\ref{sect:scaling}).
+Over this period, water discharge variations could lead to a stronger relationship with sediment transport capacity (Figure~\ref{fig:chan_types} d and Section ~\ref{sect:scaling}).
 Particularly, if water discharge evolves slowly with respect to the subglacial channel size, then it will covary with sediment transport capacity (Figure~\ref{fig:chan_types} d).
 Such a strong relationship would not represent the impact of actual shorter-term fluctuations in water discharge.
-
-Clockwise hysteresis loops between sediment concentration and water discharge suggest that sediment availability is reduced over the event scale \citep[][]{williams1989}.
-These loops are often observed in glacierized catchments and used to suggest that access to subglacial sediment has been limited \citep[e.g.][]{collins1979,willis1996,richards2003,stott2007,delaney2018}.
-Results here suggest that clockwise hysteresis could also be expected in transport-limited regimes in many subglacial environments.
-Here, the smaller channel size on the rising limb would result in greater sediment transport capacity and sediment mobilization (Figure~\ref{fig:chan_types} b).
-The subsequently larger channel size would reduce sediment transport capacity for an equivalent water discharge on the falling limb, creating clockwise hysteresis.
 
@@
 The manuscript's first objective is to establish if water discharge covaries with subglacial sediment transport capacity.
-In subglacial channels, the timing of peak water velocity and sediment discharge capacity occurs before peak water discharge during a discharge event, due to evolving channel size.
-In subaerial channels, the timing of peak sediment discharge capacity and water discharge coincide.
+In subglacial channels, the timing of peak water velocity and sediment transport capacity occurs before peak water discharge during a discharge event, due to evolving channel size.
+In subaerial channels, the timing of peak sediment transport capacity and water discharge coincide.
 Results here suggest that, even in a transport-limited subglacial system, an incoherent relationship between water and subglacial sediment discharge could be expected in most channels.
@@
 the shape of subglacial channels and the response time of subglacial channels to water discharge variations.
-The poor constraints on subglacial sediment size and sorting processes make it difficult to link shear stress to sediment discharge capacity.
-Further quantifying these processes will help to better inform the response of sediment discharge capacity to water discharge forcing in subglacial environments.
+The poor constraints on subglacial sediment size and sorting processes make it difficult to link shear stress to sediment transport capacity.
+Further quantifying these processes will help to better inform the response of sediment transport capacity to water discharge forcing in subglacial environments.
 
@@
   M.A.W. and D.F. provided data from Fieschergletscher and contributed to analysis.
-  All authors provided key inputs in writing and editing the manuscript.} %% this section is mandatory
+  All authors provided key inputs in writing and editing the manuscript.}
 
-\competinginterests{The authors declare no competing interests.} %% This section is mandatory even if you declare that no competing interests are present
+\competinginterests{The one or more of the (co-) authors is a member of The Cryosphere editorial board and is not involved in the review process.}
 
@@
   We benefited from fruitful discussions with J. Irving.
-  We thank two reviewers for constructive and critical reviews on a previous version of the manuscript.
+  We thank two reviewers for constructive and critical comments on a previous version of the manuscript.
 \end{acknowledgements}
-%\newpage
-%\appendix
-
-% \appendixtables   %% needs to be added in front of appendix tables
-
-% \appendixfigures  %% needs to be added in front of appendix figures
-
-%
-
-%% Please add \clearpage between each table and/or figure. Further guidelines on figures and tables can be found below.
-
 \bibliographystyle{copernicus}
@@
 
-%% REFERENCES
-
-%% The reference list is compiled as follows:
-
-% \begin{thebibliography}{}
-
-% \bibitem[AUTHOR(YEAR)]{LABEL1}
-%   REFERENCE 1
-
-% \bibitem[AUTHOR(YEAR)]{LABEL2}
-%   REFERENCE 2
-
-% \end{thebibliography}
-
-%% Since the Copernicus LaTeX package includes the BibTeX style file copernicus.bst,
-%% authors experienced with BibTeX only have to include the following two lines:
-%% 
-%% 
-%% \bibliography{example.bib}
-%% 
-%% URLs and DOIs can be entered in your BibTeX file as:
-%% 
-%% URL = {http://www.xyz.org/~jones/idx_g.htm}
-%% DOI = {10.5194/xyz}
-
-%% LITERATURE CITATIONS
-%% 
-%% command                        & example result
-%% \citet{jones90}|               & Jones et al. (1990)
-%% \citep{jones90}|               & (Jones et al., 1990)
-%% \citep{jones90,jones93}|       & (Jones et al., 1990, 1993)
-%% \citep[p.~32]{jones90}|        & (Jones et al., 1990, p.~32)
-%% \citep[e.g.,][]{jones90}|      & (e.g., Jones et al., 1990)
-%% \citep[e.g.,][p.~32]{jones90}| & (e.g., Jones et al., 1990, p.~32)
-%% \citeauthor{jones90}|          & Jones et al.
-%% \citeyear{jones90}|            & 1990
-
-%% FIGURES
-
-%% When figures and tables are placed at the end of the MS (article in one-column style), please add \clearpage
-%% between bibliography and first table and/or figure as well as between each table and/or figure.
-
-% The figure files should be labelled correctly with Arabic numerals (e.g. fig01.jpg, fig02.png).
-
-%% ONE-COLUMN FIGURES
-
-%% f
-% \begin{figure}[t]
-%   \includegraphics[width=8.3cm]{FILE NAME}
-%   \caption{TEXT}
-% \end{figure}
-% 
-%%% TWO-COLUMN FIGURES
-% 
-%% f
-% \begin{figure*}[t]
-%   \includegraphics[width=12cm]{FILE NAME}
-%   \caption{TEXT}
-% \end{figure*}
-% 
-% 
-%%% TABLES
-%%% 
-%%% The different columns must be seperated with a & command and should
-%%% end with \\ to identify the column brake.
-% 
-%%% ONE-COLUMN TABLE
-% 
-%% t
-% \begin{table}[t]
-%   \caption{TEXT}
-%   \begin{tabular}{column = lcr}
-      %       \tophline
-      %       
-      %       \middlehline
-      %       
-      %       \bottomhline
-      %     \end{tabular}
-      %       \belowtable{} % Table Footnotes
-      %       \end{table}
-      %       
-      %%%       TWO-COLUMN TABLE
-      %       
-      %%       t
-      %       \begin{table*}[t]
-      %       \caption{TEXT}
-      %       \begin{tabular}{column = lcr}
-      %       \tophline
-      %       
-      %       \middlehline
-      %       
-      %       \bottomhline
-      %     \end{tabular}
-      %       \belowtable{} % Table Footnotes
-      %       \end{table*}
-      %       
-      %%%       LANDSCAPE TABLE
-      %       
-      %%       t
-      %       \begin{sidewaystable*}[t]
-      %       \caption{TEXT}
-      %       \begin{tabular}{column = lcr}
-      %       \tophline
-      %       
-      %       \middlehline
-      %       
-      %       \bottomhline
-      %     \end{tabular}
-      %       \belowtable{} % Table Footnotes
-      %       \end{sidewaystable*}
-      %       
-      %       
-      %%%       MATHEMATICAL EXPRESSIONS
-      %       
-      %%%       All papers typeset by Copernicus Publications follow the math typesetting regulations
-      %%%       given by the IUPAC Green Book (IUPAC: Quantities, Units and Symbols in Physical Chemistry,
-      %%%       2nd Edn., Blackwell Science, available at: http://old.iupac.org/publications/books/gbook/green_book_2ed.pdf, 1993).
-      %%%       
-      %%%       Physical quantities/variables are typeset in italic font (t for time, T for Temperature)
-      %%%       Indices which are not defined are typeset in italic font (x, y, z, a, b, c)
-      %%%       Items/objects which are defined are typeset in roman font (Car A, Car B)
-      %%%       Descriptions/specifications which are defined by itself are typeset in roman font (abs, rel, ref, tot, net, ice)
-      %%%       Abbreviations from 2 letters are typeset in roman font (RH, LAI)
-      %%%       Vectors are identified in bold italic font using \vec{x}
-      %%%       Matrices are identified in bold roman font
-      %%%       Multiplication signs are typeset using the LaTeX commands \times (for vector products, grids, and exponential notations) or \cdot
-      %%%       The character * should not be applied as mutliplication sign
-      %       
-      %       
-      %%%       EQUATIONS
-      %       
-      %%%       Single-row equation
-      %       
-      %       \begin{equation}
-      %       
-      %       \end{equation}
-      %       
-      %%%       Multiline equation
-      %       
-      %       \begin{align}
-      %       & 3 + 5 = 8\\
-      %               & 3 + 5 = 8\\
-      %               & 3 + 5 = 8
-                        %     \end{align}
-                        %                         
-                        %                         
-                        %%%     MATRICES
-                        %                         
-                        %                         \begin{matrix}
-                        %                         x & y & z\\
-      %       x & y & z\\
-      %       x & y & z\\
-      %       \end{matrix}
-      %       
-      %       
-      %%%       ALGORITHM
-      %       
-      %       \begin{algorithm}
-      %       \caption{...}
-      %       \label{a1}
-      %       \begin{algorithmic}
-      %       ...
-      %       \end{algorithmic}
-      %       \end{algorithm}
-      %       
-      %       
-      %%%       CHEMICAL FORMULAS AND REACTIONS
-      %       
-      %%%       For formulas embedded in the text, please use \chem{}
-      %       
-      %%%       The reaction environment creates labels including the letter R, i.e. (R1), (R2), etc.
-      %       
-      %       \begin{reaction}
-      %%%         \rightarrow should be used for normal (one-way) chemical reactions
-      %%%         \rightleftharpoons should be used for equilibria
-      %%%         \leftrightarrow should be used for resonance structures
-      %       \end{reaction}
-      %       
-      %       
-      %%%       PHYSICAL UNITS
-      %%%       
-      %%%       Please use \unit{} and apply the exponential notation
-
 \end{document}
-
-      %       \ian{comeback here}
-      %       For sufficiently short periods, channel width and wetted area are fixed in the subglacial system (akin to pipe flow) so that the response of width-integrated shear stress is $w_g\, \tau_{g}\, \propto\,  Q^{2}$ \citep[Table\,~\ref{tab:Qs}; ][]{alley1997}.
-
-      %       R\"othlisberger channels in steady state with water discharge (i.e.
-channel sizes evolves at a faster rate than water discharge), the width-integrated shear stress response to water discharge is $ w_g\,\tau_{g}\, \propto\,  Q^{\sim\frac{4}{5}}$ (Table\,~\ref{tab:Qs}).
-      %       This relationship suggests that when discharge varies more slowly than channel shape subglacial channels have less variable sediment transport capacity than subaerial ones.
-      %       Yet, subglacial channels rarely operate in a steady state \citep{gimbert2016}.
-      %       However, the reduced variability in the  \icesheet{} case compared to the \alpine{} one suggests that it is closer to an equilibrium state (Figure~\ref{fig:Qw_vari}).
-      %       \ian{comeback here}
-      %       The channel width and size variations in subglacial channels presented here also occur in subaerial channels in response to water discharge \citep{phillips2016}, likely over decadal timescales or in response to individual extreme events \citep{slater2013,dean2013}.
-      %       These timescales are considerably longer than the continuous changes to the subglacial channel's response time of days.
-      %       Even so, if channel width changes occur in subaerial channels, then the values of $\alpha$ in Equation~\ref{eq:wcf} could change in time.
-      %       $\alpha$ must be less than $1$, and even this would cause the exponent on $Q$ in the  $w_g\,\tau_{g}$ relationships to remain substantially below that of pipeflow (see Sections \ref{sect:scaling}, Table~\ref{tab:Qs}, Figure~\ref{fig:multi_run}).
-
-With even smaller water discharge variations, the subglacial sediment transport capacity may behave similarly to the subaerial system.
-Note that in pipe flow conditions, maximum sediment transport would coincide with peak water discharge, unlike the evolving channel size alters the timing (Figure \ref{fig:model_outs} and \ref{fig:Qw_vari}).
-
-When sediment availability processes are considered, a glacier's sediment discharge record also represent sediment availability and bedrock erosion from water pressure variations and sliding  \citep[e.g.
-][]{iverson2012,herman2015,delaney2023}.
-The greater amount of in variability of subglacial sediment evacuation can make it difficult to isolate the drivers of sediment export with respect to hydrological and glaciological conditions \citep{delaney2024}.

### 2024-08-15 Delaneyetal2024.tex
Updates from Overleaf

--- 2024-08-08 Delaneyetal2024.tex
@@
   In subaerial channels, water discharge variations are accommodated by flow depth and width changes, along with water velocity.
-  However, in subglacial channels, water is pressurized by the ice above and grow in response to frictional heating of water flowing through them. 
-  As a result, water discharge changes mainly result in velocity variations, as the channel geometry evolves slowly (over days).
+  However, in subglacial channels,  water discharge changes mainly result in velocity variations, as the channel geometry evolves slowly (over days).
+  This processes occurs as water is pressurized by the ice above, and channels grow in response to the frictional heating of water flowing through them. 
   Here, we present formulations of sediment transport capacity in different channel types and apply subglacial and subaerial hydraulics models to hydrographs from an Alpine glacier and the Greenland Ice.
@@
   The results also indicate that high subglacial sediment transport capacities can occur across a wide range of water discharges.
-  A second set of numerical experiments show that subglacial sediment transport is highly non-linear with respect to water discharge, creating more variability in sediment transport capacity.
-  Yet, results and  formulations of subglacial sediment transport capacity show that its variability can approach that of subaerial systems when subglacial channel size is in equilibrium with water discharge.
+  A second set of numerical experiments shows that subglacial sediment transport is highly non-linear with respect to water discharge, creating more variability in sediment transport capacity.
+  Yet, results and formulations of subglacial sediment transport capacity show that its variability approaches that of subaerial systems when subglacial channel size is in equilibrium with water discharge.
   The implications of these findings are discussed in the context of sediment discharge from glaciers with different hydro-climatic forcings.
@@
 
-Despite the different physical processes between the subaerial and subglacial systems, observed water discharge and sediment export are often compared in contemporary glacierized catchments \citep[e.g.][]{willis1996,hodson1998,pearce2003,richards2003,swift2005,chu2009,tedstone2012,chu2012,overeem2017,delaney2018,swift2021,lu2022,andresen2024}.
+Despite the different physical processes between the subaerial and subglacial systems,  water discharge and sediment export are often compared in contemporary glacierized catchments \citep[e.g.][]{willis1996,hodson1998,pearce2003,richards2003,swift2005,chu2009,tedstone2012,chu2012,overeem2017,delaney2018,swift2021,lu2022,andresen2024}.
 Many of these studies discuss the disparity between water discharge and sediment export, especially given the role of sediment access or production.
@@
 These formulations extend the relationships presented in \citet{alley1997} and illustrate the sediment transport capacity behavior of different channel assumptions.
-Findings indicate differences in the relationships amongst water discharge, channel geometry, water velocity and sediment transport capacity in subglacial and subaerial channels.
+Findings indicate differences in the relationships amongst water discharge, channel geometry, water velocity, and sediment transport capacity in subglacial and subaerial channels.
 The manuscript then discusses the implications of variability in sediment transport capacity from glaciers and the interpretation of sediment transport records.
@@
 These experiments assumes  a subglacial channel with $\beta=\frac{\pi}{6}$ and a subaerial channel with $\alpha = \frac{1}{3}$ and slope of $0.02$ (Table~\ref{table:vpm}).
-Both friction factors $f_r$ and $f_f$ are tuned so that reasonable water velocities \citep[$\sim\,1- 1.5\,$\unit{m}\,\unit{s}$^{-1}$][]{werder2010b,chandler2013} occur for both \alpine{} and \icesheet{}.
+Both friction factors $f_r$ and $f_f$ are tuned so that reasonable water velocities \citep[$\sim\,1- 2\,$\unit{m}\,\unit{s}$^{-1}$][]{werder2010b,chandler2013} occur for both \alpine{} and \icesheet{}.
 To test the covariance between sediment transport capacity and water discharge, model outputs are compared to water discharge from the glaciers using Spearman rank correlation.
@@
   \caption{Model outputs from simulations using the hydrographs in panels a and e for the scenarios \alpine{} (a-d) and \icesheet{} (e-h).
-    Black (gray) lines in a and e represent the subglacial channel size (water discharge).
+    Black (gray) lines in a and e represent the water discharge (subglacial channel size).
     Blue (orange) lines represent outputs from the subglacial (subaerial) channel.
@@
 As a result, water velocity, as opposed to channel growth, accommodates increases in water discharge, and velocity variability increases.
-Higher values of $f_i$ allow faster channel growth that accommodates increase in water discharge, reducing velocity variability.
+Higher values of $f_i$ allow faster channel growth that adapts to increases in water discharge, reducing velocity variability.
 Smaller values of channel factor $\beta$, creating low and broad channels, result in more variability in width-integrated shear stress.
@@
 
+While variability in model outputs can increase with certain parameters in the subaerial model, they do not exceed variability in subglacial outputs.
 Smaller values of subaerial channel shape factor $\alpha$, or a channel cross-section shape closer to a slot canyon, result in greater variability in subaerial velocity and shear stress that can approach the variability of subglacial values (Figure~\ref{fig:params}).
 Steeper subaerial channel slopes $\nabla z_c$ result in greater variability in both shear stress and width-integrated shear stress but only approach the subglacial channel's variability in the \alpine case (Figure~\ref{fig:params}).
-Greater values of subaerial friction factors $f_p$ resulting in greater variability shear stress and width-integrated shear stress. 
-Yet, they to not exceed the variability of subglaical channels. 
+Greater values of subaerial friction factors $f_p$ result in greater variability of shear stress and width-integrated shear stress. 
 We note that these parameter values span a commonly accepted range.
@@
 
-The numerical experiments above consider the size evolution of subglacial channel and demonstrate that for these hydrographs subglacial sediment transport variability is greater than its subaerial counterpart (Section~\ref{sect:ensemble}).
-Here, we compare the sediment transport behavior of different channels types as they respond water discharge, channel shape, and hydraulic gradient. 
+The numerical experiments above consider the size evolution of subglacial channels and demonstrate that for these hydrographs subglacial sediment transport variability is greater than its subaerial counterpart with the same hydrological forcing.
+Here, we scale formulations of subglacial sediment transport capacity to compare the behavior of different channel types as they respond to water discharge, channel shape, and hydraulic gradient. 
 This yields insights into the conditions where subglacial sediment transport capacity could exhibit similar variability as subaerial channels.
@@
 However, the formulations establish the scaling of these different channel types  .
-The sediment transport capacity is calculated with a given water discharge and hydraulic gradient for three different sediment transport capacity formulas: Meyer-Peter M\"uller  \citep[MPM; ][]{meyer1948}, Engelund and Hansen \citep[EH; ][]{engelund1967}, and Bagnold \citep{bagnold1980}; additionally width-integrated shear stress is assessed, as used as proxy for sediment transport above.
+The sediment transport capacity is calculated with a given water discharge and hydraulic gradient for three different sediment transport capacity formulas: Meyer-Peter M\"uller  \citep[MPM; ][]{meyer1948}, Engelund and Hansen \citep[EH; ][]{engelund1967}, and Bagnold \citep{bagnold1980}; for comparison, with the result above additionally width-integrated shear stress is also assessed.
 
@@
 \end{equation}
-respectively, for sediment transport conditions well above the threshold of sediment motion.
+respectively, for sediment transport conditions well above the threshold of sediment motion, where the relationships hold true.
 
@@
 R-channels rarely operate in a steady state with variations in water discharge, especially during severe rain or melt is too short for the channel to reach a steady state (Figures~\ref{fig:model_outs} and \ref{fig:Qw_vari}).
-In these cases with high water discharge variability, channels can behave more like a pipe of fixed cross-section.
+With high water discharge variability, channels can behave more like a pipe of fixed cross-section.
 Here the cross-section responds to a characteristic discharge, but variations in water discharge deviate substantially from the flow conditions responsible for the channel size \citep[i.e.
@@
 Sediment exhaustion through repeatedly crossing the mobilization threshold may explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on the glaciers' basal shear stress, a proxy for bedrock erosion, rather than glacier melt \citep{overeem2017}.
-More generally, the same process could also result in sediment discharge's strong dependence on sediment production from glacial sliding, itself dependent of basal shear stress of the glacier \citep{herman2015,koppes2015}.
+More generally, the same process could also result in sediment discharge's strong dependence on sediment production from glacial sliding, itself dependent on basal shear stress of the glacier \citep{herman2015,koppes2015}.
 
-While transport-limited states likely do not occur at many glaciers \citep[e.g.][]{alley1997}, abundant sediment could persist underneath some glaciers, potentially creating a transport-limited regime \citep[e.g.][]{walter2014,stevens2022,delaney2022}.
-At these glaciers, the great variability in subglacial sediment transport capacity may make it difficult to link sediment discharge to hydrology, especially when peak events occur \citep{cowan1988,delaney2018,lu2022}.
-The results here suggest that different channel sizes with the same hydrology forcing can result in very different sediment transport capacities, making it challenging to establish the effects of individual extreme precipitation or melt events without establishing the antecedent state of the subglacial drainage system.
+While transport-limited states likely do not occur at many glaciers \citep[e.g.][]{alley1997}, abundant sediment could persist underneath some glaciers, potentially creating conditions for a transport-limited regime \citep[e.g.][]{walter2014,stevens2022,delaney2022}.
+At these glaciers, the great variability in subglacial sediment transport capacity may make it difficult to link potential transport-limited sediment discharge to hydrology, especially when peak events occur \citep{cowan1988,delaney2018,lu2022}.
+The results here show that different channel sizes with the same hydrology forcing can result in very different sediment transport capacities, making it challenging to establish the effects of individual extreme precipitation or melt events without establishing the antecedent state of the subglacial drainage system.
 
@@
 
-The co-varying relationship between sediment discharge capacity and water discharge in subaerial conditions could be noticeable $\sim20$\,\unit{km} downstream of the Leverett site.
-A strong correlation persists between sediment plume size and the Watson River's water discharge into the Kangerlussuaq fjord \citep[Figure~\ref{fig:chan_types}\,c; ][]{chu2009,mcgrath2010}.
+The co-varying relationship between sediment discharge capacity and water discharge in subaerial conditions could be noticeable $\sim20$\,\unit{km} downstream of the Leverett site where the Watson River enters the Kangerlussuaq fjord.
+A strong correlation persists between sediment plume size and the Watson River's water discharge, potentially due to the adjustment to subearial sediment transport conditions along the river's length from the glacier outlet \citep[Figure~\ref{fig:chan_types}\,c; ][]{chu2009,mcgrath2010}.
 In contrast, in marine-terminating glacier catchments, a less consistent relationship may occur between water discharge or melt extent and sediment plume size \citep{chu2012,tedstone2012}.
@@
 
-Different hydraulic gradients control the velocity and sediment transport capacity in the subglacial and subaerial cases (Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}). 
+Different hydraulic gradients control the water velocity and sediment transport capacity in the subglacial and subaerial cases (Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}). 
 The subglacial one is generally much steeper \citep{alley1997}.
@@
 
-Width-integrated shear stress is examined in the two numerical experiments as it does not have a dependence on grain size, unlike sediment transport relationships \citep[][Section \ref{sect:scaling}]{meyer1948}.
+Width-integrated shear stress is examined in the two numerical experiments as it does not have a dependence on grain size, unlike sediment transport relationships \citep[e.g.
+][Section \ref{sect:scaling}]{meyer1948}.
 This makes comparison between the two systems simpler. 
@@
 
-The manuscript's first objective is to establish if water discharge covaries with subglacial sediment transport capacity.
+The manuscript's first objective is to establish if water discharge co-varies with subglacial sediment transport capacity.
 In subglacial channels, the timing of peak water velocity and sediment discharge capacity occurs before peak water discharge during a discharge event, due to evolving channel size.
@@
   A.J.T. helped interpret findings from Leverett Glacier and guided experiment design.
-  M.A.W. and D.F. provided data from Fieschergletscher and contributed to analysis.
+  M.A.W. and D.F. provided data from Fieschergletscher and contributed to the analysis.
   All authors provided key inputs in writing and editing the manuscript.} %% this section is mandatory

### 2025-01-22 Delaneyetal2024.tex
beginning revisions

--- 2024-08-15 Delaneyetal2024.tex
@@
 As a result, water velocity, as opposed to channel growth, accommodates increases in water discharge, and velocity variability increases.
-Higher values of $f_i$ allow faster channel growth that accommodates increase in water discharge, reducing velocity variability.
+Higher values of $f_i$ allow faster channel growth that accommodates increases in water discharge, reducing velocity variability.
 Smaller values of channel factor $\beta$, creating low and broad channels, result in more variability in width-integrated shear stress.
@@
 Steeper subaerial channel slopes $\nabla z_c$ result in greater variability in both shear stress and width-integrated shear stress but only approach the subglacial channel's variability in the \alpine case (Figure~\ref{fig:params}).
-Greater values of subaerial friction factors $f_p$ resulting in greater variability shear stress and width-integrated shear stress. 
-Yet, they to not exceed the variability of subglaical channels. 
+Greater values of subaerial friction factors $f_p$ result in greater variability of shear stress and width-integrated shear stress. 
+Yet, they do not exceed the variability of subglaical channels. 
 We note that these parameter values span a commonly accepted range.
@@
 The numerical experiments above consider the size evolution of subglacial channels and demonstrate that for these hydrographs subglacial sediment transport variability is greater than its subaerial counterpart (Section~\ref{sect:ensemble}).
-Here, we compare the sediment transport behavior of different channels types as they respond water discharge, channel shape, and hydraulic gradient. 
-This yields insights into the conditions where subglacial sediment transport capacity could exhibit similar variability as subaerial channels.
+Here, we compare the sediment transport behavior of different channel types as they respond to water discharge, channel shape, and hydraulic gradient. 
+This yields insights into the conditions where subglacial sediment transport capacity could exhibit similar variability as in subaerial channels.
 
@@
 These formulations do not allow us to evaluate non-steady state R-channels presented above (Section~\ref{sect:sub_mode}, Figure~\ref{fig:chan_types} b).
-However, the formulations establish the scaling of these different channel types  .
+However, the formulations establish the scaling of these different channel typesc.
 The sediment transport capacity is calculated with a given water discharge and hydraulic gradient for three different sediment transport capacity formulas: Meyer-Peter M\"uller  \citep[MPM; ][]{meyer1948}, Engelund and Hansen \citep[EH; ][]{engelund1967}, and Bagnold \citep{bagnold1980}; additionally width-integrated shear stress is assessed, as used as proxy for sediment transport above.
@@
 
-Several aspects of the models and experiment design make the comparison of the subaerial and subglaical systems difficult. 
+Several aspects of the models and experiment design make the comparison of the subaerial and subglacial systems difficult. 
 The lumped nature of the models means that they operate independently of the upstream drainage network.

### 2025-02-03 revised_v1/response_R1.tex
small change

\documentclass[11pt]{article}
\usepackage{lscape}
\usepackage[table]{xcolor}% http://ctan.org/pkg/xcolor
\setlength{\parindent}{0pt}
\usepackage [colon]{natbib}
\usepackage {comment}
\renewcommand{\familydefault}{\sfdefault}
\usepackage[a4paper, margin=2.40cm]{geometry}
\usepackage{doi}
\usepackage{multirow}
\usepackage[T1, OT1]{fontenc}
\DeclareTextSymbolDefault{\dh}{T1}
\renewcommand{\rmdefault}{phv} % Arial
\renewcommand{\sfdefault}{phv} % Arial
\usepackage[modulo]{lineno}
\usepackage{wrapfig}
% \linenumbers
\usepackage{hyperref}
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\usepackage{amsmath}
\usepackage{float}
% \errorcontextlines=3
\usepackage[explicit]{titlesec}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{authblk}
\usepackage{setspace}
\usepackage[export]{adjustbox}
\renewcommand{\bibfont}{\footnotesize}
\usepackage{bibentry}
\usepackage{tikz}
\usepackage[inline,shortlabels]{enumitem}
\usepackage[all]{nowidow}

\newcommand{\TODO}[1]{\textbf{TODO #1}}
\newcommand{\ian}[1]{{\textbf{\color{blue}Ian says:} \color{blue} #1} }
\newcommand{\leif}[1]{{\textbf{\color{red}Leif says:} \color{red} #1} }
\newcommand{\frederic}[1]{{\textbf{\color{red}Frederic says:} \color{red} #1} }
\newcommand{\aref}[1]{\textbf{Reference #1}}
\newcommand{\alpine}{\textit{ALPINE}\,}
\newcommand{\icesheet}{\textit{ICESHEET}\,}
\title{}
\author{}

\begin{document}

\vspace{.1cm}

Best regards,

\vspace{.1cm}

Ian Delaney on behalf of all authors

\textbf{General Comments}

\begin{itemize}

\item \textbf{In my view the paper is interesting and relevant, and hence worthy of publication. I have a couple
    of general observations, followed by a list of minor suggestions.
    My general observation relates to the structure of the main findings reported by the manuscript.
    In reading the Results and Discussion I was sometimes lost as to what I was learning that was new.
    I think this was mainly because the structure lacked a clear and logical progression from one
    general finding through progressively more specific findings. This may well stem an initial
    imperfect expression of the objectives.}

  \textit{We thank the Reviewer for the positive assessment of our work and take note of the structural issues along with the organizational ones. }

\item  \textbf{Here, most of what is novel seems to fgall under Objective
    2 and not under Objective 1. Indeed, isn’t Objective 1 (‘to establish whether sub-seasonal water
    discharge can co-vary with sediment transport capacity in subglacial systems’) already established
    beyond reasonable doubt? To me, and I believe the introduction to this manuscript, it clearly
    ‘can’… (note the objective refers to ‘can it’, and not to ‘does it’ [to which the answer is also almost
    certainly ‘yes’ but perhaps with a little more room for maneuver]). I would recast the objectives
    to reflect better the new material presented in this manuscript (closer to objective 2 in fact, which
    I feel could readily be sub-divided). Having considered this, I feel the progression of the Results
    could be far more accessible to the reader and allow the manuscript to focus on a few really key
    points (which I think are currently slightly lost in the density of the presentation).
    Interpretation is presented within each Results section (e.g., see ‘due to… on L174). This
    continues throughout Results, and I would separate all of the explanation and hence
    interpretation out form the results.}

  \textit{We value this comment greatly. Upon consideration, we have chosen to rephrase  objective one as ``to establish the hydrological conditions where sediment transport capacity co-varies
    with water discharge in subglacial systems.'' 
    In our view this objective focus the attention on the differences between the ALPINE and ICESHEET scenarios. It also is practical when evaluating the differences in smooth of water discharge, as discussed in the next point.
    We appreciate the need to remove interpretation from the results section. We have done this. Additionally, Results Section 4.3 with the algebraic formulations has been moved to the discussion, as we think the pipe flow and steady state R-channel scenarios serve as useful end members through which to discuss the results.
  }

\item \textbf{There are some very important points in Results, which I would attach some key data to
    and make more explicit here and in the Conclusion and Abstract. I’m thinking for example of
    associating the (increased) hysteresis as a function of discharge with a headline \% change in
    sediment transport capacity. E.g., “thus, a more variable, but typical, Alpine Q record can
    transport up to *\% greater sediment per cumec than a less variable ice sheet-type Q” and/or “the
    greater variability in Q from Alpine than from ice sheet style Q can account by an offset of peak
    sediment discharge from peak meltwater discharge by up to * \% of the cycle involved”. At present,
    I feel the manuscript lacks this incisive illustration to bring the scale of the results home to the
    reader.}

  \textit{This is an excellent comment. We have worked to address this need for a clearer hysteresis explanation.
    One challenge with the approach recommended by the reviewer is that sediment transport capacity scales very non-linearly with water discharge, especially in subglacial systems. This is shown in Section 4.3.
    As a result, findings across hydrographs would not really be comparable even if they were scaled with a discharge quantity.
    Instead, we will add several paragraphs that to section 4.1 that discuss The role of water discharge variability on hysteresis.

  \textit{Here the aim is to evaluate the effects of smoothing hydrographs, yielding insights in to the hydrological conditions where discharge and sediment transport capacity can co-vary.
    Note that the smoothed hydrographs used for the analysis here are presented in the supplement. 
    A version of the new paragraphs is as follows.}

Water discharge and sediment transport characteristics correlate better in \icesheet{} with lower diurnal variations in water discharge compared to the \alpine{}.
To evaluate the impact of water discharge variability, we smoothed the two hydrographs across different periods and evaluated the Spearman rank correlation between sediment transport characteristics and water discharge.
We assume that a higher rank correlation indicates a reduced amount of hysteresis and behavior more similar subaerial channels (Figure~3).

Smoothing water discharge over periods longer than one day causes a substantial increase in rank correlation (Figure~5).
This increase occurs as diurnal variations are removed (Figures S1-S14).
Correlations in sediment transport characteristics in \icesheet{} remain higher than \alpine, which has more water discharge variations even when smoothed compared to \icesheet{} (Figure~S1-S14).

\begin{figure}[hbt!]
  \centering
  \includegraphics[width=0.8\linewidth]{../Fig4.png}
  \caption{Spearman rank correlation between water discharge over a smoothed period and sediment transport characteristics (a) water velocity, (b) shear stress, (c) width-integrated shear stress. We assume that higher correlation scores approach subaerial behavior.
  }
  \label{fig:corri}
\end{figure}

\item \textbf{Less importantly, but something I’d like to check, is that the fundamental conclusions are
    applicable more generally. Since they derive from modelling based on two ‘type’ datasets, there
    may be a possibility that not all key findings hold for certain other variations of subglacial
    discharge pattern. I have every reason to believe they do, but I feel the manuscript should report
    that this has been done more widely and the key results/patterns still hold. I don’t think there’s
    any need to include any such results into a revised manuscript – but I would just make the
    statement that while, for simplicity of illustration, two typical but contrasting datasets are
    explored here, the model has been run with other data sets and the key do hold more generally.}

  \textit{ We thank the reviewer for this comment as it, in part, encouraged us to add the additional section above with the water discharge smoothing.
    The other hydrological scenario we are aware of not explicitly considered here is Antarctica, where little surface melt occurs.
    Thus, we expect variations in water discharge to be minimal.
    The following sentence has been added to the conclusion:}
    ``Water discharge and sediment transport capacity covariance could be possible when water discharge varies at a slower rate than subglacial channel size, such as in Antarctica with minimal surface melt input.''

\item   \textbf{
    Incidentally, I would also include an explicit disclaimer early on that the analysis only relates to
    pressurized subglacial flow and not to open-channel flow prominent approaching the terminus of
    many valley glaciers etc.}

   \textit{ We will add the following sentence to the model implementation section:}
``We also note that the subglacial results apply to pressurize subglacial channels, not depressurized one that can occur below glacier termini \citep{perolo2018}''
\end{itemize}

\textbf{Specific Comments}

\begin{itemize}
\item \textbf{Title:}The title states the obvious from prior research and I would recast it to
  reflect more accurately the new and original contribution of this
  manuscript. Also, if the authors elect to not change the title I would change
  the existing wording to ‘transport capacities respond differently’.
  Grammatically, this is not clear given the two capacities noted earlier in the
  title, but stylistically I think ‘capacities respond’ reads better than does
  ‘capacity respond’. (I would change the whole thing anyway… Why not start
  by considering ‘Sediment transport capacity response to variations in water
  discharge in pressurized subglacial channels’)

  \textit{Excellent suggestion, title follows the reviewer's suggestion: Sediment transport capacity response to variations in water discharge in pressurized subglacial channels}

\item \textbf{3:} ‘full’ subglacial channels?

  \textit{Done.}

\item \textbf{5:} There is no need to mention ‘over days’ here (indeed, it is misleading
  anyway without quantification of how much change as a function of time).
  I would just leave it as changes slowly (or slowly relatively to variations in
  water Q).

  \textit{Done.}

\item \textbf{7:} ‘Sheet’ is missing (the authors really should have picked up on this obvious
  typo)

  \textit{Done.}

\item \textbf{8:} This hysteresis causes (no need to qualify the type of hysteresis since the
  use of ‘this’ serves the purpose)

  \textit{Done.}

\item \textbf{64:} I think there is earlier use of transport-limited (in which case the definition
  should be presented there, and not here)

  \textit{We thank the reviewer for the careful read. The definition was presented in Line 28, and thus we will remove the definition here.}

\item \textbf{75-76 and 80-81} I’d remove summary of the results of this study from the Introduction. They
  are not yet established.

  \textit{These lines will be removed.}

\item \textbf{70-71} In phrasing objective 1 (in more than one place in the manuscript),
  shouldn’t the dependent variable come first? Thus, I would write it as ‘to
  establish whether sediment transport capacity can co-vary with water
  discharge in subglacial systems’.

  \textit{Given the discussion above, the text will read: ``to establish the hydrological conditions where  sediment transport capacity co-varies with water discharge in subglacial systems''}

\item \textbf{79} Given the apparent relevance of the Alley et al. study, I feel it’s key findings
  should be presented in detail here to provide more direction to the specific
  objectives of this study.

  \textit{ We believe that this is discussed in Lines 51--54 in the introduction.}

\item \textbf{159} The parameter values are not purely random (but random within certain
  sensible ranges, no?).

  \textit{The text will now read: ``with random parameter values in a range''}

\item \textbf{171-} I get a little muddled right at the start here – not helped by the sub-heading,
  which I would recast as “Influence of subglacial channel size on the timing
  and variability of…”. The section (and others following) also combines
  results with explanation (i.e., interpretation) which I feel also adds
  unnecessarily to the difficulty in following the progression. I would separate
  results from interpretation.

  \textit{This section will be retitled:} ``Timing of subaerial and subglacial sediment transport capacity variations.''
  \textit{This discusses the role water discharge on hysteresis in water velocity or sediment transport capacity. Thus dealing with objective 1 of the manuscript. We have also noted the comment about results and interpretation. These have been removed from the manuscript}

\item \textbf{172} This narrative jumps straight in with the ambition to ‘quantify the sources
  of increased variability…’ but the nature of that increased variability has not
  yet be established or illustrated. This really needs building up more
  progressively – bringing the reader through with the first-order model
  results. Then go on to explore more and more detailed influences and
  relationships. Most of this is, in fact, in Figure 2 – but the reader is not
  guided through the fundamental relationships here.

\ian{Comeback here}

\item \textbf{173} Change rather than evolution? (elsewhere too)

  \textit{We appreciate the comment. We believe that given the nature of subglacial channels as describe in Equation 3, evolution is a more accurate description. To us, ``change'' could include a modification for any number of reasons. However,
    ``evolution'' conotates the dependence on anticedent conditions which we believe to be more accurate in this case. }

\item \textbf{Fig 2 caption} Is it really an arbitrary y axis scale? If it is truly arbitrary, does it even need
  a scale range?

  \textit{We believe that description of the insets' axis is needed as these lines have the same axis in the main plots. However, their very different values mean that in makes sense to adjust values to examine their changes during the extreme events. The text has been modified slightly to read: ``different y-axis ranges for subaerial and subglacial values.''}

\item \textbf{179-183} Is this not already established? If so, it should be presented clearly in the
  Introduction and developed as necessary here. It is also interpretation.

  \textit{ To the best of our knowledge this is the first time that this finding is being presented.
    We would welcome any citations or references that suggest otherwise. We believe that the content in this paragraph describes the response to water discharge in Figure 2. We agree that the reference to the Methods section, gave the impression of the interpretation. The paragraph will be adjusted slightly to read:}

    Peaks in subaerial model outputs occur coincident with peaks in water discharge (Figure~2).
    In the subglacial channel, peaks in model outputs generally occur when water discharge increases, but before the maximum water discharge.
    As the water discharge stabilizes at its peak, channel growth continues (Figure~2 a, e), causing water velocity and other model outputs to decrease from their peak values (Figure~2 b-d, f-h).
    Subglacial sediment transport capacity is greatest on the hydrograph's rising limb, relative to the falling limb, creating a hysteresis effect.''}

\item \textbf{Fig 3 (and Fig 4)} Shouldn’t the dependent variable be plotted on y and the independent
  variable on x? I for one at least think this way around and therefore find
  these plots a little confusing.
  Is there a need for the axis labels to be at an angle rather than parallel to
  the axis? If not, I would reorientate them parallel in all cases.

  \textit{The axes will be changed in both figures}

\item \textbf{193-4} This is interpretation.

  \textit{This sentence will be removed.}

\item \textbf{195} The reference to ‘increased variability’ needs to make it clear what property
  is being referred to. Like the previous sub-heading, this sub-heading is a
  statement of findings before those results have been presented. I would
  avoid this and present it as the less assuming: ‘Influence of…’

  \textit{The section head will be modified to read:} Variability in sediment transport capacity across a range of channel shapes, slopes, and friction values

\item \textbf{205-6} Interpretation in Results

  \textit{Removed.}

\item \textbf{210-11} Interpretation in Results

  \textit{ Sentence will be changed to:''Smaller values of channel factor $\beta$, creating low and broad channels where the channel width grows more quickly in response to water discharge increases as compared to a semi-circular channel with $\beta = \pi$ (Equation~4).
''}

\item \textbf{220 Typo (subglacial)}

  \textit{Fixed.}

\item \textbf{Section 4.3} A lot of this reads more like methods and background modelling operation
  and/or equation terms than Results of the modelling. If it is all results then
  the key findings are difficult to pull out from the density of the presentation
  – perhaps some of it could be omitted or moved to supplementary?

  \textit{We have considered the comments here, and with the editors support, think it fits best here.
    In a previous version of the manuscript, it was included in the supplementary material. However, after those reviews, we believe that some misunderstandings emerged. We believe that Tables 2 and 3 in the section are valuable results in evaluating the sensitivity of sediment transport capacity to water discharge under different assumptions.
    The importance of these findings are discussed in lines xxx yyy.
  }

  \textit{We will modify the introductory text to read: ``The numerical experiments above consider the size evolution of subglacial channels and demonstrate that for these hydrographs subglacial sediment transport variability is greater than its subaerial counterpart (Section~4.3). 
Additionally, results demonstrate the impact of water discharge variability on sediment transport capacity.
Here, we compare the sediment transport behavior of different channel types as they respond to water discharge, channel shape, and hydraulic gradient.''
}

\item \textbf{Fig 5} I find the figure too small. If larger the font would be easier to read and the
  y axes could have the property written out rather than represented by
  symbols. I’m sure there is no need for it to be this small.

  \textit{We will increase the size of the axis labels.}

\item \textbf{232} Typo? (‘typesc’)
  \textit{Removed.}

\item \textbf{235} The text: ‘…as used as proxy for…’) doesn’t make sense (or at least could be
  worded more clearly).

\textit{Text will read: ''additionally width-integrated shear stress is assessed, instead of sediment transport, as above.''}

\item \textbf{284} (maybe
  elsewhere too)
  The wording should be “greater …. than” and not ‘greater… compared
  with…).

  \textit{Done.}

\item \textbf{297} “…itself dependent of…” doesn’t make sense to me

  \textit{Text now reads: ''which is dependent''}

\item \textbf{298} I don’t believe there is sufficient evidence to make the claim: ‘While
  transport limited states likely do not occur at many glaciers’. In my own
  experience most glacier do have at least parts of the bed that are sediment-
  rich and hence have the capacity to be transport limited.

  \textit{We agree with  about the lack of data. This sentence will been removed. }

\item \textbf{302-4} I would try to pull some key quantification out here and present it as a
  headline figure to illustrate the magnitude of the effect.

\item \textbf{319} I’m not sure ‘sporadic’ is the right word here. Please check you mean this.

  \textit{We will replace ``sporadic'' with ``variable''.}

\item \textbf{321-22} I see the point being made here but it’s not black or white. Surely there is
  some control exerted by discharge; it’s just that it is complicated by the
  additional forces in the subglacial scenario. The way this is presented here
  may be technically ok but the reader could be forgiven for coming away
  with the feeling that there is no solid relationship between Q and sediment
  transport capacity is these cases – which would not be accurate. I feel it
  would be particularly helpful if the effect could be quantified in some
  generalized or headline way in terms of the likely (or maximum if you
  prefer) effect on a typical alpine and/or ice sheet hydrograph. That figure
  could then also go into the Conclusions and Abstract.

  \textit{The sentence will read: ``This hysteresis can limit water discharge as an indicator of sediment discharge capacity in these systems, especially when water discharge is highly variable and out of equilibrium with subglacial channel size.''}

\item \textbf{394} Surely the relationship can be characterized a little more usefully than
  simply to state that it is ‘incoherent’. This also relates to my point above
  relating to lines 321-22. Can this be characterized accurately or
  systematically so the reader has a little more to learn in terms of the nature
  and magnitude of the effects?

\textit{This paragraph will read: ``The manuscript's first objective is to establish the conditions where water discharge covaries with subglacial sediment transport capacity.
In subglacial channels, the timing of peak water velocity and sediment transport capacity occurs before peak water discharge during a discharge event, due to evolving channel size.
In subaerial channels, the timing of peak sediment transport capacity and water discharge coincide.
Results here suggest that, even in a transport-limited subglacial system, an variable relationship between water and subglacial sediment discharge.
Water discharge variations in both Greenland Ice Sheet and Alpine are variable enough to cause this variable relationship.
This variable relationship presents a challenge in linking hydro-climatic conditions or events to sediment export from glaciers.
Water discharge and sediment transport capacity covariance could be possible when water discharge varies at a slower rate than subglacial channel size, such as Antarctica with minimal surface melt input.
Results also suggest that water discharge records averaged over periods longer than $12$ hours from Alpine and Ice Sheet hydrographs show substantial impacts on subglacial sediment transport capacity characteristics.''}

\item \textbf{407-10} It would not do any harm here to point to advances in subglacial sediment
  tracking, which speaks to this issue. I am well familiar with Jenkins’ work –
  but there may well be others too.

  \textit{Excellent comment that we should have already accounted for. Jenkins et al. 2023 will be added and referenced.}

\end{itemize}

\bibliographystyle{apalike} 
\bibliography{../PaperLib.bib}
\end{document}

### 2025-02-03 revised_v1/response_R2.tex
small change

\documentclass[11pt]{article}
\usepackage{lscape}
\usepackage[table]{xcolor}% http://ctan.org/pkg/xcolor
\setlength{\parindent}{0pt}
\usepackage [colon]{natbib}
\usepackage {comment}
\renewcommand{\familydefault}{\sfdefault}
\usepackage[a4paper, margin=2.40cm]{geometry}
\usepackage{doi}
\usepackage{multirow}
\usepackage[T1, OT1]{fontenc}
\DeclareTextSymbolDefault{\dh}{T1}
\renewcommand{\rmdefault}{phv} % Arial
\renewcommand{\sfdefault}{phv} % Arial
\usepackage[modulo]{lineno}
\usepackage{wrapfig}
% \linenumbers
\usepackage{hyperref}
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\usepackage{amsmath}
\usepackage{float}
% \errorcontextlines=3
\usepackage[explicit]{titlesec}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{authblk}
\usepackage{setspace}
\usepackage[export]{adjustbox}
\renewcommand{\bibfont}{\footnotesize}
\usepackage{bibentry}
\usepackage{tikz}
\usepackage[inline,shortlabels]{enumitem}
\usepackage[all]{nowidow}

\newcommand{\alpine}{\textit{ALPINE}\,}
\newcommand{\icesheet}{\textit{ICESHEET}\,}
\newcommand{\TODO}[1]{\textbf{TODO #1}}
\newcommand{\ian}[1]{{\textbf{\color{blue}Ian says:} \color{blue} #1} }
\newcommand{\leif}[1]{{\textbf{\color{red}Leif says:} \color{red} #1} }
\newcommand{\frederic}[1]{{\textbf{\color{red}Frederic says:} \color{red} #1} }
\newcommand{\aref}[1]{\textbf{Reference #1}}
\title{}
\author{}

\begin{document}

\textbf{General Comments}

\begin{itemize}

\item \textbf{
    The trouble is, that while the first term on the right hand side of (3) makes sense to me, the second does not, and the reason for this is that the closure term involves the effective pressure and thus the hydraulic head, but not its gradient, so I am a bit sceptical about this, because the structure of the Nye model seems to have been altered.
    You might say that the distinction disappears in the lumping, but it seems to me that if you have an upstream head h- and a downstream head h+, the first term on the right of (3) will involve the difference, but the second will involve the average, and there are two independent quantities involved.
    I suppose you can get around this by saying, oh, the outlet head or more accurately the effective pressure is zero, and perhaps that is what is done.
    But that is a bit disingenuous, because the outflow becomes open channel flow somewhere upstream of the actual mouth of the stream.
    Now I went and looked up this Werder 2010 paper, and you can see in its equation (4) the same distinction I am making here.
    I suppose my overall view is that (I think) the Clarke paper intended to use the flow elements of resistors, etc., as ingredients of a larger scale flow path, and this lumping of a whole channel is an extremely coarse thing to do, and looks a bit like avoiding confronting the spatially-dependent physics simply because it’s too hard.
    So I think you can rescue this aspect of the model presentation, but a bit more exposition would help, in particular the figure 2or equivalent of the 2010 paper would help.}

  \textit{We thank the reviewer for this comment, as it demands clarification. The text now reads:}
  The evolution of subglacial channel size $S_g$ is given as
  \begin{linenomath*}
    \begin{equation}
      \label{eq:dS_dt}
      \frac{\partial S_g}{\partial t} = C_1 \frac{Q \Delta h}{l} - C_2 \left(h_{o}-\overline{h}\right)^n\,S_g,
    \end{equation}
  \end{linenomath*}
  \noindent where $t$ is time, $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (values in Table~1), $g$ is the acceleration due to gravity, $Q$ is water discharge, $\overline{h}=\frac{1}{2}(h+h_p)$ is the mean hydraulic head in the channel with $h_p$ being the proglacial hydraulic head equal to zero, $l$, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the mean ice overburden pressure expressed in meter water equivalent ($\rho_w$ is density of water; $\rho_i$ is density of ice), and $n$ is Glen's n \citep[usually $n=3$; ][]{glen1955}.
  The first term on the equation's right side represents the channel opening by frictional heating, while the following term represents channel closure from ice deformation.

  \textit{In order to reduce the number of variables in the original text, $h_p$ was omitted as its value is zero, as is hydraulic head at the glacier terminus. The text has been updated to clarify this point.
    Additionally, we confirm the reviewers comment, that effective pressure at the glacier terminus is zero.
    The reviewer also correctly points out that water de-pressurizes underneath the glacier, in agreement with the other reviewer, we have included a comment that we assume pressurized flow across the glacier bed, as is common in subglacial hydrology models, although this may not always be the case \citep{perolo2018}.
  }

  \textit{Lastly, we have chosen a simple model, lumped element, or channel segment model here as we believe that it most clearly and concisely explains the dynamics that we wish to discuss.
    Our strong opinion is that the experiments that we have done with the model, across a range of hydrographs, demonstrate consistently the behavior of subglacial channels.
    Therefore, we do not believe that more complex model is needed to show these processes.
    The limitation of this approach is acknowledge and discussed.
  }

\item  \textbf{And the model should come with some caveats, e. g., you could say that it is a simplistic and possibly unreliable model. To be fair, this is done in section 5.3, but that is too late.}

  \textit{The ``Methods'' section is largely meant to describe the model, as opposed to discuss its behavior or simplicity. In response to this comment, we will move the ``Experiment limitations'' to a section that directly after the ``Results'' but before the ``Discussion.''}

\item \textbf{  One of the comments made is that because the sub-aerial model has an algebraic relation between width and discharge (equation 9), the relationship between velocity and discharge is also algebraic; but this is simply a choice of the model. In reality, the width of a channel will also evolve over (long) time through processes of bank erosion, and the use of equation 9 properly only involves a long term time average of discharge, so it is misleading to use it in a time-specific way. Actually, this is admitted at line 370, along with other limitations in the models.}

  \textit{We appreciate these comments. With regard to the first point about channel width evolution in the subaerial model, this matter is confronted in Section 4.2 and 4.3 (now 5.1).
    Here we show that across a range of exponents and relationships between channel cross section and water discharge, subglacial systems are still more sensitive to water discharge variations that subaerial ones.
    This increased variability occurs across a range of $\alpha$ values that describe the relationship between channel width and water discharge as shown in Figure 6 (now 7).
    As a result, we believe that these result hold across as range of water discharge- channel width relationships, including ones where $\alpha$ evolves in time.
    The exception, shown in Section 4.3 (now 5.1), is that when water discharge evolves more slowly that channel size. Here, the behavior between subaerial and subglacial is very similar, because channel area in both channel type adjust to discharge variations.}

\item   \textbf{Two kinds of numerical experiment are described in section 3.3, and the results of these presented in section 4. Presumably the input to the model runs is the (time- varying) water discharge, but that seems not to be stated in 3.3, which I would expect, although it is stated in the caption to figure 2.
    The second half of the paper characterises the results of these experiments and there is an extended discussion. There is no doubt the authors have put a lot of effort into this, but my interest waned at this point.}

  \textit{    We have modified the first sentence of Section 3.3 to explicitly state:}
    ``Proglacial discharge records from the Fieschergletscher (scenario \alpine{}) and the Leverett Glacier (scenario \icesheet{}) are used as inputs for the models above.''

\item \textbf{  In summary, this is one of those difficult papers to judge, because the effort in- volved is quite substantial and honestly applied, but the whole philosophy of the approach is, in my view, misguided. This is a coxless boat crew who are rowing up a backwater without a rudder, and have lost the main direction of the stream.}

  \textit{With respect to the last sentence, it is true that the lead author feels this way occasionally. Regardless, we think valuable findings are within this manuscript.}

\end{itemize}

\textbf{Specific Comments}

\begin{itemize}
\item \textbf{33: I don’t know what ‘Following mass conservation’ has to do with this.}

  \textit{We will remove this phrase.}

\item \textbf{99-100: upon that of Clarke.}

  \textit{Done.}

\item \textbf{102: presumably $h_{ice}$ is the thickness of the ice, but the syntax is not clear, it could be the depth of the channel.}

  \textit{The text will read:} `` a glacier with channel length $l$, with a flat bed and a mean ice thickness of $h_{ice}$.''

\item \textbf{eq. 3: it’s probably in the Werder 2010 paper,} but I’d like to see an extra comment about where this $\frac{\Delta h}{2}$ term comes from.
  Since the bracket represents $N = pi-pw$, $pw$ the pointwise form of this second term would be $\frac{p_w}{\rho_wg}$, so it’s not obvious to me where the $\Delta h$ comes from.
  Also in this light, it is worth elaborating the confusing term ‘hydraulic head drop change’ and defining what $\Delta h$ is in terms of $pw$.
  Actually, the more I think about this, the more suspicious I become. And in fact, at least when you’re dealing with j\"okulhlaups, the discharge doesn’t vary much along the length of the channel, and the consequence of that is that neither does the effective pressure.
  Of course, that may not apply here but I think the same principle applies.
  So maybe I am promoting this to a more substantive issue
  (as indeed now further discussed earlier).

  \textit{The comment about $\frac{\Delta h}{2}$ is addressed above. We understand the point about flow accumulation and increasing water discharge along the channel. As the author points out, the implications of this are presented in the ``Experiment limitations.'' Additionally, we note that despite these effects, the results here show convincingly that the main topics in the paper still hold, i.e. sediment transport capacity is more variable in subglacial channels than subaerial ones and sediment transport capacity does not vary with water discharge if the subglacial channel size is out of equilibrium with water discharge. }

\item \textbf{114:} I had no idea what the central angle referred to, but evidently it is the angle subtended at the centre of the circular arc of the channel upper boundary; and then $\beta = 2\alpha$, where $\alpha$ is the contact angle at the channel edge, which seems a more natural quantity to use. Also I checked the algebra of equation (5), and got this formula, but with the factor of two in the numerator, not the denominator, assuming hydraulic diameter is twice hydraulic radius, the latter of which is area/perimeter, so please check this. Incidentally, I did also then check equation (7) and I agree with that, so I do think there is an error in (5).

  \textit{We thank the review for the careful assessment of our work. We have double checked an believe that there is no problem in the equation. Our proof is below:}

   \includegraphics[width=0.3\linewidth]{Hooke.png}
  Guiding equations:

  \begin{equation}
    D_h = \frac{4S_g}{P_w}~~~(I)
  \end{equation}

  \begin{equation}
    P_w = R \beta +2R \sin\frac{\beta}{2}= R(\beta +2\sin \frac{\beta}{2} ~~~(II)
  \end{equation}

  \begin{equation}
    S_g=\frac{R^2}{2}(\beta - \sin \beta)~~~(III)
  \end{equation}

  $D_h$ is hydraulic diameter, $P_w$ is wetter perimeter, $R$ is radius, $\beta$ is Hooke angle \citep{hooke1990}, and $S_g$ is the subglacial channel cross section.

  Proof:

  \begin{linenomath*}
    \begin{equation}
      \frac{D_h}{4S_g} = \frac{1}{R\beta + 2R\sin\frac{\beta}{2}}
    \end{equation}
  \end{linenomath*}

  \begin{equation}
    R = \frac{4S_g}{D_h (\beta +2 \sin \frac{\beta}{2})}, \quad \mathrm{into~} III 
  \end{equation}

  \begin{equation}
    S_g = \frac{16 S_g^2}{D_h^2 (\beta +2\sin\frac{\beta}{2})^2} ~\frac{1}{2}(\beta-\sin\beta)
  \end{equation}

  \begin{equation}
    1 = \frac{8 S_g}{D_h^2 (\beta +2\sin\frac{\beta}{2})^2} (\beta-\sin\beta)
  \end{equation}

  \begin{equation}
    \frac{D_h^2}{8} \frac{(\beta+2\sin\frac{\beta}{2})^2}{\beta - \sin \beta} = S_g
  \end{equation}

  \begin{equation}
    \frac{D_h^2 (2(\frac{\beta}{2}+\sin \frac{\beta}{2}))^2}{8(\beta - \sin \beta)} = S_g
  \end{equation}

   \begin{equation}
    \frac{D_h^2 ((\frac{\beta}{2}+\sin \frac{\beta}{2}))^2}{2(\beta - \sin \beta)} = S_g
  \end{equation}

  \begin{equation}
       S_g = \frac{D_h^2}{2} \frac{(\frac{\beta}{2}+\sin \frac{\beta}{2})^2}{(\beta - \sin \beta)}  
  \end{equation}

  This is the same as Equation 5 of our manuscript.

\end{itemize}

\bibliographystyle{apalike} 
\bibliography{../PaperLib.bib}
\end{document}

### 2025-02-03 Delaneyetal2024.tex
small change

@@
-%%Copernicus Publications Manuscript Preparation Template for LaTeX Submissions
+%% Copernicus Publications Manuscript Preparation Template for LaTeX Submissions
 %% ---------------------------------
@@
 
-\title{Subglacial and subaerial fluvial sediment transport capacity respond differently to water discharge variations}
+\title{Sediment transport capacity response to variations in water discharge in pressurized subglacial channels}
 
@@
   Sediment transport capacity in both subaerial and subglacial channels depends on the shear stress exerted across the channel bottom, which varies with water velocity and channel width.
-  In subaerial channels, water discharge variations are accommodated  by flow depth and width changes, along with water velocity.
+  In full subaerial channels, water discharge variations are accommodated  by flow depth and width changes, along with water velocity.
   However, in subglacial channels, water is pressurized by the ice above, and they grow in response to frictional heating of water flowing through them. 
-  As a result, water discharge changes mainly result in velocity variations, as the channel geometry evolves slowly (over days).
-  Here, we present formulations of sediment transport capacity in different channel types and apply subglacial and subaerial hydraulics models to hydrographs from an Alpine glacier and the Greenland Ice.
+  As a result, rapid changes in water discharge  mainly result in velocity variations, as the channel geometry evolves slowly.
+  Here, we present formulations of sediment transport capacity in different channel types and apply subglacial and subaerial hydraulics models to hydrographs from an Alpine glacier and the Greenland Ice Sheet.
   Numerical experiments show that the changing channel size results in sediment transport capacity peaking before the maximum water discharge.
-  This hysteresis in channel size causes a highly variable relationship between sediment and water discharge in a transport-limited subglacial system.
+  This hysteresis causes a highly variable relationship between sediment and water discharge in a transport-limited subglacial system.
   The results also indicate that high subglacial sediment transport capacities can occur across a wide range of water discharges.
+  Reducing water discharge variability lessens the hysteresis effects, in some cases to the point where a covarying relationship between water discharge and sediment transport capacity may occur, similar to subaerial systems.
   A second set of numerical experiments shows that subglacial sediment transport is highly non-linear with respect to water discharge, creating more variability in sediment transport capacity.
   Yet, results and  formulations of subglacial sediment transport capacity show that its variability can approach that of subaerial systems when subglacial channel size is in equilibrium with water discharge.
-  The implications of these findings are discussed in the context of sediment discharge from glaciers with different hydro-climatic forcings.
-  We also discuss the impact of different assumptions of channel behavior on sediment transport capacity.
+  We present these results in the context of different formulations of the relationship between water discharge and sediment transport capacity based assumptions of subglacial and subaerial flow regimes.
+  The implications of these findings help to evaluated sediment discharge from glaciers with different hydro-climatic forcings.
   These findings can improve the interpretation of sediment discharge records in glacierized catchments.
@@
 % 
-Following mass conservation, the mean velocity of the water flowing through a  channel is
+The mean velocity of the water flowing through a  channel is
 \begin{linenomath*}
@@
 Many of these studies discuss the disparity between water discharge and sediment export, especially given the role of sediment access or production.
-Yet, a variable relationship between water discharge and sediment export could be expected from glaciers with evolving conduit size in a transport-limited regime, where sediment transport responds to the channel's hydraulic conditions, instead of sediment availability.
+Yet, a variable relationship between water discharge and sediment export could be expected from glaciers with evolving conduit size in a transport-limited regime, instead of sediment availability.
 The response of sediment transport capacity to water discharge variations may affect the interpretation of sediment transport records in these glacierized catchments \citep[e.g.][]{ganti2016,mancini2023}.
@@
 
-This manuscript has two objectives: 1)  to establish whether sub-seasonal water discharge can co-vary with sediment transport capacity in subglacial systems, and 2) to evaluate the variability of  sediment transport capacity in subglacial channels with evolving channel size and different hydrographs compared to subaerial ones.
+This manuscript has two objectives: 1)  to establish the hydrological conditions where  sediment transport capacity co-varies with water discharge in subglacial systems,
+and 2) to evaluate the variability of  sediment transport capacity in subglacial channels with evolving channel size and different hydrographs compared to subaerial ones.
 We use numerical models to examine sediment transport capacity in both subglacial channels that evolve in size and subaerial channels that remain fixed.
 They are applied to proglacial hydrological records from an Alpine glacier in Switzerland (Fieschergletscher) and a land-terminating glacier in Greenland (Leverett Glacier).
-Model outputs demonstrate the specific processes that can drive variability in sediment discharge capacity in subglacial systems.
 We also run a model ensemble to evaluate the sediment transport capacity variability in subglacial channels compared to subaerial channels.
@@
 These formulations extend the relationships presented in \citet{alley1997} and illustrate the sediment transport capacity behavior of different channel assumptions.
-Findings indicate differences in the relationships amongst water discharge, channel geometry, water velocity and sediment transport capacity in subglacial and subaerial channels.
 The manuscript then discusses the implications of variability in sediment transport capacity from glaciers and the interpretation of sediment transport records.
@@
 The subglacial channel model accounts for the channel geometry and the water's velocity to evaluate the shear stress of water flowing across sediments underneath a glacier.
-We use a lumped hydraulics model from \citet{werder2010b}, itself based upon  \citet{clarke1996}.
+We use a lumped hydraulics model from \citet{werder2010b}, itself based upon that of \citet{clarke1996}.
 
-Here, it is assumed that the water is transported through a subglacial channel \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972} beneath a glacier with channel length $l$, with a flat bed and a mean thickness of $h_{ice}$.
+Here, it is assumed that the water is transported through a subglacial channel \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972} beneath a glacier with channel length $l$, with a flat bed and a mean ice thickness of $h_{ice}$.
 The channel size grows from melt due to frictional heating from water flow and closes due to ice creep.
@@
     \label{eq:dS_dt}
-    \frac{\partial S_g}{\partial t} = C_1 \frac{Q \Delta h}{l} - C_2 \left(h_{o}-\frac{\Delta h}{2}\right)^n\,S_g,
+    \frac{\partial S_g}{\partial t} = C_1 \frac{Q \Delta h}{l} - C_2 \left(h_{o}-\overline{h}\right)^n\,S_g,
   \end{equation}
 \end{linenomath*}
-\noindent where $t$ is time, $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (values in Table~\ref{table:vpm}), $g$ is the acceleration due to gravity, $Q$ is water discharge, $\Delta h$ is the hydraulic head drop change over $l$, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the mean ice overburden pressure expressed in meter water equivalent ($\rho_w$ is density of water; $\rho_i$ is density of ice), and $n$ is Glen's n \citep[usually $n=3$; ][]{glen1955}.
+\noindent where $t$ is time, $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (values in Table~\ref{table:vpm}), $g$ is the acceleration due to gravity, $Q$ is water discharge, $\overline{h}=\frac{1}{2}(h+h_p)$ is the mean hydraulic head in the channel with $h_p$ being the proglacial hydraulic head equal to zero, $l$, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the mean ice overburden pressure expressed in meter water equivalent ($\rho_w$ is density of water; $\rho_i$ is density of ice), and $n$ is Glen's n \citep[usually $n=3$; ][]{glen1955}.
 The first term on the equation's right side represents the channel opening by frictional heating, while the following term represents channel closure from ice deformation.
@@
 
-Note that this subaerial channel model is purely algebraic, whereas the subglacial model comprises a differential equation for the evolution of $S_g$.
-Thus, the channel size in the subaerial model has no history dependence on the discharge $Q_w$, whereas the subglacial one does (Equation~\ref{eq:dS_dt}).
+% Note that this subaerial channel model is purely algebraic, whereas the subglacial model comprises a differential equation for the evolution of $S_g$.
+% Thus, the channel size in the subaerial model has no history dependence on the discharge $Q_w$, whereas the subglacial one does (Equation~\ref{eq:dS_dt}).
 
@@
     A second line refers to the range of values examined in the parameter search.}
-  \small 
+  \footnotesize
   \begin{tabular}{ l  c  c c }
@@
     Hydraulic head &$\Delta h$&& $\mathrm{m}$\\
+    Mean hydraulic head & $\overline{h}$&& $\mathrm{m}$\\
     Hydraulic gradient &$\Psi=\frac{\Delta h}{l}$&& $\mathrm{m\, m^{-1}}$\\
@@
     Effective glacier length &$l$& $7,000$ or $26,000$&\unit{m}\\
+    Hydraulic head at terminus&$h_p$&$0$& \unit{m}\\
     Constant $1$ in Equation~3 &$C_1$&$2.2\times10^{-5}$&\unit{m}$^{-1}$\\
@@
 
-The models above are applied to proglacial discharge records from the Fieschergletscher (scenario \alpine{}) and the Leverett Glacier (scenario \icesheet{}).
+Proglacial discharge records from the Fieschergletscher (scenario \alpine{}) and the Leverett Glacier (scenario \icesheet{}) are used as inputs for the models above.
 The model outputs represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
@@
 \icesheet{}  is exemplified by thick ice  \citep[$h_{ice}$= $700$\,\unit{m}; ][]{morlighem2017}, high water discharge ($\sim\,300$\,\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge at Leverett glacier (Figure~\ref{fig:model_outs}\, e).
+We note that the subglacial results apply to pressurize subglacial channels, not depressurized one that can occur below glacier termini \citep{perolo2018}.
 
@@
 Both friction factors $f_r$ and $f_f$ are tuned so that reasonable water velocities \citep[$\sim\,1- 1.5\,$\unit{m}\,\unit{s}$^{-1}$][]{werder2010b,chandler2013} occur for both \alpine{} and \icesheet{}.
+
 To test the covariance between sediment transport capacity and water discharge, model outputs are compared to water discharge from the glaciers using Spearman rank correlation.
@@
 Rank correlation reduces the impact of the non-linear relationship between sediment transport capacity and hydrology.
+Note that this subaerial channel model is purely algebraic, whereas the subglacial model comprises a differential equation for the evolution of $S_g$.
+Thus, the channel size in the subaerial model has no history dependence on the discharge $Q_w$, whereas the subglacial one does (Equation~\ref{eq:dS_dt}).
+Therefore, rank correlation is $1$ between water discharge and sediment transport capacity subaerial systems, or subglacial ones where channel size $S_g$ evolves more slowly than water discharge.
 
@@
 
-The effects of channel shape, gradient, and roughness are established by running the model with random parameter values of channel slope and geometry factors ($\beta$,$\nabla z_c$, and $\alpha$) and friction factors ($f_r$ and $f_f$; see Table~\ref{table:vpm} for value range).
+The effects of channel shape, gradient, and roughness are established by running the model with random parameter values in a range of channel slope and geometry factors ($\beta$,$\nabla z_c$, and $\alpha$) and friction factors ($f_r$ and $f_f$; see Table~\ref{table:vpm} for value range).
 Runs with parameter combinations are accepted if their mean subglacial water velocity over the season lies between $0.5$\,\unit{m}\,\unit{s}$^{-1}$ and $2$\,\unit{m}\,\unit{s}$^{-1}$ or if subaerial water velocity lies between $0.3$\,\unit{m}\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}\,\unit{s}$^{-1}$ \citep[e.g.][]{werder2010b,magnusson2012,chandler2013}.
@@
 The routine runs until $100$ different parameter combinations for each water discharge smoothing period are accepted with the conditions described above.
-To test the effects of variability in water discharge, it is averaged over the smoothing period ranging
+To test the effects of variability in water discharge, it is averaged over the smoothing periods with the moving average from $15$ minutes to $15$ days.
 
@@
 
-\subsection{Changing subglacial channel size drives different timing and variability in sediment transport capacity}
+\subsection{Timing and covariance of subaerial and subglacial  sediment transport capacity variations}
 
 The first numerical experiment aims to quantify the sources of increased variability in the subglacial model outputs as they exhibit different seasonal evolutions and peaks (Figure~\ref{fig:model_outs}).
-Variable relationships between model outputs and water discharge emerge for the subaerial and subglacial cases due to the hysteresis in channel size in the subglacial model (Equation~\ref{eq:dS_dt}).
+Variable relationships between model outputs and water discharge emerge for the subaerial and subglacial cases as channel size evolves (Figure~\ref{fig:model_outs} a, e).
 Because subaerial channels have no history dependence, each water discharge value in the subaerial channel produces a unique water velocity, shear stress, and width-integrated shear stress (Section~\ref{sect:sub_mode}).
@@
   \centering
-  \includegraphics[width=\linewidth]{Fig2.png}
+  \includegraphics[width=0.75\linewidth]{Fig2.png}
   \caption{Model outputs from simulations using the hydrographs in panels a and e for the scenarios \alpine{} (a-d) and \icesheet{} (e-h).
@@
     Arrows show examples where variables peak in the subglacial channel before the subaerial one.
-    Insets in f-h show the peak melt event denoted by the shaded area in panels e--h, with an arbitrary y-axis.
+    Insets in f-h show the peak melt event denoted by the shaded area in panels e--h, with 
   }
@@
 Peaks in subaerial model outputs occur coincident with peaks in water discharge (Figure~\ref{fig:model_outs}).
-In the subglacial channel, peaks in model outputs generally occur when water discharge increases at the fastest rate, but before the maximum water discharge.
-As the water discharge stabilizes at its peak, channel growth continues, causing water velocity and other model outputs to decrease from their peak values (Section~\ref{sect:sub_mode}).
-As a result, subglacial sediment transport capacity is greatest on the hydrograph's rising limb, relative to the falling limb, creating a hysteresis effect.
+In the subglacial channel, peaks in model outputs generally occur when water discharge increases, but before the maximum water discharge.
+As the water discharge stabilizes at its peak, channel growth continues (Figure~\ref{fig:model_outs} a, e), causing water velocity and other model outputs to decrease from their peak values (Figure~\ref{fig:model_outs} b-d, f-h).
+Subglacial sediment transport capacity is greatest on the hydrograph's rising limb, relative to the falling limb, creating a hysteresis effect.
 
@@
 The subglacial channel's evolving width can counteract some of these effects.
-Width-integrated shear stress generally increases with water discharge, with greater rank correlation compared to water velocity or shear stress (\alpine{}, Figure~\ref{fig:Qw_vari} a--c).
+Width-integrated shear stress generally increases with water discharge, with greater rank correlation than water velocity or shear stress (\alpine{}, Figure~\ref{fig:Qw_vari} a--c).
 Yet, even the width-integrated shear stress can vary substantially relative to water discharge.
@@
 The variability in width-integrated shear stress is less pronounced in the \icesheet{} scenario, where the hydrograph has less diurnal variability (Figure~\ref{fig:Qw_vari} c, f).
-This results from the discharge variations being in closer equilibrium with subglacial channels compared to \alpine{} (Sections~\ref{sect:scaling} and~\ref{sect:ensemble}).
+% This results from the discharge variations being in closer equilibrium with subglacial channels compared to \alpine{} (Sections~\ref{sect:scaling} and~\ref{sect:ensemble}).
 
@@
   \centering
-  \includegraphics[width=0.8\linewidth]{Fig3.png}
+  \includegraphics[width=0.7\linewidth]{Fig3.png}
   \caption{
@@
 
+Water discharge and sediment transport characteristics correlate better in \icesheet{} with lower diurnal variations in water discharge compared to the \alpine{}.
+To evaluate the impact of water discharge variability, we smoothed the two hydrographs across different periods and evaluated the Spearman rank correlation between sediment transport characteristics and water discharge.
+We assume that a higher rank correlation indicates a reduced amount of hysteresis and behavior more similar subaerial channels (Figure~\ref{fig:Qw_vari}).
+
+Smoothing water discharge over periods longer than one day causes a substantial increase in rank correlation (Figure~\ref{fig:corri}).
+This increase occurs as diurnal variations are removed (Figures S1-S14).
+Correlations in sediment transport characteristics in \icesheet{} remain higher than \alpine, which has more water discharge variations even when smoothed compared to \icesheet{} (Figure~S1-S14).
+
+\begin{figure}[hbt!]
+  \centering
+  \includegraphics[width=0.8\linewidth]{Fig4.png}
+  \caption{Spearman rank correlation between water discharge over a smoothed period and sediment transport characteristics (a) water velocity, (b) shear stress, (c) width-integrated shear stress.
+We assume that higher correlation scores approach subaerial behavior.
+  }
+  \label{fig:corri}
+\end{figure}
+
 \FloatBarrier
-\subsection{Increased variability in evolving subglacial channels occurs across a range of channel shapes, slopes,  and friction values}
+\subsection{Variability in sediment transport capacity across a range of channel shapes, slopes,  and friction values}
 \label{sect:ensemble}
@@
 These smoothing timescales remove the diurnal variations in water discharge, thereby reducing variability in model outputs. 
-Velocity variability in \icesheet{} is substantially smaller than \alpine{} (Figure~\ref{fig:multi_run}\,a and \, d).
-This could result from the subglacial conduit being in closer equilibrium with water discharge.
-Variations in \icesheet{} shear stress are comparable to \alpine due to the effects of the evolution of shear stress with subglacial conduit size (Equations ~\ref{eq:dS_dt} and \ref{eq:DW};). 
+Velocity variability in \icesheet{}, with smaller relative water discharge variations, is substantially smaller than \alpine{} (Figure~\ref{fig:multi_run}\,a and \, d).
+Variations in \icesheet{} shear stress are comparable to \alpine due to the effects of changing of shear stress with subglacial conduit size (Equations ~\ref{eq:dS_dt} and \ref{eq:DW};). 
 The much greater water flux and thus subglacial channel size in \icesheet{} drives the larger width integrated shear stress variations (Figure \ref{fig:model_outs}\,c and f).
@@
   \centering
-  \includegraphics[width=0.7\linewidth]{Fig4.pdf}
+  \includegraphics[width=0.7\linewidth]{Fig5.pdf}
   \caption{Standard deviation of detrended model outputs for different smoothing times with different subglacial and subaerial channel shapes and friction factors.
@@
 Greater subglacial friction factors $f_i$ result in greater variability in the shear stress and width-integrate shear stress in the \alpine case (Figure~\ref{fig:params}).
-This result is expected given Equations~\ref{eq:tau_g}.
 Low values of $f_i$ result in slower growth rates in subglacial channels (Equations~\ref{eq:dS_dt} and \ref{eq:dh}).
-As a result, water velocity, as opposed to channel growth, accommodates increases in water discharge, and velocity variability increases.
 Higher values of $f_i$ allow faster channel growth that accommodates increases in water discharge, reducing velocity variability.
-Smaller values of channel factor $\beta$, creating low and broad channels, result in more variability in width-integrated shear stress.
-Here, the channel width can grow more quickly in response to water discharge increases as compared to a semi-circular channel with $\beta = \pi$ (Equation~ \ref{eq:dh}).
+Smaller values of channel factor $\beta$, creating low and broad channels where the channel width grows more quickly in response to water discharge increases as compared to a semi-circular channel with $\beta = \pi$ (Equation~ \ref{eq:dh}).
 
@@
 Greater values of subaerial friction factors $f_p$ result in greater variability of shear stress and width-integrated shear stress. 
-Yet, they do not exceed the variability of subglaical channels. 
+Yet, subaerial values do not exceed the variability of subglacial channels. 
 We note that these parameter values span a commonly accepted range.
@@
   \centering
-  \includegraphics[width=0.7\linewidth]{Fig5.pdf}
+  \includegraphics[width=0.8\linewidth]{Fig6.pdf}
   \caption{Parameter values compared to variability in model outputs.
@@
 \FloatBarrier
+\section{Experiment limitations}
+\FloatBarrier
+
+Several aspects of the models and experiment design make the comparison of the subaerial and subglacial systems difficult. 
+The lumped nature of the models means that they operate independently of the upstream drainage network.
+Additionally, we omit analysis of suspended sediment discharge records due to their dependence on sediment supply in capturing variations in sediment discharge, in addition to sediment transport capacity \citep[e.g.][]{delaney2019}.
+In reality, processes such as sediment access are usually important in controlling sediment export in glacierized catchments \citep[e.g.][]{herman2015,vergara2022}.
+Furthermore, meltwater can be distributed to flow through several adjacent conduits impacting the sediment transport capacity in each \citep[e.g.][]{werder2013,hewitt2019,delaney2023}.
+The lumped models used here isolate the relationship between water discharge and sediment transport capacity in subglacial and subaerial systems. 
+However, they neglect more complex, yet important, spatially distributed processes.
+
+Different hydraulic gradients control the velocity and sediment transport capacity in the subglacial and subaerial cases (Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}). 
+The subglacial one is generally much steeper \citep{alley1997}.
+This results in the shear stresses and width-integrated shear stresses across the channel bed that controls sediment transport capacity being much greater in subglacial channels (Figure~\ref{fig:model_outs}).
+The parameters' range tested in Section~\ref{sect:ensemble} covers a likely span of viable shear stresses in both subglacial and subaerial channels.
+The greater sediment transport capacity in the subglacial channels implies that sediment grain size underneath subglacial channels is likely larger than subaerial counterparts.
+Note, however, that the water velocities are similar in both channel types.
+
+The channel width and size variations in subglacial channels presented here also occur in subaerial channels in response to water discharge \citep{phillips2016}, likely over decadal timescales or in response to individual extreme events \citep[e.g.][]{slater2013,dean2013}.
+These timescales are likely considerably longer than the continuous changes to the subglacial channel's response time of days.
+Furthermore, observations from subaerial systems can show a co-varying relationship between water discharge and sediment transport in time \citep[e.g.][]{schmidt2006,pitlick2021}.
+Even so, if channel width changes occur in subaerial channels, then the values of $\alpha$ in Equation~\ref{eq:wcf} could change in time.
+$\alpha$ must be less than $1$, and even this would cause the exponent on $Q$ in the  $w_g\,\tau_{g}$ relationships to remain substantially below that of pipeflow (Section ~\ref{sect:scaling}, Table\,\ref{tab:eqs1} and \ref{tab:Qs})).
+
+Width-integrated shear stress is examined in the two numerical experiments as it does not have a dependence on grain size, unlike sediment transport relationships \citep[][Section \ref{sect:scaling}]{meyer1948}.
+This makes comparison between the two systems simpler. 
+Preferential transport of smaller sediment clasts and the input of upstream sediment impact sediment transport capacity as grain size evolves \citep[e.g.][]{gomez1983}.
+These processes are only beginning to be evaluated underneath glaciers \citep{aitken2024}, but can be important in subaerial systems.
+Subglacial sorting processes could be an additional source of variability in sediment export from glaciers with respect to water discharge, especially for bedload transport.
+
+\FloatBarrier
+\section{Discussion}
+
+\FloatBarrier
 \subsection{Sediment transport scaling in different channel types}
@@
   \centering
-  \includegraphics[width=0.7\linewidth]{Fig6.pdf}
+  \includegraphics[width=0.7\linewidth]{Fig7.pdf}
   \caption{Response of sediment transport capacity (orange line) and channel size (bottom row) to a water discharge event (blue line) under different channel conditions.
@@
 The numerical experiments above consider the size evolution of subglacial channels and demonstrate that for these hydrographs subglacial sediment transport variability is greater than its subaerial counterpart (Section~\ref{sect:ensemble}).
-Here, we compare the sediment transport behavior of different channel types as they respond to water discharge, channel shape, and hydraulic gradient. 
-This yields insights into the conditions where subglacial sediment transport capacity could exhibit similar variability as in subaerial channels.
+Additionally, results demonstrate the impact of water discharge variability on sediment transport capacity.
+Here, we discuss and compare the sediment transport behavior of different channel types as they respond to water discharge, channel shape, and hydraulic gradient, extending the formulations presented in \citet{alley1997}.
 
-Channel types examined include subaerial channels (Figure~\ref{fig:chan_types} c), steady-state R-channels \citep[Figure~\ref{fig:chan_types} d]{rothlisberger1972} and pipe-flow (i.e.
+Channel types examined include
+\begin{enumerate}
+\item subaerial channels (Figure~\ref{fig:chan_types} c),
+
+\item steady-state R-channels \citep[Figure~\ref{fig:chan_types} d]{rothlisberger1972}, where water discharge is in equilibrium with channel size.
+This scenario could be approached with a large amount of discharge smoothing (i.e.
+Figure~\ref{fig:corri}) or hydrographs with minimal variations, and
+
+\item pipe-flow (i.e.
 R-channels of fixed size that do not adjust their size to discharge conditions; Figure~\ref{fig:chan_types} a).
+This scenario could represent \alpine{} with minimal discharge smoothing and high water fluctuations compared to channel size (i.e.
+Figure~\ref{fig:model_outs} a-d).
+\end{enumerate}
 These formulations do not allow us to evaluate non-steady state R-channels presented above (Section~\ref{sect:sub_mode}, Figure~\ref{fig:chan_types} b).
-However, the formulations establish the scaling of these different channel typesc.
-The sediment transport capacity is calculated with a given water discharge and hydraulic gradient for three different sediment transport capacity formulas: Meyer-Peter M\"uller  \citep[MPM; ][]{meyer1948}, Engelund and Hansen \citep[EH; ][]{engelund1967}, and Bagnold \citep{bagnold1980}; additionally width-integrated shear stress is assessed, as used as proxy for sediment transport above.
+The sediment transport capacity is calculated with a given water discharge and hydraulic gradient for three different sediment transport capacity formulas: Meyer-Peter M\"uller  \citep[MPM; ][]{meyer1948}, Engelund and Hansen \citep[EH; ][]{engelund1967}, and Bagnold \citep{bagnold1980}; additionally width-integrated shear stress is assessed, instead of sediment transport, as above.
 
@@
 Pipe flow-like conditions occur when an R-channel is subjected to rapid discharge variations such that the channel cannot adjust its size.
+Such a case could occur over diurnal variations in the \alpine{} case, where water discharge changes rapidly with respect to channel size.
 In this case, it is assumed that the cross-sectional area $S$ is fixed, and $\Psi$ adjusts to the specified $Q$.
@@
 
-\FloatBarrier
-\section{Discussion}
-
 \subsection{Increased variability in  sediment transport capacity in subglacial systems}
@@
 
-The greater variations in shear stress in subglacial channels compared with subaerial channels that we present here may cause even greater variations in sediment transport capacity in subglacial channels than our numerical results suggest  (Figure~\ref{fig:multi_run}).
+The greater variations in shear stress in subglacial channels than subaerial channels may cause even greater variations in sediment transport capacity in subglacial channels than our numerical results suggest  (Figure~\ref{fig:multi_run}).
 In sediment transport capacity relationships, such as in \citet{meyer1948} or \citet{engelund1967}, shear stress is scaled to the power of $\frac{3}{2}$ or $\frac{5}{2}$, respectively (e.g.
@@
 Sediment exhaustion through repeatedly crossing the mobilization threshold may explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on the glaciers' basal shear stress, a proxy for bedrock erosion, rather than glacier melt \citep{overeem2017}.
-More generally, the same process could also result in sediment discharge's strong dependence on sediment production from glacial sliding, itself dependent of basal shear stress of the glacier \citep{herman2015,koppes2015}.
+More generally, the same process could also result in sediment discharge's strong dependence on sediment production from glacial sliding, which is dependent of basal shear stress of the glacier \citep{herman2015,koppes2015}.
 
-While transport-limited states likely do not occur at many glaciers \citep[e.g.][]{alley1997}, abundant sediment could persist underneath some glaciers, potentially creating a transport-limited regime \citep[e.g.][]{walter2014,stevens2022,delaney2022}.
-At these glaciers, the great variability in subglacial sediment transport capacity may make it difficult to link sediment discharge to hydrology, especially when peak events occur \citep{cowan1988,delaney2018,lu2022}.
+At glaciers with abundant sediment at the bed, in a transport-limited regime, the great variability in subglacial sediment transport capacity may make it difficult to link sediment discharge to hydrology, especially when peak events occur \citep{cowan1988,delaney2018,lu2022}.
 The results here suggest that different channel sizes with the same hydrology forcing can result in very different sediment transport capacities, making it challenging to establish the effects of individual extreme precipitation or melt events without establishing the antecedent state of the subglacial drainage system.
@@
 Catchments with reduced water discharge variability may experience less variability in sediment transport capacity, shown by the \alpine{} and \icesheet{} model outputs.
-Indeed, the more decoupled and sporadic relationship between model outputs and water discharge in \alpine{} results from the relatively larger discharge variations on sub-daily to weekly timescales (Figure~\ref{fig:Qw_vari}).
+Indeed, the more decoupled and variable relationship between model outputs and water discharge in \alpine{} results from the relatively larger discharge variations on sub-daily to weekly timescales (Figure~\ref{fig:Qw_vari}).
 This occurs as the subglacial channel's size evolves slowly compared to the variations in water discharge (Figure~\ref{fig:chan_types} b).
@@
 
-A sporadic relationship between subglacial sediment transport capacities and water discharge occurs due to hysteresis in subglacial channel size (Figure~\ref{fig:Qw_vari}).
-This hysteresis limits the use of water discharge as an indicator of sediment discharge capacity in these systems.
+A variable relationship between subglacial sediment transport capacities and water discharge occurs due to hysteresis in subglacial channel size (Figure~\ref{fig:Qw_vari}).
+This hysteresis can limit water discharge as an indicator of sediment discharge capacity in these systems, especially when water discharge is highly variable and out of equilibrium with subglacial channel size.
 As a result, characteristics such as bankfull or effective water discharge that link geomorphic work to hydro-climatic conditions could have limited meaning in evaluating subglacial sediment transport \citep{wolman1960,lenzi2006}.
@@
 
-\subsection{Experiment limitations}
-
-Several aspects of the models and experiment design make the comparison of the subaerial and subglacial systems difficult. 
-The lumped nature of the models means that they operate independently of the upstream drainage network.
-Additionally, we omit analysis of suspended sediment discharge records due to their dependence on sediment supply in capturing variations in sediment discharge, in addition to sediment transport capacity \citep[e.g.][]{delaney2019}.
-In reality, processes such as sediment access are usually important in controlling sediment export in glacierized catchments \citep[e.g.][]{herman2015,vergara2022}.
-Furthermore, meltwater can be distributed to flow through several adjacent conduits impacting the sediment transport capacity in each \citep[e.g.][]{werder2013,hewitt2019,delaney2023}.
-The lumped models used here isolate the relationship between water discharge and sediment transport capacity in subglacial and subaerial systems. 
-However, they neglect more complex, yet important, spatially distributed processes.
-
-Different hydraulic gradients control the velocity and sediment transport capacity in the subglacial and subaerial cases (Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}). 
-The subglacial one is generally much steeper \citep{alley1997}.
-This results in the shear stresses and width-integrated shear stresses across the channel bed that controls sediment transport capacity being much greater in subglacial channels (Figure~\ref{fig:model_outs}).
-The parameters' range tested in Section~\ref{sect:ensemble} covers a likely span of viable shear stresses in both subglacial and subaerial channels.
-The greater sediment transport capacity in the subglacial channels implies that sediment grain size underneath subglacial channels is likely larger than subaerial counterparts.
-Note, however, that the water velocities are similar in both channel types.
-
-The channel width and size variations in subglacial channels presented here also occur in subaerial channels in response to water discharge \citep{phillips2016}, likely over decadal timescales or in response to individual extreme events \citep[e.g.][]{slater2013,dean2013}.
-These timescales are likely considerably longer than the continuous changes to the subglacial channel's response time of days.
-Furthermore, observations from subaerial systems can show a co-varying relationship between water discharge and sediment transport in time \citep[e.g.][]{schmidt2006,pitlick2021}.
-Even so, if channel width changes occur in subaerial channels, then the values of $\alpha$ in Equation~\ref{eq:wcf} could change in time.
-$\alpha$ must be less than $1$, and even this would cause the exponent on $Q$ in the  $w_g\,\tau_{g}$ relationships to remain substantially below that of pipeflow (Section ~\ref{sect:scaling}, Table\,\ref{tab:eqs1} and \ref{tab:Qs})).
-
-Width-integrated shear stress is examined in the two numerical experiments as it does not have a dependence on grain size, unlike sediment transport relationships \citep[][Section \ref{sect:scaling}]{meyer1948}.
-This makes comparison between the two systems simpler. 
-Preferential transport of smaller sediment clasts and the input of upstream sediment impact sediment transport capacity as grain size evolves \citep[e.g.][]{gomez1983}.
-These processes are only beginning to be evaluated underneath glaciers \citep{aitken2024}, but can be important in subaerial systems.
-Subglacial sorting processes could be an additional source of variability in sediment export from glaciers with respect to water discharge, especially for bedload transport.
-
 \conclusions
@@
 
-The manuscript's first objective is to establish if water discharge covaries with subglacial sediment transport capacity.
-In subglacial channels, the timing of peak water velocity and sediment transport capacity occurs before peak water discharge during a discharge event, due to evolving channel size.
+The manuscript's first objective is to establish the conditions for  water discharge to covary with subglacial sediment transport capacity.
+In subglacial channels, the timing of peak water velocity and sediment transport capacity occurs before peak water discharge during a discharge event on hydrographs from an alpine and ice sheet catchment, due to evolving channel size.
 In subaerial channels, the timing of peak sediment transport capacity and water discharge coincide.
-Results here suggest that, even in a transport-limited subglacial system, an incoherent relationship between water and subglacial sediment discharge could be expected in most channels.
-This incoherent relationship presents a challenge in linking hydro-climatic conditions or events to sediment export from glaciers.
-Water discharge and sediment transport capacity covariance could be possible when water discharge varies at a slower rate than subglacial channel size.
-In natural environments, this appears to rarely be the case.
+Results here suggest that, even in a transport-limited subglacial system, an variable relationship between water and subglacial sediment discharge.
+Water discharge variations in both Greenland Ice Sheet and Alpine are substantial enough to cause this inconsistent relationship, even when water discharge records are smoothed.
+This variable relationship presents a challenge in linking hydro-climatic conditions or events to sediment export from glaciers.
+Water discharge and sediment transport capacity covariance could be possible when water discharge varies at a slower rate than subglacial channel size, such as Antarctica with minimal surface melt input.
+Results also suggest that water discharge records averaged over periods longer than $12$ hours from alpine and ice sheet hydrographs show substantial impacts on subglacial sediment transport capacity characteristics.
 
@@
 The poor constraints on subglacial sediment size and sorting processes make it difficult to link shear stress to sediment transport capacity.
-Further quantifying these processes will help to better inform the response of sediment transport capacity to water discharge forcing in subglacial environments.
+New methods are being developed to quantifying these processes \citep{jenkin2023},  better informing the response of sediment transport capacity to water discharge forcing in subglacial environments.
 

### 2025-02-03 Delaneyetal2024.tex
Updates from Overleaf

--- 2025-01-22 Delaneyetal2024.tex
@@
 As a result, water velocity, as opposed to channel growth, accommodates increases in water discharge, and velocity variability increases.
-Higher values of $f_i$ allow faster channel growth that accommodates increases in water discharge, reducing velocity variability.
+Higher values of $f_i$ allow faster channel growth that accommodates increase in water discharge, reducing velocity variability.
 Smaller values of channel factor $\beta$, creating low and broad channels, result in more variability in width-integrated shear stress.
@@
 Steeper subaerial channel slopes $\nabla z_c$ result in greater variability in both shear stress and width-integrated shear stress but only approach the subglacial channel's variability in the \alpine case (Figure~\ref{fig:params}).
-Greater values of subaerial friction factors $f_p$ result in greater variability of shear stress and width-integrated shear stress. 
-Yet, they do not exceed the variability of subglaical channels. 
+Greater values of subaerial friction factors $f_p$ resulting in greater variability shear stress and width-integrated shear stress. 
+Yet, they to not exceed the variability of subglaical channels. 
 We note that these parameter values span a commonly accepted range.
@@
 The numerical experiments above consider the size evolution of subglacial channels and demonstrate that for these hydrographs subglacial sediment transport variability is greater than its subaerial counterpart (Section~\ref{sect:ensemble}).
-Here, we compare the sediment transport behavior of different channel types as they respond to water discharge, channel shape, and hydraulic gradient. 
-This yields insights into the conditions where subglacial sediment transport capacity could exhibit similar variability as in subaerial channels.
+Here, we compare the sediment transport behavior of different channels types as they respond water discharge, channel shape, and hydraulic gradient. 
+This yields insights into the conditions where subglacial sediment transport capacity could exhibit similar variability as subaerial channels.
 
@@
 These formulations do not allow us to evaluate non-steady state R-channels presented above (Section~\ref{sect:sub_mode}, Figure~\ref{fig:chan_types} b).
-However, the formulations establish the scaling of these different channel typesc.
+However, the formulations establish the scaling of these different channel types  .
 The sediment transport capacity is calculated with a given water discharge and hydraulic gradient for three different sediment transport capacity formulas: Meyer-Peter M\"uller  \citep[MPM; ][]{meyer1948}, Engelund and Hansen \citep[EH; ][]{engelund1967}, and Bagnold \citep{bagnold1980}; additionally width-integrated shear stress is assessed, as used as proxy for sediment transport above.
@@
 
-Several aspects of the models and experiment design make the comparison of the subaerial and subglacial systems difficult. 
+Several aspects of the models and experiment design make the comparison of the subaerial and subglaical systems difficult. 
 The lumped nature of the models means that they operate independently of the upstream drainage network.

### 2025-02-03 revised_v1/response_R1.tex
small changes

@@
 \begin{document}
+Dear Editor,
 
-\vspace{.1cm}
+\vspace{.3cm}
+
+We thank Reviewer 1 for their detailed and constructive comments. There work, we believe, has provided feed back that has greatly improved the manuscript.
+
+The reviewer's comments are in bold, our response is in italics and quotations from the new text are in normal font.
+
+\vspace{.3cm}
 
@@
 
-\vspace{.1cm}
+\vspace{.75cm}
 
 Ian Delaney on behalf of all authors
+
+\vspace{2cm}
 
@@
 
-\ian{Comeback here}
+ \textit{Excellent comment.
+The first sentence of section will read:} ``The first numerical experiment aims to quantify the timing and covariance of the subglacial model outputs with respect to water discharge in response to different seasonal evolutions and peaks''
 
@@
     As the water discharge stabilizes at its peak, channel growth continues (Figure~2 a, e), causing water velocity and other model outputs to decrease from their peak values (Figure~2 b-d, f-h).
-    Subglacial sediment transport capacity is greatest on the hydrograph's rising limb, relative to the falling limb, creating a hysteresis effect.''}
+    Subglacial sediment transport capacity is greatest on the hydrograph's rising limb, relative to the falling limb, creating a hysteresis effect.''
 

### 2025-02-03 revised_v1/response_R2.tex
small changes

@@
 \begin{document}
+Dear Editor,
+
+\vspace{.3cm}
+
+We thank Reviewer 2 for their comments.
+We believe that in addressing them, we have clarified the model presentation.
+We note, however, that the presentations of criticisms by the reviewer was at times difficult to respond to, given the format of the review.
+Please let us know if additionally points should be addressed.
+
+The reviewer's comments are in bold, our response is in italics and quotations from the new text are in normal font.
+
+\vspace{.3cm}
+
+Best regards,
+
+\vspace{.75cm}
+
+Ian Delaney on behalf of all authors
+
+\vspace{2cm}
 
@@
 
-  \textit{We thank the reviewer for this comment, as it demands clarification. The text now reads:}
+  \textit{We thank the reviewer for this comment, as it demands clarification in our model presentation.
+In order to reduce the number of variables in the original text, $h_p$ was omitted as its value is zero, as is hydraulic head at the glacier terminus.
+The text has been updated to clarify this point.
+    Additionally, we confirm the reviewers comment, that effective pressure at the glacier terminus is zero. The text will read:}
   The evolution of subglacial channel size $S_g$ is given as
@@
 
-  \textit{In order to reduce the number of variables in the original text, $h_p$ was omitted as its value is zero, as is hydraulic head at the glacier terminus.
-The text has been updated to clarify this point.
-    Additionally, we confirm the reviewers comment, that effective pressure at the glacier terminus is zero.
+  \textit{
     The reviewer also correctly points out that water de-pressurizes underneath the glacier, in agreement with the other reviewer, we have included a comment that we assume pressurized flow across the glacier bed, as is common in subglacial hydrology models, although this may not always be the case \citep{perolo2018}.
@@
     As a result, we believe that these result hold across as range of water discharge- channel width relationships, including ones where $\alpha$ evolves in time.
-    The exception, shown in Section 4.3 (now 5.1), is that when water discharge evolves more slowly that channel size.
+    The exception, shown in Section 4.3 (now 6.1), is that when water discharge evolves more slowly that channel size.
 Here, the behavior between subaerial and subglacial is very similar, because channel area in both channel type adjust to discharge variations.}

### 2025-02-03 revised_v1/response_editor.tex
small changes

--- 2025-02-03 revised_v1/response_R2.tex
@@
 
-\newcommand{\alpine}{\textit{ALPINE}\,}
-\newcommand{\icesheet}{\textit{ICESHEET}\,}
 \newcommand{\TODO}[1]{\textbf{TODO #1}}
@@
 \newcommand{\aref}[1]{\textbf{Reference #1}}
+\newcommand{\alpine}{\textit{ALPINE}\,}
+\newcommand{\icesheet}{\textit{ICESHEET}\,}
 \title{}
@@
 \begin{document}
+Dear Editor,
 
-\textbf{General Comments}
+\vspace{.3cm}
 
-\begin{itemize}
+We thank you for taking on this manuscript and coordinating the reviews. We found both reviews very helpful for improving the manuscripts.
 
-\item \textbf{
-    The trouble is, that while the first term on the right hand side of (3) makes sense to me, the second does not, and the reason for this is that the closure term involves the effective pressure and thus the hydraulic head, but not its gradient, so I am a bit sceptical about this, because the structure of the Nye model seems to have been altered.
-    You might say that the distinction disappears in the lumping, but it seems to me that if you have an upstream head h- and a downstream head h+, the first term on the right of (3) will involve the difference, but the second will involve the average, and there are two independent quantities involved.
-    I suppose you can get around this by saying, oh, the outlet head or more accurately the effective pressure is zero, and perhaps that is what is done.
-    But that is a bit disingenuous, because the outflow becomes open channel flow somewhere upstream of the actual mouth of the stream.
-    Now I went and looked up this Werder 2010 paper, and you can see in its equation (4) the same distinction I am making here.
-    I suppose my overall view is that (I think) the Clarke paper intended to use the flow elements of resistors, etc., as ingredients of a larger scale flow path, and this lumping of a whole channel is an extremely coarse thing to do, and looks a bit like avoiding confronting the spatially-dependent physics simply because it’s too hard.
-    So I think you can rescue this aspect of the model presentation, but a bit more exposition would help, in particular the figure 2or equivalent of the 2010 paper would help.}
+\vspace{.2cm}
 
-  \textit{We thank the reviewer for this comment, as it demands clarification. The text now reads:}
-  The evolution of subglacial channel size $S_g$ is given as
-  \begin{linenomath*}
-    \begin{equation}
-      \label{eq:dS_dt}
-      \frac{\partial S_g}{\partial t} = C_1 \frac{Q \Delta h}{l} - C_2 \left(h_{o}-\overline{h}\right)^n\,S_g,
-    \end{equation}
-  \end{linenomath*}
-  \noindent where $t$ is time, $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (values in Table~1), $g$ is the acceleration due to gravity, $Q$ is water discharge, $\overline{h}=\frac{1}{2}(h+h_p)$ is the mean hydraulic head in the channel with $h_p$ being the proglacial hydraulic head equal to zero, $l$, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the mean ice overburden pressure expressed in meter water equivalent ($\rho_w$ is density of water; $\rho_i$ is density of ice), and $n$ is Glen's n \citep[usually $n=3$; ][]{glen1955}.
-  The first term on the equation's right side represents the channel opening by frictional heating, while the following term represents channel closure from ice deformation.
+Following Reviewer 1, we have modified the first objective of the manuscript to describe the conditions where covariance between sediment transport capacity and water discharge could occur.
+To develop this point further, we have conducted additional experiments with a series of smoothed hydrographs to better quantify the conditions when covariance could occur.
+Additionally, we have reorganized and modified the text in response to their comments.
+Lastly, at several points, in the review, it was mentioned that a point was already established.
+It could be a substantial oversight of the literature on our part as authors.
+However, we would be grateful if the reviewer would provide the specific references in these cases to help us fully address their comments.
 
-  \textit{In order to reduce the number of variables in the original text, $h_p$ was omitted as its value is zero, as is hydraulic head at the glacier terminus.
-The text has been updated to clarify this point.
-    Additionally, we confirm the reviewers comment, that effective pressure at the glacier terminus is zero.
-    The reviewer also correctly points out that water de-pressurizes underneath the glacier, in agreement with the other reviewer, we have included a comment that we assume pressurized flow across the glacier bed, as is common in subglacial hydrology models, although this may not always be the case \citep{perolo2018}.
-  }
+\vspace{.2cm}
 
-  \textit{Lastly, we have chosen a simple model, lumped element, or channel segment model here as we believe that it most clearly and concisely explains the dynamics that we wish to discuss.
-    Our strong opinion is that the experiments that we have done with the model, across a range of hydrographs, demonstrate consistently the behavior of subglacial channels.
-    Therefore, we do not believe that more complex model is needed to show these processes.
-    The limitation of this approach is acknowledge and discussed.
-  }
+With respect to Reviewer 2, we believe that their specific comments about the model have been addressed by generalizing the equations in questions.
+We understand their critique about the accumulation of water flow along the glacier's length.
+However, we cannot find a rational that would suggest that these findings would not over shorter channels or in a more complex (i.e. spatially distributed) model.
+To further highlight their concerns, we have moved the ``Experiment limitations'' discussion to its own section following the ``Results.''
 
-\item  \textbf{And the model should come with some caveats, e. g., you could say that it is a simplistic and possibly unreliable model. To be fair, this is done in section 5.3, but that is too late.}
+\vspace{.2cm}
+Again we are grateful for your assistance with the manuscript and hope that you find our responses compelling
 
-  \textit{The ``Methods'' section is largely meant to describe the model, as opposed to discuss its behavior or simplicity.
-In response to this comment, we will move the ``Experiment limitations'' to a section that directly after the ``Results'' but before the ``Discussion.''}
+\vspace{.5cm}
 
-\item \textbf{  One of the comments made is that because the sub-aerial model has an algebraic relation between width and discharge (equation 9), the relationship between velocity and discharge is also algebraic; but this is simply a choice of the model.
-In reality, the width of a channel will also evolve over (long) time through processes of bank erosion, and the use of equation 9 properly only involves a long term time average of discharge, so it is misleading to use it in a time-specific way.
-Actually, this is admitted at line 370, along with other limitations in the models.}
+Best regards,
 
-  \textit{We appreciate these comments. With regard to the first point about channel width evolution in the subaerial model, this matter is confronted in Section 4.2 and 4.3 (now 5.1).
-    Here we show that across a range of exponents and relationships between channel cross section and water discharge, subglacial systems are still more sensitive to water discharge variations that subaerial ones.
-    This increased variability occurs across a range of $\alpha$ values that describe the relationship between channel width and water discharge as shown in Figure 6 (now 7).
-    As a result, we believe that these result hold across as range of water discharge- channel width relationships, including ones where $\alpha$ evolves in time.
-    The exception, shown in Section 4.3 (now 5.1), is that when water discharge evolves more slowly that channel size.
-Here, the behavior between subaerial and subglacial is very similar, because channel area in both channel type adjust to discharge variations.}
+\vspace{.75cm}
 
-\item   \textbf{Two kinds of numerical experiment are described in section 3.3, and the results of these presented in section 4.
-Presumably the input to the model runs is the (time- varying) water discharge, but that seems not to be stated in 3.3, which I would expect, although it is stated in the caption to figure 2.
-    The second half of the paper characterises the results of these experiments and there is an extended discussion.
-There is no doubt the authors have put a lot of effort into this, but my interest waned at this point.}
-
-  \textit{    We have modified the first sentence of Section 3.3 to explicitly state:}
-    ``Proglacial discharge records from the Fieschergletscher (scenario \alpine{}) and the Leverett Glacier (scenario \icesheet{}) are used as inputs for the models above.''
-
-\item \textbf{  In summary, this is one of those difficult papers to judge, because the effort in- volved is quite substantial and honestly applied, but the whole philosophy of the approach is, in my view, misguided.
-This is a coxless boat crew who are rowing up a backwater without a rudder, and have lost the main direction of the stream.}
-
-  \textit{With respect to the last sentence, it is true that the lead author feels this way occasionally. Regardless, we think valuable findings are within this manuscript.}
-
-\end{itemize}
-
-\textbf{Specific Comments}
-
-\begin{itemize}
-\item \textbf{33: I don’t know what ‘Following mass conservation’ has to do with this.}
-
-  \textit{We will remove this phrase.}
-
-\item \textbf{99-100: upon that of Clarke.}
-
-  \textit{Done.}
-
-\item \textbf{102: presumably $h_{ice}$ is the thickness of the ice, but the syntax is not clear, it could be the depth of the channel.}
-
-  \textit{The text will read:} `` a glacier with channel length $l$, with a flat bed and a mean ice thickness of $h_{ice}$.''
-
-\item \textbf{eq. 3: it’s probably in the Werder 2010 paper,} but I’d like to see an extra comment about where this $\frac{\Delta h}{2}$ term comes from.
-  Since the bracket represents $N = pi-pw$, $pw$ the pointwise form of this second term would be $\frac{p_w}{\rho_wg}$, so it’s not obvious to me where the $\Delta h$ comes from.
-  Also in this light, it is worth elaborating the confusing term ‘hydraulic head drop change’ and defining what $\Delta h$ is in terms of $pw$.
-  Actually, the more I think about this, the more suspicious I become.
-And in fact, at least when you’re dealing with j\"okulhlaups, the discharge doesn’t vary much along the length of the channel, and the consequence of that is that neither does the effective pressure.
-  Of course, that may not apply here but I think the same principle applies.
-  So maybe I am promoting this to a more substantive issue
-  (as indeed now further discussed earlier).
-
-  \textit{The comment about $\frac{\Delta h}{2}$ is addressed above.
-We understand the point about flow accumulation and increasing water discharge along the channel.
-As the author points out, the implications of this are presented in the ``Experiment limitations.'' Additionally, we note that despite these effects, the results here show convincingly that the main topics in the paper still hold, i.e.
-sediment transport capacity is more variable in subglacial channels than subaerial ones and sediment transport capacity does not vary with water discharge if the subglacial channel size is out of equilibrium with water discharge.
-}
-
-\item \textbf{114:} I had no idea what the central angle referred to, but evidently it is the angle subtended at the centre of the circular arc of the channel upper boundary; and then $\beta = 2\alpha$, where $\alpha$ is the contact angle at the channel edge, which seems a more natural quantity to use.
-Also I checked the algebra of equation (5), and got this formula, but with the factor of two in the numerator, not the denominator, assuming hydraulic diameter is twice hydraulic radius, the latter of which is area/perimeter, so please check this.
-Incidentally, I did also then check equation (7) and I agree with that, so I do think there is an error in (5).
-
-  \textit{We thank the review for the careful assessment of our work. We have double checked an believe that there is no problem in the equation. Our proof is below:}
-
-   \includegraphics[width=0.3\linewidth]{Hooke.png}
-  Guiding equations:
-
-  \begin{equation}
-    D_h = \frac{4S_g}{P_w}~~~(I)
-  \end{equation}
-
-  \begin{equation}
-    P_w = R \beta +2R \sin\frac{\beta}{2}= R(\beta +2\sin \frac{\beta}{2} ~~~(II)
-  \end{equation}
-
-  \begin{equation}
-    S_g=\frac{R^2}{2}(\beta - \sin \beta)~~~(III)
-  \end{equation}
-
-  $D_h$ is hydraulic diameter, $P_w$ is wetter perimeter, $R$ is radius, $\beta$ is Hooke angle \citep{hooke1990}, and $S_g$ is the subglacial channel cross section.
-
-  Proof:
-
-  \begin{linenomath*}
-    \begin{equation}
-      \frac{D_h}{4S_g} = \frac{1}{R\beta + 2R\sin\frac{\beta}{2}}
-    \end{equation}
-  \end{linenomath*}
-
-  \begin{equation}
-    R = \frac{4S_g}{D_h (\beta +2 \sin \frac{\beta}{2})}, \quad \mathrm{into~} III 
-  \end{equation}
-
-  \begin{equation}
-    S_g = \frac{16 S_g^2}{D_h^2 (\beta +2\sin\frac{\beta}{2})^2} ~\frac{1}{2}(\beta-\sin\beta)
-  \end{equation}
-
-  \begin{equation}
-    1 = \frac{8 S_g}{D_h^2 (\beta +2\sin\frac{\beta}{2})^2} (\beta-\sin\beta)
-  \end{equation}
-
-  \begin{equation}
-    \frac{D_h^2}{8} \frac{(\beta+2\sin\frac{\beta}{2})^2}{\beta - \sin \beta} = S_g
-  \end{equation}
-
-  \begin{equation}
-    \frac{D_h^2 (2(\frac{\beta}{2}+\sin \frac{\beta}{2}))^2}{8(\beta - \sin \beta)} = S_g
-  \end{equation}
-
-   \begin{equation}
-    \frac{D_h^2 ((\frac{\beta}{2}+\sin \frac{\beta}{2}))^2}{2(\beta - \sin \beta)} = S_g
-  \end{equation}
-
-  \begin{equation}
-       S_g = \frac{D_h^2}{2} \frac{(\frac{\beta}{2}+\sin \frac{\beta}{2})^2}{(\beta - \sin \beta)}  
-  \end{equation}
-
-  This is the same as Equation 5 of our manuscript.
-
-\end{itemize}
+Ian Delaney on behalf of all authors
 

### 2025-02-03 Delaneyetal2024.tex
small changes

--- 2025-02-03 Delaneyetal2024.tex
@@
 
-The first numerical experiment aims to quantify the sources of increased variability in the subglacial model outputs as they exhibit different seasonal evolutions and peaks (Figure~\ref{fig:model_outs}).
+The first numerical experiment aims to quantify the timing and covariance of the subglacial model outputs with respect to water discharge in response to different seasonal evolutions and peaks (Figure~\ref{fig:model_outs}).
 Variable relationships between model outputs and water discharge emerge for the subaerial and subglacial cases as channel size evolves (Figure~\ref{fig:model_outs} a, e).
@@
   We benefited from fruitful discussions with J. Irving.
-  We thank two reviewers for constructive and critical comments on a previous version of the manuscript.
+  We thank B. Hubbard and an anonymous reviewer for constructive and critical comments on the manuscript.
 \end{acknowledgements}

### 2025-02-03 revised_v1/response_R1.tex
changes to share

@@
 
-  \textit{We thank the Reviewer for the positive assessment of our work and take note of the structural issues along with the organizational ones. }
+  \textit{We thank the Reviewer for their positive assessment of our work and acknowledge the structural and organizational issues they have highlighted.}
 
@@
 
-  \textit{We value this comment greatly. Upon consideration, we have chosen to rephrase  objective one as ``to establish the hydrological conditions where sediment transport capacity co-varies
-    with water discharge in subglacial systems.'' 
-    In our view this objective focus the attention on the differences between the ALPINE and ICESHEET scenarios.
-It also is practical when evaluating the differences in smooth of water discharge, as discussed in the next point.
-    We appreciate the need to remove interpretation from the results section.
-We have done this.
-Additionally, Results Section 4.3 with the algebraic formulations has been moved to the discussion, as we think the pipe flow and steady state R-channel scenarios serve as useful end members through which to discuss the results.
+  \textit{We greatly appreciate this comment.
+After careful consideration, we have chosen to rephrase Objective One as:} ``to establish the hydrological conditions under which sediment transport capacity co-varies with water discharge in subglacial systems.''
+
+\textit{In our view, this objective directs attention to the differences between the ALPINE and ICESHEET scenarios.
+It is also practical for evaluating the effects of water discharge smoothing, as discussed in the next point.}
+
+\textit{We acknowledge the importance of removing interpretation from the results section and have made the necessary revisions.
+Additionally, we will moved Section 4.3, which contains the algebraic formulations, to the discussion section, as we believe that the pipe flow and steady-state R-channel scenarios serve as useful end members for interpreting the results.
 }
@@
 
-  \textit{This is an excellent comment. We have worked to address this need for a clearer hysteresis explanation.
-    One challenge with the approach recommended by the reviewer is that sediment transport capacity scales very non-linearly with water discharge, especially in subglacial systems.
-This is shown in Section 4.3.
-    As a result, findings across hydrographs would not really be comparable even if they were scaled with a discharge quantity.
-    Instead, we will add several paragraphs that to section 4.1 that discuss The role of water discharge variability on hysteresis.
+  \textit{This is an excellent comment, and we have worked to provide a clearer explanation of hysteresis.}
+
+\textit{One challenge with the approach recommended by the reviewer is that sediment transport capacity scales highly non-linearly with water discharge, particularly in subglacial systems, as demonstrated in Section 4.3.
+Consequently, findings across different hydrographs would not be directly comparable, even if they were scaled using a discharge quantity.}
+
+\textit{Instead, we have added several paragraphs to Section 4.1 discussing the role of water discharge variability in hysteresis.}
 
@@
 
-  \textit{Excellent suggestion, title follows the reviewer's suggestion: Sediment transport capacity response to variations in water discharge in pressurized subglacial channels}
+  \textit{Excellent suggestion, title follows the reviewer's suggestion:} Sediment transport capacity response to variations in water discharge in pressurized subglacial channels
 
@@
 
-  \textit{Given the discussion above, the text will read: ``to establish the hydrological conditions where  sediment transport capacity co-varies with water discharge in subglacial systems''}
+  \textit{Given the discussion above, the text will read:} ``to establish the hydrological conditions where  sediment transport capacity co-varies with water discharge in subglacial systems''
 
@@
 
-  \textit{The text will now read: ``with random parameter values in a range''}
+  \textit{The text will now read:} ``with random parameter values in a range''
 
@@
   \textit{This discusses the role water discharge on hysteresis in water velocity or sediment transport capacity.
-Thus dealing with objective 1 of the manuscript.
+Thus dealing with Objective One of the manuscript.
 We have also noted the comment about results and interpretation.
@@
 However, their very different values mean that in makes sense to adjust values to examine their changes during the extreme events.
-The text has been modified slightly to read: ``different y-axis ranges for subaerial and subglacial values.''}
+The text will be modified slightly to read:} ``different y-axis ranges for subaerial and subglacial values.''
 
@@
 We believe that the content in this paragraph describes the response to water discharge in Figure 2.
-We agree that the reference to the Methods section, gave the impression of the interpretation.
+We agree that the reference to the ``Methods'' section, gave the impression of the interpretation.
 The paragraph will be adjusted slightly to read:}
@@
 
-  \textit{ Sentence will be changed to:''Smaller values of channel factor $\beta$, creating low and broad channels where the channel width grows more quickly in response to water discharge increases as compared to a semi-circular channel with $\beta = \pi$ (Equation~4).
-''}
+  \textit{ Sentence will be changed to:}''Smaller values of channel factor $\beta$, creating low and broad channels where the channel width grows more quickly in response to water discharge increases as compared to a semi-circular channel with $\beta = \pi$ (Equation~4).''
 
@@
 
-  \textit{We will modify the introductory text to read: ``The numerical experiments above consider the size evolution of subglacial channels and demonstrate that for these hydrographs subglacial sediment transport variability is greater than its subaerial counterpart (Section~4.3).
+  \textit{We will modify the introductory text to read}: ``The numerical experiments above consider the size evolution of subglacial channels and demonstrate that for these hydrographs subglacial sediment transport variability is greater than its subaerial counterpart (Section~4.3).
 Additionally, results demonstrate the impact of water discharge variability on sediment transport capacity.
 Here, we compare the sediment transport behavior of different channel types as they respond to water discharge, channel shape, and hydraulic gradient.''
-}
 
@@
 
-  \textit{The sentence will read: ``This hysteresis can limit water discharge as an indicator of sediment discharge capacity in these systems, especially when water discharge is highly variable and out of equilibrium with subglacial channel size.''}
+  \textit{The sentence will read:} ``This hysteresis can limit water discharge as an indicator of sediment discharge capacity in these systems, especially when water discharge is highly variable and out of equilibrium with subglacial channel size.''
 
@@
 
-\textit{This paragraph will read: ``The manuscript's first objective is to establish the conditions where water discharge covaries with subglacial sediment transport capacity.
+\textit{This paragraph will read:} ``The manuscript's first objective is to establish the conditions where water discharge covaries with subglacial sediment transport capacity.
 In subglacial channels, the timing of peak water velocity and sediment transport capacity occurs before peak water discharge during a discharge event, due to evolving channel size.
@@
 Water discharge and sediment transport capacity covariance could be possible when water discharge varies at a slower rate than subglacial channel size, such as Antarctica with minimal surface melt input.
-Results also suggest that water discharge records averaged over periods longer than $12$ hours from Alpine and Ice Sheet hydrographs show substantial impacts on subglacial sediment transport capacity characteristics.''}
+Results also suggest that water discharge records averaged over periods longer than $12$ hours from Alpine and Ice Sheet hydrographs show substantial impacts on subglacial sediment transport capacity characteristics.''
 

### 2025-02-03 revised_v1/response_editor.tex
changes to share

@@
 
-\vspace{.3cm}
-
-We thank you for taking on this manuscript and coordinating the reviews. We found both reviews very helpful for improving the manuscripts.
+We thank you for handling this manuscript and coordinating the reviews. We found both reviews very helpful in improving the manuscript.
 
@@
 
-Following Reviewer 1, we have modified the first objective of the manuscript to describe the conditions where covariance between sediment transport capacity and water discharge could occur.
-To develop this point further, we have conducted additional experiments with a series of smoothed hydrographs to better quantify the conditions when covariance could occur.
-Additionally, we have reorganized and modified the text in response to their comments.
-Lastly, at several points, in the review, it was mentioned that a point was already established.
-It could be a substantial oversight of the literature on our part as authors.
-However, we would be grateful if the reviewer would provide the specific references in these cases to help us fully address their comments.
+Following Reviewer 1’s feedback, we have revised the first objective of the manuscript to better describe the conditions under which covariance between sediment transport capacity and water discharge may occur.
+To further develop this point, we conducted additional experiments using a series of smoothed hydrographs to better quantify the conditions that lead to covariance.
+Additionally, we have reorganized and revised the text in response to their comments.
+
+At several points in the review, it was noted that a particular point had already been established.
+If this reflects a substantial oversight of the literature on our part, we sincerely appreciate the reviewer bringing it to our attention.
+However, we would be grateful if the reviewer could provide specific references in these instances to help us fully address their concerns.
 
@@
 
-With respect to Reviewer 2, we believe that their specific comments about the model have been addressed by generalizing the equations in questions.
-We understand their critique about the accumulation of water flow along the glacier's length.
-However, we cannot find a rational that would suggest that these findings would not over shorter channels or in a more complex (i.e. spatially distributed) model.
-To further highlight their concerns, we have moved the ``Experiment limitations'' discussion to its own section following the ``Results.''
+Regarding Reviewer 2, we believe that their specific comments about the model have been addressed by generalizing the equations in question.
+We acknowledge their critique concerning the accumulation of water flow along the glacier’s length.
+However, we do not see a rationale suggesting that these findings would not hold for shorter channels or within a more complex (i.e., spatially distributed) model.
+To further emphasize these concerns, we have moved the ``Experiment Limitations'' discussion to a separate section following the ``Results'' section.
 
 \vspace{.2cm}
-Again we are grateful for your assistance with the manuscript and hope that you find our responses compelling
+
+Once again, we sincerely appreciate your assistance with this manuscript and hope that you find our responses satisfactory.
 

### 2025-02-03 Delaneyetal2024.tex
changes to share

@@
   The results also indicate that high subglacial sediment transport capacities can occur across a wide range of water discharges.
-  Reducing water discharge variability lessens the hysteresis effects, in some cases to the point where a covarying relationship between water discharge and sediment transport capacity may occur, similar to subaerial systems.
+  Reducing water discharge variability by smoothing lessens the hysteresis effects, in some cases to the point where a covarying relationship between water discharge and sediment transport capacity may occur, similar to subaerial systems.
   A second set of numerical experiments shows that subglacial sediment transport is highly non-linear with respect to water discharge, creating more variability in sediment transport capacity.

### 2025-02-03 Delaneyetal2024.tex
Updates from Overleaf

--- 2025-02-03 Delaneyetal2024.tex
@@
 
-The first numerical experiment aims to quantify the timing and covariance of the subglacial model outputs with respect to water discharge in response to different seasonal evolutions and peaks (Figure~\ref{fig:model_outs}).
+The first numerical experiment aims to quantify the sources of increased variability in the subglacial model outputs as they exhibit different seasonal evolutions and peaks (Figure~\ref{fig:model_outs}).
 Variable relationships between model outputs and water discharge emerge for the subaerial and subglacial cases as channel size evolves (Figure~\ref{fig:model_outs} a, e).
@@
   We benefited from fruitful discussions with J. Irving.
-  We thank B. Hubbard and an anonymous reviewer for constructive and critical comments on the manuscript.
+  We thank two reviewers for constructive and critical comments on a previous version of the manuscript.
 \end{acknowledgements}

### 2025-02-04 revised_v1/response_R1.tex
Grammar check on Overleaf

@@
 
-\vspace{.3cm}
+\vspace{.2cm}
 
-We thank Reviewer 1 for their detailed and constructive comments. There work, we believe, has provided feed back that has greatly improved the manuscript.
+We thank Reviewer 1 for their detailed and constructive comments. Their work, we believe, has provided feedback that has greatly improved the manuscript.
+
+\vspace{.2cm}
 
@@
 
-\item  \textbf{Here, most of what is novel seems to fgall under Objective
+\item  \textbf{Here, most of what is novel seems to fall under Objective
     2 and not under Objective 1. Indeed, isn’t Objective 1 (‘to establish whether sub-seasonal water
@@
 
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.8\linewidth]{../Fig4.png}
+  \caption{Spearman rank correlation between water discharge over a smoothed period and sediment transport characteristics (a) water velocity, (b) shear stress, (c) width-integrated shear stress.
+Higher correlation scores mean reduced hysteresis and approach subaerial behavior.
+  }
+  \label{fig:corri}
+\end{figure}
+
 Water discharge and sediment transport characteristics correlate better in \icesheet{} with lower diurnal variations in water discharge compared to the \alpine{}.
@@
 Correlations in sediment transport characteristics in \icesheet{} remain higher than \alpine, which has more water discharge variations even when smoothed compared to \icesheet{} (Figure~S1-S14).
-
-\begin{figure}[hbt!]
-  \centering
-  \includegraphics[width=0.8\linewidth]{../Fig4.png}
-  \caption{Spearman rank correlation between water discharge over a smoothed period and sediment transport characteristics (a) water velocity, (b) shear stress, (c) width-integrated shear stress.
-We assume that higher correlation scores approach subaerial behavior.
-  }
-  \label{fig:corri}
-\end{figure}
 
@@
 
-  \textit{ We thank the reviewer for this comment as it, in part, encouraged us to add the additional section above with the water discharge smoothing.
+  \textit{We thank the reviewer for this comment as it, in part, encouraged us to add the additional section above with the water discharge smoothing.
     The other hydrological scenario we are aware of not explicitly considered here is Antarctica, where little surface melt occurs.
@@
    \textit{ We will add the following sentence to the model implementation section:}
-``We also note that the subglacial results apply to pressurize subglacial channels, not depressurized one that can occur below glacier termini \citep{perolo2018}''
+``We also note that the subglacial results apply to pressurized subglacial channels, not depressurized one that can occur below glacier termini \citep{perolo2018}''
 \end{itemize}
@@
 
-  \textit{Excellent suggestion, title follows the reviewer's suggestion:} Sediment transport capacity response to variations in water discharge in pressurized subglacial channels
+  \textit{Excellent suggestion, the title follows the reviewer's suggestion:} Sediment transport capacity response to variations in water discharge in pressurized subglacial channels
 
@@
 
-  \textit{ We believe that this is discussed in Lines 51--54 in the introduction.}
+  \textit{We believe that this is discussed in Lines 51--54 in the introduction.}
 
@@
   \textit{This section will be retitled:} ``Timing of subaerial and subglacial sediment transport capacity variations.''
-  \textit{This discusses the role water discharge on hysteresis in water velocity or sediment transport capacity.
+  \textit{This discusses the role of water discharge on hysteresis in water velocity or sediment transport capacity.
 Thus dealing with Objective One of the manuscript.
@@
  \textit{Excellent comment.
-The first sentence of section will read:} ``The first numerical experiment aims to quantify the timing and covariance of the subglacial model outputs with respect to water discharge in response to different seasonal evolutions and peaks''
+The first sentence of the section will read:} ``The first numerical experiment aims to quantify the timing and covariance of the subglacial model outputs with respect to water discharge in response to different seasonal evolutions and peaks''
 
@@
   \textit{We appreciate the comment.
-We believe that given the nature of subglacial channels as describe in Equation 3, evolution is a more accurate description.
+We believe that given the nature of subglacial channels, as described in Equation 3, evolution is a more accurate description.
 To us, ``change'' could include a modification for any number of reasons.
 However,
-    ``evolution'' conotates the dependence on anticedent conditions which we believe to be more accurate in this case. }
+    ``evolution'' connotates the dependence on antecedent conditions which we believe to be more accurate in this case. }
 
@@
 
-  \textit{We believe that description of the insets' axis is needed as these lines have the same axis in the main plots.
-However, their very different values mean that in makes sense to adjust values to examine their changes during the extreme events.
+  \textit{We believe that a description of the insets' axis is needed as these lines have the same axis in the main plots.
+However, their very different values mean that it makes sense to adjust values to examine their changes during the extreme melt event.
 The text will be modified slightly to read:} ``different y-axis ranges for subaerial and subglacial values.''
@@
 
-  \textit{ Sentence will be changed to:}''Smaller values of channel factor $\beta$, creating low and broad channels where the channel width grows more quickly in response to water discharge increases as compared to a semi-circular channel with $\beta = \pi$ (Equation~4).''
+  \textit{ Sentence will be changed to:}'' Smaller values of channel factor $\beta$, creating low and broad channels where the channel width grows more quickly in response to water discharge increases as compared to a semi-circular channel with $\beta = \pi$ (Equation~4).''
 
@@
 
-  \textit{We have considered the comments here, and with the editors support, think it fits best here.
+  \textit{We have considered the comments here, and with the editor's support, think it fits best in the manuscript.
+  However, we have moved it to the discussion to set up the end member cases of pipe flow and steady-state R-channels.
     In a previous version of the manuscript, it was included in the supplementary material.
@@
 We believe that Tables 2 and 3 in the section are valuable results in evaluating the sensitivity of sediment transport capacity to water discharge under different assumptions.
-    The importance of these findings are discussed in lines xxx yyy.
   }
@@
 
-  \textit{We agree with  about the lack of data. This sentence will been removed. }
+  \textit{We agree with the reviewer about the lack of data. This sentence will be removed. }
 
@@
 In subaerial channels, the timing of peak sediment transport capacity and water discharge coincide.
-Results here suggest that, even in a transport-limited subglacial system, an variable relationship between water and subglacial sediment discharge.
-Water discharge variations in both Greenland Ice Sheet and Alpine are variable enough to cause this variable relationship.
+Results here suggest that, even in a transport-limited subglacial system, a variable relationship between water and subglacial sediment discharge.
+Water discharge variations in both the Greenland Ice Sheet and Alpine cases are variable enough to cause this variable relationship.
 This variable relationship presents a challenge in linking hydro-climatic conditions or events to sediment export from glaciers.

### 2025-02-04 revised_v1/response_R2.tex
Grammar check on Overleaf

@@
 
-\vspace{.3cm}
+\vspace{.2cm}
 
@@
 We note, however, that the presentations of criticisms by the reviewer was at times difficult to respond to, given the format of the review.
-Please let us know if additionally points should be addressed.
+Please let us know if additional points should be addressed.
+
+\vspace{.2cm}
 
@@
   \textit{We thank the reviewer for this comment, as it demands clarification in our model presentation.
-In order to reduce the number of variables in the original text, $h_p$ was omitted as its value is zero, as is hydraulic head at the glacier terminus.
+To reduce the number of variables in the original text, $h_p$ was omitted as its value is zero, as is the hydraulic head at the glacier terminus.
 The text has been updated to clarify this point.
-    Additionally, we confirm the reviewers comment, that effective pressure at the glacier terminus is zero. The text will read:}
+    Additionally, we confirm the reviewer's comment, that effective pressure at the glacier terminus is zero. The text will read:}
   The evolution of subglacial channel size $S_g$ is given as
@@
     Our strong opinion is that the experiments that we have done with the model, across a range of hydrographs, demonstrate consistently the behavior of subglacial channels.
-    Therefore, we do not believe that more complex model is needed to show these processes.
-    The limitation of this approach is acknowledge and discussed.
+    Therefore, we do not believe that a more complex model is needed to show these processes.
+    The limitation of this approach is acknowledged and discussed.
   }
@@
 
-  \textit{The ``Methods'' section is largely meant to describe the model, as opposed to discuss its behavior or simplicity.
-In response to this comment, we will move the ``Experiment limitations'' to a section that directly after the ``Results'' but before the ``Discussion.''}
+  \textit{The ``Methods'' section is largely meant to describe the model, as opposed to discussing its behavior or simplicity.
+In response to this comment, we will move the ``Experiment limitations'' to a section directly after the ``Results'' but before the ``Discussion.''}
 
@@
 
-  \textit{We appreciate these comments. With regard to the first point about channel width evolution in the subaerial model, this matter is confronted in Section 4.2 and 4.3 (now 5.1).
-    Here we show that across a range of exponents and relationships between channel cross section and water discharge, subglacial systems are still more sensitive to water discharge variations that subaerial ones.
+  \textit{We appreciate these comments. With regard to the first point about channel width evolution in the subaerial model, this matter is confronted in Sections 4.2 and 4.3 (now 5.1).
+    Here we show that across a range of exponents and relationships between channel cross-section and water discharge, subglacial systems are still more sensitive to water discharge variations than subaerial ones.
     This increased variability occurs across a range of $\alpha$ values that describe the relationship between channel width and water discharge as shown in Figure 6 (now 7).
-    As a result, we believe that these result hold across as range of water discharge- channel width relationships, including ones where $\alpha$ evolves in time.
-    The exception, shown in Section 4.3 (now 6.1), is that when water discharge evolves more slowly that channel size.
-Here, the behavior between subaerial and subglacial is very similar, because channel area in both channel type adjust to discharge variations.}
+    As a result, we believe that these results hold across a range of water discharge- channel width relationships, including ones where $\alpha$ evolves in time.
+    The exception, shown in Section 4.3 (now 6.1), is that when water discharge evolves more slowly than channel size.
+Here, the behavior between subaerial and subglacial is very similar, because channel areas in both channel types adjust to discharge variations.}
 
-\item   \textbf{Two kinds of numerical experiment are described in section 3.3, and the results of these presented in section 4.
+\item   \textbf{Two kinds of numerical experiments are described in section 3.3, and the results of these presented in section 4.
 Presumably the input to the model runs is the (time- varying) water discharge, but that seems not to be stated in 3.3, which I would expect, although it is stated in the caption to figure 2.
@@
 
-  \textit{    We have modified the first sentence of Section 3.3 to explicitly state:}
+  \textit{To make it clear that water discharge is the input, we will modify the first sentence of Section 3.3 to explicitly state:}
     ``Proglacial discharge records from the Fieschergletscher (scenario \alpine{}) and the Leverett Glacier (scenario \icesheet{}) are used as inputs for the models above.''
@@
 
-\item \textbf{eq. 3: it’s probably in the Werder 2010 paper,} but I’d like to see an extra comment about where this $\frac{\Delta h}{2}$ term comes from.
+\item \textbf{eq. 3: it’s probably in the Werder 2010 paper, but I’d like to see an extra comment about where this $\frac{\Delta h}{2}$ term comes from.
   Since the bracket represents $N = pi-pw$, $pw$ the pointwise form of this second term would be $\frac{p_w}{\rho_wg}$, so it’s not obvious to me where the $\Delta h$ comes from.
@@
   So maybe I am promoting this to a more substantive issue
-  (as indeed now further discussed earlier).
+  (as indeed now further discussed earlier).}
 
@@
 
-  \textit{We thank the review for the careful assessment of our work. We have double checked an believe that there is no problem in the equation. Our proof is below:}
+  \textit{We thank the reviewer for the careful assessment of our work. We have double-checked and believe that there is no problem in the equation. Our proof is below:}
 

### 2025-02-04 revised_v1/response_editor.tex
Grammar check on Overleaf

@@
 
-We thank you for handling this manuscript and coordinating the reviews. We found both reviews very helpful in improving the manuscript.
+\vspace{.2cm}
+
+We thank you for handling this manuscript and coordinating the reviews. Both reviews were very helpful in improving the manuscript, and we have worked to address the comments as best as possible.
 
@@
 
-Following Reviewer 1’s feedback, we have revised the first objective of the manuscript to better describe the conditions under which covariance between sediment transport capacity and water discharge may occur.
+Following the feedback of reviewer 1, we have revised the first objective of the manuscript to describe better the conditions under which covariance between sediment transport capacity and water discharge may occur.
 To further develop this point, we conducted additional experiments using a series of smoothed hydrographs to better quantify the conditions that lead to covariance.
@@
 
-At several points in the review, it was noted that a particular point had already been established.
-If this reflects a substantial oversight of the literature on our part, we sincerely appreciate the reviewer bringing it to our attention.
+At several points in the review, the reviewer noted that a particular point had already been established.
+If this reflects a substantial oversight of the literature on our part, we thank the reviewer for bringing it to our attention.
 However, we would be grateful if the reviewer could provide specific references in these instances to help us fully address their concerns.
@@
 
-Regarding Reviewer 2, we believe that their specific comments about the model have been addressed by generalizing the equations in question.
+Regarding the second reviewer, we believe that their specific comments about the model have been addressed by generalizing the equations in question.
 We acknowledge their critique concerning the accumulation of water flow along the glacier’s length.
@@
 
-\bibliographystyle{apalike} 
-\bibliography{../PaperLib.bib}
 \end{document}

### 2025-02-04 Delaneyetal2024.tex
Grammar check on Overleaf

--- 2025-02-03 Delaneyetal2024.tex
@@
   Yet, results and formulations of subglacial sediment transport capacity show that its variability can approach that of subaerial systems when subglacial channel size is in equilibrium with water discharge.
-  We present these results in the context of different formulations of the relationship between water discharge and sediment transport capacity based assumptions of subglacial and subaerial flow regimes.
-  The implications of these findings help to evaluated sediment discharge from glaciers with different hydro-climatic forcings.
+  We present these results in the context of different formulations of the relationship between water discharge and sediment transport capacity-based assumptions of subglacial and subaerial flow regimes.
+  The implications of these findings help to evaluate sediment discharge from glaciers with different hydro-climatic forcings.
   These findings can improve the interpretation of sediment discharge records in glacierized catchments.
@@
 \icesheet{}  is exemplified by thick ice  \citep[$h_{ice}$= $700$\,\unit{m}; ][]{morlighem2017}, high water discharge ($\sim\,300$\,\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge at Leverett glacier (Figure~\ref{fig:model_outs}\, e).
-We note that the subglacial results apply to pressurize subglacial channels, not depressurized one that can occur below glacier termini \citep{perolo2018}.
+We note that the subglacial results apply to pressurized subglacial channels, not depressurized ones that can occur underneath glacier termini \citep{perolo2018}.
 
@@
 Note, however, that pipe flow assumptions would cause sediment discharge capacity to covary with water discharge, as is not the case in normal R-channels (Figures~\ref{fig:Qw_vari}~and~\ref{fig:chan_types}\,a\,and\,b).
-Covariance also occurs for steady state R-channels (Figure \ref{fig:chan_types} d).
+Covariance also occurs for steady-state R-channels (Figure \ref{fig:chan_types} d).
 
@@
 Sediment exhaustion through repeatedly crossing the mobilization threshold may explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on the glaciers' basal shear stress, a proxy for bedrock erosion, rather than glacier melt \citep{overeem2017}.
-More generally, the same process could also result in sediment discharge's strong dependence on sediment production from glacial sliding, which is dependent of basal shear stress of the glacier \citep{herman2015,koppes2015}.
+More generally, the same process could also result in sediment discharge's strong dependence on sediment production from glacial sliding, which is dependent on basal shear stress of the glacier \citep{herman2015,koppes2015}.
 
@@
 
-The manuscript's first objective is to establish the conditions for  water discharge to covary with subglacial sediment transport capacity.
+The manuscript's first objective is to establish the conditions where water discharge covaries with subglacial sediment transport capacity.
 In subglacial channels, the timing of peak water velocity and sediment transport capacity occurs before peak water discharge during a discharge event on hydrographs from an alpine and ice sheet catchment, due to evolving channel size.
 In subaerial channels, the timing of peak sediment transport capacity and water discharge coincide.
-Results here suggest that, even in a transport-limited subglacial system, an variable relationship between water and subglacial sediment discharge.
+Results here suggest that, even in a transport-limited subglacial system, a variable relationship between water and subglacial sediment discharge.
 Water discharge variations in both Greenland Ice Sheet and Alpine are substantial enough to cause this inconsistent relationship, even when water discharge records are smoothed.
@@
 The poor constraints on subglacial sediment size and sorting processes make it difficult to link shear stress to sediment transport capacity.
-New methods are being developed to quantifying these processes \citep{jenkin2023},  better informing the response of sediment transport capacity to water discharge forcing in subglacial environments.
+New methods are being developed to quantify these processes \citep{jenkin2023},  better informing the response of sediment transport capacity to water discharge forcing in subglacial environments.
 
@@
 
-\competinginterests{The one or more of the (co-) authors is a member of The Cryosphere editorial board and is not involved in the review process.}
+\competinginterests{One or more of the (co-) authors is a member of The Cryosphere editorial board and is not involved in the review process.}
 

### 2025-02-07 Delaneyetal2024.tex
small changes

@@
   As a result, rapid changes in water discharge mainly result in velocity variations, as the channel geometry evolves slowly.
-  Here, we present formulations of sediment transport capacity in different channel types and apply subglacial and subaerial hydraulics models to hydrographs from an Alpine glacier and the Greenland Ice Sheet.
+  Here, we present formulations of sediment transport capacity in different channel types and apply subglacial and subaerial hydraulics models to hydrographs from an Alpine glacier and a catchment of the Greenland Ice Sheet.
   Numerical experiments show that the changing channel size results in sediment transport capacity peaking before the maximum water discharge.
@@
 Despite the different physical processes between the subaerial and subglacial systems, observed water discharge and sediment export are often compared in contemporary glacierized catchments \citep[e.g.][]{willis1996,hodson1998,pearce2003,richards2003,swift2005,chu2009,tedstone2012,chu2012,overeem2017,delaney2018,swift2021,lu2022,andresen2024}.
-Many of these studies discuss the disparity between water discharge and sediment export, especially given the role of sediment access or production.
-Yet, a variable relationship between water discharge and sediment export could be expected from glaciers with evolving conduit size in a transport-limited regime, instead of sediment availability.
+Many of these studies discuss the disparity between water discharge and sediment export, especially given the role of sediment availability or production.
+Yet, a variable relationship between water discharge and sediment export could be expected from glaciers with evolving conduit size in a transport-limited regime.
 The response of sediment transport capacity to water discharge variations may affect the interpretation of sediment transport records in these glacierized catchments \citep[e.g.][]{ganti2016,mancini2023}.
@@
 The Leverett Glacier in Greenland (labeled \icesheet{}) serves as the ice sheet setting.
-Water discharge was measured roughly $2$\unit{km} downstream from the terminus ($67^\circ\,03'\,5''$ N, $50^\circ\,12'\,59''$ W), at a $5$\,\unit{min} time interval from May 28, 2012 to August 8, 2012 \citet[][Figure~\ref{fig:model_outs}\,b]{tedstone2013}.
+Water discharge was measured roughly $2$\unit{km} downstream from the terminus ($67^\circ\,03'\,5''$ N, $50^\circ\,12'\,59''$ W), at a $5$\,\unit{min} time interval from May 28, 2012 to August 8, 2012 \citep[][; Figure~\ref{fig:model_outs}\,b]{tedstone2013}.
 
@@
 The parameters' range tested in Section~\ref{sect:ensemble} covers a likely span of viable shear stresses in both subglacial and subaerial channels.
-The greater sediment transport capacity in the subglacial channels implies that sediment grain size underneath subglacial channels is likely larger than subaerial counterparts.
-Note, however, that the water velocities are similar in both channel types.
+The greater sediment transport capacity in the subglacial channels implies that sediment grain size underneath subglacial channels is likely larger than subaerial counterparts, as smaller clasts would be evacuated.
+Note, however, that the mean water velocities are similar in both channel types.
 
@@
 \label{sect:scaling}
+The numerical experiments above consider the size evolution of subglacial channels and demonstrate that for these hydrographs subglacial sediment transport variability is greater than its subaerial counterpart (Section~\ref{sect:ensemble}).
+Additionally, results demonstrate the impact of water discharge variability on sediment transport capacity.
+Here, we discuss and compare the sediment transport behavior of different channel types as they respond to water discharge, channel shape, and hydraulic gradient, extending the formulations presented in \citet{alley1997}.
 
@@
 \end{figure}
-
-The numerical experiments above consider the size evolution of subglacial channels and demonstrate that for these hydrographs subglacial sediment transport variability is greater than its subaerial counterpart (Section~\ref{sect:ensemble}).
-Additionally, results demonstrate the impact of water discharge variability on sediment transport capacity.
-Here, we discuss and compare the sediment transport behavior of different channel types as they respond to water discharge, channel shape, and hydraulic gradient, extending the formulations presented in \citet{alley1997}.
 
@@
 Catchments with reduced water discharge variability may experience less variability in sediment transport capacity, shown by the \alpine{} and \icesheet{} model outputs.
-Indeed, the more decoupled and variable relationship between model outputs and water discharge in \alpine{} results from the relatively larger discharge variations on sub-daily to weekly timescales (Figure~\ref{fig:Qw_vari}).
+Indeed, the more decoupled and variable relationship exists between model outputs and water discharge in \alpine{} results from the relatively larger discharge variations on sub-daily to weekly timescales (Figure~\ref{fig:Qw_vari}).
 This occurs as the subglacial channel's size evolves slowly compared to the variations in water discharge (Figure~\ref{fig:chan_types} b).
@@
 Results here suggest that, even in a transport-limited subglacial system, a variable relationship between water and subglacial sediment discharge.
-Water discharge variations in both Greenland Ice Sheet and Alpine are substantial enough to cause this inconsistent relationship, even when water discharge records are smoothed.
+Water discharge variations in potentially  representative Greenland Ice Sheet and Alpine catchments are substantial enough to cause this inconsistent relationship, even when water discharge records are smoothed.
 This variable relationship presents a challenge in linking hydro-climatic conditions or events to sediment export from glaciers.
@@
 
-Few observations of the subglacial environment hinder our ability to quantify processes such as
+Few observations of the subglacial environment hinders our ability to quantify processes such as
 the shape of subglacial channels and the response time of subglacial channels to water discharge variations.

### 2025-02-17 revised_v1/response_R1.tex
Changes from OL
preparing diff file

@@
 \textit{We acknowledge the importance of removing interpretation from the results section and have made the necessary revisions.
-Additionally, we will moved Section 4.3, which contains the algebraic formulations, to the discussion section, as we believe that the pipe flow and steady-state R-channel scenarios serve as useful end members for interpreting the results.
 }
@@
 We have also noted the comment about results and interpretation.
-These have been removed from the manuscript}
+These have been removed from the manuscript.}
 

### 2025-02-17 revised_v1/response_R2.tex
small changes
preparing diff file

@@
 
-  \textit{We appreciate these comments. With regard to the first point about channel width evolution in the subaerial model, this matter is confronted in Sections 4.2 and 4.3 (now 5.1).
+  \textit{We appreciate these comments. With regard to the first point about channel width evolution in the subaerial model, this matter is confronted in Sections 4.2 and 4.3.
     Here we show that across a range of exponents and relationships between channel cross-section and water discharge, subglacial systems are still more sensitive to water discharge variations than subaerial ones.
-    This increased variability occurs across a range of $\alpha$ values that describe the relationship between channel width and water discharge as shown in Figure 6 (now 7).
+    This increased variability in subglacial sediment transport capacity occurs across a range of $\alpha$ values that describe the relationship between channel width and water discharge as shown in Figure 6 (will be 7).
     As a result, we believe that these results hold across a range of water discharge- channel width relationships, including ones where $\alpha$ evolves in time.
-    The exception, shown in Section 4.3 (now 6.1), is that when water discharge evolves more slowly than channel size.
+    The exception, shown in Section 4.3, is that when water discharge evolves more slowly than channel size.
 Here, the behavior between subaerial and subglacial is very similar, because channel areas in both channel types adjust to discharge variations.}
@@
 
-  \textit{With respect to the last sentence, it is true that the lead author feels this way occasionally. Regardless, we think valuable findings are within this manuscript.}
+  \textit{With respect to the last sentence, it is true that the lead author feels this way occasionally. Regardless, we think several valuable and unique findings are within this manuscript.}
 
@@
 
+  Hydraulic diameter:
   \begin{equation}
@@
 
+  Wetter perimeter of the conduit \citep[][pg. 69]{hooke1990}:
+
   \begin{equation}
-    P_w = R \beta +2R \sin\frac{\beta}{2}= R(\beta +2\sin \frac{\beta}{2} ~~~(II)
+    P_w = R \beta +2R \sin\frac{\beta}{2}= R(\beta +2\sin \frac{\beta}{2}) ~~~(II)
   \end{equation}
 
+  Cross sectional area of the conduit \citep[][pg. 69]{hooke1990}:
   \begin{equation}
@@
     \begin{equation}
-      \frac{D_h}{4S_g} = \frac{1}{R\beta + 2R\sin\frac{\beta}{2}}
+      \frac{D_h}{4S_g} = \frac{1}{R\beta + 2R\sin\frac{\beta}{2}} \quad I \mathrm{into~} II 
     \end{equation}
@@
   \begin{equation}
-    S_g = \frac{16 S_g^2}{D_h^2 (\beta +2\sin\frac{\beta}{2})^2} ~\frac{1}{2}(\beta-\sin\beta)
+    S_g = \frac{16 S_g^2}{D_h^2 (\beta +2\sin\frac{\beta}{2})^2} ~~~\frac{1}{2}(\beta-\sin\beta)
   \end{equation}
@@
   \begin{equation}
-    1 = \frac{8 S_g}{D_h^2 (\beta +2\sin\frac{\beta}{2})^2} (\beta-\sin\beta)
+    1 = \frac{8 S_g}{D_h^2 (\beta +2\sin\frac{\beta}{2})^2} ~~~ (\beta-\sin\beta)
   \end{equation}
@@
   \begin{equation}
-    \frac{D_h^2}{8} \frac{(\beta+2\sin\frac{\beta}{2})^2}{\beta - \sin \beta} = S_g
+    \frac{D_h^2}{8} ~~~ \frac{(\beta+2\sin\frac{\beta}{2})^2}{\beta - \sin \beta} = S_g
   \end{equation}
@@
    \begin{equation}
-    \frac{D_h^2 ((\frac{\beta}{2}+\sin \frac{\beta}{2}))^2}{2(\beta - \sin \beta)} = S_g
+    \frac{D_h^2 (\frac{\beta}{2}+\sin \frac{\beta}{2})^2}{2(\beta - \sin \beta)} = S_g
   \end{equation}

### 2025-02-17 Delaneyetal2024.tex
preparing diff file

@@
 \subsection{Timing and covariance of subaerial and subglacial  sediment transport capacity variations}
-
+\label{sect:timing}
 The first numerical experiment aims to quantify the timing and covariance of the subglacial model outputs with respect to water discharge in response to different seasonal evolutions and peaks (Figure~\ref{fig:model_outs}).
@@
 \FloatBarrier
-\section{Experiment limitations}
-\FloatBarrier
-
-Several aspects of the models and experiment design make the comparison of the subaerial and subglacial systems difficult. 
-The lumped nature of the models means that they operate independently of the upstream drainage network.
-Additionally, we omit analysis of suspended sediment discharge records due to their dependence on sediment supply in capturing variations in sediment discharge, in addition to sediment transport capacity \citep[e.g.][]{delaney2019}.
-In reality, processes such as sediment access are usually important in controlling sediment export in glacierized catchments \citep[e.g.][]{herman2015,vergara2022}.
-Furthermore, meltwater can be distributed to flow through several adjacent conduits impacting the sediment transport capacity in each \citep[e.g.][]{werder2013,hewitt2019,delaney2023}.
-The lumped models used here isolate the relationship between water discharge and sediment transport capacity in subglacial and subaerial systems. 
-However, they neglect more complex, yet important, spatially distributed processes.
-
-Different hydraulic gradients control the velocity and sediment transport capacity in the subglacial and subaerial cases (Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}). 
-The subglacial one is generally much steeper \citep{alley1997}.
-This results in the shear stresses and width-integrated shear stresses across the channel bed that controls sediment transport capacity being much greater in subglacial channels (Figure~\ref{fig:model_outs}).
-The parameters' range tested in Section~\ref{sect:ensemble} covers a likely span of viable shear stresses in both subglacial and subaerial channels.
-The greater sediment transport capacity in the subglacial channels implies that sediment grain size underneath subglacial channels is likely larger than subaerial counterparts, as smaller clasts would be evacuated.
-Note, however, that the mean water velocities are similar in both channel types.
-
-The channel width and size variations in subglacial channels presented here also occur in subaerial channels in response to water discharge \citep{phillips2016}, likely over decadal timescales or in response to individual extreme events \citep[e.g.][]{slater2013,dean2013}.
-These timescales are likely considerably longer than the continuous changes to the subglacial channel's response time of days.
-Furthermore, observations from subaerial systems can show a co-varying relationship between water discharge and sediment transport in time \citep[e.g.][]{schmidt2006,pitlick2021}.
-Even so, if channel width changes occur in subaerial channels, then the values of $\alpha$ in Equation~\ref{eq:wcf} could change in time.
-$\alpha$ must be less than $1$, and even this would cause the exponent on $Q$ in the  $w_g\,\tau_{g}$ relationships to remain substantially below that of pipeflow (Section ~\ref{sect:scaling}, Table\,\ref{tab:eqs1} and \ref{tab:Qs})).
-
-Width-integrated shear stress is examined in the two numerical experiments as it does not have a dependence on grain size, unlike sediment transport relationships \citep[][Section \ref{sect:scaling}]{meyer1948}.
-This makes comparison between the two systems simpler. 
-Preferential transport of smaller sediment clasts and the input of upstream sediment impact sediment transport capacity as grain size evolves \citep[e.g.][]{gomez1983}.
-These processes are only beginning to be evaluated underneath glaciers \citep{aitken2024}, but can be important in subaerial systems.
-Subglacial sorting processes could be an additional source of variability in sediment export from glaciers with respect to water discharge, especially for bedload transport.
-
-\FloatBarrier
-\section{Discussion}
-
-\FloatBarrier
 \subsection{Sediment transport scaling in different channel types}
@@
 
-With these assumptions, the Darcy-Weisbach equation~\eqref{eq:DW} can be solved for the not-fixed quantity: $H$ for a subaerial channel, $S$ for a R-channel, and $\Psi$ for pipe-flow.
-Then, using equations~\eqref{eq:tau-omega}, the shear stress and stream power can be calculated.
-These results are summarised in Table~\ref{tab:eqs1}.
-The results of the 12 combinations of shear stress and sediment transport capacity are presented in Table~\ref{tab:Qs} and also in Table~S1 where the fractions in the exponents are approximately given by decimal numbers for ease of comparison.
-
 \begin{table}[hbt!]
@@
 
-Table~\ref{tab:Qs} shows the scaling of the proxy $w\tau$ as well as $Q_s$ of the MPM, EH, and Bagnold sediment transport formulas with respect to $Q$, $\Psi$ or $S$ for the three different channel types and  $\alpha$ values.
-Remarkably, the Bagnold formula has a negative exponent for $f$ in all but the pipe-flow channel type.
-The total transport formula EH gives a slightly stronger dependence on all variables due to the larger exponent on $\tau$ of $\frac{5}{2}$ versus  $\frac{3}{2}$ for the MPM (Equation~\ref{eq:Qs_eq}).
-Albeit, the sediment transport response in pipe flow for the Bagnold case is close to EH.
-Conversely, the sediment transport proxies $w_g\,\tau_g$ and $w_f\,\tau_f$ used previously scale only as $Q^2$ for pipe-flow, whereas the sediment transport capacity scales at least with $Q^3$.
-The exponent for the width scaling $\alpha$ only impacts the relationship between sediment transport and water discharge in the EH relation in any meaningful way.
-However, for the value of $\alpha$ around $\frac{1}{3}$ (a value appropriate for most streams), the exponent on $Q$ for EH is only slightly greater than the other transport relations.
-
-The sediment discharge capacity in the steady state R-channel scales very similarly to the subaerial channel case for all relations and virtually identically for the $\alpha=\frac{1}{3}$ case (Table~\ref{tab:Qs}).
-However, the head gradients $\Psi$ are likely higher for comparable $Q$ in an ice-sheet marginal or alpine glacier setting than in a subaerial channel \citep{alley1997}.
-Thus, sediment transport capacity is likely higher in a steady-state R-channel.
+With these assumptions, the Darcy-Weisbach equation~\eqref{eq:DW} can be solved for the not-fixed quantity: $H$ for a subaerial channel, $S$ for a R-channel, and $\Psi$ for pipe-flow.
+Then, using equations~\eqref{eq:tau-omega}, the shear stress and stream power can be calculated.
+These results are summarised in Table~\ref{tab:eqs1}.
+The results of the 12 combinations of shear stress and sediment transport capacity are presented in Table~\ref{tab:Qs} and also in Table~S1 where the fractions in the exponents are approximately given by decimal numbers for ease of comparison.
 
@@
 
-R-channels rarely operate in a steady state with variations in water discharge, especially during severe rain or melt is too short for the channel to reach a steady state (Figures~\ref{fig:model_outs} and \ref{fig:Qw_vari}).
-In these cases with high water discharge variability, channels can behave more like a pipe of fixed cross-section.
-Here the cross-section responds to a characteristic discharge, but variations in water discharge deviate substantially from the flow conditions responsible for the channel size \citep[i.e.
-diurnal water discharge variations in \alpine{} case Figure \ref{fig:Qw_vari}; e.g.][]{gimbert2016}.
-Table~\ref{tab:Qs} shows that sediment transport in pipe-flow scales much more severely with discharge; the exponent on $Q$ being between $3$ and $5$, compared to the other two channel types when that exponent is at most $\frac{5}{3}$.
-Thus fluctuations of discharge on short timescales (on the order of a day; Figure~\ref{fig:multi_run}) have the potential to cause conditions with very high sediment transport capacities.
-Alternatively, at low water discharges these channels could become depressurized and transition to subearial flow \citep{perolo2018}.
-These sediment transport capacities variations are of far higher magnitude than those of subaerial channels \citep[][]{alley1997}.
-Note, however, that pipe flow assumptions would cause sediment discharge capacity to covary with water discharge, as is not the case in normal R-channels (Figures~\ref{fig:Qw_vari}~and~\ref{fig:chan_types}\,a\,and\,b).
-Covariance also occurs for steady-state R-channels (Figure \ref{fig:chan_types} d).
+Table~\ref{tab:Qs} shows the scaling of the proxy $w\tau$ as well as $Q_s$ of the MPM, EH, and Bagnold sediment transport formulas with respect to $Q$, $\Psi$ or $S$ for the three different channel types and  $\alpha$ values.
+Sediment transport in pipe-flow scales much more severely with discharge (Table~\ref{tab:Qs}); the exponent on $Q$ being between $3$ and $5$, compared to the other two channel types when that exponent is at most $\frac{5}{3}$.
+Remarkably, the Bagnold formula has a negative exponent for $f$ in all but the pipe-flow channel type.
+The total transport formula EH gives a slightly stronger dependence on all variables due to the larger exponent on $\tau$ of $\frac{5}{2}$ versus  $\frac{3}{2}$ for the MPM (Equation~\ref{eq:Qs_eq}).
+Albeit, the sediment transport response in pipe flow for the Bagnold case is close to EH.
+Conversely, the sediment transport proxies $w_g\,\tau_g$ and $w_f\,\tau_f$ used previously scale only as $Q^2$ for pipe-flow, whereas the sediment transport capacity scales at least with $Q^3$.
+The exponent for the width scaling $\alpha$ only impacts the relationship between sediment transport and water discharge in the EH relation in any meaningful way.
+However, for the value of $\alpha$ around $\frac{1}{3}$ (a value appropriate for most streams), the exponent on $Q$ for EH is only slightly greater than the other transport relations.
 
+The sediment discharge capacity in the steady state R-channel scales very similarly to the subaerial channel case for all relations and virtually identically for the $\alpha=\frac{1}{3}$ case (Table~\ref{tab:Qs}).
+However, the head gradients $\Psi$ are likely higher for comparable $Q$ in an ice-sheet marginal or alpine glacier setting than in a subaerial channel \citep{alley1997}.
+Thus, sediment transport capacity is likely higher in a steady-state R-channel.
+
+\FloatBarrier
+\section{Experiment limitations}
+\FloatBarrier
+
+Several aspects of the models and experiment design make the comparison of the subaerial and subglacial systems difficult. 
+The lumped nature of the models means that they operate independently of the upstream drainage network.
+Additionally, we omit analysis of suspended sediment discharge records due to their dependence on sediment supply in capturing variations in sediment discharge, in addition to sediment transport capacity \citep[e.g.][]{delaney2019}.
+In reality, processes such as sediment access are usually important in controlling sediment export in glacierized catchments \citep[e.g.][]{herman2015,vergara2022}.
+Furthermore, meltwater can be distributed to flow through several adjacent conduits impacting the sediment transport capacity in each \citep[e.g.][]{werder2013,hewitt2019,delaney2023}.
+The lumped models used here isolate the relationship between water discharge and sediment transport capacity in subglacial and subaerial systems. 
+However, they neglect more complex, yet important, spatially distributed processes.
+
+Different hydraulic gradients control the velocity and sediment transport capacity in the subglacial and subaerial cases (Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}). 
+The subglacial one is generally much steeper \citep{alley1997}.
+This results in the shear stresses and width-integrated shear stresses across the channel bed that controls sediment transport capacity being much greater in subglacial channels (Figure~\ref{fig:model_outs}).
+The parameters' range tested in Section~\ref{sect:ensemble} covers a likely span of viable shear stresses in both subglacial and subaerial channels.
+The greater sediment transport capacity in the subglacial channels implies that sediment grain size underneath subglacial channels is likely larger than subaerial counterparts, as smaller clasts would be evacuated.
+Note, however, that the mean water velocities are similar in both channel types.
+
+The channel width and size variations in subglacial channels presented here also occur in subaerial channels in response to water discharge \citep{phillips2016}, likely over decadal timescales or in response to individual extreme events \citep[e.g.][]{slater2013,dean2013}.
+These timescales are likely considerably longer than the continuous changes to the subglacial channel's response time of days.
+Furthermore, observations from subaerial systems can show a co-varying relationship between water discharge and sediment transport in time \citep[e.g.][]{schmidt2006,pitlick2021}.
+Even so, if channel width changes occur in subaerial channels, then the values of $\alpha$ in Equation~\ref{eq:wcf} could change in time.
+$\alpha$ must be less than $1$, and even this would cause the exponent on $Q$ in the  $w_g\,\tau_{g}$ relationships to remain substantially below that of pipeflow (Section ~\ref{sect:scaling}, Table\,\ref{tab:eqs1} and \ref{tab:Qs})).
+
+Width-integrated shear stress is examined in the two numerical experiments as it does not have a dependence on grain size, unlike sediment transport relationships \citep[][Section \ref{sect:scaling}]{meyer1948}.
+This makes comparison between the two systems simpler. 
+Preferential transport of smaller sediment clasts and the input of upstream sediment impact sediment transport capacity as grain size evolves \citep[e.g.][]{gomez1983}.
+These processes are only beginning to be evaluated underneath glaciers \citep{aitken2024}, but can be important in subaerial systems.
+Subglacial sorting processes could be an additional source of variability in sediment export from glaciers with respect to water discharge, especially for bedload transport.
+
+\FloatBarrier
+\section{Discussion}
 \subsection{Increased variability in  sediment transport capacity in subglacial systems}
@@
 The exponent greater than $1$ magnifies sediment discharge variability beyond the variable sediment transport parameters described above (Figure~\ref{fig:multi_run}; Table~\ref{tab:Qs}).
+
+R-channels rarely operate in a steady state with variations in water discharge, especially during severe rain or melt is too short for the channel to reach a steady state (Figures~\ref{fig:model_outs} and \ref{fig:Qw_vari}).
+In these cases with high water discharge variability, channels can behave more like a pipe of fixed cross-section.
+Here the cross-section responds to a characteristic discharge, but variations in water discharge deviate substantially from the flow conditions responsible for the channel size \citep[i.e.
+diurnal water discharge variations in \alpine{} case Figure \ref{fig:Qw_vari}; e.g.][]{gimbert2016}.
+Thus fluctuations of discharge on short timescales (on the order of a day; Figure~\ref{fig:multi_run}) have the potential to cause conditions with very high sediment transport capacities.
+Alternatively, at low water discharges these channels could become depressurized and transition to subearial flow \citep{perolo2018}.
+These sediment transport capacities variations are of far higher magnitude than those of subaerial channels \citep[][]{alley1997}.
+Note, however, that pipe flow assumptions would cause sediment discharge capacity to covary with water discharge, as is not the case in normal R-channels (Figures~\ref{fig:Qw_vari}~and~\ref{fig:chan_types}\,a\,and\,b).
+Covariance also occurs for steady-state R-channels (Section \ref{sect:timing}; Figure \ref{fig:chan_types} d).
 
